### 2013-10-08 kipchak-lrec-2013/kipchak-paper.tex
new paper
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@47967 72bbbca6-d526-0410-a7d9-f06f51895060

\documentstyle{article}

\usepackage{fontspec}
\usepackage{xunicode}
\usepackage{xltxtra}

\title{Finite-state morphological transducers for three Kipchak languages}

\begin{document}

\maketitle

\begin{abstract}
Hargle, bargle.
\end{abstract}

\section{Introduction}

\section{Methodology}

\section{Evaluation}

\section{Conclusions}

\end{document}

### 2013-10-08 kipchak-lrec-2013/kipchak-paper.tex
skelington
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@47968 72bbbca6-d526-0410-a7d9-f06f51895060

@@
-\documentstyle{article}
+\documentclass{article}
 
@@
 \usepackage{xltxtra}
+\usepackage{natbib}
+
+\bibdata{kipchak-paper}
 
 \title{Finite-state morphological transducers for three Kipchak languages}
+
+\author{Jonathan North Washington \\
+Departments of Linguistics and Central Eurasian Studies\\
+Indiana University\\
+Bloomington, IN 47405 (USA)\\
+\texttt{jonwashi@indiana.edu} \and
+Ilnar Salimzyanov  \\
+Institut für Maschinelle Sprachverarbeitung \\
+Universität Stuttgart\\
+Stuttgart (Germany) \\
+\texttt{ilnar@ilnar} \and 
+Francis M. Tyers\\
+Departament de Llenguatges i Sistemes Informàtics \\  
+Universitat d'Alacant\\
+E-03071 Alacant (Spain)\\
+\texttt{ftyers@dlsi.ua.es} 
+}
 
@@
 
+\cite{washington2012}
+\cite{salimzyanov2013}
+
+\section{Languages}
+
+\subsection{Kazakh}
+
+\subsection{Tatar}
+
+\subsection{Kumyk}
+
 \section{Methodology}
@@
 
+\section{Future work}
+
 \section{Conclusions}
 
+\bibliographystyle{lrec2012}
+\bibliography{paper}
+
 \end{document}

### 2013-10-10 kipchak-lrec-2013/kipchak-paper.tex
more
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@47969 72bbbca6-d526-0410-a7d9-f06f51895060
futwork
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@47974 72bbbca6-d526-0410-a7d9-f06f51895060
some aper update
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@47980 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{natbib}
+\usepackage{multirow}
 
@@
 
+\cite{bammatov1960}
+
 \section{Methodology}
@@
 
+\begin{table}
+\begin{center}
+\begin{tabular}{lrrr}
+		\hline
+\multirow{2}{*}{\textbf{Part of speech}} & \multicolumn{3}{|c|}{\textbf{Number of stems}} \\
+                        & Kazakh & Tatar & Kumyk \\
+		\hline
+		Noun & - & - & - \\
+		Verb & - & - & - \\
+		Adjective & - & - & - \\
+		Proper noun & - & - & - \\
+		Adverb & - & - & - \\
+		Numeral & - & - & - \\
+		Conjunction & - & - & - \\
+		Postposition & - & - & - \\
+		Pronoun & - & - & - \\
+		Determiner & - & - & - \\
+		\hline
+		Total: & - & - & - \\
+		\hline
+\end{tabular}
+\end{center}
+
+\end{table}
+
+\begin{table}
+\begin{center}
+\begin{tabular}{lr}
+\textbf{Language} & - \\
+\hline
+Kazakh & - \\
+Tatar & - \\
+Kumyk & - \\
+\hline
+\end{tabular}
+ \caption{Naïve coverage}
+ \label{table:coverage}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\begin{tabular}{lrr}
+\textbf{Language} & \textbf{Precision} & \textbf{Recall} \\
+\hline
+Kazakh & - &  - \\
+Tatar & - & - \\
+Kumyk & - & - \\
+\hline
+\end{tabular}
+ \caption{Precision and recall}
+ \label{table:coverage}
+\end{center}
+\end{table}
+
 \section{Future work}
+
+%Code switching
+%More languages
 
@@
 
-\bibliographystyle{lrec2012}
-\bibliography{paper}
+\bibliographystyle{apalike}
+
+\bibliography{kipchak-paper}
 

### 2013-10-11 kipchak-lrec-2013/kipchak-paper.tex
add bibitem
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@47983 72bbbca6-d526-0410-a7d9-f06f51895060
some fixes and additions
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@48028 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{multirow}
+\usepackage{fullpage}
+\usepackage{multirow}
+
+\usepackage{booktabs}
+
+\setromanfont{DejaVu Serif}
 
@@
 
-\title{Finite-state morphological transducers for three Kipchak languages}
+\title{Finite-state morphological transducers for three Kypchak languages}
 
@@
 
+The Northwestern branch of Turkic is often referred to as the Kypchak branch, and can be divided into three subbranches.
+Kumyk is a member of the Western Kypchak group, Tatar is a member of the Northern Kypchak group, and Kazakh is a member of the south Kypchak group \citep[82-83]{histofturkic}.
+The geographic distribution of the languages is shown in map \ref{}.
+
 \cite{washington2012}
 \cite{salimzyanov2013}
+\cite{bekmanova2013}
 
@@
 
+\begin{table}
+\begin{center}
+\begin{tabular}{lrr}
+\textbf{Corpus} & \textbf{Words} & \textbf{Coverage} \\
+\hline
+wikipedia & 850K &  - \\
+Äwezov & 155K & - \\
+RFERL 2010 & 3.2M & - \\
+bible & 577K & - \\
+quran & 107K & - \\
+\hline
+\end{tabular}
+ \caption{Corpora used for coverage of Kazakh}
+ \label{table:corpora-kaz}
+\end{center}
+\end{table}
+
+\begin{table}
+\begin{center}
+\begin{tabular}{llrr}
+\toprule
+\textbf{Language} & \textbf{Corpus} & \textbf{Words} & \textbf{Coverage} \\
+\midrule
+\multirow{6}{*}{Kazakh} & wikipedia 2011& 850K &  - \\
+	& Äwezov & 155K & - \\
+	& RFERL 2010 & 3.2M & - \\
+	& bible & 577K & - \\
+	& quran & 107K & - \\\cline{2-4}
+	& average & - & 90.5\% \\
+\midrule
+\multirow{5}{*}{Tatar} & wikipedia 2013 & 128K &  - \\
+	& news 2005-2011 & 4.6M & - \\
+	& new testament & 137K & - \\
+	& quran & 165K & - \\
+	& Aytmatov & 5K & - \\\cline{2-4}
+	& average & - & 89.0\% \\
+\midrule
+\multirow{4}{*}{Kumyk} & yoldash & 287K &  - \\
+	& new testament & 154K & - \\
+	& book of Genesis & 28K & - \\\cline{2-4}
+	& average & - & 88.0\% \\
+\bottomrule
+\end{tabular}
+ \caption{Corpora used for coverage tests}
+ \label{table:corpora}
+\end{center}
+\end{table}
+
 \section{Future work}

### 2013-10-11 kipchak-lrec-2013/kipchak-paper.tex
updates and fixes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@48029 72bbbca6-d526-0410-a7d9-f06f51895060

@@
-\documentclass{article}
+\documentclass[a4paper,11pt,twocolumn]{article}
+
+\usepackage{lrec2012}
+
+\usepackage{polyglossia}
+\setdefaultlanguage[variant=australian]{english}
 
@@
 \usepackage{multirow}
-\usepackage{fullpage}
+%\usepackage{fullpage}
 \usepackage{multirow}
@@
 
-\setromanfont{DejaVu Serif}
+\setromanfont{Times New Roman}
 
@@
 
+\name{Jonathan North Washington, Ilnar Salimzyanov, Francis M. Tyers}
+
+\address{Departments of Linguistics and Central Eurasian Studies\\
+Indiana University\\
+Bloomington, IN 47405 (USA)\\
+\texttt{jonwashi@indiana.edu} \and
+Institut für Maschinelle Sprachverarbeitung \\
+Universität Stuttgart\\
+Stuttgart (Germany) \\
+\texttt{ilnar@ilnar} \and 
+Departament de Llenguatges i Sistemes Informàtics \\  
+Universitat d'Alacant\\
+E-03071 Alacant (Spain)\\
+\texttt{ftyers@dlsi.ua.es} 
+}
+
+\abstract{Hargle, bargle.\\
+\Keywords{Kazakh, Tatar, Kumyk, morphology, transducer}}
+
 \begin{document}
 
-\maketitle
-
-\begin{abstract}
-Hargle, bargle.
-\end{abstract}
+\maketitleabstract{}
 
@@
 
-\bibliographystyle{apalike}
-
+\bibliographystyle{lrec2012}
 \bibliography{kipchak-paper}

### 2013-10-14 kipchak-lrec-2013/kipchak-paper.tex
more notes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@48037 72bbbca6-d526-0410-a7d9-f06f51895060
more work
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@48079 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-\name{Jonathan North Washington, Ilnar Salimzyanov, Francis M. Tyers}
+\name{Jonathan North Washington$^\dagger$, Ilnar Salimzyanov$^\ddagger$, Francis M. Tyers$^\star$}
 
-\address{Departments of Linguistics and Central Eurasian Studies\\
+\address{$^\dagger$Departments of Linguistics and Central Eurasian Studies\\
 Indiana University\\
@@
 \texttt{jonwashi@indiana.edu} \and
-Institut für Maschinelle Sprachverarbeitung \\
+$^\ddagger$Institut für Maschinelle Sprachverarbeitung \\
 Universität Stuttgart\\
@@
 \texttt{ilnar@ilnar} \and 
-Departament de Llenguatges i Sistemes Informàtics \\  
+$^\star$Departament de Llenguatges i Sistemes Informàtics \\  
 Universitat d'Alacant\\
@@
 \cite{bammatov1960}
+\cite{olmesov2000}
 
@@
 
-\section{Evaluation}
+% Noun morphotactics -- is basically equivalent in all three, except for archiphonemes
+% Adjective categorisation
+% Adverb categorisation
+% "Irregular" harmony, caused by orthography
+% Non-finite verb form categorisation "converbs"
+
+% Handling numerals and acronyms.
+
+\subsection{Development effort}
+% how long it took
+% make a graph for the kumyk one
+
+\subsection{Statistics}
 
@@
 \begin{center}
-\begin{tabular}{lrrr}
+\begin{tabular}{|l|rrr|}
 		\hline
-\multirow{2}{*}{\textbf{Part of speech}} & \multicolumn{3}{|c|}{\textbf{Number of stems}} \\
+\multirow{2}{*}{\textbf{Part of speech}} & \multicolumn{3}{|c|}{\textbf{Number of stems}} \\ \cline{2-4}
                         & Kazakh & Tatar & Kumyk \\
@@
 \end{tabular}
+ \caption{Number of stems in each of the categories}
+ \label{table:coverage}
 \end{center}
@@
 \end{table}
+
+\section{Evaluation}
+
+We have evaluated the morphological analysers in two ways. The first was by calculating the naïve coverage and mean ambiguity 
+on freely available corpora. The second was by performing an evaluation of precision and recall on some 
+smaller, hand-validated test sets.
+
+\subsection{Corpora}
 
@@
  \label{table:coverage}
-\end{center}
-\end{table}
-
-\begin{table}
-\begin{center}
-\begin{tabular}{lrr}
-\textbf{Corpus} & \textbf{Words} & \textbf{Coverage} \\
-\hline
-wikipedia & 850K &  - \\
-Äwezov & 155K & - \\
-RFERL 2010 & 3.2M & - \\
-bible & 577K & - \\
-quran & 107K & - \\
-\hline
-\end{tabular}
- \caption{Corpora used for coverage of Kazakh}
- \label{table:corpora-kaz}
 \end{center}
@@
 
-%Code switching
-%More languages
+%Code switching - +example
+%More languages: Nogai, Bashkir, Karakalpak, Karachay-Balkar
 
 \section{Conclusions}
+
+\section*{Acknowledgements}
+
+We would like to thank the Google Code-in (2011) for supporting the development 
+of the Kazakh transducer and the Google Summer of Code (2012) for supporting the 
+development of both the Kazakh and the Tatar transducers.
 

### 2013-10-14 kipchak-lrec-2013/kipchak-paper.tex
more work
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@48081 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-We have evaluated the morphological analysers in two ways. The first was by calculating the naïve coverage and mean ambiguity 
+We have evaluated the morphological analysers in two ways.
+The first was by calculating the naïve coverage\footnote{Naïve coverage refers to the percentage of surface forms in a given corpora that receive at least one analysis.
+Forms counted by this measure may have other analyses which are not delivered by the transducer.} and mean ambiguity
 on freely available corpora. The second was by performing an evaluation of precision and recall on some 
@@
 
+% For kazakh+tatar tested the coverage of the analysers over 3 separate domains: encyclopaedic text, news and religion, 
+% as there was no wikipedia for kumyk we just tested news and religion. The corpora were obtained from:
+% Wikipedia:
+% RFE/RL
+% Yoldaš: 
+% New testaments:
+
+% kaz: kkwiki-20131006-pages-articles.xml.bz2
+% tat:
+
 \begin{table}
 \begin{center}
-\begin{tabular}{lr}
-\textbf{Language} & - \\
-\hline
-Kazakh & - \\
-Tatar & - \\
-Kumyk & - \\
-\hline
+\begin{tabular}{llrr}
+\toprule
+\textbf{Language} & \textbf{Corpus} & \textbf{Words} & \textbf{Coverage} \\
+\midrule
+\multirow{4}{*}{Kazakh} & Wikipedia 2013 &  -  &  - \\
+	& RFE/RL 2010 & 3.2M & - \\
+	& Bible & 577K & - \\\cline{2-4}
+	& Average & - & 90.5\% \\
+\midrule
+\multirow{4}{*}{Tatar} & Wikipedia 2013 & 128K &  - \\
+	& RFE/RL 2005--2011 & 4.6M & - \\
+	& New Testament & 137K & - \\\cline{2-4}
+	& Average & - & 89.0\% \\
+\midrule
+\multirow{3}{*}{Kumyk} & Yoldaš & 287K &  - \\
+	& New Testament & 154K & - \\\cline{2-4}
+	& Average & - & 90.1\% \\
+\bottomrule
 \end{tabular}
- \caption{Naïve coverage}
- \label{table:coverage}
+ \caption{Corpora used for naïve coverage tests}
+ \label{table:corpora}
 \end{center}
@@
  \label{table:coverage}
-\end{center}
-\end{table}
-
-\begin{table}
-\begin{center}
-\begin{tabular}{llrr}
-\toprule
-\textbf{Language} & \textbf{Corpus} & \textbf{Words} & \textbf{Coverage} \\
-\midrule
-\multirow{6}{*}{Kazakh} & wikipedia 2011& 850K &  - \\
-	& Äwezov & 155K & - \\
-	& RFERL 2010 & 3.2M & - \\
-	& bible & 577K & - \\
-	& quran & 107K & - \\\cline{2-4}
-	& average & - & 90.5\% \\
-\midrule
-\multirow{5}{*}{Tatar} & wikipedia 2013 & 128K &  - \\
-	& news 2005-2011 & 4.6M & - \\
-	& new testament & 137K & - \\
-	& quran & 165K & - \\
-	& Aytmatov & 5K & - \\\cline{2-4}
-	& average & - & 89.0\% \\
-\midrule
-\multirow{4}{*}{Kumyk} & yoldash & 287K &  - \\
-	& new testament & 154K & - \\
-	& book of Genesis & 28K & - \\\cline{2-4}
-	& average & - & 88.0\% \\
-\bottomrule
-\end{tabular}
- \caption{Corpora used for coverage tests}
- \label{table:corpora}
 \end{center}

### 2013-10-14 kipchak-lrec-2013/kipchak-paper.tex
more
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@48082 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 We would like to thank the Google Code-in (2011) for supporting the development 
-of the Kazakh transducer and the Google Summer of Code (2012) for supporting the 
+of the Kazakh transducer, and in particular the effort by Nathan Maxson. We 
+would also like to thank the Google Summer of Code (2012) for supporting the 
 development of both the Kazakh and the Tatar transducers.

### 2013-10-20 kipchak-lrec-2013/kipchak-paper.tex
a little work
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@48154 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \abstract{Hargle, bargle.\\
-\Keywords{Kazakh, Tatar, Kumyk, morphology, transducer}}
+	\Keywords{Kazakh, Tatar, Kumyk, morphology, transducer}
+}
 
@@
 
-The Northwestern branch of Turkic is often referred to as the Kypchak branch, and can be divided into three subbranches.
-Kumyk is a member of the Western Kypchak group, Tatar is a member of the Northern Kypchak group, and Kazakh is a member of the south Kypchak group \citep[82-83]{histofturkic}.
+This paper describes the development of morphological transducers for three closly related languages: Kazakh, Tatar, and Kumyk.
+
+These languages belong to the Northwestern branch of Turkic, which is often referred to as the Kypchak branch.
+This branch can be divided into three subbranches.
+Kumyk is a member of the Western Kypchak group, Tatar is a member of the Northern Kypchak group, and Kazakh is a member of the Southern Kypchak group \citep[82-83]{histofturkic}.
+As such, each of these three languages represents a different one of the three branches of Kypchak.
 The geographic distribution of the languages is shown in map \ref{}.
+
+In a linguistic sense, these languages have different amounts of influence from other Turkic branches (e.g., moderate Oguz (SE) influence in the Western group, slight Oguz influence in the Northern group) and from Mongolic languages (moderate influence on the Southern group, lighter in the other groups), and all have heavy influence from Persian.
 
@@
 
+%FIXME: background on how, when, and with what support were developed?
+The transducers for these languages
+
 \section{Languages}
@@
 \subsection{Kazakh}
+Kazakh has approximately 12 million speakers
 
 \subsection{Tatar}
+Tatar has approximately  million speakers
 
 \subsection{Kumyk}
+Kumyk has approximately  million speakers
 

### 2013-10-20 kipchak-lrec-2013/stems.sh
count all stems
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@48155 72bbbca6-d526-0410-a7d9-f06f51895060

@@ -1,0 +1,14 @@
+#!/bin/bash
+
+pushd $APERTIUMPATH/languages/apertium-kum/dev/
+	./countstems.sh
+popd
+
+pushd $APERTIUMPATH/languages/apertium-kaz/dev/
+	./countstems.sh
+popd
+
+pushd $APERTIUMPATH/languages/apertium-kaz/dev/
+	./countstems.sh
+popd
+

### 2013-10-20 kipchak-lrec-2013/kipchak-paper.tex
kaz and tat population estimates
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@48156 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \subsection{Kazakh}
-Kazakh has approximately 12 million speakers
+Kazakh is primarily spoken in Kazakhstan, where it is the national language, sharing official status with Russian as an official language.
+Large communities of native speakers also exist in China, neighbouring Central-Eurasian republics, and Mongolia.
+Estimates of the total number of speakers range from 8 million \citep{ethnologue} to 11 million \citet{encyclopedin} people.
 
 \subsection{Tatar}
-Tatar has approximately  million speakers
+Tatar is spoken in and around Tatarstan by approximately 5.4 million people \citep{ethnologue}.
+It is co-official with Russian in Tatarstan --- a republic within Russia.
+A majority of native speakers of both languages are bilingual in Russian.
+%, and the standard varieties our system FIXMEs are written in Cyrillic.
 
 \subsection{Kumyk}
-Kumyk has approximately  million speakers
+Kumyk has approximately 430 thousand speakers \cite{ethnologue}
 

### 2013-10-20 kipchak-lrec-2013/kipchak-paper.tex
blah
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@48157 72bbbca6-d526-0410-a7d9-f06f51895060
more
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@48158 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-In a linguistic sense, these languages have different amounts of influence from other Turkic branches (e.g., moderate Oguz (SE) influence in the Western group, slight Oguz influence in the Northern group) and from Mongolic languages (moderate influence on the Southern group, lighter in the other groups), and all have heavy influence from Persian.
+In a linguistic sense, these languages have different amounts of influence from other Turkic branches (e.g., moderate Oghuz (SE) influence in the Western group, slight Oghuz influence in the Northern group) and from Mongolic languages (moderate influence on the Southern group, lighter in the other groups), and all have heavy influence from Persian.
 
@@
 
-%FIXME: background on how, when, and with what support were developed?
+%FIXME: brief background on how, when, and with what support were developed?
 The transducers for these languages
@@
 \subsection{Kazakh}
-Kazakh is primarily spoken in Kazakhstan, where it is the national language, sharing official status with Russian as an official language.
+Kazakh /qɑzɑq/ is spoken primarily in Kazakhstan, where it is the national language, sharing official status with Russian as an official language.
 Large communities of native speakers also exist in China, neighbouring Central-Eurasian republics, and Mongolia.
-Estimates of the total number of speakers range from 8 million \citep{ethnologue} to 11 million \citet{encyclopedin} people.
+Estimates of the total number of speakers range from 8 million \citep{ethnologue} to 11 million \citep{encyclopedin} people.
 
 \subsection{Tatar}
-Tatar is spoken in and around Tatarstan by approximately 5.4 million people \citep{ethnologue}.
+Tatar /tɑtɑr/ is spoken in and around Tatarstan by approximately 5.4 million people \citep{ethnologue}.
 It is co-official with Russian in Tatarstan --- a republic within Russia.
@@
 \subsection{Kumyk}
-Kumyk has approximately 430 thousand speakers \cite{ethnologue}
+Kumyk /qumuq/ is spoken in Dagestan, a Republic of the Russia Federation, where it is co-official with a number of other languages of Dagestan \citep{ethnologue}.
+There are approximately 430 thousand speakers \citep{ethnologue}.
 

### 2013-10-20 kipchak-lrec-2013/kipchak-paper.tex
some work
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@48159 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-\cite{washington2012}
-\cite{salimzyanov2013}
 \cite{bekmanova2013}
@@
 \cite{olmesov2000}
+
+\section{Background}
+\subsection{Morphological transducers}
+The objective of a morphological transducer is twofold.
+Firstly to take surface forms (e.g., алдым) and generate all possible lexical forms, and secondly to take lexical forms (e.g.,  ал{\tt {\small <v><tv><ifi><p1><sg>}}, алд{\tt {\small <n><px1sg><nom>}}, etc.) and generate one or more surface forms.
+
+The transducers were designed based on the Helsinki Finite State Toolkit \citep{linden2011} which is a free/open-source reimplementation of the Xerox finite-state toolchain, popular in the field of morphological analysis.
+It implements both the \textbf{lexc} formalism for defining lexicons, and the \textbf{twol} and \textbf{xfst} formalisms for modeling morphophonological rules.
+It also supports other finite state transducer formalisms such as \textbf{sfst}.
+This toolkit has been chosen as it -- or the equivalent XFST -- has been widely used for other Turkic languages, such as Turkish \citep{coltekin2010}, Crimean~Tatar~\citep{altintas2001}, Turkmen \citep{tantug2006}, and Kyrgyz \citep{washington2012}, and is available under a free/open-source licence.
+
+The Tatar transducer was developed originally as part of a prototype Tatar and Bashkir machine translation system \citep{tyerswashingtonsalimzyanbattalov12} and was enhanced during a Google Summer of Code project.
+The Kazakh transducer was originally created as part of an experimental Kazakh-Tatar machine translation system in December of 2010, and was expanded during Google Code-In 2010 and 2011.
+The Kazakh-Tatar machine translation system was expanded to production-level quality as part of a Google Summer of Code project in 2012 \citep{salimzyanov2013}.
+
+%FIXME: update for the three languages
+\subsection{Description}
+The tagset consists of 127 separate tags, 19 covering the main parts of speech (noun, verb, adjective, adverb, postposition, etc.) and 108 covering morphological subcategorisation for e.g.
+case, number, person, possession, transitivity, tense-aspect-mood, etc.
+The tags are represented as multicharacter symbols, between less than `<' and greater than `>' symbols.
+The tagset is quite extensive and still not entirely stabilised, as such a full listing is not included here.
+However, the tags are listed in the source code of the transducer,\footnote{\url{https://apertium.svn.sourceforge.net/svnroot/apertium/branches/apertium-kir/}} along with comments describing their usage.
 

### 2013-10-21 kipchak-lrec-2013/kipchak-paper.tex
fix missing package
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@48162 72bbbca6-d526-0410-a7d9-f06f51895060
more background
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@48163 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{multirow}
+\usepackage[colorlinks=true,citecolor=black,linkcolor=black,urlcolor=blue]{hyperref}
 
@@
 
+%FIXME: usually star comes first, then daggers
 \name{Jonathan North Washington$^\dagger$, Ilnar Salimzyanov$^\ddagger$, Francis M. Tyers$^\star$}
@@
 \subsection{Morphological transducers}
-The objective of a morphological transducer is twofold.
-Firstly to take surface forms (e.g., алдым) and generate all possible lexical forms, and secondly to take lexical forms (e.g.,  ал{\tt {\small <v><tv><ifi><p1><sg>}}, алд{\tt {\small <n><px1sg><nom>}}, etc.) and generate one or more surface forms.
+%FIXME: need a source, maybe something by Tommi?
+The objective of a morphological transducer is twofold: firstly to take surface forms (e.g., алдым) and generate all possible lexical forms, and secondly to take lexical forms (e.g.,  ал{\tt {\small <v><tv><ifi><p1><sg>}}, алд{\tt {\small <n><px1sg><nom>}}, etc.) and generate one or more surface forms.
+Since they are implemented as finite-state automata \citet{FSTs}, they are reversible by default.
 
@@
 
-The Tatar transducer was developed originally as part of a prototype Tatar and Bashkir machine translation system \citep{tyerswashingtonsalimzyanbattalov12} and was enhanced during a Google Summer of Code project.
-The Kazakh transducer was originally created as part of an experimental Kazakh-Tatar machine translation system in December of 2010, and was expanded during Google Code-In 2010 and 2011.
-The Kazakh-Tatar machine translation system was expanded to production-level quality as part of a Google Summer of Code project in 2012 \citep{salimzyanov2013}.
+Creating a morphological transducers in the above-mentioned formalisms simply involves encoding linguistic knowledge about the language in the formalisms.
+The lexc and twol formalisms resemble linguistic formalisms, allowing the coders to work with abstractions resembling linguistic categories such lexemes, morphemes, phonemes, and even archiphonemes---as opposed to a raw FST, where input characters are translated to output characters along a graph.
+
+The three transducers discussed in this paper are for Kazakh, Tatar, and Kumyk.
+The Kazakh and Tatar transducers were originally created as part of an experimental Kazakh-Tatar machine translation system in December of 2010.
+The Kazakh transducer was expanded during Google Code-In 2010 and 2011, and the Tatar transducer was expanded as part of a prototype Tatar and Bashkir machine translation system \citep{tyerswashingtonsalimzyanbattalov12}.
+The Kazakh-Tatar machine translation system, along with the two transducers, was expanded to production-level quality as part of a Google Summer of Code project in 2012 \citep{salimzyanov2013}.
+
+The Kumyk transducer was developed starting at the beginning of October, 2013 as experiment to see how difficult it would be to extend lessons learned from the development of the Tatar and Kazakh transducers to a related language.
+This paper explores some of these lessons and how the development of the Kumyk transducer benefited from knowledge gained from the development of the Tatar and Kazakh transducers.
+
+\subsection{Description}
+
+%FIXME: keep revision number up to date
+The transducers are available / under development at \url{https://svn.code.sf.net/p/apertium/svn/languages/}, in the directories apertium-kaz, apertium-tat, and apertium-kum.
+The revision of the entire subversion repository that the numbers (stem counts, evaluation, etc.) in this paper represent is \texttt{r48137}.
 
 %FIXME: update for the three languages
-\subsection{Description}
 The tagset consists of 127 separate tags, 19 covering the main parts of speech (noun, verb, adjective, adverb, postposition, etc.) and 108 covering morphological subcategorisation for e.g.
@@
 \section{Methodology}
+% cite Hengeveld (1992) ?
 

### 2013-10-24 kipchak-lrec-2013/kipchak-paper.tex
another FIXME
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@48164 72bbbca6-d526-0410-a7d9-f06f51895060
a map of Kypchak
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@48174 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 The tagset is quite extensive and still not entirely stabilised, as such a full listing is not included here.
-However, the tags are listed in the source code of the transducer,\footnote{\url{https://apertium.svn.sourceforge.net/svnroot/apertium/branches/apertium-kir/}} along with comments describing their usage.
+However, the tags are listed in the source code of the transducer,\footnote{\url{https://apertium.svn.sourceforge.net/svnroot/apertium/branches/apertium-kir/} testing} along with comments describing their usage.
 
@@
 \multirow{4}{*}{Tatar} & Wikipedia 2013 & 128K &  - \\
-	& RFE/RL 2005--2011 & 4.6M & - \\
+	%FIXME: the News corpus isn't RFERL, pretty sure; maybe we should test just on RFERL content?
+	& News 2005--2011 & 4.6M & - \\
 	& New Testament & 137K & - \\\cline{2-4}

### 2013-10-24 kipchak-lrec-2013/kipchak-paper.tex
integrate map
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@48175 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{booktabs}
+\usepackage{jnwmaps}
 
@@
 As such, each of these three languages represents a different one of the three branches of Kypchak.
-The geographic distribution of the languages is shown in map \ref{}.
+The geographic distribution of the languages is shown in map \ref{map}.
+
+\begin{map*}
+	\label{map}
+	\includegraphics[width=\textwidth]{map/map}
+	\caption{The three branches of Kypchak (N, S, W), showing the distribution of the three languages for which transducers were developed (tat, kaz, kum).  Language codes are from ISO 639-3.}
+\end{map*}
 
@@
 %FIXME: keep revision number up to date
-The transducers are available / under development at \url{https://svn.code.sf.net/p/apertium/svn/languages/}, in the directories apertium-kaz, apertium-tat, and apertium-kum.
+The transducers are available / under development in apertium's subversion repository,\footnote{\url{https://svn.code.sf.net/p/apertium/svn/languages/}} in the directories apertium-kaz, apertium-tat, and apertium-kum.
 The revision of the entire subversion repository that the numbers (stem counts, evaluation, etc.) in this paper represent is \texttt{r48137}.
@@
 The tagset is quite extensive and still not entirely stabilised, as such a full listing is not included here.
-However, the tags are listed in the source code of the transducer,\footnote{\url{https://apertium.svn.sourceforge.net/svnroot/apertium/branches/apertium-kir/} testing} along with comments describing their usage.
+However, the tags are listed in the source code of the transducer,\footnote{no url needed?} along with comments describing their usage.
 

### 2013-10-24 kipchak-lrec-2013/kipchak-paper.tex
Fran needs to learn how to use booktabs
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@48176 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \begin{center}
-\begin{tabular}{|l|rrr|}
-		\hline
-\multirow{2}{*}{\textbf{Part of speech}} & \multicolumn{3}{|c|}{\textbf{Number of stems}} \\ \cline{2-4}
+\begin{tabular}{lrrr}
+		\toprule
+\multirow{2}{*}{\textbf{Part of speech}} & \multicolumn{3}{c}{\textbf{Number of stems}} \\ \cline{2-4}
                         & Kazakh & Tatar & Kumyk \\
-		\hline
+		\midrule
 		Noun & - & - & - \\
@@
 		Determiner & - & - & - \\
-		\hline
+		\midrule
 		Total: & - & - & - \\
-		\hline
+		\bottomrule
 \end{tabular}
@@
 	\begin{tabular}{lrr}
+	\toprule
 		\textbf{Language} & \textbf{Precision} & \textbf{Recall} \\
-\hline
+	\midrule
 		Kazakh & - &  - \\
@@
 		Kumyk & - & - \\
-\hline
+	\bottomrule
 	\end{tabular}

### 2013-10-24 kipchak-lrec-2013/kipchak-paper.tex
some cleaning up and rearranging and adding refs
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@48177 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
+\defaultfontfeatures{Mapping=tex-text}
 \setromanfont{Times New Roman}
+\newfontfamily\symbl[Scale=MatchLowercase]{FreeSerif}
 
@@
 
-These languages belong to the Northwestern branch of Turkic, which is often referred to as the Kypchak branch.
-This branch can be divided into three subbranches.
-Kumyk is a member of the Western Kypchak group, Tatar is a member of the Northern Kypchak group, and Kazakh is a member of the Southern Kypchak group \citep[82-83]{histofturkic}.
-As such, each of these three languages represents a different one of the three branches of Kypchak.
-The geographic distribution of the languages is shown in map \ref{map}.
-
-\begin{map*}
-	\label{map}
-	\includegraphics[width=\textwidth]{map/map}
-	\caption{The three branches of Kypchak (N, S, W), showing the distribution of the three languages for which transducers were developed (tat, kaz, kum).  Language codes are from ISO 639-3.}
-\end{map*}
-
-In a linguistic sense, these languages have different amounts of influence from other Turkic branches (e.g., moderate Oghuz (SE) influence in the Western group, slight Oghuz influence in the Northern group) and from Mongolic languages (moderate influence on the Southern group, lighter in the other groups), and all have heavy influence from Persian.
-
 \cite{bekmanova2013}
@@
 
+The three languages for which transducers were developed belong to the Northwestern branch of Turkic, which is often referred to as the Kypchak branch.
+This branch can be divided into three subbranches.
+Kumyk is a member of the Western Kypchak group, Tatar is a member of the Northern Kypchak group, and Kazakh is a member of the Southern Kypchak group \citep[82-83]{histofturkic}.
+As such, each of these three languages represents a different one of the three branches of Kypchak.
+The geographic distribution of the languages is shown in map \ref{map1}.
+
+\begin{map*}[htbp]
+	\includegraphics[width=\textwidth]{map/map}
+	\caption{The three branches of Kypchak (N, S, W), showing the distribution of the three languages for which transducers were developed (tat, kaz, kum).  Language codes are from ISO 639-3.}
+	\label{map1}
+\end{map*}
+
+These languages have different amounts of linguistic influence from other Turkic branches (e.g., moderate Oghuz (SE) influence in the Western group, slight Oghuz influence in the Northern group) and from Mongolic languages (moderate influence on the Southern group, lighter in the other groups), and all have heavy influence from Persian.
+
 \subsection{Kazakh}
-Kazakh /qɑzɑq/ is spoken primarily in Kazakhstan, where it is the national language, sharing official status with Russian as an official language.
+Kazakh /q{\symbl ɑ}z{\symbl ɑ}q/ is spoken primarily in Kazakhstan, where it is the national language, sharing official status with Russian as an official language.
 Large communities of native speakers also exist in China, neighbouring Central-Eurasian republics, and Mongolia.
@@
 \subsection{Tatar}
-Tatar /tɑtɑr/ is spoken in and around Tatarstan by approximately 5.4 million people \citep{ethnologue}.
+Tatar /t{\symbl ɒ}t{\symbl ɑ}r/ is spoken in and around Tatarstan by approximately 5.4 million people \citep{ethnologue}.
 It is co-official with Russian in Tatarstan --- a republic within Russia.

### 2013-10-24 kipchak-lrec-2013/kipchak-paper.tex
more work
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@48178 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-The three transducers discussed in this paper are for Kazakh, Tatar, and Kumyk.
-The Kazakh and Tatar transducers were originally created as part of an experimental Kazakh-Tatar machine translation system in December of 2010.
-The Kazakh transducer was expanded during Google Code-In 2010 and 2011, and the Tatar transducer was expanded as part of a prototype Tatar and Bashkir machine translation system \citep{tyerswashingtonsalimzyanbattalov12}.
-The Kazakh-Tatar machine translation system, along with the two transducers, was expanded to production-level quality as part of a Google Summer of Code project in 2012 \citep{salimzyanov2013}.
-
-The Kumyk transducer was developed starting at the beginning of October, 2013 as experiment to see how difficult it would be to extend lessons learned from the development of the Tatar and Kazakh transducers to a related language.
-This paper explores some of these lessons and how the development of the Kumyk transducer benefited from knowledge gained from the development of the Tatar and Kazakh transducers.
-
 \subsection{Description}
@@
 \subsection{Development effort}
-% how long it took
-% make a graph for the kumyk one
+
+The three transducers discussed in this paper are for Kazakh, Tatar, and Kumyk.
+The Kazakh and Tatar transducers were originally created as part of an experimental Kazakh-Tatar machine translation system in December of 2010.
+The Kazakh transducer was expanded during Google Code-In 2010 and 2011, and the Tatar transducer was expanded as part of a prototype Tatar and Bashkir machine translation system \citep{tyerswashingtonsalimzyanbattalov12}.
+The Kazakh-Tatar machine translation system, along with the two transducers, was expanded to production-level quality as part of a Google Summer of Code project in 2012 \citep{salimzyanov2013}.
+
+The Kumyk transducer was developed starting at the beginning of October, 2013 as experiment to see how difficult it would be to extend lessons learned from the development of the Tatar and Kazakh transducers to a related language.
+This paper explores some of these lessons and how the development of the Kumyk transducer benefited from knowledge gained from the development of the Tatar and Kazakh transducers.
+
+The morphotactics of Turkic languages are complex enough that even a linguist who is fluent in the language and has a good linguistic understanding of it may not understand how exactly all morphemes combine.
+Native speakers educated about the morphology of their languages also do not have an explicit knowledge of the complete morphotactics.
+Hence it often becomes necessary to use fieldwork methodology to elicit the full extent of the morphotactics, be this a linguist with no to little knowledge of a Turkic language working with a native speaker, or a native speaker who understands the extent of what knowledge is necessary to encode in the transducer.
+When there is no native speaker of a particular language available, the authors have found that information previously encoded about a closely related language or the intuitions of a speaker of a closely related language may be combined with the use of textual corpora to ``elicit'' information about the morphotactics of a language.
+Depending on the contents of corpus and chance, this may not result in a completely accurate model, but it is possible to be thorough.
+
+The Kazakh morphotactics were originally developed based on the Kyrgyz transducer, which was coauthored by a linguist who is fluent in and has a good linguistic knowledge of Kyrgyz together with a native speaker of Kyrgyz, and in consultation with another native speaker of Kyrgyz.
+The initial developer of the Kazakh morphotactic was the same linguist, who is fluent in and has a good linguistic understanding Kazakh.
+The morphotactics of Tatar were developed for the most part by a native Tatar speaker, who also worked to polish off the morphotactics of the Kazakh transducer.
+
+As the authors found it difficult to locate native speakers of Kumyk, the morphotactics of the Kumyk transducer were developed based on the existing encoded morphotactics of Kazakh and Tatar (and occasionally Kyrgyz), with consultation of corpora, as described above.
+
+% FIXME: how long it took
+% FIXME: make a graph for the kumyk one
 
@@
 The first was by calculating the naïve coverage\footnote{Naïve coverage refers to the percentage of surface forms in a given corpora that receive at least one analysis.
-Forms counted by this measure may have other analyses which are not delivered by the transducer.} and mean ambiguity
-on freely available corpora. The second was by performing an evaluation of precision and recall on some 
-smaller, hand-validated test sets.
+Forms counted by this measure may have other analyses which are not delivered by the transducer.} and mean ambiguity  on freely available corpora.
+The mean ambiguity measure was calculated by performing an evaluation of precision and recall on some smaller, hand-validated test sets.
 
@@
 
-% For kazakh+tatar tested the coverage of the analysers over 3 separate domains: encyclopaedic text, news and religion, 
-% as there was no wikipedia for kumyk we just tested news and religion. The corpora were obtained from:
+%FIXME: sources
+We tested the coverage of the Kazakh and Tatar analysers over three separate domains: encyclopaedic text,\footnote{The following wikipedia dumps were used: kkwiki-20131006-pages-articles.xml.bz2, FIXME.} news,\footnote{All content from RFE/RL (\url{http://www.azattyq.org/}) for 2010 was used for Kazakh, as well as all content from 2005 to 2011 on \url{http://tat.tatar-inform.ru} for Tatar.} and religion.\footnote{We used a Kazakh bible translation available from \url{https://kkitap.net/} and a Tatar translation of the New Testament available from \url{http://ibt.org.ru}}  As there is currently no wikipedia in Kumyk, we tested only news and religion.\footnote{The bible corpus is from \url{http://ibt.org.ru/} and the news corpus consists of all Kumyk from \footnote{http://sh-tavisi.etnosmi.ru/}.}
+
+%The corpora were obtained from:
 % Wikipedia:
@@
 
-% kaz: kkwiki-20131006-pages-articles.xml.bz2
+% kaz: 
 % tat:
+
+The coverage of each transducer over the various corpora is shown in table \ref{table:corpora}.
 

### 2013-10-24 kipchak-lrec-2013/kipchak-paper.tex
some citation fixes, etc.
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@48179 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-\cite{bekmanova2013}
-
 %FIXME: brief background on how, when, and with what support were developed?
@@
 	\includegraphics[width=\textwidth]{map/map}
-	\caption{The three branches of Kypchak (N, S, W), showing the distribution of the three languages for which transducers were developed (tat, kaz, kum).  Language codes are from ISO 639-3.}
+	\caption{The three subbranches of Kypchak (N, S, W), roughly divided with black lines, showing the geographic distribution of the three languages for which transducers were developed (tat, kaz, kum).
+Language codes are from ISO 639-3.}
 	\label{map1}
@@
 Large communities of native speakers also exist in China, neighbouring Central-Eurasian republics, and Mongolia.
-Estimates of the total number of speakers range from 8 million \citep{ethnologue} to 11 million \citep{encyclopedin} people.
+Estimates of the total number of speakers range from 8 million \citep{ethnologue} to 11 million \citep{nationalencyklopedin} people.
 
@@
 
-\cite{bammatov1960}
-\cite{olmesov2000}
-
 \section{Background}
-\subsection{Morphological transducers}
+\subsection{Morphological transducers, previous work}
 %FIXME: need a source, maybe something by Tommi?
 The objective of a morphological transducer is twofold: firstly to take surface forms (e.g., алдым) and generate all possible lexical forms, and secondly to take lexical forms (e.g.,  ал{\tt {\small <v><tv><ifi><p1><sg>}}, алд{\tt {\small <n><px1sg><nom>}}, etc.) and generate one or more surface forms.
-Since they are implemented as finite-state automata \citet{FSTs}, they are reversible by default.
+Since they are implemented as finite-state automata, they are reversible by default.
 
-The transducers were designed based on the Helsinki Finite State Toolkit \citep{linden2011} which is a free/open-source reimplementation of the Xerox finite-state toolchain, popular in the field of morphological analysis.
+The transducers were designed based on the Helsinki Finite State Toolkit \citep{hfst/2011} which is a free/open-source reimplementation of the Xerox finite-state toolchain, popular in the field of morphological analysis.
 It implements both the \textbf{lexc} formalism for defining lexicons, and the \textbf{twol} and \textbf{xfst} formalisms for modeling morphophonological rules.
@@
 This toolkit has been chosen as it -- or the equivalent XFST -- has been widely used for other Turkic languages, such as Turkish \citep{coltekin2010}, Crimean~Tatar~\citep{altintas2001}, Turkmen \citep{tantug2006}, and Kyrgyz \citep{washington2012}, and is available under a free/open-source licence.
+
+The authors learned of another Kazakh morphological transducer in existence \citep{bekmanova2013} only after production-ready version of our transducer was released.
+We have not yet been able to evaluate this system or compare it to ours.
 
@@
 As the authors found it difficult to locate native speakers of Kumyk, the morphotactics of the Kumyk transducer were developed based on the existing encoded morphotactics of Kazakh and Tatar (and occasionally Kyrgyz), with consultation of corpora, as described above.
+A dictionary \citep{bammatov1960} and a grammar \citep{olmesov2000} of Kumyk were also consulted as needed.
 
@@
  \caption{Number of stems in each of the categories}
- \label{table:coverage}
+ \label{table:stems}
 \end{center}

### 2013-10-24 kipchak-lrec-2013/kipchak-paper.tex
some updates and fixes and expanding
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@48180 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-\abstract{Hargle, bargle.\\
+\abstract{This paper describes the development of free/open-source finite-state morphological transducers for three Turkic languages---Kazakh, Tatar, and Kumyk---representing a language from each of the three subbranches of the Kypchak branch of Turkic.
+The finite-state toolkit used for the work is the Helsinki Finite-State Toolkit (HFST).
+This paper describes how the development of a transducer for each subsequent language took less time.
+An evaluation is presented which shows that the transducers all have production-level coverage---around 90\%---on freely available corpora of the languages, and high precision and recall over a manually verified test set.
+\\
 	\Keywords{Kazakh, Tatar, Kumyk, morphology, transducer}
@@
 
-%FIXME: update for the three languages
-The tagset consists of 127 separate tags, 19 covering the main parts of speech (noun, verb, adjective, adverb, postposition, etc.) and 108 covering morphological subcategorisation for e.g.
-case, number, person, possession, transitivity, tense-aspect-mood, etc.
-The tags are represented as multicharacter symbols, between less than `<' and greater than `>' symbols.
-The tagset is quite extensive and still not entirely stabilised, as such a full listing is not included here.
-However, the tags are listed in the source code of the transducer,\footnote{no url needed?} along with comments describing their usage.
-
 \section{Methodology}
@@
 The Kumyk transducer was developed starting at the beginning of October, 2013 as experiment to see how difficult it would be to extend lessons learned from the development of the Tatar and Kazakh transducers to a related language.
-This paper explores some of these lessons and how the development of the Kumyk transducer benefited from knowledge gained from the development of the Tatar and Kazakh transducers.
+While the Kazakh and Tatar transducers took years to reach their current production-quality (around 90\%) coverage of corpora, the Kumyk transducer only took a couple weeks to reach this level of quality.
+This paper explores how the development of the Kumyk transducer benefited from knowledge gained from the development of the Tatar and Kazakh transducers.
 
@@
 
-\subsection{Statistics}
+\subsection{Transducer contents}
+
+Each transducer's lexc source consists of lists of stems, with each stem pointing at a complex continuation lexicon containing the appropriate morphology for the type of stem.
+As the compiled transducer parses the stems and morphology, tags are returned.
+
+%FIXME: update for the three languages
+The tagset for each transducer is designed to be compatible with the others.
+Each transducer consists of about 120 separate tags, of which close to 20 cover the main parts of speech (noun, verb, adjective, adverb, postposition, interjection, etc.).
+The remaining tags cover morphological subcategorisation for e.g.\ case, number, person, possession, transitivity, tense-aspect-mood, etc.
+The tags are represented as multicharacter symbols, between less-than `<' and greater-than `>' symbols.
+The tagset is quite extensive and still not entirely stabilised, so a full listing is not included here.
+However, the tags are listed in the source code of the transducers, along with comments describing their usage.
+
+%\subsection{Statistics}
+
+Table \ref{table:stems} lists the number of stems of the primary categories in each transducer.
 
@@
 		\midrule
-		Noun & - & - & - \\
-		Verb & - & - & - \\
-		Adjective & - & - & - \\
-		Proper noun & - & - & - \\
-		Adverb & - & - & - \\
-		Numeral & - & - & - \\
-		Conjunction & - & - & - \\
-		Postposition & - & - & - \\
-		Pronoun & - & - & - \\
-		Determiner & - & - & - \\
+		Noun & 2463 & 2692 & 2588 \\
+		Verb & 1587 & 1333 & 262 \\
+		Adjective & 823 & 852 & 217 \\
+		Proper noun & 5412 & 3523 & 1444 \\
+		Adverb & 197 & 205 & 63 \\
+		Numeral & 63 & 67 & 45 \\
+		Conjunction & 52 & 50 & 15 \\
+		Postposition & 56 & 46 & 12 \\
+		Pronoun & 17 & 17 & 18 \\
+		Determiner & 42 & 37 & 9 \\
 		\midrule
-		Total: & - & - & - \\
+		Total: & 10942 & 9009 & 4687 \\
 		\bottomrule
@@
 %FIXME: sources
-We tested the coverage of the Kazakh and Tatar analysers over three separate domains: encyclopaedic text,\footnote{The following wikipedia dumps were used: kkwiki-20131006-pages-articles.xml.bz2, FIXME.} news,\footnote{All content from RFE/RL (\url{http://www.azattyq.org/}) for 2010 was used for Kazakh, as well as all content from 2005 to 2011 on \url{http://tat.tatar-inform.ru} for Tatar.} and religion.\footnote{We used a Kazakh bible translation available from \url{https://kkitap.net/} and a Tatar translation of the New Testament available from \url{http://ibt.org.ru}}  As there is currently no wikipedia in Kumyk, we tested only news and religion.\footnote{The bible corpus is from \url{http://ibt.org.ru/} and the news corpus consists of all Kumyk from \footnote{http://sh-tavisi.etnosmi.ru/}.}
+We tested the coverage of the Kazakh and Tatar analysers over three separate domains: encyclopaedic text,\footnote{The following wikipedia dumps were used: kkwiki-20131006-pages-articles.xml.bz2, FIXME.} news,\footnote{All content from \url{http://www.azattyq.org/} for 2010 was used for Kazakh, as well as all content from 2005 to 2011 on \url{http://tat.tatar-inform.ru} for Tatar.} and religion.\footnote{We used a Kazakh bible translation available from \url{https://kkitap.net/} and a Tatar translation of the New Testament available from \url{http://ibt.org.ru}}  As there is currently no wikipedia in Kumyk, we tested only news and religion.\footnote{The bible corpus is from \url{http://ibt.org.ru/} and the news corpus consists of all Kumyk content from \url{http://sh-tavisi.etnosmi.ru/}.}
 
@@
 
+\subsection{Precision and recall}
+
+Precision and recall are measures of the average accuracy of analyses provided by a morphological transducer.
+Precision represents the number of the analyses given for a form that are correct.
+Recall is the percentage of analyses that are deemed correct for a form (by comparing against a gold standard) that are provided by the transducer.
+
+To calculate precision and recall, it was necessary to create a hand-verified list of surface forms and their analyses.
+We extracted 1,000 unique surface forms at random from a news corpus for each language, and checked that they were valid words in the languages and correctly spelt.
+Where a word was incorrectly spelt or deemed not to be a form used in the language, it was discarded and a new random word selected.
+
+%FIXME: footnote with urls
+This list of surface forms was then analysed with the most recent version of the analyser, and each analysis was checked.
+Where an analysis was erroneous, it was removed; where an analysis was missing, it was added.
+This process gave us a `gold standard' morphologically analysed word list of 500 surface forms with their analyses.
+The list is publically available for each language.%\footnote{\url{https://apertium.svn.sourceforge.net/svnroot/apertium/branches/apertium-kir/eval/ref.1000.txt}}
+
+We then took the same list of surface forms and ran them through the morphological analyser once more.
+Precision was calculated as the number of analyses which were found in both the output from the morphological analyser and the gold standard, divided by the total number of analyses output by the morphological analyser.
+
+Recall was calculated as the total number of analyses found in both the output from the morphological analyser and the gold standard, divided by the number of analyses found in the morphological analyser plus the number of analyses found in the gold standard but not in the morphological analyser.
+
+The results for precision and recall will be presented in table~\ref{table:precrecall}.
+
 \begin{table}
@@
 	\caption{Precision and recall}
-	\label{table:coverage}
+	\label{table:precrecall}
 \end{center}
@@
 
+One direction for future work is to develop transducers for more languages.
+We have already constructed usable prototype transducers for three other Kypchak languages: Bashkir (N), Nogay (S), and Karakalpak (S).
+Since our ability to develop production-ready transducers is limited by availability of resources, including corpora in the languages and native speaker consultants, the Western Kypchak languages (aside from Kumyk) have been more neglected by our team.
+However, these language communities would benefit from computational tools for the language, and work on them may be bootstrapped from the existing transducers, so working on morphological transducers for these languages is also a priority.
+
 \section{Conclusions}
+
+We have described morphological transducers for three Kypchak languages---one from each branch of Kypchak---including the development process and performance of the analysers.
 

### 2013-10-24 kipchak-lrec-2013/stems.sh

@@ -1,14 +1,17 @@
 #!/bin/bash
 
 pushd $APERTIUMPATH/languages/apertium-kum/dev/
+	hfst-lexc --format foma ../apertium-kum.kum.lexc -o /dev/null
 	./countstems.sh
 popd
 
 pushd $APERTIUMPATH/languages/apertium-kaz/dev/
+	hfst-lexc --format foma ../apertium-kaz.kaz.lexc -o /dev/null
 	./countstems.sh
 popd
 
-pushd $APERTIUMPATH/languages/apertium-kaz/dev/
+pushd $APERTIUMPATH/languages/apertium-tat/dev/
+	hfst-lexc --format foma ../apertium-tat.tat.lexc -o /dev/null
 	./countstems.sh
 popd
 

### 2014-02-16 kipchak-lrec-2013/kipchak-paper.tex
submitted version
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@48181 72bbbca6-d526-0410-a7d9-f06f51895060
sadly, I'm not a t IMS yet
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@50131 72bbbca6-d526-0410-a7d9-f06f51895060
numbers
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@50288 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 Ilnar Salimzyanov  \\
-Institut für Maschinelle Sprachverarbeitung \\
-Universität Stuttgart\\
-Stuttgart (Germany) \\
-\texttt{ilnar@ilnar} \and 
+Kazan Federal University \\
+Kazan, Republic of Tatarstan\\
+(Russian Federation) \\
+\texttt{ilnar.salimzyan@gmail.com} \and 
 Francis M. Tyers\\
@@
 
-%The corpora were obtained from:
-% Wikipedia:
-% RFE/RL
-% Yoldaš: 
-% New testaments:
-
-% kaz: 
-% tat:
+% Tatar, religion: tat.nt.ibt.txt, tat.quran.nughmani.txt.bz2
+% Tatar, wp: tat.wikipedia.2013-02-25.txt.bz2
+% Tatar, news: tat.news.2005-2011_300K-sentences.txt.bz2
+% Kazakh, religion: kaz.bible.kkitap.txt.bz2, kaz.quran.altay.txt.bz2
+% Kazakh, wp: kaz.wikipedia2.2013.txt.bz2
+% Kazakh, news: kaz.rferl.2010.xml.bz2, kaz.rferl2.2012.txt
+% Kumyk, religion: kum.nt.ibt.txt, (kum.genesis.ibt.txt)
+% Kumyk, news: kum.yoldash.2013.txt, kum.articles.kumukia.txt.bz2
 
@@
 \toprule
-\textbf{Language} & \textbf{Corpus} & \textbf{Words} & \textbf{Coverage} \\
+\textbf{Language} & \textbf{Corpus} & \textbf{Words} & \textbf{Coverage} (\%) \\
 \midrule
-\multirow{4}{*}{Kazakh} & Wikipedia 2013 &  -  &  - \\
-	& RFE/RL 2010 & 3.2M & - \\
-	& Bible & 577K & - \\\cline{2-4}
+\multirow{4}{*}{Kazakh} & Wikipedia &  -  &  - \\
+	& News & 3.2M & - \\
+	& Religion & 577K & - \\\cline{2-4}
 	& Average & - & 90.5\% \\
 \midrule
-\multirow{4}{*}{Tatar} & Wikipedia 2013 & 128K &  - \\
+\multirow{4}{*}{Tatar} & Wikipedia & 128K &  - \\
 	%FIXME: the News corpus isn't RFERL, pretty sure; maybe we should test just on RFERL content?
-	& News 2005--2011 & 4.6M & - \\
-	& New Testament & 137K & - \\\cline{2-4}
+	& News & 4.6M & - \\
+	& Religion & 137K & - \\\cline{2-4}
 	& Average & - & 89.0\% \\
 \midrule
-\multirow{3}{*}{Kumyk} & Yoldaš & 287K &  - \\
-	& New Testament & 154K & - \\\cline{2-4}
-	& Average & - & 90.1\% \\
+% news = kum.yoldash.2013.txt
+% religion = kum.genesis.ibt.txt, kum.nt.ibt.txt
+\multirow{4}{*}{Kumyk} & Wikipedia & -- & -- \\
+        & News & 287K &  91.4 $\pm$ 0.7 \\
+	& Religion & 227K & 92.0 $\pm$ 1.1 \\\cline{2-4}
+	& Average & - & 91.7 $\pm$ 0.7 \\
 \bottomrule

### 2014-02-16 kipchak-lrec-2013/kipchak-paper.tex
update
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@50292 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
+% FIXME: what are transducers used for ?
+
 \section{Languages}
@@
 	\includegraphics[width=\textwidth]{map/map}
-	\caption{The three subbranches of Kypchak (N, S, W), roughly divided with black lines, showing the geographic distribution of the three languages for which transducers were developed (tat, kaz, kum).
+	\caption{The three sub-branches of Kypchak (North, South, West), roughly divided with black lines, showing the geographic distribution of the three languages for which transducers were developed (\texttt{tat}, \texttt{kaz}, \texttt{kum}).
 Language codes are from ISO 639-3.}
@@
 The Kumyk transducer was developed starting at the beginning of October, 2013 as experiment to see how difficult it would be to extend lessons learned from the development of the Tatar and Kazakh transducers to a related language.
-While the Kazakh and Tatar transducers took years to reach their current production-quality (around 90\%) coverage of corpora, the Kumyk transducer only took a couple weeks to reach this level of quality.
+While the Kazakh and Tatar transducers took years of part-time work to reach their current production-quality (around 90\%) coverage of corpora, the Kumyk transducer only took a couple weeks to reach this level of quality.
 This paper explores how the development of the Kumyk transducer benefited from knowledge gained from the development of the Tatar and Kazakh transducers.
@@
 
-The Kazakh morphotactics were originally developed based on the Kyrgyz transducer, which was coauthored by a linguist who is fluent in and has a good linguistic knowledge of Kyrgyz together with a native speaker of Kyrgyz, and in consultation with another native speaker of Kyrgyz.
+The Kazakh morphotactics were originally developed based on the Kyrgyz transducer, which was co-authored by a linguist who is fluent in and has a good linguistic knowledge of Kyrgyz together with a native speaker of Kyrgyz, and in consultation with another native speaker of Kyrgyz.
 The initial developer of the Kazakh morphotactic was the same linguist, who is fluent in and has a good linguistic understanding Kazakh.
@@
 Each transducer's lexc source consists of lists of stems, with each stem pointing at a complex continuation lexicon containing the appropriate morphology for the type of stem.
-As the compiled transducer parses the stems and morphology, tags are returned.
 
@@
 The remaining tags cover morphological subcategorisation for e.g.\ case, number, person, possession, transitivity, tense-aspect-mood, etc.
-The tags are represented as multicharacter symbols, between less-than `<' and greater-than `>' symbols.
+The tags are represented as multicharacter symbols, between less-than \texttt{<} and greater-than \texttt{>} symbols.
 The tagset is quite extensive and still not entirely stabilised, so a full listing is not included here.
@@
 %FIXME: sources
-We tested the coverage of the Kazakh and Tatar analysers over three separate domains: encyclopaedic text,\footnote{The following wikipedia dumps were used: kkwiki-20131006-pages-articles.xml.bz2, FIXME.} news,\footnote{All content from \url{http://www.azattyq.org/} for 2010 was used for Kazakh, as well as all content from 2005 to 2011 on \url{http://tat.tatar-inform.ru} for Tatar.} and religion.\footnote{We used a Kazakh bible translation available from \url{https://kkitap.net/} and a Tatar translation of the New Testament available from \url{http://ibt.org.ru}}  As there is currently no wikipedia in Kumyk, we tested only news and religion.\footnote{The bible corpus is from \url{http://ibt.org.ru/} and the news corpus consists of all Kumyk content from \url{http://sh-tavisi.etnosmi.ru/}.}
+We tested the coverage of the Kazakh and Tatar analysers over three separate domains: encyclopaedic text,\footnote{The following Wikipedia dumps were used: kkwiki-20131006-pages-articles.xml.bz2, FIXME.} news,\footnote{All content from \url{http://www.azattyq.org/} for 2010 was used for Kazakh, as well as all content from 2005 to 2011 on \url{http://tat.tatar-inform.ru} for Tatar.} and religion.\footnote{We used a Kazakh bible translation available from \url{https://kkitap.net/} and a Tatar translation of the New Testament available from \url{http://ibt.org.ru}}  As there is currently no Wikipedia in Kumyk, we tested only news and religion.\footnote{The bible corpus is from \url{http://ibt.org.ru/} and the news corpus consists of all Kumyk content from \url{http://sh-tavisi.etnosmi.ru/}.}
 
-% Tatar, religion: tat.nt.ibt.txt, tat.quran.nughmani.txt.bz2
-% Tatar, wp: tat.wikipedia.2013-02-25.txt.bz2
-% Tatar, news: tat.news.2005-2011_300K-sentences.txt.bz2
 % Kazakh, religion: kaz.bible.kkitap.txt.bz2, kaz.quran.altay.txt.bz2
@@
 % Kazakh, news: kaz.rferl.2010.xml.bz2, kaz.rferl2.2012.txt
+
+% Tatar, wp: tat.wikipedia.2013-02-25.txt.bz2
+% Tatar, news: tat.news.2005-2011_300K-sentences.txt.bz2
+% Tatar, religion: tat.nt.ibt.txt, tat.quran.nughmani.txt.bz2
+
+% Kumyk, news: kum.yoldash.2013.txt, kum.articles.kumukia.txt.bz2
 % Kumyk, religion: kum.nt.ibt.txt, (kum.genesis.ibt.txt)
-% Kumyk, news: kum.yoldash.2013.txt, kum.articles.kumukia.txt.bz2
 
@@
 \toprule
-\textbf{Language} & \textbf{Corpus} & \textbf{Words} & \textbf{Coverage} (\%) \\
+\textbf{Language} & \textbf{Corpus} & \textbf{Tokens} & \textbf{Coverage} (\%) \\
 \midrule
@@
 	& News & 3.2M & - \\
-	& Religion & 577K & - \\\cline{2-4}
-	& Average & - & 90.5\% \\
+	& Religion & 848K & 92.7 $\pm$ 1.7 \\\cline{2-4}
+	& Average & - &  \\
 \midrule
 \multirow{4}{*}{Tatar} & Wikipedia & 128K &  - \\
-	%FIXME: the News corpus isn't RFERL, pretty sure; maybe we should test just on RFERL content?
 	& News & 4.6M & - \\
-	& Religion & 137K & - \\\cline{2-4}
-	& Average & - & 89.0\% \\
+	& Religion & 205K & 88.7 $\pm$ 0.5 \\\cline{2-4}
+	& Average & - &  \\
 \midrule
@@
 Where an analysis was erroneous, it was removed; where an analysis was missing, it was added.
-This process gave us a `gold standard' morphologically analysed word list of 500 surface forms with their analyses.
+This process gave us a `gold standard' morphologically analysed word list of 1,000 surface forms with their analyses.
 The list is publically available for each language.%\footnote{\url{https://apertium.svn.sourceforge.net/svnroot/apertium/branches/apertium-kir/eval/ref.1000.txt}}
@@
 We have already constructed usable prototype transducers for three other Kypchak languages: Bashkir (N), Nogay (S), and Karakalpak (S).
-Since our ability to develop production-ready transducers is limited by availability of resources, including corpora in the languages and native speaker consultants, the Western Kypchak languages (aside from Kumyk) have been more neglected by our team.
-However, these language communities would benefit from computational tools for the language, and work on them may be bootstrapped from the existing transducers, so working on morphological transducers for these languages is also a priority.
+Since our ability to develop production-ready transducers is limited by availability of resources, including corpora in the languages and native-speaker consultants, the Western Kypchak languages (aside from Kumyk) have been more neglected by our team.
+However, these language communities would benefit from computational tools for their languages, and work on them may be bootstrapped from the existing transducers, so working on morphological transducers for these languages is also a priority.
 

### 2014-02-16 kipchak-lrec-2013/kipchak-paper.tex
update
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@50293 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \subsection{Kumyk}
-Kumyk /qumuq/ is spoken in Dagestan, a Republic of the Russia Federation, where it is co-official with a number of other languages of Dagestan \citep{ethnologue}.
+Kumyk /qumuq/ is spoken in Dagestan, a Republic of the Russian Federation, where it is co-official with a number of other languages of Dagestan \citep{ethnologue}.
 There are approximately 430 thousand speakers \citep{ethnologue}.

### 2014-02-16 kipchak-lrec-2013/kipchak-paper.tex
more
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@50296 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \newfontfamily\symbl[Scale=MatchLowercase]{FreeSerif}
+
+\newcommand{\tags}[1]{\texttt{#1}}
 
@@
 
+% Russian loans not obeying phonology N5, e.g. ё vs . самолёт
+
 \subsection{Development effort}
@@
 \end{table}
+
+% FIXME: example output
+
+\begin{table*}
+
+% kaz: ^Құдай/Құдай<n><nom>/Құдай<n><attr>/Құдай<n><nom>+е<cop><p3><pl>/Құдай<n><nom>+е<cop><p3><sg>$ ^Өзінің/Өз<prn><ref><p3><sg><gen>$ ^жаратқандарының/жарат<v><tv><ger_past><pl><px3sp><gen>/жарат<v><tv><gpr_past><subst><pl><px3sp><gen>$ ^бәріне/бәрі<prn><qnt><px3sp><dat>$ ^қарап/қара<v><tv><prc_perf>/қара<v><tv><gna_perf>/қара<v><tv><prc_perf>$^,/,<cm>$ ^өте/өт<v><tv><prc_impf>/өте<adv>/өте<v><tv><imp><p2><sg>$ ^жақсы/жақсы<adj>/жақсы<adv>/жақсы<adj><advl>/жақсы<adj><subst><nom>/жақсы<adj>+е<cop><p3><pl>/жақсы<adj>+е<cop><p3><sg>/жақсы<adj><subst><nom>+е<cop><p3><pl>/жақсы<adj><subst><nom>+е<cop><p3><sg>$ ^екенін/е<cop><ger_past><px3sp><acc>$ ^көрді/көр<vaux><ifi><p3><pl>/көр<vaux><ifi><p3><sg>/көр<v><tv><ifi><p3><pl>/көр<v><tv><ifi><p3><sg>$^./.<sent>$
+% kum: ^Аллагь/аллагь<n><nom>/аллагь<n><attr>$ ^Оьзю/оьз<prn><ref><px3sp><nom>$ ^яратгъан/ярат<v><tv><gpr_past>/ярат<v><tv><gpr_past><subst><nom>/ярат<v><tv><ger_past><nom>/ярат<v><tv><past><p3><sg>$ ^затлагъа/зат<n><pl><dat>$ ^къарап/къара<v><tv><prc_perf>/къара<v><tv><gna_perf>$^,/,<cm>$ ^олар/о<prn><dem><pl><nom>/олар<prn><pers><p3><pl><nom>$ ^бек/бек<adv>$ ^яхшы/яхшы<adj>/яхшы<adj><subst><nom>$ ^экенин/э<cop><ger_past><px3sp><acc>$ ^гёрген/гёр<v><tv><gpr_past>/гёр<v><tv><gpr_past><subst><nom>/гёр<v><tv><ger_past><nom>/гёр<v><tv><past><p3><sg>$^../..<sent>$
+
+\centering
+
+Құдай Өзінің жаратқандарының бәріне қарап, өте жақсы екенін көрді. \\
+% ходай Үзенең яратканнарының *бəріне карап, бик яхшы икәнен күрде.
+Аллагь Оьзю яратгъан затлагъа къарап, олар бек яхшы экенин гёрген. \\
+
+~\\
+
+\begin{tabular}{lll}
+ \textbf{Kazakh} & \textbf{Tatar} & \textbf{Kumyk} \\
+\hline
+ Құдай\tags{<n><nom>} &  & Аллагь\tags{<n><nom>}  \\
+ Өз\tags{<prn><ref><p3><sg><gen>} & & Оьз\tags{<prn><ref><px3sp><nom>} \\
+ жарат\tags{<v><tv><ger\_past><pl><px3sp><gen>} & & ярат\tags{<v><tv><gpr\_past>} \\
+ бәрі\tags{<prn><qnt><px3sp><dat>} & & зат\tags{<n><pl><dat>} \\
+ қара\tags{<v><tv><gna\_perf>} & & къара\tags{<v><tv><gna\_perf>} \\ 
+ ,\tags{<cm>} & & ,\tags{<cm>} \\
+ ---      & & олар\tags{<prn><pers><p3><pl><nom>} \\
+ өт\tags{<adv>} & & бек\tags{<adv>} \\
+ жақсы\tags{<adj>} & & яхшы\tags{<adj>} \\
+ е\tags{<cop><ger\_past><px3sp><acc>} & & э\tags{<cop><ger\_past><px3sp><acc>} \\
+ көр\tags{<v><tv><ifi><p3><sg>} & & гёр\tags{<v><tv><past><p3><sg>} \\  
+ .\tags{<sent>} & & .\tags{<sent>} \\
+\hline
+\end{tabular}
+
+\label{table:exoutput}
+\caption{An example of the output of each of the morphological transducers for the same sentence.}
+
+\end{table*}
 

### 2014-03-14 kipchak-lrec-2013/kipchak-paper.tex
data from kaz.bible.kkitap + kaz.quran.altay catted together and split into 10 chunks
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@50414 72bbbca6-d526-0410-a7d9-f06f51895060
some new coverage numbers
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@50418 72bbbca6-d526-0410-a7d9-f06f51895060
update
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@50545 72bbbca6-d526-0410-a7d9-f06f51895060
add numbers for kazakh
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@50598 72bbbca6-d526-0410-a7d9-f06f51895060
add ack aida
g

@@
 %FIXME: brief background on how, when, and with what support were developed?
+% In Apertium project, Turkic transducers were developed either as part of Turkic-to-Turkic machine translators, or they started as a "port" (in particular of morphotactics) of already existing Turkic transducers.
+
 The transducers for these languages
@@
 
+\subsection{Categorisation and tagset}
+
+% FIXME: we should change this title
+\subsection{Technolinguistic issues} % hidden phonology issues? ...?
+There is a class of phenomena encountered during the development of these transducers, united by the fact that the morphophonology needs information about the stem's phonological form that is not provided by the orthographical representation.
+These phenomena were left by \citet{washington2012} as future work, but have been extensively implemented in the transducers described by the current paper.
+
+Types of these phenomena include acronyms, ``irregular'' vowel harmony, Russian loanwords, and numerals.  A brief overview of each follows, along with examples, followed by details of our solution.
+
+%FIXME: description of each one, with examples
+
+The best way we have found to handle these phenomena in a two-layer FST like those that we use is to add phonological information at the end of stems that need it in the form of special ``abstract letters'' at the morphotactic stage that trigger phonological processes at the morphophonological stage and are deleted.
+
+%FIXME: example of implementation
+
+Since \texttt{twol} performs all operations in parallel, rules triggered by these must make reference to abstract letters that exist on the input tape but are null on the output tape.
+This complicates existing rules that have been designed to ignore null characters on the output tape.
+The \texttt{twol} rules resulting from the combination of processes simultaneously triggered by and ignoring output-null characters are rather ...
+uh, \textit{intricate} ...
+but they seem to work okay?
+Just don't poke them too hard.
+%FIXME
+
 \section{Evaluation}
@@
 \midrule
-\multirow{4}{*}{Kazakh} & Wikipedia &  -  &  - \\
+\multirow{4}{*}{Kazakh} & Wikipedia & 18.2M &  85.1 ± 1.33 \\
 	& News & 3.2M & - \\
-	& Religion & 848K & 92.7 $\pm$ 1.7 \\\cline{2-4}
+	& Religion & 849K & 92.7 $\pm$ 1.57 \\\cline{2-4}
 	& Average & - &  \\
@@
 
+% tat:
+%Precision: 95.0339558573854
+%Recall: 85.65416985462892
+%F-score: 90.10060362173037
+
+% kaz:
+% Precision: 98.61202908129543
+% Recall: 57.986785853089785
+% F-score: 73.02985805188449
+
 \begin{table}
@@
 	\midrule
-		Kazakh & - &  - \\
-		Tatar & - & - \\
+		Kazakh & 98.61 &  57.98 \\
+		Tatar & 95.03 & 85.65 \\
 		Kumyk & - & - \\
@@
 
+The authors would also like to express their gratitude to Aida Sundetova and X 
+for assistance in evaluating precision an recall.
+
 \bibliographystyle{lrec2012}

### 2014-03-14 kipchak-lrec-2013/kipchak-paper.tex
moar FIXME for §4.4
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@50726 72bbbca6-d526-0410-a7d9-f06f51895060
mor echaanges
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@50735 72bbbca6-d526-0410-a7d9-f06f51895060

--- 2014-02-16 kipchak-lrec-2013/kipchak-paper.tex
@@
 \usepackage[colorlinks=true,citecolor=black,linkcolor=black,urlcolor=blue]{hyperref}
-
+\usepackage[small,bf]{caption}
 \usepackage{booktabs}
@@
 Francis M. Tyers\\
-Departament de Llenguatges i Sistemes Informàtics \\  
-Universitat d'Alacant\\
-E-03071 Alacant (Spain)\\
-\texttt{ftyers@dlsi.ua.es} 
+HSL-fakultehta, \\
+UiT Norgga árktalaš universitehta, \\
+9019 Romsa (Norway)\\
+\texttt{francis.tyers@uit.no} 
 }
@@
 \texttt{ilnar@ilnar} \and 
-$^\star$Departament de Llenguatges i Sistemes Informàtics \\  
-Universitat d'Alacant\\
-E-03071 Alacant (Spain)\\
-\texttt{ftyers@dlsi.ua.es} 
+$^\star$HSL-fakultehta \\  
+UiT Norgga árktalaš universitehta \\
+9019 Romsa (Norway)\\
+\texttt{francis.tyers@uit.no} 
 }
 
-\abstract{This paper describes the development of free/open-source finite-state morphological transducers for three Turkic languages---Kazakh, Tatar, and Kumyk---representing a language from each of the three subbranches of the Kypchak branch of Turkic.
+\abstract{This paper describes the development of free/open-source finite-state morphological transducers for three Turkic languages---Kazakh, Tatar, and Kumyk---representing one language from each of the three subbranches of the Kypchak branch of Turkic.
 The finite-state toolkit used for the work is the Helsinki Finite-State Toolkit (HFST).
-This paper describes how the development of a transducer for each subsequent language took less time.
-An evaluation is presented which shows that the transducers all have production-level coverage---around 90\%---on freely available corpora of the languages, and high precision and recall over a manually verified test set.
+This paper describes how the development of a transducer for each subsequent language took less development time.
+An evaluation is presented which shows that the transducers all have a reasonable coverage---around 90\%---on freely available corpora of the languages, and high precision over a manually verified test set.
 \\
@@
 
+A finite-state transducer which models the morphology of a language has a wide variety of usages. For example,
+for creating proofing tools such as spellcheckers and grammarcheckers, and for machine translation.
+
 \section{Languages}
 
-The three languages for which transducers were developed belong to the Northwestern branch of Turkic, which is often referred to as the Kypchak branch.
+The three languages for which transducers were developed belong to the Northwestern branch of the Turkic family, which is often referred to as the Kypchak branch.
 This branch can be divided into three subbranches.
 Kumyk is a member of the Western Kypchak group, Tatar is a member of the Northern Kypchak group, and Kazakh is a member of the Southern Kypchak group \citep[82-83]{histofturkic}.
-As such, each of these three languages represents a different one of the three branches of Kypchak.
+As such, each of these three languages represents one of the three branches of Kypchak.
 The geographic distribution of the languages is shown in map \ref{map1}.
@@
 	\includegraphics[width=\textwidth]{map/map}
-	\caption{The three sub-branches of Kypchak (North, South, West), roughly divided with black lines, showing the geographic distribution of the three languages for which transducers were developed (\texttt{tat}, \texttt{kaz}, \texttt{kum}).
-Language codes are from ISO 639-3.}
+	\caption{The three sub-branches of Kypchak (North, South, West), roughly divided with black lines, showing the geographic distribution of the three languages for which transducers were developed (ISO639-3: \texttt{tat}, \texttt{kaz}, \texttt{kum}).}
 	\label{map1}
@@
 \subsection{Kazakh}
-Kazakh /q{\symbl ɑ}z{\symbl ɑ}q/ is spoken primarily in Kazakhstan, where it is the national language, sharing official status with Russian as an official language.
+Kazakh /q{\symbl ɑ}z{\symbl ɑ}q/ is spoken primarily in Kazakhstan, where it is the national language, where it is co-official with Russian.
 Large communities of native speakers also exist in China, neighbouring Central-Eurasian republics, and Mongolia.
-Estimates of the total number of speakers range from 8 million \citep{ethnologue} to 11 million \citep{nationalencyklopedin} people.
+Ethnologue estimates
+the total number of speakers to be around 8 million \citep{ethnologue}.
+
+%Estimates of the total number of speakers range from 8 million \citep{ethnologue} to 11 million \citep{nationalencyklopedin} people.
 
 \subsection{Tatar}
-Tatar /t{\symbl ɒ}t{\symbl ɑ}r/ is spoken in and around Tatarstan by approximately 5.4 million people \citep{ethnologue}.
-It is co-official with Russian in Tatarstan --- a republic within Russia.
+Tatar /t{\symbl ɒ}t{\symbl ɑ}r/ is spoken in and around Tatarstan, by approximately 5.4 million people \citep{ethnologue}.
+It is co-official with Russian in Tatarstan --- a republic of the Russian Federation.
 A majority of native speakers of both languages are bilingual in Russian.
@@
 \subsection{Kumyk}
-Kumyk /qumuq/ is spoken in Dagestan, a Republic of the Russian Federation, where it is co-official with a number of other languages of Dagestan \citep{ethnologue}.
+Kumyk /qumuq/ is spoken in Dagestan, a republic of the Russian Federation, where it is co-official with a number of other national languages \citep{ethnologue}.
 There are approximately 430 thousand speakers \citep{ethnologue}.
@@
 The objective of a morphological transducer is twofold: firstly to take surface forms (e.g., алдым) and generate all possible lexical forms, and secondly to take lexical forms (e.g.,  ал{\tt {\small <v><tv><ifi><p1><sg>}}, алд{\tt {\small <n><px1sg><nom>}}, etc.) and generate one or more surface forms.
-Since they are implemented as finite-state automata, they are reversible by default.
+As they are implemented as finite-state transducers, they are reversible by default.
+For more information on usin finite-state transducers for morphological analysis and generation, the reader is referred to \cite{beesley2003}.
 
@@
 
-The authors learned of another Kazakh morphological transducer in existence \citep{bekmanova2013} only after production-ready version of our transducer was released.
+The authors learnt of another Kazakh morphological transducer in existence \citep{bekmanova2013} only after this paper was submitted and our transducer was released.
 We have not yet been able to evaluate this system or compare it to ours.
 
-Creating a morphological transducers in the above-mentioned formalisms simply involves encoding linguistic knowledge about the language in the formalisms.
-The lexc and twol formalisms resemble linguistic formalisms, allowing the coders to work with abstractions resembling linguistic categories such lexemes, morphemes, phonemes, and even archiphonemes---as opposed to a raw FST, where input characters are translated to output characters along a graph.
+Creating morphological transducers in the above-mentioned formalisms involves encoding linguistic knowledge about the language in the formalisms.
+The lexc and twol formalisms resemble linguistic formalisms, allowing the coders to work with abstractions resembling linguistic categories such lexemes, morphemes, phonemes, and even archiphonemes.
 
@@
 The Kumyk transducer was developed starting at the beginning of October, 2013 as experiment to see how difficult it would be to extend lessons learned from the development of the Tatar and Kazakh transducers to a related language.
-While the Kazakh and Tatar transducers took years of part-time work to reach their current production-quality (around 90\%) coverage of corpora, the Kumyk transducer only took a couple weeks to reach this level of quality.
-This paper explores how the development of the Kumyk transducer benefited from knowledge gained from the development of the Tatar and Kazakh transducers.
+While the Kazakh and Tatar transducers took around six months work to reach their current coverage level, the Kumyk transducer only took a couple weeks to reach this level of quality (see Figure~\ref{fig:kumcov}).
+This paper explores how the development of the Kumyk transducer benefitted from knowledge gained from the development of the Tatar and Kazakh transducers.
 
-The morphotactics of Turkic languages are complex enough that even a linguist who is fluent in the language and has a good linguistic understanding of it may not understand how exactly all morphemes combine.
+The morphotactics\footnote{The morphotactics of a language is the way in which morphemes can be combined to create words.} of Turkic languages are complex enough that even a linguist who is fluent in the language and has a good linguistic understanding of it may not understand how exactly all morphemes combine.
 Native speakers educated about the morphology of their languages also do not have an explicit knowledge of the complete morphotactics.
-Hence it often becomes necessary to use fieldwork methodology to elicit the full extent of the morphotactics, be this a linguist with no to little knowledge of a Turkic language working with a native speaker, or a native speaker who understands the extent of what knowledge is necessary to encode in the transducer.
+Hence it often becomes necessary to use fieldwork methodology to elicit the full extent of the morphotactics, be this a linguist with little to no knowledge of a Turkic language working with a native speaker, or a native speaker who understands the extent of what knowledge is necessary to encode in the transducer.
 When there is no native speaker of a particular language available, the authors have found that information previously encoded about a closely related language or the intuitions of a speaker of a closely related language may be combined with the use of textual corpora to ``elicit'' information about the morphotactics of a language.
@@
 
-The Kazakh morphotactics were originally developed based on the Kyrgyz transducer, which was co-authored by a linguist who is fluent in and has a good linguistic knowledge of Kyrgyz together with a native speaker of Kyrgyz, and in consultation with another native speaker of Kyrgyz.
-The initial developer of the Kazakh morphotactic was the same linguist, who is fluent in and has a good linguistic understanding Kazakh.
-The morphotactics of Tatar were developed for the most part by a native Tatar speaker, who also worked to polish off the morphotactics of the Kazakh transducer.
+The Kazakh morphotactics were originally developed based on the Kyrgyz transducer, which was co-authored by the first author, who is fluent in and has a good linguistic knowledge of Kyrgyz and two native speakers of Kyrgyz.
+The first author also developed the Kazakh morphotactics who is fluent in and has a good linguistic understanding Kazakh.
+The morphotactics of Tatar were developed for the most part by the third author, a native speaker of Tatar, who also worked to polish off the morphotactics of the Kazakh transducer.
 
-As the authors found it difficult to locate native speakers of Kumyk, the morphotactics of the Kumyk transducer were developed based on the existing encoded morphotactics of Kazakh and Tatar (and occasionally Kyrgyz), with consultation of corpora, as described above.
+%As the authors found it difficult to locate native speakers of Kumyk, the morphotactics of the Kumyk transducer were developed based on the existing encoded morphotactics of Kazakh and Tatar (and occasionally Kyrgyz), with consultation of corpora, as described above.
 A dictionary \citep{bammatov1960} and a grammar \citep{olmesov2000} of Kumyk were also consulted as needed.
+
+In order to create the Kumyk transducer, we approached one of the major parts of speech at a time. We started with the nouns, copying the 
+continuation lexica\footnote{A continuation lexicon is a set of morphemes, for example, in the Turkic languages there is a continuation lexicon for `case' which includes the possible case suffixes.} (nominal morphotactics) from the Kazakh transducer.
+The suffixes were then replaced with the Kumyk suffixes according to the grammar \citep{olmesov2000}.
+Where the grammar was not explicit regarding a suffix form, we looked the corpus for possible forms and at their contexts.
+The same process was done for verbs.
 
@@
 % FIXME: make a graph for the kumyk one
+
+\begin{figure}
+%\centering
+\hspace{-25pt}
+    \rotatebox{-90}{
+\includegraphics[width=0.4\textwidth]{graphs/kumyk-yoldash-coverage}
+}
+
+\caption{Number of stems and coverage of corpus over time. Time is measured in days 
+starting from the 2nd October 2013. The graph shows that to reach 80\% coverage, took
+around a week, and to reach 90\% took another week.}
+\label{fig:kumcov}
+\end{figure}
 
@@
 \end{tabular}
- \caption{Number of stems in each of the categories}
+ \caption{Number of stems in each of the categories. The large number of noun stems as compared
+    to other parts of speech in the Kumyk transducer can be explained by the relative ease of 
+    categorising nouns as opposed to adjectives, adverbs and verbs. Adding a noun essentially involves
+    choosing between loan word and native word. Where adding stems from the other main categories 
+    requires more in depth categorisation.}
  \label{table:stems}
@@
 
+\subsection{Categorisation and tagset}
+
+The open categories in Turkic languages can be broadly split into two groups: nominals
+and verbals. The nominals group can be further split into nouns, adjectives and adverbs.
+We define nouns as stems that can be subjects of a finite verb, adjectives as stems that principally qualify 
+nouns, and adverbs as stems that principally modify verbs.
+
+Most of the stems in the lexicon can be used, with no extra suffixes in any of these functions, for 
+example the adjective жақсы in Kazakh can be used as attributively, жақсы китап `good book', adverbially
+жақсы ойнадыңыз `(you) played well', and substantively (as a noun) `жақсы келетін еді' `(the) good (ones) would come'. This
+is a productive process and relevant to the morphotactics as an adjective used substantively may take the 
+full range of case and possessive suffixes, and adjectives used adverbially will take a different set of 
+clitics to adjectives used substantively and attributively.
+
+However, this is not completely productive. A grammar which allows all adjectives to function
+as nouns and adverbs will overgenerate (and overanalyse). However, most grammars do not 
+mention that there are exceptions to this productive process.
+
+Other analysers for Turkic languages---such as \texttt{TRmorph} \citep{coltekin2010}---approach this 
+problem by positing a zero-derivation rule by which any noun can be `derived' into an adjective or 
+an adverb, any adjective can be `derived' into a noun or adverb, etc.
+
+Our approach is to describe this rather in terms of function (not unlike \cite{hengeveld92}). We 
+posit that nominals can be used either substantively, attributively or adverbially. Each part of speech 
+has a `default' function: nouns are by default substantive, adjectives are by default attributive, 
+and adverbs, adverbial. When they are used outside this default use they receive a tag to mark 
+their function. Where these functions are ambiguous, then they may be disambiguated 
+in context.
+
+The advantage of this approach as opposed to simply allowing one part of speech to derive into
+another is that it allows us to be more principled, and decreases overgeneration,
+for example, nouns used attributively will not take the whole adjective continuation (e.g.
+they will not take comparison, and cannot later substantivise). At the same time, it prevents 
+having to have two lexemes, one for each usage.
+
+There are certain lexemes which require more explicit categorisation, for example, the word 
+for `today'\footnote{Tatar: бүген; Kazakh: бүгін; Kumyk: бугюн. The word is a contraction of the words for `this' and `day'.} 
+cannot be used attributively without the presence of an extra morpheme -KI. Compound nouns
+using possessive morphology, such as Kazakh ауа райы `weather' are not well-formed
+without a possessive suffix.
+
+%\begin{tabular}{lccc}
+%          &  Substantive & Attributive & Adverbial \\ 
+%Noun      &     X        &    ?        &  ?\\ 
+%Adjective &     ?        &   X         &  ? \\ 
+%Adverb    &     ?        &   ?         &  X \\ 
+%\end{tabular}
+
+% FIXME: we should change this title
+\subsection{Technolinguistic and orthographical issues} % hidden phonology issues? ...?
+
+There is a class of phenomena encountered during the development of these transducers, united by the fact that the morphophonology needs information about the stem's phonological form that is not provided by the orthographic representation.
+These phenomena were left by \citet{washington2012} as future work, but have been extensively implemented in the transducers described by the current paper.
+
+Types of these phenomena include acronyms, ``irregular'' vowel harmony, Russian loanwords, and numerals.  A brief overview of each follows, along with examples, followed by details of our solution.
+
+%FIXME: description of each one, with examples
+\begin{itemize}
+ \item \textbf{Ambiguous and `irregular' harmony}: In Tatar, certain characters in the alphabet which represent vowels can be 
+   ambiguous between front vowels, such as е. For example: дәрес `lesson', дәресләр `lessons'; ел `year', 
+   еллар `years'. 
+ \item \textbf{Loan words}: The character ё in Russian words is pronounced as /jo/ where in native Kumyk words 
+   it represents the sound /ø/. These two pronunciations will select for different suffixes, /jo/ taking
+   back vowel suffixes самолётлар `aeroplanes', and /ø/ taking front vowel suffixes сёзлер `words'.
+ \item \textbf{Acronyms and numerals}: Acronyms and numerals are challenging as they will often be pronounced out loud
+   in their non-abbreviated forms, for example in Kazakh, 30-дан `from thirty', 5-тен `from 5'. The ablative
+   suffix -DAn alters for the phonology, but in the numeral string there is no indication of how they should alter.
+\end{itemize}
+
+Dealing with these phenomena would not be necessary if we were setting out to develop a simple computational
+model of the phonology of the language. However, a wide-coverage morphological analyser and generator needs to be 
+able to deal with all phenomena that are found in corpora.
+
+Regarding numerals, work has been done for Finnish in the finite-state framework \cite{karttunen06}, however
+this relies on converting all numerals to their fully spelt out form, which would involve complex operations 
+on the transducer.
+
+% FIXME: cyrillic in \texttt
+Our solution is to add phonological information at the end of morphemes that need it in the form of special ``abstract letters'' that trigger phonological processes at the morphophonological stage and are deleted.
+For example, the string \texttt{5<num><subst><abl>} would have
+the morphotactic representation 5-\{э\}\{с\}>\{D\}\{A\}н, where \{э\} and \{с\} stand for phonological triggers, and \{D\}\{A\}н is the morphotactic representation of the ablative.
+The symbol \{э\} signals front harmony, and \{с\} signals final unvoiced consonant.
+So, the rules that operate on \{D\} and \{A\} will be able to do the forward voicing assimilation \{D\} surfacing as т /t/ and the front harmony \{A\} surfacing as е /e/.
+
+%FIXME: example of implementation
+
+%FIXME: "performs all PHONOLOGY operations in parallel\footnote{the formalism is like SPE, but the rules are applied in parallel, somewhat like constraints in OT}" ← clarify
+%Since \texttt{twol} performs all operations in parallel, rules triggered by these must make reference to abstract letters that exist on the input tape but are null on the output tape.
+This complicates existing rules that have been designed to ignore null characters on the output tape.
+The \texttt{twol} rules resulting from the combination of processes simultaneously triggered by and ignoring output-null characters are rather ...
+uh, \textit{intricate} ...
+but they seem to work okay?
+Just don't poke them too hard.
+%FIXME
+
 \section{Evaluation}
@@
 %FIXME: sources
-We tested the coverage of the Kazakh and Tatar analysers over three separate domains: encyclopaedic text,\footnote{The following Wikipedia dumps were used: kkwiki-20131006-pages-articles.xml.bz2, FIXME.} news,\footnote{All content from \url{http://www.azattyq.org/} for 2010 was used for Kazakh, as well as all content from 2005 to 2011 on \url{http://tat.tatar-inform.ru} for Tatar.} and religion.\footnote{We used a Kazakh bible translation available from \url{https://kkitap.net/} and a Tatar translation of the New Testament available from \url{http://ibt.org.ru}}  As there is currently no Wikipedia in Kumyk, we tested only news and religion.\footnote{The bible corpus is from \url{http://ibt.org.ru/} and the news corpus consists of all Kumyk content from \url{http://sh-tavisi.etnosmi.ru/}.}
+We tested the coverage of the Kazakh and Tatar analysers over three separate domains: encyclopaedic text,\footnote{The following Wikipedia dumps were used: kkwiki-20131006-pages-articles.xml.bz2.} %FIXME
+ news,\footnote{All content from \url{http://www.azattyq.org/} for 2010 was used for Kazakh, as well as all content from 2005 to 2011 on \url{http://tat.tatar-inform.ru} for Tatar.} and religion.\footnote{We used a Kazakh bible translation available from \url{https://kkitap.net/} and a Tatar translation of the New Testament available from \url{http://ibt.org.ru}}  As there is currently no Wikipedia in Kumyk, we tested only news and religion.\footnote{The bible corpus is from \url{http://ibt.org.ru/} and the news corpus consists of all Kumyk content from \url{http://sh-tavisi.etnosmi.ru/}.}
 
@@
 \midrule
-\multirow{4}{*}{Kazakh} & Wikipedia &  -  &  - \\
+\multirow{4}{*}{Kazakh} & Wikipedia & 18.2M &  85.1 $\pm$ 1.33 \\
 	& News & 3.2M & - \\
-	& Religion & 848K & 92.7 $\pm$ 1.7 \\\cline{2-4}
+	& Religion & 849K & 92.7 $\pm$ 1.57 \\\cline{2-4}
 	& Average & - &  \\
@@
 
-The results for precision and recall will be presented in table~\ref{table:precrecall}.
+The results for precision and recall are be presented in table~\ref{table:precrecall}.
+
+The low recall for Kazakh can be explained by the fact that the corpus is much bigger, giving 
+more hapax words and proper names. There were 403 unknown words out of 1,000 in the Kazakh list.
+Of these 403, 160 were proper nouns, and 148 were common nouns. The lower precision for the 
+Tatar transducer can be partly explained by the less transparent orthography of Tatar.
+
+% tat:
+%Precision: 95.0339558573854
+%Recall: 85.65416985462892
+%F-score: 90.10060362173037
+
+% kaz:
+% Precision: 98.61202908129543
+% Recall: 57.986785853089785
+% F-score: 73.02985805188449
 
@@
 One direction for future work is to develop transducers for more languages.
-We have already constructed usable prototype transducers for three other Kypchak languages: Bashkir (N), Nogay (S), and Karakalpak (S).
-Since our ability to develop production-ready transducers is limited by availability of resources, including corpora in the languages and native-speaker consultants, the Western Kypchak languages (aside from Kumyk) have been more neglected by our team.
-However, these language communities would benefit from computational tools for their languages, and work on them may be bootstrapped from the existing transducers, so working on morphological transducers for these languages is also a priority.
+We have already constructed usable prototype transducers for three other Kypchak languages: Bashkir (North), Nogay (South), and Karakalpak (South).
+Since our ability to develop transducers is limited by availability of resources, including corpora in the languages and native-speaker consultants, the Western Kypchak languages (aside from Kumyk) have been more neglected by our team.
+However, these language communities would benefit from computational tools (such as spellcheckers) for their languages, and work on them may be bootstrapped from the existing transducers, so working on morphological transducers for these languages is also a priority.
+
+The principle obstacle to increasing coverage of the lexicons is the categorisation of stems.
+Future work would be investigating ways of automatically categorising stems by subcategory.
+For example, verbs stems by transitivity; adjective stems by if they can be used adverbially or substantively; etc.
 
@@
 We have described morphological transducers for three Kypchak languages---one from each branch of Kypchak---including the development process and performance of the analysers.
+The development of the third transducer (for a related language) was substantially quicker than the first two as a result of being able to reuse large portions of the morphotactic description from the first two transducers.
 
@@
 
+The authors would also like to express their gratitude to Aida Sundetova and Ağarhim Sultanmuradov
+for assistance in evaluating precision and recall.
+
 \bibliographystyle{lrec2012}

### 2014-03-14 kipchak-lrec-2013/kipchak-paper.tex
add Tatar example in Table 2
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@50736 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 % kaz: ^Құдай/Құдай<n><nom>/Құдай<n><attr>/Құдай<n><nom>+е<cop><p3><pl>/Құдай<n><nom>+е<cop><p3><sg>$ ^Өзінің/Өз<prn><ref><p3><sg><gen>$ ^жаратқандарының/жарат<v><tv><ger_past><pl><px3sp><gen>/жарат<v><tv><gpr_past><subst><pl><px3sp><gen>$ ^бәріне/бәрі<prn><qnt><px3sp><dat>$ ^қарап/қара<v><tv><prc_perf>/қара<v><tv><gna_perf>/қара<v><tv><prc_perf>$^,/,<cm>$ ^өте/өт<v><tv><prc_impf>/өте<adv>/өте<v><tv><imp><p2><sg>$ ^жақсы/жақсы<adj>/жақсы<adv>/жақсы<adj><advl>/жақсы<adj><subst><nom>/жақсы<adj>+е<cop><p3><pl>/жақсы<adj>+е<cop><p3><sg>/жақсы<adj><subst><nom>+е<cop><p3><pl>/жақсы<adj><subst><nom>+е<cop><p3><sg>$ ^екенін/е<cop><ger_past><px3sp><acc>$ ^көрді/көр<vaux><ifi><p3><pl>/көр<vaux><ifi><p3><sg>/көр<v><tv><ifi><p3><pl>/көр<v><tv><ifi><p3><sg>$^./.<sent>$
+% tat: ^Аллаһ/Аллаһ<n><nom>/Аллаһ<n><attr>/Аллаһ<n><nom>+и<cop><p3><sg>$ ^Үзе/Үз<prn><ref><p3><sg><nom>/Үз<prn><ref><p3><sg><nom>+и<cop><p3><sg>$ ^яраткан/ярат<v><tv><gpr_past>/яраткан<adj>/ярат<v><tv><ger_past><nom>/яраткан<adj><advl>/ярат<v><tv><past><p3><sg>/ярат<v><tv><gpr_past><subst><nom>/яраткан<adj><subst><nom>/яраткан<adj>+и<cop><p3><sg>/яраткан<adj><subst><nom>+и<cop><p3><sg>$ ^нәрсәләргә/нәрсә<prn><itg><pl><dat>$ ^карап/кара<vaux><prc_perf>/кара<vaux><gna_perf>/кара<v><tv><prc_perf>/кара<v><tv><gna_perf>$^,/,<cm>$ ^аларның/ул<prn><dem><pl><gen>/алар<prn><pers><p3><pl><gen>$ ^бик/бик<adv>/бик<n><nom>/бик<n><attr>/бик<n><nom>+и<cop><p3><sg>$ ^яхшы/яхшы<ij>/яхшы<adj>/яхшы<adj><advl>/яхшы<adj><subst><nom>/яхшы<adj>+и<cop><p3><sg>/яхшы<adj><subst><nom>+и<cop><p3><sg>$ ^икәнен/и<cop><ger_past><px3sp><acc>$ ^күрде/күр<v><tv><ifi><p3><sg>$^./.<sent>$
 % kum: ^Аллагь/аллагь<n><nom>/аллагь<n><attr>$ ^Оьзю/оьз<prn><ref><px3sp><nom>$ ^яратгъан/ярат<v><tv><gpr_past>/ярат<v><tv><gpr_past><subst><nom>/ярат<v><tv><ger_past><nom>/ярат<v><tv><past><p3><sg>$ ^затлагъа/зат<n><pl><dat>$ ^къарап/къара<v><tv><prc_perf>/къара<v><tv><gna_perf>$^,/,<cm>$ ^олар/о<prn><dem><pl><nom>/олар<prn><pers><p3><pl><nom>$ ^бек/бек<adv>$ ^яхшы/яхшы<adj>/яхшы<adj><subst><nom>$ ^экенин/э<cop><ger_past><px3sp><acc>$ ^гёрген/гёр<v><tv><gpr_past>/гёр<v><tv><gpr_past><subst><nom>/гёр<v><tv><ger_past><nom>/гёр<v><tv><past><p3><sg>$^../..<sent>$
@@
 Құдай Өзінің жаратқандарының бәріне қарап, өте жақсы екенін көрді. \\
-% ходай Үзенең яратканнарының *бəріне карап, бик яхшы икәнен күрде.
+% (close to kaz) Ходай Үзенең яратканнарының барысына карап, бик яхшы икәнен күрде. \\
+% (from the actual translation of the Bible) Аллаһ үзе яраткан нәрсәләрне күздән кичерде, һәм, менә, һәр нәрсәнең гаять яхшы икәнлеген күрде. \\
+Аллаһ Үзе яраткан нәрсәләргә карап, аларның бик яхшы икәнен күрде. \\
 Аллагь Оьзю яратгъан затлагъа къарап, олар бек яхшы экенин гёрген. \\
@@
 \hline
- Құдай\tags{<n><nom>} &  & Аллагь\tags{<n><nom>}  \\
- Өз\tags{<prn><ref><p3><sg><gen>} & & Оьз\tags{<prn><ref><px3sp><nom>} \\
- жарат\tags{<v><tv><ger\_past><pl><px3sp><gen>} & & ярат\tags{<v><tv><gpr\_past>} \\
- бәрі\tags{<prn><qnt><px3sp><dat>} & & зат\tags{<n><pl><dat>} \\
- қара\tags{<v><tv><gna\_perf>} & & къара\tags{<v><tv><gna\_perf>} \\ 
+ Құдай\tags{<n><nom>} & Аллаһ\tags{<n><nom>} & Аллагь\tags{<n><nom>}  \\
+ Өз\tags{<prn><ref><p3><sg><gen>} & Үз\tags{<prn><ref><p3><sg><nom>} & Оьз\tags{<prn><ref><px3sp><nom>} \\
+ жарат\tags{<v><tv><ger\_past><pl><px3sp><gen>} & ярат\tags{<v><tv><gpr\_past>} & ярат\tags{<v><tv><gpr\_past>} \\
+ бәрі\tags{<prn><qnt><px3sp><dat>} & нәрсә\tags{<prn><itg><pl><dat>} & зат\tags{<n><pl><dat>} \\
+ қара\tags{<v><tv><gna\_perf>} & кара\tags{<v><tv><gna\_perf>} & къара\tags{<v><tv><gna\_perf>} \\ 
 ,\tags{<cm>} & & ,\tags{<cm>} \\
- ---      & & олар\tags{<prn><pers><p3><pl><nom>} \\
- өт\tags{<adv>} & & бек\tags{<adv>} \\
- жақсы\tags{<adj>} & & яхшы\tags{<adj>} \\
- е\tags{<cop><ger\_past><px3sp><acc>} & & э\tags{<cop><ger\_past><px3sp><acc>} \\
- көр\tags{<v><tv><ifi><p3><sg>} & & гёр\tags{<v><tv><past><p3><sg>} \\  
- .\tags{<sent>} & & .\tags{<sent>} \\
+ ---      & алар\tags{<prn><pers><p3><pl><gen>} & олар\tags{<prn><pers><p3><pl><nom>} \\
+ өт\tags{<adv>} & бик\tags{<adv>} & бек\tags{<adv>} \\
+ жақсы\tags{<adj>} & яхшы\tags{<adj>} & яхшы\tags{<adj>} \\
+ е\tags{<cop><ger\_past><px3sp><acc>} & и\tags{<cop><ger\_past><px3sp><acc>} & э\tags{<cop><ger\_past><px3sp><acc>} \\
+ көр\tags{<v><tv><ifi><p3><sg>} & күр\tags{<v><tv><past><p3><sg>} & гёр\tags{<v><tv><past><p3><sg>} \\  
+ .\tags{<sent>} & .\tags{<sent>} & .\tags{<sent>} \\
 \hline

### 2014-03-14 kipchak-lrec-2013/kipchak-paper.tex
table
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@50738 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
+%
+%\textbf{Kazakh:} Құдай Өзінің жаратқандарының бәріне қарап, өте жақсы екенін көрді. \\
+%% (close to kaz) Ходай Үзенең яратканнарының барысына карап, бик яхшы икәнен күрде. \\
+%% (from the actual translation of the Bible) Аллаһ үзе яраткан нәрсәләрне күздән кичерде, һәм, менә, һәр нәрсәнең гаять яхшы икәнлеген күрде. \\
+%\textbf{Tatar:} Аллаһ Үзе яраткан нәрсәләргә карап, аларның бик яхшы икәнен күрде. \\
+%\textbf{Kumyk:} Аллагь Оьзю яратгъан затлагъа къарап, олар бек яхшы экенин гёрген. \\
+%
 \centering
 
-Құдай Өзінің жаратқандарының бәріне қарап, өте жақсы екенін көрді. \\
-% (close to kaz) Ходай Үзенең яратканнарының барысына карап, бик яхшы икәнен күрде. \\
-% (from the actual translation of the Bible) Аллаһ үзе яраткан нәрсәләрне күздән кичерде, һәм, менә, һәр нәрсәнең гаять яхшы икәнлеген күрде. \\
-Аллаһ Үзе яраткан нәрсәләргә карап, аларның бик яхшы икәнен күрде. \\
-Аллагь Оьзю яратгъан затлагъа къарап, олар бек яхшы экенин гёрген. \\
-
-~\\
-
+\scalebox{0.80}{
 \begin{tabular}{lll}
  \textbf{Kazakh} & \textbf{Tatar} & \textbf{Kumyk} \\
+\hline
+ Құдай Өзінің жаратқандарының  &  Аллаһ Үзе яраткан нәрсәләргә карап, & Аллагь Оьзю яратгъан затлагъа\\
+ бәріне қарап, өте жақсы екенін көрді. &  аларның бик яхшы икәнен күрде. & къарап, олар бек яхшы экенин гёрген. \\
 \hline
@@
 \end{tabular}
-
+}
 \label{table:exoutput}

### 2014-03-14 kipchak-lrec-2013/kipchak-paper.tex
ambiguous harmony
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@50739 72bbbca6-d526-0410-a7d9-f06f51895060

@@
  \item \textbf{Ambiguous and `irregular' harmony}: In Tatar, certain characters in the alphabet which represent vowels can be 
-   ambiguous between front vowels, such as е. For example: дәрес `lesson', дәресләр `lessons'; ел `year', 
+   ambiguous between a vowel and a 'yoticised' vowel, such as е. For example: дәрес `lesson', дәресләр `lessons'; ел `year', 
    еллар `years'. 

### 2014-03-14 kipchak-lrec-2013/kipchak-paper.tex
trrrrrrrkic
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@50740 72bbbca6-d526-0410-a7d9-f06f51895060

--- 2014-03-14 kipchak-lrec-2013/kipchak-paper.tex
@@
 	\includegraphics[width=\textwidth]{map/map}
-	\caption{The three sub-branches of Kypchak (North, South, West), roughly divided with black lines, showing the geographic distribution of the three languages for which transducers were developed (ISO639-3: \texttt{tat}, \texttt{kaz}, \texttt{kum}).}
+	\caption{The three sub-branches of Kypchak (North, South, West), roughly divided with black lines, showing the geographic distribution of the three languages for which transducers were developed: Tatar (\texttt{tat}), Kazakh (\texttt{kaz}), and Kumyk (\texttt{kum}).}
 	\label{map1}
@@
 %\centering
-\hspace{-25pt}
+\hspace{-15pt}
     \rotatebox{-90}{
@@
  Құдай\tags{<n><nom>} & Аллаһ\tags{<n><nom>} & Аллагь\tags{<n><nom>}  \\
- Өз\tags{<prn><ref><p3><sg><gen>} & Үз\tags{<prn><ref><p3><sg><nom>} & Оьз\tags{<prn><ref><px3sp><nom>} \\
+ Өз\tags{<prn><ref><px3sp><gen>} & Үз\tags{<prn><ref><px3sp><nom>} & Оьз\tags{<prn><ref><px3sp><nom>} \\
  жарат\tags{<v><tv><ger\_past><pl><px3sp><gen>} & ярат\tags{<v><tv><gpr\_past>} & ярат\tags{<v><tv><gpr\_past>} \\
@@
 \label{table:exoutput}
-\caption{An example of the output of each of the morphological transducers for the same sentence.}
+\caption{An example of the output of each of the morphological transducers for the same 
+  sentence (``And God saw every thing that he had made, and, behold, it was 
+  very good.'', Genesis 1:31). The sentence may be very roughly glossed as ``God own created things to looking,
+  their good being saw''. The output has been 
+  abbreviated to only show the appropriate tag in context, 
+  the actual output would be ambiguous.}
 
@@
  \item \textbf{Ambiguous and `irregular' harmony}: In Tatar, certain characters in the alphabet which represent vowels can be 
-   ambiguous between front vowels, such as е. For example: дәрес `lesson', дәресләр `lessons'; ел `year', 
-   еллар `years'. 
- \item \textbf{Loan words}: The character ё in Russian words is pronounced as /jo/ where in native Kumyk words 
-   it represents the sound /ø/. These two pronunciations will select for different suffixes, /jo/ taking
+   ambiguous between a vowel and a `yoticised' vowel, such as е, which is front or back. For 
+   example: дәрес `lesson', дәресләр `lessons'; ел `year', еллар `years'. 
+ \item \textbf{Loan words}: In Kumyk, the character ё in Russian words is pronounced as in Russian where in native words 
+   it represents the sound /ø/. These two pronunciations will select for different suffixes, the Russian pronunciation taking
    back vowel suffixes самолётлар `aeroplanes', and /ø/ taking front vowel suffixes сёзлер `words'.

### 2014-03-14 kipchak-lrec-2013/kipchak-paper.tex
some minor fixes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@50741 72bbbca6-d526-0410-a7d9-f06f51895060

@@
  ---      & алар\tags{<prn><pers><p3><pl><gen>} & олар\tags{<prn><pers><p3><pl><nom>} \\
- өт\tags{<adv>} & бик\tags{<adv>} & бек\tags{<adv>} \\
+ өте\tags{<adv>} & бик\tags{<adv>} & бек\tags{<adv>} \\
  жақсы\tags{<adj>} & яхшы\tags{<adj>} & яхшы\tags{<adj>} \\
@@
   sentence (``And God saw every thing that he had made, and, behold, it was 
-  very good.'', Genesis 1:31). The sentence may be very roughly glossed as ``God own created things to looking,
-  their good being saw''. The output has been 
+  very good.'', Genesis 1:31). The sentence may be very roughly glossed as ``God own-his created things/everything-at looking,
+  their very good being saw''. The output has been 
   abbreviated to only show the appropriate tag in context, 
@@
 Most of the stems in the lexicon can be used, with no extra suffixes in any of these functions, for 
-example the adjective жақсы in Kazakh can be used as attributively, жақсы китап `good book', adverbially
+example the adjective жақсы in Kazakh can be used as attributively, жақсы кітап `good book', adverbially
 жақсы ойнадыңыз `(you) played well', and substantively (as a noun) `жақсы келетін еді' `(the) good (ones) would come'. This
@@
 
-Regarding numerals, work has been done for Finnish in the finite-state framework \cite{karttunen06}, however
-this relies on converting all numerals to their fully spelt out form, which would involve complex operations 
+Regarding numerals, work has been done for Finnish in the finite-state framework \citep{karttunen06}; however, this relies on converting all numerals to their fully spelt out form, which would involve complex operations
 on the transducer.
@@
 Our solution is to add phonological information at the end of morphemes that need it in the form of special ``abstract letters'' that trigger phonological processes at the morphophonological stage and are deleted.
-For example, the string \texttt{5<num><subst><abl>} would have
-the morphotactic representation 5-\{э\}\{с\}>\{D\}\{A\}н, where \{э\} and \{с\} stand for phonological triggers, and \{D\}\{A\}н is the morphotactic representation of the ablative.
-The symbol \{э\} signals front harmony, and \{с\} signals final unvoiced consonant.
+For example, the string \texttt{5<num><subst><abl>}\footnote{The resulting forms, 5-тен\super{\texttt{kaz}}, 5-тән\super{\texttt{tat}}, and 5-ден\super{\texttt{kum}} would be pronounced бестен\super{\texttt{kaz}}, биштән\super{\texttt{tat}}, and бешден\super{\texttt{kum}}.} would have
+the morphotactic representation 5-\{э\}\{с\}>\{D\}\{A\}н, where \{э\} and \{с\} stand for phonological triggers, and \{D\}\{A\}н is the representation of the ablative morpheme at the morphotactic level.
+The symbol \{э\} signals that the following vowel needs to harmonise to a front unrounded vowel, and \{с\} signals that there is a final voiceless consonant.
 So, the rules that operate on \{D\} and \{A\} will be able to do the forward voicing assimilation \{D\} surfacing as т /t/ and the front harmony \{A\} surfacing as е /e/.

### 2014-03-14 kipchak-lrec-2013/kipchak-paper.tex
irregular harmony
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@50751 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \newfontfamily\symbl[Scale=MatchLowercase]{FreeSerif}
+\newfontfeature{IPA}{+mgrk}
+\newfontfamily\qipa[IPA,Scale=MatchLowercase]{FreeSerif}
 
@@
 %FIXME: description of each one, with examples
-\begin{itemize}
- \item \textbf{Ambiguous and `irregular' harmony}: In Tatar, certain characters in the alphabet which represent vowels can be 
-   ambiguous between a vowel and a `yoticised' vowel, such as е, which is front or back. For 
-   example: дәрес `lesson', дәресләр `lessons'; ел `year', еллар `years'. 
+\begin{description}
+ \item{\textbf{Ambiguous and `irregular' harmony}:} In Tatar, the `yoticised' vowel letters ‹е›, ‹я›, and ‹ю› each ambiguously represent a set of sounds.
+For example, ‹е› can represent the sounds of /j/ followed the sound of ‹э› (as in ‹егетләр› `boys') or ‹ы› (as in ‹еллар› `years'), and is also the non-initial orthographic variant of ‹э› (as in ‹дәресләр› `lessons').
+In Kumyk, ‹ё› and ‹ю› are used intervocalically to represent front rounded vowels (as in ‹гёзлер› `eyes' and ‹гюнлер› `days'), but word-initially can represent ‹й› followed by either a front (as in ‹юреклер› `hearts', ‹ёнкюлер› `darlings') or a back (as in ‹юлдузлар› `stars', ёллар `roads') vowel.
+In Kazakh, the vowels /{\qipa ɘ}/ ‹і› and /ə/ ‹ы› followed by a glide /w/ ‹у› or /j/ ‹й› are written ‹у› and ‹и›, depending on the glide; for example.
+The character ‹ю› is used in turn to represent ‹й› followed by either multi-phoneme value of ‹у›.
+%FIXME: good Kazakh examples
  \item \textbf{Loan words}: In Kumyk, the character ё in Russian words is pronounced as in Russian where in native words 
    it represents the sound /ø/. These two pronunciations will select for different suffixes, the Russian pronunciation taking
-   back vowel suffixes самолётлар `aeroplanes', and /ø/ taking front vowel suffixes сёзлер `words'.
+   back vowel suffixes самолётлар `aeroplanes', and /ø/ taking front vowel suffixes сёзлер `words'.  %FIXME: -ия in Tatar
  \item \textbf{Acronyms and numerals}: Acronyms and numerals are challenging as they will often be pronounced out loud
@@
    suffix -DAn alters for the phonology, but in the numeral string there is no indication of how they should alter.
-\end{itemize}
+\end{description}
 
@@
 Our solution is to add phonological information at the end of morphemes that need it in the form of special ``abstract letters'' that trigger phonological processes at the morphophonological stage and are deleted.
-For example, the string \texttt{5<num><subst><abl>}\footnote{The resulting forms, 5-тен\super{\texttt{kaz}}, 5-тән\super{\texttt{tat}}, and 5-ден\super{\texttt{kum}} would be pronounced бестен\super{\texttt{kaz}}, биштән\super{\texttt{tat}}, and бешден\super{\texttt{kum}}.} would have
-the morphotactic representation 5-\{э\}\{с\}>\{D\}\{A\}н, where \{э\} and \{с\} stand for phonological triggers, and \{D\}\{A\}н is the representation of the ablative morpheme at the morphotactic level.
+For example, the string \texttt{5<num><subst><abl>}\footnote{The resulting forms, 5-тен\super{\texttt{kaz}}, 5-тән\super{\texttt{tat}}, and 5-ден\super{\texttt{kum}} would be pronounced бестен\super{\texttt{kaz}}, биштән\super{\texttt{tat}}, and бешден\super{\texttt{kum}}.} has
+the morphotactic representation 5\{э\}\{с\}>-\{D\}\{A\}н, where \{э\} and \{с\} stand for phonological triggers, > represents a morpheme boundary, and \{D\}\{A\}н is the representation of the ablative morpheme at the morphotactic level.
 The symbol \{э\} signals that the following vowel needs to harmonise to a front unrounded vowel, and \{с\} signals that there is a final voiceless consonant.
-So, the rules that operate on \{D\} and \{A\} will be able to do the forward voicing assimilation \{D\} surfacing as т /t/ and the front harmony \{A\} surfacing as е /e/.
+So, with 5\{э\}\{с\} in the lexicon, and a consistent continuation lexicon specifying the underlying form of case affixes, the rules that operate on \{D\} and \{A\} are able to do produce the correct output form in each language, avoiding the incorrect default *5-дан.
 

### 2014-03-18 kipchak-lrec-2013/kipchak-paper.tex
better map, 2014 stylesheet, introduction
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@50925 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-\usepackage{lrec2012}
+\usepackage{lrec2014}
 
@@
 
-\author{Jonathan North Washington \\
-Departments of Linguistics and Central Eurasian Studies\\
-Indiana University\\
-Bloomington, IN 47405 (USA)\\
-\texttt{jonwashi@indiana.edu} \and
-Ilnar Salimzyanov  \\
-Kazan Federal University \\
-Kazan, Republic of Tatarstan\\
-(Russian Federation) \\
-\texttt{ilnar.salimzyan@gmail.com} \and 
-Francis M. Tyers\\
-HSL-fakultehta, \\
-UiT Norgga árktalaš universitehta, \\
-9019 Romsa (Norway)\\
-\texttt{francis.tyers@uit.no} 
-}
+%\author{Jonathan North Washington \\
+%Departments of Linguistics and Central Eurasian Studies\\
+%Indiana University\\
+%Bloomington, IN 47405 (USA)\\
+%\texttt{jonwashi@indiana.edu} \and
+%Ilnar Salimzyanov  \\
+%Kazan Federal University \\
+%Kazan, Republic of Tatarstan\\
+%(Russian Federation) \\
+%\texttt{ilnar.salimzyan@gmail.com} \and 
+%Francis M. Tyers\\
+%5HSL-fakultehta, \\
+%UiT Norgga árktalaš universitehta, \\
+%9019 Romsa (Norway)\\
+%\texttt{francis.tyers@uit.no} 
+%}
 
@@
 
-This paper describes the development of morphological transducers for three closly related languages: Kazakh, Tatar, and Kumyk.
+This paper describes the development of free/open-source morphological transducers for three closly related languages: Kazakh, Tatar, and Kumyk.
+Morphological transducers are computational models of the languages' morphology, and are used to output morphological analyses from word forms and vice-versa.
+
+These transducers were all developed as part of the Apertium project, which is aimed at creating rule-based MT systems for lesser resourced languages.
+As such, the transducers were developed with the intent that they be used as morphological analysers and generators in RBMT systems.
+However, this sort of finite-state transducer has a wide variety of usages, including for creating proofing tools such as spellcheckers and grammarcheckers.
+
+%the potential uses of morphological transducers is much broader; for example, we have found that they are easily converted to spell checkers.
+
+All three transducers were designed as ``ports'' of Apertium's Kyrgyz morphological transducer \citep{washington2012}.
+The Tatar and Kazakh transducers are used in a release-quality Kazakh-Tatar MT system \citep{salimzyanov2013}.
 
@@
 
-The transducers for these languages
+%The transducers for these languages
 
@@
 
-A finite-state transducer which models the morphology of a language has a wide variety of usages. For example,
-for creating proofing tools such as spellcheckers and grammarcheckers, and for machine translation.
+This paper overviews the languages (§\ref{sec:lgs}), provides background on these morphological transducers and related previous work (§\ref{sec:bg}), describes the development effort and contents of the transducers (§\ref{sec:methodology}), evaluates the effectiveness of these transducers (§\ref{sec:evaluation}), and summarises the findings (§\ref{sec:conclusions}).
 
-\section{Languages}
+\section{Languages}\label{sec:lgs}
 
 The three languages for which transducers were developed belong to the Northwestern branch of the Turkic family, which is often referred to as the Kypchak branch.
-This branch can be divided into three subbranches.
-Kumyk is a member of the Western Kypchak group, Tatar is a member of the Northern Kypchak group, and Kazakh is a member of the Southern Kypchak group \citep[82-83]{histofturkic}.
+This branch can be divided into three subbranches.\footnote{It is the belief of the authors of this paper that Kyrgyz constitutes a fourth branch.}  Kumyk is a member of the Western Kypchak group, Tatar is a member of the Northern Kypchak group, and Kazakh is a member of the Southern Kypchak group \citep[82-83]{histofturkic}.
 As such, each of these three languages represents one of the three branches of Kypchak.
@@
 
-\section{Background}
+\section{Background}\label{sec:bg}
 \subsection{Morphological transducers, previous work}
@@
 
-\section{Methodology}
+\section{Methodology}\label{sec:methodology}
 % cite Hengeveld (1992) ?
@@
 
-\section{Evaluation}
+\section{Evaluation}\label{sec:evaluation}
 
@@
 
-\section{Conclusions}
+\section{Conclusions}\label{sec:conclusions}
 

### 2014-03-18 kipchak-lrec-2013/kipchak-paper.tex
some minor edits, §4.4.1
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@50927 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 The three languages for which transducers were developed belong to the Northwestern branch of the Turkic family, which is often referred to as the Kypchak branch.
-This branch can be divided into three subbranches.\footnote{It is the belief of the authors of this paper that Kyrgyz constitutes a fourth branch.}  Kumyk is a member of the Western Kypchak group, Tatar is a member of the Northern Kypchak group, and Kazakh is a member of the Southern Kypchak group \citep[82-83]{histofturkic}.
-As such, each of these three languages represents one of the three branches of Kypchak.
+This branch can be divided into three subbranches.
+Kumyk is a member of the Western Kypchak group, Tatar is a member of the Northern Kypchak group, and Kazakh is a member of the Southern Kypchak group \citep[82-83]{histofturkic}.\footnote{It is the belief of the authors of this paper that Kyrgyz constitutes a fourth branch.}  As such, each of these three languages represents one of the three branches of Kypchak.
 The geographic distribution of the languages is shown in map \ref{map1}.
@@
 
-%FIXME: description of each one, with examples
-\begin{description}
- \item{\textbf{Ambiguous and `irregular' harmony}:} In Tatar, the `yoticised' vowel letters ‹е›, ‹я›, and ‹ю› each ambiguously represent a set of sounds.
-For example, ‹е› can represent the sounds of /j/ followed the sound of ‹э› (as in ‹егетләр› `boys') or ‹ы› (as in ‹еллар› `years'), and is also the non-initial orthographic variant of ‹э› (as in ‹дәресләр› `lessons').
-In Kumyk, ‹ё› and ‹ю› are used intervocalically to represent front rounded vowels (as in ‹гёзлер› `eyes' and ‹гюнлер› `days'), but word-initially can represent ‹й› followed by either a front (as in ‹юреклер› `hearts', ‹ёнкюлер› `darlings') or a back (as in ‹юлдузлар› `stars', ёллар `roads') vowel.
+\subsubsection{Ambiguous characters}
+In Tatar, the `yoticised' vowel letters ‹е›, ‹я›, and ‹ю› each ambiguously represent a set of sounds.
+For example, ‹е› is the non-initial orthographic variant of ‹э› (as in ‹дәресләр› `lessons'), but can also represent /j/ followed by the phoneme of ‹э› (as in ‹егетләр› `boys') or ‹ы› (as in ‹еллар› `years').
+In Kumyk, ‹ё› and ‹ю› are used intervocalically to represent front rounded vowels (as in ‹гёзлер› `eyes' and ‹гюнлер› `days'), but word-initially can represent ‹й› followed by either a front vowel (as in ‹юреклер› `hearts', ‹ёнкюлер› `darlings') or a back vowel (as in ‹юлдузлар› `stars', ‹ёллар› `roads').
 In Kazakh, the vowels /{\qipa ɘ}/ ‹і› and /ə/ ‹ы› followed by a glide /w/ ‹у› or /j/ ‹й› are written ‹у› and ‹и›, depending on the glide; for example.
 The character ‹ю› is used in turn to represent ‹й› followed by either multi-phoneme value of ‹у›.
-%FIXME: good Kazakh examples
- \item \textbf{Loan words}: In Kumyk, the character ё in Russian words is pronounced as in Russian where in native words 
+For example, ‹қиюда› /qəjəwdɑ/ `in the process of chopping down' and ‹киюде› /kɘjɘwdi̯ɘ/ `in the process of getting dressed' have back- and front-harmonising values, respectively, for ‹и› and ‹ю›.
+
+These problems can often be solved by building the context into the morphophonological rules, though the addition of new contexts and exceptions to existing contexts results in rather complex \texttt{twol} rules.
+
+\subsubsection{Loan words}
+%FIXME!!! 
+In Kumyk, the character ё in Russian words is pronounced as in Russian where in native words 
    it represents the sound /ø/. These two pronunciations will select for different suffixes, the Russian pronunciation taking
    back vowel suffixes самолётлар `aeroplanes', and /ø/ taking front vowel suffixes сёзлер `words'.  %FIXME: -ия in Tatar
- \item \textbf{Acronyms and numerals}: Acronyms and numerals are challenging as they will often be pronounced out loud
+
+\subsubsection{Acronyms and numerals}
+%FIXME!!!
+Acronyms and numerals are challenging as they will often be pronounced out loud
    in their non-abbreviated forms, for example in Kazakh, 30-дан `from thirty', 5-тен `from 5'. The ablative
    suffix -DAn alters for the phonology, but in the numeral string there is no indication of how they should alter.
-\end{description}
 

### 2014-03-18 kipchak-lrec-2013/kipchak-paper.tex
minor fixes to §4.4.1
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@50928 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 For example, ‹е› is the non-initial orthographic variant of ‹э› (as in ‹дәресләр› `lessons'), but can also represent /j/ followed by the phoneme of ‹э› (as in ‹егетләр› `boys') or ‹ы› (as in ‹еллар› `years').
+
 In Kumyk, ‹ё› and ‹ю› are used intervocalically to represent front rounded vowels (as in ‹гёзлер› `eyes' and ‹гюнлер› `days'), but word-initially can represent ‹й› followed by either a front vowel (as in ‹юреклер› `hearts', ‹ёнкюлер› `darlings') or a back vowel (as in ‹юлдузлар› `stars', ‹ёллар› `roads').
+
 In Kazakh, the vowels /{\qipa ɘ}/ ‹і› and /ə/ ‹ы› followed by a glide /w/ ‹у› or /j/ ‹й› are written ‹у› and ‹и›, depending on the glide; for example.
 The character ‹ю› is used in turn to represent ‹й› followed by either multi-phoneme value of ‹у›.
-For example, ‹қиюда› /qəjəwdɑ/ `in the process of chopping down' and ‹киюде› /kɘjɘwdi̯ɘ/ `in the process of getting dressed' have back- and front-harmonising values, respectively, for ‹и› and ‹ю›.
+For example, ‹қиюда› /qəjəwd{\qipa ɑ}/ `in the process of chopping down' and ‹киюде› /k{\qipa ɘ}j{\qipa ɘ}wd{\qipa i̯ɘ}/ `in the process of getting dressed' have back- and front-harmonising values, respectively, for ‹и› and ‹ю›.
 

### 2014-03-20 kipchak-lrec-2013/kipchak-paper.tex
hmm did not commi this
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@50967 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-\abstract{This paper describes the development of free/open-source finite-state morphological transducers for three Turkic languages---Kazakh, Tatar, and Kumyk---representing one language from each of the three subbranches of the Kypchak branch of Turkic.
+\abstract{This paper describes the development of free/open-source finite-state morphological transducers for three Turkic languages---Kazakh, Tatar, and Kumyk---representing one language from each of the three sub-branches of the Kypchak branch of Turkic.
 The finite-state toolkit used for the work is the Helsinki Finite-State Toolkit (HFST).
@@
 %FIXME: need a source, maybe something by Tommi?
-The objective of a morphological transducer is twofold: firstly to take surface forms (e.g., алдым) and generate all possible lexical forms, and secondly to take lexical forms (e.g.,  ал{\tt {\small <v><tv><ifi><p1><sg>}}, алд{\tt {\small <n><px1sg><nom>}}, etc.) and generate one or more surface forms.
+The objective of a morphological transducer is twofold: firstly to take surface forms (e.g., алдым) and generate all possible lexical forms, and secondly to take lexical forms (e.g.,  ал{\tt {\small <v><tv><ifi><p1><sg>}}, алд{\tt {\small <n><px1sg><nom>}},\footnote{For a description of the tags used throughout this paper, please see Appendix~\ref{app:glossary}.} etc.) and generate one or more surface forms.
 As they are implemented as finite-state transducers, they are reversible by default.
@@
 
-\scalebox{0.80}{
+\scalebox{0.81}{
 \begin{tabular}{lll}
@@
  қара\tags{<v><tv><gna\_perf>} & кара\tags{<v><tv><gna\_perf>} & къара\tags{<v><tv><gna\_perf>} \\ 
-,\tags{<cm>} & & ,\tags{<cm>} \\
+,\tags{<cm>} & ,\tags{<cm>} & ,\tags{<cm>} \\
  ---      & алар\tags{<prn><pers><p3><pl><gen>} & олар\tags{<prn><pers><p3><pl><nom>} \\
@@
   abbreviated to only show the appropriate tag in context, 
-  the actual output would be ambiguous.}
+  the actual output would be ambiguous. Refer to Appendix~\ref{app:glossary} for tag descriptions.}
 
@@
 
+\appendix
+
+\section{Glossary of symbols}
+\label{app:glossary}
+\centering
+\begin{tabular}{|c|l|}
+\hline
+ \textbf{Symbol} & \textbf{Description}\\
+\hline
+ \tags{n} & Noun  \\
+ \tags{v} & Verb  \\
+ \tags{adj} & Adjective  \\
+ \tags{adv} & Adverb  \\
+ \tags{prn} & Pronoun  \\
+ \tags{cm} &  Comma \\
+ \tags{cop} & Copula  \\
+ \tags{ifi} & Past definite  \\
+ \tags{past} & Past  \\
+ \tags{ger\_past} & Past gerund \\
+ \tags{gna\_past} & Past verbal adverb  \\
+ \tags{gpr\_past} & Past verbal adjective  \\
+ \tags{itg} & Interrogative  \\
+ \tags{nom} &  Nominative \\
+ \tags{gen} & Genitive  \\
+ \tags{dat} & Dative  \\
+ \tags{p1} & First person  \\
+ \tags{p3} & Third person  \\
+ \tags{px1sg} & First person, singular possessive  \\
+ \tags{px3sp} & Third person possessive  \\
+ \tags{pers} & Personal  \\
+ \tags{qnt} & Quantifier  \\
+ \tags{ref} & Reflexive  \\
+ \tags{sg} & Singular  \\
+ \tags{pl} &  Plural \\
+ \tags{tv} & Transitive  \\
+ \tags{sent} & End-of-sentence marker  \\
+
+\hline
+
+\end{tabular}
 \end{document}

### 2014-03-20 kipchak-lrec-2013/kipchak-paper.tex
updates
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@50999 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 The finite-state toolkit used for the work is the Helsinki Finite-State Toolkit (HFST).
-This paper describes how the development of a transducer for each subsequent language took less development time.
+This paper describes how the development of a transducer for each subsequent closely-related language took less development time.
 An evaluation is presented which shows that the transducers all have a reasonable coverage---around 90\%---on freely available corpora of the languages, and high precision over a manually verified test set.
@@
 As such, the transducers were developed with the intent that they be used as morphological analysers and generators in RBMT systems.
-However, this sort of finite-state transducer has a wide variety of usages, including for creating proofing tools such as spellcheckers and grammarcheckers.
+However, this sort of finite-state transducer has a wide variety of usages, including for creating proofing tools such as spellcheckers and grammarcheckers, morphological annotation for linguistic research, creating resources for learners and CALL, and lemmatised dictionary lookup.
 
@@
 All three transducers were designed as ``ports'' of Apertium's Kyrgyz morphological transducer \citep{washington2012}.
-The Tatar and Kazakh transducers are used in a release-quality Kazakh-Tatar MT system \citep{salimzyanov2013}.
+The Tatar and Kazakh transducers are used in a Kazakh--Tatar MT system \citep{salimzyanov2013}.
 
@@
 % In Apertium project, Turkic transducers were developed either as part of Turkic-to-Turkic machine translators, or they started as a "port" (in particular of morphotactics) of already existing Turkic transducers.
-
-%The transducers for these languages
-
-% FIXME: what are transducers used for ?
 
@@
 The three languages for which transducers were developed belong to the Northwestern branch of the Turkic family, which is often referred to as the Kypchak branch.
-This branch can be divided into three subbranches.
-Kumyk is a member of the Western Kypchak group, Tatar is a member of the Northern Kypchak group, and Kazakh is a member of the Southern Kypchak group \citep[82-83]{histofturkic}.\footnote{It is the belief of the authors of this paper that Kyrgyz constitutes a fourth branch.}  As such, each of these three languages represents one of the three branches of Kypchak.
-The geographic distribution of the languages is shown in map \ref{map1}.
+This branch can be divided into three sub-branches.
+Kumyk is a member of the Western Kypchak group, Tatar is a member of the Northern Kypchak group, and Kazakh is a member of the Southern Kypchak group \citep[82-83]{histofturkic}.\footnote{It is the belief of the authors of this paper that Kyrgyz constitutes a fourth branch.}   The geographic distribution of the languages is shown in map~\ref{map1}.
+%As such, each of these three languages represents one of the three branches of Kypchak.
 
@@
 	\includegraphics[width=\textwidth]{map/map}
-	\caption{The three sub-branches of Kypchak (North, South, West), roughly divided with black lines, showing the geographic distribution of the three languages for which transducers were developed: Tatar (\texttt{tat}), Kazakh (\texttt{kaz}), and Kumyk (\texttt{kum}).}
+	\caption{The three sub-branches of Kypchak (North, South, West), roughly divided with black lines, showing the geographic distribution of the three languages for which transducers were developed: Tatar (\texttt{tat}), Kazakh (\texttt{kaz}), and Kumyk (\texttt{kum}).
+The other codes represent Bashkir (\texttt{bak}), Kyrgyz (\texttt{kir}), Karakalpak (\texttt{kaa}), Nogay (\texttt{nog}), Karachay-Balkar (\texttt{krc}), Urum (\texttt{uum}), Crimean Tatar (\texttt{crh}) and Karaim (\texttt{kdr}).}
 	\label{map1}
@@
 
-These languages have different amounts of linguistic influence from other Turkic branches (e.g., moderate Oghuz (SE) influence in the Western group, slight Oghuz influence in the Northern group) and from Mongolic languages (moderate influence on the Southern group, lighter in the other groups), and all have heavy influence from Persian.
+These languages display different amounts of linguistic influence from other Turkic branches (e.g., moderate Oghuz (SE) influence in the Western group, slight Oghuz influence in the Northern group) and from Mongolic languages (moderate influence on the Southern group, lighter in the other groups), and all display heavy influence from Persian.
 
 \subsection{Kazakh}
-Kazakh /q{\symbl ɑ}z{\symbl ɑ}q/ is spoken primarily in Kazakhstan, where it is the national language, where it is co-official with Russian.
+Kazakh /q{\symbl ɑ}z{\symbl ɑ}q/ is spoken primarily in Kazakhstan, and it is the national language, where it is co-official with Russian.
 Large communities of native speakers also exist in China, neighbouring Central-Eurasian republics, and Mongolia.
@@
 \subsection{Tatar}
-Tatar /t{\symbl ɒ}t{\symbl ɑ}r/ is spoken in and around Tatarstan, by approximately 5.4 million people \citep{ethnologue}.
-It is co-official with Russian in Tatarstan --- a republic of the Russian Federation.
-A majority of native speakers of both languages are bilingual in Russian.
+Tatar /t{\symbl ɒ}t{\symbl ɑ}r/ is spoken in and around Tatarstan, --- a republic of the Russian Federation, where it is co-official with Russian.
+The majority of native speakers are bilingual in Russian.
+Tatar is spoken by approximately 5.4 million people \citep{ethnologue}.
 %, and the standard varieties our system FIXMEs are written in Cyrillic.
@@
 \section{Background}\label{sec:bg}
-\subsection{Morphological transducers, previous work}
+\subsection{Morphological transducers}
 %FIXME: need a source, maybe something by Tommi?
@@
 As they are implemented as finite-state transducers, they are reversible by default.
-For more information on usin finite-state transducers for morphological analysis and generation, the reader is referred to \cite{beesley2003}.
+For more information on using finite-state transducers for morphological analysis and generation, the reader is referred to \cite{beesley2003}.
 
 The transducers were designed based on the Helsinki Finite State Toolkit \citep{hfst/2011} which is a free/open-source reimplementation of the Xerox finite-state toolchain, popular in the field of morphological analysis.
-It implements both the \textbf{lexc} formalism for defining lexicons, and the \textbf{twol} and \textbf{xfst} formalisms for modeling morphophonological rules.
-It also supports other finite state transducer formalisms such as \textbf{sfst}.
-This toolkit has been chosen as it -- or the equivalent XFST -- has been widely used for other Turkic languages, such as Turkish \citep{coltekin2010}, Crimean~Tatar~\citep{altintas2001}, Turkmen \citep{tantug2006}, and Kyrgyz \citep{washington2012}, and is available under a free/open-source licence.
+It implements both the \texttt{lexc} formalism for defining lexicons, and the \texttt{twol} and \texttt{xfst} formalisms for modeling morphophonological rules.
+It also supports other finite state transducer formalisms such as \texttt{sfst}.
+This toolkit has been chosen as it -- or the equivalent XFST \cite{beesley2003} -- has been widely used for other Turkic languages, such as Turkish \citep{coltekin2010}, Crimean~Tatar~\citep{altintas2001}, Turkmen \citep{tantug2006}, and Kyrgyz \citep{washington2012}, and is available under a free/open-source licence.
 
 The authors learnt of another Kazakh morphological transducer in existence \citep{bekmanova2013} only after this paper was submitted and our transducer was released.
-We have not yet been able to evaluate this system or compare it to ours.
+The system is unfortunately not freely available so we have not been able to evaluate this system or compare it to ours.
 
 Creating morphological transducers in the above-mentioned formalisms involves encoding linguistic knowledge about the language in the formalisms.
-The lexc and twol formalisms resemble linguistic formalisms, allowing the coders to work with abstractions resembling linguistic categories such lexemes, morphemes, phonemes, and even archiphonemes.
+The \texttt{lexc} and \texttt{twol} formalisms resemble linguistic formalisms, allowing the coders to work with abstractions resembling linguistic categories such as lexemes, morphemes, phonemes, and even archiphonemes.
 
-\subsection{Description}
+%\subsection{Description}
 
@@
 
+% FIXME: This section needs to be totally rewritten I think.
+
 The three transducers discussed in this paper are for Kazakh, Tatar, and Kumyk.
-The Kazakh and Tatar transducers were originally created as part of an experimental Kazakh-Tatar machine translation system in December of 2010.
+The Kazakh and Tatar transducers were originally created as part of an experimental Kazakh--Tatar machine translation system in December of 2010.
 The Kazakh transducer was expanded during Google Code-In 2010 and 2011, and the Tatar transducer was expanded as part of a prototype Tatar and Bashkir machine translation system \citep{tyerswashingtonsalimzyanbattalov12}.
-The Kazakh-Tatar machine translation system, along with the two transducers, was expanded to production-level quality as part of a Google Summer of Code project in 2012 \citep{salimzyanov2013}.
+The Kazakh--Tatar machine translation system, along with the two transducers, was expanded as part of a Google Summer of Code project in 2012 \citep{salimzyanov2013}.
 
-The Kumyk transducer was developed starting at the beginning of October, 2013 as experiment to see how difficult it would be to extend lessons learned from the development of the Tatar and Kazakh transducers to a related language.
-While the Kazakh and Tatar transducers took around six months work to reach their current coverage level, the Kumyk transducer only took a couple weeks to reach this level of quality (see Figure~\ref{fig:kumcov}).
-This paper explores how the development of the Kumyk transducer benefitted from knowledge gained from the development of the Tatar and Kazakh transducers.
+The Kumyk transducer was developed starting at the beginning of October, 2013 as an experiment to see how difficult it would be to extend lessons learned from the development of the Tatar and Kazakh transducers to a related language.
+While the Kazakh and Tatar transducers took around six months work to reach their current coverage level, the Kumyk transducer only took a couple of weeks to reach the same level of coverage (see Figure~\ref{fig:kumcov}).
+This article explores how the development of the Kumyk transducer benefitted from knowledge gained from the development of the Tatar and Kazakh transducers.
 
@@
 Where the grammar was not explicit regarding a suffix form, we looked the corpus for possible forms and at their contexts.
-The same process was done for verbs.
+The same process was undertaken for verbs.
 
 % FIXME: how long it took
-% FIXME: make a graph for the kumyk one
 
@@
 
-\caption{Number of stems and coverage of corpus over time. Time is measured in days 
+\caption{Number of stems and coverage of corpus over time for the Kumyk transducer. Time is measured in days 
 starting from the 2nd October 2013. The graph shows that to reach 80\% coverage, took
@@
 
-Each transducer's lexc source consists of lists of stems, with each stem pointing at a complex continuation lexicon containing the appropriate morphology for the type of stem.
+Each transducer's \texttt{lexc} source consists of lists of stems, with each stem pointing at a complex continuation lexicon containing the appropriate morphology for the type of stem.
 
@@
     categorising nouns as opposed to adjectives, adverbs and verbs. Adding a noun essentially involves
-    choosing between loan word and native word. Where adding stems from the other main categories 
+    choosing between loan word and native word. Adding stems from the other main categories 
     requires more in depth categorisation.}
@@
   sentence (``And God saw every thing that he had made, and, behold, it was 
-  very good.'', Genesis 1:31). The sentence may be very roughly glossed as ``God own-his created things/everything-at looking,
+  very good.'', Genesis 1:31). The sentence may be very roughly glossed as ``God own-his created things-to looking,
   their very good being saw''. The output has been 
-  abbreviated to only show the appropriate tag in context, 
-  the actual output would be ambiguous. Refer to Appendix~\ref{app:glossary} for tag descriptions.}
+  abbreviated to only show the appropriate tag sequence in context, the actual output would give multiple interpretations
+  that would require further disambiguation. Refer to Appendix~\ref{app:glossary} for tag descriptions.}
 
@@
 as nouns and adverbs will overgenerate (and overanalyse). However, most grammars do not 
-mention that there are exceptions to this productive process.
+mention that there are exceptions to this process.
 
@@
 
-There are certain lexemes which require more explicit categorisation, for example, the word 
+There are certain lexemes that require more explicit categorisation, for example, the word 
 for `today'\footnote{Tatar: бүген; Kazakh: бүгін; Kumyk: бугюн. The word is a contraction of the words for `this' and `day'.} 
@@
 using possessive morphology, such as Kazakh ауа райы `weather' are not well-formed
-without a possessive suffix.
+without a possessive suffix. %FIXME: Explain the compound 'weather'.
 
@@
 There is a class of phenomena encountered during the development of these transducers, united by the fact that the morphophonology needs information about the stem's phonological form that is not provided by the orthographic representation.
-These phenomena were left by \citet{washington2012} as future work, but have been extensively implemented in the transducers described by the current paper.
+These phenomena were left by \citet{washington2012} for future work, but have been extensively implemented in the transducers described in the current paper.
 
@@
 
-Regarding numerals, work has been done for Finnish in the finite-state framework \citep{karttunen06}; however, this relies on converting all numerals to their fully spelt out form, which would involve complex operations
+Regarding numerals, work has been done for Finnish in the finite-state framework by \citep{karttunen06}; however, this relies on converting all numerals to their fully spelt out form, which would involve complex operations
 on the transducer.
@@
 We have evaluated the morphological analysers in two ways.
-The first was by calculating the naïve coverage\footnote{Naïve coverage refers to the percentage of surface forms in a given corpora that receive at least one analysis.
-Forms counted by this measure may have other analyses which are not delivered by the transducer.} and mean ambiguity  on freely available corpora.
-The mean ambiguity measure was calculated by performing an evaluation of precision and recall on some smaller, hand-validated test sets.
+The first was by calculating the naïve coverage and mean ambiguity on freely available corpora.
+Naïve coverage refers to the percentage of surface forms in a given corpora that receive at least one analysis.
+Forms counted by this measure may have other analyses which are not delivered by the transducer.
+The mean ambiguity measure was calculated as the average number of analyses returned per token in the corpus.
+
+%by performing an evaluation of precision and recall on some smaller, hand-validated test sets.
 
@@
 
-The results for precision and recall are be presented in table~\ref{table:precrecall}.
+The results for precision and recall are presented in table~\ref{table:precrecall}.
 
 The low recall for Kazakh can be explained by the fact that the corpus is much bigger, giving 
-more hapax words and proper names. There were 403 unknown words out of 1,000 in the Kazakh list.
+more hapax words and proper names. There were 403 out-of-vocabulary words out of 1,000 in the Kazakh list.
 Of these 403, 160 were proper nouns, and 148 were common nouns. The lower precision for the 
@@
 	\toprule
-		\textbf{Language} & \textbf{Precision} & \textbf{Recall} \\
+		\textbf{Language} & \textbf{Precision (\%)} & \textbf{Recall (\%)} \\
 	\midrule
@@
 	\end{tabular}
-	\caption{Precision and recall}
+	\caption{Precision and recall presented as percentages}
 	\label{table:precrecall}
@@
 Future work would be investigating ways of automatically categorising stems by subcategory.
-For example, verbs stems by transitivity; adjective stems by if they can be used adverbially or substantively; etc.
+For example, verbs stems by transitivity; adjective stems by whether  they can be used adverbially or substantively; etc.
 

### 2014-03-21 kipchak-lrec-2013/kipchak-paper.tex
some work for the night
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@51022 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-In Kumyk, ‹ё› and ‹ю› are used intervocalically to represent front rounded vowels (as in ‹гёзлер› `eyes' and ‹гюнлер› `days'), but word-initially can represent ‹й› followed by either a front vowel (as in ‹юреклер› `hearts', ‹ёнкюлер› `darlings') or a back vowel (as in ‹юлдузлар› `stars', ‹ёллар› `roads').
+In Kumyk, ‹ё› and ‹ю› are used between consonants to represent front rounded vowels (as in ‹гёзлер› `eyes' and ‹гюнлер› `days'), but word-initially can represent ‹й› followed by either a front vowel (as in ‹юреклер› `hearts', ‹ёнкюлер› `darlings') or a back vowel (as in ‹юлдузлар› `stars', ‹ёллар› `roads').
 
@@
 \subsubsection{Loan words}
-%FIXME!!! 
-In Kumyk, the character ё in Russian words is pronounced as in Russian where in native words 
-   it represents the sound /ø/. These two pronunciations will select for different suffixes, the Russian pronunciation taking
-   back vowel suffixes самолётлар `aeroplanes', and /ø/ taking front vowel suffixes сёзлер `words'.  %FIXME: -ия in Tatar
+In Kazakh, Tatar, and Kumyk, most Russian loanwords are spelled using Russian orthography, despite incompatibilities between Russian letter-to-phoneme mapping and that of the Turkic languages.
+
+In Kumyk, the character ‹ё› is used between consonants to represent a mid front rounded vowel (e.g., ‹сёзлер› `words'); however, in Russian words, ‹ё› is used after a consonant to represent the sound of ‹о› while indicating that the previous consonant is palatalised (e.g., ‹самолётлар› `aeroplanes').
+
+%FIXME: gloss `scientists' correct??
+In Tatar, the character ‹и› normally represents a high front vowel (e.g., ‹галимләр› `scientists'), but in Russian words it harmonises as a back vowel (e.g., ‹артистлар› `artists').
+
+In Kazakh, ‹е› is a front diphthong (e.g., ‹елдің› `counter-\gmk{gen}'), but in Russian 
+% FIXME: -ев in Kazakh
 
@@
 \centering
-\begin{tabular}{|c|l|}
-\hline
+\begin{tabular}{cl}
+\toprule
  \textbf{Symbol} & \textbf{Description}\\
-\hline
+\midrule
  \tags{n} & Noun  \\
@@
 
-\hline
+\bottomrule
 

### 2014-03-21 kipchak-lrec-2013/kipchak-paper.tex
§4 is getting there...
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@51042 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \newfontfamily\qipa[IPA,Scale=MatchLowercase]{FreeSerif}
+\newfontfamily\qgmk[Scale=MatchLowercase,Letters=SmallCaps]{Linux Libertine O}
 
+\newcommand{\gmk}[1]{{\qgmk #1}}
 \newcommand{\tags}[1]{\texttt{#1}}
@@
 There are certain lexemes that require more explicit categorisation, for example, the word 
-for `today'\footnote{Tatar: бүген; Kazakh: бүгін; Kumyk: бугюн. The word is a contraction of the words for `this' and `day'.} 
+for `today'\footnote{Tatar: бүген; Kazakh: бүгін; Kumyk: бугюн. The word is etymologically composed of the words for `this' and `day'.} 
 cannot be used attributively without the presence of an extra morpheme -KI. Compound nouns
@@
 
-% FIXME: we should change this title
-\subsection{Technolinguistic and orthographical issues} % hidden phonology issues? ...?
+% FIXME: how's the title now?
+\subsection{Orthography-phonology mapping issues}
 
 There is a class of phenomena encountered during the development of these transducers, united by the fact that the morphophonology needs information about the stem's phonological form that is not provided by the orthographic representation.
-These phenomena were left by \citet{washington2012} for future work, but have been extensively implemented in the transducers described in the current paper.
+These phenomena were left by \citet{washington2012} for future work, but many of them been implemented in the transducers described in the current paper.
 
-Types of these phenomena include acronyms, ``irregular'' vowel harmony, Russian loanwords, and numerals.  A brief overview of each follows, along with examples, followed by details of our solution.
+Types of these phenomena include characters used to represent multiple phonological forms, loanwords, and acronyms and numerals.
+A brief overview of each follows, along with examples and details of our solution.
 
@@
 
-\subsubsection{Loan words}
+\subsubsection{Loanwords}
 In Kazakh, Tatar, and Kumyk, most Russian loanwords are spelled using Russian orthography, despite incompatibilities between Russian letter-to-phoneme mapping and that of the Turkic languages.
@@
 
-%FIXME: gloss `scientists' correct??
-In Tatar, the character ‹и› normally represents a high front vowel (e.g., ‹галимләр› `scientists'), but in Russian words it harmonises as a back vowel (e.g., ‹артистлар› `artists').
+In Tatar, the character ‹и› normally represents a high front vowel (e.g., ‹галимнәр› `scientists'), but in Russian words it harmonises as a back vowel (e.g., ‹артистлар› `artists').
 
-In Kazakh, ‹е› is a front diphthong (e.g., ‹елдің› `counter-\gmk{gen}'), but in Russian 
-% FIXME: -ев in Kazakh
+In Kazakh, ‹е› is a front diphthong (e.g., ‹елдің› `country-\gmk{gen}'), but in e.g., family names with Russian morphology, it harmonises as a back vowel (e.g., ‹Назарбаевтың› `Nazarbayev-\gmk{gen}').
+
+Our solution for these consistent exceptions is to make a separate continuation lexicon for e.g., Russian nouns.
+This continuation lexicon adds a character that the phonology deletes, but that triggers phonological rules.
+
+Since \texttt{twol} ``phonological'' rules can be understood to be applied all at one stage (as opposed to being ordered), rules triggered by these characters must make reference to abstract characters that exist on the input tape but are null on the output tape.
+This complicates existing rules that have been designed to ignore null characters on the output tape.
+Hence, the \texttt{twol} rules resulting from the combination of processes simultaneously triggered by and ignoring output-null characters can quickly become quite unwieldy.
 
@@
 
+This solution has the same issue as loanwords above.
 %FIXME: "performs all PHONOLOGY operations in parallel\footnote{the formalism is like SPE, but the rules are applied in parallel, somewhat like constraints in OT}" ← clarify

### 2014-03-22 kipchak-lrec-2013/kipchak-paper.tex
Kazakh coverage numbers, some other edits
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@51052 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 This branch can be divided into three sub-branches.
-Kumyk is a member of the Western Kypchak group, Tatar is a member of the Northern Kypchak group, and Kazakh is a member of the Southern Kypchak group \citep[82-83]{histofturkic}.\footnote{It is the belief of the authors of this paper that Kyrgyz constitutes a fourth branch.}   The geographic distribution of the languages is shown in map~\ref{map1}.
+Kumyk is a member of the Western Kypchak group, Tatar is a member of the Northern Kypchak group, and Kazakh is a member of the Southern Kypchak group \citep[82-83]{histofturkic}.\footnote{It is the professional opinion of the authors of this paper that Kyrgyz constitutes a fourth branch.}   The geographic distribution of the languages is shown in map~\ref{map1}.
 %As such, each of these three languages represents one of the three branches of Kypchak.
@@
 
-We have evaluated the morphological analysers in two ways.
+%FIXME: revisions
+We have evaluated the morphological analysers\footnote{The versions of the analysers used were: r50547 for Kazakh, r××××× for Tatar, and r××××× for Kumyk.} in two ways.
 The first was by calculating the naïve coverage and mean ambiguity on freely available corpora.
@@
 We tested the coverage of the Kazakh and Tatar analysers over three separate domains: encyclopaedic text,\footnote{The following Wikipedia dumps were used: kkwiki-20131006-pages-articles.xml.bz2.} %FIXME
- news,\footnote{All content from \url{http://www.azattyq.org/} for 2010 was used for Kazakh, as well as all content from 2005 to 2011 on \url{http://tat.tatar-inform.ru} for Tatar.} and religion.\footnote{We used a Kazakh bible translation available from \url{https://kkitap.net/} and a Tatar translation of the New Testament available from \url{http://ibt.org.ru}}  As there is currently no Wikipedia in Kumyk, we tested only news and religion.\footnote{The bible corpus is from \url{http://ibt.org.ru/} and the news corpus consists of all Kumyk content from \url{http://sh-tavisi.etnosmi.ru/}.}
+ news,\footnote{All content from \url{http://www.azattyq.org/} for 2010 was used for Kazakh, as well as all content from 2005 to 2011 on \url{http://tat.tatar-inform.ru} for Tatar.} and religion.\footnote{We used a Kazakh bible translation available from \url{https://kkitap.net/} and Quran translation available from \url{http://kuran.kz/}, as well as Tatar and Kumyk translations of the New Testament available from \url{http://ibt.org.ru}.}  As there is currently no Wikipedia in Kumyk, we tested only news and religion.\footnote{The bible corpus is from \url{http://ibt.org.ru/} and the news corpus consists of all Kumyk content from \url{http://sh-tavisi.etnosmi.ru/}.}
 
@@
 \midrule
-\multirow{4}{*}{Kazakh} & Wikipedia & 18.2M &  85.1 $\pm$ 1.33 \\
-	& News & 3.2M & - \\
-	& Religion & 849K & 92.7 $\pm$ 1.57 \\\cline{2-4}
+\multirow{4}{*}{Kazakh} & Wikipedia & 18.2M & 85.61 $\pm$ 1.37 \\
+	& News & 3.8M & 92.12 $\pm$ 2.72  \\
+	& Religion & 851K & 92.49 $\pm$ 1.66 \\\cline{2-4}
 	& Average & - &  \\

### 2014-03-22 kipchak-lrec-2013/kipchak-paper.tex
18.2M words from wc -w → 25.6M analyses from apertium
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@51053 72bbbca6-d526-0410-a7d9-f06f51895060
more coverage, etc.
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@51054 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 %FIXME: usually star comes first, then daggers
-\name{Jonathan North Washington$^\dagger$, Ilnar Salimzyanov$^\ddagger$, Francis M. Tyers$^\star$}
+\name{Jonathan North Washington$^\star$, Ilnar Salimzyanov$^\dagger$, Francis M. Tyers$^\ddagger$}
 
-\address{$^\dagger$Departments of Linguistics and Central Eurasian Studies\\
+\address{$^\star$Departments of Linguistics and Central Eurasian Studies\\
 Indiana University\\
@@
 \texttt{jonwashi@indiana.edu} \and
-$^\ddagger$Institut für Maschinelle Sprachverarbeitung \\
-Universität Stuttgart\\
-Stuttgart (Germany) \\
-\texttt{ilnar@ilnar} \and 
-$^\star$HSL-fakultehta \\  
+$^\dagger$Kazan Federal University \\
+Kazan, Republic of Tatarstan\\
+(Russian Federation) \\
+\texttt{ilnar.salimzyan@gmail.com} \and 
+$^\ddagger$HSL-fakultehta \\  
 UiT Norgga árktalaš universitehta \\
@@
 %FIXME: revisions
-We have evaluated the morphological analysers\footnote{The versions of the analysers used were: r50547 for Kazakh, r××××× for Tatar, and r××××× for Kumyk.} in two ways.
+We have evaluated the morphological analysers\footnote{The versions of the analysers used were: r50547 for Kazakh, r50260 for Tatar, and r××××× for Kumyk.} in two ways.
 The first was by calculating the naïve coverage and mean ambiguity on freely available corpora.
@@
 %FIXME: sources
-We tested the coverage of the Kazakh and Tatar analysers over three separate domains: encyclopaedic text,\footnote{The following Wikipedia dumps were used: kkwiki-20131006-pages-articles.xml.bz2.} %FIXME
- news,\footnote{All content from \url{http://www.azattyq.org/} for 2010 was used for Kazakh, as well as all content from 2005 to 2011 on \url{http://tat.tatar-inform.ru} for Tatar.} and religion.\footnote{We used a Kazakh bible translation available from \url{https://kkitap.net/} and Quran translation available from \url{http://kuran.kz/}, as well as Tatar and Kumyk translations of the New Testament available from \url{http://ibt.org.ru}.}  As there is currently no Wikipedia in Kumyk, we tested only news and religion.\footnote{The bible corpus is from \url{http://ibt.org.ru/} and the news corpus consists of all Kumyk content from \url{http://sh-tavisi.etnosmi.ru/}.}
+We tested the coverage of the Kazakh and Tatar analysers over three separate domains: encyclopaedic text,\footnote{The following Wikipedia dumps were used: \texttt{kkwiki-20131006-pages-articles.xml.bz2} and \texttt{ttwiki-20130225-pages-articles.xml.bz2}.}
+ news,\footnote{All content from \url{http://www.azattyq.org/} for 2010 was used for Kazakh, as well as all content from 2005 to 2011 on \url{http://tat.tatar-inform.ru} for Tatar.} and religion.\footnote{We used a Kazakh bible translation available from \url{https://kkitap.net/} and Quran translation available from \url{http://kuran.kz/}, and a Tatar translation of the New Testament available from \url{http://ibt.org.ru}.}  As there is currently no Wikipedia in Kumyk, we tested only news and religion.\footnote{The bible corpus is from \url{http://ibt.org.ru/} and the news corpus consists of all Kumyk content from \url{http://sh-tavisi.etnosmi.ru/}.}
 
@@
 \midrule
-\multirow{4}{*}{Kazakh} & Wikipedia & 18.2M & 85.61 $\pm$ 1.37 \\
+\multirow{4}{*}{Kazakh} & Wikipedia & 25.6M & 85.61 $\pm$ 1.37 \\
 	& News & 3.8M & 92.12 $\pm$ 2.72  \\
 	& Religion & 851K & 92.49 $\pm$ 1.66 \\\cline{2-4}
-	& Average & - &  \\
+	& Average & - & 90.07 $\pm$ 1.91 \\
 \midrule
-\multirow{4}{*}{Tatar} & Wikipedia & 128K &  - \\
+\multirow{4}{*}{Tatar} & Wikipedia & 159K & 86.35 $\pm$ 2.17 \\
 	& News & 4.6M & - \\

### 2014-03-22 kipchak-lrec-2013/kipchak-paper.tex
more coverage, etc.
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@51055 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 We tested the coverage of the Kazakh and Tatar analysers over three separate domains: encyclopaedic text,\footnote{The following Wikipedia dumps were used: \texttt{kkwiki-20131006-pages-articles.xml.bz2} and \texttt{ttwiki-20130225-pages-articles.xml.bz2}.}
- news,\footnote{All content from \url{http://www.azattyq.org/} for 2010 was used for Kazakh, as well as all content from 2005 to 2011 on \url{http://tat.tatar-inform.ru} for Tatar.} and religion.\footnote{We used a Kazakh bible translation available from \url{https://kkitap.net/} and Quran translation available from \url{http://kuran.kz/}, and a Tatar translation of the New Testament available from \url{http://ibt.org.ru}.}  As there is currently no Wikipedia in Kumyk, we tested only news and religion.\footnote{The bible corpus is from \url{http://ibt.org.ru/} and the news corpus consists of all Kumyk content from \url{http://sh-tavisi.etnosmi.ru/}.}
+ news,\footnote{All content from \url{http://www.azattyq.org/} for 2010 was used for Kazakh, as well as all content from 2005 to 2011 on \url{http://tat.tatar-inform.ru} for Tatar.} and religion.\footnote{We used a Kazakh bible translation available from \url{https://kkitap.net/} and Quran translation available from \url{http://kuran.kz/}, as well as a Tatar translation of the New Testament available from \url{http://ibt.org.ru} and Quran translation available from \url{http://tanzil.net}.}  As there is currently no Wikipedia in Kumyk, we tested only news and religion.\footnote{The bible corpus is from \url{http://ibt.org.ru/} and the news corpus consists of all Kumyk content from \url{http://sh-tavisi.etnosmi.ru/}.}
 
@@
 \multirow{4}{*}{Tatar} & Wikipedia & 159K & 86.35 $\pm$ 2.17 \\
-	& News & 4.6M & - \\
-	& Religion & 205K & 88.7 $\pm$ 0.5 \\\cline{2-4}
+	& News & 5.2M & 89.75 $\pm$ 0.07 \\
+	& Religion & 382K & 91.25 $\pm$ 2.55 \\\cline{2-4}
 	& Average & - &  \\

### 2014-03-22 kipchak-lrec-2013/kipchak-paper.tex
tatar averages
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@51056 72bbbca6-d526-0410-a7d9-f06f51895060
Kumyk coverage numbers
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@51057 72bbbca6-d526-0410-a7d9-f06f51895060

--- 2014-03-22 kipchak-lrec-2013/kipchak-paper.tex
@@
 %FIXME: revisions
-We have evaluated the morphological analysers\footnote{The versions of the analysers used were: r50547 for Kazakh, r50260 for Tatar, and r××××× for Kumyk.} in two ways.
+We have evaluated the morphological analysers\footnote{The versions of the analysers used were: r50547 for Kazakh, r50260 for Tatar, and r50300 for Kumyk.} in two ways.
 The first was by calculating the naïve coverage and mean ambiguity on freely available corpora.
@@
 We tested the coverage of the Kazakh and Tatar analysers over three separate domains: encyclopaedic text,\footnote{The following Wikipedia dumps were used: \texttt{kkwiki-20131006-pages-articles.xml.bz2} and \texttt{ttwiki-20130225-pages-articles.xml.bz2}.}
- news,\footnote{All content from \url{http://www.azattyq.org/} for 2010 was used for Kazakh, as well as all content from 2005 to 2011 on \url{http://tat.tatar-inform.ru} for Tatar.} and religion.\footnote{We used a Kazakh bible translation available from \url{https://kkitap.net/} and Quran translation available from \url{http://kuran.kz/}, and a Tatar translation of the New Testament available from \url{http://ibt.org.ru}.}  As there is currently no Wikipedia in Kumyk, we tested only news and religion.\footnote{The bible corpus is from \url{http://ibt.org.ru/} and the news corpus consists of all Kumyk content from \url{http://sh-tavisi.etnosmi.ru/}.}
+ news,\footnote{All content from \url{http://www.azattyq.org/} for 2010 was used for Kazakh, as well as all content from 2005 to 2011 on \url{http://tat.tatar-inform.ru} for Tatar.} and religion.\footnote{We used a Kazakh bible translation available from \url{https://kkitap.net/} and Quran translation available from \url{http://kuran.kz/}, as well as a Tatar translation of the New Testament available from \url{http://ibt.org.ru} and Quran translation available from \url{http://tanzil.net}.}  As there is currently no Wikipedia in Kumyk, we tested only news and religion.\footnote{The bible corpus consists of Genesis and the New Testament, as available from \url{http://ibt.org.ru/}, and the news corpus consists of all Kumyk content from \url{http://sh-tavisi.etnosmi.ru/}.}
 
@@
 	& Religion & 851K & 92.49 $\pm$ 1.66 \\\cline{2-4}
-	& Average & - & 90.07 $\pm$ 1.91 \\
+	& Average & -- & 90.07 $\pm$ 1.91 \\
 \midrule
 \multirow{4}{*}{Tatar} & Wikipedia & 159K & 86.35 $\pm$ 2.17 \\
-	& News & 4.6M & - \\
-	& Religion & 205K & 88.7 $\pm$ 0.5 \\\cline{2-4}
-	& Average & - &  \\
+	& News & 5.2M & 89.75 $\pm$ 0.07 \\
+	& Religion & 382K & 91.25 $\pm$ 2.55 \\\cline{2-4}
+	& Average & -- & 89.12 $\pm$ 1.60 \\
 \midrule
@@
 \multirow{4}{*}{Kumyk} & Wikipedia & -- & -- \\
-        & News & 287K &  91.4 $\pm$ 0.7 \\
-	& Religion & 227K & 92.0 $\pm$ 1.1 \\\cline{2-4}
-	& Average & - & 91.7 $\pm$ 0.7 \\
+        & News & 286K &  91.10 $\pm$ 0.86 \\
+	& Religion & 227K & 92.47 $\pm$ 1.03 \\\cline{2-4}
+	& Average & -- & 91.78 $\pm$ 0.94 \\
 \bottomrule

### 2014-03-22 kipchak-lrec-2013/kipchak-paper.tex
updated numbers of stems
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@51058 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 	\includegraphics[width=\textwidth]{map/map}
-	\caption{The three sub-branches of Kypchak (North, South, West), roughly divided with black lines, showing the geographic distribution of the three languages for which transducers were developed: Tatar (\texttt{tat}), Kazakh (\texttt{kaz}), and Kumyk (\texttt{kum}).
-The other codes represent Bashkir (\texttt{bak}), Kyrgyz (\texttt{kir}), Karakalpak (\texttt{kaa}), Nogay (\texttt{nog}), Karachay-Balkar (\texttt{krc}), Urum (\texttt{uum}), Crimean Tatar (\texttt{crh}) and Karaim (\texttt{kdr}).}
+	\caption{The three sub-branches of Kypchak (North, South, West), roughly divided with black lines, showing the geographic distribution of the three languages for which transducers were developed.
+The Kypchak languages shown on the map are Tatar (\texttt{tat}), Kazakh (\texttt{kaz}), and Kumyk (\texttt{kum}).
+The other codes represent Bashkir (\texttt{bak}), Kyrgyz (\texttt{kir}), Karakalpak (\texttt{kaa}), Nogay (\texttt{nog}), Karachay-Balkar (\texttt{krc}), Urum (\texttt{uum}), Crimean Tatar (\texttt{crh}), and Karaim (\texttt{kdr}).}
 	\label{map1}
@@
 		\midrule
-		Noun & 2463 & 2692 & 2588 \\
-		Verb & 1587 & 1333 & 262 \\
-		Adjective & 823 & 852 & 217 \\
-		Proper noun & 5412 & 3523 & 1444 \\
-		Adverb & 197 & 205 & 63 \\
-		Numeral & 63 & 67 & 45 \\
-		Conjunction & 52 & 50 & 15 \\
-		Postposition & 56 & 46 & 12 \\
-		Pronoun & 17 & 17 & 18 \\
-		Determiner & 42 & 37 & 9 \\
+		Noun & 2640 & 2795 & 2568 \\
+		Verb & 1470 & 1143 & 386 \\
+		Adjective & 754 & 816 & 219 \\
+		Proper noun & 5701 & 5361 & 1443 \\
+		Adverb & 171 & 177 & 63 \\
+		Numeral & 63 & 63 & 44 \\
+		Conjunction & 46 & 45 & 13 \\
+		Postposition & 50 & 43 & 12 \\
+		Pronoun & 32 & 28 & 17 \\
+		Determiner & 39 & 34 & 9 \\
 		\midrule
-		Total: & 10942 & 9009 & 4687 \\
+		Total: & 11224 & 10737 & 4845 \\
 		\bottomrule
@@
 We tested the coverage of the Kazakh and Tatar analysers over three separate domains: encyclopaedic text,\footnote{The following Wikipedia dumps were used: \texttt{kkwiki-20131006-pages-articles.xml.bz2} and \texttt{ttwiki-20130225-pages-articles.xml.bz2}.}
- news,\footnote{All content from \url{http://www.azattyq.org/} for 2010 was used for Kazakh, as well as all content from 2005 to 2011 on \url{http://tat.tatar-inform.ru} for Tatar.} and religion.\footnote{We used a Kazakh bible translation available from \url{https://kkitap.net/} and Quran translation available from \url{http://kuran.kz/}, as well as a Tatar translation of the New Testament available from \url{http://ibt.org.ru} and Quran translation available from \url{http://tanzil.net}.}  As there is currently no Wikipedia in Kumyk, we tested only news and religion.\footnote{The bible corpus consists of Genesis and the New Testament, as available from \url{http://ibt.org.ru/}, and the news corpus consists of all Kumyk content from \url{http://sh-tavisi.etnosmi.ru/}.}
+ news,\footnote{All content from \url{http://www.azattyq.org/} for 2010 was used for Kazakh, as well as all content from 2005 to 2011 on \url{http://tat.tatar-inform.ru/} for Tatar.} and religion.\footnote{We used a Kazakh bible translation available from \url{https://kkitap.net/} and Quran translation available from \url{http://kuran.kz/}, as well as a Tatar translation of the New Testament available from \url{http://ibt.org.ru/} and Quran translation available from \url{http://tanzil.net/}.}  As there is currently no Wikipedia in Kumyk, we tested only news and religion.\footnote{The bible corpus consists of Genesis and the New Testament, as available from \url{http://ibt.org.ru/}, and the news corpus consists of all Kumyk content from \url{http://sh-tavisi.etnosmi.ru/}.}
 

### 2014-03-22 kipchak-lrec-2013/kipchak-paper.tex
§4.2 and 4.3 proofed
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@51059 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \newfontfamily\qipa[IPA,Scale=MatchLowercase]{FreeSerif}
-\newfontfamily\qgmk[Scale=MatchLowercase,Letters=SmallCaps]{Linux Libertine O}
-
-\newcommand{\gmk}[1]{{\qgmk #1}}
+%\newfontfamily\qgmk[Scale=MatchLowercase,Letters=SmallCaps]{Linux Libertine O}
+%
+%\newcommand{\gmk}[1]{{\qgmk #1}}
 \newcommand{\tags}[1]{\texttt{#1}}
@@
 жақсы ойнадыңыз `(you) played well', and substantively (as a noun) `жақсы келетін еді' `(the) good (ones) would come'. This
-is a productive process and relevant to the morphotactics as an adjective used substantively may take the 
+is a productive process and relevant to the morphotactics, as an adjective used substantively may take the 
 full range of case and possessive suffixes, and adjectives used adverbially will take a different set of 
-clitics to adjectives used substantively and attributively.
+clitics than adjectives used substantively and attributively.
 
-However, this is not completely productive. A grammar which allows all adjectives to function
-as nouns and adverbs will overgenerate (and overanalyse). However, most grammars do not 
+However, this is not completely productive, as not all adjectives can be used substantively or adverbially. A grammar which allows all adjectives to function
+as nouns and adverbs will overgenerate (and overanalyse). However, most grammatical descriptions of these languages do not 
 mention that there are exceptions to this process.
@@
 has a `default' function: nouns are by default substantive, adjectives are by default attributive, 
-and adverbs, adverbial. When they are used outside this default use they receive a tag to mark 
-their function. Where these functions are ambiguous, then they may be disambiguated 
+and adverbs adverbial. When they are used outside this default use they receive a tag to mark 
+their function. Where these functions are ambiguous, they may be disambiguated 
 in context.
@@
 The advantage of this approach as opposed to simply allowing one part of speech to derive into
-another is that it allows us to be more principled, and decreases overgeneration,
+another is that it allows us to be more principled, and decreases overgeneration;
 for example, nouns used attributively will not take the whole adjective continuation (e.g.
@@
 for `today'\footnote{Tatar: бүген; Kazakh: бүгін; Kumyk: бугюн. The word is etymologically composed of the words for `this' and `day'.} 
-cannot be used attributively without the presence of an extra morpheme -KI. Compound nouns
-using possessive morphology, such as Kazakh ауа райы `weather' are not well-formed
-without a possessive suffix. %FIXME: Explain the compound 'weather'.
+cannot be used attributively in Kypchak languages without the presence of an extra morpheme -KI. Compound nouns
+using possessive morphology, such as Kazakh ауа райы `weather', do not take additional morphology the same way other substantives in the language do, due to the presence of a final possessive morpheme (e.g., ‹ауа райынан› `weather-\tags{abl}', not *‹ауа райыдан›).
+%FIXME: Explain the compound 'weather'.
 
@@
 
-In Kazakh, ‹е› is a front diphthong (e.g., ‹елдің› `country-\gmk{gen}'), but in e.g., family names with Russian morphology, it harmonises as a back vowel (e.g., ‹Назарбаевтың› `Nazarbayev-\gmk{gen}').
+In Kazakh, ‹е› is a front diphthong (e.g., ‹елдің› `country-\tags{gen}'), but in e.g., family names with Russian morphology, it harmonises as a back vowel (e.g., ‹Назарбаевтың› `Nazarbayev-\tags{gen}').
 
@@
  \tags{dat} & Dative  \\
+ \tags{abl} & Ablative \\
  \tags{p1} & First person  \\

### 2014-03-22 kipchak-lrec-2013/kipchak-paper.tex
some cleaning up
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@51060 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 %FIXME: keep revision number up to date
-The transducers are available / under development in apertium's subversion repository,\footnote{\url{https://svn.code.sf.net/p/apertium/svn/languages/}} in the directories apertium-kaz, apertium-tat, and apertium-kum.
-The revision of the entire subversion repository that the numbers (stem counts, evaluation, etc.) in this paper represent is \texttt{r48137}.
+The transducers are available / under development in apertium's subversion repository,\footnote{{\scriptsize \url{https://svn.code.sf.net/p/apertium/svn/languages/}}} in the directories apertium-kaz, apertium-tat, and apertium-kum, and compiled versions may be run at \url{http://turkic.apertium.org/}.
+The revision of each module in the subversion repository that the numbers in this paper (stem counts, evaluation, etc.) represent is \texttt{r50547} for Kazakh, \texttt{r50260} for Tatar, and \texttt{r50300} for Kumyk.
 
@@
 
-\caption{Number of stems and coverage of corpus over time for the Kumyk transducer. Time is measured in days 
-starting from the 2nd October 2013. The graph shows that to reach 80\% coverage, took
+\caption{Number of stems and coverage of a corpus over time for the Kumyk transducer. Time is measured in days 
+starting from the 2nd October 2013. The graph shows that to reach 80\% coverage took
 around a week, and to reach 90\% took another week.}
@@
 %FIXME: revisions
-We have evaluated the morphological analysers\footnote{The versions of the analysers used were: r50547 for Kazakh, r50260 for Tatar, and r50300 for Kumyk.} in two ways.
+We have evaluated the morphological analysers in two ways.
 The first was by calculating the naïve coverage and mean ambiguity on freely available corpora.
@@
 
-We would like to thank the Google Code-in (2011) for supporting the development 
+We would like to thank the Google Code-in (2011) for supporting the original development 
 of the Kazakh transducer, and in particular the effort by Nathan Maxson. We 
@@
 
-The authors would also like to express their gratitude to Aida Sundetova and Ağarhim Sultanmuradov
+The authors would also like to express their gratitude to Aida Sundetova 
+%and Ağarhim Sultanmuradov
 for assistance in evaluating precision and recall.

### 2014-03-22 kipchak-lrec-2013/kipchak-paper.tex
яхшы example ;)
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@51062 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 Most of the stems in the lexicon can be used, with no extra suffixes in any of these functions, for 
-example the adjective жақсы in Kazakh can be used as attributively, жақсы кітап `good book', adverbially
-жақсы ойнадыңыз `(you) played well', and substantively (as a noun) `жақсы келетін еді' `(the) good (ones) would come'. This
+example the adjective ‹яхшы› in Tatar can be used as attributively (e.g., ‹яхшы аучы› `good hunter'), adverbially
+(e.g., ‹яхшы бөгелә› `bends well'), and substantively, i.e. as a noun (e.g., ‹яхшы белән› `with a good [person]’). This
 is a productive process and relevant to the morphotactics, as an adjective used substantively may take the 
@@
 
-Types of these phenomena include characters used to represent multiple phonological forms, loanwords, and acronyms and numerals.
+Types of these phenomena include single characters used to represent multiple phonological forms, loanwords, and acronyms and numerals.
 A brief overview of each follows, along with examples and details of our solution.

### 2014-03-22 kipchak-lrec-2013/kipchak-paper.tex
update
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@51069 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 Other analysers for Turkic languages---such as \texttt{TRmorph} \citep{coltekin2010}---approach this 
-problem by positing a zero-derivation rule by which any noun can be `derived' into an adjective or 
+problem by allowing a zero-derivation by which any noun can be `derived' into an adjective or 
 an adverb, any adjective can be `derived' into a noun or adverb, etc.
@@
 cannot be used attributively in Kypchak languages without the presence of an extra morpheme -KI. Compound nouns
-using possessive morphology, such as Kazakh ауа райы `weather', do not take additional morphology the same way other substantives in the language do, due to the presence of a final possessive morpheme (e.g., ‹ауа райынан› `weather-\tags{abl}', not *‹ауа райыдан›).
+using possessive morphology, such as Kazakh ауа райы `weather' (lit.
+`air condition'), do not take additional morphology the same way other substantives in the language do, due to the presence of a final possessive morpheme (e.g., ‹ауа райынан› `weather-{\sc abl}', not *‹ауа райыдан›).
 %FIXME: Explain the compound 'weather'.

### 2014-03-22 kipchak-lrec-2013/kipchak-paper.tex
improve placement of graph
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@51077 72bbbca6-d526-0410-a7d9-f06f51895060
add numbers of kumyk
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@51095 72bbbca6-d526-0410-a7d9-f06f51895060
smallcaps
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@51096 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 %\centering
-\hspace{-15pt}
+\hspace{-35pt}
     \rotatebox{-90}{
-\includegraphics[width=0.4\textwidth]{graphs/kumyk-yoldash-coverage}
+\includegraphics[width=0.38\textwidth]{graphs/kumyk-yoldash-coverage}
 }
+~\\
+~\\
 
@@
 
-In Kazakh, ‹е› is a front diphthong (e.g., ‹елдің› `country-\tags{gen}'), but in e.g., family names with Russian morphology, it harmonises as a back vowel (e.g., ‹Назарбаевтың› `Nazarbayev-\tags{gen}').
+In Kazakh, ‹е› is a front diphthong (e.g., ‹елдің› `country-\tags{gen}'), but in e.g., family names with Russian morphology, it harmonises as a back vowel (e.g., ‹Назарбаевтың› `Nazarbayev-{\sc gen}').
 
@@
 		Tatar & 95.03 & 85.65 \\
-		Kumyk & - & - \\
+		Kumyk & 96.57 & 69.11 \\
 	\bottomrule

### 2014-03-22 kipchak-lrec-2013/kipchak-paper.tex
gmk
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@51097 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \newfontfamily\qipa[IPA,Scale=MatchLowercase]{FreeSerif}
-%\newfontfamily\qgmk[Scale=MatchLowercase,Letters=SmallCaps]{Linux Libertine O}
+\newfontfamily\qgmk[Scale=MatchLowercase,Letters=SmallCaps]{Linux Libertine O}
 %
-%\newcommand{\gmk}[1]{{\qgmk #1}}
+\newcommand{\gmk}[1]{{\qgmk #1}}
 \newcommand{\tags}[1]{\texttt{#1}}
@@
 using possessive morphology, such as Kazakh ауа райы `weather' (lit.
-`air condition'), do not take additional morphology the same way other substantives in the language do, due to the presence of a final possessive morpheme (e.g., ‹ауа райынан› `weather-{\sc abl}', not *‹ауа райыдан›).
+`air condition'), do not take additional morphology the same way other substantives in the language do, due to the presence of a final possessive morpheme (e.g., ‹ауа райынан› `weather-\gmk{abl}', not *‹ауа райыдан›).
 %FIXME: Explain the compound 'weather'.
@@
 
-In Kazakh, ‹е› is a front diphthong (e.g., ‹елдің› `country-\tags{gen}'), but in e.g., family names with Russian morphology, it harmonises as a back vowel (e.g., ‹Назарбаевтың› `Nazarbayev-{\sc gen}').
+In Kazakh, ‹е› is a front diphthong (e.g., ‹елдің› `country-\tags{gen}'), but in e.g., family names with Russian morphology, it harmonises as a back vowel (e.g., ‹Назарбаевтың› `Nazarbayev-\gmk{gen}').
 

### 2014-03-22 kipchak-lrec-2013/kipchak-paper.tex
minor stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@51098 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \subsubsection{Acronyms and numerals}
-%FIXME!!!
 Acronyms and numerals are challenging as they will often be pronounced out loud
-   in their non-abbreviated forms, for example in Kazakh, 30-дан `from thirty', 5-тен `from 5'. The ablative
+   in their non-abbreviated forms, for example in Kazakh, ‹30-дан› `from thirty' would be spelled out ‹отыздан›, and 5-тен `from 5' would be spelled out ‹бестен›. The ablative
    suffix -DAn alters for the phonology, but in the numeral string there is no indication of how they should alter.
@@
 The symbol \{э\} signals that the following vowel needs to harmonise to a front unrounded vowel, and \{с\} signals that there is a final voiceless consonant.
-So, with 5\{э\}\{с\} in the lexicon, and a consistent continuation lexicon specifying the underlying form of case affixes, the rules that operate on \{D\} and \{A\} are able to do produce the correct output form in each language, avoiding the incorrect default *5-дан.
+So, with 5\{э\}\{с\}\footnote{Note that there is not one extra symbol for every sound of the language, but a much reduced set.
+E.g., \{с\} is used to represent all voiceless consonants.} in the lexicon, and a consistent continuation lexicon specifying the underlying form of case affixes, the rules that operate on \{D\} and \{A\} are able to do produce the correct output form in each language, avoiding the incorrect default *5-дан.
 
@@
 This solution has the same issue as loanwords above.
-%FIXME: "performs all PHONOLOGY operations in parallel\footnote{the formalism is like SPE, but the rules are applied in parallel, somewhat like constraints in OT}" ← clarify
-%Since \texttt{twol} performs all operations in parallel, rules triggered by these must make reference to abstract letters that exist on the input tape but are null on the output tape.
-This complicates existing rules that have been designed to ignore null characters on the output tape.
-The \texttt{twol} rules resulting from the combination of processes simultaneously triggered by and ignoring output-null characters are rather ...
-uh, \textit{intricate} ...
-but they seem to work okay?
-Just don't poke them too hard.
-%FIXME
 
@@
 
-%FIXME: revisions
 We have evaluated the morphological analysers in two ways.
@@
 
-%FIXME: sources
 We tested the coverage of the Kazakh and Tatar analysers over three separate domains: encyclopaedic text,\footnote{The following Wikipedia dumps were used: \texttt{kkwiki-20131006-pages-articles.xml.bz2} and \texttt{ttwiki-20130225-pages-articles.xml.bz2}.}

### 2014-03-22 kipchak-lrec-2013/kipchak-paper.tex
something about kum
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@51100 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 Tatar transducer can be partly explained by the less transparent orthography of Tatar.
+Due to limited access to native Kumyk speakers, the authors performed precision themselves based on their admittedly incomplete knowledge of the language.
+Recall is low for Kumyk due mostly to some systematic gaps in the morphotactics, such as lack of attributive analyses in several continuation lexica.
 

### 2014-03-22 kipchak-lrec-2013/kipchak-paper.tex
try this for graphics position
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@51102 72bbbca6-d526-0410-a7d9-f06f51895060
update
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@51103 72bbbca6-d526-0410-a7d9-f06f51895060

--- 2014-03-22 kipchak-lrec-2013/kipchak-paper.tex
@@
 \begin{figure}
-%\centering
-\hspace{-35pt}
+\centering
+%\hspace{-35pt}
     \rotatebox{-90}{
-\includegraphics[width=0.38\textwidth]{graphs/kumyk-yoldash-coverage}
+\includegraphics[width=0.38\textwidth,bb=0 50 554 770]{graphs/kumyk-yoldash-coverage}
 }
-~\\
-~\\
+%~\\
+%~\\
 
@@
 Where an analysis was erroneous, it was removed; where an analysis was missing, it was added.
-This process gave us a `gold standard' morphologically analysed word list of 1,000 surface forms with their analyses.
+This process gave us a `gold standard' morphologically analysed word list of 1,000 (Kazakh and Tatar) and 500 (Kumyk) surface forms with their analyses.
 The list is publically available for each language.%\footnote{\url{https://apertium.svn.sourceforge.net/svnroot/apertium/branches/apertium-kir/eval/ref.1000.txt}}
+
+The lower number of forms for Kumyk was due to the limited access to native speakers of Kumyk. 
+The process for the Kumyk evaluation was also slightly different.
+They were first checked by a non-expert native speaker of Kumyk and then the result was proofread by the authors with the help of a dictionary.
 
@@
 Tatar transducer can be partly explained by the less transparent orthography of Tatar.
+
+%Due to limited access to native Kumyk speakers, the authors performed precision themselves based on their admittedly incomplete knowledge of the language.  
+Recall is low for Kumyk due mostly to some systematic gaps in the morphotactics, such as lack of attributive analyses in several continuation lexica.
+There are also a higher number of missing verb stems (over half of all missing analyses were verb forms) in the Kumyk results than in the Kazakh and Tatar, where missing verb stems accounted for 10\% and 8\% respectively.
 
@@
 The authors would also like to express their gratitude to Aida Sundetova 
-%and Ağarhim Sultanmuradov
+and Ağarhim Sultanmuradov
 for assistance in evaluating precision and recall.

### 2014-03-22 kipchak-lrec-2013/kipchak-paper.tex
playing with space and stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@51104 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \subsection{Kazakh}
-Kazakh /q{\symbl ɑ}z{\symbl ɑ}q/ is spoken primarily in Kazakhstan, and it is the national language, where it is co-official with Russian.
+Kazakh /q{\symbl ɑ}z{\symbl ɑ}q/ is spoken primarily in Kazakhstan, where it is the national language and is co-official with Russian.
 Large communities of native speakers also exist in China, neighbouring Central-Eurasian republics, and Mongolia.
@@
 It also supports other finite state transducer formalisms such as \texttt{sfst}.
-This toolkit has been chosen as it -- or the equivalent XFST \cite{beesley2003} -- has been widely used for other Turkic languages, such as Turkish \citep{coltekin2010}, Crimean~Tatar~\citep{altintas2001}, Turkmen \citep{tantug2006}, and Kyrgyz \citep{washington2012}, and is available under a free/open-source licence.
+This toolkit has been chosen as it -- or the equivalent XFST \citep{beesley2003} -- has been widely used for other Turkic languages, such as Turkish \citep{coltekin2010}, Crimean~Tatar~\citep{altintas2001}, Turkmen \citep{tantug2006}, and Kyrgyz \citep{washington2012}, and is available under a free/open-source licence.
 
@@
 \begin{figure}
-\centering
+%\centering
 %\hspace{-35pt}
-    \rotatebox{-90}{
-\includegraphics[width=0.38\textwidth,bb=0 50 554 770]{graphs/kumyk-yoldash-coverage}
-}
+%\hspace{-1em}
+%    \rotatebox{-90}{
+%\hspace{-15pt}
+\hspace{-10pt}
+\includegraphics[width=0.37\textwidth,bb=50 50 554 770,angle=270]{graphs/kumyk-yoldash-coverage}
+%}
 %~\\
@@
 \end{center}
-
+\vspace{-2em}
 \end{table}
@@
 \end{center}
+\vspace{-1.5em}
 \end{table}

### 2014-03-22 kipchak-lrec-2013/kipchak-paper.tex
slight wording change
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@51105 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 This branch can be divided into three sub-branches.
-Kumyk is a member of the Western Kypchak group, Tatar is a member of the Northern Kypchak group, and Kazakh is a member of the Southern Kypchak group \citep[82-83]{histofturkic}.\footnote{It is the professional opinion of the authors of this paper that Kyrgyz constitutes a fourth branch.}   The geographic distribution of the languages is shown in map~\ref{map1}.
+Kumyk is a member of the Western Kypchak group, Tatar is a member of the Northern Kypchak group, and Kazakh is a member of the Southern Kypchak group \citep[82-83]{histofturkic}.\footnote{It is the professional opinion of the authors of this paper that Kyrgyz constitutes a fourth branch of Kypchak.}   The geographic distribution of the languages is shown in map~\ref{map1}.
 %As such, each of these three languages represents one of the three branches of Kypchak.

### 2014-03-22 kipchak-lrec-2013/kipchak-paper.tex
a little bit more fooling with space
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@51106 72bbbca6-d526-0410-a7d9-f06f51895060
some more minor stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@51107 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \label{fig:kumcov}
+\vspace{-0.5em}
 \end{figure}
@@
 
-In Kazakh, ‹е› is a front diphthong (e.g., ‹елдің› `country-\tags{gen}'), but in e.g., family names with Russian morphology, it harmonises as a back vowel (e.g., ‹Назарбаевтың› `Nazarbayev-\gmk{gen}').
+In Kazakh, ‹е› is a front diphthong (e.g., ‹елдің› `country-\gmk{gen}'), but in e.g., family names with Russian morphology, it harmonises as a back vowel (e.g., ‹Назарбаевтың› `Nazarbayev-\gmk{gen}').
 
@@
 Where an analysis was erroneous, it was removed; where an analysis was missing, it was added.
-This process gave us a `gold standard' morphologically analysed word list of 1,000 (Kazakh and Tatar) and 500 (Kumyk) surface forms with their analyses.
-The list is publically available for each language.%\footnote{\url{https://apertium.svn.sourceforge.net/svnroot/apertium/branches/apertium-kir/eval/ref.1000.txt}}
+This process gave us a `gold standard' morphologically analysed word list of 1,000 forms for Kazakh and Tatar and 500 forms for Kumyk, along with their analyses.
+The list is publically available for each language in Apertium's svn repository.%\footnote{\url{https://apertium.svn.sourceforge.net/svnroot/apertium/branches/apertium-kir/eval/ref.1000.txt}}
 

### 2014-03-22 kipchak-lrec-2013/kipchak-paper.tex
apertium-kaz/tat/kum in \texttt{}
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@51108 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 %FIXME: keep revision number up to date
-The transducers are available / under development in apertium's subversion repository,\footnote{{\scriptsize \url{https://svn.code.sf.net/p/apertium/svn/languages/}}} in the directories apertium-kaz, apertium-tat, and apertium-kum, and compiled versions may be run at \url{http://turkic.apertium.org/}.
+The transducers are available / under development in apertium's subversion repository,\footnote{{\scriptsize \url{https://svn.code.sf.net/p/apertium/svn/languages/}}} in the directories \texttt{apertium-kaz}, \texttt{apertium-tat}, and \texttt{apertium-kum}, and compiled versions may be run at \url{http://turkic.apertium.org/}.
 The revision of each module in the subversion repository that the numbers in this paper (stem counts, evaluation, etc.) represent is \texttt{r50547} for Kazakh, \texttt{r50260} for Tatar, and \texttt{r50300} for Kumyk.
@@
 %\centering
-%\hspace{-35pt}
+\hspace{-35pt}
 %\hspace{-1em}

### 2014-03-23 kipchak-lrec-2013/kipchak-paper.tex
another version
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@51110 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-%FIXME: keep revision number up to date
 The transducers are available / under development in apertium's subversion repository,\footnote{{\scriptsize \url{https://svn.code.sf.net/p/apertium/svn/languages/}}} in the directories \texttt{apertium-kaz}, \texttt{apertium-tat}, and \texttt{apertium-kum}, and compiled versions may be run at \url{http://turkic.apertium.org/}.
@@
 
-The Kazakh morphotactics were originally developed based on the Kyrgyz transducer, which was co-authored by the first author, who is fluent in and has a good linguistic knowledge of Kyrgyz and two native speakers of Kyrgyz.
+The Kazakh morphotactics were originally developed based on the Kyrgyz transducer, which was co-authored by the first author---who is fluent in and has a good linguistic knowledge of both Kyrgyz and Kazakh---and two native speakers of Kyrgyz.
+The morphotactics of Tatar were developed for the most part by the second author, a native speaker of Tatar, who also worked to polish off the morphotactics of the Kazakh transducer
+%The Kazakh morphotactics were originally developed based on the Kyrgyz transducer, which was co-authored by the first author, who is fluent in and has a good linguistic knowledge of Kyrgyz, along with two native speakers of Kyrgyz.
 The first author also developed the Kazakh morphotactics who is fluent in and has a good linguistic understanding Kazakh.
@@
 The same process was undertaken for verbs.
+Phonological rules were developed as needed.
 
@@
 %\centering
-\hspace{-35pt}
+%\hspace{-35pt}
 %\hspace{-1em}
@@
 \begin{tabular}{lll}
+\toprule
  \textbf{Kazakh} & \textbf{Tatar} & \textbf{Kumyk} \\
-\hline
+\midrule
  Құдай Өзінің жаратқандарының  &  Аллаһ Үзе яраткан нәрсәләргә карап, & Аллагь Оьзю яратгъан затлагъа\\
  бәріне қарап, өте жақсы екенін көрді. &  аларның бик яхшы икәнен күрде. & къарап, олар бек яхшы экенин гёрген. \\
-\hline
+\midrule
  Құдай\tags{<n><nom>} & Аллаһ\tags{<n><nom>} & Аллагь\tags{<n><nom>}  \\
@@
  .\tags{<sent>} & .\tags{<sent>} & .\tags{<sent>} \\
-\hline
+\bottomrule
 \end{tabular}
@@
   that would require further disambiguation. Refer to Appendix~\ref{app:glossary} for tag descriptions.}
-
+\vspace{-1.5em}
 \end{table*}
@@
 Types of these phenomena include single characters used to represent multiple phonological forms, loanwords, and acronyms and numerals.
-A brief overview of each follows, along with examples and details of our solution.
+A brief overview of each follows, along with examples and details of our solutions.
 
@@
 \end{center}
-\vspace{-1.5em}
+\vspace{-2.5em}
 \end{table}

### 2014-03-23 kipchak-lrec-2013/kipchak-paper.tex
another version
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@51111 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-\abstract{This paper describes the development of free/open-source finite-state morphological transducers for three Turkic languages---Kazakh, Tatar, and Kumyk---representing one language from each of the three sub-branches of the Kypchak branch of Turkic.
+\abstract{This paper describes the development of free/open-source finite-state morphological transducers for three Turkic languages---Kazakh, Tatar, and Kumyk---representing one language from each of the three commonly distinguished sub-branches of the Kypchak branch of Turkic.
 The finite-state toolkit used for the work is the Helsinki Finite-State Toolkit (HFST).
@@
 
-These transducers were all developed as part of the Apertium project, which is aimed at creating rule-based MT systems for lesser resourced languages.
+These transducers were all developed as part of the Apertium project, which is aimed at creating rule-based machine translation (RBMT) systems for lesser resourced languages.
 As such, the transducers were developed with the intent that they be used as morphological analysers and generators in RBMT systems.

### 2014-03-23 kipchak-lrec-2013/kipchak-paper.tex
minor wording and spacing fixes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@51112 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-This paper overviews the languages (§\ref{sec:lgs}), provides background on these morphological transducers and related previous work (§\ref{sec:bg}), describes the development effort and contents of the transducers (§\ref{sec:methodology}), evaluates the effectiveness of these transducers (§\ref{sec:evaluation}), and summarises the findings (§\ref{sec:conclusions}).
+This paper gives an overview of the languages (§\ref{sec:lgs}), provides background on these morphological transducers and related previous work (§\ref{sec:bg}), describes the development effort and contents of the transducers (§\ref{sec:methodology}), evaluates the effectiveness of these transducers (§\ref{sec:evaluation}), and summarises the findings (§\ref{sec:conclusions}).
 
@@
 
-Our solution for these consistent exceptions is to make a separate continuation lexicon for e.g., Russian nouns.
-This continuation lexicon adds a character that the phonology deletes, but that triggers phonological rules.
+Our solution for these consistent exceptions is to make a separate continuation lexicon for e.g., Russian nouns, which adds a character which the phonology deletes, but which triggers various phonological rules.
 

### 2014-04-07 kipchak-lrec-2013/kipchak-paper.tex
fixed some typos and stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@51769 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 	\caption{The three sub-branches of Kypchak (North, South, West), roughly divided with black lines, showing the geographic distribution of the three languages for which transducers were developed.
-The Kypchak languages shown on the map are Tatar (\texttt{tat}), Kazakh (\texttt{kaz}), and Kumyk (\texttt{kum}).
-The other codes represent Bashkir (\texttt{bak}), Kyrgyz (\texttt{kir}), Karakalpak (\texttt{kaa}), Nogay (\texttt{nog}), Karachay-Balkar (\texttt{krc}), Urum (\texttt{uum}), Crimean Tatar (\texttt{crh}), and Karaim (\texttt{kdr}).}
+The Kypchak languages this paper describes are Tatar (\texttt{tat}), Kazakh (\texttt{kaz}), and Kumyk (\texttt{kum}).
+The other codes on the map represent Bashkir (\texttt{bak}), Kyrgyz (\texttt{kir}), Karakalpak (\texttt{kaa}), Nogay (\texttt{nog}), Karachay-Balkar (\texttt{krc}), Urum (\texttt{uum}), Crimean Tatar (\texttt{crh}), and Karaim (\texttt{kdr}).}
 	\label{map1}
@@
 \subsection{Tatar}
-Tatar /t{\symbl ɒ}t{\symbl ɑ}r/ is spoken in and around Tatarstan, --- a republic of the Russian Federation, where it is co-official with Russian.
+Tatar /t{\symbl ɒ}t{\symbl ɑ}r/ is spoken in and around Tatarstan, a republic of the Russian Federation, where it is co-official with Russian.
 The majority of native speakers are bilingual in Russian.
@@
 The Kazakh morphotactics were originally developed based on the Kyrgyz transducer, which was co-authored by the first author---who is fluent in and has a good linguistic knowledge of both Kyrgyz and Kazakh---and two native speakers of Kyrgyz.
-The morphotactics of Tatar were developed for the most part by the second author, a native speaker of Tatar, who also worked to polish off the morphotactics of the Kazakh transducer
+The morphotactics of Tatar were developed for the most part by the second author, a native speaker of Tatar, who also worked to polish off the morphotactics of the Kazakh transducer.
 %The Kazakh morphotactics were originally developed based on the Kyrgyz transducer, which was co-authored by the first author, who is fluent in and has a good linguistic knowledge of Kyrgyz, along with two native speakers of Kyrgyz.
@@
 In order to create the Kumyk transducer, we approached one of the major parts of speech at a time. We started with the nouns, copying the 
-continuation lexica\footnote{A continuation lexicon is a set of morphemes, for example, in the Turkic languages there is a continuation lexicon for `case' which includes the possible case suffixes.} (nominal morphotactics) from the Kazakh transducer.
+continuation lexica\footnote{A continuation lexicon is a set of morphemes; for example, in the Turkic languages there is a continuation lexicon for `case' which includes the possible case suffixes.} (nominal morphotactics) from the Kazakh transducer.
 The suffixes were then replaced with the Kumyk suffixes according to the grammar \citep{olmesov2000}.
@@
 %\centering
-%\hspace{-35pt}
+%\hspace{-30pt}
 %\hspace{-1em}
 %    \rotatebox{-90}{
-%\hspace{-15pt}
-\hspace{-10pt}
+\hspace{-30pt}
+%\hspace{-10pt}
 \includegraphics[width=0.37\textwidth,bb=50 50 554 770,angle=270]{graphs/kumyk-yoldash-coverage}
@@
 
+\vspace{1.2em}
 \caption{Number of stems and coverage of a corpus over time for the Kumyk transducer. Time is measured in days 
@@
 \label{fig:kumcov}
-\vspace{-0.5em}
+\vspace{-1.8em}
 \end{figure}
@@
     choosing between loan word and native word. Adding stems from the other main categories 
-    requires more in depth categorisation.}
+    requires more in-depth categorisation.}
  \label{table:stems}
@@
   their very good being saw''. The output has been 
-  abbreviated to only show the appropriate tag sequence in context, the actual output would give multiple interpretations
+  abbreviated to only show the appropriate tag sequence in context; the actual output would give multiple interpretations
   that would require further disambiguation. Refer to Appendix~\ref{app:glossary} for tag descriptions.}
@@
 Our approach is to describe this rather in terms of function (not unlike \cite{hengeveld92}). We 
-posit that nominals can be used either substantively, attributively or adverbially. Each part of speech 
-has a `default' function: nouns are by default substantive, adjectives are by default attributive, 
+posit that e.g., nominals can be used either substantively, attributively or adverbially. Each part of speech 
+hence has a `default' function: nouns are by default substantive, adjectives are by default attributive, 
 and adverbs adverbial. When they are used outside this default use they receive a tag to mark 
@@
 The character ‹ю› is used in turn to represent ‹й› followed by either multi-phoneme value of ‹у›.
-For example, ‹қиюда› /qəjəwd{\qipa ɑ}/ `in the process of chopping down' and ‹киюде› /k{\qipa ɘ}j{\qipa ɘ}wd{\qipa i̯ɘ}/ `in the process of getting dressed' have back- and front-harmonising values, respectively, for ‹и› and ‹ю›.
+For example, ‹қиюда› /qəjəwd{\qipa ɑ}/ `in the process of chopping down' and ‹киюде› /k{\qipa ɘ}j{\qipa ɘ}wd{\qipa i̯ɘ}/ `in the process of getting dressed' have back- and front-harmonising values, respectively, for ‹и› and ‹ю›.\footnote{Note also that the orthography suggests the presence of an additional /j/ in these words.}
 
@@
 Our solution is to add phonological information at the end of morphemes that need it in the form of special ``abstract letters'' that trigger phonological processes at the morphophonological stage and are deleted.
-For example, the string \texttt{5<num><subst><abl>}\footnote{The resulting forms, 5-тен\super{\texttt{kaz}}, 5-тән\super{\texttt{tat}}, and 5-ден\super{\texttt{kum}} would be pronounced бестен\super{\texttt{kaz}}, биштән\super{\texttt{tat}}, and бешден\super{\texttt{kum}}.} has
+For example, the string \texttt{5<num><subst><abl>}\footnote{The resulting forms, 5-тен\super{\texttt{kaz}}, 5-тән\super{\texttt{tat}}, and 5-ден\super{\texttt{kum}} would be spelled out ‹бестен›\super{\texttt{kaz}}, ‹биштән›\super{\texttt{tat}}, and ‹бешден›\super{\texttt{kum}}.} has
 the morphotactic representation 5\{э\}\{с\}>-\{D\}\{A\}н, where \{э\} and \{с\} stand for phonological triggers, > represents a morpheme boundary, and \{D\}\{A\}н is the representation of the ablative morpheme at the morphotactic level.
@@
 The authors would also like to express their gratitude to Aida Sundetova 
-and Ağarhim Sultanmuradov
+and Ağarahim Sultanmuradov
 for assistance in evaluating precision and recall.

### 2014-05-08 codeswitch-cltw-2014/codeswitch.tex
new dir
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52619 72bbbca6-d526-0410-a7d9-f06f51895060

%
% File coling2014.tex
%
% Contact: jwagner@computing.dcu.ie
%%
%% Based on the style files for ACL-2014, which were, in turn,
%% Based on the style files for ACL-2013, which were, in turn,
%% Based on the style files for ACL-2012, which were, in turn,
%% based on the style files for ACL-2011, which were, in turn, 
%% based on the style files for ACL-2010, which were, in turn, 
%% based on the style files for ACL-IJCNLP-2009, which were, in turn,
%% based on the style files for EACL-2009 and IJCNLP-2008...

%% Based on the style files for EACL 2006 by 
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

\documentclass[11pt]{article}
\usepackage{coling2014}
\usepackage{times}
\usepackage{url}
\usepackage{latexsym}

%\setlength\titlebox{5cm}

% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.

\title{Instructions for COLING-2014 Proceedings}

\author{First Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  {\tt email@domain} \\\And
  Second Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  {\tt email@domain} \\}

\date{}

\begin{document}
\maketitle
\begin{abstract}
  This document contains the instructions for preparing a camera-ready
  manuscript for the proceedings of COLING-2014. The document itself
  conforms to its own specifications, and is therefore an example of
  what your manuscript should look like. These instructions should be
  used for both papers submitted for review and for final versions of
  accepted papers.  Authors are asked to conform to all the directions
  reported in this document.
\end{abstract}

\section{Credits}

This document has been adapted from the instructions for the ACL-2014
proceedings compiled by Alexander Koller and Yusuke Miyao,
which are, in turn, based on the instructions for earlier ACL proceedings,
including those for ACL-2012 by Maggie Li and Michael
White, those from ACL-2010 by Jing-Shing Chang and Philipp Koehn,
those for ACL-2008 by Johanna D. Moore, Simone Teufel, James Allan,
and Sadaoki Furui, those for ACL-2005 by Hwee Tou Ng and Kemal
Oflazer, those for ACL-2002 by Eugene Charniak and Dekang Lin, and
earlier ACL and EACL formats. Those versions were written by several
people, including John Chen, Henry S. Thompson and Donald
Walker. Additional elements were taken from the formatting
instructions of the {\em International Joint Conference on Artificial
  Intelligence}.

\section{Introduction}
\label{intro}

%
% The following footnote without marker is needed for the camera-ready
% version of the paper.
% Comment out the instructions (first text) and uncomment the 8 lines
% under "final paper" for your variant of English.
% 
\blfootnote{
    %
    % for review submission
    %
    \hspace{-0.65cm}  % space normally used by the marker
    Place licence statement here for the camera-ready version, see
    Section~\ref{licence} of the instructions for preparing a
    manuscript.
    %
    % % final paper: en-uk version (to license, a licence)
    %
    % \hspace{-0.65cm}  % space normally used by the marker
    % This work is licensed under a Creative Commons 
    % Attribution 4.0 International Licence.
    % Page numbers and proceedings footer are added by
    % the organisers.
    % Licence details:
    % \url{http://creativecommons.org/licenses/by/4.0/}
    % 
    % % final paper: en-us version (to licence, a license)
    %
    % \hspace{-0.65cm}  % space normally used by the marker
    % This work is licenced under a Creative Commons 
    % Attribution 4.0 International License.
    % Page numbers and proceedings footer are added by
    % the organizers.
    % License details:
    % \url{http://creativecommons.org/licenses/by/4.0/}
}

The following instructions are directed to authors of papers submitted
to COLING-2014 or accepted for publication in its proceedings. All
authors are required to adhere to these specifications. Authors are
required to provide a Portable Document Format (PDF) version of their
papers. \textbf{The proceedings are designed for printing on A4
  paper.}

Authors from countries in which access to word-processing systems is
limited should contact the publication chairs,
Joachim Wagner, Liadh Kelly and Lorraine Goeuriot
(\texttt{firstname.lastname @computing.dcu.ie} respectively),
as soon as possible.

We will make additional instructions available at
\url{http://www.coling-2014.org/call-for-papers.php}. Please check
this website regularly.

\section{General Instructions}

Manuscripts must be in single-column format.
The title, authors' names and complete
addresses
must be centred at the top of the first page, and
any full-width figures or tables (see the guidelines in
Subsection~\ref{ssec:first}). {\bf Type single-spaced.}  Start all
pages directly under the top margin. See the guidelines later
regarding formatting the first page.  The manuscript should be
printed single-sided and its length
should not exceed the maximum page limit described in Section~\ref{sec:length}.
Do not number the pages.

\subsection{Electronically-available resources}

We strongly prefer that you prepare your PDF files using \LaTeX{} with
the official COLING 2014 style file (coling2014.sty) and bibliography style
(acl.bst). These files are available at
\url{http://www.coling-2014.org/call-for-papers.php}.
You will also find the document
you are currently reading (coling2014.pdf) and its \LaTeX{} source code
(coling2014.tex) on this website.

You can alternatively use Microsoft Word to produce your PDF file. In
this case, we strongly recommend the use of the Word template file
(coling2014.dot) on the COLING 2014 website. If you have an option, we
recommend that you use the \LaTeX2e{} version. If you will be
  using the Microsoft Word template, we suggest that you anonymise
  your source file so that the pdf produced does not retain your
  identity.  This can be done by removing any personal information
from your source document properties.

\subsection{Format of Electronic Manuscript}
\label{sect:pdf}

For the production of the electronic manuscript you must use Adobe's
Portable Document Format (PDF). PDF files are usually produced from
\LaTeX{} using the \textit{pdflatex} command. If your version of
\LaTeX{} produces Postscript files, you can convert these into PDF
using \textit{ps2pdf} or \textit{dvipdf}. On Windows, you can also use
Adobe Distiller to generate PDF.

Please make sure that your PDF file includes all the necessary fonts
(especially tree diagrams, symbols, and fonts with Asian
characters). When you print or create the PDF file, there is usually
an option in your printer setup to include none, all or just
non-standard fonts.  Please make sure that you select the option of
including ALL the fonts. \textbf{Before sending it, test your PDF by
  printing it from a computer different from the one where it was
  created.} Moreover, some word processors may generate very large PDF
files, where each page is rendered as an image. Such images may
reproduce poorly. In this case, try alternative ways to obtain the
PDF. One way on some systems is to install a driver for a postscript
printer, send your document to the printer specifying ``Output to a
file'', then convert the file to PDF.

It is of utmost importance to specify the \textbf{A4 format} (21 cm
x 29.7 cm) when formatting the paper. When working with
{\tt dvips}, for instance, one should specify {\tt -t a4}.

If you cannot meet the above requirements
for the
production of your electronic submission, please contact the
publication chairs as soon as possible.

\subsection{Layout}
\label{ssec:layout}

Format manuscripts with a single column to a page, in the manner these
instructions are formatted. The exact dimensions for a page on A4
paper are:

\begin{itemize}
\item Left and right margins: 2.5 cm
\item Top margin: 2.5 cm
\item Bottom margin: 2.5 cm
\item Width: 16.0 cm
\item Height: 24.7 cm
\end{itemize}

\noindent Papers should not be submitted on any other paper size.
If you cannot meet the above requirements for
the production of your electronic submission, please contact the
publication chairs above as soon as possible.

\subsection{Fonts}

For reasons of uniformity, Adobe's {\bf Times Roman} font should be
used. In \LaTeX2e{} this is accomplished by putting

\begin{quote}
\begin{verbatim}
\usepackage{times}
\usepackage{latexsym}
\end{verbatim}
\end{quote}
in the preamble. If Times Roman is unavailable, use {\bf Computer
  Modern Roman} (\LaTeX2e{}'s default).  Note that the latter is about
  10\% less dense than Adobe's Times Roman font.

The {\bf Times New Roman} font, which is configured for us in the
Microsoft Word template (coling2014.dot) and which some Linux
distributions offer for installation, can be used as well.

\begin{table}[h]
\begin{center}
\begin{tabular}{|l|rl|}
\hline \bf Type of Text & \bf Font Size & \bf Style \\ \hline
paper title & 15 pt & bold \\
author names & 12 pt & bold \\
author affiliation & 12 pt & \\
the word ``Abstract'' & 12 pt & bold \\
section titles & 12 pt & bold \\
document text & 11 pt  &\\
captions & 11 pt & \\
sub-captions & 9 pt & \\
abstract text & 10 pt & \\
bibliography & 10 pt & \\
footnotes & 9 pt & \\
\hline
\end{tabular}
\end{center}
\caption{\label{font-table} Font guide. }
\end{table}

\subsection{The First Page}
\label{ssec:first}

Centre the title, author's name(s) and affiliation(s) across
the page.
Do not use footnotes for affiliations. Do not include the
paper ID number assigned during the submission process.

{\bf Title}: Place the title centred at the top of the first page, in
a 15 pt bold font. (For a complete guide to font sizes and styles,
see Table~\ref{font-table}) Long titles should be typed on two lines
without a blank line intervening. Approximately, put the title at 2.5
cm from the top of the page, followed by a blank line, then the
author's names(s), and the affiliation on the following line. Do not
use only initials for given names (middle initials are allowed). Do
not format surnames in all capitals (e.g., use ``Schlangen'' not
``SCHLANGEN'').  Do not format title and section headings in all
capitals as well except for proper names (such as ``BLEU'') that are
conventionally in all capitals.  The affiliation should contain the
author's complete address, and if possible, an electronic mail
address. Start the body of the first page 7.5 cm from the top of the
page.

The title, author names and addresses should be completely identical
to those entered to the electronical paper submission website in order
to maintain the consistency of author information among all
publications of the conference. If they are different, the publication
chairs may resolve the difference without consulting with you; so it
is in your own interest to double-check that the information is
consistent.

{\bf Abstract}: Type the abstract between addresses and main body.
The width of the abstract text should be
smaller than main body by about 0.6 cm on each side.
Centre the word {\bf Abstract} in a 12 pt bold
font above the body of the abstract. The abstract should be a concise
summary of the general thesis and conclusions of the paper. It should
be no longer than 200 words. The abstract text should be in 10 pt font.

{\bf Text}: Begin typing the main body of the text immediately after
the abstract, observing the single-column format as shown in 
the present document. Do not include page numbers.

{\bf Indent} when starting a new paragraph. Use 11 pt for text and 
subsection headings, 12 pt for section headings and 15 pt for
the title.

{\bf Licence}: Include a licence statement as an unmarked (unnumbered)
footnote on the first page of the final, camera-ready paper.
See Section~\ref{licence} below for details and motivation.

\subsection{Sections}

{\bf Headings}: Type and label section and subsection headings in the
style shown on the present document.  Use numbered sections (Arabic
numerals) in order to facilitate cross references. Number subsections
with the section number and the subsection number separated by a dot,
in Arabic numerals. Do not number subsubsections.

{\bf Citations}: Citations within the text appear in parentheses
as~\cite{Gusfield:97} or, if the author's name appears in the text
itself, as Gusfield~\shortcite{Gusfield:97}.  Append lowercase letters
to the year in cases of ambiguity.  Treat double authors as
in~\cite{Aho:72}, but write as in~\cite{Chandra:81} when more than two
authors are involved. Collapse multiple citations as
in~\cite{Gusfield:97,Aho:72}. Also refrain from using full citations
as sentence constituents. We suggest that instead of
\begin{quote}
  ``\cite{Gusfield:97} showed that ...''
\end{quote}
you use
\begin{quote}
``Gusfield \shortcite{Gusfield:97}   showed that ...''
\end{quote}

If you are using the provided \LaTeX{} and Bib\TeX{} style files, you
can use the command \verb|\newcite| to get ``author (year)'' citations.

As reviewing will be double-blind, the submitted version of the papers
should not include the authors' names and affiliations. Furthermore,
self-references that reveal the author's identity, e.g.,
\begin{quote}
``We previously showed \cite{Gusfield:97} ...''  
\end{quote}
should be avoided. Instead, use citations such as 
\begin{quote}
``Gusfield \shortcite{Gusfield:97}
previously showed ... ''
\end{quote}

\textbf{Please do not use anonymous citations} and do not include
any of the following when submitting your paper for review:
acknowledgements, project names, grant numbers, and names or URLs of
resources or tools that have only been made publicly available in
the last 3 weeks or are about to be made public.
Papers that do not
conform to these requirements may be rejected without review.
These details can, however, be included in the camera-ready, final paper.

\textbf{References}: Gather the full set of references together under
the heading {\bf References}; place the section before any Appendices,
unless they contain references. Arrange the references alphabetically
by first author, rather than by order of occurrence in the text.
Provide as complete a citation as possible, using a consistent format,
such as the one for {\em Computational Linguistics\/} or the one in the 
{\em Publication Manual of the American 
Psychological Association\/}~\cite{APA:83}.  Use of full names for
authors rather than initials is preferred.  A list of abbreviations
for common computer science journals can be found in the ACM 
{\em Computing Reviews\/}~\cite{ACM:83}.

The \LaTeX{} and Bib\TeX{} style files provided roughly fit the
American Psychological Association format, allowing regular citations, 
short citations and multiple citations as described above.

{\bf Appendices}: Appendices, if any, directly follow the text and the
references (but see above).  Letter them in sequence and provide an
informative title: {\bf Appendix A. Title of Appendix}.

\subsection{Footnotes}

{\bf Footnotes}: Put footnotes at the bottom of the page and use 9 pt
text. They may be numbered or referred to by asterisks or other
symbols.\footnote{This is how a footnote should appear.} Footnotes
should be separated from the text by a line.\footnote{Note the line
separating the footnotes from the text.}

\subsection{Graphics}

{\bf Illustrations}: Place figures, tables, and photographs in the
paper near where they are first discussed, rather than at the end, if
possible. 
Colour
illustrations are discouraged, unless you have verified that  
they will be understandable when printed in black ink.

{\bf Captions}: Provide a caption for every illustration; number each one
sequentially in the form:  ``Figure 1. Caption of the Figure.'' ``Table 1.
Caption of the Table.''  Type the captions of the figures and 
tables below the body, using 11 pt text.

Narrow graphics together with the single-column format may lead to
large empty spaces,
see for example the wide margins on both sides of Table~\ref{font-table}.
If you have multiple graphics with related content, it may be
preferable to combine them in one graphic.
You can identify the sub-graphics with sub-captions below the
sub-graphics numbered (a), (b), (c) etc.\ and using 9 pt text.
The \LaTeX{} packages wrapfig, subfig, subtable and/or subcaption
may be useful.

\subsection{Licence Statement}
\label{licence}

To avoid the need for sending signed copyright hand-over forms to the
organisers or publishers, we require that authors license their
camera-ready papers under a
Creative Commons Attribution 4.0 International Licence
(CC-BY).
This means that authors (or their employers) retain copyright but
grant everybody 
the right to adapt and re-distribute their paper 
as long as the authors are credited and modifications listed,
see
\url{http://creativecommons.org/licenses/by/4.0/} for the
licence terms.

Depending on whether you use American or British English in your
paper, please include one of the following as an unmarked
(unnumbered) footnote on page 1 of your paper.
The \LaTeX{} style file (coling2014.sty) adds a command
\texttt{blfootnote} for this purpose, and usage of the command is
prepared in the \LaTeX{} source code (coling2014.tex) at the start
of Section~\ref{intro} ``Introduction''.

\begin{itemize}
    %
    % % final paper: en-uk version (to license, a licence)
    %
    \item  This work is licensed under a Creative Commons 
           Attribution 4.0 International Licence.
           Page numbers and proceedings footer are added by
           the organisers.
           Licence details:
           \url{http://creativecommons.org/licenses/by/4.0/}
    %
    % % final paper: en-us version (to licence, a license)
    %
    \item  This work is licenced under a Creative Commons 
           Attribution 4.0 International License.
           Page numbers and proceedings footer are added by
           the organizers.
           License details:
           \url{http://creativecommons.org/licenses/by/4.0/}
\end{itemize}

We cannot accept the restrictions ``NonCommercial'' (NC) and
``ShareAlike'' (SA) as they would
limit our options for an on-demand printing service
(we hope to find a partner for this after the conference)
and the restriction ``NoDerivatives'' (ND) as it may impede our
ability to add page numbers and a footer, as well as our ability
to combine papers into volumes.
We intend to publish the proceedings before the conference on
the ACL Anthology and to hand out a memory stick copy at the
conference.
Mentioning in the footnote that we, the organisers, add page numbers
and proceedings footer is a requirement stemming
from the CC-BY licence terms.

%
% We are not doing the following for Coling 2014. Maybe next time.
%
% \section{XML conversion and supported \LaTeX\ packages}
% 
% ACL 2014 innovates over earlier years in that we will attempt to
% automatically convert your \LaTeX\ source files to machine-readable
% XML with semantic markup. This will facilitate future research that
% uses the ACL proceedings themselves as a corpus.
% 
% We encourage you to submit a ZIP file of your \LaTeX\ sources along
% with the camera-ready version of your paper. We will then convert them
% to XML automatically, using the LaTeXML tool
% (\url{http://dlmf.nist.gov/LaTeXML}). LaTeXML has \emph{bindings} for
% a number of \LaTeX\ packages, including the ACL 2014 stylefile. These
% bindings allow LaTeXML to render the commands from these packages
% correctly in XML. For best results, we encourage you to use the
% packages that are officially supported by LaTeXML, listed at
% \url{http://dlmf.nist.gov/LaTeXML/manual/included.bindings}

\section{Translation of non-English Terms}

It is also advised to supplement non-English characters and terms
with appropriate transliterations and/or translations
since not all readers understand all such characters and terms.
Inline transliteration or translation can be represented in
the order of: original-form transliteration ``translation''.

\section{Length of Submission}
\label{sec:length}

The maximum submission length is 8 pages (A4), plus two extra pages for
references. Authors of accepted papers will be given additional space in
the camera-ready version to reflect space needed for changes stemming
from reviewers comments.

Papers that do not
conform to the specified length and formatting requirements may be
rejected without review.

\section*{Acknowledgements}

The acknowledgements should go immediately before the references.  Do
not number the acknowledgements section. Do not include this section
when submitting your paper for review.

% include your own bib file like this:
%\bibliographystyle{acl}
%\bibliography{acl2014}

\begin{thebibliography}{}

\bibitem[\protect\citename{Aho and Ullman}1972]{Aho:72}
Alfred~V. Aho and Jeffrey~D. Ullman.
\newblock 1972.
\newblock {\em The Theory of Parsing, Translation and Compiling}, volume~1.
\newblock Prentice-{Hall}, Englewood Cliffs, NJ.

\bibitem[\protect\citename{{American Psychological Association}}1983]{APA:83}
{American Psychological Association}.
\newblock 1983.
\newblock {\em Publications Manual}.
\newblock American Psychological Association, Washington, DC.

\bibitem[\protect\citename{{Association for Computing Machinery}}1983]{ACM:83}
{Association for Computing Machinery}.
\newblock 1983.
\newblock {\em Computing Reviews}, 24(11):503--512.

\bibitem[\protect\citename{Chandra \bgroup et al.\egroup }1981]{Chandra:81}
Ashok~K. Chandra, Dexter~C. Kozen, and Larry~J. Stockmeyer.
\newblock 1981.
\newblock Alternation.
\newblock {\em Journal of the Association for Computing Machinery},
  28(1):114--133.

\bibitem[\protect\citename{Gusfield}1997]{Gusfield:97}
Dan Gusfield.
\newblock 1997.
\newblock {\em Algorithms on Strings, Trees and Sequences}.
\newblock Cambridge University Press, Cambridge, UK.

\end{thebibliography}

\end{document}

### 2014-05-08 codeswitch-cltw-2014/codeswitch.tex
update
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52621 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-\title{Instructions for COLING-2014 Proceedings}
+\title{Subsegmental language detection in Celtic language text}
 
-\author{First Author \\
+\author{Akshay \\
   Affiliation / Address line 1 \\
@@
   {\tt email@domain} \\\And
-  Second Author \\
-  Affiliation / Address line 1 \\
-  Affiliation / Address line 2 \\
-  Affiliation / Address line 3 \\
-  {\tt email@domain} \\}
+  Francis M. Tyers \\
+  Giellatekno / CLEAR \\
+  UiT Norgga \'arktala\v{s} universitehta  \\
+  9017 Romsa (Norway) \\
+  {\tt francis.tyers@uit.no} \\}
 

### 2014-05-08 codeswitch-cltw-2014/codeswitch.tex
add more
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52622 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \begin{abstract}
-  This document contains the instructions for preparing a camera-ready
-  manuscript for the proceedings of COLING-2014. The document itself
-  conforms to its own specifications, and is therefore an example of
-  what your manuscript should look like. These instructions should be
-  used for both papers submitted for review and for final versions of
-  accepted papers.  Authors are asked to conform to all the directions
-  reported in this document.
+  This paper describes an experiment to detect the language of subsegments
+  of text in three Celtic languages: Breton, Irish and Welsh. 
 \end{abstract}
-
-\section{Credits}
-
-This document has been adapted from the instructions for the ACL-2014
-proceedings compiled by Alexander Koller and Yusuke Miyao,
-which are, in turn, based on the instructions for earlier ACL proceedings,
-including those for ACL-2012 by Maggie Li and Michael
-White, those from ACL-2010 by Jing-Shing Chang and Philipp Koehn,
-those for ACL-2008 by Johanna D. Moore, Simone Teufel, James Allan,
-and Sadaoki Furui, those for ACL-2005 by Hwee Tou Ng and Kemal
-Oflazer, those for ACL-2002 by Eugene Charniak and Dekang Lin, and
-earlier ACL and EACL formats. Those versions were written by several
-people, including John Chen, Henry S. Thompson and Donald
-Walker. Additional elements were taken from the formatting
-instructions of the {\em International Joint Conference on Artificial
-  Intelligence}.
 
@@
 
-\section{General Instructions}
-
-Manuscripts must be in single-column format.
-The title, authors' names and complete
-addresses
-must be centred at the top of the first page, and
-any full-width figures or tables (see the guidelines in
-Subsection~\ref{ssec:first}). {\bf Type single-spaced.}  Start all
-pages directly under the top margin. See the guidelines later
-regarding formatting the first page.  The manuscript should be
-printed single-sided and its length
-should not exceed the maximum page limit described in Section~\ref{sec:length}.
-Do not number the pages.
-
-\subsection{Electronically-available resources}
-
-We strongly prefer that you prepare your PDF files using \LaTeX{} with
-the official COLING 2014 style file (coling2014.sty) and bibliography style
-(acl.bst). These files are available at
-\url{http://www.coling-2014.org/call-for-papers.php}.
-You will also find the document
-you are currently reading (coling2014.pdf) and its \LaTeX{} source code
-(coling2014.tex) on this website.
-
-You can alternatively use Microsoft Word to produce your PDF file. In
-this case, we strongly recommend the use of the Word template file
-(coling2014.dot) on the COLING 2014 website. If you have an option, we
-recommend that you use the \LaTeX2e{} version. If you will be
-  using the Microsoft Word template, we suggest that you anonymise
-  your source file so that the pdf produced does not retain your
-  identity.  This can be done by removing any personal information
-from your source document properties.
-
-\subsection{Format of Electronic Manuscript}
-\label{sect:pdf}
-
-For the production of the electronic manuscript you must use Adobe's
-Portable Document Format (PDF). PDF files are usually produced from
-\LaTeX{} using the \textit{pdflatex} command. If your version of
-\LaTeX{} produces Postscript files, you can convert these into PDF
-using \textit{ps2pdf} or \textit{dvipdf}. On Windows, you can also use
-Adobe Distiller to generate PDF.
-
-Please make sure that your PDF file includes all the necessary fonts
-(especially tree diagrams, symbols, and fonts with Asian
-characters). When you print or create the PDF file, there is usually
-an option in your printer setup to include none, all or just
-non-standard fonts.  Please make sure that you select the option of
-including ALL the fonts. \textbf{Before sending it, test your PDF by
-  printing it from a computer different from the one where it was
-  created.} Moreover, some word processors may generate very large PDF
-files, where each page is rendered as an image. Such images may
-reproduce poorly. In this case, try alternative ways to obtain the
-PDF. One way on some systems is to install a driver for a postscript
-printer, send your document to the printer specifying ``Output to a
-file'', then convert the file to PDF.
-
-It is of utmost importance to specify the \textbf{A4 format} (21 cm
-x 29.7 cm) when formatting the paper. When working with
-{\tt dvips}, for instance, one should specify {\tt -t a4}.
-
-If you cannot meet the above requirements
-for the
-production of your electronic submission, please contact the
-publication chairs as soon as possible.
-
-\subsection{Layout}
-\label{ssec:layout}
-
-Format manuscripts with a single column to a page, in the manner these
-instructions are formatted. The exact dimensions for a page on A4
-paper are:
-
-\begin{itemize}
-\item Left and right margins: 2.5 cm
-\item Top margin: 2.5 cm
-\item Bottom margin: 2.5 cm
-\item Width: 16.0 cm
-\item Height: 24.7 cm
-\end{itemize}
-
-\noindent Papers should not be submitted on any other paper size.
-If you cannot meet the above requirements for
-the production of your electronic submission, please contact the
-publication chairs above as soon as possible.
-
-\subsection{Fonts}
-
-For reasons of uniformity, Adobe's {\bf Times Roman} font should be
-used. In \LaTeX2e{} this is accomplished by putting
-
-\begin{quote}
-\begin{verbatim}
-\usepackage{times}
-\usepackage{latexsym}
-\end{verbatim}
-\end{quote}
-in the preamble. If Times Roman is unavailable, use {\bf Computer
-  Modern Roman} (\LaTeX2e{}'s default).  Note that the latter is about
-  10\% less dense than Adobe's Times Roman font.
-
-The {\bf Times New Roman} font, which is configured for us in the
-Microsoft Word template (coling2014.dot) and which some Linux
-distributions offer for installation, can be used as well.
-
-\begin{table}[h]
-\begin{center}
-\begin{tabular}{|l|rl|}
-\hline \bf Type of Text & \bf Font Size & \bf Style \\ \hline
-paper title & 15 pt & bold \\
-author names & 12 pt & bold \\
-author affiliation & 12 pt & \\
-the word ``Abstract'' & 12 pt & bold \\
-section titles & 12 pt & bold \\
-document text & 11 pt  &\\
-captions & 11 pt & \\
-sub-captions & 9 pt & \\
-abstract text & 10 pt & \\
-bibliography & 10 pt & \\
-footnotes & 9 pt & \\
-\hline
-\end{tabular}
-\end{center}
-\caption{\label{font-table} Font guide. }
-\end{table}
-
-\subsection{The First Page}
-\label{ssec:first}
-
-Centre the title, author's name(s) and affiliation(s) across
-the page.
-Do not use footnotes for affiliations. Do not include the
-paper ID number assigned during the submission process.
-
-{\bf Title}: Place the title centred at the top of the first page, in
-a 15 pt bold font. (For a complete guide to font sizes and styles,
-see Table~\ref{font-table}) Long titles should be typed on two lines
-without a blank line intervening. Approximately, put the title at 2.5
-cm from the top of the page, followed by a blank line, then the
-author's names(s), and the affiliation on the following line. Do not
-use only initials for given names (middle initials are allowed). Do
-not format surnames in all capitals (e.g., use ``Schlangen'' not
-``SCHLANGEN'').  Do not format title and section headings in all
-capitals as well except for proper names (such as ``BLEU'') that are
-conventionally in all capitals.  The affiliation should contain the
-author's complete address, and if possible, an electronic mail
-address. Start the body of the first page 7.5 cm from the top of the
-page.
-
-The title, author names and addresses should be completely identical
-to those entered to the electronical paper submission website in order
-to maintain the consistency of author information among all
-publications of the conference. If they are different, the publication
-chairs may resolve the difference without consulting with you; so it
-is in your own interest to double-check that the information is
-consistent.
-
-{\bf Abstract}: Type the abstract between addresses and main body.
-The width of the abstract text should be
-smaller than main body by about 0.6 cm on each side.
-Centre the word {\bf Abstract} in a 12 pt bold
-font above the body of the abstract. The abstract should be a concise
-summary of the general thesis and conclusions of the paper. It should
-be no longer than 200 words. The abstract text should be in 10 pt font.
-
-{\bf Text}: Begin typing the main body of the text immediately after
-the abstract, observing the single-column format as shown in 
-the present document. Do not include page numbers.
-
-{\bf Indent} when starting a new paragraph. Use 11 pt for text and 
-subsection headings, 12 pt for section headings and 15 pt for
-the title.
-
-{\bf Licence}: Include a licence statement as an unmarked (unnumbered)
-footnote on the first page of the final, camera-ready paper.
-See Section~\ref{licence} below for details and motivation.
-
-\subsection{Sections}
-
-{\bf Headings}: Type and label section and subsection headings in the
-style shown on the present document.  Use numbered sections (Arabic
-numerals) in order to facilitate cross references. Number subsections
-with the section number and the subsection number separated by a dot,
-in Arabic numerals. Do not number subsubsections.
-
-{\bf Citations}: Citations within the text appear in parentheses
-as~\cite{Gusfield:97} or, if the author's name appears in the text
-itself, as Gusfield~\shortcite{Gusfield:97}.  Append lowercase letters
-to the year in cases of ambiguity.  Treat double authors as
-in~\cite{Aho:72}, but write as in~\cite{Chandra:81} when more than two
-authors are involved. Collapse multiple citations as
-in~\cite{Gusfield:97,Aho:72}. Also refrain from using full citations
-as sentence constituents. We suggest that instead of
-\begin{quote}
-  ``\cite{Gusfield:97} showed that ...''
-\end{quote}
-you use
-\begin{quote}
-``Gusfield \shortcite{Gusfield:97}   showed that ...''
-\end{quote}
-
-If you are using the provided \LaTeX{} and Bib\TeX{} style files, you
-can use the command \verb|\newcite| to get ``author (year)'' citations.
-
-As reviewing will be double-blind, the submitted version of the papers
-should not include the authors' names and affiliations. Furthermore,
-self-references that reveal the author's identity, e.g.,
-\begin{quote}
-``We previously showed \cite{Gusfield:97} ...''  
-\end{quote}
-should be avoided. Instead, use citations such as 
-\begin{quote}
-``Gusfield \shortcite{Gusfield:97}
-previously showed ... ''
-\end{quote}
-
-\textbf{Please do not use anonymous citations} and do not include
-any of the following when submitting your paper for review:
-acknowledgements, project names, grant numbers, and names or URLs of
-resources or tools that have only been made publicly available in
-the last 3 weeks or are about to be made public.
-Papers that do not
-conform to these requirements may be rejected without review.
-These details can, however, be included in the camera-ready, final paper.
-
-\textbf{References}: Gather the full set of references together under
-the heading {\bf References}; place the section before any Appendices,
-unless they contain references. Arrange the references alphabetically
-by first author, rather than by order of occurrence in the text.
-Provide as complete a citation as possible, using a consistent format,
-such as the one for {\em Computational Linguistics\/} or the one in the 
-{\em Publication Manual of the American 
-Psychological Association\/}~\cite{APA:83}.  Use of full names for
-authors rather than initials is preferred.  A list of abbreviations
-for common computer science journals can be found in the ACM 
-{\em Computing Reviews\/}~\cite{ACM:83}.
-
-The \LaTeX{} and Bib\TeX{} style files provided roughly fit the
-American Psychological Association format, allowing regular citations, 
-short citations and multiple citations as described above.
-
-{\bf Appendices}: Appendices, if any, directly follow the text and the
-references (but see above).  Letter them in sequence and provide an
-informative title: {\bf Appendix A. Title of Appendix}.
-
-\subsection{Footnotes}
-
-{\bf Footnotes}: Put footnotes at the bottom of the page and use 9 pt
-text. They may be numbered or referred to by asterisks or other
-symbols.\footnote{This is how a footnote should appear.} Footnotes
-should be separated from the text by a line.\footnote{Note the line
-separating the footnotes from the text.}
-
-\subsection{Graphics}
-
-{\bf Illustrations}: Place figures, tables, and photographs in the
-paper near where they are first discussed, rather than at the end, if
-possible. 
-Colour
-illustrations are discouraged, unless you have verified that  
-they will be understandable when printed in black ink.
-
-{\bf Captions}: Provide a caption for every illustration; number each one
-sequentially in the form:  ``Figure 1. Caption of the Figure.'' ``Table 1.
-Caption of the Table.''  Type the captions of the figures and 
-tables below the body, using 11 pt text.
-
-Narrow graphics together with the single-column format may lead to
-large empty spaces,
-see for example the wide margins on both sides of Table~\ref{font-table}.
-If you have multiple graphics with related content, it may be
-preferable to combine them in one graphic.
-You can identify the sub-graphics with sub-captions below the
-sub-graphics numbered (a), (b), (c) etc.\ and using 9 pt text.
-The \LaTeX{} packages wrapfig, subfig, subtable and/or subcaption
-may be useful.
-
-\subsection{Licence Statement}
-\label{licence}
-
-To avoid the need for sending signed copyright hand-over forms to the
-organisers or publishers, we require that authors license their
-camera-ready papers under a
-Creative Commons Attribution 4.0 International Licence
-(CC-BY).
-This means that authors (or their employers) retain copyright but
-grant everybody 
-the right to adapt and re-distribute their paper 
-as long as the authors are credited and modifications listed,
-see
-\url{http://creativecommons.org/licenses/by/4.0/} for the
-licence terms.
-
-Depending on whether you use American or British English in your
-paper, please include one of the following as an unmarked
-(unnumbered) footnote on page 1 of your paper.
-The \LaTeX{} style file (coling2014.sty) adds a command
-\texttt{blfootnote} for this purpose, and usage of the command is
-prepared in the \LaTeX{} source code (coling2014.tex) at the start
-of Section~\ref{intro} ``Introduction''.
-
-\begin{itemize}
-    %
-    % % final paper: en-uk version (to license, a licence)
-    %
-    \item  This work is licensed under a Creative Commons 
-           Attribution 4.0 International Licence.
-           Page numbers and proceedings footer are added by
-           the organisers.
-           Licence details:
-           \url{http://creativecommons.org/licenses/by/4.0/}
-    %
-    % % final paper: en-us version (to licence, a license)
-    %
-    \item  This work is licenced under a Creative Commons 
-           Attribution 4.0 International License.
-           Page numbers and proceedings footer are added by
-           the organizers.
-           License details:
-           \url{http://creativecommons.org/licenses/by/4.0/}
-\end{itemize}
-
-We cannot accept the restrictions ``NonCommercial'' (NC) and
-``ShareAlike'' (SA) as they would
-limit our options for an on-demand printing service
-(we hope to find a partner for this after the conference)
-and the restriction ``NoDerivatives'' (ND) as it may impede our
-ability to add page numbers and a footer, as well as our ability
-to combine papers into volumes.
-We intend to publish the proceedings before the conference on
-the ACL Anthology and to hand out a memory stick copy at the
-conference.
-Mentioning in the footnote that we, the organisers, add page numbers
-and proceedings footer is a requirement stemming
-from the CC-BY licence terms.
-
-%
-% We are not doing the following for Coling 2014. Maybe next time.
-%
-% \section{XML conversion and supported \LaTeX\ packages}
-% 
-% ACL 2014 innovates over earlier years in that we will attempt to
-% automatically convert your \LaTeX\ source files to machine-readable
-% XML with semantic markup. This will facilitate future research that
-% uses the ACL proceedings themselves as a corpus.
-% 
-% We encourage you to submit a ZIP file of your \LaTeX\ sources along
-% with the camera-ready version of your paper. We will then convert them
-% to XML automatically, using the LaTeXML tool
-% (\url{http://dlmf.nist.gov/LaTeXML}). LaTeXML has \emph{bindings} for
-% a number of \LaTeX\ packages, including the ACL 2014 stylefile. These
-% bindings allow LaTeXML to render the commands from these packages
-% correctly in XML. For best results, we encourage you to use the
-% packages that are officially supported by LaTeXML, listed at
-% \url{http://dlmf.nist.gov/LaTeXML/manual/included.bindings}
-
-\section{Translation of non-English Terms}
-
-It is also advised to supplement non-English characters and terms
-with appropriate transliterations and/or translations
-since not all readers understand all such characters and terms.
-Inline transliteration or translation can be represented in
-the order of: original-form transliteration ``translation''.
-
-\section{Length of Submission}
-\label{sec:length}
-
-The maximum submission length is 8 pages (A4), plus two extra pages for
-references. Authors of accepted papers will be given additional space in
-the camera-ready version to reflect space needed for changes stemming
-from reviewers comments.
-
-Papers that do not
-conform to the specified length and formatting requirements may be
-rejected without review.
-
 \section*{Acknowledgements}
 
-The acknowledgements should go immediately before the references.  Do
-not number the acknowledgements section. Do not include this section
-when submitting your paper for review.
+We thank Kevin Scannell...
 
 % include your own bib file like this:
-%\bibliographystyle{acl}
-%\bibliography{acl2014}
-
-\begin{thebibliography}{}
-
-\bibitem[\protect\citename{Aho and Ullman}1972]{Aho:72}
-Alfred~V. Aho and Jeffrey~D. Ullman.
-\newblock 1972.
-\newblock {\em The Theory of Parsing, Translation and Compiling}, volume~1.
-\newblock Prentice-{Hall}, Englewood Cliffs, NJ.
-
-\bibitem[\protect\citename{{American Psychological Association}}1983]{APA:83}
-{American Psychological Association}.
-\newblock 1983.
-\newblock {\em Publications Manual}.
-\newblock American Psychological Association, Washington, DC.
-
-\bibitem[\protect\citename{{Association for Computing Machinery}}1983]{ACM:83}
-{Association for Computing Machinery}.
-\newblock 1983.
-\newblock {\em Computing Reviews}, 24(11):503--512.
-
-\bibitem[\protect\citename{Chandra \bgroup et al.\egroup }1981]{Chandra:81}
-Ashok~K. Chandra, Dexter~C. Kozen, and Larry~J. Stockmeyer.
-\newblock 1981.
-\newblock Alternation.
-\newblock {\em Journal of the Association for Computing Machinery},
-  28(1):114--133.
-
-\bibitem[\protect\citename{Gusfield}1997]{Gusfield:97}
-Dan Gusfield.
-\newblock 1997.
-\newblock {\em Algorithms on Strings, Trees and Sequences}.
-\newblock Cambridge University Press, Cambridge, UK.
-
-\end{thebibliography}
+\bibliographystyle{acl}
+\bibliography{codeswitch}
 

### 2014-05-08 codeswitch-cltw-2014/codeswitch.tex
hmm
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52623 72bbbca6-d526-0410-a7d9-f06f51895060
add seg
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52624 72bbbca6-d526-0410-a7d9-f06f51895060
testing
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52625 72bbbca6-d526-0410-a7d9-f06f51895060
minocha-try2
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52626 72bbbca6-d526-0410-a7d9-f06f51895060
start
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52631 72bbbca6-d526-0410-a7d9-f06f51895060

@@
+%Try1: akshay minocha
 %
@@
 
-\author{Akshay \\
-  Affiliation / Address line 1 \\
-  Affiliation / Address line 2 \\
-  Affiliation / Address line 3 \\
-  {\tt email@domain} \\\And
+\author{Akshay Minocha \\
+  IIIT Hyderabad  \\
+  {\tt akshay.minocha@students.iiit.ac.in} \\\And
   Francis M. Tyers \\
@@
 \label{intro}
+
+\cite{lyu2006}
 
@@
 as soon as possible.
-
+Testing
 We will make additional instructions available at
@@
 this website regularly.
+
+\section{Evaluation}
+
+\section{Conclusions}
 

### 2014-05-08 codeswitch-cltw-2014/codeswitch.tex
intro added v1
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52636 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \label{intro}
-
+With the increase in multilingualism and the adoption of ‘another’ acceptable language apart from the usual creates a change in the data.
+This change can be looked upon as bad when we are faced with the burden of upgrading systems to adapt to it, on the other hand such change is
+unavoidable and has to be welcomed.
+The good part is that the user or the speaker feels more comfortable in expressing their opinions.
+Such change directly requires us to upgrade our current systems \cite{fung2008multilingual} so that the processing or the analysis of the data is in no way compromised.\\
+   Such code switching is more common in less formal ways of communication like on chats, forums and social media.
+Our motivation is to build a detection scheme which can sub-segment such sentences to the primary and secondary language tags and then prepare this to be processed further by modules which make use of this information to analyse useful information from this data.
+Specially in the case of languages like Irish, Breton and Welsh where the native speakers are very less, these people reach out to others and most of their posts are code-switched.
+Rather than promoting research based on monolingual technologies we would like to adopt to the trend and try to help other researchers in understanding this data in a better way.
 \cite{lyu2006}

### 2014-05-08 codeswitch-cltw-2014/codeswitch.tex
some more
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52639 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 Rather than promoting research based on monolingual technologies we would like to adopt to the trend and try to help other researchers in understanding this data in a better way.
+   In Ireland, and the surrounding areas a significant number of speakers use Welsh, Breton, Irish, English and French.
+About 14.6\% of the population of Wales considers themselves to be fluent in Welsh(2011) \footnote{http://ons.gov.uk/ons/rel/census/2011-census/key-statistics-for-unitary-authorities-in-wales/stb-2011-census-key-statistics-for-wales.html#tab---Proficiency-in-Welsh}.
+In the Republic of Ireland 94,000 people use it as a primary language outside the education system \footnote{http://www.cso.ie/en/media/csoie/census/documents/census2011pdr/Pdf\%208\%20Tables.pdf}.
+It is noticed that the Irish and the Welsh people often use English with their respective languages while they code-switch while Bretons have a natural French influence.
+Hence, in our experiments we plan to analyse the results for code switching and detection of segments in the following bilingual pairs - Welsh-English, Irish-English and Breton-French,  thus identifying the primary and secondary languages in the data.
 \cite{lyu2006}
@@
 %
+
 \blfootnote{

### 2014-05-09 codeswitch-cltw-2014/codeswitch.tex
evaluation
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52652 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \section{Evaluation}
+For the Evaluation procedure, we follow the footsteps of the CoNLL-2000 shared task: dividing text into syntactically related non-overlapping groups of words.
+This chunking mechanism \cite{tjong2003introduction} is very similar to ours, in terms of words which only belong to one category (here, language), and also evaluation based on the phrase structure present in the data.
+The chunks here are such that they belong to only one language.
+For example, the sentence below, shows the chunks as group of words in square brackets and their tagged language at the beginning in braces - \\
+
+[ \{br\} ur pegsun nevez da dapout e montroulez d'ar 15 a viz kerzu ] [\{fr\} nouvel autocollant dispo à #morlaix le 15 décembre ! ] \\
+
+For our current task, each of the texts have been limited to two languages i.e. the primary and the secondary language. Hence the type of chunking tags are limited to these.
 

### 2014-05-09 codeswitch-cltw-2014/codeswitch.tex
evaluation+
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52658 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-For our current task, each of the texts have been limited to two languages i.e. the primary and the secondary language. Hence the type of chunking tags are limited to these.
+For our current task, each of the texts have been limited to two languages i.e.
+the primary and the secondary language.
+Hence the type of chunking tags are limited to these.
+The evaluation statics shown in the Table mention two values for each of the experiment conducted on the three bilingual language datasets.
+The first, is the percentage of correctly detected phrases, which is the precision and the second is the number of phrases in the data that were found by the chunker, which is the recall.
 

### 2014-05-09 codeswitch-cltw-2014/codeswitch.tex
evaluation+
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52659 72bbbca6-d526-0410-a7d9-f06f51895060

--- 2014-05-09 codeswitch-cltw-2014/codeswitch.tex
@@
 
-For our current task, each of the texts have been limited to two languages i.e. the primary and the secondary language. Hence the type of chunking tags are limited to these.
+For our current task, each of the texts have been limited to two languages i.e.
+the primary and the secondary language.
+Hence the type of chunking tags are limited to these.
+The evaluation statics shown in the Table mention two values for each of the experiment conducted on the three bilingual language datasets.
+The first, is the percentage of correctly detected phrases, which is the overall precision and the second is the number of phrases in the data that were found by the chunker, which is the overall recall.
 

### 2014-05-09 codeswitch-cltw-2014/codeswitch.tex
add table
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52661 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{latexsym}
+\usepackage{alltt}
+\usepackage[small,bf]{caption}
+\usepackage{multirow}
 
@@
   IIIT Hyderabad  \\
-  {\tt akshay.minocha@students.iiit.ac.in} \\\And
+  Hyderabad (India) \\
+  {\small {\tt akshay.minocha@students.iiit.ac.in}} \\\And
   Francis M. Tyers \\
@@
   9017 Romsa (Norway) \\
-  {\tt francis.tyers@uit.no} \\}
+  {\small {\tt francis.tyers@uit.no}} \\}
 
@@
 Rather than promoting research based on monolingual technologies we would like to adopt to the trend and try to help other researchers in understanding this data in a better way.
-   In Ireland, and the surrounding areas a significant number of speakers use Welsh, Breton, Irish, English and French.
+
+%   In Ireland, and the surrounding areas a significant number of speakers use Welsh, Breton, Irish, English and French.
 About 14.6\% of the population of Wales considers themselves to be fluent in Welsh(2011) \footnote{http://ons.gov.uk/ons/rel/census/2011-census/key-statistics-for-unitary-authorities-in-wales/stb-2011-census-key-statistics-for-wales.html#tab---Proficiency-in-Welsh}.
@@
 \cite{lyu2006}
+
+\begin{figure}
+
+\end{figure}
+
+% issues:
+%% proper nouns
+%% other language interjections (fáilte), diolch, diolch yn fawr
+%% spelling
+%% abbreviations
 
@@
 
-The following instructions are directed to authors of papers submitted
-to COLING-2014 or accepted for publication in its proceedings. All
-authors are required to adhere to these specifications. Authors are
-required to provide a Portable Document Format (PDF) version of their
-papers. \textbf{The proceedings are designed for printing on A4
-  paper.}
-
-Authors from countries in which access to word-processing systems is
-limited should contact the publication chairs,
-Joachim Wagner, Liadh Kelly and Lorraine Goeuriot
-(\texttt{firstname.lastname @computing.dcu.ie} respectively),
-as soon as possible.
-Testing
-We will make additional instructions available at
-\url{http://www.coling-2014.org/call-for-papers.php}. Please check
-this website regularly.
-
 \section{Evaluation}
@@
 
-[ \{br\} ur pegsun nevez da dapout e montroulez d'ar 15 a viz kerzu ] [\{fr\} nouvel autocollant dispo à #morlaix le 15 décembre ! ] \\
+\begin{figure}
+\begin{small}
+\begin{alltt}
+@steffdafydd [\textbf{cy} gorfod cael bach o tan] [\textbf{en} though init]
+[\textbf{en} omg] [\textbf{cy} mar cwn bach yn] [\textbf{en} black and tan] [\textbf{cy} a popeth,] [\textbf{en} even cuter!!]
+\end{alltt}
+\end{small}
+\label{fig:tweets}
+\caption{Example of text from a microblogging site chunked}
+\end{figure}
 
@@
 The first, is the percentage of correctly detected phrases, which is the overall precision and the second is the number of phrases in the data that were found by the chunker, which is the overall recall.
+
+\section{Results}
+
+\begin{table}
+\begin{center}
+\begin{tabular}{|lc|rr|rr|rr|}
+\hline
+            & & \multicolumn{2}{c}{Irish---English} & \multicolumn{2}{c}{Welsh---English} & \multicolumn{2}{c}{Breton---French}  \\
+\hline
+                                        &         &  Irish &       English & Welsh & English & Breton & French \\
+\multirow{2}{*}{\texttt{baseline}} &  $p$ &  2.5 & 0.0        & 0.0 & 0.0 & 0.0 & 0.0 \\
+                                       & $r$  & 2.5   & 0.0       & 0.0 & 0.0 & 0.0 & 0.0 \\
+\hline
+\multirow{2}{*}{\texttt{langid2}} &  $p$ &  2.5 & 0.0        & 0.0 & 0.0 & 0.0 & 0.0 \\
+                                       & $r$  & 2.5   & 0.0       & 0.0 & 0.0 & 0.0 & 0.0 \\
+\hline
+\multirow{2}{*}{\texttt{langid3}} &  $p$ &  2.5 & 0.0        & 0.0 & 0.0 & 0.0 & 0.0 \\
+                                       & $r$  & 2.5   & 0.0       & 0.0 & 0.0 & 0.0 & 0.0 \\
+\hline
+\multirow{2}{*}{\texttt{wordlist}} &  $p$ &  2.5 & 0.0        & 0.0 & 0.0 & 0.0 & 0.0 \\
+                                       & $r$  & 2.5   & 0.0       & 0.0 & 0.0 & 0.0 & 0.0 \\
+\hline
+\multirow{2}{*}{\texttt{bigram}} &  $p$ &  2.5 & 0.0        & 0.0 & 0.0 & 0.0 & 0.0 \\
+                                       & $r$  & 2.5   & 0.0       & 0.0 & 0.0 & 0.0 & 0.0 \\
+\hline
+\multirow{2}{*}{\texttt{wordlist+bigram}} &  $p$ &  2.5 & 0.0        & 0.0 & 0.0 & 0.0 & 0.0 \\
+                                       & $r$  & 2.5   & 0.0       & 0.0 & 0.0 & 0.0 & 0.0 \\
+\hline
+\end{tabular}
+\end{center}
+
+\end{table}
 

### 2014-05-09 codeswitch-cltw-2014/codeswitch.tex
update numbers
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52662 72bbbca6-d526-0410-a7d9-f06f51895060
baseline
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52664 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 The first, is the percentage of correctly detected phrases, which is the overall precision and the second is the number of phrases in the data that were found by the chunker, which is the overall recall.
+\\
+
+The different techniques used for the sub-segmenting task are described as follows -
+
+1.
+Baseline - This is the most naive method of classification, we used the language identification tool langid.py \cite{lui2012langid} on the whole dataset and labelled all the individual lines according to this single major classification.
+\\
 
@@
 \hline
-\multirow{2}{*}{\texttt{wordlist}} &  $p$ &  2.5 & 0.0        & 0.0 & 0.0 & 0.0 & 0.0 \\
-                                       & $r$  & 2.5   & 0.0       & 0.0 & 0.0 & 0.0 & 0.0 \\
+\multirow{2}{*}{\texttt{wordlist}} &  $p$ &  32.50 & 28.57        & 26.69 & 40.91 & 57.41 & 33.96 \\
+                                       & $r$  & 23.64   & 26.09       & 26.03 & 33.75 & 47.69 & 33.33 \\
 \hline
@@
 
+(CHECKLATER-FORMATTING)
+1. Baseline - Negligible precision and recall show that the task of detecting sub-segments in complex and there needs to be more improvement in the methodology.
+
 \section{Conclusions}

### 2014-05-09 codeswitch-cltw-2014/codeswitch.tex
langid
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52668 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \\
+2.
+Langid character bigram and trigram prediction - After restricting the predicted languages to be amongst the two in the dataset, we used the character bi-gram and then tri-gram probabilities to predict the detected language for each token, some heuristics like in Methodology were followed which included that the span of a segment or a code-switched phrase is more than a word, the social media non-language entities such as the hashtags, usernames and urls do not make a difference.
+\\
 

### 2014-05-09 codeswitch-cltw-2014/codeswitch.tex
word lists
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52671 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \\
+3.
+Word based prediction - This is a simple heuristic which was designed on the basis of the the most common words in the wordlists of the languages which are in question.
+After checking each word against one of the word lists, it was associated to that particular language.
+In case of a confusion, for example, when the word exists in both the word lists or not even in one, the option of continuing with the previous span was taken and the previous selected tag was labelled, thus increasing the chunk.
+\\
 

### 2014-05-09 codeswitch-cltw-2014/codeswitch.tex
word lists backoff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52673 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \\
+4. Character bigram - This method is in line with what has been described in the Methodology section \\
+5.
+Word based Prediction with Character backoff - A better way to predict the spans of the sub-segments of the text is to include the two methods of word and character based techniques as described above.
+In case of the word being present in only one of the two monolingual word lists the classification is simple, but in case of a confusion as described in 4.
+above, a character bigram backoff was introduced to help us disambiguate the language label.
+This method works well because the earlier heuristic approach of just labelling the word with the label of the span which is expanding, would mean less code-switch detection and more shift towards baseline.
+\\
 

### 2014-05-09 codeswitch-cltw-2014/codeswitch.tex
word lists backoff+
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52674 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 2.
-Langid character bigram and trigram prediction - After restricting the predicted languages to be amongst the two in the dataset, we used the character bi-gram and then tri-gram probabilities to predict the detected language for each token, some heuristics like in Methodology were followed which included that the span of a segment or a code-switched phrase is more than a word, the social media non-language entities such as the hashtags, usernames and urls do not make a difference.
+Langid character bigram and trigram prediction - After restricting the predicted languages to be amongst the two in the dataset, we used the character bi-gram and then tri-gram probabilities to predict the detected language for each token, some rules like in Methodology were followed which included that the span of a segment or a code-switched phrase is more than a word, the social media non-language entities such as the hashtags, usernames and urls do not make a difference.
 \\

### 2014-05-09 codeswitch-cltw-2014/codeswitch.tex
accuracy
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52675 72bbbca6-d526-0410-a7d9-f06f51895060
add more tables
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52676 72bbbca6-d526-0410-a7d9-f06f51895060
update
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52677 72bbbca6-d526-0410-a7d9-f06f51895060
add table doc stats
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52679 72bbbca6-d526-0410-a7d9-f06f51895060
results+
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52681 72bbbca6-d526-0410-a7

@@
 
+As described in the evaluation the techniques were used on a dataset collected over a period of time, for the three language pairs\footnote{\url{http://indigenoustweets.blogspot.in/2013/12/mapping-celtic-twittersphere.html}}.
+The statics of the same are described in the Table below.
+//
+
+\begin{table}
+\begin{center}
+\begin{tabular}{|l|r|r|r|r}
+\hline
+\textbf{System} &  \multicolumn{3}{c|}{\textbf{Statistics} (\%)} \\
+\hline
+       &   lang(1) Tokens & lang(2) Tokens & Phrases of lang(1) & Phrases of lang(2) \\
+\hline
+\texttt{Irish(1)---English(2)} & 332 & 379 & 40 & 42 \\
+\hline
+\texttt{Welsh(1)---English(2)} & 419 & 378 & 64 & 66 \\
+\hline
+\texttt{Breton(1)---English(2)} & 388 & 379 & 54 & 53 \\
+\hline
+\end{tabular}
+\end{center}
+\label{table:accuracy}
+\caption{Document Statistics of the annotated data used.}
+\end{table}
+
 \begin{table}
@@
 
+\begin{table}
+\begin{center}
+\begin{tabular}{|l|r|r|r|}
+\hline
+\textbf{System} &  \multicolumn{3}{c|}{\textbf{Accuracy} (\%)} \\
+\hline
+       &   Irish---English & Welsh---English & Breton---French \\ 
+\hline
+\texttt{baseline} & 42.76 & 42.16 & 44.07 \\
+\hline
+\texttt{langid2} & 50.35 & 42.28 & 45.11  \\
+\hline
+\texttt{langid3} & 57.24 & 45.92 & 43.16 \\
+\hline
+\texttt{wordlist} & 79.75 & 74.28 & 83.96 \\
+\hline
+\texttt{bigram} & 81.29 & 65.62 & 76.79 \\
+\hline
+\texttt{wordlist+bigram} & 85.79 & 72.40 & 88.79 \\
+\hline
+\end{tabular}
+\end{center}
+\label{table:accuracy}
+\caption{Accuracy of the systems over the three language pairs}
+\end{table}
+
 (CHECKLATER-FORMATTING)
-1. Baseline - Negligible precision and recall show that the task of detecting sub-segments in complex and there needs to be more improvement in the methodology.
+
+The Results clearly show that the baseline has a negligible precision and recall, this means that the task of detecting sub-segments is complex and there needs to be more improvement in the technique.
+You can see that langid.py is better suited for document classification and that when such varied content is introduced the result is an improvement but not promising.
+In case of the character bigram prediction technique and the word based chunking method, the results show a significant increase in precision and recall values from the rest.
+The last in the list is the word based prediction with character backoff, which performs collectively better than the others with a high accurracy margin.
+\\
 

### 2014-05-09 codeswitch-cltw-2014/codeswitch.tex
table correction
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52682 72bbbca6-d526-0410-a7d9-f06f51895060
update
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52685 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 % Contact: jwagner@computing.dcu.ie
-%%
-%% Based on the style files for ACL-2014, which were, in turn,
-%% Based on the style files for ACL-2013, which were, in turn,
-%% Based on the style files for ACL-2012, which were, in turn,
-%% based on the style files for ACL-2011, which were, in turn, 
-%% based on the style files for ACL-2010, which were, in turn, 
-%% based on the style files for ACL-IJCNLP-2009, which were, in turn,
-%% based on the style files for EACL-2009 and IJCNLP-2008...
-
-%% Based on the style files for EACL 2006 by 
-%%e.agirre@ehu.es or Sergi.Balari@uab.es
-%% and that of ACL 08 by Joakim Nivre and Noah Smith
 
@@
 \label{intro}
-With the increase in multilingualism and the adoption of ‘another’ acceptable language apart from the usual creates a change in the data.
-This change can be looked upon as bad when we are faced with the burden of upgrading systems to adapt to it, on the other hand such change is
-unavoidable and has to be welcomed.
-The good part is that the user or the speaker feels more comfortable in expressing their opinions.
-Such change directly requires us to upgrade our current systems \cite{fung2008multilingual} so that the processing or the analysis of the data is in no way compromised.\\
-   Such code switching is more common in less formal ways of communication like on chats, forums and social media.
-Our motivation is to build a detection scheme which can sub-segment such sentences to the primary and secondary language tags and then prepare this to be processed further by modules which make use of this information to analyse useful information from this data.
-Specially in the case of languages like Irish, Breton and Welsh where the native speakers are very less, these people reach out to others and most of their posts are code-switched.
-Rather than promoting research based on monolingual technologies we would like to adopt to the trend and try to help other researchers in understanding this data in a better way.
 
-%   In Ireland, and the surrounding areas a significant number of speakers use Welsh, Breton, Irish, English and French.
-About 14.6\% of the population of Wales considers themselves to be fluent in Welsh(2011) \footnote{http://ons.gov.uk/ons/rel/census/2011-census/key-statistics-for-unitary-authorities-in-wales/stb-2011-census-key-statistics-for-wales.html#tab---Proficiency-in-Welsh}.
-In the Republic of Ireland 94,000 people use it as a primary language outside the education system \footnote{http://www.cso.ie/en/media/csoie/census/documents/census2011pdr/Pdf\%208\%20Tables.pdf}.
-It is noticed that the Irish and the Welsh people often use English with their respective languages while they code-switch while Bretons have a natural French influence.
-Hence, in our experiments we plan to analyse the results for code switching and detection of segments in the following bilingual pairs - Welsh-English, Irish-English and Breton-French,  thus identifying the primary and secondary languages in the data.
-\cite{lyu2006}
+Determining the language of a piece of text is one of the first steps that must be taken
+before proceeding with further computational processing. This task has received a substantial amount of
+attention in recent years (X,Y,Z). However, previous research has on the whole assumed
+that a given text will be in a single language. When dealing with text from formal domains,
+this may be the case --- although there are exceptions --- such as quotations embedded in
+the text in another language. But when dealing with informal text, particularly in languages
+where the speech community is predominantly bi- or multi-lingual, this assumption may not hold.
 
-\begin{figure}
+There are several ways in which an informal text may contain parts in different languages, such as
+code switching, quotations, named entities, interjections, translations, etc. Some of these are presented
+in Table~\ref{table:examples}
 
-\end{figure}
+\begin{table}
+\begin{tabular}{ll}
+ \textbf{Code switching}: & \\
+\textbf{Quotations}: & And then he said ` ' \\
+ \textbf{Named entities}: &``[Dr Jekyll] ha [Mr Hyde] embannet gant [Éditions Aber]'' \\
+ \textbf{Interjections}: & ``Hey, that's great, [diolch yn fawr!]'' \\
+ \textbf{Translations}: & ``Bloavezh mat d'an holl ! [Bonne année à tous !]'' \\
+\end{tabular}
+\label{table:examples}
+\caption{Some examples of text segments containing more than one language}
+\end{table}
+
+The work presented in this paper was motivated by the problems in normalising non-standard input
+for the Celtic languages as a precursor to machine translation. When applying a normalisation 
+strategy to a piece of text, it is necessary to first know the language of the piece of text you 
+are applying it to.
+
+The remainder of the paper is laid out as follows. In section~\ref{sec:method} we describe the problem
+in more detail and look at relevant prior work before proposing a novel method of sub-sentential
+language detection. Section~\ref{sec:eval} describes the evaluation methodology. Then in section~\ref{sec:results}
+we present the results of our method and compare it against several other possible methods. Finally, section~\ref{sec:conclusions}
+presents future work and conclusions.
+
+\section{Methodology}
+\label{sec:method}
 
@@
 \section{Evaluation}
+\label{sec:eval}
 For the Evaluation procedure, we follow the footsteps of the CoNLL-2000 shared task: dividing text into syntactically related non-overlapping groups of words.
@@
 \section{Results}
-
+\label{sec:results}
 As described in the evaluation the techniques were used on a dataset collected over a period of time, for the three language pairs\footnote{\url{http://indigenoustweets.blogspot.in/2013/12/mapping-celtic-twittersphere.html}}.
@@
 \hline
-\textbf{System} &  \multicolumn{3}{c|}{\textbf{Statistics} (\%)} \\
+\multirow{2}{*}{\textbf{Pair}} & \multirow{2}{*}{\textbf{Language}}  & \multicolumn{2}{c|}{\textbf{Statistics} (\%)} \\\cline{3-4}
+                &  &  Tokens & Segments \\
 \hline
-       &   lang(1) Tokens & lang(2) Tokens & Phrases of lang(1) & Phrases of lang(2) \\
+\multirow{2}{*}{Irish---English} & Irish & 332 & 40 \\
+                                 & English & 379 & 42 \\
 \hline
-\texttt{Irish(1)---English(2)} & 332 & 379 & 40 & 42 \\
+\multirow{2}{*}{Welsh---English} & Welsh & 419 & 64 \\
+                                 & English & 378 & 66  \\
 \hline
-\texttt{Welsh(1)---English(2)} & 419 & 378 & 64 & 66 \\
-\hline
-\texttt{Breton(1)---English(2)} & 388 & 379 & 54 & 53 \\
+\multirow{2}{*}{Breton---French} & Breton & 388 & 54 \\
+                                 & French & 379 & 53  \\
 \hline
@@
 \section{Conclusions}
+\label{sec:conclusions}
 

### 2014-05-09 codeswitch-cltw-2014/codeswitch.tex
updates
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52686 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 %\setlength\titlebox{5cm}
-
-% You can expand the titlebox if you need extra space
-% to show all the authors. Please do not make the titlebox
-% smaller than 5cm (the original size); we will check this
-% in the camera-ready version and ask you to change it back.
 
@@
 \textbf{Quotations}: & And then he said ` ' \\
- \textbf{Named entities}: &``[Dr Jekyll] ha [Mr Hyde] embannet gant [Éditions Aber]'' \\
- \textbf{Interjections}: & ``Hey, that's great, [diolch yn fawr!]'' \\
- \textbf{Translations}: & ``Bloavezh mat d'an holl ! [Bonne année à tous !]'' \\
+ \textbf{Named entities}: & [Dr Jekyll] ha [Mr Hyde] embannet gant [\'{E}ditions Aber] \\
+ \textbf{Interjections}: & Hey, that's great, [diolch yn fawr!] \\
+ \textbf{Translations}: & Bloavezh mat d'an holl ! [Bonne ann\'{e} \`{a} tous !] \\
 \end{tabular}

### 2014-05-09 codeswitch-cltw-2014/codeswitch.tex
perfect takk kps
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52687 72bbbca6-d526-0410-a7d9-f06f51895060
+
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52688 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \begin{tabular}{ll}
- \textbf{Code switching}: & \\
+ \textbf{Code switching}: & You're a [Meiriceánach, cén fáth] are you [foghlaim Gaeilge?!] \\
 \textbf{Quotations}: & And then he said ` ' \\
@@
 \label{sec:eval}
+
 For the Evaluation procedure, we follow the footsteps of the CoNLL-2000 shared task: dividing text into syntactically related non-overlapping groups of words.
@@
 The different techniques used for the sub-segmenting task are described as follows - 
-
 1.

### 2014-05-09 codeswitch-cltw-2014/codeswitch.tex
numbers
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52689 72bbbca6-d526-0410-a7d9-f06f51895060
update2
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52690 72bbbca6-d526-0410-a7d9-f06f51895060
some notes for future work
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52695 72bbbca6-d526-0410-a7d9-f06f51895060
some notes for future work
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52696 72bbbca6-d526-0410-a7d9-f06f51895060
update
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52698 72bbb

@@
 \begin{tabular}{ll}
- \textbf{Code switching}: & You're a [Meiriceánach, cén fáth] are you [foghlaim Gaeilge?!] \\
+ \textbf{Code switching}: & You're a [Meirice\'{a}nach, c\'{e}n f\'{a}th] are you [foghlaim Gaeilge?!] \\
 \textbf{Quotations}: & And then he said ` ' \\
@@
                                           &      &  Irish &  English & Welsh  & English & Breton & French \\ 
-\multirow{2}{*}{\texttt{baseline}} &  $p$ &  2.5 & 0.0        & 0.0 & 0.0 & 0.0 & 0.0 \\
-                                       & $r$  & 2.5   & 0.0       & 0.0 & 0.0 & 0.0 & 0.0 \\
+\multirow{2}{*}{\texttt{baseline}}        &  $p$ &  2.50   & 0.0      & 0.0   & 0.0 & 0.0 & 0.0 \\
+                                          & $r$  & 2.56    & 0.0      & 0.0   & 0.0 & 0.0 & 0.0 \\
 \hline
-\multirow{2}{*}{\texttt{langid2}} &  $p$ &  2.5 & 0.0        & 0.0 & 0.0 & 0.0 & 0.0 \\
-                                       & $r$  & 2.5   & 0.0       & 0.0 & 0.0 & 0.0 & 0.0 \\
+\multirow{2}{*}{\texttt{langid2}}         &  $p$ &  0.00   & 7.14     & 0.0   & 18.18    & 0.0 & 28.30 \\
+                                          & $r$  & 0.00    & 5.00     & 0.0   & 14.12 & 0.0 & 18.99 \\
 \hline
-\multirow{2}{*}{\texttt{langid3}} &  $p$ &  2.5 & 0.0        & 0.0 & 0.0 & 0.0 & 0.0 \\
-                                       & $r$  & 2.5   & 0.0       & 0.0 & 0.0 & 0.0 & 0.0 \\
+\multirow{2}{*}{\texttt{langid3}}         &  $p$ &  5.00   & 14.29    & 0.0   & 21.21 & 1.85 & 20.75 \\
+                                          & $r$  & 5.41    & 8.45     & 0.0   & 14.58 & 1.92 & 12.36 \\
 \hline
-\multirow{2}{*}{\texttt{wordlist}} &  $p$ &  32.50 & 28.57        & 26.69 & 40.91 & 57.41 & 33.96 \\
-                                       & $r$  & 23.64   & 26.09       & 26.03 & 33.75 & 47.69 & 33.33 \\
+\multirow{2}{*}{\texttt{wordlist}}        &  $p$ &  32.50 & 28.57     & 26.69 & {\bf 40.91} & 57.41 & 33.96 \\
+                                          & $r$  & 23.64  & 26.09     & {\bf 26.03} & {\bf 33.75} & 47.69 & 33.33 \\
 \hline
-\multirow{2}{*}{\texttt{bigram}} &  $p$ &  2.5 & 0.0        & 0.0 & 0.0 & 0.0 & 0.0 \\
-                                       & $r$  & 2.5   & 0.0       & 0.0 & 0.0 & 0.0 & 0.0 \\
+\multirow{2}{*}{\texttt{bigram}}          &  $p$ &  32.50   & 35.71   & 23.44 & 19.70  & 57.41 & 52.83 \\
+                                          & $r$  & 22.41    & 26.79   & 15.31 & 16.67 & 41.33 & 37.84 \\
 \hline
-\multirow{2}{*}{\texttt{wordlist+bigram}} &  $p$ &  2.5 & 0.0        & 0.0 & 0.0 & 0.0 & 0.0 \\
-                                       & $r$  & 2.5   & 0.0       & 0.0 & 0.0 & 0.0 & 0.0 \\
+\multirow{2}{*}{\texttt{wordlist+bigram}} &  $p$ &  {\bf 52.50}   & {\bf 50.00}   & {\bf 32.81} & 31.82 & {\bf 70.37} & {\bf 67.92} \\
+                                          & $r$  & {\bf 38.18}    & {\bf 43.75}   & 24.14 & 25.61 & {\bf 57.58} & {\bf 57.14} \\
 \hline
@@
 \end{center}
+\label{table:precisionrecall}
+\caption{Precision, $p$ and recall, $r$ for the systems by language.}
 
@@
 \label{table:accuracy}
-\caption{Accuracy of the systems over the three language pairs}
+\caption{Accuracy of the systems over the three language pairs. The accuracy measures how often a token
+  was assigned to the right language, independent of span.}
 \end{table}
@@
 
+% future work
+
+%% bigger and better test data, and distributable
+%% better evaluation -- significance testing
+%%  evaluation by application
+%% try it with scots gaelic
+%% 
+%% mikel's idea
+%% higher order character n-gram models
+
 \section*{Acknowledgements}

### 2014-05-09 codeswitch-cltw-2014/codeswitch.tex
related work
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52712 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 presents future work and conclusions.
+
+\section{Related Work}
+\label{sec:relwork}
+
+There has been significant research on code-switching and segment detection problems.
+The good thing about multilingual code switched data is that there are differences in the words of different languages - phonologically, morphologically, lexically and syntactically.
+At least that is the case with the pairs of languages we have chosen according to their usage in common day scenario.
+\\
+
+Before the Internet era, this study was focussed more on speech data.
+Initial research on this code switching was done for Marathi and English \cite{joshi1982processing} which included notable work about the formalisms and parsing.
+\\
+
+It is seen that language modelling techniques have shown promise earlier, such as in \cite{yu2013identification} the experiment on Mandarin-Taiwanese code-switched sentences show a high accuracy in terms of detecting code-switched sentences.
+\\
+
+Our idea of going to character level inspection has been inspired both by the language modelling techniques used by the community as well as the acoustic modelling research.
+From \cite{chan2004detection} which has made use of the bi-phone probabilities and calculated them to measure a confidence metric, to \cite{lyu2006language} which has made use of a classification method, named syllable-based duration classification, which uses the tonal syllable properties along with the speech signals to help predict the code switch points.
+\cite{yeong2010language} uses syllable structure information, and increases our confidence is using the character n-gram approach for detection of the code switching points, their study was for the Malay-English, identifying the language of the words in the code switched data set.\\
 

### 2014-05-09 codeswitch-cltw-2014/codeswitch.tex
formatting
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52714 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \begin{center}
-\begin{tabular}{|lc|rr|rr|rr|}
+\begin{tabular}{|lc|r|r|r|r|r|r|}
 \hline
-            & & \multicolumn{2}{c}{Irish---English} & \multicolumn{2}{c}{Welsh---English} & \multicolumn{2}{c}{Breton---French}  \\
+\multirow{2}{*}{\textbf{System}}            & & \multicolumn{2}{c}{\textbf{Irish---English}} & \multicolumn{2}{|c|}{\textbf{Welsh---English}} & \multicolumn{2}{c|}{\textbf{Breton---French}}  \\\cline{3-8}
+                                          &      &  Irish &  English & Welsh  & English & Breton & French \\
 \hline 
-                                          &      &  Irish &  English & Welsh  & English & Breton & French \\ 
 \multirow{2}{*}{\texttt{baseline}}        &  $p$ &  2.50   & 0.0      & 0.0   & 0.0 & 0.0 & 0.0 \\
@@
 \hline
-\texttt{wordlist} & 79.75 & 74.28 & 83.96 \\
+\texttt{wordlist} & 79.75 & \textbf{74.28} & 83.96 \\
 \hline
@@
 \hline
-\texttt{wordlist+bigram} & 85.79 & 72.40 & 88.79 \\
+\texttt{wordlist+bigram} & \textbf{85.79} & 72.40 & \textbf{88.79} \\
 \hline

### 2014-05-09 codeswitch-cltw-2014/codeswitch.tex
update
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52715 72bbbca6-d526-0410-a7d9-f06f51895060
ref+
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52719 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 before proceeding with further computational processing. This task has received a substantial amount of
-attention in recent years (X,Y,Z). However, previous research has on the whole assumed
+attention in recent years (\cite{cavnar1994n},\cite{lui2012langid}). However, previous research has on the whole assumed
 that a given text will be in a single language. When dealing with text from formal domains,
@@
 \hline
-\textbf{System} &  \multicolumn{3}{c|}{\textbf{Accuracy} (\%)} \\
-\hline
+\multirow{2}{*}{\textbf{System}} &  \multicolumn{3}{c|}{\textbf{Accuracy} (\%)} \\\cline{2-4}
        &   Irish---English & Welsh---English & Breton---French \\ 

### 2014-05-09 codeswitch-cltw-2014/codeswitch.tex
methodology
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52721 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \label{sec:method}
+We use the character n-gram approach along with some heuristics to mainly define the first task at hand.
+We would like to both predict the code switched points but looking at the surrounding structure also decide the inclusion of them into the current or the next segment.
+\\
+
+Alphabet n-gram approach -
+
+To predict code switch points we needed some raw text to train our character language models on.
+For English and French, we used the help of the Europarl \cite{koehn2005europarl}, for Breton, Welsh and Irish, data from credible sources on web were fetched.
+The size of the dataset for all the language sources was about 1.5million words.
+\\
+Building the character language models was undertaken using IRSTLM \cite{federico08a}, the input characters are to be introduced in the format of token separated by space and blank spaces replaced by ‘_’.
+For example, “sláinte” would be broken down into a set which is \{‘_ s’, ‘s l’, ‘l á’, ‘á i’, ‘i n’, ‘n t’, ‘t e’, ‘e _’\}.Then the probability for the alphabet sequence is calculated for each language pair, to be used in the experiment.
+Improved kneser-ney is used for smoothing, and this smoothing approach would be useful for higher-grams.
+\cite{kneser1995improved} \\
 

### 2014-05-09 codeswitch-cltw-2014/codeswitch.tex
methodology
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52722 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \cite{kneser1995improved} \\
+
+Once we get the probabilities, our task at hand is to take in the data and process it such that the  sub-segments of code-switched data are detected in it.
+We collected data from twitter over a period of time and this was filtered into three sets which had been identified to be Irish, Welsh and Breton.
+To be able to compare with the language model built above in the character n-gram approach, this data is to be processed like it aswell.
+\\
+
+Since, there were instances of non-standard input in the Twitter data, certain terminologies were escaped, such as mentions (starting with ‘@’), hashtags (#xyz), Re-Tweet (RT) and hyperlinks.\\
+This processed data is then given as input to our module which uses heuristics to come up with sub-segment information. \\
+
+Code-Switch Chunking - \\
 

### 2014-05-09 codeswitch-cltw-2014/codeswitch.tex
methodology+
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52723 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 Code-Switch Chunking - \\
+
+This section describes the rules taken into account while performing the chunking of the input data according to the code-switch prediction. \\
+
+A flag is set at the beginning, and this moves forward to the position in the text upto where confident chunking has been performed.
+For example, in case of a continuous prediction there would be no change and since the predicted tag of the new word corresponds to that already of the expanding chunk, this word is also included in the same chunk.
+In case of a doubt, the flag is set at the same position, and no final decision on labelling the word is performed, to have more confidence on the code-switch, the process moves on to the next token, if it corresponds to that of the previous chunk, then both these undecided words are labelled and included in the chunk, otherwise a change is noted and a new chunk is created with the changed label.
+In this process, we are trying to identify the sub-segments that we need, thus doing two tasks together and optimally, one is the prediction and second is the decision on chunking.
+\\
 

### 2014-05-09 codeswitch-cltw-2014/codeswitch.tex
fix formatting
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52727 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \\
-Building the character language models was undertaken using IRSTLM \cite{federico08a}, the input characters are to be introduced in the format of token separated by space and blank spaces replaced by ‘_’.
-For example, “sláinte” would be broken down into a set which is \{‘_ s’, ‘s l’, ‘l á’, ‘á i’, ‘i n’, ‘n t’, ‘t e’, ‘e _’\}.Then the probability for the alphabet sequence is calculated for each language pair, to be used in the experiment.
+Building the character language models was undertaken using IRSTLM \cite{federico08a}, the input characters are to be introduced in the format of token separated by space and blank spaces replaced by `\_'.
+For example, `sl\'{a}inte' would be broken down into a set which is \{`\_ s', `s l', `l \'{a}', `\'{a} i', `i n', `n t', `t e', `e \_'\}.Then the probability for the alphabet sequence is calculated for each language pair, to be used in the experiment.
 Improved kneser-ney is used for smoothing, and this smoothing approach would be useful for higher-grams.
@@
 
-Since, there were instances of non-standard input in the Twitter data, certain terminologies were escaped, such as mentions (starting with ‘@’), hashtags (#xyz), Re-Tweet (RT) and hyperlinks.\\
+Since, there were instances of non-standard input in the Twitter data, certain terminologies were escaped, such as mentions (starting with ‘@’), hashtags (\#xyz), Re-Tweet (RT) and hyperlinks.\\
 This processed data is then given as input to our module which uses heuristics to come up with sub-segment information. \\

### 2014-05-09 codeswitch-cltw-2014/codeswitch.tex
methodology+
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52732 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 Once we get the probabilities, our task at hand is to take in the data and process it such that the  sub-segments of code-switched data are detected in it.
-We collected data from twitter over a period of time and this was filtered into three sets which had been identified to be Irish, Welsh and Breton.
+We collected data from twitter over a period of time \footnote{\url{http://indigenoustweets.blogspot.in/2013/12/mapping-celtic-twittersphere.html}} and this was filtered into three sets which had been identified to be Irish, Welsh and Breton.
 To be able to compare with the language model built above in the character n-gram approach, this data is to be processed like it aswell.
@@
 \label{sec:results}
-As described in the evaluation the techniques were used on a dataset collected over a period of time, for the three language pairs\footnote{\url{http://indigenoustweets.blogspot.in/2013/12/mapping-celtic-twittersphere.html}}.
-The statics of the same are described in the Table below.
+As described in the ~\ref{sec:method} the data collected from Twitter for the three language pairs, was processed using the techniques mentioned.
+The statics of the same are described in ~\ref{table:accuracy}.
 //
@@
 
-(CHECKLATER-FORMATTING)
-
 The Results clearly show that the baseline has a negligible precision and recall, this means that the task of detecting sub-segments is complex and there needs to be more improvement in the technique.

### 2014-05-09 codeswitch-cltw-2014/codeswitch.tex
some revisions and additions here and there
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52733 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
+Word based prediction -
+
+ This is a simple heuristic which was designed on the basis of the the most common words in the wordlists of the languages which are in question.
+After checking each word against one of the two word lists, it is associated to that particular language.
+In case of a confusion, for example, when the word exists in both the word lists or not in either, the option of continuing with the previous span was taken and the previous selected tag was labelled, thus increasing the chunk.
+\\
+
+Word based Prediction with Character backoff -
+
+  A better way to predict the spans of the sub-segments of the text is to include the two methods of word and character based techniques as described above.
+In case of the word being present in only one of the two monolingual word lists the classification is simple, but in case of a confusion as described in Word based prediction.
+above, a character bigram backoff was introduced to help us disambiguate the language label.
+This method works well because the earlier heuristic approach of just labelling the word with the label of the span which is expanding, would mean less code-switch detection and more shift towards baseline.
+\\
+
 % issues:
@@
 
-The different techniques used for the sub-segmenting task are described as follows - 
-1.
-Baseline - This is the most naive method of classification, we used the language identification tool langid.py \cite{lui2012langid} on the whole dataset and labelled all the individual lines according to this single major classification.
+Apart from the techniques discussed in ~/ref{sec:method}, some baselines are also used to give a comparative view of how well all the mechanism perform.
+
+Baseline \\
+
+ This is the most naive method of classification, we used the language identification tool langid.py \cite{lui2012langid} on the whole dataset and labelled all the individual lines according to this single major classification.
 \\
-2.
-Langid character bigram and trigram prediction - After restricting the predicted languages to be amongst the two in the dataset, we used the character bi-gram and then tri-gram probabilities to predict the detected language for each token, some rules like in Methodology were followed which included that the span of a segment or a code-switched phrase is more than a word, the social media non-language entities such as the hashtags, usernames and urls do not make a difference.
-\\
-3.
-Word based prediction - This is a simple heuristic which was designed on the basis of the the most common words in the wordlists of the languages which are in question.
-After checking each word against one of the word lists, it was associated to that particular language.
-In case of a confusion, for example, when the word exists in both the word lists or not even in one, the option of continuing with the previous span was taken and the previous selected tag was labelled, thus increasing the chunk.
-\\
-4. Character bigram - This method is in line with what has been described in the Methodology section \\
-5.
-Word based Prediction with Character backoff - A better way to predict the spans of the sub-segments of the text is to include the two methods of word and character based techniques as described above.
-In case of the word being present in only one of the two monolingual word lists the classification is simple, but in case of a confusion as described in 4.
-above, a character bigram backoff was introduced to help us disambiguate the language label.
-This method works well because the earlier heuristic approach of just labelling the word with the label of the span which is expanding, would mean less code-switch detection and more shift towards baseline.
+
+Langid character bigram and trigram prediction - \\
+
+After restricting the predicted languages to be amongst the two in the dataset, we used the character bi-gram and then tri-gram probabilities to predict the detected language for each token, some rules like in ~\ref{sec:method} were followed which included that the span of a segment or a code-switched phrase is more than a word, the social media non-language entities such as the hashtags, usernames and urls do not make a difference.
 \\

### 2014-05-09 codeswitch-cltw-2014/codeswitch.tex
subsections+
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52734 72bbbca6-d526-0410-a7d9-f06f51895060
+
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52735 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-Alphabet n-gram approach -
+\subsection{Alphabet n-gram approach}
+\label{alphan}
 
@@
 
-Code-Switch Chunking - \\
+\subsection{Code-Switch Chunking}
+\label{cschunking}
 
@@
 
-Word based prediction -
+\subsection{Word based prediction}
+\label{wbased}
 
@@
 
-Word based Prediction with Character backoff -
+\subsection{Word based Prediction with Character backoff}
+\label{wbasedbackoff}
 
   A better way to predict the spans of the sub-segments of the text is to include the two methods of word and character based techniques as described above.
-In case of the word being present in only one of the two monolingual word lists the classification is simple, but in case of a confusion as described in Word based prediction.
+In case of the word being present in only one of the two monolingual word lists the classification is simple, but in case of a confusion as described in ~/ref{wbasedbackoff}.
 above, a character bigram backoff was introduced to help us disambiguate the language label.
@@
 
-Baseline \\
+\subsection{Baseline}
+\label{baseline}
 
@@
 
-Langid character bigram and trigram prediction - \\
+\subsection{Langid character bigram and trigram prediction}
+\label{langidstuff}
 

### 2014-05-09 codeswitch-cltw-2014/codeswitch.tex
formatting
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52736 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \cite{kneser1995improved} \\
-
 Once we get the probabilities, our task at hand is to take in the data and process it such that the  sub-segments of code-switched data are detected in it.
@@
 \\
-
 Since, there were instances of non-standard input in the Twitter data, certain terminologies were escaped, such as mentions (starting with ‘@’), hashtags (\#xyz), Re-Tweet (RT) and hyperlinks.\\
@@
 \label{wbased}
-
+\\
 This is a simple heuristic which was designed on the basis of the the most common words in the wordlists of the languages which are in question.
@@
 \label{wbasedbackoff}
-
+\\
 A better way to predict the spans of the sub-segments of the text is to include the two methods of word and character based techniques as described above.
@@
 \\
-
 Apart from the techniques discussed in ~/ref{sec:method}, some baselines are also used to give a comparative view of how well all the mechanism perform.
@@
 \label{sec:results}
-As described in the ~\ref{sec:method} the data collected from Twitter for the three language pairs, was processed using the techniques mentioned.
+As described in ~\ref{sec:method} the data collected from Twitter for the three language pairs, was processed using the techniques mentioned.
 The statics of the same are described in ~\ref{table:accuracy}.

### 2014-05-09 codeswitch-cltw-2014/codeswitch.tex
update
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52739 72bbbca6-d526-0410-a7d9-f06f51895060

--- 2014-05-09 codeswitch-cltw-2014/codeswitch.tex
@@
 before proceeding with further computational processing. This task has received a substantial amount of
-attention in recent years (\cite{cavnar1994n},\cite{lui2012langid}). However, previous research has on the whole assumed
+attention in recent years \cite{cavnar1994n,lui2012langid}. However, previous research has on the whole assumed
 that a given text will be in a single language. When dealing with text from formal domains,
@@
 code switching, quotations, named entities, interjections, translations, etc. Some of these are presented
-in Table~\ref{table:examples}
-
-\begin{table}
-\begin{tabular}{ll}
- \textbf{Code switching}: & You're a [Meirice\'{a}nach, c\'{e}n f\'{a}th] are you [foghlaim Gaeilge?!] \\
-\textbf{Quotations}: & And then he said ` ' \\
- \textbf{Named entities}: & [Dr Jekyll] ha [Mr Hyde] embannet gant [\'{E}ditions Aber] \\
- \textbf{Interjections}: & Hey, that's great, [diolch yn fawr!] \\
- \textbf{Translations}: & Bloavezh mat d'an holl ! [Bonne ann\'{e} \`{a} tous !] \\
-\end{tabular}
-\label{table:examples}
-\caption{Some examples of text segments containing more than one language}
-\end{table}
+in Table~\ref{table:examples}.
 
@@
 
+\begin{table}[h]
+\begin{tabular}{ll}
+ \textbf{Code switching}: & You're a [Meirice\'{a}nach, c\'{e}n f\'{a}th] are you [foghlaim Gaeilge?!] \\
+\textbf{Quotations}: & And then he said ` ' \\
+ \textbf{Named entities}: & [Dr Jekyll] ha [Mr Hyde] embannet gant [\'{E}ditions Aber] \\
+ \textbf{Interjections}: & Hey, that's great, [diolch yn fawr!] \\
+ \textbf{Translations}: & Bloavezh mat d'an holl ! [Bonne ann\'{e} \`{a} tous !] \\
+\end{tabular}
+\caption{Some examples of text segments containing more than one language}
+\label{table:examples}
+\end{table}
+
 \section{Related Work}
@@
 
-There has been significant research on code-switching and segment detection problems.
-The good thing about multilingual code switched data is that there are differences in the words of different languages - phonologically, morphologically, lexically and syntactically.
-At least that is the case with the pairs of languages we have chosen according to their usage in common day scenario.
-\\
+Code-switching and segment detection problems have been the subject of previous research. A good deal of work
+has been done on detecting code-switched segments in speech data \cite{chan2004detection,lyu2006language}.
+It is seen that language modelling techniques have shown promise earlier, such as in \cite{yu2013identification} 
+the experiment on Mandarin-Taiwanese code-switched sentences show a high accuracy in terms of detecting code-switched sentences. \\
 
-Before the Internet era, this study was focussed more on speech data.
-Initial research on this code switching was done for Marathi and English \cite{joshi1982processing} which included notable work about the formalisms and parsing.
-\\
+\newcite{chan2004detection} has made use of the bi-phone probabilities and calculated them to measure a confidence metric, 
+to \cite{lyu2006language} which has made use of a classification method, named syllable-based duration classification, 
+which uses the tonal syllable properties along with the speech signals to help predict the code switch points. \newcite{yeong2010language} 
+uses syllable structure information to identify words in code-switched text in Malay-English, however they did not 
+recognise segments in running text, only identifying individual words.
 
-It is seen that language modelling techniques have shown promise earlier, such as in \cite{yu2013identification} the experiment on Mandarin-Taiwanese code-switched sentences show a high accuracy in terms of detecting code-switched sentences.
-\\
-
-Our idea of going to character level inspection has been inspired both by the language modelling techniques used by the community as well as the acoustic modelling research.
-From \cite{chan2004detection} which has made use of the bi-phone probabilities and calculated them to measure a confidence metric, to \cite{lyu2006language} which has made use of a classification method, named syllable-based duration classification, which uses the tonal syllable properties along with the speech signals to help predict the code switch points.
-\cite{yeong2010language} uses syllable structure information, and increases our confidence is using the character n-gram approach for detection of the code switching points, their study was for the Malay-English, identifying the language of the words in the code switched data set.\\
+%Our idea of going to character level inspection has been inspired both by the language modelling techniques used by the community as well as the acoustic modelling research.
 
@@
 \label{sec:method}
-We use the character n-gram approach along with some heuristics to mainly define the first task at hand.
-We would like to both predict the code switched points but looking at the surrounding structure also decide the inclusion of them into the current or the next segment.
-\\
+We use the character n-gram approach along with some heuristics which are relevant to our problem domain of identifying
+segments for subsequent processing. We would like 
+to both predict the code switched points but looking at the surrounding structure also decide the inclusion of them 
+into the current or the next segment. \\
 
@@
 
-To predict code switch points we needed some raw text to train our character language models on.
-For English and French, we used the help of the Europarl \cite{koehn2005europarl}, for Breton, Welsh and Irish, data from credible sources on web were fetched.
-The size of the dataset for all the language sources was about 1.5million words.
-\\
-Building the character language models was undertaken using IRSTLM \cite{federico08a}, the input characters are to be introduced in the format of token separated by space and blank spaces replaced by `\_'.
-For example, `sl\'{a}inte' would be broken down into a set which is \{`\_ s', `s l', `l \'{a}', `\'{a} i', `i n', `n t', `t e', `e \_'\}.Then the probability for the alphabet sequence is calculated for each language pair, to be used in the experiment.
-Improved kneser-ney is used for smoothing, and this smoothing approach would be useful for higher-grams.
-\cite{kneser1995improved} \\
+We first built character language models using IRSTLM \cite{federico08a} for the five languages in question. For English
+and French a model was trained using the EuroParl corpus \cite{koehn2005europarl}. 
+For Breton, Welsh and Irish we used corpora of text crawled from the web. For each of the languages we sampled 1.5 million words. \\
+In order to build a character language model as opposed to the standard word-ngram based model we replaced spaces with the underscore 
+symbol `\_', and then placed a space chracter between each character. For example, `sl\'{a}inte' would be broken down into a sequence 
+of \{`\_ s', `s l', `l \'{a}', `\'{a} i', `i n', `n t', `t e', `e \_'\}. Then the probability for the alphabet sequence is calculated 
+for each language to be used in the experiment. 
+%Improved Kneser-Ney is used for smoothing, and this smoothing approach would be useful 
+%for higher-grams. \cite{kneser1995improved} \\
 
-Once we get the probabilities, our task at hand is to take in the data and process it such that the  sub-segments of code-switched data are detected in it.
-We collected data from twitter over a period of time \footnote{\url{http://indigenoustweets.blogspot.in/2013/12/mapping-celtic-twittersphere.html}} and this was filtered into three sets which had been identified to be Irish, Welsh and Breton.
-To be able to compare with the language model built above in the character n-gram approach, this data is to be processed like it aswell.
-\\
-
-Since, there were instances of non-standard input in the Twitter data, certain terminologies were escaped, such as mentions (starting with ‘@’), hashtags (\#xyz), Re-Tweet (RT) and hyperlinks.\\
-This processed data is then given as input to our module which uses heuristics to come up with sub-segment information. \\
-
-\subsection{Code-Switch Chunking}
+\subsection{Sequence chunking}
 \label{cschunking}
 
-This section describes the rules taken into account while performing the chunking of the input data according to the code-switch prediction. \\
+This section describes the heuristics taken into account while performing the chunking of the input data according to 
+the sequences of segments in different languages.
 
-A flag is set at the beginning, and this moves forward to the position in the text upto where confident chunking has been performed.
-For example, in case of a continuous prediction there would be no change and since the predicted tag of the new word corresponds to that already of the expanding chunk, this word is also included in the same chunk.
-In case of a doubt, the flag is set at the same position, and no final decision on labelling the word is performed, to have more confidence on the code-switch, the process moves on to the next token, if it corresponds to that of the previous chunk, then both these undecided words are labelled and included in the chunk, otherwise a change is noted and a new chunk is created with the changed label.
-In this process, we are trying to identify the sub-segments that we need, thus doing two tasks together and optimally, one is the prediction and second is the decision on chunking.
-\\
+A flag is set at the beginning of the input text, and this moves forward to the position in the text upto where confident 
+chunking has been performed. For example, in case of a continuous prediction there would be no change and since the 
+predicted tag of the new word corresponds to that already of the expanding chunk, this word is also included in the 
+same chunk. In case of a doubt, the flag is set at the same position, and no final decision on labelling the word is 
+performed, to have more confidence on the code-switch, the process moves on to the next token, if it corresponds to 
+that of the previous chunk, then both these undecided words are labelled and included in the chunk, otherwise a change 
+is noted and a new chunk is created with the changed label. In this process, we are trying to identify the sub-segments 
+that we need, thus performing both tasks at once, the first being language prediction and second being the chunking decision.
 
-\subsection{Word based prediction}
+\subsection{Word-based prediction}
 \label{wbased}
 
- This is a simple heuristic which was designed on the basis of the the most common words in the wordlists of the languages which are in question.
-After checking each word against one of the two word lists, it is associated to that particular language.
-In case of a confusion, for example, when the word exists in both the word lists or not in either, the option of continuing with the previous span was taken and the previous selected tag was labelled, thus increasing the chunk.
-\\
+This is a simple heuristic which was designed on the basis of the the most common words in the wordlists of the languages 
+which are in question. After checking each word against both of the word lists,\footnote{For the wordlists we used the \emph{aspell}
+wordlists widely available on Unix systems.} it is associated to one language or another. In 
+case of a conflict, for example, when the word exists in both wordlists, or in the case that it is unknown to both, the option 
+of continuing with the previous span was taken and the previous selected tag was labelled, thus increasing the chunk.
 
-\subsection{Word based Prediction with Character backoff}
+\subsection{Word-based prediction with character backoff}
 \label{wbasedbackoff}
 
-  A better way to predict the spans of the sub-segments of the text is to include the two methods of word and character based techniques as described above.
-In case of the word being present in only one of the two monolingual word lists the classification is simple, but in case of a confusion as described in ~/ref{wbasedbackoff}.
-above, a character bigram backoff was introduced to help us disambiguate the language label.
-This method works well because the earlier heuristic approach of just labelling the word with the label of the span which is expanding, would mean less code-switch detection and more shift towards baseline.
-\\
+A better way to predict the spans of the sub-segments of the text is to include the two methods of word and character-based techniques 
+as described above. In case of the word being present in only one of the two monolingual word lists the classification is simple, 
+but in case of a conflict, a character bigram backoff was introduced to help us disambiguate the language label.
+This method works well because the earlier heuristic approach of just labelling the word with the label of the span which is expanding, would mean less code-switch detection and more shift towards the majority class.
 
@@
 
-For the Evaluation procedure, we follow the footsteps of the CoNLL-2000 shared task: dividing text into syntactically related non-overlapping groups of words.
-This chunking mechanism \cite{tjong2003introduction} is very similar to ours, in terms of words which only belong to one category (here, language), and also evaluation based on the phrase structure present in the data.
-The chunks here are such that they belong to only one language.
-For example, the sentence below, shows the chunks as group of words in square brackets and their tagged language at the beginning in braces - \\
+We hand-annotated a small evaluation set from a selection of posts to a popular microblogging site\footnote{\url{http://indigenoustweets.blogspot.in/2013/12/mapping-celtic-twittersphere.html}}.
+The posts `tweets' were filtered into three sets which had been identified as Irish, Welsh and Breton.
+Certain tokens were escaped from the data, such as mentioned (which are preceded by an \texttt{@} symbol), subject tags `hashtags' which are preceded by a \texttt{\#} symbol), hyperlinks and the sequence \texttt{rt} which stands for `re-tweet'.
+An example of the content of our corpus after annotation is given in Figure~\ref{fig:tweets}.
+All of the tweets had at least one instance of code-switching.
+
+For the Evaluation procedure, we follow the footsteps of the CoNLL-2000 shared task on language-independent 
+named entity recognition: dividing text into syntactically related non-overlapping groups of words. This 
+chunking mechanism \cite{tjong2003introduction} is very similar to ours, in terms of words which only belong to 
+one category (here, language), and also evaluation based on the segment structure present in the data. The chunks 
+here are such that they belong to only one language.
 
@@
 \begin{alltt}
+[\textbf{en} You're a] [\textbf{ga} Meirice\'{a}nach, c\'{e}n f\'{a}th] [\textbf{en} are you] [\textbf{ga} foghlaim Gaeilge?!] 
 @steffdafydd [\textbf{cy} gorfod cael bach o tan] [\textbf{en} though init] 
@@
 \end{small}
+\caption{Example of text from a microblogging site chunked}
 \label{fig:tweets}
-\caption{Example of text from a microblogging site chunked}
 \end{figure}
 
-For our current task, each of the texts have been limited to two languages i.e.
-the primary and the secondary language.
-Hence the type of chunking tags are limited to these.
-The evaluation statics shown in the Table mention two values for each of the experiment conducted on the three bilingual language datasets.
-The first, is the percentage of correctly detected phrases, which is the overall precision and the second is the number of phrases in the data that were found by the chunker, which is the overall recall.
-\\
+For our current task, each of the texts have been limited to two languages i.e. the primary language (the Celtic language) and the 
+secondary language (the `majority' language). Hence the type of chunking tags are limited to these. The evaluation statistics shown 
+in Tables~\ref{table:precisionrecall}~and~\ref{table:accuracy} mention two values for each of the experiment conducted on the three bilingual 
+language datasets. The first, is the percentage of correctly detected phrases, which is the overall precision and the second is the 
+number of phrases in the data that were found by the chunker, which is the overall recall. \\
 
-Apart from the techniques discussed in ~/ref{sec:method}, some baselines are also used to give a comparative view of how well all the mechanism perform.
+Apart from the techniques discussed in section~\ref{sec:method}, some baselines are also used to give a comparative view of how well all the mechanism perform.
 
@@
 
- This is the most naive method of classification, we used the language identification tool langid.py \cite{lui2012langid} on the whole dataset and labelled all the individual lines according to this single major classification.
+This is the most na\"{i}ve method of classification, we used the language identification tool \texttt{langid.py} \cite{lui2012langid} on the whole dataset and labelled all the individual lines according to this single majority classification.
+As no chunking is performed we can expect that the precision and recall will be very low.
+However it does provide a reasonable baseline for the per-word accuracy.
 \\
 
-\subsection{Langid character bigram and trigram prediction}
+\subsection{\texttt{langid} character bigram and trigram prediction}
 \label{langidstuff}
 
-After restricting the predicted languages to be amongst the two in the dataset, we used the character bi-gram and then tri-gram probabilities to predict the detected language for each token, some rules like in ~\ref{sec:method} were followed which included that the span of a segment or a code-switched phrase is more than a word, the social media non-language entities such as the hashtags, usernames and urls do not make a difference.
-\\
+After restricting the predicted languages to be amongst the two in the dataset, we used the character bigram and then trigram probabilities to predict the detected language for each token, some rules like in section~\ref{sec:method} were followed which included that the span of a segment or a code-switched phrase is more than a word, the social media non-language entities such as the hashtags, usernames and urls do not make a difference.
 
@@
 As described in the ~\ref{sec:method} the data collected from Twitter for the three language pairs, was processed using the techniques mentioned.
-The statics of the same are described in ~\ref{table:accuracy}.
+The statistics of the same are given in Table~\ref{table:datastats}.
 //
@@
 \end{center}
-\label{table:accuracy}
-\caption{Document Statistics of the annotated data used.}
+\caption{Document statistics of the annotated data used. }
+\label{table:datastats}
 \end{table}
@@
 \end{center}
+\caption{Precision, $p$ and recall, $r$ for the systems by language.}
 \label{table:precisionrecall}
-\caption{Precision, $p$ and recall, $r$ for the systems by language.}
 
@@
 \end{center}
-\label{table:accuracy}
 \caption{Accuracy of the systems over the three language pairs. The accuracy measures how often a token
   was assigned to the right language, independent of span.}
+\label{table:accuracy}
 \end{table}
 
-The Results clearly show that the baseline has a negligible precision and recall, this means that the task of detecting sub-segments is complex and there needs to be more improvement in the technique.
-You can see that langid.py is better suited for document classification and that when such varied content is introduced the result is an improvement but not promising.
-In case of the character bigram prediction technique and the word based chunking method, the results show a significant increase in precision and recall values from the rest.
-The last in the list is the word based prediction with character backoff, which performs collectively better than the others with a high accurracy margin.
-\\
+In terms of per-word accuracy, the baseline performs as expected, given that half of the words in the test set are from a given language. 
+The \texttt{langid.py} based methods allow for segmenting the text into chunks, but favour the majority language.\footnote{This could be because of the default models being used.} While the precision and recall are low for the remaining models, we see that we are able to improve
+the performance by combining the wordlist based model with a character bigram model. And what is more, we are able to begin to not only 
+identify particular words in a language, but also segments.
 
@@
 \label{sec:conclusions}
+
+This paper has presented a very preliminary investigation into subsegment language identification in Celtic language texts. We have
+proposed a model that chunks input text into segments, and performs language identification on these segments at the same time. Precision and
+recall are low, leaving a lot of room for further work. Some of our plans are as follows. We would like to annotate more test data,
+at least 100 tweets in each language. We would also like to attempt our method using higher order character n-gram models for backoff, and
+n-gram word language models for detection.
 
@@
 
-We thank Kevin Scannell...
+We thank Kevin Scannell for help with providing data from Twitter and for useful comments and suggestions.
 

### 2014-05-09 codeswitch-cltw-2014/codeswitch.tex
update
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52740 72bbbca6-d526-0410-a7d9-f06f51895060
update
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52741 72bbbca6-d526-0410-a7d9-f06f51895060
update
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52742 72bbbca6-d526-0410-a7d9-f06f51895060
+1
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52743 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \begin{table}[h]
+\begin{center}
 \begin{tabular}{ll}
  \textbf{Code switching}: & You're a [Meirice\'{a}nach, c\'{e}n f\'{a}th] are you [foghlaim Gaeilge?!] \\
-\textbf{Quotations}: & And then he said ` ' \\
+\textbf{Quotations}: & The anthem starts with the words [`Mae hen wlad fy nhadau\ldots'] \\
  \textbf{Named entities}: & [Dr Jekyll] ha [Mr Hyde] embannet gant [\'{E}ditions Aber] \\
@@
 \label{table:examples}
+\end{center}
 \end{table}
@@
 has been done on detecting code-switched segments in speech data \cite{chan2004detection,lyu2006language}.
-It is seen that language modelling techniques have shown promise earlier, such as in \cite{yu2013identification} 
+It is seen that language modelling techniques have shown promise earlier, such as in \newcite{yu2013identification} 
 the experiment on Mandarin-Taiwanese code-switched sentences show a high accuracy in terms of detecting code-switched sentences. \\

### 2014-05-09 codeswitch-cltw-2014/codeswitch.tex
abstract
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52745 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \begin{abstract}
-  This paper describes an experiment to detect the language of subsegments
-  of text in three Celtic languages: Breton, Irish and Welsh. 
+  Bilingual communities contribute majorly to the code-switching phenomena, where speakers produce sentences containing words and expressions from a second language.
+With the increase in the data on the web, specially on informal platforms like that of the social media, it becomes increasingly important to detect such irregularities and handle them appropriately.
+In this paper we propose ways of dealing with code-switching and detecting sub-segments for texts in three Celtic languages: Breton, Irish and Welsh.
 \end{abstract}

### 2014-05-09 codeswitch-cltw-2014/codeswitch.tex
anonymise
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@52746 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-\author{Akshay Minocha \\
-  IIIT Hyderabad  \\
-  Hyderabad (India) \\
-  {\small {\tt akshay.minocha@students.iiit.ac.in}} \\\And
-  Francis M. Tyers \\
-  Giellatekno / CLEAR \\
-  UiT Norgga \'arktala\v{s} universitehta  \\
-  9017 Romsa (Norway) \\
-  {\small {\tt francis.tyers@uit.no}} \\}
+\author{}
+%\author{Akshay Minocha \\
+%  IIIT Hyderabad  \\
+%  Hyderabad (India) \\
+%  {\small {\tt akshay.minocha@students.iiit.ac.in}} \\\And
+%  Francis M. Tyers \\
+%  Giellatekno / CLEAR \\
+%  UiT Norgga \'arktala\v{s} universitehta  \\
+%  9017 Romsa (Norway) \\
+%  {\small {\tt francis.tyers@uit.no}} \\}
 
@@
 to both predict the code switched points but looking at the surrounding structure also decide the inclusion of them 
-into the current or the next segment. \\
+into the current or the next segment.\footnote{The software used in this experiment is available online at: \url{removed for review}.}
 
@@
 
-We thank Kevin Scannell for help with providing data from Twitter and for useful comments and suggestions.
+%We thank Kevin Scannell for help with providing data from Twitter and for useful comments and suggestions.
 

### 2014-05-17 codeswitch-cltw-2014/codeswitch.tex
a bnuch of poster updates
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@53027 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 [\textbf{en} You're a] [\textbf{ga} Meirice\'{a}nach, c\'{e}n f\'{a}th] [\textbf{en} are you] [\textbf{ga} foghlaim Gaeilge?!] 
-@steffdafydd [\textbf{cy} gorfod cael bach o tan] [\textbf{en} though init] 
+@afaltomkins [\textbf{cy} gorfod cael bach o tan] [\textbf{en} though init] 
 [\textbf{en} omg] [\textbf{cy} mar cwn bach yn] [\textbf{en} black and tan] [\textbf{cy} a popeth,] [\textbf{en} even cuter!!] 

### 2014-05-21 kipchak-lrec-2013/kipchak-paper.tex
added missing examples
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@53322 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-In Kazakh, the vowels /{\qipa ɘ}/ ‹і› and /ə/ ‹ы› followed by a glide /w/ ‹у› or /j/ ‹й› are written ‹у› and ‹и›, depending on the glide; for example.
+In Kazakh, the vowels /{\qipa ɘ}/ ‹і› and /ə/ ‹ы› followed by a glide /w/ ‹у› or /j/ ‹й› are written ‹у› and ‹и›, depending on the glide; e.g.\ ‹қи› /qəj/ `chop down', ‹ки› /k{\qipa ɘ}j/ `get dressed'.
 The character ‹ю› is used in turn to represent ‹й› followed by either multi-phoneme value of ‹у›.
 For example, ‹қиюда› /qəjəwd{\qipa ɑ}/ `in the process of chopping down' and ‹киюде› /k{\qipa ɘ}j{\qipa ɘ}wd{\qipa i̯ɘ}/ `in the process of getting dressed' have back- and front-harmonising values, respectively, for ‹и› and ‹ю›.\footnote{Note also that the orthography suggests the presence of an additional /j/ in these words.}
+%би /b{\qipa ɘ}j/ `dance', қидым /qəjdəm/ `I chopped', су /səw/ `water', білу /b{\qipa ɘ}l{\qipa ɘ}w/ `knowing'
 

### 2014-06-11 kypchak-morphfest-2014/abstract.tex
beginnings of morphfest poster
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@54153 72bbbca6-d526-0410-a7d9-f06f51895060

\documentclass[letterpaper,11pt,onecolumn]{article}

%\usepackage{lrec2012}

\usepackage{polyglossia}
\setdefaultlanguage[variant=australian]{english}

\usepackage{fontspec}
\usepackage{xunicode}
\usepackage{xltxtra}
\usepackage{natbib}
\usepackage{multirow}
%\usepackage{fullpage}
\usepackage{multirow}
\usepackage[colorlinks=true,citecolor=black,linkcolor=black,urlcolor=blue]{hyperref}

\usepackage{booktabs}
\usepackage[in]{fullpage}
%\usepackage{jnwmaps}

\defaultfontfeatures{Mapping=tex-text}
\setromanfont{Times New Roman}
\newfontfamily\symbl[Scale=MatchLowercase]{FreeSerif}

\newcommand{\tags}[1]{\texttt{#1}}

\bibdata{kipchak-paper}

\begin{document}
\thispagestyle{empty}

\begin{center}
	\textbf{Designing finite-state morphological transducers for Kypchak languages}
\end{center}

This paper describes the development of free/open-source finite-state morphological transducers for several Northwestern Turkic (or Kypchak) languages.  The transducers function to provide morphological analyses of surface forms as well as produce surface forms from morphological information, and can be repurposed as spell-checkers for these under-resourced languages.  The finite-state toolkit used for the work is the Helsinki Finite-State Toolkit (HFST) \citep{hfst/2011}.\\

Work is ongoing on morphological transducers for seven Kypchak languages: three Southern Kypchak languages (Kazakh, Karakalpak, and Nogay), two Northern Kypchak languages (Tatar and Bashqort), one Western Kypchak language (Kumyk), and Kyrgyz.  This paper describes how the development of a transducer for each subsequent language took less time than the previous one.\\

An evaluation is presented for the transducer of one language from each of the four branches of Kypchak (Kazakh, Tatar, Kumyk, and Kyrgyz).  Each transducer has production-level coverage---around 90\%---on freely available corpora of the languages, and high precision (the number of the analyses given for a form that are correct) and recall (the percentage of possible correct analyses for a form that are provided by the transducer) over a manually verified test set.\\

The computational implementation of the morphotactics and morphophonology of the languages is described.  Emphasis is placed on the types of differences that exist between the morphologies of these languages, as well as on previously undocumented morphological patterns in these languages that were discovered through the development of the transducers.

\bibliographystyle{lrec2012}
\bibliography{kipchak-paper}

\end{document}

### 2014-06-12 kypchak-morphfest-2014/poster.tex
some initial work on the poster: rotated, sized (but need to check whether it will fit on the poster displays), added some Kyrgyz stuff, work on header, etc.
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@54160 72bbbca6-d526-0410-a7d9-f06f51895060

\documentclass[fontscale=0.3,landscape,paperwidth=72in,paperheight=36in]{baposter}  % fontscale=0.285, dvipdfm

\usepackage{setspace}
\usepackage{multicol}
\usepackage{multirow}

%\usepackage{tikz}
%\usepackage{pgfbaselayers}
%\pgfdeclarelayer{background}
%\pgfdeclarelayer{foreground}
%\pgfsetlayers{background,main,foreground}

%%% Color Definitions %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\definecolor{bordercol}{RGB}{40,40,40}
%\definecolor{bordercol}{RGB}{100,0,0}
\definecolor{bordercol}{RGB}{0,0,0}
%\definecolor{headercol1}{RGB}{186,215,230}
%\definecolor{headercol2}{RGB}{80,80,80}
%\definecolor{headercol1}{RGB}{153,0,0}
%\definecolor{headercol2}{RGB}{153,0,0}
\definecolor{headercolone}{RGB}{6,130,22}
\definecolor{headercoltwo}{RGB}{6,130,22}
%\definecolor{boxcolor}{RGB}{186,215,230}
%\definecolor{boxcolor}{RGB}{110,230,209}
%\definecolor{boxcolor}{RGB}{143,255,236}
%\definecolor{boxcolor}{RGB}{171,255,241}
\definecolor{boxcolor}{RGB}{176,232,223}
\definecolor{headerfontcol}{RGB}{255,255,255}
%\definecolor{boxcolor}{RGB}{186,215,230}
%\definecolor{boxcolor}{RGB}{226,200,200}
%\newcommand{\hilitetwo}[1]{{\addfontfeature{Color=FF000099}#1}}
\newcommand{\hilitetwo}[1]{{\addfontfeature{Color=99333399}#1}}
%\newcommand{\hiliteone}[1]{{\addfontfeature{Color=33993399}#1}}
\newcommand{\hiliteone}[1]{{\addfontfeature{Color=06821699}#1}}
\newcommand{\hilitegrey}[1]{{\addfontfeature{Color=77777799}#1}}

%%% Utility functions %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% Save space in lists. Use this after the opening of the list %%%%%%%%%%%%%%%%
\newcommand{\compresslist}{
        \setlength{\itemsep}{1pt}
        \setlength{\parskip}{0pt}
        \setlength{\parsep}{0pt}
}

\usepackage{polyglossia}
\setdefaultlanguage[variant=australian]{english}

\usepackage{expex}

\usepackage{fontspec}
\defaultfontfeatures{PunctuationSpace=3,Scale=MatchLowercase,Mapping=tex-text}
\newfontfeature{IPA}{+mgrk}
%\setromanfont[IPA]{FreeSerif}
%\setromanfont[IPA,Scale=0.8]{CMU Serif}
\setromanfont[IPA]{Liberation Serif}
%\setromanfont[IPA]{Times New Roman}
%\setromanfont[IPA]{NimbusRomNo9L-Regu}
%\setromanfont{Times New Roman}
%\setmonofont[IPA]{Liberation Mono}
\setmonofont[IPA]{DejaVu Sans Mono}
%\renewcommand{\sfdefault}{phv}
%\renewcommand{\rmdefault}{ptm}
%\renewcommand{\ttdefault}{pcr}
\usepackage[small,bf]{caption}
%\newfontfamily\qipa[IPA]{NimbusRomNo9L-Regu}
\newfontfamily\qipa[IPA,Scale=MatchLowercase]{FreeSerif}
\newfontfamily\qgmk[IPA,Scale=0.8]{CMU Serif}
\newfontfamily\litamono[IPA,Scale=0.5]{DejaVu Sans Mono}

\newcommand{\tilda}{{\qipa ∼}}

\newcommand{\tags}[1]{\hiliteone{#1}}
\newcommand{\tag}[1]{\hilitegrey{<}\hiliteone{#1}\hilitegrey{>}}

%\newfontfamily\htwo[IPA,Scale=1.2]{FreeSans}}
%\newfontfamily\htwofont[IPA,Scale=1]{CMU Sans Serif}
%\newfontfamily\htwofont[IPA,Scale=1,Color=333344FF]{CMU Sans Serif}
\newfontfamily\htwofont[IPA,Scale=1,Color=111111FF]{CMU Sans Serif}
%\newfontfamily\titlefont[IPA,Scale=0.52]{CMU Serif}
\newfontfamily\titlefont[IPA,Scale=0.7]{CMU Serif}
%\newfontfamily\titlefont[IPA,Scale=0.55]{DejaVu Serif}

%\newcommand{\htwo}[1]{::: {\htwofont #1} :::}%\hrule} %\hline\\}
%\newcommand{\htwo}[1]{{\htwofont \textbf{:::#1:::}}}%\hrule} %\hline\\}
\newcommand{\htwo}[1]{{\htwofont \textbf{\dotfill{}#1\dotfill{}}}}

\usepackage{graphicx}  % [dvipdfm]

%\definecolor{MyGray}{rgb}{0.96,0.97,0.98}
%\newenvironment{codebox}{%
%   \begin{lrbox}{\@tempboxa}\begin{minipage}{\columnwidth}}{\end{minipage}\end{lrbox}%
%   \colorbox{MyGray}{\usebox{\@tempboxa}}
%}
%\newcommand{\codeex}[1]{\begin{codebox}#1\end{codebox}}

%\definecolor{grey}{rgb}{0.96,0.97,0.98}
\definecolor{grey}{rgb}{0.91,0.91,0.91}
\newcommand{\codeex}[1]{
   \fbox{\colorbox{grey}{
         \begin{minipage}[t]{0.91\textwidth}{\litamono
            #1
         }\end{minipage}
      }
   }
}

\newcommand{\blank}[1]{\underline{\hspace{#1}}}
% FIXME: Breaks baposter
%\usepackage[novoc,fdf2alif]{arabxetex}
%\newfontfamily\uighurfont[Script=Arabic,Scale=1.5]{Lateef}
%\newfontfamily\arabicfont[Script=Arabic,Scale=1.5]{Lateef}

\usepackage{natbib}

%\makeatletter
%\renewenvironment{thebibliography}[1][]{\htwo{\bibname}}
%     {\section*{\bibname}% <-- this line was changed from \chapter* to \section*
%	  \htwo{\bibname}
%      \@mkboth{\MakeUppercase\bibname}{\MakeUppercase\bibname}%
%      \list{\@biblabel{\@arabic\c@enumiv}}%
%           {\settowidth\labelwidth{\@biblabel{#1}}%
%            \leftmargin\labelwidth
%            \advance\leftmargin\labelsep
%            \@openbib@code
%            \usecounter{enumiv}%
%            \let\p@enumiv\@empty
%            \renewcommand\theenumiv{\@arabic\c@enumiv}}%
%      \sloppy
%      \clubpenalty4000
%      \@clubpenalty \clubpenalty
%      \widowpenalty4000%
%      \sfcode`\.\@m}
%     {\def\@noitemerr
%       {\@latex@warning{Empty `thebibliography' environment}}%
%      \endlist}%}
%}
%\makeatother
%\usepackage[in]{fullpage}
\usepackage[colorlinks=true,citecolor=black,linkcolor=black,urlcolor=black]{hyperref}

\usepackage{subfigure}
\usepackage{booktabs}

%%\bibpunct{(}{)}{;}{A}{,}{,}
%\bibdata{paper}

\newcommand{\citemultileft}[1]{(\citeauthor{#1}, \citeyear{#1}}
\newcommand{\citemultimid}[1]{\citeauthor{#1}, \citeyear{#1}}
\newcommand{\citemultiright}[1]{\citeauthor{#1}, \citeyear{#1})}
\newcommand{\citetwoyears}[2]{\citeauthor{#1} (\citeyear{#1} and \citeyear{#2})}

% for glosses
\newcommand{\eng}[1]{`{\em #1}'}
%dammit, sc doesn't seem to be working
\newcommand{\gmk}[1]{{\qgmk \textsc{#1}}}

\usepackage{enumitem}
\setlist{nolistsep,leftmargin=*}
\newenvironment{itemise}[1]{
        \begin{itemize}\setlength{\leftmargin}{-4em}\setlength{\itemsep}{-0.2em}
        \vspace{-0.5em}
        #1
}{
        \end{itemize}
        \vspace{-2pt}
}

%\newcommand{\h2}[1]{{\big

\begin{document}
	% To get it to be A0 consistently on all machines..
	%\special{papersize=1189mm,841mm}
	%\special{papersize=841mm,1189mm}
	\special{papersize=72in,36in}
	\setlength{\pdfpageheight}{\paperheight}
	\setlength{\pdfpagewidth}{\paperwidth}

%%% Setting Background Image %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\background{
%	\includegraphics[width=0.99\textwidth]{flagkg2}
%}
	\background{{
%		\begin{tikzpicture}[remember picture,overlay]%
%			\draw (current page.north west)+(-4em,2em) node[anchor=north west] {\includegraphics[width=1.1\textwidth,height=1.1\textheight]{flagkg2}};
%		\end{tikzpicture}%
%%		\includegraphics[width=0.99\textwidth]{flagkg2}
%%			\draw (current page.north west)+(-2em,2em) node[anchor=north west] {\includegraphics[width=0.9\textwidth]{flagkg2}};
	}}

	\begin{poster}{
			grid=false,
			%eyecatcher=false,
			borderColor=bordercol,
			headerColorOne=headercolone,
			headerColorTwo=headercoltwo,
			headerFontColor=white,
			% Only simple background color used, no shading, so boxColorTwo isn't necessary
			boxColorOne=boxcolor,
			headershape=roundedright,
  headerborder=open,
  headerheight=0.1\textheight,
  %headershape=roundedright,
  %headershade=plain,
  %headerfont=\Large\textsf, %Sans Serif
			%headerfont=\Large\sf\bf,
			textborder=rectangle,
			%background=plain,
			%background=user,
			background=none,
			headerborder=open,
			boxshade=plain,
			textborder=roundedleft,
			columns=6,
		}{
			%\hspace{-2em}\includegraphics[height=6.5em,bb=0 0 242 203]{apertium}
			%\hspace{-2em}\includegraphics[height=6.5em,bb=0 0 612 792]{apertium2}
			%\hspace{-2em}\includegraphics[height=6.5em,bb=0 0 203 242]{apertium3}
			%\hspace{-2em}\includegraphics[height=6.5em,bb=0 0 242 203]{apertium3a}
			%\hspace{-2em}\includegraphics[height=6.5em,bb=0 0 242 203]{apertium4}
			%\hspace{-2em}
			\includegraphics[angle=90,height=8em]{apertium5a}%\hspace{-2em}
			%Eye Catcher, empty if option eyecatcher=false - unused
			%\begin{minipage}[t]{0.2\textwidth}
				\hspace{1cm}\includegraphics[height=6.5em]{iu_tab}
				\hspace{1cm}\includegraphics[height=6.5em]{kfu}
				\hspace{1cm}\includegraphics[height=6.5em]{uitlogo}
			%\end{minipage}
		}{
			{\vspace{0pt}\hspace{1cm}
			{\titlefont \textsc{Designing finite-state morphological transducers\\for Kypchak languages}}}
		}{
			%Jonathan North Washington {\small (Indiana University, \texttt{jonwashi@indiana.edu})}, Mirlan Ipasov {\small (International Ataturk Alatoo University, \texttt{mipasov@gmail.com})}, Francis M.\ Tyers {\small (Universitat d'Alacant, \texttt{ftyers@dlsi.ua.es})}

			\vspace{-0.5em}
			\begin{center}
			{\begin{minipage}[t]{12em}
				\begin{spacing}{0.4}
					{Jonathan North Washington}\\
					{\footnotesize Indiana University\\\texttt{jonwashi@indiana.edu}}
				\end{spacing}
			\end{minipage}
			\begin{minipage}[t]{11em}
				\begin{spacing}{0.4}
					{Ilnar Salimzyanov}\\
					{\footnotesize Казан (Идел буе) федераль университеты \\\texttt{ilnar.salimzyan@gmail.com}}
				\end{spacing}
			\end{minipage}
			\begin{minipage}[t]{9em}
				\begin{spacing}{0.4}
					{Francis M.\ Tyers}\\
					{\footnotesize UiT Norgga Árktalaš Universitehta \\\texttt{francis.tyers@uit.no}}
				\end{spacing}
			\end{minipage}}
			\begin{minipage}[t]{7em}
				\vspace{-0.7em}
				\begin{spacing}{0.4}
					{\footnotesize Special thanks to:}\\
					{\small Tolgonay Kubatova}\\
					{\small Aida Sundetova}\\
					%{\footnotesize \texttt{sun27aida@gmail.com}}
					%{\small Ağarahim Sultanmuradov}\\
					{\small Ağarahim Sultanmuradov}\\
					%{\small Google Summer of Code}\\
					%{\small Google Code-In}
					%{\small GSoC, GCI}
				\end{spacing}
			\end{minipage}
			\end{center}

			%\begin{minipage}[c]{12em}
			%	\centering
			%	\begin{spacing}{0.4}
			%		Jonathan Washington\\
			%		Indiana University\\
			%		\texttt{jonwashi@indiana.edu}
			%	\end{spacing}
			%\end{minipage}
			%\begin{minipage}[c]{18em}
			%	\centering
			%	\begin{spacing}{0.4}
			%		Mirlan Ipasov\\
			%		International Ataturk Alatoo University\\
			%		\texttt{mipasov@gmail.com}
			%	\end{spacing}
			%\end{minipage}
			%\begin{minipage}[c]{10em}
			%	\centering
			%	\begin{spacing}{0.4}
			%		Francis M.\ Tyers\\
			%		Universitat d'Alacant\\
			%		\texttt{ftyers@dlsi.ua.es}
			%	\end{spacing}
			%\end{minipage}

			%Jonathan Washington, Mirlan Ipasov, Francis M.\ Tyers\\
			%Indiana University, Universitat d'Alacant, International Ataturk Alatoo University\\
			%\texttt{jonwashi@indiana.edu}, \texttt{mipasov@gmail.com}, \texttt{ftyers@dlsi.ua.es}
		}{
			%University Logo(s)
			%{\begin{minipage}{19em}
				%\hfill
				%\includegraphics[height=6.5em]{apertium}
				%\includegraphics[height=6.5em]{flags/flags}
				\includegraphics[height=6.5em]{Flag_of_Dagestan}
				\hspace{1cm}\includegraphics[height=6.5em]{Flag_of_Tatarstan}
				\hspace{1cm}\includegraphics[height=6.5em]{Flag_of_Kyrgyzstan}
				\hspace{1cm}\includegraphics[height=6.5em]{Flag_of_Kazakhstan}
			%\end{minipage}}
		}

%		\headerbox{Overview}{name=overview,column=0,row=1}{
%			\begin{spacing}{0.8}
%This paper describes the development of a free/open-source finite-state morphological transducer for Kyrgyz. The transducer 
%has been developed for morphological generation for use within a prototype Turkish$\rightarrow$Kyrgyz machine translation system, but has also been extensively tested for analysis. The finite-state toolkit used for the work was the Helsinki Finite-State 
%Toolkit (HFST). The paper describes some issues in Kyrgyz morphology, the development of the tool, some linguistic issues encountered and how they were dealt with, and 
%which issues are left to resolve. An evaluation is presented which shows that the transducer has medium-level coverage, between 82\% and 87\% on two freely available corpora of Kyrgyz, and high precision and recall over a manually verified test set.
%			\end{spacing}\vspace{-0.25ex}
%		}

	\headerbox{Kypchak languages}{name=kypchaklgs,column=0,row=0}{
		\vspace{0.5ex}
		\includegraphics[width=\textwidth]{map}
		\vspace{-1.1em}  % -1.25em
		\begin{itemize}
			\item Turkic languages (SOV, agglutinative, vowel harmony)
		\end{itemize}

	\scalebox{0.85}{
		\begin{tabular}{lllll}
			\toprule
				 & \textbf{Kyrgyz} & \textbf{Kazakh} & \textbf{Tatar} & \textbf{Kumyk} \\
				 & /{\qipa qɯɾʁɯz}/ & /{\qipa qɑzɑq}/ & /{\qipa tɒtɑɾ}/ & /{\qipa qumuq}/ \\
				classification & Eastern & Southern & Northern & Western \\
			\midrule
				\multicolumn{4}{l}{population of speakers} \\
			\midrule
				number & 3M & 8M-12M & 5.4M & 430K \\
				primary & Kyrgyzstan & Kazakhstan & Tatarstan & Dagestan \\
				secondary & China, etc. & China, Mongolia & Bashqortostan & --- \\
			\midrule
				\multicolumn{4}{l}{external influences} \\
			\midrule
				Mongolic & moderate & moderate & light & light \\
				Oghuz & — & — & light & moderate \\
				Persian & heavy & heavy & heavy & heavy \\
				Russian & heavy & heavy & heavy & heavy \\
			\bottomrule
		\end{tabular}
	}
%			\begin{itemize}
%				\item Similar historically to Southern Altay
%				\item Similar by convergence to Kazakh, Uzbek
%			\end{itemize}
%			\item Spoken in
%			\begin{itemize}
%				\item Kyrgyzstan, as co-official language\\
%				\textit{high levels of bilingualism with Russian}
%				\item China, Tajikistan, Uzbekistan
%			\end{itemize}
%			\item Over 3 million native speakers\\(estimate based on data from Ethnologue) %\citep{lewis2009}
%			\item Our transducer based on written Kyrgyz of the former Soviet Union (literary and colloquial standards)
%		\end{itemize}
	}

	\headerbox{Morphological transducers}{name=morphtrans,below=kypchaklgs}{
		\htwo{Morphological transducers}
		\begin{itemize}
			\item Take a surface form, and produce valid lexical form(s)
			\item Take a lexical form, and produce valid surface form(s)
		\end{itemize}
		\scalebox{0.96}{
			`алдым' \hilitetwo{↔} \texttt{ал{\small <v><tv><ifi><p1><sg>}}, \texttt{алд{\small <n><px1sg><nom>}}
		}

		\htwo{Transducers for Turkic languages}
		\begin{itemize}
			\item Turkish (\hiliteone{Çöltekin, 2010 \& 2014}; \hilitetwo{Oflazer, 1994})
			\item Crimean Tatar (\hilitetwo{Altıntaş, 2001})
			\item Turkmen (\hilitetwo{Tantuğ et al., 2006})
			\item Kyrgyz (\hiliteone{Washington et al., 2012})%\\
			\item Kazakh (\hilitetwo{Бекманова \& Махимов, 2013})
			%\vspace{1pt}\hfill{} \hiliteone{GPL (=free and open)}!
			\item our Kazakh, Tatar, Kumyk: all \hiliteone{GPL (=free and open)}!
		\end{itemize}

		\htwo{Framework: HFST}
		\begin{itemize}
			\item Reimplements Xerox FST formalisms ({\tt lexc} \& {\tt twol})
			\item Also provides a wrapper around popular free/open-source FST toolkits: SFST, OpenFST, and Foma
		\end{itemize}
	}
	\headerbox{Development effort}{name=devel,below=morphtrans}{
		\begin{itemize}
			\item Kumyk transducer based on Kazakh, Tatar transducers
			\item \tilda{}1 week to reach 80\% coverage, +1 week to reach 90\%
		\end{itemize}
		\vspace*{-2.5em}
		\hspace*{-2em}\includegraphics[width=0.9\textwidth,angle=270]{../kipchak-lrec-2013/graphs/kumyk-yoldash-cov}
		\vspace*{-2em}
	}

%		\headerbox{Available Turkic Transducers}{name=turkictrans,below=morphtrans}{
%* Other Turkic morphological transducers\\
%		}
%		\headerbox{Framework: HFST / Apertium}{name=framework,below=turkictrans}{
%* HFST/Apertium
%		}

%			\htwo{Morphological \& orthographical words}
%
%			\begin{itemize}
%				\item{} {\qipa өнүктүрөбүзбү ?} \eng{will we develop [it]?}\\
%				\texttt{ өнүк{\small <v><tv><caus><aor><p1><pl>+}бы{\small <qst>}}
%				\item{} {\qipa келатсаң} \eng{if you come}\\
%				\texttt{кел{\tt {\small <v><iv><prt\_impf>}}+жат{\small <vaux><gna\_cnd><p2><sg>}}
%			\end{itemize}
%
%
%			\codeex{
%			{\small
%			\texttt{LEXICON N-INFL-3PX-COMPOUND\\
%			\vspace{-0.5ex}\%<n\%>:\%>\%\{S\%\}\%\{I\%\}\%\{n\%\} GEN-POS ;}\\
%			
%			\texttt{LEXICON Nouns\\
%			\vspace{-0.5ex}аба\%\ ырайы:аба\%\ ырай\ N-INFL-3PX-COMPOUND ;\\\vspace{-0.5ex}\hspace{0.1pt}\hfill{}!\ "weather"\\
%			\vspace{-0.5ex}чакыруу\%\ кагазы:чакыруу\%\ кагаз\ N-INFL-3PX-\\\vspace{-0.5ex}\hspace{0.1pt}\hfill{}COMPOUND ;\ !\ "invitation"}}
%			}
%				
%		}

					\renewcommand{\arraystretch}{0.65}

		\headerbox{Example output}{name=exout,span=2,column=1,row=0}{

		\htwo{Gloss}

		\pex[aboveexskip=0pt,belowexskip=3pt,aboveglftskip=0pt]\label{ex:sent}
			\begingl
				\gla[everygla={\qipa }] Кудай Өзү жаратканынын баарына карап, {} өтө жакшы экенин көрдү. //
				\glb[everyglb={\qipa }] Құдай Өзінің жаратқандарының бәріне қарап, {} өте жақсы екенін көрді. //
				\glb[everyglb={\qipa }] Аллаһ Үзе яраткан нәрсәләргә карап, аларның бик яхшы икәнен күрде. //
				\glb[everyglb={\qipa }] Аллагь Оьзю яратгъан затлагъа къарап, олар бек яхшы экенин гёрген. //
				\glb God own-his created [everything/thing-s]-to looked.at, they/their very good being saw. // 
				\glft \eng{God looked at everything he had created and saw that it was very good. (Bible, Genesis 1:31)} //
			\endgl
		\xe

		\htwo{Output}
\scalebox{0.81}{
\begin{tabular}{llll}
\toprule
 \textbf{Kyrgyz} (\textbf{kir}) & \textbf{Kazakh} (\textbf{kaz}) & \textbf{Tatar} (\textbf{tat}) & \textbf{Kumyk} (\textbf{kum}) \\
\midrule
\texttt{Кудай Өзү жаратканынын баарына} & \texttt{Құдай Өзінің жаратқандарының бәріне} & \texttt{Аллаһ Үзе яраткан нәрсәләргә карап,} & \texttt{Аллагь Оьзю яратгъан затлагъа} \\
\texttt{карап, өтө жакшы экенин көрдү.} & \texttt{қарап, өте жақсы екенін көрді.} & \texttt{аларның бик яхшы икәнен күрде.} & \texttt{къарап, олар бек яхшы экенин гёрген.} \\
\midrule
	\texttt{Кудай\tag{n}\tag{nom}} & \texttt{Құдай\tag{n}\tag{nom}} & \texttt{Аллаһ\tag{n}\tag{nom}} & \texttt{Аллагь\tag{n}\tag{nom}} \\
	\texttt{Өз\tag{prn}\tag{ref}\tag{px3sp}\tag{nom}} & \texttt{Өз\tag{prn}\tag{ref}\tag{px3sp}\tag{gen}} & \texttt{Үз\tag{prn}\tag{ref}\tag{px3sp}\tag{nom}} & \texttt{Оьз\tag{prn}\tag{ref}\tag{px3sp}\tag{nom}} \\
	\texttt{жарат\tag{v}\tag{tv}\tag{ger\_past}\tag{px3sp}\tag{gen}} & \texttt{жарат\tag{v}\tag{tv}\tag{ger\_past}\tag{pl}\tag{px3sp}\tag{gen}} & \texttt{ярат\tag{v}\tag{tv}\tag{gpr\_past}} & \texttt{ярат\tag{v}\tag{tv}\tag{gpr\_past}} \\
	\texttt{баары\tag{prn}\tag{qnt}\tag{px3sp}\tag{dat}} & \texttt{бәрі\tag{prn}\tag{qnt}\tag{px3sp}\tag{dat}} & \texttt{нәрсә\tag{n}\tag{pl}\tag{dat}} & \texttt{зат\tag{n}\tag{pl}\tag{dat}} \\
	\texttt{кара\tag{v}\tag{tv}\tag{gna\_perf}} & \texttt{қара\tag{v}\tag{tv}\tag{gna\_perf}} & \texttt{кара\tag{v}\tag{tv}\tag{gna\_perf}} & \texttt{къара\tag{v}\tag{tv}\tag{gna\_perf}} \\ 
	\texttt{,\tag{cm}} & \texttt{,\tag{cm}} & \texttt{,\tag{cm}} & \texttt{,\tag{cm}} \\
	 --- & ---      & \texttt{алар\tag{prn}\tag{pers}\tag{p3}\tag{pl}\tag{gen}} & \texttt{олар\tag{prn}\tag{pers}\tag{p3}\tag{pl}\tag{nom}} \\
	\texttt{өтө\tag{adv}} & \texttt{өте\tag{adv}} & \texttt{бик\tag{adv}} & \texttt{бек\tag{adv}} \\
	\texttt{жакшы\tag{adj}} & \texttt{жақсы\tag{adj}} & \texttt{яхшы\tag{adj}} & \texttt{яхшы\tag{adj}} \\
	\texttt{э\tag{cop}\tag{ger\_past}\tag{px3sp}\tag{acc}} & \texttt{е\tag{cop}\tag{ger\_past}\tag{px3sp}\tag{acc}} & \texttt{и\tag{cop}\tag{ger\_past}\tag{px3sp}\tag{acc}} & \texttt{э\tag{cop}\tag{ger\_past}\tag{px3sp}\tag{acc}} \\
	\texttt{көр\tag{v}\tag{tv}\tag{ifi}\tag{p3}\tag{sg}} & \texttt{көр\tag{v}\tag{tv}\tag{ifi}\tag{p3}\tag{sg}} & \texttt{күр\tag{v}\tag{tv}\tag{past}\tag{p3}\tag{sg}} & \texttt{гёр\tag{v}\tag{tv}\tag{past}\tag{p3}\tag{sg}} \\  
	\texttt{.\tag{sent}} & \texttt{.\tag{sent}} & \texttt{.\tag{sent}} & \texttt{.\tag{sent}} \\
\bottomrule
\end{tabular}
}

		\htwo{Tagset}

\vspace{-1.5em}
\setlength{\columnsep}{-1em}
\setlength{\topsep}{-\parskip}
%\spacing{0.8}
\begin{multicols}{4}
\begin{tabbing}
  \texttt{{\small \tag{n}}}\hspace{1.5em} \=  Noun\\[-0.5ex]
%  \texttt{{\small <np>}} \> Proper noun\\[-0.5ex]
  \texttt{{\small \tag{v}}} \>  Verb\\[-0.5ex]
  \texttt{{\small \tag{prn}}} \> Pronoun \\[-0.5ex]
  \texttt{{\small \tag{det}}} \>  Determiner\\[-0.5ex]
%  \texttt{{\small <cnjcoo>}} \> Coord. conjunct.\\[-0.5ex]
%  \texttt{{\small <cnjadv>}} \> Adv. conjunct.\\[-0.5ex]
  \texttt{{\small \tag{adj}}} \> Adjective\\[-0.5ex]
  \texttt{{\small \tag{adv}}} \> Adverb\\[-0.5ex]
%  \texttt{{\small <vaux>}} \> Auxiliary verb\\[-0.5ex]
%  \texttt{{\small <cop>}} \> Copula\\[-0.5ex]
  \texttt{{\small \tag{iv}}} \> Intransitive\\[-0.5ex]
  \texttt{{\small \tag{tv}}} \> Transitive\\[-0.5ex]
%  \texttt{{\small <p2>}} \> Second person\\[-0.5ex]
  \texttt{{\small \tag{p3}}} \> Third person\\[-0.5ex]
%  \texttt{{\small <ant>}} \> Anthroponym\\[-0.5ex]
%  \texttt{{\small <dem>}} \> Demonstrative\\[-0.5ex]
%  \texttt{{\small <m>}} \> Masculine\\[-0.5ex]
%  \texttt{{\small <sg>}} \> Singular\\[-0.5ex]
  \texttt{{\small \tag{pl}}} \> Plural\\[-0.5ex]
  \texttt{{\small \tag{nom}}} \> `Nominative'\\[-0.5ex]
  \texttt{{\small \tag{gen}}} \> Genitive\\[-0.5ex]
  \texttt{{\small \tag{acc}}} \> Accusative\\[-0.5ex]
  \texttt{{\small \tag{dat}}} \> Dative\\[-0.5ex]
  \texttt{{\small \tag{qnt}}} \> Quantifier \\[-0.5ex]
%  \texttt{{\small \tag{itg}}} \> Interrogative \\[-0.5ex]
  \texttt{{\small \tag{ref}}} \> Reflexive \\[-0.5ex]
  \texttt{{\small \tag{pers}}} \> Personal \\[-0.5ex]
%  \texttt{{\small <loc>}} \> Locative\\[-0.5ex]
%  \texttt{{\small <px3sg>}} \> 3rd person poss. (Singular)\\[-0.5ex]
  \texttt{{\small \tag{cm}}} \> Comma\\[-0.5ex]
  \texttt{{\small \tag{sent}}} \> Sentence\\[-0.5ex]
%  \texttt{{\small <px3pl>}} \> 3rd person poss. (Plural)\\[-0.5ex]
%  \texttt{{\small <neg>}} \> Negative\\[-0.5ex]
%  \texttt{{\small <aor>}} \> Aorist\\[-0.5ex]
  \texttt{{\small \tag{past}}} \> Past (General) \\[-0.5ex]
  \texttt{{\small \tag{ifi}}} \> Past \\[-0.5ex]
  	\> \hspace{0.5em}(Eyewitness/Recent) \\[-0.5ex]
  \\[-0.5ex]
  \\[-0.5ex]
  \texttt{{\small \tag{px3sp}}} \hspace{1.1em} \= 3rd person poss.\\[-0.5ex]
  	\> \hspace{0.5em}(Singular/Plural)\\[-0.5ex]
%  \texttt{{\small <imp>}} \> Imperative\\[-0.5ex]
  \texttt{{\small \tag{gna\_perf}}} \> Verbal adverb \\[-0.5ex]
  	\> \hspace{0.5em}(Perfect)\\[-0.5ex]
  \texttt{{\small \tag{gpr\_past}}} \> Verbal adjective \\[-0.5ex]
  	\> \hspace{0.5em}(Past)\\[-0.5ex]
  \texttt{{\small \tag{ger\_past}}} \> Verbal noun (Past)
%  \texttt{{\small <prc\_impf>}} \> Participle (Imperfect)\\[-0.5ex]
%  \texttt{{\small <prc\_irre>}} \> Participle (Irrealis)\\[-0.5ex]
%  \texttt{{\small <prc\_real>}} \> Participle (Realis)\\[-0.5ex]

\end{tabbing}
\end{multicols}

		}

		\headerbox{Categorisation}{name=lexc,below=exout,column=1}{
		%\headerbox{Morphophonology}{name=morphophon,below=morphtrans}{
			\begin{itemize}
				\item Other Turkic transducers:\ \texttt{0}-derivation (\hilitetwo{overgenerates})
				\item Our approach:\ categorisation (e.g., adjectives, below)
			\end{itemize}

		\scalebox{0.63}{
			\begin{tabular}{lllll}
				\toprule
				\textbf{Type} & \textbf{Gloss} & {\small \texttt{\tag{adj}(\tag{comp})}} & {\small \texttt{\tag{adj}(\tag{comp})\tag{subst}}} & {\small \texttt{\tag{adj}(\tag{comp})\tag{advl}}} \\
				\midrule
				\textbf{A1} & \eng{good} & {\qipa яхшы (яхшырак)} & {\qipa яхшы (яхшырак)} & {\qipa яхшы (яхшырак)} \\
				\textbf{A2} & \eng{old} & {\qipa иске (искерәк)} & {\qipa иске (искерәк)} & {\qipa — (—)} \\
				\textbf{A3} & \eng{dead} & {\qipa үле (—)} & {\qipa үле (—)} & {\qipa — (—)} \\
				\textbf{A4} & \eng{basic} & {\qipa төп (—)} & {\qipa — (—)} & {\qipa — (—)} \\
				\bottomrule
			\end{tabular}
		}
	}

		\headerbox{Evaluation}{name=coverage,span=1,column=3,row=0}{

			\htwo{Number of stems}
			%\vspace{-1.5em}
			\vspace{-0.5em}
			\begin{center}
			\begin{tabular}{lrrr}
				\toprule
				\multirow{2}{*}{\textbf{Part of speech}} & \multicolumn{3}{c}{\textbf{Number of stems}} \\ \cmidrule{2-4}
				& \textbf{Kazakh} & \textbf{Tatar} & \textbf{Kumyk} \\
				\midrule
				Noun & 2640 & 2795 & 2568 \\
				Verb & 1470 & 1143 & 386 \\
				Adjective & 754 & 816 & 219 \\
				Proper noun & 5701 & 5361 & 1443 \\
				Adverb & 171 & 177 & 63 \\
				Numeral & 63 & 63 & 44 \\
				Conjunction & 46 & 45 & 13 \\
				Postposition & 50 & 43 & 12 \\
				Pronoun & 32 & 28 & 17 \\
				Determiner & 39 & 34 & 9 \\
				\midrule
				Total: & \hiliteone{11224} & \hiliteone{10737} & \hiliteone{4845} \\
				\bottomrule
			\end{tabular}
			\end{center}
			%\vspace{2em}
			\vspace{-0.5em}

			\htwo{Test corpora}

%		\scalebox{0.8}{
%			\begin{tabular}{lll}
%				\toprule
%				\textbf{type} & \textbf{lang} & \textbf{contents} \\ % & \textbf{origin} \\
%				\midrule
%				Encyclopædic & kaz & Wikipedia \\ % wpdump & 20131006 \\
%					& tat & Wikipedia \\ % wpdump & 20130225 \\
%					& kum & — \\ % — & — \\
%				\midrule
%				News & kaz & RFE/RL (azattyq.org) \\ % RFE/RL & azattyq.org 2010 \\
%					& tat & tat.tatar-inform.ru \\ % Татар-информ & tat.tatar-inform.ru 2005-2011 \\
%					& kum & Ёлдаш (yoldash.etnosmi.ru) \\ % Ёлдаш & yoldash.etnosmi.ru \\
%				\midrule
%				Religion & kaz & Quran + Bible \\ % quran + bible & kkitap.net, kuran.kz \\
%					& tat & Quran + New Testament \\ % quran + nt & ibt.org.ru, tanzil.net \\
%					& kum & Genesis + New Testament \\ % genesis + nt & ibt.org.ru \\
%				\bottomrule
%			\end{tabular}
%		}
		\scalebox{0.78}{
			\begin{tabular}{llll}
				\toprule
					& \textbf{Wikipedia} & \textbf{News} & \textbf{Religion} \\
				\midrule
					\texttt{\textbf{kaz}} & Wikipedia & azattyk.org & Quran + Bible \\
					\texttt{\textbf{tat}} & Wikipedia & tat.tatar-inform.ru & Quran + New Testament \\
					\texttt{\textbf{kum}} & --- & yoldash.etnosmi.ru & Genesis + New Testament \\
				\bottomrule
			\end{tabular}

%			\begin{tabular}{llll}
%				\toprule
%					& \textbf{Kazakh} & \textbf{Tatar} & \textbf{Kumyk} \\
%				\midrule
%					Wikipedia & Wikipedia & Wikipedia & --- \\
%					News & azattyq.org & tat.tatar-inform.ru & yoldash.etnosmi.ru \\
%					Religion & Quran + Bible & Quran + New Testament & Genesis + New Testament \\
%				\bottomrule
%			\end{tabular}
		}\vspace{0.55em}

%			\begin{itemize}
%				\item split into 10 equal parts; coverage calculated over each separately; standard deviation of mean calculated
%			\end{itemize}
			%both split into 10 equal parts; coverage calculated over each separately; standard deviation of the mean was calculated\\
			\htwo{Evaluation measures}
			\begin{itemize}
				\item \textbf{Naïve coverage} - percentage of surface forms in a given corpus receiving ≥ 1 analysis%\\(surface forms may have missing analyses)
				\item \textbf{Mean ambiguity} - average number of analyses for each surface form found in analysed corpus
				\item \textbf{Precision} - of a form's analyses, \% correct
				\item \textbf{Recall} - \% of analyses provided by transducer that are correct for a form, by comparing against a gold standard
			\end{itemize}
			%\htwo{Coverage results}\\
			\htwo{Evaluation results}
%\centering
%\vspace{1pt}\hspace{2em}
			%%\begin{centering}
			%{\large
			%\begin{itemize}
			%	\item each corpus split into 10 equal parts; coverage calculated over each separately; st.\ dev.\ of mean calculated
			%\end{itemize}

		\scalebox{0.89}{
			\begin{tabular}{llrrr}
				\toprule
				 & \textbf{Corpus} & \textbf{Tokens} & \textbf{Coverage} (\%) & \textbf{Amb.} \\
				\midrule
				\multirow{4}{*}{\textbf{Kazakh}} & Wikipedia & 25.6M & 85.61 $\pm$ 1.37 & 2.43 \\
					& News & 3.8M & 92.12 $\pm$ 2.72  & 2.88 \\
					& Religion & 851K & 92.49 $\pm$ 1.66  & 2.63 \\\cmidrule{2-5}
				 (r50547) & Average &  & \hiliteone{90.07} $\pm$ 1.91  & 2.64\\
				\midrule
				\multirow{4}{*}{\textbf{Tatar}} & Wikipedia & 159K & 86.35 $\pm$ 2.17  & 2.24\\
					& News & 5.2M & 89.75 $\pm$ 0.07  & 2.30 \\
					& Religion & 382K & 91.25 $\pm$ 2.55 & 2.24 \\\cmidrule{2-5}
				 (r50260) & Average &  & \hiliteone{89.12} $\pm$ 1.60 & 2.26 \\
				\midrule
				\multirow{3}{*}{\textbf{Kumyk}} & News & 286K &  91.10 $\pm$ 0.86  & 1.53 \\
					& Religion & 227K & 92.47 $\pm$ 1.03  & 1.53 \\\cmidrule{2-5}
				 (r50300) & Average &  & \hiliteone{91.78} $\pm$ 0.94 & 1.53 \\
				\bottomrule
			\end{tabular}\\
		}
			%}
			%%\end{centering}
%		}

%		\headerbox{Precision \& recall}{name=precrecall,column=1,below=coverage}{

%			\htwo{Two measures}
%			\htwo{Precision \& recall}
			\vspace{0.425em}
			\begin{itemize}
				\item {}{\footnotesize selected \& proofed unique random surface forms from news corpora}
			\end{itemize}

			\vspace{-1em}
			\begin{center}
				\begin{tabular}{lrrr}
				\toprule
					\textbf{Language} & \textbf{Forms} & \textbf{Precision} (\%) & \textbf{Recall} (\%) \\
				\midrule
					Kazakh & 1000 & 98.61 &  57.98 \\
					Tatar & 1000 & 95.03 & 85.65 \\
					Kumyk & 500 & 96.57 & 69.11 \\
				\bottomrule
				\end{tabular}
			\end{center}

			%\vspace{0.49em}
%			\begin{itemize}
%				\item selected \& proofed unique random surface forms from news corpus
%				\item \textbf{Precision} (of a form's analyses \% correct): \hfill \hiliteone{\textbf{97.32\%}}
%				\item \textbf{Recall} (percentage of analyses provided by the transducer that are correct for a form, by comparing against a gold standard): \hfill \hiliteone{\textbf{94.56\%}}
%			\end{itemize}

%			\htwo{Results}
%
%			\vspace{-0.5em}
%			\begin{center}
%				\begin{tabular}{ll}
%					\toprule
%					Precision & Recall \\
%					\midrule 
%					97.32\% & 94.56\%  \\
%					\bottomrule
%				\end{tabular}
%			\end{center}
		}

		\headerbox{Orthography-phonology mapping issues}{name=twol,column=2,below=exout}{
			\htwo{Ambiguous characters}
			\begin{itemize}
				\item Have front- and back-vowel readings in native words
			\end{itemize}

		\scalebox{0.92}{
			\begin{tabular}{llll}
				\toprule
					& \textbf{letters} & \textbf{values} & \textbf{examples} \\
				\midrule
					\multirow{2}{*}{\texttt{\textbf{kaz}}} & \multirow{2}{*}{{\qipa и, у, ю}} & {\qipa /əj, əw, jəw/} & {\qipa қ\hiliteone{ию}д\hilitetwo{а}} \eng{chopping down} \\
						&	& {\qipa /ɘj, ɘw, jɘw/} &  {\qipa к\hiliteone{ию}д\hilitetwo{е}} \eng{getting dressed} \\
				\midrule

					\multirow{3}{*}{\texttt{\textbf{tat}}} & \multirow{3}{*}{{\qipa е}} & {\qipa э / C \_} & {\qipa дәр\hiliteone{е}сл\hilitetwo{ә}р} \eng{lessons} \\
						&	& {\qipa /j/+ы} & {\qipa \hiliteone{е}лл\hilitetwo{а}р} \eng{years} \\
						&	& {\qipa /j/+э} & {\qipa ег\hiliteone{е}тл\hilitetwo{ә}р} \eng{boys} \\
				\midrule

					\multirow{6}{*}{\texttt{\textbf{kum}}} & \multirow{6}{*}{{\qipa ё, ю}} & \multirow{2}{*}{{\qipa /ø, y/ / C \_}} & {\qipa г\hiliteone{ю}нл\hilitetwo{е}р} \eng{days} \\
						&	&	& {\qipa г\hiliteone{ё}зл\hilitetwo{е}р} \eng{eyes} \\
						&	& \multirow{2}{*}{{\qipa /jø, jy/}} & {\qipa \hiliteone{ю}рекл\hilitetwo{е}р} \eng{hearts}\\
						&	&	& {\qipa \hiliteone{ё}нкюл\hilitetwo{е}р} \eng{darlings} \\
						&	& \multirow{2}{*}{{\qipa /jo, ju/}} & {\qipa \hiliteone{ю}лдузл\hilitetwo{а}р} \eng{stars} \\
						&	&	& {\qipa \hiliteone{ё}лл\hilitetwo{а}р} \eng{roads} \\
				\bottomrule
			\end{tabular}
		}\vspace{0.5em}
			\begin{itemize}
				\item solution:\ hairy \texttt{twol} rules cover majority of examples
				\item unaccounted-for words get a harmony-forcing character
				\item adjust rules for harmony-forcing characters
			\end{itemize}

			\htwo{Loanwords}
				\begin{itemize}
					\item Letters that represent front vowels in native words may represent ``back'' vowels in Russian words
				\end{itemize}
			\scalebox{0.86}{
				\begin{tabular}{lll}
					\toprule
						 & \textbf{native word example} & \textbf{Russian word example} \\
					\midrule
						\texttt{\textbf{kaz}} & {\qipa \hiliteone{е}лд\hilitetwo{і}ң} \eng{country's} & {\qipa Назарба\hiliteone{е}вт\hilitetwo{ы}ң} \eng{Nazarbayev's} \\
						\texttt{\textbf{tat}} & {\qipa гал\hiliteone{и}мн\hilitetwo{ә}р} \eng{scientists} & {\qipa арт\hiliteone{и}стл\hilitetwo{а}р} \eng{artists} \\
						\texttt{\textbf{kum}} & {\qipa с\hiliteone{ё}зл\hilitetwo{е}р} \eng{words} & {\qipa самол\hiliteone{ё}тл\hilitetwo{а}р} \eng{airplanes} \\
					\bottomrule
				\end{tabular}
			}
			\begin{itemize}
				\item solution:\ separate continuation lexicon (messy rules)
			\end{itemize}
				\codeex{
					\begin{spacing}{0.9}
					\texttt{LEXICON N1-RUS\\
					:\%\{☭\%\} N1 ;\\
					\vspace*{-0.5em}\\
					LEXICON Nouns\\
					артист:артист N1-RUS ; ! "artist"\\
					галим:галим N1 ; ! "scientist"}\end{spacing}
					\vspace{0em}
				}\\

			\htwo{Acronyms and numerals}
				\begin{itemize}
					\item \texttt{twol} rules handle phonology for spelt-out words\\
					{\qipa от\hiliteone{ыз}\hilitetwo{да}н} \eng{from thirty}, {\qipa б\hiliteone{ес}\hilitetwo{те}н} \eng{from five}
					\item no phonological triggers available in numerals (etc.)\\
					{\qipa 30-\hilitetwo{да}н} \eng{from 30}, {\qipa 5-\hilitetwo{те}н} \eng{from 5}
					\item solution:\ phonology-triggering characters
				\end{itemize}
				\codeex{
					\begin{spacing}{0.9}
					\texttt{4:4\%\{э\%\}\%\{с\%\}           NUM-DIGIT ; ! "төрт"\\
					5:5\%\{э\%\}\%\{с\%\}           NUM-DIGIT ; ! "бес"\\
					3\%0:3\%0\%\{а\%\}\%\{з\%\}       NUM-DIGIT ; ! "отыз"}\end{spacing}
					\vspace{0em}
				}\\

			\htwo{A resulting messy \texttt{twol} rule}
				\codeex{
					\begin{spacing}{0.4}
					RdYotVow = ё ю Ё Ю ; \\
					AbstractVow = \%\{а\%\} \%\{э\%\} \%\{ү\%\} \%\{о\%\} ; \\
					\\
					"{A} front unrounded harmony" \\
\%\{A\%\}:е <=> \hfill{} [ [ :FrontVow | [ :Vow :ь ] ] :Cns :Cns* ]/:0 \_ ; \\
\hspace*{0.1em}\hfill{}                        [ [ :RdVow :ь ] :Cns :Cns* ]/:0 \_ ; \\
\hspace*{0.1em}\hfill{}     [ [ [ \textbackslash{}[ .\#.\ | :Vow ] ] :RdYotVow ] :Cns :Cns* ]/:0 \_ ; \\
\hspace*{0.1em}\hfill{}      [ :RdYotVow й:0 :RdYotVow :Cns :Cns* ]/[ :0 - й:0 ] \_ ; \\ % ! кююню \\
\hspace*{0.1em}\hfill{} [[\%\{э\%\}:0|\%\{ү\%\}:0] :Cns* ]/[[ :0 - AbstractVow: ] | \%-: ]* \_ ; \\
\hspace*{4em}        except \\
\hspace*{0.1em}\hfill{}       [ :RdYotVow :Cns* \%\{☭\%\}:0 :Cns* ]/[ :0 - \%\{☭\%\}:0 ] \_ ; \\ % ! -ёрлар in Russian words \\
\hspace*{0.1em}\hfill{}                          [ :Cns :р \%\{☭\%\}: \%>: :Cns* ]/:0 \_ ; \\% ! центр and тигр take back vowels \\
\hspace*{0.1em}\hfill{}         [ [ :Vow - :RdYotVow ] :RdYotVow :Cns :Cns* ]/:0 \_ ; \\% ! for some reason the rule above doesn't exclude this condition  e.g., боюнгъа \\
\hspace*{0.1em}\hfill{}                          [ :Vow ]/[ [ :0 - й:0 ] | \%>: ] \_ ;\\
			\vspace{-1em}
					\end{spacing}
				}

		}

%			\htwo{Vowel harmony}
%			\begin{itemize}
%				\item two basic vowel archiphonemes:
%				\begin{itemize}
%					\item high vowel, \texttt{\{I\}}; [±back] and [±round] from prev.\ V\\
%		\begin{tabular}{ccccc}
%			\toprule
%			after & result & & after & result \\
%			\midrule
%			{\qipa и} & {\qipa и} & & {\qipa ы} & {\qipa ы} \\
%			{\qipa ү} & {\qipa ү} & & {\qipa у} & {\qipa у} \\
%			{\qipa е} & {\qipa и} & & {\qipa а} & {\qipa ы} \\
%			{\qipa ө} & {\qipa ү} & & {\qipa о} & {\qipa у} \\
%			\bottomrule
%		\end{tabular}
%					\item low vowel, \texttt{\{A\}}; [±back] and [±round] from prev.\ V\\
%					with one important exception: after \hilitetwo{у}\\
%		\begin{tabular}{ccccc}
%			\toprule
%			after & result & & after & result \\
%			\midrule
%			{\qipa и} & {\qipa е} & & {\qipa ы} & {\qipa а} \\
%			{\qipa ү} & {\qipa ө} & & {\qipa \hilitetwo{у}} & {\qipa \hilitetwo{а}} \\
%			{\qipa е} & {\qipa е} & & {\qipa а} & {\qipa а} \\
%			{\qipa ө} & {\qipa ө} & & {\qipa о} & {\qipa о} \\
%			\bottomrule
%		\end{tabular}
%
%
%				\end{itemize}
%				\item take backness and roundness of previous vowel\\
%				exception: 
%
%			\end{itemize}

%			\htwo{Desonorisation}
%\begin{itemize}
%	\item {\texttt \{N\}} desonorises to д after a consonant\\
%		алма-\hiliteone{\{N\}}\{I\} → алма\hiliteone{н}ы \eng{apple--\gmk{ACC}} \\
%		сыр-\hiliteone{\{N\}}\{I\} → сыр\hiliteone{д}ы \eng{secret--\gmk{ACC}}
%	\item {\texttt \{L\}} desonorises to д after cons.\ of sonority ≤ /l/ \\
%		сыр-\hiliteone{\{L\}}\{A\}р → сыр\hiliteone{л}ар \eng{secret--\gmk{PL}} \\
%		кыз-\hiliteone{\{L\}}\{A\}р → кыз\hiliteone{д}ар \eng{girl--\gmk{PL}} \\
%\end{itemize}
%
%	\vspace{-0.5ex}
%\codeex{
%			\texttt{"L Desonorisation"\\
%			\%\{L\%\}:д <=> :VoicedLowSonCns \%>:\ \blank{1em} ;}\\
%			
%			\texttt{"N Desonorisation"\\
%			\%\{N\%\}:д <=> :VoicedCns \%>:\ \blank{1em} ;}
%}
%
%			\htwo{Lenition}
%\begin{itemize}
%	\item Turn {\texttt \{y\}} into a harmonised high vowel when a vowel doesn't follow the following consonant:\\
%	мур\{у\}н → мурун \eng{nose}\\
%	мур\{у\}н+\{I\}м → мурдум \eng{my nose}\\
%\end{itemize}
%	\vspace{-0.5ex}
%\codeex{
%{\small \texttt{\%\{y\%\}:Vy <=> [ :LastVowel :Cns* :Cns ]/[:0] \blank{1em}\\
%	\vspace{-0.5ex}\hspace{1pt}\hfill{} [ :Cns [ .\#.\ | :Cns ] ]/[ :0 | \%>:]\ ;\\
%  \vspace{-0.5ex}\hspace{1ex}where  Vy  in  (  и  ү  и  и  ү  ы  ы  у  у  ы  у  у )\\
%  \vspace{-0.5ex}LastVowel  in  (  и  ү  е  э  ө  я  а  ё  о  ы  ю  у )\\
%  \vspace{-0.5ex}\hspace{12ex}         matched ;}}
%}
%
%			%\htwo{Nouns ending in /рн/}
%
%			\htwo{й+vowel letters}
%			\begin{itemize}
%				\item{} [ а о у ] become [ я ё ю ] after й and й deletes
%				\item й incorporated into the context of many rules
%				\item + separate rules to change the characters
%				\item + a rule to delete the original й\\
%			\end{itemize}
%	\vspace{-1ex}
%
%				\codeex{
%					\texttt{"Deletion of й before yoticised vowels"\\
%					й:0 <=> \blank{1em} [ :YotVow ]/[ :0 | \%>:\ ]\ ;}
%				}
%		}

		\headerbox{Ongoing and future work}{name=nedostatki,column=3,below=coverage}{
			\begin{itemize}
%				\item case changes for words with one root\\
%					\hiliteone{Ф}инландия \eng{Finland}, \hiliteone{ф}инландиялык \eng{Finnish}
%				\item phonol.\ (vowel harmony, desonorisation) with abbrevs.\\
%					АКШ [акышы] \eng{USA} → АКШ\hiliteone{н}ын / *АКШ\hiliteone{т}ын
%				\item vowel harmony with numbers\\
%					{\qipa 100 [жүз] → 100дүн [жүздүн] / *100нын}
%				\item compound verbs (esp.\ ones with changeable parts)
%				\item gerunds with mono-syllabic V-final verbs\\
%					{\qipa иште- \eng{work} → иштеш / иштөө \eng{working}}\\
%					{\qipa же- \eng{eat} → жеш / *жөө \eng{eating}}
				\item Disambiguation, more stems
				\item Machine translation between these languages
				\item Other Turkic lgs.:\ Nogay, Bashqort, Uzbek, Chuvash
			\end{itemize}
			%\vspace{0.98em}
		}

		\headerbox{Further information}{name=getting,column=4}{
			\begin{itemize}
				\item Part of Apertium Turkic project:\\
				{\small \url{http://wiki.apertium.org/wiki/Apertium\_Turkic}}
				\item Transducers available live at \href{http://turkic.apertium.org/}{\texttt{turkic.apertium.org}}
				\item Source code available from apertium's svn repo
				%info at {\small \url{http://wiki.apertium.org/wiki/apertium-kir}}

				\item Turkic RBMT mailing list (>25 subscribers):\\
				\texttt{apertium-turkic@lists.sourceforge.net}\\
				\vspace{-0.5ex}Feel free to post in any language!\\\vspace{-2.5ex}
				\item See our paper in the LREC 2014 proceedings
				\item And feel free to contact the authors any time!
			\end{itemize}
			%\vspace{2.8ex}
		}

%		\headerbox{References}{name=references,column=2,below=nedostatki}{
%			%\begin{thebibliography}{1}\itemsep=-0.01em
%			\setlength{\baselineskip}{0.4em}
%		}

	\end{poster}
\end{document}

### 2014-06-13 kypchak-morphfest-2014/poster.tex
better logos, more content
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@54176 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 			%\hspace{-2em}
-			\includegraphics[angle=90,height=8em]{apertium5a}%\hspace{-2em}
+			\includegraphics[angle=90,height=6.5em]{apertium5b}%\hspace{-2em}
 			%Eye Catcher, empty if option eyecatcher=false - unused
@@
 		\vspace{0.5ex}
-		\includegraphics[width=\textwidth]{map}
+		\includegraphics[width=\textwidth]{map/map}
 		\vspace{-1.1em}  % -1.25em
@@
 
-	\scalebox{0.85}{
+	\scalebox{0.79}{
 		\begin{tabular}{lllll}
@@
 			\item Turkmen (\hilitetwo{Tantuğ et al., 2006})
-			\item Kyrgyz (\hiliteone{Washington et al., 2012})%\\
 			\item Kazakh (\hilitetwo{Бекманова \& Махимов, 2013})
 			%\vspace{1pt}\hfill{} \hiliteone{GPL (=free and open)}!
-			\item our Kazakh, Tatar, Kumyk: all \hiliteone{GPL (=free and open)}!
+			\item our Kyrgyz, Kazakh, Tatar, Kumyk:\ all \hiliteone{GPL (=free and open)}!
 		\end{itemize}
@@
 				\glb God own-his created [everything/thing-s]-to looked.at, they/their very good being saw. // 
-				\glft \eng{God looked at everything he had created and saw that it was very good. (Bible, Genesis 1:31)} //
+				\glft \eng{God looked at everything he had created and saw that it was very good.} (Bible, Genesis 1:31) //
 			\endgl
@@
 		\htwo{Output}
-\scalebox{0.81}{
+\scalebox{0.68}{
 \begin{tabular}{llll}
@@
 %\spacing{0.8}
-\begin{multicols}{4}
+\begin{multicols}{5}
 \begin{tabbing}
@@
   \texttt{{\small \tag{pl}}} \> Plural\\[-0.5ex]
+  \texttt{{\small \tag{ref}}} \> Reflexive \\[-0.5ex]
+  \texttt{{\small \tag{pers}}} \> Personal \\[-0.5ex]
   \texttt{{\small \tag{nom}}} \> `Nominative'\\[-0.5ex]
@@
 %  \texttt{{\small \tag{itg}}} \> Interrogative \\[-0.5ex]
-  \texttt{{\small \tag{ref}}} \> Reflexive \\[-0.5ex]
-  \texttt{{\small \tag{pers}}} \> Personal \\[-0.5ex]
 %  \texttt{{\small <loc>}} \> Locative\\[-0.5ex]
@@
   \texttt{{\small \tag{cm}}} \> Comma\\[-0.5ex]
-  \texttt{{\small \tag{sent}}} \> Sentence\\[-0.5ex]
+  \texttt{{\small \tag{sent}}} \hspace{0.2em} \= Sentence\\[-0.5ex]
 %  \texttt{{\small <px3pl>}} \> 3rd person poss. (Plural)\\[-0.5ex]
@@
   	\> \hspace{0.5em}(Eyewitness/Recent) \\[-0.5ex]
-  \\[-0.5ex]
-  \\[-0.5ex]
-  \texttt{{\small \tag{px3sp}}} \hspace{1.1em} \= 3rd person poss.\\[-0.5ex]
+%  \\[-0.5ex]
+%  \\[-0.5ex]
+  \texttt{{\small \tag{px3sp}}} \> 3rd person poss.\\[-0.5ex]
   	\> \hspace{0.5em}(Singular/Plural)\\[-0.5ex]
 %  \texttt{{\small <imp>}} \> Imperative\\[-0.5ex]
-  \texttt{{\small \tag{gna\_perf}}} \> Verbal adverb \\[-0.5ex]
+  \texttt{{\small \tag{gna\_perf}}} \hspace{0.05em} \= Verbal adverb \\[-0.5ex]
   	\> \hspace{0.5em}(Perfect)\\[-0.5ex]
@@
   	\> \hspace{0.5em}(Past)\\[-0.5ex]
-  \texttt{{\small \tag{ger\_past}}} \> Verbal noun (Past)
+  \texttt{{\small \tag{ger\_past}}} \> Verbal noun \\[-0.5ex]
+	\> \hspace{0.5em} (Past)
 %  \texttt{{\small <prc\_impf>}} \> Participle (Imperfect)\\[-0.5ex]
@@
 			\begin{itemize}
+				\item Example:\ morphologically distinct adjective classes (not documented elsewhere)
 				\item Other Turkic transducers:\ \texttt{0}-derivation (\hilitetwo{overgenerates})
-				\item Our approach:\ categorisation (e.g., adjectives, below)
+				\item Our approach:\ categorisation (generates/analyses only correct forms)
 			\end{itemize}
@@
 %		}
-		\scalebox{0.78}{
+		\scalebox{0.77}{
 			\begin{tabular}{llll}
@@
 						&	& {\qipa /j/+ы} & {\qipa \hiliteone{е}лл\hilitetwo{а}р} \eng{years} \\
-						&	& {\qipa /j/+э} & {\qipa ег\hiliteone{е}тл\hilitetwo{ә}р} \eng{boys} \\
+						&	& {\qipa /j/+э} & {\qipa \hiliteone{е}гетл\hilitetwo{ә}р} \eng{boys} \\
 				\midrule
@@
 				\item Machine translation between these languages
-				\item Other Turkic lgs.:\ Nogay, Bashqort, Uzbek, Chuvash
+				\item Bring our other Kypchak transducers to comparable performance:\\
+				Qaraqalpaq, Bashqort, Nogay, Crimean Tatar
+				\item Other Turkic lgs.:\ Uzbek, Uyghur, Chuvash, Yakut, Tuvan, etc.
 			\end{itemize}
@@
 				\vspace{-0.5ex}Feel free to post in any language!\\\vspace{-2.5ex}
-				\item See our paper in the LREC 2014 proceedings
+				\item See our papers in LREC proceedings\\
+				(2012:\ Kyrgyz, 2014:\ Kazakh, Tatar, Kumyk)
 				\item And feel free to contact the authors any time!

### 2014-06-16 kypchak-morphfest-2014/poster.tex
some more Kyrgyz numbers and stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@54480 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 	}
-	\headerbox{Development effort}{name=devel,below=morphtrans}{
+
+	\headerbox{Development effort}{name=devel,below=adjectives,column=1}{
+
 		\begin{itemize}
@@
 
-		\headerbox{Categorisation}{name=lexc,below=exout,column=1}{
+		\headerbox{Categorisation}{name=adjectives,below=exout,column=1}{
 		%\headerbox{Morphophonology}{name=morphophon,below=morphtrans}{
@@
 			\begin{center}
-			\begin{tabular}{lrrr}
+			\begin{tabular}{lrrrr}
 				\toprule
-				\multirow{2}{*}{\textbf{Part of speech}} & \multicolumn{3}{c}{\textbf{Number of stems}} \\ \cmidrule{2-4}
-				& \textbf{Kazakh} & \textbf{Tatar} & \textbf{Kumyk} \\
+				\multirow{2}{*}{\textbf{Part of speech}} & \multicolumn{4}{c}{\textbf{Number of stems}} \\ \cmidrule{2-5}
+				& \textbf{Kyrgyz} & \textbf{Kazakh} & \textbf{Tatar} & \textbf{Kumyk} \\
 				\midrule
-				Noun & 2640 & 2795 & 2568 \\
-				Verb & 1470 & 1143 & 386 \\
-				Adjective & 754 & 816 & 219 \\
-				Proper noun & 5701 & 5361 & 1443 \\
-				Adverb & 171 & 177 & 63 \\
-				Numeral & 63 & 63 & 44 \\
-				Conjunction & 46 & 45 & 13 \\
-				Postposition & 50 & 43 & 12 \\
-				Pronoun & 32 & 28 & 17 \\
-				Determiner & 39 & 34 & 9 \\
+				Noun & 4582 & 2640 & 2795 & 2568 \\
+				Verb & 1193 & 1470 & 1143 & 386 \\
+				Adjective & 1211 & 754 & 816 & 219 \\
+				Proper noun & 5887 & 5701 & 5361 & 1443 \\
+				Adverb & 312 & 171 & 177 & 63 \\
+				Numeral & 66 & 63 & 63 & 44 \\
+				Conjunction & 77 & 46 & 45 & 13 \\
+				Postposition & 50 & 50 & 43 & 12 \\
+				Pronoun & 51 & 32 & 28 & 17 \\
+				Determiner & 64 & & 39 & 34 & 9 \\
 				\midrule
-				Total: & \hiliteone{11224} & \hiliteone{10737} & \hiliteone{4845} \\
+				Total: & \hiliteone{13749} & \hiliteone{11224} & \hiliteone{10737} & \hiliteone{4845} \\
 				\bottomrule
@@
 				\midrule
-					\texttt{\textbf{kaz}} & Wikipedia & azattyk.org & Quran + Bible \\
-					\texttt{\textbf{tat}} & Wikipedia & tat.tatar-inform.ru & Quran + New Testament \\
-					\texttt{\textbf{kum}} & --- & yoldash.etnosmi.ru & Genesis + New Testament \\
+					\texttt{\textbf{Kyrgyz}} & Wikipedia & azattyk.org & Quran + Bible \\
+					\texttt{\textbf{Kazakh}} & Wikipedia & azattyq.org & Quran + Bible \\
+					\texttt{\textbf{Tatar}} & Wikipedia & tat.tatar-inform.ru & Quran + New Testament \\
+					\texttt{\textbf{Kumyk}} & --- & yoldash.etnosmi.ru & Genesis + New Testament \\
 				\bottomrule
@@
 				 & \textbf{Corpus} & \textbf{Tokens} & \textbf{Coverage} (\%) & \textbf{Amb.} \\
+				\midrule
+
+				\multirow{4}{*}{\textbf{Kyrgyz}} & Wikipedia & M &  $\pm$  &  \\
+					&  &  &  $\pm$  &  \\\cmidrule{2-5}
+				 (r54474) & Average &  & \hiliteone{} $\pm$  &  \\
 				\midrule
@@
 
-		\headerbox{Ongoing and future work}{name=nedostatki,column=3,below=coverage}{
+		\headerbox{Ongoing and future work}{name=nedostatki,column=5}{
 			\begin{itemize}
@@
 				\item Disambiguation, more stems
+				\item Clean up Kyrgyz transducer
 				\item Machine translation between these languages
@@
 
-		\headerbox{Further information}{name=getting,column=4}{
+		\headerbox{Further information}{name=getting,column=0,below=morphtrans}{
 			\begin{itemize}

### 2014-06-16 kypchak-morphfest-2014/poster.tex
more work
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@54486 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 		\begin{itemize}
+			\item Efficient (in speed \& size) models of a language's morphology
 			\item Take a surface form, and produce valid lexical form(s)
@@
 
-	\headerbox{Development effort}{name=devel,below=adjectives,column=1}{
+	\headerbox{Development effort}{name=devel,column=4,row=0}{
 
@@
 
-		\headerbox{Evaluation}{name=coverage,span=1,column=3,row=0}{
+		\headerbox{Evaluation}{name=coverage,span=1,column=5,row=0}{
 
@@
 					\texttt{\textbf{Kyrgyz}} & Wikipedia & azattyk.org & Quran + Bible \\
+					% 2013-05-30, 2010, 
 					\texttt{\textbf{Kazakh}} & Wikipedia & azattyq.org & Quran + Bible \\
@@
 
-				\multirow{4}{*}{\textbf{Kyrgyz}} & Wikipedia & M &  $\pm$  &  \\
-					&  &  &  $\pm$  &  \\\cmidrule{2-5}
+				\multirow{4}{*}{\textbf{Kyrgyz}} & Wikipedia & M & 84.51 $\pm$ 2.27 &  \\
+					& News &  & 91.43 $\pm$ 0.51 &  \\
+					& Religion &  &  $\pm$  &  \\\cmidrule{2-5}
 				 (r54474) & Average &  & \hiliteone{} $\pm$  &  \\
@@
 				\midrule
+					Kyrgyz & 500 &  &  \\
 					Kazakh & 1000 & 98.61 &  57.98 \\
@@
 
-		\headerbox{Orthography-phonology mapping issues}{name=twol,column=2,below=exout}{
+		\headerbox{Orthography-phonology mapping issues}{name=twol,column=3}{ %,below=exout}{
 			\htwo{Ambiguous characters}
@@
 
+		\headerbox{Ongoing and future work}{name=nedostatki,column=5,below=coverage}{
+			\begin{itemize}
+%				\item case changes for words with one root\\
+%					\hiliteone{Ф}инландия \eng{Finland}, \hiliteone{ф}инландиялык \eng{Finnish}
+%				\item phonol.\ (vowel harmony, desonorisation) with abbrevs.\\
+%					АКШ [акышы] \eng{USA} → АКШ\hiliteone{н}ын / *АКШ\hiliteone{т}ын
+%				\item vowel harmony with numbers\\
+%					{\qipa 100 [жүз] → 100дүн [жүздүн] / *100нын}
+%				\item compound verbs (esp.\ ones with changeable parts)
+%				\item gerunds with mono-syllabic V-final verbs\\
+%					{\qipa иште- \eng{work} → иштеш / иштөө \eng{working}}\\
+%					{\qipa же- \eng{eat} → жеш / *жөө \eng{eating}}
+				\item Disambiguation, more stems
+				\item Clean up Kyrgyz transducer
+				\item Machine translation between these languages
+				\item Bring other Kypchak transducers to comparable performance:\\
+				Qaraqalpaq, Bashqort, Nogay, Crimean Tatar
+				\item Other Turkic lgs.:\ Uzbek, Uyghur, Chuvash, Yakut, Tuvan, etc.
+			\end{itemize}
+			%\vspace{0.98em}
+		}
+
+		\headerbox{Further information}{name=getting,column=0,below=morphtrans}{
+			\begin{itemize}
+				\item Part of Apertium Turkic project:\\
+				{\small \url{http://wiki.apertium.org/wiki/Apertium\_Turkic}}
+				\item Transducers available live at \href{http://turkic.apertium.org/}{\texttt{turkic.apertium.org}}
+				\item Source code available from apertium's svn repo
+				%info at {\small \url{http://wiki.apertium.org/wiki/apertium-kir}}
+
+				\item Turkic RBMT mailing list (>25 subscribers):\\
+				\texttt{apertium-turkic@lists.sourceforge.net}\\
+				\vspace{-0.5ex}Feel free to post in any language!\\\vspace{-2.5ex}
+				\item See our papers in LREC proceedings\\
+				(2012:\ Kyrgyz, 2014:\ Kazakh, Tatar, Kumyk)
+				\item And feel free to contact the authors any time!
+			\end{itemize}
+			%\vspace{2.8ex}
+		}
+
+%		\headerbox{References}{name=references,column=2,below=nedostatki}{
+%			%\begin{thebibliography}{1}\itemsep=-0.01em
+%			\setlength{\baselineskip}{0.4em}
+%		}
+
+		\headerbox{Morphophonology}{name=morphophon,column=1,below=exout}{
 %			\htwo{Vowel harmony}
@@
 
-%			\htwo{Desonorisation}
-%\begin{itemize}
-%	\item {\texttt \{N\}} desonorises to д after a consonant\\
-%		алма-\hiliteone{\{N\}}\{I\} → алма\hiliteone{н}ы \eng{apple--\gmk{ACC}} \\
-%		сыр-\hiliteone{\{N\}}\{I\} → сыр\hiliteone{д}ы \eng{secret--\gmk{ACC}}
-%	\item {\texttt \{L\}} desonorises to д after cons.\ of sonority ≤ /l/ \\
-%		сыр-\hiliteone{\{L\}}\{A\}р → сыр\hiliteone{л}ар \eng{secret--\gmk{PL}} \\
-%		кыз-\hiliteone{\{L\}}\{A\}р → кыз\hiliteone{д}ар \eng{girl--\gmk{PL}} \\
-%\end{itemize}
-%
-%	\vspace{-0.5ex}
-%\codeex{
-%			\texttt{"L Desonorisation"\\
-%			\%\{L\%\}:д <=> :VoicedLowSonCns \%>:\ \blank{1em} ;}\\
-%			
-%			\texttt{"N Desonorisation"\\
-%			\%\{N\%\}:д <=> :VoicedCns \%>:\ \blank{1em} ;}
-%}
-%
-%			\htwo{Lenition}
-%\begin{itemize}
-%	\item Turn {\texttt \{y\}} into a harmonised high vowel when a vowel doesn't follow the following consonant:\\
-%	мур\{у\}н → мурун \eng{nose}\\
-%	мур\{у\}н+\{I\}м → мурдум \eng{my nose}\\
-%\end{itemize}
-%	\vspace{-0.5ex}
-%\codeex{
-%{\small \texttt{\%\{y\%\}:Vy <=> [ :LastVowel :Cns* :Cns ]/[:0] \blank{1em}\\
-%	\vspace{-0.5ex}\hspace{1pt}\hfill{} [ :Cns [ .\#.\ | :Cns ] ]/[ :0 | \%>:]\ ;\\
-%  \vspace{-0.5ex}\hspace{1ex}where  Vy  in  (  и  ү  и  и  ү  ы  ы  у  у  ы  у  у )\\
-%  \vspace{-0.5ex}LastVowel  in  (  и  ү  е  э  ө  я  а  ё  о  ы  ю  у )\\
-%  \vspace{-0.5ex}\hspace{12ex}         matched ;}}
-%}
-%
+			\htwo{Desonorisation (kaz \& kir)}
+				\begin{itemize}
+					\item {\texttt \{N\}} desonorises to д after a consonant\\
+						алма-\hiliteone{\{N\}}\{I\} → алма\hiliteone{н}ы \eng{apple--\gmk{ACC}} \\
+						сыр-\hiliteone{\{N\}}\{I\} → сыр\hiliteone{д}ы \eng{secret--\gmk{ACC}}
+					\item {\texttt \{L\}} desonorises to д after cons.\ of sonority ≤ /l/ \\
+						сыр-\hiliteone{\{L\}}\{A\}р → сыр\hiliteone{л}ар \eng{secret--\gmk{PL}} \\
+						кыз-\hiliteone{\{L\}}\{A\}р → кыз\hiliteone{д}ар \eng{girl--\gmk{PL}} \\
+				\end{itemize}
+
+				\vspace{-0.5ex}
+				\codeex{
+					\texttt{"L Desonorisation"\\
+						\%\{L\%\}:д <=> :VoicedLowSonCns \%>:\ \blank{1em} ;}\\
+
+					\texttt{"N Desonorisation"\\
+						\%\{N\%\}:д <=> :VoicedCns \%>:\ \blank{1em} ;}
+				}
+
+			\htwo{Epenthesis}
+				\begin{itemize}
+					\item Turn {\texttt \{y\}} into a harmonised high vowel when a vowel doesn't follow the following consonant:\\
+					мур\{у\}н → мурун \eng{nose}\\
+					мур\{у\}н+\{I\}м → мурдум \eng{my nose}\\
+				\end{itemize}
+					\vspace{-0.5ex}
+\codeex{
+{\small \texttt{\%\{y\%\}:Vy <=> [ :LastVowel :Cns* :Cns ]/[:0] \blank{1em}\\
+	\vspace{-0.5ex}\hspace{1pt}\hfill{} [ :Cns [ .\#.\ | :Cns ] ]/[ :0 | \%>:]\ ;\\
+  \vspace{-0.5ex}\hspace{1ex}where  Vy  in  (  и  ү  и  и  ү  ы  ы  у  у  ы  у  у )\\
+  \vspace{-0.5ex}LastVowel  in  (  и  ү  е  э  ө  я  а  ё  о  ы  ю  у )\\
+  \vspace{-0.5ex}\hspace{12ex}         matched ;}}
+}
+
 %			%\htwo{Nouns ending in /рн/}
@@
 %				}
-%		}
-
-		\headerbox{Ongoing and future work}{name=nedostatki,column=5}{
-			\begin{itemize}
-%				\item case changes for words with one root\\
-%					\hiliteone{Ф}инландия \eng{Finland}, \hiliteone{ф}инландиялык \eng{Finnish}
-%				\item phonol.\ (vowel harmony, desonorisation) with abbrevs.\\
-%					АКШ [акышы] \eng{USA} → АКШ\hiliteone{н}ын / *АКШ\hiliteone{т}ын
-%				\item vowel harmony with numbers\\
-%					{\qipa 100 [жүз] → 100дүн [жүздүн] / *100нын}
-%				\item compound verbs (esp.\ ones with changeable parts)
-%				\item gerunds with mono-syllabic V-final verbs\\
-%					{\qipa иште- \eng{work} → иштеш / иштөө \eng{working}}\\
-%					{\qipa же- \eng{eat} → жеш / *жөө \eng{eating}}
-				\item Disambiguation, more stems
-				\item Clean up Kyrgyz transducer
-				\item Machine translation between these languages
-				\item Bring our other Kypchak transducers to comparable performance:\\
-				Qaraqalpaq, Bashqort, Nogay, Crimean Tatar
-				\item Other Turkic lgs.:\ Uzbek, Uyghur, Chuvash, Yakut, Tuvan, etc.
-			\end{itemize}
-			%\vspace{0.98em}
-		}
-
-		\headerbox{Further information}{name=getting,column=0,below=morphtrans}{
-			\begin{itemize}
-				\item Part of Apertium Turkic project:\\
-				{\small \url{http://wiki.apertium.org/wiki/Apertium\_Turkic}}
-				\item Transducers available live at \href{http://turkic.apertium.org/}{\texttt{turkic.apertium.org}}
-				\item Source code available from apertium's svn repo
-				%info at {\small \url{http://wiki.apertium.org/wiki/apertium-kir}}
-
-				\item Turkic RBMT mailing list (>25 subscribers):\\
-				\texttt{apertium-turkic@lists.sourceforge.net}\\
-				\vspace{-0.5ex}Feel free to post in any language!\\\vspace{-2.5ex}
-				\item See our papers in LREC proceedings\\
-				(2012:\ Kyrgyz, 2014:\ Kazakh, Tatar, Kumyk)
-				\item And feel free to contact the authors any time!
-			\end{itemize}
-			%\vspace{2.8ex}
-		}
-
-%		\headerbox{References}{name=references,column=2,below=nedostatki}{
-%			%\begin{thebibliography}{1}\itemsep=-0.01em
-%			\setlength{\baselineskip}{0.4em}
-%		}
+		}
+
+		\headerbox{Morphotactics}{name=lexc,column=2,below=exout}{
+			\htwo{Morphological \& orthographical words}
+
+			\begin{itemize}
+				\item{} {\qipa өнүктүрөбүзбү ?} \eng{will we develop [it]?}\\
+				\texttt{ өнүк{\small <v><tv><caus><aor><p1><pl>+}бы{\small <qst>}}
+				\item{} {\qipa келатсаң} \eng{if you come}\\
+				\texttt{кел{\tt {\small <v><iv><prc\_impf>}}+жат{\small <vaux><gna\_cnd><p2><sg>}}
+			\end{itemize}
+
+			%\htwo{Irregular negatives of finite verb forms}
+
+			\htwo{Irregular [noun + possessive + case] forms}
+
+			\begin{itemize}
+				\item Some combinations of possessive and case morphemes are non-canonical (i.e., not formed simply by concatenation):
+
+			\begin{center}
+			\vspace{-0.5em}
+{\small
+{\qipa		
+	\begin{tabular}{llllll}
+		\toprule
+		case & form & 1SG & 2SG & 3SP  \\
+		\midrule
+		nom & — & -(I)м & -(I)ң & -(S)I  \\
+		acc & -NI & -(I)мдI & -(I)ңдI & -\textbf{(S)Iн}  \\
+		gen & -NIн & -(I)мдIн & -(I)ңдIн & -(S)IнIн  \\
+		loc & -DA & -(I)мдA & -(I)ңдA & -\textbf{(S)IндA}  \\
+		abl & -DAн & -(I)мдAн, & -(I)ндAн, & -\textbf{(S)IнAн}  \\
+		~   &     & -\textbf{(I)мAн} & -\textbf{(I)ңAн} \\
+		dat & -GA & -\textbf{(I)мA} & -\textbf{(I)ңA} & -\textbf{(S)IнA} \\
+		\bottomrule
+	\end{tabular}
+}}
+			\end{center}
+
+%				\item Trade-off:
+%				\begin{itemize}
+					\item morphophon.\ complicateder, morphotactics simpler
+					\item underlying form used: \texttt{\{S\}\{I\}\{n\}}
+					\item phonological rules delete \texttt{\{n\}}, \texttt{\{S\}} by context
+%				\end{itemize}
+			\end{itemize}
+
+			\htwo{Noun-noun compounds}
+
+			\begin{itemize}
+				\item one type of N-N compunds: N2 has \texttt{<px3>} and related morphology
+			\end{itemize}
+			\vspace{0.6ex}
+			\codeex{
+			{\small
+			\texttt{LEXICON N-INFL-3PX-COMPOUND\\
+			\vspace{-0.5ex}\%<n\%>:\%>\%\{S\%\}\%\{I\%\}\%\{n\%\} GEN-POS ;}\\
+
+			\texttt{LEXICON Nouns\\
+			\vspace{-0.5ex}аба\%\ ырайы:аба\%\ ырай\ N-INFL-3PX-COMPOUND ;\\\vspace{-0.5ex}\hspace{0.1pt}\hfill{}!\ "weather"\\
+			\vspace{-0.5ex}чакыруу\%\ кагазы:чакыруу\%\ кагаз\ N-INFL-3PX-\\\vspace{-0.5ex}\hspace{0.1pt}\hfill{}COMPOUND ;\ !\ "invitation"}}
+			}
+
+			\vspace{0.5ex}
+			%чакыруу кагазы
+		}
 

### 2014-06-17 kypchak-morphfest-2014/poster.tex
some more numbers and stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@54508 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 			%\begin{minipage}[t]{0.2\textwidth}
+				\hspace{1cm}\includegraphics[height=6.5em]{22888536}
 				\hspace{1cm}\includegraphics[height=6.5em]{iu_tab}
@@
 		\scalebox{0.96}{
-			`алдым' \hilitetwo{↔} \texttt{ал{\small <v><tv><ifi><p1><sg>}}, \texttt{алд{\small <n><px1sg><nom>}}
+			`алдым' \hilitetwo{↔} \texttt{ал\tag{v}\tag{tv}\tag{ifi}\tag{p1}\tag{sg}}, \texttt{алд\tag{n}\tag{px1sg}\tag{nom}}
 		}
@@
 				Pronoun & 51 & 32 & 28 & 17 \\
-				Determiner & 64 & & 39 & 34 & 9 \\
+				Determiner & 64 & 39 & 34 & 9 \\
 				\midrule
@@
 %		}
-		\scalebox{0.77}{
+		\scalebox{0.8}{
 			\begin{tabular}{llll}
@@
 				\midrule
-					\texttt{\textbf{Kyrgyz}} & Wikipedia & azattyk.org & Quran + Bible \\
+					\textbf{Kyrgyz} & Wikipedia & azattyk.org & Bible \\
 					% 2013-05-30, 2010, 
-					\texttt{\textbf{Kazakh}} & Wikipedia & azattyq.org & Quran + Bible \\
-					\texttt{\textbf{Tatar}} & Wikipedia & tat.tatar-inform.ru & Quran + New Testament \\
-					\texttt{\textbf{Kumyk}} & --- & yoldash.etnosmi.ru & Genesis + New Testament \\
+					\textbf{Kazakh} & Wikipedia & azattyq.org & Quran + Bible \\
+					\textbf{Tatar} & Wikipedia & tat.tatar-inform.ru & Quran + New Testament \\
+					\textbf{Kumyk} & --- & yoldash.etnosmi.ru & Genesis + New Testament \\
 				\bottomrule
@@
 			%\end{itemize}
-
-		\scalebox{0.89}{
+			\vspace{-0.5em}
+\begin{center}
+		\scalebox{0.91}{
 			\begin{tabular}{llrrr}
@@
 
-				\multirow{4}{*}{\textbf{Kyrgyz}} & Wikipedia & M & 84.51 $\pm$ 2.27 &  \\
-					& News &  & 91.43 $\pm$ 0.51 &  \\
-					& Religion &  &  $\pm$  &  \\\cmidrule{2-5}
-				 (r54474) & Average &  & \hiliteone{} $\pm$  &  \\
+				\multirow{4}{*}{\textbf{Kyrgyz}} & Wikipedia & 5.3M & 84.51 $\pm$ 2.27 & 3.56 \\
+					& News & 4.1M & 91.43 $\pm$ 0.51 & 4.19 \\
+					& Religion & 215K & 91.66 $\pm$ 1.81 & 3.99 \\\cmidrule{2-5}
+				 (r54474) & Average &  & \hiliteone{89.20} $\pm$ 3.48 &  \\
 				\midrule
@@
 		}
+\end{center}
 			%}
@@
 %			\htwo{Precision \& recall}
-			\vspace{0.425em}
-			\begin{itemize}
-				\item {}{\footnotesize selected \& proofed unique random surface forms from news corpora}
+			%\vspace{0.425em}
+			\vspace{-0.5em}
+			\begin{itemize}
+				\item {}{\small selected \& proofed unique random surface forms from news corpora}
 			\end{itemize}
@@
 %					{\qipa же- \eng{eat} → жеш / *жөө \eng{eating}}
-				\item Disambiguation, more stems
-				\item Clean up Kyrgyz transducer
+				\item Disambiguation, more stems, clean up transducers
 				\item Machine translation between these languages
@@
 				Qaraqalpaq, Bashqort, Nogay, Crimean Tatar
-				\item Other Turkic lgs.:\ Uzbek, Uyghur, Chuvash, Yakut, Tuvan, etc.
+				\item Other Turkic lgs:\ Uzbek, Uyghur, Chuvash, Yakut, Tuvan, etc.
 			\end{itemize}
@@
 				\item{} {\qipa өнүктүрөбүзбү ?} \eng{will we develop [it]?}\\
-				\texttt{ өнүк{\small <v><tv><caus><aor><p1><pl>+}бы{\small <qst>}}
+				\texttt{өнүк\tag{v}\tag{tv}\tag{caus}\tag{aor}\tag{p1}\tag{pl}+бы\tag{qst}}
 				\item{} {\qipa келатсаң} \eng{if you come}\\
-				\texttt{кел{\tt {\small <v><iv><prc\_impf>}}+жат{\small <vaux><gna\_cnd><p2><sg>}}
+				\texttt{кел\tag{v}\tag{iv}\tag{prc\_impf}+жат\tag{vaux}\tag{gna\_cnd}\tag{p2}\tag{sg}}
 			\end{itemize}
@@
 			\begin{itemize}
-				\item one type of N-N compunds: N2 has \texttt{<px3>} and related morphology
+				\item one type of N-N compunds:\ N2 has \tag{px3} and related morphology
 			\end{itemize}
@@
 			\vspace{-0.5ex}аба\%\ ырайы:аба\%\ ырай\ N-INFL-3PX-COMPOUND ;\\\vspace{-0.5ex}\hspace{0.1pt}\hfill{}!\ "weather"\\
-			\vspace{-0.5ex}чакыруу\%\ кагазы:чакыруу\%\ кагаз\ N-INFL-3PX-\\\vspace{-0.5ex}\hspace{0.1pt}\hfill{}COMPOUND ;\ !\ "invitation"}}
+			\vspace{-0.5ex}чакыруу\%\ кагазы:чакыруу\%\ кагаз\ N-INFL-3PX-COMPOUND\\\vspace{-0.5ex}\hspace{0.1pt}\hfill{};\ !\ "invitation"}}
 			}

### 2014-06-17 kypchak-morphfest-2014/poster.tex
some work on the poster
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@54509 72bbbca6-d526-0410-a7d9-f06f51895060
progress, some adjustment of the colours, etc.
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@54525 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 %\definecolor{headercol2}{RGB}{153,0,0}
+%\definecolor{headercol}{RGB}{6,130,22}
+%\definecolor{headercolone}{RGB}{6,130,22}
 \definecolor{headercolone}{RGB}{6,130,22}
@@
 %\definecolor{boxcolor}{RGB}{171,255,241}
-\definecolor{boxcolor}{RGB}{176,232,223}
+%definecolor{boxcolor}{RGB}{176,232,223}
+\definecolor{boxcolor}{RGB}{176,208,223}
 \definecolor{headerfontcol}{RGB}{255,255,255}
@@
 \newcommand{\tag}[1]{\hilitegrey{<}\hiliteone{#1}\hilitegrey{>}}
+\newcommand{\archiphon}[1]{\hilitegrey{\{}\hiliteone{#1}\hilitegrey{\}}}
 
@@
 
-	\scalebox{0.79}{
+	\scalebox{0.78}{
 		\begin{tabular}{lllll}
@@
 			\item Take a surface form, and produce valid lexical form(s)
-			\item Take a lexical form, and produce valid surface form(s)
-		\end{itemize}
-		\scalebox{0.96}{
-			`алдым' \hilitetwo{↔} \texttt{ал\tag{v}\tag{tv}\tag{ifi}\tag{p1}\tag{sg}}, \texttt{алд\tag{n}\tag{px1sg}\tag{nom}}
-		}
+			\item Take a lexical form, and produce valid surface form(s)\\
+		%\end{itemize}
+		\scalebox{0.98}{
+			алдым \hilitetwo{↔} \texttt{ал\tag{v}\tag{tv}\tag{ifi}\tag{p1}\tag{sg}}, \texttt{алд\tag{n}\tag{px1sg}\tag{nom}}
+		}
+		\end{itemize}
 
@@
 		\end{itemize}
+		\htwo{Approach}
+		\begin{itemize}
+			\item morphotactics implemented in \texttt{lexc}
+			\item morphophonology implemented in \texttt{twol}
+			\item compiled separately; compose-intersected to single transducer\\
+			алдым \hilitetwo{↔} \texttt{ал\hilitegrey{>}\archiphon{D}\archiphon{I}\hilitegrey{>}м} \hilitetwo{↔} \texttt{ал\tag{v}\tag{tv}\tag{ifi}\tag{p1}\tag{sg}}
+		\end{itemize}
+
 	}
@@
 
-		\headerbox{Categorisation}{name=adjectives,below=exout,column=1}{
+		\headerbox{Categorisation}{name=adjectives,column=4,below=devel}{
 		%\headerbox{Morphophonology}{name=morphophon,below=morphtrans}{
@@
 
-		\scalebox{0.63}{
+		\scalebox{0.72}{
 			\begin{tabular}{lllll}
@@
 					& Religion & 215K & 91.66 $\pm$ 1.81 & 3.99 \\\cmidrule{2-5}
-				 (r54474) & Average &  & \hiliteone{89.20} $\pm$ 3.48 &  \\
+				 (r54474) & Average &  & \hiliteone{89.20} $\pm$ 3.48 & 3.91 \\
 				\midrule
@@
 
-		\scalebox{0.92}{
+		%\scalebox{0.92}{
 			\begin{tabular}{llll}
@@
 			\end{tabular}
-		}\vspace{0.5em}
+		%}
+		\vspace{0.5em}
 			\begin{itemize}
@@
 				\end{itemize}
-			\scalebox{0.86}{
+			\scalebox{0.92}{
 				\begin{tabular}{lll}
@@
 					{\qipa от\hiliteone{ыз}\hilitetwo{да}н} \eng{from thirty}, {\qipa б\hiliteone{ес}\hilitetwo{те}н} \eng{from five}
-					\item no phonological triggers available in numerals (etc.)\\
+					\item no phonological triggers available in numerals
+					\item (incorrect phonological triggers in acronyms)\\
 					{\qipa 30-\hilitetwo{да}н} \eng{from 30}, {\qipa 5-\hilitetwo{те}н} \eng{from 5}
 					\item solution:\ phonology-triggering characters
+					\item simplified:\ e.g., \archiphon{с} for all voiceless ostruents
 				\end{itemize}
@@
 \hspace*{0.1em}\hfill{}         [ [ :Vow - :RdYotVow ] :RdYotVow :Cns :Cns* ]/:0 \_ ; \\% ! for some reason the rule above doesn't exclude this condition  e.g., боюнгъа \\
-\hspace*{0.1em}\hfill{}                          [ :Vow ]/[ [ :0 - й:0 ] | \%>: ] \_ ;\\
-			\vspace{-1em}
-					\end{spacing}
-				}
+\hspace*{0.1em}\hfill{}                          [ :Vow ]/[ [ :0 - й:0 ] | \%>: ] \_ ;\end{spacing}
+			\vspace{1em}
+
+				}
+
+			\htwo{Related issues}
+				\begin{itemize}
+					\item Stem-final й combines with following а, о, у to form я, ё, ю
+					\item 
+				\end{itemize}
 
@@
 			\begin{itemize}
-				\item Part of Apertium Turkic project:\\
+				\item Part of \textbf{Apertium Turkic} project:\\
 				{\small \url{http://wiki.apertium.org/wiki/Apertium\_Turkic}}
-				\item Transducers available live at \href{http://turkic.apertium.org/}{\texttt{turkic.apertium.org}}
-				\item Source code available from apertium's svn repo
+				\item Transducers available \textbf{live} at \href{http://turkic.apertium.org/}{\texttt{turkic.apertium.org}}
+				\item \textbf{Source code} available from Apertium's svn repo
 				%info at {\small \url{http://wiki.apertium.org/wiki/apertium-kir}}
 
-				\item Turkic RBMT mailing list (>25 subscribers):\\
+				\item Turkic RBMT \textbf{mailing list} (>25 subscribers):\\
 				\texttt{apertium-turkic@lists.sourceforge.net}\\
@@
 					\item Turn {\texttt \{y\}} into a harmonised high vowel when a vowel doesn't follow the following consonant:\\
-					мур\{у\}н → мурун \eng{nose}\\
-					мур\{у\}н+\{I\}м → мурдум \eng{my nose}\\
+					мур\archiphon{у}н → мурун \eng{nose}\\
+					мур\archiphon{у}н+\archiphon{I}м → мурдум \eng{my nose}\\
 				\end{itemize}
@@
 \codeex{
+			\begin{spacing}{0.9}
 {\small \texttt{\%\{y\%\}:Vy <=> [ :LastVowel :Cns* :Cns ]/[:0] \blank{1em}\\
@@
   \vspace{-0.5ex}LastVowel  in  (  и  ү  е  э  ө  я  а  ё  о  ы  ю  у )\\
-  \vspace{-0.5ex}\hspace{12ex}         matched ;}}
+  \vspace{-0.5ex}\hspace{12ex}         matched ;}}\vspace{-1.5em}
+  			\end{spacing}
 }
@@
 			\begin{itemize}
-				\item Some combinations of possessive and case morphemes are non-canonical (i.e., not formed simply by concatenation):
+				\item Some combinations of possessive + case morphemes are unpredicted (i.e., not formed simply by concatenation and application of phonology):
+			\end{itemize}
 
@@
 		\midrule
-		nom & — & -(I)м & -(I)ң & -(S)I  \\
-		acc & -NI & -(I)мдI & -(I)ңдI & -\textbf{(S)Iн}  \\
-		gen & -NIн & -(I)мдIн & -(I)ңдIн & -(S)IнIн  \\
-		loc & -DA & -(I)мдA & -(I)ңдA & -\textbf{(S)IндA}  \\
-		abl & -DAн & -(I)мдAн, & -(I)ндAн, & -\textbf{(S)IнAн}  \\
+		nominative & — & -(I)м & -(I)ң & -(с)I  \\
+		accusative & -NI & -(I)мдI & -(I)ңдI & -\textbf{(с)Iн}  \\
+		genitive & -NIн & -(I)мдIн & -(I)ңдIн & -(с)IнIн  \\
+		locative & -DA & -(I)мдA & -(I)ңдA & -\textbf{(с)IндA}  \\
+		ablative & -DAн & -(I)мдAн, & -(I)ндAн, & -\textbf{(с)IнAн}  \\
 		~   &     & -\textbf{(I)мAн} & -\textbf{(I)ңAн} \\
-		dat & -GA & -\textbf{(I)мA} & -\textbf{(I)ңA} & -\textbf{(S)IнA} \\
+		dative & -GA & -\textbf{(I)мA} & -\textbf{(I)ңA} & -\textbf{(с)IнA} \\
 		\bottomrule
-	\end{tabular}
+	\end{tabular}\\
 }}
-			\end{center}
+	{\footnotesize A,I,N,D,G have various allophones; (I) null after vowels; (с) null after cons.}\vspace{-1em}\end{center}
 
 %				\item Trade-off:
-%				\begin{itemize}
-					\item morphophon.\ complicateder, morphotactics simpler
-					\item underlying form used: \texttt{\{S\}\{I\}\{n\}}
-					\item phonological rules delete \texttt{\{n\}}, \texttt{\{S\}} by context
+				\begin{itemize}
+					\item underlying \tag{px3sp} form used:\ \texttt{\archiphon{s}\archiphon{I}\archiphon{n}}
+					\item phonological rules delete \texttt{\archiphon{s}}, \texttt{\archiphon{n}} by context
+					\item morphophonology more complicated, morphotactics simpler
 %				\end{itemize}
@@
 			\begin{itemize}
-				\item one type of N-N compunds:\ N2 has \tag{px3} and related morphology
+				\item a N-N compund type:\ N2 has \tag{px3} and related morphology\\
+				e.g., \texttt{аба ырайы\tag{n}\tag{loc}:\ \hiliteone{аба ырайында}, *\hilitetwo{аба ырайыда}}
 			\end{itemize}
@@
 			\codeex{
+			\begin{spacing}{0.9}
 			{\small
@@
 			\vspace{-0.5ex}аба\%\ ырайы:аба\%\ ырай\ N-INFL-3PX-COMPOUND ;\\\vspace{-0.5ex}\hspace{0.1pt}\hfill{}!\ "weather"\\
-			\vspace{-0.5ex}чакыруу\%\ кагазы:чакыруу\%\ кагаз\ N-INFL-3PX-COMPOUND\\\vspace{-0.5ex}\hspace{0.1pt}\hfill{};\ !\ "invitation"}}
-			}
-
-			\vspace{0.5ex}
+			\vspace{-0.5ex}чакыруу\%\ кагазы:чакыруу\%\ кагаз\ N-INFL-3PX-COMPOUND\\\vspace{-0.5ex}\hspace{0.1pt}\hfill{};\ !\ "invitation"}}\end{spacing}
+			\vspace{0.5ex}}
+
+			%\vspace{0.5ex}
 			%чакыруу кагазы

### 2014-06-17 kypchak-morphfest-2014/poster.tex
not much left
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@54540 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 	\headerbox{Kypchak languages}{name=kypchaklgs,column=0,row=0}{
-		\vspace{0.5ex}
+		\vspace{0.1ex}
 		\includegraphics[width=\textwidth]{map/map}
-		\vspace{-1.1em}  % -1.25em
+		\vspace{-1.25em}  % -1.25em
 		\begin{itemize}
@@
 			\item Take a surface form, and produce valid lexical form(s)
-			\item Take a lexical form, and produce valid surface form(s)\\
+			\item Take a lexical form, and produce valid surface form(s)\vspace{-0.1ex}\\
 		%\end{itemize}
@@
 			\item morphophonology implemented in \texttt{twol}
-			\item compiled separately; compose-intersected to single transducer\\
-			алдым \hilitetwo{↔} \texttt{ал\hilitegrey{>}\archiphon{D}\archiphon{I}\hilitegrey{>}м} \hilitetwo{↔} \texttt{ал\tag{v}\tag{tv}\tag{ifi}\tag{p1}\tag{sg}}
+			\item compiled separately; compose-intersected to single transducer\vspace{-0.1ex}\\
+			алдым\hfill{}\hilitetwo{↔}\hfill{}\texttt{ал\hilitegrey{>}\archiphon{D}\archiphon{I}\hilitegrey{>}м}\hfill{}\hilitetwo{↔}\hfill{}\texttt{ал\tag{v}\tag{tv}\tag{ifi}\tag{p1}\tag{sg}}\vspace{-0.1ex}\\
+			алдым\hfill{}\hilitetwo{↔}\hfill{}\texttt{алд\hilitegrey{>}\archiphon{I}м}\hfill{}\hilitetwo{↔}\hfill{}\texttt{алд\tag{n}\tag{px1sg}\tag{nom}}
 		\end{itemize}
@@
 	\headerbox{Development effort}{name=devel,column=4,row=0}{
-
-		\begin{itemize}
+	\scalebox{0.92}{
+		\begin{tabular}{lllll}
+			\toprule
+			 & \textbf{Kyrgyz} & \textbf{Kazakh} & \textbf{Tatar} & \textbf{Kumyk} \\
+			\midrule
+			begun & Apr.\ 2011 & Dec.\ 2010 & Dec.\ 2011 & Oct.\ 2013 \\
+			80\% cov.\ & Aug.?\ 2011 & Aug.\ 2012 & Aug.\ 2012 & Oct.\ 2013 \\\midrule
+			time & 4 months & 19 months & 7 months & 1 week \\
+			\bottomrule
+		\end{tabular}
+	}
+		\begin{itemize}
+			%\item Kyrgyz and Kazakh transducers begun in 2011
+			%\item Kyrgyz transducer reached 80\% coverage summer of 2011
+			\item Kazakh transducer based on Kyrgyz transducer
 			\item Kumyk transducer based on Kazakh, Tatar transducers
@@
 		%\headerbox{Morphophonology}{name=morphophon,below=morphtrans}{
-			\begin{itemize}
-				\item Example:\ morphologically distinct adjective classes (not documented elsewhere)
+			\htwo{Adjectives}
+			\begin{itemize}
+				\item morphologically distinct adjective classes
+				\item most sources:\ adjectives can be used substantively and adverbially
 				\item Other Turkic transducers:\ \texttt{0}-derivation (\hilitetwo{overgenerates})
-				\item Our approach:\ categorisation (generates/analyses only correct forms)
-			\end{itemize}
-
-		\scalebox{0.72}{
+				\item but not all adjectives have all of the following:\\
+				comparative forms, substantive readings, adverbial readings
+				\item Our approach:\ categorisation 
+				\item only correct forms are analysed and generated
+			\end{itemize}
+
+		\scalebox{0.7}{
 			\begin{tabular}{lllll}
@@
 \begin{center}
-		\scalebox{0.91}{
+		\scalebox{0.94}{
 			\begin{tabular}{llrrr}
@@
 			%\vspace{0.425em}
-			\vspace{-0.5em}
+			\vspace{-0.4em}
 			\begin{itemize}
@@
 			\begin{center}
+			\scalebox{1.05}{
 				\begin{tabular}{lrrr}
@@
 				\end{tabular}
+			}
 			\end{center}
@@
 
+			\vspace{-0.75ex}
 			\htwo{Acronyms and numerals}
@@
 					{\qipa от\hiliteone{ыз}\hilitetwo{да}н} \eng{from thirty}, {\qipa б\hiliteone{ес}\hilitetwo{те}н} \eng{from five}
-					\item no phonological triggers available in numerals
-					\item (incorrect phonological triggers in acronyms)\\
+					\item no phonological triggers available in numerals\vspace{-0.2em}\\
+					 (incorrect phonological triggers in acronyms)\\
 					{\qipa 30-\hilitetwo{да}н} \eng{from 30}, {\qipa 5-\hilitetwo{те}н} \eng{from 5}
@@
 				}\\
+
+			\vspace{-0.75ex}
+			\htwo{й + vowel letters}
+				\begin{itemize}
+					%\item Stem-final й combines with following а, о, у to form я, ё, ю
+					\item{} [ а о у ] become [ я ё ю ] after й and й deletes
+					\item й incorporated into the context of many rules
+					\item additional rules to change the characters and delete original й
+				\end{itemize}
+%	\vspace{-1ex}
+%
+				\codeex{
+					\begin{spacing}{0.9}
+					\texttt{"Deletion of й before yoticised vowels"\\
+					й:0 <=> \blank{1em} [ :YotVow ]/[ :0 | \%>:\ ]\ ;}
+					\end{spacing}
+					\vspace{0em}
+				}
+%					\item 
+%				\end{itemize}
 
@@
 				}
-
-			\htwo{Related issues}
-				\begin{itemize}
-					\item Stem-final й combines with following а, о, у to form я, ё, ю
-					\item 
-				\end{itemize}
-
 		}
@@
 				\item Part of \textbf{Apertium Turkic} project:\\
-				{\small \url{http://wiki.apertium.org/wiki/Apertium\_Turkic}}
+				\url{http://wiki.apertium.org/wiki/Apertium\_Turkic}
 				\item Transducers available \textbf{live} at \href{http://turkic.apertium.org/}{\texttt{turkic.apertium.org}}
@@
 					\item underlying \tag{px3sp} form used:\ \texttt{\archiphon{s}\archiphon{I}\archiphon{n}}
-					\item phonological rules delete \texttt{\archiphon{s}}, \texttt{\archiphon{n}} by context
+					\item \texttt{\archiphon{s}} and \texttt{\archiphon{n}} default to с and н; rules map to null by context
 					\item morphophonology more complicated, morphotactics simpler

### 2014-06-18 kypchak-morphfest-2014/poster.tex
precision and recall definition fixes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@54593 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 				\item \textbf{Mean ambiguity} - average number of analyses for each surface form found in analysed corpus
-				\item \textbf{Precision} - of a form's analyses, \% correct
-				\item \textbf{Recall} - \% of analyses provided by transducer that are correct for a form, by comparing against a gold standard
+%				Precision is the probability that an analysis provided by the transducer is accurate.
+%				Recall is the probability that some correct analysis is among those provided.
+				\item \textbf{Precision} - probability that a provided analysis is accurate
+				\item \textbf{Recall} - probability that a certain correct analysis is among those provided by the transducer
+				%\item \textbf{Precision} - of a form's analyses, \% correct
+				%\item \textbf{Recall} - \% of analyses provided by transducer that are correct for a form, by comparing against a gold standard
 			\end{itemize}

### 2014-06-18 kypchak-morphfest-2014/poster.tex
valid
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@54594 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 %				Recall is the probability that some correct analysis is among those provided.
-				\item \textbf{Precision} - probability that a provided analysis is accurate
-				\item \textbf{Recall} - probability that a certain correct analysis is among those provided by the transducer
+				\item \textbf{Precision} - probability that a provided analysis is valid
+				\item \textbf{Recall} - probability that a certain valid analysis is among those provided by the transducer
 				%\item \textbf{Precision} - of a form's analyses, \% correct

### 2014-06-19 kypchak-morphfest-2014/poster.tex
last night's work
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@54635 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 				\item Our approach:\ categorisation 
-				\item only correct forms are analysed and generated
+				\item if properly categorised, only correct forms are analysed and generated
 			\end{itemize}
@@
 		}
+
+		\htwo{Adverbs}
+			\begin{itemize}
+				\item Certain time adverbs have special attributive and genetive forms\\
+				бүгүн, быйыл, кече, жана, 
+			\end{itemize}
+
 	}
@@
 
+		\headerbox{Other uses}{name=uses,column=1,below=morphophon}{
+			\begin{itemize}
+				\item HFST transducers are trivially converted to \textbf{spell checkers}
+				\item Segmenter, e.g.\ {\qipa көрүшкөндөрүмдөнсүңбү} :\\
+				 {\small \texttt{көр\hilitegrey{>}\{I\}ш\hilitegrey{>}\{G\}\{A\}н\hilitegrey{>}\{L\}\{A\}р\hilitegrey{>}\{I\}м\hilitegrey{>}\{D\}\{A\}н\hilitegrey{>}с\{I\}ң\hilitegrey{>}\{B\}\{I\}}}
+			\end{itemize}
+		}
+
 		\headerbox{Morphotactics}{name=lexc,column=2,below=exout}{

### 2014-06-20 kypchak-morphfest-2014/poster.tex
a complete draft
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@54688 72bbbca6-d526-0410-a7d9-f06f51895060

--- 2014-06-18 kypchak-morphfest-2014/poster.tex
@@
 			%\item Kyrgyz transducer reached 80\% coverage summer of 2011
+			\item (various periods of intermission, various rewrites)
 			\item Kazakh transducer based on Kyrgyz transducer
-			\item Kumyk transducer based on Kazakh, Tatar transducers
-			\item \tilda{}1 week to reach 80\% coverage, +1 week to reach 90\%
+			\item Kyrgyz transducer currently being rewritten based on insights gained while writing other Turkic transducers
+			\item Kumyk transducer based on Kazakh, Tatar transducers:\\
+			\tilda{}1 week to reach 80\% coverage, +1 week to reach 90\%
 		\end{itemize}
@@
 				\item morphologically distinct adjective classes
-				\item most sources:\ adjectives can be used substantively and adverbially
+				\item most sources claim:\ adjectives can be used substantively and adverbially
 				\item Other Turkic transducers:\ \texttt{0}-derivation (\hilitetwo{overgenerates})
@@
 			\end{tabular}
+		}
+
+		\htwo{Adverbs}
+			\begin{itemize}
+				\item Certain adverbs have special attributive and ablative forms
+				\item Mostly time adverbs
+				%\item Mostly time adverbs, e.g.\\
+				%{\qipa бүгүн} \eng{today}, {\qipa быйыл} \eng{this year}, {\qipa кечээ} \eng{yesterday}, {\qipa жана} \eng{just now}
+				%\item Attributive forms:\\
+				%{\qipa бүгүнкү} \eng{today's}, {\qipa быйылкы} \eng{this year's}, {\qipa кечээги} \eng{yesterday's}, {\qipa жанагы} \eng{just now}
+				%\item Ablative forms:\\
+				%{\qipa бүгүнтөн} \eng{from today}, {\qipa быйылтан} \eng{from this year}, {\qipa кечээтен} \eng{from yesterday}, {\qipa жанатан} \eng{from just now}
+				\item Some also have noun readings:\ regular ablative, other cases:
+				%{\qipa бүгүндөн} \eng{from today}, {\qipa быйылдан} \eng{from this year}
+			\end{itemize}
+			\scalebox{0.83}{
+			\begin{tabular}{lllll}
+				\toprule
+				word & {\qipa бүгүн} & {\qipa быйыл} & {\qipa кечээ} &  {\qipa жана}\\
+				gloss & \eng{today} & \eng{this year} & \eng{yesterday} & \eng{just now} \\
+				\midrule
+				\tag{attr} form & {\qipa бүгүнкү} & {\qipa быйылкы} & {\qipa кечээги} &  {\qipa жанагы}\\
+				\tag{adv}\tag{abl} form & {\qipa бүгүнтөн} & {\qipa быйылтан} & {\qipa кечээтен} &  {\qipa жанатан}\\
+				\tag{n}\tag{abl} form & {\qipa бүгүндөн} & {\qipa быйылдан} & — & — \\
+				\bottomrule
+			\end{tabular}
+			}\vspace{0.2em}
+				\codeex{
+					\begin{spacing}{0.8}
+					\texttt{LEXICON ADV-WITH-KI-ABL\\
+\\
+ADV ;\\
+ADV-KI ;\\
+ADV-ABL ;
+					}
+					\end{spacing}
+					\vspace{0em}
 				}
@@
 				\midrule
-					Kyrgyz & 500 &  &  \\
+					Kyrgyz & 200 & 90.77 & 69.15 \\
 					Kazakh & 1000 & 98.61 &  57.98 \\
@@
 
+		\headerbox{Other uses}{name=uses,column=1,below=morphophon}{
+			\begin{itemize}
+				\item HFST transducers are trivially converted to \textbf{spell checkers}
+				\item Segmenter, e.g.\ {\qipa көргөзгөндөрдөнсүңбү} :\\
+				 {\small \texttt{көр\hilitegrey{>}\{G\}\{A\}з\hilitegrey{>}\{G\}\{A\}н\hilitegrey{>}\{L\}\{A\}р\hilitegrey{>}\{D\}\{A\}н\hilitegrey{>}с\{I\}ң\hilitegrey{>}\{B\}\{I\}}}
+
+				 %көр\hilitegrey{>}\{I\}ш\hilitegrey{>}\{G\}\{A\}н\hilitegrey{>}\{L\}\{A\}р\hilitegrey{>}\{I\}м\hilitegrey{>}\{D\}\{A\}н\hilitegrey{>}с\{I\}ң\hilitegrey{>}\{B\}\{I\}}}
+			\end{itemize}
+		}
+
 		\headerbox{Morphotactics}{name=lexc,column=2,below=exout}{

### 2014-06-27 codeswitch-cltw-2014/codeswitch_camera_ready.tex
aff.
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@55017 72bbbca6-d526-0410-a7d9-f06f51895060
3.1 correction review-1.2
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@55018 72bbbca6-d526-0410-a7d9-f06f51895060

--- 2014-05-17 codeswitch-cltw-2014/codeswitch.tex
@@
 \author{}
-%\author{Akshay Minocha \\
-%  IIIT Hyderabad  \\
-%  Hyderabad (India) \\
-%  {\small {\tt akshay.minocha@students.iiit.ac.in}} \\\And
-%  Francis M. Tyers \\
-%  Giellatekno / CLEAR \\
-%  UiT Norgga \'arktala\v{s} universitehta  \\
-%  9017 Romsa (Norway) \\
-%  {\small {\tt francis.tyers@uit.no}} \\}
+\author{Akshay Minocha \\
+  IIIT Hyderabad  \\
+  Hyderabad (India) \\
+  {\small {\tt akshay.minocha@students.iiit.ac.in}} \\\And
+  Francis M. Tyers \\
+  Giellatekno / CLEAR \\
+  UiT Norgga \'arktala\v{s} universitehta  \\
+  9017 Romsa (Norway) \\
+  {\small {\tt francis.tyers@uit.no}} \\}
 
@@
 and French a model was trained using the EuroParl corpus \cite{koehn2005europarl}. 
-For Breton, Welsh and Irish we used corpora of text crawled from the web. For each of the languages we sampled 1.5 million words. \\
+For Breton, Welsh and Irish we used corpora of text crawled from the web.
+For each of the languages we sampled 1.5 million words.
+The accuracy of the results saturate after the size of the dataset is around 1 million, and also since our collected corpus for Irish was around 1.5 million words we chose the limit of the sample to be this which also ensures no bias in the sampling of the languages.
+\\
 In order to build a character language model as opposed to the standard word-ngram based model we replaced spaces with the underscore 
-symbol `\_', and then placed a space chracter between each character. For example, `sl\'{a}inte' would be broken down into a sequence 
-of \{`\_ s', `s l', `l \'{a}', `\'{a} i', `i n', `n t', `t e', `e \_'\}. Then the probability for the alphabet sequence is calculated 
-for each language to be used in the experiment. 
+symbol `\_', and then placed a space character between each character.
+Punctuations and non letters are also part of this language model as we didn't mean to preserve the identity of the language being used in the original form.
+For example, `sl\'{a}inte' would be broken down into a sequence
+of \{`\_ s', `s l', `l \'{a}', `\'{a} i', `i n', `n t', `t e', `e \_'\}.
+
+%Then the probability for the alphabet sequence is calculated 
+%for each language to be used in the experiment. 
 %Improved Kneser-Ney is used for smoothing, and this smoothing approach would be useful 

### 2014-06-27 codeswitch-cltw-2014/codeswitch_camera_ready.tex
evaluation review 1.4
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@55020 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 Certain tokens were escaped from the data, such as mentioned (which are preceded by an \texttt{@} symbol), subject tags `hashtags' which are preceded by a \texttt{\#} symbol), hyperlinks and the sequence \texttt{rt} which stands for `re-tweet'.
-An example of the content of our corpus after annotation is given in Figure~\ref{fig:tweets}.
+An example of the content of our corpus after hand annotation is given in Figure~\ref{fig:tweets}.
 All of the tweets had at least one instance of code-switching.
@@
 \end{small}
-\caption{Example of text from a microblogging site chunked}
+\caption{Example of text from a microblogging site chunked manually.}
 \label{fig:tweets}

### 2014-06-27 codeswitch-cltw-2014/codeswitch_camera_ready.tex
evaluation review 2.7
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@55021 72bbbca6-d526-0410-a7d9-f06f51895060
review 2.2
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@55022 72bbbca6-d526-0410-a7d9-f06f51895060

@@
-%Try1: akshay minocha
+%Try2: akshay minocha
 %
@@
 has been done on detecting code-switched segments in speech data \cite{chan2004detection,lyu2006language}.
-It is seen that language modelling techniques have shown promise earlier, such as in \newcite{yu2013identification} 
+It is seen that language modelling techniques have shown promise earlier, such as in \newcite{yu2013identification}, 
 the experiment on Mandarin-Taiwanese code-switched sentences show a high accuracy in terms of detecting code-switched sentences. \\
 
-\newcite{chan2004detection} has made use of the bi-phone probabilities and calculated them to measure a confidence metric, 
+In \newcite{chan2004detection} the authors have made use of the bi-phone probabilities and calculated them to measure a confidence metric, 
 to \cite{lyu2006language} which has made use of a classification method, named syllable-based duration classification, 
-which uses the tonal syllable properties along with the speech signals to help predict the code switch points. \newcite{yeong2010language} 
-uses syllable structure information to identify words in code-switched text in Malay-English, however they did not 
+which uses the tonal syllable properties along with the speech signals to help predict the code switch points. In \newcite{yeong2010language} 
+ the authors use syllable structure information to identify words in code-switched text in Malay-English, however they did not 
 recognise segments in running text, only identifying individual words.
@@
 We hand-annotated a small evaluation set from a selection of posts to a popular microblogging site\footnote{\url{http://indigenoustweets.blogspot.in/2013/12/mapping-celtic-twittersphere.html}}.
-The posts `tweets' were filtered into three sets which had been identified as Irish, Welsh and Breton.
+The posts `tweets' were filtered into three sets which had been identified as Irish, Welsh and Breton using the langid tool.
 Certain tokens were escaped from the data, such as mentioned (which are preceded by an \texttt{@} symbol), subject tags `hashtags' which are preceded by a \texttt{\#} symbol), hyperlinks and the sequence \texttt{rt} which stands for `re-tweet'.

### 2014-06-27 codeswitch-cltw-2014/codeswitch_camera_ready.tex
review 2.8
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@55023 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-After restricting the predicted languages to be amongst the two in the dataset, we used the character bigram and then trigram probabilities to predict the detected language for each token, some rules like in section~\ref{sec:method} were followed which included that the span of a segment or a code-switched phrase is more than a word, the social media non-language entities such as the hashtags, usernames and urls do not make a difference.
+Restricting the predicted languages to be amongst the two in the dataset was a decision made to identify the language pairs in a better fashion, both by our techniques and by the competitive tools too.
+Then we used the character bigram and then trigram probabilities to predict the detected language for each token.
+Rules in section~\ref{sec:method} were followed for the text processing and chunking part of the method.
+%which included that the span of a segment or a code-switched phrase is more than a word, the social media non-language entities such as the hashtags, usernames and urls do not make a difference.
 

### 2014-06-27 codeswitch-cltw-2014/codeswitch_camera_ready.tex
review 2.10 table correction
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@55024 72bbbca6-d526-0410-a7d9-f06f51895060
review 2.10 table correction
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@55025 72bbbca6-d526-0410-a7d9-f06f51895060
+
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@55026 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 Restricting the predicted languages to be amongst the two in the dataset was a decision made to identify the language pairs in a better fashion, both by our techniques and by the competitive tools too.
-Then we used the character bigram and then trigram probabilities to predict the detected language for each token.
+Then we used the character bigram and then character trigram probabilities to predict the detected language for each token.
 Rules in section~\ref{sec:method} were followed for the text processing and chunking part of the method.
@@
 \label{sec:results}
-As described in the ~\ref{sec:method} the data collected from Twitter for the three language pairs, was processed using the techniques mentioned.
+As described in ~\ref{sec:method} the data collected from Twitter for the three language pairs, was processed using the techniques mentioned.
 The statistics of the same are given in Table~\ref{table:datastats}.
@@
 \hline
-\multirow{2}{*}{\texttt{langid2}}         &  $p$ &  0.00   & 7.14     & 0.0   & 18.18    & 0.0 & 28.30 \\
+\multirow{2}{*}{\texttt{langid-2character}}         &  $p$ &  0.00   & 7.14     & 0.0   & 18.18    & 0.0 & 28.30 \\
                                           & $r$  & 0.00    & 5.00     & 0.0   & 14.12 & 0.0 & 18.99 \\
 \hline
-\multirow{2}{*}{\texttt{langid3}}         &  $p$ &  5.00   & 14.29    & 0.0   & 21.21 & 1.85 & 20.75 \\
+\multirow{2}{*}{\texttt{langid-3character}}         &  $p$ &  5.00   & 14.29    & 0.0   & 21.21 & 1.85 & 20.75 \\
                                           & $r$  & 5.41    & 8.45     & 0.0   & 14.58 & 1.92 & 12.36 \\
@@
 \hline
-\multirow{2}{*}{\texttt{bigram}}          &  $p$ &  32.50   & 35.71   & 23.44 & 19.70  & 57.41 & 52.83 \\
+\multirow{2}{*}{\texttt{character bigram}}          &  $p$ &  32.50   & 35.71   & 23.44 & 19.70  & 57.41 & 52.83 \\
                                           & $r$  & 22.41    & 26.79   & 15.31 & 16.67 & 41.33 & 37.84 \\
 \hline
-\multirow{2}{*}{\texttt{wordlist+bigram}} &  $p$ &  {\bf 52.50}   & {\bf 50.00}   & {\bf 32.81} & 31.82 & {\bf 70.37} & {\bf 67.92} \\
+\multirow{2}{*}{\texttt{wordlist+character bigram}} &  $p$ &  {\bf 52.50}   & {\bf 50.00}   & {\bf 32.81} & 31.82 & {\bf 70.37} & {\bf 67.92} \\
                                           & $r$  & {\bf 38.18}    & {\bf 43.75}   & 24.14 & 25.61 & {\bf 57.58} & {\bf 57.14} \\
@@
 \hline
-\texttt{langid2} & 50.35 & 42.28 & 45.11  \\
+\texttt{langid-2character} & 50.35 & 42.28 & 45.11  \\
 \hline
-\texttt{langid3} & 57.24 & 45.92 & 43.16 \\
+\texttt{langid-3character} & 57.24 & 45.92 & 43.16 \\
 \hline
@@
 \hline
-\texttt{bigram} & 81.29 & 65.62 & 76.79 \\
+\texttt{character bigram} & 81.29 & 65.62 & 76.79 \\
 \hline
-\texttt{wordlist+bigram} & \textbf{85.79} & 72.40 & \textbf{88.79} \\
+\texttt{wordlist+character bigram} & \textbf{85.79} & 72.40 & \textbf{88.79} \\
 \hline

### 2014-06-27 codeswitch-cltw-2014/codeswitch_camera_ready.tex
some fixes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@55029 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \begin{abstract}
-  Bilingual communities contribute majorly to the code-switching phenomena, where speakers produce sentences containing words and expressions from a second language.
+  This paper describes an experiment to perform language identification on a sub-sentence basis. The typical
+  case of language identification is to detect the language of documents or sentences. However, it may be the 
+  case that a single sentence or segment contains more than one language. This is especially the case in texts
+  where \emph{code switching} occurs.
+
+%  Bilingual communities contribute majorly to the code-switching phenomena, where speakers produce sentences containing words and expressions from a second language.
 With the increase in the data on the web, specially on informal platforms like that of the social media, it becomes increasingly important to detect such irregularities and handle them appropriately.
@@
 There are several ways in which an informal text may contain parts in different languages, such as
-code switching, quotations, named entities, interjections, translations, etc. Some of these are presented
-in Table~\ref{table:examples}.
+code switching, quotations, named entities, interjections, translations --- either of the full text, or 
+part of the text, comments, etc. Some of these are presented in Table~\ref{table:examples}.
 
@@
 \label{sec:method}
+
+\subsection{Corpus}
+
+\begin{table}
+\begin{center}
+\begin{tabular}{|l|r|r|r|r}
+\hline
+\multirow{2}{*}{\textbf{Pair}} & \multirow{2}{*}{\textbf{Language}}  & \multicolumn{2}{c|}{\textbf{Statistics} (\%)} \\\cline{3-4}
+                &  &  Tokens & Segments \\
+\hline
+\multirow{2}{*}{Irish---English} & Irish & 332 & 40 \\
+                                 & English & 379 & 42 \\
+\hline
+\multirow{2}{*}{Welsh---English} & Welsh & 419 & 64 \\
+                                 & English & 378 & 66  \\
+\hline
+\multirow{2}{*}{Breton---French} & Breton & 388 & 54 \\
+                                 & French & 379 & 53  \\
+\hline
+\end{tabular}
+\end{center}
+\caption{Document statistics of the annotated data used. }
+\label{table:datastats}
+\end{table}
+
+We hand-annotated a small evaluation set from a selection of posts to a popular microblogging site.\footnote{\url{http://indigenoustweets.blogspot.in/2013/12/mapping-celtic-twittersphere.html}} The \emph{tweets} (microblog posts) were filtered into three sets which had been identified as Irish, Welsh and Breton using the \emph{langid} tool \cite{lui2012langid}.
+From these, we manually selected between 40 and 50 tweets for each language pair.
+Statistics on the number of segments and tokens is presented in Table~\ref{table:datastats}.
+Certain tokens were escaped from the data, such as the `mentioned' character (\texttt{@} symbol), subject tags `hashtags' which are preceded by a \texttt{\#} symbol, hyperlinks and the sequence \texttt{rt} which stands for `re-tweet'.
+An example of the content of our corpus after hand annotation is given in Figure~\ref{fig:tweets}.
+All of the tweets had at least one instance of code-switching.
+
+\begin{figure}
+\begin{small}
+\begin{alltt}
+[\textbf{en} You're a] [\textbf{ga} Meirice\'{a}nach, c\'{e}n f\'{a}th] [\textbf{en} are you] [\textbf{ga} foghlaim Gaeilge?!] 
+@afaltomkins [\textbf{cy} gorfod cael bach o tan] [\textbf{en} though init] 
+[\textbf{en} omg] [\textbf{cy} mar cwn bach yn] [\textbf{en} black and tan] [\textbf{cy} a popeth,] [\textbf{en} even cuter!!] 
+\end{alltt}
+\end{small}
+\caption{Example of text from a microblogging site chunked manually.}
+\label{fig:tweets}
+\end{figure}
+
+\subsection{Alphabet n-gram approach}
+\label{alphan}
+
 We use the character n-gram approach along with some heuristics which are relevant to our problem domain of identifying
@@
 
-\subsection{Alphabet n-gram approach}
-\label{alphan}
-
 We first built character language models using IRSTLM \cite{federico08a} for the five languages in question. For English
@@
 For each of the languages we sampled 1.5 million words.
-The accuracy of the results saturate after the size of the dataset is around 1 million, and also since our collected corpus for Irish was around 1.5 million words we chose the limit of the sample to be this which also ensures no bias in the sampling of the languages.
+The accuracy of the results saturate after the size of the dataset is around 1 million, and also since our collected corpus for Breton was around 1.5 million words we chose the limit of the sample to be this which also ensures no bias in the sampling of the languages.
 \\
 In order to build a character language model as opposed to the standard word-ngram based model we replaced spaces with the underscore 
-symbol `\_', and then placed a space character between each character.
-Punctuations and non letters are also part of this language model as we didn't mean to preserve the identity of the language being used in the original form.
-For example, `sl\'{a}inte' would be broken down into a sequence
-of \{`\_ s', `s l', `l \'{a}', `\'{a} i', `i n', `n t', `t e', `e \_'\}.
+symbol `\_', and then placed a space character between each character. Punctuation and non-letter characters are also 
+part of this language model. For example, the word `sl\'{a}inte!' would be broken down into a sequence 
+of \{`\_ s', `s l', `l \'{a}', `\'{a} i', `i n', `n t', `t e', `e !', `! \_'\}.
 
@@
 
-We hand-annotated a small evaluation set from a selection of posts to a popular microblogging site\footnote{\url{http://indigenoustweets.blogspot.in/2013/12/mapping-celtic-twittersphere.html}}.
-The posts `tweets' were filtered into three sets which had been identified as Irish, Welsh and Breton using the langid tool.
-Certain tokens were escaped from the data, such as mentioned (which are preceded by an \texttt{@} symbol), subject tags `hashtags' which are preceded by a \texttt{\#} symbol), hyperlinks and the sequence \texttt{rt} which stands for `re-tweet'.
-An example of the content of our corpus after hand annotation is given in Figure~\ref{fig:tweets}.
-All of the tweets had at least one instance of code-switching.
-
-For the Evaluation procedure, we follow the footsteps of the CoNLL-2000 shared task on language-independent 
+or the Evaluation procedure, we follow the footsteps of the CoNLL-2000 shared task on language-independent 
 named entity recognition: dividing text into syntactically related non-overlapping groups of words. This 
@@
 here are such that they belong to only one language.
-
-\begin{figure}
-\begin{small}
-\begin{alltt}
-[\textbf{en} You're a] [\textbf{ga} Meirice\'{a}nach, c\'{e}n f\'{a}th] [\textbf{en} are you] [\textbf{ga} foghlaim Gaeilge?!] 
-@afaltomkins [\textbf{cy} gorfod cael bach o tan] [\textbf{en} though init] 
-[\textbf{en} omg] [\textbf{cy} mar cwn bach yn] [\textbf{en} black and tan] [\textbf{cy} a popeth,] [\textbf{en} even cuter!!] 
-\end{alltt}
-\end{small}
-\caption{Example of text from a microblogging site chunked manually.}
-\label{fig:tweets}
-\end{figure}
 
@@
 However it does provide a reasonable baseline for the per-word accuracy.
-\\
 
-\subsection{\texttt{langid} character bigram and trigram prediction}
+\subsection{\texttt{langid} character trigram prediction}
 \label{langidstuff}
 
-Restricting the predicted languages to be amongst the two in the dataset was a decision made to identify the language pairs in a better fashion, both by our techniques and by the competitive tools too.
-Then we used the character bigram and then character trigram probabilities to predict the detected language for each token.
-Rules in section~\ref{sec:method} were followed for the text processing and chunking part of the method.
-%which included that the span of a segment or a code-switched phrase is more than a word, the social media non-language entities such as the hashtags, usernames and urls do not make a difference.
+Restricting the predicted languages to be amongst the two in the dataset was a decision made to identify the language pairs in a better fashion, both by our techniques and by the other tools.
+Then we used the character trigram probabilities to predict the detected language for each token.
+Trigrams were chosen after experimenting with 1--5 grams.
+The heuristics in section~\ref{sec:method} were followed for the text processing and chunking part of the method.
 
@@
 The statistics of the same are given in Table~\ref{table:datastats}.
-//
-
-\begin{table}
-\begin{center}
-\begin{tabular}{|l|r|r|r|r}
-\hline
-\multirow{2}{*}{\textbf{Pair}} & \multirow{2}{*}{\textbf{Language}}  & \multicolumn{2}{c|}{\textbf{Statistics} (\%)} \\\cline{3-4}
-                &  &  Tokens & Segments \\
-\hline
-\multirow{2}{*}{Irish---English} & Irish & 332 & 40 \\
-                                 & English & 379 & 42 \\
-\hline
-\multirow{2}{*}{Welsh---English} & Welsh & 419 & 64 \\
-                                 & English & 378 & 66  \\
-\hline
-\multirow{2}{*}{Breton---French} & Breton & 388 & 54 \\
-                                 & French & 379 & 53  \\
-\hline
-\end{tabular}
-\end{center}
-\caption{Document statistics of the annotated data used. }
-\label{table:datastats}
-\end{table}
 
@@
                                           & $r$  & 2.56    & 0.0      & 0.0   & 0.0 & 0.0 & 0.0 \\
-\hline
-\multirow{2}{*}{\texttt{langid-2character}}         &  $p$ &  0.00   & 7.14     & 0.0   & 18.18    & 0.0 & 28.30 \\
-                                          & $r$  & 0.00    & 5.00     & 0.0   & 14.12 & 0.0 & 18.99 \\
 \hline
@@
 \hline
-\texttt{langid-2character} & 50.35 & 42.28 & 45.11  \\
-\hline
 \texttt{langid-3character} & 57.24 & 45.92 & 43.16 \\
@@
 This paper has presented a very preliminary investigation into subsegment language identification in Celtic language texts. We have
-proposed a model that chunks input text into segments, and performs language identification on these segments at the same time. Precision and
+proposed a model that chunks input text into segments and performs language identification on these segments at the same time. Precision and
 recall are low, leaving a lot of room for further work. Some of our plans are as follows. We would like to annotate more test data,
@@
 
-%We thank Kevin Scannell for help with providing data from Twitter and for useful comments and suggestions.
+We thank Kevin Scannell for help with providing data from Twitter and for useful comments and suggestions. This work
+was undertaken as part of the Apertium project in the 2014 Google Summer of Code.
 

### 2014-06-27 codeswitch-cltw-2014/codeswitch_camera_ready.tex
update
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@55030 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \label{sec:method}
+
+As the number of possible languages for each segment is in theory the set of all the world's written languages, we take 
+a decision to simplify the task by only looking at texts in the Celtic language and the corresponding majority language spoken
+where the Celtic language is spoken. That is, we looked at detecting between Irish and English, Welsh and English 
+and Breton and French.
 
@@
 
-Restricting the predicted languages to be amongst the two in the dataset was a decision made to identify the language pairs in a better fashion, both by our techniques and by the other tools.
-Then we used the character trigram probabilities to predict the detected language for each token.
+For this system we used the character trigram probabilities to predict the detected language for each token.
 Trigrams were chosen after experimenting with 1--5 grams.

### 2014-06-27 codeswitch-cltw-2014/codeswitch_camera_ready.tex
bib
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@55037 72bbbca6-d526-0410-a7d9-f06f51895060
+
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@55041 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-Apart from the techniques discussed in section~\ref{sec:method}, some baselines are also used to give a comparative view of how well all the mechanism perform.
+Apart from the techniques discussed in section~\ref{sec:method}, some baselines are also used to give a comparative view of how well all the mechanisms perform.
 
@@
 \bibliographystyle{acl}
-\bibliography{codeswitch}
+\bibliography{codeswitch_camera_ready}
 

### 2014-06-28 codeswitch-cltw-2014/codeswitch_camera_ready.tex
cc4 licence
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@55049 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 presents future work and conclusions.
+\blfootnote{
+    %
+    % for review submission
+    %
+    %\hspace{-0.65cm}  % space normally used by the marker
+    %Place licence statement here for the camera-ready version, see
+    %Section~\ref{licence} of the instructions for preparing a
+    %manuscript.
+    %
+     % final paper: en-uk version (to license, a licence)
+
+     \hspace{-0.65cm}  % space normally used by the marker
+     This work is licensed under a Creative Commons 
+     Attribution 4.0 International Licence.
+     Page numbers and proceedings footer are added by
+     the organisers.
+     Licence details:
+     \url{http://creativecommons.org/licenses/by/4.0/}
+
+    % % final paper: en-us version (to licence, a license)
+    %
+    % \hspace{-0.65cm}  % space normally used by the marker
+    % This work is licenced under a Creative Commons 
+    % Attribution 4.0 International License.
+    % Page numbers and proceedings footer are added by
+    % the organizers.
+    % License details:
+    % \url{http://creativecommons.org/licenses/by/4.0/}
+}
 
@@
 
-\blfootnote{
-    %
-    % for review submission
-    %
-    \hspace{-0.65cm}  % space normally used by the marker
-    Place licence statement here for the camera-ready version, see
-    Section~\ref{licence} of the instructions for preparing a
-    manuscript.
-    %
-    % % final paper: en-uk version (to license, a licence)
-    %
-    % \hspace{-0.65cm}  % space normally used by the marker
-    % This work is licensed under a Creative Commons 
-    % Attribution 4.0 International Licence.
-    % Page numbers and proceedings footer are added by
-    % the organisers.
-    % Licence details:
-    % \url{http://creativecommons.org/licenses/by/4.0/}
-    % 
-    % % final paper: en-us version (to licence, a license)
-    %
-    % \hspace{-0.65cm}  % space normally used by the marker
-    % This work is licenced under a Creative Commons 
-    % Attribution 4.0 International License.
-    % Page numbers and proceedings footer are added by
-    % the organizers.
-    % License details:
-    % \url{http://creativecommons.org/licenses/by/4.0/}
-}
-
 \section{Evaluation}

### 2014-06-28 codeswitch-cltw-2014/codeswitch_camera_ready.tex
conclusions
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@55051 72bbbca6-d526-0410-a7d9-f06f51895060
chunking re written .. pseducode needs to be added
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@55052 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 the sequences of segments in different languages. 
+'Chunks' represent the evaluated tuples of segments and their labelled language, buffer is the expanding segment.
+ \textit{lang\_predict} corressponds to any function which is used to determine the language of the token. The flag 
+variable helps in implementing the heuristic of minimum segment size while labelling chunks.
 
-A flag is set at the beginning of the input text, and this moves forward to the position in the text upto where confident 
-chunking has been performed. For example, in case of a continuous prediction there would be no change and since the 
-predicted tag of the new word corresponds to that already of the expanding chunk, this word is also included in the 
-same chunk. In case of a doubt, the flag is set at the same position, and no final decision on labelling the word is 
-performed, to have more confidence on the code-switch, the process moves on to the next token, if it corresponds to 
-that of the previous chunk, then both these undecided words are labelled and included in the chunk, otherwise a change 
-is noted and a new chunk is created with the changed label. In this process, we are trying to identify the sub-segments 
-that we need, thus performing both tasks at once, the first being language prediction and second being the chunking decision.
+%A flag is set at the beginning of the input text, and this moves forward to the position in the text upto where confident 
+%chunking has been performed. For example, in case of a continuous prediction there would be no change and since the 
+%predicted tag of the new word corresponds to that already of the expanding chunk, this word is also included in the 
+%same chunk. In case of a doubt, the flag is set at the same position, and no final decision on labelling the word is 
+%performed, to have more confidence on the code-switch, the process moves on to the next token, if it corresponds to 
+%that of the previous chunk, then both these undecided words are labelled and included in the chunk, otherwise a change 
+%is noted and a new chunk is created with the changed label. In this process, we are trying to identify the sub-segments 
+%that we need, thus performing both tasks at once, the first being language prediction and second being the chunking decision.
 
@@
 recall are low, leaving a lot of room for further work. Some of our plans are as follows. We would like to annotate more test data,
-at least 100 tweets in each language. We would also like to attempt our method using higher order character n-gram models for backoff, and
-n-gram word language models for detection.
+at least 100 tweets in each language.
+Although In \cite{king2013labeling}, the authors label on a per word level, yet we would like to include supervised methods and features talked about in this research to improve our efficiency while dealing with segments.
+We would also like to attempt our method using higher order character n-gram models for backoff, and n-gram word language models for detection.
 

### 2014-06-28 codeswitch-cltw-2014/codeswitch_camera_ready.tex
updated psedocode
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@55053 72bbbca6-d526-0410-a7d9-f06f51895060
more changes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@55055 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{multirow}
+
+\usepackage{amsmath}
+\usepackage{algorithm}
+\usepackage[noend]{algpseudocode}
 
@@
 \label{intro}
-
+\vspace{-0.1cm}
 Determining the language of a piece of text is one of the first steps that must be taken
@@
 \begin{center}
+\vspace{-0.3cm}
 \begin{tabular}{ll}
@@
 \end{table}
+\vspace{-0.6cm}
 
@@
 It is seen that language modelling techniques have shown promise earlier, such as in \newcite{yu2013identification}, 
-the experiment on Mandarin-Taiwanese code-switched sentences show a high accuracy in terms of detecting code-switched sentences. \\
+the experiment on Mandarin-Taiwanese sentences show a high accuracy in terms of detecting code-switched sentences. \\
 
@@
 For Breton, Welsh and Irish we used corpora of text crawled from the web. 
-For each of the languages we sampled 1.5 million words.
+%For each of the languages we sampled 1.5 million words.
 The accuracy of the results saturate after the size of the dataset is around 1 million, and also since our collected corpus for Breton was around 1.5 million words we chose the limit of the sample to be this which also ensures no bias in the sampling of the languages.
@@
 variable helps in implementing the heuristic of minimum segment size while labelling chunks.
+
+\begin{algorithm}
+\caption{Chunking algorithm}\label{euclid}
+\begin{algorithmic}[1]
+%\Procedure{}{}
+\State $\textit{buffer} = [\ ] \text{\ \ \  \#undecided expanding window of chunk}$
+\State $\textit{chunks} = [\ ] \text{\ \ \  \#decided labelled segment}$
+\State $\textit{buffer\_language} \gets \textit{lang\_predict}\text{(sentence[0])} \text{\ \ \ \#Language of first word}$
+\State $\textit{flag} \gets \text{0}$
+%\State $\text{for \ } \textit{word \ } \text{in \ } \textit{sentence:} $
+\For{\texttt{word in sentence}}
+\If {$ \textit{lang\_predict(word)} \text{==} \textit{buffer\_language} $}
+        \If {$ \textit{flag} \text{=\ =\ 1} $}
+        \State $\textit{buffer} += [\textit{word\_d,word}] $
+        \State $\textit{flag} \gets \text{0}$
+        \Else
+        \State $\textit{buffer} += [\textit{word}] $
+        \EndIf
+        \EndIf
+\If {$ \textit{lang\_prediction(word)} \text{!=} \textit{buffer\_langugae} $}
+        \If {$ \textit{flag} \text{=\ =\ 0} $}
+        \State $\textit{flag} \gets \text{1}$
+        \State $\textit{word\_d} \gets \textit{word} $
+        \State $\text{continue}$
+
+        \Else
+        \State $\textit{chunk} \text{+= \ }  \textit{[(buffer,buffer\_language)]} $
+        %\State $\textit{buffer} \gets [] $
+        \State $\textit{buffer} = [\textit{word\_d,word}] $
+        \State $\textit{buffer\_language} = \gets \textit{lang\_predict}\text{(word)}$% \text{\#Language of new expanding chunk}$
+        \State $\textit{flag} \gets \text{0}$
+        \EndIf
+
+\EndIf
+\EndFor
+\If {$\textit{length(buffer)} \text{!=0\ } $}
+        \State $\textit{chunk} \text{+= \ }  \text{[\textit{[(buffer,buffer\_language)]}} $
+
+\EndIf
+%\EndProcedure
+\end{algorithmic}
+\end{algorithm}
 
@@
 \section{Conclusions}
+%\vspace{-0.2cm}
 \label{sec:conclusions}

### 2014-06-28 codeswitch-cltw-2014/codeswitch_camera_ready.tex
more changes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@55056 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-%
-% The following footnote without marker is needed for the camera-ready
-% version of the paper.
-% Comment out the instructions (first text) and uncomment the 8 lines
-% under "final paper" for your variant of English.
-%
-
 \section{Evaluation}
@@
 
-or the Evaluation procedure, we follow the footsteps of the CoNLL-2000 shared task on language-independent 
+For the Evaluation procedure, we follow the footsteps of the CoNLL-2000 shared task on language-independent 
 named entity recognition: dividing text into syntactically related non-overlapping groups of words. This 

### 2014-06-28 codeswitch-cltw-2014/codeswitch_camera_ready.tex
more changes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@55057 72bbbca6-d526-0410-a7d9-f06f51895060
more changes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@55060 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{algorithm}
+\usepackage{algcompatible}
 \usepackage[noend]{algpseudocode}
@@
 
-There are several ways in which an informal text may contain parts in different languages, such as
-code switching, quotations, named entities, interjections, translations --- either of the full text, or 
-part of the text, comments, etc. Some of these are presented in Table~\ref{table:examples}.
+%Such as
+%code switching, quotations, named entities, interjections, translations --- either of the full text, or 
+%part of the text, comments, etc.
 
@@
 are applying it to.
+
+There are several ways in which an informal text may contain parts in different languages.
+Some of these are presented in Table~\ref{table:examples}.
 
@@
 the sequences of segments in different languages. 
-'Chunks' represent the evaluated tuples of segments and their labelled language, buffer is the expanding segment.
+In Figure ~\ref{fig:chunkal}, `chunks' represent the list of evaluated tuples of segments and their labelled language, `buffer' is the expanding segment.
  \textit{lang\_predict} corressponds to any function which is used to determine the language of the token. The flag 
 variable helps in implementing the heuristic of minimum segment size while labelling chunks.
-
 \begin{algorithm}
-\caption{Chunking algorithm}\label{euclid}
+\caption*{\textsc{}}
+%\caption{Chunking algorithm}\label{euclid}
+{\fontsize{10}{10}\selectfont
 \begin{algorithmic}[1]
-%\Procedure{}{}
-\State $\textit{buffer} = [\ ] \text{\ \ \  \#undecided expanding window of chunk}$
-\State $\textit{chunks} = [\ ] \text{\ \ \  \#decided labelled segment}$
-\State $\textit{buffer\_language} \gets \textit{lang\_predict}\text{(sentence[0])} \text{\ \ \ \#Language of first word}$
-\State $\textit{flag} \gets \text{0}$
+%\Statex{}%Procedure{}{}
+\Statex $\textit{buffer} = [\ ] \text{\ \ \  \#undecided expanding window of chunk}$
+\Statex $\textit{chunks} = [\ ] \text{\ \ \  \#decided labelled segment}$
+\Statex $\textit{buffer\_language} \gets \textit{lang\_predict}\text{(sentence[0])} \text{\ \ \ \#Language of first word}$
+\Statex $\textit{flag} \gets \text{0}$
 %\State $\text{for \ } \textit{word \ } \text{in \ } \textit{sentence:} $
@@
         \If {$ \textit{flag} \text{=\ =\ 1} $}
-        \State $\textit{buffer} += [\textit{word\_d,word}] $
-        \State $\textit{flag} \gets \text{0}$
+        \Statex $\textit{buffer} += [\textit{word\_d,word}] $
+        \Statex $\textit{flag} \gets \text{0}$
         \Else
-        \State $\textit{buffer} += [\textit{word}] $
+        \Statex $\textit{buffer} += [\textit{word}] $
         \EndIf
@@
         \If {$ \textit{flag} \text{=\ =\ 0} $}
-        \State $\textit{flag} \gets \text{1}$
-        \State $\textit{word\_d} \gets \textit{word} $
-        \State $\text{continue}$
+        \Statex $\textit{flag} \gets \text{1}$
+        \Statex $\textit{word\_d} \gets \textit{word} $
+        \Statex $\text{continue}$
 
         \Else
-        \State $\textit{chunk} \text{+= \ }  \textit{[(buffer,buffer\_language)]} $
+        \Statex $\textit{chunk} \text{+= \ }  \textit{[(buffer,buffer\_language)]} $
         %\State $\textit{buffer} \gets [] $
-        \State $\textit{buffer} = [\textit{word\_d,word}] $
-        \State $\textit{buffer\_language} = \gets \textit{lang\_predict}\text{(word)}$% \text{\#Language of new expanding chunk}$
-        \State $\textit{flag} \gets \text{0}$
+        \Statex $\textit{buffer} = [\textit{word\_d,word}] $
+        \Statex $\textit{buffer\_language} = \gets \textit{lang\_predict}\text{(word)}$% \text{\#Language of new expanding chunk}$
+        \Statex $\textit{flag} \gets \text{0}$
         \EndIf
@@
 \If {$\textit{length(buffer)} \text{!=0\ } $}
-        \State $\textit{chunk} \text{+= \ }  \text{[\textit{[(buffer,buffer\_language)]}} $
+        \Statex $\textit{chunk} \text{+= \ }  \text{[\textit{[(buffer,buffer\_language)]}} $
 
@@
 %\EndProcedure
-\end{algorithmic}
+%\Statex{}
+\end{algorithmic}}
 \end{algorithm}
-
+\vspace{-0.4cm}
+\begin{figure}[hb]
+\def\@fs@post{}
+\caption{Chunking Algorithm}
+\label{fig:chunkal}
+\end{figure}
 %A flag is set at the beginning of the input text, and this moves forward to the position in the text upto where confident 
@@
 
+\vspace{-0.6cm}
 \subsection{Word-based prediction}
@@
 
+\vspace{-0.4cm}
+
 \section{Evaluation}
@@
 language datasets. The first, is the percentage of correctly detected phrases, which is the overall precision and the second is the 
-number of phrases in the data that were found by the chunker, which is the overall recall. \\
+number of phrases in the data that were found by the chunker, which is the overall recall.
 

### 2014-06-28 codeswitch-cltw-2014/codeswitch_camera_ready.tex
more changes+
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@55061 72bbbca6-d526-0410-a7d9-f06f51895060

--- 2014-06-28 codeswitch-cltw-2014/codeswitch_camera_ready.tex
@@
 It is seen that language modelling techniques have shown promise earlier, such as in \newcite{yu2013identification}, 
-the experiment on Mandarin-Taiwanese sentences show a high accuracy in terms of detecting code-switched sentences. \\
+the experiment on Mandarin-Taiwanese sentences show a high accuracy in terms of detecting code-switched sentences.
 
@@
 the sequences of segments in different languages. 
-'Chunks' represent the evaluated tuples of segments and their labelled language, buffer is the expanding segment.
+In Figure ~\ref{fig:chunkal}, `chunks' represent the list of evaluated tuples of segments and their labelled language, `buffer' is the expanding segment.
  \textit{lang\_predict} corressponds to any function which is used to determine the language of the token. The flag 
 variable helps in implementing the heuristic of minimum segment size while labelling chunks.
-
 \begin{algorithm}
-\caption{Chunking algorithm}\label{euclid}
+\caption*{\textsc{}}
+%\caption{Chunking algorithm}\label{euclid}
+{\fontsize{10}{10}\selectfont
 \begin{algorithmic}[1]
-%\Procedure{}{}
+%\Statex{}%Procedure{}{}
 \State $\textit{buffer} = [\ ] \text{\ \ \  \#undecided expanding window of chunk}$

### 2014-06-28 codeswitch-cltw-2014/codeswitch_camera_ready.tex
more changes--
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@55071 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \label{intro}
-\vspace{-0.1cm}
 Determining the language of a piece of text is one of the first steps that must be taken
@@
 where the speech community is predominantly bi- or multi-lingual, this assumption may not hold.
-
 %Such as
@@
 \end{table}
-\vspace{-0.6cm}
+\vspace{-0.65cm}
 
@@
 \begin{table}
+\vspace{-0.65cm}
 \begin{center}
@@
 \end{small}
+\vspace{-0.3cm}
 \caption{Example of text from a microblogging site chunked manually.}
@@
 For Breton, Welsh and Irish we used corpora of text crawled from the web.
+To ensure no bias and also since our dataset for Breton was around 1.5 million, we sampled the same size of data for all the five languages.
 %For each of the languages we sampled 1.5 million words.
-The accuracy of the results saturate after the size of the dataset is around 1 million, and also since our collected corpus for Breton was around 1.5 million words we chose the limit of the sample to be this which also ensures no bias in the sampling of the languages.
+%The accuracy of the results saturate after the size of the dataset is around 1 million, and also since our collected corpus for Breton was around 1.5 million words we chose the limit of the sample to be this which also ensures no bias in the sampling of the languages.
 \\
-In order to build a character language model as opposed to the standard word-ngram based model we replaced spaces with the underscore 
+In order to build a character language model %as opposed to the standard word-ngram based model %
+we replaced spaces with the underscore 
 symbol `\_', and then placed a space character between each character. Punctuation and non-letter characters are also 
@@
 
-\vspace{-0.6cm}
+\vspace{-0.8cm}
 \subsection{Word-based prediction}
@@
 
-\vspace{-0.4cm}
+\vspace{-0.2cm}
 
@@
 \label{sec:results}
-As described in ~\ref{sec:method} the data collected from Twitter for the three language pairs, was processed using the techniques mentioned.
+As described in Section~\ref{sec:method} the data collected from Twitter for the three language pairs, was processed using the techniques mentioned.
 The statistics of the same are given in Table~\ref{table:datastats}.
@@
 \end{center}
+\vspace{-0.4cm}
 \caption{Precision, $p$ and recall, $r$ for the systems by language.}

### 2014-06-28 codeswitch-cltw-2014/codeswitch_camera_ready.tex
more and more
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@55072 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 %\caption{Chunking algorithm}\label{euclid}
-{\fontsize{10}{10}\selectfont
+{\fontsize{9}{9}\selectfont
 \begin{algorithmic}[1]
@@
 \end{algorithm}
-\vspace{-0.4cm}
+\vspace{-0.6cm}
 \begin{figure}[hb]
@@
 
-This is a simple heuristic which was designed on the basis of the the most common words in the wordlists of the languages 
-which are in question. After checking each word against both of the word lists,\footnote{For the wordlists we used the \emph{aspell}
+Designed keeping in mind importance of the most common words, this procedure included 
+%This is a simple heuristic which was designed on the basis of the the most common words in the wordlists of the languages 
+%which are in question. After 
+checking each word against both of the word lists in question,\footnote{For the wordlists we used the \emph{aspell}
 wordlists widely available on Unix systems.} it is associated to one language or another. In 
@@
 
-For our current task, each of the texts have been limited to two languages i.e. the primary language (the Celtic language) and the 
-secondary language (the `majority' language). Hence the type of chunking tags are limited to these. The evaluation statistics shown 
+%For our current task, each of the texts have been limited to two languages i.e. the primary language (the Celtic language) and the 
+%secondary language. % (the `majority' language).
+%Hence the type of chunking tags are limited to these. 
+The evaluation statistics shown 
 in Tables~\ref{table:precisionrecall}~and~\ref{table:accuracy} mention two values for each of the experiment conducted on the three bilingual 
@@
 
-This is the most na\"{i}ve method of classification, we used the language identification tool \texttt{langid.py} \cite{lui2012langid} on the whole dataset and labelled all the individual lines according to this single majority classification.
+%This is the most na\"{i}ve method of classification, 
+We used the language identification tool \texttt{langid.py} \cite{lui2012langid} on the whole dataset and labelled all the individual lines according to this majority classification.
 As no chunking is performed we can expect that the precision and recall will be very low.
@@
 \section{Results}
+\vspace{-0.152cm}
 \label{sec:results}
@@
 \end{center}
+\vspace{-0.3cm}
 \caption{Accuracy of the systems over the three language pairs. The accuracy measures how often a token
@@
 \end{table}
-
 In terms of per-word accuracy, the baseline performs as expected, given that half of the words in the test set are from a given language. 
-The \texttt{langid.py} based methods allow for segmenting the text into chunks, but favour the majority language.\footnote{This could be because of the default models being used.} While the precision and recall are low for the remaining models, we see that we are able to improve
-the performance by combining the wordlist based model with a character bigram model. And what is more, we are able to begin to not only 
+The \texttt{langid.py} based methods allow for segmenting the text into chunks, but favour the majority language.\footnote{This could be because of the default models being used.} 
+%While the precision and recall are low for the remaining models, we see that we are able to improve 
+The performance by combining the wordlist based model with a character bigram model. And what is more, we are able to begin to not only 
 identify particular words in a language, but also segments.
@@
 \section{Conclusions}
-%\vspace{-0.2cm}
 \label{sec:conclusions}

### 2014-06-28 codeswitch-cltw-2014/codeswitch_camera_ready.tex
more and more and more
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@55074 72bbbca6-d526-0410-a7d9-f06f51895060

@@
   Francis M. Tyers \\
-  Giellatekno / CLEAR \\
+  Giellatekno \\ %/CLEAR \\
   UiT Norgga \'arktala\v{s} universitehta  \\
@@
 for the Celtic languages as a precursor to machine translation. When applying a normalisation 
-strategy to a piece of text, it is necessary to first know the language of the piece of text you 
-are applying it to.
-
-There are several ways in which an informal text may contain parts in different languages.
-Some of these are presented in Table~\ref{table:examples}.
+strategy to a piece of text, it is necessary to first know the language of the piece of text you are applying it to.
+%Table~\ref{table:examples} shows some examples of how code switching is present in text
+%There are several ways in which an informal text may contain parts in different languages.
+%Some of these are presented in Table~\ref{table:examples}.
 
@@
 
-\begin{table}[h]
-\begin{center}
-\vspace{-0.3cm}
-\begin{tabular}{ll}
-\textbf{Code switching}: & You're a [Meirice\'{a}nach, c\'{e}n f\'{a}th] are you [foghlaim Gaeilge?!] \\
-\textbf{Quotations}: & The anthem starts with the words [`Mae hen wlad fy nhadau\ldots'] \\
-\textbf{Named entities}: & [Dr Jekyll] ha [Mr Hyde] embannet gant [\'{E}ditions Aber] \\
-\textbf{Interjections}: & Hey, that's great, [diolch yn fawr!] \\
-\textbf{Translations}: & Bloavezh mat d'an holl ! [Bonne ann\'{e} \`{a} tous !] \\
-\end{tabular}
-\caption{Some examples of text segments containing more than one language}
-\label{table:examples}
-\end{center}
-\end{table}
-\vspace{-0.65cm}
+%\begin{table}[h]
+%\begin{center}
+%\vspace{-0.3cm}
+%%\begin{tabular}{ll}
+%\textbf{Code switching}: & You're a [Meirice\'{a}nach, c\'{e}n f\'{a}th] are you [foghlaim Gaeilge?!] \\
+%\textbf{Quotations}: & The anthem starts with the words [`Mae hen wlad fy nhadau\ldots'] \\
+%\textbf{Named entities}: & [Dr Jekyll] ha [Mr Hyde] embannet gant [\'{E}ditions Aber] \\
+%\textbf{Interjections}: & Hey, that's great, [diolch yn fawr!] \\
+%\textbf{Translations}: & Bloavezh mat d'an holl ! [Bonne ann\'{e} \`{a} tous !] \\
+%\end{tabular}
+%\caption{Some examples of text segments containing more than one language}
+%\label{table:examples}
+%\end{center}
+%\end{table}
+%\vspace{-0.65cm}
 
@@
 In \newcite{chan2004detection} the authors have made use of the bi-phone probabilities and calculated them to measure a confidence metric, 
-to \cite{lyu2006language} which has made use of a classification method, named syllable-based duration classification, 
+to \cite{lyu2006language} which has made use of named syllable-based duration classification, 
 which uses the tonal syllable properties along with the speech signals to help predict the code switch points. In \newcite{yeong2010language} 
@@
 \begin{table}
-\vspace{-0.65cm}
 \begin{center}
@@
 to both predict the code switched points but looking at the surrounding structure also decide the inclusion of them 
-into the current or the next segment.\footnote{The software used in this experiment is available online at: \url{removed for review}.}
+into the current or the next segment.
+%\footnote{The software used in this experiment is available online at: \url{removed for review}.}
 
@@
 
-This section describes the heuristics taken into account while performing the chunking of the input data according to 
-the sequences of segments in different languages. 
+This section describes the way we apply heuristics to segment and label the input text.
+%the heuristics taken into account while performing the chunking of the input data according to 
+%the sequences of segments in different languages. 
 In Figure ~\ref{fig:chunkal}, `chunks' represent the list of evaluated tuples of segments and their labelled language, `buffer' is the expanding segment.
@@
 \end{algorithm}
-\vspace{-0.6cm}
 \begin{figure}[hb]
@@
 
-\vspace{-0.8cm}
+\vspace{-0.7cm}
 \subsection{Word-based prediction}
+\vspace{-0.132cm}
 \label{wbased}
@@
 \subsection{Word-based prediction with character backoff}
+\vspace{-0.152cm}
 \label{wbasedbackoff}
 
-A better way to predict the spans of the sub-segments of the text is to include the two methods of word and character-based techniques 
-as described above. In case of the word being present in only one of the two monolingual word lists the classification is simple, 
+%A better way to predict the spans of the sub-segments of the text is to include the two methods of word and character-based techniques 
+%as described above.
+In case of the word being present in only one of the two monolingual word lists the classification is simple, 
 but in case of a conflict, a character bigram backoff was introduced to help us disambiguate the language label.
@@
 
-\vspace{-0.2cm}
-
 \section{Evaluation}
+\vspace{-0.152cm}
 \label{sec:eval}
@@
 \section{Results}
-\vspace{-0.152cm}
 \label{sec:results}
@@
 In terms of per-word accuracy, the baseline performs as expected, given that half of the words in the test set are from a given language. 
-The \texttt{langid.py} based methods allow for segmenting the text into chunks, but favour the majority language.\footnote{This could be because of the default models being used.} 
+The \texttt{langid.py} based methods allow for segmenting the text into chunks, but favour the majority language.
+%\footnote{This could be because of the default models being used.} 
 %While the precision and recall are low for the remaining models, we see that we are able to improve 
@@
 \section{Conclusions}
+\vspace{-0.152cm}
 \label{sec:conclusions}
@@
 proposed a model that chunks input text into segments and performs language identification on these segments at the same time. Precision and
-recall are low, leaving a lot of room for further work. Some of our plans are as follows. We would like to annotate more test data,
-at least 100 tweets in each language.
+recall are low, leaving a lot of room for further work.
+% Some of our plans are as follows. We would like to annotate more test data,
+%at least 100 tweets in each language.
  Although In \cite{king2013labeling}, the authors label on a per word level, yet we would like to include supervised methods and features talked about in this research to improve our efficiency while dealing with segments.
-We would also like to attempt our method using higher order character n-gram models for backoff, and n-gram word language models for detection.
+We would also like to attempt our method using higher order character n-gram models for backoff, and n-gram word language models for detection and on more annotated data.
 

### 2014-06-28 codeswitch-cltw-2014/codeswitch_camera_ready.tex
almost there
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@55075 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \end{table}
-In terms of per-word accuracy, the baseline performs as expected, given that half of the words in the test set are from a given language. 
-The \texttt{langid.py} based methods allow for segmenting the text into chunks, but favour the majority language.
+%In terms of per-word accuracy, the baseline performs as expected, given that half of the words in the test set are from a given language. 
+%The \texttt{langid.py} based methods allow for segmenting the text into chunks, but favour the majority language.
 %\footnote{This could be because of the default models being used.} 
-%While the precision and recall are low for the remaining models, we see that we are able to improve 
+While the precision and recall are low for the remaining models, we see that we are able to improve 
 The performance by combining the wordlist based model with a character bigram model. And what is more, we are able to begin to not only 

### 2014-07-02 codeswitch-cltw-2014/codeswitch_camera_ready.tex
maybe there
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@55076 72bbbca6-d526-0410-a7d9-f06f51895060
FINAL
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@55078 72bbbca6-d526-0410-a7d9-f06f51895060
comments from teresa
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@55176 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \section{Introduction}
+\vspace{-0.132cm}
 \label{intro}
@@
 
-The remainder of the paper is laid out as follows. In section~\ref{sec:method} we describe the problem
+The remainder of the paper is laid out as follows. In Section~\ref{sec:method} we describe the problem
 in more detail and look at relevant prior work before proposing a novel method of sub-sentential
-language detection. Section~\ref{sec:eval} describes the evaluation methodology. Then in section~\ref{sec:results}
-we present the results of our method and compare it against several other possible methods. Finally, section~\ref{sec:conclusions}
+language detection. Section~\ref{sec:eval} describes the evaluation methodology. Then in Section~\ref{sec:results}
+we present the results of our method and compare it against several other possible methods. Finally, Section~\ref{sec:conclusions}
 presents future work and conclusions.
@@
 \section{Related Work}
+\vspace{-0.132cm}
 \label{sec:relwork}
@@
 \section{Methodology}
+\vspace{-0.132cm}
 \label{sec:method}
@@
 \subsection{Corpus}
+\vspace{-0.152cm}
 
@@
 \end{small}
-\vspace{-0.3cm}
 \caption{Example of text from a microblogging site chunked manually.}
@@
 \subsection{Alphabet n-gram approach}
+\vspace{-0.152cm}
 \label{alphan}
@@
 \subsection{Sequence chunking}
+\vspace{-0.152cm}
 \label{cschunking}
@@
 In Figure ~\ref{fig:chunkal}, `chunks' represent the list of evaluated tuples of segments and their labelled language, `buffer' is the expanding segment.
- \textit{lang\_predict} corressponds to any function which is used to determine the language of the token. The flag 
+ \textit{lang\_predict} corresponds to any function which is used to determine the language of the token. The flag 
 variable helps in implementing the heuristic of minimum segment size while labelling chunks.
@@
 %\Statex{}%Procedure{}{}
-\State $\textit{buffer} = [\ ] \text{\ \ \  \#undecided expanding window of chunk}$
-\State $\textit{chunks} = [\ ] \text{\ \ \  \#decided labelled segment}$
+\State $\textit{buffer} = [\ ] \text{\ \ \  \#Undecided expanding window of chunk}$
+\State $\textit{chunks} = [\ ] \text{\ \ \  \#Decided labelled segment}$
 \State $\textit{buffer\_language} \gets \textit{lang\_predict}\text{(sentence[0])} \text{\ \ \ \#Language of first word}$
@@
         \EndIf
-\If {$ \textit{lang\_prediction(word)} \text{!=} \textit{buffer\_langugae} $}
+\If {$ \textit{lang\_prediction(word)} \text{!=} \textit{buffer\_language} $}
         \If {$ \textit{flag} \text{=\ =\ 0} $}
@@
 \If {$\textit{length(buffer)} \text{!=0\ } $}
-        \State $\textit{chunk} \text{+= \ }  \text{[\textit{[(buffer,buffer\_language)]}} $
+        \State $\textit{chunk} \text{+= \ }  \text{[\textit{(buffer,buffer\_language)}]} $
 
@@
 \end{algorithm}
+\vspace{-0.4cm}
 \begin{figure}[hb]
@@
 \end{figure}
+\vspace{-0.4cm}
 %A flag is set at the beginning of the input text, and this moves forward to the position in the text upto where confident 
@@
 
-\vspace{-0.7cm}
 \subsection{Word-based prediction}
-\vspace{-0.132cm}
+\vspace{-0.152cm}
 \label{wbased}
@@
 but in case of a conflict, a character bigram backoff was introduced to help us disambiguate the language label. 
-This method works well because the earlier heuristic approach of just labelling the word with the label of the span which is expanding, would mean less code-switch detection and more shift towards the majority class.
+%This method works well because the earlier heuristic approach of just labelling the word with the label of the span which is expanding, would mean less code-switch detection and more shift towards the majority class.
 
@@
 
+\vspace{-0.132cm}
 \section{Evaluation}
-\vspace{-0.152cm}
+\vspace{-0.132cm}
 \label{sec:eval}
@@
 
-Apart from the techniques discussed in section~\ref{sec:method}, some baselines are also used to give a comparative view of how well all the mechanisms perform.
+Apart from the techniques discussed in Section~\ref{sec:method}, some baselines are also used to give a comparative view of how well all the mechanisms perform.
 
 \subsection{Baseline}
+\vspace{-0.132cm}
 \label{baseline}
@@
 \subsection{\texttt{langid} character trigram prediction}
+\vspace{-0.132cm}
 \label{langidstuff}
@@
 Trigrams were chosen after experimenting with 1--5 grams.
-The heuristics in section~\ref{sec:method} were followed for the text processing and chunking part of the method.
+The heuristics in Section~\ref{sec:method} were followed for the text processing and chunking part of the method.
 
@@
 \end{center}
-\vspace{-0.3cm}
 \caption{Accuracy of the systems over the three language pairs. The accuracy measures how often a token
@@
 
+\vspace{-0.132cm}
 \section{Conclusions}
-\vspace{-0.152cm}
+\vspace{-0.132cm}
 \label{sec:conclusions}
@@
 %at least 100 tweets in each language.
- Although In \cite{king2013labeling}, the authors label on a per word level, yet we would like to include supervised methods and features talked about in this research to improve our efficiency while dealing with segments.
+ Although \newcite{king2013labeling} label on a per word level, yet we would like to include supervised methods and features talked about in this research to improve our efficiency while dealing with segments.
 We would also like to attempt our method using higher order character n-gram models for backoff, and n-gram word language models for detection and on more annotated data.
@@
 \section*{Acknowledgements}
+\vspace{-0.132cm}
 

### 2014-07-02 codeswitch-cltw-2014/codeswitch_camera_ready.tex
included email and removed small
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@55179 72bbbca6-d526-0410-a7d9-f06f51895060

@@
   Hyderabad (India) \\
-  {\small {\tt akshay.minocha@students.iiit.ac.in}} \\\And
+  {\tt akshay.minocha@students.iiit.ac.in} \\\And
   Francis M. Tyers \\
@@
   9017 Romsa (Norway) \\
-  {\small {\tt francis.tyers@uit.no}} \\}
+  {\tt francis.tyers@uit.no} \\}
 

### 2014-08-22 assim-eamt-2015/tat-rus/tables.tex
Added Tatar-Russian evaluation results
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@56968 72bbbca6-d526-0410-a7d9-f06f51895060

\documentclass[11pt, oneside]{article}   	% use "amsart" instead of "article" for AMSLaTeX format
\usepackage{geometry}                		% See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   		% ... or a4paper or a5paper or ... 
%\geometry{landscape}                		% Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    		% Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}				% Use pdf, png, jpg, or eps§ with pdflatex; use eps in DVI mode
								% TeX will automatically convert eps --> pdf in pdflatex		
\usepackage{amssymb}
%\date{}							% Activate to display a given date or no date
\usepackage{float}
\restylefloat{table}

\begin{document}
\section{General results}

\begin{table}[H]
\centering
\begin{tabular}{|r |*{4}{c}|}
\hline
 &\multicolumn{4}{c|}{Evaluation mode}\\
\hline
Gap percentage & No hint & Source hint & MT hint & MT and source hint\\
\hline
10\%&40.34 +/- 39.19&49.03 +/- 43.06&58.15 +/- 42.62&48.13 +/- 42.77\\
20\%&37.72 +/- 32.24&43.46 +/- 32.07&56.88 +/- 33.36&59.99 +/- 29.47\\
30\%&33.73 +/- 28.74&37.03 +/- 27.98&54.25 +/- 26.09&52.95 +/- 25.64\\
\hline
\end{tabular}
\caption {Gaps successfully filled (\%), exactly as in reference, mean and standard deviation} \label{tab:title} 
\end{table}

\begin{table}[H]
\centering
\begin{tabular}{|r |*{4}{c}|}
\hline
 &\multicolumn{4}{c|}{Evaluation mode}\\
\hline
Gap percentage & No hint & Source hint & MT hint & MT and source hint\\
\hline
10\%&42.68&51.27&63.03&48.50\\
20\%&38.71&46.24&58.18&62.00\\
30\%&36.67&36.70&51.04&55.27\\
\hline
\end{tabular}
\caption {Gaps successfully filled (\%), exactly as in reference, average} \label{tab:title} 
\end{table}

\begin{table}[H]
\centering
\begin{tabular}{|r |*{4}{c}|}
\hline
 &\multicolumn{4}{c|}{Evaluation mode}\\
\hline
Gap percentage & No hint & Source hint & MT hint & MT and source hint\\
\hline
10\%&45.96 +/- 40.86&53.52 +/- 43.57&63.76 +/- 41.15&57.02 +/- 41.57\\
20\%&38.56 +/- 32.11&45.60 +/- 31.23&59.77 +/- 33.10&64.93 +/- 30.95\\
30\%&35.45 +/- 29.76&40.00 +/- 28.16&55.83 +/- 26.13&58.77 +/- 28.07\\
\hline
\end{tabular}
\caption {Gaps successfully filled (\%), using a synonym list, mean and standard deviation} \label{tab:title} 
\end{table}

\begin{table}[H]
\centering
\begin{tabular}{|r |*{4}{c}|}
\hline
 &\multicolumn{4}{c|}{Evaluation mode}\\
\hline
Gap percentage & No hint & Source hint & MT hint & MT and source hint\\
\hline
10\%&46.34&54.43&66.06&53.89\\
20\%&39.43&48.39&60.73&65.33\\
30\%&38.39&39.66&52.59&59.90\\
\hline
\end{tabular}
\caption {Gaps successfully filled (\%), using a synonym list, average} \label{tab:title} 
\end{table}

\begin{table}[H]
\centering
\begin{tabular}{|r |*{4}{c}|}
\hline
 &\multicolumn{4}{c|}{Evaluation mode}\\
\hline
Gap percentage & No hint & Source hint & MT hint & MT and source hint\\
\hline
10\%&1:44 +/- 7:18&0:50 +/- 1:05&1:14 +/- 1:52&1:26 +/- 1:10\\
20\%&0:57 +/- 1:22&1:19 +/- 2:34&1:39 +/- 3:02&3:26 +/- 8:44\\
30\%&1:20 +/- 1:50&1:14 +/- 1:12&2:05 +/- 6:36&2:36 +/- 6:26\\
\hline
\end{tabular}
\caption {Time taken to fill the gaps in one sentence} \label{tab:title} 
\end{table}

\begin{table}[H]
\centering
\begin{tabular}{|r |*{1}{c}|}
\hline
Gap percentage & Number of holes\\
\hline
10\%&272\\
20\%&396\\
30\%&724\\
\hline
\end{tabular}
\caption {Total number of holes to be filled} \label{tab:title}
\end{table}

\begin{table}[H]
\centering
\begin{tabular}{|r |*{4}{c}|}
\hline
 &\multicolumn{4}{c|}{Evaluation mode}\\
\hline
Gap percentage & No hint & Source hint & MT hint & MT ans Source hint\\
\hline
10\%&0.517 &0.711 &0.459 &0.598\\
20\%&0.261 &0.473 &0.667 &0.740\\
30\%&0.412 &0.411 &0.581 &0.534\\
\hline
\end{tabular}
\caption {Krippendorff Alpha measure} \label{tab:title}
\end{table}

\section{Results by topic}
\subsection{Casual conversations}

\begin{table}[H]
\centering
\begin{tabular}{|r |*{4}{c}|}
\hline
 &\multicolumn{4}{c|}{Evaluation mode}\\
\hline
Gap percentage & No hint & Source hint & MT hint & MT and source hint\\
\hline
10\%&46.97 +/- 44.28&53.03 +/- 47.58&48.48 +/- 45.20&51.52 +/- 46.85\\
20\%&35.56 +/- 31.18&37.22 +/- 29.17&59.44 +/- 35.92&63.89 +/- 29.21\\
30\%&33.54 +/- 31.47&34.58 +/- 27.22&56.11 +/- 23.82&59.17 +/- 28.05\\
\hline
\end{tabular}
\caption {Gaps successfully filled (\%), exactly as in reference, mean and standard deviation} \label{tab:title} 
\end{table}

\begin{table}[H]
\centering
\begin{tabular}{|r |*{4}{c}|}
\hline
 &\multicolumn{4}{c|}{Evaluation mode}\\
\hline
Gap percentage & No hint & Source hint & MT hint & MT and source hint\\
\hline
10\%&50.00&57.50&48.78&50.00\\
20\%&35.44&39.24&59.21&65.79\\
30\%&37.63&37.50&53.85&59.09\\
\hline
\end{tabular}
\caption {Gaps successfully filled (\%), exactly as in reference, average} \label{tab:title} 
\end{table}

\begin{table}[H]
\centering
\begin{tabular}{|r |*{4}{c}|}
\hline
 &\multicolumn{4}{c|}{Evaluation mode}\\
\hline
Gap percentage & No hint & Source hint & MT hint & MT and source hint\\
\hline
10\%&53.03 +/- 44.28&62.12 +/- 46.11&63.64 +/- 43.12&63.64 +/- 44.84\\
20\%&37.78 +/- 30.94&41.11 +/- 27.88&63.33 +/- 35.85&73.33 +/- 31.51\\
30\%&37.71 +/- 33.32&39.44 +/- 27.10&59.58 +/- 23.89&70.28 +/- 31.12\\
\hline
\end{tabular}
\caption {Gaps successfully filled (\%), using a synonym list, mean and standard deviation} \label{tab:title} 
\end{table}

\begin{table}[H]
\centering
\begin{tabular}{|r |*{4}{c}|}
\hline
 &\multicolumn{4}{c|}{Evaluation mode}\\
\hline
Gap percentage & No hint & Source hint & MT hint & MT and source hint\\
\hline
10\%&54.76&65.00&60.98&59.52\\
20\%&37.97&44.30&63.16&73.68\\
30\%&41.94&41.67&58.24&68.18\\
\hline
\end{tabular}
\caption {Gaps successfully filled (\%), using a synonym list, average} \label{tab:title} 
\end{table}

\begin{table}[H]
\centering
\begin{tabular}{|r |*{4}{c}|}
\hline
 &\multicolumn{4}{c|}{Evaluation mode}\\
\hline
Gap percentage & No hint & Source hint & MT hint & MT and source hint\\
\hline
10\%&3:35 +/- 11:43&0:45 +/- 1:00&1:03 +/- 1:14&1:28 +/- 1:16\\
20\%&0:33 +/- 0:31&0:47 +/- 0:41&0:38 +/- 0:36&0:52 +/- 1:28\\
30\%&1:09 +/- 2:21&0:49 +/- 0:42&0:40 +/- 0:29&0:53 +/- 0:54\\
\hline
\end{tabular}
\caption {Time taken to fill the gaps in one sentence} \label{tab:title} 
\end{table}

\begin{table}[H]
\centering
\begin{tabular}{|r |*{1}{c}|}
\hline
Gap percentage & Number of holes\\
\hline
10\%&60\\
20\%&124\\
30\%&184\\
\hline
\end{tabular}
\caption {Total number of holes to be filled} \label{tab:title}
\end{table}

\begin{table}[H]
\centering
\begin{tabular}{|r |*{4}{c}|}
\hline
 &\multicolumn{4}{c|}{Evaluation mode}\\
\hline
Gap percentage & No hint & Source hint & MT hint & MT ans Source hint\\
\hline
10\%&0.411 &0.601 &0.426 &0.504\\
20\%&0.235 &0.371 &0.570 &0.712\\
30\%&0.121 &0.459 &0.846 &0.310\\
\hline
\end{tabular}
\caption {Krippendorff Alpha measure} \label{tab:title}
\end{table}

\subsection{Legal texts}

\begin{table}[H]
\centering
\begin{tabular}{|r |*{4}{c}|}
\hline
 &\multicolumn{4}{c|}{Evaluation mode}\\
\hline
Gap percentage & No hint & Source hint & MT hint & MT and source hint\\
\hline
10\%&29.10 +/- 29.64&38.77 +/- 37.82&68.21 +/- 34.99&53.02 +/- 40.31\\
20\%&48.31 +/- 33.52&50.28 +/- 33.55&65.76 +/- 24.16&60.57 +/- 24.98\\
30\%&33.68 +/- 28.16&38.49 +/- 27.81&48.30 +/- 28.87&47.39 +/- 26.29\\
\hline
\end{tabular}
\caption {Gaps successfully filled (\%), exactly as in reference, mean and standard deviation} \label{tab:title} 
\end{table}

\begin{table}[H]
\centering
\begin{tabular}{|r |*{4}{c}|}
\hline
 &\multicolumn{4}{c|}{Evaluation mode}\\
\hline
Gap percentage & No hint & Source hint & MT hint & MT and source hint\\
\hline
10\%&35.21&43.48&67.11&52.00\\
20\%&43.31&52.46&63.57&61.74\\
30\%&36.17&35.16&41.32&53.76\\
\hline
\end{tabular}
\caption {Gaps successfully filled (\%), exactly as in reference, average} \label{tab:title} 
\end{table}

\begin{table}[H]
\centering
\begin{tabular}{|r |*{4}{c}|}
\hline
 &\multicolumn{4}{c|}{Evaluation mode}\\
\hline
Gap percentage & No hint & Source hint & MT hint & MT and source hint\\
\hline
10\%&32.95 +/- 34.88&38.77 +/- 37.82&68.21 +/- 34.99&53.02 +/- 40.31\\
20\%&48.31 +/- 33.52&50.28 +/- 33.55&65.76 +/- 24.16&60.57 +/- 24.98\\
30\%&34.15 +/- 28.73&40.05 +/- 27.37&48.30 +/- 28.87&50.06 +/- 26.25\\
\hline
\end{tabular}
\caption {Gaps successfully filled (\%), using a synonym list, mean and standard deviation} \label{tab:title} 
\end{table}

\begin{table}[H]
\centering
\begin{tabular}{|r |*{4}{c}|}
\hline
 &\multicolumn{4}{c|}{Evaluation mode}\\
\hline
Gap percentage & No hint & Source hint & MT hint & MT and source hint\\
\hline
10\%&38.03&43.48&67.11&52.00\\
20\%&43.31&52.46&63.57&61.74\\
30\%&37.23&37.36&41.32&56.07\\
\hline
\end{tabular}
\caption {Gaps successfully filled (\%), using a synonym list, average} \label{tab:title} 
\end{table}

\begin{table}[H]
\centering
\begin{tabular}{|r |*{4}{c}|}
\hline
 &\multicolumn{4}{c|}{Evaluation mode}\\
\hline
Gap percentage & No hint & Source hint & MT hint & MT and source hint\\
\hline
10\%&0:49 +/- 1:13&1:04 +/- 1:21&1:17 +/- 1:01&1:34 +/- 1:03\\
20\%&1:32 +/- 2:02&2:28 +/- 4:18&3:26 +/- 4:41&8:22 +/- 13:46\\
30\%&1:13 +/- 1:36&1:19 +/- 1:21&3:32 +/- 10:24&2:57 +/- 4:00\\
\hline
\end{tabular}
\caption {Time taken to fill the gaps in one sentence} \label{tab:title} 
\end{table}

\begin{table}[H]
\centering
\begin{tabular}{|r |*{1}{c}|}
\hline
Gap percentage & Number of holes\\
\hline
10\%&132\\
20\%&244\\
30\%&284\\
\hline
\end{tabular}
\caption {Total number of holes to be filled} \label{tab:title}
\end{table}

\begin{table}[H]
\centering
\begin{tabular}{|r |*{4}{c}|}
\hline
 &\multicolumn{4}{c|}{Evaluation mode}\\
\hline
Gap percentage & No hint & Source hint & MT hint & MT ans Source hint\\
\hline
10\%&0.543 &0.481 &0.402 &0.642\\
20\%&0.317 &0.453 &0.631 &0.652\\
30\%&0.488 &0.301 &0.426 &0.532\\
\hline
\end{tabular}
\caption {Krippendorff Alpha measure} \label{tab:title}
\end{table}

\subsection{News}

\begin{table}[H]
\centering
\begin{tabular}{|r |*{4}{c}|}
\hline
 &\multicolumn{4}{c|}{Evaluation mode}\\
\hline
Gap percentage & No hint & Source hint & MT hint & MT and source hint\\
\hline
10\%&42.78 +/- 38.41&54.02 +/- 40.51&59.77 +/- 43.66&40.00 +/- 38.87\\
20\%&28.97 +/- 28.79&44.44 +/- 32.51&43.70 +/- 34.86&54.46 +/- 33.34\\
30\%&33.98 +/- 26.52&37.65 +/- 28.77&59.83 +/- 22.90&53.70 +/- 20.22\\
\hline
\end{tabular}
\caption {Gaps successfully filled (\%), exactly as in reference, mean and standard deviation} \label{tab:title} 
\end{table}

\begin{table}[H]
\centering
\begin{tabular}{|r |*{4}{c}|}
\hline
 &\multicolumn{4}{c|}{Evaluation mode}\\
\hline
Gap percentage & No hint & Source hint & MT hint & MT and source hint\\
\hline
10\%&47.06&57.14&68.75&42.00\\
20\%&34.25&43.59&47.14&58.67\\
30\%&36.72&38.28&61.72&54.69\\
\hline
\end{tabular}
\caption {Gaps successfully filled (\%), exactly as in reference, average} \label{tab:title} 
\end{table}

\begin{table}[H]
\centering
\begin{tabular}{|r |*{4}{c}|}
\hline
 &\multicolumn{4}{c|}{Evaluation mode}\\
\hline
Gap percentage & No hint & Source hint & MT hint & MT and source hint\\
\hline
10\%&49.44 +/- 39.08&57.47 +/- 42.12&59.77 +/- 43.66&53.33 +/- 37.86\\
20\%&28.97 +/- 28.79&46.53 +/- 31.99&48.56 +/- 35.40&59.33 +/- 33.89\\
30\%&34.81 +/- 27.01&40.51 +/- 30.10&61.49 +/- 22.11&58.14 +/- 22.42\\
\hline
\end{tabular}
\caption {Gaps successfully filled (\%), using a synonym list, mean and standard deviation} \label{tab:title} 
\end{table}

\begin{table}[H]
\centering
\begin{tabular}{|r |*{4}{c}|}
\hline
 &\multicolumn{4}{c|}{Evaluation mode}\\
\hline
Gap percentage & No hint & Source hint & MT hint & MT and source hint\\
\hline
10\%&50.98&61.22&68.75&52.00\\
20\%&34.25&46.15&52.86&64.00\\
30\%&37.50&41.41&63.28&59.38\\
\hline
\end{tabular}
\caption {Gaps successfully filled (\%), using a synonym list, average} \label{tab:title} 
\end{table}

\begin{table}[H]
\centering
\begin{tabular}{|r |*{4}{c}|}
\hline
 &\multicolumn{4}{c|}{Evaluation mode}\\
\hline
Gap percentage & No hint & Source hint & MT hint & MT and source hint\\
\hline
10\%&0:31 +/- 0:20&0:42 +/- 0:49&1:25 +/- 2:49&1:17 +/- 1:09\\
20\%&0:50 +/- 0:53&0:49 +/- 0:42&0:56 +/- 0:39&1:05 +/- 0:33\\
30\%&1:38 +/- 1:28&1:31 +/- 1:17&1:40 +/- 1:12&3:53 +/- 10:27\\
\hline
\end{tabular}
\caption {Time taken to fill the gaps in one sentence} \label{tab:title} 
\end{table}

\begin{table}[H]
\centering
\begin{tabular}{|r |*{1}{c}|}
\hline
Gap percentage & Number of holes\\
\hline
10\%&80\\
20\%&148\\
30\%&256\\
\hline
\end{tabular}
\caption {Total number of holes to be filled} \label{tab:title}
\end{table}

\begin{table}[H]
\centering
\begin{tabular}{|r |*{4}{c}|}
\hline
 &\multicolumn{4}{c|}{Evaluation mode}\\
\hline
Gap percentage & No hint & Source hint & MT hint & MT ans Source hint\\
\hline
10\%&0.537 &0.917 &0.574 &0.598\\
20\%&0.143 &0.644 &0.850 &0.928\\
30\%&0.513 &0.456 &0.532 &0.644\\
\hline
\end{tabular}
\caption {Krippendorff Alpha measure} \label{tab:title}
\end{table}

\end{document}

### 2014-11-02 rusaccent-nodalida-2015/rusaccent.tex
dir for one nodalida paper
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@57712 72bbbca6-d526-0410-a7d9-f06f51895060

%
% File nodalida2015.tex
%
% Contact beata.megyesi@lingfil.uu.se
%
% Based on the instruction file for EACL 2014
% which in turn was based on the instruction files for previous 
% ACL and EACL conferences.

\documentclass[11pt]{article}
\usepackage{nodalida2015}
%\usepackage{times}
\usepackage{mathptmx}
%\usepackage{txfonts}
\usepackage{url}
\usepackage{latexsym}
\special{papersize=210mm,297mm} % to avoid having to use "-t a4" with dvips 
%\setlength\titlebox{6.5cm}  % You can expand the title box if you really have to

\title{Automatic accent annotation for Russian}

\author{Robert Reynolds \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  {\tt email@uit.no} \\\And
  Francis M. Tyers \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  {\tt email@uit.no} \\}

\date{}

\begin{document}
\maketitle
\begin{abstract}
  This document contains the instructions for preparing a camera-ready
  manuscript for the proceedings of Nodalida-2015. The document itself
  conforms to its own specifications, and is therefore an example of
  what your manuscript should look like. These instructions should be
  used for both papers submitted for review and for final versions of
  accepted papers.  Authors are asked to conform to all the directions
  reported in this document.
\end{abstract}

\section{Credits}

This document has been adapted from the instructions for earlier
(E)ACL-proceedings, including those for EACL-2014 by Gosse Bouma
and Yannick Parmentier, those for ACL-2013 by Roberto Navigli
and Jing-Shing Chang, those for ACL-2012 by Maggie Li and Michael
White, those for ACL-2010 by Jing-Shing Chang and Philipp Koehn,
those for ACL-2008 by Joakim Nivre and Noah Smith, 
those for EACL-2006 by Eneko Agirre and Sergi Balari, 
those for ACL-2005 by Hwee Tou Ng and Kemal Oflazer, those for
ACL-2002 by Eugene Charniak and Dekang Lin, and earlier ACL and EACL
formats. Those versions were written by several people, including John
Chen, Henry S. Thompson and Donald Walker. Additional elements were
taken from the formatting instructions of the {\em International Joint
  Conference on Artificial Intelligence}.

\section{Introduction}
% das: removed reference to PostScript

The following instructions are directed to authors of papers submitted
to Nodalida-2015 or accepted for publication in its proceedings. 
All submissions must conform to the NoDaLiDa 2015 style guides, as summarized
and exemplified in this document. All authors are required to adhere to these specifications. 
Templates are available from the conference web site for \LaTeX\
(recommended), or MS~Word. If for some reason you have a problem producing this format, please contact the
Programme Chair (Beata Megyesi; \texttt{beata.megyesi@lingfil.uu.se}) as early as
possible. The NoDaLiDa Programme Committee reserves the right not to publish papers not
conforming to the standard format.

Submission and reviewing will be on-line, managed by the EasyChair system.
The only accepted format for submitted papers is PDF.
NoDaLiDa does not allow submission of supplementary materials.
All submissions must be uploaded to EasyChair by the submission deadlines;
submissions received after that time will not be accepted.
To minimize the risk of network congestion, we encourage authors to upload
their submissions as early as possible. 
Improved versions of the submissions may be continuously uploaded until the
final deadline.

On acceptance of a submission, precise instructions will be given on how to
send the camera ready version; following reviewers' feedback, such instructions
may include specific requests for modification.
Camera-ready versions not conforming to requests for revision made by the
Programme Committee will be considered as unsent, and hence will not be
included in the proceedings or presented at the conference.

\section{General Instructions}

Manuscripts must be in two-column format.  Exceptions to the
two-column format include the title, authors' names and complete
addresses, which must be centered at the top of the first page, and
any full-width figures or tables (see the guidelines in
Subsection~\ref{ssec:fonts}). {\bf Type single-spaced.}  Start all
pages directly under the top margin. See the guidelines later
regarding formatting the first page (Subsection~\ref{ssec:first}).
The manuscript should be printed single-sided and its length should
not exceed the maximum page limit described in
Section~\ref{sec:length}.  {\bf Do not number the pages.}

\subsection{Electronically-available resources}

Nodalida-2015 provides this description in \LaTeX2e ({\small {\tt
    nodalida2015.tex}}) and PDF format ({\small {\tt nodalida2015.pdf}}),
along with the \LaTeX2e style file used to format it ({\small {\tt
    nodalida2015.sty}}) and an ACL bibliography style ({\small {\tt
    acl.bst}}) in case you want to use the Bib\TeX\ reference management
software. These files are all available at
\url{http://stp.lingfil.uu.se/~bea/nodalida15/}. A Microsoft Word template file ({\small
  {\tt nodalida2015.docx}}) is also available at the same URL. We strongly
recommend the use of these style files, which have been appropriately
tailored for the Nodalida-2015 proceedings. If you have an option, we
recommend that you use the \LaTeX2e version. \textbf{If you will be
  using the Microsoft Word template, we suggest that you anonymize
  your source file so that the pdf produced does not retain your
  identity.}  This can be done by removing any personal information
from your source document properties.

\subsection{Format of Electronic Manuscript}
\label{sect:pdf}

For the production of the electronic manuscript you must use Adobe's
Portable Document Format (PDF). 
%% Add instructions about using pdflatex
This format can be generated directly from \LaTeX2e files or from
postscript ones. On Linux/Unix-like systems, you can use {\tt
  pdflatex} to generate a PDF file from \LaTeX2e files, or {\tt
  ps2pdf} to convert from postscript to PDF. 
%%
In Microsoft Windows, you can use Adobe's Distiller, or if you have
\texttt{cygwin} installed, you can use \texttt{dvipdf} or
\texttt{ps2pdf}. Note that some word processing programs generate PDF
which may not include all the necessary fonts (esp. tree diagrams,
symbols). When you print or create the PDF file, there is usually an
option in your printer setup to include none, all or just non-standard
fonts.  Please make sure that you select the option of including ALL
the fonts. {\em Before sending it, test your PDF by printing it from a
  computer different from the one where it was created.} Moreover,
some word processors may generate very large postscript/PDF files,
where each page is rendered as an image. Such images may reproduce
poorly. In this case, try alternative ways to obtain the postscript
and/or PDF. One way on some systems is to install a driver for a
postscript printer, send your document to the printer specifying
``Output to a file'', then convert the file to PDF.
Please keep in mind that it is of utmost importance to use \textbf{A4 format}.

%%Removed by YP (use of the \special command above instead):
%% It is of utmost importance to specify the \textbf{A4 format} (21 cm x
%% 29.7 cm) when formatting the paper. 
%% When working with {\tt dvips}, for instance, one should specify {\tt -t a4}.

Print-outs of the PDF file on A4 paper should be identical to the
hardcopy version. If you cannot meet the above requirements about the
production of your electronic submission, please contact the
program chair above as soon as possible.

\subsection{Layout}
\label{ssec:layout}

Format manuscripts two columns to a page, in the manner these
instructions are formatted. The exact dimensions for a page on A4
paper are:

\begin{itemize}
\item Left and right margins: 2.5 cm
\item Top margin: 2.5 cm
\item Bottom margin: 2.5 cm
\item Column width: 7.7 cm
\item Column height: 24.7 cm
\item Gap between columns: 0.6 cm
\end{itemize}

\noindent Papers should not be submitted on any other paper size.
 If you cannot meet the above requirements about the production of your electronic submission, 
please contact the program chair above as soon as possible.

\subsection{Fonts}
\label{ssec:fonts}
For reasons of uniformity, Adobe's {\bf Times Roman} font should be
used. In \LaTeX2e{} this is accomplished by putting

\begin{quote}
\begin{verbatim}
\usepackage{times}
\usepackage{latexsym}
\end{verbatim}
\end{quote}
in the preamble. If Times Roman is unavailable, use {\bf Computer
  Modern Roman} (\LaTeX2e{}'s default).  Note that the latter is about
  10\% less dense than Adobe's Times Roman font.

\begin{table}[h]
\begin{center}
\begin{tabular}{|l|rl|}
\hline \bf Type of Text & \bf Font Size & \bf Style \\ \hline
paper title & 15 pt & bold \\
author names & 12 pt & bold \\
author affiliation & 12 pt & \\
the word ``Abstract'' & 12 pt & bold \\
section titles & 12 pt & bold \\
document text & 11 pt  &\\
captions & 11 pt & \\
abstract text & 10 pt & \\
bibliography & 10 pt & \\
footnotes & 9 pt & \\
\hline
\end{tabular}
\end{center}
\caption{\label{font-table} Font guide. }
\end{table}

\subsection{The First Page}
\label{ssec:first}

Center the title, author's name(s) and affiliation(s) across both
columns. Do not use footnotes for affiliations. Do not include the
paper ID number assigned during the submission process. Use the
two-column format only when you begin the abstract.

{\bf Title}: Place the title centered at the top of the first page, in
a 15-point bold font. (For a complete guide to font sizes and styles, 
see Table~\ref{font-table}.) Long titles should be typed on two lines without
a blank line intervening. Approximately, put the title at 2.5 cm from
the top of the page, followed by a blank line, then the author's
names(s), and the affiliation on the following line. Do not use only
initials for given names (middle initials are allowed). Do not format surnames
in all capitals (e.g., use ``Schlangen'' not ``SCHLANGEN'').
Do not format title and section headings in all capitals as well
except for proper names (such as ``BLEU'') that are conventionally
in all capitals.
The affiliation should contain the author's complete address, and if
possible, an electronic mail address. Leave about 2 cm between the
affiliation and the body of the first page.
The title, author names and addresses should be completely
identical to those entered to the electronical paper submission
website in order to maintain the consistency of author information
among all publications of the conference.

{\bf Abstract}: Type the abstract at the beginning of the first
column. The width of the abstract text should be smaller than the
width of the columns for the text in the body of the paper by about
0.6 cm on each side. Center the word {\bf Abstract} in a 12 point bold
font above the body of the abstract. The abstract should be a concise
summary of the general thesis and conclusions of the paper. It should
be no longer than 200 words. The abstract text should be in 10 point font.

{\bf Text}: Begin typing the main body of the text immediately after
the abstract, observing the two-column format as shown in 
the present document. Do not include page numbers.

{\bf Indent} when starting a new paragraph. Use 11 points for text and 
subsection headings, 12 points for section headings and 15 points for
the title.

\subsection{Sections}

{\bf Headings}: Type and label section and subsection headings in the
style shown on the present document.  Use numbered sections (Arabic
numerals) in order to facilitate cross references. Number subsections
with the section number and the subsection number separated by a dot,
in Arabic numerals. Do not number subsubsections.

{\bf Citations}: Citations within the text appear
in parentheses as~\cite{Gusfield:97} or, if the author's name appears in
the text itself, as Gusfield~\shortcite{Gusfield:97}. 
Append lowercase letters to the year in cases of ambiguity.  
Treat double authors as in~\cite{Aho:72}, but write as
 in~\cite{Chandra:81} when more than two authors are involved. Collapse multiple citations as
in~\cite{Gusfield:97,Aho:72}. Also refrain from using full citations as sentence constituents. We
suggest that instead of
\begin{quote}
  ``\cite{Gusfield:97} showed that ...''
\end{quote}
you use
\begin{quote}
``Gusfield \shortcite{Gusfield:97}   showed that ...''
\end{quote}

If you are using the provided \LaTeX{} and Bib\TeX{} style files, you
can use the command \verb|\newcite| to get ``author (year)'' citations.

As reviewing will be double-blind, the submitted version of the papers should not include the
authors' names and affiliations. Furthermore, self-references that
reveal the author's identity, e.g.,
\begin{quote}
``We previously showed \cite{Gusfield:97} ...''  
\end{quote}
should be avoided. Instead, use citations such as 
\begin{quote}
``Gusfield \shortcite{Gusfield:97}
previously showed ... ''
\end{quote}

\textbf{Please do not  use anonymous citations} and  do not include acknowledgements 
when submitting your paper. Papers that do not conform
to these requirements may be rejected without review.

\textbf{References}: Gather the full set of references together under
the heading {\bf References}; place the section before any Appendices,
unless they contain references. Arrange the references alphabetically
by first author, rather than by order of occurrence in the text.
Provide as complete a citation as possible, using a consistent format,
such as the one for {\em Computational Linguistics\/} or the one in the 
{\em Publication Manual of the American 
Psychological Association\/}~\cite{APA:83}.  Use of full names for
authors rather than initials is preferred.  A list of abbreviations
for common computer science journals can be found in the ACM 
{\em Computing Reviews\/}~\cite{ACM:83}.

The \LaTeX{} and Bib\TeX{} style files provided roughly fit the
American Psychological Association format, allowing regular citations, 
short citations and multiple citations as described above.

{\bf Appendices}: Appendices, if any, directly follow the text and the
references (but see above).  Letter them in sequence and provide an
informative title: {\bf Appendix A. Title of Appendix}.

\textbf{Acknowledgement} section should appear in accepted manuscripts
only. It should go as a last section immediately
before the references.  Do not number the acknowledgement section.

\subsection{Footnotes}

{\bf Footnotes}: Put footnotes at the bottom of the page and use 9
points text. They may be numbered or referred to by asterisks or other
symbols.\footnote{This is how a footnote should appear.} Footnotes
should be separated from the text by a line.\footnote{Note the line
separating the footnotes from the text.}

\subsection{Graphics}

{\bf Illustrations}: Place figures, tables, and photographs in the
paper near where they are first discussed, rather than at the end, if
possible.  Wide illustrations may run across both columns.  Color
illustrations are allowed, provided you have verified that  
they will be understandable when printed in black ink.

{\bf Captions}: Provide a caption for every illustration; number each one
sequentially in the form:  ``Figure 1. Caption of the Figure.'' ``Table 1.
Caption of the Table.''  Type the captions of the figures and 
tables below the body, using 11 point text.

\section{Translation of non-English Terms}

It is also advised to supplement non-English characters and terms
with appropriate transliterations and/or translations
since not all readers understand all such characters and terms.
Inline transliteration or translation can be represented in
the order of: original-form transliteration ``translation''.

\section{Length of Submission}
\label{sec:length}

Long papers may consist of up to 8 pages of content (excluding
references) and short papers may consist of up to 
four (4) pages plus 2 pages for references in the proceedings.  Papers that do not conform to the
specified length and formatting requirements are subject to be
rejected without review.

%\section{Other Issues}

%Papers that had software and/or dataset submitted for the review
%process should also submit it with the camera-ready paper. Besides,
%the software and/or dataset should not be anonymous.

% Please note that the publications of Nodalida-2015 will be publicly
% available at ACL Anthology (http://aclweb.org/anthology-new/) in May,
% 2015. Since some of the authors may have plans to file patents related to their papers in
% the conference, we are reminding authors that May 12th, 2015 may be
% considered to be the official publication date, i.e., the opening
% day of the conference.

\section*{Acknowledgments}

Do not number the acknowledgment section. Do not include this section
when submitting your paper for review.

% If you use BibTeX with a bib file named eacl2014.bib, 
% you should add the following two lines:
\bibliographystyle{acl}
\bibliography{rusaccent}

\end{document}

### 2014-12-09 rusaccent-nodalida-2015/rusaccent.tex
the next draft will suck less
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@57952 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
+%\usepackage[T2A,T1]{fontenc}
+%\usepackage[russian,english]{babel}
+%\newcommand{\rus}[1]{\foreignlanguage{russian}{#1}}
+
 \title{Automatic accent annotation for Russian}
 
-\author{Robert Reynolds \\
+\author{Author1 \\
   Affiliation / Address line 1 \\
@@
   Affiliation / Address line 3 \\
-  {\tt email@uit.no} \\\And
-  Francis M. Tyers \\
+  {\tt email@domain.com} \\\And
+  Author2 \\
   Affiliation / Address line 1 \\
@@
   Affiliation / Address line 3 \\
-  {\tt email@uit.no} \\}
+  {\tt email@domain.com} \\}
 
@@
 
-This document has been adapted from the instructions for earlier
-(E)ACL-proceedings, including those for EACL-2014 by Gosse Bouma
-and Yannick Parmentier, those for ACL-2013 by Roberto Navigli
-and Jing-Shing Chang, those for ACL-2012 by Maggie Li and Michael
-White, those for ACL-2010 by Jing-Shing Chang and Philipp Koehn,
-those for ACL-2008 by Joakim Nivre and Noah Smith, 
-those for EACL-2006 by Eneko Agirre and Sergi Balari, 
-those for ACL-2005 by Hwee Tou Ng and Kemal Oflazer, those for
-ACL-2002 by Eugene Charniak and Dekang Lin, and earlier ACL and EACL
-formats. Those versions were written by several people, including John
-Chen, Henry S. Thompson and Donald Walker. Additional elements were
-taken from the formatting instructions of the {\em International Joint
-  Conference on Artificial Intelligence}.
+The template had a credits section for the history of the template, but we don't need
+a Credits section, right?
 
 \section{Introduction}
-% das: removed reference to PostScript
+
+Russian word stress is variable, which means that the position of the stress in a word is not 
+generally predictable. Furthermore, the inflecting word classes include complex patterns of
+shifting stress. Lexical stress and its attendant vowel reduction are a prominent feature of
+spoken Russian, but standard written Russian does not typically mark word stress. Students
+of Russian frequently struggle to know how to pronounce written words, so many
+textbooks mark stress. Even texts intended for native speakers
+sometimes mark words that cannot be disambiguated through context.
+
+Word stress ambiguity in Russian arises both between lexemes and within inflectional
+paradigms. Stress ambiguity between lexemes includes pairs of the type
+????? \emph{zámok} `castle' and ????? \emph{zamók} 'lock'. For every inflected form of 
+these two words, they are both morphosyntactically and orthographically identical, their
+meaning differentiated only by stress placement. This kind of ambiguity is relatively uncommon
+in Russian. This paper focuses on stress ambiguities which are connected to morphosyntactic
+differences. ADD EXAMPLE HERE with ????????, ????????, ????????, AND ??????/??????.
+ADD ANOTHER EXAMPLE WITH the paradigms of ???? and ??????.
+
+\subsection{Related Work}
+
+Automatic stress placement in Russian is similar to Diacritic Restoration, a task
+which has received increasing interest over the last 20 years. Generally speaking, diacritics
+disambiguate otherwise homographic wordforms, so missing diacritics can
+complicate many NLP tasks, such as text-to-speech and speech-to-text. For example, speakers 
+of Czech and French sometimes type emails and other communications
+without standard diacritics. In order to generate speech from these texts, they must first be 
+normalized by restoring diacritics. A slightly different
+situation arises with Arabic and Hebrew, since standard orthography lacks vowel diacritics
+altogether. For such languages, the `restoration' of vowel diacritics results in less
+ambiguity than in standard orthography. For languages with inherently ambiguous orthography,
+it may be more precise to refer to this task as `diacritic enhancement', since it produces
+text that is less ambiguous than the standard language. In this sense, Russian orthography
+is similar to Arabic and Hebrew.
 

### 2014-12-17 rusaccent-nodalida-2015/rusaccent.tex
More on intro
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58078 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \documentclass[11pt]{article}
-\usepackage{nodalida2015}
-%\usepackage{times}
-\usepackage{mathptmx}
+\usepackage[T2A,T1]{fontenc}
+\usepackage[utf8]{inputenc}
+\usepackage[russian,english]{babel}
+
+\usepackage{times}
+\usepackage{latexsym}
+\usepackage{fixltx2e} %allows subscripts
+%\usepackage{mathptmx}
 %\usepackage{txfonts}
 \usepackage{url}
-\usepackage{latexsym}
 \special{papersize=210mm,297mm} % to avoid having to use "-t a4" with dvips 
@@
 
-%\usepackage[T2A,T1]{fontenc}
-%\usepackage[russian,english]{babel}
-%\newcommand{\rus}[1]{\foreignlanguage{russian}{#1}}
+\usepackage{nodalida2015}
+
+\usepackage{linguex}
+
+\newcommand{\rus}[1]{\foreignlanguage{russian}{#1}}
+\newcommand{\ft}[1]{\marginpar{\scriptsize F: #1}}
+\newcommand{\rr}[1]{\marginpar{\scriptsize R: #1}}
 
@@
 
-\section{Credits}
-
-The template had a credits section for the history of the template, but we don't need
-a Credits section, right?
+%\section{Credits}
+%
+%The template had a credits section for the history of the template, but we don't need
+%a Credits section, right?
 
@@
 
-Russian word stress is variable, which means that the position of the stress in a word is not 
+Russian word stress is variable, which means that the stress position in a given word is not 
 generally predictable. Furthermore, the inflecting word classes include complex patterns of
 shifting stress. Lexical stress and its attendant vowel reduction are a prominent feature of
-spoken Russian, but standard written Russian does not typically mark word stress. Students
-of Russian frequently struggle to know how to pronounce written words, so many
-textbooks mark stress. Even texts intended for native speakers
-sometimes mark words that cannot be disambiguated through context.
+spoken Russian, but standard written Russian does not typically mark word stress.\footnote{
+Texts intended for native speakers sometimes mark stress on words that cannot be 
+disambiguated through context.} Without information about lexical stress position, 
+correctly converting written Russian text to speech is impossible. This is true both for humans
+(e.g. foreign language students) and computers (e.g. text-to-speech or automatic speech 
+recognition).
 
-Word stress ambiguity in Russian arises both between lexemes and within inflectional
-paradigms. Stress ambiguity between lexemes includes pairs of the type
-????? \emph{zámok} `castle' and ????? \emph{zamók} 'lock'. For every inflected form of 
-these two words, they are both morphosyntactically and orthographically identical, their
-meaning differentiated only by stress placement. This kind of ambiguity is relatively uncommon
-in Russian. This paper focuses on stress ambiguities which are connected to morphosyntactic
-differences. ADD EXAMPLE HERE with ????????, ????????, ????????, AND ??????/??????.
-ADD ANOTHER EXAMPLE WITH the paradigms of ???? and ??????.
+We identify three different types of word stress ambiguity. \emph{Intraparadigmatic} ambiguity refers
+to homographic wordforms belonging to the same lexeme, as shown in \ref{ex:intrahom}.
 
-\subsection{Related Work}
+\ex. Intraparadigmatic homographs \label{ex:intrahom}
+\a. \rus{т\'{е}ла} \emph{t\'{e}la} `body.\textsubscript{SG-GEN}' \label{ex:bodySGGEN}
+\b. \rus{тел\'{а}} \emph{tel\'{a}} `body.\textsubscript{PL-NOM}' \label{ex:bodyPLNOM}
+
+The remaining two types of stress ambiguity occur between lexemes. \emph{Morphosyntactically
+congruent homographs} are homographs that belong to separate lexemes, and whose morphosyntactic values are identical, as shown in Table~\ref{table:MScongruent}.
+
+\begin{table}
+\begin{center}
+\begin{tabular}{c|ll}
+\bf Number-Case & \bf `castle' & \bf `lock' \\
+\hline
+\texttt{SG-NOM} & \rus{з\'{a}мок} & \rus{зам\'{о}к} \\
+\texttt{SG-GEN} & \rus{з\'{a}мка} & \rus{замк\'{а}} \\
+\texttt{PL-NOM} & \rus{з\'{a}мки} & \rus{замк\'{и}} \\
+\texttt{PL-GEN} & \rus{з\'{a}мков} & \rus{замк\'{о}в} \\
+... & ... & ... \\
+\end{tabular}
+\end{center}
+\caption{ \label{table:MScongruent} Morphosyntactically congruent homographs. }
+\end{table}
+
+\emph{Morphosyntactically incongruent homographs} are homographs that belong to separate
+lexemes, and whose morphosyntactic values are different, as shown in 
+\ref{ex:MSincongruent}.
+
+\ex. Morphosyntactically incongruent homographs \label{ex:MSincongruent} \rr{more interesting examples? \rus{нашей или было или дорога} }
+\a. \rus{попад\'{а}ть} \emph{popadát'} `fall.\textsubscript{INF-IPFV}' \\
+     \rus{поп\'{а}дать} \emph{popádat'} `fall.\textsubscript{INF-PFV}'
+\b. \rus{узнаёт} \emph{uznaët} `finds out.\textsubscript{PRS-IPFV-3SG}' \\
+     \rus{узн\'{а}ет} \emph{uznáet} `finds out.\textsubscript{FUT-PFV-3SG}'
+
+\rr{another relevant issue is whether the ambiguity can be resolved at all.
+Maybe this is something we should bring up in the discussion section.
+For example, when is stress marked in authentic by-native-for-native texts?
+A computer should not be expected to predict the stress on such words (although a computer could possibly identify words that should be marked for natives.)}
+Different types of stress position ambiguity are overcome by different information. Detailed 
+part-of-speech tagging can determine the stress of intraparadigmatic homographs and
+morphosyntactically incongruent homographs, whereas word sense disambiguation can 
+determine the stress position of morphosyntactically congruent homographs. In the case of
+running text in Russian, most stress placement ambiguity is rooted in morphosyntactic
+ambiguity, so in this study we focus on the role of part-of-speech tagging in improving
+automatic stress placement, leaving word sense disambiguation to future work.
+
+\subsection{Background and Task Definition}
 
@@
 complicate many NLP tasks, such as text-to-speech and speech-to-text. For example, speakers 
-of Czech and French sometimes type emails and other communications
+of Czech may type emails and other communications
 without standard diacritics. In order to generate speech from these texts, they must first be 
@@
 ambiguity than in standard orthography. For languages with inherently ambiguous orthography,
-it may be more precise to refer to this task as `diacritic enhancement', since it produces
+it may be more precise to refer to this as `diacritic enhancement', since it produces
 text that is less ambiguous than the standard language. In this sense, Russian orthography
-is similar to Arabic and Hebrew.
+is similar to Arabic and Hebrew, since its vowel qualities are underspecified in
+standard orthography.
 
-The following instructions are directed to authors of papers submitted
-to Nodalida-2015 or accepted for publication in its proceedings. 
-All submissions must conform to the NoDaLiDa 2015 style guides, as summarized
-and exemplified in this document. All authors are required to adhere to these specifications. 
-Templates are available from the conference web site for \LaTeX\
-(recommended), or MS~Word. If for some reason you have a problem producing this format, please contact the
-Programme Chair (Beata Megyesi; \texttt{beata.megyesi@lingfil.uu.se}) as early as
-possible. The NoDaLiDa Programme Committee reserves the right not to publish papers not
-conforming to the standard format.
+Many studies 
+of Russian text-to-speech and automatic speech recognition make note of the 
+difficulties caused by the shortcomings of their stress-marking resources (e.g. CITE Krivnova 1998).
+\cite{Xomicevic_2008} developed a set of heuristics for guessing stress placement on 
+unknown words in Russian. More recently, CITE Hall \& Sproat (2013) trained a maximum entropy
+model on a dictionary of Russian words, achieving 83.9\% accuracy for stress on unknown words 
+and 98.7\% on known words. One major shortcoming of evaluating on wordlists is 
+that they do not reflect wordform frequency and opportunities for disambiguation.
+Many of the most problematic stress ambiguities in Russian occur in 
+high-frequency wordforms, so evaluations based on wordlists - as opposed to running
+text - have inflated results. An added benefit of working with running text
+is the possibility of disambiguating homographs based on syntactic context.
 
-Submission and reviewing will be on-line, managed by the EasyChair system.
-The only accepted format for submitted papers is PDF.
-NoDaLiDa does not allow submission of supplementary materials.
-All submissions must be uploaded to EasyChair by the submission deadlines;
-submissions received after that time will not be accepted.
-To minimize the risk of network congestion, we encourage authors to upload
-their submissions as early as possible. 
-Improved versions of the submissions may be continuously uploaded until the
-final deadline.
+So far, the few studies related to automatic stress placement in Russian have been aimed at
+text-to-speech and speech recognition technology. However, the target application of our stress
+annotator is a different domain: language learning. Since standard Russian does not contain 
+stress-markings, learners are frequently unable to pronounce unknown words correctly without
+referencing a dictionary or similar resources. In the context
+of language learning, marking stress incorrectly is arguably worse than not marking it
+at all. Because of this, we want our stress annotator to be able to abstain from marking
+stress on words that it is unable to resolve with high confidence.
 
-On acceptance of a submission, precise instructions will be given on how to
-send the camera ready version; following reviewers' feedback, such instructions
-may include specific requests for modification.
-Camera-ready versions not conforming to requests for revision made by the
-Programme Committee will be considered as unsent, and hence will not be
-included in the proceedings or presented at the conference.
+\section{Automatic Stress Placement}
 
-\section{General Instructions}
+State-of-the-art part-of-speech and morphological analysis in Russian is based on finite-state
+technology. We developed free and open-source finite-state tools with the ability to analyze and
+generate stressed wordforms. A finite-state morphological analyzer generates all possible readings
+of each wordform, and the constraint grammar 
+to analyze running text, and to generate stressed wordforms.
 

### 2015-01-05 rusaccent-nodalida-2015/rusaccent.tex
more
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58230 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-\title{Automatic accent annotation for Russian}
+\title{Automatic word stress annotation of Russian running text (max 8 pages)}
 
@@
 
-\date{}
+\date{2015}
 
@@
 \begin{abstract}
-  This document contains the instructions for preparing a camera-ready
-  manuscript for the proceedings of Nodalida-2015. The document itself
-  conforms to its own specifications, and is therefore an example of
-  what your manuscript should look like. These instructions should be
-  used for both papers submitted for review and for final versions of
-  accepted papers.  Authors are asked to conform to all the directions
-  reported in this document.
+  We evaluate the effectiveness of new tools that we developed for automatically annotating 
+  word stress in Russian running text. Given an adequate lexicon with specified stress, the
+  primary obstacle for correct stress placement is disambiguating homographic forms. Using a
+  Constraint Grammar, we achieve a X\% improvement over the baseline undisambiguated performance.
 \end{abstract}
@@
 
-Russian word stress is variable, which means that the stress position in a given word is not 
-generally predictable. Furthermore, the inflecting word classes include complex patterns of
-shifting stress. Lexical stress and its attendant vowel reduction are a prominent feature of
-spoken Russian, but standard written Russian does not typically mark word stress.\footnote{
+Russian word stress is variable. The inflecting word classes include complex patterns of
+shifting stress, and a lexeme's stress pattern cannot be predicted from surface forms.
+Although lexical stress and its attendant vowel reduction are a prominent feature of
+spoken Russian, standard written Russian does not typically mark word stress.\footnote{
 Texts intended for native speakers sometimes mark stress on words that cannot be 
-disambiguated through context.} Without information about lexical stress position, 
+disambiguated through context, although this is somewhat rare.} Without information about lexical 
+stress position, 
 correctly converting written Russian text to speech is impossible. This is true both for humans
-(e.g. foreign language students) and computers (e.g. text-to-speech or automatic speech 
-recognition).
+(e.g. foreign language students) and computers (e.g. text-to-speech).
 
-We identify three different types of word stress ambiguity. \emph{Intraparadigmatic} ambiguity refers
+We identify three different types of word stress ambiguity. \emph{Intraparadigmatic} ambiguity 
+refers
 to homographic wordforms belonging to the same lexeme, as shown in \ref{ex:intrahom}.
@@
 The remaining two types of stress ambiguity occur between lexemes. \emph{Morphosyntactically
-congruent homographs} are homographs that belong to separate lexemes, and whose morphosyntactic values are identical, as shown in Table~\ref{table:MScongruent}.
+congruent homographs} are homographs that belong to separate lexemes, and whose morphosyntactic 
+values are identical, as shown in Table~\ref{table:MScongruent}.
 
@@
 \texttt{PL-GEN} & \rus{з\'{a}мков} & \rus{замк\'{о}в} \\
-... & ... & ... \\
+... & \hskip 1em ... & \hskip 1em ... \\
 \end{tabular}
@@
 
-\ex. Morphosyntactically incongruent homographs \label{ex:MSincongruent} \rr{more interesting examples? \rus{нашей или было или дорога} }
+\ex. Morphosyntactically incongruent homographs \label{ex:MSincongruent}
 \a. \rus{попад\'{а}ть} \emph{popadát'} `fall.\textsubscript{INF-IPFV}' \\
@@
     \rus{узн\'{а}ет} \emph{uznáet} `finds out.\textsubscript{FUT-PFV-3SG}'
+\c. \rus{н\'{а}шей} \emph{nášej} `our.\textsubscript{F-SG-GEN/DAT/LOC...}'\\
+    \rus{наш\'{е}й} \emph{našéj} `sew on.\textsubscript{IMP-2SG}'
+\d. \rus{дор\'{о}га} \emph{doróga} `road.\textsubscript{N-F-SG-NOM}'\\
+    \rus{дорог\'{а}} \emph{dorogá} `dear.\textsubscript{ADJ-F-SG-PRED}'
 
-\rr{another relevant issue is whether the ambiguity can be resolved at all.
-Maybe this is something we should bring up in the discussion section.
-For example, when is stress marked in authentic by-native-for-native texts?
-A computer should not be expected to predict the stress on such words (although a computer could possibly identify words that should be marked for natives.)}
-Different types of stress position ambiguity are overcome by different information. Detailed 
+\rr{another relevant issue is whether the ambiguity can be resolved at all. Maybe this is 
+something we should bring up in the discussion section. For example, when is stress marked in 
+authentic by-native-for-native texts? A computer should not be expected to predict the stress on 
+such words (although a computer could possibly identify words that should be marked for 
+natives.)} It should be noted that not all morphosyntactic ambiguity leads to stress ambiguity. For example,
+all inflecting modifiers in Russian have the same form for F-SG-GEN, F-SG-LOC, F-SG-DAT and 
+F-SG-INS. We refer to this as \emph{stress-irrelevant} ambiguity, since all readings have
+the same stress placement. Stress-irrelevant ambiguity can be intraparadigmatic and/or interlexical.
+
+Different types of stress ambiguity are overcome by different information. Detailed 
 part-of-speech tagging can determine the stress of intraparadigmatic homographs and
@@
 
-Many studies 
-of Russian text-to-speech and automatic speech recognition make note of the 
-difficulties caused by the shortcomings of their stress-marking resources (e.g. CITE Krivnova 1998).
+Many studies of Russian text-to-speech and automatic speech recognition make note of the 
+difficulties caused by the shortcomings of their stress-marking resources (e.g. CITE Krivnova 
+1998), but very few studies have targeted stress-marking itself, and they tend to
+focus on place stress on unknown words, with almost no attention to the problems that
+arise with known words. For example, 
 \cite{Xomicevic_2008} developed a set of heuristics for guessing stress placement on 
@@
 and 98.7\% on known words. One major shortcoming of evaluating on wordlists is 
-that they do not reflect wordform frequency and opportunities for disambiguation.
+that they do not reflect wordform frequency.
 Many of the most problematic stress ambiguities in Russian occur in 
-high-frequency wordforms, so evaluations based on wordlists - as opposed to running
-text - have inflated results. An added benefit of working with running text
-is the possibility of disambiguating homographs based on syntactic context.
+high-frequency wordforms\rr{might it be worth computing the difference in stress
+ambiguity between wordform lists and running text? i.e. what percentage of tokens have ambiguous
+stress.}, so evaluations based on wordlists - as opposed to running
+text - have inflated results. Furthermore, working with running text allows for the
+possibility of disambiguating homographs based on syntactic context.
 
-So far, the few studies related to automatic stress placement in Russian have been aimed at
-text-to-speech and speech recognition technology. However, the target application of our stress
+So far, the implicit target application the few studies related to automatic stress placement
+in Russian has been text-to-speech and speech recognition technology. However, the target 
+application of our stress
 annotator is a different domain: language learning. Since standard Russian does not contain 
@@
 State-of-the-art part-of-speech and morphological analysis in Russian is based on finite-state
-technology. We developed free and open-source finite-state tools with the ability to analyze and
-generate stressed wordforms. A finite-state morphological analyzer generates all possible readings
-of each wordform, and the constraint grammar 
-to analyze running text, and to generate stressed wordforms.
+technology.\footnote{add footnote or citations for Dialog, mystem, etc.} We developed free
+and open-source finite-state tools capable of analyzing and
+generating stressed wordforms. A finite-state transducer\footnote{a two-level 
+morphology, xfst/hfst, blah, blah, blah} generates all possible readings
+of each wordform, and a constraint grammar\footnote{vislcg3, blah blah} then removes
+some extra readings based on syntactic context. The ultimate success of our stress placement
+system depends on the performance of the Constraint Grammar. Ideally, the Constraint Grammar would
+successfully remove all but the correct reading for each token, but in practice some
+tokens still have more than one reading remaining. Therefore,
+we also evaluate various approaches to deal with the remaining ambiguity,
+as described in the following paragraph and Table~\ref{tab:conditions}.
 
-Manuscripts must be in two-column format.  Exceptions to the
-two-column format include the title, authors' names and complete
-addresses, which must be centered at the top of the first page, and
-any full-width figures or tables (see the guidelines in
-Subsection~\ref{ssec:fonts}). {\bf Type single-spaced.}  Start all
-pages directly under the top margin. See the guidelines later
-regarding formatting the first page (Subsection~\ref{ssec:first}).
-The manuscript should be printed single-sided and its length should
-not exceed the maximum page limit described in
-Section~\ref{sec:length}.  {\bf Do not number the pages.}
+\begin{table*}[t]
+  \centering
+  \begin{tabular}{r|c|c}
+    & \rus{кость}-N-F-SG-GEN\hskip 1em\rus{к\'{о}сти} & \rus{кость}-N-F-SG-GEN\hskip 1em\rus{к\'{о}сти} \\
+    & \rus{кость}-N-F-SG-DAT\hskip 1em\rus{к\'{о}сти} & \rus{кость}-N-F-SG-DAT\hskip 1em\rus{к\'{о}сти} \\
+    & \rus{костить}-V-IPFV-IMP\hskip 1em\rus{кост\'{и}} & \\
+    \hline
+    {\small {\tt bare}} & \rus{кости} & \rus{кости} \\
+    {\small {\tt safe}} & \rus{кости} & \rus{к\'{о}сти} \\
+    {\small {\tt guess}} & \rus{к\'{о}сти} or \rus{кост\'{и}} & \rus{к\'{о}сти}\\
+    {\small {\tt guessFreq}} & \rus{к\'{о}сти} & \rus{к\'{о}сти}
+  \end{tabular}
+  \caption{Example output of each condition, given a particular set of readings}
+  \label{tab:conditions}
+\end{table*}
+
+The {\small {\tt bare}} approach is to not mark stress on words with more than one reading.
+The {\small {\tt safe}} approach is to mark stress only on
+tokens whose morphosyntactic ambiguity is stress-irrelevant. The {\small {\tt guess}}
+approach is to randomly select one of the available readings.
+The {\small {\tt guessFreq}} approach is to select the reading that is most frequent
+in our corpus. If none of the readings is found in the corpus, then it selects the reading
+with the tagset that is most frequent in our corpus. Note that for tokens with
+stress-irrelevant ambiguity, {\small {\tt guess}} and {\small {\tt guessFreq}} produce
+the same result as the {\small {\tt safe}} method.
+
+For our baseline, we take the output of our morphological analyzer (without the Constraint 
+Grammar) in combination with {\small {\tt bare}}, {\small {\tt safe}}, {\small {\tt guess}} and {\small {\tt guessFreq}}. The purpose of this study is to compare the
 

### 2015-01-07 rusaccent-nodalida-2015/rusaccent.tex
more
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58258 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-We identify three different types of word stress ambiguity. \emph{Intraparadigmatic} ambiguity 
-refers
+We identify three different types of word stress ambiguity. \emph{Intraparadigmatic}
+stress ambiguity refers
 to homographic wordforms belonging to the same lexeme, as shown in \ref{ex:intrahom}.
@@
 The remaining two types of stress ambiguity occur between lexemes. \emph{Morphosyntactically
-congruent homographs} are homographs that belong to separate lexemes, and whose morphosyntactic 
-values are identical, as shown in Table~\ref{table:MScongruent}.
+congruent} stress ambiguity occurs between homographs that belong to separate lexemes, 
+and whose morphosyntactic values are identical, as shown in Table~\ref{table:MScongruent}.
 
@@
 
-\emph{Morphosyntactically incongruent homographs} are homographs that belong to separate
-lexemes, and whose morphosyntactic values are different, as shown in 
+\emph{Morphosyntactically incongruent} stress ambiguity occurs between homographs
+that belong to separate lexemes, and whose morphosyntactic values are different, as shown in 
 \ref{ex:MSincongruent}.
@@
 such words (although a computer could possibly identify words that should be marked for 
-natives.)} It should be noted that not all morphosyntactic ambiguity leads to stress ambiguity. For example,
-all inflecting modifiers in Russian have the same form for F-SG-GEN, F-SG-LOC, F-SG-DAT and 
+natives.)} It should be noted that not all morphosyntactic ambiguity leads to stress ambiguity. For example, all inflecting modifiers have the same form for F-SG-GEN, F-SG-LOC, F-SG-DAT and 
 F-SG-INS. We refer to this as \emph{stress-irrelevant} ambiguity, since all readings have
@@
 disambiguate otherwise homographic wordforms, so missing diacritics can
-complicate many NLP tasks, such as text-to-speech and speech-to-text. For example, speakers 
+complicate many NLP tasks, such as text-to-speech. For example, speakers 
 of Czech may type emails and other communications
@@
 1998), but very few studies have targeted stress-marking itself, and they tend to
-focus on place stress on unknown words, with almost no attention to the problems that
+focus on placing stress on unknown words, with almost no attention to the problems that
 arise with known words. For example, 
@@
 So far, the implicit target application the few studies related to automatic stress placement
-in Russian has been text-to-speech and speech recognition technology. However, the target 
+in Russian has been text-to-speech and automatic speech recognition. However, the target 
 application of our stress
@@
 
-State-of-the-art part-of-speech and morphological analysis in Russian is based on finite-state
-technology.\footnote{add footnote or citations for Dialog, mystem, etc.} We developed free
-and open-source finite-state tools capable of analyzing and
-generating stressed wordforms. A finite-state transducer\footnote{a two-level 
+State-of-the-art morphological analysis in Russian is based on finite-state
+technology.\footnote{add footnote or citations for Dialog, mystem, etc.} To our knowledge,
+no open-source resources are available for analyzing and generating stressed wordforms.
+Therefore, we developed free and open-source finite-state tools capable of analyzing and
+generating stressed wordforms. A finite-state transducer\footnote{Using two-level 
 morphology, xfst/hfst, blah, blah, blah} generates all possible readings
@@
   \end{tabular}
-  \caption{Example output of each condition, given a particular set of readings}
+  \caption{Example output of each stress placement approach, given a particular set of readings}
   \label{tab:conditions}
@@
 For our baseline, we take the output of our morphological analyzer (without the Constraint 
-Grammar) in combination with {\small {\tt bare}}, {\small {\tt safe}}, {\small {\tt guess}} and {\small {\tt guessFreq}}. The purpose of this study is to compare the
+Grammar) in combination with the {\small {\tt bare}}, {\small {\tt safe}}, 
+{\small {\tt guess}} and {\small {\tt guessFreq}} approaches.
 
-\subsection{Electronically-available resources}
+\subsection{Stress Corpus}
 
-Nodalida-2015 provides this description in \LaTeX2e ({\small {\tt
-    nodalida2015.tex}}) and PDF format ({\small {\tt nodalida2015.pdf}}),
-along with the \LaTeX2e style file used to format it ({\small {\tt
-    nodalida2015.sty}}) and an ACL bibliography style ({\small {\tt
-    acl.bst}}) in case you want to use the Bib\TeX\ reference management
-software. These files are all available at
-\url{http://stp.lingfil.uu.se/~bea/nodalida15/}. A Microsoft Word template file ({\small
-  {\tt nodalida2015.docx}}) is also available at the same URL. We strongly
-recommend the use of these style files, which have been appropriately
-tailored for the Nodalida-2015 proceedings. If you have an option, we
-recommend that you use the \LaTeX2e version. \textbf{If you will be
-  using the Microsoft Word template, we suggest that you anonymize
-  your source file so that the pdf produced does not retain your
-  identity.}  This can be done by removing any personal information
-from your source document properties.
+Obtaining a suitable corpus for this study Russian texts with marked word stress are relatively rare, except in materials
+for language learners, which are predominantly proprietary.
 

### 2015-01-14 rusaccent-nodalida-2015/rusaccent.tex
results\!
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58322 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \newcommand{\rus}[1]{\foreignlanguage{russian}{#1}}
-\newcommand{\ft}[1]{\marginpar{\scriptsize F: #1}}
-\newcommand{\rr}[1]{\marginpar{\scriptsize R: #1}}
+
+\newcommand{\ft}[1]{\marginpar{\scriptsize F: #1}} % Fran's comments
+\newcommand{\rr}[1]{\marginpar{\scriptsize R: #1}} % Rob's comments
+\newcommand{\dm}[1]{\marginpar{\scriptsize D: #1}} % Detmar's comments
+\newcommand{\lj}[1]{\marginpar{\scriptsize L: #1}} % Laura's comments
 
@@
 \begin{abstract}
-  We evaluate the effectiveness of new tools that we developed for automatically annotating 
+  We evaluate the effectiveness of tools we developed for automatically annotating 
   word stress in Russian running text. Given an adequate lexicon with specified stress, the
@@
 
-\section{Automatic Stress Placement}
-
-State-of-the-art morphological analysis in Russian is based on finite-state
-technology.\footnote{add footnote or citations for Dialog, mystem, etc.} To our knowledge,
-no open-source resources are available for analyzing and generating stressed wordforms.
-Therefore, we developed free and open-source finite-state tools capable of analyzing and
-generating stressed wordforms. A finite-state transducer\footnote{Using two-level 
-morphology, xfst/hfst, blah, blah, blah} generates all possible readings
-of each wordform, and a constraint grammar\footnote{vislcg3, blah blah} then removes
-some extra readings based on syntactic context. The ultimate success of our stress placement
-system depends on the performance of the Constraint Grammar. Ideally, the Constraint Grammar would
-successfully remove all but the correct reading for each token, but in practice some
-tokens still have more than one reading remaining. Therefore,
-we also evaluate various approaches to deal with the remaining ambiguity,
-as described in the following paragraph and Table~\ref{tab:conditions}.
-
 \begin{table*}[t]
@@
 
+\subsection{Stress Corpus}
+
+Russian texts with marked word stress are relatively rare, except in materials
+for language learners, which are predominantly proprietary. Our corpus was collected from
+free texts on Russian language-learning websites.\footnote{For a list of URL's, please 
+contact the authors.} This small corpus (N=7689) is representative of texts 
+that learners of Russian are likely to encounter in their studies. These texts 
+include excerpts from well-known literary works, as well as dialogs and 
+prose and individual sentences that were written for learners.
+
+Unfortunately, the general practice for marking stress in Russian is to
+\emph{not} mark stress on monosyllabic tokens, effectively assuming that
+all monosyllabics are stressed. However, this approach 
+is not well-motivated. Many words -- both monosyllabic and multisyllabic -- are 
+unstressed, especially among prepositions, conjunctions, and particles.
+\rr{multisyllabics that can be unstressed: \rus{через, между}} Furthermore, there are many
+high-frequency monosyllabic homographs that can be either stressed or unstressed, 
+depending on their part of speech, or particular collocations. Clearly, for
+such words, one cannot assume that they are stressed on the basis of their syllable count.
+
+Based on these considerations, we built our tools to mark stress without regard to
+the number of syllables in a word. However, because our gold corpus texts do not
+mark stress on monosyllabic words, we cannot evaluate our annotation of those words.
+
+\section{Automatic Stress Placement}
+
+State-of-the-art morphological analysis in Russian is based on finite-state
+technology.\footnote{add footnote or citations for Dialog, mystem, etc.} To our knowledge,
+no open-source resources are available for analyzing and generating stressed wordforms.
+Therefore, we developed free and open-source finite-state tools capable of analyzing and
+generating stressed wordforms. A finite-state transducer\footnote{Using two-level 
+morphology, xfst/hfst, blah, blah, blah} (FST) generates all possible morphosyntactic readings
+of each wordform, and a constraint grammar\footnote{vislcg3, blah blah} then removes
+some extra readings based on syntactic context. The ultimate success of our stress placement
+system depends on the performance of the Constraint Grammar. Ideally, the Constraint Grammar would
+successfully remove all but the correct reading for each token, but in practice some
+tokens still have more than one reading remaining. Therefore,
+we also evaluate various approaches to deal with the remaining ambiguity,
+as described in the following paragraph and Table~\ref{tab:conditions}.
+
 The {\small {\tt bare}} approach is to not mark stress on words with more than one reading.
@@
 tokens whose morphosyntactic ambiguity is stress-irrelevant. The {\small {\tt guess}}
-approach is to randomly select one of the available readings.
-The {\small {\tt guessFreq}} approach is to select the reading that is most frequent
-in our corpus. If none of the readings is found in the corpus, then it selects the reading
-with the tagset that is most frequent in our corpus. Note that for tokens with
+approach is to randomly select one of the available readings, and abstain if no readings
+are available. The {\small {\tt guessFreq}} approach is to select the reading that is most frequent.
+\footnote{Frequency data taken from a separate hand-disambiguated corpus.} If none of the readings is found in the corpus, then it selects the reading
+with the tag sequence (lemma removed) that is most frequent in our corpus. If the tag sequence is not found
+in our corpus, then it backs off to the  {\small {\tt guess}} algorithm. Note that for tokens with
 stress-irrelevant ambiguity, {\small {\tt guess}} and {\small {\tt guessFreq}} produce
 the same result as the {\small {\tt safe}} method.
+
+So far, the approaches discussed are
+dependent on the availability of readings from the FST. Two final approaches address unknown tokens.
+The {\small {\tt guessSyll}} and {\small {\tt guessFreqSyll}} approaches are identical to 
+{\small {\tt guess}} and {\small {\tt guessFreq}}, respectively,
+except that for unknown tokens, they places the stress near the end of the stem.\footnote{Determining 
+which vowel is stem-final
+is itself a difficult task. Our algorithm merely places the stress on the last vowel that is 
+followed by a consonant.} \rr{Explain why stem-final is a good guess (and cite someone?)}
+\rr{Discuss the success of the guessSyll approach in the discussion section.}
 
@@
 
-\subsection{Stress Corpus}
+We also compare our outcomes with the RussianGram\footnote{http://russiangram.com/} plugin 
+for the Google Chrome web browser. RussianGram is not open-source, so we
+can only guess what technologies support the service. In any case, it provides a
+meaningful reference point for the success of each of the methods described above.
 
-Obtaining a suitable corpus for this study Russian texts with marked word stress are relatively rare, except in materials
-for language learners, which are predominantly proprietary.
+\section{Results}
 
-\subsection{Format of Electronic Manuscript}
-\label{sect:pdf}
+\begin{table*}[t]
+  \centering
+  \begin{tabular}{r | r r r | r r }
+    approach & correct\% & error\% & abstention\% & totTry\% & totFail\% \\
+    \hline
+    \hline
+    {\small {\tt noCGbare}} & 30.43 & 0.17 & 69.39 & 30.61 & 69.57 \\
+    {\small {\tt noCGsafe}} & 90.07 & 0.49 & 9.44 & 90.56 & 9.93 \\
+    {\small {\tt noCGguess}} & 94.34 & 3.36 & 2.30 & 97.70 & 5.66 \\
+    {\small {\tt noCGguessFreq}} & 95.53 & 2.59 & 1.88 & 98.12 & 4.47 \\
+    {\small {\tt noCGguessSyll}} & 94.99 & 4.05 & 0.96 & 99.04 & 5.01 \\
+    {\small {\tt noCGguessFreqSyll}} & 95.83 & 3.46 & 0.72 & 99.28 & 4.17 \\
+    \hline
+    {\small {\tt CGbare}} & 45.78 & 0.44 & 53.78 & 46.22 & 54.22 \\
+    {\small {\tt CGsafe}} & 93.21 & 0.74 & 6.05 & 93.95 & 6.79 \\
+    {\small {\tt CGguess}} & 95.50 & 2.59 & 1.90 & 98.10 & 4.50 \\
+    {\small {\tt CGguessFreq}} & 95.73 & 2.40 & 1.88 & 98.12 & 4.27 \\
+    {\small {\tt CGguessSyll}} & 95.92 & 3.33 & 0.74 & 99.26 & 4.08 \\
+    {\small {\tt CGguessFreqSyll}} & 96.15 & 3.14 & 0.72 & 99.28 & 3.85 \\
+    \hline
+    {\small {\tt RussianGram}} & 90.09 & 0.79 & 9.12 & 90.88 & 9.91
+  \end{tabular}
+  \caption{Results of stress placement task evaluation. (N = 4048)}
+  \label{tab:results}
+\end{table*}
 
-For the production of the electronic manuscript you must use Adobe's
-Portable Document Format (PDF). 
-%% Add instructions about using pdflatex
-This format can be generated directly from \LaTeX2e files or from
-postscript ones. On Linux/Unix-like systems, you can use {\tt
-  pdflatex} to generate a PDF file from \LaTeX2e files, or {\tt
-  ps2pdf} to convert from postscript to PDF. 
-%%
-In Microsoft Windows, you can use Adobe's Distiller, or if you have
-\texttt{cygwin} installed, you can use \texttt{dvipdf} or
-\texttt{ps2pdf}. Note that some word processing programs generate PDF
-which may not include all the necessary fonts (esp. tree diagrams,
-symbols). When you print or create the PDF file, there is usually an
-option in your printer setup to include none, all or just non-standard
-fonts.  Please make sure that you select the option of including ALL
-the fonts. {\em Before sending it, test your PDF by printing it from a
-  computer different from the one where it was created.} Moreover,
-some word processors may generate very large postscript/PDF files,
-where each page is rendered as an image. Such images may reproduce
-poorly. In this case, try alternative ways to obtain the postscript
-and/or PDF. One way on some systems is to install a driver for a
-postscript printer, send your document to the printer specifying
-``Output to a file'', then convert the file to PDF.
-Please keep in mind that it is of utmost importance to use \textbf{A4 format}.
+Single-syllable words and words that are unstressed in the gold standard were
+not evaluated for reasons discussed above, leaving 4048 tokens for evaluation.
+Recall was computed by testing whether stressed 
+multisyllabic wordforms in the gold standard stress corpus were found in the 
+finite-state transducer. Out of 4048 tokens, 3949 were found in the FST, which is equal to 97.55\%.
+This number represents the ceiling for rule-based methods relying on the FST.
+Higher scores are only achievable by expanding the FST or by using syllable-guessing algorithms.
+After running the Constraint Grammar, 3937 tokens were found in the FST, which
+indicates that the Constraint Grammar ruled out 0.20\% of the correct forms.
+By inspecting these errors, we can improve the Constraint Grammar.
 
-%%Removed by YP (use of the \special command above instead):
-%% It is of utmost importance to specify the \textbf{A4 format} (21 cm x
-%% 29.7 cm) when formatting the paper. 
-%% When working with {\tt dvips}, for instance, one should specify {\tt -t a4}.
+Results were compiled for each of the nine conditions discussed above: 
+\emph{without} the Constraint Grammar ({\small {\tt noCG}}) X 4 approaches, 
+\emph{with} the Constraint Grammar ({\small {\tt CG}}) X 4 approaches, 
+and RussianGram ({\small {\tt RG}}). 
+Failures to correctly place stress were split into two categories: errors and abstentions. 
+If the stress tool outputs a stressed wordform, and it is incorrect, then it is
+counted as an `error'. If the stress tool outputs an unstressed wordform, then
+it is counted as an `abstention'. The sum of the errors and abstentions is referred
+to as `totFail'. The sum of correct attempts and `errors' is referred to as `totTry'.
+Results are given in Table~\ref{tab:results}.
 
-Print-outs of the PDF file on A4 paper should be identical to the
-hardcopy version. If you cannot meet the above requirements about the
-production of your electronic submission, please contact the
-program chair above as soon as possible.
+WE NEED TO COMPUTE SIGNIFICANCE ON THESE RESULTS
 
-\subsection{Layout}
-\label{ssec:layout}
+\section{Discussion}
 
-Format manuscripts two columns to a page, in the manner these
-instructions are formatted. The exact dimensions for a page on A4
-paper are:
+%\section*{Acknowledgments}
+%
+%Do not number the acknowledgment section. Do not include this section
+%when submitting your paper for review.
 
-\begin{itemize}
-\item Left and right margins: 2.5 cm
-\item Top margin: 2.5 cm
-\item Bottom margin: 2.5 cm
-\item Column width: 7.7 cm
-\item Column height: 24.7 cm
-\item Gap between columns: 0.6 cm
-\end{itemize}
-
-\noindent Papers should not be submitted on any other paper size.
- If you cannot meet the above requirements about the production of your electronic submission, 
-please contact the program chair above as soon as possible.
-
-\subsection{Fonts}
-\label{ssec:fonts}
-For reasons of uniformity, Adobe's {\bf Times Roman} font should be
-used. In \LaTeX2e{} this is accomplished by putting
-
-\begin{quote}
-\begin{verbatim}
-\usepackage{times}
-\usepackage{latexsym}
-\end{verbatim}
-\end{quote}
-in the preamble. If Times Roman is unavailable, use {\bf Computer
-  Modern Roman} (\LaTeX2e{}'s default).  Note that the latter is about
-  10\% less dense than Adobe's Times Roman font.
-
-\begin{table}[h]
-\begin{center}
-\begin{tabular}{|l|rl|}
-\hline \bf Type of Text & \bf Font Size & \bf Style \\ \hline
-paper title & 15 pt & bold \\
-author names & 12 pt & bold \\
-author affiliation & 12 pt & \\
-the word ``Abstract'' & 12 pt & bold \\
-section titles & 12 pt & bold \\
-document text & 11 pt  &\\
-captions & 11 pt & \\
-abstract text & 10 pt & \\
-bibliography & 10 pt & \\
-footnotes & 9 pt & \\
-\hline
-\end{tabular}
-\end{center}
-\caption{\label{font-table} Font guide. }
-\end{table}
-
-\subsection{The First Page}
-\label{ssec:first}
-
-Center the title, author's name(s) and affiliation(s) across both
-columns. Do not use footnotes for affiliations. Do not include the
-paper ID number assigned during the submission process. Use the
-two-column format only when you begin the abstract.
-
-{\bf Title}: Place the title centered at the top of the first page, in
-a 15-point bold font. (For a complete guide to font sizes and styles, 
-see Table~\ref{font-table}.) Long titles should be typed on two lines without
-a blank line intervening. Approximately, put the title at 2.5 cm from
-the top of the page, followed by a blank line, then the author's
-names(s), and the affiliation on the following line. Do not use only
-initials for given names (middle initials are allowed). Do not format surnames
-in all capitals (e.g., use ``Schlangen'' not ``SCHLANGEN'').
-Do not format title and section headings in all capitals as well
-except for proper names (such as ``BLEU'') that are conventionally
-in all capitals.
-The affiliation should contain the author's complete address, and if
-possible, an electronic mail address. Leave about 2 cm between the
-affiliation and the body of the first page.
-The title, author names and addresses should be completely
-identical to those entered to the electronical paper submission
-website in order to maintain the consistency of author information
-among all publications of the conference.
-
-{\bf Abstract}: Type the abstract at the beginning of the first
-column. The width of the abstract text should be smaller than the
-width of the columns for the text in the body of the paper by about
-0.6 cm on each side. Center the word {\bf Abstract} in a 12 point bold
-font above the body of the abstract. The abstract should be a concise
-summary of the general thesis and conclusions of the paper. It should
-be no longer than 200 words. The abstract text should be in 10 point font.
-
-{\bf Text}: Begin typing the main body of the text immediately after
-the abstract, observing the two-column format as shown in 
-the present document. Do not include page numbers.
-
-{\bf Indent} when starting a new paragraph. Use 11 points for text and 
-subsection headings, 12 points for section headings and 15 points for
-the title.
-
-\subsection{Sections}
-
-{\bf Headings}: Type and label section and subsection headings in the
-style shown on the present document.  Use numbered sections (Arabic
-numerals) in order to facilitate cross references. Number subsections
-with the section number and the subsection number separated by a dot,
-in Arabic numerals. Do not number subsubsections.
-
-{\bf Citations}: Citations within the text appear
-in parentheses as~\cite{Gusfield:97} or, if the author's name appears in
-the text itself, as Gusfield~\shortcite{Gusfield:97}. 
-Append lowercase letters to the year in cases of ambiguity.  
-Treat double authors as in~\cite{Aho:72}, but write as
- in~\cite{Chandra:81} when more than two authors are involved. Collapse multiple citations as
-in~\cite{Gusfield:97,Aho:72}. Also refrain from using full citations as sentence constituents. We
-suggest that instead of
-\begin{quote}
-  ``\cite{Gusfield:97} showed that ...''
-\end{quote}
-you use
-\begin{quote}
-``Gusfield \shortcite{Gusfield:97}   showed that ...''
-\end{quote}
-
-If you are using the provided \LaTeX{} and Bib\TeX{} style files, you
-can use the command \verb|\newcite| to get ``author (year)'' citations.
-
-As reviewing will be double-blind, the submitted version of the papers should not include the
-authors' names and affiliations. Furthermore, self-references that
-reveal the author's identity, e.g.,
-\begin{quote}
-``We previously showed \cite{Gusfield:97} ...''  
-\end{quote}
-should be avoided. Instead, use citations such as 
-\begin{quote}
-``Gusfield \shortcite{Gusfield:97}
-previously showed ... ''
-\end{quote}
-
-\textbf{Please do not  use anonymous citations} and  do not include acknowledgements 
-when submitting your paper. Papers that do not conform
-to these requirements may be rejected without review.
-
-\textbf{References}: Gather the full set of references together under
-the heading {\bf References}; place the section before any Appendices,
-unless they contain references. Arrange the references alphabetically
-by first author, rather than by order of occurrence in the text.
-Provide as complete a citation as possible, using a consistent format,
-such as the one for {\em Computational Linguistics\/} or the one in the 
-{\em Publication Manual of the American 
-Psychological Association\/}~\cite{APA:83}.  Use of full names for
-authors rather than initials is preferred.  A list of abbreviations
-for common computer science journals can be found in the ACM 
-{\em Computing Reviews\/}~\cite{ACM:83}.
-
-The \LaTeX{} and Bib\TeX{} style files provided roughly fit the
-American Psychological Association format, allowing regular citations, 
-short citations and multiple citations as described above.
-
-{\bf Appendices}: Appendices, if any, directly follow the text and the
-references (but see above).  Letter them in sequence and provide an
-informative title: {\bf Appendix A. Title of Appendix}.
-
-\textbf{Acknowledgement} section should appear in accepted manuscripts
-only. It should go as a last section immediately
-before the references.  Do not number the acknowledgement section.
-
-\subsection{Footnotes}
-
-{\bf Footnotes}: Put footnotes at the bottom of the page and use 9
-points text. They may be numbered or referred to by asterisks or other
-symbols.\footnote{This is how a footnote should appear.} Footnotes
-should be separated from the text by a line.\footnote{Note the line
-separating the footnotes from the text.}
-
-\subsection{Graphics}
-
-{\bf Illustrations}: Place figures, tables, and photographs in the
-paper near where they are first discussed, rather than at the end, if
-possible.  Wide illustrations may run across both columns.  Color
-illustrations are allowed, provided you have verified that  
-they will be understandable when printed in black ink.
-
-{\bf Captions}: Provide a caption for every illustration; number each one
-sequentially in the form:  ``Figure 1. Caption of the Figure.'' ``Table 1.
-Caption of the Table.''  Type the captions of the figures and 
-tables below the body, using 11 point text.
-
-\section{Translation of non-English Terms}
-
-It is also advised to supplement non-English characters and terms
-with appropriate transliterations and/or translations
-since not all readers understand all such characters and terms.
-Inline transliteration or translation can be represented in
-the order of: original-form transliteration ``translation''.
-
-\section{Length of Submission}
-\label{sec:length}
-
-Long papers may consist of up to 8 pages of content (excluding
-references) and short papers may consist of up to 
-four (4) pages plus 2 pages for references in the proceedings.  Papers that do not conform to the
-specified length and formatting requirements are subject to be
-rejected without review.
-
-%\section{Other Issues}
-
-%Papers that had software and/or dataset submitted for the review
-%process should also submit it with the camera-ready paper. Besides,
-%the software and/or dataset should not be anonymous.
-
-% Please note that the publications of Nodalida-2015 will be publicly
-% available at ACL Anthology (http://aclweb.org/anthology-new/) in May,
-% 2015. Since some of the authors may have plans to file patents related to their papers in
-% the conference, we are reminding authors that May 12th, 2015 may be
-% considered to be the official publication date, i.e., the opening
-% day of the conference.
-
-\section*{Acknowledgments}
-
-Do not number the acknowledgment section. Do not include this section
-when submitting your paper for review.
-
-% If you use BibTeX with a bib file named eacl2014.bib, 
-% you should add the following two lines:
 \bibliographystyle{acl}

### 2015-01-14 engkaz-eamt-2015/engkaz-eamt-2015.tex
add art
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58326 72bbbca6-d526-0410-a7d9-f06f51895060

--- 2015-01-14 rusaccent-nodalida-2015/rusaccent.tex
@@
-%
-% File nodalida2015.tex
-%
-% Contact beata.megyesi@lingfil.uu.se
-%
-% Based on the instruction file for EACL 2014
-% which in turn was based on the instruction files for previous 
-% ACL and EACL conferences.
+\documentclass[a4paper,11pt]{article}
 
-\documentclass[11pt]{article}
-\usepackage[T2A,T1]{fontenc}
-\usepackage[utf8]{inputenc}
-\usepackage[russian,english]{babel}
-
-\usepackage{times}
-\usepackage{latexsym}
-\usepackage{fixltx2e} %allows subscripts
-%\usepackage{mathptmx}
-%\usepackage{txfonts}
 \usepackage{url}
-\special{papersize=210mm,297mm} % to avoid having to use "-t a4" with dvips 
-%\setlength\titlebox{6.5cm}  % You can expand the title box if you really have to
-
-\usepackage{nodalida2015}
-
-\usepackage{linguex}
-
-\newcommand{\rus}[1]{\foreignlanguage{russian}{#1}}
-
-\newcommand{\ft}[1]{\marginpar{\scriptsize F: #1}} % Fran's comments
-\newcommand{\rr}[1]{\marginpar{\scriptsize R: #1}} % Rob's comments
-\newcommand{\dm}[1]{\marginpar{\scriptsize D: #1}} % Detmar's comments
-\newcommand{\lj}[1]{\marginpar{\scriptsize L: #1}} % Laura's comments
-
-\title{Automatic word stress annotation of Russian running text (max 8 pages)}
-
-\author{Author1 \\
-  Affiliation / Address line 1 \\
-  Affiliation / Address line 2 \\
-  Affiliation / Address line 3 \\
-  {\tt email@domain.com} \\\And
-  Author2 \\
-  Affiliation / Address line 1 \\
-  Affiliation / Address line 2 \\
-  Affiliation / Address line 3 \\
-  {\tt email@domain.com} \\}
-
-\date{2015}
 
 \begin{document}
-\maketitle
-\begin{abstract}
-  We evaluate the effectiveness of tools we developed for automatically annotating 
-  word stress in Russian running text. Given an adequate lexicon with specified stress, the
-  primary obstacle for correct stress placement is disambiguating homographic forms. Using a
-  Constraint Grammar, we achieve a X\% improvement over the baseline undisambiguated performance.
-\end{abstract}
-
-%\section{Credits}
-%
-%The template had a credits section for the history of the template, but we don't need
-%a Credits section, right?
 
@@
 
-Russian word stress is variable. The inflecting word classes include complex patterns of
-shifting stress, and a lexeme's stress pattern cannot be predicted from surface forms.
-Although lexical stress and its attendant vowel reduction are a prominent feature of
-spoken Russian, standard written Russian does not typically mark word stress.\footnote{
-Texts intended for native speakers sometimes mark stress on words that cannot be 
-disambiguated through context, although this is somewhat rare.} Without information about lexical 
-stress position, 
-correctly converting written Russian text to speech is impossible. This is true both for humans
-(e.g. foreign language students) and computers (e.g. text-to-speech).
-
-We identify three different types of word stress ambiguity. \emph{Intraparadigmatic}
-stress ambiguity refers
-to homographic wordforms belonging to the same lexeme, as shown in \ref{ex:intrahom}.
-
-\ex. Intraparadigmatic homographs \label{ex:intrahom}
-\a. \rus{т\'{е}ла} \emph{t\'{e}la} `body.\textsubscript{SG-GEN}' \label{ex:bodySGGEN}
-\b. \rus{тел\'{а}} \emph{tel\'{a}} `body.\textsubscript{PL-NOM}' \label{ex:bodyPLNOM}
-
-The remaining two types of stress ambiguity occur between lexemes. \emph{Morphosyntactically
-congruent} stress ambiguity occurs between homographs that belong to separate lexemes, 
-and whose morphosyntactic values are identical, as shown in Table~\ref{table:MScongruent}.
-
-\begin{table}
-\begin{center}
-\begin{tabular}{c|ll}
-\bf Number-Case & \bf `castle' & \bf `lock' \\
-\hline
-\texttt{SG-NOM} & \rus{з\'{a}мок} & \rus{зам\'{о}к} \\
-\texttt{SG-GEN} & \rus{з\'{a}мка} & \rus{замк\'{а}} \\
-\texttt{PL-NOM} & \rus{з\'{a}мки} & \rus{замк\'{и}} \\
-\texttt{PL-GEN} & \rus{з\'{a}мков} & \rus{замк\'{о}в} \\
-... & \hskip 1em ... & \hskip 1em ... \\
-\end{tabular}
-\end{center}
-\caption{ \label{table:MScongruent} Morphosyntactically congruent homographs. }
-\end{table}
-
-\emph{Morphosyntactically incongruent} stress ambiguity occurs between homographs
-that belong to separate lexemes, and whose morphosyntactic values are different, as shown in 
-\ref{ex:MSincongruent}.
-
-\ex. Morphosyntactically incongruent homographs \label{ex:MSincongruent}
-\a. \rus{попад\'{а}ть} \emph{popadát'} `fall.\textsubscript{INF-IPFV}' \\
-    \rus{поп\'{а}дать} \emph{popádat'} `fall.\textsubscript{INF-PFV}'
-\b. \rus{узнаёт} \emph{uznaët} `finds out.\textsubscript{PRS-IPFV-3SG}' \\
-    \rus{узн\'{а}ет} \emph{uznáet} `finds out.\textsubscript{FUT-PFV-3SG}'
-\c. \rus{н\'{а}шей} \emph{nášej} `our.\textsubscript{F-SG-GEN/DAT/LOC...}'\\
-    \rus{наш\'{е}й} \emph{našéj} `sew on.\textsubscript{IMP-2SG}'
-\d. \rus{дор\'{о}га} \emph{doróga} `road.\textsubscript{N-F-SG-NOM}'\\
-    \rus{дорог\'{а}} \emph{dorogá} `dear.\textsubscript{ADJ-F-SG-PRED}'
-
-\rr{another relevant issue is whether the ambiguity can be resolved at all. Maybe this is 
-something we should bring up in the discussion section. For example, when is stress marked in 
-authentic by-native-for-native texts? A computer should not be expected to predict the stress on 
-such words (although a computer could possibly identify words that should be marked for 
-natives.)} It should be noted that not all morphosyntactic ambiguity leads to stress ambiguity. For example, all inflecting modifiers have the same form for F-SG-GEN, F-SG-LOC, F-SG-DAT and 
-F-SG-INS. We refer to this as \emph{stress-irrelevant} ambiguity, since all readings have
-the same stress placement. Stress-irrelevant ambiguity can be intraparadigmatic and/or interlexical.
-
-Different types of stress ambiguity are overcome by different information. Detailed 
-part-of-speech tagging can determine the stress of intraparadigmatic homographs and
-morphosyntactically incongruent homographs, whereas word sense disambiguation can 
-determine the stress position of morphosyntactically congruent homographs. In the case of
-running text in Russian, most stress placement ambiguity is rooted in morphosyntactic
-ambiguity, so in this study we focus on the role of part-of-speech tagging in improving
-automatic stress placement, leaving word sense disambiguation to future work.
-
-\subsection{Background and Task Definition}
-
-Automatic stress placement in Russian is similar to Diacritic Restoration, a task
-which has received increasing interest over the last 20 years. Generally speaking, diacritics
-disambiguate otherwise homographic wordforms, so missing diacritics can
-complicate many NLP tasks, such as text-to-speech. For example, speakers 
-of Czech may type emails and other communications
-without standard diacritics. In order to generate speech from these texts, they must first be 
-normalized by restoring diacritics. A slightly different
-situation arises with Arabic and Hebrew, since standard orthography lacks vowel diacritics
-altogether. For such languages, the `restoration' of vowel diacritics results in less
-ambiguity than in standard orthography. For languages with inherently ambiguous orthography,
-it may be more precise to refer to this as `diacritic enhancement', since it produces
-text that is less ambiguous than the standard language. In this sense, Russian orthography
-is similar to Arabic and Hebrew, since its vowel qualities are underspecified in
-standard orthography.
-
-Many studies of Russian text-to-speech and automatic speech recognition make note of the 
-difficulties caused by the shortcomings of their stress-marking resources (e.g. CITE Krivnova 
-1998), but very few studies have targeted stress-marking itself, and they tend to
-focus on placing stress on unknown words, with almost no attention to the problems that
-arise with known words. For example, 
-\cite{Xomicevic_2008} developed a set of heuristics for guessing stress placement on 
-unknown words in Russian. More recently, CITE Hall \& Sproat (2013) trained a maximum entropy
-model on a dictionary of Russian words, achieving 83.9\% accuracy for stress on unknown words 
-and 98.7\% on known words. One major shortcoming of evaluating on wordlists is 
-that they do not reflect wordform frequency.
-Many of the most problematic stress ambiguities in Russian occur in 
-high-frequency wordforms\rr{might it be worth computing the difference in stress
-ambiguity between wordform lists and running text? i.e. what percentage of tokens have ambiguous
-stress.}, so evaluations based on wordlists - as opposed to running
-text - have inflated results. Furthermore, working with running text allows for the
-possibility of disambiguating homographs based on syntactic context.
-
-So far, the implicit target application the few studies related to automatic stress placement
-in Russian has been text-to-speech and automatic speech recognition. However, the target 
-application of our stress
-annotator is a different domain: language learning. Since standard Russian does not contain 
-stress-markings, learners are frequently unable to pronounce unknown words correctly without
-referencing a dictionary or similar resources. In the context
-of language learning, marking stress incorrectly is arguably worse than not marking it
-at all. Because of this, we want our stress annotator to be able to abstain from marking
-stress on words that it is unable to resolve with high confidence.
-
-\begin{table*}[t]
-  \centering
-  \begin{tabular}{r|c|c}
-    & \rus{кость}-N-F-SG-GEN\hskip 1em\rus{к\'{о}сти} & \rus{кость}-N-F-SG-GEN\hskip 1em\rus{к\'{о}сти} \\
-    & \rus{кость}-N-F-SG-DAT\hskip 1em\rus{к\'{о}сти} & \rus{кость}-N-F-SG-DAT\hskip 1em\rus{к\'{о}сти} \\
-    & \rus{костить}-V-IPFV-IMP\hskip 1em\rus{кост\'{и}} & \\
-    \hline
-    {\small {\tt bare}} & \rus{кости} & \rus{кости} \\
-    {\small {\tt safe}} & \rus{кости} & \rus{к\'{о}сти} \\
-    {\small {\tt guess}} & \rus{к\'{о}сти} or \rus{кост\'{и}} & \rus{к\'{о}сти}\\
-    {\small {\tt guessFreq}} & \rus{к\'{о}сти} & \rus{к\'{о}сти}
-  \end{tabular}
-  \caption{Example output of each stress placement approach, given a particular set of readings}
-  \label{tab:conditions}
-\end{table*}
-
-\subsection{Stress Corpus}
-
-Russian texts with marked word stress are relatively rare, except in materials
-for language learners, which are predominantly proprietary. Our corpus was collected from
-free texts on Russian language-learning websites.\footnote{For a list of URL's, please 
-contact the authors.} This small corpus (N=7689) is representative of texts 
-that learners of Russian are likely to encounter in their studies. These texts 
-include excerpts from well-known literary works, as well as dialogs and 
-prose and individual sentences that were written for learners.
-
-Unfortunately, the general practice for marking stress in Russian is to
-\emph{not} mark stress on monosyllabic tokens, effectively assuming that
-all monosyllabics are stressed. However, this approach 
-is not well-motivated. Many words -- both monosyllabic and multisyllabic -- are 
-unstressed, especially among prepositions, conjunctions, and particles.
-\rr{multisyllabics that can be unstressed: \rus{через, между}} Furthermore, there are many
-high-frequency monosyllabic homographs that can be either stressed or unstressed, 
-depending on their part of speech, or particular collocations. Clearly, for
-such words, one cannot assume that they are stressed on the basis of their syllable count.
-
-Based on these considerations, we built our tools to mark stress without regard to
-the number of syllables in a word. However, because our gold corpus texts do not
-mark stress on monosyllabic words, we cannot evaluate our annotation of those words.
-
-\section{Automatic Stress Placement}
-
-State-of-the-art morphological analysis in Russian is based on finite-state
-technology.\footnote{add footnote or citations for Dialog, mystem, etc.} To our knowledge,
-no open-source resources are available for analyzing and generating stressed wordforms.
-Therefore, we developed free and open-source finite-state tools capable of analyzing and
-generating stressed wordforms. A finite-state transducer\footnote{Using two-level 
-morphology, xfst/hfst, blah, blah, blah} (FST) generates all possible morphosyntactic readings
-of each wordform, and a constraint grammar\footnote{vislcg3, blah blah} then removes
-some extra readings based on syntactic context. The ultimate success of our stress placement
-system depends on the performance of the Constraint Grammar. Ideally, the Constraint Grammar would
-successfully remove all but the correct reading for each token, but in practice some
-tokens still have more than one reading remaining. Therefore,
-we also evaluate various approaches to deal with the remaining ambiguity,
-as described in the following paragraph and Table~\ref{tab:conditions}.
-
-The {\small {\tt bare}} approach is to not mark stress on words with more than one reading.
-The {\small {\tt safe}} approach is to mark stress only on
-tokens whose morphosyntactic ambiguity is stress-irrelevant. The {\small {\tt guess}}
-approach is to randomly select one of the available readings, and abstain if no readings
-are available. The {\small {\tt guessFreq}} approach is to select the reading that is most frequent.
-\footnote{Frequency data taken from a separate hand-disambiguated corpus.} If none of the readings is found in the corpus, then it selects the reading
-with the tag sequence (lemma removed) that is most frequent in our corpus. If the tag sequence is not found
-in our corpus, then it backs off to the  {\small {\tt guess}} algorithm. Note that for tokens with
-stress-irrelevant ambiguity, {\small {\tt guess}} and {\small {\tt guessFreq}} produce
-the same result as the {\small {\tt safe}} method.
-
-So far, the approaches discussed are
-dependent on the availability of readings from the FST. Two final approaches address unknown tokens.
-The {\small {\tt guessSyll}} and {\small {\tt guessFreqSyll}} approaches are identical to 
-{\small {\tt guess}} and {\small {\tt guessFreq}}, respectively,
-except that for unknown tokens, they places the stress near the end of the stem.\footnote{Determining 
-which vowel is stem-final
-is itself a difficult task. Our algorithm merely places the stress on the last vowel that is 
-followed by a consonant.} \rr{Explain why stem-final is a good guess (and cite someone?)}
-\rr{Discuss the success of the guessSyll approach in the discussion section.}
-
-For our baseline, we take the output of our morphological analyzer (without the Constraint 
-Grammar) in combination with the {\small {\tt bare}}, {\small {\tt safe}}, 
-{\small {\tt guess}} and {\small {\tt guessFreq}} approaches.
-
-We also compare our outcomes with the RussianGram\footnote{http://russiangram.com/} plugin 
-for the Google Chrome web browser. RussianGram is not open-source, so we
-can only guess what technologies support the service. In any case, it provides a
-meaningful reference point for the success of each of the methods described above.
-
-\section{Results}
-
-\begin{table*}[t]
-  \centering
-  \begin{tabular}{r | r r r | r r }
-    approach & correct\% & error\% & abstention\% & totTry\% & totFail\% \\
-    \hline
-    \hline
-    {\small {\tt noCGbare}} & 30.43 & 0.17 & 69.39 & 30.61 & 69.57 \\
-    {\small {\tt noCGsafe}} & 90.07 & 0.49 & 9.44 & 90.56 & 9.93 \\
-    {\small {\tt noCGguess}} & 94.34 & 3.36 & 2.30 & 97.70 & 5.66 \\
-    {\small {\tt noCGguessFreq}} & 95.53 & 2.59 & 1.88 & 98.12 & 4.47 \\
-    {\small {\tt noCGguessSyll}} & 94.99 & 4.05 & 0.96 & 99.04 & 5.01 \\
-    {\small {\tt noCGguessFreqSyll}} & 95.83 & 3.46 & 0.72 & 99.28 & 4.17 \\
-    \hline
-    {\small {\tt CGbare}} & 45.78 & 0.44 & 53.78 & 46.22 & 54.22 \\
-    {\small {\tt CGsafe}} & 93.21 & 0.74 & 6.05 & 93.95 & 6.79 \\
-    {\small {\tt CGguess}} & 95.50 & 2.59 & 1.90 & 98.10 & 4.50 \\
-    {\small {\tt CGguessFreq}} & 95.73 & 2.40 & 1.88 & 98.12 & 4.27 \\
-    {\small {\tt CGguessSyll}} & 95.92 & 3.33 & 0.74 & 99.26 & 4.08 \\
-    {\small {\tt CGguessFreqSyll}} & 96.15 & 3.14 & 0.72 & 99.28 & 3.85 \\
-    \hline
-    {\small {\tt RussianGram}} & 90.09 & 0.79 & 9.12 & 90.88 & 9.91
-  \end{tabular}
-  \caption{Results of stress placement task evaluation. (N = 4048)}
-  \label{tab:results}
-\end{table*}
-
-Single-syllable words and words that are unstressed in the gold standard were
-not evaluated for reasons discussed above, leaving 4048 tokens for evaluation.
-Recall was computed by testing whether stressed 
-multisyllabic wordforms in the gold standard stress corpus were found in the 
-finite-state transducer. Out of 4048 tokens, 3949 were found in the FST, which is equal to 97.55\%.
-This number represents the ceiling for rule-based methods relying on the FST.
-Higher scores are only achievable by expanding the FST or by using syllable-guessing algorithms.
-After running the Constraint Grammar, 3937 tokens were found in the FST, which
-indicates that the Constraint Grammar ruled out 0.20\% of the correct forms.
-By inspecting these errors, we can improve the Constraint Grammar.
-
-Results were compiled for each of the nine conditions discussed above: 
-\emph{without} the Constraint Grammar ({\small {\tt noCG}}) X 4 approaches, 
-\emph{with} the Constraint Grammar ({\small {\tt CG}}) X 4 approaches, 
-and RussianGram ({\small {\tt RG}}). 
-Failures to correctly place stress were split into two categories: errors and abstentions. 
-If the stress tool outputs a stressed wordform, and it is incorrect, then it is
-counted as an `error'. If the stress tool outputs an unstressed wordform, then
-it is counted as an `abstention'. The sum of the errors and abstentions is referred
-to as `totFail'. The sum of correct attempts and `errors' is referred to as `totTry'.
-Results are given in Table~\ref{tab:results}.
-
-WE NEED TO COMPUTE SIGNIFICANCE ON THESE RESULTS
-
-\section{Discussion}
-
-%\section*{Acknowledgments}
-%
-%Do not number the acknowledgment section. Do not include this section
-%when submitting your paper for review.
-
-\bibliographystyle{acl}
-\bibliography{rusaccent}
-
 \end{document}

### 2015-01-15 engkaz-eamt-2015/engkaz-eamt-2015.tex
update
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58327 72bbbca6-d526-0410-a7d9-f06f51895060
update
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58328 72bbbca6-d526-0410-a7d9-f06f51895060
added sections
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58330 72bbbca6-d526-0410-a7d9-f06f51895060
introduction, some parts will be edited later
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58363 72bbbca6-d526-0410-a7d9-f06f51895060

@@
-\documentclass[a4paper,11pt]{article}
+\documentclass[a4paper,twocolumn,11pt]{article}
 
 \usepackage{url}
+
+\newcommand{\com}[1]{\marginpar{\scriptsize #1}}
+
+\title{A free/open-source machine translation system for English to Kazakh}
+
+% Authors:
+% Aida, Mikel, Fran, Inari, ... ?
 
@@
 
+SOMETHING ABOUT ENGLISH-to-KAZAKH trnslation problem??
+
+This paper describes work in progress of development of machine translation system for English-to-Kazakh, which developed by using Apertium free/open-source machine translation platform (Forcada et al.
+2011, http://www.apertium.org).
+
+\section{The Apertium platform}
+Apertium (Forcada et al.
+2011, http://www.apertium.org) is a free/open-source rule-based machine translation (MT) platform that was launched in 2005 by the Universitat d’Alacant.
+Though it was initially aimed at translating between closely related languages, it was later extended to be able to deal with unrelated languages.
+All of the components of the platform (MT engine, developer’s tools, and linguistic data for an increasing number of language pairs) are licensed under the free/open-source GNU General Public License (GPL, versions 2 and 3) and are available to everyone interested in the website.
+FIGURE construction of Apertium
+WILL EDIT THIS
+	Deformatter.
+It separates the text to be translated from the formatting tags.
+Formatting tags are encapsulated as “superblanks” that are placed between words in such a way that the remaining modules see them as regular blanks.
+	Morphological analyser.
+For each surface form (that is, for each lexical unit as it appears in the text), the morphological analyser generates one or more lexical forms composed of: lemma (dictionary or citation form), lexical category (or part-of-speech), and inflection information.
+The morphological analyser executes a finite-state transducer generated by compiling a morphological dictionary for the source language.
+Lexical units containing more than one word (multiword lexical units) are analyzed as a single lexical unit .
+Morphological analyser uses a finite state transducer based on two-level rules (in the case of Kazakh, apertium-kaz.kaz.lexc, apertium-kaz.kaz.twol).
+This module therefore separates lexemes and processes morphological analysis, and then returns possible lexical forms.
+	Part-of-speech (POS) tagger.
+Apertium's POS tagger is based on a statistical model based on hidden Markov models which processes the result of the application of  on constraint-grammar rules (Karlsson 2005), which areused to discard some analyses  using simple rules ( written in apertium-kaz.kaz.rlx) based on context.
+For example, consider the morphological analysis of word қара:
+^қара/қара<adj>/қара<adj><advl>/қара<adj><subst><nom>/қара<v><tv><imp><p2><sg>/қара<adj>+е<cop><p3><pl>/қара<adj>+е<cop><p3><sg>/қара<adj><subst><nom>+е<cop><p3><pl>/қара<adj><subst><nom>+е<cop><p3><sg>
+This word is ambiguous and has 6 meanings.
+Many  surface forms are ambiguous, which means that these words have more than one POS and therefore more than one possible translation.
+After  this module, all words have only one morphological analysis.
+	Lexical transfer.
+This module uses a bilingual dictionary (apertium-eng-kaz.eng-kaz.dix) which has very simple structure [7].
+The module reads each source-language lexical form and finds one or more corresponding target-language lexical forms.
+Multiword units are translated as a single word.
+	Lexical selection.
+It uses rules that select for those lexical words having many translations,  one of the translations in the target language according to context.
+All rules are written in file apertium-eng-kaz.kaz-eng.lrx .
+	Structural transfer.
+This module identifies sequences of lexical forms  (phrases or segments), which need syntactical processing (handling of number, prepositions, etc.) to be translated.
+It uses  files with rules, which specify  the syntactic transformation as a cascaded process.
+Transfer rules, which transform lexical-form sequences into a new sequences for the  target language, perform the work in this module.
+
+Figure-1. The Apertium machine translation pipeline 
+	 Morphological generator.
+From the sequence of target-language lexical forms produced by the structural transfer, it generates a corresponding sequence of target language surface forms.
+The morphological generator executes a finite-state transducer generated by compiling a morphological dictionary for the target language.
+	Post-generator.
+It takes care of some minor orthographical operations in the target language (for instance, it generates the English form cannot from can and not).
+This module is generated from file with rules which are very   similar in format to dictionary files.
+	Reformatter. It places format tags back into the text so that its format is preserved.
+
+\section{Linguistic Data(dictionaries,rules for eng-kaz)}
+
+In Apertium is used XML-based formats for linguistic data, it includes bilingual and monolingual dictionaries, structural transfer, lexical rules and rules for part-of-speech tagging.
+
+Dictionaries
+
+\section{Results}
+
+\section{Evaluation}
+
+\com{task-based evaluation (WER, assim, TM-patching?) -FMT}
+
+\com{Compare with GT + other systems -FMT}
+
+\section*{Acknowledgements}
+
 \end{document}

### 2015-01-16 rusaccent-nodalida-2015/rusaccent.tex
Results finished. Some bibtex updated.
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58374 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-\rr{another relevant issue is whether the ambiguity can be resolved at all. Maybe this is 
-something we should bring up in the discussion section. For example, when is stress marked in 
-authentic by-native-for-native texts? A computer should not be expected to predict the stress on 
-such words (although a computer could possibly identify words that should be marked for 
-natives.)} It should be noted that not all morphosyntactic ambiguity leads to stress ambiguity. For example, all inflecting modifiers have the same form for F-SG-GEN, F-SG-LOC, F-SG-DAT and 
+It should be noted that not all morphosyntactic ambiguity leads to stress ambiguity. For example, all inflecting modifiers have the same form for F-SG-GEN, F-SG-LOC, F-SG-DAT and 
 F-SG-INS. We refer to this as \emph{stress-irrelevant} ambiguity, since all readings have
@@
 Many of the most problematic stress ambiguities in Russian occur in 
-high-frequency wordforms\rr{might it be worth computing the difference in stress
-ambiguity between wordform lists and running text? i.e. what percentage of tokens have ambiguous
+high-frequency wordforms\rr{compute the difference in stress
+ambiguity between wordform lists and running text, i.e. what percentage of tokens have ambiguous
 stress.}, so evaluations based on wordlists - as opposed to running
@@
 
-Based on these considerations, we built our tools to mark stress without regard to
-the number of syllables in a word. However, because our gold corpus texts do not
+Based on these considerations, we built our tools to mark stress on every
+word, both monosyllabic and multisyllabic. However, because our gold corpus texts do not
 mark stress on monosyllabic words, we cannot evaluate our annotation of those words.
@@
 Therefore, we developed free and open-source finite-state tools capable of analyzing and
-generating stressed wordforms. A finite-state transducer\footnote{Using two-level 
+generating stressed wordforms, based on the well-known 
+\emph{Grammatical Dictionary of Russian} \cite{Zaliznjak-77}. A finite-state 
+transducer\footnote{Using two-level 
 morphology, xfst/hfst, blah, blah, blah} (FST) generates all possible morphosyntactic readings
@@
 followed by a consonant.} \rr{Explain why stem-final is a good guess (and cite someone?)}
-\rr{Discuss the success of the guessSyll approach in the discussion section.}
 
@@
   \begin{tabular}{r | r r r | r r }
-    approach & correct\% & error\% & abstention\% & totTry\% & totFail\% \\
+    approach & accuracy\% & error\% & abstention\% & totTry\% & totFail\% \\
     \hline
@@
 not evaluated for reasons discussed above, leaving 4048 tokens for evaluation.
-Recall was computed by testing whether stressed 
-multisyllabic wordforms in the gold standard stress corpus were found in the 
-finite-state transducer. Out of 4048 tokens, 3949 were found in the FST, which is equal to 97.55\%.
-This number represents the ceiling for rule-based methods relying on the FST.
-Higher scores are only achievable by expanding the FST or by using syllable-guessing algorithms.
-After running the Constraint Grammar, 3937 tokens were found in the FST, which
-indicates that the Constraint Grammar ruled out 0.20\% of the correct forms.
-By inspecting these errors, we can improve the Constraint Grammar.
+Since our approach is lexicon-based, some of our results should be interpreted
+with respect to how many of the stressed wordforms in the gold standerd corpus
+can be found in the finite-state transducer. We refer to this measure as 
+\emph{recall}. Out of 4048 tokens, 3949 were 
+found in the FST, which is equal to 97.55\%.
+This number represents the ceiling for methods relying on the FST.
+Higher scores are only achievable by expanding the FST or by using syllable-
+guessing algorithms. After running the Constraint Grammar, recall was 97.35\%, 
+a reduction of 0.20\%.
 
-Results were compiled for each of the nine conditions discussed above: 
-\emph{without} the Constraint Grammar ({\small {\tt noCG}}) X 4 approaches, 
-\emph{with} the Constraint Grammar ({\small {\tt CG}}) X 4 approaches, 
-and RussianGram ({\small {\tt RG}}). 
-Failures to correctly place stress were split into two categories: errors and abstentions. 
+Results were compiled for each of the 13 approaches discussed above: 
+\emph{without} the Constraint Grammar ({\small {\tt noCG}}) x 6 approaches, 
+\emph{with} the Constraint Grammar ({\small {\tt CG}}) x 6 approaches, 
+and RussianGram ({\small {\tt RussianGram}}). 
+Each token was categorised as either Failures to correctly place stress were 
+split into two categories: errors and abstentions. 
 If the stress tool outputs a stressed wordform, and it is incorrect, then it is
@@
 
-WE NEED TO COMPUTE SIGNIFICANCE ON THESE RESULTS
+\rr{Are differences significant?}
+
+The {\small {\tt noCGbare}} approach achieves a baseline accuracy of
+30.43\%, so roughly two thirds of the tokens in our corpus are
+morphosyntactically ambiguous. The error rate of 0.17\% primarily represents 
+forms whose stress position varies from speaker to speaker (e.g.
+\rus{зав\'{и}л\'{и}сь} \emph{zav\'{i}l\'{i}s'}), errors in the gold corpus (e.g. 
+\rus{вер\'{и}м} \emph{ver\'{i}m}) . The {\small {\tt noCGsafe}} approach
+achieves a 60\% improvement in accuracy (90.07\%), which means that 89.39\% of 
+morphosyntactic ambiguity on our corpus is stress-irrelevant. Interestingly, the 
+RussianGram web service achieves results that are very close 
+to the {\small {\tt noCGsafe}} approach, so it would appear that it 
+utilizes a similar approach.
+
+Since the ceiling recall for the FST is 97.55\%, and since the 
+{\small {\tt noCGsafe}} approach achieves 90.07\%, the maximum improvement 
+that a Constraint Grammar could theoretically achieve is 7.48\%. A comparison
+of {\small {\tt noCGsafe}} and {\small {\tt noCGsafe}} reveals an improvement
+of 3.14\%\footnote{one of the authors considers it significant that the CG 
+achieves an improvement roughly equivalent to $\pi$. :)}, which is about 42\%
+of the way to the ceiling.\rr{...while raising the error rate 0.25\%}
+
+The {\small {\tt CGguess}} and {\small {\tt CGguessFreq}} approaches are also 
+limited by the 97.55\% ceiling from the FST, and their accuracies achieve
+improvements of 2.29\% and 2.52\%, respectively, over {\small {\tt CGsafe}}.
+However, these gains come at the cost of error rates as much as 3.5 times 
+higher than {\small {\tt CGsafe}}: +1.85\% and +1.66\%, respectively. It is
+not surprising that {\small {\tt CGguessFreq}} has higher accuracy and a lower
+error rate than {\small {\tt CGguess}}, since frequency-based guesses are by
+definition more likely to occur. The frequency data were taken from a very 
+small corpus, and it is likely that frequency from a larger corpus would yield 
+better results.
+
+The two {\small {\tt Syll}} approaches were designed to make a blind guess on 
+every wordform that is not found in the FST, which would ideally result in an
+abstention rate of 0\%. However, the abstention rates of approximately 0.7\%
+are the result of the fact that some words in the FST, especially proper nouns, 
+had not been assigned stress. Because the FST outputs a form -- albeit unstressed 
+-- the {\small {\tt Syll}} algorithm is not called. This means that 
+{\small {\tt Syll}} is only guessing on about 2\% of the tokens. The improvement
+on overall accuracy from {\small {\tt CGguessFreq}} to 
+{\small {\tt CGguessFreqSyll}} is 0.42\%, which means that the 
+{\small {\tt Syll}} was accurate 21\% of the time.
+
+\begin{itemize}
+\item Examine errors and categorize/refute them.
+\end{itemize}
 
@@
 
+One of the main points of this paper is to highlight the importance of 
+syntactic context in the Russian wordstress placement task. \cite{hall_sproat_russian_2013} 
+achieve 98.7\% accuracy on stress placement of individual wordforms. To 
+understand this result, it should be remembered that Russian has a considerable
+number of words that are ambiguous with regard to stress. Based on the surface 
+forms in our FST (which is based on the same dictionary used for 
+\cite{hall_sproat_russian_2013}), we calculate that only 29~518 (1.05\%) 
+of the 2~804~492 wordforms contained in our FST are stress-ambiguous. \rr{These 
+stress-ambiguous wordforms have an average of 2.01 possible stress positions.} 
+Compare this with the finding of our experiment, that in running 
+text, at least 7.5\% of the tokens are stress-ambiguous. This is because many of 
+the lexemes with shifting stress patterns are high-frequency. Any system that
+cannot leverage syntactic context to disambiguate these wordforms cannot 
+achieve accuracy above 92.5\% without making an higher number of errors.
+Since the task of wordstress placement is virtually always performed on
+running text, it seems prudent to make use of surrounding information.
+
+We have shown that our Russian Constraint Grammar is able to
+
+Limitations
+
+Another relevant issue is whether the ambiguity can be resolved at all. For example, when is stress marked in 
+authentic by-native-for-native texts? A computer should not be expected to predict the stress on 
+such words (although a computer could possibly identify words that should be marked for 
+natives.)
+
+Future work:
+Two sections: known words and unknown words
+\begin{itemize}
+\item New domains and reading levels
+\item As with any lexicon- and rule-based system, those components can be expanded and improved, especially considering the short span over which they were developed.
+\item Integrate a more principled unknown word-guessing method (homicevic)
+\item Compare and/or integrate with stochastic methods. So far stochastic
+methods have only been used for guessing stress on unknown words.
+\item
+
+\end{itemize}
 %\section*{Acknowledgments}

### 2015-01-16 rusaccent-nodalida-2015/rusaccent.tex
More conclusions and some more bibtex.
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58375 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 Many studies of Russian text-to-speech and automatic speech recognition make note of the 
-difficulties caused by the shortcomings of their stress-marking resources (e.g. CITE Krivnova 
-1998), but very few studies have targeted stress-marking itself, and they tend to
+difficulties caused by the shortcomings of their stress-marking resources (e.g. \cite{krivnova_1998}), but very few studies have targeted stress-marking itself, and they tend to
 focus on placing stress on unknown words, with almost no attention to the problems that
@@
 \cite{Xomicevic_2008} developed a set of heuristics for guessing stress placement on 
-unknown words in Russian. More recently, CITE Hall \& Sproat (2013) trained a maximum entropy
+unknown words in Russian. More recently, \cite{hall_sproat_russian_2013} trained a maximum entropy
 model on a dictionary of Russian words, achieving 83.9\% accuracy for stress on unknown words 
@@
 generating stressed wordforms, based on the well-known 
-\emph{Grammatical Dictionary of Russian} \cite{Zaliznjak-77}. A finite-state 
-transducer\footnote{Using two-level 
+\emph{Grammatical Dictionary of Russian} \cite{Zaliznjak-77}. A Finite-State 
+Transducer\footnote{Using two-level 
 morphology, xfst/hfst, blah, blah, blah} (FST) generates all possible morphosyntactic readings
@@
 Compare this with the finding of our experiment, that in running 
-text, at least 7.5\% of the tokens are stress-ambiguous. This is because many of 
+text, at least 7.5\% of the tokens are stress-ambiguous. This is at least partly 
+because many of 
 the lexemes with shifting stress patterns are high-frequency. Any system that
@@
 running text, it seems prudent to make use of surrounding information.
+The experiment described in this paper demonstrates that a Constraint Grammar
+can effectively improve the accuracy of a stress placement system without 
+significantly raising the error rate. Our Russian Constraint Grammar is
+under continual development, so we expect higher accuracy in the future.
 
-We have shown that our Russian Constraint Grammar is able to
+The experiment was limited by properties of the gold standard corpus, including
+its size, genre distribution, and quality. Ideally, a gold standard corpus for Russian wordstress should mark 
+whether monosyllabic words are stressed, and of course its annotations should
+be accurate. Our gold corpus represents a broad variety of text genres, which 
+makes our results more generalizable, but a larger corpus would allow for
+evaluating each genre individually. For example, the vast majority of words with
+shifting stress are of Slavic origins, so we expect a genre such as technical
+writing to have a lower proportion of words with stress ambiguity, since it
+uses a more modern lexicon that has not inherited the idiosyncrasies that lead
+to wordstress ambiguity.
 
-Limitations
+In addition to genre, it seems likely that text
+complexity affects the difficulty of the stress placement task, so subcorpora at
+various reading levels would shed light on those interactions as well.
 
-Another relevant issue is whether the ambiguity can be resolved at all. For example, when is stress marked in 
-authentic by-native-for-native texts? A computer should not be expected to predict the stress on 
-such words (although a computer could possibly identify words that should be marked for 
-natives.)
+The syllable-guessing algorithm used in this experiment was overly simplistic,
+and it was not surprising that it was
 
-Future work:
+Future work/potential applications:
 Two sections: known words and unknown words
 \begin{itemize}
-\item New domains and reading levels
 \item As with any lexicon- and rule-based system, those components can be expanded and improved, especially considering the short span over which they were developed.
@@
 methods have only been used for guessing stress on unknown words.
-\item
+\item Another relevant issue is whether the ambiguity can be resolved at all.
+For example, when is stress marked in authentic by-native-for-native texts?
+A computer should not be expected to predict the stress on such words (although a computer could possibly identify words that should be marked for
+natives.)
 

### 2015-01-17 engkaz-eamt-2015/engkaz-eamt-2015.tex
edited about Apertium
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58381 72bbbca6-d526-0410-a7d9-f06f51895060

--- 2015-01-14 engkaz-eamt-2015/engkaz-eamt-2015.tex
@@
 
+SOMETHING ABOUT ENGLISH-to-KAZAKH trnslation problem??
+
+This paper describes work in progress of development of machine translation system for English-to-Kazakh, which developed by using Apertium free/open-source machine translation platform (Forcada et al.
+2011, http://www.apertium.org).
+
+\section{The Apertium platform}
+Apertium (Forcada et al.
+2011, http://www.apertium.org) is a free/open-source rule-based machine translation (MT) platform that was built in 2005 by the Universitat d’Alacant.
+At first, it was initially aimed to translate texts between closely related languages, then it was extended to deal with unrelated languages.
+This platform has next components: machine translation engine, developer’s tools, and linguistic data for an increasing number of language pairs and they are licensed under the free/open-source GNU General Public License (GPL, versions 2 and 3).
+
+FIGURE construction of Apertium
+WILL EDIT THIS
+	De-formatter.
+Separates the text to be translated from the formatting tags.
+Formatting tags are encapsulated in brackets so they are treated as “superblanks” that are placed between words in such a way that the remaining modules see them as regular blanks.
+	Morphological analyser.
+For each surface form the morphological analyser generates one or more lexical forms, which consist of: lemma (dictionary or citation form), lexical category (or part-of-speech), and inflection information.
+For example, the morphological analyser will deliver next sentence "I played" as showed below:
+^I/I<prn><subj><p1><mf><sg>$ 
+^played/play<vblex><past>/play<vblex><pp>$
+
+In this example each source form has been analysed as one or more lexical forms: "I" is analysed into "I", where lexical category is subject pronoun(<prn><subj>) and first person(p1), could be masculine or feminine(mf), singular(sg); "played" is analysed into lexical verb(vblex) "play" with past simple tense(past) or it could be past participle(pp), so it has two analysis.
+Each analysis of word are separated by "^" and "" symbols, and for one word each lexical forms are delimited by "/", and tags(<...>) show grammatical attributes of lexical form.
+ For Kazakh language finite state transducer based on two-level rules (in the case of Kazakh, apertium-kaz.kaz.lexc, apertium-kaz.kaz.twol).
+This module therefore separates lexemes and processes morphological analysis, and then returns possible lexical forms.
+	Part-of-speech (POS) tagger.
+This module chooses one of lexical forms of ambiguous word.
+As can be seen from the previous example, the morphological analyser could deliver more than one lexical form and choosing wrong one could produce errors in translation.
+POS tagger is based on a statistical model, which based on hidden Markov models and it has been trained on source language texts, which processes the result of the application of  on constraint-grammar rules (Karlsson 2005), which areused to discard some analyses  using simple rules (written in apertium-eng-kaz.eng-kaz.rlx) based on context.
+For example, consider the morphological analysis of word "book":
+^book/book<n><sg>
+/book<vblex><inf>
+/book<vblex><pres>
+This word has 3 lexical forms, and depends on context, ont of them will be chosen by rules or tagger. After  this module, all words have only one morphological analysis.
+	Lexical transfer.
+It reads each lexical form of source language and delivers a corresponding target language lexical form.
+This module uses a bilingual dictionary (apertium-eng-kaz.eng-kaz.dix) which has very simple structure(1).
+For English-Kazakh language pair, each lexical form of English word is translated into Kazakh as follows:
+^I<prn><subj><p1><mf><sg> /мен<prn><pers><subj><p1><mf><sg>$ 
+^play<vblex><past>/ойна<v><tv><past>$
+ Multiword units are translated as a single word.
+	Lexical selection.
+It uses rules for lexical words, which have many translations, to select one of the translations in the target language according to context.
+All rules are written in file apertium-eng-kaz.eng-kaz.lrx.
+	Structural transfer.
+This module uses pattern matching to identify sequences of lexical forms  (phrases or segments), which need syntactical processing to translate grammatical differences between two languages (handling of number, gender, etc.).
+It uses  files with rules, which specify  the syntactic transformation such as word reorderings, lexical changes such as changes in prepositions and agreement between target language lexical forms.
+Transfer rules can produce new sequences for the  target language, for instance, preposition-noun rule is used to built sequence, where for noun is chosen right case, which depends on preposition:
+in garden  -  ^бақша<n><sg><PXD><loc
+
+	 Morphological generator.
+Delivers the sequence of target-language lexical forms, produced by the structural transfer, to a corresponding sequence of target language surface forms.
+The morphological generator executes a finite-state transducer generated by compiling a morphological dictionary for the target language.
+	Post-generator.
+It performs some minor orthographical operations in the target language (for instance, it generates the English form cannot from can and not).
+This module is generated from file with rules which are very   similar in format to dictionary files.
+	Reformatter. It places format tags back into the text so that its format is preserved.
+
+\section{Linguistic Data(dictionaries,rules for eng-kaz)}
+
+In Apertium is used XML-based formats for linguistic data, it includes bilingual and monolingual dictionaries, structural transfer, lexical rules and rules for part-of-speech tagging.
+
+Dictionaries
+
+\section{Results}
+
+\section{Evaluation}
+
+\com{task-based evaluation (WER, assim, TM-patching?) -FMT}
+
+\com{Compare with GT + other systems -FMT}
+
+\section*{Acknowledgements}
+
+\section{References}
+1 Сундетова А.М., Кәрібаева А.С., АПЕРТИУМ ПЛАТФОРМАСЫНДАҒЫ АҒЫЛШЫН-ҚАЗАҚ МАШИНАЛЫҚ АУДАРМАШЫ ҮШІН ЕКІТІЛДІ СӨЗДІКТІ ҚҰРУ.
+Матер.
+межд.
+научн.-практ.
+конф.
+«Применение информационно-коммуникационных технологий в образовании и науке», посвященной 50-летию Департамента информационно-коммуникационных технологий и 40-летию кафедры «Информационные системы» КазНУ им.
+аль-Фараби.
+22 ноября 2013г.
+– Алматы: Қазақ Университеті, 2013.
+– С.53-57.
+
 \end{document}

### 2015-01-19 rusaccent-nodalida-2015/rusaccent.tex
more
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58450 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{linguex}
+\usepackage{needspace}
 
@@
 \begin{abstract}
-  We evaluate the effectiveness of tools we developed for automatically annotating 
-  word stress in Russian running text. Given an adequate lexicon with specified stress, the
-  primary obstacle for correct stress placement is disambiguating homographic forms. Using a
-  Constraint Grammar, we achieve a X\% improvement over the baseline undisambiguated performance.
+  We evaluate the effectiveness of finite-state tools we developed for 
+  automatically annotating word stress in Russian running text. To our knowledge,
+  this is the first study to empirically evaluate the results of this task. 
+  Given an adequate 
+  lexicon with specified stress, the primary obstacle for correct stress 
+  placement is disambiguating homographic wordforms. Using a Constraint Grammar, 
+  we achieve 93.21\% accuracy, a 3.14\% improvement over the baseline
+  undisambiguated performance. 
+  Combining our Constraint Grammar with frequency data and a simple
+  algorithm for unknown word stress position guessing, we achieve 96.15\%
+  accuracy. These results highlight the need morphosyntactic disambiguation
+  in the wordstress placement task for Russian, and set a standard for future
+  research on this task.
 \end{abstract}
@@
 
-Russian word stress is variable. The inflecting word classes include complex patterns of
-shifting stress, and a lexeme's stress pattern cannot be predicted from surface forms.
-Although lexical stress and its attendant vowel reduction are a prominent feature of
-spoken Russian, standard written Russian does not typically mark word stress.\footnote{
-Texts intended for native speakers sometimes mark stress on words that cannot be 
-disambiguated through context, although this is somewhat rare.} Without information about lexical 
-stress position, 
-correctly converting written Russian text to speech is impossible. This is true both for humans
-(e.g. foreign language students) and computers (e.g. text-to-speech).
+Russian word stress is variable and phonemic. The inflecting word classes include complex 
+patterns of shifting stress, and a lexeme's stress pattern cannot be predicted 
+from surface forms. Lexical stress and its attendant vowel reduction are 
+a prominent feature of spoken Russian, and the incorrect placement of stress
+can render speech incomprehensible. However, standard written Russian does not 
+typically mark word stress.\footnote{Texts intended for native speakers sometimes 
+mark stress on words that cannot be disambiguated through context, although this 
+is rare.} Without information about lexical stress position, 
+correctly converting written Russian text to speech is impossible. This is true 
+both for humans (e.g. foreign language students) and computers (e.g. text-to-
+speech).
 
-We identify three different types of word stress ambiguity. \emph{Intraparadigmatic}
-stress ambiguity refers
-to homographic wordforms belonging to the same lexeme, as shown in \ref{ex:intrahom}.
+We identify three different types of word stress ambiguity.
+\emph{Intraparadigmatic} stress ambiguity refers
+to homographic wordforms belonging to the same lexeme, as shown in 
+\ref{ex:intrahom}.
 
+\needspace{4\baselineskip}
 \ex. Intraparadigmatic homographs \label{ex:intrahom}
-\a. \rus{т\'{е}ла} \emph{t\'{e}la} `body.\textsubscript{SG-GEN}' \label{ex:bodySGGEN}
-\b. \rus{тел\'{а}} \emph{tel\'{a}} `body.\textsubscript{PL-NOM}' \label{ex:bodyPLNOM}
+\a. \rus{т\'{е}ла} \emph{t\'{e}la} `body.\textsubscript{SG-GEN}' 
+    \label{ex:bodySGGEN}
+\b. \rus{тел\'{а}} \emph{tel\'{a}} `body.\textsubscript{PL-NOM}' 
+    \label{ex:bodyPLNOM}
 
-The remaining two types of stress ambiguity occur between lexemes. \emph{Morphosyntactically
-congruent} stress ambiguity occurs between homographs that belong to separate lexemes, 
-and whose morphosyntactic values are identical, as shown in Table~\ref{table:MScongruent}.
+The remaining two types of stress ambiguity occur between lexemes. 
+\emph{Morphosyntactically congruent} stress ambiguity occurs between homographs 
+that belong to separate lexemes, 
+and whose morphosyntactic values are identical, as shown in
+Table~\ref{table:MScongruent}. This kind of stress ambiguity is relatively rare.
+Resolving this ambiguity would require the use of  Word Sense Disambiguation, which is
+outside the scope of this paper.
 
@@
 
-It should be noted that not all morphosyntactic ambiguity leads to stress ambiguity. For example, all inflecting modifiers have the same form for F-SG-GEN, F-SG-LOC, F-SG-DAT and 
-F-SG-INS. We refer to this as \emph{stress-irrelevant} ambiguity, since all readings have
-the same stress placement. Stress-irrelevant ambiguity can be intraparadigmatic and/or interlexical.
+It should be noted that not all morphosyntactic ambiguity leads to stress 
+ambiguity. For example, all inflecting modifiers have the same form for F-SG-GEN, 
+F-SG-LOC, F-SG-DAT and F-SG-INS. We refer to this as \emph{stress-irrelevant} 
+ambiguity, since all readings have
+the same stress placement. Stress-irrelevant ambiguity can be intraparadigmatic 
+and/or interlexical.
 
-Different types of stress ambiguity are overcome by different information. Detailed 
-part-of-speech tagging can determine the stress of intraparadigmatic homographs and
-morphosyntactically incongruent homographs, whereas word sense disambiguation can 
-determine the stress position of morphosyntactically congruent homographs. In the case of
-running text in Russian, most stress placement ambiguity is rooted in morphosyntactic
-ambiguity, so in this study we focus on the role of part-of-speech tagging in improving
+Different types of stress ambiguity are overcome by different information. 
+Detailed part-of-speech tagging can determine the stress of intraparadigmatic 
+homographs and morphosyntactically incongruent homographs, whereas word sense 
+disambiguation can determine the stress position of morphosyntactically congruent 
+homographs. In the case of running text in Russian, most stress placement 
+ambiguity is rooted in morphosyntactic ambiguity, so in this study we focus on 
+the role of part-of-speech tagging in improving
 automatic stress placement, leaving word sense disambiguation to future work.
@@
 Automatic stress placement in Russian is similar to Diacritic Restoration, a task
-which has received increasing interest over the last 20 years. Generally speaking, diacritics
-disambiguate otherwise homographic wordforms, so missing diacritics can
-complicate many NLP tasks, such as text-to-speech. For example, speakers 
-of Czech may type emails and other communications
-without standard diacritics. In order to generate speech from these texts, they must first be 
-normalized by restoring diacritics. A slightly different
-situation arises with Arabic and Hebrew, since standard orthography lacks vowel diacritics
-altogether. For such languages, the `restoration' of vowel diacritics results in less
-ambiguity than in standard orthography. For languages with inherently ambiguous orthography,
-it may be more precise to refer to this as `diacritic enhancement', since it produces
-text that is less ambiguous than the standard language. In this sense, Russian orthography
-is similar to Arabic and Hebrew, since its vowel qualities are underspecified in
-standard orthography.
+which has received increasing interest over the last 20 years. Generally 
+speaking, diacritics disambiguate otherwise homographic wordforms, so missing 
+diacritics can complicate many NLP tasks, such as text-to-speech. For example, 
+speakers of Czech may type emails and other communications
+without standard diacritics. In order to generate speech from these texts, they 
+must first be normalized by restoring diacritics. A slightly different
+situation arises with Arabic and Hebrew, since standard orthography lacks vowel 
+diacritics altogether. For such languages, the `restoration' of vowel diacritics 
+results in less ambiguity than in standard orthography. For languages with 
+inherently ambiguous orthography,
+it may be more precise to refer to this as `diacritic enhancement', since it 
+produces text that is less ambiguous than the standard language. In this sense, 
+Russian orthography is similar to Arabic and Hebrew, since its vowel qualities 
+are underspecified in standard orthography.
 
@@
 
-So far, the implicit target application the few studies related to automatic stress placement
+So far, the implicit target application of the few studies related to automatic stress placement
 in Russian has been text-to-speech and automatic speech recognition. However, the target 
@@
 So far, the approaches discussed are
-dependent on the availability of readings from the FST. Two final approaches address unknown tokens.
-The {\small {\tt guessSyll}} and {\small {\tt guessFreqSyll}} approaches are identical to 
-{\small {\tt guess}} and {\small {\tt guessFreq}}, respectively,
-except that for unknown tokens, they places the stress near the end of the stem.\footnote{Determining 
+dependent on the availability of readings from the FST. Two final approaches 
+address unknown tokens.
+The {\small {\tt guessSyll}} and {\small {\tt guessFreqSyll}} approaches are 
+identical to {\small {\tt guess}} and {\small {\tt guessFreq}}, respectively,
+except that the {\small {\tt Syll}} approaches attempt to place stress on
+unknown tokens (i.e. tokens not found in the FST). Empirical research with
+Russian native speakers reading nonce words in running text concludes that
+the default stress position in Russian is the last syllable of the stem, or
+in other words, the last vowel that is not part of an inflectional ending 
+\cite{Crosswhite.ea-03}. Therefore, the {\small {\tt Syll}} approaches 
+place the stress near the end of the stem. Determining 
 which vowel is stem-final
-is itself a difficult task. Our algorithm merely places the stress on the last vowel that is 
-followed by a consonant.} \rr{Explain why stem-final is a good guess (and cite someone?)}
+is itself a difficult task. Our algorithm merely places the stress on the last 
+vowel that is followed by a consonant. Although more sophisticated algorithms
+are necessary to accurately locate the stem boundary, this is not the central
+focus of this study.
 
-For our baseline, we take the output of our morphological analyzer (without the Constraint 
-Grammar) in combination with the {\small {\tt bare}}, {\small {\tt safe}}, 
-{\small {\tt guess}} and {\small {\tt guessFreq}} approaches.
-
-We also compare our outcomes with the RussianGram\footnote{http://russiangram.com/} plugin 
+For our baseline, we take the output of our morphological analyzer (without the 
+Constraint Grammar) in combination with the {\small {\tt bare}}, 
+{\small {\tt safe}}, {\small {\tt guess}} and {\small {\tt guessFreq}} 
+approaches. We also compare our outcomes with the 
+RussianGram\footnote{http://russiangram.com/} plugin 
 for the Google Chrome web browser. RussianGram is not open-source, so we
 can only guess what technologies support the service. In any case, it provides a
-meaningful reference point for the success of each of the methods described above.
+meaningful reference point for the success of each of the methods described 
+above.
 
@@
 counted as an `error'. If the stress tool outputs an unstressed wordform, then
-it is counted as an `abstention'. The sum of the errors and abstentions is referred
-to as `totFail'. The sum of correct attempts and `errors' is referred to as `totTry'.
-Results are given in Table~\ref{tab:results}.
+it is counted as an `abstention'. The sum of the errors and abstentions is 
+referred to as `totFail'. The sum of correct attempts and `errors' is referred to 
+as `totTry'. Results are given in Table~\ref{tab:results}.
 
@@
 
-\begin{itemize}
-\item Examine errors and categorize/refute them.
-\end{itemize}
+\rr{Examine errors and categorize/refute them.}
 
@@
 One of the main points of this paper is to highlight the importance of 
-syntactic context in the Russian wordstress placement task. \cite{hall_sproat_russian_2013} 
-achieve 98.7\% accuracy on stress placement of individual wordforms. To 
+syntactic context in the Russian wordstress placement task. 
+\newcite{hall_sproat_russian_2013} achieve 98.7\% accuracy on stress placement of 
+individual wordforms. To 
 understand this result, it should be remembered that Russian has a considerable
@@
 forms in our FST (which is based on the same dictionary used for 
-\cite{hall_sproat_russian_2013}), we calculate that only 29~518 (1.05\%) 
+\cite{hall_sproat_russian_2013}\rr{need to cite this without parentheses}), we 
+calculate that only 29~518 (1.05\%) 
 of the 2~804~492 wordforms contained in our FST are stress-ambiguous. \rr{These 
@@
 
+We are unaware of any other studies of Russian wordstress placement in running
+text. The results of our experiment are promising, but many questions remain
+unanswered.
 The experiment was limited by properties of the gold standard corpus, including
-its size, genre distribution, and quality. Ideally, a gold standard corpus for Russian wordstress should mark 
+its size, genre distribution, and quality. Ideally, a gold standard corpus for 
+Russian wordstress should mark 
 whether monosyllabic words are stressed, and of course its annotations should
@@
 
-In addition to genre, it seems likely that text
-complexity affects the difficulty of the stress placement task, so subcorpora at
-various reading levels would shed light on those interactions as well.
+In addition to genre, it is also likely that text
+complexity affects the difficulty of the stress placement task. The distribution
+of different kinds of syntactic constructions vary with text complexity 
+\cite{Vajjala.Meurers-12}, and
+we expect that the effectiveness of the Constraint Grammar will be affected by
+those differences. used in would shed light on those interactions as well.
 
-The syllable-guessing algorithm used in this experiment was overly simplistic,
-and it was not surprising that it was
+The syllable-guessing algorithm {\small {\tt Syll}} used in this experiment was 
+overly simplistic, and it was not surprising that it was only moderately 
+successful. More rigorous approaches have been suggested in other studies.
+One approach -- exemplified by \newcite{Church-85}, 
+\newcite{Williams-87} and \newcite{Xomicevic_2008} -- is to 
+attempt to parse the unknown token by matching known prefixes and suffixes.
+Another approach -- exemplified by Pearson.ea-00, Webster-04, Dou.ea-09, and 
+\newcite{hall_sproat_russian_2013} -- is based on machine-learning. For unknown 
+words, \newcite{hall_sproat_russian_2013} achieve an accuracy of 83.9\%.
+It should be noted that their model was trained on a full list of Russian words,
+which is not representative of the words that would be unknown to a system like
+ours. Most of the complicated wordstress patterns are a closed 
+class.\footnote{The growing number of masculine nouns with shifting stress are 
+the exception to this generalization.} Future machine-learning work on Russian 
+wordstress placement in running text should limit the training data to word
+classes that are likely to be similar to unknown tokens, such as those with
+productive derivational affixes.
 

### 2015-01-19 finnormal-nodalida-2015/finnormal2015.tex
add finnormal stuffs
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58451 72bbbca6-d526-0410-a7d9-f06f51895060

%
% File nodalida2015.tex
%
% Contact beata.megyesi@lingfil.uu.se
%
% Based on the instruction file for EACL 2014
% which in turn was based on the instruction files for previous 
% ACL and EACL conferences.

\documentclass[11pt]{article}
\usepackage{nodalida2015}
%\usepackage{times}
\usepackage{mathptmx}
%\usepackage{txfonts}
\usepackage{url}
\usepackage{latexsym}
\special{papersize=210mm,297mm} % to avoid having to use "-t a4" with dvips 
%\setlength\titlebox{6.5cm}  % You can expand the title box if you really have to

\title{Automatic conversion of colloquial Finnish to standard Finnish}

\author{Francis M. Tyers\\
  UiT Norgga \'{a}rktala\v{s} universitehta, \\
  N-9015 Norway \\
  {\tt francis.tyers@uit.no} \\\And
  Inari Listenmaa\\
  Chalmers tekniska h\"{o}gskola, \\
  Sweden \\
  {\tt inari@cse.chalmers.se} \\}

\date{}

\begin{document}
\maketitle
\begin{abstract}
  This paper presents an unsupervised method for converting between colloquial Finnish
  and standard Finnish. The method relies upon a small number of orthographical rules
  combined with a large language model of standard Finnish for ranking the possible 
  conversions. Aside from this contribution, the paper also presents an evaluation
  corpus consisting of aligned sentences in colloquial Finnish, orthographically-standardised
  colloquial Finnish and standard Finnish. The methods we present outperforms the baseline
  of simply treating colloquial Finnish as standard Finnish and offers promise for the adaptation
  of language-technology tools created for standard Finnish to colloquial Finnish.
\end{abstract}

\section{Introduction}

Most language technology tools are designed or trained based on standard language 
forms, where they exist. The application of these tools to non-standard
language can cause a substantial decrease in quality. For example in machine translation,
or part-of-speech tagging. CITEXXX

For language-technology researchers working on non-standard forms of language, there are 
two clear options: either create new tools to process non-standard text,
  or create tools to preprocess non-standard text, standardising it to be subsequently processed
 by existing tools.

This paper presents a method for converting colloquial Finnish to standard Finnish and 
a parallel corpus for evaluation.

\section{Related work}

There are a number of areas of research related to the task of text normalisation. Text
proofing tools, such as spelling and grammar checkers CITEXXX can be used to encourage adherence
to particular orthographic or grammatical norms. Accent and diacritic restoration CITEXXX is similar
in that it aims to bring text closer to standard orthography in order to facilitate treatment by 
automatic tools. Another related area is machine translation between different written norms of 
the same language. For example between Norwegian Bokmål and Norwegian Nynorsk CITEXXX.

\newcite{scannell2014} presents a method for normalising pre-standardised text in Irish to 
the modern standard. The method relies on

\subsection{Colloquial Finnish}

 \emph{suomen puhekieli}

\section{Corpus}

\begin{table*}
  \centering
  \begin{tabular}{|l|l|l|}
  \hline
  \textbf{Colloquial} & \textbf{Normalised} & \textbf{Standardised} \\
  \hline
  mut siis, oli siell\"{a} jotain vaikka     & mutta siis, oli siell\"{a} jotain vaikka     & mutta oli siell\"{a} jotain     \\
  puheteknologiamatskuja.  & puheteknologiamateriaaleja. & puheteknologiamateriaaleja.  \\
  \hline
  oon tosi hy\"{o}dyllinen assari     & ole todella hy\"{o}dyllinen assistentti & ole todella hy\"{o}dyllinen assistentti \\
  t\"{a}n vikan minuutin.           & t\"{a}m\"{a}n viimeisen minuutin.            & t\"{a}m\"{a}n viimeisen minuutin. \\
  \hline
  \ldots & \ldots & \ldots \\
  \hline
  \end{tabular}
  \caption{Example sentences from the parallel corpus.}
  \label{table:corpexample}
\end{table*}

Our evaluation corpus was created by manually translating texts in colloquial Finnish
to standard Finnish.\footnote{The corpus is freely available and published under the 
CreativeCommons {\sc cc-by-sa} 3.0 licence.} The texts were extracts from internet 
relay chat (IRC) conversations. We performed the conversion process in two steps, the first
step involved simple orthographic normalisation, for 
example \emph{oon} $\rightarrow$ \emph{olen} `I am'. Syntactic and stylistic conversions 
were not applied at this stage. The second conversion step normalised the text
both orthographically and syntactically/stylistically. Table~\ref{table:corpexample} presents
an excerpt from each of the three parts of the corpus.

The corpus was split into three parts, testing, development and training. The testing
and development portions contain 1,000 words each, with the remaining X,000 for training
supervised models.\footnote{The corpus is split into XX files of 500 words each. Files 01--02 
  were used for development; 03--04 for testing and 05--XX for training.}

\section{Methodology}

\subsection{Unsupervised normalisation}

For the unsupervised normalisation we applied a set of regular-expression based 
replace rules to the input text to produce all the possible candidate sentences 
in standard Finnish and then used a target-language model to rank the possible 
candidates. The candidate with the highest rank was selected as the normalised sentence.
For the target-language model we used the Finnish side of the English--Finnish EuroParl
parallel corpus \cite{koehn2005}.

We developed two sets of rules for the unsupervised normalisation approach:

\begin{itemize}
  \item 273 rules from \newcite{karlsson2008}'s grammar of Finnish (\S95--97)
  \item Rules from the development corpus.
\end{itemize}

%Rules: mä -> minä, oon -> olen, ?*nkaa -> ?*n kanssa
%
%Input: Mä oon Inarinkaa
%Step 1: Mä oon Inarinkaa
%        Minä oon Inarinkaa
%Step 2: Mä oon Inarinkaa
%        Minä oon Inarinkaa
%        Mä olen Inarinkaa
%        Minä olen Inarinkaa
%Step 3: Mä oon Inarinkaa
%        Minä oon Inarinkaa
%        Mä olen Inarinkaa
%        Minä olen Inarinkaa
%        Mä oon Inarin kanssa
%        Minä oon Inarin kanssa
%        Mä olen Inarin kanssa
%        Minä olen Inarin kanssa
%Rank:
%	-13.4885 ← minä olen Inarin kanssa
%	-12.7176 ← minä oon Inarin kanssa
%	-12.4408 ← mä olen Inarin kanssa
%	-9.2045 ← mä oon Inarin kanssa
%	-8.8651 ← minä olen Inarinkaa
%	-8.0941 ← minä oon Inarinkaa
%	-7.8174 ← mä olen Inarinkaa
%	-4.5811 ← mä oon Inarinkaa
%Output: Minä olen Inarin kanssa

\subsection{Statistical machine translation}

The statistical-machine translation approaches were implemented using the 
Moses toolkit \cite{koehn2007}. The training set up was that used for
the baseline system in the  in the WMT shared tasks
on machine translation.\footnote{\url{http://www.statmt.org/wmt11/baseline.html}}

The target-language model used was the same as in the unsupervised experiments.

For both approaches we trained two systems, the first used the \emph{normalised}
part of the corpus as the target language; the second used the \emph{standardised} 
part of the corpus as the target language.

\subsubsection{Phrase-based}

\subsubsection{Character-based}

\section{Results}

\begin{table}
  \centering
  \begin{tabular}{|l|r|r|r|}
     \hline
    \textbf{System} & \textbf{PER} & \textbf{WER} & \textbf{BLEU} \\
     \hline
     Baseline & 46.12 & 48.04 & 26.31 \\
     \hline
     rules-1 & 38.27 & 41.19 & 32.65 \\
     rules-2 & - & - & - \\
     \hline
     CBMT-cn & - & - & - \\
     CBMT-cs & - & - & - \\
     PBMT-cn & - & - & - \\
     PBMT-cs & - & - & - \\
     \hline
  \end{tabular}
  \label{table:results}
  \caption{Results...}
\end{table}

\section{Future work}

% weight rules 
% use system in pipeline (e.g. translation)
%

\section{Conclusions}

We have presented a parallel corpus of colloquial Finnish and standard Finnish --
to our knowledge the first of its kind -- and a method for converting colloquial
Finnish to standard Finnish.

%The main contribution of this paper is a parallel corpus of colloquial Finnish
%and standard Finnish

\section*{Acknowledgments}

Joonas Kyl\"{a}m\"{a}

% If you use BibTeX with a bib file named eacl2014.bib, 
% you should add the following two lines:
\bibliographystyle{acl}
\bibliography{finnormal2015.bib}

\end{document}

### 2015-01-19 rusaccent-nodalida-2015/rusaccent.tex
Changes from Laura's comments
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58457 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 a prominent feature of spoken Russian, and the incorrect placement of stress
-can render speech incomprehensible. However, standard written Russian does not 
+can render speech almost incomprehensible. However, standard written Russian does not 
 typically mark word stress.\footnote{Texts intended for native speakers sometimes 
@@
 Table~\ref{table:MScongruent}. This kind of stress ambiguity is relatively rare.
-Resolving this ambiguity would require the use of  Word Sense Disambiguation, which is
-outside the scope of this paper.
+Resolving this ambiguity would require the use of technologies such as Word Sense 
+Disambiguation.
 
@@
 \emph{Morphosyntactically incongruent} stress ambiguity occurs between homographs
-that belong to separate lexemes, and whose morphosyntactic values are different, as shown in 
-\ref{ex:MSincongruent}.
+that belong to separate lexemes, and whose morphosyntactic values are different, 
+as shown in \ref{ex:MSincongruent}.
 
@@
 
-Different types of stress ambiguity are overcome by different information. 
-Detailed part-of-speech tagging can determine the stress of intraparadigmatic 
-homographs and morphosyntactically incongruent homographs, whereas word sense 
-disambiguation can determine the stress position of morphosyntactically congruent 
-homographs. In the case of running text in Russian, most stress placement 
-ambiguity is rooted in morphosyntactic ambiguity, so in this study we focus on 
+In the case of running 
+text in Russian, most stress placement 
+ambiguity is rooted in intraparadigmatic and morphosyntactically incongruent 
+ambiguity. Detailed part-of-speech tagging can help determine the stress of 
+these forms, since each alternative stress placement is tied to a different
+tag sequence. In this study we focus on 
 the role of part-of-speech tagging in improving
-automatic stress placement, leaving word sense disambiguation to future work.
+automatic stress placement, leaving morphosyntactically congruent stress 
+ambiguity to future work. \rr{this paragraph is choppy.}
 
@@
 \cite{Xomicevic_2008} developed a set of heuristics for guessing stress placement on 
-unknown words in Russian. More recently, \cite{hall_sproat_russian_2013} trained a maximum entropy
-model on a dictionary of Russian words, achieving 83.9\% accuracy for stress on unknown words 
-and 98.7\% on known words. One major shortcoming of evaluating on wordlists is 
+unknown words in Russian. More recently, \cite{hall_sproat_russian_2013} trained 
+a maximum entropy model on a dictionary of Russian words, and evaluated on
+wordlists containing `known' and `unknown' wordforms.\footnote{ADD FOOTNOTE
+EXPLAINING THEIR DEFINITION OF KNOWN AND UNKNOWN.} Their model 
+achieved 98.7\% accuracy on known words, and 83.9\% accuracy on unknown words. 
+One major shortcoming of evaluating on wordlists is 
 that they do not reflect wordform frequency.
 Many of the most problematic stress ambiguities in Russian occur in 
-high-frequency wordforms\rr{compute the difference in stress
-ambiguity between wordform lists and running text, i.e. what percentage of tokens have ambiguous
-stress.}, so evaluations based on wordlists - as opposed to running
-text - have inflated results. Furthermore, working with running text allows for the
+high-frequency wordforms, so evaluations based on wordlists - as opposed to 
+running text - have inflated results (see discussion in 
+Section~\ref{discussion}). Furthermore, working with running text allows for the
 possibility of disambiguating homographs based on syntactic context.
 
-So far, the implicit target application of the few studies related to automatic stress placement
-in Russian has been text-to-speech and automatic speech recognition. However, the target 
-application of our stress
-annotator is a different domain: language learning. Since standard Russian does not contain 
-stress-markings, learners are frequently unable to pronounce unknown words correctly without
-referencing a dictionary or similar resources. In the context
-of language learning, marking stress incorrectly is arguably worse than not marking it
-at all. Because of this, we want our stress annotator to be able to abstain from marking
-stress on words that it is unable to resolve with high confidence.
+So far, the implicit target application of the few studies related to automatic 
+stress placement in Russian has been text-to-speech and automatic speech 
+recognition. However, the target application of our stress
+annotator is a different domain: language learning. Since standard Russian does 
+not contain stress-markings, learners are frequently unable to pronounce unknown 
+words correctly without referencing a dictionary or similar resources. In the 
+context of language learning, marking stress incorrectly is arguably worse than 
+not marking it at all. Because of this, we want our stress annotator to be able 
+to abstain from marking stress on words that it is unable to resolve with high 
+confidence.
 
@@
     {\small {\tt safe}} & \rus{кости} & \rus{к\'{о}сти} \\
-    {\small {\tt guess}} & \rus{к\'{о}сти} or \rus{кост\'{и}} & \rus{к\'{о}сти}\\
+    {\small {\tt guess}} & \rus{к\'{о}сти} ($p$=0.67) or \rus{кост\'{и}} ($p$=0.33) & \rus{к\'{о}сти}\\
     {\small {\tt guessFreq}} & \rus{к\'{о}сти} & \rus{к\'{о}сти}
   \end{tabular}
-  \caption{Example output of each stress placement approach, given a particular set of readings}
+  \caption{Example output of each stress placement approach, given a particular set of readings for the token \rus{кости} \emph{kosti}.}
   \label{tab:conditions}
@@
 Russian texts with marked word stress are relatively rare, except in materials
-for language learners, which are predominantly proprietary. Our corpus was collected from
-free texts on Russian language-learning websites.\footnote{For a list of URL's, please 
-contact the authors.} This small corpus (N=7689) is representative of texts 
+for language learners, which are predominantly proprietary. Our corpus was 
+collected from
+free texts on Russian language-learning websites.\footnote{URLs can be found in
+the TROLLing data repository (ADD URL).} This small corpus (N=7689) is 
+representative of texts 
 that learners of Russian are likely to encounter in their studies. These texts 
@@
 unstressed, especially among prepositions, conjunctions, and particles.
-\rr{multisyllabics that can be unstressed: \rus{через, между}} Furthermore, there are many
+\rr{multisyllabics that can be unstressed: \rus{через, между}} Furthermore, there 
+are many
 high-frequency monosyllabic homographs that can be either stressed or unstressed, 
 depending on their part of speech, or particular collocations. Clearly, for
-such words, one cannot assume that they are stressed on the basis of their syllable count.
+such words, one cannot assume that they are stressed on the basis of their 
+syllable count.
 
 Based on these considerations, we built our tools to mark stress on every
-word, both monosyllabic and multisyllabic. However, because our gold corpus texts do not
-mark stress on monosyllabic words, we cannot evaluate our annotation of those words.
+word, both monosyllabic and multisyllabic. However, because our gold corpus texts 
+do not mark stress on monosyllabic words, we cannot evaluate our annotation of 
+those words.
 
@@
 State-of-the-art morphological analysis in Russian is based on finite-state
-technology.\footnote{add footnote or citations for Dialog, mystem, etc.} To our knowledge,
-no open-source resources are available for analyzing and generating stressed wordforms.
-Therefore, we developed free and open-source finite-state tools capable of analyzing and
-generating stressed wordforms, based on the well-known 
-\emph{Grammatical Dictionary of Russian} \cite{Zaliznjak-77}. A Finite-State 
-Transducer\footnote{Using two-level 
-morphology, xfst/hfst, blah, blah, blah} (FST) generates all possible morphosyntactic readings
-of each wordform, and a constraint grammar\footnote{vislcg3, blah blah} then removes
-some extra readings based on syntactic context. The ultimate success of our stress placement
-system depends on the performance of the Constraint Grammar. Ideally, the Constraint Grammar would
-successfully remove all but the correct reading for each token, but in practice some
+technology.\footnote{add footnote or citations for Dialog, mystem, etc.} To our 
+knowledge, no open-source resources are available for analyzing and generating 
+stressed wordforms. Therefore, we developed free and open-source finite-state 
+tools capable of analyzing and generating stressed wordforms, based on the
+well-known \emph{Grammatical Dictionary of Russian} \cite{Zaliznjak-77}.
+A Finite-State Transducer\footnote{Using two-level 
+morphology, xfst/hfst, blah, blah, blah} (FST) generates all possible 
+morphosyntactic readings of each wordform, and a constraint grammar\footnote{vislcg3, blah blah} then removes
+some extra readings based on syntactic context. The ultimate success of our 
+stress placement system depends on the performance of the Constraint Grammar. 
+Ideally, the Constraint Grammar would successfully remove all but the correct 
+reading for each token, but in practice some
 tokens still have more than one reading remaining. Therefore,
@@
 
-The {\small {\tt bare}} approach is to not mark stress on words with more than one reading.
+The {\small {\tt bare}} approach is to not mark stress on words with more than 
+one reading. Since both sets of readings in Table~\ref{tab:conditions} have more
+than one reading, {\small {\tt bare}} does not output a stressed form.
+
 The {\small {\tt safe}} approach is to mark stress only on
-tokens whose morphosyntactic ambiguity is stress-irrelevant. The {\small {\tt guess}}
-approach is to randomly select one of the available readings, and abstain if no readings
-are available. The {\small {\tt guessFreq}} approach is to select the reading that is most frequent.
-\footnote{Frequency data taken from a separate hand-disambiguated corpus.} If none of the readings is found in the corpus, then it selects the reading
-with the tag sequence (lemma removed) that is most frequent in our corpus. If the tag sequence is not found
-in our corpus, then it backs off to the  {\small {\tt guess}} algorithm. Note that for tokens with
-stress-irrelevant ambiguity, {\small {\tt guess}} and {\small {\tt guessFreq}} produce
-the same result as the {\small {\tt safe}} method.
+tokens whose morphosyntactic ambiguity is stress-irrelevant. In 
+Table~\ref{tab:conditions}, the first column has readings that result in two
+different stress positions, so {\small {\tt safe}} does not output a stressed
+form. However, in the second column, both readings have the same stress position, 
+so {\small {\tt safe}} outputs that stress position.
+
+The {\small {\tt guess}} approach is to randomly select one of the available 
+readings, and abstain if no readings are available. In the first column of
+Table~\ref{tab:conditions}, a random selection means that \rus{к\'{о}сти} 
+\emph{k\'{o}sti} is twice as likely as \rus{кост\'{и}} \emph{kost\'{i}}. The
+second column of Table~\ref{tab:conditions} contains stress-irrelevant
+ambiguity, so a random selection of a reading has the same result as the
+{\small {\tt safe}} approach.
+
+The {\small {\tt guessFreq}} 
+approach is to select the reading that is most frequent, with frequency data
+taken from a separate hand-disambiguated corpus. If none of the readings is 
+found in the corpus, then {\small {\tt guessFreq}} selects the reading
+with the tag sequence (lemma removed) that is most frequent in our corpus. If the 
+tag sequence is not found
+in our frequency list, then {\small {\tt guessFreq}} backs off to the 
+{\small {\tt guess}} algorithm. In the first column of
+Table~\ref{tab:conditions}, {\small {\tt guessFreq}} selects \rus{к\'{о}сти} 
+\emph{k\'{o}sti} because the tag sequence N-F-SG-GEN is more frequent than the
+other alternatives. Note 
+that for tokens with stress-irrelevant ambiguity (e.g. the second column of
+Table~\ref{tab:conditions}), {\small {\tt guess}} and 
+{\small {\tt guessFreq}} produce the same result as the {\small {\tt safe}} 
+method.
 
@@
 identical to {\small {\tt guess}} and {\small {\tt guessFreq}}, respectively,
-except that the {\small {\tt Syll}} approaches attempt to place stress on
-unknown tokens (i.e. tokens not found in the FST). Empirical research with
+except that the -{\small {\tt Syll}} approaches attempt to place stress on
+unknown tokens (i.e. tokens not found in the FST) by guessing which 
+\emph{syllable} is stressed. Empirical research with
 Russian native speakers reading nonce words in running text concludes that
@@
 in other words, the last vowel that is not part of an inflectional ending 
-\cite{Crosswhite.ea-03}. Therefore, the {\small {\tt Syll}} approaches 
+\cite{Crosswhite.ea-03}. Therefore, the -{\small {\tt Syll}} approaches 
 place the stress near the end of the stem. Determining 
@@
 Since our approach is lexicon-based, some of our results should be interpreted
-with respect to how many of the stressed wordforms in the gold standerd corpus
-can be found in the finite-state transducer. We refer to this measure as 
-\emph{recall}. Out of 4048 tokens, 3949 were 
+with respect to how many of the stressed wordforms in the gold standard corpus
+can be found in the output of the finite-state transducer. We refer to this 
+measure as \emph{recall}. Out of 4048 tokens, 3949 were 
 found in the FST, which is equal to 97.55\%.
@@
 guessing algorithms. After running the Constraint Grammar, recall was 97.35\%, 
-a reduction of 0.20\%.
+a reduction of 0.20\%. Apparently the Constraint Grammar erroneously removes some 
+correct readings, or else there are errors in the gold corpus.
 
@@
 \emph{with} the Constraint Grammar ({\small {\tt CG}}) x 6 approaches, 
-and RussianGram ({\small {\tt RussianGram}}). 
-Each token was categorised as either Failures to correctly place stress were 
-split into two categories: errors and abstentions. 
+and RussianGram ({\small {\tt RussianGram}}). Results are given in 
+Table~\ref{tab:results}. Each token was categorised as either an accurate output, 
+or one of two categories of failures: errors and abstentions. 
 If the stress tool outputs a stressed wordform, and it is incorrect, then it is
 counted as an `error'. If the stress tool outputs an unstressed wordform, then
-it is counted as an `abstention'. The sum of the errors and abstentions is 
-referred to as `totFail'. The sum of correct attempts and `errors' is referred to 
-as `totTry'. Results are given in Table~\ref{tab:results}.
+it is counted as an `abstention'.
+
+The right-most columns in 
+Table~\ref{tab:results} combine values of the basic categories. The term `totTry' 
+refers to the sum of the accuracy and error rate. This figure represents the 
+proportion of tokens on which our system output a stressed wordform. In the case 
+of {\small {\tt noCGbare}}, the accuracy\% (30.43) and error\% (0.17) sum to the 
+totTry value of 30.61. The term `totFail' refers to the sum of error rate 
+and abstention rate, which is the 
+proportion of tokens for which the system failed to output the correct stressed 
+form. In the case of {\small {\tt noCGbare}}, the error\% (0.17) 
+and abstention\% (69.39) sum to the (rounded) totFail value of of 69.57.
 
@@
 that a Constraint Grammar could theoretically achieve is 7.48\%. A comparison
-of {\small {\tt noCGsafe}} and {\small {\tt noCGsafe}} reveals an improvement
-of 3.14\%\footnote{one of the authors considers it significant that the CG 
-achieves an improvement roughly equivalent to $\pi$. :)}, which is about 42\%
-of the way to the ceiling.\rr{...while raising the error rate 0.25\%}
+of {\small {\tt noCGsafe}} and {\small {\tt CGsafe}} reveals an improvement
+of 3.14\%, which is about 42\% of the way to the ceiling recall.\rr{...while raising the error rate 0.25\%}
 
@@
 {\small {\tt CGguessFreqSyll}} is 0.42\%, which means that the 
-{\small {\tt Syll}} was accurate 21\% of the time.
+{\small {\tt Syll}} guess was accurate 21\% of the time.
 
@@
 
-\section{Discussion}
+\section{Discussion} \label{discussion}
 
@@
 number of words that are ambiguous with regard to stress. Based on the surface 
-forms in our FST (which is based on the same dictionary used for 
-\cite{hall_sproat_russian_2013}\rr{need to cite this without parentheses}), we 
-calculate that only 29~518 (1.05\%) 
+forms in our FST -- which is based on the same dictionary used for 
+\newcite{hall_sproat_russian_2013} -- 
+we calculate that only 29~518 (1.05\%) 
 of the 2~804~492 wordforms contained in our FST are stress-ambiguous. \rr{These 
@@
 Compare this with the finding of our experiment, that in running 
-text, at least 7.5\% of the tokens are stress-ambiguous. This is at least partly 
-because many of 
+text, at least 7.5\% of the tokens are stress-ambiguous. Stress ambiguity is
+therefore 7 times more prevalent in our corpus of running text than it is in 
+wordform dictionaries. This is at least partly because many of 
 the lexemes with shifting stress patterns are high-frequency. Any system that
@@
 we expect that the effectiveness of the Constraint Grammar will be affected by
-those differences. used in would shed light on those interactions as well.
+those differences.
 
+The resources needed for machine-learning --
+such as a large corpus of Russian running text with marked stress -- are simply
+not available at this time. Even so, lexicon- and rule-based approaches have
+some advantages over machine-learning approaches. For example, we are able to
+abstain from marking stress on tokens whose morphosyntactic ambiguity cannot be
+adequately resolved. In language-learning applications, this reduces the 
+likelihood of learners being exposed to incorrect wordforms, and accepting them
+as authoritative. Such circumstances can lead to considerable frustration and
+lack of trust in the learning tool. However, machine-learning does seem well-
+suited to
+placing stress on unknown words, since morphosyntactic analysis is problematic.
 The syllable-guessing algorithm {\small {\tt Syll}} used in this experiment was 
-overly simplistic, and it was not surprising that it was only moderately 
-successful. More rigorous approaches have been suggested in other studies.
-One approach -- exemplified by \newcite{Church-85}, 
-\newcite{Williams-87} and \newcite{Xomicevic_2008} -- is to 
-attempt to parse the unknown token by matching known prefixes and suffixes.
-Another approach -- exemplified by Pearson.ea-00, Webster-04, Dou.ea-09, and 
-\newcite{hall_sproat_russian_2013} -- is based on machine-learning. For unknown 
-words, \newcite{hall_sproat_russian_2013} achieve an accuracy of 83.9\%.
-It should be noted that their model was trained on a full list of Russian words,
-which is not representative of the words that would be unknown to a system like
-ours. Most of the complicated wordstress patterns are a closed 
-class.\footnote{The growing number of masculine nouns with shifting stress are 
-the exception to this generalization.} Future machine-learning work on Russian 
-wordstress placement in running text should limit the training data to word
-classes that are likely to be similar to unknown tokens, such as those with
+overly simplistic, and so it was not surprising that it was only moderately 
+successful. More rigorous rule-based approaches have been suggested in other 
+studies \cite{Church-85,Williams-87,Xomicevic_2008}. For example, 
+\newcite{Xomicevic_2008} attempt to parse the unknown token by matching known 
+prefixes and suffixes.
+
+Other studies have applied machine-learning to guessing stress of unknown words
+\cite{Pearson.ea-00,Webster-04,Dou.ea-09,hall_sproat_russian_2013}. For example,
+for unknown words, \newcite{hall_sproat_russian_2013} achieve an accuracy of 
+83.9\%. Implementing a similar model in a rule-based system like ours would allow 
+for some improvements to this result. Their model was trained on a full list of 
+Russian words, which is not representative of the words that would be unknown to 
+a system like ours. Most of the complicated wordstress patterns are a closed 
+class\footnote{The growing number of masculine nouns with shifting stress 
+(\emph{d\'{o}ktor$\sim$doktor\'{a}}) are an exception to this generalization.},
+so we could exclude closed classes of words from the training data, leaving only
+word classes that are likely to be similar to unknown tokens, such as those with 
 productive derivational affixes.
@@
 \item As with any lexicon- and rule-based system, those components can be expanded and improved, especially considering the short span over which they were developed.
-\item Integrate a more principled unknown word-guessing method (homicevic)
-\item Compare and/or integrate with stochastic methods. So far stochastic
-methods have only been used for guessing stress on unknown words.
 \item Another relevant issue is whether the ambiguity can be resolved at all.

### 2015-01-19 engkaz-eamt-2015/engkaz-eamt-2015.tex
dictionary
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58459 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 Dictionaries
+Dictionaries are used in lexical processing and morphological analysing, there are use two types: morphological(monolingual) and bilingual.
 
+Monolingual dictionaries
+
+English dictionary uses to determine lexical forms of each surface forms, it contains of list of all lexical units, definition of source language alphabet, grammatical categories, which is used to specify lexical forms, for example, like noun, verb, gender, etc., each categories have paradigms, they describes groups of correspondences between parts of surface forms and lexical forms.
+
+For Kazakh side, is used morphological transducers, which are based on the Helsinki
+Finite State Toolkit (Linden et al., 2011).
+It uses formalisms lexc for defining lexicons, twol formalism for morphophonological rules.
+Lexc formalism defines different word classes and subclasses.
+And twol determines phonological transformation like vowel harmony,desonorisation.
+For example, depend on preceding vowel, plural form "TAR" could be TER by harmonising with vowels on word: kitap+TAR, but mektep+TER.
+
+Bilingual dictionary
+
+Kazakh dictionary uses
 \section{Results}

### 2015-01-19 rusaccent-nodalida-2015/rusaccent.tex
typos
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58462 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 Constraint Grammar) in combination with the {\small {\tt bare}}, 
-{\small {\tt safe}}, {\small {\tt guess}} and {\small {\tt guessFreq}} 
-approaches. We also compare our outcomes with the 
-RussianGram\footnote{http://russiangram.com/} plugin 
-for the Google Chrome web browser. RussianGram is not open-source, so we
+{\small {\tt safe}}, {\small {\tt guess}}, {\small {\tt guessFreq}}, 
+{\small {\tt guessSyll}}, and {\small {\tt guessFreqSyll}} approaches. We also 
+compare our outcomes with the RussianGram\footnote{http://russiangram.com/} 
+plugin for the Google Chrome web browser. RussianGram is not open-source, so we
 can only guess what technologies support the service. In any case, it provides a
@@
 Russian words, which is not representative of the words that would be unknown to 
-a system like ours. Most of the complicated wordstress patterns are a closed 
-class\footnote{The growing number of masculine nouns with shifting stress 
+a system like ours. Most of the complicated wordstress patterns are closed 
+classes\footnote{The growing number of masculine nouns with shifting stress 
 (\emph{d\'{o}ktor$\sim$doktor\'{a}}) are an exception to this generalization.},
@@
 Future work/potential applications:
-Two sections: known words and unknown words
 \begin{itemize}
@@
 For example, when is stress marked in authentic by-native-for-native texts?
-A computer should not be expected to predict the stress on such words (although a computer could possibly identify words that should be marked for
-natives.)
-
+A computer should not be expected to predict the stress on such words (although a computer could possibly identify words that should be marked for natives.)
 \end{itemize}

### 2015-01-20 rusaccent-nodalida-2015/rusaccent.tex
boldfaced important numbers in results table
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58476 72bbbca6-d526-0410-a7d9-f06f51895060

@@
     {\small {\tt noCGbare}} & 30.43 & 0.17 & 69.39 & 30.61 & 69.57 \\
-    {\small {\tt noCGsafe}} & 90.07 & 0.49 & 9.44 & 90.56 & 9.93 \\
+    {\small {\tt noCGsafe}} & \bf{90.07} & 0.49 & 9.44 & 90.56 & 9.93 \\
     {\small {\tt noCGguess}} & 94.34 & 3.36 & 2.30 & 97.70 & 5.66 \\
@@
     {\small {\tt CGbare}} & 45.78 & 0.44 & 53.78 & 46.22 & 54.22 \\
-    {\small {\tt CGsafe}} & 93.21 & 0.74 & 6.05 & 93.95 & 6.79 \\
+    {\small {\tt CGsafe}} & \bf{93.21} & 0.74 & 6.05 & 93.95 & 6.79 \\
     {\small {\tt CGguess}} & 95.50 & 2.59 & 1.90 & 98.10 & 4.50 \\
-    {\small {\tt CGguessFreq}} & 95.73 & 2.40 & 1.88 & 98.12 & 4.27 \\
+    {\small {\tt CGguessFreq}} & \bf{95.73} & \bf{2.40} & 1.88 & 98.12 & 4.27 \\
     {\small {\tt CGguessSyll}} & 95.92 & 3.33 & 0.74 & 99.26 & 4.08 \\
-    {\small {\tt CGguessFreqSyll}} & 96.15 & 3.14 & 0.72 & 99.28 & 3.85 \\
+    {\small {\tt CGguessFreqSyll}} & \bf{96.15} & \bf{3.14} & 0.72 & 99.28 & 3.85 \\
     \hline
-    {\small {\tt RussianGram}} & 90.09 & 0.79 & 9.12 & 90.88 & 9.91
+    {\small {\tt RussianGram}} & \bf{90.09} & 0.79 & 9.12 & 90.88 & 9.91
   \end{tabular}

### 2015-01-20 rusaccent-nodalida-2015/rusaccent.tex
revised some of the Discussion section
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58492 72bbbca6-d526-0410-a7d9-f06f51895060

--- 2015-01-19 rusaccent-nodalida-2015/rusaccent.tex
@@
 
-\subsection{Background and Task Definition}
+\subsection{Background and Task Definition} \label{background}
 
@@
 form. In the case of {\small {\tt noCGbare}}, the error\% (0.17) 
-and abstention\% (69.39) sum to the (rounded) totFail value of of 69.57.
+and abstention\% (69.39) sum to the totFail value of of 69.57 (rounded).
 
@@
 RussianGram web service achieves results that are very close 
-to the {\small {\tt noCGsafe}} approach, so it would appear that it 
-utilizes a similar approach.
+to the {\small {\tt noCGsafe}} approach.
 
@@
 One of the main points of this paper is to highlight the importance of 
-syntactic context in the Russian wordstress placement task. 
-\newcite{hall_sproat_russian_2013} achieve 98.7\% accuracy on stress placement of 
-individual wordforms. To 
-understand this result, it should be remembered that Russian has a considerable
-number of words that are ambiguous with regard to stress. Based on the surface 
+syntactic context in the Russian wordstress placement task. If your intended 
+application has a low tolerance for error, the 
+{\small {\tt noCGsafe}} approach represents the highest accuracy that is possible 
+without leveraging syntactic information for disambiguation (90.07\%). In other words, a 
+system that is blind to morphosyntax and contextual disambiguation cannot
+significantly outperform {\small {\tt noCGsafe}}. It would appear that this is
+the method used by {\small {\tt RussianGram}}, since its results are so similar
+to {\small {\tt noCGsafe}}. Indeed, this result can be achieved most efficiently
+without any part-of-speech tagging, but through simple dictionary lookup.
+
+We noted in Section~\ref{background} that \newcite{hall_sproat_russian_2013}
+achieved 98.7\% accuracy on stress placement for individual wordforms 
+in a list (i.e. \emph{without} syntax). This result is 8.63\% higher than
+{\small {\tt noCGsafe}}, but it is also a fundamentally different task.
+Based on the surface 
 forms in our FST -- which is based on the same dictionary used for 
@@
 of the 2~804~492 wordforms contained in our FST are stress-ambiguous. \rr{These 
-stress-ambiguous wordforms have an average of 2.01 possible stress positions.} 
-Compare this with the finding of our experiment, that in running 
-text, at least 7.5\% of the tokens are stress-ambiguous. Stress ambiguity is
+stress-ambiguous wordforms have an average of 2.01 possible stress positions,
+so a perfect stress placement tool has a 50/50 chance of getting them right.
+This means that a perfect stress placement tool could achieve 99.48\% on their
+task.} In our corpus of running text, at least 7.5\% of the tokens are stress-
+ambiguous. Stress ambiguity is
 therefore 7 times more prevalent in our corpus of running text than it is in 
-wordform dictionaries. This is at least partly because many of 
-the lexemes with shifting stress patterns are high-frequency. Any system that
-cannot leverage syntactic context to disambiguate these wordforms cannot 
-achieve accuracy above 92.5\% without making an higher number of errors.
+our wordform dictionary.
 Since the task of wordstress placement is virtually always performed on
@@
 
-We are unaware of any other studies of Russian wordstress placement in running
-text. The results of our experiment are promising, but many questions remain
-unanswered.
+We are unaware of any other empirical evaluations of Russian wordstress placement 
+in running text. The results of our experiment are promising, but many questions 
+remain unanswered.
 The experiment was limited by properties of the gold standard corpus, including
@@
 writing to have a lower proportion of words with stress ambiguity, since it
-uses a more modern lexicon that has not inherited the idiosyncrasies that lead
-to wordstress ambiguity.
+uses a higher proportion of borrowed words.
 

### 2015-01-21 rusaccent-nodalida-2015/rusaccent.tex
more edits
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58512 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 from surface forms. Lexical stress and its attendant vowel reduction are 
-a prominent feature of spoken Russian, and the incorrect placement of stress
+a prominent feature of spoken Russian; the incorrect placement of stress
 can render speech almost incomprehensible. However, standard written Russian does not 
 typically mark word stress.\footnote{Texts intended for native speakers sometimes 
-mark stress on words that cannot be disambiguated through context, although this 
-is rare.} Without information about lexical stress position, 
+mark stress on words that cannot be disambiguated through context. Theoretically, 
+a perfect wordstress placement method could help an author identify
+tokens which should be stressed: any token that cannot
+be disambiguated by syntactic or semantic means should be marked for stress.} 
+Without information about lexical stress position, 
 correctly converting written Russian text to speech is impossible. This is true 
@@
 
-In the case of running 
-text in Russian, most stress placement 
+In the case of running text in Russian, most stress placement 
 ambiguity is rooted in intraparadigmatic and morphosyntactically incongruent 
@@
 the role of part-of-speech tagging in improving
-automatic stress placement, leaving morphosyntactically congruent stress 
-ambiguity to future work. \rr{this paragraph is choppy.}
+automatic stress placement. We leave morphosyntactically congruent stress 
+ambiguity to future work because of a combination of two reasons, 1) it is by far the least common type of stress ambiguity, and 2) disambiguating morphosyntactically congruent stress requires word sense disambiguation, a fundamentally different technology from the other approaches of this study.
 
@@
 a maximum entropy model on a dictionary of Russian words, and evaluated on
-wordlists containing `known' and `unknown' wordforms.\footnote{ADD FOOTNOTE
-EXPLAINING THEIR DEFINITION OF KNOWN AND UNKNOWN.} Their model 
+wordlists containing `known' and `unknown' wordforms.\footnote{\newcite{hall_sproat_russian_2013} randomly selected their training
+and test data from a list of wordforms, and so a number of lexemes had wordforms in both the training and test data.
+Wordforms in the test data whose sibling wordforms from the same lexeme were in the training set were categorised as
+`known' wordforms.} Their model 
 achieved 98.7\% accuracy on known words, and 83.9\% accuracy on unknown words. 
@@
 free texts on Russian language-learning websites.\footnote{URLs can be found in
-the TROLLing data repository (ADD URL).} This small corpus (N=7689) is 
+the X data repository (URL will be added in the final version).} This small corpus (N=7689) is 
 representative of texts 
@@
 unstressed, especially among prepositions, conjunctions, and particles.
-\rr{multisyllabics that can be unstressed: \rus{через, между}} Furthermore, there 
-are many
+% multisyllabics that can be unstressed: \rus{через, между} 
+Furthermore, there are many
 high-frequency monosyllabic homographs that can be either stressed or unstressed, 
@@
 writing to have a lower proportion of words with stress ambiguity, since it
-uses a higher proportion of borrowed words.
+uses a higher proportion of borrowed words, calques, and neologisms with simple
+stress patterns.
 
@@
 
-Future work/potential applications:
-\begin{itemize}
-\item As with any lexicon- and rule-based system, those components can be expanded and improved, especially considering the short span over which they were developed.
-\item Another relevant issue is whether the ambiguity can be resolved at all.
-For example, when is stress marked in authentic by-native-for-native texts?
-A computer should not be expected to predict the stress on such words (although a computer could possibly identify words that should be marked for natives.)
-\end{itemize}
+\section{Conclusions}
+
+We have demonstrated the effectiveness of using a Constraint Grammar to improve
+the results of a Russian wordstress placement task in running text by resolving
+42\% of the ambiguity in our gold corpus. We showed
+that stress ambiguity is 7 times more prevalent in our corpus of running text
+than it is in our lexicon, suggesting the importance of context-based
+disambiguation for this task.  As with any 
+lexicon- and rule-based system, the lexicon and rules can be expanded and 
+improved, but our initial results are promising, especially considering the short 
+timespan over which they were developed.
+
+As this is the first empirical study of its kind, we also discussed 
+methodological limitations and possibilities for subsequent research, including
+collecting stressed corpora of varying text complexity and/or genre, as well as
+implementing and/or adapting established wordstress-guessing methods for unknown 
+words.
+
 %\section*{Acknowledgments}

### 2015-01-21 rusaccent-nodalida-2015/rusaccent.tex
Added more citations in bibtex and cleaned up footnotes. Formatting is now submittable. Need to remove marginpars.
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58514 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-\subsection{Background and Task Definition} \label{background}
+\subsection{Background and task definition} \label{background}
 
@@
 
-\subsection{Stress Corpus}
+\subsection{Stress corpus}
 
@@
 
-\section{Automatic Stress Placement}
+\section{Automatic stress placement}
 
 State-of-the-art morphological analysis in Russian is based on finite-state
-technology.\footnote{add footnote or citations for Dialog, mystem, etc.} To our 
-knowledge, no open-source resources are available for analyzing and generating 
+technology\cite{Nozhov-03,Segalovich-03}. To our knowledge, no existing 
+open-source resources are available for analyzing and generating 
 stressed wordforms. Therefore, we developed free and open-source finite-state 
@@
 A Finite-State Transducer\footnote{Using two-level 
-morphology, xfst/hfst, blah, blah, blah} (FST) generates all possible 
-morphosyntactic readings of each wordform, and a constraint grammar\footnote{vislcg3, blah blah} then removes
-some extra readings based on syntactic context. The ultimate success of our 
+morphology \cite{Koskenniemi-84}, implemented in xfst \cite{Beesley.Karttunen-03} and hfst \cite{hfst-11}} (FST) generates all possible 
+morphosyntactic readings of each wordform, and a constraint 
+grammar\footnote{Implemented using vislcg3 constraint grammar parser (http://beta.visl.sdu.dk/cg3.html).}
+\cite{Karlsson-90,Karlsson.Voutilainen.ea-95} then removes
+some readings based on syntactic context. The ultimate success of our 
 stress placement system depends on the performance of the Constraint Grammar. 
@@
 approach is to select the reading that is most frequent, with frequency data
-taken from a separate hand-disambiguated corpus. If none of the readings is 
+taken from a separate hand-disambiguated corpus. If none of the readings are 
 found in the corpus, then {\small {\tt guessFreq}} selects the reading
@@
 ambiguous. Stress ambiguity is
-therefore 7 times more prevalent in our corpus of running text than it is in 
+therefore seven times more prevalent in our corpus of running text than it is in 
 our wordform dictionary.
@@
 42\% of the ambiguity in our gold corpus. We showed
-that stress ambiguity is 7 times more prevalent in our corpus of running text
+that stress ambiguity is seven times more prevalent in our corpus of running text
 than it is in our lexicon, suggesting the importance of context-based

### 2015-01-21 finnormal-nodalida-2015/finnormal2015.tex
some results from the unsupervised stuffs
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58452 72bbbca6-d526-0410-a7d9-f06f51895060
corpus stats
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58515 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-\author{Francis M. Tyers\\
-  UiT Norgga \'{a}rktala\v{s} universitehta, \\
-  N-9015 Norway \\
-  {\tt francis.tyers@uit.no} \\\And
-  Inari Listenmaa\\
-  Chalmers tekniska h\"{o}gskola, \\
-  Sweden \\
-  {\tt inari@cse.chalmers.se} \\}
-
+\author{}
+%
+%Francis M. Tyers\\
+%  UiT Norgga \'{a}rktala\v{s} universitehta, \\
+%  N-9015 Norway \\
+%  {\tt francis.tyers@uit.no} \\\And
+%  Inari Listenmaa\\
+%  Chalmers tekniska h\"{o}gskola, \\
+%  Sweden \\
+%  {\tt inari@cse.chalmers.se} \\}
+%
 \date{}
@@
 \section{Corpus}
+
+\begin{table}
+  \centering
+ \begin{tabular}{|l|r|}
+    \hline
+    \textbf{Section} & \textbf{Tokens} \\
+    \hline
+    \texttt{train} & 2,538 \\
+    \texttt{test} & 1,012 \\
+    \texttt{dev} & 1,003 \\
+    \hline
+ \end{tabular} 
+  \caption{Statistics on sentences from the parallel corpus.}
+  \label{table:corpsize}
+\end{table}
 
@@
 \begin{itemize}
-  \item 273 rules from \newcite{karlsson2008}'s grammar of Finnish (\S95--97)
-  \item Rules from the development corpus.
+  \item 273 rules from \newcite{karlsson2008}'s grammar of Finnish (\S95--97). The 
+    rules took around one hour to implement.
+  \item 98 rules written by examining the development corpus, these rules also
+    took approximately one hour to implement.
 \end{itemize}
@@
      rules-1 & 38.27 & 41.19 & 32.65 \\
-     rules-2 & - & - & - \\
+     rules-2 & 38.17 & 35.25 & 36.41 \\
+     rules-c & 36.56 & 39.68 & 34.68 \\
      \hline

### 2015-01-22 finnormal-nodalida-2015/finnormal2015.tex
stuffs
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58516 72bbbca6-d526-0410-a7d9-f06f51895060
smt done, now cbmt
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58517 72bbbca6-d526-0410-a7d9-f06f51895060
smt infos
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58522 72bbbca6-d526-0410-a7d9-f06f51895060
add numbers for unigram cbmt
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58525 72bbbca6-d526-0410-a7d9-f06f51895060
probably want to try unigram and bigram and take the best
git-svn-id: https://svn.code.sf.n

@@
 automatic tools. Another related area is machine translation between different written norms of 
-the same language. For example between Norwegian Bokmål and Norwegian Nynorsk CITEXXX.
+the same language. For example between Norwegian Bokmål and Norwegian Nynorsk \cite{unhammer2009}.
 
@@
 The corpus was split into three parts, testing, development and training. The testing
-and development portions contain 1,000 words each, with the remaining X,000 for training
+and development portions contain approximately 1,000 words each, with the remaining 2,538 for training
 supervised models.\footnote{The corpus is split into XX files of 500 words each. Files 01--02 
@@
 %
-%Input: Mä oon Inarinkaa
-%Step 1: Mä oon Inarinkaa
-%        Minä oon Inarinkaa
-%Step 2: Mä oon Inarinkaa
-%        Minä oon Inarinkaa
-%        Mä olen Inarinkaa
-%        Minä olen Inarinkaa
-%Step 3: Mä oon Inarinkaa
-%        Minä oon Inarinkaa
-%        Mä olen Inarinkaa
-%        Minä olen Inarinkaa
-%        Mä oon Inarin kanssa
-%        Minä oon Inarin kanssa
-%        Mä olen Inarin kanssa
-%        Minä olen Inarin kanssa
+%Input: Mä oon Tomminkaa
+%Step 1: Mä oon Tomminkaa
+%        Minä oon Tomminkaa
+%Step 2: Mä oon Tomminkaa
+%        Minä oon Tomminkaa
+%        Mä olen Tomminkaa
+%        Minä olen Tomminkaa
+%Step 3: Mä oon Tomminkaa
+%        Minä oon Tomminkaa
+%        Mä olen Tomminkaa
+%        Minä olen Tomminkaa
+%        Mä oon Tommin kanssa
+%        Minä oon Tommin kanssa
+%        Mä olen Tommin kanssa
+%        Minä olen Tommin kanssa
 %Rank:
-%	-13.4885 ← minä olen Inarin kanssa
-%	-12.7176 ← minä oon Inarin kanssa
-%	-12.4408 ← mä olen Inarin kanssa
-%	-9.2045 ← mä oon Inarin kanssa
-%	-8.8651 ← minä olen Inarinkaa
-%	-8.0941 ← minä oon Inarinkaa
-%	-7.8174 ← mä olen Inarinkaa
-%	-4.5811 ← mä oon Inarinkaa
-%Output: Minä olen Inarin kanssa
+%	-13.4885 ← minä olen Tommin kanssa
+%	-12.7176 ← minä oon Tommin kanssa
+%	-12.4408 ← mä olen Tommin kanssa
+%	-9.2045 ← mä oon Tommin kanssa
+%	-8.8651 ← minä olen Tomminkaa
+%	-8.0941 ← minä oon Tomminkaa
+%	-7.8174 ← mä olen Tomminkaa
+%	-4.5811 ← mä oon Tomminkaa
+%Output: Minä olen Tommin kanssa
 
@@
 
-The target-language model used was the same as in the unsupervised experiments.
+The target-language model corpus used was the same as in the unsupervised experiments.
+
+We trained models based on two approaches, the first being phrase-based machine translation
+CITEXXX and the second on character-based machine translation \cite{nakov2012}.
 
@@
 
-\subsubsection{Phrase-based}
+\subsubsection{Character-based}
 
-\subsubsection{Character-based}
+\newcite{nakov2012}
 
@@
      \hline
-     CBMT-cn & - & - & - \\
-     CBMT-cs & - & - & - \\
-     PBMT-cn & - & - & - \\
-     PBMT-cs & - & - & - \\
+     CBMT$^1$-cn & 43.09 & 48.34 & 33.55 \\
+     CBMT$^1$-cs & 46.22 & 52.27 & 29.21 \\
+     CBMT$^2$-cn & - & - & - \\
+     CBMT$^2$-cs & - & - & - \\
+     PBMT-cn & 28.05 & 35.42 & 48.37 \\
+     PBMT-cs & 27.95 & 36.13 & 46.76 \\
      \hline

### 2015-01-22 engkaz-eamt-2015/engkaz-eamt-2015.tex
bilingual dictionary
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58532 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-Kazakh dictionary uses
+Bilingual dictionary provides correspondences between lexical forms of source and target lanugages.
+Each entry consists of English word on the left side with part of speech and corresponding Kazakh word on the right side.
+Some entries could be ambiguous by part of speech, or just have many translations, which wil be chosen by lexical selection rules depend on context.
+This dictionary currently contains 13130 stems.
+
 \section{Results}

### 2015-01-22 finnormal-nodalida-2015/finnormal2015.tex
update table
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58530 72bbbca6-d526-0410-a7d9-f06f51895060
add citations to two papers
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58533 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{url}
+\usepackage{multirow}
 \usepackage{latexsym}
@@
   \centering
- \begin{tabular}{|l|r|}
+ \begin{tabular}{|l|r|r|r|}
     \hline
-    \textbf{Section} & \textbf{Tokens} \\
+    \multirow{2}{*}{\textbf{Section}} & \multicolumn{3}{|c|}{\textbf{Tokens}} \\\cline{2-4}
+                                      & \texttt{train} & \texttt{test} & \texttt{dev} \\
     \hline
-    \texttt{train} & 2,538 \\
-    \texttt{test} & 1,012 \\
-    \texttt{dev} & 1,003 \\
+    Colloquial                           & 2,538 & 1,012 & 1,003 \\ 
+    Normalised                           & 2,538 & 1,012 & 1,003 \\ 
+    Standardised                         & 2,476 & 991 & 1,000 \\
     \hline
@@
   \hline
-  oon tosi hy\"{o}dyllinen assari     & ole todella hy\"{o}dyllinen assistentti & ole todella hy\"{o}dyllinen assistentti \\
+  oon tosi hy\"{o}dyllinen assari     & olen todella hy\"{o}dyllinen assistentti & olen todella hy\"{o}dyllinen assistentti \\
   t\"{a}n vikan minuutin.           & t\"{a}m\"{a}n viimeisen minuutin.            & t\"{a}m\"{a}n viimeisen minuutin. \\
@@
 
+CBMT: \cite{TiedemannEAMT2009, TiedemannEACL2012}
+
 For both approaches we trained two systems, the first used the \emph{normalised}
@@
 part of the corpus as the target language.
+
+The idea behind this was that the \emph{normalised} part of the corpus would be
 
@@
 
-Joonas Kyl\"{a}m\"{a}
+Joonas Kylm\"{a}l\"{a}
 

### 2015-01-23 finnormal-nodalida-2015/finnormal2015.tex
bokmål
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58536 72bbbca6-d526-0410-a7d9-f06f51895060
more stuffs on the paper
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58538 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 automatic tools. Another related area is machine translation between different written norms of 
-the same language. For example between Norwegian Bokmål and Norwegian Nynorsk \cite{unhammer2009}.
+the same language. For example between Norwegian Bokm\aa{}l and Norwegian Nynorsk \cite{unhammer2009}.
 
@@
 We trained models based on two approaches, the first being phrase-based machine translation
-CITEXXX and the second on character-based machine translation \cite{nakov2012}.
-
-CBMT: \cite{TiedemannEAMT2009, TiedemannEACL2012}
+CITEXXX and the second on character-based machine translation \cite{nakov2012,TiedemannEAMT2009}.
 
@@
 
-\newcite{nakov2012}
+\newcite{nakov2012} present a method of statistical machine translation on the character level between related languages that takes advantage of phrase-based machine translation architecture.
+The method relies on preprocessing the input and output by inserting spaces in between the characters of words, for example the string `M\"{a} meen Helsinkiin' would become `M \"{a} \$ m e e n \$ H e l s i n k i i n' with a unigram model, or `M\"{a} \$ me en \$ He ls in ki in' with a bigram model.
 

### 2015-01-23 finnormal-nodalida-2015/finnormal2015.tex
more stuffs
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58539 72bbbca6-d526-0410-a7d9-f06f51895060
more stuffs
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58540 72bbbca6-d526-0410-a7d9-f06f51895060
add todo list
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58541 72bbbca6-d526-0410-a7d9-f06f51895060
changed examples; more text and cite in 4.2.1
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58549 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
- \emph{suomen puhekieli}
+\cite{viinikka2013}
+
+``Yleensä kirjakieli ymmärretään samaksi kuin yleiskieli: kielenhuollon suositusten mukaiseksi standardoiduksi kieleksi, jonka sanasto on yleistajuista.
+Puhekieli puolestaan mielletään kirjakieltä arkisemmaksi ja tyyliltään vapaamuotoisemmaksi kielimuodoksi, jonka ei tarvitse noudattaa kirjakielen sääntöjä (esim.
+mie, hänen koira, me kirjotetaan vs.
+minä, hänen koiransa, me kirjoitamme).''
 
@@
   \hline
-  mut siis, oli siell\"{a} jotain vaikka     & mutta siis, oli siell\"{a} jotain vaikka     & mutta oli siell\"{a} jotain     \\
-  puheteknologiamatskuja.  & puheteknologiamateriaaleja. & puheteknologiamateriaaleja.  \\
+  tai emm\"{a} tii\"{a} olikse erikseen joku & tai en\#min\"{a} tied\"{a} oliko\#se
+  erikseen jokin & tai en min\"{a} tied\"{a} oliko erikseen \\
+  nuorisoalennus & nuorisoalennus & nuorisoalennus \\
   \hline
-  oon tosi hy\"{o}dyllinen assari     & olen todella hy\"{o}dyllinen assistentti & olen todella hy\"{o}dyllinen assistentti \\
-  t\"{a}n vikan minuutin.           & t\"{a}m\"{a}n viimeisen minuutin.            & t\"{a}m\"{a}n viimeisen minuutin. \\
+ toistaseks tullu kaks kysymyst\"{a} &toistaiseksi tullut kaksi kysymyst\"{a} & toistaseksi on tullut kaksi kysymyst\"{a} \\
   \hline
@@
 
+After preprocessing, the corpora are processed as with the phrase-based system, with the difference that the language model order is increased from 5 to 10-grams.
+
 \section{Results}
@@
   \end{tabular}
+  \caption{Results...}
   \label{table:results}
-  \caption{Results...}
 \end{table}
+
+Table~\ref{table:results} presents the results of the experiments.
 

### 2015-01-23 engkaz-eamt-2015/engkaz-eamt-2015.tex
about cg
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58557 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
+Rules
+Disambiguation rules
+
+This kind of rules solves the problem of part-of-speech(POS) ambiguity by Constraint Grammar(CG) (Karlsson et al., 1995).
+The output of morphological analyser could be ambiguous, and in this case CG rules choose correct analysis from multiple analyses.
+From English to Kazakh POS ambiguity could be only in English side, but also some choices depends on Kazakh grammar.
+
+Lexicon selection rules
+
 \section{Results}

### 2015-01-23 finnormal-nodalida-2015/finnormal2015.tex
add more stuffs to 2.1
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58558 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{nodalida2015}
+\usepackage{expex}
 %\usepackage{times}
@@
 %  Sweden \\
-%  {\tt inari@cse.chalmers.se} \\}
+%  {\tt inari@chalmers.se} \\}
 %
@@
 
-\cite{viinikka2013}
+Colloquial Finnish differs from standard Finnish in some key features:
 
-``Yleensä kirjakieli ymmärretään samaksi kuin yleiskieli: kielenhuollon suositusten mukaiseksi standardoiduksi kieleksi, jonka sanasto on yleistajuista.
-Puhekieli puolestaan mielletään kirjakieltä arkisemmaksi ja tyyliltään vapaamuotoisemmaksi kielimuodoksi, jonka ei tarvitse noudattaa kirjakielen sääntöjä (esim.
-mie, hänen koira, me kirjotetaan vs.
-minä, hänen koiransa, me kirjoitamme).''
+Morphology: dialectal variation, shortening and merging together morphemes, e.g. \emph{en min\"{a}} `not I' $\rightarrow$ \emph{emm\"{a}} (`I don't').
+Syntax: due to the spontaneous nature of colloquial language, syntactic structures have stuffs that are common to spoken language, such as ......
+
+Sentence 1 is written in colloquial Finnish, and Sentence 2 is how it would be expressed in a standard way.
+
+\ex
+\begingl
+\gla nelonenkin oli vaan silleen et fonotaksista p\"{a}\"{a}ttelin //
+\glb fourth-{\sc also} was just like.that that phonotactics.{\sc ela} I.deduced //
+% \glft `I deduced also the fourth from phonotactics' //
+\endgl
+\xe
+
+\ex
+\begingl
+\gla p\"{a}\"{a}ttelin nelj\"{a}nnenkin fonotaksista //
+\glb I.deduced fourth-{\sc also} phonotactics.{\sc ela} //
+% \glft `I deduced also the fourth from phonotactics' //
+\endgl
+\xe
+
+\cite{viinikka2013} describe the common meaning of the terms colloquial  (\emph{puhekieli}) and standard (\emph{yleiskieli} or \emph{kirjakieli}) Finnish: standard language follows the recommendations of kielenhuolto and is unified in morphology and vocabulary; colloquial language is everyday and free and shows local and idiolectal variation and doesn't need to follow the rules of kielenhuolto.
+% ``Yleens\"{a} kirjakieli ymm\"{a}rret\"{a}\"{a}n samaksi kuin yleiskieli: kielenhuollon suositusten mukaiseksi standardoiduksi kieleksi, jonka sanasto on yleistajuista.
+Puhekieli puolestaan miellet\"{a}\"{a}n kirjakielt\"{a} arkisemmaksi ja tyylilt\"{a}\"{a}n vapaamuotoisemmaksi kielimuodoksi, jonka ei tarvitse noudattaa kirjakielen s\"{a}\"{a}ntöj\"{a} (esim.
+mie, h\"{a}nen koira, me kirjotetaan vs.
+min\"{a}, h\"{a}nen koiransa, me kirjoitamme).''
 
@@
   \hline
- toistaseks tullu kaks kysymyst\"{a} &toistaiseksi tullut kaksi kysymyst\"{a} & toistaseksi on tullut kaksi kysymyst\"{a} \\
+ toistaseks tullu kaks kysymyst\"{a} &toistaiseksi tullut kaksi kysymyst\"{a} & toistaiseksi on tullut kaksi kysymyst\"{a} \\
   \hline
@@
 
-%Rules: mä -> minä, oon -> olen, ?*nkaa -> ?*n kanssa
+%Rules: m\"{a} -> min\"{a}, oon -> olen, ?*nkaa -> ?*n kanssa
 %
-%Input: Mä oon Tomminkaa
-%Step 1: Mä oon Tomminkaa
-%        Minä oon Tomminkaa
-%Step 2: Mä oon Tomminkaa
-%        Minä oon Tomminkaa
-%        Mä olen Tomminkaa
-%        Minä olen Tomminkaa
-%Step 3: Mä oon Tomminkaa
-%        Minä oon Tomminkaa
-%        Mä olen Tomminkaa
-%        Minä olen Tomminkaa
-%        Mä oon Tommin kanssa
-%        Minä oon Tommin kanssa
-%        Mä olen Tommin kanssa
-%        Minä olen Tommin kanssa
+%Input: M\"{a} oon Tomminkaa
+%Step 1: M\"{a} oon Tomminkaa
+%        Min\"{a} oon Tomminkaa
+%Step 2: M\"{a} oon Tomminkaa
+%        Min\"{a} oon Tomminkaa
+%        M\"{a} olen Tomminkaa
+%        Min\"{a} olen Tomminkaa
+%Step 3: M\"{a} oon Tomminkaa
+%        Min\"{a} oon Tomminkaa
+%        M\"{a} olen Tomminkaa
+%        Min\"{a} olen Tomminkaa
+%        M\"{a} oon Tommin kanssa
+%        Min\"{a} oon Tommin kanssa
+%        M\"{a} olen Tommin kanssa
+%        Min\"{a} olen Tommin kanssa
 %Rank:
-%	-13.4885 ← minä olen Tommin kanssa
-%	-12.7176 ← minä oon Tommin kanssa
-%	-12.4408 ← mä olen Tommin kanssa
-%	-9.2045 ← mä oon Tommin kanssa
-%	-8.8651 ← minä olen Tomminkaa
-%	-8.0941 ← minä oon Tomminkaa
-%	-7.8174 ← mä olen Tomminkaa
-%	-4.5811 ← mä oon Tomminkaa
-%Output: Minä olen Tommin kanssa
+%	-13.4885 ← min\"{a} olen Tommin kanssa
+%	-12.7176 ← min\"{a} oon Tommin kanssa
+%	-12.4408 ← m\"{a} olen Tommin kanssa
+%	-9.2045 ← m\"{a} oon Tommin kanssa
+%	-8.8651 ← min\"{a} olen Tomminkaa
+%	-8.0941 ← min\"{a} oon Tomminkaa
+%	-7.8174 ← m\"{a} olen Tomminkaa
+%	-4.5811 ← m\"{a} oon Tomminkaa
+%Output: Min\"{a} olen Tommin kanssa
 

### 2015-01-25 finnormal-nodalida-2015/finnormal2015.tex
more stuff to 2.1
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58576 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
+\cite{viinikka2013} describe the common meaning of the terms colloquial (\emph{puhekieli}) and standard (\emph{yleiskieli} or \emph{kirjakieli}) Finnish: standard language follows the recommendations of kielenhuolto and is unified in morphology and vocabulary; colloquial language shows local and idiolectal variation, and has structures that are characteristic to spoken variety,  such as discourse particles and incomplete clauses.
+Prototypically, colloquial language is spoken, everyday and spontaneous, and standard is written, planned and structured.
+
 Colloquial Finnish differs from standard Finnish in some key features:
 
-Morphology: dialectal variation, shortening and merging together morphemes, e.g. \emph{en min\"{a}} `not I' $\rightarrow$ \emph{emm\"{a}} (`I don't').
-Syntax: due to the spontaneous nature of colloquial language, syntactic structures have stuffs that are common to spoken language, such as ......
+Morphology: dialectal variation, shortening and merging of morphemes, e.g. \emph{en min\"{a}} `not I' $\rightarrow$ \emph{emm\"{a}} (`I don't').
 
-Sentence 1 is written in colloquial Finnish, and Sentence 2 is how it would be expressed in a standard way.
+Syntax: due to the spontaneous nature of colloquial language, syntactic structures are common for spoken language.
+The following example is from our data set, first sentence is the original colloquial version.
+The gloss shows the actual word-by-word translation, and the translation shows a similar style and register in English.
 
@@
 \begingl
-\gla nelonenkin oli vaan silleen et fonotaksista p\"{a}\"{a}ttelin //
-\glb fourth-{\sc also} was just like.that that phonotactics.{\sc ela} I.deduced //
-% \glft `I deduced also the fourth from phonotactics' //
+\gla seiskakin oli vaan silleen et fonotaksista p\"{a}\"{a}ttelin //
+\glb seven-{\sc also} was just like.that that phonotactics.{\sc ela} I.deduced //
+\glft `also the seventh, it was like, I just deduced it from phonotactics' //
 \endgl
@@
 
+Sentence 2 is translated in many ways.
+On the lexical level, the word \emph{seiska} is a colloquial name for `number 7', and in a literary standard it is replaced by the ordinal \emph{seitsemäs} `seventh'.
+The syntactic structure of the sentence is markedly spoken; the word \emph{seiska} is topicalised, and the main information ``deduced from phonotactics'' is a subordinate clause.
+The translation below is shorter and more precise, leaving just the main information.
+
 \ex
 \begingl
-\gla p\"{a}\"{a}ttelin nelj\"{a}nnenkin fonotaksista //
-\glb I.deduced fourth-{\sc also} phonotactics.{\sc ela} //
-% \glft `I deduced also the fourth from phonotactics' //
+\gla p\"{a}\"{a}ttelin seitsem\"{a}nnenkin fonotaksista //
+\glb I.deduced seventh-{\sc also} phonotactics.{\sc ela} //
+\glft `I deduced also the seventh from phonotactics' //
 \endgl
@@
 
-\cite{viinikka2013} describe the common meaning of the terms colloquial  (\emph{puhekieli}) and standard (\emph{yleiskieli} or \emph{kirjakieli}) Finnish: standard language follows the recommendations of kielenhuolto and is unified in morphology and vocabulary; colloquial language is everyday and free and shows local and idiolectal variation and doesn't need to follow the rules of kielenhuolto.
+For the partial version, we didn't touch the syntactic structure, but changed morphology and vocabulary. Below is a version that shows the original and the partial.
+
+\ex
+\begingl
+\gla seiskakin oli vaan silleen et fonotaksista p\"{a}\"{a}ttelin //
+\glb seitsemäskin oli vain sill\"{a}\#lailla ett\"{a} fonotaksista p\"{a}\"{a}ttelin //
+%\glb seventh-{\sc also} was just like.that that phonotactics.{\sc ela} I.deduced //
+% \glft `also the seventh, it was like, I just deduced it from phonotactics' //
+\endgl
+\xe
+
 % ``Yleens\"{a} kirjakieli ymm\"{a}rret\"{a}\"{a}n samaksi kuin yleiskieli: kielenhuollon suositusten mukaiseksi standardoiduksi kieleksi, jonka sanasto on yleistajuista.

### 2015-01-26 finnormal-nodalida-2015/finnormal2015.tex
cleanup
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58579 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-\cite{viinikka2013} describe the common meaning of the terms colloquial (\emph{puhekieli}) and standard (\emph{yleiskieli} or \emph{kirjakieli}) Finnish: standard language follows the recommendations of kielenhuolto and is unified in morphology and vocabulary; colloquial language shows local and idiolectal variation, and has structures that are characteristic to spoken variety,  such as discourse particles and incomplete clauses.
-Prototypically, colloquial language is spoken, everyday and spontaneous, and standard is written, planned and structured.
+\cite{viinikka2013} describe the common meaning of the terms colloquial (\emph{puhekieli}) and standard (\emph{yleiskieli} or \emph{kirjakieli}) Finnish: standard language is regulated by kielenhuolto and is unified in morphology and vocabulary; colloquial language shows local and idiolectal variation, and has structures that are characteristic to spoken variety,  such as discourse particles and incomplete clauses.
 
@@
 
-Morphology: dialectal variation, shortening and merging of morphemes, e.g. \emph{en min\"{a}} `not I' $\rightarrow$ \emph{emm\"{a}} (`I don't').
+Morphology shows dialectal variation, shortening and merging of morphemes, e.g. \emph{en min\"{a}} `not I' $\rightarrow$ \emph{emm\"{a}} (`I don't').
 
 Syntax: due to the spontaneous nature of colloquial language, syntactic structures are common for spoken language. 
-The following example is from our data set, first sentence is the original colloquial version.
-The gloss shows the actual word-by-word translation, and the translation shows a similar style and register in English.
+Prototypically, colloquial language is spoken and spontaneous, and standard is written, planned and structured.
+We illustrate the differences with the following example from our data set.
+Sentence 1 is the original colloquial version.
+The gloss shows the actual word-by-word translation, and the translation shows similar style and register in English.
 
@@
 
-Sentence 2 is translated in many ways.
-On the lexical level, the word \emph{seiska} is a colloquial name for `number 7', and in a literary standard it is replaced by the ordinal \emph{seitsemäs} `seventh'.
-The syntactic structure of the sentence is markedly spoken; the word \emph{seiska} is topicalised, and the main information ``deduced from phonotactics'' is a subordinate clause.
-The translation below is shorter and more precise, leaving just the main information.
+For the partial version, we didn't touch the syntactic structure, but changed morphology and vocabulary.
+On the lexical level, the word \emph{seiska} `number 7' is colloquial style, and in the standard translation it is replaced by the ordinal \emph{seitsem\"{a}s} `seventh'.
+Other changes in the partial version are common morphological or phonological, phenomena, such as reducing the diphthong in \emph{vaan} $\rightarrow$ \emph{vain} or shortening in  \emph{et} $\rightarrow$ \emph{että}.
+The original sentence and the partial translation are shown below, aligned word by word.
+
+\ex
+\begingl
+\gla seiskakin oli vaan silleen et fonotaksista p\"{a}\"{a}ttelin //
+\glb seitsem\"{a}skin oli vain sill\"{a}\#lailla ett\"{a} fonotaksista p\"{a}\"{a}ttelin //
+%\glb seventh-{\sc also} was just like.that that phonotactics.{\sc ela} I.deduced //
+% \glft `also the seventh, it was like, I just deduced it from phonotactics' //
+\endgl
+\xe
+
+ The syntactic structure of the original sentence is markedly spoken; the word \emph{seiska} is topicalised, and the main information ``deduced from phonotactics'' is in a subordinate clause. 
+The translation into  is shorter and more precise, leaving just the main information.
 
@@
 \glft `I deduced also the seventh from phonotactics' //
-\endgl
-\xe
-
-For the partial version, we didn't touch the syntactic structure, but changed morphology and vocabulary. Below is a version that shows the original and the partial.
-
-\ex
-\begingl
-\gla seiskakin oli vaan silleen et fonotaksista p\"{a}\"{a}ttelin //
-\glb seitsemäskin oli vain sill\"{a}\#lailla ett\"{a} fonotaksista p\"{a}\"{a}ttelin //
-%\glb seventh-{\sc also} was just like.that that phonotactics.{\sc ela} I.deduced //
-% \glft `also the seventh, it was like, I just deduced it from phonotactics' //
 \endgl

### 2015-01-26 finnormal-nodalida-2015/finnormal2015.tex
some fixes to examples
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58580 72bbbca6-d526-0410-a7d9-f06f51895060

@@
   of simply treating colloquial Finnish as standard Finnish and offers promise for the adaptation
-  of language-technology tools created for standard Finnish to colloquial Finnish.
+  of language-technology tools created for standard Finnish to colloquial Finnish. To this end
+  the paper also presents preliminary results which show promise for using normalisation in 
+  the machine translation task.
 \end{abstract}
@@
 
-For the partial version, we didn't touch the syntactic structure, but changed morphology and vocabulary.
+For the partial version, we did not change the syntactic structure, but changed only morphology and vocabulary.
 On the lexical level, the word \emph{seiska} `number 7' is colloquial style, and in the standard translation it is replaced by the ordinal \emph{seitsem\"{a}s} `seventh'.
@@
     \hline
-    Colloquial                           & 2,538 & 1,012 & 1,003 \\ 
-    Normalised                           & 2,538 & 1,012 & 1,003 \\ 
-    Standardised                         & 2,476 & 991 & 1,000 \\
+    Colloquial                           & 5,103 & 1,012 & 1,003 \\ 
+    Normalised                           & 5,105 & 1,012 & 1,003 \\ 
+    Standardised                         & 4,982 & 991 & 1,000 \\
     \hline
@@
   \hline
-  tai emm\"{a} tii\"{a} olikse erikseen joku & tai en\#min\"{a} tied\"{a} oliko\#se
-  erikseen jokin & tai en min\"{a} tied\"{a} oliko erikseen \\
-  nuorisoalennus & nuorisoalennus & nuorisoalennus \\
+  tai emm\"{a} tii\"{a} olikse erikseen & tai en\#min\"{a} tied\"{a} oliko\#se erikseen & tai en min\"{a} tied\"{a} oliko erikseen \\
+  joku nuorisoalennus                             & jokin nuorisoalennus & nuorisoalennus \\
   \hline
- toistaseks tullu kaks kysymyst\"{a} &toistaiseksi tullut kaksi kysymyst\"{a} & toistaiseksi on tullut kaksi kysymyst\"{a} \\
+   toistaseks tullu        & toistaiseksi tullut  & toistaiseksi on tullut  \\
+   kaks kysymyst\"{a}       & kaksi kysymyst\"{a} & kaksi kysymyst\"{a} \\
   \hline
@@
 The corpus was split into three parts, testing, development and training. The testing
-and development portions contain approximately 1,000 words each, with the remaining 2,538 for training
+and development portions contain approximately 1,000 words each, with the remaining approximately 5,000 words for training
 supervised models.\footnote{The corpus is split into XX files of 500 words each. Files 01--02 
-  were used for development; 03--04 for testing and 05--XX for training.}
+  were used for development; 03--04 for testing and 05--14 for training.}
 
-\section{Methodology}
+\section{Experiments}
 
@@
 
+\subsection{End-to-end translation}
+
+The idea of the end-to-end translation experiment was to evaluate how well the different 
+normalisation strategies worked in combination with another language technology tool. To
+evaluate this, we took the test corpus and translated it to English.
+
+We then translated the corpus with each of the best performing methods of each class, and 
+calculated the {\sc bleu} score.
+
 \section{Results}
@@
      \hline
-     CBMT$^1$-cn & 43.09 & 48.34 & 33.55 \\
-     CBMT$^1$-cs & 46.22 & 52.27 & 29.21 \\
-     CBMT$^2$-cn & - & - & - \\
-     CBMT$^2$-cs & - & - & - \\
+     CBMT-cn & 43.09 & 48.34 & 33.55 \\
+     CBMT-cs & 46.22 & 52.27 & 29.21 \\
      PBMT-cn & 28.05 & 35.42 & 48.37 \\
@@
   \end{tabular}
-  \caption{Results...}
-  \label{table:results}
+  \caption{Results for the normalisation task}
+  \label{table:results-norm}
 \end{table}
 
-Table~\ref{table:results} presents the results of the experiments.
+Tables~\ref{table:results-norm} and \ref{table:results-trad} presents the results of the experiments.
+
+\begin{table}
+  \centering
+  \begin{tabular}{|l|r|r|r|}
+     \hline
+    \textbf{System} & \textbf{PER} & \textbf{WER} & \textbf{BLEU} \\
+     \hline
+     Colloquial & - & - & 11.95 \\
+     Normalised & - & - & 21.81 \\
+     Standardised & - & - & 17.62 \\
+     \hline
+     rules-2 & - & - & 12.28 \\
+     cbmt-cn & - & - & 12.35 \\
+     pbmt-cn & - & - & 17.90 \\
+     \hline
+  \end{tabular}
+  \caption{Results for the translation task}
+  \label{table:results-trad}
+\end{table}
 
@@
 
+% expand test corpus -- up to 10,000 words
+% add other colloquial genres
+% apply to other NLP tasks
 % weight rules 
-% use system in pipeline (e.g. translation)
 %
@@
 We have presented a parallel corpus of colloquial Finnish and standard Finnish --
-to our knowledge the first of its kind -- and a method for converting colloquial
-Finnish to standard Finnish.
+to our knowledge the first of its kind -- and an evaluation of methods for 
+converting colloquial Finnish to standard Finnish.
+
+We have shown that converting from colloquial Finnish to standard Finnish substantially
+helps with the machine translation task.
 

### 2015-01-26 finnormal-nodalida-2015/finnormal2015.tex
more stuffs
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58581 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-%Rules: m\"{a} -> min\"{a}, oon -> olen, ?*nkaa -> ?*n kanssa
-%
-%Input: M\"{a} oon Tomminkaa
-%Step 1: M\"{a} oon Tomminkaa
-%        Min\"{a} oon Tomminkaa
-%Step 2: M\"{a} oon Tomminkaa
-%        Min\"{a} oon Tomminkaa
-%        M\"{a} olen Tomminkaa
-%        Min\"{a} olen Tomminkaa
-%Step 3: M\"{a} oon Tomminkaa
-%        Min\"{a} oon Tomminkaa
-%        M\"{a} olen Tomminkaa
-%        Min\"{a} olen Tomminkaa
-%        M\"{a} oon Tommin kanssa
-%        Min\"{a} oon Tommin kanssa
-%        M\"{a} olen Tommin kanssa
-%        Min\"{a} olen Tommin kanssa
-%Rank:
-%	-13.4885 ← min\"{a} olen Tommin kanssa
-%	-12.7176 ← min\"{a} oon Tommin kanssa
-%	-12.4408 ← m\"{a} olen Tommin kanssa
-%	-9.2045 ← m\"{a} oon Tommin kanssa
-%	-8.8651 ← min\"{a} olen Tomminkaa
-%	-8.0941 ← min\"{a} oon Tomminkaa
-%	-7.8174 ← m\"{a} olen Tomminkaa
-%	-4.5811 ← m\"{a} oon Tomminkaa
-%Output: Min\"{a} olen Tommin kanssa
+\begin{table*}
+  \centering
+  \begin{tabular}{|lrl|l|}
+    \hline
+    \textbf{Rules:} & & 1. m\"{a} $\rightarrow$ min\"{a} \\
+     ~              & & 2. oon $\rightarrow$ olen \\
+     ~              & & 3. (?+)nkaa $\rightarrow$ ...n kanssa \\
+    \hline
+  \end{tabular}
+~\\
+~\\
+  \begin{tabular}{|l|r|l|l|}
+    \hline
+    \textbf{Input:} & \multicolumn{3}{l|}{M\"{a} oon Inarinkaa `I am with Inari'.} \\
+    \hline
+    \multirow{2}{*}{Step 1} & & M\"{a} oon Inarinkaa & apply rule 1 \\
+                            & & Min\"{a} oon Inarinkaa & \\
+    \hline
+    \multirow{2}{*}{Step 2} & & M\"{a} oon Inarinkaa & apply rule 2 \\
+                            &   & Min\"{a} oon Inarinkaa & \\
+                            &   & M\"{a} olen Inarinkaa & \\
+                            &   & Min\"{a} olen Inarinkaa & \\
+    \hline
+    \multirow{2}{*}{Step 3} &   & M\"{a} oon Inarinkaa   & apply rule 3\\
+                            &   & Min\"{a} oon Inarinkaa  &\\
+                            &   & M\"{a} olen Inarinkaa  &\\
+                            &   & Min\"{a} olen Inarinkaa  &\\
+                            &   & M\"{a} oon Inarin kanssa  &\\
+                            &   & Min\"{a} oon Inarin kanssa  &\\
+                            &   & M\"{a} olen Inarin kanssa  &\\
+                            &   & Min\"{a} olen Inarin kanssa  &\\
+    \hline
+    \multirow{2}{*}{Step 4} & -4.5811 &  Min\"{a} olen Inarin kanssa & rank candidates \\
+                            & -7.8174 &  Min\"{a} oon Inarin kanssa & \\
+                            & -8.0941 &  M\"{a} olen Inarin kanssa & \\
+                            & -8.8651&  M\"{a} oon Inarin kanssa & \\
+                            & -9.2045 &  Min\"{a} olen Inarinkaa & \\
+                            & -12.4408 &  Min\"{a} oon Inarinkaa & \\
+                            & -12.7176 &  M\"{a} olen Inarinkaa & \\
+                            & -13.4885 &  M\"{a} oon Inarinkaa & \\
+
+    \hline
+    \textbf{Output:} & \multicolumn{3}{l|}{Min\"{a} olen Inarin kanssa} \\
+    \hline
+  \end{tabular}
+  \caption{Example trace of the unsupervised normalisation method. Rules are applied in order to each of 
+     the possible candidate translations in turn. The candidates are then ranked using an $n$-gram language model 
+     of standard Finnish and either an $n$-best list
+     or the best candidate is output.}
+  \label{table:unsup-trace}
+\end{table*}
 

### 2015-01-26 finnormal-nodalida-2015/finnormal2015.tex
more stuffs
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58582 72bbbca6-d526-0410-a7d9-f06f51895060
more references
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58583 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 proofing tools, such as spelling and grammar checkers CITEXXX can be used to encourage adherence
-to particular orthographic or grammatical norms. Accent and diacritic restoration CITEXXX is similar
+to particular orthographic or grammatical norms. Accent and diacritic restoration --- for example in \cite{scannell2011} --- is similar
 in that it aims to bring text closer to standard orthography in order to facilitate treatment by 
@@
      ~              & & 2. oon $\rightarrow$ olen \\
-     ~              & & 3. (?+)nkaa $\rightarrow$ ...n kanssa \\
+     ~              & & 3. (?+)nkaa $\rightarrow$ \textbackslash1n kanssa \\
     \hline

### 2015-01-26 finnormal-nodalida-2015/finnormal2015.tex
more stuffs
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58584 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 proofing tools, such as spelling and grammar checkers CITEXXX can be used to encourage adherence
-to particular orthographic or grammatical norms. Accent and diacritic restoration --- for example in \cite{scannell2011} --- is similar
+to particular orthographic or grammatical norms. Accent and diacritic restoration --- for example in \newcite{scannell2011} --- is similar
 in that it aims to bring text closer to standard orthography in order to facilitate treatment by 

### 2015-01-26 finnormal-nodalida-2015/finnormal2015.tex
more fixes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58585 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-\cite{viinikka2013} describe the common meaning of the terms colloquial (\emph{puhekieli}) and standard (\emph{yleiskieli} or \emph{kirjakieli}) Finnish: standard language is regulated by kielenhuolto and is unified in morphology and vocabulary; colloquial language shows local and idiolectal variation, and has structures that are characteristic to spoken variety,  such as discourse particles and incomplete clauses.
+\newcite{viinikka2013} describe the common meaning of the terms colloquial (\emph{puhekieli}) and standard (\emph{yleiskieli} or \emph{kirjakieli}) Finnish: standard language is regulated by kielenhuolto and is unified in morphology and vocabulary; colloquial language shows local and idiolectal variation, and has structures that are characteristic to spoken variety,  such as discourse particles and incomplete clauses.
 

### 2015-01-26 finnormal-nodalida-2015/finnormal2015.tex
fix bigram example
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58586 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \newcite{nakov2012} present a method of statistical machine translation on the character level between related languages that takes advantage of phrase-based machine translation architecture.
-The method relies on preprocessing the input and output by inserting spaces in between the characters of words, for example the string `M\"{a} meen Helsinkiin' would become `M \"{a} \$ m e e n \$ H e l s i n k i i n' with a unigram model, or `M\"{a} \$ me en \$ He ls in ki in' with a bigram model.
+The method relies on preprocessing the input and output by inserting spaces in between the characters of words, for example the string `m\"{a} meen Helsinkiin' would become `m \"{a} \$ m e e n \$ H e l s i n k i i n' with a unigram model, or `m\"{a} \"{a}\$ \$m me ee en n\$ \$H He el ls si in nk ki ii in' with a bigram model.
 

### 2015-01-26 finnormal-nodalida-2015/finnormal2015.tex
add example
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58587 72bbbca6-d526-0410-a7d9-f06f51895060
+1 ref
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58588 72bbbca6-d526-0410-a7d9-f06f51895060
one more thing
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58589 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 language can cause a substantial decrease in quality. For example in machine translation,
-or part-of-speech tagging. CITEXXX
+parsing and part-of-speech tagging \cite{eisenstein2013}.
 
@@
   \hline
-  \ldots & \ldots & \ldots \\
+  ja sit 2009 just ennenku & ja sitten 2009 juuri ennen\#kuin  & ja sitten 2009 juuri ennen kuin  \\ 
+menin Japaniin & menin Japaniin & menin Japaniin
+ \\
   \hline

### 2015-01-26 rusaccent-nodalida-2015/rusaccent.tex
Final submissiona draft. Marginpars removed. I read over it and did not find any incomplete or badly worded sentences.
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58590 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \newcommand{\rr}[1]{\marginpar{\scriptsize R: #1}} % Rob's comments
+\newcommand{\cw}[1]{\marginpar{\scriptsize W: #1}} % Cute wife's comments
 \newcommand{\dm}[1]{\marginpar{\scriptsize D: #1}} % Detmar's comments
@@
 
-\title{Automatic word stress annotation of Russian running text (max 8 pages)}
+\title{Automatic word stress annotation of Russian unrestricted text}
 
@@
   We evaluate the effectiveness of finite-state tools we developed for 
-  automatically annotating word stress in Russian running text. To our knowledge,
-  this is the first study to empirically evaluate the results of this task. 
-  Given an adequate 
-  lexicon with specified stress, the primary obstacle for correct stress 
-  placement is disambiguating homographic wordforms. Using a Constraint Grammar, 
-  we achieve 93.21\% accuracy, a 3.14\% improvement over the baseline
-  undisambiguated performance. 
-  Combining our Constraint Grammar with frequency data and a simple
-  algorithm for unknown word stress position guessing, we achieve 96.15\%
-  accuracy. These results highlight the need morphosyntactic disambiguation
+  automatically annotating word stress in Russian unrestricted text. To our 
+  knowledge, this is the first study to empirically evaluate the results of this 
+  task. Given an adequate lexicon with specified stress, the primary obstacle for 
+  correct stress placement is disambiguating homographic wordforms. Using a 
+  Constraint Grammar, we achieve 93.21\% accuracy with minimal errors, a 3.14\% 
+  improvement over the baseline undisambiguated performance with minimal errors. 
+  For applications with a higher threshold for errors, we combined our Constraint 
+  Grammar with frequency-based guessing and a simple
+  algorithm for unknown word stress position guessing, thereby achieving 96.15\%
+  accuracy. These results highlight the need for morphosyntactic disambiguation
   in the wordstress placement task for Russian, and set a standard for future
@@
 mark stress on words that cannot be disambiguated through context. Theoretically, 
-a perfect wordstress placement method could help an author identify
-tokens which should be stressed: any token that cannot
+a perfect wordstress placement system could help an author identify
+tokens which should be stressed for natives: any token that cannot
 be disambiguated by syntactic or semantic means should be marked for stress.} 
@@
 
-\needspace{4\baselineskip}
+\needspace{4\baselineskip} % keeps example all together (no orphan line) 
 \ex. Intraparadigmatic homographs \label{ex:intrahom}
@@
 The remaining two types of stress ambiguity occur between lexemes. 
+\emph{Morphosyntactically incongruent} stress ambiguity occurs between homographs
+that belong to separate lexemes, and whose morphosyntactic values are different, 
+as shown in \ref{ex:MSincongruent}.
+
+\ex. Morphosyntactically incongruent homographs \label{ex:MSincongruent}
+\a. \rus{попад\'{а}ть} \emph{popadát'} `fall.\textsubscript{INF-IPFV}' \\
+    \rus{поп\'{а}дать} \emph{popádat'} `fall.\textsubscript{INF-PFV}'
+\b. \rus{узнаёт} \emph{uznaët} `finds out.\textsubscript{PRS-IPFV-3SG}' \\
+    \rus{узн\'{а}ет} \emph{uznáet} `finds out.\textsubscript{FUT-PFV-3SG}'
+\c. \rus{н\'{а}шей} \emph{nášej} `our.\textsubscript{F-SG-GEN/DAT/LOC...}'\\
+    \rus{наш\'{е}й} \emph{našéj} `sew on.\textsubscript{IMP-2SG}'
+\d. \rus{дор\'{о}га} \emph{doróga} `road.\textsubscript{N-F-SG-NOM}'\\
+    \rus{дорог\'{а}} \emph{dorogá} `dear.\textsubscript{ADJ-F-SG-PRED}'
+
 \emph{Morphosyntactically congruent} stress ambiguity occurs between homographs 
@@
 and whose morphosyntactic values are identical, as shown in
-Table~\ref{table:MScongruent}. This kind of stress ambiguity is relatively rare.
-Resolving this ambiguity would require the use of technologies such as Word Sense 
-Disambiguation.
+Table~\ref{table:MScongruent}. This kind of stress ambiguity is relatively rare, and resolving this ambiguity would require the use of technologies such as Word Sense Disambiguation.
 
@@
 
-\emph{Morphosyntactically incongruent} stress ambiguity occurs between homographs
-that belong to separate lexemes, and whose morphosyntactic values are different, 
-as shown in \ref{ex:MSincongruent}.
-
-\ex. Morphosyntactically incongruent homographs \label{ex:MSincongruent}
-\a. \rus{попад\'{а}ть} \emph{popadát'} `fall.\textsubscript{INF-IPFV}' \\
-    \rus{поп\'{а}дать} \emph{popádat'} `fall.\textsubscript{INF-PFV}'
-\b. \rus{узнаёт} \emph{uznaët} `finds out.\textsubscript{PRS-IPFV-3SG}' \\
-    \rus{узн\'{а}ет} \emph{uznáet} `finds out.\textsubscript{FUT-PFV-3SG}'
-\c. \rus{н\'{а}шей} \emph{nášej} `our.\textsubscript{F-SG-GEN/DAT/LOC...}'\\
-    \rus{наш\'{е}й} \emph{našéj} `sew on.\textsubscript{IMP-2SG}'
-\d. \rus{дор\'{о}га} \emph{doróga} `road.\textsubscript{N-F-SG-NOM}'\\
-    \rus{дорог\'{а}} \emph{dorogá} `dear.\textsubscript{ADJ-F-SG-PRED}'
-
 It should be noted that not all morphosyntactic ambiguity leads to stress 
-ambiguity. For example, all inflecting modifiers have the same form for F-SG-GEN, 
-F-SG-LOC, F-SG-DAT and F-SG-INS. We refer to this as \emph{stress-irrelevant} 
-ambiguity, since all readings have
+ambiguity. For example, all inflecting modifiers have the identical form for 
+F-SG-GEN, F-SG-LOC, F-SG-DAT and F-SG-INS. We refer to this as 
+\emph{stress-irrelevant} ambiguity, since all readings have
 the same stress placement. Stress-irrelevant ambiguity can be intraparadigmatic 
@@
 
-In the case of running text in Russian, most stress placement 
+In the case of unrestricted text in Russian, most stress placement 
 ambiguity is rooted in intraparadigmatic and morphosyntactically incongruent 
-ambiguity. Detailed part-of-speech tagging can help determine the stress of 
-these forms, since each alternative stress placement is tied to a different
-tag sequence. In this study we focus on 
-the role of part-of-speech tagging in improving
+ambiguity. Detailed part-of-speech tagging with morphosyntactic analysis
+can help determine the stress of these forms, since each alternative stress 
+placement is tied to a different tag sequence. In this study we focus on 
+the role of detailed part-of-speech tagging in improving
 automatic stress placement. We leave morphosyntactically congruent stress 
-ambiguity to future work because of a combination of two reasons, 1) it is by far the least common type of stress ambiguity, and 2) disambiguating morphosyntactically congruent stress requires word sense disambiguation, a fundamentally different technology from the other approaches of this study.
+ambiguity to future work because it is by far the least common type of stress
+%\rr{Exactly how common is congruent stress ambiguity?}
+ambiguity, and disambiguating morphosyntactically congruent stress requires 
+a fundamentally different technology from the other approaches of this study
+(i.e. Word Sense Disambiguation).
 
@@
 without standard diacritics. In order to generate speech from these texts, they 
-must first be normalized by restoring diacritics. A slightly different
-situation arises with Arabic and Hebrew, since standard orthography lacks vowel 
-diacritics altogether. For such languages, the `restoration' of vowel diacritics 
-results in less ambiguity than in standard orthography. For languages with 
-inherently ambiguous orthography,
+must first be normalized by restoring diacritics.
+
+A slightly different situation arises with languages whose standard orthography 
+is underspecified, like vowel quality in Arabic or Hebrew. For such languages, 
+the `restoration' of vowel diacritics results in less ambiguity than in standard 
+orthography. For languages with inherently ambiguous orthography,
 it may be more precise to refer to this as `diacritic enhancement', since it 
@@
 Many studies of Russian text-to-speech and automatic speech recognition make note of the 
-difficulties caused by the shortcomings of their stress-marking resources (e.g. \cite{krivnova_1998}), but very few studies have targeted stress-marking itself, and they tend to
-focus on placing stress on unknown words, with almost no attention to the problems that
-arise with known words. For example, 
-\cite{Xomicevic_2008} developed a set of heuristics for guessing stress placement on 
-unknown words in Russian. More recently, \cite{hall_sproat_russian_2013} trained 
-a maximum entropy model on a dictionary of Russian words, and evaluated on
-wordlists containing `known' and `unknown' wordforms.\footnote{\newcite{hall_sproat_russian_2013} randomly selected their training
-and test data from a list of wordforms, and so a number of lexemes had wordforms in both the training and test data.
-Wordforms in the test data whose sibling wordforms from the same lexeme were in the training set were categorised as
-`known' wordforms.} Their model 
-achieved 98.7\% accuracy on known words, and 83.9\% accuracy on unknown words. 
-One major shortcoming of evaluating on wordlists is 
-that they do not reflect wordform frequency.
+difficulties caused by the shortcomings of their stress-marking resources (e.g. 
+\newcite{krivnova_1998}), but very few studies have targeted stress-marking 
+itself, and they focus on placing stress on unknown words, with almost no 
+attention to the problems that arise with known words. For example, 
+\newcite{Xomicevic_2008} developed a set of heuristics for guessing stress 
+placement on unknown words in Russian. More recently, 
+\newcite{hall_sproat_russian_2013} trained a maximum entropy model on a 
+dictionary of Russian words, and evaluated on wordlists containing `known' and 
+`unknown' wordforms.\footnote{\newcite{hall_sproat_russian_2013} randomly 
+selected their training and test data from a list of wordforms, and so a number 
+of lexemes had wordforms in both the training and test data. Wordforms in the 
+test data whose sibling wordforms from the same lexeme were in the training set 
+were categorised as `known' wordforms.} Their model achieved 98.7\% accuracy on 
+known words, and 83.9\% accuracy on unknown words.
+
+One major shortcoming of evaluating on wordlists is that they do not reflect 
+wordform frequency.
 Many of the most problematic stress ambiguities in Russian occur in 
@@
 Russian texts with marked word stress are relatively rare, except in materials
-for language learners, which are predominantly proprietary. Our corpus was 
+for language learners, which are predominantly proprietary. Our gold corpus was 
 collected from
@@
 Furthermore, there are many
+%\rr{examples}
 high-frequency monosyllabic homographs that can be either stressed or unstressed, 
@@
 
+Similarly, some compound Russian words have secondary stress, but this is rarely 
+marked, if at all, even in educational materials. Therefore, even though our 
+tools are built to mark secondary stress, we cannot evaluate secondary stress
+marks, since they are absent in our gold corpus.
+
+In order to test our wordstress placement system, we removed all stress marks
+from the gold corpus, then marked stress on the plain text version, and then
+compared with the original.
+
 \section{Automatic stress placement}
@@
 State-of-the-art morphological analysis in Russian is based on finite-state
-technology\cite{Nozhov-03,Segalovich-03}. To our knowledge, no existing 
+technology \cite{Nozhov-03,Segalovich-03}. To our knowledge, no existing 
 open-source resources are available for analyzing and generating 
@@
 well-known \emph{Grammatical Dictionary of Russian} \cite{Zaliznjak-77}.
-A Finite-State Transducer\footnote{Using two-level 
-morphology \cite{Koskenniemi-84}, implemented in xfst \cite{Beesley.Karttunen-03} and hfst \cite{hfst-11}} (FST) generates all possible 
-morphosyntactic readings of each wordform, and a constraint 
-grammar\footnote{Implemented using vislcg3 constraint grammar parser (http://beta.visl.sdu.dk/cg3.html).}
+Our Finite-State Transducer\footnote{Using two-level 
+morphology \cite{Koskenniemi-84}, implemented in both xfst
+\cite{Beesley.Karttunen-03} and hfst \cite{hfst-11}} (FST) generates all possible 
+morphosyntactic readings of each wordform, and our Constraint 
+Grammar\footnote{Implemented using vislcg3 constraint grammar parser
+(http://beta.visl.sdu.dk/cg3.html).}
 \cite{Karlsson-90,Karlsson.Voutilainen.ea-95} then removes
@@
 we also evaluate various approaches to deal with the remaining ambiguity,
-as described in the following paragraph and Table~\ref{tab:conditions}.
+as described below.
 
@@
 The {\small {\tt guess}} approach is to randomly select one of the available 
-readings, and abstain if no readings are available. In the first column of
+readings. In the first column of
 Table~\ref{tab:conditions}, a random selection means that \rus{к\'{о}сти} 
@@
 
-So far, the approaches discussed are
-dependent on the availability of readings from the FST. Two final approaches 
-address unknown tokens.
-The {\small {\tt guessSyll}} and {\small {\tt guessFreqSyll}} approaches are 
-identical to {\small {\tt guess}} and {\small {\tt guessFreq}}, respectively,
-except that the -{\small {\tt Syll}} approaches attempt to place stress on
-unknown tokens (i.e. tokens not found in the FST) by guessing which 
-\emph{syllable} is stressed. Empirical research with
-Russian native speakers reading nonce words in running text concludes that
-the default stress position in Russian is the last syllable of the stem, or
-in other words, the last vowel that is not part of an inflectional ending 
-\cite{Crosswhite.ea-03}. Therefore, the -{\small {\tt Syll}} approaches 
-place the stress near the end of the stem. Determining 
-which vowel is stem-final
-is itself a difficult task. Our algorithm merely places the stress on the last 
-vowel that is followed by a consonant. Although more sophisticated algorithms
-are necessary to accurately locate the stem boundary, this is not the central
-focus of this study.
+So far, the approaches discussed are dependent on the availability of readings 
+from the FST. The focus of our study is on disambiguation of known 
+words, but we also wanted to guess the stress of unknown tokens in order to 
+establish some kind of accuracy maximum for applications that are more tolerant 
+of higher error rates. To this end, we selected a simple guessing method for 
+unknown words. A recent study by \newcite{Lavitskaya.Kabak-14} concludes that 
+Russian has default final stress in consonant-final words, and penultimate stress 
+in vowel-final words.\footnote{There is some controversy over how to define 
+default stress in Russian. \newcite{Crosswhite.ea-03} conclude that the default 
+stress position in Russian is the last syllable of the stem.} Based on this 
+conclusion, our syllable-guessing method places the stress on the last vowel that 
+is followed by a consonant.\footnote{More sophisticated heuristics and machine-
+learning methods have been suggested, but unknown words are not the central focus 
+of this study, and so a simple method was judged to be sufficient.} This method 
+is applied in two approaches, {\small {\tt guessSyll}} 
+and {\small {\tt guessFreqSyll}}, which are otherwise identical to 
+{\small {\tt guess}} and {\small {\tt guessFreq}}, respectively.
 
@@
 above.
-
-\section{Results}
 
@@
 
+\section{Results}
+
 Single-syllable words and words that are unstressed in the gold standard were
@@
 This number represents the ceiling for methods relying on the FST.
-Higher scores are only achievable by expanding the FST or by using syllable-
-guessing algorithms. After running the Constraint Grammar, recall was 97.35\%, 
-a reduction of 0.20\%. Apparently the Constraint Grammar erroneously removes some 
-correct readings, or else there are errors in the gold corpus.
+Higher scores are only achievable by expanding the FST or by using 
+syllable-guessing algorithms. After running the Constraint Grammar, recall was 
+97.35\%, a reduction of 0.20\%. Apparently the Constraint Grammar erroneously 
+removes some correct readings, or else there are errors in the gold corpus.
 
@@
 counted as an `error'. If the stress tool outputs an unstressed wordform, then
-it is counted as an `abstention'.
+it is counted as an `abstention'. Abstentions can be result of either unknown
+wordforms, or known wordforms with unspecified stress.
 
-The right-most columns in 
+The two right-most columns in 
 Table~\ref{tab:results} combine values of the basic categories. The term `totTry' 
-refers to the sum of the accuracy and error rate. This figure represents the 
+refers to the sum of the accuracy and error rate. This number represents the 
 proportion of tokens on which our system output a stressed wordform. In the case 
@@
 and abstention\% (69.39) sum to the totFail value of of 69.57 (rounded). 
-
-\rr{Are differences significant?}
+%\rr{Are differences significant?}
 
@@
 forms whose stress position varies from speaker to speaker (e.g.
-\rus{зав\'{и}л\'{и}сь} \emph{zav\'{i}l\'{i}s'}), errors in the gold corpus (e.g. 
-\rus{вер\'{и}м} \emph{ver\'{i}m}) . The {\small {\tt noCGsafe}} approach
+\rus{зав\'{и}л\'{и}сь} \emph{zav\'{i}l\'{i}s'}), or errors in the gold corpus 
+(e.g. \rus{вер\'{и}м} \emph{ver\'{i}m}).
+
+The {\small {\tt noCGsafe}} approach
 achieves a 60\% improvement in accuracy (90.07\%), which means that 89.39\% of 
@@
 of {\small {\tt noCGsafe}} and {\small {\tt CGsafe}} reveals an improvement
-of 3.14\%, which is about 42\% of the way to the ceiling recall.\rr{...while raising the error rate 0.25\%}
+of 3.14\%, which is about 42\% of the way to the ceiling recall.
+%\rr{...while raising the error rate 0.25\%}
 
@@
 {\small {\tt Syll}} guess was accurate 21\% of the time.
-
-\rr{Examine errors and categorize/refute them.}
+%\rr{Examine errors and categorize/refute them.}
 
@@
 we calculate that only 29~518 (1.05\%) 
-of the 2~804~492 wordforms contained in our FST are stress-ambiguous. \rr{These 
-stress-ambiguous wordforms have an average of 2.01 possible stress positions,
-so a perfect stress placement tool has a 50/50 chance of getting them right.
-This means that a perfect stress placement tool could achieve 99.48\% on their
-task.} In our corpus of running text, at least 7.5\% of the tokens are stress-
-ambiguous. Stress ambiguity is
-therefore seven times more prevalent in our corpus of running text than it is in 
-our wordform dictionary.
+of the 2~804~492 wordforms contained in our FST are stress-ambiguous.
+%\rr{These stress-ambiguous wordforms have an average of 2.01 possible stress positions, so a perfect stress placement tool has a 50/50 chance of getting them right.
+This means that a perfect stress placement tool could achieve 99.48\% on their task.}
+In our corpus of unrestricted text, at least 7.5\% of the tokens are 
+stress-ambiguous. Stress ambiguity is therefore seven times more prevalent in our 
+corpus of unrestricted text than it is in our wordform dictionary.
 Since the task of wordstress placement is virtually always performed on
-running text, it seems prudent to make use of surrounding information.
+running text, it seems prudent to make use of surrounding contextual information.
 The experiment described in this paper demonstrates that a Constraint Grammar
@@
 We are unaware of any other empirical evaluations of Russian wordstress placement 
-in running text. The results of our experiment are promising, but many questions 
-remain unanswered.
+in unrestricted text. The results of our experiment are promising, but many 
+questions remain unanswered.
 The experiment was limited by properties of the gold standard corpus, including
@@
 makes our results more generalizable, but a larger corpus would allow for
-evaluating each genre individually. For example, the vast majority of words with
-shifting stress are of Slavic origins, so we expect a genre such as technical
-writing to have a lower proportion of words with stress ambiguity, since it
-uses a higher proportion of borrowed words, calques, and neologisms with simple
-stress patterns.
+evaluating each genre individually. For example, the vast majority of Russian
+words with shifting stress are of Slavic origins, so we expect a genre such as 
+technical writing to have a lower proportion of words with stress ambiguity, 
+since it uses a higher proportion of borrowed words, calques, and neologisms with 
+simple stress patterns.
 
@@
 of different kinds of syntactic constructions vary with text complexity 
-\cite{Vajjala.Meurers-12}, and
-we expect that the effectiveness of the Constraint Grammar will be affected by
-those differences.
+\cite{Vajjala.Meurers-12}, and so we expect that the effectiveness of the 
+Constraint Grammar will be affected by those differences.
 
-The resources needed for machine-learning --
-such as a large corpus of Russian running text with marked stress -- are simply
+The resources needed for machine-learning -- such as a large corpus of Russian 
+unrestricted text with marked stress -- are simply
 not available at this time. Even so, lexicon- and rule-based approaches have
@@
 as authoritative. Such circumstances can lead to considerable frustration and
-lack of trust in the learning tool. However, machine-learning does seem well-
-suited to
-placing stress on unknown words, since morphosyntactic analysis is problematic.
-The syllable-guessing algorithm {\small {\tt Syll}} used in this experiment was 
-overly simplistic, and so it was not surprising that it was only moderately 
-successful. More rigorous rule-based approaches have been suggested in other 
-studies \cite{Church-85,Williams-87,Xomicevic_2008}. For example, 
+lack of trust in the learning tool. However, in error-tolerant applications, 
+machine-learning does seem well-suited to placing stress on unknown words, since 
+morphosyntactic analysis is problematic.
+
+The syllable-guessing algorithm {\small {\tt Syll}} used in this 
+experiment was overly simplistic, and so it was not surprising that it was only 
+moderately successful. More rigorous rule-based approaches have been suggested in 
+other studies \cite{Church-85,Williams-87,Xomicevic_2008}. For example, 
 \newcite{Xomicevic_2008} attempt to parse the unknown token by matching known 
@@
 for unknown words, \newcite{hall_sproat_russian_2013} achieve an accuracy of 
-83.9\%. Implementing a similar model in a rule-based system like ours would allow 
-for some improvements to this result. Their model was trained on a full list of 
+83.9\%. Their model was trained on a full list of 
 Russian words, which is not representative of the words that would be unknown to 
-a system like ours. Most of the complicated wordstress patterns are closed 
+a system like ours, so we could modify their approach to fit our application. 
+Most of the complicated wordstress patterns are closed 
 classes\footnote{The growing number of masculine nouns with shifting stress 
-(\emph{d\'{o}ktor$\sim$doktor\'{a}}) are an exception to this generalization.},
+(\rus{д\'{о}ктор}$\sim$\rus{доктор\'{а}} \emph{d\'{o}ktor$\sim$doktor\'{a}} 
+`doctor'$\sim$`doctors') are one exception to this generalization.},
 so we could exclude closed classes of words from the training data, leaving only
@@
 productive derivational affixes.
+%\rr{What about using mystem to evaluate, convert to our tags, then generate?}
 
@@
 We have demonstrated the effectiveness of using a Constraint Grammar to improve
-the results of a Russian wordstress placement task in running text by resolving
-42\% of the ambiguity in our gold corpus. We showed
+the results of a Russian wordstress placement task in unrestricted text by 
+resolving 42\% of the stress ambiguity in our gold corpus. We showed
 that stress ambiguity is seven times more prevalent in our corpus of running text

### 2015-01-26 finnormal-nodalida-2015/finnormal2015.tex
some more improvements, describe scannell's experiment
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58591 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-This paper presents a method for converting colloquial Finnish to standard Finnish and 
-a parallel corpus for evaluation.
+This paper evaluates a number of methods for converting colloquial Finnish to standard Finnish and describes a parallel corpus for evaluation.
 
@@
 \newcite{scannell2014} presents a method for normalising pre-standardised text in Irish to 
-the modern standard. The method relies on
+the modern standard. The method relies on a translation model consisting of word-to-word 
+correspondences in addition to spelling rules. Each word-to-word mapping has the same
+conditional probability and a penalty is assigned to each spelling rule application. Decoding
+works by processing the source sentence word-for-word left-to-right, keeping track of the 
+possible `hypothesis' translations and their probabilities, and when the end of 
+sentence is reached, the most probable is output.
 
@@
   \end{tabular}
-  \caption{Example sentences from the parallel corpus.}
+  \caption{Example sentences from the parallel corpus. The \texttt{\#} mark represents a missing
+    word boundary.}
   \label{table:corpexample}
@@
 and development portions contain approximately 1,000 words each, with the remaining approximately 5,000 words for training
-supervised models.\footnote{The corpus is split into XX files of 500 words each. Files 01--02 
-  were used for development; 03--04 for testing and 05--14 for training.}
+supervised models.\footnote{The corpus is split into 14 files of 500 words each. Files 01--02 
+  were used for development; 03--04 for testing and 05--14 for training.} Table~\ref{table:corpsize} gives statistics on the number of words in each section.
 

### 2015-01-26 finnormal-nodalida-2015/finnormal2015.tex
add ref for zens
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58592 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 We trained models based on two approaches, the first being phrase-based machine translation
-CITEXXX and the second on character-based machine translation \cite{nakov2012,TiedemannEAMT2009}.
+\cite{zens2002} and the second on character-based machine translation \cite{nakov2012,TiedemannEAMT2009}.
 

### 2015-01-26 finnormal-nodalida-2015/finnormal2015.tex
remove kielenhuolto
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58593 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-\newcite{viinikka2013} describe the common meaning of the terms colloquial (\emph{puhekieli}) and standard (\emph{yleiskieli} or \emph{kirjakieli}) Finnish: standard language is regulated by kielenhuolto and is unified in morphology and vocabulary; colloquial language shows local and idiolectal variation, and has structures that are characteristic to spoken variety,  such as discourse particles and incomplete clauses.
+\newcite{viinikka2013} describe the common meaning of the terms colloquial (\emph{puhekieli}) and standard (\emph{yleiskieli} or \emph{kirjakieli}) Finnish: standard language is regulated and is unified in morphology and vocabulary; colloquial language shows local and idiolectal variation, and has structures that are characteristic to spoken variety,  such as discourse particles and incomplete clauses.
 

### 2015-01-26 finnormal-nodalida-2015/finnormal2015.tex
more text
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58594 72bbbca6-d526-0410-a7d9-f06f51895060
update text
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58595 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
+The rules included both simple one-to-one (`m\"{a}' $\rightarrow$ `min\"{a}') and one-to-many (`emm\"{a}' $\rightarrow$ `en min\"{a}') word correspondences,
+and also regular expression substitutions which could match a prefix or a suffix (`(?+)nkaa' $\rightarrow$ `\textbackslash1n kanssa').
+
 \begin{table*}
@@
 
-The idea behind this was that the \emph{normalised} part of the corpus would be
+The idea behind this was that the \emph{normalised} part of the corpus would be closer
+to the original colloquial text than the \emph{standardised} part, making it easier
+to learn the alignment model.
 
@@
 
-The idea of the end-to-end translation experiment was to evaluate how well the different 
-normalisation strategies worked in combination with another language technology tool. To
-evaluate this, we took the test corpus and translated it to English.
+In order to evaluate how well the different normalisation strategies worked in 
+combination with another language technology tool, we performed an end-to-end experiment
+involving machine translation.
+To evaluate this, we took the test corpus and translated it to English.
 

### 2015-01-26 finnormal-nodalida-2015/finnormal2015.tex
add ref to table
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58596 72bbbca6-d526-0410-a7d9-f06f51895060
small changes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58597 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 On the lexical level, the word \emph{seiska} `number 7' is colloquial style, and in the standard translation it is replaced by the ordinal \emph{seitsem\"{a}s} `seventh'.
-Other changes in the partial version are common morphological or phonological, phenomena, such as reducing the diphthong in \emph{vaan} $\rightarrow$ \emph{vain} or shortening in  \emph{et} $\rightarrow$ \emph{että}.
+Other changes in the partial version are common morphological or phonological, phenomena, such as reducing the diphthong in \emph{vain} $\rightarrow$ \emph{vaan} or shortening in  \emph{ett\"{a}} $\rightarrow$ \emph{et}.
 The original sentence and the partial translation are shown below, aligned word by word.
@@
 
+Table~\ref{table:unsup-trace} gives an example trace of the system on a simple sentence
+using three replace rules.
+
 \begin{table*}
@@
     \hline
-    \textbf{Input:} & \multicolumn{3}{l|}{M\"{a} oon Inarinkaa `I am with Inari'.} \\
+    \textbf{Input:} & \multicolumn{3}{l|}{M\"{a} oon Tomminkaa `I am with Tommi'.} \\
     \hline
-    \multirow{2}{*}{Step 1} & & M\"{a} oon Inarinkaa & apply rule 1 \\
-                            & & Min\"{a} oon Inarinkaa & \\
+    \multirow{2}{*}{Step 1} & & M\"{a} oon Tomminkaa & apply rule 1 \\
+                            & & Min\"{a} oon Tomminkaa & \\
     \hline
-    \multirow{2}{*}{Step 2} & & M\"{a} oon Inarinkaa & apply rule 2 \\
-                            &   & Min\"{a} oon Inarinkaa & \\
-                            &   & M\"{a} olen Inarinkaa & \\
-                            &   & Min\"{a} olen Inarinkaa & \\
+    \multirow{2}{*}{Step 2} & & M\"{a} oon Tomminkaa & apply rule 2 \\
+                            &   & Min\"{a} oon Tomminkaa & \\
+                            &   & M\"{a} olen Tomminkaa & \\
+                            &   & Min\"{a} olen Tomminkaa & \\
     \hline
-    \multirow{2}{*}{Step 3} &   & M\"{a} oon Inarinkaa   & apply rule 3\\
-                            &   & Min\"{a} oon Inarinkaa  &\\
-                            &   & M\"{a} olen Inarinkaa  &\\
-                            &   & Min\"{a} olen Inarinkaa  &\\
-                            &   & M\"{a} oon Inarin kanssa  &\\
-                            &   & Min\"{a} oon Inarin kanssa  &\\
-                            &   & M\"{a} olen Inarin kanssa  &\\
-                            &   & Min\"{a} olen Inarin kanssa  &\\
+    \multirow{2}{*}{Step 3} &   & M\"{a} oon Tomminkaa   & apply rule 3\\
+                            &   & Min\"{a} oon Tomminkaa  &\\
+                            &   & M\"{a} olen Tomminkaa  &\\
+                            &   & Min\"{a} olen Tomminkaa  &\\
+                            &   & M\"{a} oon Tommin kanssa  &\\
+                            &   & Min\"{a} oon Tommin kanssa  &\\
+                            &   & M\"{a} olen Tommin kanssa  &\\
+                            &   & Min\"{a} olen Tommin kanssa  &\\
     \hline
-    \multirow{2}{*}{Step 4} & -4.5811 &  Min\"{a} olen Inarin kanssa & rank candidates \\
-                            & -7.8174 &  Min\"{a} oon Inarin kanssa & \\
-                            & -8.0941 &  M\"{a} olen Inarin kanssa & \\
-                            & -8.8651&  M\"{a} oon Inarin kanssa & \\
-                            & -9.2045 &  Min\"{a} olen Inarinkaa & \\
-                            & -12.4408 &  Min\"{a} oon Inarinkaa & \\
-                            & -12.7176 &  M\"{a} olen Inarinkaa & \\
-                            & -13.4885 &  M\"{a} oon Inarinkaa & \\
+    \multirow{2}{*}{Step 4} & -4.5811 &  Min\"{a} olen Tommin kanssa & rank candidates \\
+                            & -7.8174 &  Min\"{a} oon Tommin kanssa & \\
+                            & -8.0941 &  M\"{a} olen Tommin kanssa & \\
+                            & -8.8651&  M\"{a} oon Tommin kanssa & \\
+                            & -9.2045 &  Min\"{a} olen Tomminkaa & \\
+                            & -12.4408 &  Min\"{a} oon Tomminkaa & \\
+                            & -12.7176 &  M\"{a} olen Tomminkaa & \\
+                            & -13.4885 &  M\"{a} oon Tomminkaa & \\
 
     \hline
-    \textbf{Output:} & \multicolumn{3}{l|}{Min\"{a} olen Inarin kanssa} \\
+    \textbf{Output:} & \multicolumn{3}{l|}{Min\"{a} olen Tommin kanssa} \\
     \hline
@@
 
-Tables~\ref{table:results-norm} and \ref{table:results-trad} presents the results of the experiments.
+Tables~\ref{table:results-norm} and \ref{table:results-trad} present the results of the experiments.
 

### 2015-01-26 finnormal-nodalida-2015/finnormal2015.tex
future work
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58598 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
+Although a reasonable size for a test corpus, the corpus is still too small for 
+wide-coverage experiments. We intend to expand the corpus size to at least 10,000
+words. Another weakness of the corpus is that it contains text from a single author,
+we would ideally like to add texts from other authors and other colloquial genres. Although
+the challenge here is finding text that is both free of privacy issues and available
+to release under a free/open-source licence.
+
+As for the methods, we would like to follow \newcite{scannell2014} in incorporating
+`translation' probabilities into our normalisation model. Our current model relies 
+exclusively on the target-language model probability, however some rules may be more
+reliable or probable than others.
+
 \section{Conclusions}

### 2015-01-26 finnormal-nodalida-2015/finnormal2015.tex
removed for review
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58599 72bbbca6-d526-0410-a7d9-f06f51895060
add source
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58600 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 There are a number of areas of research related to the task of text normalisation. Text
-proofing tools, such as spelling and grammar checkers CITEXXX can be used to encourage adherence
+proofing tools, such as spelling and grammar checkers \cite{kukich1992} can be used to encourage adherence
 to particular orthographic or grammatical norms. Accent and diacritic restoration --- for example in \newcite{scannell2011} --- is similar
@@
 
-Joonas Kylm\"{a}l\"{a}
+\textit{Removed for review}
+%Joonas Kylm\"{a}l\"{a}
 
@@
 \bibliographystyle{acl}
-\bibliography{finnormal2015.bib}
+\bibliography{finnormal2015}
 

### 2015-01-26 finnormal-nodalida-2015/finnormal2015.tex
small cleanup
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58601 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-\newcite{viinikka2013} describe the common meaning of the terms colloquial (\emph{puhekieli}) and standard (\emph{yleiskieli} or \emph{kirjakieli}) Finnish: standard language is regulated and is unified in morphology and vocabulary; colloquial language shows local and idiolectal variation, and has structures that are characteristic to spoken variety,  such as discourse particles and incomplete clauses.
+\newcite{viinikka2013} describe the common meaning of the terms colloquial (\emph{puhekieli}) and standard (\emph{yleiskieli} or \emph{kirjakieli}) Finnish: standard language is unified in morphology and vocabulary, following the regulations of a language board; colloquial language shows local and idiolectal variation, and has structures that are characteristic to spoken variety,  such as discourse particles and incomplete clauses.
 
-Colloquial Finnish differs from standard Finnish in some key features:
+% Colloquial Finnish differs from standard Finnish in some key features:
 
-Morphology shows dialectal variation, shortening and merging of morphemes, e.g. \emph{en min\"{a}} `not I' $\rightarrow$ \emph{emm\"{a}} (`I don't').
+% Morphology shows dialectal variation, shortening and merging of morphemes, e.g. \emph{en min\"{a}} `not I' $\rightarrow$ \emph{emm\"{a}} (`I don't').
 
-Syntax: due to the spontaneous nature of colloquial language, syntactic structures are common for spoken language. 
-Prototypically, colloquial language is spoken and spontaneous, and standard is written, planned and structured.
+% Syntax: due to the spontaneous nature of colloquial language, syntactic structures are common for spoken language. 
+%Prototypically, colloquial language is spoken and spontaneous, and standard is written, planned and structured.
+
 We illustrate the differences with the following example from our data set.
@@
 
-For the partial version, we did not change the syntactic structure, but changed only morphology and vocabulary.
+For the partial version, we changed only morphology and vocabulary.
 On the lexical level, the word \emph{seiska} `number 7' is colloquial style, and in the standard translation it is replaced by the ordinal \emph{seitsem\"{a}s} `seventh'.

### 2015-01-26 finnormal-nodalida-2015/finnormal2015.tex
improve table
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58602 72bbbca6-d526-0410-a7d9-f06f51895060

@@
   \centering
-  \begin{tabular}{|lrl|l|}
-    \hline
-    \textbf{Rules:} & & 1. m\"{a} $\rightarrow$ min\"{a} \\
-     ~              & & 2. oon $\rightarrow$ olen \\
-     ~              & & 3. (?+)nkaa $\rightarrow$ \textbackslash1n kanssa \\
-    \hline
-  \end{tabular}
-~\\
-~\\
   \begin{tabular}{|l|r|l|l|}
@@
     \multirow{2}{*}{Step 1} & & M\"{a} oon Tomminkaa & apply rule 1 \\
-                            & & Min\"{a} oon Tomminkaa & \\
+                            & & Min\"{a} oon Tomminkaa & m\"{a} $\rightarrow$ min\"{a} \\
     \hline
     \multirow{2}{*}{Step 2} & & M\"{a} oon Tomminkaa & apply rule 2 \\
-                            &   & Min\"{a} oon Tomminkaa & \\
+                            &   & Min\"{a} oon Tomminkaa & oon $\rightarrow$ olen \\
                             &   & M\"{a} olen Tomminkaa & \\
@@
     \multirow{2}{*}{Step 3} &   & M\"{a} oon Tomminkaa   & apply rule 3\\
-                            &   & Min\"{a} oon Tomminkaa  &\\
+                            &   & Min\"{a} oon Tomminkaa  & (?+)nkaa $\rightarrow$ \textbackslash1n kanssa \\
                             &   & M\"{a} olen Tomminkaa  &\\

### 2015-01-26 finnormal-nodalida-2015/finnormal2015.tex
add more stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58603 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-This paper evaluates a number of methods for converting colloquial Finnish to standard Finnish and describes a parallel corpus for evaluation.
+This paper evaluates a number of methods for converting colloquial Finnish to 
+standard Finnish and describes a parallel corpus for evaluation.
 
@@
 \begin{itemize}
-  \item 273 rules from \newcite{karlsson2008}'s grammar of Finnish (\S95--97). The 
+  \item \texttt{rules-1}: 273 rules from \newcite{karlsson2008}'s grammar of Finnish (\S95--97). The 
     rules took around one hour to implement.
-  \item 98 rules written by examining the development corpus, these rules also
+  \item \texttt{rules-2}: 98 rules written by examining the development corpus, these rules also
     took approximately one hour to implement.
@@
     \hline
-    \multirow{2}{*}{Step 1} & & M\"{a} oon Tomminkaa & apply rule 1 \\
+    \multirow{2}{*}{Step 1} & & M\"{a} oon Tomminkaa & apply rule 1: \\
                             & & Min\"{a} oon Tomminkaa & m\"{a} $\rightarrow$ min\"{a} \\
     \hline
-    \multirow{2}{*}{Step 2} & & M\"{a} oon Tomminkaa & apply rule 2 \\
+    \multirow{2}{*}{Step 2} & & M\"{a} oon Tomminkaa & apply rule 2: \\
                             &   & Min\"{a} oon Tomminkaa & oon $\rightarrow$ olen \\
@@
     \hline
-    \multirow{2}{*}{Step 3} &   & M\"{a} oon Tomminkaa   & apply rule 3\\
+    \multirow{2}{*}{Step 3} &   & M\"{a} oon Tomminkaa   & apply rule 3:\\
                             &   & Min\"{a} oon Tomminkaa  & (?+)nkaa $\rightarrow$ \textbackslash1n kanssa \\
@@
 
-We trained models based on two approaches, the first being phrase-based machine translation
-\cite{zens2002} and the second on character-based machine translation \cite{nakov2012,TiedemannEAMT2009}.
+We trained models based on two approaches, the first being phrase-based machine translation (PBMT, \newcite{zens2002}) and the second on character-based machine translation (CBMT, \newcite{nakov2012,TiedemannEAMT2009}).
 
@@
 involving machine translation.
-To evaluate this, we took the test corpus and translated it to English.
-
-We then translated the corpus with each of the best performing methods of each class, and 
-calculated the {\sc bleu} score.
+To evaluate this, we took the colloquial portion of the test corpus and manually 
+translated it to English. For each of the best-performing systems we first passed
+the colloquial text through, and then translated the output to English using
+a widely-used online machine translation engine with Finnish to English. We compared
+the output to translating the text to English without the standardiser.
 
@@
      rules-1 & 38.27 & 41.19 & 32.65 \\
-     rules-2 & 38.17 & 35.25 & 36.41 \\
+     rules-2 & \textbf{38.17} & \textbf{35.25} & \textbf{36.41} \\
      rules-c & 36.56 & 39.68 & 34.68 \\
@@
      CBMT-cs & 46.22 & 52.27 & 29.21 \\
-     PBMT-cn & 28.05 & 35.42 & 48.37 \\
+     PBMT-cn & \textbf{28.05} & \textbf{35.42} & \textbf{48.37} \\
      PBMT-cs & 27.95 & 36.13 & 46.76 \\
@@
   \end{tabular}
-  \caption{Results for the normalisation task}
+  \caption{Results for the normalisation task. The system \texttt{rules-c} is the combination
+     of the rules in \texttt{rules-1} and \texttt{rules-2}. The figures in bold are the 
+     best results for unsupervised and supervised methods.}
   \label{table:results-norm}
@@
 Tables~\ref{table:results-norm} and \ref{table:results-trad} present the results of the experiments.
+
+% unsupervised does as well as CBMT, and and half the way towards pbmt
+
+% supervised systems trained on the normalised corpus outperform those trained on the standardised corpus (WHY?)
+
+% the rules written from the development portion of the corpus outperform the rules from the grammar
 
@@
      \hline
-     Colloquial & - & - & 11.95 \\
-     Normalised & - & - & 21.81 \\
-     Standardised & - & - & 17.62 \\
+     Colloquial   & 45.31 & 71.28 & 11.95 \\
+     Normalised   & 34.63 & 62.34 & 21.81 \\
+     Standardised & 38.67 & 63.35 & 17.62 \\
      \hline
-     rules-2 & - & - & 12.28 \\
-     cbmt-cn & - & - & 12.35 \\
-     pbmt-cn & - & - & 17.90 \\
+     rules-2      & 42.71 & 70.42 & 12.28 \\
+     \hline
+     cbmt-cn      & 44.44 & 74.17 & 12.35 \\
+     pbmt-cn      & 38.38 & 66.81 & 17.90 \\
      \hline
   \end{tabular}
-  \caption{Results for the translation task}
+  \caption{Results for the translation task. The first three rows are the results 
+    from translating the sections of the corpus. As with the }
   \label{table:results-trad}
 \end{table}
+
+% example translations
+
+%colloq: ooh what onks there snow
+%pbmt: ooh what if there's snow
+%ref: ooh what, is there snow?
 

### 2015-01-26 finnormal-nodalida-2015/finnormal2015.tex
updates
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58604 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{multirow}
+\usepackage[small,bf]{caption}
 \usepackage{latexsym}
@@
 
-We trained models based on two approaches, the first being phrase-based machine translation (PBMT, \newcite{zens2002}) and the second on character-based machine translation (CBMT, \newcite{nakov2012,TiedemannEAMT2009}).
+We trained models based on two approaches, the first being phrase-based machine translation (PBMT, \newcite{zens2002}) and the second on character-based machine translation (CBMT, \newcite{nakov2012}; \newcite{TiedemannEAMT2009}).
 
@@
      cbmt-cn      & 44.44 & 74.17 & 12.35 \\
-     pbmt-cn      & 38.38 & 66.81 & 17.90 \\
+     pbmt-cn      & \textbf{38.38} & \textbf{66.81} & \textbf{17.90} \\
      \hline

### 2015-01-26 finnormal-nodalida-2015/finnormal2015.tex
some improvements
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58605 72bbbca6-d526-0410-a7d9-f06f51895060

@@
   \centering
-  \begin{tabular}{|l|r|l|l|}
+  \begin{tabular}{|l|l|r|l|}
     \hline
@@
     \hline
-    \multirow{2}{*}{Step 1} & & M\"{a} oon Tomminkaa & apply rule 1: \\
-                            & & Min\"{a} oon Tomminkaa & m\"{a} $\rightarrow$ min\"{a} \\
+    \multirow{2}{*}{Step 1}  & M\"{a} oon Tomminkaa & & apply rule 1: \\
+                             & Min\"{a} oon Tomminkaa & & m\"{a} $\rightarrow$ min\"{a} \\
     \hline
-    \multirow{2}{*}{Step 2} & & M\"{a} oon Tomminkaa & apply rule 2: \\
-                            &   & Min\"{a} oon Tomminkaa & oon $\rightarrow$ olen \\
-                            &   & M\"{a} olen Tomminkaa & \\
-                            &   & Min\"{a} olen Tomminkaa & \\
+    \multirow{2}{*}{Step 2}  & M\"{a} oon Tomminkaa & & apply rule 2: \\
+                             & Min\"{a} oon Tomminkaa & & oon $\rightarrow$ olen \\
+                             & M\"{a} olen Tomminkaa & & \\
+                             & Min\"{a} olen Tomminkaa & & \\
     \hline
-    \multirow{2}{*}{Step 3} &   & M\"{a} oon Tomminkaa   & apply rule 3:\\
-                            &   & Min\"{a} oon Tomminkaa  & (?+)nkaa $\rightarrow$ \textbackslash1n kanssa \\
-                            &   & M\"{a} olen Tomminkaa  &\\
-                            &   & Min\"{a} olen Tomminkaa  &\\
-                            &   & M\"{a} oon Tommin kanssa  &\\
-                            &   & Min\"{a} oon Tommin kanssa  &\\
-                            &   & M\"{a} olen Tommin kanssa  &\\
-                            &   & Min\"{a} olen Tommin kanssa  &\\
+    \multirow{2}{*}{Step 3}  & M\"{a} oon Tomminkaa   & & apply rule 3:\\
+                             & Min\"{a} oon Tomminkaa &  & (?+)nkaa $\rightarrow$ \textbackslash1n kanssa \\
+                             & M\"{a} olen Tomminkaa  & &\\
+                             & Min\"{a} olen Tomminkaa  & &\\
+                             & M\"{a} oon Tommin kanssa  & &\\
+                             & Min\"{a} oon Tommin kanssa  & &\\
+                             & M\"{a} olen Tommin kanssa  & &\\
+                             & Min\"{a} olen Tommin kanssa  & &\\
     \hline
-    \multirow{2}{*}{Step 4} & -4.5811 &  Min\"{a} olen Tommin kanssa & rank candidates \\
-                            & -7.8174 &  Min\"{a} oon Tommin kanssa & \\
-                            & -8.0941 &  M\"{a} olen Tommin kanssa & \\
-                            & -8.8651&  M\"{a} oon Tommin kanssa & \\
-                            & -9.2045 &  Min\"{a} olen Tomminkaa & \\
-                            & -12.4408 &  Min\"{a} oon Tomminkaa & \\
-                            & -12.7176 &  M\"{a} olen Tomminkaa & \\
-                            & -13.4885 &  M\"{a} oon Tomminkaa & \\
+    \multirow{2}{*}{Step 4} &  Min\"{a} olen Tommin kanssa & -4.5811 & rank candidates \\
+                            &  Min\"{a} oon Tommin kanssa & -7.8174 & \\
+                            &  M\"{a} olen Tommin kanssa & -8.0941 & \\
+                            &  M\"{a} oon Tommin kanssa & -8.8651 & \\
+                            &  Min\"{a} olen Tomminkaa & -9.2045 & \\
+                            &  Min\"{a} oon Tomminkaa & -12.4408 & \\
+                            &  M\"{a} olen Tomminkaa & -12.7176 & \\
+                            &  M\"{a} oon Tomminkaa & -13.4885 & \\
 
@@
 
-% unsupervised does as well as CBMT, and and half the way towards pbmt
+The baseline was made by calculating the metric scores between the standardised `reference' 
+and the colloquial input. The results show that all conversion methods outperform the baseline.
+
+Our unsupervised method performs similarly to the character-based machine translation systems. Both 
+the character-based systems and the unsupervised system achieve around half of the performance of the 
+supervised phrase-based system.
+
+Out of the unsupervised systems, the set of rules which was created by examining the development corpus
+outperforms both the set of rules from the grammar and the combined rules.
 
 % supervised systems trained on the normalised corpus outperform those trained on the standardised corpus (WHY?)
-
-% the rules written from the development portion of the corpus outperform the rules from the grammar
 
@@
 
+%! ooh mitä onks siellä lunta
+%ref: ooh what, is there snow?
 %colloq: ooh what onks there snow
+%book: what! Is there snow?
 %pbmt: ooh what if there's snow
-%ref: ooh what, is there snow?
+%rules: ooh what if there's snow
+
+%! mä vaa protestoin et ei siinä mitää tavurajaa oo i:den välissä
+%ref: I was just protesting that there's no syllable break between the two i's
+%colloq: I VAA you do not protest it makes no tavurajaa oo i: in between
+%book: I just protest, that's not anything tavurajaa no i in between
+%pbmt: I only protest that there is no tavurajaa in between the i-den
+%rules: I VAA you do not protest it makes no tavurajaa no i in between
+
+%! (kyse oli siis vaan jostain 3 rivistä loopin sisällä)
+% ref: (it was only like 3 lines inside a loop)
+% colloq: (it was, therefore, but for some three lines within the loop)
+% book: (it was, therefore, only three to a row within a loop)
+% pbmt: (it was, therefore, only some three-line within the loop)
+% rules: (it was Siisi only some 3 lines of the loop inside)
 

### 2015-01-26 finnormal-nodalida-2015/finnormal2015.tex
more fixes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58606 72bbbca6-d526-0410-a7d9-f06f51895060

@@
     \multirow{2}{*}{Step 1}  & M\"{a} oon Tomminkaa & & apply rule 1: \\
-                             & Min\"{a} oon Tomminkaa & & m\"{a} $\rightarrow$ min\"{a} \\
+                             & Min\"{a} oon Tomminkaa & & ~~~m\"{a} $\rightarrow$ min\"{a} \\
     \hline
     \multirow{2}{*}{Step 2}  & M\"{a} oon Tomminkaa & & apply rule 2: \\
-                             & Min\"{a} oon Tomminkaa & & oon $\rightarrow$ olen \\
+                             & Min\"{a} oon Tomminkaa & & ~~~oon $\rightarrow$ olen \\
                              & M\"{a} olen Tomminkaa & & \\
@@
     \multirow{2}{*}{Step 3}  & M\"{a} oon Tomminkaa   & & apply rule 3:\\
-                             & Min\"{a} oon Tomminkaa &  & (?+)nkaa $\rightarrow$ \textbackslash1n kanssa \\
+                             & Min\"{a} oon Tomminkaa &  & ~~~(?+)nkaa $\rightarrow$ \textbackslash1n kanssa \\
                              & M\"{a} olen Tomminkaa  & &\\
@@
 
-The target-language model corpus used was the same as in the unsupervised experiments.
+The target-language model corpus, trained using KenLM \cite{heafield2011}, used was the same as in the unsupervised experiments.
 

### 2015-01-26 finnormal-nodalida-2015/finnormal2015.tex
partial -> normalised
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58607 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-For the partial version, we changed only morphology and vocabulary.
+For the normalised version, we changed only morphology and vocabulary.
 On the lexical level, the word \emph{seiska} `number 7' is colloquial style, and in the standard translation it is replaced by the ordinal \emph{seitsem\"{a}s} `seventh'.
-Other changes in the partial version are common morphological or phonological, phenomena, such as reducing the diphthong in \emph{vain} $\rightarrow$ \emph{vaan} or shortening in  \emph{ett\"{a}} $\rightarrow$ \emph{et}.
-The original sentence and the partial translation are shown below, aligned word by word.
+Other changes in the normalised version are common morphological or phonological, phenomena, such as reducing the diphthong in \emph{vain} $\rightarrow$ \emph{vaan} or shortening in  \emph{ett\"{a}} $\rightarrow$ \emph{et}.
+The original sentence and the normalised translation are shown below, aligned word by word.
 

### 2015-01-26 finnormal-nodalida-2015/finnormal2015.tex
fix
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58608 72bbbca6-d526-0410-a7d9-f06f51895060
updates
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58609 72bbbca6-d526-0410-a7d9-f06f51895060

@@
  The syntactic structure of the original sentence is markedly spoken; the word \emph{seiska} is topicalised, and the main information ``deduced from phonotactics'' is in a subordinate clause. 
-The translation into  is shorter and more precise, leaving just the main information.
+The translation into standard Finnish is shorter and more precise, leaving just the main information.
 
@@
 Moses toolkit \cite{koehn2007}. The training set up was that used for
-the baseline system in the  in the WMT shared tasks
+the baseline system in the WMT shared tasks
 on machine translation.\footnote{\url{http://www.statmt.org/wmt11/baseline.html}}
@@
   \end{tabular}
-  \caption{Results for the translation task. The first three rows are the results 
+  \caption{Results for the Finnish to English translation task. The first three rows are the results 
     from translating the sections of the corpus. As with the }

### 2015-01-26 finnormal-nodalida-2015/finnormal2015.tex
more stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58610 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 % supervised systems trained on the normalised corpus outperform those trained on the standardised corpus (WHY?)
+% could be because of better word alignment
 
@@
   \caption{Results for the Finnish to English translation task. The first three rows are the results 
-    from translating the sections of the corpus. As with the }
+    from translating the sections of the corpus.}
   \label{table:results-trad}
@@
 
-% example translations
+The systems for normalisation are able to improve out-of-vocabulary rates in 
+many cases, most likely as the online statistical system that we used is trained
+on more formal texts. Frequent contractions such as \emph{onks?} `is there?',
+\emph{oo} `not' and \emph{vaa} `only' are found untranslated in the output.
 
@@
 We have shown that converting from colloquial Finnish to standard Finnish substantially
-helps with the machine translation task.
-
-%The main contribution of this paper is a parallel corpus of colloquial Finnish
-%and standard Finnish
+helps with the Finnish to English machine translation task.
 

### 2015-01-26 finnormal-nodalida-2015/finnormal2015.tex
more stuffs
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58611 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-For the normalised version, we changed only morphology and vocabulary.
+For the normalised version,\footnote{The normalised version is converted orthographically and lexically, but not syntactically or stylistically.} we changed only morphology and vocabulary.
 On the lexical level, the word \emph{seiska} `number 7' is colloquial style, and in the standard translation it is replaced by the ordinal \emph{seitsem\"{a}s} `seventh'.
@@
 
-% expand test corpus -- up to 10,000 words
-% add other colloquial genres
-% apply to other NLP tasks
-% weight rules 
-%
-
 Although a reasonable size for a test corpus, the corpus is still too small for 
@@
 As for the methods, we would like to follow \newcite{scannell2014} in incorporating
-`translation' probabilities into our normalisation model. Our current model relies 
+`translation' probabilities into our unsupervised normalisation model. Our current model relies 
 exclusively on the target-language model probability, however some rules may be more

### 2015-01-26 finnormal-nodalida-2015/finnormal2015.tex
+1
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58613 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-% supervised systems trained on the normalised corpus outperform those trained on the standardised corpus (WHY?)
-% could be because of better word alignment
+It is interesting to note that supervised systems trained on the normalised section of the corpus outperform
+those trained on the standardised corpus. One explanation for this could be that the corpus size is small, so 
+that the word alignments are not as reliable on the standardised corpus which is by nature not a word-for-word 
+conversion.
 
@@
 on more formal texts. Frequent contractions such as \emph{onks?} `is there?',
-\emph{oo} `not' and \emph{vaa} `only' are found untranslated in the output.
+\emph{oo} `not' and \emph{vaa} `only' are found untranslated in the output, but 
+are easily converted by the systems.
 
 %! ooh mitä onks siellä lunta
+%! mitä! Onko siellä lunta?
 %ref: ooh what, is there snow?
@@
 %! mä vaa protestoin et ei siinä mitää tavurajaa oo i:den välissä
+%! minä vain protestoin, että ei siinä mitään tavurajaa ole i:den välissä
 %ref: I was just protesting that there's no syllable break between the two i's
@@
 %! (kyse oli siis vaan jostain 3 rivistä loopin sisällä)
+%! (kyse oli siis vain 3:sta rivistä loopin sisällä)
 % ref: (it was only like 3 lines inside a loop)

### 2015-01-26 finnormal-nodalida-2015/finnormal2015.tex
update
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58614 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 language can cause a substantial decrease in quality. For example in machine translation,
-parsing and part-of-speech tagging \cite{eisenstein2013}.
+parsing and part-of-speech tagging \cite{eisenstein2013}. Non-standard language can 
+have different orthographic conventions, along with different morphology, syntax and stylistics.
 
@@
      \hline
-     CBMT-cn & 43.09 & 48.34 & 33.55 \\
-     CBMT-cs & 46.22 & 52.27 & 29.21 \\
-     PBMT-cn & \textbf{28.05} & \textbf{35.42} & \textbf{48.37} \\
-     PBMT-cs & 27.95 & 36.13 & 46.76 \\
+     {\sc cbmt}-cn & 43.09 & 48.34 & 33.55 \\
+     {\sc cbmt}-cs & 46.22 & 52.27 & 29.21 \\
+     {\sc pbmt}-cn & \textbf{28.05} & \textbf{35.42} & \textbf{48.37} \\
+     {\sc pbmt}-cs & 27.95 & 36.13 & 46.76 \\
      \hline
@@
      \hline
-     Colloquial   & 45.31 & 71.28 & 11.95 \\
-     Normalised   & 34.63 & 62.34 & 21.81 \\
-     Standardised & 38.67 & 63.35 & 17.62 \\
+     Colloquial &  41.02  &  69.57  & 12.11 \\
+     Normalised &  30.73  &  59.73  & 22.56 \\
+     Standardised &  33.88  &  61.61  & 19.49 \\
      \hline
-     rules-2      & 42.71 & 70.42 & 12.28 \\
+     rules-2 &  37.94  &  69.35  & 12.92 \\
      \hline
-     cbmt-cn      & 44.44 & 74.17 & 12.35 \\
-     pbmt-cn      & \textbf{38.38} & \textbf{66.81} & \textbf{17.90} \\
+     {\sc cbmt}-cn &  65.59  &  86.25  & 13.06 \\
+     {\sc pbmt}-cn &  \textbf{35.69}  &  \textbf{65.14}  & \textbf{17.75} \\
      \hline

### 2015-02-02 2015-eamt-melsa/2015-eamt-melsa.tex
add dir
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58643 72bbbca6-d526-0410-a7d9-f06f51895060

%
% File eamt15.tex, an adaptation of eamt14.tex (which was a copy of
% eamt12.tex)
%
% Contact: eamt2015@dlsi.ua.es

%%% To ease future customizations, various replaceables have been paramaterized
%%% as listed in the newcommands section

\documentclass[11pt]{article}
\usepackage{eamt15}
\usepackage{times}
\usepackage{latexsym}
\setlength\titlebox{6.5cm}    % Expanding the titlebox
%%% YOUR PACKAGES BELOW THIS LINE %%%

\newcommand{\confname}{EAMT 2015}
\newcommand{\website}{http://www.eamt2015.org/}
\newcommand{\contactname}{the conference chairs (Felipe
  S\'anchez-Mart\'inez, Gema Ram\'irez-S\'anchez and Fred Hollowood)}
\newcommand{\contactemail}{eamt2015@dlsi.ua.es}
\newcommand{\conffilename}{eamt15}
\newcommand{\downloadsite}{http://www.eamt2015.org/}
\newcommand{\paperlength}{$8$ (eight)}
\newcommand{\shortpaperlength}{$4$ (four)}

\title{Instructions for \confname\ Proceedings}

\author{First Author\\
  Affiliation / Address line 1\\
  Affiliation / Address line 2\\
  Affiliation / Address line 3\\
  {\tt email@domain}  \And
  Second Author\\
  Affiliation / Address line 1\\
  Affiliation / Address line 2\\
  Affiliation / Address line 3\\
  {\tt email@domain}}

\date{}

\begin{document}
\maketitle
\begin{abstract}
   ... 
\end{abstract}

\section{Introduction}

\cite{koehn2005}

\bibliographystyle{apalike}

\bibliography{2015-eamt-melsa}

\end{document}

### 2015-02-02 2013-lrec-turkspell/turk-spell-lrec2013.tex
move name
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58646 72bbbca6-d526-0410-a7d9-f06f51895060

\documentclass[a5paper]{article}

\usepackage{amssymb}
\usepackage{polyglossia}
\usepackage[shortcuts]{extdash}

\usepackage{fontspec}
\usepackage{xunicode}
\usepackage{xltxtra}

\usepackage{url}
\usepackage{hyperref}
\usepackage[obeyDraft]{todonotes}

\usepackage{graphicx}
\usepackage{geometry}

\setmainfont[Mapping=tex-text]{Liberation Serif}

\title{Finite-State Spell-Checking for Turkic Minority Languages}
\author{Francis M. Tyers \and Jonathan North Washington \and Ilnar Salimzyanov \and Tommi A. Pirinen}

\date{\today}

\begin{document}

\maketitle

\begin{abstract}
    In this article there's spell-chekcers and Turkic languages
\end{abstract}

\section{Introduction}
\label{sec:introduction}

Spell-checking is one of the most commonly used and most basic pieces of
software for language support. It is used by all platforms from regular
computers to tablets and mobile phones, and majority of software from office
suites to web browsers. Furthermore, the role of spell-checking softwares has
been increasing in modern platforms, as limited input systems require more
assistance from spell-checker to predict users input than traditional keyboard
inputs. The contemporary systems for spell-checking are mostly based on
word-form lists and statistics, and some basic form of suffix stripping and
word-form compounding. For language modeling in general there are well known
and more established models for use in morphologically complex languages. In
this article we use well-established finite-state spell-checking
technology~\cite{pirinen2012compiling} in conjunction with Turkic minority
languages and compare the results against traditional word-form list
approaches.

\section{Methods}

Finite-state spell-checking~\cite{agata2002typographical} is well-established
method for handling spell-checking of morphologically complexer languages.
Specifically weighted finite-state spell-checking has been studied and extended
in last years~\cite{pirinen2010finitestate}. In this article we use those
methods to build finite-state spell-checkers for Türkich languages. The
extension of finite-state spell-checkers by hunspell error models is done
as described iñ~\cite{pirinen2010building}.

\section{Data}

For the language models we have selected freely available open source dictionaries
from Apertium project\footnote{\url{}}, and translated them to finite-state
spell-checking systems as described in~\cite{pirinen2012compiling}. The
errors have been discovered from Wikipedia using methods described 
in~\cite{max2010mining}. The hunspell data has been acquired from the internet.

\section{Evaluation}

To evaluate our finite-state spell-checkers we have applied them and hunspell
to our hand-picked error corpus and measured the positions of correct
corrections as shown in table~\ref{table:precision}

We have also briefly measured speed and memory usage of the spell-checkers when
applied to large sections of running text containing both correctly written
strings and mistakes. The results are in table~\ref{table:efficiency}

\section{Discussion}

We have shown that spell-checking works nicely.

\subsection{Error Analysis}

It is interesting that... because edit distance of 2 does not work very well
for Turkic languages.

\subsection{Future Work}

More betterer error models should be built.

\section{Conclusion}

In this article we studied the use of finite-state spell-checking as an
alternative to other contemporary spell-checking methods for turkic langauges.
we have shown that it works better and is faster.

\bibliographystyle{unsrt}
\bibliography{turkspell}

\end{document}
% vim: set spell:

### 2015-02-02 2015-nodalida-finnormal/finnormal2015.tex
move dir
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58647 72bbbca6-d526-0410-a7d9-f06f51895060

--- 2015-01-26 finnormal-nodalida-2015/finnormal2015.tex
@@
 on more formal texts. Frequent contractions such as \emph{onks?} `is there?',
-\emph{oo} `not' and \emph{vaa} `only' are found untranslated in the output, but 
+\emph{oo} `be.{\sc conneg}'\footnote{The word \emph{oo} is the negative form of the verb \emph{olla} `to be' in Finnish.} and \emph{vaa} `only' are found untranslated in the output, but 
 are easily converted by the systems.

### 2015-02-03 2015-nodalida-ruscg/ruscg.tex
new paper
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58651 72bbbca6-d526-0410-a7d9-f06f51895060

--- 2015-01-21 rusaccent-nodalida-2015/rusaccent.tex
@@
 
-\title{Automatic word stress annotation of Russian running text (max 8 pages)}
+\title{A preliminary constraint grammar for Russian}
 
@@
 \begin{abstract}
-  We evaluate the effectiveness of finite-state tools we developed for 
-  automatically annotating word stress in Russian running text. To our knowledge,
-  this is the first study to empirically evaluate the results of this task. 
-  Given an adequate 
-  lexicon with specified stress, the primary obstacle for correct stress 
-  placement is disambiguating homographic wordforms. Using a Constraint Grammar, 
-  we achieve 93.21\% accuracy, a 3.14\% improvement over the baseline
-  undisambiguated performance. 
-  Combining our Constraint Grammar with frequency data and a simple
-  algorithm for unknown word stress position guessing, we achieve 96.15\%
-  accuracy. These results highlight the need morphosyntactic disambiguation
-  in the wordstress placement task for Russian, and set a standard for future
-  research on this task.
+  Hargle bargle
 \end{abstract}
-
-%\section{Credits}
-%
-%The template had a credits section for the history of the template, but we don't need
-%a Credits section, right?
 
@@
 
-Russian word stress is variable and phonemic. The inflecting word classes include complex 
-patterns of shifting stress, and a lexeme's stress pattern cannot be predicted 
-from surface forms. Lexical stress and its attendant vowel reduction are 
-a prominent feature of spoken Russian; the incorrect placement of stress
-can render speech almost incomprehensible. However, standard written Russian does not 
-typically mark word stress.\footnote{Texts intended for native speakers sometimes 
-mark stress on words that cannot be disambiguated through context. Theoretically, 
-a perfect wordstress placement method could help an author identify
-tokens which should be stressed: any token that cannot
-be disambiguated by syntactic or semantic means should be marked for stress.} 
-Without information about lexical stress position, 
-correctly converting written Russian text to speech is impossible. This is true 
-both for humans (e.g. foreign language students) and computers (e.g. text-to-
-speech).
-
-We identify three different types of word stress ambiguity.
-\emph{Intraparadigmatic} stress ambiguity refers
-to homographic wordforms belonging to the same lexeme, as shown in 
-\ref{ex:intrahom}.
-
-\needspace{4\baselineskip}
-\ex. Intraparadigmatic homographs \label{ex:intrahom}
-\a. \rus{т\'{е}ла} \emph{t\'{e}la} `body.\textsubscript{SG-GEN}' 
-    \label{ex:bodySGGEN}
-\b. \rus{тел\'{а}} \emph{tel\'{a}} `body.\textsubscript{PL-NOM}' 
-    \label{ex:bodyPLNOM}
-
-The remaining two types of stress ambiguity occur between lexemes. 
-\emph{Morphosyntactically congruent} stress ambiguity occurs between homographs 
-that belong to separate lexemes, 
-and whose morphosyntactic values are identical, as shown in
-Table~\ref{table:MScongruent}. This kind of stress ambiguity is relatively rare.
-Resolving this ambiguity would require the use of technologies such as Word Sense 
-Disambiguation.
-
-\begin{table}
-\begin{center}
-\begin{tabular}{c|ll}
-\bf Number-Case & \bf `castle' & \bf `lock' \\
-\hline
-\texttt{SG-NOM} & \rus{з\'{a}мок} & \rus{зам\'{о}к} \\
-\texttt{SG-GEN} & \rus{з\'{a}мка} & \rus{замк\'{а}} \\
-\texttt{PL-NOM} & \rus{з\'{a}мки} & \rus{замк\'{и}} \\
-\texttt{PL-GEN} & \rus{з\'{a}мков} & \rus{замк\'{о}в} \\
-... & \hskip 1em ... & \hskip 1em ... \\
-\end{tabular}
-\end{center}
-\caption{ \label{table:MScongruent} Morphosyntactically congruent homographs. }
-\end{table}
-
-\emph{Morphosyntactically incongruent} stress ambiguity occurs between homographs
-that belong to separate lexemes, and whose morphosyntactic values are different, 
-as shown in \ref{ex:MSincongruent}.
-
-\ex. Morphosyntactically incongruent homographs \label{ex:MSincongruent}
-\a. \rus{попад\'{а}ть} \emph{popadát'} `fall.\textsubscript{INF-IPFV}' \\
-    \rus{поп\'{а}дать} \emph{popádat'} `fall.\textsubscript{INF-PFV}'
-\b. \rus{узнаёт} \emph{uznaët} `finds out.\textsubscript{PRS-IPFV-3SG}' \\
-    \rus{узн\'{а}ет} \emph{uznáet} `finds out.\textsubscript{FUT-PFV-3SG}'
-\c. \rus{н\'{а}шей} \emph{nášej} `our.\textsubscript{F-SG-GEN/DAT/LOC...}'\\
-    \rus{наш\'{е}й} \emph{našéj} `sew on.\textsubscript{IMP-2SG}'
-\d. \rus{дор\'{о}га} \emph{doróga} `road.\textsubscript{N-F-SG-NOM}'\\
-    \rus{дорог\'{а}} \emph{dorogá} `dear.\textsubscript{ADJ-F-SG-PRED}'
-
-It should be noted that not all morphosyntactic ambiguity leads to stress 
-ambiguity. For example, all inflecting modifiers have the same form for F-SG-GEN, 
-F-SG-LOC, F-SG-DAT and F-SG-INS. We refer to this as \emph{stress-irrelevant} 
-ambiguity, since all readings have
-the same stress placement. Stress-irrelevant ambiguity can be intraparadigmatic 
-and/or interlexical.
-
-In the case of running text in Russian, most stress placement 
-ambiguity is rooted in intraparadigmatic and morphosyntactically incongruent 
-ambiguity. Detailed part-of-speech tagging can help determine the stress of 
-these forms, since each alternative stress placement is tied to a different
-tag sequence. In this study we focus on 
-the role of part-of-speech tagging in improving
-automatic stress placement. We leave morphosyntactically congruent stress 
-ambiguity to future work because of a combination of two reasons, 1) it is by far the least common type of stress ambiguity, and 2) disambiguating morphosyntactically congruent stress requires word sense disambiguation, a fundamentally different technology from the other approaches of this study.
-
-\subsection{Background and Task Definition} \label{background}
-
-Automatic stress placement in Russian is similar to Diacritic Restoration, a task
-which has received increasing interest over the last 20 years. Generally 
-speaking, diacritics disambiguate otherwise homographic wordforms, so missing 
-diacritics can complicate many NLP tasks, such as text-to-speech. For example, 
-speakers of Czech may type emails and other communications
-without standard diacritics. In order to generate speech from these texts, they 
-must first be normalized by restoring diacritics. A slightly different
-situation arises with Arabic and Hebrew, since standard orthography lacks vowel 
-diacritics altogether. For such languages, the `restoration' of vowel diacritics 
-results in less ambiguity than in standard orthography. For languages with 
-inherently ambiguous orthography,
-it may be more precise to refer to this as `diacritic enhancement', since it 
-produces text that is less ambiguous than the standard language. In this sense, 
-Russian orthography is similar to Arabic and Hebrew, since its vowel qualities 
-are underspecified in standard orthography.
-
-Many studies of Russian text-to-speech and automatic speech recognition make note of the 
-difficulties caused by the shortcomings of their stress-marking resources (e.g. \cite{krivnova_1998}), but very few studies have targeted stress-marking itself, and they tend to
-focus on placing stress on unknown words, with almost no attention to the problems that
-arise with known words. For example, 
-\cite{Xomicevic_2008} developed a set of heuristics for guessing stress placement on 
-unknown words in Russian. More recently, \cite{hall_sproat_russian_2013} trained 
-a maximum entropy model on a dictionary of Russian words, and evaluated on
-wordlists containing `known' and `unknown' wordforms.\footnote{\newcite{hall_sproat_russian_2013} randomly selected their training
-and test data from a list of wordforms, and so a number of lexemes had wordforms in both the training and test data.
-Wordforms in the test data whose sibling wordforms from the same lexeme were in the training set were categorised as
-`known' wordforms.} Their model 
-achieved 98.7\% accuracy on known words, and 83.9\% accuracy on unknown words. 
-One major shortcoming of evaluating on wordlists is 
-that they do not reflect wordform frequency.
-Many of the most problematic stress ambiguities in Russian occur in 
-high-frequency wordforms, so evaluations based on wordlists - as opposed to 
-running text - have inflated results (see discussion in 
-Section~\ref{discussion}). Furthermore, working with running text allows for the
-possibility of disambiguating homographs based on syntactic context.
-
-So far, the implicit target application of the few studies related to automatic 
-stress placement in Russian has been text-to-speech and automatic speech 
-recognition. However, the target application of our stress
-annotator is a different domain: language learning. Since standard Russian does 
-not contain stress-markings, learners are frequently unable to pronounce unknown 
-words correctly without referencing a dictionary or similar resources. In the 
-context of language learning, marking stress incorrectly is arguably worse than 
-not marking it at all. Because of this, we want our stress annotator to be able 
-to abstain from marking stress on words that it is unable to resolve with high 
-confidence.
-
-\begin{table*}[t]
-  \centering
-  \begin{tabular}{r|c|c}
-    & \rus{кость}-N-F-SG-GEN\hskip 1em\rus{к\'{о}сти} & \rus{кость}-N-F-SG-GEN\hskip 1em\rus{к\'{о}сти} \\
-    & \rus{кость}-N-F-SG-DAT\hskip 1em\rus{к\'{о}сти} & \rus{кость}-N-F-SG-DAT\hskip 1em\rus{к\'{о}сти} \\
-    & \rus{костить}-V-IPFV-IMP\hskip 1em\rus{кост\'{и}} & \\
-    \hline
-    {\small {\tt bare}} & \rus{кости} & \rus{кости} \\
-    {\small {\tt safe}} & \rus{кости} & \rus{к\'{о}сти} \\
-    {\small {\tt guess}} & \rus{к\'{о}сти} ($p$=0.67) or \rus{кост\'{и}} ($p$=0.33) & \rus{к\'{о}сти}\\
-    {\small {\tt guessFreq}} & \rus{к\'{о}сти} & \rus{к\'{о}сти}
-  \end{tabular}
-  \caption{Example output of each stress placement approach, given a particular set of readings for the token \rus{кости} \emph{kosti}.}
-  \label{tab:conditions}
-\end{table*}
-
-\subsection{Stress Corpus}
-
-Russian texts with marked word stress are relatively rare, except in materials
-for language learners, which are predominantly proprietary. Our corpus was 
-collected from
-free texts on Russian language-learning websites.\footnote{URLs can be found in
-the X data repository (URL will be added in the final version).} This small corpus (N=7689) is 
-representative of texts 
-that learners of Russian are likely to encounter in their studies. These texts 
-include excerpts from well-known literary works, as well as dialogs and 
-prose and individual sentences that were written for learners.
-
-Unfortunately, the general practice for marking stress in Russian is to
-\emph{not} mark stress on monosyllabic tokens, effectively assuming that
-all monosyllabics are stressed. However, this approach 
-is not well-motivated. Many words -- both monosyllabic and multisyllabic -- are 
-unstressed, especially among prepositions, conjunctions, and particles.
-% multisyllabics that can be unstressed: \rus{через, между} 
-Furthermore, there are many
-high-frequency monosyllabic homographs that can be either stressed or unstressed, 
-depending on their part of speech, or particular collocations. Clearly, for
-such words, one cannot assume that they are stressed on the basis of their 
-syllable count.
-
-Based on these considerations, we built our tools to mark stress on every
-word, both monosyllabic and multisyllabic. However, because our gold corpus texts 
-do not mark stress on monosyllabic words, we cannot evaluate our annotation of 
-those words.
-
-\section{Automatic Stress Placement}
-
-State-of-the-art morphological analysis in Russian is based on finite-state
-technology.\footnote{add footnote or citations for Dialog, mystem, etc.} To our 
-knowledge, no open-source resources are available for analyzing and generating 
-stressed wordforms. Therefore, we developed free and open-source finite-state 
-tools capable of analyzing and generating stressed wordforms, based on the
-well-known \emph{Grammatical Dictionary of Russian} \cite{Zaliznjak-77}.
-A Finite-State Transducer\footnote{Using two-level 
-morphology, xfst/hfst, blah, blah, blah} (FST) generates all possible 
-morphosyntactic readings of each wordform, and a constraint grammar\footnote{vislcg3, blah blah} then removes
-some extra readings based on syntactic context. The ultimate success of our 
-stress placement system depends on the performance of the Constraint Grammar. 
-Ideally, the Constraint Grammar would successfully remove all but the correct 
-reading for each token, but in practice some
-tokens still have more than one reading remaining. Therefore,
-we also evaluate various approaches to deal with the remaining ambiguity,
-as described in the following paragraph and Table~\ref{tab:conditions}.
-
-The {\small {\tt bare}} approach is to not mark stress on words with more than 
-one reading. Since both sets of readings in Table~\ref{tab:conditions} have more
-than one reading, {\small {\tt bare}} does not output a stressed form.
-
-The {\small {\tt safe}} approach is to mark stress only on
-tokens whose morphosyntactic ambiguity is stress-irrelevant. In 
-Table~\ref{tab:conditions}, the first column has readings that result in two
-different stress positions, so {\small {\tt safe}} does not output a stressed
-form. However, in the second column, both readings have the same stress position, 
-so {\small {\tt safe}} outputs that stress position.
-
-The {\small {\tt guess}} approach is to randomly select one of the available 
-readings, and abstain if no readings are available. In the first column of
-Table~\ref{tab:conditions}, a random selection means that \rus{к\'{о}сти} 
-\emph{k\'{o}sti} is twice as likely as \rus{кост\'{и}} \emph{kost\'{i}}. The
-second column of Table~\ref{tab:conditions} contains stress-irrelevant
-ambiguity, so a random selection of a reading has the same result as the
-{\small {\tt safe}} approach.
-
-The {\small {\tt guessFreq}} 
-approach is to select the reading that is most frequent, with frequency data
-taken from a separate hand-disambiguated corpus. If none of the readings is 
-found in the corpus, then {\small {\tt guessFreq}} selects the reading
-with the tag sequence (lemma removed) that is most frequent in our corpus. If the 
-tag sequence is not found
-in our frequency list, then {\small {\tt guessFreq}} backs off to the 
-{\small {\tt guess}} algorithm. In the first column of
-Table~\ref{tab:conditions}, {\small {\tt guessFreq}} selects \rus{к\'{о}сти} 
-\emph{k\'{o}sti} because the tag sequence N-F-SG-GEN is more frequent than the
-other alternatives. Note 
-that for tokens with stress-irrelevant ambiguity (e.g. the second column of
-Table~\ref{tab:conditions}), {\small {\tt guess}} and 
-{\small {\tt guessFreq}} produce the same result as the {\small {\tt safe}} 
-method.
-
-So far, the approaches discussed are
-dependent on the availability of readings from the FST. Two final approaches 
-address unknown tokens.
-The {\small {\tt guessSyll}} and {\small {\tt guessFreqSyll}} approaches are 
-identical to {\small {\tt guess}} and {\small {\tt guessFreq}}, respectively,
-except that the -{\small {\tt Syll}} approaches attempt to place stress on
-unknown tokens (i.e. tokens not found in the FST) by guessing which 
-\emph{syllable} is stressed. Empirical research with
-Russian native speakers reading nonce words in running text concludes that
-the default stress position in Russian is the last syllable of the stem, or
-in other words, the last vowel that is not part of an inflectional ending 
-\cite{Crosswhite.ea-03}. Therefore, the -{\small {\tt Syll}} approaches 
-place the stress near the end of the stem. Determining 
-which vowel is stem-final
-is itself a difficult task. Our algorithm merely places the stress on the last 
-vowel that is followed by a consonant. Although more sophisticated algorithms
-are necessary to accurately locate the stem boundary, this is not the central
-focus of this study.
-
-For our baseline, we take the output of our morphological analyzer (without the 
-Constraint Grammar) in combination with the {\small {\tt bare}}, 
-{\small {\tt safe}}, {\small {\tt guess}}, {\small {\tt guessFreq}}, 
-{\small {\tt guessSyll}}, and {\small {\tt guessFreqSyll}} approaches. We also 
-compare our outcomes with the RussianGram\footnote{http://russiangram.com/} 
-plugin for the Google Chrome web browser. RussianGram is not open-source, so we
-can only guess what technologies support the service. In any case, it provides a
-meaningful reference point for the success of each of the methods described 
-above.
-
-\section{Results}
-
-\begin{table*}[t]
-  \centering
-  \begin{tabular}{r | r r r | r r }
-    approach & accuracy\% & error\% & abstention\% & totTry\% & totFail\% \\
-    \hline
-    \hline
-    {\small {\tt noCGbare}} & 30.43 & 0.17 & 69.39 & 30.61 & 69.57 \\
-    {\small {\tt noCGsafe}} & 90.07 & 0.49 & 9.44 & 90.56 & 9.93 \\
-    {\small {\tt noCGguess}} & 94.34 & 3.36 & 2.30 & 97.70 & 5.66 \\
-    {\small {\tt noCGguessFreq}} & 95.53 & 2.59 & 1.88 & 98.12 & 4.47 \\
-    {\small {\tt noCGguessSyll}} & 94.99 & 4.05 & 0.96 & 99.04 & 5.01 \\
-    {\small {\tt noCGguessFreqSyll}} & 95.83 & 3.46 & 0.72 & 99.28 & 4.17 \\
-    \hline
-    {\small {\tt CGbare}} & 45.78 & 0.44 & 53.78 & 46.22 & 54.22 \\
-    {\small {\tt CGsafe}} & 93.21 & 0.74 & 6.05 & 93.95 & 6.79 \\
-    {\small {\tt CGguess}} & 95.50 & 2.59 & 1.90 & 98.10 & 4.50 \\
-    {\small {\tt CGguessFreq}} & 95.73 & 2.40 & 1.88 & 98.12 & 4.27 \\
-    {\small {\tt CGguessSyll}} & 95.92 & 3.33 & 0.74 & 99.26 & 4.08 \\
-    {\small {\tt CGguessFreqSyll}} & 96.15 & 3.14 & 0.72 & 99.28 & 3.85 \\
-    \hline
-    {\small {\tt RussianGram}} & 90.09 & 0.79 & 9.12 & 90.88 & 9.91
-  \end{tabular}
-  \caption{Results of stress placement task evaluation. (N = 4048)}
-  \label{tab:results}
-\end{table*}
-
-Single-syllable words and words that are unstressed in the gold standard were
-not evaluated for reasons discussed above, leaving 4048 tokens for evaluation.
-Since our approach is lexicon-based, some of our results should be interpreted
-with respect to how many of the stressed wordforms in the gold standard corpus
-can be found in the output of the finite-state transducer. We refer to this 
-measure as \emph{recall}. Out of 4048 tokens, 3949 were 
-found in the FST, which is equal to 97.55\%.
-This number represents the ceiling for methods relying on the FST.
-Higher scores are only achievable by expanding the FST or by using syllable-
-guessing algorithms. After running the Constraint Grammar, recall was 97.35\%, 
-a reduction of 0.20\%. Apparently the Constraint Grammar erroneously removes some 
-correct readings, or else there are errors in the gold corpus.
-
-Results were compiled for each of the 13 approaches discussed above: 
-\emph{without} the Constraint Grammar ({\small {\tt noCG}}) x 6 approaches, 
-\emph{with} the Constraint Grammar ({\small {\tt CG}}) x 6 approaches, 
-and RussianGram ({\small {\tt RussianGram}}). Results are given in 
-Table~\ref{tab:results}. Each token was categorised as either an accurate output, 
-or one of two categories of failures: errors and abstentions. 
-If the stress tool outputs a stressed wordform, and it is incorrect, then it is
-counted as an `error'. If the stress tool outputs an unstressed wordform, then
-it is counted as an `abstention'.
-
-The right-most columns in 
-Table~\ref{tab:results} combine values of the basic categories. The term `totTry' 
-refers to the sum of the accuracy and error rate. This figure represents the 
-proportion of tokens on which our system output a stressed wordform. In the case 
-of {\small {\tt noCGbare}}, the accuracy\% (30.43) and error\% (0.17) sum to the 
-totTry value of 30.61. The term `totFail' refers to the sum of error rate 
-and abstention rate, which is the 
-proportion of tokens for which the system failed to output the correct stressed 
-form. In the case of {\small {\tt noCGbare}}, the error\% (0.17) 
-and abstention\% (69.39) sum to the totFail value of of 69.57 (rounded).
-
-\rr{Are differences significant?}
-
-The {\small {\tt noCGbare}} approach achieves a baseline accuracy of
-30.43\%, so roughly two thirds of the tokens in our corpus are
-morphosyntactically ambiguous. The error rate of 0.17\% primarily represents 
-forms whose stress position varies from speaker to speaker (e.g.
-\rus{зав\'{и}л\'{и}сь} \emph{zav\'{i}l\'{i}s'}), errors in the gold corpus (e.g. 
-\rus{вер\'{и}м} \emph{ver\'{i}m}) . The {\small {\tt noCGsafe}} approach
-achieves a 60\% improvement in accuracy (90.07\%), which means that 89.39\% of 
-morphosyntactic ambiguity on our corpus is stress-irrelevant. Interestingly, the 
-RussianGram web service achieves results that are very close 
-to the {\small {\tt noCGsafe}} approach.
-
-Since the ceiling recall for the FST is 97.55\%, and since the 
-{\small {\tt noCGsafe}} approach achieves 90.07\%, the maximum improvement 
-that a Constraint Grammar could theoretically achieve is 7.48\%. A comparison
-of {\small {\tt noCGsafe}} and {\small {\tt CGsafe}} reveals an improvement
-of 3.14\%, which is about 42\% of the way to the ceiling recall.\rr{...while raising the error rate 0.25\%}
-
-The {\small {\tt CGguess}} and {\small {\tt CGguessFreq}} approaches are also 
-limited by the 97.55\% ceiling from the FST, and their accuracies achieve
-improvements of 2.29\% and 2.52\%, respectively, over {\small {\tt CGsafe}}.
-However, these gains come at the cost of error rates as much as 3.5 times 
-higher than {\small {\tt CGsafe}}: +1.85\% and +1.66\%, respectively. It is
-not surprising that {\small {\tt CGguessFreq}} has higher accuracy and a lower
-error rate than {\small {\tt CGguess}}, since frequency-based guesses are by
-definition more likely to occur. The frequency data were taken from a very 
-small corpus, and it is likely that frequency from a larger corpus would yield 
-better results.
-
-The two {\small {\tt Syll}} approaches were designed to make a blind guess on 
-every wordform that is not found in the FST, which would ideally result in an
-abstention rate of 0\%. However, the abstention rates of approximately 0.7\%
-are the result of the fact that some words in the FST, especially proper nouns, 
-had not been assigned stress. Because the FST outputs a form -- albeit unstressed 
--- the {\small {\tt Syll}} algorithm is not called. This means that 
-{\small {\tt Syll}} is only guessing on about 2\% of the tokens. The improvement
-on overall accuracy from {\small {\tt CGguessFreq}} to 
-{\small {\tt CGguessFreqSyll}} is 0.42\%, which means that the 
-{\small {\tt Syll}} guess was accurate 21\% of the time.
-
-\rr{Examine errors and categorize/refute them.}
-
-\section{Discussion} \label{discussion}
-
-One of the main points of this paper is to highlight the importance of 
-syntactic context in the Russian wordstress placement task. If your intended 
-application has a low tolerance for error, the 
-{\small {\tt noCGsafe}} approach represents the highest accuracy that is possible 
-without leveraging syntactic information for disambiguation (90.07\%). In other words, a 
-system that is blind to morphosyntax and contextual disambiguation cannot
-significantly outperform {\small {\tt noCGsafe}}. It would appear that this is
-the method used by {\small {\tt RussianGram}}, since its results are so similar
-to {\small {\tt noCGsafe}}. Indeed, this result can be achieved most efficiently
-without any part-of-speech tagging, but through simple dictionary lookup.
-
-We noted in Section~\ref{background} that \newcite{hall_sproat_russian_2013}
-achieved 98.7\% accuracy on stress placement for individual wordforms 
-in a list (i.e. \emph{without} syntax). This result is 8.63\% higher than
-{\small {\tt noCGsafe}}, but it is also a fundamentally different task.
-Based on the surface 
-forms in our FST -- which is based on the same dictionary used for 
-\newcite{hall_sproat_russian_2013} -- 
-we calculate that only 29~518 (1.05\%) 
-of the 2~804~492 wordforms contained in our FST are stress-ambiguous. \rr{These 
-stress-ambiguous wordforms have an average of 2.01 possible stress positions,
-so a perfect stress placement tool has a 50/50 chance of getting them right.
-This means that a perfect stress placement tool could achieve 99.48\% on their
-task.} In our corpus of running text, at least 7.5\% of the tokens are stress-
-ambiguous. Stress ambiguity is
-therefore 7 times more prevalent in our corpus of running text than it is in 
-our wordform dictionary.
-Since the task of wordstress placement is virtually always performed on
-running text, it seems prudent to make use of surrounding information.
-The experiment described in this paper demonstrates that a Constraint Grammar
-can effectively improve the accuracy of a stress placement system without 
-significantly raising the error rate. Our Russian Constraint Grammar is
-under continual development, so we expect higher accuracy in the future.
-
-We are unaware of any other empirical evaluations of Russian wordstress placement 
-in running text. The results of our experiment are promising, but many questions 
-remain unanswered.
-The experiment was limited by properties of the gold standard corpus, including
-its size, genre distribution, and quality. Ideally, a gold standard corpus for 
-Russian wordstress should mark 
-whether monosyllabic words are stressed, and of course its annotations should
-be accurate. Our gold corpus represents a broad variety of text genres, which 
-makes our results more generalizable, but a larger corpus would allow for
-evaluating each genre individually. For example, the vast majority of words with
-shifting stress are of Slavic origins, so we expect a genre such as technical
-writing to have a lower proportion of words with stress ambiguity, since it
-uses a higher proportion of borrowed words, calques, and neologisms with simple
-stress patterns.
-
-In addition to genre, it is also likely that text
-complexity affects the difficulty of the stress placement task. The distribution
-of different kinds of syntactic constructions vary with text complexity 
-\cite{Vajjala.Meurers-12}, and
-we expect that the effectiveness of the Constraint Grammar will be affected by
-those differences.
-
-The resources needed for machine-learning --
-such as a large corpus of Russian running text with marked stress -- are simply
-not available at this time. Even so, lexicon- and rule-based approaches have
-some advantages over machine-learning approaches. For example, we are able to
-abstain from marking stress on tokens whose morphosyntactic ambiguity cannot be
-adequately resolved. In language-learning applications, this reduces the 
-likelihood of learners being exposed to incorrect wordforms, and accepting them
-as authoritative. Such circumstances can lead to considerable frustration and
-lack of trust in the learning tool. However, machine-learning does seem well-
-suited to
-placing stress on unknown words, since morphosyntactic analysis is problematic.
-The syllable-guessing algorithm {\small {\tt Syll}} used in this experiment was 
-overly simplistic, and so it was not surprising that it was only moderately 
-successful. More rigorous rule-based approaches have been suggested in other 
-studies \cite{Church-85,Williams-87,Xomicevic_2008}. For example, 
-\newcite{Xomicevic_2008} attempt to parse the unknown token by matching known 
-prefixes and suffixes.
-
-Other studies have applied machine-learning to guessing stress of unknown words
-\cite{Pearson.ea-00,Webster-04,Dou.ea-09,hall_sproat_russian_2013}. For example,
-for unknown words, \newcite{hall_sproat_russian_2013} achieve an accuracy of 
-83.9\%. Implementing a similar model in a rule-based system like ours would allow 
-for some improvements to this result. Their model was trained on a full list of 
-Russian words, which is not representative of the words that would be unknown to 
-a system like ours. Most of the complicated wordstress patterns are closed 
-classes\footnote{The growing number of masculine nouns with shifting stress 
-(\emph{d\'{o}ktor$\sim$doktor\'{a}}) are an exception to this generalization.},
-so we could exclude closed classes of words from the training data, leaving only
-word classes that are likely to be similar to unknown tokens, such as those with 
-productive derivational affixes.
-
 \section{Conclusions}
-
-We have demonstrated the effectiveness of using a Constraint Grammar to improve
-the results of a Russian wordstress placement task in running text by resolving
-42\% of the ambiguity in our gold corpus. We showed
-that stress ambiguity is 7 times more prevalent in our corpus of running text
-than it is in our lexicon, suggesting the importance of context-based
-disambiguation for this task.  As with any 
-lexicon- and rule-based system, the lexicon and rules can be expanded and 
-improved, but our initial results are promising, especially considering the short 
-timespan over which they were developed.
-
-As this is the first empirical study of its kind, we also discussed 
-methodological limitations and possibilities for subsequent research, including
-collecting stressed corpora of varying text complexity and/or genre, as well as
-implementing and/or adapting established wordstress-guessing methods for unknown 
-words.
 
@@
 \bibliographystyle{acl}
-\bibliography{rusaccent}
+\bibliography{ruscg}
 

### 2015-02-09 2015-eamt-melsa/2015-eamt-melsa.tex
glue in melsa paper
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58776 72bbbca6-d526-0410-a7d9-f06f51895060

@@
-%
-% File eamt15.tex, an adaptation of eamt14.tex (which was a copy of
-% eamt12.tex)
-%
-% Contact: eamt2015@dlsi.ua.es
-
-%%% To ease future customizations, various replaceables have been paramaterized
-%%% as listed in the newcommands section
+% test
 
@@
 %%% YOUR PACKAGES BELOW THIS LINE %%%
+
+\usepackage{amsmath}
+\usepackage{amsfonts}
+\usepackage{multirow}
+\usepackage{url}
+\usepackage{natbib}
+\usepackage{color}
+\usepackage[small,bf]{caption}
+
+\usepackage[backgroundcolor=lightgray]{todonotes}
+%\usepackage[disable]{todonotes}
 
@@
 
-\title{Instructions for \confname\ Proceedings}
+% \newcommand{\comment}[1]{{\small{\texttt{$\clubsuit$#1$\clubsuit$}}}}
+%\newcommand{\comment}[1]{\marginpar{\scriptsize\sf \textcolor{blue}{#1}}}
+\newcommand{\comment}[1]{\todo{#1}}
 
-\author{First Author\\
-  Affiliation / Address line 1\\
-  Affiliation / Address line 2\\
-  Affiliation / Address line 3\\
-  {\tt email@domain}  \And
-  Second Author\\
-  Affiliation / Address line 1\\
-  Affiliation / Address line 2\\
-  Affiliation / Address line 3\\
-  {\tt email@domain}}
+  \title{Unsupervised training of maximum-entropy models for lexical
+    selection in rule-based machine translation}
+
+%\author{}
+
+\author{
+ Francis M.\ Tyers \\
+ HSL-fakultehta, \\
+ UiT Norgga \'{a}rktala\v{s} universitehta, \\
+ N-9018 Romsa \\
+ \and
+ Felipe S\'anchez-Mart\'inez \\
+ Dept.\ Lleng.\ i Sist.\ Inform., \\
+ Universitat d'Alacant,\\
+ E-03071 Alacant\\
+ \and
+ Mikel L.\ Forcada \\
+ Dept.\ Lleng.\ i Sist.\ Inform., \\
+ Universitat d'Alacant,\\
+ E-03071 Alacant\\
+}
 
@@
 \maketitle
+
 \begin{abstract}
-   ... 
+  This article presents a method of training maximum-entropy models to
+  perform lexical selection in a rule-based machine translation
+  system. The training method described is unsupervised; that is, it
+  does not require any annotated corpus. The method uses
+  source-language monolingual corpora, an existing machine translation
+  (MT) system and a statistical target-language model. Using the
+  existing MT system, the sentences in the source-language corpus are
+  translated in all possible ways according to the different
+  translation equivalents in the bilingual dictionary of the system.
+%  taking into account all of the lexical selection possibilities.
+  These translations are then scored on the target-language model and
+  the scores are normalised to provide fractional counts for training
+  a source-language maximum-entropy lexical-selection model. We show
+  that this model can perform equally well, or better, than using the
+  target-language model directly for lexical selection, at a
+  substantially reduced computational cost.
 \end{abstract}
@@
 
-\cite{koehn2005}
+Corpus-based MT has been the primary research direction in the field 
+of machine translation in recent years. However, RBMT systems are 
+still being developed, and there are many successful commercial 
+and non-commercial systems. One reason for the continued development
+of RBMT systems is that in order to be successful, corpus-based MT requires parallel corpora in
+the order of tens of millions of words. Although for some language
+pairs these exist, they only exist for a fraction of the 
+world's languages.
+
+An RBMT system typically consists of an analysis component,\footnote{For example
+a morphological or syntactic analyser.} a transfer
+component and a generation component. As part of the transfer component
+it is necessary to make choices regarding words in the source language
+which may have more than one translation in the target language.
+
+Lexical selection is the task of choosing, for a given source-language
+word, the most adequate translation in the target language among a
+known set of alternatives.  The task is related to the task of
+word-sense disambiguation \citep{ide98a}. However, it is different to
+word-sense disambiguation in that lexical selection is a bilingual
+problem, not a monolingual problem: its aim is to find the most
+adequate translation, not the most adequate sense. Thus, it is not
+necessary to choose between a series of fine-grained senses if all
+these senses result in the same final translation; however, it may be
+necessary to choose a different translation for the same sense, for
+example in a collocation.
+% It could also be the case that a single sense could have more than
+% one possible translation (for example synonyms).
+
+%\comment{FSM: A more general introduction to MT and RBMT should be
+%  provided. The necessity of the method should also be justified (lack
+%  of parallel corpora), probably at the same time RBMT is claimed to
+%  be a need for most language pairs with no parallel corpora
+%  available.}
+
+\subsection{Prior work}
+\comment{This section should be ironed out a bit.}  In their
+pioneering work, \cite{dagan1994word} used the term
+\emph{word sense disambiguation} to refer to what is actually target
+word selection in machine translation; they used a parser to identify
+some syntactic relations such as subject--object or subject--verb.
+After generating all the possible translations for a given
+input sentence using an ambiguous bilingual dictionary, they extract the syntactic
+tuples from the target language and count the frequency in a previously-trained target-language 
+model of tuples.  They use maximum-likelihood estimation to calculate the probability of a given 
+target language tuple being the translation of a given source language tuple, with
+an automatically determined threshold for confidence.
+
+Then \cite{berger1996} illustrated the use of maximum-entropy classifiers
+on the specific problem of lexical selection in IBM-style word-based
+statistical machine translation. Other authors
+\citep{carbonell06,melero07a} have used target-language models to rank
+the translations resulting from all possible combinations of lexical
+selections.  Nowadays, in state-of-the-art phrase-based statistical
+machine translation \citep{koehn2009statistical}, lexical selection is
+taken care of by a combination of the translation model and the
+language model. The translation model provides probabilities of
+translation between words or word sequences (often referred to as
+\emph{phrases}) in the source and target language. The target-language
+model provides probabilities of word sequences in the target language.
+As regards rule-based machine translation, \cite{sanchez07a} proposed 
+to use information from the target language to train a statistical 
+lexical-selection module but did not provide any experiments. 
+\todo[inline]{FMT: integrate \citep{sanchez07a} somewhere here}
+\cite{marechek10} trained a maximum-entropy lexical selector for their TectoMT rule-based system
+using a bilingual corpus.  More recently, \cite{tyers12a} presented a
+method of lexical selection for rule-based machine translation based
+on rules which select or remove translations in fixed-length contexts,
+along with a training method for learning the rules from a
+word-aligned parallel corpus. The principal disadvantage with this
+method is that it relies on a parallel corpus for the language pair,
+something which is not always available.
+
+The remainder of the article is organised as follows: 
+\todo[inline]{FMT: Add article structure}
+
+\section{Method}
+
+%\todo[inline]{FSM: In the experiments a development corpus is
+%  mentioned. For what is it used? This should be explain in this
+%  section.}
+
+\comment{Improve this} Lexical selection in this paper considers for
+each word a simple source-language context made up of neighbouring
+lemma+part-of-speech combinations. Contexts considered include up to
+two words to the left and up to two words to the right of the word to
+be translated. 
+%\comment{FSM: Later, a development corpus is
+%  mentioned. In this section it should be mentioned what it is used
+% for.}
+
+Let the probability of a translation $t$ being the translation of a
+word $s$ in an SL context $c$ be $p_s(t|c)$. In principle, this value
+could be estimated directly from the available corpora for every
+combination of $(s, t, c)$.  This would however present two questions:
+(1) how should the relevant contexts be chosen? and (2) what should be
+done when the translations of words are not found in the corpus?  A
+maximum-entropy model answers both of these questions. It allows the
+contexts that we consider to be linguistically interesting to be
+defined \emph{a priori} and then integrate these seamlessly into a
+probabilistic model \citep{Manning99b}.  In answer to the second
+question, a maximum-entropy model assumes nothing about what is not
+present in the training data. That is, if there is no information in
+the training data, then it assumes that all outcomes ---that is, all
+possible translations--- are equally likely.  As previously mentioned,
+the principle of maximum entropy has been applied to the problem of
+lexical selection before; in particular, \cite{berger1996} cast the
+problem of lexical selection in statistical MT as a classification
+problem. They learn a separate maximum-entropy classifier for each SL
+word form,\comment{[:we do too, integrate this later]} using SL context to
+distinguish between possible translations. These classifiers are then
+incorporated into the translation model of their word-based SMT
+system.\comment{[how is our work different?]}  In their approach, a
+classifier consists of a set of binary feature functions and
+corresponding weights for each feature. 
+In both \cite{berger1996} and our method, features are defined in the
+form $h_k^{s}(t, c)$, where $t$
+is a translation, and $c$ is an SL context. One difference is that \cite{berger1996} take $s$, $t$
+and $c$ to be based on word forms, whereas in our method they are based
+on lemma forms.
+A feature where the Spanish word \emph{pez} (`fish' as a living animal) is seen as the
+translation of \emph{arrain} (`fish') in the context \emph{arrain
+  handi} `big fish' would therefore be defined as:
+%\comment{FSM: I have introduced the index $k$
+%  because several features are defined for a SL word $s$.} 
+\begin{equation}
+h_k^{\mathrm{arrain}}(t, c) = \left \{ \begin{matrix}
+                 ~~1 & \mbox{if } t~ =~ \textit{pez}~ \mathrm{and}~ \textit{handi}~ \mathrm{follows}~ \textit{arrain}   \\
+                 ~~0 & \mbox{otherwise} \end{matrix} \right.
+\label{eq:feat-func}
+\end{equation}
+This feature considers a context of zero words to the left of the
+problem word and one word to the right of it.
+
+As a result of training, each feature in the classifier is assigned a
+weight $\lambda_k^s$, and combining these weights of active features
+as in equation~(\ref{eq:max-ent-sum}) yields the probability of a
+translation $t$ for word $s$ in context $c$.
+
+\begin{equation}
+p_s(t|c) = \frac{1}{Z^s(c)} \exp{ \sum_{k=1}^{n_F} \lambda_k^s h_k^s(t, c)}
+\label{eq:max-ent-sum}
+\end{equation}
+
+In this equation, $Z^s(c)$ is a normalising constant. Thus, the most
+probable translation \(t^\star\) can be found using
+% equation~(\ref{eq:max-ent-max}).
+
+\begin{equation}
+  t^\star=\underset{t \in T_s(s)}{\arg\max} ~ p_s(t|c) = \underset{t \in T_s(s)}{\arg\max} \sum_{k=1}^{n_F} \lambda_k^s h_k^s(t, c) .
+\label{eq:max-ent-max}
+\end{equation}
+
+The approaches by \cite{berger1996} and by \cite{marechek10} cited
+above both take advantage of a parallel corpus to collect counts of
+contexts and translations in order to train the maximum-entropy
+model. However, parallel corpora are not available for the majority of
+the world's written languages. In this section we describe an
+unsupervised method to learn the model using only monolingual corpora
+and the components from an existing machine translation system.
+
+The input to our method consists of a collection of samples,
+$\mathcal{G} = (S, G)$, where $S = (s_1, s_2, \ldots s_{|S|})$ is a
+sequence of source-language words, and $G = \{g_1, g_2, \ldots
+g_{|G|}\}$ is a set of possible \emph{lexical-selection paths}. A
+lexical-selection path $g = (t_1, t_2, \ldots, t_{|s|})$ is a sequence
+of lexical-selection choices of those SL words, where, \(t_i=T(s_i)\)
+and \(T(\cdot)\) is the set of all translations for a source
+word. This is produced in the first stages of rule-based machine
+translation, just after morphological analysis, part-of-speech
+tagging, and bilingual dictionary lookup, and before any structural
+transfer takes place (we will call this \emph{pre-lexsel}). In our
+model, it is after these first stages that lexical selection
+(\emph{lexsel}) occurs. After lexical selection, structural transfer
+and generation take place; a function $\tau(g_i,S)$ represents the
+result of these last stages, which we will call \emph{post-lexsel},
+and returns a finished translation of a specific lexical-selection
+path $g_i$ of sentence $S$. Schematically:
+\begin{displaymath}
+  \label{eq:steps}
+  S \to \mbox{\framebox{\parbox{0.8cm}{\textit{pre- lexsel}}}} \to (\{g_i\}_{i=1}^{i=|G|},S) \to \mbox{\framebox{\parbox{1.0cm}{\textit{\textbf{lexsel}}}}} \to (g^\star,S) \to \mbox{\framebox{\parbox{0.8cm}{\textit{post- lexsel}}}} \to \tau(g^\star,S)
+\end{displaymath}
+
+%Taking our previous example for the parallel corpus, Table~\ref{table:ex-monocorp} 
+%presents an example of five possible samples.
+
+As our method is unsupervised, and therefore the occurrences of
+specific lexical selection events \((s,t,c)\) cannot be counted, a
+target language model \(P_{TL}(\cdot)\) is used to compute a value for
+the fractional count for disambiguation path \(g_i\), $p(g_i|S)$ after
+suitable normalisation:
+\begin{equation}
+ \label{eq:normalising}
+   p(g_i|s) = \frac{P_{TL}(\tau(g_i,S))}{\sum_{g_i \in G} P_{TL}(\tau(g_i,S))}
+\end{equation}
+~\\
+The maximum-entropy model is trained instead using the fractional count
+$p(g_i|S)$ for the events $(s,t,c)$ found in \(g_i\), that is, when in
+\(g_i\) $t$ is the translation for \(s\) in context \(c\). That is, as
+if event $(s,t,c)$ had been seen a fractional number $p(g_i,s)$ of
+times. We prune features occurring less than a certain number of times 
+in the corpus, using a development corpus to guide pruning (see section~\ref{sec:eval-ref-results}).
+\comment{Reword this to avoid repetition} The method used here
+for lexical selection is analogous to the method used by
+\cite{sanchez08b} to train a hidden-Markov-model-based part-of-speech
+tagger in a rule-based machine translation system. 
+Table~\ref{tb:workedexample} shows how a collection of samples
+\(\mathcal{G}\) for the translation of Basque sentences containing the
+word \emph{arrain} into Spanish.\footnote{The word \emph{arrain} `fish' has two translations
+in Spanish, the first \emph{pez} refers to the living animal, whereas
+\emph{pescado} refers to the animal as food.} 
+%\comment{FSM: The correct translations
+%  should be clearly identified in the table.}
+%\todo[inline]{FSM: There is an important difference between the
+%  approach in \cite{sanchez08b} and the one in this paper: we do not
+%  have to face the problem of free rides. Should this be mentioned
+%  here?}
+
+\begin{table}
+ \begin{center}
+  \begin{tabular}{|c|l|r|}
+    \hline
+       & \textbf{Sentence} & $p(g_i|S)$ \\
+    \hline % Los pescados grandes peque\~{n}o come
+     \(S_1\) & \emph{Arrain handiak txikia jaten du}  &  \\
+                        $\tau(g_{11},S_1)$ & El pez grande se come el peque\~{n}o &  0.830 \\
+                        $\tau(g_{12},S_1)$ & El pescado grande se come el peque\~{n}o & 0.134 \\
+                        $\tau(g_{13},S_1)$ & El pescado capaz se come el peque\~{n}o & 0.034\\
+                        $\tau(g_{14},S_1)$ & El pez capaz se come el peque\~{n}o  & 0.001\\
+    \hline % El padre el pescado nos ha preparado para cenar
+     \(S_2\) & \emph{Aitak arraina prestatu digu afaltzeko} &  \\
+                        $\tau(g_{21},S_2)$ & Padre nos ha preparado pescado para cenar & 0.846 \\
+                        $\tau(g_{22},S_2)$ & Pap\'{a} nos ha preparado pescado para cenar & 0.134 \\
+                        $\tau(g_{23},S_2)$ & Padre nos ha preparado pez para cenar & 0.015 \\
+                        $\tau(g_{24},S_2)$ & Pap\'{a} nos ha preparado pez para cenar & 0.005 \\
+    \hline % Aqu\'{\i} el pescado muy es dulce
+    $S_3$  & \emph{Hemen arraina oso goxoa da}  & \\
+                        $\tau(g_{31},S_3)$ & Aqu\'{\i} el pescado es muy rico & 0.912 \\
+                        $\tau(g_{32},S_3)$ & Aqu\'{\i} el pescado es muy suave & 0.067 \\
+                        $\tau(g_{33},S_3)$ & Aqu\'{\i} el pez es muy rico & 0.011 \\
+                        $\tau(g_{34},S_3)$ & Aqu\'{\i} el pez es muy suave & 0.008 \\
+                        $\tau(g_{35},S_3)$ & Aqu\'{\i} el pez es muy dulce & 0.001 \\
+                        $\tau(g_{36},S_3)$ & Aqu\'{\i} el pescado es muy dulce & 0.001 \\
+    \hline % Un pescado grande del mar es
+    \(S_4\) & \emph{Itsasoko arrain handi bat da} & \\
+                        $\tau(g_{41},S_4)$ & Es un pez grande del mar & 0.595 \\
+                        $\tau(g_{42},S_4)$ & Es un pescado grande del mar & 0.403 \\
+                        $\tau(g_{43},S_4)$ & Es un pez capaz del mar & 0.001 \\
+                        $\tau(g_{44},S_4)$ & Es un pescado capaz del mar & 0.001 \\
+    \hline % El pescado grande por esos serv\'{\i}an para alimentar
+    $S_5$ &  \emph{Arrain handiak horiez baliatzen ziren elikatzeko} &  \\
+                        $\tau(g_{51},S_5)$ & Los peces grandes se nutr\'{\i}an de esos & 0.862 \\
+                        $\tau(g_{52},S_5)$ & Los pescados grandes se nutr\'{\i}an de esos & 0.121 \\
+                        $\tau(g_{53},S_5)$ & Los peces grandes se alimentaban de esos & 0.012 \\
+                        $\tau(g_{54},S_5)$ & Los peces capaces se alimentaban de esos & 0.001 \\
+                        $\tau(g_{55},S_5)$ & Los pescados grandes se alimentaban de esos & 0.001 \\
+                        $\tau(g_{56},S_5)$ & Los pescados capaces se alimentaban de esos & 0.001 \\
+                        $\tau(g_{57},S_6)$ & Los peces capaces se nutr\'{\i}an de esos & 0.001 \\
+                        $\tau(g_{58},S_7)$ & Los pescados capaces se nutr\'{\i}an de esos & 0.001 \\
+    \hline
+
+  \end{tabular}
+  \caption{A sample of source-language sentences, the output of the
+    possible lexical-selection paths after scoring on a
+    target-language model. The scores are normalised as fractional
+    counts, which are used as inputs to the maximum-entropy training
+    process. In all sentences, the resultant translation with the highest share of the 
+    probability mass is the most adequate.}
+   \label{tb:workedexample}
+ \end{center}
+\end{table}
+
+\section{Experimental setting}
+
+This section describes the training and evaluation settings used in
+the remainder of this paper. The primary motivation behind the
+evaluation is that it should be automatic, meaningful, and be performed over a test set which
+is large enough to be representative. It should evaluate both
+performance on the specific subtask of lexical selection, and on the
+whole translation task. Evaluating lexical-selection performance is an
+intrinsic module-based evaluation.  It measures how well the lexical
+selection module disambiguates the lexical-transfer
+output as compared to a gold-standard corpus. The lexical 
+transfer output is the result if looking up the translations of the 
+source-language \emph{lexical forms} --- combinations of lemma and morphological
+tags --- in the bilingual dictionary.
+
+The
+whole translation task evaluation is an extrinsic evaluation, which
+tests how much the system improves as regards final translation
+quality in a real system.
+
+The lexical-selection module should be as language-independent as
+possible. To that end, the language pairs tested should show as wide a
+variety of linguistic phenomena. It is also important that the
+methodology be as applicable to lesser-resourced and marginalised
+languages, as to major languages.
+
+This section begins with a short description of the Apertium 
+platform \citep{forcada2011apertium}. This is followed by 
+an overview of each of the language
+pairs chosen for the evaluation. The corpora to be used for training
+and evaluation will subsequently be described, along with the method
+used for annotating them. This is followed by a description of the
+performance measures to be used in the evaluation, and the reference
+results using these metrics for each of the language pairs.
+
+\subsection{Apertium}
+\label{sec:apertium}
+
+Apertium is a free/open-source rule-based machine translation platform, 
+it comprises an engine, a toolbox and data to build rule-based machine
+translation systems. Translation is implemented as a pipeline consisting
+of the following modules: morphological analysis, morphological disambiguation,
+lexical transfer, lexical selection, structural transfer and morphological
+generation. The modules of the system for this method are the \emph{lexical transfer}
+module which provides for each source language lexical form a set of possible
+target language forms and the \emph{structural transfer} and \emph{morphological generation}
+modules which for each pair of source language lexical form and translation 
+produce a final translation in the target language.
+
+\subsection{Language pairs}
+\label{sec:eval-systems}
+
+Evaluation will be performed using four language pairs. These pairs
+have been selected as they include languages with different
+morphological complexity, and different amounts of resources available
+--- although for all pairs there is a parallel corpus available, which
+is used for evaluation (see Section \ref{ss:measures}). Table~\ref{table:lang-pair-stats} 
+presents statistics about the development status of each of the language 
+pairs chosen, including the number of bilingual dictionary entries, coverage of
+test corpora and translational ambiguity of the bilingual dictionary. Translation
+ambiguity is provided for both all words, and for only ambiguous open-category 
+words. In the experiments, lexical-selection rules are learnt for the open categories
+of adjective, noun and verb.
+
+\begin{table}
+  \centering
+  \begin{tabular}{|l|r|r|r|r|r|}
+    \hline
+    \multirow{2}{*}{\textbf{Pair}} & \textbf{Lexical} & \multirow{2}{*}{\textbf{Ambig.}} & \textbf{Na\"{i}ve}& \textbf{Total} & \textbf{Ambig. of}\\\
+                          & \textbf{items} &                            & \textbf{coverage} & \textbf{ambig.} & \textbf{ambig. words}\\
+    \hline
+    \texttt{br-fr} & 27,690 & 235  &  94.47 \%  & 1.02 & 2.5 \\
+    \texttt{mk-en} & 33,229 & 1,323 & 92.17 \%  & 1.37 & 2.8 \\
+    \texttt{eu-es} & 19,323 & 455 & 91.70 \% & 1.08 & 2.4 \\
+    \texttt{en-es} & 30,186 & 822 & 98.08 \% & 1.05 & 2.2 \\
+    \hline
+  \end{tabular}
+  \caption{Statistics about development status of language pairs. The second
+    column gives the number of entries in the bilingual dictionaries of the systems;
+    the third column gives the total number of source language words with more 
+    than one translation in the target language; the fourth column gives the coverage
+    of the dictionaries on the test corpus (see section~\ref{sec:eval-corpus}); the fifth
+    column gives the total mean ambiguity for open classes (nouns, verbs, adjectives);
+    and the final column gives the mean ambiguity for ambiguous words in open classes.}
+  \label{table:lang-pair-stats}
+\end{table}
+
+%\todo[inline]{FSM: More information about the language pairs needs to be
+%  given: naive coverage, average lexical ambiguity over all words and
+%  over the ambiguous ones (should this information be given for the
+%  different lexical categories?). This information would allow the
+%  reader to have an idea of how hard is the lexical selection problem
+%  in each case.}
+
+%\todo[inline]{FSM: If I remember correctly, we only deal with lexical
+%  ambiguity in certain lexical categories. This is not mentioned
+%  anywhere.}
+
+\begin{description}
+\item[Breton--French:] Development of the Breton--French pair has been
+  described in two articles \citep{tyers09b,tyers10b}. The bilingual
+  dictionaries were not built with polysemy in mind from the outset,
+  but some entries were added later to start work on lexical
+  selection. The version used in this paper is SVN revision
+  41375.\footnote{\url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-br-fr}}
+  This is the only available machine translation system between Breton
+  and French. It has been developed part-time over a number of years.
+
+\item[Macedonian--English:] The Macedonian--English pair
+  in Apertium was created specifically for the purposes of running
+  lexical-selection experiments. The resources reused from other pairs
+  were the English morphological dictionary from the Icelandic and
+  English pair \citep{brandt11}, the Macedonian morphological
+  dictionary and constraint grammar from the Macedonian and Bulgarian
+  pair \citep{rangelov11}, and the SETimes parallel corpus
+  \citep{tyers10}.  The work was carried out over a period of eight
+  days, and consisted of creating the bilingual dictionary and
+  transfer rules. The bilingual dictionary was created by tagging both
+  sides of the parallel corpus, word-aligning them with {\sc giza++}
+  \citep{och03a} and extracting the probabilistic lexicon. Entries
+  from this lexicon were checked manually according to frequency and
+  included in the bilingual dictionary of the machine translation
+  system. The most probable entry was checked to ensure that it was 
+  an adequate translation, and if so marked as the 
+  default.\footnote{A bilingual dictionary in Apertium
+    \citep{forcada2011apertium} may contain more than one possible
+    target-language translation for a given source-language
+    word. Where there is more than one possible translation, the
+    person who is writing the dictionary chooses the most general or
+    most frequent translation among the set of possible translations
+    and marks it as the \emph{linguistic default} translation.} As a
+  result of attempting to include all possible translations, the
+  average number of translations per word is much higher (see table~\ref{table:lang-pair-stats})
+than in other pairs.  The transfer rules were written
+  by hand to treat translation problems in the corpus. The version of
+  the software used in this paper is SVN revision
+  41476.\footnote{\url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-mk-en}}.
+\item[Basque--Spanish:] The development of the Basque to Spanish pair
+  is described in \cite{ginesti09}. The bilingual dictionary was taken
+  from the free/open-source Matxin system \citep{mayor2011}. 
+  Alternative translations were included in the bilingual
+  dictionary.% , but non-default translations were marked with a
+%  direction restriction\comment{FSM: This needs to be carefully
+%    explained to avoid confusing the reader.}.\footnote{The bilingual
+%    dictionaries in Apertium are bidirectional. They contain entries
+%    which can be used for translating both from language $A$ to
+%    language $B$ and vice versa. A direction restriction is used to
+%    limit the entry to being used for translation in one translation
+%    direction.} For lexical-selection experiments these direction
+%  restrictions were removed.
+
+\item[English--Spanish:] The English--Spanish pair was developed from
+  a combination of the English--Catalan and Spanish--Catalan pairs
+  over a period of around 3--5 months. The pair 
+%has been used in other
+%  lexical selection experiments \citep{sanchez07a}\comment{FSM: Sure?
+%    It was on en-ca and in that paper no experimental results are
+%    shown. This paper would be better mentioned in the introduction
+%    (it mentions the use of TL info for lexical selection.},
+  contains a number of entries in the bilingual dictionary with
+  more than one translation. The most-frequent translation, as judged
+  by the language pair developer, is marked as the \emph{default}
+  translation and non-default translations were added according to
+  their frequency. The version used in this paper is SVN revision
+  41387.\footnote{\url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-en-es}}
+\end{description}
+
+\subsection{Performance measures}
+\label{ss:measures}
+
+This section describes the measures that will be used to evaluate the
+performance of the lexical selection method proposed here: a
+\emph{lexical selection performance} measure and a \emph{machine
+  translation performance} measure. The main difference between the
+evaluation measures is that the lexical selection error rate magnifies
+the problem of lexical selection, by restricting the evaluation to
+this feature. Other features of the MT system, such as the transfer rules
+and morphological generation are not taken into account.
+
+\label{sec:eval-metrics}
+\subsubsection{Lexical-selection performance}
+This is an intrinsic module-based evaluation of the performance of the
+lexical-selection module.  It measures how well the lexical-selection
+module disambiguates the output of the lexical-transfer module as
+compared to a gold-standard corpus. For this task, we define a metric,
+the lexical-selection error rate ({\sc ler}),
+% see section~\ref{sec:ler}. 
+\begin{figure}
+\begin{center}
+\setlength\tabcolsep{5.2pt}%
+ \begin{tabular}{|l|llllll|}
+   \hline
+   \multicolumn{7}{|c|}{L'estiu \'{e}s una estaci\'{o} llarga}\\
+   \hline
+   $S$        & \emph{el} & \emph{estiu} & \emph{ser} & \emph{un} & \emph{estaci\'{o}} & \emph{llarg} \\
+   \hline
+   $T_s(s_i)$ & \{the\} & \{summer\} & \{be\} & \{a\} & \{station, season\} &  \{long, lengthy\} \\
+   \hline 
+   $T_r(s_i)$ & \{the\} & \{summer\} & \{be\} & \{a\} &  \{season\} & \{long\} \\
+   \hline
+   $T_t(s_i)$ & \{the\} & \{summer\} & \{be\} & \{a\} &  \{station\} & \{long\} \\
+   \hline
+ \end{tabular}
+ \caption{An example input sentence in Catalan and the three sets of
+   English translations used for calculating the lexical-selection
+   error rate. The source sentence $S = (s_1, s_2, \ldots, s_{|S|})$
+   has two ambiguous words, \emph{estaci\'{o}} and \emph{llarg}. There is one difference
+   between the reference set $T_r(s_i)$ and the test set $T_t(s_i)$ of
+   translations; as a result, the error rate for this sentence is
+   50\%.}
+\label{fig:ler-example}
+\end{center}
+\end{figure}
+
+%\todo[inline]{FSM: LER is measured on lemmas rather than on surface
+%  forms but this is not mentioned anywhere, otherwise the example in
+%  Fig. \ref{fig:ler-example} is not ok.}
+
+The lexical-selection error rate is the fraction of times the given
+system chooses a translation for a word which is not found in an
+annotated reference. The process uses a source-language sentence, $S =
+(s_1, s_2, \ldots, s_{|S|})$ and three functions. The first function,
+$T_s(s_i)$, returns all possible translations of $s_i$ according to
+the bilingual dictionary. The second function, $T_t(s_i)$, returns the
+translations of $s_i$ selected by the lexical-selection module:
+$T_t(s_i) \subseteq T_s(s_i)$; and usually $|T_t(s_i)| = 1$. If the
+lexical-selection module returns more than one translation, the first
+translation is selected. \comment{FSM: How many times does this
+  happens?. I would move this last sentence to an appendix.}  The
+function $T_r(s_i)$ returns the set of reference translations which
+are acceptable for $s_i$ in sentence $S$.\footnote{Depending on how
+  the reference is built, the set returned by $T_r(s_i)$ may be
+  incomplete in the sense that it may not include all possible
+  acceptable translations.} For a
+single sentence, we define the lexical selection error rate (LER) of
+that sentence as
+\begin{equation}
+\mathrm{LER} = \frac{\sum_{i=1}^{|S|} \mathrm{amb}(s_i) \mathrm{diff}(T_r(s_i), T_t(s_i)}{\sum_{i=1}^{|S|} \mathrm{amb}(s_i)},
+\label{eq:ler}
+\end{equation}
+where
+\begin{equation}
+\mathrm{amb}(s_i) = \left \{ \begin{matrix}
+                 1 & \mbox{if } |T_s(s_i)| > 1   \\
+                 0 & \mbox{otherwise} \end{matrix} \right.
+\label{eq:amb}
+\end{equation}
+ tests if a word is ambiguous, and the  function 
+\begin{equation}
+\mathrm{diff}(T_r(s_i), T_t(s_i)) = \left \{ \begin{matrix}
+                 1 & \mbox{if } T_r(s_i) \cap T_t(s_i) = \emptyset   \\
+                 0 & \mbox{otherwise} \end{matrix} \right.
+\label{eq:diff}
+\end{equation}
+states that there is a difference if the intersection between the set
+of reference translations $T_r(s_i)$ and the set of translations from
+the lexical selection module $T_t(s_i)$ is empty.  Recall that,
+although $T_t(s_i)$ returns a set, this set will be a singleton, as
+when the lexical-selection module returns more than one translation
+the first will be selected.\footnote{In practice this does not happen
+as each ambiguous word has a \emph{default} translation.}
+
+%\todo[inline]{FSM: What happens to LER when $\sum_{i=1}^{|S|}
+%  \mathrm{amb}(s_i)=0.0$?}
+
+%\todo[inline]{FSM: LER is defined at the sentence level. It should be
+%  defined at the corpus level, otherwise we would be treating short
+%  sentences as the long ones, and short sentences are less likely to
+%  contain ambiguous words (unfair).}
+
+The table in Figure~\ref{fig:ler-example} gives an
+overview of the inputs. In the description it is assumed that the
+reference translation has been annotated by hand. However, hand
+annotation is a time consuming process, and was not possible. A
+description of how the reference was built is given in
+section~\ref{sec:eval-corpus}.
+
+\subsubsection{Machine translation performance} 
+\comment{This part of the text needs compressing and merging} This is
+an extrinsic evaluation, which ideally would test how much the system
+improves as regards an approximate measurement of final translation
+quality in a real system. For this task, we use the widely-used {\sc
+  bleu} metric (see Section~\ref{sec:eval-metrics}). This is not ideal
+for evaluating the task of a lexical selection module as the
+performance of the module will depend greatly on (a) the coverage of
+the bilingual dictionaries of the RBMT system in question, and (b) the
+number of reference translations. It is also worth noting that
+successful lexical selections may not lead to successful translations
+due to inadequate transfer of morphological features. For example, if
+we translate the Spanish phrase \emph{La gente le dice que no venga}
+`The people tell him not to come', where \emph{decir} can be
+translated as `say' or `tell' in English, and our machine translation
+system generates `The people \emph{tells} him not to come', it would
+be counted the same as if it were `The people \emph{says} him not to
+come', although the lexical selection made was more
+adequate. Generation errors --- where a word is found in the bilingual
+dictionary, but not in the target language morphological dictionary
+--- may also lead to the same problem.  The {\sc bleu} metric \citep{papineni02} is
+included only as it is commonly used to evaluate MT systems. Other
+metrics with similar performance to {\sc bleu} such as {\sc meteor}
+\citep{lavie09} and {\sc nist} \citep{doddington02} are not included
+as we have chosen to focus on lexical-selection performance.
+
+%\todo[inline]{FSM: If we focus on lexical-selection performance may be
+%  we should not give BLEU. Anyway, I think we can give BLEU and PER
+%  (which does not take order into account) but computed using lemmas
+%  instead of surface forms. }
+
+%The {\sc bleu} (Bilingual Evaluation Understudy, \cite{papineni02})
+%metric is an algorithm for evaluating the quality of text which has
+%been machine-translated from one natural language to another. Quality
+%is considered to be the correspondence between a machine's output and
+%that of a human. The central idea is that the closer a machine
+%translation is to a human translation, the better it is
+%\citep{papineni02}.  BLEU was one of the first metrics that was
+%reported to achieve a high correlation with human judgements of
+%quality \citep{papineni02,coughlin03} and remains one of the most
+%popular automated metrics.\comment{FSM: Why such a lengthy description
+%  of BLEU? This description (if it is to be kept, I prefer not) should
+%  appear before in this section. Same applies to the next paragraph.}
+%
+%Scores are calculated over a whole document --- a set of sentences ---
+%by comparing them with a set of reference
+%translations. Intelligibility or grammatical adequacy are not
+%explicitly taken into account. BLEU is designed to approximate human
+%judgement at a corpus level, and performs badly if used to evaluate
+%the quality of individual sentences. The metric also does not
+%correlate with human judgements when ranking systems based on
+%different technologies and is recommended for tracking improvements in
+%performance over different configurations of the same system
+%\citep{callison06}. Additionally, \cite{denkowski12} report that it
+%does not detect post-edition operations that improve translation
+%quality.
+
+\subsubsection{Confidence intervals}
+Confidence intervals for both metrics will be calculated through the
+\emph{bootstrap resampling} method~\citep{efron94} as described
+by~\citet{koehn04} and \citet{zhang04}.\footnote{Broadly speaking,
+  this involves iteratively taking a random sample (with replacements)
+  from the test set and computing the score for that sample. The
+  scores are sorted and, presuming we want the 95\% confidence
+  interval, the top- and bottom-2.5\% are removed. The highest and
+  lowest remaining are given as the interval. For more detail, refer
+  to the references provided.} In all cases, bootstrap resampling will
+be carried out for 1,000 iterations. Where the $p = 0.05$ confidence
+intervals overlap, we will also perform paired bootstrap resampling
+\citep{koehn04}.
+
+\subsection{Corpora}
+\label{sec:eval-corpus}
+
+For creating the test corpora, and for providing a source language corpus 
+for training, and a target language corpus for scoring, we used four parallel corpora: 
+\begin{itemize}
+\item \textbf{Ofis ar Brezhoneg} (\texttt{OAB}): This parallel corpus
+  of Breton and French has been collected specifically for
+  lexical-selection experiments from translations produced by
+  \emph{Ofis ar Brezhoneg} `The Office of the Breton
+  language'.\footnote{In 2010 the administrative status was changed
+    and it was renamed \emph{Ofis Publik ar Brezhoneg} `Public office
+    of the Breton language'. The corpus was published before this date
+    so we use the original name.}  It contains some parallel data
+  previously described in \cite{tyers09b}, and also some new data. The
+  whole corpus has been made available online through the {\sc opus}
+  collection.\footnote{\url{http://opus.lingfil.uu.se/OfisPublik.php}.}
+\item \textbf{South-East European Times} (\texttt{SETimes}): Described
+  in \cite{tyers10}, this corpus is a multilingual corpus of the
+  Balkan languages in the news domain. The Macedonian and English
+  portion will be used.
+\item \textbf{Open Data Euskadi} (\texttt{OpenData}): This is a Basque
+  and Spanish parallel corpus made from the translation memories of
+  the \emph{Herri Arduralaritzaren Euskal Erakundea} `Basque Institute
+  of Public
+  Administration'.\footnote{\url{http://opendata.euskadi.net/w79-contdata/es/contenidos/ds_recursos_linguisticos/memorias_traduccion/es_izo/memorias_traduccion_izo.html}}
+\item \textbf{European Parliament Proceedings} (\texttt{EuroParl}):
+  Described by \cite{koehn05a}, this is a multilingual corpus of the
+  European Union official languages. We are using the English--Spanish
+  data from version 7 of the corpus.
+\end{itemize}
+
+There are a number of approaches to creating evaluation corpora for
+lexical selection in the literature. \cite{vickrey05} use a parallel
+corpus to make annotated test and training sets for experiments in
+lexical selection applied to a simplified translation problem in
+SMT. They use word alignments from {\sc giza++} \citep{och03a} to
+annotate source language words with their translations from the
+reference translation in the parallel corpus. One disadvantage of this
+method is that only one translation is annotated per source language
+word, meaning that accuracies may be lower because of missing
+translations --- this happens when the system chooses a translation
+which is adequate, but is not found in the reference translation. A
+second disadvantage is that the word alignments may not be 100\%
+reliable, which decreases the accuracy of the annotated corpus. An
+alternative method is described by \cite{zinovjeva2000}, who manually
+tags a selection of ambiguous words in English sentences with their
+translation in Swedish. The size of the annotated corpus is not
+mentioned. \cite{specia2005a} use a parallel corpus, but eschew using
+word alignment in favour of using a bilingual dictionary and heuristic
+rules based on part-of-speech and relative position. The corpus was
+reviewed manually before being used in further experiments
+\citep{specia2005}.
+
+\todo[inline]{FSM: I would explain what comes in the following
+  paragraph (the need for an evaluation corpus and how it is built) in
+  advance. Most probably where the measures are defined.}
+
+The ideal situation would be to have, as described by
+\cite{zinovjeva2000}, a hand-annotated evaluation corpus for testing
+the performance of the lexical-selection module. That is, the output
+of the lexical transfer module which had been disambiguated by hand by
+one or more human annotators. As this did not exist for any language
+pair, we decided to automatically annotate a test set for each
+language pair using a process similar to that described by
+\cite{vickrey05}. The decision was made to automatically annotate
+instead of hand annotate for a number of reasons. Firstly, the time
+involved in hand-annotating approximately 30 pages of text, per
+language pair, would be prohibitive. Secondly, linguists were not
+available for all of the language pairs. One disadvantage
+to this method is that the set of
+translations for each ambiguous word may not be complete as it will
+only include the translation which is found in the corresponding
+sentence in the parallel corpus. Further disadvantages are that 
+not all sentence pairs in the parallel corpus will have translations
+that are able to be generated by the MT system (for
+example the reference translation has a word which is not found in
+the bilingual dictionaries on the MT system).
+
+\begin{table}
+\begin{center}
+ \begin{tabular}{|l|r|r||r|r|r|r|}
+    \hline  
+    %Pair           & Lines & Words & Extracted & \texttt{train} & \texttt{test} & \texttt{dev} & No. ambig & Av. ambig \\ 
+    {\bf Pair}      & {\bf Lines} & {\bf Extract.} & \texttt{train} & \texttt{other} & {\bf No. amb} & {\bf Av. amb} \\ 
+    \hline  
+     %\texttt{br-fr} & 57,305  & 702,328  & 4,668 & 2,668 & 1,000  & 1,000 & 603  & 1.07 \\
+     \texttt{br-fr} & 57,305  & 4,668 & 2,668 & 2,000  & 603  & 3.06 \\
+    \hline  
+     %\texttt{mk-en} & 190,493  & 4,259,338  & 19,747  & 17,747  & 1,000 & 1,000 & 13,134  & 1.86  \\
+     \texttt{mk-en} & 190,493  & 19,747  & 17,747  & 2,000 & 13,134  & 3.06  \\
+    \hline
+     %\texttt{eu-es} & 765,115  & 10,190,079  & 87,907  & 85,907   & 1,000 & 1,000 & 1,806  & 1.30 \\
+     \texttt{eu-es} & 765,115  & 87,907  & 85,907  & 2,000 & 1,806  & 3.11 \\
+    \hline
+     %\texttt{en-es} & 1,467,708 & 30,154,098  & 312,162 & 310,162 & 1,000 & 1,000 & 2,082  & 1.08 \\
+     \texttt{en-es} & 1,467,708 & 312,162 & 310,162 & 2,000 & 2,082  & 2.28 \\
+    \hline  
+ \end{tabular}
+ \caption{Statistics about the source corpora. The column {\bf other}
+   gives the number of sentences reserved for testing (1,000) and
+   development (1,000). The column {\bf no.\ amb} gives the number of
+   unique tokens with more than one possible translation. The column
+   {\bf av.\ amb} gives the average number of translations per ambiguous
+   word. This is calculated by looking up each word in the corpus in
+   the bilingual dictionary of the MT system and dividing the total
+   number of translation by the number of words. Both {\bf av.\ amb}
+   and {\bf no.\ amb} are calculated over the whole corpus.}
+ \label{table:input-corp}
+\end{center}
+\end{table}
+
+%\todo[inline]{FSM: In Table \ref{table:input-corp} it would be good to
+%have the average ambiguity per ambiguous words, not all words, since
+%LER only focuses no ambiguous words.}
+
+The annotation process proceeds as follows: First we word-align the
+corpus to extract a set of word alignments, which are correspondences
+between words in sentences in the source side of the parallel corpus
+and those in the target side. Any aligner may be used, but in this
+paper we use {\sc giza++} \citep{och03a}.\footnote{The exact
+  configuration of {\sc giza++} used is equivalent to running the {\sc
+    Moses} toolkit \citep{koehn07} in default configuration up to step
+  three of training.}  We then use
+these alignments along with the bilingual dictionary of the MT system
+in question to extract only those sentences where: there is at least
+one ambiguous word; that ambiguous word is aligned to a single word in
+the target language; and the word it is aligned to in the target
+language is found in the bilingual dictionary of the MT
+system. Sentences where there are no ambiguous words (approximately
+90\%, see Table~\ref{table:input-corp}) are discarded. The source side
+of the extracted sentence is then passed through the lexical transfer
+module, which returns all the possible translations, and for each
+ambiguous word, the translation is selected which is found aligned in
+the reference.
+
+\begin{table}
+\begin{center}
+ \begin{tabular}{|l|r|r|r|r|r|}
+    \hline 
+    {\bf Pair}      & {\bf Lines} & {\bf SL words} & {\bf TL words} & {\bf Amb. words} & {\bf \% ambig} \\ 
+    \hline 
+     \texttt{br-fr} & 1,000 & 13,854  & 13,878  & 1,163 & 8.39   \\
+    \hline 
+     \texttt{mk-en} & 1,000 & 13,441 & 14,228  & 3,872 & 28.80 \\
+    \hline 
+     \texttt{eu-es} & 1,000 & 7,967 & 11,476  & 1,360  & 17.07 \\
+    \hline 
+     \texttt{en-es} & 1,000 & 19,882 & 20,944  & 1,469  & 7.38 \\
+    \hline 
+ \end{tabular} 
+\end{center}
+ \caption{Statistics about the test corpora. The columns \textbf{amb.\ words} and \textbf{\%~ambig} gives 
+    the number of word with more than one translation and the percentage of SL words which have 
+    more than one translation respectively.}
+  \label{table:test-corp}
+\end{table}
+
+After this process, we selected 1,000 sentence pairs at random for
+testing (\texttt{test}), 1,000 for development (\texttt{dev})\footnote{The development corpus  
+  was used for checking the value for frequency pruning of features.} and 
+left the remainder for training. This is a smaller
+number of sentences than used in typical evaluations (such as the WMT
+series, \cite{callisonburch2012}), but was motivated by the fact that
+the smallest corpus (for Breton--French) after sentence discarding was
+only little over 4,000 sentences long. Table~\ref{table:input-corp}
+gives statistics about the size of the input corpora, and how many
+sentences were left after processing for testing, training and
+development. Table~\ref{table:test-corp} gives information about the
+test corpora.
+
+\subsection{Reference systems}
+
+We compare our method to the following reference (or baseline) systems:
+
+\begin{itemize}
+\item \textbf{Linguist-chosen defaults}. \comment{FSM: This should go
+    to the section explaining Apertium.}  A bilingual dictionary in an
+  Apertium language pair contains correspondences between lexical
+  forms. The dictionaries allow many lexical forms to translate to one
+  lexical form. For example, \emph{ahizpa} and \emph{arreba} in Basque
+  both translate to \emph{hermana} `sister' in Spanish.\footnote{The
+    difference being that \emph{ahizpa} is a sister of a female, and
+    \emph{arreba} is a sister of a male.} But a single lexical form
+  may not have more than one translation without further
+  processing. If there are many possible translations of a lexical
+  form, then one must be marked as the \emph{default} translation. For
+  example, \emph{pisu} in Basque translates to \emph{peso} `weight'
+  and \emph{piso} `flat' in Spanish. If the translation correspondence
+  \emph{pisu $\rightarrow$ peso} is perceived as the most frequent or
+  the most general, it is marked as the default by the linguist who is
+  creating the bilingual dictionary.
+
+\item \textbf{Oracle}.  The results for the oracle system are those
+  achieved by passing the automatically annotated reference
+  translation through the rest of the modules of the MT system. This
+  is included to give an idea of the upper bound for the performance
+  of the lexical-selection module. If all of the lexical choices were
+  made in accordance with the reference, this is the result that would
+  be achieved.
+
+\item \textbf{Target language model} (TLM). One method of lexical
+  selection is to use the existing machine translation system to
+  generate all the possible translations for an input sentence, and
+  then score these translations \emph{on-line} on a model of the
+  target language. The highest scoring sentence is then picked as
+  output. This is the method used by \cite{carbonell06} and
+  \cite{melero07a}.  We used a five-gram language model of surface
+  forms generated from the target-language side of the parallel
+  corpus. %\comment{FSM: Why not a much larger corpus? (I reviewer may ask)}
+%(see section~\ref{sec:eval-corpus}). 
+  This was the best performing system when we compared different
+  approaches, but is impractical for real-world MT because the number
+  of translations to perform grows exponentially with the length of
+  the sentence.  It represents the best that can be achieved using the
+  current systems without access to a parallel corpus. In our work, it
+  was implemented with the IRST language modelling toolkit
+  \cite{federico08a}.
+\end{itemize}
+
+\section{Results}
+\label{sec:eval-ref-results}
+
+As we are working with binary features, we use the implementation of
+generalised iterative scaling available in the {\sc
+  yasmet}\footnote{\url{http://www-i6.informatik.rwth-aachen.de/web/Software/YASMET.html};
+  the compilable version we used is available as part of the Apertium
+  \texttt{lex-tools} package,
+  \url{http://downloads.sourceforge.net/project/apertium/apertium-lex-tools/apertium-lex-tools-0.1.0.tar.gz}.}
+to calculate the feature weights.  After learning the feature sets and
+weights, we compute the evaluation measures described in
+section~\ref{ss:measures}. There is an option to remove features which
+occur less than a certain number of times in the training corpus. This
+is referred to as the feature pruning frequency threshold --- features
+occuring less than threshold are discarded. The value was set
+experimentally. Values of between two and seven were tested, and the
+ones which provided the best improvement on the development corpus
+were selected; they happen to come close to the rule-of-thumb value of
+five that \citet[p.~596]{Manning99b} find to be effective.
+Table~\ref{table:maxent-features} shows the number of features that
+have eventually been sed for each language pair.
+
+%\section{Results}
+
+\begin{table}
+ \begin{center}
+    \begin{tabular}{|l|r|r|}
+        \hline
+        \textbf{Pair} & \textbf{Pruned} & \textbf{\# features} \\\hline
+        \hline
+        {\bf br-fr}  & $< 5$              & 5,277    \\ 
+        \hline
+        {\bf mk-en} & $< 7$              & 205,494 \\ 
+        \hline
+        {\bf eu-es} & $< 7$              & 196,024\\
+        \hline
+        {\bf en-es} & $< 7$              & 195,605\\
+        \hline
+    \end{tabular}
+ \end{center}
+ \caption{Number of features (rules) in each rule set and pruning frequency.}
+ \label{table:maxent-features}
+\end{table}
+
+ Evaluation results are presented in 
+%tables~\ref{table:eval-para-results} and \ref{table:eval-mono-results}.
+ table \ref{table:eval-mono-results}, which compares the results of
+ the new approach with respect to the \emph{default behaviour} (the
+ linguist-chosen defaults,\footnote{The high error rate
+   for the Breton--French pair may be as a result of having the
+   linguistic defaults tuned to a different domain than that of the
+   corpus.}  with respect to the \emph{oracle} (which represents the
+ upper bound to performance), and with respect to the results obtained
+ by using the target-language model online, for each of the language
+ pairs in Apertium with respect to our two evaluation metrics.
+ Significant improvements with respect to the results obtained using
+ the target-language model online are apparent with the Breton--French
+ ---the pair with the least data--- and the English--Spanish language
+ pairs. In the remaining cases, the maximum-entropy method comes close
+ to the target-language model performance, at a much smaller
+ computational cost.\comment{FSM: This needs to be proved, and for
+   this computation times needs to be provided.}
+
+%\begin{table}
+%\begin{tabular}{lrrrr}
+%          & \texttt{br--fr} & \texttt{mk--en} & \texttt{eu-es} & \texttt{en-es} \\
+%\hline
+%\textbf{Ling}   & 7,100         & 13,700       & 7,500        & 10,200  \\
+%\textbf{TLM}    & 630          & ---             & 300         & 940 \\
+%\textbf{MaxEnt} & 5,800  & 30           & 2,400        & 3,300  \\
+%\hline
+%\end{tabular}
+%\caption{Speed of the systems in words/second.}
+%\end{table}
+
+Improvements with
+ respect to the target-language-model performance are very likely due
+ to the effective use the maximum-entropy model makes of information
+ about the relevant source-language contexts and their translations,
+ through the weighting of features representing those source-language
+ contexts across the whole training corpus. \comment{First attempt at
+   an explanation. We should have an explanation for this that does
+   not act as referee magnet}
+
+\begin{table}
+ \begin{center}
+
+  \begin{tabular}{|l|l|r|r|r||r|}
+    \hline
+    \multirow{2}{*}{{\bf Pair}}  & \multirow{2}{*}{{\bf Metric}} & \multicolumn{4}{|c|}{{\bf System}} \\ \cline{3-6}
+                                 &              & {\tt Ling} & {\tt TLM} & \texttt{MaxEnt} & \texttt{Oracle} \\
+    \hline % tlm-alig: [44.5, 50.2]
+    \multirow{2}{*}{{\bf br-fr}} & {\sc ler} (\%)     & [54.8, 60.7] & [44.2, 50.5]  & {\bf [40.8, 46.9]} & [0.0, 0.0]      \\ 
+                                 & {\sc bleu} (\%)    & [14.5, 16.4] & [15.4, 17.3]  & {\bf [14.8, 16.6]} & [16.7, 18.6]     \\ 
+    \hline % tlm-alig: 
+    \multirow{2}{*}{{\bf mk-en}} & {\sc ler} (\%)     & [28.8, 32.6] & [26.8, 30.5]  & [25.2, 28.8] & [0.0, 0.0]    \\ 
+                                 & {\sc bleu} (\%)    & [28.6, 31.0] & [30.7, 32.3]  & [29.1, 31.5] & [30.9, 33.3]    \\ 
+    \hline % tlm-alig: 
+    \multirow{2}{*}{{\bf eu-es}} & {\sc ler} (\%)      & [43.6, 48.8] & [38.8, 44.2]  & [40.9, 46.2] & [0.0, 0.0]     \\ 
+                                 & {\sc bleu} (\%)     & [10.1, 12.0] & [10.6, 12.6]  & [10.3, 12.2] & [11.5, 13.5]     \\ 
+    \hline % tlm-alig: [10.3, 13.9]
+    \multirow{2}{*}{{\bf en-es}} & {\sc ler} (\%)      & [20.5, 24.9] & [15.1, 18.9]  & {\bf [10.4, 13.8]} & [0.0, 0.0]     \\ 
+                                 & {\sc bleu} (\%)     & [21.5, 23.4] & [21.9, 23.8]  & {\bf [22.2, 24.1]} & [22.8, 24.7]     \\ 
+    \hline
+  \end{tabular}
+
+ \end{center}
+ \caption{LER and BLEU scores with 95\% confidence intervals for the reference
+   systems on the test corpora. The \texttt{max-ent} system has been trained using 
+   fractional counts. The results in bold face show statistically significant
+   improvements for the maximum-entropy model compared to the target language model 
+   according to pair-bootstrap resampling.}
+  \label{table:eval-mono-results}
+\end{table}
+
+\todo[inline]{FSM: I think we should better discuss the results and try to see if we disambiguate all the categories the same way.
+Think about this and also about reporting the number of times not feature gets activated.
+It could be that your method is better but you do not seen in the figures because most of the times there is no context to disambiguate}
+
+\todo[inline]{FSM: The results need a more in-depth discussion. For
+  instance, why is that LER improves and BLEU gets worse at the same
+  time for br-fr?}
+
+\todo[inline]{FSM: I would further analyse the results obtained by
+  giving LER for different lexical categories. It may happen that we
+  are better at resolving the ambiguity of nouns that the ambiguity of
+  verbs, just to give an example.}
+
+\section{Conclusions}
+
+\comment{First try at a set of conclusions by Mikel} This paper has
+presented a method to perform lexical selection in a rule-based
+machine translation system, and one that can be trained in an
+unsupervised way, that is, without the need for an annotated corpus,
+in this case a word-aligned bilingual corpus: one just needs a
+source-language corpus, a statistical target-language model, and the
+rule-based machine translation system. The input to the method is
+simply the part-of-speech tagged source text in which each word is
+annotated with all the translations provided by the bilingual
+dictionary in the system: this makes it applicable to almost any
+rule-based machine translation system. The system uses a
+maximum-entropy formalism for lexical selection as \cite{berger1996}
+and \cite{marechek10}, but instead of counting actual lexical
+selection events in an annotated corpus, it counts fractional
+occurrences of these events as estimated by a target-language
+model. The method is evaluated both intrinsically (just looking at the
+actual lexical selection events) and extrinsically (measuring the
+quality of machine translation). Results on four language pairs using
+the Apertium \citep{forcada2011apertium} machine translation system
+show that the method obtains similar or better results than those
+expensively obtained by scoring an exponential number of lexical
+selections for each sentence using the target-language model online.
+
+\todo[inline]{FSM: The fact that, to the best of our knowledge, this
+  is the first time fractional counts are used with MaxEnt should be
+  stressed.}
+
+\section*{Acknowledgements}
+
+The authors thank Filip Petkovski for his help in running some of the
+experiments. They also acknowledge partial support from the Ministry
+of Science and Innovation of the Spanish Government through project
+TIN2009-14009-C02-01, and the NILS mobility grant (Abel Predoc
+Research Grant) coordinated by the Universidad Complutense de Madrid.
 
 \bibliographystyle{apalike}
-
 \bibliography{2015-eamt-melsa}

### 2015-02-09 2015-eamt-melsa/2015-eamt-melsa.tex
fix up some formatting issues
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58777 72bbbca6-d526-0410-a7d9-f06f51895060

@@
  N-9018 Romsa \\
- \and
+ \And
  Felipe S\'anchez-Mart\'inez \\
@@
  E-03071 Alacant\\
- \and
+ \And
  Mikel L.\ Forcada \\
@@
 
-\begin{table}
+\begin{table*}
  \begin{center}
@@
  \end{center}
-\end{table}
+\end{table*}
 
@@
 
-\begin{table}
+\begin{table*}
   \centering
@@
   \label{table:lang-pair-stats}
-\end{table}
+\end{table*}
 
@@
 % see section~\ref{sec:ler}. 
-\begin{figure}
+\begin{figure*}
 \begin{center}
@@
 \end{center}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{table}
+\begin{table*}
 \begin{center}
@@
 \end{center}
-\end{table}
+\end{table*}
 
@@
   \label{table:test-corp}
-\end{table}
+\end{table*}
 
@@
   \label{table:eval-mono-results}
-\end{table}
+\end{table*}
 

### 2015-02-09 2015-eamt-melsa/2015-eamt-melsa.tex
12 pages down from 13... 4 pages to cull
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58778 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-\comment{Improve this} Lexical selection in this paper considers for
+%\comment{Improve this} 
+Lexical selection in this paper considers for
 each word a simple source-language context made up of neighbouring
@@
 and returns a finished translation of a specific lexical-selection
-path $g_i$ of sentence $S$. Schematically:
+path $g_i$ of sentence $S$. Figure~\ref{fig:lexselschema} shows
+ this process schematically.
+\begin{figure*}
 \begin{displaymath}
@@
 \end{displaymath}
+ \caption{A schema of the training process}
+ \label{fig:lexselschema}
+\end{figure*}
 
@@
 
-\todo[inline]{FSM: I would explain what comes in the following
-  paragraph (the need for an evaluation corpus and how it is built) in
-  advance. Most probably where the measures are defined.}
+%! \todo[inline]{FSM: I would explain what comes in the following
+%!  paragraph (the need for an evaluation corpus and how it is built) in
+%!  advance. Most probably where the measures are defined.}
 
@@
 \begin{itemize}
-\item \textbf{Linguist-chosen defaults}. \comment{FSM: This should go
-    to the section explaining Apertium.}  A bilingual dictionary in an
+\item \textbf{Linguist-chosen defaults}. 
+%! \comment{FSM: This should go to the section explaining Apertium.}  
+  A bilingual dictionary in an
   Apertium language pair contains correspondences between lexical
@@
  to the target-language model performance, at a much smaller
- computational cost.\comment{FSM: This needs to be proved, and for
-   this computation times needs to be provided.}
+ computational cost.
+
+%! \comment{FSM: This needs to be proved, and for
+%!    this computation times needs to be provided.}
 
@@
  through the weighting of features representing those source-language
- contexts across the whole training corpus. \comment{First attempt at
-   an explanation. We should have an explanation for this that does
-   not act as referee magnet}
+ contexts across the whole training corpus.
+
+%! \comment{First attempt at
+%!    an explanation. We should have an explanation for this that does
+%!    not act as referee magnet}
 
@@
 
+%!
 \todo[inline]{FSM: I think we should better discuss the results and try to see if we disambiguate all the categories the same way.
@@
 It could be that your method is better but you do not seen in the figures because most of the times there is no context to disambiguate}
-
-\todo[inline]{FSM: The results need a more in-depth discussion. For
-  instance, why is that LER improves and BLEU gets worse at the same
-  time for br-fr?}
-
-\todo[inline]{FSM: I would further analyse the results obtained by
-  giving LER for different lexical categories. It may happen that we
-  are better at resolving the ambiguity of nouns that the ambiguity of
-  verbs, just to give an example.}
-
+%! 
+%! \todo[inline]{FSM: The results need a more in-depth discussion. For
+%!   instance, why is that LER improves and BLEU gets worse at the same
+%!   time for br-fr?}
+%! 
+%! \todo[inline]{FSM: I would further analyse the results obtained by
+%!   giving LER for different lexical categories. It may happen that we
+%!   are better at resolving the ambiguity of nouns that the ambiguity of
+%!   verbs, just to give an example.}
+%! 
 \section{Conclusions}
 
-\comment{First try at a set of conclusions by Mikel} This paper has
+%! \comment{First try at a set of conclusions by Mikel} 
+This paper has
 presented a method to perform lexical selection in a rule-based
@@
 
-\todo[inline]{FSM: The fact that, to the best of our knowledge, this
-  is the first time fractional counts are used with MaxEnt should be
-  stressed.}
+%! \todo[inline]{FSM: The fact that, to the best of our knowledge, this
+%!   is the first time fractional counts are used with MaxEnt should be
+%!   stressed.}
 
@@
 
-The authors thank Filip Petkovski for his help in running some of the
-experiments. They also acknowledge partial support from the Ministry
-of Science and Innovation of the Spanish Government through project
-TIN2009-14009-C02-01, and the NILS mobility grant (Abel Predoc
-Research Grant) coordinated by the Universidad Complutense de Madrid.
-
+\emph{Removed for review}
+
+%The authors thank Filip Petkovski for his help in running some of the
+%experiments. They also acknowledge partial support from the Ministry
+%of Science and Innovation of the Spanish Government through project
+%TIN2009-14009-C02-01, and the NILS mobility grant (Abel Predoc
+%Research Grant) coordinated by the Universidad Complutense de Madrid.
+%
 \bibliographystyle{apalike}

### 2015-02-09 2015-eamt-melsa/2015-eamt-melsa.tex
some reference culling
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58779 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 statistical machine translation. Other authors
-\citep{carbonell06,melero07a} have used target-language models to rank
+\citep{melero07a} have used target-language models to rank %@@@ carbonell06
 the translations resulting from all possible combinations of lexical
@@
 --- may also lead to the same problem.  The {\sc bleu} metric \citep{papineni02} is
-included only as it is commonly used to evaluate MT systems. Other
-metrics with similar performance to {\sc bleu} such as {\sc meteor}
-\citep{lavie09} and {\sc nist} \citep{doddington02} are not included
-as we have chosen to focus on lexical-selection performance.
+included only as it is commonly used to evaluate MT systems.
+
+%@@@ Other
+%@@@ metrics with similar performance to {\sc bleu} such as {\sc meteor}
+%@@@ \citep{lavie09} and {\sc nist} \citep{doddington02} are not included
+%@@@ as we have chosen to focus on lexical-selection performance.
 
@@
   target language. The highest scoring sentence is then picked as
-  output. This is the method used by \cite{carbonell06} and
+  output. This is the method used by %@@@ \cite{carbonell06} and
   \cite{melero07a}.  We used a five-gram language model of surface

### 2015-02-09 2015-eamt-melsa/2015-eamt-melsa.tex
some trimming
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58783 72bbbca6-d526-0410-a7d9-f06f51895060

@@
   translation equivalents in the bilingual dictionary of the system.
-%  taking into account all of the lexical selection possibilities.
   These translations are then scored on the target-language model and
@@
 \subsection{Prior work}
-\comment{This section should be ironed out a bit.}  In their
+%@@@ \comment{This section should be ironed out a bit.}  
+In their
 pioneering work, \cite{dagan1994word} used the term
@@
 lexical-selection module but did not provide any experiments. 
-\todo[inline]{FMT: integrate \citep{sanchez07a} somewhere here}
+%@@@ \todo[inline]{FMT: integrate \citep{sanchez07a} somewhere here}
 \cite{marechek10} trained a maximum-entropy lexical selector for their TectoMT rule-based system
@@
 along with a training method for learning the rules from a
-word-aligned parallel corpus. The principal disadvantage with this
-method is that it relies on a parallel corpus for the language pair,
-something which is not always available.
+word-aligned parallel corpus.
+
+% The principal disadvantage with this
+% method is that it relies on a parallel corpus for the language pair,
+% something which is not always available.
 

### 2015-02-09 2015-eamt-melsa/2015-eamt-melsa.tex
more pruning 12->11 pages
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58790 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 As a result of training, each feature in the classifier is assigned a
-weight $\lambda_k^s$, and combining these weights of active features
+weight $\lambda_k^s$. Combining these weights of active features
 as in equation~(\ref{eq:max-ent-sum}) yields the probability of a
@@
 in the corpus, using a development corpus to guide pruning (see section~\ref{sec:eval-ref-results}).
-\comment{Reword this to avoid repetition} The method used here
+%@@@ \comment{Reword this to avoid repetition} 
+The method used here
 for lexical selection is analogous to the method used by
@@
 \(\mathcal{G}\) for the translation of Basque sentences containing the
-word \emph{arrain} into Spanish.\footnote{The word \emph{arrain} `fish' has two translations
+word \emph{arrain}.\footnote{The word \emph{arrain} `fish' has two translations
 in Spanish, the first \emph{pez} refers to the living animal, whereas
@@
 output as compared to a gold-standard corpus. The lexical 
-transfer output is the result if looking up the translations of the 
-source-language \emph{lexical forms} --- combinations of lemma and morphological
+transfer output is the result of looking up the translations of the 
+source-language \emph{lexical forms} --- combinations of lemma and 
 tags --- in the bilingual dictionary.
@@
 The lexical-selection module should be as language-independent as
-possible. To that end, the language pairs tested should show as wide a
+possible. To that end, the language pairs tested show a wide
 variety of linguistic phenomena. It is also important that the
 methodology be as applicable to lesser-resourced and marginalised
-languages, as to major languages.
+languages as to major languages.
 
@@
   41375.\footnote{\url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-br-fr}}
-  This is the only available machine translation system between Breton
-  and French. It has been developed part-time over a number of years.
+%  This is the only available machine translation system between Breton
+%  and French. It has been developed part-time over a number of years.
 
@@
   in Apertium was created specifically for the purposes of running
-  lexical-selection experiments. The resources reused from other pairs
-  were the English morphological dictionary from the Icelandic and
-  English pair \citep{brandt11}, the Macedonian morphological
-  dictionary and constraint grammar from the Macedonian and Bulgarian
-  pair \citep{rangelov11}, and the SETimes parallel corpus
-  \citep{tyers10}.  The work was carried out over a period of eight
-  days, and consisted of creating the bilingual dictionary and
-  transfer rules. The bilingual dictionary was created by tagging both
-  sides of the parallel corpus, word-aligning them with {\sc giza++}
-  \citep{och03a} and extracting the probabilistic lexicon. Entries
-  from this lexicon were checked manually according to frequency and
-  included in the bilingual dictionary of the machine translation
-  system. The most probable entry was checked to ensure that it was 
+  lexical-selection experiments. 
+% The resources reused from other pairs
+%  were the English morphological dictionary from the Icelandic and
+%  English pair \citep{brandt11}, the Macedonian morphological
+%  dictionary and constraint grammar from the Macedonian and Bulgarian
+%  pair \citep{rangelov11}, and the SETimes parallel corpus
+  The lexical resources for the pair were tuned to the SETimes parallel corpus
+  \citep{tyers10}.  
+% The work was carried out over a period of eight
+%  days, and consisted of creating the bilingual dictionary and
+%  transfer rules. The bilingual dictionary was created by tagging both
+%  sides of the parallel corpus, word-aligning them with {\sc giza++}
+%  \citep{och03a} and extracting the probabilistic lexicon. Entries
+%  from this lexicon were checked manually according to frequency and
+%  included in the bilingual dictionary of the machine translation
+%  system. 
+  The most probable entry from automatic word alignment of this 
+  corpus using {\sc giza++} \citep{och03a} was checked to ensure that it was 
   an adequate translation, and if so marked as the 
@@
   average number of translations per word is much higher (see table~\ref{table:lang-pair-stats})
-than in other pairs.  The transfer rules were written
-  by hand to treat translation problems in the corpus. The version of
-  the software used in this paper is SVN revision
-  41476.\footnote{\url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-mk-en}}.
+than in other pairs.  
+% The transfer rules were written
+%  by hand to treat translation problems in the corpus. The version of
+%  the software used in this paper is SVN revision
+%  41476.\footnote{\url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-mk-en}}.
 \item[Basque--Spanish:] The development of the Basque to Spanish pair
-  is described in \cite{ginesti09}. The bilingual dictionary was taken
-  from the free/open-source Matxin system \citep{mayor2011}. 
+  is described in \cite{ginesti09}. 
+% The bilingual dictionary was taken
+%  from the free/open-source Matxin system \citep{mayor2011}. 
   Alternative translations were included in the bilingual
@@
   by the language pair developer, is marked as the \emph{default}
-  translation and non-default translations were added according to
-  their frequency. The version used in this paper is SVN revision
-  41387.\footnote{\url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-en-es}}
+  translation and non-default translations added according to
+  frequency. 
+% The version used in this paper is SVN revision
+%  41387.\footnote{\url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-en-es}}
 \end{description}
@@
    between the reference set $T_r(s_i)$ and the test set $T_t(s_i)$ of
-   translations; as a result, the error rate for this sentence is
+   translations; thus, the error rate for this sentence is
    50\%.}
@@
 lexical-selection module returns more than one translation, the first
-translation is selected. \comment{FSM: How many times does this
-  happens?. I would move this last sentence to an appendix.}  The
+translation is selected. 
+%! \comment{FSM: How many times does this
+%!  happens?. I would move this last sentence to an appendix.}  
+The
 function $T_r(s_i)$ returns the set of reference translations which
@@
 \emph{bootstrap resampling} method~\citep{efron94} as described
-by~\citet{koehn04} and \citet{zhang04}.\footnote{Broadly speaking,
-  this involves iteratively taking a random sample (with replacements)
-  from the test set and computing the score for that sample. The
-  scores are sorted and, presuming we want the 95\% confidence
-  interval, the top- and bottom-2.5\% are removed. The highest and
-  lowest remaining are given as the interval. For more detail, refer
-  to the references provided.} In all cases, bootstrap resampling will
+by~\citet{koehn04}.%\footnote{Broadly speaking, % and \citet{zhang04}
+%  this involves iteratively taking a random sample (with replacements)
+%  from the test set and computing the score for that sample. The
+%  scores are sorted and, presuming we want the 95\% confidence
+%  interval, the top- and bottom-2.5\% are removed. The highest and
+%  lowest remaining are given as the interval. For more detail, refer
+%  to the references provided.} 
+In all cases, bootstrap resampling will
 be carried out for 1,000 iterations. Where the $p = 0.05$ confidence
@@
   previously described in \cite{tyers09b}, and also some new data. The
-  whole corpus has been made available online through the {\sc opus}
-  collection.\footnote{\url{http://opus.lingfil.uu.se/OfisPublik.php}.}
+  corpus has been made available online through {\sc opus}.\footnote{\url{http://opus.lingfil.uu.se/OfisPublik.php}.}
 \item \textbf{South-East European Times} (\texttt{SETimes}): Described
@@
   European Union official languages. We are using the English--Spanish
-  data from version 7 of the corpus.
+  data from version 7.
 \end{itemize}

### 2015-02-09 2015-eamt-melsa/2015-eamt-melsa.tex
comment out worked example... one page less
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58791 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 tagger in a rule-based machine translation system. 
-Table~\ref{tb:workedexample} shows how a collection of samples
-\(\mathcal{G}\) for the translation of Basque sentences containing the
-word \emph{arrain}.\footnote{The word \emph{arrain} `fish' has two translations
-in Spanish, the first \emph{pez} refers to the living animal, whereas
-\emph{pescado} refers to the animal as food.} 
+%Table~\ref{tb:workedexample} shows how a collection of samples
+%\(\mathcal{G}\) for the translation of Basque sentences containing the
+%word \emph{arrain}.\footnote{The word \emph{arrain} `fish' has two translations
+%in Spanish, the first \emph{pez} refers to the living animal, whereas
+%\emph{pescado} refers to the animal as food.} 
 %\comment{FSM: The correct translations
@@
 %  here?}
-
-\begin{table*}
- \begin{center}
-  \begin{tabular}{|c|l|r|}
-    \hline
-       & \textbf{Sentence} & $p(g_i|S)$ \\
-    \hline % Los pescados grandes peque\~{n}o come
-     \(S_1\) & \emph{Arrain handiak txikia jaten du}  &  \\
-                        $\tau(g_{11},S_1)$ & El pez grande se come el peque\~{n}o &  0.830 \\
-                        $\tau(g_{12},S_1)$ & El pescado grande se come el peque\~{n}o & 0.134 \\
-                        $\tau(g_{13},S_1)$ & El pescado capaz se come el peque\~{n}o & 0.034\\
-                        $\tau(g_{14},S_1)$ & El pez capaz se come el peque\~{n}o  & 0.001\\
-    \hline % El padre el pescado nos ha preparado para cenar
-     \(S_2\) & \emph{Aitak arraina prestatu digu afaltzeko} &  \\
-                        $\tau(g_{21},S_2)$ & Padre nos ha preparado pescado para cenar & 0.846 \\
-                        $\tau(g_{22},S_2)$ & Pap\'{a} nos ha preparado pescado para cenar & 0.134 \\
-                        $\tau(g_{23},S_2)$ & Padre nos ha preparado pez para cenar & 0.015 \\
-                        $\tau(g_{24},S_2)$ & Pap\'{a} nos ha preparado pez para cenar & 0.005 \\
-    \hline % Aqu\'{\i} el pescado muy es dulce
-    $S_3$  & \emph{Hemen arraina oso goxoa da}  & \\
-                        $\tau(g_{31},S_3)$ & Aqu\'{\i} el pescado es muy rico & 0.912 \\
-                        $\tau(g_{32},S_3)$ & Aqu\'{\i} el pescado es muy suave & 0.067 \\
-                        $\tau(g_{33},S_3)$ & Aqu\'{\i} el pez es muy rico & 0.011 \\
-                        $\tau(g_{34},S_3)$ & Aqu\'{\i} el pez es muy suave & 0.008 \\
-                        $\tau(g_{35},S_3)$ & Aqu\'{\i} el pez es muy dulce & 0.001 \\
-                        $\tau(g_{36},S_3)$ & Aqu\'{\i} el pescado es muy dulce & 0.001 \\
-    \hline % Un pescado grande del mar es
-    \(S_4\) & \emph{Itsasoko arrain handi bat da} & \\
-                        $\tau(g_{41},S_4)$ & Es un pez grande del mar & 0.595 \\
-                        $\tau(g_{42},S_4)$ & Es un pescado grande del mar & 0.403 \\
-                        $\tau(g_{43},S_4)$ & Es un pez capaz del mar & 0.001 \\
-                        $\tau(g_{44},S_4)$ & Es un pescado capaz del mar & 0.001 \\
-    \hline % El pescado grande por esos serv\'{\i}an para alimentar
-    $S_5$ &  \emph{Arrain handiak horiez baliatzen ziren elikatzeko} &  \\
-                        $\tau(g_{51},S_5)$ & Los peces grandes se nutr\'{\i}an de esos & 0.862 \\
-                        $\tau(g_{52},S_5)$ & Los pescados grandes se nutr\'{\i}an de esos & 0.121 \\
-                        $\tau(g_{53},S_5)$ & Los peces grandes se alimentaban de esos & 0.012 \\
-                        $\tau(g_{54},S_5)$ & Los peces capaces se alimentaban de esos & 0.001 \\
-                        $\tau(g_{55},S_5)$ & Los pescados grandes se alimentaban de esos & 0.001 \\
-                        $\tau(g_{56},S_5)$ & Los pescados capaces se alimentaban de esos & 0.001 \\
-                        $\tau(g_{57},S_6)$ & Los peces capaces se nutr\'{\i}an de esos & 0.001 \\
-                        $\tau(g_{58},S_7)$ & Los pescados capaces se nutr\'{\i}an de esos & 0.001 \\
-    \hline
-
-  \end{tabular}
-  \caption{A sample of source-language sentences, the output of the
-    possible lexical-selection paths after scoring on a
-    target-language model. The scores are normalised as fractional
-    counts, which are used as inputs to the maximum-entropy training
-    process. In all sentences, the resultant translation with the highest share of the 
-    probability mass is the most adequate.}
-   \label{tb:workedexample}
- \end{center}
-\end{table*}
+% 
+% \begin{table*}
+%  \begin{center}
+%   \begin{tabular}{|c|l|r|}
+%     \hline
+%        & \textbf{Sentence} & $p(g_i|S)$ \\
+%     \hline % Los pescados grandes peque\~{n}o come
+%      \(S_1\) & \emph{Arrain handiak txikia jaten du}  &  \\
+%                         $\tau(g_{11},S_1)$ & El pez grande se come el peque\~{n}o &  0.830 \\
+%                         $\tau(g_{12},S_1)$ & El pescado grande se come el peque\~{n}o & 0.134 \\
+%                         $\tau(g_{13},S_1)$ & El pescado capaz se come el peque\~{n}o & 0.034\\
+%                         $\tau(g_{14},S_1)$ & El pez capaz se come el peque\~{n}o  & 0.001\\
+%     \hline % El padre el pescado nos ha preparado para cenar
+%      \(S_2\) & \emph{Aitak arraina prestatu digu afaltzeko} &  \\
+%                         $\tau(g_{21},S_2)$ & Padre nos ha preparado pescado para cenar & 0.846 \\
+%                         $\tau(g_{22},S_2)$ & Pap\'{a} nos ha preparado pescado para cenar & 0.134 \\
+%                         $\tau(g_{23},S_2)$ & Padre nos ha preparado pez para cenar & 0.015 \\
+%                         $\tau(g_{24},S_2)$ & Pap\'{a} nos ha preparado pez para cenar & 0.005 \\
+%     \hline % Aqu\'{\i} el pescado muy es dulce
+%     $S_3$  & \emph{Hemen arraina oso goxoa da}  & \\
+%                         $\tau(g_{31},S_3)$ & Aqu\'{\i} el pescado es muy rico & 0.912 \\
+%                         $\tau(g_{32},S_3)$ & Aqu\'{\i} el pescado es muy suave & 0.067 \\
+%                         $\tau(g_{33},S_3)$ & Aqu\'{\i} el pez es muy rico & 0.011 \\
+%                         $\tau(g_{34},S_3)$ & Aqu\'{\i} el pez es muy suave & 0.008 \\
+%                         $\tau(g_{35},S_3)$ & Aqu\'{\i} el pez es muy dulce & 0.001 \\
+%                         $\tau(g_{36},S_3)$ & Aqu\'{\i} el pescado es muy dulce & 0.001 \\
+%     \hline % Un pescado grande del mar es
+%     \(S_4\) & \emph{Itsasoko arrain handi bat da} & \\
+%                         $\tau(g_{41},S_4)$ & Es un pez grande del mar & 0.595 \\
+%                         $\tau(g_{42},S_4)$ & Es un pescado grande del mar & 0.403 \\
+%                         $\tau(g_{43},S_4)$ & Es un pez capaz del mar & 0.001 \\
+%                         $\tau(g_{44},S_4)$ & Es un pescado capaz del mar & 0.001 \\
+%     \hline % El pescado grande por esos serv\'{\i}an para alimentar
+%     $S_5$ &  \emph{Arrain handiak horiez baliatzen ziren elikatzeko} &  \\
+%                         $\tau(g_{51},S_5)$ & Los peces grandes se nutr\'{\i}an de esos & 0.862 \\
+%                         $\tau(g_{52},S_5)$ & Los pescados grandes se nutr\'{\i}an de esos & 0.121 \\
+%                         $\tau(g_{53},S_5)$ & Los peces grandes se alimentaban de esos & 0.012 \\
+%                         $\tau(g_{54},S_5)$ & Los peces capaces se alimentaban de esos & 0.001 \\
+%                         $\tau(g_{55},S_5)$ & Los pescados grandes se alimentaban de esos & 0.001 \\
+%                         $\tau(g_{56},S_5)$ & Los pescados capaces se alimentaban de esos & 0.001 \\
+%                         $\tau(g_{57},S_6)$ & Los peces capaces se nutr\'{\i}an de esos & 0.001 \\
+%                         $\tau(g_{58},S_7)$ & Los pescados capaces se nutr\'{\i}an de esos & 0.001 \\
+%     \hline
+%   
+%   \end{tabular}
+%   \caption{A sample of source-language sentences, the output of the
+%     possible lexical-selection paths after scoring on a
+%     target-language model. The scores are normalised as fractional
+%     counts, which are used as inputs to the maximum-entropy training
+%     process. In all sentences, the resultant translation with the highest share of the 
+%     probability mass is the most adequate.}
+%    \label{tb:workedexample}
+%  \end{center}
+% \end{table*}
 
@@
 \emph{bootstrap resampling} method~\citep{efron94} as described
-by~\citet{koehn04}.%\footnote{Broadly speaking, % and \citet{zhang04}
+by~\citet{koehn04}. %\footnote{Broadly speaking, % and \citet{zhang04}
 %  this involves iteratively taking a random sample (with replacements)
@@
 translation in Swedish. The size of the annotated corpus is not
-mentioned. \cite{specia2005a} use a parallel corpus, but eschew using
-word alignment in favour of using a bilingual dictionary and heuristic
-rules based on part-of-speech and relative position. The corpus was
-reviewed manually before being used in further experiments
-\citep{specia2005}.
+mentioned. 
+%\cite{specia2005a} use a parallel corpus, but eschew using
+%word alignment in favour of using a bilingual dictionary and heuristic
+%rules based on part-of-speech and relative position. The corpus was
+%reviewed manually before being used in further experiments
+%\citep{specia2005}.
 

### 2015-02-10 2015-eamt-melsa/2015-eamt-melsa.tex
down to 9
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58800 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-The remainder of the article is organised as follows: 
-\todo[inline]{FMT: Add article structure}
+%The remainder of the article is organised as follows: 
+%\todo[inline]{FMT: Add article structure}
 
@@
 problem. They learn a separate maximum-entropy classifier for each SL
-word form,\comment{[:we do too, integrate this later]} using SL context to
+word form,
+%\comment{[:we do too, integrate this later]} 
+using SL context to
 distinguish between possible translations. These classifiers are then
 incorporated into the translation model of their word-based SMT
-system.\comment{[how is our work different?]}  In their approach, a
+system.
+%\comment{[how is our work different?]}  
+In their approach, a
 classifier consists of a set of binary feature functions and
@@
 --- although for all pairs there is a parallel corpus available, which
-is used for evaluation (see Section \ref{ss:measures}). Table~\ref{table:lang-pair-stats} 
-presents statistics about the development status of each of the language 
-pairs chosen, including the number of bilingual dictionary entries, coverage of
-test corpora and translational ambiguity of the bilingual dictionary. Translation
-ambiguity is provided for both all words, and for only ambiguous open-category 
-words. In the experiments, lexical-selection rules are learnt for the open categories
-of adjective, noun and verb.
-
-\begin{table*}
-  \centering
-  \begin{tabular}{|l|r|r|r|r|r|}
-    \hline
-    \multirow{2}{*}{\textbf{Pair}} & \textbf{Lexical} & \multirow{2}{*}{\textbf{Ambig.}} & \textbf{Na\"{i}ve}& \textbf{Total} & \textbf{Ambig. of}\\\
-                          & \textbf{items} &                            & \textbf{coverage} & \textbf{ambig.} & \textbf{ambig. words}\\
-    \hline
-    \texttt{br-fr} & 27,690 & 235  &  94.47 \%  & 1.02 & 2.5 \\
-    \texttt{mk-en} & 33,229 & 1,323 & 92.17 \%  & 1.37 & 2.8 \\
-    \texttt{eu-es} & 19,323 & 455 & 91.70 \% & 1.08 & 2.4 \\
-    \texttt{en-es} & 30,186 & 822 & 98.08 \% & 1.05 & 2.2 \\
-    \hline
-  \end{tabular}
-  \caption{Statistics about development status of language pairs. The second
-    column gives the number of entries in the bilingual dictionaries of the systems;
-    the third column gives the total number of source language words with more 
-    than one translation in the target language; the fourth column gives the coverage
-    of the dictionaries on the test corpus (see section~\ref{sec:eval-corpus}); the fifth
-    column gives the total mean ambiguity for open classes (nouns, verbs, adjectives);
-    and the final column gives the mean ambiguity for ambiguous words in open classes.}
-  \label{table:lang-pair-stats}
-\end{table*}
+is used for evaluation (see Section \ref{ss:measures}). 
+% Table~\ref{table:lang-pair-stats} 
+%presents statistics about the development status of each of the language 
+%pairs chosen, including the number of bilingual dictionary entries, coverage of
+%test corpora and translational ambiguity of the bilingual dictionary. Translation
+%ambiguity is provided for both all words, and for only ambiguous open-category 
+%words. In the experiments, lexical-selection rules are learnt for the open categories
+%of adjective, noun and verb.
+
+% \begin{table*}
+%   \centering
+%   \begin{tabular}{|l|r|r|r|r|r|}
+%     \hline
+%     \multirow{2}{*}{\textbf{Pair}} & \textbf{Lexical} & \multirow{2}{*}{\textbf{Ambig.}} & \textbf{Na\"{i}ve}& \textbf{Total} & \textbf{Ambig. of}\\\
+%                           & \textbf{items} &                            & \textbf{coverage} & \textbf{ambig.} & \textbf{ambig. words}\\
+%     \hline
+%     \texttt{br-fr} & 27,690 & 235  &  94.47 \%  & 1.02 & 2.5 \\
+%     \texttt{mk-en} & 33,229 & 1,323 & 92.17 \%  & 1.37 & 2.8 \\
+%     \texttt{eu-es} & 19,323 & 455 & 91.70 \% & 1.08 & 2.4 \\
+%     \texttt{en-es} & 30,186 & 822 & 98.08 \% & 1.05 & 2.2 \\
+%     \hline
+%   \end{tabular}
+%   \caption{Statistics about development status of language pairs. The second
+%     column gives the number of entries in the bilingual dictionaries of the systems;
+%     the third column gives the total number of source language words with more 
+%     than one translation in the target language; the fourth column gives the coverage
+%     of the dictionaries on the test corpus (see section~\ref{sec:eval-corpus}); the fifth
+%     column gives the total mean ambiguity for open classes (nouns, verbs, adjectives);
+%     and the final column gives the mean ambiguity for ambiguous words in open classes.}
+%   \label{table:lang-pair-stats}
+% \end{table*}
 
@@
 \item[Breton--French:] Development of the Breton--French pair has been
-  described in two articles \citep{tyers09b,tyers10b}. The bilingual
+  described in \citep{tyers10b}. The bilingual % tyers09b
   dictionaries were not built with polysemy in mind from the outset,
@@
   result of attempting to include all possible translations, the
-  average number of translations per word is much higher (see table~\ref{table:lang-pair-stats})
+  average number of translations per word is much higher % (see table~\ref{table:lang-pair-stats})
 than in other pairs.  
@@
 \subsubsection{Machine translation performance} 
-\comment{This part of the text needs compressing and merging} This is
+%! \comment{This part of the text needs compressing and merging} 
+This is
 an extrinsic evaluation, which ideally would test how much the system
@@
 successful lexical selections may not lead to successful translations
-due to inadequate transfer of morphological features. For example, if
-we translate the Spanish phrase \emph{La gente le dice que no venga}
-`The people tell him not to come', where \emph{decir} can be
-translated as `say' or `tell' in English, and our machine translation
-system generates `The people \emph{tells} him not to come', it would
-be counted the same as if it were `The people \emph{says} him not to
-come', although the lexical selection made was more
-adequate. Generation errors --- where a word is found in the bilingual
-dictionary, but not in the target language morphological dictionary
---- may also lead to the same problem.  The {\sc bleu} metric \citep{papineni02} is
+due to inadequate transfer of morphological features. 
+% For example, if
+% we translate the Spanish phrase \emph{La gente le dice que no venga}
+% `The people tell him not to come', where \emph{decir} can be
+% translated as `say' or `tell' in English, and our machine translation
+% system generates `The people \emph{tells} him not to come', it would
+% be counted the same as if it were `The people \emph{says} him not to
+% come', although the lexical selection made was more
+% adequate. Generation errors --- where a word is found in the bilingual
+% dictionary, but not in the target language morphological dictionary
+% --- may also lead to the same problem.  
+The {\sc bleu} metric \citep{papineni02} is
 included only as it is commonly used to evaluate MT systems.
@@
 language pair using a process similar to that described by
-\cite{vickrey05}. The decision was made to automatically annotate
-instead of hand annotate for a number of reasons. Firstly, the time
-involved in hand-annotating approximately 30 pages of text, per
-language pair, would be prohibitive. Secondly, linguists were not
-available for all of the language pairs. One disadvantage
-to this method is that the set of
-translations for each ambiguous word may not be complete as it will
-only include the translation which is found in the corresponding
-sentence in the parallel corpus. Further disadvantages are that 
-not all sentence pairs in the parallel corpus will have translations
-that are able to be generated by the MT system (for
-example the reference translation has a word which is not found in
-the bilingual dictionaries on the MT system).
+\cite{vickrey05}.
+
+% The decision was made to automatically annotate
+% instead of hand annotate for a number of reasons. Firstly, the time
+% involved in hand-annotating approximately 30 pages of text, per
+% language pair, would be prohibitive. Secondly, linguists were not
+% available for all of the language pairs. One disadvantage
+% to this method is that the set of
+% translations for each ambiguous word may not be complete as it will
+% only include the translation which is found in the corresponding
+% sentence in the parallel corpus. Further disadvantages are that 
+% not all sentence pairs in the parallel corpus will have translations
+% that are able to be generated by the MT system (for
+% example the reference translation has a word which is not found in
+% the bilingual dictionaries on the MT system).
 
@@
   \emph{pisu $\rightarrow$ peso} is perceived as the most frequent or
-  the most general, it is marked as the default by the linguist who is
-  creating the bilingual dictionary.
+  the most general, it is marked as the default by the author of the dictionary. % linguist who is
+%  creating the bilingual dictionary.
 
@@
   is included to give an idea of the upper bound for the performance
-  of the lexical-selection module. If all of the lexical choices were
-  made in accordance with the reference, this is the result that would
-  be achieved.
+  of the lexical-selection module. % If all of the lexical choices were
+%  made in accordance with the reference, this is the result that would
+%  be achieved.
 
@@
   output. This is the method used by %@@@ \cite{carbonell06} and
-  \cite{melero07a}.  We used a five-gram language model of surface
-  forms generated from the target-language side of the parallel
-  corpus. %\comment{FSM: Why not a much larger corpus? (I reviewer may ask)}
-%(see section~\ref{sec:eval-corpus}). 
-  This was the best performing system when we compared different
-  approaches, but is impractical for real-world MT because the number
-  of translations to perform grows exponentially with the length of
-  the sentence.  It represents the best that can be achieved using the
-  current systems without access to a parallel corpus. In our work, it
-  was implemented with the IRST language modelling toolkit
-  \cite{federico08a}.
+  \cite{melero07a}.  
+% We used a five-gram language model of surface
+%   forms generated from the target-language side of the parallel
+%   corpus. %\comment{FSM: Why not a much larger corpus? (I reviewer may ask)}
+% %(see section~\ref{sec:eval-corpus}). 
+%   This was the best performing system when we compared different
+%   approaches, but is impractical for real-world MT because the number
+%   of translations to perform grows exponentially with the length of
+%   the sentence.  It represents the best that can be achieved using the
+%   current systems without access to a parallel corpus. In our work, it
+%   was implemented with the IRST language modelling toolkit
+%   \cite{federico08a}.
 \end{itemize}

### 2015-02-10 2015-nodalida-ruscg/ruscg.tex
add a cite
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58652 72bbbca6-d526-0410-a7d9-f06f51895060
add layout
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58656 72bbbca6-d526-0410-a7d9-f06f51895060
error categorisation
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58669 72bbbca6-d526-0410-a7d9-f06f51895060
rule types
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58673 72bbbca6-d526-0410-a7d9-f06f51895060
add comment
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58676 72bbbca6-d526-04

@@
 \newcommand{\rr}[1]{\marginpar{\scriptsize R: #1}} % Rob's comments
-\newcommand{\cw}[1]{\marginpar{\scriptsize W: #1}} % Cute wife's comments
-\newcommand{\dm}[1]{\marginpar{\scriptsize D: #1}} % Detmar's comments
-\newcommand{\lj}[1]{\marginpar{\scriptsize L: #1}} % Laura's comments
 
@@
 
-\author{Author1 \\
-  Affiliation / Address line 1 \\
-  Affiliation / Address line 2 \\
-  Affiliation / Address line 3 \\
-  {\tt email@domain.com} \\\And
-  Author2 \\
-  Affiliation / Address line 1 \\
-  Affiliation / Address line 2 \\
-  Affiliation / Address line 3 \\
-  {\tt email@domain.com} \\}
+\author{Francis M. Tyers \\
+  HSL-fakultehta, \\
+  UiT Norgga árktalaš universitehta, \\
+  N-9018 Romsa \\
+  {\tt francis.tyers@uit.no} \\\And
+  Robert Reynolds \\
+  HSL-fakultehta, \\
+  UiT Norgga árktalaš universitehta, \\
+  N-9018 Romsa \\
+  {\tt robert.reynolds@uit.no} \\}
 
+% \author{Author1 \\
+%   Affiliation / Address line 1 \\
+%   Affiliation / Address line 2 \\
+%   Affiliation / Address line 3 \\
+%   {\tt email@domain.com} \\\And
+%   Author2 \\
+%   Affiliation / Address line 1 \\
+%   Affiliation / Address line 2 \\
+%   Affiliation / Address line 3 \\
+%   {\tt email@domain.com} \\}
+% 
 \date{2015}
@@
 \begin{abstract}
-  Hargle bargle
+ This paper presents preliminary work on a constraint
+ grammar based disambiguator for Russian. Russian is
+ a Slavic language with a high degree of both in-category
+ and out-category homonymy in the inflectional system.
+ The pipeline consists of a finite-state morphological
+ analyser and constraint grammar. The constraint 
+ grammar is tuned to be high recall at the expense of 
+ low precision.
 \end{abstract}
@@
 \section{Introduction}
+
+\cite{Karlsson-90}
+
+\section{Review of literature}
+% ROB + FRAN
+% different CGs 
+% faroese
+% estonian
+%
+
+\cite{trosterud2009}
+
+\section{Ambiguity in Russian}
+
+\section{Pipeline}
+
+\subsection{Morphological analyser}
+% ROB
+
+\subsection{Disambiguation rules}
+
+\begin{table}
+  \begin{tabular}{lrrr}
+    \hline
+         & {\sc select} & {\sc remove} & {\sc map} \\
+    \hline
+  \end{tabular}
+\end{table}
+
+\section{Development process}
+
+\subsection{Combining with a statistical tagger}
+
+\section{Evaluation}
+
+\begin{table}
+  \begin{tabular}{|l|r|r|}
+    \hline
+    Missing lexeme & &  \\          % безбалластный, FALSENEG: Кий<np><ant><m><sg><ins> ['кий<n><m><nn><sg><ins>']
+    Partial lexeme & &  \\          % франко-норманнское
+    Missing analysis & &  \\        % 
+    Equally valid analysis & & \\   % большой<adj><comp><pred> ['больше<adv>'] (?)
+    Correct reading removed & &  \\ % 
+    \hline
+  \end{tabular}
+
+\end{table}
+
+% unknown words from test corpus, categorisation
+
+% rule error analysis
+
+\section{Future work}
+
+% func analysis
+% dep analysis
 

### 2015-02-10 2015-eamt-melsa/2015-eamt-melsa.tex
some more compressing
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58809 72bbbca6-d526-0410-a7d9-f06f51895060
trimming
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58810 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 for training, and a target language corpus for scoring, we used four parallel corpora: 
+\addtolength{\itemsep}{-0.15in}
 \begin{itemize}
@@
   \emph{Ofis ar Brezhoneg} `The Office of the Breton
-  language'.\footnote{In 2010 the administrative status was changed
-    and it was renamed \emph{Ofis Publik ar Brezhoneg} `Public office
-    of the Breton language'. The corpus was published before this date
-    so we use the original name.}  It contains some parallel data
+  language'.%\footnote{In 2010 the administrative status was changed
+%    and it was renamed \emph{Ofis Publik ar Brezhoneg} `Public office
+%    of the Breton language'. The corpus was published before this date
+%    so we use the original name.}  
+  It contains some parallel data
   previously described in \cite{tyers09b}, and also some new data. The
-  corpus has been made available online through {\sc opus}.\footnote{\url{http://opus.lingfil.uu.se/OfisPublik.php}.}
+  corpus has been made available online through {\sc opus}.\footnote{\url{http://opus.lingfil.uu.se}}
 \item \textbf{South-East European Times} (\texttt{SETimes}): Described
@@
 \begin{itemize}
+\addtolength{\itemsep}{-0.05in}
 \item \textbf{Linguist-chosen defaults}. 
@@
 \bibliographystyle{apalike}
+\small{
 \bibliography{2015-eamt-melsa}
+}
 

### 2015-02-10 2015-eamt-melsa/2015-eamt-melsa.tex
more pruning
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58812 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \begin{description}
+\addtolength{\itemsep}{-0.10in}
+
 \item[Breton--French:] Development of the Breton--French pair has been
@@
 for training, and a target language corpus for scoring, we used four parallel corpora: 
-\addtolength{\itemsep}{-0.15in}
 \begin{itemize}
+\addtolength{\itemsep}{-0.10in}
 \item \textbf{Ofis ar Brezhoneg} (\texttt{OAB}): This parallel corpus
@@
 
-\begin{table*}
-\begin{center}
- \begin{tabular}{|l|r|r|r|r|r|}
-    \hline 
-    {\bf Pair}      & {\bf Lines} & {\bf SL words} & {\bf TL words} & {\bf Amb. words} & {\bf \% ambig} \\ 
-    \hline 
-     \texttt{br-fr} & 1,000 & 13,854  & 13,878  & 1,163 & 8.39   \\
-    \hline 
-     \texttt{mk-en} & 1,000 & 13,441 & 14,228  & 3,872 & 28.80 \\
-    \hline 
-     \texttt{eu-es} & 1,000 & 7,967 & 11,476  & 1,360  & 17.07 \\
-    \hline 
-     \texttt{en-es} & 1,000 & 19,882 & 20,944  & 1,469  & 7.38 \\
-    \hline 
- \end{tabular} 
-\end{center}
- \caption{Statistics about the test corpora. The columns \textbf{amb.\ words} and \textbf{\%~ambig} gives 
-    the number of word with more than one translation and the percentage of SL words which have 
-    more than one translation respectively.}
-  \label{table:test-corp}
-\end{table*}
+%  \begin{table*}
+%  \begin{center}
+%   \begin{tabular}{|l|r|r|r|r|r|}
+%      \hline 
+%      {\bf Pair}      & {\bf Lines} & {\bf SL words} & {\bf TL words} & {\bf Amb. words} & {\bf \% ambig} \\ 
+%      \hline 
+%       \texttt{br-fr} & 1,000 & 13,854  & 13,878  & 1,163 & 8.39   \\
+%      \hline 
+%       \texttt{mk-en} & 1,000 & 13,441 & 14,228  & 3,872 & 28.80 \\
+%      \hline 
+%       \texttt{eu-es} & 1,000 & 7,967 & 11,476  & 1,360  & 17.07 \\
+%      \hline 
+%       \texttt{en-es} & 1,000 & 19,882 & 20,944  & 1,469  & 7.38 \\
+%      \hline 
+%   \end{tabular} 
+%  \end{center}
+%   \caption{Statistics about the test corpora. The columns \textbf{amb.\ words} and \textbf{\%~ambig} gives 
+%      the number of word with more than one translation and the percentage of SL words which have 
+%      more than one translation respectively.}
+%    \label{table:test-corp}
+%  \end{table*}
 
@@
 sentences were left after processing for testing, training and
-development. Table~\ref{table:test-corp} gives information about the
-test corpora.
+development. %Table~\ref{table:test-corp} gives information about the
+%test corpora.
 
@@
   then score these translations \emph{on-line} on a model of the
-  target language. The highest scoring sentence is then picked as
+  target language. The highest scoring sentence is then 
   output. This is the method used by %@@@ \cite{carbonell06} and

### 2015-02-10 2015-eamt-melsa/2015-eamt-melsa.tex
trimming
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58813 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 %    so we use the original name.}  
-  It contains some parallel data
-  previously described in \cite{tyers09b}, and also some new data. The
-  corpus has been made available online through {\sc opus}.\footnote{\url{http://opus.lingfil.uu.se}}
+%  It contains some parallel data
+%  previously described in \cite{tyers09b}, and also some new data. 
+The corpus has been made available online through {\sc opus}.\footnote{\url{http://opus.lingfil.uu.se}}
 \item \textbf{South-East European Times} (\texttt{SETimes}): Described
@@
 tags a selection of ambiguous words in English sentences with their
-translation in Swedish. The size of the annotated corpus is not
-mentioned. 
+translation in Swedish. %The size of the annotated corpus is not
+%mentioned. 
 %\cite{specia2005a} use a parallel corpus, but eschew using
@@
 Improvements with
- respect to the target-language-model performance are very likely due
+ respect to the target-language-model performance are likely due
  to the effective use the maximum-entropy model makes of information
@@
  through the weighting of features representing those source-language
- contexts across the whole training corpus.
+ contexts across the whole corpus.
 

### 2015-02-10 2015-eamt-assim/eng-kaz/tables-updated.tex
move dir
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58814 72bbbca6-d526-0410-a7d9-f06f51895060

--- 2014-08-22 assim-eamt-2015/tat-rus/tables.tex
@@
 \usepackage{amssymb}
+
 %\date{}							% Activate to display a given date or no date
-\usepackage{float}
-\restylefloat{table}
 
 \begin{document}
-\section{General results}
+%\section{}
+%\subsection{}
 
-\begin{table}[H]
-\centering
-\begin{tabular}{|r |*{4}{c}|}
-\hline
- &\multicolumn{4}{c|}{Evaluation mode}\\
-\hline
-Gap percentage & No hint & Source hint & MT hint & MT and source hint\\
-\hline
-10\%&40.34 +/- 39.19&49.03 +/- 43.06&58.15 +/- 42.62&48.13 +/- 42.77\\
-20\%&37.72 +/- 32.24&43.46 +/- 32.07&56.88 +/- 33.36&59.99 +/- 29.47\\
-30\%&33.73 +/- 28.74&37.03 +/- 27.98&54.25 +/- 26.09&52.95 +/- 25.64\\
-\hline
-\end{tabular}
-\caption {Gaps successfully filled (\%), exactly as in reference, mean and standard deviation} \label{tab:title} 
-\end{table}
-
-\begin{table}[H]
-\centering
-\begin{tabular}{|r |*{4}{c}|}
-\hline
- &\multicolumn{4}{c|}{Evaluation mode}\\
-\hline
-Gap percentage & No hint & Source hint & MT hint & MT and source hint\\
-\hline
-10\%&42.68&51.27&63.03&48.50\\
-20\%&38.71&46.24&58.18&62.00\\
-30\%&36.67&36.70&51.04&55.27\\
-\hline
-\end{tabular}
-\caption {Gaps successfully filled (\%), exactly as in reference, average} \label{tab:title} 
-\end{table}
-
-\begin{table}[H]
-\centering
-\begin{tabular}{|r |*{4}{c}|}
-\hline
- &\multicolumn{4}{c|}{Evaluation mode}\\
-\hline
-Gap percentage & No hint & Source hint & MT hint & MT and source hint\\
-\hline
-10\%&45.96 +/- 40.86&53.52 +/- 43.57&63.76 +/- 41.15&57.02 +/- 41.57\\
-20\%&38.56 +/- 32.11&45.60 +/- 31.23&59.77 +/- 33.10&64.93 +/- 30.95\\
-30\%&35.45 +/- 29.76&40.00 +/- 28.16&55.83 +/- 26.13&58.77 +/- 28.07\\
-\hline
-\end{tabular}
-\caption {Gaps successfully filled (\%), using a synonym list, mean and standard deviation} \label{tab:title} 
-\end{table}
-
-\begin{table}[H]
-\centering
-\begin{tabular}{|r |*{4}{c}|}
-\hline
- &\multicolumn{4}{c|}{Evaluation mode}\\
-\hline
-Gap percentage & No hint & Source hint & MT hint & MT and source hint\\
-\hline
-10\%&46.34&54.43&66.06&53.89\\
-20\%&39.43&48.39&60.73&65.33\\
-30\%&38.39&39.66&52.59&59.90\\
-\hline
-\end{tabular}
-\caption {Gaps successfully filled (\%), using a synonym list, average} \label{tab:title} 
-\end{table}
-
-\begin{table}[H]
-\centering
-\begin{tabular}{|r |*{4}{c}|}
-\hline
- &\multicolumn{4}{c|}{Evaluation mode}\\
-\hline
-Gap percentage & No hint & Source hint & MT hint & MT and source hint\\
-\hline
-10\%&1:44 +/- 7:18&0:50 +/- 1:05&1:14 +/- 1:52&1:26 +/- 1:10\\
-20\%&0:57 +/- 1:22&1:19 +/- 2:34&1:39 +/- 3:02&3:26 +/- 8:44\\
-30\%&1:20 +/- 1:50&1:14 +/- 1:12&2:05 +/- 6:36&2:36 +/- 6:26\\
-\hline
-\end{tabular}
-\caption {Time taken to fill the gaps in one sentence} \label{tab:title} 
-\end{table}
-
-\begin{table}[H]
-\centering
-\begin{tabular}{|r |*{1}{c}|}
-\hline
-Gap percentage & Number of holes\\
-\hline
-10\%&272\\
-20\%&396\\
-30\%&724\\
-\hline
-\end{tabular}
-\caption {Total number of holes to be filled} \label{tab:title}
-\end{table}
-
-\begin{table}[H]
+\begin{table}
 \centering
@@
 \hline
-10\%&0.517 &0.711 &0.459 &0.598\\
-20\%&0.261 &0.473 &0.667 &0.740\\
-30\%&0.412 &0.411 &0.581 &0.534\\
-\hline
-\end{tabular}
-\caption {Krippendorff Alpha measure} \label{tab:title}
-\end{table}
-
-\section{Results by topic}
-\subsection{Casual conversations}
-
-\begin{table}[H]
-\centering
-\begin{tabular}{|r |*{4}{c}|}
-\hline
- &\multicolumn{4}{c|}{Evaluation mode}\\
-\hline
-Gap percentage & No hint & Source hint & MT hint & MT and source hint\\
-\hline
-10\%&46.97 +/- 44.28&53.03 +/- 47.58&48.48 +/- 45.20&51.52 +/- 46.85\\
-20\%&35.56 +/- 31.18&37.22 +/- 29.17&59.44 +/- 35.92&63.89 +/- 29.21\\
-30\%&33.54 +/- 31.47&34.58 +/- 27.22&56.11 +/- 23.82&59.17 +/- 28.05\\
+10\%&40.48 +/- 45.30&30.16 +/- 35.86&38.10 +/- 46.05&35.33 +/- 38.68\\
+20\%&54.63 +/- 27.53&39.19 +/- 20.37&24.79 +/- 19.16&48.78 +/- 32.65\\
+30\%&23.87 +/- 24.95&35.13 +/- 30.18&30.41 +/- 34.73&43.07 +/- 32.71\\
 \hline
@@
 
-\begin{table}[H]
-\centering
-\begin{tabular}{|r |*{4}{c}|}
-\hline
- &\multicolumn{4}{c|}{Evaluation mode}\\
-\hline
-Gap percentage & No hint & Source hint & MT hint & MT and source hint\\
-\hline
-10\%&50.00&57.50&48.78&50.00\\
-20\%&35.44&39.24&59.21&65.79\\
-30\%&37.63&37.50&53.85&59.09\\
-\hline
-\end{tabular}
-\caption {Gaps successfully filled (\%), exactly as in reference, average} \label{tab:title} 
-\end{table}
-
-\begin{table}[H]
-\centering
-\begin{tabular}{|r |*{4}{c}|}
-\hline
- &\multicolumn{4}{c|}{Evaluation mode}\\
-\hline
-Gap percentage & No hint & Source hint & MT hint & MT and source hint\\
-\hline
-10\%&53.03 +/- 44.28&62.12 +/- 46.11&63.64 +/- 43.12&63.64 +/- 44.84\\
-20\%&37.78 +/- 30.94&41.11 +/- 27.88&63.33 +/- 35.85&73.33 +/- 31.51\\
-30\%&37.71 +/- 33.32&39.44 +/- 27.10&59.58 +/- 23.89&70.28 +/- 31.12\\
-\hline
-\end{tabular}
-\caption {Gaps successfully filled (\%), using a synonym list, mean and standard deviation} \label{tab:title} 
-\end{table}
-
-\begin{table}[H]
-\centering
-\begin{tabular}{|r |*{4}{c}|}
-\hline
- &\multicolumn{4}{c|}{Evaluation mode}\\
-\hline
-Gap percentage & No hint & Source hint & MT hint & MT and source hint\\
-\hline
-10\%&54.76&65.00&60.98&59.52\\
-20\%&37.97&44.30&63.16&73.68\\
-30\%&41.94&41.67&58.24&68.18\\
-\hline
-\end{tabular}
-\caption {Gaps successfully filled (\%), using a synonym list, average} \label{tab:title} 
-\end{table}
-
-\begin{table}[H]
-\centering
-\begin{tabular}{|r |*{4}{c}|}
-\hline
- &\multicolumn{4}{c|}{Evaluation mode}\\
-\hline
-Gap percentage & No hint & Source hint & MT hint & MT and source hint\\
-\hline
-10\%&3:35 +/- 11:43&0:45 +/- 1:00&1:03 +/- 1:14&1:28 +/- 1:16\\
-20\%&0:33 +/- 0:31&0:47 +/- 0:41&0:38 +/- 0:36&0:52 +/- 1:28\\
-30\%&1:09 +/- 2:21&0:49 +/- 0:42&0:40 +/- 0:29&0:53 +/- 0:54\\
-\hline
-\end{tabular}
-\caption {Time taken to fill the gaps in one sentence} \label{tab:title} 
-\end{table}
-
-\begin{table}[H]
-\centering
-\begin{tabular}{|r |*{1}{c}|}
-\hline
-Gap percentage & Number of holes\\
-\hline
-10\%&60\\
-20\%&124\\
-30\%&184\\
-\hline
-\end{tabular}
-\caption {Total number of holes to be filled} \label{tab:title}
-\end{table}
-
-\begin{table}[H]
+\begin{table}
 \centering
@@
 \hline
-10\%&0.411 &0.601 &0.426 &0.504\\
-20\%&0.235 &0.371 &0.570 &0.712\\
-30\%&0.121 &0.459 &0.846 &0.310\\
-\hline
-\end{tabular}
-\caption {Krippendorff Alpha measure} \label{tab:title}
-\end{table}
-
-\subsection{Legal texts}
-
-\begin{table}[H]
-\centering
-\begin{tabular}{|r |*{4}{c}|}
-\hline
- &\multicolumn{4}{c|}{Evaluation mode}\\
-\hline
-Gap percentage & No hint & Source hint & MT hint & MT and source hint\\
-\hline
-10\%&29.10 +/- 29.64&38.77 +/- 37.82&68.21 +/- 34.99&53.02 +/- 40.31\\
-20\%&48.31 +/- 33.52&50.28 +/- 33.55&65.76 +/- 24.16&60.57 +/- 24.98\\
-30\%&33.68 +/- 28.16&38.49 +/- 27.81&48.30 +/- 28.87&47.39 +/- 26.29\\
-\hline
-\end{tabular}
-\caption {Gaps successfully filled (\%), exactly as in reference, mean and standard deviation} \label{tab:title} 
-\end{table}
-
-\begin{table}[H]
-\centering
-\begin{tabular}{|r |*{4}{c}|}
-\hline
- &\multicolumn{4}{c|}{Evaluation mode}\\
-\hline
-Gap percentage & No hint & Source hint & MT hint & MT and source hint\\
-\hline
-10\%&35.21&43.48&67.11&52.00\\
-20\%&43.31&52.46&63.57&61.74\\
-30\%&36.17&35.16&41.32&53.76\\
+10\%&31.43&27.50&34.15&30.43\\
+20\%&51.72&36.23&22.54&43.84\\
+30\%&21.85&32.58&25.56&35.63\\
 \hline
@@
 
-\begin{table}[H]
-\centering
-\begin{tabular}{|r |*{4}{c}|}
-\hline
- &\multicolumn{4}{c|}{Evaluation mode}\\
-\hline
-Gap percentage & No hint & Source hint & MT hint & MT and source hint\\
-\hline
-10\%&32.95 +/- 34.88&38.77 +/- 37.82&68.21 +/- 34.99&53.02 +/- 40.31\\
-20\%&48.31 +/- 33.52&50.28 +/- 33.55&65.76 +/- 24.16&60.57 +/- 24.98\\
-30\%&34.15 +/- 28.73&40.05 +/- 27.37&48.30 +/- 28.87&50.06 +/- 26.25\\
-\hline
-\end{tabular}
-\caption {Gaps successfully filled (\%), using a synonym list, mean and standard deviation} \label{tab:title} 
-\end{table}
-
-\begin{table}[H]
-\centering
-\begin{tabular}{|r |*{4}{c}|}
-\hline
- &\multicolumn{4}{c|}{Evaluation mode}\\
-\hline
-Gap percentage & No hint & Source hint & MT hint & MT and source hint\\
-\hline
-10\%&38.03&43.48&67.11&52.00\\
-20\%&43.31&52.46&63.57&61.74\\
-30\%&37.23&37.36&41.32&56.07\\
-\hline
-\end{tabular}
-\caption {Gaps successfully filled (\%), using a synonym list, average} \label{tab:title} 
-\end{table}
-
-\begin{table}[H]
-\centering
-\begin{tabular}{|r |*{4}{c}|}
-\hline
- &\multicolumn{4}{c|}{Evaluation mode}\\
-\hline
-Gap percentage & No hint & Source hint & MT hint & MT and source hint\\
-\hline
-10\%&0:49 +/- 1:13&1:04 +/- 1:21&1:17 +/- 1:01&1:34 +/- 1:03\\
-20\%&1:32 +/- 2:02&2:28 +/- 4:18&3:26 +/- 4:41&8:22 +/- 13:46\\
-30\%&1:13 +/- 1:36&1:19 +/- 1:21&3:32 +/- 10:24&2:57 +/- 4:00\\
-\hline
-\end{tabular}
-\caption {Time taken to fill the gaps in one sentence} \label{tab:title} 
-\end{table}
-
-\begin{table}[H]
-\centering
-\begin{tabular}{|r |*{1}{c}|}
-\hline
-Gap percentage & Number of holes\\
-\hline
-10\%&132\\
-20\%&244\\
-30\%&284\\
-\hline
-\end{tabular}
-\caption {Total number of holes to be filled} \label{tab:title}
-\end{table}
-
-\begin{table}[H]
+\begin{table}
 \centering
@@
 \hline
-10\%&0.543 &0.481 &0.402 &0.642\\
-20\%&0.317 &0.453 &0.631 &0.652\\
-30\%&0.488 &0.301 &0.426 &0.532\\
-\hline
-\end{tabular}
-\caption {Krippendorff Alpha measure} \label{tab:title}
-\end{table}
-
-\subsection{News}
-
-\begin{table}[H]
-\centering
-\begin{tabular}{|r |*{4}{c}|}
-\hline
- &\multicolumn{4}{c|}{Evaluation mode}\\
-\hline
-Gap percentage & No hint & Source hint & MT hint & MT and source hint\\
-\hline
-10\%&42.78 +/- 38.41&54.02 +/- 40.51&59.77 +/- 43.66&40.00 +/- 38.87\\
-20\%&28.97 +/- 28.79&44.44 +/- 32.51&43.70 +/- 34.86&54.46 +/- 33.34\\
-30\%&33.98 +/- 26.52&37.65 +/- 28.77&59.83 +/- 22.90&53.70 +/- 20.22\\
-\hline
-\end{tabular}
-\caption {Gaps successfully filled (\%), exactly as in reference, mean and standard deviation} \label{tab:title} 
-\end{table}
-
-\begin{table}[H]
-\centering
-\begin{tabular}{|r |*{4}{c}|}
-\hline
- &\multicolumn{4}{c|}{Evaluation mode}\\
-\hline
-Gap percentage & No hint & Source hint & MT hint & MT and source hint\\
-\hline
-10\%&47.06&57.14&68.75&42.00\\
-20\%&34.25&43.59&47.14&58.67\\
-30\%&36.72&38.28&61.72&54.69\\
-\hline
-\end{tabular}
-\caption {Gaps successfully filled (\%), exactly as in reference, average} \label{tab:title} 
-\end{table}
-
-\begin{table}[H]
-\centering
-\begin{tabular}{|r |*{4}{c}|}
-\hline
- &\multicolumn{4}{c|}{Evaluation mode}\\
-\hline
-Gap percentage & No hint & Source hint & MT hint & MT and source hint\\
-\hline
-10\%&49.44 +/- 39.08&57.47 +/- 42.12&59.77 +/- 43.66&53.33 +/- 37.86\\
-20\%&28.97 +/- 28.79&46.53 +/- 31.99&48.56 +/- 35.40&59.33 +/- 33.89\\
-30\%&34.81 +/- 27.01&40.51 +/- 30.10&61.49 +/- 22.11&58.14 +/- 22.42\\
+10\%&40.48 +/- 45.30&42.06 +/- 38.70&42.86 +/- 44.42&39.33 +/- 37.38\\
+20\%&63.89 +/- 28.87&49.78 +/- 19.89&36.01 +/- 24.61&48.78 +/- 32.65\\
+30\%&32.08 +/- 27.38&49.42 +/- 25.61&45.01 +/- 37.62&47.04 +/- 33.64\\
 \hline
@@
 
-\begin{table}[H]
-\centering
-\begin{tabular}{|r |*{4}{c}|}
-\hline
- &\multicolumn{4}{c|}{Evaluation mode}\\
-\hline
-Gap percentage & No hint & Source hint & MT hint & MT and source hint\\
-\hline
-10\%&50.98&61.22&68.75&52.00\\
-20\%&34.25&46.15&52.86&64.00\\
-30\%&37.50&41.41&63.28&59.38\\
-\hline
-\end{tabular}
-\caption {Gaps successfully filled (\%), using a synonym list, average} \label{tab:title} 
-\end{table}
-
-\begin{table}[H]
-\centering
-\begin{tabular}{|r |*{4}{c}|}
-\hline
- &\multicolumn{4}{c|}{Evaluation mode}\\
-\hline
-Gap percentage & No hint & Source hint & MT hint & MT and source hint\\
-\hline
-10\%&0:31 +/- 0:20&0:42 +/- 0:49&1:25 +/- 2:49&1:17 +/- 1:09\\
-20\%&0:50 +/- 0:53&0:49 +/- 0:42&0:56 +/- 0:39&1:05 +/- 0:33\\
-30\%&1:38 +/- 1:28&1:31 +/- 1:17&1:40 +/- 1:12&3:53 +/- 10:27\\
-\hline
-\end{tabular}
-\caption {Time taken to fill the gaps in one sentence} \label{tab:title} 
-\end{table}
-
-\begin{table}[H]
-\centering
-\begin{tabular}{|r |*{1}{c}|}
-\hline
-Gap percentage & Number of holes\\
-\hline
-10\%&80\\
-20\%&148\\
-30\%&256\\
-\hline
-\end{tabular}
-\caption {Total number of holes to be filled} \label{tab:title}
-\end{table}
-
-\begin{table}[H]
+\begin{table}
 \centering
@@
 \hline
-10\%&0.537 &0.917 &0.574 &0.598\\
-20\%&0.143 &0.644 &0.850 &0.928\\
-30\%&0.513 &0.456 &0.532 &0.644\\
+10\%&31.43&40.00&39.02&34.78\\
+20\%&60.34&47.83&33.80&43.84\\
+30\%&31.09&44.94&36.56&41.38\\
+\hline
+\end{tabular}
+\caption {Gaps successfully filled (\%), using a synonym list, average} \label{tab:title} 
+\end{table}
+
+\begin{table}
+\centering
+\begin{tabular}{|r |*{4}{c}|}
+\hline
+ &\multicolumn{4}{c|}{Evaluation mode}\\
+\hline
+Gap percentage & No hint & Source hint & MT hint & MT ans Source hint\\
+\hline
+10\%&0:47 +/- 0:35&1:47 +/- 1:36&1:34 +/- 1:03&3:18 +/- 2:35\\
+20\%&1:13 +/- 1:02&2:24 +/- 2:21&2:19 +/- 2:07&1:31 +/- 1:21\\
+30\%&4:20 +/- 11:28&2:01 +/- 1:49&3:21 +/- 4:18&2:00 +/- 1:30\\
+\hline
+\end{tabular}
+\caption {Time taken to fill the gaps in one sentence (m:s)} \label{tab:title} 
+\end{table}
+
+\begin{table}
+\centering
+\begin{tabular}{|r |*{1}{c}|}
+\hline
+Gap percentage & Number of holes\\
+\hline
+10\%&22\\
+20\%&43\\
+30\%&52\\
+\hline
+\end{tabular}
+\caption {Total number of holes to be filled by one evaluator.} \label{tab:title}
+\end{table}
+
+\begin{table}
+\centering
+\begin{tabular}{|r |*{4}{c}|}
+\hline
+ &\multicolumn{4}{c|}{Evaluation mode}\\
+\hline
+Gap percentage & No hint & Source hint & MT hint & MT ans Source hint\\
+\hline
+10\%&0.812&0.000&0.516&0.143\\
+20\%&0.638&0.282&0.425&0.507\\
+30\%&0.521&0.348&0.795&0.413\\
 \hline

### 2015-02-10 2015-eamt-assim/eu-es/tables.tex
move dir
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58814 72bbbca6-d526-0410-a7d9-f06f51895060

--- 2015-02-10 2015-eamt-assim/eng-kaz/tables-updated.tex
@@
 
-%\date{}							% Activate to display a given date or no date
-
+\title{Basque-Spanish Apertium translation system evaluation}
 \begin{document}
-%\section{}
 %\subsection{}
@@
 \hline
-Gap percentage & No hint & Source hint & MT hint & MT ans Source hint\\
+Gap percentage & MT and source hint & MT hint & No hint & Source hint\\
 \hline
-10\%&40.48 +/- 45.30&30.16 +/- 35.86&38.10 +/- 46.05&35.33 +/- 38.68\\
-20\%&54.63 +/- 27.53&39.19 +/- 20.37&24.79 +/- 19.16&48.78 +/- 32.65\\
-30\%&23.87 +/- 24.95&35.13 +/- 30.18&30.41 +/- 34.73&43.07 +/- 32.71\\
+10\%&47.90 +/- 35.16&58.04 +/- 27.96&34.65 +/- 37.69&33.84 +/- 36.86\\
+20\%&46.30 +/- 25.14&52.73 +/- 23.73&26.52 +/- 27.64&30.16 +/- 28.46\\
+30\%&38.71 +/- 22.71&28.89 +/- 19.54&16.26 +/- 17.35&20.92 +/- 18.75\\
 \hline
 \end{tabular}
-\caption {Gaps successfully filled (\%), exactly as in reference, mean and standard deviation} \label{tab:title} 
+\caption {Gaps successfully filled (\%), exactly as in reference.} \label{tab:title} 
 \end{table}
@@
 \hline
-Gap percentage & No hint & Source hint & MT hint & MT ans Source hint\\
+Gap percentage & MT and source hint & MT hint & No hint & Source hint\\
 \hline
-10\%&31.43&27.50&34.15&30.43\\
-20\%&51.72&36.23&22.54&43.84\\
-30\%&21.85&32.58&25.56&35.63\\
+10\%&62.19 +/- 32.07&58.04 +/- 27.96&48.79 +/- 39.84&39.90 +/- 38.96\\
+20\%&64.81 +/- 30.03&70.18 +/- 26.98&30.91 +/- 30.22&31.37 +/- 28.42\\
+30\%&48.13 +/- 26.08&39.55 +/- 23.90&18.07 +/- 18.15&25.94 +/- 19.67\\
 \hline
 \end{tabular}
-\caption {Gaps successfully filled (\%), exactly as in reference, average} \label{tab:title} 
+\caption {Gaps successfully filled (\%), using a synonym list.} \label{tab:title} 
 \end{table}
@@
 \hline
-Gap percentage & No hint & Source hint & MT hint & MT ans Source hint\\
+Gap percentage & MT and source hint & MT hint & No hint & Source hint\\
 \hline
-10\%&40.48 +/- 45.30&42.06 +/- 38.70&42.86 +/- 44.42&39.33 +/- 37.38\\
-20\%&63.89 +/- 28.87&49.78 +/- 19.89&36.01 +/- 24.61&48.78 +/- 32.65\\
-30\%&32.08 +/- 27.38&49.42 +/- 25.61&45.01 +/- 37.62&47.04 +/- 33.64\\
+10\%&49.38&55.70&35.06&32.47\\
+20\%&48.23&55.64&30.14&32.31\\
+30\%&41.64&29.97&17.67&20.14\\
 \hline
 \end{tabular}
-\caption {Gaps successfully filled (\%), using a synonym list, mean and standard deviation} \label{tab:title} 
+\caption {Gaps successfully filled (\%), exactly as in reference.} \label{tab:title}
 \end{table}
@@
 \hline
-Gap percentage & No hint & Source hint & MT hint & MT ans Source hint\\
+Gap percentage & MT and source hint & MT hint & No hint & Source hint\\
 \hline
-10\%&31.43&40.00&39.02&34.78\\
-20\%&60.34&47.83&33.80&43.84\\
-30\%&31.09&44.94&36.56&41.38\\
+10\%&58.02&55.70&45.45&35.06\\
+20\%&65.96&68.42&33.56&33.85\\
+30\%&49.84&37.28&18.93&24.23\\
 \hline
 \end{tabular}
-\caption {Gaps successfully filled (\%), using a synonym list, average} \label{tab:title} 
+\caption {Gaps successfully filled (\%), using a synonym list.} \label{tab:title} 
+\end{table}
+
+\begin{table}
+\centering
+\begin{tabular}{|r |*{1}{c}|}
+\hline
+Gap percentage & Number of holes\\
+\hline
+10\%&94\\
+20\%&181\\
+30\%&272\\
+\hline
+\end{tabular}
+\caption {Total number of holes to be filled.} \label{tab:title}
 \end{table}
@@
 \hline
-Gap percentage & No hint & Source hint & MT hint & MT ans Source hint\\
+Gap percentage & MT and source hint & MT hint & No hint & Source hint\\
 \hline
-10\%&0:47 +/- 0:35&1:47 +/- 1:36&1:34 +/- 1:03&3:18 +/- 2:35\\
-20\%&1:13 +/- 1:02&2:24 +/- 2:21&2:19 +/- 2:07&1:31 +/- 1:21\\
-30\%&4:20 +/- 11:28&2:01 +/- 1:49&3:21 +/- 4:18&2:00 +/- 1:30\\
+10\%&1:27 +/- 1:27&1:47 +/- 1:57&0:46 +/- 0:44&1:06 +/- 0:38\\
+20\%&1:31 +/- 1:19&1:24 +/- 1:17&1:22 +/- 1:17&1:27 +/- 1:02\\
+30\%&3:54 +/- 5:02&2:50 +/- 1:51&2:47 +/- 2:07&2:36 +/- 1:22\\
 \hline
@@
 \caption {Time taken to fill the gaps in one sentence (m:s)} \label{tab:title} 
-\end{table}
-
-\begin{table}
-\centering
-\begin{tabular}{|r |*{1}{c}|}
-\hline
-Gap percentage & Number of holes\\
-\hline
-10\%&22\\
-20\%&43\\
-30\%&52\\
-\hline
-\end{tabular}
-\caption {Total number of holes to be filled by one evaluator.} \label{tab:title}
 \end{table}
@@
 \hline
-Gap percentage & No hint & Source hint & MT hint & MT ans Source hint\\
+Gap percentage & MT and source hint & MT hint & No hint & Source hint\\
 \hline
-10\%&0.812&0.000&0.516&0.143\\
-20\%&0.638&0.282&0.425&0.507\\
-30\%&0.521&0.348&0.795&0.413\\
+10\%&0.496&0.517&0.124&0.400\\
+20\%&0.714&0.700&0.275&0.358\\
+30\%&0.559&0.430&0.300&0.406\\
 \hline

### 2015-02-10 2015-eamt-assim/2015-eamt-assim.tex
add tex files
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58815 72bbbca6-d526-0410-a7d9-f06f51895060

%
% File eamt15.tex, an adaptation of eamt14.tex (which was a copy of
% eamt12.tex)
%
% Contact: eamt2015@dlsi.ua.es

%%% To ease future customizations, various replaceables have been paramaterized
%%% as listed in the newcommands section

\documentclass[11pt]{article}
\usepackage{eamt15}
\usepackage{times}
\usepackage{latexsym}
\setlength\titlebox{6.5cm}    % Expanding the titlebox
%%% YOUR PACKAGES BELOW THIS LINE %%%

\usepackage{multirow}

\newcommand{\confname}{EAMT 2015}
\newcommand{\website}{http://www.eamt2015.org/}
\newcommand{\contactname}{the conference chairs (Felipe
  S\'anchez-Mart\'inez, Gema Ram\'irez-S\'anchez and Fred Hollowood)}
\newcommand{\contactemail}{eamt2015@dlsi.ua.es}
\newcommand{\conffilename}{eamt15}
\newcommand{\downloadsite}{http://www.eamt2015.org/}
\newcommand{\paperlength}{$8$ (eight)}
\newcommand{\shortpaperlength}{$4$ (four)}

\title{A gap-filling approach to evaluation of machine translation for assimilation}

\author{First Author\\
  Affiliation / Address line 1\\
  Affiliation / Address line 2\\
  Affiliation / Address line 3\\
  {\tt email@domain}  \And
  Second Author\\
  Affiliation / Address line 1\\
  Affiliation / Address line 2\\
  Affiliation / Address line 3\\
  {\tt email@domain}}

\date{}

\begin{document}
\maketitle
\begin{abstract}
   ... 
\end{abstract}

\section{Introduction}

As suggested by [Church, Hovy 1993], modern machine translation (MT) systems may
be divided into two broad categories according to their purpose: post-editing systems, e.g.
Google Translate 1 , and assimilation systems, e.g. Apertium 2 . Both kinds may be
systematically evaluated, either to control for quality in the development process or to
compare the systems. The evaluation methods, however, differ depending on the type of
the system. The methods are described in detail in [Church, Hovy 1993].
In this work I perform the evaluation of an assimilation MT system called Apertium.
Apertium is an open-source community-maintained project that provides translation for
39 language pairs. The goal of this study is to assess whether machine translation
performed by this system assists users in understanding the gist of texts in an unknown
language. The evaluation is performed in three separate experiments for the following
translation directions: Basque to Spanish, English to Kazakh, and Tatar to Russian.
I build on the methodology established in [Trosterund, Unhammer 2012] and
\cite{oregan2013}, which are the works describing previous evaluations of
Apertium language pairs. [O’Regan, Forcada 2013] proposes the gap-filling task as a
means of assessing participant’s understanding of text. The same methodology is used in
this work and described in detail in the section below.

\section{Experiment design}

\subsection{Methodology}

This experiment is based on an idea that understanding of text may be assessed through
gap-filling questions. If the participant is able to restore principal information that had
been left out from the text, they are assumed to have understood it.
During the course of the experiment, the participants are offered sentences in their native
language (the target language of the pair), with some key words left out. Each sentence
has a counterpart in the source language of the pair. I use four different task settings:
\begin{itemize}
\item  Target sentence only. The participants are asked to fill the gaps not being given
any context. This task serves as a baseline score and as an indicator of phrases
that can be completed using common knowledge or language intuition (e.g.
idioms and strong collocations);
\item  Target sentence and source sentence. In addition to the question, the participants
are offered the same sentence in the source language of the pair. By setup, the
participants have no command of this language, however, it may help them to fill
in proper nouns or loan words;
\item  Target sentence and machine-translated hint. In addition to the task, the
participants see the source sentence translated via Apertium. This type of task is
used for measuring the contribution of machine translation to understanding the
gist of text;
\item  Target sentence and both the source and machine-translated hint. This task is
added mostly for completing the possible combination of task types, and to check
for any unexpected insights, should they arise.
\end{itemize}

\subsection{Material}

The data are collected separately for each experiment. A set of sentences is drawn from
parallel corpora. The source sentences are translated into target language via Apertium.
The target sentences are processed with a special script that removes significant words.
For the current experiments, the word is removed if it falls into a list of allowed parts of
speech:
\begin{itemize}
\item nouns;
\item proper nouns;
\item adjectives;
\item adverbs;
\item lexical verbs.
\end{itemize}
Parts of speech are determined automatically with Apertium morphological analyser. To
rule out homonymy, the set of all possible POS attributions of a given word is checked
against the POS list. If all of the possible word analyses match the parts of speech listed
above, it is considered a candidate for removal.
Each sentence in the set is evaluated in three “gap densities” (ratio of gaps to words in the
sentence): 10\%, 20\% and 30\%. After the script finds all possible keyword candidates in
the sentence, it removes the specified percentage of words, evenly distributing the gaps
across the sentence.
To exclude headlines and also to allow at least one gap in 10\%-gap sentences, only the
sentences over 10 words long are used in the tasks.
The sentences for each experiment are drawn from various parallel text sources:
\begin{enumerate}
\item  For Basque-Spanish, from the corpus of legal texts Memorias de traducción del
Servicio Oficial de Traductores del IVAP 3
\item  For English-Kazakh, from the official website of the President of the Republic of
Kazakhstan 4
\item  For Tatar-Russian, from the following sources on three different topics:
  \begin{enumerate}
    \item  Casual conversations, from a textbook 5 of spoken Tatar;
    \item  Legal texts, from the Constitution and laws 6 of Tatarstan;
    \item  News, from the President of Tatarstan website 7 .
  \end{enumerate}
\end{enumerate}

Each set features 36 pairs of sentences. For the first two experiments the pairs are drawn
randomly from the corpora, for Tatar-Russian, compiled by hand by the developer of the
language pair. After processing, the material is transformed into XML used for uploading
into the evaluation system (see section 2.4).

\subsection{Participants}

For each experiment we called for native speakers of L2 of the language pair (i.e.
Spanish, Kazakh, and Russian) who had no command of L1 of the pair (Basque, English
and Tatar, respectively). We had 11 participants in Basque-Spanish evaluation, 8 in
English-Kazakh, and 28 in Tatar-Russian (although not everyone has completed their task
in full, see discussion).

\subsection{Procedure}

The evaluations take place online, in a system 8 designed specifically for various MT
evaluation tasks. We adapted the code to accommodate for the gap-filling tasks. The
tasks are uploaded into the system and manually distributed between the participants by
the following rules:
\begin{enumerate}
\item  Each participant evaluates every sentence (understood as a succession of words),
a total of 36;
\item  Each participant evaluates 9 sentences in each of the four modes (section 2.1);
\item  All sentences of the set are evaluated with 10, 20 and 30\% of words removed;
\item  Each sentence-mode-percentage combination is evaluated by at least three
\end{enumerate}
The participants are given the instructions in their native language above each task. The
tasks are split into smaller sets for the participants’ convenience. The instructions are to

read all the available hints and fill each gap with one suitable word, guessing if unsure.
Participants’ answers are recorded and marked correct or incorrect automatically. In
addition, the time taken to fill the gaps in one sentence is recorded.
Since gap-filling task requires open answers, it is possible that the participants may
provide words that fit the gaps well, but do not match the original answer. To account for
these cases, we process all the answers to detect possible synonyms. An answer is
considered a candidate synonym if given by two or more evaluators and does not match
the answer key. We record each candidate synonym along with the answer key and the
context sentence. Based on this data, a native speaker of the target language decides
whether the candidate synonym is indeed synonymic to the answer key in the given
context. We then check participants’ results against the compiled synonym list and
increase scores where appropriate.

\section{Expected results}

We expect that, for tasks with MT assistance, the users will fill more gaps correctly,
which would correspond to better understanding of text. Therefore, the percentage of
correct answers should be higher for MT and two-hint kinds of tasks than for tasks with
source sentence as a hint and no hint at all. If the percentage were lower, we would
expect an error in experiment design.

\section{Results}

The experiment results are presented in the tables below. The first table of each set shows
the proportion of correct answers depending on evaluation mode and gap density. The
percentage is spread across all evaluators, that is, we divide the number of correct
answers given by all evaluators by the number of questions answered by all evaluators.
The second table of the set shows average time it took to fill the gaps in one sentence. To
reduce the noise from participants who were distracted during evaluation, when
calculating times we remove all the results over 6 minutes (the mode is approximately
two minutes).

\begin{table}
  \begin{tabular}{|l|r|r|r|r|}
    \hline
    \multirow{2}{*}{\textbf{Density}} & \multicolumn{4}{|c|}{Evaluation mode} \\\cline{2-5}
                                            & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
    10\%                                    &   58.02            & 55.70       & 35.06        & 45.45        \\
    20\%                                    &   65.96            & 68.42       & 33.85        & 33.56        \\
    30\%                                    &   49.84            & 37.28       & 24.23        & 18.93        \\
    \hline
  \end{tabular}
  \caption{Basque--Spanish: Gaps successfully filled (\%), using a synonym list}
  \label{table:res-eus-spa}
\end{table}

\begin{table}
  \begin{tabular}{|l|r|r|r|r|}
    \hline
    \multirow{2}{*}{\textbf{Density}} & \multicolumn{4}{|c|}{Evaluation mode} \\\cline{2-5}
                                            & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
    10\%                                    & 34.78              & 39.02       & 40.00        & 31.43        \\
    20\%                                    & 43.84              & 33.80       & 47.83        & 60.34        \\
    30\%                                    & 41.38              & 36.56       & 44.94        & 31.09        \\
    \hline
  \end{tabular}
  \caption{English--Kazakh: Gaps successfully filled (\%), using a synonym list}
  \label{table:res-eng-kaz}
\end{table}

\begin{table}
  \begin{tabular}{|l|r|r|r|r|}
    \hline
    \multirow{2}{*}{\textbf{Density}} & \multicolumn{4}{|c|}{Evaluation mode} \\\cline{2-5}
                                            & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
    10\%                                    & 53.89              & 66.06       & 54.43        & 46.34        \\
    20\%                                    & 65.33              & 60.73       & 48.39        & 39.43        \\
    30\%                                    & 59.90              & 52.59       & 39.66        & 38.39        \\
    \hline
  \end{tabular}
  \caption{Tatar--Russian: Gaps successfully filled (\%), using a synonym list}
  \label{table:res-tat-rus}
\end{table}

\section{Discussion}

The following trends may be derived from the data above:
\begin{enumerate}
\item  In general, tasks with MT assistance score higher than task without MT
assistance;
\item  The more words are removed from context, the more difficult it is to fill the gaps
(see below for exceptions);
\item  Task without hints take less time to complete than others.
\end{enumerate}
The first two trends align with our expectations. Higher scores on MT tasks indicate
better understanding of text. The large number of gaps in the sentence makes it more
difficult to predict the answer based on the context, and also leaves more room for
translation mistakes. We would expect that tasks without hints take the largest amount of
time rather than that the smallest, since the participant must come up with their own
answer unassisted. However, in the no-hint task the participant reads only one sentence,
as opposed to two or three in other tasks, which may be the source of difference.
In Basque-Spanish and Tatar-Russian these trends are defined reasonably well. In
English-Kazakh, however, the scores for source-hint task (a parallel sentence in English
for assistance) are systematically higher than for other types of tasks. In addition, a 20\%no-hint task stands out with 60\% accuracy, which is substantially higher than any other
scores for this experiment.
Having analyzed the tasks and the responses for English-Kazakh, I have found the
following possible reasons of the obtained results:
\begin{enumerate}
\item  The experiment was planned for 11 evaluators, although only 8 attempted the
tasks and 6 completed them in full. Because of this, the sentences were evaluated
once or twice, and some were not attempted at all. To reliably compare results
across different categories, all tasks must be completed;
\item  Due to an error during experiment design, some sentences are not repeated across
categories, which also suggests that we may not compare these percentages;
\item  The outlier score of 60\% is made up by three sentences. Two of these scored
unusually high. It appeared that one of them is a news headline, and the other
contains clichés such as “express condolences”. There may also be interference
from other tasks of the set, because some of the sentences are drawn from the
same news article. We would expect the same sentences to contribute to higher
percentage in other task modes, but they were not completed there;
\item  It may be the case that the participants have at least some command of English,
which would have helped them to derive extra information in the source-hint
\end{enumerate}
It remains questionable whether we can compare results for different gap densities. In the
Basque-Spanish and Tatar-Russian experiments the 10, 20 and 30\% sets were comprised
of the same sentences. However, in each case different words were removed. It appears
that some content words are easier to fill than the others. This may explain why in
Basque-Spanish the 20\% MT tasks are completed with better accuracy than 10\% tasks.
It may be worth noting that the majority of the participants reported feeling frustrated in
the course of their work, especially while working on the no-hint tasks. 9 out of 51
participants quit the experiment before completing it.

\section{Future work}
The experiment may be repeated for any language pair (provided a parallel corpus) and
any machine translation system. To enhance the results, the procedure may be amended
in the following ways:
\begin{enumerate}
\item  Reduce the number of questions per participant. In the above experiments each
participant filled from 110 to 187 gaps, divided into small groups. Reducing the
amount of work may increase task completion rate;
\item  Increase the number of participants per experiment. The more evaluators work on
every question, the more reliable the data;
\item  Design the tasks with different gap densities in such a way that e.g. the 20\% task
contains all the gaps of the 10\% task, plus any additional ones;
\item  Draw sentences from large corpora. This will reduce the influence of previous
tasks, which may provide the context for other tasks in the set.
\end{enumerate}

\bibliographystyle{apalike}

\bibliography{2015-eamt-assim}

\end{document}

### 2015-02-10 2015-eamt-assim/2015-eamt-assim.tex
import stuffs
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58816 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-As suggested by [Church, Hovy 1993], modern machine translation (MT) systems may
+As suggested by \newcite{church93}, modern machine translation (MT) systems may
 be divided into two broad categories according to their purpose: post-editing systems, e.g.
@@
 compare the systems. The evaluation methods, however, differ depending on the type of
-the system. The methods are described in detail in [Church, Hovy 1993].
+the system. The methods are described in detail in \newcite{church93}.
 In this work I perform the evaluation of an assimilation MT system called Apertium.
@@
 translation directions: Basque to Spanish, English to Kazakh, and Tatar to Russian.
-I build on the methodology established in [Trosterund, Unhammer 2012] and
-\cite{oregan2013}, which are the works describing previous evaluations of
-Apertium language pairs. [O’Regan, Forcada 2013] proposes the gap-filling task as a
+I build on the methodology established in \newcite{trosterud12} and
+\newcite{oregan2013}, which are the works describing previous evaluations of
+Apertium language pairs. \newcite{oregan2013} proposes the gap-filling task as a
 means of assessing participant’s understanding of text. The same methodology is used in

### 2015-02-11 2015-eamt-melsa/2015-eamt-melsa.tex
8 pages \o/
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58848 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 %@@@ \comment{This section should be ironed out a bit.}  
-In their
-pioneering work, \cite{dagan1994word} used the term
+%In their pioneering work, 
+\cite{dagan1994word} used the term
 \emph{word sense disambiguation} to refer to what is actually target
@@
 tuples from the target language and count the frequency in a previously-trained target-language 
-model of tuples.  They use maximum-likelihood estimation to calculate the probability of a given 
-target language tuple being the translation of a given source language tuple, with
-an automatically determined threshold for confidence.
+model of tuples.  They use maximum-likelihood estimation to calculate the probability that a given 
+target language tuple is the translation of a given source language tuple, with
+an automatically determined confidence threshold.
 
@@
 model provides probabilities of word sequences in the target language.
-As regards rule-based machine translation, \cite{sanchez07a} proposed 
-to use information from the target language to train a statistical 
-lexical-selection module but did not provide any experiments. 
+%As regards rule-based machine translation, \cite{sanchez07a} proposed 
+%to use information from the target language to train a statistical 
+%lexical-selection module but did not provide any experiments. 
 %@@@ \todo[inline]{FMT: integrate \citep{sanchez07a} somewhere here}
@@
 alternative method is described by \cite{zinovjeva2000}, who manually
-tags a selection of ambiguous words in English sentences with their
+tags ambiguous words in English sentences with their % a selection of
 translation in Swedish. %The size of the annotated corpus is not
@@
 
-The ideal situation would be to have, as described by
-\cite{zinovjeva2000}, a hand-annotated evaluation corpus for testing
-the performance of the lexical-selection module. That is, the output
-of the lexical transfer module which had been disambiguated by hand by
-one or more human annotators. As this did not exist for any language
-pair, we decided to automatically annotate a test set for each
-language pair using a process similar to that described by
-\cite{vickrey05}.
+%The ideal situation would be to have, as described by
+%\cite{zinovjeva2000}, a hand-annotated evaluation corpus for testing
+%the performance of the lexical-selection module. That is, the output
+%of the lexical transfer module which had been disambiguated by hand by
+%one or more human annotators. As this did not exist for any language
+%pair, we decided to automatically annotate a test set for each
+%language pair using a process similar to that described by
+%\cite{vickrey05}. 
+%
+Ideally we would have had a hand-annotated evaluation corpus, as
+described by \cite{zinovjeva2000}, but as this did not exist, we 
+decided to automatically annotate a test set using a process similar to 
+that described by \cite{vickrey05}.
 
@@
   was used for checking the value for frequency pruning of features.} and 
-left the remainder for training. This is a smaller
-number of sentences than used in typical evaluations (such as the WMT
-series, \cite{callisonburch2012}), but was motivated by the fact that
-the smallest corpus (for Breton--French) after sentence discarding was
-only little over 4,000 sentences long. Table~\ref{table:input-corp}
+left the remainder for training. 
+%This is a smaller
+%number of sentences than used in typical evaluations (such as the WMT
+%series, \cite{callisonburch2012}), but was motivated by the fact that
+%the smallest corpus (for Breton--French) after sentence discarding was
+%only little over 4,000 sentences long. 
+Table~\ref{table:input-corp}
 gives statistics about the size of the input corpora, and how many
@@
   forms. The dictionaries allow many lexical forms to translate to one
-  lexical form. For example, \emph{ahizpa} and \emph{arreba} in Basque
-  both translate to \emph{hermana} `sister' in Spanish.\footnote{The
-    difference being that \emph{ahizpa} is a sister of a female, and
-    \emph{arreba} is a sister of a male.} But a single lexical form
+  lexical form. 
+%For example, \emph{ahizpa} and \emph{arreba} in Basque
+%  both translate to \emph{hermana} `sister' in Spanish.\footnote{The
+%    difference being that \emph{ahizpa} is a sister of a female, and
+%    \emph{arreba} is a sister of a male.} 
+  But a single lexical form
   may not have more than one translation without further
   processing. If there are many possible translations of a lexical
-  form, then one must be marked as the \emph{default} translation. For
-  example, \emph{pisu} in Basque translates to \emph{peso} `weight'
-  and \emph{piso} `flat' in Spanish. If the translation correspondence
-  \emph{pisu $\rightarrow$ peso} is perceived as the most frequent or
-  the most general, it is marked as the default by the author of the dictionary. % linguist who is
+  form, then one must be marked as the \emph{default} translation. 
+%For
+%  example, \emph{pisu} in Basque translates to \emph{peso} `weight'
+%  and \emph{piso} `flat' in Spanish. If the translation correspondence
+%  \emph{pisu $\rightarrow$ peso} is perceived as the most frequent or
+%  the most general, it is marked as the default by the author of the dictionary. % linguist who is
 %  creating the bilingual dictionary.

### 2015-02-11 2015-eamt-melsa/2015-eamt-melsa.tex
some magic got us 10 lines
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58853 72bbbca6-d526-0410-a7d9-f06f51895060
some comments
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58879 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage[small,bf]{caption}
+\usepackage{microtype} % a.k.a. conference paper magic
 
@@
 \author{
- Francis M.\ Tyers \\
- HSL-fakultehta, \\
- UiT Norgga \'{a}rktala\v{s} universitehta, \\
- N-9018 Romsa \\
+\phantom{Francis M.\ Tyers} \\
+\phantom{HSL-fakultehta,} \\
+\phantom{UiT Norgga \'{a}rktala\v{s} universitehta,} \\
+\phantom{N-9018 Romsa} \\
  \And
- Felipe S\'anchez-Mart\'inez \\
- Dept.\ Lleng.\ i Sist.\ Inform., \\
- Universitat d'Alacant,\\
- E-03071 Alacant\\
+\phantom{Felipe S\'anchez-Mart\'inez} \\
+\phantom{Dept.\ Lleng.\ i Sist.\ Inform.,} \\
+\phantom{Universitat d'Alacant,}\\
+\phantom{E-03071 Alacant}\\
  \And
- Mikel L.\ Forcada \\
- Dept.\ Lleng.\ i Sist.\ Inform., \\
- Universitat d'Alacant,\\
- E-03071 Alacant\\
+\phantom{Mikel L.\ Forcada} \\
+\phantom{Dept.\ Lleng.\ i Sist.\ Inform.,} \\
+\phantom{Universitat d'Alacant,}\\
+\phantom{E-03071 Alacant}\\
 }
@@
 \begin{document}
+
 \maketitle
+
+\renewcommand{\baselinestretch}{0.97} % the undetectable 3% squeeze
 
@@
 Corpus-based MT has been the primary research direction in the field 
-of machine translation in recent years. However, RBMT systems are 
+of machine translation in recent years. However, rule based MT (RBMT) systems are 
 still being developed, and there are many successful commercial 
@@
 
-An RBMT system typically consists of an analysis component,\footnote{For example
+An RBMT system typically consists of an analysis component,\footnote{Such as
 a morphological or syntactic analyser.} a transfer
@@
 
-Lexical selection is the task of choosing, for a given source-language
+\emph{Lexical selection} is the task of choosing, for a given source-language
 word, the most adequate translation in the target language among a
@@
 necessary to choose between a series of fine-grained senses if all
-these senses result in the same final translation; however, it may be
-necessary to choose a different translation for the same sense, for
+these senses result in the same final translation; however, it may sometimes
+be necessary to choose a different translation for the same sense, for
 example in a collocation.
@@
 
-Then \cite{berger1996} illustrated the use of maximum-entropy classifiers
+Later, \cite{berger1996} illustrated the use of maximum-entropy classifiers
 on the specific problem of lexical selection in IBM-style word-based
@@
 the translations resulting from all possible combinations of lexical
-selections.  Nowadays, in state-of-the-art phrase-based statistical
+selections.\todo{METIS too?}  Nowadays, in state-of-the-art phrase-based statistical
 machine translation \citep{koehn2009statistical}, lexical selection is
@@
 \cite{marechek10} trained a maximum-entropy lexical selector for their TectoMT rule-based system
-using a bilingual corpus.  More recently, \cite{tyers12a} presented a
+using a bilingual corpus.\todo{Say a bit more about TectoMT?}  More recently, \cite{tyers12a} presented a
 method of lexical selection for rule-based machine translation based
@@
 
-Let the probability of a translation $t$ being the translation of a
+Let the probability of a word $t$ being the translation of a
 word $s$ in an SL context $c$ be $p_s(t|c)$. In principle, this value
@@
 (1) how should the relevant contexts be chosen? and (2) what should be
-done when the translations of words are not found in the corpus?  A
+done when the translations of words are not found in the corpus?
+\todo{Wording? When $(s,t,c)$ not found in the corpus?}  A
 maximum-entropy model answers both of these questions. It allows the
@@
 h_k^{\mathrm{arrain}}(t, c) = \left \{ \begin{matrix}
-                 ~~1 & \mbox{if } t~ =~ \textit{pez}~ \mathrm{and}~ \textit{handi}~ \mathrm{follows}~ \textit{arrain}   \\
+                 ~~1 & \mbox{if } 
+\left\{ \begin{array}{c}
+t~ =~ \textit{pez}~ \\ \mathrm{and}~ \\ \textit{handi}~ \mathrm{follows}~ \textit{arrain} \end{array} \right. \\
                  ~~0 & \mbox{otherwise} \end{matrix} \right.
@@
 g_{|G|}\}$ is a set of possible \emph{lexical-selection paths}. A
-lexical-selection path $g = (t_1, t_2, \ldots, t_{|s|})$ is a sequence
-of lexical-selection choices of those SL words, where, \(t_i=T(s_i)\)
-and \(T(\cdot)\) is the set of all translations for a source
+lexical-selection path $g = (t_1, t_2, \ldots, t_{|S|})$ is a sequence
+of lexical-selection choices of those SL words, where, \(t_i=T_s(s_i)\)
+and \(T_s(\cdot)\) \todo{Check change \(T(\cdot)\to T_s(\cdot)\) for consistency with later notation} is the set of all translations for a source
 word. This is produced in the first stages of rule-based machine
@@
 path $g_i$ of sentence $S$. Figure~\ref{fig:lexselschema} shows
- this process schematically.
+ this process schematically.\todo{Can we try to bring the figure closer?}
 \begin{figure*}
@@
 specific lexical selection events \((s,t,c)\) cannot be counted, a
-target language model \(P_{TL}(\cdot)\) is used to compute a value for
+target language model \(P_\mathrm{TL}(\cdot)\) is used to compute a value for
 the fractional count for disambiguation path \(g_i\), $p(g_i|S)$ after
@@
  \label{eq:normalising}
-   p(g_i|s) = \frac{P_{TL}(\tau(g_i,S))}{\sum_{g_i \in G} P_{TL}(\tau(g_i,S))}
+   p(g_i|s) = \frac{P_\mathrm{TL}(\tau(g_i,S))}{\sum_{g_i \in G} P_\mathrm{TL}(\tau(g_i,S))}
 \end{equation}
@@
 \(g_i\) $t$ is the translation for \(s\) in context \(c\). That is, as
-if event $(s,t,c)$ had been seen a fractional number $p(g_i,s)$ of
-times. We prune features occurring less than a certain number of times 
+if event $(s,t,c)$ had been seen a fractional number $p(g_i|s)$ of
+times. We prune features\todo{features or contexts?} occurring less than a certain number of times 
 in the corpus, using a development corpus to guide pruning (see section~\ref{sec:eval-ref-results}).
@@
 whole translation task. Evaluating lexical-selection performance is an
-intrinsic module-based evaluation.  It measures how well the lexical
+\emph{intrinsic} module-based evaluation.  It measures how well the lexical
 selection module disambiguates the lexical-transfer
@@
 The
-whole translation task evaluation is an extrinsic evaluation, which
+whole translation task evaluation is an \emph{extrinsic} evaluation, which
 tests how much the system improves as regards final translation
@@
 lexical transfer, lexical selection, structural transfer and morphological
-generation. The modules of the system for this method are the \emph{lexical transfer}
+generation. The relevant \todo{OK?}  modules of the system for this method are the \emph{lexical transfer}
 module which provides for each source language lexical form a set of possible
@@
 modules which for each pair of source language lexical form and translation 
-produce a final translation in the target language.
+produce a final translation in the target language. \todo{The wording of this last sentence is unclear. Identify pre-lexsel, lexsel and post-lexsel as before}
 
@@
 morphological complexity, and different amounts of resources available
---- although for all pairs there is a parallel corpus available, which
+--- although for all pairs there is at least \todo{check "at least" here} a parallel corpus available, which
 is used for evaluation (see Section \ref{ss:measures}). 
@@
     most frequent translation among the set of possible translations
-    and marks it as the \emph{linguistic default} translation.} As a
+    and marks it as the \emph{linguistic default} translation.}\todo{Doesn't this happen in br--fr too?} As a
   result of attempting to include all possible translations, the
@@
   translation and non-default translations added according to
-  frequency. 
+  frequency.\todo{give SVN rev. for all?}\todo{Default is explained for some LPs only, why?} 
 % The version used in this paper is SVN revision
@@
   translation performance} measure. The main difference between the
-evaluation measures is that the lexical selection error rate magnifies
+evaluation measures is that the lexical selection error rate focuses on 
 the problem of lexical selection, by restricting the evaluation to
@@
 The lexical-selection error rate is the fraction of times the given
-system chooses a translation for a word which is not found in an
+system chooses a translation for a word which is not  the one found in an
 annotated reference. The process uses a source-language sentence, $S =
@@
 The table in Figure~\ref{fig:ler-example} gives an
-overview of the inputs. In the description it is assumed that the
+overview of the inputs. \todo{Could we add "diff()" and "amb()" if we had some space?} In the description it is assumed that the
 reference translation has been annotated by hand. However, hand
-annotation is a time consuming process, and was not possible. A
+annotation is a time-consuming process, and was not possible. A
 description of how the reference was built is given in
@@
   \emph{Ofis ar Brezhoneg} `The Office of the Breton
-  language'.%\footnote{In 2010 the administrative status was changed
+  language'. %\footnote{In 2010 the administrative status was changed
 %    and it was renamed \emph{Ofis Publik ar Brezhoneg} `Public office
@@
 %  previously described in \cite{tyers09b}, and also some new data. 
-The corpus has been made available online through {\sc opus}.\footnote{\url{http://opus.lingfil.uu.se}}
+The corpus has recently been made available online through {\sc opus}.\footnote{\url{http://opus.lingfil.uu.se}}
 \item \textbf{South-East European Times} (\texttt{SETimes}): Described
   in \cite{tyers10}, this corpus is a multilingual corpus of the
-  Balkan languages in the news domain. The Macedonian and English
-  portion will be used.
+  Balkan languages (and English) in the news domain. The Macedonian and English
+  portion will be used.\todo{Not in OPUS?}
 \item \textbf{Open Data Euskadi} (\texttt{OpenData}): This is a Basque
@@
   of Public
-  Administration'.\footnote{\url{http://opendata.euskadi.net/w79-contdata/es/contenidos/ds_recursos_linguisticos/memorias_traduccion/es_izo/memorias_traduccion_izo.html}}
+  Administration'.\footnote{\url{http://tinyurl.com/eu-es-tm}}\todo{shortened URL}
 \item \textbf{European Parliament Proceedings} (\texttt{EuroParl}):
@@
   European Union official languages. We are using the English--Spanish
-  data from version 7.
+  data from version 7.\todo{URL?}
 \end{itemize}
@@
 these alignments along with the bilingual dictionary of the MT system
-in question to extract only those sentences where: there is at least
-one ambiguous word; that ambiguous word is aligned to a single word in
-the target language; and the word it is aligned to in the target
+in question to extract only those sentences where: (a) there is at least
+one ambiguous word; (b) that ambiguous word is aligned to a single word in
+the target language; and (c) the word it is aligned to in the target
 language is found in the bilingual dictionary of the MT
@@
 sentences were left after processing for testing, training and
-development. %Table~\ref{table:test-corp} gives information about the
+development.
+\todo{Why do ambiguity numbers change so much from the paper?
+Having devel and test in different columns would make the caption smaller and the table would still fit}%Table~\ref{table:test-corp} gives information about the
 %test corpora.
+\todo{``av. amb'' is a corpus average or a dictionary average?}
 
@@
   output. This is the method used by %@@@ \cite{carbonell06} and
-  \cite{melero07a}.  
+  \cite{melero07a}.  \todo{Also METIS?}
 % We used a five-gram language model of surface
@@
 were selected; they happen to come close to the rule-of-thumb value of
-five that \citet[p.~596]{Manning99b} find to be effective.
+five that \citet[p.~596]{Manning99b} found to be effective.
 Table~\ref{table:maxent-features} shows the number of features that
-have eventually been sed for each language pair.
+have eventually been used for each language pair.
 
@@
  to the target-language model performance, at a much smaller
- computational cost.
+ computational cost.\todo{I miss a brief separate discussion of BLEU and LER, which are reported separately!}
 
@@
 unsupervised way, that is, without the need for an annotated corpus,
-in this case a word-aligned bilingual corpus: one just needs a
+(in this case a word-aligned bilingual corpus): one just needs a
 source-language corpus, a statistical target-language model, and the
@@
 rule-based machine translation system. The system uses a
-maximum-entropy formalism for lexical selection as \cite{berger1996}
+maximum-entropy formalism for lexical selection, as \cite{berger1996}
 and \cite{marechek10}, but instead of counting actual lexical
@@
 %The authors thank Filip Petkovski for his help in running some of the
-%experiments. They also acknowledge partial support from the Ministry
-%of Science and Innovation of the Spanish Government through project
-%TIN2009-14009-C02-01, and the NILS mobility grant (Abel Predoc
+%experiments. They also acknowledge partial support from the 
+%Spanish Government through projects 
+%TIN2009-14009-C02-01 and TIN2012-32615, the European Commission through project
+%Abu-MaTran (FP7-PEOPLE, 324414) (and the NILS mobility grant (Abel Predoc
 %Research Grant) coordinated by the Universidad Complutense de Madrid.

### 2015-02-12 2015-eamt-melsa/2015-eamt-melsa.tex
added in an old table, but went a bit over :(
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58894 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
+%%% THIS IS IN A WEIRD PLACE BUT WE WANT IT ON PAGE 3 -FMT
+\begin{figure*}
+\begin{displaymath}
+  \label{eq:steps}
+  S \to \mbox{\framebox{\parbox{0.8cm}{\textit{pre- lexsel}}}} \to (\{g_i\}_{i=1}^{i=|G|},S) \to \mbox{\framebox{\parbox{1.0cm}{\textit{\textbf{lexsel}}}}} \to (g^\star,S) \to \mbox{\framebox{\parbox{0.8cm}{\textit{post- lexsel}}}} \to \tau(g^\star,S)
+\end{displaymath}
+ \caption{A schema of the training process}
+ \label{fig:lexselschema}
+\end{figure*}
+
 As a result of training, each feature in the classifier is assigned a
@@
 path $g_i$ of sentence $S$. Figure~\ref{fig:lexselschema} shows
- this process schematically.\todo{Can we try to bring the figure closer?}
-\begin{figure*}
-\begin{displaymath}
-  \label{eq:steps}
-  S \to \mbox{\framebox{\parbox{0.8cm}{\textit{pre- lexsel}}}} \to (\{g_i\}_{i=1}^{i=|G|},S) \to \mbox{\framebox{\parbox{1.0cm}{\textit{\textbf{lexsel}}}}} \to (g^\star,S) \to \mbox{\framebox{\parbox{0.8cm}{\textit{post- lexsel}}}} \to \tau(g^\star,S)
-\end{displaymath}
- \caption{A schema of the training process}
- \label{fig:lexselschema}
-\end{figure*}
+ this process schematically. %%%\todo{Can we try to bring the figure closer?}
 
@@
 if event $(s,t,c)$ had been seen a fractional number $p(g_i|s)$ of
-times. We prune features\todo{features or contexts?} occurring less than a certain number of times 
+times. We prune features \todo{features or contexts?} occurring less than a certain number of times 
 in the corpus, using a development corpus to guide pruning (see section~\ref{sec:eval-ref-results}).
@@
 whole translation task evaluation is an \emph{extrinsic} evaluation, which
-tests how much the system improves as regards final translation
+tests how the system improves as regards final translation
 quality in a real system.
@@
 lexical transfer, lexical selection, structural transfer and morphological
-generation. The relevant \todo{OK?}  modules of the system for this method are the \emph{lexical transfer}
+generation. The relevant modules of the system for this method are the \emph{lexical transfer}
 module which provides for each source language lexical form a set of possible
@@
 morphological complexity, and different amounts of resources available
---- although for all pairs there is at least \todo{check "at least" here} a parallel corpus available, which
-is used for evaluation (see Section \ref{ss:measures}). 
+--- although for all pairs there is a parallel corpus available
+for evaluation (see Section \ref{ss:measures}). 
 % Table~\ref{table:lang-pair-stats} 
@@
     most frequent translation among the set of possible translations
-    and marks it as the \emph{linguistic default} translation.}\todo{Doesn't this happen in br--fr too?} As a
+    and marks it as the \emph{linguistic default} translation.} As a
   result of attempting to include all possible translations, the
@@
 % The transfer rules were written
-%  by hand to treat translation problems in the corpus. The version of
-%  the software used in this paper is SVN revision
-%  41476.\footnote{\url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-mk-en}}.
+%  by hand to treat translation problems in the corpus. 
+The version of
+  the software used in this paper is SVN revision
+  41476.\footnote{\url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-mk-en}}
 \item[Basque--Spanish:] The development of the Basque to Spanish pair
@@
   Alternative translations were included in the bilingual
-  dictionary.% , but non-default translations were marked with a
+  dictionary.\todo{SVN rev}
+% , but non-default translations were marked with a
 %  direction restriction\comment{FSM: This needs to be carefully
@@
   translation and non-default translations added according to
-  frequency.\todo{give SVN rev. for all?}\todo{Default is explained for some LPs only, why?} 
-% The version used in this paper is SVN revision
-%  41387.\footnote{\url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-en-es}}
+  frequency.\todo{Default is explained for some LPs only, why?} 
+ The version used in this paper is SVN revision
+  41387.\footnote{\url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-en-es}}
 \end{description}
@@
 \subsubsection{Confidence intervals}
-Confidence intervals for both metrics will be calculated through the
-\emph{bootstrap resampling} method~\citep{efron94} as described
+Confidence intervals for both metrics will be calculated through 
+\emph{bootstrap resampling}~\citep{efron94} as described
 by~\citet{koehn04}. %\footnote{Broadly speaking, % and \citet{zhang04}
@@
   Balkan languages (and English) in the news domain. The Macedonian and English
-  portion will be used.\todo{Not in OPUS?}
+  part will be used.%\todo{Not in OPUS?}
 \item \textbf{Open Data Euskadi} (\texttt{OpenData}): This is a Basque
@@
   of Public
-  Administration'.\footnote{\url{http://tinyurl.com/eu-es-tm}}\todo{shortened URL}
+  Administration'.\footnote{\url{http://tinyurl.com/eu-es-tm}}
 \item \textbf{European Parliament Proceedings} (\texttt{EuroParl}):
@@
   European Union official languages. We are using the English--Spanish
-  data from version 7.\todo{URL?}
+  data from version 7.\footnote{\url{http://www.statmt.org/europarl/}}
 \end{itemize}
@@
 
-%  \begin{table*}
-%  \begin{center}
-%   \begin{tabular}{|l|r|r|r|r|r|}
-%      \hline 
-%      {\bf Pair}      & {\bf Lines} & {\bf SL words} & {\bf TL words} & {\bf Amb. words} & {\bf \% ambig} \\ 
-%      \hline 
-%       \texttt{br-fr} & 1,000 & 13,854  & 13,878  & 1,163 & 8.39   \\
-%      \hline 
-%       \texttt{mk-en} & 1,000 & 13,441 & 14,228  & 3,872 & 28.80 \\
-%      \hline 
-%       \texttt{eu-es} & 1,000 & 7,967 & 11,476  & 1,360  & 17.07 \\
-%      \hline 
-%       \texttt{en-es} & 1,000 & 19,882 & 20,944  & 1,469  & 7.38 \\
-%      \hline 
-%   \end{tabular} 
-%  \end{center}
-%   \caption{Statistics about the test corpora. The columns \textbf{amb.\ words} and \textbf{\%~ambig} gives 
-%      the number of word with more than one translation and the percentage of SL words which have 
-%      more than one translation respectively.}
-%    \label{table:test-corp}
-%  \end{table*}
+  \begin{table}
+  \begin{center}
+   \begin{tabular}{|l|r|r|r|r|}
+      \hline 
+      %{\bf Pair}      & {\bf Lines} & {\bf SL words} & {\bf TL words} & {\bf Amb. words} & {\bf \% ambig} \\ 
+      {\bf Pair}      & {\bf SL} & {\bf TL} & {\bf Amb.} & {\bf \% amb.} \\ 
+      \hline 
+       \texttt{br-fr} & 13,854  & 13,878  & 1,163 & 8.39   \\
+      \hline 
+       \texttt{mk-en} & 13,441 & 14,228  & 3,872 & 28.80 \\
+      \hline 
+       \texttt{eu-es} & 7,967 & 11,476  & 1,360  & 17.07 \\
+      \hline 
+       \texttt{en-es} & 19,882 & 20,944  & 1,469  & 7.38 \\
+      \hline 
+   \end{tabular} 
+  \end{center}
+   \caption{Statistics about the test corpora. The columns \textbf{SL} and \textbf{TL} give the number
+      of tokens in the source and target languages respectively. The columns \textbf{amb.\ words} and \textbf{\%~ambig} gives 
+      the number of word with more than one translation and the percentage of SL words which have 
+      more than one translation respectively.}
+    \label{table:test-corp}
+  \end{table}
 
@@
 \todo{Why do ambiguity numbers change so much from the paper?
-Having devel and test in different columns would make the caption smaller and the table would still fit}%Table~\ref{table:test-corp} gives information about the
-%test corpora.
+Having devel and test in different columns would make the caption smaller and the table would still fit}Table~\ref{table:test-corp} gives information about the
+test corpora.
 \todo{``av. amb'' is a corpus average or a dictionary average?}
@@
         \hline
-        \textbf{Pair} & \textbf{Pruned} & \textbf{\# features} \\\hline
-        \hline
-        {\bf br-fr}  & $< 5$              & 5,277    \\ 
-        \hline
-        {\bf mk-en} & $< 7$              & 205,494 \\ 
-        \hline
-        {\bf eu-es} & $< 7$              & 196,024\\
-        \hline
-        {\bf en-es} & $< 7$              & 195,605\\
+        \textbf{Pair} & \textbf{Pruned} & \textbf{\# features} \\
+        \hline
+        {\tt br-fr}  & $< 5$              & 5,277    \\ 
+        \hline
+        {\tt mk-en} & $< 7$              & 205,494 \\ 
+        \hline
+        {\tt eu-es} & $< 7$              & 196,024\\
+        \hline
+        {\tt en-es} & $< 7$              & 195,605\\
         \hline

### 2015-02-15 2015-eamt-melsa/2015-eamt-melsa.tex
some improvements
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58925 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 transfer output is the result of looking up the translations of the 
-source-language \emph{lexical forms} --- combinations of lemma and 
+source-language \emph{lexical forms} --- lemmas and  % combinations of lemma and tags
 tags --- in the bilingual dictionary.
@@
    \hline
+   $\textrm{amb}(s_i)$ & 0     & 0          & 0      & 0     & 1            & 1 \\
+   \hline
+   $\mathrm{diff}(T_r(s_i), T_t(s_i))$ & 0     & 0          & 0      & 0     & 1            & 0 \\
+   \hline
  \end{tabular}
@@
 are acceptable for $s_i$ in sentence $S$.\footnote{Depending on how
-  the reference is built, the set returned by $T_r(s_i)$ may be
-  incomplete in the sense that it may not include all possible
+  the reference is built, the set returned by $T_r(s_i)$ % may be
+  % incomplete in the sense that it 
+  may not include all possible
   acceptable translations.} For a
@@
 The table in Figure~\ref{fig:ler-example} gives an
-overview of the inputs. \todo{Could we add "diff()" and "amb()" if we had some space?} In the description it is assumed that the
+overview of the inputs. In the description it is assumed that the
 reference translation has been annotated by hand. However, hand
@@
 
-For creating the test corpora, and for providing a source language corpus 
+For creating the test corpora, providing a source language corpus 
 for training, and a target language corpus for scoring, we used four parallel corpora: 
@@
 \begin{center}
- \begin{tabular}{|l|r|r||r|r|r|r|}
+ \begin{tabular}{|l|r|r||r|r|r|r|r|}
     \hline  
     %Pair           & Lines & Words & Extracted & \texttt{train} & \texttt{test} & \texttt{dev} & No. ambig & Av. ambig \\ 
-    {\bf Pair}      & {\bf Lines} & {\bf Extract.} & \texttt{train} & \texttt{other} & {\bf No. amb} & {\bf Av. amb} \\ 
-    \hline  
-     %\texttt{br-fr} & 57,305  & 702,328  & 4,668 & 2,668 & 1,000  & 1,000 & 603  & 1.07 \\
-     \texttt{br-fr} & 57,305  & 4,668 & 2,668 & 2,000  & 603  & 3.06 \\
+    {\bf Pair}      & {\bf Lines} & {\bf Extract.} & \texttt{train} & \texttt{dev} & \texttt{test} & {\bf No. amb} & {\bf Av. amb} \\ 
+    \hline  
+     %\texttt{br-fr} & 57,305  & 702,328  & 4,668 & 2,668 & & 1,000  & 1,000 & 603  & 1.07 \\
+     \texttt{br-fr} & 57,305  & 4,668 & 2,668 & 1,000 & 1,000  & 603  & 3.06 \\
     \hline  
      %\texttt{mk-en} & 190,493  & 4,259,338  & 19,747  & 17,747  & 1,000 & 1,000 & 13,134  & 1.86  \\
-     \texttt{mk-en} & 190,493  & 19,747  & 17,747  & 2,000 & 13,134  & 3.06  \\
+     \texttt{mk-en} & 190,493  & 19,747  & 17,747  & 1,000 & 1,000 & 13,134  & 3.06  \\
     \hline
      %\texttt{eu-es} & 765,115  & 10,190,079  & 87,907  & 85,907   & 1,000 & 1,000 & 1,806  & 1.30 \\
-     \texttt{eu-es} & 765,115  & 87,907  & 85,907  & 2,000 & 1,806  & 3.11 \\
+     \texttt{eu-es} & 765,115  & 87,907  & 85,907  & 1,000 & 1,000 & 1,806  & 3.11 \\
     \hline
      %\texttt{en-es} & 1,467,708 & 30,154,098  & 312,162 & 310,162 & 1,000 & 1,000 & 2,082  & 1.08 \\
-     \texttt{en-es} & 1,467,708 & 312,162 & 310,162 & 2,000 & 2,082  & 2.28 \\
+     \texttt{en-es} & 1,467,708 & 312,162 & 310,162 & 1,000 & 1,000 & 2,082  & 2.28 \\
     \hline  
  \end{tabular}
- \caption{Statistics about the source corpora. The column {\bf other}
-   gives the number of sentences reserved for testing (1,000) and
-   development (1,000). The column {\bf no.\ amb} gives the number of
+ \caption{Statistics about the source corpora. The column {\bf no.\ amb} gives the number of
    unique tokens with more than one possible translation. The column
@@
 sentences were left after processing for testing, training and
-development.
-\todo{Why do ambiguity numbers change so much from the paper?
-Having devel and test in different columns would make the caption smaller and the table would still fit}Table~\ref{table:test-corp} gives information about the
+development. \todo{Why do ambiguity numbers change so much from the paper?}Table~\ref{table:test-corp} gives information about the
 test corpora.
-\todo{``av. amb'' is a corpus average or a dictionary average?}
 
@@
  \end{center}
- \caption{Number of features (rules) in each rule set and pruning frequency.}
+ \caption{Features in each rule set and pruning frequency.}
  \label{table:maxent-features}

### 2015-02-15 2015-eamt-melsa/2015-eamt-melsa.tex
some fixes from the plane
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58927 72bbbca6-d526-0410-a7d9-f06f51895060

@@
   default.\footnote{A bilingual dictionary in Apertium
-    \citep{forcada2011apertium} may contain more than one possible
+    \citep{forcada2011apertium} may contain $>1$ possible
     target-language translation for a given source-language
-    word. Where there is more than one possible translation, the
-    person who is writing the dictionary chooses the most general or
+    word. Where there is $>1$ possible translation, the
+    dictionary writer chooses the most general or
     most frequent translation among the set of possible translations
@@
   Alternative translations were included in the bilingual
-  dictionary.\todo{SVN rev}
+  dictionary. The version used is SVN revision 44846.\footnote{\url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-eu-es}}
 % , but non-default translations were marked with a
@@
   translation through the rest of the modules of the MT system. This
-  is included to give an idea of the upper bound for the performance
+  is included to show the upper bound for the performance
   of the lexical-selection module. % If all of the lexical choices were

### 2015-02-15 2015-eamt-melsa/2015-eamt-melsa.tex
take care of comments
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58928 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 the translations resulting from all possible combinations of lexical
-selections.\todo{METIS too?}  Nowadays, in state-of-the-art phrase-based statistical
+selections. Nowadays, in state-of-the-art phrase-based statistical % \todo{METIS too?}
 machine translation \citep{koehn2009statistical}, lexical selection is
@@
   output. This is the method used by %@@@ \cite{carbonell06} and
-  \cite{melero07a}.  \todo{Also METIS?}
+  \cite{melero07a}.%  \todo{Also METIS?}
 % We used a five-gram language model of surface

### 2015-02-17 2015-eamt-melsa/2015-eamt-melsa.tex
some misc. plane improvements
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58934 72bbbca6-d526-0410-a7d9-f06f51895060
fixes from felipe
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58978 72bbbca6-d526-0410-a7d9-f06f51895060

@@
-% test
-
 \documentclass[11pt]{article}
@@
 
-\newcommand{\confname}{EAMT 2015}
-\newcommand{\website}{http://www.eamt2015.org/}
-\newcommand{\contactname}{the conference chairs (Felipe
-  S\'anchez-Mart\'inez, Gema Ram\'irez-S\'anchez and Fred Hollowood)}
-\newcommand{\contactemail}{eamt2015@dlsi.ua.es}
-\newcommand{\conffilename}{eamt15}
-\newcommand{\downloadsite}{http://www.eamt2015.org/}
-\newcommand{\paperlength}{$8$ (eight)}
-\newcommand{\shortpaperlength}{$4$ (four)}
-
 % \newcommand{\comment}[1]{{\small{\texttt{$\clubsuit$#1$\clubsuit$}}}}
@@
     selection in rule-based machine translation}
-
-%\author{}
 
@@
   does not require any annotated corpus. The method uses
-  source-language monolingual corpora, an existing machine translation
-  (MT) system and a statistical target-language model. Using the
-  existing MT system, the sentences in the source-language corpus are
+  source-language monolingual corpora, the machine translation (MT)
+  system in which the models are integrated and a statistical
+  target-language model. Using the MT system, \todo{FSM: Check
+    rewording! I have removed existing MT from this and previous
+    sentence.} the sentences in the source-language corpus are
   translated in all possible ways according to the different
@@
   the scores are normalised to provide fractional counts for training
-  a source-language maximum-entropy lexical-selection model. We show
-  that this model can perform equally well, or better, than using the
-  target-language model directly for lexical selection, at a
+  source-language maximum-entropy lexical-selection models. We show
+  that these models can perform equally well, or better, than using
+  the target-language model directly for lexical selection, at a
   substantially reduced computational cost.
@@
 Corpus-based MT has been the primary research direction in the field
-of machine translation in recent years. However, rule based MT (RBMT) systems are 
-still being developed, and there are many successful commercial 
-and non-commercial systems. One reason for the continued development
-of RBMT systems is that in order to be successful, corpus-based MT requires parallel corpora in
-the order of tens of millions of words. Although for some language
-pairs these exist, they only exist for a fraction of the 
-world's languages.
-
-An RBMT system typically consists of an analysis component,\footnote{Such as
-a morphological or syntactic analyser.} a transfer
-component and a generation component. As part of the transfer component
-it is necessary to make choices regarding words in the source language
-which may have more than one translation in the target language.
-
-\emph{Lexical selection} is the task of choosing, for a given source-language
-word, the most adequate translation in the target language among a
-known set of alternatives.  The task is related to the task of
-word-sense disambiguation \citep{ide98a}. However, it is different to
-word-sense disambiguation in that lexical selection is a bilingual
-problem, not a monolingual problem: its aim is to find the most
-adequate translation, not the most adequate sense. Thus, it is not
-necessary to choose between a series of fine-grained senses if all
-these senses result in the same final translation; however, it may sometimes
-be necessary to choose a different translation for the same sense, for
-example in a collocation.
+of machine translation in recent years. However, rule-based MT (RBMT)
+systems are still being developed, and there are many successful
+commercial and non-commercial systems. One reason for the continued
+development of RBMT systems is that in order to be successful,
+corpus-based MT requires parallel corpora in the order of tens of
+millions of words. Although for some language pairs these exist, they
+only exist for a fraction of the world's languages.
+
+An RBMT system typically consists of an analysis
+component,\footnote{Such as a morphological or syntactic analyser.} a
+transfer component and a generation component. As part of the transfer
+component it is necessary to make choices regarding words in the
+source language which may have more than one translation in the target
+language.
+
+\emph{Lexical selection} is the task of choosing, for a given
+source-language word, the most adequate translation in the target
+language among a known set of alternatives.  The task is related to
+the task of word-sense disambiguation \citep{ide98a}. However, it is
+different to word-sense disambiguation in that lexical selection is a
+bilingual problem, not a monolingual problem: its aim is to find the
+most adequate translation, not the most adequate sense. Thus, it is
+not necessary to choose between a series of fine-grained senses if all
+these senses result in the same final translation; however, it may
+sometimes be necessary to choose a different translation for the same
+sense, for example in a collocation.
 % It could also be the case that a single sense could have more than
@@
 %In their pioneering work, 
-\cite{dagan1994word} used the term
-\emph{word sense disambiguation} to refer to what is actually target
-word selection in machine translation; they used a parser to identify
-some syntactic relations such as subject--object or subject--verb.
-After generating all the possible translations for a given
-input sentence using an ambiguous bilingual dictionary, they extract the syntactic
-tuples from the target language and count the frequency in a previously-trained target-language 
-model of tuples.  They use maximum-likelihood estimation to calculate the probability that a given 
-target language tuple is the translation of a given source language tuple, with
-an automatically determined confidence threshold.
-
-Later, \cite{berger1996} illustrated the use of maximum-entropy classifiers
-on the specific problem of lexical selection in IBM-style word-based
-statistical machine translation. Other authors
-\citep{melero07a} have used target-language models to rank %@@@ carbonell06
+\cite{dagan1994word} used the term \emph{word sense disambiguation} to
+refer to what is actually target word selection in machine
+translation; they used a parser to identify some syntactic relations
+such as subject--object or subject--verb.  After generating all the
+possible translations for a given input sentence using an ambiguous
+bilingual dictionary, they extract the syntactic tuples from the
+target language and count the frequency in a previously-trained
+target-language model of tuples.  They use maximum-likelihood
+estimation to calculate the probability that a given target language
+tuple is the translation of a given source language tuple, with an
+automatically determined confidence threshold.
+
+Later, \cite{berger1996} illustrated the use of maximum-entropy
+classifiers on the specific problem of lexical selection in IBM-style
+word-based statistical machine translation. Other authors
+\citep{melero07a} have used target-language models to
+rank %@@@ carbonell06
 the translations resulting from all possible combinations of lexical
-selections. Nowadays, in state-of-the-art phrase-based statistical % \todo{METIS too?}
+selections. Nowadays, in state-of-the-art phrase-based
+statistical % \todo{METIS too?}
 machine translation \citep{koehn2009statistical}, lexical selection is
@@
 model provides probabilities of word sequences in the target language.
-%As regards rule-based machine translation, \cite{sanchez07a} proposed 
-%to use information from the target language to train a statistical 
-%lexical-selection module but did not provide any experiments. 
-%@@@ \todo[inline]{FMT: integrate \citep{sanchez07a} somewhere here}
-\cite{marechek10} trained a maximum-entropy lexical selector for their TectoMT rule-based system
-using a bilingual corpus.\todo{Say a bit more about TectoMT?}  More recently, \cite{tyers12a} presented a
-method of lexical selection for rule-based machine translation based
-on rules which select or remove translations in fixed-length contexts,
-along with a training method for learning the rules from a
-word-aligned parallel corpus.
-
-% The principal disadvantage with this
-% method is that it relies on a parallel corpus for the language pair,
-% something which is not always available.
-
-%The remainder of the article is organised as follows: 
-%\todo[inline]{FMT: Add article structure}
+% As regards rule-based machine translation, \cite{sanchez07a}
+% proposed to use information from the target language to train a
+% statistical lexical-selection module but did not provide any
+% experiments.  @@@ \todo[inline]{FMT: integrate \citep{sanchez07a}
+% somewhere here}
+\cite{marechek10} trained a maximum-entropy lexical selector for their
+TectoMT rule-based system using a bilingual corpus.\todo{Say a bit
+  more about TectoMT? FSM: Yes, please!}  More recently,
+\cite{tyers12a} presented a method of lexical selection for rule-based
+machine translation based on rules which select or remove translations
+in fixed-length contexts, along with a training method for learning
+the rules from a word-aligned parallel corpus.
+
+% The principal disadvantage with this method is that it relies on a
+% parallel corpus for the language pair, something which is not always
+% available.
+
+% The remainder of the article is organised as follows:
+% \todo[inline]{FMT: Add article structure}
 
@@
 % equation~(\ref{eq:max-ent-max}).
-
 \begin{equation}
-  t^\star=\underset{t \in T_s(s)}{\arg\max} ~ p_s(t|c) = \underset{t \in T_s(s)}{\arg\max} \sum_{k=1}^{n_F} \lambda_k^s h_k^s(t, c) .
+  t^\star=\underset{t \in T_s(s)}{\arg\max} ~ p_s(t|c) = \underset{t \in T_s(s)}{\arg\max} \sum_{k=1}^{n_F} \lambda_k^s h_k^s(t, c) ,
 \label{eq:max-ent-max}
 \end{equation}
+where $T_s(s)$ is the set of possible translations for SL word $s$.
 
@@
 above both take advantage of a parallel corpus to collect counts of
-contexts and translations in order to train the maximum-entropy
-model. However, parallel corpora are not available for the majority of
-the world's written languages. In this section we describe an
-unsupervised method to learn the model using only monolingual corpora
-and the components from an existing machine translation system.
+contexts and translations in order to train maximum-entropy
+models. However, parallel corpora are not available for the majority
+of the world's written languages. In this section we describe an
+unsupervised method to learn the models using only monolingual corpora
+and the components from the RBMT system in which they are
+used. \todo{FSM: Check rewording!}
 
@@
 $\mathcal{G} = (S, G)$, where $S = (s_1, s_2, \ldots s_{|S|})$ is a
-sequence of source-language words, and $G = \{g_1, g_2, \ldots
-g_{|G|}\}$ is a set of possible \emph{lexical-selection paths}. A
-lexical-selection path $g = (t_1, t_2, \ldots, t_{|S|})$ is a sequence
-of lexical-selection choices of those SL words, where, \(t_i=T_s(s_i)\)
-and \(T_s(\cdot)\) \todo{Check change \(T(\cdot)\to T_s(\cdot)\) for consistency with later notation} is the set of all translations for a source
-word. This is produced in the first stages of rule-based machine
-translation, just after morphological analysis, part-of-speech
-tagging, and bilingual dictionary lookup, and before any structural
-transfer takes place (we will call this \emph{pre-lexsel}). In our
-model, it is after these first stages that lexical selection
-(\emph{lexsel}) occurs. After lexical selection, structural transfer
-and generation take place; a function $\tau(g_i,S)$ represents the
-result of these last stages, which we will call \emph{post-lexsel},
-and returns a finished translation of a specific lexical-selection
-path $g_i$ of sentence $S$. Figure~\ref{fig:lexselschema} shows
- this process schematically. %%%\todo{Can we try to bring the figure closer?}
+sequence of SL words, and $G = \{g_1, g_2, \ldots g_{|G|}\}$ is a set
+of possible \emph{lexical-selection paths}. A lexical-selection path
+$g = (t_1, t_2, \ldots, t_{|S|})$ is a sequence of lexical-selection
+choices of those SL words, where \(t_i \in
+T_s(s_i)\). % and \(T_s(\cdot)\)
+\todo{Check change \(T(\cdot)\to T_s(\cdot)\) for consistency with
+  later notation. FSM: Check \(t_i \in T_s(s_i)\); before it was \(t_i
+  = T_s(s_i)\)} %is the set of all
+%translations for a source word. 
+This is produced in the first stages of RBMT, just after morphological
+analysis, part-of-speech tagging, and bilingual dictionary lookup, and
+before any structural transfer takes place (we will call this
+\emph{pre-lexsel}). In our model, it is after these first stages that
+lexical selection (\emph{lexsel}) occurs. After lexical selection,
+structural transfer and generation take place; a function
+$\tau(g_i,S)$ represents the result of these last stages, which we
+will call \emph{post-lexsel}, and returns a finished translation of a
+specific lexical-selection path $g_i$ of sentence
+$S$. Figure~\ref{fig:lexselschema} shows this process
+schematically. %%%\todo{Can we try to bring the figure closer?}
 
@@
  \label{eq:normalising}
-   p(g_i|s) = \frac{P_\mathrm{TL}(\tau(g_i,S))}{\sum_{g_i \in G} P_\mathrm{TL}(\tau(g_i,S))}
+   p(g_i|S) = \frac{P_\mathrm{TL}(\tau(g_i,S))}{\sum_{g_i \in G} P_\mathrm{TL}(\tau(g_i,S))}
 \end{equation}
 ~\\
-The maximum-entropy model is trained instead using the fractional count
-$p(g_i|S)$ for the events $(s,t,c)$ found in \(g_i\), that is, when in
-\(g_i\) $t$ is the translation for \(s\) in context \(c\). That is, as
-if event $(s,t,c)$ had been seen a fractional number $p(g_i|s)$ of
-times. We prune features \todo{features or contexts?} occurring less than a certain number of times 
-in the corpus, using a development corpus to guide pruning (see section~\ref{sec:eval-ref-results}).
+The maximum-entropy model is trained instead using the fractional
+count $p(g_i|S)$ for the events $(s,t,c)$ found in \(g_i\), that is,
+when in \(g_i\) $t$ is the translation for \(s\) in context
+\(c\). That is, as if event $(s,t,c)$ had been seen a fractional
+number $p(g_i|S)$ of times. We prune features (contexts)
+\todo{features or contexts?} occurring less than a certain number of
+times in the corpus, using a development corpus to guide pruning (see
+section~\ref{sec:eval-ref-results}).
 %@@@ \comment{Reword this to avoid repetition} 
-The method used here
-for lexical selection is analogous to the method used by
-\cite{sanchez08b} to train a hidden-Markov-model-based part-of-speech
-tagger in a rule-based machine translation system. 
+The method used here for lexical selection is analogous to the method
+used by \cite{sanchez08b} to train a hidden-Markov-model-based
+part-of-speech tagger in a rule-based machine translation system.
 %Table~\ref{tb:workedexample} shows how a collection of samples
@@
 
-Apertium is a free/open-source rule-based machine translation platform, 
-it comprises an engine, a toolbox and data to build rule-based machine
-translation systems. Translation is implemented as a pipeline consisting
-of the following modules: morphological analysis, morphological disambiguation,
-lexical transfer, lexical selection, structural transfer and morphological
-generation. The relevant modules of the system for this method are the \emph{lexical transfer}
-module which provides for each source language lexical form a set of possible
-target language forms and the \emph{structural transfer} and \emph{morphological generation}
-modules which for each pair of source language lexical form and translation 
-produce a final translation in the target language. \todo{The wording of this last sentence is unclear. Identify pre-lexsel, lexsel and post-lexsel as before}
+Apertium is a free/open-source RBMT platform, it comprises an engine,
+a toolbox and data to build RBMT systems. Translation is implemented
+as a pipeline consisting of the following modules: morphological
+analysis, morphological disambiguation, lexical transfer, lexical
+selection, structural transfer and morphological generation. The
+relevant modules of the system for this method are the \emph{lexical
+  transfer} module which provides for each source language lexical
+form a set of possible target language lexical forms and the
+\emph{structural transfer} and \emph{morphological generation} modules
+which for each pair of source language lexical form and translation
+produce a final translation in the target language. \todo{The wording
+  of this last sentence is unclear. Identify pre-lexsel, lexsel and
+  post-lexsel as before}
 
@@
 %  system. 
-  The most probable entry from automatic word alignment of this 
-  corpus using {\sc giza++} \citep{och03a} was checked to ensure that it was 
-  an adequate translation, and if so marked as the 
-  default.\footnote{A bilingual dictionary in Apertium
-    \citep{forcada2011apertium} may contain $>1$ possible
-    target-language translation for a given source-language
-    word. Where there is $>1$ possible translation, the
-    dictionary writer chooses the most general or
-    most frequent translation among the set of possible translations
-    and marks it as the \emph{linguistic default} translation.} As a
-  result of attempting to include all possible translations, the
-  average number of translations per word is much higher % (see table~\ref{table:lang-pair-stats})
+  The most probable entry from automatic word alignment of this corpus
+  using {\sc giza++} \citep{och03a} was checked to ensure that it was
+  an adequate translation, and if so marked as the default.\footnote{A
+    bilingual dictionary in Apertium \citep{forcada2011apertium} may
+    contain several target-language translation for a given SL
+    word. Where there is more than one possible translation, the
+    dictionary writer chooses the most general or most frequent
+    translation among the set of possible translations and marks it as
+    the \emph{linguistic default} translation.} As a result of
+  attempting to include all possible translations, the average number
+  of translations per word is much
+  higher % (see table~\ref{table:lang-pair-stats})
   than in other pairs.
@@
   a combination of the English--Catalan and Spanish--Catalan pairs
-  over a period of around 3--5 months. The pair 
+  over a period of around 3--5 months.\todo{FSM: 3, 4 or 5?} The pair
 %has been used in other
@@
 %    (it mentions the use of TL info for lexical selection.},
-  contains a number of entries in the bilingual dictionary with
-  more than one translation. The most-frequent translation, as judged
-  by the language pair developer, is marked as the \emph{default}
-  translation and non-default translations added according to
-  frequency.\todo{Default is explained for some LPs only, why?} 
- The version used in this paper is SVN revision
+  contains a number of entries in the bilingual dictionary with more
+  than one translation. The most-frequent translation, as judged by
+  the language pair developer, is marked as the \emph{default}
+  translation and non-default translations were added according to
+  frequency. \todo{Default is explained for some LPs only, why?}  The
+  version used in this paper is SVN revision
   41387.\footnote{\url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-en-es}}
@@
 performance of the lexical selection method proposed here: a
-\emph{lexical selection performance} measure and a \emph{machine
-  translation performance} measure. The main difference between the
-evaluation measures is that the lexical selection error rate focuses on 
-the problem of lexical selection, by restricting the evaluation to
-this feature. Other features of the MT system, such as the transfer rules
-and morphological generation are not taken into account.
+(intrinsic) \emph{lexical selection performance} measure and an
+(extrinsic) \emph{machine translation performance} measure. \todo{FSM:
+I moved some stuff from here to the LER subsection.}
 
@@
 compared to a gold-standard corpus. For this task, we define a metric,
-the lexical-selection error rate ({\sc ler}),
+the lexical-selection error rate ({\sc ler}), that focuses on the
+problem of lexical selection by restricting the evaluation to this
+feature; other features of the MT system, such as the transfer rules
+and morphological generation, are not taken into account.
+
 % see section~\ref{sec:ler}. 
@@
 \begin{equation}
-\mathrm{LER} = \frac{\sum_{i=1}^{|S|} \mathrm{amb}(s_i) \mathrm{diff}(T_r(s_i), T_t(s_i)}{\sum_{i=1}^{|S|} \mathrm{amb}(s_i)},
+\mathrm{LER} = \frac{\sum_{i=1}^{|S|} \mathrm{amb}(s_i)\; \mathrm{diff}(T_r(s_i), T_t(s_i))}{\sum_{i=1}^{|S|} \mathrm{amb}(s_i)},
 \label{eq:ler}
@@
 when the lexical-selection module returns more than one translation
-the first will be selected.\footnote{In practice this does not happen
+the first one will be selected.\footnote{In practice this does not happen
 as each ambiguous word has a \emph{default} translation.}
@@
 %! \comment{This part of the text needs compressing and merging} 
-This is
-an extrinsic evaluation, which ideally would test how much the system
-improves as regards an approximate measurement of final translation
-quality in a real system. For this task, we use the widely-used {\sc
-  bleu} metric (see Section~\ref{sec:eval-metrics}). This is not ideal
-for evaluating the task of a lexical selection module as the
-performance of the module will depend greatly on (a) the coverage of
-the bilingual dictionaries of the RBMT system in question, and (b) the
-number of reference translations. It is also worth noting that
-successful lexical selections may not lead to successful translations
-due to inadequate transfer of morphological features. 
+This is an extrinsic evaluation, which ideally would test how much the
+system improves as regards an approximate measurement of final
+translation quality in a real system. For this task, we use the
+widely-used {\sc bleu} metric (see
+Section~\ref{sec:eval-metrics}). This is not ideal for evaluating the
+task of a lexical selection module as the performance of the module
+will depend greatly on (a) the coverage of the bilingual dictionaries
+of the RBMT system in question, and (b) the number of reference
+translations. It is also worth noting that successful lexical
+selections may not lead to successful translations due to inadequate
+transfer of morphological features.
 % For example, if
@@
   part will be used.%\todo{Not in OPUS?}
-\item \textbf{Open Data Euskadi} (\texttt{OpenData}): This is a Basque
-  and Spanish parallel corpus made from the translation memories of
-  the \emph{Herri Arduralaritzaren Euskal Erakundea} `Basque Institute
-  of Public
+\item \textbf{Open Data Euskadi} (\texttt{OpenData}): This is a
+  Basque--Spanish parallel corpus made from the translation
+  memories of the \emph{Herri Arduralaritzaren Euskal Erakundea}
+  `Basque Institute of Public
   Administration'.\footnote{\url{http://tinyurl.com/eu-es-tm}}
@@
   After this process, we selected 1,000 sentence pairs at random for
-testing (\texttt{test}), 1,000 for development (\texttt{dev})\footnote{The development corpus  
-  was used for checking the value for frequency pruning of features.} and 
-left the remainder for training. 
+  testing (\texttt{test}), 1,000 for development
+  (\texttt{dev})\footnote{The development corpus was used for checking
+    the value for frequency pruning of features.} and left the
+  remainder for training.
 %This is a smaller
@@
 %only little over 4,000 sentences long. 
-Table~\ref{table:input-corp}
-gives statistics about the size of the input corpora, and how many
-sentences were left after processing for testing, training and
-development. \todo{Why do ambiguity numbers change so much from the paper?}Table~\ref{table:test-corp} gives information about the
-test corpora.
+  Table~\ref{table:input-corp} gives statistics about the size of the
+  input corpora, and how many sentences were left after processing for
+  testing, training and development. \todo{Why do ambiguity numbers
+    change so much from the paper?} Table~\ref{table:test-corp} gives
+  information about the test corpora. \todo{FSM: One may wonder which
+    would be the results when the TL side of the training corpus is
+    used to train the models in the usual way.}
 
@@
   then score these translations \emph{on-line} on a model of the
-  target language. The highest scoring sentence is then 
-  output. This is the method used by %@@@ \cite{carbonell06} and
+  target language. The highest scoring sentence is then output. This
+  is the method used by %@@@ \cite{carbonell06} and
   \cite{melero07a}.% \todo{Also METIS?}
@@
 weights, we compute the evaluation measures described in
-section~\ref{ss:measures}. There is an option to remove features which
-occur less than a certain number of times in the training corpus. This
-is referred to as the feature pruning frequency threshold --- features
-occuring less than threshold are discarded. The value was set
+section~\ref{ss:measures}. There is an option to remove features
+(contexts)\todo{FSM: Features are contexts, aren't they?} which occur
+less than a certain number of times in the training corpus. This is
+referred to as the feature pruning frequency threshold --- features
+occuring less than the threshold are discarded. The value was set
 experimentally. Values of between two and seven were tested, and the
@@
 
-Improvements with
- respect to the target-language-model performance are likely due
- to the effective use the maximum-entropy model makes of information
- about the relevant source-language contexts and their translations,
- through the weighting of features representing those source-language
- contexts across the whole corpus.
+ Improvements with respect to the target-language-model performance
+ are likely due to the effective use that the maximum-entropy model
+ makes of information about the relevant source-language contexts and
+ their translations, through the weighting of features representing
+ those source-language contexts across the whole corpus.
 
@@
     \multirow{2}{*}{{\bf br-fr}} & {\sc ler} (\%)     & [54.8, 60.7] & [44.2, 50.5]  & {\bf [40.8, 46.9]} & [0.0, 0.0]      \\ 
-                                 & {\sc bleu} (\%)    & [14.5, 16.4] & [15.4, 17.3]  & {\bf [14.8, 16.6]} & [16.7, 18.6]     \\ 
+                                 & {\sc bleu} (\%)    & [14.5, 16.4] & [15.4, 17.3]  & [14.8, 16.6] & [16.7, 18.6]     \\ 
     \hline % tlm-alig: 
@@
 %! \comment{First try at a set of conclusions by Mikel} 
-This paper has
-presented a method to perform lexical selection in a rule-based
-machine translation system, and one that can be trained in an
-unsupervised way, that is, without the need for an annotated corpus,
-(in this case a word-aligned bilingual corpus): one just needs a
-source-language corpus, a statistical target-language model, and the
-rule-based machine translation system. The input to the method is
-simply the part-of-speech tagged source text in which each word is
-annotated with all the translations provided by the bilingual
-dictionary in the system: this makes it applicable to almost any
-rule-based machine translation system. The system uses a
+This paper has presented a method to perform lexical selection in
+RBMT, and one that can be trained in an unsupervised way, that is,
+without the need for an annotated corpus, (in this case a word-aligned
+bilingual corpus): one just needs a source-language corpus, a
+statistical target-language model, and the RBMT system itself. The
+input to the method is simply the part-of-speech tagged source text in
+which each word is annotated with all the translations provided by the
+bilingual dictionary in the system: this makes it applicable to almost
+any rule-based machine translation system. The system uses a
 maximum-entropy formalism for lexical selection, as \cite{berger1996}

### 2015-02-18 2015-eamt-assim/2015-eamt-assim.tex
Added the abstract, intro, methodology, setup and bibliography
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58988 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{multirow}
+\usepackage{natbib}
 
@@
 
-\title{A gap-filling approach to evaluation of machine translation for assimilation}
+\title{Evaluating machine translation for assimilation}
 
-\author{First Author\\
-  Affiliation / Address line 1\\
-  Affiliation / Address line 2\\
-  Affiliation / Address line 3\\
-  {\tt email@domain}  \And
-  Second Author\\
-  Affiliation / Address line 1\\
-  Affiliation / Address line 2\\
-  Affiliation / Address line 3\\
-  {\tt email@domain}}
+\author{Ekaterina Ageeva\\
+  School of Linguistics\\
+  Higher School of Economics\\
+  Moscow, Russia\\
+  {\tt evageeva\_2@edu.hse.ru}\\[2ex]
+  \textbf{Francis M. Tyers}\\
+  HSL-fakultetet\\ 
+  UiT Norgga\'{a}rktala\v{s} universitehta \\
+  9017 Romsa, Norway \\
+  {\tt francis.tyers@uit.no}
+  \And
+  Mikel L. Forcada\\
+  Dept. Llenguatges i Sistemes Inform\`{a}tics\\
+  Universitat d'Alacant, Spain \\
+  {\tt mlf@dlsi.ua.es}  \\[2ex]
+  \textbf{Juan Antonio P\'{e}rez-Ortiz} \\
+  Dept. Llenguatges i Sistemes Inform\`{a}tics\\
+  Universitat d'Alacant, Spain \\
+  {\tt japerez@dlsi.ua.es}
+}
 
@@
 \maketitle
+%Katya: has not it first been used in trosterud12?
 \begin{abstract}
-   ... 
+This paper provides additional observations on the viability of a strategy proposed in 2012 for evaluation of machine translation for assimilation purposes.
+The evaluation method is applied to three translation directions.
+To reduce preparation time, an automatic open-source task management system is introduced.
 \end{abstract}
@@
 
-As suggested by \newcite{church93}, modern machine translation (MT) systems may
+As suggested by \citet{church93}, modern machine translation (MT) systems may
 be divided into two broad categories according to their purpose: post-editing systems, e.g.
-Google Translate 1 , and assimilation systems, e.g. Apertium 2 . Both kinds may be
-systematically evaluated, either to control for quality in the development process or to
-compare the systems. The evaluation methods, however, differ depending on the type of
-the system. The methods are described in detail in \newcite{church93}.
-In this work I perform the evaluation of an assimilation MT system called Apertium.
-Apertium is an open-source community-maintained project that provides translation for
-39 language pairs. The goal of this study is to assess whether machine translation
-performed by this system assists users in understanding the gist of texts in an unknown
-language. The evaluation is performed in three separate experiments for the following
-translation directions: Basque to Spanish, English to Kazakh, and Tatar to Russian.
-I build on the methodology established in \newcite{trosterud12} and
-\newcite{oregan2013}, which are the works describing previous evaluations of
-Apertium language pairs. \newcite{oregan2013} proposes the gap-filling task as a
-means of assessing participant’s understanding of text. The same methodology is used in
-this work and described in detail in the section below.
+Google Translate\footnote{translate.google.com}, and assimilation systems, e.g.
+Apertium\footnote{apertium.org}.
+Both kinds may be systematically evaluated, either to control for quality in the development process or to compare the systems.
+Importantly, according to \citet{church93}, the evaluation methods must closely consider the system's primary purpose.
 
-\section{Experiment design}
+%\Katya: I'd like a citation for the "most massive application", could you suggest something?
+Despite the fact that assimilation (or gisting) is currently the most massive application of MT, few methodologies are established for assimilation evaluation of MT.
+The methods include post-editing and comparison by bilingual experts \citep{ginesti09}, and multiple choice tests \citep{jones07}, \citep{trosterud12} These approaches are often costly and prone to subjectivity, see the discussion in \citep{oregan13}.
+As an alternative, the modification of Cloze test was introduced for assimilation evaluation, first in \citep{trosterud12} as a supplementary, and then in \citep{oregan13} as a stand-alone method.
+Prior to this, Cloze tests have been used to evaluate raw MT quality, see e.g.
+\citep{vanslype79}, \citep{somers00}.
+As a principal difference, \citet{trosterud12} and \citet{oregan13} propose to fill the gaps in the reference (human) translation.
 
-\subsection{Methodology}
+The gap-filling method has been successfully used to evaluate the Basque-English Apertium language pair.
+In this work we extend the evaluation to three more language pairs: Basque-Spanish, English-Kazakh and Tatar-Russian.
+For the latter two pairs, the evaluation served as a quality check in the period of active development during the Google Summer of Code'14 program.
+In addition to evaluating, we explore new aspects of the experiment: the correlation between evaluators' scores, and the influence of the texts' linguistic domain.
+To facilitate the evaluation, we introduce an automated task creation system.
+This system is integrated into the Appraise MT evaluation platform \citep{federmann12}; the code is open source and is available on Github\footnote{https://github.com/Sereni/Appraise}.
 
-This experiment is based on an idea that understanding of text may be assessed through
-gap-filling questions. If the participant is able to restore principal information that had
-been left out from the text, they are assumed to have understood it.
-During the course of the experiment, the participants are offered sentences in their native
-language (the target language of the pair), with some key words left out. Each sentence
-has a counterpart in the source language of the pair. I use four different task settings:
+We anticipate that MT systems under investigation will contribute to the users' understanding of text.
+We also expect to see different results depending on text domain and on the number of words left out from the task.
+
+The paper is organized as follows: in section~\ref{sec:methodology} we describe the gap-filling method for assimilation evaluation: the task layout, the choice of words, and how the tasks are generated.
+Section~\ref{sec:setup} introduces the experimental material, the evaluators, the distribution of tasks and the evaluation procedure.
+In section~\ref{sec:results} we describe and discuss the experiment results.
+Finally, section~\ref{sec:conclusion} draws conclusions.
+
+\section{Methodology}
+\label{sec:methodology}
+
+This section discusses the reasoning behind the gap-filling method and task structure.
+The gap-filling method of evaluating machine translation for assimilation purposes is based on the following notion: readers' understanding of a text correlates with the number of keywords they are able to correctly restore in this text.
+Therefore, the base of an assimilation task is a (reference) sentence, where some of the keywords are blacked out, or removed.
+The sentence is produced by a human (as opposed to machine-translated), and it is in the language known to evaluators, which is also the ~\em{target language }\em  of the evaluated pair.
+
+%katya: I'm not sure this is intended behavior: unless I "close" the em tag, it keeps making everything after it italic. I thought brackets should limit this...
+The additional elements of the task are what we call hints, or extra sentences that help the participant to understand the main sentence.
+There are two types of hints: first, the \em{source}\em, which is the same sentence as \em{reference}\em, also human-produced, but in the source language of the pair.
+The second type is the \em{machine-translated }\em hint, which comes from the machine translation of the \em{source }\em sentence.
+
+%katya: I don't really have a good explanation why we used source+mt tasks, except for being exhaustive. could you suggest anything?
+In the course of the experiment, the following hint combinations are offered:
 \begin{itemize}
-\item  Target sentence only. The participants are asked to fill the gaps not being given
+\item  Reference sentence only. The participants are asked to fill the gaps not being given
 any context. This task serves as a baseline score and as an indicator of phrases
@@
 idioms and strong collocations);
-\item  Target sentence and source sentence. In addition to the question, the participants
-are offered the same sentence in the source language of the pair. By setup, the
-participants have no command of this language, however, it may help them to fill
+\item  Reference sentence and source sentence. By setup, the
+participants have no command of the source language, however, it may help them to fill
 in proper nouns or loan words;
-\item  Target sentence and machine-translated hint. In addition to the task, the
+\item  Reference sentence and machine-translated hint. In addition to the task, the
 participants see the source sentence translated via Apertium. This type of task is
@@
 gist of text;
-\item  Target sentence and both the source and machine-translated hint. This task is
-added mostly for completing the possible combination of task types, and to check
-for any unexpected insights, should they arise.
+\item  Reference sentence and both the source and machine-translated hint. This task is
+added for completing the possible combination of task types, and to check
+for unexpected insights, should they arise.
 \end{itemize}
 
-\subsection{Material}
+In order to prepare the evaluation questions, we determine and remove the keywords in the reference sentences.
+We consider two parameters: the list of allowed parts of speech (POS), and the number of gaps relative to sentence length ("gap density").
+For the evaluations described in this paper we use gap densities of 10, 20 and 30 percent, and the following parts of speech: noun, proper noun, adjective, adverb and lexical verb (as opposed to auxiliary verb).
 
-The data are collected separately for each experiment. A set of sentences is drawn from
-parallel corpora. The source sentences are translated into target language via Apertium.
-The target sentences are processed with a special script that removes significant words.
-For the current experiments, the word is removed if it falls into a list of allowed parts of
-speech:
-\begin{itemize}
-\item nouns;
-\item proper nouns;
-\item adjectives;
-\item adverbs;
-\item lexical verbs.
-\end{itemize}
-Parts of speech are determined automatically with Apertium morphological analyser. To
-rule out homonymy, the set of all possible POS attributions of a given word is checked
-against the POS list. If all of the possible word analyses match the parts of speech listed
-above, it is considered a candidate for removal.
-Each sentence in the set is evaluated in three “gap densities” (ratio of gaps to words in the
-sentence): 10\%, 20\% and 30\%. After the script finds all possible keyword candidates in
-the sentence, it removes the specified percentage of words, evenly distributing the gaps
-across the sentence.
-To exclude headlines and also to allow at least one gap in 10\%-gap sentences, only the
-sentences over 10 words long are used in the tasks.
-The sentences for each experiment are drawn from various parallel text sources:
+For each sentence, the list of candidate keywords is prepared.
+It constitutes all the words that fall into the allowed POS list.
+The number of gaps in the sentence is calculated based on sentence length and specified gap density.
+Finally, the required number of keywords is selected from the candidate list in such a manner that the gaps are distributed evenly throughout the sentence.
+
+Keyword removal is one of the most time-consuming steps in task preparation.
+In our setup, the above procedure is performed by a script integrated into the task generation pipeline.
+Parts of speech are determined with Apertium morphological analyzer.
+To control for homonymy, we only allow the word into the candidate list if all of its possible part of speech attributions are on the POS list.
+
+Having prepared the sentence sets, we assemble them into XML formatted for the Appraise platform.
+
+%To exclude headlines and also to allow at least one gap in 10\%-gap sentences, only the
+%sentences over 10 words long are used in the tasks.
+%The sentences for each experiment are drawn from various parallel text sources:
+%\begin{enumerate}
+%\item  For Basque-Spanish, from the corpus of legal texts Memorias de traducci\'on del
+%Servicio Oficial de Traductores del IVAP 3
+%\item  For English-Kazakh, from the official website of the President of the Republic of
+%Kazakhstan 4
+%\item  For Tatar-Russian, from the following sources on three different topics:
+%  \begin{enumerate}
+%    \item  Casual conversations, from a textbook 5 of spoken Tatar;
+%    \item  Legal texts, from the Constitution and laws 6 of Tatarstan;
+%    \item  News, from the President of Tatarstan website 7 .
+%  \end{enumerate}
+%\end{enumerate}
+%
+%Each set features 36 pairs of sentences. For the first two experiments the pairs are drawn
+%randomly from the corpora, for Tatar-Russian, compiled by hand by the developer of the
+%language pair. After processing, the material is transformed into XML used for uploading
+%into the evaluation system (see section 2.4).
+
+\section{Experimental set-up}
+\label{sec:setup}
+
+In this section we will discuss the evaluators, the evaluation procedure, and the tasks in more detail.
+
+For each experiment we called for native speakers of L2 of the language pair (i.e.
+Spanish, Kazakh, and Russian) who had no command of L1 of the pair (Basque, English
+and Tatar, respectively). 11 evaluators participated in the in Basque-Spanish experiment, 8 in
+English-Kazakh, and 28 in Tatar-Russian (although not everyone has completed their task
+in full, see discussion).
+
+%katya: maybe I should put this into the previous section?
+By design, our gap-filling tasks require a human translation (reference) of source sentences.
+Calling for a human translator, however, would significantly increase the resources needed for evaluation.
+We therefore use parallel text sources, which provide the same sentence in two languages simultaneously:
 \begin{enumerate}
-\item  For Basque-Spanish, from the corpus of legal texts Memorias de traducción del
-Servicio Oficial de Traductores del IVAP 3
+\item  For Basque-Spanish, from the corpus of legal texts Memorias de traducci\'on del
+Servicio Oficial de Traductores del IVAP\footnote{http://opendata.euskadi.net/memorias-de-traduccion-del-servicio-oficial-de-traductores-del-ivap/w79-contdata/es/};
 \item  For English-Kazakh, from the official website of the President of the Republic of
-Kazakhstan 4
+Kazakhstan\footnote{http://www.akorda.kz/};
 \item  For Tatar-Russian, from the following sources on three different topics:
   \begin{enumerate}
-    \item  Casual conversations, from a textbook 5 of spoken Tatar;
-    \item  Legal texts, from the Constitution and laws 6 of Tatarstan;
-    \item  News, from the President of Tatarstan website 7 .
+  %whoops, what is the best way to make it understand cyrillic?
+    \item  Casual conversations, from a textbook\footnote{????????
+?.?.
+?
+???????
+????????
+??-????????.
+??????: ?????????
+??.
+???-??, 1994.
+?
+320 ?.
+ISBN 5?298?00463?6 (???.
+219, 220, 232, 233, 234)} of spoken Tatar;
+    \item  Legal texts, from the Constitution and laws\footnote{http://tatarstan.ru} of Tatarstan;
+    \item  News, from the President of Tatarstan website\footnote{http://president.tatarstan.ru/}.
   \end{enumerate}
@@
 randomly from the corpora, for Tatar-Russian, compiled by hand by the developer of the
-language pair. After processing, the material is transformed into XML used for uploading
-into the evaluation system (see section 2.4).
-
-\subsection{Participants}
-
-For each experiment we called for native speakers of L2 of the language pair (i.e.
-Spanish, Kazakh, and Russian) who had no command of L1 of the pair (Basque, English
-and Tatar, respectively). We had 11 participants in Basque-Spanish evaluation, 8 in
-English-Kazakh, and 28 in Tatar-Russian (although not everyone has completed their task
-in full, see discussion).
+language pair.
 
@@
 
-The evaluations take place online, in a system 8 designed specifically for various MT
+The evaluations take place online, in a system called Appraise \citep{federmann12}, which is designed specifically for various MT
 evaluation tasks. We adapted the code to accommodate for the gap-filling tasks. The
@@
 a total of 36;
-\item  Each participant evaluates 9 sentences in each of the four modes (section 2.1);
+\item  Each participant evaluates 9 sentences in each of the four modes (see section~\ref{sec:methodology};
 \item  All sentences of the set are evaluated with 10, 20 and 30\% of words removed;
-\item  Each sentence-mode-percentage combination is evaluated by at least three
+\item  Each sentence-mode-percentage combination is evaluated by at least three participants.
 \end{enumerate}
 The participants are given the instructions in their native language above each task. The
-tasks are split into smaller sets for the participants’ convenience. The instructions are to
+tasks are split into smaller sets for the participants'۪ convenience. The instructions are the following: read all the available hints and fill each gap with one suitable word, guessing if unsure.
+Participants'۪ answers are recorded and marked correct or incorrect automatically. In
+addition, the time taken to fill the gaps in one sentence is recorded.
 
-read all the available hints and fill each gap with one suitable word, guessing if unsure.
-Participants’ answers are recorded and marked correct or incorrect automatically. In
-addition, the time taken to fill the gaps in one sentence is recorded.
-Since gap-filling task requires open answers, it is possible that the participants may
+This variety of the gap-filling task requires open answers, and it is therefore possible that the participants may
 provide words that fit the gaps well, but do not match the original answer. To account for
-these cases, we process all the answers to detect possible synonyms. An answer is
-considered a candidate synonym if given by two or more evaluators and does not match
+these cases, we process all the answers to detect possible synonyms (a method suggested by \citet{oregan13}. An answer is
+considered a candidate synonym if it is given by two or more evaluators, and it does not match
 the answer key. We record each candidate synonym along with the answer key and the
@@
 whether the candidate synonym is indeed synonymic to the answer key in the given
-context. We then check participants’ results against the compiled synonym list and
-increase scores where appropriate.
+context. We then check participants'۪ results against the compiled synonym list and
+increase scores where appropriate. Candidate synonym extraction and score update is performed automatically.
+
+% the sections below are from the report and have not been edited yet.
 
@@
 unusually high. It appeared that one of them is a news headline, and the other
-contains clichés such as “express condolences”. There may also be interference
+contains clich\'es such as "express condolences". There may also be interference
 from other tasks of the set, because some of the sentences are drawn from the

### 2015-02-18 2015-eamt-melsa/2015-eamt-melsa.tex
more
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58990 72bbbca6-d526-0410-a7d9-f06f51895060

@@
   system in which the models are integrated and a statistical
-  target-language model. Using the MT system, \todo{FSM: Check
-    rewording! I have removed existing MT from this and previous
-    sentence.} the sentences in the source-language corpus are
+  target-language model. Using the MT system, the sentences in 
+  the source-language corpus are
   translated in all possible ways according to the different

### 2015-02-18 2015-eamt-assim/2015-eamt-assim.tex
some updates
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58994 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{natbib}
+\usepackage{color}
 
@@
 \newcommand{\shortpaperlength}{$4$ (four)}
+
+\newcommand{\comment}[1]{\marginpar{\scriptsize\sf \textcolor{blue}{#1}}}
 
@@
 Importantly, according to \citet{church93}, the evaluation methods must closely consider the system's primary purpose.
+\comment{FMT: Apertium can also be used for postedition, e.g. spanish-catalan, and some Google translate directions aren't suitable for that. Perhaps reword or remove this}
 
-%\Katya: I'd like a citation for the "most massive application", could you suggest something?
+\comment{EA: I'd like a citation for the ``most massive application'', could you suggest something?}
 Despite the fact that assimilation (or gisting) is currently the most massive application of MT, few methodologies are established for assimilation evaluation of MT.
@@
 In this work we extend the evaluation to three more language pairs: Basque-Spanish, English-Kazakh and Tatar-Russian.
-For the latter two pairs, the evaluation served as a quality check in the period of active development during the Google Summer of Code'14 program.
+For the latter two pairs, the evaluation served as a quality check in the period of active development during the Google Summer of Code'14 programme.
 In addition to evaluating, we explore new aspects of the experiment: the correlation between evaluators' scores, and the influence of the texts' linguistic domain.
@@
 Therefore, the base of an assimilation task is a (reference) sentence, where some of the keywords are blacked out, or removed.
-The sentence is produced by a human (as opposed to machine-translated), and it is in the language known to evaluators, which is also the ~\em{target language }\em  of the evaluated pair.
+The sentence is produced by a human (as opposed to machine-translated), and it is in the language known to evaluators, which is also the ~\emph{target language }  of the evaluated pair.
 
-%katya: I'm not sure this is intended behavior: unless I "close" the em tag, it keeps making everything after it italic. I thought brackets should limit this...
 The additional elements of the task are what we call hints, or extra sentences that help the participant to understand the main sentence.
-There are two types of hints: first, the \em{source}\em, which is the same sentence as \em{reference}\em, also human-produced, but in the source language of the pair.
-The second type is the \em{machine-translated }\em hint, which comes from the machine translation of the \em{source }\em sentence.
+There are two types of hints: first, the \emph{source}, which is the same sentence as \emph{reference}, also human-produced, but in the source language of the pair.
+The second type is the \emph{machine-translated } hint, which comes from the machine translation of the \emph{source } sentence.
 
-%katya: I don't really have a good explanation why we used source+mt tasks, except for being exhaustive. could you suggest anything?
+\comment{EA: I don't really have a good explanation why we used source+mt tasks, except for being exhaustive. could you suggest anything?}
 In the course of the experiment, the following hint combinations are offered:
@@
 In order to prepare the evaluation questions, we determine and remove the keywords in the reference sentences.
-We consider two parameters: the list of allowed parts of speech (POS), and the number of gaps relative to sentence length ("gap density").
+We consider two parameters: the list of allowed parts of speech (POS), and the number of gaps relative to sentence length (``gap density'').
 For the evaluations described in this paper we use gap densities of 10, 20 and 30 percent, and the following parts of speech: noun, proper noun, adjective, adverb and lexical verb (as opposed to auxiliary verb).
@@
 
-%katya: maybe I should put this into the previous section?
+\comment{EA: maybe I should put this into the previous section?}
 By design, our gap-filling tasks require a human translation (reference) of source sentences.
@@
   \begin{enumerate}
-  %whoops, what is the best way to make it understand cyrillic?
-    \item  Casual conversations, from a textbook\footnote{????????
-?.?.
-?
-???????
-????????
-??-????????.
-??????: ?????????
-??.
-???-??, 1994.
-?
-320 ?.
-ISBN 5?298?00463?6 (???.
-219, 220, 232, 233, 234)} of spoken Tatar;
+\comment{EA: whoops, what is the best way to make it understand cyrillic?}
+    \item  Casual conversations, from a textbook\footnote{FIXME, 1994. ? 320 ?. ISBN 5?298?00463?6 (???. 219, 220, 232, 233, 234)} of spoken Tatar;
     \item  Legal texts, from the Constitution and laws\footnote{http://tatarstan.ru} of Tatarstan;
@@
 The participants are given the instructions in their native language above each task. The
-tasks are split into smaller sets for the participants'۪ convenience. The instructions are the following: read all the available hints and fill each gap with one suitable word, guessing if unsure.
-Participants'۪ answers are recorded and marked correct or incorrect automatically. In
+tasks are split into smaller sets for the participants' convenience. The instructions are the following: read all the available hints and fill each gap with one suitable word, guessing if unsure.
+Participants' answers are recorded and marked correct or incorrect automatically. In
 addition, the time taken to fill the gaps in one sentence is recorded.
@@
 whether the candidate synonym is indeed synonymic to the answer key in the given
-context. We then check participants'۪ results against the compiled synonym list and
+context. We then check participants' results against the compiled synonym list and
 increase scores where appropriate. Candidate synonym extraction and score update is performed automatically.
+\comment{FMT: How long did making the synonym list take, how many synonyms were added? More details here.}
 
-% the sections below are from the report and have not been edited yet.
+\comment{EA: the sections below are from the report and have not been edited yet.}
 
@@
 The experiment results are presented in the tables below. The first table of each set shows
-the proportion of correct answers depending on evaluation mode and gap density. The
+the proportion of correct answers depending on evaluation mode and gap density.\footnote{FMT: We should probably integrate the other tables, and refer to them using labels.} The
 percentage is spread across all evaluators, that is, we divide the number of correct
@@
 unusually high. It appeared that one of them is a news headline, and the other
-contains clich\'es such as "express condolences". There may also be interference
+contains clich\'es such as ``express condolences''. There may also be interference
 from other tasks of the set, because some of the sentences are drawn from the

### 2015-02-18 2015-eamt-melsa/2015-eamt-melsa.tex
some removal of comments
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58991 72bbbca6-d526-0410-a7d9-f06f51895060
more fixes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58996 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \cite{marechek10} trained a maximum-entropy lexical selector for their
-TectoMT rule-based system using a bilingual corpus.\todo{Say a bit
-  more about TectoMT? FSM: Yes, please!}  More recently,
-\cite{tyers12a} presented a method of lexical selection for rule-based
+dependency-grammar-based transfer system TectoMT using a bilingual corpus.
+More recently, \cite{tyers12a} presented a method of lexical selection for rule-based
 machine translation based on rules which select or remove translations
@@
 (1) how should the relevant contexts be chosen? and (2) what should be
-done when the translations of words are not found in the corpus?
-\todo{Wording? When $(s,t,c)$ not found in the corpus?}  A
-maximum-entropy model answers both of these questions. It allows the
+done when $(s, t, c)$ is not found in the corpus?
+A maximum-entropy model answers both of these questions. It allows the
 contexts that we consider to be linguistically interesting to be
@@
 and the components from the RBMT system in which they are
-used. \todo{FSM: Check rewording!}
+used.
 
@@
 T_s(s_i)\). % and \(T_s(\cdot)\)
-\todo{Check change \(T(\cdot)\to T_s(\cdot)\) for consistency with
-  later notation. FSM: Check \(t_i \in T_s(s_i)\); before it was \(t_i
-  = T_s(s_i)\)} %is the set of all
+%is the set of all
 %translations for a source word. 
@@
 \(c\). That is, as if event $(s,t,c)$ had been seen a fractional
-number $p(g_i|S)$ of times. We prune features (contexts)
-\todo{features or contexts?} occurring less than a certain number of
+number $p(g_i|S)$ of times. We prune $(s, t, c)$
+occurring less than a certain number of
 times in the corpus, using a development corpus to guide pruning (see
@@
 analysis, morphological disambiguation, lexical transfer, lexical
-selection, structural transfer and morphological generation. The
-relevant modules of the system for this method are the \emph{lexical
-  transfer} module which provides for each source language lexical
-form a set of possible target language lexical forms and the
-\emph{structural transfer} and \emph{morphological generation} modules
-which for each pair of source language lexical form and translation
-produce a final translation in the target language. \todo{The wording
-  of this last sentence is unclear. Identify pre-lexsel, lexsel and
-  post-lexsel as before}
+selection, structural transfer and morphological generation. 
+%The relevant modules of the system for this method are the \emph{lexical
+%  transfer} module which provides for each source language lexical
+%form a set of possible target language lexical forms and the
+%\emph{structural transfer} and \emph{morphological generation} modules
+%which for each pair of source language lexical form and translation
+%produce a final translation in the target language. \todo{The wording
+%  of this last sentence is unclear. Identify pre-lexsel, lexsel and
+%  post-lexsel as before}
 
@@
   a combination of the English--Catalan and Spanish--Catalan pairs
-  over a period of around 3--5 months.\todo{FSM: 3, 4 or 5?} The pair
+  over a period of around four months. The pair
 %has been used in other
@@
   contains a number of entries in the bilingual dictionary with more
-  than one translation. The most-frequent translation, as judged by
-  the language pair developer, is marked as the \emph{default}
-  translation and non-default translations were added according to
-  frequency. \todo{Default is explained for some LPs only, why?}  The
+  than one translation. %The most-frequent translation, as judged by
+%  the language pair developer, is marked as the \emph{default}
+%  translation and non-default translations were added according to
+%  frequency. \todo{Default is explained for some LPs only, why?}  
+  The
   version used in this paper is SVN revision
@@
 (intrinsic) \emph{lexical selection performance} measure and an
-(extrinsic) \emph{machine translation performance} measure. \todo{FSM:
-I moved some stuff from here to the LER subsection.}
+(extrinsic) \emph{machine translation performance} measure.
 
@@
 
+  %% FMT: INCLUDE AVERAGE AMBIGUITY OVER AMBIGUOUS WORDS
   \begin{table}
@@
     change so much from the paper?} Table~\ref{table:test-corp} gives
-  information about the test corpora. \todo{FSM: One may wonder which
-    would be the results when the TL side of the training corpus is
-    used to train the models in the usual way.}
+  information about the test corpora. 
+%! \todo{FSM: One may wonder which
+%!    would be the results when the TL side of the training corpus is
+%!    used to train the models in the usual way.}
 
@@
 weights, we compute the evaluation measures described in
-section~\ref{ss:measures}. There is an option to remove features
-(contexts)\todo{FSM: Features are contexts, aren't they?} which occur
+section~\ref{ss:measures}. There is an option to remove events 
+$(s, t, c)$ which occur
 less than a certain number of times in the training corpus. This is
@@
 %Research Grant) coordinated by the Universidad Complutense de Madrid.
-%
+
 \bibliographystyle{apalike}

### 2015-02-18 2015-eamt-melsa/2015-eamt-melsa.tex
fixes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58997 72bbbca6-d526-0410-a7d9-f06f51895060

@@
  pairs. In the remaining cases, the maximum-entropy method comes close
- to the target-language model performance, at a much smaller
- computational cost.\todo{I miss a brief separate discussion of BLEU and LER, which are reported separately!}
+ to the target-language model performance in terms of similar or better BLEU and LER scores, at a much smaller
+ computational cost.
+% \todo{I miss a brief separate discussion of BLEU and LER, which are reported separately!}
 

### 2015-02-18 2015-nodalida-ruscg/ruscg.tex
stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58852 72bbbca6-d526-0410-a7d9-f06f51895060
more stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@58998 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{url}
+\usepackage[small,bf]{caption}
 \special{papersize=210mm,297mm} % to avoid having to use "-t a4" with dvips 
@@
 
+% objectives:
+%% make a free/open-source morphology for russian based on Z's dictionary that includes stress
+
 \section{Review of literature}
@@
 
+\cite{yablonsky1999russian}
+
 \subsection{Disambiguation rules}
@@
 \begin{table}
+  \centering
   \begin{tabular}{lrrr}
@@
          & {\sc select} & {\sc remove} & {\sc map} \\
+    Safe & & & \\
+    Safe heuristic  & & & \\
+    Heuristic &  & & \\
+    Syntax labelling &  &   & \\ 
     \hline
   \end{tabular}
+  \caption{Rules in the grammar are separated into four sections. }
 \end{table}
@@
 
+\begin{table*}
+  \centering
+  \begin{tabular}{|l|r|r|r|r|}
+    \hline
+    \textbf{Domain} & \textbf{Tokens} & \textbf{Precision} & \textbf{Recall} & \textbf{Ambig. solved} \\
+    \hline
+    Wikipedia       & 7,857      & 0.506        & 0.996    & 44.92\%  \\
+    News            & 642        & 0.471        & 0.990    & 41.60\%  \\
+    Literature      & 1,652      & 0.473        & 0.984    & 42.95\%  \\
+    \hline
+    Total           & 10,150     & 0.498        &  0.994   &  44.39\% \\
+    \hline
+  \end{tabular}
+  \caption{Results for the test corpora.}
+\end{table*}
+
 \begin{table}
+  \centering
   \begin{tabular}{|l|r|r|}
+
     \hline
-    Missing lexeme & &  \\          % безбалластный, FALSENEG: Кий<np><ant><m><sg><ins> ['кий<n><m><nn><sg><ins>']
-    Partial lexeme & &  \\          % франко-норманнское
+    Error in original        &   & 2 \\
+    Missing lexeme           &   & 8  \\          % безбалластный, FALSENEG: Кий<np><ant><m><sg><ins> ['кий<n><m><nn><sg><ins>']
+    Partial lexeme           &   & 2  \\          % франко-норманнское, рельс
     Missing analysis         &   &    \\        % 
     Equally valid analysis   &   &    \\   % большой<adj><comp><pred> ['больше<adv>'] (?)
-    Correct reading removed & &  \\ % 
+    Correct reading removed  &   & 34   \\ % 
+    \hline
+                             &   & 46 \\
     \hline
   \end{tabular}
-
+  \caption{Error analysis of false negatives}
 \end{table}

### 2015-02-18 2015-eamt-kazkir/2015-eamt-kazkir.tex
stuffs
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59003 72bbbca6-d526-0410-a7d9-f06f51895060

%
% File eamt15.tex, an adaptation of eamt14.tex (which was a copy of
% eamt12.tex)
%
% Contact: eamt2015@dlsi.ua.es

%%% To ease future customizations, various replaceables have been paramaterized
%%% as listed in the newcommands section

\documentclass[11pt]{article}
\usepackage{eamt15}
\usepackage{times}
\usepackage{latexsym}
\setlength\titlebox{6.5cm}    % Expanding the titlebox
%%% YOUR PACKAGES BELOW THIS LINE %%%

\usepackage{multirow}
\usepackage{natbib}
\usepackage{color}
\usepackage{url}

\newcommand{\confname}{EAMT 2015}
\newcommand{\website}{http://www.eamt2015.org/}
\newcommand{\contactname}{the conference chairs (Felipe
  S\'anchez-Mart\'inez, Gema Ram\'irez-S\'anchez and Fred Hollowood)}
\newcommand{\contactemail}{eamt2015@dlsi.ua.es}
\newcommand{\conffilename}{eamt15}
\newcommand{\downloadsite}{http://www.eamt2015.org/}
\newcommand{\paperlength}{$8$ (eight)}
\newcommand{\shortpaperlength}{$4$ (four)}

\newcommand{\comment}[1]{\marginpar{\scriptsize\sf \textcolor{blue}{#1}}}

\title{A comparison of approaches to Kazakh to Kyrgyz machine translation}

\author{
  Jonathan North Washington\\
  Departments of Linguistics\\
  and Central Eurasian Studies\\
  Indiana University\\
  Bloomington, Indiana 47405 USA\\
  \texttt{jonwashi@indiana.edu}  \And
  \textbf{Francis M. Tyers}\\
  HSL-fakultetet\\ 
  UiT Norgga\'{a}rktala\v{s} universitehta \\
  9017 Romsa, Norway \\
  {\tt francis.tyers@uit.no}
}

\date{}

\begin{document}
\maketitle

\begin{abstract}
...
\end{abstract}

\section{Introduction}

\section{Previous work}
\label{sec:prev}

Within the Apertium project, work on several MT systems between Turkic languages has been started (Turkish--Kyrgyz, Azeri--Turkish, Tatar--Bashkir), but the Kazakh--Tatar system described by the present study is the closest to production-ready of them.  Among these systems is a prototype Tatar--Bashkir machine translation system \citep{tyerswashingtonsalimzyanbattalov12}; due to the closeness of these languages, it proved to provide high accuracy in its translations, but being a prototype system by design, had relatively low coverage.

Besides these systems, several previous works on making machine translation systems between Turkic languages 
exist, although to our knowledge none are publicly available except for the Turkish--Azerbaijani pair available through Google Translate.\footnote{\url{http://translate.google.com}}
Some MT systems have been reported that translate between Turkish and other Turkic languages, 
including Turkish--Crimean Tatar \citep{altintas01},
Turkish--Azerbaijani \citep{hamzaoglu93}, Turkish--Tatar \citep{suleymanov08}, and
Turkish--Turkmen \citep{tantug07}, though none of these have been released to a public audience.

\bibliographystyle{apalike}
\bibliography{2015-eamt-kazkir}

\end{document}

### 2015-02-18 2015-eamt-kazkir/2015-eamt-kazkir.tex
fixes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59005 72bbbca6-d526-0410-a7d9-f06f51895060

@@
   HSL-fakultetet\\ 
-  UiT Norgga\'{a}rktala\v{s} universitehta \\
+  UiT Norgga \'{a}rktala\v{s} universitehta \\
   9017 Romsa, Norway \\
@@
 
-Within the Apertium project, work on several MT systems between Turkic languages has been started (Turkish--Kyrgyz, Azeri--Turkish, Tatar--Bashkir), but the Kazakh--Tatar system described by the present study is the closest to production-ready of them.
-Among these systems is a prototype Tatar--Bashkir machine translation system \citep{tyerswashingtonsalimzyanbattalov12}; due to the closeness of these languages, it proved to provide high accuracy in its translations, but being a prototype system by design, had relatively low coverage.
+Within the Apertium project, work on several MT systems between Turkic languages has been started (Kazakh--Tatar, Tatar--Bashkir, Turkish--Kyrgyz, and Azeri--Turkish).
+There are two published works describing these systems, the first is \cite{tyerswashingtonsalimzyanbattalov12} which describes a Tatar--Bashkir machine translation system; due to the closeness of these languages, it proved to provide high accuracy in its translations, but being a prototype system by design, had relatively low coverage.
+The second is \cite{tyerswashingtonsalimzyan13} which describes a high-coverage Kazakh--Tatar system.
 
@@
 
+\section*{Acknowledgements}
+
+\emph{Anonymised for review}
+
 \bibliographystyle{apalike}

### 2015-02-18 2015-eamt-kazkir/2015-eamt-kazkir.tex
hargle
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59006 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
+This paper presents a comparison of approaches to machine translation
+from Kazakh to Kyrgyz.
+The paper will be laid out as follows: Section~\ref{sec:prev}\ gives a short review of some previous work in the area of Turkic--Turkic language machine translation; Section~\ref{sec:lang} introduces Kazakh and Kyrgyz and compares their grammar;
+Section~\ref{sec:sys}\ describes the systems we built and the tools used to construct them;
+Section~\ref{sec:eval}\ gives an evaluation of the systems; and
+finally Section~\ref{sec:conc}\ describes our aims for future work and some
+concluding remarks.
+
 \section{Previous work}

### 2015-02-19 2015-eamt-assim/2015-eamt-assim.tex
more comments, some format fixes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59015 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 To reduce preparation time, an automatic open-source task management system is introduced.
+\comment{FMT: Abstract should mirror/summarise the article, e.g. include something about conclusions too}
 \end{abstract}
@@
 Despite the fact that assimilation (or gisting) is currently the most massive application of MT, few methodologies are established for assimilation evaluation of MT.
-The methods include post-editing and comparison by bilingual experts \citep{ginesti09}, and multiple choice tests \citep{jones07}, \citep{trosterud12} These approaches are often costly and prone to subjectivity, see the discussion in \citep{oregan13}.
+The methods include post-editing and comparison by bilingual experts \citep{ginesti09}, and multiple choice tests \citep{jones07,trosterud12} These approaches are often costly and prone to subjectivity, see the discussion in \cite{oregan13}.
 As an alternative, the modification of Cloze test was introduced for assimilation evaluation, first in \citep{trosterud12} as a supplementary, and then in \citep{oregan13} as a stand-alone method.
@@
 
-The paper is organized as follows: in section~\ref{sec:methodology} we describe the gap-filling method for assimilation evaluation: the task layout, the choice of words, and how the tasks are generated.
+The paper is organised as follows: in section~\ref{sec:methodology} we describe the gap-filling method for assimilation evaluation: the task layout, the choice of words, and how the tasks are generated.
 Section~\ref{sec:setup} introduces the experimental material, the evaluators, the distribution of tasks and the evaluation procedure.
@@
 \label{sec:methodology}
+\comment{FMT: This could benefit from an example i think, something like table1? }
 
-This section discusses the reasoning behind the gap-filling method and task structure.
-The gap-filling method of evaluating machine translation for assimilation purposes is based on the following notion: readers' understanding of a text correlates with the number of keywords they are able to correctly restore in this text.
-Therefore, the base of an assimilation task is a (reference) sentence, where some of the keywords are blacked out, or removed.
-The sentence is produced by a human (as opposed to machine-translated), and it is in the language known to evaluators, which is also the ~\emph{target language }  of the evaluated pair.
+\begin{table*}
+  \begin{tabular}{|l|l|}
+     \hline
+     \textbf{Ref}   & Ayudas económicas para el tratamiento de toxicomanías en comunidades terapéuticas no concertadas. \\
+     \textbf{Task}   & Ayudas económicas para el \{ \} de toxicomanías en comunidades terapéuticas no concertadas. \\
+     \textbf{Src} & Komunitate terapeutiko itundu gabeetan toxikomaniak tratatzeko diru-laguntzak ematea. \\
+     \textbf{MT}     & Comunidad terapéutico pactar gabeetan toxikomaniak las-ayudas de dinero para tratar dar. \\
+     \hline
+  \end{tabular}
+  \caption{An example of a set of sentences.} 
+  \label{table:example}
+\end{table*}
 
-The additional elements of the task are what we call hints, or extra sentences that help the participant to understand the main sentence.
-There are two types of hints: first, the \emph{source}, which is the same sentence as \emph{reference}, also human-produced, but in the source language of the pair.
-The second type is the \emph{machine-translated } hint, which comes from the machine translation of the \emph{source } sentence.
+This section discusses the reasoning behind the gap-filling method and task structure. The gap-filling method 
+of evaluating machine translation for assimilation purposes is based on the following notion: a reader's understanding 
+of a given text correlates with the number of keywords they are able to correctly restore in the text. Therefore, the 
+base of an assimilation task is a (reference) sentence, where some of the keywords are blacked out, or removed. The sentence 
+is produced by a human (as opposed to machine-translated), and it is in the language known to evaluators, which is also 
+the \emph{target language} of the machine translation system.
+
+The additional elements of the task are what we call hints, or extra sentences that help the participant to understand 
+the main sentence. There are two types of hints: first, the \emph{source}, which is the same sentence as \emph{reference}, 
+also human-produced, but in the source language of the pair. The second type is the \emph{machine-translated} hint, which 
+comes from the machine translation of the \emph{source } sentence.
 
@@
 In the course of the experiment, the following hint combinations are offered:
-\begin{itemize}
-\item  Reference sentence only. The participants are asked to fill the gaps not being given
+\begin{description}
+\item[Reference sentence only:] The participants are asked to fill the gaps not being given
 any context. This task serves as a baseline score and as an indicator of phrases
@@
 idioms and strong collocations);
-\item  Reference sentence and source sentence. By setup, the
+\item[Reference sentence and source sentence:] By setup, the
 participants have no command of the source language, however, it may help them to fill
 in proper nouns or loan words;
-\item  Reference sentence and machine-translated hint. In addition to the task, the
+\item[Reference sentence and MT hint:] In addition to the task, the
 participants see the source sentence translated via Apertium. This type of task is
@@
 gist of text;
-\item  Reference sentence and both the source and machine-translated hint. This task is
+\item[Reference sentence and both hints:] This task is
 added for completing the possible combination of task types, and to check
 for unexpected insights, should they arise.
-\end{itemize}
+\end{description}
 
-In order to prepare the evaluation questions, we determine and remove the keywords in the reference sentences.
-We consider two parameters: the list of allowed parts of speech (POS), and the number of gaps relative to sentence length (``gap density'').
-For the evaluations described in this paper we use gap densities of 10, 20 and 30 percent, and the following parts of speech: noun, proper noun, adjective, adverb and lexical verb (as opposed to auxiliary verb).
+In order to prepare the evaluation questions, we determine and remove the keywords in the 
+reference sentences. We consider two parameters: the list of allowed parts of speech (POS), and the 
+number of gaps relative to sentence length (``gap density''). For the evaluations described in 
+this paper we use gap densities of 10, 20 and 30 percent, and the following parts of speech: noun, 
+proper noun, adjective, adverb and lexical verb (as opposed to auxiliary verb).
 
-For each sentence, the list of candidate keywords is prepared.
-It constitutes all the words that fall into the allowed POS list.
-The number of gaps in the sentence is calculated based on sentence length and specified gap density.
-Finally, the required number of keywords is selected from the candidate list in such a manner that the gaps are distributed evenly throughout the sentence.
+For each sentence, the list of candidate keywords is prepared. It constitutes all the words that 
+fall into the allowed POS list. The number of gaps in the sentence is calculated based on 
+sentence length and specified gap density. Finally, the required number of keywords is selected 
+from the candidate list in such a manner that the gaps are distributed evenly throughout the sentence.
 
@@
 In our setup, the above procedure is performed by a script integrated into the task generation pipeline.
-Parts of speech are determined with Apertium morphological analyzer.
+Parts of speech are determined with Apertium morphological analyser.
 To control for homonymy, we only allow the word into the candidate list if all of its possible part of speech attributions are on the POS list.
@@
 
-The evaluations take place online, in a system called Appraise \citep{federmann12}, which is designed specifically for various MT
-evaluation tasks. We adapted the code to accommodate for the gap-filling tasks. The
-tasks are uploaded into the system and manually distributed between the participants by
-the following rules:
+The evaluations take place online, in a system called Appraise \citep{federmann12}, which is 
+designed specifically for various MT evaluation tasks. We adapted the code to 
+accommodate for the gap-filling tasks. The tasks are uploaded into the system and 
+manually distributed between the participants by the following rules:
 \begin{enumerate}
@@
 \end{enumerate}
+
 The participants are given the instructions in their native language above each task. The
-tasks are split into smaller sets for the participants' convenience. The instructions are the following: read all the available hints and fill each gap with one suitable word, guessing if unsure.
+tasks are split into smaller sets for the participants' convenience. The instructions are the 
+following: read all the available hints and fill each gap with one suitable word, guessing if unsure.
 Participants' answers are recorded and marked correct or incorrect automatically. In
@@
 increase scores where appropriate. Candidate synonym extraction and score update is performed automatically.
-\comment{FMT: How long did making the synonym list take, how many synonyms were added? More details here.}
+\comment{FMT: How long did making the synonym list take, how many synonyms were added? More details here. Also perhaps an example?}
 
@@
 \section{Results}
+\label{sec:results}
 
 The experiment results are presented in the tables below. The first table of each set shows
-the proportion of correct answers depending on evaluation mode and gap density.\footnote{FMT: We should probably integrate the other tables, and refer to them using labels.} The
+the proportion of correct answers depending on evaluation mode and gap density.\comment{FMT: We should probably integrate the other tables, and refer to them using labels.} The
 percentage is spread across all evaluators, that is, we divide the number of correct
@@
 
-\section{Discussion}
-
+\section{Conclusions}
+\label{sec:conclusion}
 The following trends may be derived from the data above:
@@
 scores for this experiment.
-Having analyzed the tasks and the responses for English-Kazakh, I have found the
+Having analysed the tasks and the responses for English-Kazakh, I have found the
 following possible reasons of the obtained results:
@@
 \section{Future work}
+\label{sec:futurework}
 The experiment may be repeated for any language pair (provided a parallel corpus) and

### 2015-02-19 2015-eamt-assim/2015-eamt-assim.tex
+1 comment
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59016 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 the \emph{target language} of the machine translation system.
-
+\comment{FMT: It would also be nice to have a screenshot here}
 The additional elements of the task are what we call hints, or extra sentences that help the participant to understand 

### 2015-02-19 2015-nodalida-ruscg/ruscg.tex
organise
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59019 72bbbca6-d526-0410-a7d9-f06f51895060
average is better
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59020 72bbbca6-d526-0410-a7d9-f06f51895060
more stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59021 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
+Our purposes: Processing stressed wordforms and open-source machine translation
+
 \cite{Karlsson-90}
@@
 % estonian
-%
+
+State-of-the-art morphological analysis in Russian is primarily based on finite-
+state technology \cite{Nozhov-03,Segalovich-03}, although some success has 
+also been found with machine-learning approaches \cite{sharoff08lrec-mocky}. 
+However, no existing Russian morphology tools were suitable for our needs, so
+we have 
+The resources described in this paper Therefore, we developed free and 
+open-source finite-state 
+tools capable of analyzing and generating stressed wordforms, based on the
+well-known \emph{Grammatical Dictionary of Russian} \cite{Zaliznjak-77}.
+Our Finite-State Transducer\footnote{Using two-level 
+morphology \cite{Koskenniemi-84}, implemented in both xfst
+\cite{Beesley.Karttunen-03} and hfst \cite{hfst-11}} (FST) generates all possible 
+morphosyntactic readings of each wordform, and our Constraint 
+Grammar\footnote{Implemented using vislcg3 constraint grammar parser
+(http://beta.visl.sdu.dk/cg3.html).}
+\cite{Karlsson-90,Karlsson.Voutilainen.ea-95} then removes
+some readings based on syntactic context.
 
@@
     Wikipedia       & 7,857      & 0.506        & 0.996    & 44.92\%  \\
+    Literature      & 1,652      & 0.473        & 0.984    & 42.95\%  \\
     News            & 642        & 0.471        & 0.990    & 41.60\%  \\
-    Literature      & 1,652      & 0.473        & 0.984    & 42.95\%  \\
     \hline
-    Total           & 10,150     & 0.498        &  0.994   &  44.39\% \\
+    \textbf{Average}& 10,150     & 0.498        &  0.994   &  44.39\% \\
     \hline

### 2015-02-19 2015-eamt-kazkir/2015-eamt-kazkir.tex
harg
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59007 72bbbca6-d526-0410-a7d9-f06f51895060
more
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59008 72bbbca6-d526-0410-a7d9-f06f51895060
lll
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59009 72bbbca6-d526-0410-a7d9-f06f51895060
hargle
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59018 72bbbca6-d526-0410-a7d9-f06f51895060
stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59025 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
+\section{Kazakh and Kyrgyz}
+
+\section{Approaches}
+
+\subsection{Rule-based machine translation}
+
+\cite{washingtonsalimzyantyers14}
+\cite{washingtonipasovtyers12}
+
+\subsection{Corpus-based machine translation}
+
+We trained models based on two approaches, the first being phrase-based machine translation (PBMT, \cite{zens2002}) and the second on character-based machine translation (CBMT, \cite{nakov2012}; \cite{TiedemannEAMT2009}).
+
+\subsection{Corpus}
+
+\subsubsection{Phrase-based}
+
+\subsubsection{Character-based }
+
+\cite{nakov2012} present a method of statistical machine translation on the character level between related languages that takes advantage of phrase-based machine translation architecture. 
+%The method relies on preprocessing the input and output by inserting spaces in between the characters of words, for example the string `m\"{a} meen Helsinkiin' would become `m \"{a} \$ m e e n \$ H e l s i n k i i n' with a unigram model, or `m\"{a} \"{a}\$ \$m me ee en n\$ \$H He el ls si in nk ki ii in' with a bigram model.
+
+%After preprocessing, the corpora are processed as with the phrase-based system, with the difference that the language model order is increased from 5 to 10-grams.
+
+\section{Evaluation}
+
+\subsection{Coverage}
+
+\begin{table}
+  \begin{tabular}{|l|r|r|r|r|}
+    \hline
+    \textbf{Corpus} & \textbf{Tokens} & \textbf{RBMT} & \textbf{PBSMT} & \textbf{CBMT} \\
+    \hline
+    NT              & 144,946      &      &          &       \\
+    Akorda          & 382,716       &      &          &       \\
+    Wikipedia       & 27,802,493       &      &          &       \\
+    \hline
+  \end{tabular}
+  \caption{Vocabulary coverage for the three systems}
+  \label{table:coverage}
+\end{table}
+
+\subsection{Translation quality}
+
+\section{Future work}
+
+\section{Conclusions}
+
 \section*{Acknowledgements}

### 2015-02-19 2015-eamt-kazkir/2015-eamt-kazkir.tex
harg
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59026 72bbbca6-d526-0410-a7d9-f06f51895060
stuffs
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59027 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 Besides these systems, several previous works on making machine translation systems between Turkic languages 
-exist, although to our knowledge none are publicly available except for the Turkish--Azerbaijani pair available through Google Translate.\footnote{\url{http://translate.google.com}}
+exist, although to our knowledge none are publicly available except for the Turkic--Turkic pairs (including Azerbaijani, Kazakh, Turkish, and Uzbek) available through Google Translate.\footnote{\url{http://translate.google.com}}
 Some MT systems have been reported that translate between Turkish and other Turkic languages, 
@@
 
+\begin{table}
+  \begin{tabular}{|l|r|r|r|}
+    \hline
+    \textbf{System} & \textbf{PER} & \textbf{WER} & \textbf{BLEU} \\ 
+    \hline
+    RBMT            &              &              &               \\
+    PBSMT            &              &              &               \\
+    CBMT            &              &              &               \\
+    \hline  
+  \end{tabular}
+\end{table}
+
 \section{Future work}

### 2015-02-20 2015-eamt-assim/2015-eamt-assim.tex
Made suggested edits to the first 3 sections
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59029 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \maketitle
-%Katya: has not it first been used in trosterud12?
+\comment{EA: has not it first been used in trosterud12?}
 \begin{abstract}
@@
 \comment{FMT: Abstract should mirror/summarise the article, e.g. include something about conclusions too}
+\comment{EA: will add conclusions after writing the section on them}
 \end{abstract}
+
+\comment{EA: I noticed some spelling edits, e.g.
+organize -> -se, program -> -mme.
+Is it important that we use British spelling?
+If so, may I count on someone to go through the paper and correct it?
+I am not sure I can use British spelling consistently.}
 
@@
 As suggested by \citet{church93}, modern machine translation (MT) systems may
-be divided into two broad categories according to their purpose: post-editing systems, e.g.
-Google Translate\footnote{translate.google.com}, and assimilation systems, e.g.
-Apertium\footnote{apertium.org}.
+be divided into two broad categories according to their purpose: post-editing systems and assimilation systems.
 Both kinds may be systematically evaluated, either to control for quality in the development process or to compare the systems.
@@
 \comment{FMT: Apertium can also be used for postedition, e.g. spanish-catalan, and some Google translate directions aren't suitable for that. Perhaps reword or remove this}
+\comment{EA: removed}
 
@@
      \hline
-     \textbf{Ref}   & Ayudas económicas para el tratamiento de toxicomanías en comunidades terapéuticas no concertadas. \\
-     \textbf{Task}   & Ayudas económicas para el \{ \} de toxicomanías en comunidades terapéuticas no concertadas. \\
+     \textbf{Ref}   & Ayudas econ\'{o}micas para el tratamiento de toxicoman\'{i}­as en comunidades terap\'{e}uticas no concertadas. \\
+     \textbf{Task}   & Ayudas econ\'{o}micas para el \{ \} de toxicoman\'{i}­as en comunidades terap\'{e}uticas no concertadas. \\
      \textbf{Src} & Komunitate terapeutiko itundu gabeetan toxikomaniak tratatzeko diru-laguntzak ematea. \\
-     \textbf{MT}     & Comunidad terapéutico pactar gabeetan toxikomaniak las-ayudas de dinero para tratar dar. \\
+     \textbf{MT}     & Comunidad terap\'{e}utico pactar gabeetan toxikomaniak las-ayudas de dinero para tratar dar. \\
      \hline
@@
 a total of 36;
-\item  Each participant evaluates 9 sentences in each of the four modes (see section~\ref{sec:methodology};
+\item  Each participant evaluates 9 sentences in each of the four modes (see section~\ref{sec:methodology});
 \item  All sentences of the set are evaluated with 10, 20 and 30\% of words removed;
@@
 increase scores where appropriate. Candidate synonym extraction and score update is performed automatically.
+
+\begin{table*}
+  \begin{tabular}{|l|l|}
+     \hline
+     \textbf{Sentence:}   & Aprender a jugar y divertirse en el agua sin asumir riesgos. \\
+     \textbf{Key:}   & asumir \\
+     \textbf{Synonym:} & correr \\
+     \hline
+  \end{tabular}
+  \caption{An example candidate synonym.} 
+  \label{table:syn}
+\end{table*}
+\comment{EA: added}
+The synonym lists for Basque-Spanish, English-Kazakh and Tatar-Russian contain 52, 38 and 25 words, respectively.
+Time taken to compile one list depends on the number of candidate synonyms, and in our case was approximately 30 minutes.
 \comment{FMT: How long did making the synonym list take, how many synonyms were added? More details here. Also perhaps an example?}

### 2015-02-20 2015-eamt-assim/2015-eamt-assim.tex
Added result tables
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59037 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \comment{EA: I noticed some spelling edits, e.g.
-organize -> -se, program -> -mme.
+organize - -se, program - -mme.
 Is it important that we use British spelling?
@@
 As suggested by \citet{church93}, modern machine translation (MT) systems may
-be divided into two broad categories according to their purpose: post-editing systems and assimilation systems.
-Both kinds may be systematically evaluated, either to control for quality in the development process or to compare the systems.
+be divided into two broad categories according to their purpose: post-editing and assimilation systems.
+Both kinds may be evaluated, either to control for quality in the development process or to compare the systems.
 Importantly, according to \citet{church93}, the evaluation methods must closely consider the system's primary purpose.
@@
 context. We then check participants' results against the compiled synonym list and
-increase scores where appropriate. Candidate synonym extraction and score update is performed automatically.
+increase scores where appropriate. On average, the scores improved by three percent in all evaluation modes. Candidate synonym extraction and score update is performed automatically.
 
@@
 
-\section{Expected results}
+%\section{Expected results}
+%
+%We expect that, for tasks with MT assistance, the users will fill more gaps correctly,
+%which would correspond to better understanding of text. Therefore, the percentage of
+%correct answers should be higher for MT and two-hint kinds of tasks than for tasks with
+%source sentence as a hint and no hint at all. If the percentage were lower, we would
+%expect an error in experiment design.
 
-We expect that, for tasks with MT assistance, the users will fill more gaps correctly,
-which would correspond to better understanding of text. Therefore, the percentage of
-correct answers should be higher for MT and two-hint kinds of tasks than for tasks with
-source sentence as a hint and no hint at all. If the percentage were lower, we would
-expect an error in experiment design.
-
-\section{Results}
+\section{Results and discussion}
 \label{sec:results}
 
-The experiment results are presented in the tables below. The first table of each set shows
+The experiment results are presented in the tables below. Tables \ref{table:res-eus-spa}, \ref{table:res-eng-kaz} and \ref{table:res-tat-rus} show
 the proportion of correct answers depending on evaluation mode and gap density.\comment{FMT: We should probably integrate the other tables, and refer to them using labels.} The
@@
 answers given by all evaluators by the number of questions answered by all evaluators.
-The second table of the set shows average time it took to fill the gaps in one sentence. To
+Tables \ref{table:time-eus-spa}, \ref{table:time-eng-kaz} and \ref{table:time-tat-rus} show average time it took to fill the gaps in one sentence. To
 reduce the noise from participants who were distracted during evaluation, when
@@
 
+\comment{EA: and what would we like to say about those?
+The results depend heavily on the Tatar-Russian system.
+Do we have to provide info on what vocabulary it handles better, what it has and has not been trained on?
+Or do we just say that casual texts are easier to fill than legal texts?}
+For Tatar--Russian language pair the experiment was conducted in three parts, each part on texts from different domains.
+The results by domains are displayed in tables \ref{table:rus-casual}, \ref{table:rus-legal}, and \ref{table:rus-news}.
+
+% results with synonyms
 \begin{table}
@@
 \end{table}
+
+% begin time tables
+\comment{EA: maybe remove SD from time tables? is it informative?}
+\begin{table*}
+\begin{tabular}{|r |*{4}{c}|}
+\hline
+ &\multicolumn{4}{c|}{\textbf{Evaluation mode}}\\
+\hline
+\textbf{Gap percentage} & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint}\\
+10\%&1:17 +/- 1:05&1:33 +/- 1:22&1:06 +/- 0:38&0:46 +/- 0:44\\
+20\%&1:23 +/- 1:04&1:14 +/- 0:58&1:27 +/- 1:02&1:22 +/- 1:17\\
+30\%&2:22 +/- 1:10&2:33 +/- 1:26&2:36 +/- 1:22&2:25 +/- 1:30\\
+\hline
+\end{tabular}
+\caption {Basque--Spanish: Time taken to fill the gaps in one sentence, mean and standard deviation.}
+\label{table:time-eus-spa} 
+\end{table*}
+
+\begin{table*}
+\begin{tabular}{|r |*{4}{c}|}
+\hline
+  &\multicolumn{4}{c|}{\textbf{Evaluation mode}}\\
+\hline
+\textbf{Gap percentage} & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
+10\%&2:31 +/- 1:29&1:34 +/- 1:03&1:47 +/- 1:36&0:47 +/- 0:35\\
+20\%&1:17 +/- 0:51&2:05 +/- 1:57&2:10 +/- 2:14&1:13 +/- 1:02\\
+30\%&2:00 +/- 1:30&1:46 +/- 1:48&1:41 +/- 1:02&1:44 +/- 1:34\\
+\hline
+\end{tabular}
+\caption {English--Kazakh: Time taken to fill the gaps in one sentence, mean and standard deviation.}
+\label{table:time-eng-kaz} 
+\end{table*}
+
+\begin{table*}
+\begin{tabular}{|r |*{4}{c}|}
+\hline
+  &\multicolumn{4}{c|}{\textbf{Evaluation mode}}\\
+\hline
+\textbf{Gap percentage} & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
+10\%&1:26 +/- 1:10&1:05 +/- 1:05&0:45 +/- 0:52&0:31 +/- 0:22\\
+20\%&1:16 +/- 2:06&1:12 +/- 1:44&0:57 +/- 0:59&0:50 +/- 0:55\\
+30\%&1:19 +/- 1:15&1:14 +/- 1:11&1:10 +/- 1:04&1:01 +/- 0:53\\
+\hline
+\end{tabular}
+\caption {Tatar--Russian: Time taken to fill the gaps in one sentence, mean and standard deviation.}
+\label{table:time-tat-rus} 
+\end{table*}
+
+% results by text mode in tat-rus, with synonyms
+
+\begin{table*}
+\begin{tabular}{|r |*{4}{c}|}
+\hline
+  &\multicolumn{4}{c|}{\textbf{Evaluation mode}}\\
+\hline
+\textbf{Gap percentage} & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
+10\%&59.52&60.98&65.00&54.76\\
+20\%&73.68&63.16&44.30&37.97\\
+30\%&68.18&58.24&41.67&41.94\\
+\hline
+\end{tabular}
+\caption {Tatar--Russian, casual conversations: Gaps successfully filled (\%), using a synonym list, average} \label{table:rus-casual} 
+\end{table*}
+
+\begin{table*}
+\begin{tabular}{|r |*{4}{c}|}
+\hline
+  &\multicolumn{4}{c|}{\textbf{Evaluation mode}}\\
+\hline
+\textbf{Gap percentage} & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
+10\%&52.00&67.11&43.48&38.03\\
+20\%&61.74&63.57&52.46&43.31\\
+30\%&56.07&41.32&37.36&37.23\\
+\hline
+\end{tabular}
+\caption {Tatar--Russian, legal texts: Gaps successfully filled (\%), using a synonym list, average} \label{table:rus-legal} 
+\end{table*}
+
+\begin{table*}
+\begin{tabular}{|r |*{4}{c}|}
+\hline
+  &\multicolumn{4}{c|}{\textbf{Evaluation mode}}\\
+\hline
+\textbf{Gap percentage} & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
+10\%&52.00&68.75&61.22&50.98\\
+20\%&64.00&52.86&46.15&34.25\\
+30\%&59.38&63.28&41.41&37.50\\
+\hline
+\end{tabular}
+\caption {Tatar--Russian, news: Gaps successfully filled (\%), using a synonym list, average} \label{table:rus-news} 
+\end{table*}
 

### 2015-02-21 2015-eamt-assim/2015-eamt-assim.tex
remove dealt-with comment
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59039 72bbbca6-d526-0410-a7d9-f06f51895060
First draft complete
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59059 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 To reduce preparation time, an automatic open-source task management system is introduced.
+The evaluation results suggest that the gap-filling task reflects users' understanding of text, and may be used to measure MT quality for assimilation purposes.
 \comment{FMT: Abstract should mirror/summarise the article, e.g. include something about conclusions too}
-\comment{EA: will add conclusions after writing the section on them}
+\comment{EA: added}
 \end{abstract}
@@
 fall into the allowed POS list. The number of gaps in the sentence is calculated based on 
-sentence length and specified gap density. Finally, the required number of keywords is selected 
+sentence length and specified gap density. All reference sentences are over 10 words in length. Finally, the required number of keywords is selected 
 from the candidate list in such a manner that the gaps are distributed evenly throughout the sentence.
@@
 Having prepared the sentence sets, we assemble them into XML formatted for the Appraise platform.
-
-%To exclude headlines and also to allow at least one gap in 10\%-gap sentences, only the
-%sentences over 10 words long are used in the tasks.
-%The sentences for each experiment are drawn from various parallel text sources:
-%\begin{enumerate}
-%\item  For Basque-Spanish, from the corpus of legal texts Memorias de traducci\'on del
-%Servicio Oficial de Traductores del IVAP 3
-%\item  For English-Kazakh, from the official website of the President of the Republic of
-%Kazakhstan 4
-%\item  For Tatar-Russian, from the following sources on three different topics:
-%  \begin{enumerate}
-%    \item  Casual conversations, from a textbook 5 of spoken Tatar;
-%    \item  Legal texts, from the Constitution and laws 6 of Tatarstan;
-%    \item  News, from the President of Tatarstan website 7 .
-%  \end{enumerate}
-%\end{enumerate}
-%
-%Each set features 36 pairs of sentences. For the first two experiments the pairs are drawn
-%randomly from the corpora, for Tatar-Russian, compiled by hand by the developer of the
-%language pair. After processing, the material is transformed into XML used for uploading
-%into the evaluation system (see section 2.4).
 
@@
 \item  All sentences of the set are evaluated with 10, 20 and 30\% of words removed;
-\item  Each sentence-mode-percentage combination is evaluated by at least three participants.
+\item  Each sentence-mode-percentage combination is evaluated by more than one participant.
 \end{enumerate}
@@
 
-\comment{EA: the sections below are from the report and have not been edited yet.}
-
-%\section{Expected results}
-%
-%We expect that, for tasks with MT assistance, the users will fill more gaps correctly,
-%which would correspond to better understanding of text. Therefore, the percentage of
-%correct answers should be higher for MT and two-hint kinds of tasks than for tasks with
-%source sentence as a hint and no hint at all. If the percentage were lower, we would
-%expect an error in experiment design.
-
 \section{Results and discussion}
@@
 
+% expected results
+We expect to see the following trend: scores received in different task modes inside one gap density are observed, from highest to lowest, in tasks with MT and source hint, then MT hint only, then source hint only, and finally, no hint.
+We also expect that with the increase in gap density, time taken to fill the gaps also increases, and percentage of correct answers decreases.
+
+% a few notes on time
+For Basque--Spanish and Tatar--Russian, the latter trend holds: the average time taken to fill the gaps increases and the average percentage of correct answers decreases as the relative number of gaps goes up.
+The larger number of gaps in the sentence makes it more
+difficult to predict the answer based on the context, and also leaves more room for
+translation mistakes.
+Exploring different percentage-mode combinations, we may note that the 10\% no-hint tasks take the least time to complete.
+We would have expected longer completion time, since the participant must come up with their own answer unassisted.
+However, in the no-hint task the participant reads only one sentence, as opposed to two or three in other tasks.
+Also, the number of gaps in 10\%-gap tasks is low, as it never exceeds three.
+We found that, as opposed to trying to devise the best word for no-hint gaps, the participants often resorted to filing these gaps with random words, which takes little time.
+\comment{EA: anything else I could write about times?}
+
+% this section talks about percentages in eu-es and tat-rus, and how everything is nice.
+We will now discuss the percentage of correct answers based on task type.
+In general, tasks with MT hints score higher than tasks without MT hints.
+This aligns well with our expectations and suggests than machine translation helps to understand the provided text.
+In addition, tasks with source hint are completed better than tasks without hints, and the same relation holds between MT+source and MT-only types of tasks.
+
+Two records break this trend: the no-hint 10\% sentences in Basque--Spanish, which scored significantly higher than the source-hint in the same category, and MT+source 10\% sentences in Tatar--Russian, which we would have expected to score higher than the corresponding MT task.
+In the first case, this is largely due to the use of synonyms list.
+Before taking synonyms into account, the scores were 32 and 35 percent for source and no-hint tasks, respectively.
+This still shows a small difference in favor of no-hint tasks.
+However, the latter percentage increases significantly after we extend the answer list with synonyms.
+Such an increase suggests that, in this case, the content words were restored by sense rather than through strong collocation.
+The second pattern, low scores in Tatar--Russian 10\% MT+source, does not stem from the task content.
+Instead, it is the result of the fixed order of tasks: the participants have always been given MT+source 10\% sentences first, followed by other task types.
+The participants have not received any training tasks before the main evaluations.
+Therefore, it is possible that the accommodation period is responsible for lower-than-expected scores in this mode of evaluation.
+
+It remains questionable whether we can compare results for different gap densities. In the
+Basque-Spanish and Tatar-Russian experiments the 10, 20 and 30\% sets were comprised
+of the same sentences. However, in each case different words were removed. It appears
+that some content words are easier to fill than the others. This may explain why in
+Basque-Spanish the 20\% MT tasks are completed with better accuracy than 10\% tasks.
+
+% a little on categories
 \comment{EA: and what would we like to say about those?
@@
 Or do we just say that casual texts are easier to fill than legal texts?}
-For Tatar--Russian language pair the experiment was conducted in three parts, each part on texts from different domains.
+For Tatar--Russian language pair the participants were offered texts from three different domains (in equal proportions): casual conversations, legal texts and news.
 The results by domains are displayed in tables \ref{table:rus-casual}, \ref{table:rus-legal}, and \ref{table:rus-news}.
+The system used in the evaluation has been targeted to translate texts from all three of the domains.
+Taking into consideration the above discussion of 10\% MT+Source tasks, we observe similar results across the three categories.
+Note that, in casual and news texts, the source sentences paired with MT significantly improve participants' performance compared to MT-only task mode.
+This may be due to the fact that many words are borrowed from Russian into Tatar, and are in fact understood by Russian speakers.
+
+% this section talks about eng-kaz, where everything is not so nice 
+\comment{EA: considering all those flaws in setup, should we report Kazakh results at all?}
+In the above discussion of results we do not take into account the data from the English--Kazakh language pair.
+It remains questionable whether we may draw reliable conclusions from this part of the experiment.
+By design, the evaluation tasks are distributed evenly among the participants, in such a manner that each sentence is evaluated in all modes or gap densities by several people.
+The English--Kazakh evaluations have been set up for 11 participants, however, only 8 participants attempted their tasks, and 6 of them completed it in full.
+As a result, the evaluation became unbalanced: out of 4 sets of sentences, three have been evaluated by three, two and one participant, respectively, and the fourth set has been partially completed by only one participant.
+A total of 40 (about 10\%) gaps have not been filled by any of the participants.
+
+We should also note the outlier score of 60\% for the 20\% no-hint task.
+It is made up by three sentences, two of which scored unusually high.
+One of these sentences is a news headline, and the other contains clich\'es such as ``express condolences''.
+We would have expected the same sentences to contribute to percentages in other evaluation modes, but they have not been completed there.
+
+Finally, we would like to comment on the relatively higher scores on tasks with source hints (columns 2 and 4 of Table \ref{table:res-eng-kaz}.
+By experiment design, the participants must have no command of the source language of the pair, which is, in this case, English.
+In Kazakhstan, English is a mandatory part of the middle school curriculum, and the students start learning it at the age of 10.
+It therefore seems likely that our participants, aged 23-26, had at least some command of English, which may have contributed to their understanding of source-hint tasks.
+
+% participants and their suffering
+It is worth noting that many participants reported feeling frustrated in
+the course of evaluations, especially while working on the no-hint tasks. 6 out of 49
+participants quit the experiment before completing it.
+Considering the importance of receiving the full set of evaluations, we must address the issue of participant motivation in the upcoming experiments.
 
@@
 \label{sec:conclusion}
-The following trends may be derived from the data above:
+We have conducted assimilation evaluation of three Apertium translation directions: Basque--Spanish, English--Kazakh and Tatar--Russian.
+The results suggest that this evaluation method reflects the contribution of MT to users' understanding of text.
+The version of the toolkit used in this experiment may be downloaded from our repository\footnote{https://github.com/Sereni/Appraise/tree/1e9d735faee64d1b97fb343ab111ace6a64509d7}.
+
+The experiment may be repeated for any language pair (provided a parallel corpus) and
+any machine translation system. Based on our experience, we would like to suggest the following amendments to the procedure:
 \begin{enumerate}
-\item  In general, tasks with MT assistance score higher than task without MT
-assistance;
-\item  The more words are removed from context, the more difficult it is to fill the gaps
-(see below for exceptions);
-\item  Task without hints take less time to complete than others.
+\item If the experiment is conducted with different gap densities, each consecutive set of sentences should be derived from the set with lower gap density.
+To put another way, the sentence with e.g.
+20\% gaps should contain all the gaps of the same sentence in 10\% mode, plus any additional gaps;
+\item Unless the evaluation is targeted at a specific text domain, it may be beneficial to include a stylistic variety of texts in the initial corpus.
+Neighboring sentences on the same topic may assist the users in gap-filling tasks;
+\item If possible, increase the number of evaluators, or reduce the number of questions per participant.
+In the above experiments each participant filled from 110 to 187 gaps, divided into small groups.
+Reducing the amount of work may increase task completion rate;
+\item To account for the adaptation period, provide training tasks before the main evaluations take place.
 \end{enumerate}
-The first two trends align with our expectations. Higher scores on MT tasks indicate
-better understanding of text. The large number of gaps in the sentence makes it more
-difficult to predict the answer based on the context, and also leaves more room for
-translation mistakes. We would expect that tasks without hints take the largest amount of
-time rather than that the smallest, since the participant must come up with their own
-answer unassisted. However, in the no-hint task the participant reads only one sentence,
-as opposed to two or three in other tasks, which may be the source of difference.
-In Basque-Spanish and Tatar-Russian these trends are defined reasonably well. In
-English-Kazakh, however, the scores for source-hint task (a parallel sentence in English
-for assistance) are systematically higher than for other types of tasks. In addition, a 20\%no-hint task stands out with 60\% accuracy, which is substantially higher than any other
-scores for this experiment.
-Having analysed the tasks and the responses for English-Kazakh, I have found the
-following possible reasons of the obtained results:
-\begin{enumerate}
-\item  The experiment was planned for 11 evaluators, although only 8 attempted the
-tasks and 6 completed them in full. Because of this, the sentences were evaluated
-once or twice, and some were not attempted at all. To reliably compare results
-across different categories, all tasks must be completed;
-\item  Due to an error during experiment design, some sentences are not repeated across
-categories, which also suggests that we may not compare these percentages;
-\item  The outlier score of 60\% is made up by three sentences. Two of these scored
-unusually high. It appeared that one of them is a news headline, and the other
-contains clich\'es such as ``express condolences''. There may also be interference
-from other tasks of the set, because some of the sentences are drawn from the
-same news article. We would expect the same sentences to contribute to higher
-percentage in other task modes, but they were not completed there;
-\item  It may be the case that the participants have at least some command of English,
-which would have helped them to derive extra information in the source-hint
-\end{enumerate}
-It remains questionable whether we can compare results for different gap densities. In the
-Basque-Spanish and Tatar-Russian experiments the 10, 20 and 30\% sets were comprised
-of the same sentences. However, in each case different words were removed. It appears
-that some content words are easier to fill than the others. This may explain why in
-Basque-Spanish the 20\% MT tasks are completed with better accuracy than 10\% tasks.
-It may be worth noting that the majority of the participants reported feeling frustrated in
-the course of their work, especially while working on the no-hint tasks. 9 out of 51
-participants quit the experiment before completing it.
 
-\section{Future work}
-\label{sec:futurework}
-The experiment may be repeated for any language pair (provided a parallel corpus) and
-any machine translation system. To enhance the results, the procedure may be amended
-in the following ways:
-\begin{enumerate}
-\item  Reduce the number of questions per participant. In the above experiments each
-participant filled from 110 to 187 gaps, divided into small groups. Reducing the
-amount of work may increase task completion rate;
-\item  Increase the number of participants per experiment. The more evaluators work on
-every question, the more reliable the data;
-\item  Design the tasks with different gap densities in such a way that e.g. the 20\% task
-contains all the gaps of the 10\% task, plus any additional ones;
-\item  Draw sentences from large corpora. This will reduce the influence of previous
-tasks, which may provide the context for other tasks in the set.
-\end{enumerate}
+\section{Acknowledgments}
+
+This work has been partly funded by the Spanish Ministerio de Economía y Competitividad through project TIN2012-32615 and by the Google Summer of Code programme.
+We would like to thank the volunteers who participated in the evaluations.
 

### 2015-02-23 2015-eamt-assim/2015-eamt-assim.tex
some fixes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59086 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 The evaluation results suggest that the gap-filling task reflects users' understanding of text, and may be used to measure MT quality for assimilation purposes.
-\comment{FMT: Abstract should mirror/summarise the article, e.g. include something about conclusions too}
-\comment{EA: added}
 \end{abstract}
-
-\comment{EA: I noticed some spelling edits, e.g.
-organize - -se, program - -mme.
-Is it important that we use British spelling?
-If so, may I count on someone to go through the paper and correct it?
-I am not sure I can use British spelling consistently.}
 
@@
 Importantly, according to \citet{church93}, the evaluation methods must closely consider the system's primary purpose.
-\comment{FMT: Apertium can also be used for postedition, e.g. spanish-catalan, and some Google translate directions aren't suitable for that. Perhaps reword or remove this}
-\comment{EA: removed}
 
@@
 \begin{table*}
+  \centering
   \begin{tabular}{|l|l|}

### 2015-02-26 engkaz-eamt-2015/engkaz-eamt-2015.tex
c/p some of the text from the .odt into the .tex file
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59140 72bbbca6-d526-0410-a7d9-f06f51895060

@@
-\documentclass[a4paper,twocolumn,11pt]{article}
+% File eamt15.tex, an adaptation of eamt14.tex (which was a copy of
+% eamt12.tex)
+%
+% Contact: eamt2015@dlsi.ua.es
 
+%%% To ease future customizations, various replaceables have been paramaterized
+%%% as listed in the newcommands section
+
+\documentclass[11pt]{article}
+\usepackage{eamt15}
+\usepackage{times}
+\usepackage{latexsym}
+\setlength\titlebox{6.5cm}    % Expanding the titlebox
+%%% YOUR PACKAGES BELOW THIS LINE %%%
 \usepackage{url}
+
+\usepackage{multirow}
+
+\newcommand{\confname}{EAMT 2015}
+\newcommand{\website}{http://www.eamt2015.org/}
+\newcommand{\contactname}{the conference chairs (Felipe
+  S\'anchez-Mart\'inez, Gema Ram\'irez-S\'anchez and Fred Hollowood)}
+\newcommand{\contactemail}{eamt2015@dlsi.ua.es}
+\newcommand{\conffilename}{eamt15}
+\newcommand{\downloadsite}{http://www.eamt2015.org/}
+\newcommand{\paperlength}{$8$ (eight)}
+\newcommand{\shortpaperlength}{$4$ (four)}
 
@@
 
+\date{}
+
 \begin{document}
+\maketitle
+
+\begin{abstract}
+..
+\end{abstract}
 
@@
 
-SOMETHING ABOUT ENGLISH-to-KAZAKH trnslation problem??
-
+This paper presents a shallow-transfer rule-based machine translation system from English into Kazakh language.
+First of all, English language is a West Germanic language and Kazakh belong to group of Turkic languages.
+Therefore, the complex agglutinative morphology of Kazakh language is very different from morphologically not too complex language like English; on the other hand, Turkic language morphology shows clear morphotactics (ordering of morphemes), its morphophonology shows complex phonological changes to due to interactions between neighboring morphemes (vowel harmony, sonorization, etc.) many of which are explicitly represented in writing.
+On the other hand, there are many differences between the syntax of Turkic languages and English:.subject-adverb–object–verb order (compare subject–verb–object-adverb in English), use of postpositions (compare prepositions in English), head-final syntax with modifiers and specifiers always preceding the modified/specified (normally following in English), overt case marking allowing for a rather free ordering of arguments (versus a more fixed order in English), lack of definite articles (extensively used in English), verbal-noun-centered structures where English uses modal verbs (must, have to, want to) or verbal-noun or verbal-adjective-centered constructions where English has subordinate clauses using finite verbs with relatives or subordinating conjunctions (the book which I read, the place where I saw him, before he came), lack of a parallel of the English verb have, as used for possession, etc.
+For an account (in Russian) of syntax differences between English and Kazakh, see Печерских & Амангельдина (2012).
 This paper describes work in progress of development of machine translation system for English-to-Kazakh, which developed by using Apertium free/open-source machine translation platform (Forcada et al.
@@
 
-FIGURE construction of Apertium
-WILL EDIT THIS
-	De-formatter.
+\begin{figure*}[htbp]
+\begin{center}
+ \includegraphics[width=0.8\textwidth]{architecture.pdf}
+\end{center}
+\caption{The pipeline architecture of the Apertium system.}
+\label{fig:modules}
+%\vspace{-1em}
+\end{figure*}
+
+\begin{itemize}
+\item 	De-formatter.
 Separates the text to be translated from the formatting tags.
 Formatting tags are encapsulated in brackets so they are treated as “superblanks” that are placed between words in such a way that the remaining modules see them as regular blanks.
-	Morphological analyser.
+\item	Morphological analyser.
 For each surface form the morphological analyser generates one or more lexical forms, which consist of: lemma (dictionary or citation form), lexical category (or part-of-speech), and inflection information.
@@
 This module therefore separates lexemes and processes morphological analysis, and then returns possible lexical forms.
-	Part-of-speech (POS) tagger.
+\item	Part-of-speech (POS) tagger.
 This module chooses one of lexical forms of ambiguous word.
@@
 This word has 3 lexical forms, and depends on context, ont of them will be chosen by rules or tagger. After  this module, all words have only one morphological analysis.
-	Lexical transfer.
+\item	Lexical transfer.
 It reads each lexical form of source language and delivers a corresponding target language lexical form.
@@
  Multiword units are translated as a single word.
-	Lexical selection.
+\item Lexical selection.
 It uses rules for lexical words, which have many translations, to select one of the translations in the target language according to context.
 All rules are written in file apertium-eng-kaz.eng-kaz.lrx.
-	Structural transfer.
+\item Structural transfer.
 This module uses pattern matching to identify sequences of lexical forms  (phrases or segments), which need syntactical processing to translate grammatical differences between two languages (handling of number, gender, etc.).
@@
 
-	 Morphological generator.
+\item	 Morphological generator.
 Delivers the sequence of target-language lexical forms, produced by the structural transfer, to a corresponding sequence of target language surface forms.
 The morphological generator executes a finite-state transducer generated by compiling a morphological dictionary for the target language.
-	Post-generator.
+\item Post-generator.
 It performs some minor orthographical operations in the target language (for instance, it generates the English form cannot from can and not).
 This module is generated from file with rules which are very   similar in format to dictionary files.
-	Reformatter. It places format tags back into the text so that its format is preserved.
+\item Reformatter. It places format tags back into the text so that its format is preserved.
 
-\section{Linguistic Data(dictionaries,rules for eng-kaz)}
+\section{Linguistic data}
 
-In Apertium is used XML-based formats for linguistic data, it includes bilingual and monolingual dictionaries, structural transfer, lexical rules and rules for part-of-speech tagging.
+Apertium uses XML-based formats for linguistic data, which include bilingual and monolingual dictionaries, structural transfer, lexical rules and rules for part-of-speech tagging.
 
-Dictionaries
+\subsection{Dictionaries}
+
 Dictionaries are used in lexical processing and morphological analysing, there are use two types: morphological(monolingual) and bilingual.
 
-Monolingual dictionaries
+\subsubsection{Monolingual dictionaries}
 
@@
 
-Bilingual dictionary
+\subsubsection{Bilingual dictionary}
 
@@
 
-Rules
-Disambiguation rules
+\subsection{Rules}
+
+\subsection{Disambiguation rules}
 
@@
 
-Lexicon selection rules
+\subsection{Lexicon selection rules }
+
+Lexical selection rules choose on of the alternative translations of one word depend on the context.
+This rules are written in apertium-eng-kaz.eng-kaz.lrx.
+Alternative translations are defined in bilingual dictionary by multiple entries of one word.
+For exmaple, adjective “beautiful” could be translated as “сұлу”, if following noun is subject: “beautiful girl” - “сұлу қыз”; or “әдемі”, “көркем”, if the noun means place or : “beautiful mountains” - “әдемі таулар”.
+In the following table could be seen some lexical selection rules:
+
+\begin{table*}
+ \begin{tabular}{|l|l|l|l|}
+    \hline
+    \textbf{Word} & \text{Translation} & \textbf{Condition} & \textbf{Example} \\
+    \hline
+    \multirow{2}{*}{residence} & мекен & -- & I live in a residence. -- (Мен) мекенде өмір сүремін. \\
+                               & резиденция & \_ of \_ president & I see the residence of the president -- (Мен) президенттің резиденциясын көремін \\
+
+    \hline
+ \end{tabular}
+
+\end{table*}
+
+\subsection{Transfer rules}
+
+For English–Kazakh system this stage uses three stages to improve the granularity of structural transfer rules (each one has its own rules file):
+\begin{itemize}
+\item A first round of transformations (“chunker”) detects source language(SL) LF patterns and generates the corresponding sequences of TL LFs grouped in chunks representing simple constituents such as noun phrases, prepositional phrases, etc.
+\item The second round (“interchunk”) reads patterns of chunks and produces a new sequence of chunks.
+This is the module where one can attempt to perform some longer-range reordering operations, interchunk agreement, case selection, etc.
+\item The third round (“postchunk”) transfers chunk-level tags to the lexical forms they contain and whose lexical-form-level tags are linked (through a referencing systems) to chunk-level tags (for instance, case determined for a noun phrase is transferred to the main noun), and removes all grouping information to generate the desired sequence of TL LFs.
+\end{itemize}
+
+\subsection{English-to-Kazakh structural transfer}
+
+The structural transfer module in Apertium processes the stream of source-language lexical form – target-language lexical form pairs (SL LF–TL LF pairs) and transforms it into a new sequence of TL LFs after a series of structural transfer operations specified in a set of rules: reordering, elimination or insertion of TL LFs, agreement, etc.
+This section describes the current structural transfer in Apertium-eng-kaz (revision ..., ).
+English to Kazakh chunker rules (file apertium-eng-kaz.eng-kaz.t1x), inter-chunk rules (file apertium-eng-kaz.eng-kaz.t2x) and additional clean-up stage(file apertium-eng-kaz.eng-kaz.t4x) will be described in detail next 3 sections.
+
+\subsubsection{The English-to-Kazakh chunker}
+
+In the first round of structural transfer (the “chunker”, rules written in the apertium-eng-kaz.eng-kaz.t1x file), rules do some local operations to cut sentences into chunks, like short noun phrases, adjective phrases, verb phrases and adpositional phrases (that is, prepositional phrases in English and postpositional phrases in Kazakh).
+Chunking rules, of which there are currently 168, identify 8 kinds of chunks and translate them into Kazakh as much as possible, leaving some minor operations to be performed in later stages of structural transfer (for instance, the case of noun phrases).
+Table 3 describes what kind of noun phrases could appear in sentences and how it could be detected.
+
+%% EXAMPLES ?
+
+\begin{description}
+\item[Noun phrases].
+Consider the following example: the chunker identifies the English sequence the large book  (determiner–adjective–noun) as a noun-phrase chunk.
+It translates into Kazakh, and assigns it four chunk-level tags: number (set to singular), person (set to 3rd), possessor (to be determined, as the noun кітап ('book') could receive a 3rd-person possesive ending (кітабы) later if the context were, for instance, the large book of animals, аңдардың үлкен кітабы), and case (to be determined as it could be, for instance, accusative in I saw the large book, Мен үлкен кітапты көрдім ).
+Also noun form of the verb, for instance,  gerund form is determined as NP phrase(Her playing is good – Оның ойыны жақсы).
+\item[Prepositional phrases]. English prepositional phrase is translated into Kazakh a postpositional phrase, there are three possible outcomes with different cases:
+\begin{itemize}
+\item Translating the locative “-{D}{A}”,1 ablative “-{D}{A}н”, etc.,except the genitive “-{N}{I}ң” cases: [PP [P from ] [NP the five cars] ] → [PP [NP бес көлік] [P -тен ] ]
+\item Genitive -NIņ, which will be marked GenP:  [PP [P of ] [NP the beautiful garden] ] → [GenP [NP әдемі бақша] [P -ның] ] 
+\item 3 With noun such as аст, үст, etc. which takes possessive from main noun:  [PP [P under ] [NP the garden] ] → [PP [NP [GenP [NP бақша ][P -ның]] [NP астын]] [P -да ] ]
+\end{itemize}
+\item[Verb phrases]. Translation verb phrases into Kazakh  are not always be straightforward, for instance: 
+present simple and future are rendered using the same tense in Kazakh (I play → Мен ойнаймын;  I will play → Мен ойнаймын), this ambiguity problem may be solved by lexical selection, which will choose right tense by watching context;
+tenses expressing continued activity, such as the English present continuous or past continuous (I am playing, I was playing), have to be detected and mapped onto sets of two lexical units (Мен ойнап жатырмын, Мен ойнап отырдым) where the main verb is found in the -п participle form (ойнап), and a suitable finite form  (жатырмын, отырдым) of an auxiliary verb (жатыр, отыр) is used to express number and person agreement1,
+modal verbs are translated by adding adjectives, which means “necessary” (керек) or “proper” (жөн), and a form of the copula (absent in present tense); the subject receives the genitive or dative case;
+negative verbs for present continuous and past tense are translated by adding negative word “емес” for present perfect continuous( He has not been playing - Ол ойнап отырған емес ) and  “жоқ” for past tense and present continuous(see examples above).
+\item[Translation of adjectival phrases].
+In Kazakh noun phrases, adjectives come before nouns and do not show any agreement with nouns.
+Adjectives can also appear in separate adjective phrases.
+Here are some examples:
+\begin{itemize}
+\item (1)  The adjective alone, marked AdjP: [AdjP big ] → [AdjP үлкен] 
+\item (2)  Comparative adjective phrases  (English more + adjective, or adjective-[e]r); the Kazakh translation chooses the comparative suffix “-{I}р{A}{K}”: [AdjP  more  interesting ]  → [AdjP қызығырақ]
+\item (3)  For superlative adjective phrases  “the most + adjective”  or “adjective-[e]st”, translation is built using “ең” + adjective:  [SupP  the biggest ]  → [SupP ең үлкен] [SupP  the most important ]   → [SupP ең маңызды]
+\end{itemize}
+Superlative adjective phrases have some properties of noun phrases (such as receiving possessive morphemes when modified by a genitive phrase: the most beautiful of people → адамдардың ең әдемісі); one could say that they are treated as NPs with an implied noun.
+\end{description}
+
+\subsubsection{English-to-Kazakh inter-chunk processing}
+
+The second round of structural transfer (the “interchunk” rules written in the apertium-eng-kaz.eng-kaz.t2x file) is currently performed by a proof-of-concept set of 90 rules, representative of following operations:
+\begin{itemize}
+\item Inter-chunk agreement between subject noun phrase and verb phrase in their person and number.
+\item Assigning case to noun phrases (which are generated without case by the chunker): for instance, accusative case for objects (I bought the table → Мен үстелді  сатып алдым), genitive case for obligatory constructs (I must see → менің көруім керек),  dative case for the verb to need (I need a doctor →Маған дәрігер керек), locative case for possession (I have a book → Менде кітап бар), etc.
+\item Reordering: placing of object before verb ( I[1] bought[2] the table[3] → Мен[1] үстелді[3] сатып алдым[2]),placing of prepositional pharses before the verb  (They[1] played[2] on top of the tree [3] → Олар[1] ағаштың үстінде[3] ойнады [2]), etc.
+\item Adding question particle ма/ме/ба/бе in the end of question, if question mark chunk “?” detected (Did<VP_ques> you<NP> watch<VP> last film<NP> ?<Q_m>→ Сіз<NP> соңғы фильмді<NP> көрдіңіз<VP> бе<ques> ?<Q_m>).
+\item Place possessive for long noun + noun + noun structures(The university of city of Kazakhstan - Қазақстан  қаласының университеті).
+\end{itemize}
+The set of rules has to be extended, as many combinations of the above phenomena are still not covered.
+
+\subsubsection{Postchunk}
+
+Additional cleaning stage was created for removing extra tags which did not determine through the chunker and interchunk steps(rules are written in  apertium-eng-kaz.eng-kaz.t4x file), for instance, noun was not act as object and did not get accusative case, so not determined tag was 'CD' after “cleaning” it will be set on nominative case; at this step, also carried out agreement between noun or adjective and copula “е”(Мен дәрігер+е<p1 person singualr>), vowel and consonants agreement between last word of question and question particle(Сіз келдіңіз + ме?
+- Сіз келдіңіз + бе?).
 
@@
 
+Current system can translate simple sentences and questions. Bilingual English-kazakh dictionary contains of  13135 entries, “chunker” rules are 168 and interchunk rules are 148. 
+Evaluating was tested against  revision 58998 in the Apertium SVN1.
+Lexical coverage is calculated over EuroParl1, SETimes2, NewsCommentary3, Wikipedia4(last check 04/06/2014).
+As can be seen from  Table 5, the higher coverage is of :
+
+\begin{table}
+  \begin{tabular}{|l|r|}
+    \hline
+    \textbf{Corpus} & \textbf{Coverage} (\%) \\
+    \hline
+    Wikipedia & 84.67 \\
+    EuroParl & 97.95 \\
+    NewsCommentary & 96.27 \\
+    SETimes & 97.90 \\
+    \hline
+  \end{tabular}
+\end{table}
+
 \section{Evaluation}
+
+\begin{table}
+  \begin{tabular}{|l|r|r|}
+    \hline
+    \textbf{System} & \textbf{BLEU} & \textbf{WER} \\
+    \hline
+    \texttt{apertium-eng-kaz} & 44.23 & 42.88 \\
+    Google & 62.00 & 25.92 \\
+    Sanasoft & 3.09 & 99.72 \\
+    Trident & 5.06 & 92.86 \\
+    \hline
+  \end{tabular}
+\end{table}
+
+Evaluating the output of the system was measured with BLEU (Papineni et al.
+2002) and Word Error Rate metric(Levenstein, 1966).
+Was chosen text small English text, and by postediting output of English-kazakh system, was build parallel text.
 
@@
 \com{Compare with GT + other systems -FMT}
+
+\section{Conclusions}
+
+The current English-Kazakh machine translation system already successfully translates noun-phrases, verb-phrases, prepositional-phrases, and adjectival-phrases, and contains a good vocabulary for testing purposes.
 

### 2015-02-26 engkaz-eamt-2015/engkaz-eamt-2015.tex
some fixes to formatting
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59141 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{multirow}
+
+\usepackage{polyglossia}
+\setdefaultlanguage[variant=australian]{english}
+\usepackage{latexsym}
+%\usepackage[small,bf]{caption}
+\usepackage[bf]{caption}
+\usepackage{xltxtra}
+\usepackage{times}
+\usepackage{fontspec}
+\defaultfontfeatures{Scale=MatchLowercase,Mapping=tex-text}
+\usepackage{listings}
+
+\newfontfeature{IPA}{+mgrk}
+%\setromanfont[IPA]{FreeSerif}
+%\setromanfont[Scale=0.9]{Times New Roman}
+\setromanfont{Times New Roman}
+\newfontfamily\qipa[IPA,Scale=MatchLowercase]{FreeSerif}
+\newfontfamily\qipb[IPA,Scale=MatchLowercase]{Junicode}
+\setmonofont[Scale=0.7]{DejaVu Sans Mono}
+\newfontfamily\smallertt[Scale=0.55]{DejaVu Sans Mono}
+\newfontfamily\smallerrm[Scale=0.85]{Times New Roman}
+
+%\fontspec[FakeBold=2.5]{Times New Roman}
+%\DeclareFontShape{EU1}{TimesNewRoman(0)}{m}{sc}{<->ssub * TimesNewRoman(1)/m/sc}{}
 
@@
 On the other hand, there are many differences between the syntax of Turkic languages and English:.subject-adverb–object–verb order (compare subject–verb–object-adverb in English), use of postpositions (compare prepositions in English), head-final syntax with modifiers and specifiers always preceding the modified/specified (normally following in English), overt case marking allowing for a rather free ordering of arguments (versus a more fixed order in English), lack of definite articles (extensively used in English), verbal-noun-centered structures where English uses modal verbs (must, have to, want to) or verbal-noun or verbal-adjective-centered constructions where English has subordinate clauses using finite verbs with relatives or subordinating conjunctions (the book which I read, the place where I saw him, before he came), lack of a parallel of the English verb have, as used for possession, etc.
-For an account (in Russian) of syntax differences between English and Kazakh, see Печерских & Амангельдина (2012).
+For an account (in Russian) of syntax differences between English and Kazakh, see Печерских \& Амангельдина (2012).
 This paper describes work in progress of development of machine translation system for English-to-Kazakh, which developed by using Apertium free/open-source machine translation platform (Forcada et al.
@@
 For example, the morphological analyser will deliver next sentence "I played" as showed below:
+\begin{verbatim}
 ^I/I<prn><subj><p1><mf><sg>$ 
 ^played/play<vblex><past>/play<vblex><pp>$
-
+\end{verbatim}
 In this example each source form has been analysed as one or more lexical forms: "I" is analysed into "I", where lexical category is subject pronoun(<prn><subj>) and first person(p1), could be masculine or feminine(mf), singular(sg); "played" is analysed into lexical verb(vblex) "play" with past simple tense(past) or it could be past participle(pp), so it has two analysis.
-Each analysis of word are separated by "^" and "" symbols, and for one word each lexical forms are delimited by "/", and tags(<...>) show grammatical attributes of lexical form.
+Each analysis of word are separated by \texttt{\^{}} and \texttt{\$} symbols, and for one word each lexical forms are delimited by "/", and tags(<...>) show grammatical attributes of lexical form.
  For Kazakh language finite state transducer based on two-level rules (in the case of Kazakh, apertium-kaz.kaz.lexc, apertium-kaz.kaz.twol).
@@
 For example, consider the morphological analysis of word "book":
+\begin{verbatim}
 ^book/book<n><sg>
 /book<vblex><inf>
-/book<vblex><pres>
+/book<vblex><pres>$
+\end{verbatim}
 This word has 3 lexical forms, and depends on context, ont of them will be chosen by rules or tagger. After  this module, all words have only one morphological analysis.
@@
 For English-Kazakh language pair, each lexical form of English word is translated into Kazakh as follows:
-^I<prn><subj><p1><mf><sg> /мен<prn><pers><subj><p1><mf><sg>$ 
+\begin{verbatim}
+^I<prn><subj><p1><mf><sg>/мен<prn><pers><subj><p1><mf><sg>$ 
 ^play<vblex><past>/ойна<v><tv><past>$
+\end{verbatim}
  Multiword units are translated as a single word.
@@
 Transfer rules can produce new sequences for the  target language, for instance, preposition-noun rule is used to built sequence, where for noun is chosen right case, which depends on preposition:
-in garden  -  ^бақша<n><sg><PXD><loc
-
+\begin{verbatim}
+in garden  -  ^бақша<n><sg><PXD><loc>$
+ \end{verbatim} 
 \item	 Morphological generator.
@@
 \item Reformatter. It places format tags back into the text so that its format is preserved.
+\end{itemize}
 
@@
 \begin{table*}
+ \centering
  \begin{tabular}{|l|l|l|l|}
     \hline
-    \textbf{Word} & \text{Translation} & \textbf{Condition} & \textbf{Example} \\
+    \textbf{Word} & \textbf{Translation} & \textbf{Condition} & \textbf{Example} \\
     \hline
@@
                                & резиденция & \_ of \_ president & I see the residence of the president -- (Мен) президенттің резиденциясын көремін \\
+    \hline
+    \multirow{2}{*}{boot} & бәтеңке & -- & I bought boots. -- Мен бәтеңкелерді сатып алдым \\
+                          & жүксалғыш & \_ of \_ car &  He opened the boot of the car  -- Ол машинаның жүксалғышын ашты. \\
+    \hline
+    \multirow{2}{*}{anything} & бір нәрсе & -- & I  see anything -- Мен бір нәрсе көремін \\
+                              & еш нәрсе & not \_[verb] & I do not do anything - Мен еш нәрсе істемеймін. \\
 
@@
  \end{tabular}
-
+  \caption{Example lexical selection rules}
 \end{table*}
@@
 \item Reordering: placing of object before verb ( I[1] bought[2] the table[3] → Мен[1] үстелді[3] сатып алдым[2]),placing of prepositional pharses before the verb  (They[1] played[2] on top of the tree [3] → Олар[1] ағаштың үстінде[3] ойнады [2]), etc.
-\item Adding question particle ма/ме/ба/бе in the end of question, if question mark chunk “?” detected (Did<VP_ques> you<NP> watch<VP> last film<NP> ?<Q_m>→ Сіз<NP> соңғы фильмді<NP> көрдіңіз<VP> бе<ques> ?<Q_m>).
+\item Adding question particle ма/ме/ба/бе in the end of question, if question mark chunk “?” detected (Did<VP\_ques> you<NP> watch<VP> last film<NP> ?<Q\_m> $\rightarrow$ Сіз<NP> соңғы фильмді<NP> көрдіңіз<VP> бе<ques> ?<Q\_m>).
 \item Place possessive for long noun + noun + noun structures(The university of city of Kazakhstan - Қазақстан  қаласының университеті).

### 2015-02-27 2015-eamt-assim/2015-eamt-assim.tex
Some intermediate corrections
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59153 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{color}
+\usepackage{graphicx}
 
@@
 
-\title{Evaluating machine translation for assimilation}
+\title{Evaluating machine translation for assimilation via a gap-filling task}
 
@@
 be divided into two broad categories according to their purpose: post-editing and assimilation systems.
+The output of the former is intended to be transformed into text comparable to human translation; the latter systems' goal is to enhance user's comprehension of text.
 Both kinds may be evaluated, either to control for quality in the development process or to compare the systems.
@@
 
-\comment{EA: I'd like a citation for the ``most massive application'', could you suggest something?}
-Despite the fact that assimilation (or gisting) is currently the most massive application of MT, few methodologies are established for assimilation evaluation of MT.
-The methods include post-editing and comparison by bilingual experts \citep{ginesti09}, and multiple choice tests \citep{jones07,trosterud12} These approaches are often costly and prone to subjectivity, see the discussion in \cite{oregan13}.
-As an alternative, the modification of Cloze test was introduced for assimilation evaluation, first in \citep{trosterud12} as a supplementary, and then in \citep{oregan13} as a stand-alone method.
-Prior to this, Cloze tests have been used to evaluate raw MT quality, see e.g.
-\citep{vanslype79}, \citep{somers00}.
-As a principal difference, \citet{trosterud12} and \citet{oregan13} propose to fill the gaps in the reference (human) translation.
+\comment{EA: I'd like a citation for the ``most massive application'', could you suggest something? It is from Mikel's paper, but no reference there -- should I cite Mikel?}
+Despite the fact that assimilation (or gisting) is currently the most frequent application of MT, few methodologies are established for assimilation evaluation of MT.
+The methods include post-editing and comparison by bilingual experts \citep{ginesti09}, and multiple choice tests \citep{jones07,trosterud12}.
+These approaches are often costly and prone to subjectivity, see the discussion in \cite{oregan13}.
+As an alternative, the modification of Cloze \citep{taylor53} test was introduced for assimilation evaluation, first by \citep{trosterud12} as a supplementary, and then by \citep{oregan13} as a stand-alone method.
+Prior to this, Cloze tests have been used to evaluate raw MT quality \citep{vanslype79}, \citep{somers00}.
+As a principal difference, \citet{trosterud12} and \citet{oregan13} propose to fill gaps in the reference (human) translation.
+The gap-filling task models how well users comprehend the key points of the text.
+Thus, the method does not directly evaluate the quality of machine-produced text, but rather its usefulness in understanding the meaning of the original text.
 
+\comment{EA: I understand we selected tat-rus and eng-kaz because they were in development. why did we select eu-es?}
 The gap-filling method has been successfully used to evaluate the Basque-English Apertium language pair.
-In this work we extend the evaluation to three more language pairs: Basque-Spanish, English-Kazakh and Tatar-Russian.
-For the latter two pairs, the evaluation served as a quality check in the period of active development during the Google Summer of Code'14 programme.
+In this work we extend the evaluation to two more language pairs: Basque-Spanish and Tatar-Russian.
+For the latter pair, the evaluation served as a quality check in the period of active development during the Google Summer of Code'14 programme.
 In addition to evaluating, we explore new aspects of the experiment: the correlation between evaluators' scores, and the influence of the texts' linguistic domain.
-To facilitate the evaluation, we introduce an automated task creation system.
+To facilitate the evaluation, we introduce an automated system which creates task sets from parallel corpora given a range of parameters (number of gaps in a sentence, hint type, gap filler, etc.), checks evaluators' answers, and calculates and formats generalized results.
 This system is integrated into the Appraise MT evaluation platform \citep{federmann12}; the code is open source and is available on Github\footnote{https://github.com/Sereni/Appraise}.
 
-We anticipate that MT systems under investigation will contribute to the users' understanding of text.
-We also expect to see different results depending on text domain and on the number of words left out from the task.
+We anticipate that the assessed MT systems will contribute to the users' understanding of text, that is, the evaluation will show better results in MT-assisted task type.
+We also expect to see different results depending on such parameters as text domain and the number of gaps in a sentence in relation to sentence length.
 
@@
 In section~\ref{sec:results} we describe and discuss the experiment results.
-Finally, section~\ref{sec:conclusion} draws conclusions.
+Finally, section~\ref{sec:conclusion} draws some conclusions.
 
@@
 \label{sec:methodology}
-\comment{FMT: This could benefit from an example i think, something like table1? }
 
@@
 
+\begin{table*}
+  \begin{tabular}{|l|l|}
+     \hline
+     \textbf{10\%}   & Ayudas econ\'{o}micas para el tratamiento de toxicoman\'{i}­as en comunidades terap\'{e}uticas no concertadas. \\
+     \textbf{20\%}   & Ayudas econ\'{o}micas para el \{ \} de toxicoman\'{i}­as en comunidades terap\'{e}uticas no concertadas. \\
+     \textbf{30\%} & Komunitate terapeutiko itundu gabeetan toxikomaniak tratatzeko diru-laguntzak ematea. \\
+     \hline
+  \end{tabular}
+  \caption{Example of different gap percentage settings.} 
+  \label{table:percentage}
+\end{table*}
+
 This section discusses the reasoning behind the gap-filling method and task structure. The gap-filling method 
-of evaluating machine translation for assimilation purposes is based on the following notion: a reader's understanding 
-of a given text correlates with the number of keywords they are able to correctly restore in the text. Therefore, the 
-base of an assimilation task is a (reference) sentence, where some of the keywords are blacked out, or removed. The sentence 
+of evaluating machine translation for assimilation purposes is based on the following hypothesis: a reader's understanding 
+of a given text correlates with the number of words they are able to correctly restore in the text. Therefore, the 
+base of an assimilation task is a (reference) sentence, where some of the words are blacked out, or removed. The sentence \comment{EA: I'm sorry, I'm not sure what else I could explain here...}
 is produced by a human (as opposed to machine-translated), and it is in the language known to evaluators, which is also 
 the \emph{target language} of the machine translation system.
-\comment{FMT: It would also be nice to have a screenshot here}
 The additional elements of the task are what we call hints, or extra sentences that help the participant to understand 
-the main sentence. There are two types of hints: first, the \emph{source}, which is the same sentence as \emph{reference}, 
+the main sentence. There are two types of hints: first, the \emph{source}, which is semantically identical to \emph{reference}, 
 also human-produced, but in the source language of the pair. The second type is the \emph{machine-translated} hint, which 
-comes from the machine translation of the \emph{source } sentence.
+comes from the machine translation of the \emph{source} sentence. Table~\ref{table:example} shows a sample task, and Figure~\ref{figure:screenshot} shows the task in the online evaluation environment.
+
+\comment{EA: any way to fix placement of the screenshot?}
+\begin{figure*}
+  \centering
+\includegraphics[resolution=144]{appraise-scr}
+\label{figure:screenshot}
+ \caption{An example set of sentences in an online environment.}
+\end{figure*}
 
@@
 \begin{description}
-\item[Reference sentence only:] The participants are asked to fill the gaps not being given
-any context. This task serves as a baseline score and as an indicator of phrases
+\item[Reference sentence only:] The participants are asked to fill the gaps without being given
+any context. This task serves as a baseline score and as an indicator of gaps
 that can be completed using common knowledge or language intuition (e.g.
@@
 \item[Reference sentence and MT hint:] In addition to the task, the
-participants see the source sentence translated via Apertium. This type of task is
+participants see the source sentence translated via the MT system Apertium \citep{forcada11}. This type of task is
 used for measuring the contribution of machine translation to understanding the
@@
 
-For each sentence, the list of candidate keywords is prepared. It constitutes all the words that 
+For each sentence, the list of candidate keywords is prepared. It is composed of all the words that 
 fall into the allowed POS list. The number of gaps in the sentence is calculated based on 
@@
 
-This work has been partly funded by the Spanish Ministerio de Economía y Competitividad through project TIN2012-32615 and by the Google Summer of Code programme.
+This work has been partly funded by the Spanish Ministerio de Econom{\'i}a y Competitividad through project TIN2012-32615 and by the Google Summer of Code programme.
 We would like to thank the volunteers who participated in the evaluations.

### 2015-02-28 2015-eamt-assim/2015-eamt-assim.tex
Second draft
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59161 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 This paper provides additional observations on the viability of a strategy proposed in 2012 for evaluation of machine translation for assimilation purposes.
-The evaluation method is applied to three translation directions.
-To reduce preparation time, an automatic open-source task management system is introduced.
+The strategy involves human evaluators, who are asked to restore key words (fill gaps) in sentences.
+The evaluation method is applied to two translation directions.
+To reduce the amount of time required to prepare tasks and analyze results, an open-source task management system is introduced.
 The evaluation results suggest that the gap-filling task reflects users' understanding of text, and may be used to measure MT quality for assimilation purposes.
@@
 To facilitate the evaluation, we introduce an automated system which creates task sets from parallel corpora given a range of parameters (number of gaps in a sentence, hint type, gap filler, etc.), checks evaluators' answers, and calculates and formats generalized results.
-This system is integrated into the Appraise MT evaluation platform \citep{federmann12}; the code is open source and is available on Github\footnote{https://github.com/Sereni/Appraise}.
+This system is integrated into the Appraise MT evaluation platform \citep{federmann12}; the code is open source and is available on Github\footnote{}.
 
@@
   \end{tabular}
-  \caption{An example of a set of sentences.} 
+  \caption{An example group of sentences showing the gapped sentence and hint types. Reference, MT and task sentences are in Spanish, the source sentence is in Basque.} 
   \label{table:example}
@@
 
-\begin{table*}
-  \begin{tabular}{|l|l|}
-     \hline
-     \textbf{10\%}   & Ayudas econ\'{o}micas para el tratamiento de toxicoman\'{i}­as en comunidades terap\'{e}uticas no concertadas. \\
-     \textbf{20\%}   & Ayudas econ\'{o}micas para el \{ \} de toxicoman\'{i}­as en comunidades terap\'{e}uticas no concertadas. \\
-     \textbf{30\%} & Komunitate terapeutiko itundu gabeetan toxikomaniak tratatzeko diru-laguntzak ematea. \\
-     \hline
-  \end{tabular}
-  \caption{Example of different gap percentage settings.} 
-  \label{table:percentage}
-\end{table*}
+\comment{EA: Francis, when I save cyrillic in article body, it gets replaced by ??. I committed a separate file with the table and the footnote that has the text.}
+
+%\begin{table*}
+%  \begin{tabular}{|l|l|}
+%     \hline
+%     \textbf{10\%}   & ???????? ???????; ??? ????? { } ????? 7 ?????????, ????? ?????? ??? ????? 100 ??????. \\
+%     \textbf{20\%}   & ???????? ???????; ??? ????? { } ????? 7 ?????????, ????? { } ??? ????? 100 { }. \\
+%     \textbf{30\%} & ???????? ???????; ??? ????? { } ????? 7 { }, ????? ?????? { } ????? 100 { }. \\
+%     \hline
+%  \end{tabular}
+%  \caption{Example of different gap percentage settings.} 
+%  \label{table:percentage}
+%\end{table*}
 
@@
 also human-produced, but in the source language of the pair. The second type is the \emph{machine-translated} hint, which 
-comes from the machine translation of the \emph{source} sentence. Table~\ref{table:example} shows a sample task, and Figure~\ref{figure:screenshot} shows the task in the online evaluation environment.
+comes from the machine translation of the \emph{source} sentence.
+Table~\ref{table:example} shows a sample task, and Figure~\ref{figure:screenshot} shows the task in the online evaluation environment.\comment{EA: for some reason, the reference to screenshot is figure 2, while in the caption it says figure 1 (correctly)}
 
-\comment{EA: any way to fix placement of the screenshot?}
 \begin{figure*}
@@
 from the candidate list in such a manner that the gaps are distributed evenly throughout the sentence.
+We start at a random word in a sentence and check whether it is a keyword.
+If yes, we remove it, and move \emph{n} words forward, going back to the beginning of sentence if necessary.
+The step length \emph{n} is the sentence length divided by the desired number of gaps.
+If the word is not a keyword, or has already been removed, we look at the next word instead.
+The process is repeated until the designated number of words has been removed, or until there are no more words in the keyword list.
 
 Keyword removal is one of the most time-consuming steps in task preparation.
+It normally requires human effort, because we would like to determine the words that contribute the most to understanding the text as opposed to removing random words.
 In our setup, the above procedure is performed by a script integrated into the task generation pipeline.
@@
 To control for homonymy, we only allow the word into the candidate list if all of its possible part of speech attributions are on the POS list.
+For example, if we only allow nouns on the word list, and the word "fly" receives two possible part of speech attributions from the tagger, noun and verb, it is not considered for the candidate list.
 
@@
 
-For each experiment we called for native speakers of L2 of the language pair (i.e.
-Spanish, Kazakh, and Russian) who had no command of L1 of the pair (Basque, English
-and Tatar, respectively). 11 evaluators participated in the in Basque-Spanish experiment, 8 in
-English-Kazakh, and 28 in Tatar-Russian (although not everyone has completed their task
+For each experiment we called for native speakers of target language of the language pair (i.e.
+Spanish and Russian) who had no command of source language of the pair (Basque
+and Tatar, respectively). 11 evaluators participated in the in Basque-Spanish experiment, and 28 in Tatar-Russian (although not everyone completed the task
 in full, see discussion).
@@
 \begin{enumerate}
-\item  For Basque-Spanish, from the corpus of legal texts Memorias de traducci\'on del
-Servicio Oficial de Traductores del IVAP\footnote{http://opendata.euskadi.net/memorias-de-traduccion-del-servicio-oficial-de-traductores-del-ivap/w79-contdata/es/};
-\item  For English-Kazakh, from the official website of the President of the Republic of
-Kazakhstan\footnote{http://www.akorda.kz/};
+\item  For Basque-Spanish, from the corpus of legal texts ``Memorias de traducci\'on del
+Servicio Oficial de Traductores del IVAP''\footnote{http://opendata.euskadi.net/memorias-de-traduccion-del-servicio-oficial-de-traductores-del-ivap/w79-contdata/es/};
 \item  For Tatar-Russian, from the following sources on three different topics:
@@
 \comment{EA: whoops, what is the best way to make it understand cyrillic?}
-    \item  Casual conversations, from a textbook\footnote{FIXME, 1994. ? 320 ?. ISBN 5?298?00463?6 (???. 219, 220, 232, 233, 234)} of spoken Tatar;
+    \item  Casual conversations, from a textbook
+    %\footnote{???????? ?.?. ? ??????? ???????? ??-????????. ??????: ????????? ??. ???-??, 1994. ? 320 ?. ISBN 5?298?00463?6 (???. 219, 220, 232, 233, 234)} of spoken Tatar;
     \item  Legal texts, from the Constitution and laws\footnote{http://tatarstan.ru} of Tatarstan;
@@
 
-Each set features 36 pairs of sentences. For the first two experiments the pairs are drawn
+Each set features 36 pairs of sentences. For the Basque-Spanish experiment the pairs were drawn
 randomly from the corpora, for Tatar-Russian, compiled by hand by the developer of the
-language pair.
+language pair in Apertium.
 
@@
 
-The evaluations take place online, in a system called Appraise \citep{federmann12}, which is 
+The evaluations took place online, in a system called Appraise \citep{federmann12}, which is 
 designed specifically for various MT evaluation tasks. We adapted the code to 
-accommodate for the gap-filling tasks. The tasks are uploaded into the system and 
+accommodate for the gap-filling tasks. The tasks were uploaded into the system and 
 manually distributed between the participants by the following rules:
@@
 \item  All sentences of the set are evaluated with 10, 20 and 30\% of words removed;
+\item Each participant may encounter a given sentence in only one of the percentage variations;
 \item  Each sentence-mode-percentage combination is evaluated by more than one participant.
@@
 
-The participants are given the instructions in their native language above each task. The
-tasks are split into smaller sets for the participants' convenience. The instructions are the 
+The participants are given the instructions in their native language; these instructions are repeated above each task in the evaluation system.
+For the participants' convenience, the body of questions is split into smaller groups which allow indefinite breaks between them.
+The instructions are the
 following: read all the available hints and fill each gap with one suitable word, guessing if unsure.
@@
 considered a candidate synonym if it is given by two or more evaluators, and it does not match
-the answer key. We record each candidate synonym along with the answer key and the
-context sentence. Based on this data, a native speaker of the target language decides
-whether the candidate synonym is indeed synonymic to the answer key in the given
+the answer word. We record each candidate synonym along with the answer key and the
+context sentence (see Table~\ref{table:syn} for an example). Based on this data, a native speaker of the target language decides
+whether the candidate synonym is an acceptable replacement to the answer key in the given
 context. We then check participants' results against the compiled synonym list and
-increase scores where appropriate. On average, the scores improved by three percent in all evaluation modes. Candidate synonym extraction and score update is performed automatically.
+increase scores where appropriate.
+On average, the scores improved by three percent in all evaluation modes.
+Candidate synonyms are extracted automatically from the evaluators' responses,  and each individual score is automatically updated according to the synonym list.
 
@@
 \comment{EA: added}
-The synonym lists for Basque-Spanish, English-Kazakh and Tatar-Russian contain 52, 38 and 25 words, respectively.
+The synonym lists for Basque-Spanish and Tatar-Russian contain 52 and 25 words, respectively.
 Time taken to compile one list depends on the number of candidate synonyms, and in our case was approximately 30 minutes.
@@
 
-The experiment results are presented in the tables below. Tables \ref{table:res-eus-spa}, \ref{table:res-eng-kaz} and \ref{table:res-tat-rus} show
-the proportion of correct answers depending on evaluation mode and gap density.\comment{FMT: We should probably integrate the other tables, and refer to them using labels.} The
-percentage is spread across all evaluators, that is, we divide the number of correct
-answers given by all evaluators by the number of questions answered by all evaluators.
-Tables \ref{table:time-eus-spa}, \ref{table:time-eng-kaz} and \ref{table:time-tat-rus} show average time it took to fill the gaps in one sentence. To
-reduce the noise from participants who were distracted during evaluation, when
+The results are presented in this section. Tables \ref{table:res-eus-spa} and \ref{table:res-tat-rus} show
+the proportion and standard deviation of correct answers depending on evaluation mode and gap density.\comment{FMT: We should probably integrate the other tables, and refer to them using labels.} The evaluators' correct answer percentage is averaged over the number of evaluators.
+In addition to the percentage of correct answers we kept a record of time taken to fill the gaps in one sentence.
+To reduce the noise from participants who were distracted during evaluation, when
 calculating times we remove all the results over 6 minutes (the mode is approximately
-two minutes).
+two minutes). Typical time taken to complete one question varies from under one minute for tasks without hints and few gaps, to approximately two minutes for tasks with more hints and gaps.
 
@@
 % a few notes on time
-For Basque--Spanish and Tatar--Russian, the latter trend holds: the average time taken to fill the gaps increases and the average percentage of correct answers decreases as the relative number of gaps goes up.
+The latter trend holds: the average time taken to fill the gaps increases and the average percentage of correct answers decreases as the relative number of gaps goes up.
 The larger number of gaps in the sentence makes it more
@@
 We would have expected longer completion time, since the participant must come up with their own answer unassisted.
-However, in the no-hint task the participant reads only one sentence, as opposed to two or three in other tasks.
+However, in the no-hint task the participant is required to read only one (reference) sentence, as opposed to two or three (reference and hints) in other tasks.
 Also, the number of gaps in 10\%-gap tasks is low, as it never exceeds three.
@@
 This aligns well with our expectations and suggests than machine translation helps to understand the provided text.
-In addition, tasks with source hint are completed better than tasks without hints, and the same relation holds between MT+source and MT-only types of tasks.
+In addition, tasks with source hints are completed better than tasks without hints, and the same relation holds between MT+source and MT-only types of tasks.
 
@@
 However, the latter percentage increases significantly after we extend the answer list with synonyms.
-Such an increase suggests that, in this case, the content words were restored by sense rather than through strong collocation.
+Such an increase suggests that, in this case, the content words were restored by
+\comment{EA: to JA: I am not sure if this qualifies as common sense.
+What I meant is that people correctly guessed the meaning of the word in place, and expressed it in their own manner.
+Changed it to semantic context, is this a better rephrasing?}
+semantic context rather than through strong collocation.
 The second pattern, low scores in Tatar--Russian 10\% MT+source, does not stem from the task content.
@@
 
-It remains questionable whether we can compare results for different gap densities. In the
-Basque-Spanish and Tatar-Russian experiments the 10, 20 and 30\% sets were comprised
+It remains questionable whether we can compare results for different gap densities. The 10, 20 and 30\% sets were comprised
 of the same sentences. However, in each case different words were removed. It appears
@@
 Basque-Spanish the 20\% MT tasks are completed with better accuracy than 10\% tasks.
-
-% a little on categories
-\comment{EA: and what would we like to say about those?
-The results depend heavily on the Tatar-Russian system.
-Do we have to provide info on what vocabulary it handles better, what it has and has not been trained on?
-Or do we just say that casual texts are easier to fill than legal texts?}
-For Tatar--Russian language pair the participants were offered texts from three different domains (in equal proportions): casual conversations, legal texts and news.
-The results by domains are displayed in tables \ref{table:rus-casual}, \ref{table:rus-legal}, and \ref{table:rus-news}.
-The system used in the evaluation has been targeted to translate texts from all three of the domains.
-Taking into consideration the above discussion of 10\% MT+Source tasks, we observe similar results across the three categories.
-Note that, in casual and news texts, the source sentences paired with MT significantly improve participants' performance compared to MT-only task mode.
-This may be due to the fact that many words are borrowed from Russian into Tatar, and are in fact understood by Russian speakers.
-
-% this section talks about eng-kaz, where everything is not so nice 
-\comment{EA: considering all those flaws in setup, should we report Kazakh results at all?}
-In the above discussion of results we do not take into account the data from the English--Kazakh language pair.
-It remains questionable whether we may draw reliable conclusions from this part of the experiment.
-By design, the evaluation tasks are distributed evenly among the participants, in such a manner that each sentence is evaluated in all modes or gap densities by several people.
-The English--Kazakh evaluations have been set up for 11 participants, however, only 8 participants attempted their tasks, and 6 of them completed it in full.
-As a result, the evaluation became unbalanced: out of 4 sets of sentences, three have been evaluated by three, two and one participant, respectively, and the fourth set has been partially completed by only one participant.
-A total of 40 (about 10\%) gaps have not been filled by any of the participants.
-
-We should also note the outlier score of 60\% for the 20\% no-hint task.
-It is made up by three sentences, two of which scored unusually high.
-One of these sentences is a news headline, and the other contains clich\'es such as ``express condolences''.
-We would have expected the same sentences to contribute to percentages in other evaluation modes, but they have not been completed there.
-
-Finally, we would like to comment on the relatively higher scores on tasks with source hints (columns 2 and 4 of Table \ref{table:res-eng-kaz}.
-By experiment design, the participants must have no command of the source language of the pair, which is, in this case, English.
-In Kazakhstan, English is a mandatory part of the middle school curriculum, and the students start learning it at the age of 10.
-It therefore seems likely that our participants, aged 23-26, had at least some command of English, which may have contributed to their understanding of source-hint tasks.
 
@@
 
-% results with synonyms
-\begin{table}
+\subsection{Annotator agreement}
+After obtaining the results we calculated Krippendorff's alpha measure to represent annotator agreement, shown in Tables~\ref{table:alpha-eus-spa} and~\ref{table:alpha-tat-rus}.
+We selected this measure because of its compatibility with more than two annotators per task and the missing data (not all the gaps were evaluated).
+To calculate Krippendorff's alpha we used an algorithm implementation by Thomas Grill\footnote{http://grrrr.org}, dividing the answers in each gap into two categories: correct and incorrect.
+The previously obtained synonym lists were taken into account, i.e.
+if the two answers are different but both correct, they fall into one category.
+The measure was calculated separately for each hint and percentage combination.
+
+As the MT and MT+source hints are introduced, the agreement between annotators increases (measures closer to 1): the annotators are more consistently correct or incorrect in each given sentence.\comment{EA: I cite Juan Antonio here, because I am not completely sure what alpha shows about our experiment (lack of knowledge on my part)}  The agreement measure for the same sentences without hints is closer to zero, which contributes to the reliability of our methodology.
+We note the outlier score in Tatar--Russian 10\% source tasks, which has the most contribution from the news texts (agreement of 0.917).
+This set of sentences contains many loan words, which have similar form in Tatar and Russian (e.g.
+president, minister, championship), and are understood by Russian speakers.
+The gaps with loan words have mostly been filled correctly, while there was some disagreement in other gaps.
+
+\begin{table*}
+\centering
   \begin{tabular}{|l|r|r|r|r|}
@@
                                             & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
-    10\%                                    &   58.02            & 55.70       & 35.06        & 45.45        \\
-    20\%                                    &   65.96            & 68.42       & 33.85        & 33.56        \\
-    30\%                                    &   49.84            & 37.28       & 24.23        & 18.93        \\
+\hline
+10\%&0.496&0.517&0.400&0.124\\
+20\%&0.714&0.700&0.358&0.275\\
+30\%&0.559&0.430&0.406&0.300\\
+
 \hline
 \end{tabular}
-  \caption{Basque--Spanish: Gaps successfully filled (\%), using a synonym list}
-  \label{table:res-eus-spa}
-\end{table}
+\caption {Tatar--Russian: Krippendorff Alpha measure} 
+\label{table:alpha-eus-spa}
+\end{table*}
 
-\begin{table}
+\begin{table*}
+\centering
   \begin{tabular}{|l|r|r|r|r|}
@@
                                             & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
-    10\%                                    & 34.78              & 39.02       & 40.00        & 31.43        \\
-    20\%                                    & 43.84              & 33.80       & 47.83        & 60.34        \\
-    30\%                                    & 41.38              & 36.56       & 44.94        & 31.09        \\
+\hline
+10\%&0.598&0.459 &0.711 &0.517 \\
+20\%&0.740&0.667 &0.473 &0.261 \\
+30\%&0.534&0.581 &0.411 &0.412 \\
 \hline
 \end{tabular}
-  \caption{English--Kazakh: Gaps successfully filled (\%), using a synonym list}
-  \label{table:res-eng-kaz}
-\end{table}
+\caption {Tatar--Russian: Krippendorff Alpha measure} 
+\label{table:alpha-tat-rus}
+\end{table*}
 
-\begin{table}
+% a little on categories
+\subsection{Results for different domains}
+\comment{EA: and what would we like to say about those?
+The results depend heavily on the Tatar-Russian system.
+Do we have to provide info on what vocabulary it handles better, what it has and has not been trained on?
+Or do we just say that casual texts are easier to fill than legal texts?}
+For Tatar--Russian language pair the participants were offered texts from three different domains (in equal proportions): casual conversations, legal texts and news.
+The results by domains are displayed in tables \ref{table:rus-casual}, \ref{table:rus-legal}, and \ref{table:rus-news}.
+The MT system used in the evaluation has been targeted to translate texts from all three of the domains.
+Taking into consideration the above discussion of 10\% MT+Source tasks, we observe similar results across the three categories.
+Note that, in casual and news texts, the source sentences paired with MT significantly improve participants' performance compared to MT-only task mode.
+This may be due to the fact that many words are borrowed from Russian into Tatar, and are in fact understood by Russian speakers.
+
+% results with synonyms
+
+\begin{table*}
+\centering
   \begin{tabular}{|l|r|r|r|r|}
@@
                                             & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
-    10\%                                    & 53.89              & 66.06       & 54.43        & 46.34        \\
-    20\%                                    & 65.33              & 60.73       & 48.39        & 39.43        \\
-    30\%                                    & 59.90              & 52.59       & 39.66        & 38.39        \\
+\hline
+10\%&62.19 +/- 32.07&58.04 +/- 27.96&39.90 +/- 38.96&48.79 +/- 39.84\\
+20\%&64.81 +/- 30.03&70.18 +/- 26.98&31.37 +/- 28.42&30.91 +/- 30.22\\
+30\%&48.13 +/- 26.08&39.55 +/- 23.90&25.94 +/- 19.67&18.07 +/- 18.15\\
 \hline
 \end{tabular}
-  \caption{Tatar--Russian: Gaps successfully filled (\%), using a synonym list}
-  \label{table:res-tat-rus}
-\end{table}
-
-% begin time tables
-\comment{EA: maybe remove SD from time tables? is it informative?}
-\begin{table*}
-\begin{tabular}{|r |*{4}{c}|}
-\hline
- &\multicolumn{4}{c|}{\textbf{Evaluation mode}}\\
-\hline
-\textbf{Gap percentage} & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint}\\
-10\%&1:17 +/- 1:05&1:33 +/- 1:22&1:06 +/- 0:38&0:46 +/- 0:44\\
-20\%&1:23 +/- 1:04&1:14 +/- 0:58&1:27 +/- 1:02&1:22 +/- 1:17\\
-30\%&2:22 +/- 1:10&2:33 +/- 1:26&2:36 +/- 1:22&2:25 +/- 1:30\\
-\hline
-\end{tabular}
-\caption {Basque--Spanish: Time taken to fill the gaps in one sentence, mean and standard deviation.}
-\label{table:time-eus-spa} 
+\caption {Basque--Spanish: Average number of gaps successfully filled (\%), using a synonym list} \label{table:res-eus-spa}
 \end{table*}
@@
 \begin{table*}
-\begin{tabular}{|r |*{4}{c}|}
+\centering
+  \begin{tabular}{|l|r|r|r|r|}
     \hline
-  &\multicolumn{4}{c|}{\textbf{Evaluation mode}}\\
+    \multirow{2}{*}{\textbf{Density}} & \multicolumn{4}{|c|}{Evaluation mode} \\\cline{2-5}
+                                            & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
 \hline
-\textbf{Gap percentage} & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
-10\%&2:31 +/- 1:29&1:34 +/- 1:03&1:47 +/- 1:36&0:47 +/- 0:35\\
-20\%&1:17 +/- 0:51&2:05 +/- 1:57&2:10 +/- 2:14&1:13 +/- 1:02\\
-30\%&2:00 +/- 1:30&1:46 +/- 1:48&1:41 +/- 1:02&1:44 +/- 1:34\\
+10\%&63.76 +/- 41.15&57.02 +/- 41.57&53.52 +/- 43.57&45.96 +/- 40.86\\
+20\%&59.77 +/- 33.10&64.93 +/- 30.95&45.60 +/- 31.23&38.56 +/- 32.11\\
+30\%&55.83 +/- 26.13&58.77 +/- 28.07&40.00 +/- 28.16&35.45 +/- 29.76\\
     \hline
   \end{tabular}
-\caption {English--Kazakh: Time taken to fill the gaps in one sentence, mean and standard deviation.}
-\label{table:time-eng-kaz} 
-\end{table*}
-
-\begin{table*}
-\begin{tabular}{|r |*{4}{c}|}
-\hline
-  &\multicolumn{4}{c|}{\textbf{Evaluation mode}}\\
-\hline
-\textbf{Gap percentage} & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
-10\%&1:26 +/- 1:10&1:05 +/- 1:05&0:45 +/- 0:52&0:31 +/- 0:22\\
-20\%&1:16 +/- 2:06&1:12 +/- 1:44&0:57 +/- 0:59&0:50 +/- 0:55\\
-30\%&1:19 +/- 1:15&1:14 +/- 1:11&1:10 +/- 1:04&1:01 +/- 0:53\\
-\hline
-\end{tabular}
-\caption {Tatar--Russian: Time taken to fill the gaps in one sentence, mean and standard deviation.}
-\label{table:time-tat-rus} 
+  \caption{Tatar--Russian: Average number of gaps successfully filled (\%), using a synonym list}
+  \label{table:res-tat-rus}
 \end{table*}
@@
 \textbf{Gap percentage} & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
-10\%&59.52&60.98&65.00&54.76\\
-20\%&73.68&63.16&44.30&37.97\\
-30\%&68.18&58.24&41.67&41.94\\
+10\%&63.64 +/- 44.84&63.64 +/- 43.12&62.12 +/- 46.11&53.03 +/- 44.28\\
+20\%&73.33 +/- 31.51&63.33 +/- 35.85&41.11 +/- 27.88&37.78 +/- 30.94\\
+30\%&70.28 +/- 31.12&59.58 +/- 23.89&39.44 +/- 27.10&37.71 +/- 33.32\\
 \hline
 \end{tabular}
-\caption {Tatar--Russian, casual conversations: Gaps successfully filled (\%), using a synonym list, average} \label{table:rus-casual} 
+\caption {Tatar--Russian, casual conversations: Average number of gaps successfully filled (\%), using a synonym list} \label{table:rus-casual} 
 \end{table*}
@@
 \textbf{Gap percentage} & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
-10\%&52.00&67.11&43.48&38.03\\
-20\%&61.74&63.57&52.46&43.31\\
-30\%&56.07&41.32&37.36&37.23\\
+10\%&53.02 +/- 40.31&68.21 +/- 34.99&38.77 +/- 37.82&32.95 +/- 34.88\\
+20\%&60.57 +/- 24.98&65.76 +/- 24.16&50.28 +/- 33.55&48.31 +/- 33.52\\
+30\%&50.06 +/- 26.25&48.30 +/- 28.87&40.05 +/- 27.37&34.15 +/- 28.73\\
 \hline
 \end{tabular}
-\caption {Tatar--Russian, legal texts: Gaps successfully filled (\%), using a synonym list, average} \label{table:rus-legal} 
+\caption {Tatar--Russian, legal texts: Average number of gaps successfully filled (\%), using a synonym list} \label{table:rus-legal} 
 \end{table*}
@@
 \textbf{Gap percentage} & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
-10\%&52.00&68.75&61.22&50.98\\
-20\%&64.00&52.86&46.15&34.25\\
-30\%&59.38&63.28&41.41&37.50\\
+10\%&53.33 +/- 37.86&59.77 +/- 43.66&57.47 +/- 42.12&49.44 +/- 39.08\\
+20\%&59.33 +/- 33.89&48.56 +/- 35.40&46.53 +/- 31.99&28.97 +/- 28.79\\
+30\%&58.14 +/- 22.42&61.49 +/- 22.11&40.51 +/- 30.10&34.81 +/- 27.01\\
 \hline
 \end{tabular}
-\caption {Tatar--Russian, news: Gaps successfully filled (\%), using a synonym list, average} \label{table:rus-news} 
+\caption {Tatar--Russian, news: Average number of gaps successfully filled (\%), using a synonym list} \label{table:rus-news} 
 \end{table*}
@@
 \label{sec:conclusion}
-We have conducted assimilation evaluation of three Apertium translation directions: Basque--Spanish, English--Kazakh and Tatar--Russian.
+We have conducted assimilation evaluation of two Apertium translation directions: Basque--Spanish and Tatar--Russian.
 The results suggest that this evaluation method reflects the contribution of MT to users' understanding of text.
-The version of the toolkit used in this experiment may be downloaded from our repository\footnote{https://github.com/Sereni/Appraise/tree/1e9d735faee64d1b97fb343ab111ace6a64509d7}.
+The version of the toolkit used in this
+\comment{EA to JA: by our repository I mean my github, is that ok?}
+experiment may be downloaded from our repository\footnote{}.
+
+In addition to the mentioned language pairs, some preliminary experiments were carried out for the English--Kazakh pair.
+It remains questionable whether we may draw reliable conclusions from this part of the experiment.
+By design, the evaluation tasks are distributed evenly among the participants, in such a manner that each sentence is evaluated in all modes or gap densities by several people.
+The English--Kazakh evaluations have been set up for 11 participants, however, only 8 participants attempted their tasks, and 6 of them completed it in full.
+As a result, the evaluation became unbalanced: out of 4 sets of sentences, three have been evaluated by three, two and one participant, respectively, and the fourth set has been partially completed by only one participant.
+A total of 40 (about 10\%) gaps have not been filled by any of the participants.
+Therefore, we do not report the English--Kazakh evaluation results here.
 

### 2015-02-28 2015-eamt-assim/2015-eamt-assim.tex
Added Krippendorff reference
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59164 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \subsection{Annotator agreement}
-After obtaining the results we calculated Krippendorff's alpha measure to represent annotator agreement, shown in Tables~\ref{table:alpha-eus-spa} and~\ref{table:alpha-tat-rus}.
+After obtaining the results we calculated Krippendorff's alpha~\citep{krippendorff70} measure to represent annotator agreement, shown in Tables~\ref{table:alpha-eus-spa} and~\ref{table:alpha-tat-rus}.
 We selected this measure because of its compatibility with more than two annotators per task and the missing data (not all the gaps were evaluated).

### 2015-03-02 2015-eamt-assim/2015-eamt-assim.tex
formatting
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59178 72bbbca6-d526-0410-a7d9-f06f51895060
fix cyrillic
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59179 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{eamt15}
+\setlength\titlebox{6.5cm}    % Expanding the titlebox
+%%% YOUR PACKAGES BELOW THIS LINE %%%
+\usepackage[T2A,T1]{fontenc}
+\usepackage[utf8]{inputenc}
+\usepackage[russian,english]{babel}
+
+\usepackage{url}
 \usepackage{times}
 \usepackage{latexsym}
-\setlength\titlebox{6.5cm}    % Expanding the titlebox
-%%% YOUR PACKAGES BELOW THIS LINE %%%
 
@@
 \usepackage{color}
+\usepackage[small,bf]{caption}
 \usepackage{graphicx}
@@
 \newcommand{\comment}[1]{\marginpar{\scriptsize\sf \textcolor{blue}{#1}}}
+
+\newcommand{\rus}[1]{\foreignlanguage{russian}{#1}}
 
@@
      \hline
-     \textbf{Ref}   & Ayudas econ\'{o}micas para el tratamiento de toxicoman\'{i}­as en comunidades terap\'{e}uticas no concertadas. \\
-     \textbf{Task}   & Ayudas econ\'{o}micas para el \{ \} de toxicoman\'{i}­as en comunidades terap\'{e}uticas no concertadas. \\
+     \textbf{Ref}   & Ayudas econ\'{o}micas para el tratamiento de toxicoman\'{i}as en comunidades terap\'{e}uticas no concertadas. \\
+     \textbf{Task}   & Ayudas econ\'{o}micas para el \{ \} de toxicoman\'{i}as en comunidades terap\'{e}uticas no concertadas. \\
      \textbf{Src} & Komunitate terapeutiko itundu gabeetan toxikomaniak tratatzeko diru-laguntzak ematea. \\
@@
 
-%\begin{table*}
-%  \begin{tabular}{|l|l|}
-%     \hline
-%     \textbf{10\%}   & ???????? ???????; ??? ????? { } ????? 7 ?????????, ????? ?????? ??? ????? 100 ??????. \\
-%     \textbf{20\%}   & ???????? ???????; ??? ????? { } ????? 7 ?????????, ????? { } ??? ????? 100 { }. \\
-%     \textbf{30\%} & ???????? ???????; ??? ????? { } ????? 7 { }, ????? ?????? { } ????? 100 { }. \\
-%     \hline
-%  \end{tabular}
-%  \caption{Example of different gap percentage settings.} 
-%  \label{table:percentage}
-%\end{table*}
+\begin{table*}
+  \begin{tabular}{|l|l|}
+     \hline
+     \textbf{10\%}   & \rus{Примерно полчаса; вам нужно выйти через 7 \{ \}, потом пройти ещё около 100 метров.} \\
+     \textbf{20\%}   & \rus{\{ \} полчаса; вам нужно \{ \} через 7 остановок, \{ \} пройти ещё около 100 метров.} \\
+     \textbf{30\%} & \rus{Примерно полчаса; вам нужно \{ \} через 7 \{ \}, потом пройти \{ \} около 100 \{ \}.} \\
+     \hline
+  \end{tabular}
+  \caption{Example of different gap percentage settings.} 
+  \label{table:percentage}
+\end{table*}
 
@@
 also human-produced, but in the source language of the pair. The second type is the \emph{machine-translated} hint, which 
-comes from the machine translation of the \emph{source} sentence.
-Table~\ref{table:example} shows a sample task, and Figure~\ref{figure:screenshot} shows the task in the online evaluation environment.\comment{EA: for some reason, the reference to screenshot is figure 2, while in the caption it says figure 1 (correctly)}
+comes from the machine translation of the \emph{source} sentence. Table~\ref{table:example} shows a sample task, and Figure~\ref{figure:screenshot} shows the task in the online evaluation environment.
 
@@
   \centering
-\includegraphics[resolution=144]{appraise-scr}
+\includegraphics[width=0.9\textwidth,resolution=144]{appraise-scr}
+ \caption{An example set of sentences in an online environment.}
 \label{figure:screenshot}
- \caption{An example set of sentences in an online environment.}
 \end{figure*}
@@
 \item  For Basque-Spanish, from the corpus of legal texts ``Memorias de traducci\'on del
-Servicio Oficial de Traductores del IVAP''\footnote{http://opendata.euskadi.net/memorias-de-traduccion-del-servicio-oficial-de-traductores-del-ivap/w79-contdata/es/};
+Servicio Oficial de Traductores del IVAP''\footnote{\url{http://tinyurl.com/tm-ivap}};
 \item  For Tatar-Russian, from the following sources on three different topics:
   \begin{enumerate}
-\comment{EA: whoops, what is the best way to make it understand cyrillic?}
-    \item  Casual conversations, from a textbook
-    %\footnote{???????? ?.?. ? ??????? ???????? ??-????????. ??????: ????????? ??. ???-??, 1994. ? 320 ?. ISBN 5?298?00463?6 (???. 219, 220, 232, 233, 234)} of spoken Tatar;
-    \item  Legal texts, from the Constitution and laws\footnote{http://tatarstan.ru} of Tatarstan;
-    \item  News, from the President of Tatarstan website\footnote{http://president.tatarstan.ru/}.
+    \item  Casual conversations, from a textbook\footnote{\rus{Литвинов И.Л.
+Я начинаю говорить по-татарски.
+Казань: Татарское кн.
+изд-во, 1994.
+— 320 с.
+ISBN 5—298—00463—6 (стр.
+219, 220, 232, 233, 234)}} of spoken Tatar;
+    \item  Legal texts, from the Constitution and laws\footnote{\url{http://tatarstan.ru}} of Tatarstan;
+    \item  News, from the President of Tatarstan website\footnote{\url{http://president.tatarstan.ru/}}.
   \end{enumerate}
@@
 \begin{table*}
+
+\centering
 \begin{tabular}{|r |*{4}{c}|}
@@
 \begin{table*}
+\centering
 \begin{tabular}{|r |*{4}{c}|}

### 2015-03-02 engkaz-eamt-2015/engkaz-eamt-2015.tex
fixes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59143 72bbbca6-d526-0410-a7d9-f06f51895060
fixes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59182 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 Therefore, the complex agglutinative morphology of Kazakh language is very different from morphologically not too complex language like English; on the other hand, Turkic language morphology shows clear morphotactics (ordering of morphemes), its morphophonology shows complex phonological changes to due to interactions between neighboring morphemes (vowel harmony, sonorization, etc.) many of which are explicitly represented in writing.
+
 On the other hand, there are many differences between the syntax of Turkic languages and English:.subject-adverb–object–verb order (compare subject–verb–object-adverb in English), use of postpositions (compare prepositions in English), head-final syntax with modifiers and specifiers always preceding the modified/specified (normally following in English), overt case marking allowing for a rather free ordering of arguments (versus a more fixed order in English), lack of definite articles (extensively used in English), verbal-noun-centered structures where English uses modal verbs (must, have to, want to) or verbal-noun or verbal-adjective-centered constructions where English has subordinate clauses using finite verbs with relatives or subordinating conjunctions (the book which I read, the place where I saw him, before he came), lack of a parallel of the English verb have, as used for possession, etc.
 For an account (in Russian) of syntax differences between English and Kazakh, see Печерских \& Амангельдина (2012).
+
 This paper describes work in progress of development of machine translation system for English-to-Kazakh, which developed by using Apertium free/open-source machine translation platform (Forcada et al.
@@
     Google & 62.00 & 25.92 \\
-    Sanasoft & 3.09 & 99.72 \\
-    Trident & 5.06 & 92.86 \\
+    Sanasoft & 21.00 & 74.52 \\
     \hline
@@
 
-\com{task-based evaluation (WER, assim, TM-patching?) -FMT}
+As can be seen,  coverage of dictionary is good, average score is 94\%.
+Our system outperforms other available RBMT system, but fall short of the state-of-the-art performance.
+However, unlike the state of the art, which uses unavailable corpora, our system is free and open and can be improved.
 
-\com{Compare with GT + other systems -FMT}
+MT systems have some mistakes in translations, which was shown in <...> brackets.
+Google MT system has BLEU score higher, but in translation separate phrases, it has common error: do not assign right possessive and case.
+Sanasoft system has more errors with choosing not right translations or do not translate word from English.
 
@@
 
+We have presented the rule-based MT system from English to Kazakh.
 The current English-Kazakh machine translation system already successfully translates noun-phrases, verb-phrases, prepositional-phrases, and adjectival-phrases, and contains a good vocabulary for testing purposes.
+We plan to continue developing English-Kazakh pair, going to extend the coverage to 98% on the corpora, which we used before.
+Additionally, we will improve the quality of translation by adding more rules, such like constraint grammar rules, structural transfer and lexical rules.
+The future plan is to use the created data with other open-source MT system with Turkic languages or system with Kazakh language on right side to make transfer systems between the Turkic or other language pairs.
+Related work is currently ongoing with Russian-Kazakh.
+Our system is available as free/open-source software and the whole system may be downloaded from sourceforge(apertium-eng-kaz1, apertium-kaz2).
 
 \section*{Acknowledgements}
+
+MLF thanks the Kazakh state program for the attraction of foreign scholars and Prof.
+Ualsher Tukeyev for supporting his visit to the Kazakh National University, where part of this work was carried out.
+We also thank Prof.
+Tukeyev for his valuable input.
 

### 2015-03-02 engkaz-eamt-2015/engkaz-eamt-2015.tex
conversion
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59183 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{fontspec}
+\usepackage{natbib}
 \defaultfontfeatures{Scale=MatchLowercase,Mapping=tex-text}
@@
 
-This paper describes work in progress of development of machine translation system for English-to-Kazakh, which developed by using Apertium free/open-source machine translation platform (Forcada et al.
-2011, http://www.apertium.org).
+This paper describes work in progress of development of machine translation system for English-to-Kazakh, which developed by using Apertium free/open-source machine translation platform \citep{forcada11}.
+%(http://www.apertium.org).
 
 \section{The Apertium platform}
-Apertium (Forcada et al.
-2011, http://www.apertium.org) is a free/open-source rule-based machine translation (MT) platform that was built in 2005 by the Universitat d’Alacant.
+Apertium \citep{forcada11} is a free/open-source rule-based machine translation (MT) platform that was built in 2005 by the Universitat d’Alacant.
 At first, it was initially aimed to translate texts between closely related languages, then it was extended to deal with unrelated languages.
@@
 Separates the text to be translated from the formatting tags.
-Formatting tags are encapsulated in brackets so they are treated as “superblanks” that are placed between words in such a way that the remaining modules see them as regular blanks.
+Formatting tags are encapsulated in brackets so they are treated as ``superblanks'' that are placed between words in such a way that the remaining modules see them as regular blanks.
 \item	Morphological analyser.
@@
 Alternative translations are defined in bilingual dictionary by multiple entries of one word.
-For exmaple, adjective “beautiful” could be translated as “сұлу”, if following noun is subject: “beautiful girl” - “сұлу қыз”; or “әдемі”, “көркем”, if the noun means place or : “beautiful mountains” - “әдемі таулар”.
+For exmaple, adjective ``beautiful'' could be translated as ``сұлу'', if following noun is subject: ``beautiful girl'' - ``сұлу қыз''; or ``әдемі'', ``көркем'', if the noun means place or : ``beautiful mountains'' - ``әдемі таулар''.
 In the following table could be seen some lexical selection rules:
@@
 \begin{itemize}
-\item A first round of transformations (“chunker”) detects source language(SL) LF patterns and generates the corresponding sequences of TL LFs grouped in chunks representing simple constituents such as noun phrases, prepositional phrases, etc.
-\item The second round (“interchunk”) reads patterns of chunks and produces a new sequence of chunks.
+\item A first round of transformations (``chunker'') detects source language(SL) LF patterns and generates the corresponding sequences of TL LFs grouped in chunks representing simple constituents such as noun phrases, prepositional phrases, etc.
+\item The second round (``interchunk'') reads patterns of chunks and produces a new sequence of chunks.
 This is the module where one can attempt to perform some longer-range reordering operations, interchunk agreement, case selection, etc.
-\item The third round (“postchunk”) transfers chunk-level tags to the lexical forms they contain and whose lexical-form-level tags are linked (through a referencing systems) to chunk-level tags (for instance, case determined for a noun phrase is transferred to the main noun), and removes all grouping information to generate the desired sequence of TL LFs.
+\item The third round (``postchunk'') transfers chunk-level tags to the lexical forms they contain and whose lexical-form-level tags are linked (through a referencing systems) to chunk-level tags (for instance, case determined for a noun phrase is transferred to the main noun), and removes all grouping information to generate the desired sequence of TL LFs.
 \end{itemize}
@@
 
-In the first round of structural transfer (the “chunker”, rules written in the apertium-eng-kaz.eng-kaz.t1x file), rules do some local operations to cut sentences into chunks, like short noun phrases, adjective phrases, verb phrases and adpositional phrases (that is, prepositional phrases in English and postpositional phrases in Kazakh).
+In the first round of structural transfer (the ``chunker'', rules written in the apertium-eng-kaz.eng-kaz.t1x file), rules do some local operations to cut sentences into chunks, like short noun phrases, adjective phrases, verb phrases and adpositional phrases (that is, prepositional phrases in English and postpositional phrases in Kazakh).
 Chunking rules, of which there are currently 168, identify 8 kinds of chunks and translate them into Kazakh as much as possible, leaving some minor operations to be performed in later stages of structural transfer (for instance, the case of noun phrases).
@@
 \begin{itemize}
-\item Translating the locative “-{D}{A}”,1 ablative “-{D}{A}н”, etc.,except the genitive “-{N}{I}ң” cases: [PP [P from ] [NP the five cars] ] → [PP [NP бес көлік] [P -тен ] ]
+\item Translating the locative ``-{D}{A}'',1 ablative ``-{D}{A}н'', etc.,except the genitive ``-{N}{I}ң'' cases: [PP [P from ] [NP the five cars] ] → [PP [NP бес көлік] [P -тен ] ]
 \item Genitive -NIņ, which will be marked GenP:  [PP [P of ] [NP the beautiful garden] ] → [GenP [NP әдемі бақша] [P -ның] ] 
@@
 tenses expressing continued activity, such as the English present continuous or past continuous (I am playing, I was playing), have to be detected and mapped onto sets of two lexical units (Мен ойнап жатырмын, Мен ойнап отырдым) where the main verb is found in the -п participle form (ойнап), and a suitable finite form  (жатырмын, отырдым) of an auxiliary verb (жатыр, отыр) is used to express number and person agreement1,
-modal verbs are translated by adding adjectives, which means “necessary” (керек) or “proper” (жөн), and a form of the copula (absent in present tense); the subject receives the genitive or dative case;
-negative verbs for present continuous and past tense are translated by adding negative word “емес” for present perfect continuous( He has not been playing - Ол ойнап отырған емес ) and  “жоқ” for past tense and present continuous(see examples above).
+modal verbs are translated by adding adjectives, which means ``necessary'' (керек) or ``proper'' (жөн), and a form of the copula (absent in present tense); the subject receives the genitive or dative case;
+negative verbs for present continuous and past tense are translated by adding negative word ``емес'' for present perfect continuous( He has not been playing - Ол ойнап отырған емес ) and  ``жоқ'' for past tense and present continuous(see examples above).
 \item[Translation of adjectival phrases].
@@
 \item (1)  The adjective alone, marked AdjP: [AdjP big ] → [AdjP үлкен] 
-\item (2)  Comparative adjective phrases  (English more + adjective, or adjective-[e]r); the Kazakh translation chooses the comparative suffix “-{I}р{A}{K}”: [AdjP  more  interesting ]  → [AdjP қызығырақ]
-\item (3)  For superlative adjective phrases  “the most + adjective”  or “adjective-[e]st”, translation is built using “ең” + adjective:  [SupP  the biggest ]  → [SupP ең үлкен] [SupP  the most important ]   → [SupP ең маңызды]
+\item (2)  Comparative adjective phrases  (English more + adjective, or adjective-[e]r); the Kazakh translation chooses the comparative suffix ``-{I}р{A}{K}'': [AdjP  more  interesting ]  → [AdjP қызығырақ]
+\item (3)  For superlative adjective phrases  ``the most + adjective''  or ``adjective-[e]st'', translation is built using ``ең'' + adjective:  [SupP  the biggest ]  → [SupP ең үлкен] [SupP  the most important ]   → [SupP ең маңызды]
 \end{itemize}
@@
 
-The second round of structural transfer (the “interchunk” rules written in the apertium-eng-kaz.eng-kaz.t2x file) is currently performed by a proof-of-concept set of 90 rules, representative of following operations:
+The second round of structural transfer (the ``interchunk'' rules written in the apertium-eng-kaz.eng-kaz.t2x file) is currently performed by a proof-of-concept set of 90 rules, representative of following operations:
 \begin{itemize}
@@
 \item Reordering: placing of object before verb ( I[1] bought[2] the table[3] → Мен[1] үстелді[3] сатып алдым[2]),placing of prepositional pharses before the verb  (They[1] played[2] on top of the tree [3] → Олар[1] ағаштың үстінде[3] ойнады [2]), etc.
-\item Adding question particle ма/ме/ба/бе in the end of question, if question mark chunk “?” detected (Did<VP\_ques> you<NP> watch<VP> last film<NP> ?<Q\_m> $\rightarrow$ Сіз<NP> соңғы фильмді<NP> көрдіңіз<VP> бе<ques> ?<Q\_m>).
+\item Adding question particle ма/ме/ба/бе in the end of question, if question mark chunk ``?'' detected (Did<VP\_ques> you<NP> watch<VP> last film<NP> ?<Q\_m> $\rightarrow$ Сіз<NP> соңғы фильмді<NP> көрдіңіз<VP> бе<ques> ?<Q\_m>).
 \item Place possessive for long noun + noun + noun structures(The university of city of Kazakhstan - Қазақстан  қаласының университеті).
@@
 
-Additional cleaning stage was created for removing extra tags which did not determine through the chunker and interchunk steps(rules are written in  apertium-eng-kaz.eng-kaz.t4x file), for instance, noun was not act as object and did not get accusative case, so not determined tag was 'CD' after “cleaning” it will be set on nominative case; at this step, also carried out agreement between noun or adjective and copula “е”(Мен дәрігер+е<p1 person singualr>), vowel and consonants agreement between last word of question and question particle(Сіз келдіңіз + ме?
+Additional cleaning stage was created for removing extra tags which did not determine through the chunker and interchunk steps(rules are written in  apertium-eng-kaz.eng-kaz.t4x file), for instance, noun was not act as object and did not get accusative case, so not determined tag was 'CD' after ``cleaning'' it will be set on nominative case; at this step, also carried out agreement between noun or adjective and copula ``е''(Мен дәрігер+е<p1 person singualr>), vowel and consonants agreement between last word of question and question particle(Сіз келдіңіз + ме?
 - Сіз келдіңіз + бе?).
@@
 
-Current system can translate simple sentences and questions. Bilingual English-kazakh dictionary contains of  13135 entries, “chunker” rules are 168 and interchunk rules are 148. 
+Current system can translate simple sentences and questions. Bilingual English-kazakh dictionary contains of  13135 entries, ``chunker'' rules are 168 and interchunk rules are 148. 
 Evaluating was tested against  revision 58998 in the Apertium SVN1.
@@
 \begin{table}
-  \begin{tabular}{|l|r|}
+  \centering
+  \begin{tabular}{|l|r|r|}
     \hline
-    \textbf{Corpus} & \textbf{Coverage} (\%) \\
+    \textbf{Corpus} & \textbf{Tokens} & \textbf{Coverage} (\%) \\
     \hline
-    Wikipedia & 84.67 \\
-    EuroParl & 97.95 \\
-    NewsCommentary & 96.27 \\
-    SETimes & 97.90 \\
+    SETimes & 5.1 million & 97.90 \\
+    NewsCommentary & 6.5 million & 96.27 \\
+    EuroParl & 54.5 million & 97.95 \\
+    Wikipedia & 1.83 billion & 84.67 \\
     \hline
   \end{tabular}
+  \caption{Vocabulary coverage of the English--Kazakh system over four available corpora.}
 \end{table}
@@
 \begin{table}
+  \centering
   \begin{tabular}{|l|r|r|}
@@
   \end{tabular}
+  \caption{Metric results for the three systems compared.}
+  \label{table:metrics}
 \end{table}
@@
 
-\section{References}
-1 Сундетова А.М., Кәрібаева А.С., АПЕРТИУМ ПЛАТФОРМАСЫНДАҒЫ АҒЫЛШЫН-ҚАЗАҚ МАШИНАЛЫҚ АУДАРМАШЫ ҮШІН ЕКІТІЛДІ СӨЗДІКТІ ҚҰРУ.
+%\section{References}
+%1 Сундетова А.М., Кәрібаева А.С., АПЕРТИУМ ПЛАТФОРМАСЫНДАҒЫ АҒЫЛШЫН-ҚАЗАҚ МАШИНАЛЫҚ АУДАРМАШЫ ҮШІН ЕКІТІЛДІ СӨЗДІКТІ ҚҰРУ.
 Матер.
@@
 
+\bibliographystyle{apalike}
+
+\bibliography{2015-eamt-engkaz}
+
 \end{document}

### 2015-03-02 engkaz-eamt-2015/engkaz-eamt-2015.tex
more stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59185 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \begin{abstract}
-..
+This paper presents rule-based machine translation from English to Kazakh. Main differences 
+between two languages in syntax and morphology are presented, and showed how machine translation 
+system was designed to solve these challenges. We showed an evaluation of system coverage and 
+translation quality and future work.
 \end{abstract}
@@
 
-This paper presents a shallow-transfer rule-based machine translation system from English into Kazakh language.
-First of all, English language is a West Germanic language and Kazakh belong to group of Turkic languages.
-Therefore, the complex agglutinative morphology of Kazakh language is very different from morphologically not too complex language like English; on the other hand, Turkic language morphology shows clear morphotactics (ordering of morphemes), its morphophonology shows complex phonological changes to due to interactions between neighboring morphemes (vowel harmony, sonorization, etc.) many of which are explicitly represented in writing.
+This paper presents a shallow-transfer rule-based machine translation system from English 
+into Kazakh language. First of all, English language is a West Germanic language and Kazakh 
+belong to group of Turkic languages. Therefore, the complex agglutinative morphology of Kazakh 
+language is very different from morphologically not too complex language like English; on the other 
+hand, Turkic language morphology shows clear morphotactics (ordering of morphemes), its morphophonology 
+shows complex phonological changes to due to interactions between neighboring morphemes (vowel harmony, 
+sonorisation, etc.) many of which are explicitly represented in writing.
 
-On the other hand, there are many differences between the syntax of Turkic languages and English:.subject-adverb–object–verb order (compare subject–verb–object-adverb in English), use of postpositions (compare prepositions in English), head-final syntax with modifiers and specifiers always preceding the modified/specified (normally following in English), overt case marking allowing for a rather free ordering of arguments (versus a more fixed order in English), lack of definite articles (extensively used in English), verbal-noun-centered structures where English uses modal verbs (must, have to, want to) or verbal-noun or verbal-adjective-centered constructions where English has subordinate clauses using finite verbs with relatives or subordinating conjunctions (the book which I read, the place where I saw him, before he came), lack of a parallel of the English verb have, as used for possession, etc.
-For an account (in Russian) of syntax differences between English and Kazakh, see Печерских \& Амангельдина (2012).
+On the other hand, there are many differences between the syntax of Turkic languages 
+and English: subject-adverb–object–verb order (compare subject–verb–object-adverb in English), use of 
+postpositions (compare prepositions in English), head-final syntax with modifiers and specifiers 
+always preceding the modified/specified (normally following in English), overt case marking allowing 
+for a rather free ordering of arguments (versus a more fixed order in English), lack of definite 
+articles (extensively used in English), verbal-noun-centered structures where English uses modal 
+verbs (must, have to, want to) or verbal-noun or verbal-adjective-centred constructions where 
+English has subordinate clauses using finite verbs with relatives or subordinating 
+conjunctions (the book which I read, the place where I saw him, before he came), lack of a parallel 
+of the English verb \emph{have}, as used for possession, etc. For an account (in Russian) of 
+syntactic differences between English and Kazakh, see \cite{pecherskih2012}.
 
-This paper describes work in progress of development of machine translation system for English-to-Kazakh, which developed by using Apertium free/open-source machine translation platform \citep{forcada11}.
-%(http://www.apertium.org).
+We currently build rule-based machine translation system, because for statistical machine translation 
+system we need large sentence-aligned parallel corpora. There are no freely available, large parallel 
+corpora for the English-Kazakh pair. English-Kazakh statistical MT system was built by Google,\footnote{\url{http://translate.google.com}} 
+other work in this language pair was made by building a small parallel corpus and using Moses \citep{assylbekov14}, 
+and existing commercial systems for English to Kazakh (Sanasoft\footnote{\url{http://www.sanasoft.kz/c/ru/node/47}} and Trident\footnote{\url{http://www.translate.ua/us/on-line}}) both appear to be rule-based.
+Rule-based machine translation, in which developers
+write up dictionaries and grammatical rules that are applied by an engine, in some 
+unusual case will produce better translation.
+
+This paper describes work in progress of development of machine translation system 
+for English-to-Kazakh, which developed by using Apertium free/open-source machine 
+translation platform \citep{forcada11}. %(http://www.apertium.org).
 
 \section{The Apertium platform}
-Apertium \citep{forcada11} is a free/open-source rule-based machine translation (MT) platform that was built in 2005 by the Universitat d’Alacant.
-At first, it was initially aimed to translate texts between closely related languages, then it was extended to deal with unrelated languages.
-This platform has next components: machine translation engine, developer’s tools, and linguistic data for an increasing number of language pairs and they are licensed under the free/open-source GNU General Public License (GPL, versions 2 and 3).
+Apertium \citep{forcada11} is a free/open-source rule-based machine translation (MT) platform that 
+has been developed since 2005 at the Universitat d'Alacant. At first, it was initially aimed at 
+translating texts between closely related languages, then it was extended to deal with less-related 
+languages. This platform has next components: machine translation engine, developer's tools, and linguistic 
+data for an increasing number of language pairs and they are licensed under the 
+free/open-source GNU General Public License (GPL, versions 2 and 3).
 
@@
 
-\begin{itemize}
-\item 	De-formatter.
-Separates the text to be translated from the formatting tags.
-Formatting tags are encapsulated in brackets so they are treated as ``superblanks'' that are placed between words in such a way that the remaining modules see them as regular blanks.
-\item	Morphological analyser.
-For each surface form the morphological analyser generates one or more lexical forms, which consist of: lemma (dictionary or citation form), lexical category (or part-of-speech), and inflection information.
-For example, the morphological analyser will deliver next sentence "I played" as showed below:
+\begin{description}
+\item[De-formatter:] Separates the text to be translated from the formatting tags.  Formatting tags are 
+  encapsulated in brackets so they are treated as ``superblanks'' that are placed between words in 
+  such a way that the remaining modules see them as regular blanks.  
+\item[Morphological analyser:] For each surface form the morphological analyser generates one or more 
+  lexical forms, which consist of: lemma (dictionary or citation form), lexical category (or part-of-speech), 
+  and inflection information. For example, the morphological analyser will deliver the following analysis for 
+  the sentence ``I played''.
 \begin{verbatim}
@@
 \end{verbatim}
-In this example each source form has been analysed as one or more lexical forms: "I" is analysed into "I", where lexical category is subject pronoun(<prn><subj>) and first person(p1), could be masculine or feminine(mf), singular(sg); "played" is analysed into lexical verb(vblex) "play" with past simple tense(past) or it could be past participle(pp), so it has two analysis.
-Each analysis of word are separated by \texttt{\^{}} and \texttt{\$} symbols, and for one word each lexical forms are delimited by "/", and tags(<...>) show grammatical attributes of lexical form.
- For Kazakh language finite state transducer based on two-level rules (in the case of Kazakh, apertium-kaz.kaz.lexc, apertium-kaz.kaz.twol).
-This module therefore separates lexemes and processes morphological analysis, and then returns possible lexical forms.
-\item	Part-of-speech (POS) tagger.
-This module chooses one of lexical forms of ambiguous word.
-As can be seen from the previous example, the morphological analyser could deliver more than one lexical form and choosing wrong one could produce errors in translation.
-POS tagger is based on a statistical model, which based on hidden Markov models and it has been trained on source language texts, which processes the result of the application of  on constraint-grammar rules (Karlsson 2005), which areused to discard some analyses  using simple rules (written in apertium-eng-kaz.eng-kaz.rlx) based on context.
-For example, consider the morphological analysis of word "book":
+  In this example each source form has been analysed as one or more lexical forms: ``I'' is analysed 
+  into ``I'', where lexical category is subject pronoun (\texttt{<prn><subj>}) and first person (\texttt{<p1>}), could be 
+  masculine or feminine (\texttt{<mf>}), singular (\texttt{<sg>}); ``played'' is analysed into lexical verb (\texttt{<vblex>}) 
+  ``play'' with past simple tense (\texttt{<past>}) or it could be past participle (\texttt{<pp>}), so it has two 
+  analyses. Each analysis of word is separated by \texttt{\^{}} and \texttt{\$} symbols, and for one word each lexical 
+  forms are delimited by ``/'', and tags(<...>) show grammatical attributes of lexical form. 
+\item[Part-of-speech (POS) tagger:] This module chooses one of lexical forms of ambiguous word. As can be 
+  seen from the previous example, the morphological analyser could deliver more than one lexical form and 
+  choosing wrong one could produce errors in translation. POS tagger is based on a statistical model, which 
+  based on hidden Markov models and it has been trained on source language texts, which processes the 
+  result of the application of  on constraint-grammar rules \citep{karlsson95}, which are used to discard some 
+  analyses using rules based on context. For example, consider the morphological analysis of word ``book'':
 \begin{verbatim}
@@
 \end{verbatim}
-This word has 3 lexical forms, and depends on context, ont of them will be chosen by rules or tagger. After  this module, all words have only one morphological analysis.
-\item	Lexical transfer.
-It reads each lexical form of source language and delivers a corresponding target language lexical form.
-This module uses a bilingual dictionary (apertium-eng-kaz.eng-kaz.dix) which has very simple structure(1).
-For English-Kazakh language pair, each lexical form of English word is translated into Kazakh as follows:
+  This word has 3 lexical forms, and depends on context, ont of them will be chosen by rules or tagger. After  this module, 
+  all words have only one morphological analysis.
+\item[Lexical transfer:] It reads each lexical form of source language and delivers a corresponding target language 
+  lexical form. This module uses a bilingual dictionary. For the English-Kazakh language pair, each lexical form of 
+  English word is translated into Kazakh as follows:
 \begin{verbatim}
@@
 \end{verbatim}
- Multiword units are translated as a single word.
-\item Lexical selection.
-It uses rules for lexical words, which have many translations, to select one of the translations in the target language according to context.
-All rules are written in file apertium-eng-kaz.eng-kaz.lrx.
-\item Structural transfer.
-This module uses pattern matching to identify sequences of lexical forms  (phrases or segments), which need syntactical processing to translate grammatical differences between two languages (handling of number, gender, etc.).
-It uses  files with rules, which specify  the syntactic transformation such as word reorderings, lexical changes such as changes in prepositions and agreement between target language lexical forms.
-Transfer rules can produce new sequences for the  target language, for instance, preposition-noun rule is used to built sequence, where for noun is chosen right case, which depends on preposition:
+ Multiword units can be translated as a single word.
+\item[ Lexical selection:] It uses rules for lexical words, which have many translations, to select one of the translations 
+  in the target language according to context. The system is described in \cite{tyers12a}.
+\item [Structural transfer:] This module uses pattern matching to identify sequences of lexical forms  (phrases or 
+  segments), which need syntactical processing to translate grammatical differences between two languages (handling of 
+  number, gender, etc.). It uses  files with rules, which specify  the syntactic transformation such as word reorderings, 
+  lexical changes such as changes in prepositions and agreement between target language lexical forms. Transfer rules can 
+  produce new sequences for the  target language, for instance, preposition-noun rule is used to built sequence, where for 
+  noun is chosen right case, which depends on preposition: 
 \begin{verbatim}
-in garden  -  ^бақша<n><sg><PXD><loc>$
+in garden  --  ^бақша<n><sg><PXD><loc>$
  \end{verbatim} 
-\item	 Morphological generator.
-Delivers the sequence of target-language lexical forms, produced by the structural transfer, to a corresponding sequence of target language surface forms.
-The morphological generator executes a finite-state transducer generated by compiling a morphological dictionary for the target language.
-\item Post-generator.
-It performs some minor orthographical operations in the target language (for instance, it generates the English form cannot from can and not).
-This module is generated from file with rules which are very   similar in format to dictionary files.
-\item Reformatter. It places format tags back into the text so that its format is preserved.
-\end{itemize}
+\item[Morphological generator:] Delivers the sequence of target-language lexical forms, produced by the structural transfer, 
+  to a corresponding sequence of target language surface forms. The morphological generator executes a finite-state transducer 
+  generated by compiling a morphological dictionary for the target language.     
+  For Kazakh language a finite-state transducer based on two-level rules \citep{washingtonsalimzyantyers14} is used for generation. 
+\item[ Post-generator:] Performs some minor orthographical operations in the target language (for instance, it generates the 
+  English form cannot from can and not). This module is generated from file with rules which are very   similar in format to dictionary files. 
+\item[ Reformatter:] It places format tags back into the text so that its format is preserved.
+\end{description}
 
@@
 
-Apertium uses XML-based formats for linguistic data, which include bilingual and monolingual dictionaries, structural transfer, lexical rules and rules for part-of-speech tagging.
+Apertium uses XML-based formats for linguistic data, which include bilingual and monolingual dictionaries, structural transfer and
+lexical selection rules. % and rules for part-of-speech tagging.
 
@@
 
-Dictionaries are used in lexical processing and morphological analysing, there are use two types: morphological(monolingual) and bilingual.
+Dictionaries are used in lexical processing and morphological analysis, there are use two types: morphological (monolingual) and bilingual.
 
@@
 
-English dictionary uses to determine lexical forms of each surface forms, it contains of list of all lexical units, definition of source language alphabet, grammatical categories, which is used to specify lexical forms, for example, like noun, verb, gender, etc., each categories have paradigms, they describes groups of correspondences between parts of surface forms and lexical forms.
+English dictionary uses to determine lexical forms of each surface forms, it contains of list of all lexical units, definition 
+of source-language alphabet, grammatical categories, which is used to specify lexical forms, for example, like noun, verb, 
+gender, etc., each categories have paradigms, they describes groups of correspondences between parts of surface forms and lexical forms.
 
-For Kazakh side, is used morphological transducers, which are based on the Helsinki
-Finite State Toolkit (Linden et al., 2011).
-It uses formalisms lexc for defining lexicons, twol formalism for morphophonological rules.
-Lexc formalism defines different word classes and subclasses.
-And twol determines phonological transformation like vowel harmony,desonorisation.
-For example, depend on preceding vowel, plural form "TAR" could be TER by harmonising with vowels on word: kitap+TAR, but mektep+TER.
+For Kazakh side, is used morphological transducers, which are based on the Helsinki Finite State Toolkit \citep{hfst/2011}. It 
+uses formalisms lexc for defining lexicons, twol formalism for morphophonological rules. Lexc formalism defines different 
+word classes and subclasses. And twol determines phonological transformation like vowel harmony, desonorisation. For example, 
+depend on preceding vowel, plural form \emph{-LAr} could become -тар or -тер depending on vowel harmony: kitap+tar, but mektep+ter.
 
@@
 
-Bilingual dictionary provides correspondences between lexical forms of source and target lanugages.
-Each entry consists of English word on the left side with part of speech and corresponding Kazakh word on the right side.
-Some entries could be ambiguous by part of speech, or just have many translations, which wil be chosen by lexical selection rules depend on context.
-This dictionary currently contains 13130 stems.
+Bilingual dictionary provides correspondences between lexical forms of source and target lanugages. Each entry 
+consists of English word on the left side with part of speech and corresponding Kazakh word on the right side. 
+Some entries could be ambiguous by part of speech, or just have many translations, which will be chosen by 
+lexical-selection rules depending on context. This dictionary currently contains 13,130 stems \citep{sundetova13a}.
 
@@
 
-This kind of rules solves the problem of part-of-speech(POS) ambiguity by Constraint Grammar(CG) (Karlsson et al., 1995).
-The output of morphological analyser could be ambiguous, and in this case CG rules choose correct analysis from multiple analyses.
-From English to Kazakh POS ambiguity could be only in English side, but also some choices depends on Kazakh grammar.
+This kind of rules solves the problem of part-of-speech (POS) ambiguity by Constraint Grammar (CG, \cite{karlsson95}). The 
+output of morphological analyser could be ambiguous, and in this case CG rules choose correct analysis from 
+multiple analyses. From English to Kazakh POS ambiguity could be only in English side, but also some choice also
+depend on Kazakh grammar.
 
@@
 
-Lexical selection rules choose on of the alternative translations of one word depend on the context.
-This rules are written in apertium-eng-kaz.eng-kaz.lrx.
-Alternative translations are defined in bilingual dictionary by multiple entries of one word.
-For exmaple, adjective ``beautiful'' could be translated as ``сұлу'', if following noun is subject: ``beautiful girl'' - ``сұлу қыз''; or ``әдемі'', ``көркем'', if the noun means place or : ``beautiful mountains'' - ``әдемі таулар''.
-In the following table could be seen some lexical selection rules:
+Lexical-selection rules choose one of the alternative translations of one word depend on the context. The rule 
+formalism is described in \cite{tyers12a}. Alternative translations are defined in bilingual dictionary by 
+multiple entries of one word. For exmaple, adjective ``beautiful'' could be translated as ``сұлу'', if 
+following noun is subject: ``beautiful girl'' -- ``сұлу қыз''; or ``әдемі'', ``көркем'', if the noun means 
+place or: ``beautiful mountains'' -- ``әдемі таулар''.  Table~\ref{table:lexsel} demonstrates some 
+lexical-selection rules:
 
@@
     \hline
-    \multirow{2}{*}{residence} & мекен & -- & I live in a residence. -- (Мен) мекенде өмір сүремін. \\
-                               & резиденция & \_ of \_ president & I see the residence of the president -- (Мен) президенттің резиденциясын көремін \\
+    \multirow{4}{*}{residence} & \multirow{2}{*}{мекен}      & \multirow{2}{*}{--} & I live in a residence. \\
+                               &                             &                     & (Мен) мекенде өмір сүремін. \\
+                               & \multirow{2}{*}{резиденция} & \multirow{2}{*}{\_ of \_ president} & I see the residence of the president \\ 
+                               &                             &                                     & (Мен) президенттің резиденциясын көремін \\
     \hline
-    \multirow{2}{*}{boot} & бәтеңке & -- & I bought boots. -- Мен бәтеңкелерді сатып алдым \\
-                          & жүксалғыш & \_ of \_ car &  He opened the boot of the car  -- Ол машинаның жүксалғышын ашты. \\
+    \multirow{4}{*}{boot} & \multirow{2}{*}{бәтеңке} & -- & I bought boots. \\
+                          &                          &     &  Мен бәтеңкелерді сатып алдым \\
+
+                          & \multirow{2}{*}{жүксалғыш} & \_ of \_ car &  He opened the boot of the car  \\ 
+                          &           &              & Ол машинаның жүксалғышын ашты. \\
     \hline
-    \multirow{2}{*}{anything} & бір нәрсе & -- & I  see anything -- Мен бір нәрсе көремін \\
-                              & еш нәрсе & not \_[verb] & I do not do anything - Мен еш нәрсе істемеймін. \\
+    \multirow{4}{*}{anything} & \multirow{2}{*}{бір нәрсе} & -- & I see anything \\ 
+                              &           &    & Мен бір нәрсе көремін \\
+                              & \multirow{2}{*}{еш нәрсе} & not \_[verb] & I do not do anything \\ 
+                              &          &             &  Мен еш нәрсе істемеймін. \\
 
@@
  \end{tabular}
-  \caption{Example lexical selection rules}
+  \caption{Example lexical-selection rules}
+  \label{table:lexsel}
 \end{table*}
@@
 
-For English–Kazakh system this stage uses three stages to improve the granularity of structural transfer rules (each one has its own rules file):
+For English–Kazakh system this stage uses three stages to improve the granularity of structural-transfer 
+rules \citep{sundetova13b}:
+
 \begin{itemize}
-\item A first round of transformations (``chunker'') detects source language(SL) LF patterns and generates the corresponding sequences of TL LFs grouped in chunks representing simple constituents such as noun phrases, prepositional phrases, etc.
-\item The second round (``interchunk'') reads patterns of chunks and produces a new sequence of chunks.
-This is the module where one can attempt to perform some longer-range reordering operations, interchunk agreement, case selection, etc.
-\item The third round (``postchunk'') transfers chunk-level tags to the lexical forms they contain and whose lexical-form-level tags are linked (through a referencing systems) to chunk-level tags (for instance, case determined for a noun phrase is transferred to the main noun), and removes all grouping information to generate the desired sequence of TL LFs.
+\item A first round of transformations (``chunker'') detects source language (SL) LF patterns and generates the 
+  corresponding sequences of TL LFs grouped in chunks representing simple constituents such as noun phrases, prepositional phrases, etc. 
+\item The second round (``interchunk'') reads patterns of chunks and produces a new sequence of chunks. This is the 
+  module where one can attempt to perform some longer-range reordering operations, interchunk agreement, case selection, etc. 
+\item The third round (``postchunk'') transfers chunk-level tags to the lexical forms they contain and whose lexical-form-level 
+  tags are linked (through a referencing systems) to chunk-level tags (for instance, case determined for a noun phrase is 
+  transferred to the main noun), and removes all grouping information to generate the desired sequence of TL LFs.
 \end{itemize}
 
-\subsection{English-to-Kazakh structural transfer}
+The structural transfer module in Apertium processes the stream of source-language lexical form -- target-language lexical 
+form pairs (SL LF–TL LF pairs) and transforms it into a new sequence of TL LFs after a series of structural transfer 
+operations specified in a set of rules: reordering, elimination or insertion of TL LFs, agreement, etc.
 
-The structural transfer module in Apertium processes the stream of source-language lexical form – target-language lexical form pairs (SL LF–TL LF pairs) and transforms it into a new sequence of TL LFs after a series of structural transfer operations specified in a set of rules: reordering, elimination or insertion of TL LFs, agreement, etc.
-This section describes the current structural transfer in Apertium-eng-kaz (revision ..., ).
-English to Kazakh chunker rules (file apertium-eng-kaz.eng-kaz.t1x), inter-chunk rules (file apertium-eng-kaz.eng-kaz.t2x) and additional clean-up stage(file apertium-eng-kaz.eng-kaz.t4x) will be described in detail next 3 sections.
+This section describes the current structural transfer in \texttt{apertium-eng-kaz} (revision ..., ). English to Kazakh 
+chunker rules, interchunk rules and additional clean-up stage will be described in detail next 3 sections.
 
-\subsubsection{The English-to-Kazakh chunker}
+\subsubsection{Chunker}
 
-In the first round of structural transfer (the ``chunker'', rules written in the apertium-eng-kaz.eng-kaz.t1x file), rules do some local operations to cut sentences into chunks, like short noun phrases, adjective phrases, verb phrases and adpositional phrases (that is, prepositional phrases in English and postpositional phrases in Kazakh).
-Chunking rules, of which there are currently 168, identify 8 kinds of chunks and translate them into Kazakh as much as possible, leaving some minor operations to be performed in later stages of structural transfer (for instance, the case of noun phrases).
-Table 3 describes what kind of noun phrases could appear in sentences and how it could be detected.
+In the first round of structural transfer rules do some local operations to cut sentences into chunks, like short noun 
+phrases, adjective phrases, verb phrases and adpositional phrases (that is, prepositional phrases in English and 
+postpositional phrases in Kazakh).
+Chunking rules, of which there are currently 168, identify 8 kinds of chunks and translate them into Kazakh as much 
+as possible, leaving some minor operations to be performed in later stages of structural transfer (for instance, the 
+case of noun phrases). Table~\ref{table:nps} describes what kind of noun phrases could appear in sentences and how it could be detected.
+
+\begin{table*}
+  \centering
+  \begin{tabular}{|l|l|l|}
+  \hline
+  \textbf{Structure} & \textbf{Description} & \textbf{Examples} \\
+  \hline
+  \texttt{n} & Noun & house \\
+  \texttt{adj + n} & Adjective + noun & long road \\
+  \texttt{det + n} & Determiner + noun & an apple, the chance \\
+  \texttt{num + n} & Numeral + noun & five pieces \\
+  \texttt{num + adj + n} & Numeral + adjective + noun & five big pieces \\
+  \texttt{det + num + adj + n} & Determiner + numeral + adjective + noun & the five big pieces \\
+  \hline
+  \end{tabular}
+  \caption{Example of the most common structures of noun phrases}
+  \label{table:nps}
+\end{table*}
+
+\begin{table*}
+  \centering
+  \begin{tabular}{|l|l|l|l|}
+    \hline
+    \textbf{Phrase type} & \textbf{Description} & \textbf{Examples} & \textbf{Chunk tags} \\
+    \hline
+    \texttt{NP} & Noun phrase           &  & \\
+    \texttt{VP} & Verb phrase           & & \\
+    \texttt{PP} & Postpositional phrase & & \\
+    \texttt{GenP} & Genitive phrase  & & \\
+    \texttt{AdjP} & Adjective phrase  & & \\
+    \texttt{SupP} & Superlative phrase & & \\
+    \texttt{Q\_mark} & Question mark  & & \\
+    \hline
+  \end{tabular}
+  \caption{Phrase types in the English-Kazakh chunker}
+  \label{table:phrases} 
+\end{table*}
+
+As it is shown above, the most probable sequence of world classes are built phrases and 
+for English-Kazakh chunker there are used next kind of phrases:
 
@@
 \begin{description}
-\item[Noun phrases].
-Consider the following example: the chunker identifies the English sequence the large book  (determiner–adjective–noun) as a noun-phrase chunk.
-It translates into Kazakh, and assigns it four chunk-level tags: number (set to singular), person (set to 3rd), possessor (to be determined, as the noun кітап ('book') could receive a 3rd-person possesive ending (кітабы) later if the context were, for instance, the large book of animals, аңдардың үлкен кітабы), and case (to be determined as it could be, for instance, accusative in I saw the large book, Мен үлкен кітапты көрдім ).
-Also noun form of the verb, for instance,  gerund form is determined as NP phrase(Her playing is good – Оның ойыны жақсы).
-\item[Prepositional phrases]. English prepositional phrase is translated into Kazakh a postpositional phrase, there are three possible outcomes with different cases:
+\item[Noun phrases:] Consider the following example: the chunker identifies the English sequence 
+  the large book  (determiner–adjective–noun) as a noun-phrase chunk. It translates into 
+  Kazakh, and assigns it four chunk-level tags: number (set to singular), person (set to 3rd), 
+  possessor (to be determined, as the noun кітап ('book') could receive a 3rd-person possesive 
+  ending (кітабы) later if the context were, for instance, the large book of animals, аңдардың үлкен кітабы), 
+  and case (to be determined as it could be, for instance, accusative in I saw the large book, Мен үлкен кітапты көрдім ). Also 
+  noun form of the verb, for instance,  gerund form is determined as NP phrase (Her playing is good – Оның ойыны жақсы).
+\item[Prepositional phrases:] English prepositional phrases are translated into Kazakh as postpositional phrases, 
+  there are three possible outcomes with different cases:
    \begin{itemize}
-\item Translating the locative ``-{D}{A}'',1 ablative ``-{D}{A}н'', etc.,except the genitive ``-{N}{I}ң'' cases: [PP [P from ] [NP the five cars] ] → [PP [NP бес көлік] [P -тен ] ]
-\item Genitive -NIņ, which will be marked GenP:  [PP [P of ] [NP the beautiful garden] ] → [GenP [NP әдемі бақша] [P -ның] ] 
-\item 3 With noun such as аст, үст, etc. which takes possessive from main noun:  [PP [P under ] [NP the garden] ] → [PP [NP [GenP [NP бақша ][P -ның]] [NP астын]] [P -да ] ]
+    \item Translating the locative ``-{D}{A}'',\footnote{{D} can be д or т, and {A} can be е or а, depending on the 
+        phonological context.} ablative ``-{D}{A}н'', etc., except the genitive ``-{N}{I}ң'' cases: 
+        [PP [P from ] [NP the five cars] ] $\rightarrow$ [PP [NP бес көлік] [P -тен ] ]
+    \item Genitive -{N}{I}ң, which will be marked GenP:  
+        [PP [P of ] [NP the beautiful garden] ] $\rightarrow$ [GenP [NP әдемі бақша] [P -ның] ] 
+    \item With nouns such as аст `under', үст `on', etc. which takes possessive from main noun:  
+        [PP [P under ] [NP the garden] ] $\rightarrow$ [PP [NP [GenP [NP бақша ][P -ның]] [NP астын]] [P -да ] ]
     \end{itemize}
-\item[Verb phrases]. Translation verb phrases into Kazakh  are not always be straightforward, for instance: 
-present simple and future are rendered using the same tense in Kazakh (I play → Мен ойнаймын;  I will play → Мен ойнаймын), this ambiguity problem may be solved by lexical selection, which will choose right tense by watching context;
-tenses expressing continued activity, such as the English present continuous or past continuous (I am playing, I was playing), have to be detected and mapped onto sets of two lexical units (Мен ойнап жатырмын, Мен ойнап отырдым) where the main verb is found in the -п participle form (ойнап), and a suitable finite form  (жатырмын, отырдым) of an auxiliary verb (жатыр, отыр) is used to express number and person agreement1,
-modal verbs are translated by adding adjectives, which means ``necessary'' (керек) or ``proper'' (жөн), and a form of the copula (absent in present tense); the subject receives the genitive or dative case;
-negative verbs for present continuous and past tense are translated by adding negative word ``емес'' for present perfect continuous( He has not been playing - Ол ойнап отырған емес ) and  ``жоқ'' for past tense and present continuous(see examples above).
-\item[Translation of adjectival phrases].
-In Kazakh noun phrases, adjectives come before nouns and do not show any agreement with nouns.
-Adjectives can also appear in separate adjective phrases.
-Here are some examples:
+\item[Verb phrases:] Translation verb phrases into Kazakh  are not always be straightforward, for instance: 
+  present simple and future are rendered using the same tense in Kazakh (I play $\rightarrow$ Мен ойнаймын;  I 
+  will play $\rightarrow$ Мен ойнаймын), this ambiguity problem may be solved by lexical selection, 
+  which will choose right tense by watching context; tenses expressing continued activity, such as the 
+  English present continuous or past continuous (I am playing, I was playing), have to be detected and 
+  mapped onto sets of two lexical units (Мен ойнап жатырмын, Мен ойнап отырдым) where the main verb is 
+  found in the -п participle form (ойнап), and a suitable finite form  (жатырмын, отырдым) of an auxiliary 
+  verb (жатыр, отыр) is used to express number and person agreement,\footnote{Actually, Kazakh language uses four auxiliary verbs: жатыр ('lie', 
+      used when the activity takes a long time), отыр ('sit', used when the activity appears to be done in a sitting position), тұр ('stand', 
+      when the takes a short time), and жүр (when the activity repeats regularly). Choosing the most adequate auxiliary verb is 
+      hard without a semantic analysis, which is not easily available in Apertium. Our current choice (an approximation) 
+      is жатыр ('lie') for the present continuous and отыр ('sit') for the past continuous.} 
+  modal verbs are translated by adding adjectives, which means ``necessary'' (керек) or ``proper'' (жөн), 
+  and a form of the copula (absent in present tense); the subject receives the genitive or dative case; negative 
+  verbs for present continuous and past tense are translated by adding negative word ``емес'' for present 
+  perfect continuous (He has not been playing -- Ол ойнап отырған емес) and  ``жоқ'' for past tense and 
+  present continuous (see examples above).
+\item[Adjectival phrases:] In Kazakh noun phrases, adjectives come before nouns and do not show any agreement 
+  with nouns.  Adjectives can also appear in separate adjective phrases. Here are some examples:
 \begin{itemize}
-\item (1)  The adjective alone, marked AdjP: [AdjP big ] → [AdjP үлкен] 
-\item (2)  Comparative adjective phrases  (English more + adjective, or adjective-[e]r); the Kazakh translation chooses the comparative suffix ``-{I}р{A}{K}'': [AdjP  more  interesting ]  → [AdjP қызығырақ]
-\item (3)  For superlative adjective phrases  ``the most + adjective''  or ``adjective-[e]st'', translation is built using ``ең'' + adjective:  [SupP  the biggest ]  → [SupP ең үлкен] [SupP  the most important ]   → [SupP ең маңызды]
+\item The adjective alone, marked AdjP: [AdjP big ] $\rightarrow$ [AdjP үлкен] 
+\item Comparative adjective phrases  (English more + adjective, or adjective-[e]r); the Kazakh translation 
+   chooses the comparative suffix ``-{I}р{A}{K}'': [AdjP  more  interesting ]  $\rightarrow$ [AdjP қызығырақ]
+\item For superlative adjective phrases  ``the most + adjective''  or ``adjective-[e]st'', translation 
+   is built using ``ең'' + adjective:  [SupP  the biggest ]  $\rightarrow$ [SupP ең үлкен] 
+   [SupP  the most important]   $\rightarrow$ [SupP ең маңызды]
 \end{itemize}
-Superlative adjective phrases have some properties of noun phrases (such as receiving possessive morphemes when modified by a genitive phrase: the most beautiful of people → адамдардың ең әдемісі); one could say that they are treated as NPs with an implied noun.
+Superlative adjective phrases have some properties of noun phrases (such as receiving possessive morphemes when 
+modified by a genitive phrase: the most beautiful of people $\rightarrow$ адамдардың ең әдемісі); one could 
+say that they are treated as NPs with an implied noun.
 \end{description}
 
-\subsubsection{English-to-Kazakh inter-chunk processing}
+\begin{table*}
+  \centering
+  \begin{tabular}{|l|l|l|l|}
+    \hline
+    \multicolumn{4}{|c|}{\textbf{Affirmative forms}} \\
+    \hline
+    \textbf{English form} & \textbf{Example} & \textbf{Morphemes} & \textbf{Translation} \\
+    \hline
+    \multirow{2}{*}{Present simple} & \multirow{2}{*}{I play} & \multirow{2}{*}{ойна + <aorist> + <pers.>} & Мен ойнаймын \\
+                                    &                         &                                            & I play \\
+    \hline
+    \multirow{2}{*}{Present continuous} & \multirow{2}{*}{I am playing} & ойна + <perf. part.>  & Мен ойнап жатырмын \\
+                                    &                         &   жатыр + <pres.> + <pers.>     & I playing lie \\
+    \hline
+    \multirow{2}{*}{Past continuous} & \multirow{2}{*}{I was playing} & ойна + <perf. part.>  & Мен ойнап отырдым \\
+                                    &                         & отыр + д{I} <past>  + <pers.>     & I playing sat \\
+    \hline
+    \multirow{2}{*}{Obligation} & \multirow{2}{*}{I must go} & Мен  + <gen.>  & Менің баруым керек \\
+                                    &                         & бар + <ger.> +  <px1sg> + керек    & My going necessary [is] \\
+    \hline
+    \multirow{2}{*}{Should} & \multirow{2}{*}{I should go} & Мен  + <gen.>  & Менің Менің барғаным жөн \\
+                                    &                         & бар + <past.ger.> +  <px1sg> + жөн & My was.going proper [is] \\
+    \hline
+    \multicolumn{4}{|c|}{\textbf{Negative forms}} \\
+    \hline
+    \multirow{2}{*}{Present cont.} & \multirow{2}{*}{I am not playing} & ойна + <perf. part.>  & Мен ойнап жатқан жоқпын \\
+                                    &                         &   жатыр + <pres.> + <pers.>     & I playing lying not \\
+    \hline
+    \multirow{2}{*}{Ability} & \multirow{2}{*}{I can not play} & ойна + <impf. part.>  & Мен ойнай алмаймын \\
+                                    &                         &   ал + <neg.> + <aor.> + <p1>     & I playing can.not \\
+    \hline
+    \multirow{3}{*}{Present perf. cont. } & \multirow{3}{*}{You have not been playing} & ойна + <perf. part.>  & Сіз ойнап отырған емессі \\
+                                    &                         & +  отыр + <past>  & You playing sat not  \\
+                                    &                         & + eмес <cop.neg> + <p2> & \\
+    \hline
 
-The second round of structural transfer (the ``interchunk'' rules written in the apertium-eng-kaz.eng-kaz.t2x file) is currently performed by a proof-of-concept set of 90 rules, representative of following operations:
+  \end{tabular}
+  \caption{Examples of Verb phrase translations}
+  \label{table:vps}
+\end{table*}
+
+\subsubsection{Interchunk processing}
+
+The second round of structural transfer is currently performed by a proof-of-concept set of 90 rules, representative of following operations:
 \begin{itemize}
 \item Inter-chunk agreement between subject noun phrase and verb phrase in their person and number.
-\item Assigning case to noun phrases (which are generated without case by the chunker): for instance, accusative case for objects (I bought the table → Мен үстелді  сатып алдым), genitive case for obligatory constructs (I must see → менің көруім керек),  dative case for the verb to need (I need a doctor →Маған дәрігер керек), locative case for possession (I have a book → Менде кітап бар), etc.
-\item Reordering: placing of object before verb ( I[1] bought[2] the table[3] → Мен[1] үстелді[3] сатып алдым[2]),placing of prepositional pharses before the verb  (They[1] played[2] on top of the tree [3] → Олар[1] ағаштың үстінде[3] ойнады [2]), etc.
-\item Adding question particle ма/ме/ба/бе in the end of question, if question mark chunk ``?'' detected (Did<VP\_ques> you<NP> watch<VP> last film<NP> ?<Q\_m> $\rightarrow$ Сіз<NP> соңғы фильмді<NP> көрдіңіз<VP> бе<ques> ?<Q\_m>).
-\item Place possessive for long noun + noun + noun structures(The university of city of Kazakhstan - Қазақстан  қаласының университеті).
+\item Assigning case to noun phrases (which are generated without case by the chunker): for instance, 
+  accusative case for objects (I bought the table $\rightarrow$ Мен үстелді  сатып алдым), genitive case for 
+  obligatory constructs (I must see $\rightarrow$ менің көруім керек),  dative case for the verb to 
+  need (I need a doctor $\rightarrow$Маған дәрігер керек), locative case for 
+  possession (I have a book $\rightarrow$ Менде кітап бар), etc.
+\item Reordering: placing of object before verb ( I[1] bought[2] the table[3] $\rightarrow$ Мен[1] үстелді[3] сатып алдым[2]), 
+  placing of prepositional pharses before the verb  (They[1] played[2] on top of the tree [3] $\rightarrow$ Олар[1] ағаштың үстінде[3] ойнады [2]), etc.
+\item Adding question word ма/ме/ба/бе in the end of question, if question mark chunk ``?'' is detected (Did<VP\_ques> you<NP> watch<VP> last film<NP> ?<Q\_m> $\rightarrow$ Сіз<NP> соңғы фильмді<NP> көрдіңіз<VP> бе<ques> ?<Q\_m>).
+\item Place possessive for long noun + noun + noun structures(The university of city of Kazakhstan -- Қазақстан қаласының университеті).
 \end{itemize}
+
 The set of rules has to be extended, as many combinations of the above phenomena are still not covered.
@@
 
-Additional cleaning stage was created for removing extra tags which did not determine through the chunker and interchunk steps(rules are written in  apertium-eng-kaz.eng-kaz.t4x file), for instance, noun was not act as object and did not get accusative case, so not determined tag was 'CD' after ``cleaning'' it will be set on nominative case; at this step, also carried out agreement between noun or adjective and copula ``е''(Мен дәрігер+е<p1 person singualr>), vowel and consonants agreement between last word of question and question particle(Сіз келдіңіз + ме?
-- Сіз келдіңіз + бе?).
+Additional cleaning stage was created for removing extra tags which did not determine through the chunker and 
+interchunk steps, for instance, noun was not acting as object and did not get accusative case, so not determined 
+tag was \texttt{<CD>} after ``cleaning'' it will be set on nominative case; at this step, also carried out 
+agreement between noun or adjective and copula ``е''(Мен дәрігер+е<p1 person singualr>), vowel and consonants 
+agreement between last word of question and question particle (Сіз келдіңіз + ме? - Сіз келдіңіз + бе?).
 
@@
 
-Current system can translate simple sentences and questions. Bilingual English-kazakh dictionary contains of  13135 entries, ``chunker'' rules are 168 and interchunk rules are 148. 
-Evaluating was tested against  revision 58998 in the Apertium SVN1.
-Lexical coverage is calculated over EuroParl1, SETimes2, NewsCommentary3, Wikipedia4(last check 04/06/2014).
-As can be seen from  Table 5, the higher coverage is of :
+Current system can translate simple sentences and questions. The English-Kazakh bilingual dictionary 
+contains 13,135 entries, ``chunker'' rules are 168 and interchunk rules are 148. 
+Evaluating was tested against  revision 58998 in the Apertium SVN1. Lexical coverage is calculated 
+over EuroParl,\footnote{\url{http://www.statmt.org/europarl/v7/es-en.tgz}} SETimes,\footnote{\url{http://nlp.ffzg.hr/data/corpora/setimes/setimes.en-tr.txt.tgz}} NewsCommentary,\footnote{\url{http://www.statmt.org/wmt13/training-parallel-nc-v8.tgz}} Wikipedia.\footnote{\url{http://dumps.wikimedia.org/enwiki/20140402/enwiki-20140402-pages-articles.xml.bz2}; last checked 04/06/2014}
+Table~\ref{table:coverage} presents the size of the corpora and the vocabulary coverage of the system for the given corpus.
 
@@
     \hline
-    SETimes & 5.1 million & 97.90 \\
-    NewsCommentary & 6.5 million & 96.27 \\
-    EuroParl & 54.5 million & 97.95 \\
-    Wikipedia & 1.83 billion & 84.67 \\
+    SETimes & 5.1 mil. & 97.90 \\
+    NewsCommentary & 6.5 mil. & 96.27 \\
+    EuroParl & 54.5 mil. & 97.95 \\
+    Wikipedia & 1.8 bil. & 84.67 \\
     \hline
@@
   \caption{Vocabulary coverage of the English--Kazakh system over four available corpora.}
+  \label{table:coverage}
 \end{table}
@@
 
-Evaluating the output of the system was measured with BLEU (Papineni et al.
-2002) and Word Error Rate metric(Levenstein, 1966).
-Was chosen text small English text, and by postediting output of English-kazakh system, was build parallel text.
+Evaluating the output of the system was measured with BLEU \citep{papineni02} and Word Error Rate 
+metric \citep{levenshtein/1966}. Was chosen text small English text, and by postediting output 
+of English-Kazakh system, was build parallel text.
 
-As can be seen,  coverage of dictionary is good, average score is 94\%.
-Our system outperforms other available RBMT system, but fall short of the state-of-the-art performance.
-However, unlike the state of the art, which uses unavailable corpora, our system is free and open and can be improved.
+As can be seen,  coverage of dictionary is good, average score is 94\%. Our system outperforms the other 
+available RBMT system, but fall short of the state-of-the-art performance. However, unlike the state of 
+the art, which uses unavailable corpora, our system is free and open and can be improved.
 
-MT systems have some mistakes in translations, which was shown in <...> brackets.
-Google MT system has BLEU score higher, but in translation separate phrases, it has common error: do not assign right possessive and case.
-Sanasoft system has more errors with choosing not right translations or do not translate word from English.
+MT systems have some mistakes in translations, which was shown in <...> brackets. Google MT system 
+has a higher BLEU score, but in translation separate phrases, it has common error: it does not assign 
+the right possessive and case. Sanasoft system has more errors with choosing not right translations 
+and out of vocabulary words.
 
@@
 The current English-Kazakh machine translation system already successfully translates noun-phrases, verb-phrases, prepositional-phrases, and adjectival-phrases, and contains a good vocabulary for testing purposes.
-We plan to continue developing English-Kazakh pair, going to extend the coverage to 98% on the corpora, which we used before.
+We plan to continue developing English-Kazakh pair, going to extend the coverage to 98\% on the corpora, which we used before.
 Additionally, we will improve the quality of translation by adding more rules, such like constraint grammar rules, structural transfer and lexical rules.
 The future plan is to use the created data with other open-source MT system with Turkic languages or system with Kazakh language on right side to make transfer systems between the Turkic or other language pairs.
-Related work is currently ongoing with Russian-Kazakh.
+Related work is currently ongoing with Russian--Kazakh.
 Our system is available as free/open-source software and the whole system may be downloaded from sourceforge(apertium-eng-kaz1, apertium-kaz2).

### 2015-03-02 engkaz-eamt-2015/engkaz-eamt-2015.tex
more stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59186 72bbbca6-d526-0410-a7d9-f06f51895060
more
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59187 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{fontspec}
+\usepackage{footnote}
 \usepackage{natbib}
@@
 
+\begin{savenotes}
 \begin{table*}
@@
     \hline
-    \texttt{NP} & Noun phrase           &  & \\
-    \texttt{VP} & Verb phrase           & & \\
-    \texttt{PP} & Postpositional phrase & & \\
-    \texttt{GenP} & Genitive phrase  & & \\
-    \texttt{AdjP} & Adjective phrase  & & \\
-    \texttt{SupP} & Superlative phrase & & \\
-    \texttt{Q\_mark} & Question mark  & & \\
+    \multirow{5}{*}{\texttt{NP}} & \multirow{5}{*}{Noun phrase}    & noun & number*, person, \\
+                                 &                                 & determiner-noun & possessor*, case\\
+                                 &                                 & numeral-noun & \\
+                                 &                                 & adjective-noun & \\
+                                 &                                 & determiner-adjective-noun & \\
     \hline
+    \multirow{5}{*}{\texttt{VP}} & \multirow{5}{*}{Verb phrase}    & finite-verb & number, person, \\
+                                 &                                 & have-participle & tense/mood, possessive*, \\
+                                 &                                 & be-gerund & negation* \\
+                                 &                                 & must-infinitive & \\ 
+                                 &                                 & want\_to-infinitive & \\ 
+    \hline
+    \multirow{5}{*}{\texttt{PP}} & \multirow{5}{*}{Postpositional phrase} & preposition-noun & number, person, \\
+                                 &                                        & preposition-determiner-noun & possessor*, case \\
+                                 &                                        & preposition-numeral-noun & \\
+                                 &                                        & preposition-adjective-noun & \\
+                                 &                                        & preposition-determiner-adjective-noun & \\
+    \hline
+    \multirow{2}{*}{\texttt{GenP}} & \multirow{2}{*}{Genitive phrase}  & As with \texttt{PP} except & number, person, \\
+                                   &                                   & with the preposition `of'. & possessor*, case \\
+    \hline
+    \multirow{2}{*}{\texttt{AdjP}} & \multirow{2}{*}{Adjective phrase}  & adjective & n/a \\
+                                   &                                    & ``more''--adjective & \\
+    \hline
+    \multirow{2}{*}{\texttt{SupP}} & \multirow{2}{*}{Superlative phrase} & adjective--``est'' & possessor*,\footnote{Treated as a noun phrase with an implied noun (``the largest [book]'')} case \\
+                                   &                                     & the--``most''--adjective & \\ 
+    \hline
+%    \texttt{Q\_mark} & Question mark  & & \\
+%    \hline
   \end{tabular}
@@
 \end{table*}
+\end{savenotes}
 
@@
    \begin{itemize}
+    \item Genitive -{N}{I}ң,\footnote{Upper case letters in braces (such as {N}) represent hypothetical archiphonemes (actually archigraphemes) that are realised as phonemes (actually graphemes) after morphophonological rules have been applied.
+For instance, in the genitive ending “-{N}{I}ң”, the archiphoneme {N} may be realised as т, д, or н and the archiphoneme {I} may be realised as  і  or ы depending on the previous phonological context.
+} which will be marked GenP:
+        [PP [P of ] [NP the beautiful garden] ] $\rightarrow$ [GenP [NP әдемі бақша] [P -ның] ] 
     \item Translating the locative ``-{D}{A}'',\footnote{{D} can be д or т, and {A} can be е or а, depending on the 
@@
         [PP [P from ] [NP the five cars] ] $\rightarrow$ [PP [NP бес көлік] [P -тен ] ]
-    \item Genitive -{N}{I}ң, which will be marked GenP:  
-        [PP [P of ] [NP the beautiful garden] ] $\rightarrow$ [GenP [NP әдемі бақша] [P -ның] ] 
     \item With nouns such as аст `under', үст `on', etc. which takes possessive from main noun:  
@@
 
-\section{Results}
+\section{Evaluation and results}
 
@@
 \end{table}
-
-\section{Evaluation}
 

### 2015-03-02 engkaz-eamt-2015/engkaz-eamt-2015.tex
more
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59188 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \section{The Apertium platform}
-Apertium \citep{forcada11} is a free/open-source rule-based machine translation (MT) platform that 
-has been developed since 2005 at the Universitat d'Alacant. At first, it was initially aimed at 
-translating texts between closely related languages, then it was extended to deal with less-related 
-languages. This platform has next components: machine translation engine, developer's tools, and linguistic 
-data for an increasing number of language pairs and they are licensed under the 
-free/open-source GNU General Public License (GPL, versions 2 and 3).
 
@@
 
+Apertium \citep{forcada11} is a free/open-source rule-based machine translation (MT) platform that 
+has been developed since 2005 at the Universitat d'Alacant. At first, it was initially aimed at 
+translating texts between closely related languages, then it was extended to deal with less-related 
+languages. This platform has next components: machine translation engine, developer's tools, and linguistic 
+data for an increasing number of language pairs and they are licensed under the 
+free/open-source GNU General Public License (GPL, versions 2 and 3).
+
 \begin{description}
@@
   lexical forms, which consist of: lemma (dictionary or citation form), lexical category (or part-of-speech), 
-  and inflection information. For example, the morphological analyser will deliver the following analysis for 
-  the sentence ``I played''.
-\begin{verbatim}
-^I/I<prn><subj><p1><mf><sg>$ 
-^played/play<vblex><past>/play<vblex><pp>$
-\end{verbatim}
-  In this example each source form has been analysed as one or more lexical forms: ``I'' is analysed 
-  into ``I'', where lexical category is subject pronoun (\texttt{<prn><subj>}) and first person (\texttt{<p1>}), could be 
-  masculine or feminine (\texttt{<mf>}), singular (\texttt{<sg>}); ``played'' is analysed into lexical verb (\texttt{<vblex>}) 
-  ``play'' with past simple tense (\texttt{<past>}) or it could be past participle (\texttt{<pp>}), so it has two 
-  analyses. Each analysis of word is separated by \texttt{\^{}} and \texttt{\$} symbols, and for one word each lexical 
-  forms are delimited by ``/'', and tags(<...>) show grammatical attributes of lexical form. 
+  and inflection information. 
+%  In this example each source form has been analysed as one or more lexical forms: ``I'' is analysed 
+%  into ``I'', where lexical category is subject pronoun (\texttt{<prn><subj>}) and first person (\texttt{<p1>}), could be 
+%  masculine or feminine (\texttt{<mf>}), singular (\texttt{<sg>}); ``played'' is analysed into lexical verb (\texttt{<vblex>}) 
+%  ``play'' with past simple tense (\texttt{<past>}) or it could be past participle (\texttt{<pp>}), so it has two 
+%  analyses. Each analysis of word is separated by \texttt{\^{}} and \texttt{\$} symbols, and for one word each lexical 
+%  forms are delimited by ``/'', and tags(<...>) show grammatical attributes of lexical form. 
 \item[Part-of-speech (POS) tagger:] This module chooses one of lexical forms of ambiguous word. As can be 
@@
   result of the application of  on constraint-grammar rules \citep{karlsson95}, which are used to discard some 
-  analyses using rules based on context. For example, consider the morphological analysis of word ``book'':
-\begin{verbatim}
-^book/book<n><sg>
-/book<vblex><inf>
-/book<vblex><pres>$
-\end{verbatim}
-  This word has 3 lexical forms, and depends on context, ont of them will be chosen by rules or tagger. After  this module, 
-  all words have only one morphological analysis.
+  analyses using rules based on context. 
+%For example, consider the morphological analysis of word ``book'':
+%\begin{verbatim}
+%^book/book<n><sg>
+%/book<vblex><inf>
+%/book<vblex><pres>$
+%\end{verbatim}
+%  This word has 3 lexical forms, and depends on context, ont of them will be chosen by rules or tagger. 
+After  this module, all words have only one morphological analysis.
 \item[Lexical transfer:] It reads each lexical form of source language and delivers a corresponding target language 
-  lexical form. This module uses a bilingual dictionary. For the English-Kazakh language pair, each lexical form of 
-  English word is translated into Kazakh as follows:
-\begin{verbatim}
-^I<prn><subj><p1><mf><sg>/мен<prn><pers><subj><p1><mf><sg>$ 
-^play<vblex><past>/ойна<v><tv><past>$
-\end{verbatim}
+  lexical form. This module uses a bilingual dictionary. 
+%For the English-Kazakh language pair, each lexical form of 
+%  English word is translated into Kazakh as follows:
+%\begin{verbatim}
+%^I<prn><subj><p1><mf><sg>/мен<prn><pers><subj><p1><mf><sg>$ 
+%^play<vblex><past>/ойна<v><tv><past>$
+%\end{verbatim}
  Multiword units can be translated as a single word.
@@
   number, gender, etc.). It uses  files with rules, which specify  the syntactic transformation such as word reorderings, 
-  lexical changes such as changes in prepositions and agreement between target language lexical forms. Transfer rules can 
-  produce new sequences for the  target language, for instance, preposition-noun rule is used to built sequence, where for 
-  noun is chosen right case, which depends on preposition: 
-\begin{verbatim}
-in garden  --  ^бақша<n><sg><PXD><loc>$
- \end{verbatim} 
+  lexical changes such as changes in prepositions and agreement between target language lexical forms. %Transfer rules can 
+%  produce new sequences for the  target language, for instance, preposition-noun rule is used to build a sequence, where for 
+%  noun is chosen right case, which depends on preposition.
+%\begin{verbatim}
+%in garden  --  ^бақша<n><sg><PXD><loc>$
+% \end{verbatim} 
 \item[Morphological generator:] Delivers the sequence of target-language lexical forms, produced by the structural transfer, 
@@
 tag was \texttt{<CD>} after ``cleaning'' it will be set on nominative case; at this step, also carried out 
-agreement between noun or adjective and copula ``е''(Мен дәрігер+е<p1 person singualr>), vowel and consonants 
+agreement between noun or adjective and copula ``е''(Мен дәрігер+е<p1 person singular>), vowel and consonants 
 agreement between last word of question and question particle (Сіз келдіңіз + ме? - Сіз келдіңіз + бе?).
@@
 Evaluating the output of the system was measured with BLEU \citep{papineni02} and Word Error Rate 
-metric \citep{levenshtein/1966}. Was chosen text small English text, and by postediting output 
-of English-Kazakh system, was build parallel text.
+metric \citep{levenshtein/1966}. We chose a short text in English, and by postediting output 
+of English-Kazakh system, was build parallel text. The output of each of the machine
+translation systems was posteditted independently to avoid biasing one particular system.
 
@@
 
-We have presented the rule-based MT system from English to Kazakh.
-The current English-Kazakh machine translation system already successfully translates noun-phrases, verb-phrases, prepositional-phrases, and adjectival-phrases, and contains a good vocabulary for testing purposes.
-We plan to continue developing English-Kazakh pair, going to extend the coverage to 98\% on the corpora, which we used before.
-Additionally, we will improve the quality of translation by adding more rules, such like constraint grammar rules, structural transfer and lexical rules.
-The future plan is to use the created data with other open-source MT system with Turkic languages or system with Kazakh language on right side to make transfer systems between the Turkic or other language pairs.
-Related work is currently ongoing with Russian--Kazakh.
-Our system is available as free/open-source software and the whole system may be downloaded from sourceforge(apertium-eng-kaz1, apertium-kaz2).
+We have presented the rule-based MT system from English to Kazakh. The current English-Kazakh machine translation 
+system already successfully translates noun-phrases, verb-phrases, prepositional-phrases, and adjectival-phrases, and 
+contains a good vocabulary for testing purposes. 
+We plan to continue developing English-Kazakh pair, going to extend the coverage to 98\% on the corpora, 
+which we used before. Additionally, we will improve the quality of translation by adding more rules, such like 
+constraint grammar rules, structural transfer and lexical rules. The future plan is to use the created data with other 
+open-source MT system with Turkic languages or system with Kazakh language on right side to make transfer systems 
+between the Turkic or other language pairs. Related work is currently ongoing with Russian--Kazakh. 
+Our system is available as free/open-source software and the whole system may be downloaded 
+from SourceForge.\footnote{\url{https://svn.code.sf.net/p/apertium/svn/incubator/apertium-eng-kaz}}
 
@@
 
-%\section{References}
-%1 Сундетова А.М., Кәрібаева А.С., АПЕРТИУМ ПЛАТФОРМАСЫНДАҒЫ АҒЫЛШЫН-ҚАЗАҚ МАШИНАЛЫҚ АУДАРМАШЫ ҮШІН ЕКІТІЛДІ СӨЗДІКТІ ҚҰРУ.
-Матер.
-межд.
-научн.-практ.
-конф.
-«Применение информационно-коммуникационных технологий в образовании и науке», посвященной 50-летию Департамента информационно-коммуникационных технологий и 40-летию кафедры «Информационные системы» КазНУ им.
-аль-Фараби.
-22 ноября 2013г.
-– Алматы: Қазақ Университеті, 2013.
-– С.53-57.
-
 \bibliographystyle{apalike}

### 2015-03-02 engkaz-eamt-2015/engkaz-eamt-2015.tex
a bit over
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59189 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
+\begin{table*}
+  \centering
+  \begin{tabular}{|l|l|l|l|}
+    \hline 
+    \textbf{Structure} & \textbf{English} & \textbf{System} & \textbf{Translation} \\
+    \hline 
+    \multirow{3}{*}{Noun phrases} & \multirow{3}{*}{My difficult exercises} & Apertium & Менің қиын жаттығуларым \\
+                                  &                                         & Sanasoft & Менің қиын жаттықтырып *жатыр \\ % <not right POS>
+                                  &                                         & Google   & Менің қиын *жаттығулар \\ % <no 1 person possessive>
+    \hline 
+    \multirow{3}{*}{Adpositional phrases}  & \multirow{3}{*}{Under three big trees}  & Apertium & үш үлкен ағаштың астында \\
+                                           &                                         & Sanasoft & *Three үлкен *ағаштар астында \\ % <unk> <no genitive>
+                                           &                                         & Google   & үш үлкен *ағаштарды астында \\ % <no genitive>
+    \hline 
+    \multirow{3}{*}{Adjective phrases}    & \multirow{3}{*}{The most beautiful}      & Apertium & ең әдемі \\ 
+                                          &                                          & Sanasoft & *Көпшілік әдемі \\ % <not right translation> 
+                                          &                                          & Google   & ең әдемі \\
+    \hline
+    \multirow{3}{*}{Modal verbs}  & \multirow{3}{*}{I must pay}             & Apertium & Мен жүргізе аламын \\
+                                  &                                         & Sanasoft & Мен must pay \\ % <no translation>
+                                  &                                         & Google   & *Мен *төлеуі тиіс \\ % <no genitive> <no possessive>
+    \hline 
+  \end{tabular}
+
+  \caption{Qualitative evaluation}
+  \label{table:qualeval}
+\end{table*}
+
 \section{Conclusions}

### 2015-03-02 2015-eamt-assim/2015-eamt-assim.tex
fix
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59180 72bbbca6-d526-0410-a7d9-f06f51895060
Fixed table captions
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59192 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-\comment{EA: Francis, when I save cyrillic in article body, it gets replaced by ??. I committed a separate file with the table and the footnote that has the text.}
-
 \begin{table*}
   \begin{tabular}{|l|l|}
+  \hline
+%@@@  \textbf{Src (Tatar)} & \rus{???? ?????? ???????; ????? 7 ??????????? ??? ??????? ?????, ????? ??? ????? 100 ???? ??????? ???????.}
+%  \hline
+%  \textbf{MT (Russian)} & \rus{???????? ??????? ????; ??? 7 ??????????? ??????? ?????????? ?????? , ????? ??? ????? 100 ???? ??????? ????.}
      \hline
@@
 \includegraphics[width=0.9\textwidth,resolution=144]{appraise-scr}
- \caption{An example set of sentences in an online environment.}
+ \caption{An example set of sentences in an online environment. The task is 30\% Russian legal text.}
 \label{figure:screenshot}
@@
   \end{tabular}
-  \caption{An example candidate synonym.} 
+  \caption{An example candidate synonym. The original word \emph{asumir} may be replaced with \emph{correr} in this context.} 
   \label{table:syn}
@@
 \end{tabular}
-\caption {Tatar--Russian: Krippendorff Alpha measure} 
+\caption {Basque--Spanish: Krippendorff Alpha measure} 
 \label{table:alpha-eus-spa}

### 2015-03-02 engkaz-eamt-2015/engkaz-eamt-2015.tex
+2 examples
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59196 72bbbca6-d526-0410-a7d9-f06f51895060

@@
     \hline 
-    \multirow{3}{*}{Noun phrases} & \multirow{3}{*}{My difficult exercises} & Apertium & Менің қиын жаттығуларым \\
+    \multirow{6}{*}{Noun phrases} & \multirow{3}{*}{My difficult exercises} & Apertium & Менің қиын жаттығуларым \\
                                   &                                         & Sanasoft & Менің қиын жаттықтырып *жатыр \\ % <not right POS>
                                   &                                         & Google   & Менің қиын *жаттығулар \\ % <no 1 person possessive>
+                                  & \multirow{3}{*}{Conan Doyle}            & Apertium & *Дойлдың Конан \\
+                                  &                                         & Sanasoft & *Conan *Doyle \\ % <not right POS>
+                                  &                                         & Google   & Конан Дойл \\ % <no 1 person possessive>
     \hline 
@@
                                   &                                         & Google   & *Мен *төлеуі тиіс \\ % <no genitive> <no possessive>
+    \hline 
+    \multirow{3}{*}{Questions}    & \multirow{3}{*}{Is it right?}           & Apertium & *Екен *дұрыстың ол? \\ % extra word + extra case 
+                                  &                                         & Sanasoft & Бұл *түзуi? \\ % <not right translation>
+                                  &                                         & Google   & Бұл дұрыс па? \\
     \hline 

### 2015-03-02 engkaz-eamt-2015/engkaz-eamt-2015.tex
+1 example
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59197 72bbbca6-d526-0410-a7d9-f06f51895060

@@
     \hline
-    \multirow{3}{*}{Modal verbs}  & \multirow{3}{*}{I must pay}             & Apertium & Мен жүргізе аламын \\
+    \multirow{6}{*}{Modal verbs}  & \multirow{3}{*}{I must pay}             & Apertium & Мен жүргізе аламын \\
                                   &                                         & Sanasoft & Мен must pay \\ % <no translation>
                                   &                                         & Google   & *Мен *төлеуі тиіс \\ % <no genitive> <no possessive>
+                                  & \multirow{3}{*}{It must be James}       & Apertium & Бұл Джеймс *болады \\ % <should be "болуы керек"> 
+                                  &                                         & Sanasoft & *Оған Джеймс *must *болып *жатыр  \\ % <no translation>
+                                  &                                         & Google   & Джеймс болуы тиіс \\ % <no genitive> <no possessive>
     \hline 

### 2015-03-04 2015-eamt-assim/2015-eamt-assim.tex
update
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59193 72bbbca6-d526-0410-a7d9-f06f51895060
ref
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59194 72bbbca6-d526-0410-a7d9-f06f51895060
a modified version
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59209 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{graphicx}
+\usepackage{microtype} % This saves some space and makes nice typesetting choices
 
@@
 \begin{document}
+
 \maketitle
+\renewcommand{\baselinestretch}{0.97} % the undetectable 3% squeeze
 \comment{EA: has not it first been used in trosterud12?}
 \begin{abstract}
-This paper provides additional observations on the viability of a strategy proposed in 2012 for evaluation of machine translation for assimilation purposes.
-The strategy involves human evaluators, who are asked to restore key words (fill gaps) in sentences.
-The evaluation method is applied to two translation directions.
-To reduce the amount of time required to prepare tasks and analyze results, an open-source task management system is introduced.
-The evaluation results suggest that the gap-filling task reflects users' understanding of text, and may be used to measure MT quality for assimilation purposes.
+  This paper provides additional observations on the viability of a
+  strategy independently proposed in 2012 and 2013 for evaluation of
+  machine translation for assimilation purposes. The strategy involves
+  human evaluators, who are asked to restore key words (to fill gaps) in
+  reference translations. The evaluation method is applied to two translation
+  directions.\comment{MLF: directions \(\to\) language pairs?} To reduce the amount of time required to prepare tasks
+  and analyze results, an open-source task management system is
+  introduced. The evaluation results suggest that the gap-filling task
+  reflects users' understanding of text,\comment{MLF: ``suggests''?
+I think this is what we assume to interpret our results: how about ``shows''?
+We don't show correlation with other evaluation.} and may be used to measure MT
+  quality for assimilation purposes.
 \end{abstract}
@@
 \comment{EA: I'd like a citation for the ``most massive application'', could you suggest something? It is from Mikel's paper, but no reference there -- should I cite Mikel?}
-Despite the fact that assimilation (or gisting) is currently the most frequent application of MT, few methodologies are established for assimilation evaluation of MT.
-The methods include post-editing and comparison by bilingual experts \citep{ginesti09}, and multiple choice tests \citep{jones07,trosterud12}.
-These approaches are often costly and prone to subjectivity, see the discussion in \cite{oregan13}.
-As an alternative, the modification of Cloze \citep{taylor53} test was introduced for assimilation evaluation, first by \citep{trosterud12} as a supplementary, and then by \citep{oregan13} as a stand-alone method.
-Prior to this, Cloze tests have been used to evaluate raw MT quality \citep{vanslype79}, \citep{somers00}.
-As a principal difference, \citet{trosterud12} and \citet{oregan13} propose to fill gaps in the reference (human) translation.
-The gap-filling task models how well users comprehend the key points of the text.
-Thus, the method does not directly evaluate the quality of machine-produced text, but rather its usefulness in understanding the meaning of the original text.
+\comment{MLF: I added a footnote, please reduce and move citation to the reference}
+Despite the fact that, as a result of widespread usage of online MT, assimilation (or gisting) is currently the most frequent application of MT,\footnote{“In a given day we [Google] translate roughly as much text as you’d find in 1 million books.
+To put it another way: what all the professional human translators in the world produce in a year, our system translates in roughly a single day.”, Franz J.
+Och, \url{http://tinyurl.com/och-gisting}} few methodologies are established for assimilation evaluation of MT.
+The methods include post-editing and comparison by bilingual experts \citep{ginesti09},\comment{MLF: should we also cite the original WMT reference cited by Ginestí et al.?} and multiple choice tests \citep{jones07,trosterud12}.
+These approaches are often costly and prone to subjectivity: see the discussion in \cite{oregan13}.
+As an alternative, the modification of \emph{cloze}\comment{\emph{cloze} stands for \emph{closure}.} testing \citep{taylor53} was introduced for assimilation evaluation, first by \citep{trosterud12} as a supplementary, and then by \citep{oregan13} as a stand-alone method.
+Prior to this, cloze tests have been used to evaluate raw MT quality \citep{vanslype79,somers00}.
+While these authors ask informant to fill gaps in MT output, \citet{trosterud12} and \citet{oregan13} propose to fill gaps in the reference (human) translation.
+The gap-filling task models how well users comprehend the key points of the text, as it is roughly equivalent with answering questions.\comment{MLF: check if you agree with this} Thus, the method does not directly evaluate the quality of machine-produced text, but rather its usefulness in understanding the meaning of the original text.
 
 \comment{EA: I understand we selected tat-rus and eng-kaz because they were in development. why did we select eu-es?}
-The gap-filling method has been successfully used to evaluate the Basque-English Apertium language pair.
-In this work we extend the evaluation to two more language pairs: Basque-Spanish and Tatar-Russian.
-For the latter pair, the evaluation served as a quality check in the period of active development during the Google Summer of Code'14 programme.
+The gap-filling method has been successfully used to evaluate the Basque--English Apertium language pair.
+In this work we extend the evaluation to two more language pairs: Basque--Spanish and Tatar-Russian.
+For the latter pair, the evaluation served as a quality check in the period of active development during the Google Summer of Code 2014 programme.
 In addition to evaluating, we explore new aspects of the experiment: the correlation between evaluators' scores, and the influence of the texts' linguistic domain.
-To facilitate the evaluation, we introduce an automated system which creates task sets from parallel corpora given a range of parameters (number of gaps in a sentence, hint type, gap filler, etc.), checks evaluators' answers, and calculates and formats generalized results.
-This system is integrated into the Appraise MT evaluation platform \citep{federmann12}; the code is open source and is available on Github\footnote{}.
+To facilitate the evaluation, we introduce an automated system which creates task sets from parallel corpora given a range of parameters (number of gaps in a sentence, hint type, gap filler,\comment{MLF: gap filler or gap marker?} etc.), checks evaluators' answers, and calculates and reports generalized results.
+This system is integrated into the Appraise MT evaluation platform \citep{federmann12}; the code is open source and is available on GitHub.\footnote{\phantom{\url{https://github.com/Sereni/Appraise}}}\comment{MLF: check if phantom URL is ok}
 
-We anticipate that the assessed MT systems will contribute to the users' understanding of text, that is, the evaluation will show better results in MT-assisted task type.
-We also expect to see different results depending on such parameters as text domain and the number of gaps in a sentence in relation to sentence length.
+We anticipate that the assessed MT systems will contribute to the users' understanding of text, that is, the evaluation will show better results in MT-assisted task type.\comment{MLF: this last sentence is very hard to follow} We also expect to see different results depending on such parameters as text domain and the number of gaps in a sentence in relation to sentence length.
 
@@
 \begin{table*}
-  \begin{tabular}{|l|l|}
-  \hline
+  \begin{tabular}{|l|p{15cm}|}
 %@@@  \textbf{Src (Tatar)} & \rus{???? ?????? ???????; ????? 7 ??????????? ??? ??????? ?????, ????? ??? ????? 100 ???? ??????? ???????.}
 %  \hline
-%  \textbf{MT (Russian)} & \rus{???????? ??????? ????; ??? 7 ??????????? ??????? ?????????? ?????? , ????? ??? ????? 100 ???? ??????? ????.}
+%  \textbf{MT (Russian)} & \rus{Половина сказать часы; вам 7 тукталыштан поздний опуститься нужный , после его снова 100 метр сказать идти.} \\
      \hline
+     \textbf{Ref.} & \rus{Примерно полчаса; вам нужно выйти через 7 остановок, потом пройти ещё около 100 метров.} \\\hline
      \textbf{10\%}   & \rus{Примерно полчаса; вам нужно выйти через 7 \{ \}, потом пройти ещё около 100 метров.} \\
@@
   \end{tabular}
-  \caption{Example of different gap percentage settings.} 
+  \caption{Example of different gap percentage settings for a Russian reference sentence.} 
   \label{table:percentage}
 \end{table*}
+\comment{MLF the gap percentage table 2 is not referenced in the text; also, i have changed it to have the \textbf{Ref} row in the table?}
 
@@
 The additional elements of the task are what we call hints, or extra sentences that help the participant to understand 
-the main sentence. There are two types of hints: first, the \emph{source}, which is semantically identical to \emph{reference}, 
+the main sentence. There are two types of hints: first, the \emph{source}, which is semantically equivalent to the \emph{reference}, 
 also human-produced, but in the source language of the pair. The second type is the \emph{machine-translated} hint, which 
-comes from the machine translation of the \emph{source} sentence. Table~\ref{table:example} shows a sample task, and Figure~\ref{figure:screenshot} shows the task in the online evaluation environment.
+comes from the machine translation of the \emph{source} sentence.
+Table~\ref{table:example} shows a sample task, and Figure~\ref{figure:screenshot} shows the task in the online evaluation environment.\comment{Use the highest possible resolution for the figure so that aliasing is imperceptible}
 
@@
 \includegraphics[width=0.9\textwidth,resolution=144]{appraise-scr}
- \caption{An example set of sentences in an online environment. The task is 30\% Russian legal text.}
+ \caption{An example set of sentences in an online environment. The task is Russian legal text with 30\% gaps.}
 \label{figure:screenshot}
@@
 \comment{EA: I don't really have a good explanation why we used source+mt tasks, except for being exhaustive. could you suggest anything?}
-In the course of the experiment, the following hint combinations are offered:
+In the course of the experiment, following \cite{oregan13}, the following hint combinations are offered:
 \begin{description}
@@
 that can be completed using common knowledge or language intuition (e.g.
-idioms and strong collocations);
+idioms and strong collocations).
 \item[Reference sentence and source sentence:] By setup, the
 participants have no command of the source language, however, it may help them to fill
-in proper nouns or loan words;
+in proper nouns or loan words.
 \item[Reference sentence and MT hint:] In addition to the task, the
@@
 used for measuring the contribution of machine translation to understanding the
-gist of text;
+gist of text.
 \item[Reference sentence and both hints:] This task is
@@
 
-In order to prepare the evaluation questions, we determine and remove the keywords in the 
-reference sentences. We consider two parameters: the list of allowed parts of speech (POS), and the 
+In order to prepare the evaluation questions, we determine and remove the keywords\comment{MLF: decide, \emph{key words} or \emph{keywords} everywhere} in the 
+reference sentences. We consider two parameters: the list of allowed parts of speech (PoS), and the 
 number of gaps relative to sentence length (``gap density''). For the evaluations described in 
@@
 For each sentence, the list of candidate keywords is prepared. It is composed of all the words that 
-fall into the allowed POS list. The number of gaps in the sentence is calculated based on 
-sentence length and specified gap density. All reference sentences are over 10 words in length. Finally, the required number of keywords is selected 
+fall into the allowed PoS list. The number of gaps in the sentence is calculated based on 
+sentence length and specified gap density. All reference sentences are longer than 10 words. Finally, the required number of keywords is selected 
 from the candidate list in such a manner that the gaps are distributed evenly throughout the sentence.
@@
 If the word is not a keyword, or has already been removed, we look at the next word instead.
-The process is repeated until the designated number of words has been removed, or until there are no more words in the keyword list.
+The process is repeated until the designated number of words has been removed, or until there are no more words in the keyword list.\comment{MLF: O'Regan and Forcada did something lots simpler: why not here?
+It is probably a bit late to change this}
 
@@
 It normally requires human effort, because we would like to determine the words that contribute the most to understanding the text as opposed to removing random words.
-In our setup, the above procedure is performed by a script integrated into the task generation pipeline.
+\comment{MLF: the procedure described so far seems automatic.
+What's manual and time consuming?} In our setup, the above procedure is performed by a script integrated into the task generation pipeline.
 Parts of speech are determined with Apertium morphological analyser.
-To control for homonymy, we only allow the word into the candidate list if all of its possible part of speech attributions are on the POS list.
+To control for homonymy, we only allow the word into the candidate list if all of its possible part of speech attributions are on the PoS list.
 For example, if we only allow nouns on the word list, and the word "fly" receives two possible part of speech attributions from the tagger, noun and verb, it is not considered for the candidate list.
@@
 Spanish and Russian) who had no command of source language of the pair (Basque
-and Tatar, respectively). 11 evaluators participated in the in Basque-Spanish experiment, and 28 in Tatar-Russian (although not everyone completed the task
+and Tatar, respectively). Eleven evaluators participated in the in Basque--Spanish experiment, and 28 in Tatar-Russian (although not everyone completed the task
 in full, see discussion).
@@
 \begin{enumerate}
-\item  For Basque-Spanish, from the corpus of legal texts ``Memorias de traducci\'on del
-Servicio Oficial de Traductores del IVAP''\footnote{\url{http://tinyurl.com/tm-ivap}};
+\item  For Basque--Spanish, from the corpus of legal texts ``Memorias de traducci\'on del
+Servicio Oficial de Traductores del IVAP'';\footnote{\url{http://tinyurl.com/tm-ivap}}
 \item  For Tatar-Russian, from the following sources on three different topics:
@@
 
-Each set features 36 pairs of sentences. For the Basque-Spanish experiment the pairs were drawn
-randomly from the corpora, for Tatar-Russian, compiled by hand by the developer of the
+Each set features 36 pairs of sentences. For the Basque--Spanish experiment the pairs were drawn
+randomly from the corpora, for Tatar--Russian, compiled by hand by the developer of the
 language pair in Apertium.
@@
 The evaluations took place online, in a system called Appraise \citep{federmann12}, which is 
-designed specifically for various MT evaluation tasks. We adapted the code to 
+designed specifically for various MT evaluation tasks. We adapted the code of Appraise to 
 accommodate for the gap-filling tasks. The tasks were uploaded into the system and 
@@
 a total of 36;
-\item  Each participant evaluates 9 sentences in each of the four modes (see section~\ref{sec:methodology});
-\item  All sentences of the set are evaluated with 10, 20 and 30\% of words removed;
-\item Each participant may encounter a given sentence in only one of the percentage variations;
-\item  Each sentence-mode-percentage combination is evaluated by more than one participant.
+\item  each participant evaluates 9 sentences in each of the four modes (see section~\ref{sec:methodology});
+\item  all sentences of the set are evaluated with 10, 20 and 30\% of words removed;
+\item each participant may encounter a given sentence in only one of the percentage variations;
+\item  each sentence-mode-percentage combination is evaluated by more than one participant.
 \end{enumerate}
@@
 The participants are given the instructions in their native language; these instructions are repeated above each task in the evaluation system.
-For the participants' convenience, the body of questions is split into smaller groups which allow indefinite breaks between them.
+For the participants' convenience, the body of questions is split into smaller groups which allow indefinite time breaks\comment{MLF: time break OK?
+Possibility: \emph{multiple sessions}.} between them.
 The instructions are the
@@
 provide words that fit the gaps well, but do not match the original answer. To account for
-these cases, we process all the answers to detect possible synonyms (a method suggested by \citet{oregan13}. An answer is
+these cases, we process all the answers to detect possible synonyms (a method suggested by \cite{oregan13}). An answer is
 considered a candidate synonym if it is given by two or more evaluators, and it does not match
@@
 \comment{EA: added}
-The synonym lists for Basque-Spanish and Tatar-Russian contain 52 and 25 words, respectively.
-Time taken to compile one list depends on the number of candidate synonyms, and in our case was approximately 30 minutes.
+The synonym lists for Basque--Spanish and Tatar--Russian contain 52 and 25 words, respectively.
+Time taken to compile each list depends on the number of candidate synonyms, and in our case was approximately 30 minutes.
 \comment{FMT: How long did making the synonym list take, how many synonyms were added? More details here. Also perhaps an example?}
@@
 To reduce the noise from participants who were distracted during evaluation, when
-calculating times we remove all the results over 6 minutes (the mode is approximately
+calculating times we remove all the results over 6 minutes (the statistical mode is approximately
 two minutes). Typical time taken to complete one question varies from under one minute for tasks without hints and few gaps, to approximately two minutes for tasks with more hints and gaps.
@@
 % expected results
-We expect to see the following trend: scores received in different task modes inside one gap density are observed, from highest to lowest, in tasks with MT and source hint, then MT hint only, then source hint only, and finally, no hint.
-We also expect that with the increase in gap density, time taken to fill the gaps also increases, and percentage of correct answers decreases.
+We expect to see the following trend: scores received in different task modes inside one gap density should be observed, from highest to lowest, in tasks with MT and source hint, then MT hint only, then source hint only, and finally, no hint.
+We also expect that with the increase in gap density, time taken to fill the gaps should also increase, and percentage of correct answers should decrease.
 
@@
 difficult to predict the answer based on the context, and also leaves more room for
-translation mistakes.
+mistakes.
 Exploring different percentage-mode combinations, we may note that the 10\% no-hint tasks take the least time to complete.
@@
 The second pattern, low scores in Tatar--Russian 10\% MT+source, does not stem from the task content.
-Instead, it is the result of the fixed order of tasks: the participants have always been given MT+source 10\% sentences first, followed by other task types.
-The participants have not received any training tasks before the main evaluations.
+Instead, it is the result of the fixed order of tasks: the participants have always been given MT+source 10\% sentences first, followed by other task types.\comment{MLF: Why?
+Not random?
+\cite{oregan13} did random!
+Why was this changed?} The participants have not received any training tasks before the main evaluations.
 Therefore, it is possible that the accommodation period is responsible for lower-than-expected scores in this mode of evaluation.
@@
 that some content words are easier to fill than the others. This may explain why in
-Basque-Spanish the 20\% MT tasks are completed with better accuracy than 10\% tasks.
+Basque--Spanish the 20\% MT tasks are completed with better accuracy than 10\% tasks.\comment{MLF: I think this shows that our statistics are weak, should we say something to excuse ourselves?}
 
@@
 
-\subsection{Annotator agreement}
+\subsection{Annotator agreement}\comment{MLF cannot get these tables to appear numbered later than the other ones!}
 After obtaining the results we calculated Krippendorff's alpha~\citep{krippendorff70} measure to represent annotator agreement, shown in Tables~\ref{table:alpha-eus-spa} and~\ref{table:alpha-tat-rus}.
-We selected this measure because of its compatibility with more than two annotators per task and the missing data (not all the gaps were evaluated).
-To calculate Krippendorff's alpha we used an algorithm implementation by Thomas Grill\footnote{http://grrrr.org}, dividing the answers in each gap into two categories: correct and incorrect.
-The previously obtained synonym lists were taken into account, i.e.
-if the two answers are different but both correct, they fall into one category.
-The measure was calculated separately for each hint and percentage combination.
-
-As the MT and MT+source hints are introduced, the agreement between annotators increases (measures closer to 1): the annotators are more consistently correct or incorrect in each given sentence.\comment{EA: I cite Juan Antonio here, because I am not completely sure what alpha shows about our experiment (lack of knowledge on my part)}  The agreement measure for the same sentences without hints is closer to zero, which contributes to the reliability of our methodology.
-We note the outlier score in Tatar--Russian 10\% source tasks, which has the most contribution from the news texts (agreement of 0.917).
-This set of sentences contains many loan words, which have similar form in Tatar and Russian (e.g.
-president, minister, championship), and are understood by Russian speakers.
-The gaps with loan words have mostly been filled correctly, while there was some disagreement in other gaps.
-
 \begin{table*}
@@
 \end{table*}
+We selected this measure because of its compatibility with more than two annotators per task and the missing data (not all the gaps were evaluated).
+To calculate Krippendorff's alpha we used an algorithm implementation by Thomas Grill,\footnote{http://grrrr.org}\comment{MLF: Give a direct reference: I could not find the algorithm in such a wide website.
+Use tinyurl if it is too long.} dividing the answers in each gap into two categories: correct and incorrect.
+The previously obtained synonym lists were taken into account, i.e.
+if the two answers are different but both correct, they fall into one category.
+The measure was calculated separately for each hint and percentage combination.
+
+As the MT and MT+source hints are introduced, the agreement between annotators increases (measures closer to 1): the annotators are more consistently correct or incorrect in each given sentence.\comment{EA: I cite Juan Antonio here, because I am not completely sure what alpha shows about our experiment (lack of knowledge on my part)}  The agreement measure for the same sentences without hints is closer to zero, which contributes to the reliability of our methodology.
+We note the outlier score in Tatar--Russian 10\% source tasks, which has the most contribution from the news texts (agreement of 0.917).
+This set of sentences contains many loan words, which have similar form in Tatar and Russian (e.g.
+president, minister, championship), and are understood by Russian speakers.
+The gaps with loan words have mostly been filled correctly, while there was some disagreement in other gaps.
 
@@
 \comment{EA: and what would we like to say about those?
-The results depend heavily on the Tatar-Russian system.
+The results depend heavily on the Tatar--Russian system.
 Do we have to provide info on what vocabulary it handles better, what it has and has not been trained on?
 Or do we just say that casual texts are easier to fill than legal texts?}
-For Tatar--Russian language pair the participants were offered texts from three different domains (in equal proportions): casual conversations, legal texts and news.
+For the Tatar--Russian language pair the participants were offered texts from three different domains (in equal proportions): casual conversations, legal texts and news.
 The results by domains are displayed in tables \ref{table:rus-casual}, \ref{table:rus-legal}, and \ref{table:rus-news}.
@@
 \hline
-10\%&62.19 +/- 32.07&58.04 +/- 27.96&39.90 +/- 38.96&48.79 +/- 39.84\\
-20\%&64.81 +/- 30.03&70.18 +/- 26.98&31.37 +/- 28.42&30.91 +/- 30.22\\
-30\%&48.13 +/- 26.08&39.55 +/- 23.90&25.94 +/- 19.67&18.07 +/- 18.15\\
+10\%&62.19 \(\pm\) 32.07&58.04 \(\pm\) 27.96&39.90 \(\pm\) 38.96&48.79 \(\pm\) 39.84\\
+20\%&64.81 \(\pm\) 30.03&70.18 \(\pm\) 26.98&31.37 \(\pm\) 28.42&30.91 \(\pm\) 30.22\\
+30\%&48.13 \(\pm\) 26.08&39.55 \(\pm\) 23.90&25.94 \(\pm\) 19.67&18.07 \(\pm\) 18.15\\
 \hline
@@
 \hline
-10\%&63.76 +/- 41.15&57.02 +/- 41.57&53.52 +/- 43.57&45.96 +/- 40.86\\
-20\%&59.77 +/- 33.10&64.93 +/- 30.95&45.60 +/- 31.23&38.56 +/- 32.11\\
-30\%&55.83 +/- 26.13&58.77 +/- 28.07&40.00 +/- 28.16&35.45 +/- 29.76\\
+10\%&63.76 \(\pm\) 41.15&57.02 \(\pm\) 41.57&53.52 \(\pm\) 43.57&45.96 \(\pm\) 40.86\\
+20\%&59.77 \(\pm\) 33.10&64.93 \(\pm\) 30.95&45.60 \(\pm\) 31.23&38.56 \(\pm\) 32.11\\
+30\%&55.83 \(\pm\) 26.13&58.77 \(\pm\) 28.07&40.00 \(\pm\) 28.16&35.45 \(\pm\) 29.76\\
     \hline
@@
 \textbf{Gap percentage} & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
-10\%&63.64 +/- 44.84&63.64 +/- 43.12&62.12 +/- 46.11&53.03 +/- 44.28\\
-20\%&73.33 +/- 31.51&63.33 +/- 35.85&41.11 +/- 27.88&37.78 +/- 30.94\\
-30\%&70.28 +/- 31.12&59.58 +/- 23.89&39.44 +/- 27.10&37.71 +/- 33.32\\
+10\%&63.64 \(\pm\) 44.84&63.64 \(\pm\) 43.12&62.12 \(\pm\) 46.11&53.03 \(\pm\) 44.28\\
+20\%&73.33 \(\pm\) 31.51&63.33 \(\pm\) 35.85&41.11 \(\pm\) 27.88&37.78 \(\pm\) 30.94\\
+30\%&70.28 \(\pm\) 31.12&59.58 \(\pm\) 23.89&39.44 \(\pm\) 27.10&37.71 \(\pm\) 33.32\\
 \hline
@@
 \textbf{Gap percentage} & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
-10\%&53.02 +/- 40.31&68.21 +/- 34.99&38.77 +/- 37.82&32.95 +/- 34.88\\
-20\%&60.57 +/- 24.98&65.76 +/- 24.16&50.28 +/- 33.55&48.31 +/- 33.52\\
-30\%&50.06 +/- 26.25&48.30 +/- 28.87&40.05 +/- 27.37&34.15 +/- 28.73\\
+10\%&53.02 \(\pm\) 40.31&68.21 \(\pm\) 34.99&38.77 \(\pm\) 37.82&32.95 \(\pm\) 34.88\\
+20\%&60.57 \(\pm\) 24.98&65.76 \(\pm\) 24.16&50.28 \(\pm\) 33.55&48.31 \(\pm\) 33.52\\
+30\%&50.06 \(\pm\) 26.25&48.30 \(\pm\) 28.87&40.05 \(\pm\) 27.37&34.15 \(\pm\) 28.73\\
 \hline
@@
 \textbf{Gap percentage} & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
-10\%&53.33 +/- 37.86&59.77 +/- 43.66&57.47 +/- 42.12&49.44 +/- 39.08\\
-20\%&59.33 +/- 33.89&48.56 +/- 35.40&46.53 +/- 31.99&28.97 +/- 28.79\\
-30\%&58.14 +/- 22.42&61.49 +/- 22.11&40.51 +/- 30.10&34.81 +/- 27.01\\
+10\%&53.33 \(\pm\) 37.86&59.77 \(\pm\) 43.66&57.47 \(\pm\) 42.12&49.44 \(\pm\) 39.08\\
+20\%&59.33 \(\pm\) 33.89&48.56 \(\pm\) 35.40&46.53 \(\pm\) 31.99&28.97 \(\pm\) 28.79\\
+30\%&58.14 \(\pm\) 22.42&61.49 \(\pm\) 22.11&40.51 \(\pm\) 30.10&34.81 \(\pm\) 27.01\\
 \hline
@@
 \comment{EA to JA: by our repository I mean my github, is that ok?}
-experiment may be downloaded from our repository\footnote{}.
+experiment may be downloaded from our repository.\footnote{}\comment{MLF: Do the same as in the other GitHUb comment above}
 
@@
 
-The experiment may be repeated for any language pair (provided a parallel corpus) and
+The experiment may easily be repeated for any language pair (provided a parallel corpus) and
 any machine translation system. Based on our experience, we would like to suggest the following amendments to the procedure:
@@
 \item If the experiment is conducted with different gap densities, each consecutive set of sentences should be derived from the set with lower gap density.
-To put another way, the sentence with e.g.
-20\% gaps should contain all the gaps of the same sentence in 10\% mode, plus any additional gaps;
+To put another way, the sentence with e.g.\ 20\% gaps should contain all the gaps of the same sentence in 10\% mode, plus any additional gaps;\comment{MLF: no reason is given for this, I am not ready to agree.
+Consider removing this}
 \item Unless the evaluation is targeted at a specific text domain, it may be beneficial to include a stylistic variety of texts in the initial corpus.
-Neighboring sentences on the same topic may assist the users in gap-filling tasks;
+Neighboring sentences on the same topic may assist the users in gap-filling tasks;\comment{wasn't this already reported by \cite{oregan13}?}
 \item If possible, increase the number of evaluators, or reduce the number of questions per participant.
 In the above experiments each participant filled from 110 to 187 gaps, divided into small groups.
-Reducing the amount of work may increase task completion rate;
+Reducing the amount of work may increase task completion rate;\comment{MLF's crazy idea: recaptchas?}
 \item To account for the adaptation period, provide training tasks before the main evaluations take place.
@@
 
-This work has been partly funded by the Spanish Ministerio de Econom{\'i}a y Competitividad through project TIN2012-32615 and by the Google Summer of Code programme.
+This work has been partly funded by the Spanish Ministerio de Econom{\'i}a y Competitividad through project TIN2012-32615 and by the Google Summer of Code 2014 programme.
 We would like to thank the volunteers who participated in the evaluations.

### 2015-03-04 engkaz-eamt-2015/engkaz-eamt-2015.tex
making changew
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59212 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{listings}
+\usepackage{microtype}
 
@@
 \begin{abstract}
-This paper presents rule-based machine translation from English to Kazakh. Main differences 
-between two languages in syntax and morphology are presented, and showed how machine translation 
-system was designed to solve these challenges. We showed an evaluation of system coverage and 
-translation quality and future work.
+This paper presents the current state of development a shallow-transfer rule-based machine translation (MT) system from English to Kazakh. The main syntactic and morphological differences 
+between the two languages are presented, and it is showed how the MT 
+system was designed to tackle these challenges. We show an evaluation of system coverage and 
+translation quality and outline and future work.
 \end{abstract}
@@
 
-This paper presents a shallow-transfer rule-based machine translation system from English 
-into Kazakh language. First of all, English language is a West Germanic language and Kazakh 
-belong to group of Turkic languages. Therefore, the complex agglutinative morphology of Kazakh 
-language is very different from morphologically not too complex language like English; on the other 
-hand, Turkic language morphology shows clear morphotactics (ordering of morphemes), its morphophonology 
-shows complex phonological changes to due to interactions between neighboring morphemes (vowel harmony, 
+This paper presents a shallow-transfer rule-based machine translation (MT) system from English 
+(a West Germanic language) to Kazakh (a Turkic language)
+The complex agglutinative morphology of Kazakh 
+is in contrast with the simple morphology of English.
+Turkic language morphology shows clear morphotactics (ordering of morphemes), but with 
+complex phonological changes (morphophonology) to due to morpheme contact (vowel harmony, 
 sonorisation, etc.) many of which are explicitly represented in writing.
@@
 On the other hand, there are many differences between the syntax of Turkic languages 
-and English: subject-adverb–object–verb order (compare subject–verb–object-adverb in English), use of 
-postpositions (compare prepositions in English), head-final syntax with modifiers and specifiers 
-always preceding the modified/specified (normally following in English), overt case marking allowing 
-for a rather free ordering of arguments (versus a more fixed order in English), lack of definite 
+and English: subject–object–verb order (vs.\ subject–verb–object in English), use of 
+postpositions (vs.\ prepositions in English), head-final syntax with modifiers and specifiers 
+always preceding the modified/specified (vs.\ following in English), overt case marking allowing 
+for a rather free ordering of arguments (vs.\ a more fixed order in English), lack of definite 
 articles (extensively used in English), verbal-noun-centered structures where English uses modal 
-verbs (must, have to, want to) or verbal-noun or verbal-adjective-centred constructions where 
+verbs (\emph{must}, \emph{have to}, \emph{want to}) or verbal-noun or verbal-adjective-centred constructions where 
 English has subordinate clauses using finite verbs with relatives or subordinating 
-conjunctions (the book which I read, the place where I saw him, before he came), lack of a parallel 
-of the English verb \emph{have}, as used for possession, etc. For an account (in Russian) of 
+conjunctions (\emph{the book which I read}, \emph{the place where I saw him}, \emph{before he came}), lack of a parallel 
+of the English lexical verb \emph{to have}, etc. For an account (in Russian) of 
 syntactic differences between English and Kazakh, see \cite{pecherskih2012}.

### 2015-03-04 engkaz-eamt-2015/engkaz-eamt-2015.tex
Mikel's first ironing
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59213 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{eamt15}
-\usepackage{times}
+%\usepackage{times}
 \usepackage{latexsym}
@@
 \usepackage{polyglossia}
+\usepackage{todonotes}
 \setdefaultlanguage[variant=australian]{english}
@@
 \newcommand{\com}[1]{\marginpar{\scriptsize #1}} 
-
+    \setcounter{secnumdepth}{4}
 \title{A free/open-source machine translation system for English to Kazakh}
@@
 \maketitle 
+\renewcommand{\baselinestretch}{0.97} % the undetectable 3% squeeze
 
-\begin{abstract}
+\begin{abstract}\todo{Decide authors, and put them in phantom boxes}
 This paper presents the current state of development a shallow-transfer rule-based machine translation (MT) system from English to Kazakh. The main syntactic and morphological differences 
@@
 
-We currently build rule-based machine translation system, because for statistical machine translation 
-system we need large sentence-aligned parallel corpora. There are no freely available, large parallel 
-corpora for the English-Kazakh pair. English-Kazakh statistical MT system was built by Google,\footnote{\url{http://translate.google.com}} 
-other work in this language pair was made by building a small parallel corpus and using Moses \citep{assylbekov14}, 
-and existing commercial systems for English to Kazakh (Sanasoft\footnote{\url{http://www.sanasoft.kz/c/ru/node/47}} and Trident\footnote{\url{http://www.translate.ua/us/on-line}}) both appear to be rule-based.
+We set out to build a rule-based machine translation system, because for statistical machine translation 
+system we need large sentence-aligned parallel corpora, which are not freely available
+for the English--Kazakh pair. An English--Kazakh statistical MT system has however recently been built by Google,\footnote{\url{http://translate.google.com}}; 
+other work in this language pair was performed by building a small parallel corpus and using Moses \citep{assylbekov14}.
+Existing commercial systems for English to Kazakh (Sanasoft\footnote{\url{http://www.sanasoft.kz/c/ru/node/47}} and Trident\footnote{\url{http://www.translate.ua/us/on-line}}) both appear to be rule-based.
 Rule-based machine translation, in which developers 
@@
 unusual case will produce better translation.
-
+\todo{What does the last sentence mean? Clarify!}
 This paper describes work in progress of development of machine translation system 
-for English-to-Kazakh, which developed by using Apertium free/open-source machine 
+for English--Kazakh using the Apertium free/open-source machine 
 translation platform \citep{forcada11}. %(http://www.apertium.org).
@@
 \label{fig:modules}
-%\vspace{-1em}
+\vspace{-1em}
 \end{figure*}
@@
 Apertium \citep{forcada11} is a free/open-source rule-based machine translation (MT) platform that 
-has been developed since 2005 at the Universitat d'Alacant. At first, it was initially aimed at 
+has been developed since 2005, started at the Universitat d'Alacant. At first, it was initially aimed at 
 translating texts between closely related languages, then it was extended to deal with less-related 
-languages. This platform has next components: machine translation engine, developer's tools, and linguistic 
+languages. This platform has the following components: machine translation engine, developer's tools, and linguistic 
 data for an increasing number of language pairs and they are licensed under the 
 free/open-source GNU General Public License (GPL, versions 2 and 3).
-
-\begin{description}
+\todo{If space allowed, it would be nice to add a running example as in the Apertium paper}
+\begin{description}\itemsep 0ex
 \item[De-formatter:] Separates the text to be translated from the formatting tags.  Formatting tags are 
@@
   such a way that the remaining modules see them as regular blanks.  
-\item[Morphological analyser:] For each surface form the morphological analyser generates one or more 
-  lexical forms, which consist of: lemma (dictionary or citation form), lexical category (or part-of-speech), 
+\item[Morphological analyser:] For each surface form  (SF) the morphological analyser generates one or more 
+  lexical forms (LF), which consist of: lemma (dictionary or citation form), lexical category (or part-of-speech), 
   and inflection information. 
@@
 %  forms are delimited by ``/'', and tags(<...>) show grammatical attributes of lexical form. 
-\item[Part-of-speech (POS) tagger:] This module chooses one of lexical forms of ambiguous word. As can be 
-  seen from the previous example, the morphological analyser could deliver more than one lexical form and 
-  choosing wrong one could produce errors in translation. POS tagger is based on a statistical model, which 
-  based on hidden Markov models and it has been trained on source language texts, which processes the 
-  result of the application of  on constraint-grammar rules \citep{karlsson95}, which are used to discard some 
-  analyses using rules based on context. 
+\item[Part-of-speech (POS) tagger:] This module chooses one of the LF of an ambiguous SF. 
+%As can be 
+%  seen from the previous example, the morphological analyser could deliver more than one LF and 
+%  choosing wrong one could produce errors in translation. 
+\todo{I have removed some text and  the statistical PoS tagging stuff, as it is not used}
+%POS tagger is based on a statistical model, which 
+%  based on hidden Markov models and it has been trained on source language texts, which processes the 
+%  result of the application of  on constraint-grammar rules \citep{karlsson95}, which are used to discard some 
+%  analyses using rules based on context. 
 %For example, consider the morphological analysis of word ``book'':
@@
 %  This word has 3 lexical forms, and depends on context, ont of them will be chosen by rules or tagger. 
-After  this module, all words have only one morphological analysis.
-\item[Lexical transfer:] It reads each lexical form of source language and delivers a corresponding target language 
-  lexical form. This module uses a bilingual dictionary. 
-%For the English-Kazakh language pair, each lexical form of 
+%After  this module, all words have only one morphological analysis.
+\item[Lexical transfer:] It reads each source-language LF and delivers the corresponding target-language 
+  LFs. This module uses a bilingual dictionary. 
+%For the English--Kazakh language pair, each lexical form of 
 %  English word is translated into Kazakh as follows:
@@
  Multiword units can be translated as a single word.
-\item[ Lexical selection:] It uses rules for lexical words, which have many translations, to select one of the translations 
-  in the target language according to context. The system is described in \cite{tyers12a}.
-\item [Structural transfer:] This module uses pattern matching to identify sequences of lexical forms  (phrases or 
-  segments), which need syntactical processing to translate grammatical differences between two languages (handling of 
-  number, gender, etc.). It uses  files with rules, which specify  the syntactic transformation such as word reorderings, 
+\item[ Lexical selection:] It uses rules to select one of the target-language LFs, as described in \cite{tyers12a}.
+\item [Structural transfer:] This module uses pattern matching to identify sequences of LF (phrases or 
+  segments), which need syntactical processing 
+%to deal 
+%with grammatical differences between two languages (handling of 
+%  number, gender, etc.). 
+It uses files with rules, which specify  the syntactic transformation such as word reorderings, 
   lexical changes such as changes in prepositions and agreement between target language lexical forms. %Transfer rules can 
@@
 % \end{verbatim} 
-\item[Morphological generator:] Delivers the sequence of target-language lexical forms, produced by the structural transfer, 
-  to a corresponding sequence of target language surface forms. The morphological generator executes a finite-state transducer 
-  generated by compiling a morphological dictionary for the target language.     
-  For Kazakh language a finite-state transducer based on two-level rules \citep{washingtonsalimzyantyers14} is used for generation. 
-\item[ Post-generator:] Performs some minor orthographical operations in the target language (for instance, it generates the 
-  English form cannot from can and not). This module is generated from file with rules which are very   similar in format to dictionary files. 
+\item[Morphological generator:] Transforms the sequence of target-language LFs forms, produced by the structural transfer, 
+  to a corresponding sequence of target-language SFs. 
+% The morphological generator executes a finite-state transducer 
+For Kazakh, a finite-state transducer based on two-level rules \citep{washingtonsalimzyantyers14} is used for generation. 
+\item[Post-generator:] Performs some minor orthographical operations in the target language. This module is not used in the English--Kazakh system.\todo{Or is it?}
 \item[ Reformatter:] It places format tags back into the text so that its format is preserved.
@@
 
-Apertium uses XML-based formats for linguistic data, which include bilingual and monolingual dictionaries, structural transfer and
-lexical selection rules. % and rules for part-of-speech tagging.
+Apertium mainly uses XML-based formats for linguistic data, which include bilingual and monolingual dictionaries, structural transfer and
+lexical selection rules.
 
@@
 
-Dictionaries are used in lexical processing and morphological analysis, there are use two types: morphological (monolingual) and bilingual.
+Dictionaries are used in lexical processing: monolingual dictionaries
+for morphological analysis of English and generation of Kazakh and
+bilingual dictionaries for English--Kazakh lexical transfer.
 
-\subsubsection{Monolingual dictionaries}
+The English dictionary is used to determine lexical forms of each surface
+forms, it contains of list of all lexical units, definition of
+source-language alphabet, grammatical categories, which is used to
+specify lexical forms, for example, like noun, verb, gender, etc.;
+each category has paradigms describing groups of correspondences
+between parts of SFs and LFs.
 
-English dictionary uses to determine lexical forms of each surface forms, it contains of list of all lexical units, definition 
-of source-language alphabet, grammatical categories, which is used to specify lexical forms, for example, like noun, verb, 
-gender, etc., each categories have paradigms, they describes groups of correspondences between parts of surface forms and lexical forms.
+The Kazakh dictionary is turned into a
+morphological transducer, using the Helsinki Finite
+State Toolkit \citep{hfst/2011}. It uses the \texttt{lexc} formalism
+for defining lexicons through word classes and subclasses, and the
+\texttt{twol} formalism for morphophonological rules such as vowel
+harmony, desonorization, nasalization, etc.  Depending on the
+preceding vowels and consonants, the plural suffix \emph{-LAr}\footnote{Uppercase Latin letters are used for archiphonemes (actually archigraphemes) that are realised as phonemes (actually graphemes) after morphophonological rules have been applied} could
+become \emph{-тар}, \emph{-тер}, \emph{-lar}, \emph{-ler}, \emph{-dar}, \emph{-ler} depending on vowel harmony: kitap+tar,
+mektep+ter, etc. \todo{Aida, please typeset this example in Cyrillic; braces are not used here, but are used later; homogeneize notation in paper.}
 
-For Kazakh side, is used morphological transducers, which are based on the Helsinki Finite State Toolkit \citep{hfst/2011}. It 
-uses formalisms lexc for defining lexicons, twol formalism for morphophonological rules. Lexc formalism defines different 
-word classes and subclasses. And twol determines phonological transformation like vowel harmony, desonorisation. For example, 
-depend on preceding vowel, plural form \emph{-LAr} could become -тар or -тер depending on vowel harmony: kitap+tar, but mektep+ter.
+%\subsubsection{Bilingual dictionary}
 
-\subsubsection{Bilingual dictionary}
-
-Bilingual dictionary provides correspondences between lexical forms of source and target lanugages. Each entry 
-consists of English word on the left side with part of speech and corresponding Kazakh word on the right side. 
-Some entries could be ambiguous by part of speech, or just have many translations, which will be chosen by 
-lexical-selection rules depending on context. This dictionary currently contains 13,130 stems \citep{sundetova13a}.
+The English--Kazakh bilingual dictionary provides correspondences between English LFs and Kazakh LFs. Ambiguity is allowed: one of the LFs will be chosen by 
+lexical-selection rules depending on context. This dictionary currently contains 13,130 stems \citep{sundetova13a}.\todo{later you say 13,135}
 
@@
 
-\subsection{Disambiguation rules}
+\subsubsection{Disambiguation rules}
 
-This kind of rules solves the problem of part-of-speech (POS) ambiguity by Constraint Grammar (CG, \cite{karlsson95}). The 
-output of morphological analyser could be ambiguous, and in this case CG rules choose correct analysis from 
-multiple analyses. From English to Kazakh POS ambiguity could be only in English side, but also some choice also
-depend on Kazakh grammar.
+This kind of rules solves the problem of part-of-speech (PoS) ambiguity using the  Constraint Grammar (CG, \cite{karlsson95}) formalism.
+CG rules choose one of the LFs obtained for each SF.
+%From English to Kazakh POS ambiguity could be only in English side, but also some choice also
+%depend on Kazakh grammar.\todo{I removed an unclear sentence here}
 
-\subsection{Lexicon selection rules }
+\subsubsection{Lexical selection rules }
 
-Lexical-selection rules choose one of the alternative translations of one word depend on the context. The rule 
-formalism is described in \cite{tyers12a}. Alternative translations are defined in bilingual dictionary by 
-multiple entries of one word. For exmaple, adjective ``beautiful'' could be translated as ``сұлу'', if 
-following noun is subject: ``beautiful girl'' -- ``сұлу қыз''; or ``әдемі'', ``көркем'', if the noun means 
-place or: ``beautiful mountains'' -- ``әдемі таулар''.  Table~\ref{table:lexsel} demonstrates some 
-lexical-selection rules:
-
+Lexical-selection rules choose one of the alternative target-language LFs corresponding to one source-language LF, as described in \cite{tyers12a}.
+Alternative translations are defined in bilingual dictionary by
+multiple entries for each source-language LF. For example, the adjective \emph{beautiful} could be translated as \emph{сұлу}, if 
+the following noun is a person: \emph{beautiful girl} \(\to\) \emph{сұлу қыз}; or as \emph{әдемі} or  \emph{көркем}, if the following noun means 
+place or \todo{or what?} \emph{beautiful mountains} \(\to\) \emph{әдемі таулар}.  Table~\ref{table:lexsel} shows some 
+lexical-selection rules:\todo{but this rule is not in the table!!!}
 \begin{table*}
@@
     \hline
-    \textbf{Word} & \textbf{Translation} & \textbf{Condition} & \textbf{Example} \\
+    \textbf{SL word \(w\)} & \textbf{TL word} & \textbf{Context} & \textbf{Example} \\
     \hline
-    \multirow{4}{*}{residence} & \multirow{2}{*}{мекен}      & \multirow{2}{*}{--} & I live in a residence. \\
+    \multirow{4}{*}{residence} & \multirow{2}{*}{мекен}      & \multirow{2}{*}{default} & I live in a residence. \\
                                &                             &                     & (Мен) мекенде өмір сүремін. \\
-                               & \multirow{2}{*}{резиденция} & \multirow{2}{*}{\_ of \_ president} & I see the residence of the president \\ 
+                               & \multirow{2}{*}{резиденция} & \multirow{2}{*}{\(w\)\_ of \_ president} & I see the residence of the president \\ 
                                &                             &                                     & (Мен) президенттің резиденциясын көремін \\
     \hline
-    \multirow{4}{*}{boot} & \multirow{2}{*}{бәтеңке} & -- & I bought boots. \\
+    \multirow{4}{*}{boot} & \multirow{2}{*}{бәтеңке} & default & I bought boots. \\
                           &                          &     &  Мен бәтеңкелерді сатып алдым \\
 
-                          & \multirow{2}{*}{жүксалғыш} & \_ of \_ car &  He opened the boot of the car  \\ 
+                          & \multirow{2}{*}{жүксалғыш} & \(w\) \_ of \_ car &  He opened the boot of the car  \\ 
                           &           &              & Ол машинаның жүксалғышын ашты. \\
     \hline
-    \multirow{4}{*}{anything} & \multirow{2}{*}{бір нәрсе} & -- & I see anything \\ 
+    \multirow{4}{*}{anything} & \multirow{2}{*}{бір нәрсе} & default & I see anything  
+\\ 
                               &           &    & Мен бір нәрсе көремін \\
@@
 \end{table*}
+\todo{I do not understand the \emph{I see anything} example. Does it mean \emph{anything} as \emph{whatever}? I do not know how to put the \(w\) in this example}@
 
-\subsection{Transfer rules}
+\subsubsection{Transfer rules}
 
-For English–Kazakh system this stage uses three stages to improve the granularity of structural-transfer 
-rules \citep{sundetova13b}:
+For English--Kazakh transfer is performed in three stages \citep{sundetova13b}:
 
@@
 
-This section describes the current structural transfer in \texttt{apertium-eng-kaz} (revision ..., ). English to Kazakh 
-chunker rules, interchunk rules and additional clean-up stage will be described in detail next 3 sections.
+This section describes the current structural transfer in \texttt{apertium-eng-kaz} (revision ..., ).\todo{Complete! 58998? Move to results?} English--Kazakh 
+chunker rules, interchunk rules and an additional clean-up stage will be described in detail in the following 3 sections.
 
-\subsubsection{Chunker}
-
-In the first round of structural transfer rules do some local operations to cut sentences into chunks, like short noun 
+\paragraph{Chunker:}\todo{This should have 4 numerals, but I don't know how to do this}
+In the first round of structural transfer, rules segment sentences into chunks, like short noun 
 phrases, adjective phrases, verb phrases and adpositional phrases (that is, prepositional phrases in English and 
 postpositional phrases in Kazakh).
-Chunking rules, of which there are currently 168, identify 8 kinds of chunks and translate them into Kazakh as much 
-as possible, leaving some minor operations to be performed in later stages of structural transfer (for instance, the 
-case of noun phrases). Table~\ref{table:nps} describes what kind of noun phrases could appear in sentences and how it could be detected.
+Chunking rules, of which there are currently 168, identify 8 kinds of chunks and translate them into equivalent Kazakh chunks, leaving some adaptations to be performed in later stages of structural transfer (for instance, the
+morphological case of noun phrases).
+Table~\ref{table:nps} describes what kind of noun phrases could appear in sentences and how it could be detected.\todo{This table does not add much.
+the first and second column of the table are essentially the same: it would be more interesting to give the Kazakh sequence of PoS tags, perhaps with an example}
 
@@
     \hline
-    \multirow{2}{*}{\texttt{SupP}} & \multirow{2}{*}{Superlative phrase} & adjective--``est'' & possessor*,\footnote{Treated as a noun phrase with an implied noun (``the largest [book]'')} case \\
+    \multirow{2}{*}{\texttt{SupP}} & \multirow{2}{*}{Superlative phrase} & adjective--``est'' & possessor*, case \\
                                    &                                     & the--``most''--adjective & \\ 
@@
 \end{savenotes}
-
 As it is shown above, the most probable sequence of world classes are built phrases and 
@@
   Kazakh, and assigns it four chunk-level tags: number (set to singular), person (set to 3rd), 
-  possessor (to be determined, as the noun кітап ('book') could receive a 3rd-person possesive 
-  ending (кітабы) later if the context were, for instance, the large book of animals, аңдардың үлкен кітабы), 
-  and case (to be determined as it could be, for instance, accusative in I saw the large book, Мен үлкен кітапты көрдім ). Also 
-  noun form of the verb, for instance,  gerund form is determined as NP phrase (Her playing is good – Оның ойыны жақсы).
+  possessor (to be determined, as the noun \emph{кітап} ('book') could receive a 3rd-person possesive 
+  ending (\emph{кітабы}) later if the context were, for instance, \emph{the large book of animals}, \emph{аңдардың үлкен кітабы}), 
+  and case (to be determined as it could be, for instance, accusative in \emph{I saw the large book}, \emph{Мен үлкен кітапты көрдім} ). Also, the
+  noun form of the verb, for instance,  gerund form is determined as NP phrase (\emph{Her playing is good} – \emph{Оның ойыны жақсы}).\todo{Make sure all English and Kazakh examples are in italics}
 \item[Prepositional phrases:] English prepositional phrases are translated into Kazakh as postpositional phrases, 
-  there are three possible outcomes with different cases:
+  there are three possible outcomes with different cases:\todo{remove braces for consistency with the previous notation for archiphonemes: avoid repeating the explanation in the footnote.}
    \begin{itemize}
@@
         [PP [P of ] [NP the beautiful garden] ] $\rightarrow$ [GenP [NP әдемі бақша] [P -ның] ] 
-    \item Translating the locative ``-{D}{A}'',\footnote{{D} can be д or т, and {A} can be е or а, depending on the 
-        phonological context.} ablative ``-{D}{A}н'', etc., except the genitive ``-{N}{I}ң'' cases: 
-        [PP [P from ] [NP the five cars] ] $\rightarrow$ [PP [NP бес көлік] [P -тен ] ]
-    \item With nouns such as аст `under', үст `on', etc. which takes possessive from main noun:  
-        [PP [P under ] [NP the garden] ] $\rightarrow$ [PP [NP [GenP [NP бақша ][P -ның]] [NP астын]] [P -да ] ]
+    \item Translating into any of the morphological cases: locative ``-{D}{A}'',\footnote{{D} can be д or т, and {A} can be е or а, depending on the 
+        phonological context.} ablative ``-{D}{A}н'', etc., except the genitive ``-{N}{I}ң'' case: 
+        [PP [P \emph{from} ] [NP \emph{the five cars}] ] $\rightarrow$ [PP [NP \emph{бес көлік}] [P \emph{-тен} ] ]
+    \item Using postpositional constructs based on positional nouns such as \emph{аст} (`under'), \emph{үст} (`on'), etc., which take the possessive from the main noun:  
+        [PP [P \emph{under} ] [NP \emph{the garden}] ] $\rightarrow$ [PP [NP [GenP [NP \emph{бақша} ][P \emph{-ның}]] [NP \emph{аст\textbf{ын}}]] [P \emph{-да} ] ]
     \end{itemize}
-\item[Verb phrases:] Translation verb phrases into Kazakh  are not always be straightforward, for instance: 
-  present simple and future are rendered using the same tense in Kazakh (I play $\rightarrow$ Мен ойнаймын;  I 
-  will play $\rightarrow$ Мен ойнаймын), this ambiguity problem may be solved by lexical selection, 
-  which will choose right tense by watching context; tenses expressing continued activity, such as the 
+\item[Verb phrases:] Translation of English verb phrases into Kazakh  are not always be straightforward, for instance: 
+  % present simple and future are rendered using the same tense in Kazakh (I play $\rightarrow$ Мен ойнаймын;  I 
+  % will play $\rightarrow$ Мен ойнаймын), this ambiguity problem may be solved by lexical selection, 
+  % which will choose right tense by watching context; 
+\todo{I removed text that pertained to kaz--eng}
+tenses expressing continued activity, such as the 
   English present continuous or past continuous (I am playing, I was playing), have to be detected and 
@@
       is жатыр ('lie') for the present continuous and отыр ('sit') for the past continuous.} 
-  modal verbs are translated by adding adjectives, which means ``necessary'' (керек) or ``proper'' (жөн), 
+  modal verbs are translated by adding adjectives, which means ``necessary'' (керек) or ``proper'' (жөн), \todo{All examples in italics.
+This part is hard to read.
+The choice of auxilliary is not justified.}
   and a form of the copula (absent in present tense); the subject receives the genitive or dative case; negative 
@@
 \begin{itemize}
-\item The adjective alone, marked AdjP: [AdjP big ] $\rightarrow$ [AdjP үлкен] 
+\item The adjective alone, marked AdjP: [AdjP \emph{big} ] $\rightarrow$ [AdjP \emph{үлкен}] 
 \item Comparative adjective phrases  (English more + adjective, or adjective-[e]r); the Kazakh translation 
-   chooses the comparative suffix ``-{I}р{A}{K}'': [AdjP  more  interesting ]  $\rightarrow$ [AdjP қызығырақ]
+   chooses the comparative suffix ``-{I}р{A}{K}'': [AdjP  \emph{more  interesting} ]  $\rightarrow$ [AdjP \emph{қызығ\emph{ырақ}}]
 \item For superlative adjective phrases  ``the most + adjective''  or ``adjective-[e]st'', translation 
-   is built using ``ең'' + adjective:  [SupP  the biggest ]  $\rightarrow$ [SupP ең үлкен] 
-   [SupP  the most important]   $\rightarrow$ [SupP ең маңызды]
+   is built using \emph{ең} + adjective:  [SupP  the \emph{biggest} ]  $\rightarrow$ [SupP \emph{ең үлкен}] 
+   [SupP  \emph{the most important}]   $\rightarrow$ [SupP \emph{ең маңызды}]
 \end{itemize}
 Superlative adjective phrases have some properties of noun phrases (such as receiving possessive morphemes when 
-modified by a genitive phrase: the most beautiful of people $\rightarrow$ адамдардың ең әдемісі); one could 
-say that they are treated as NPs with an implied noun.
+modified by a genitive phrase: \emph{the most beautiful of people} $\rightarrow$ \emph{адамдардың ең әдемі\textbf{сі}}); they are treated as a noun phrase with an implied noun (``the largest [book]'').
+\todo{The table of phrase types is not referred to in the text; also, the meaning of ``*'' is not explained}
+
 \end{description}
@@
 
-\subsubsection{Interchunk processing}
+\paragraph{Interchunk processing:}
+\todo{Table~\ref{table:vps} should be typeset with examples in italics and making sure that the margins are respected}
 
-The second round of structural transfer is currently performed by a proof-of-concept set of 90 rules, representative of following operations:
+The second round of structural transfer is currently performed by a %proof-of-concept 
+set of 90 rules, representative of following operations:\todo{I have removed \emph{proof-of-concept} here}
 \begin{itemize}
@@
 \item Assigning case to noun phrases (which are generated without case by the chunker): for instance, 
-  accusative case for objects (I bought the table $\rightarrow$ Мен үстелді  сатып алдым), genitive case for 
-  obligatory constructs (I must see $\rightarrow$ менің көруім керек),  dative case for the verb to 
-  need (I need a doctor $\rightarrow$Маған дәрігер керек), locative case for 
-  possession (I have a book $\rightarrow$ Менде кітап бар), etc.
-\item Reordering: placing of object before verb ( I[1] bought[2] the table[3] $\rightarrow$ Мен[1] үстелді[3] сатып алдым[2]), 
-  placing of prepositional pharses before the verb  (They[1] played[2] on top of the tree [3] $\rightarrow$ Олар[1] ағаштың үстінде[3] ойнады [2]), etc.
-\item Adding question word ма/ме/ба/бе in the end of question, if question mark chunk ``?'' is detected (Did<VP\_ques> you<NP> watch<VP> last film<NP> ?<Q\_m> $\rightarrow$ Сіз<NP> соңғы фильмді<NP> көрдіңіз<VP> бе<ques> ?<Q\_m>).
-\item Place possessive for long noun + noun + noun structures(The university of city of Kazakhstan -- Қазақстан қаласының университеті).
+  accusative case for objects (\emph{I bought the table} $\rightarrow$ \emph{Мен үстелді  сатып алдым}), genitive case for 
+  obligatory constructs (\emph{I must see} $\rightarrow$ \emph{менің көруім керек}),  dative case for the verb \emph{to 
+  need} (\emph{I need a doctor} $\rightarrow$ \emph{Маған дәрігер керек}), locative case for 
+  possession (\emph{I have a book} $\rightarrow$ \emph{Менде кітап бар}), etc.\ 
+\item Reordering: placing of object before verb ( I[1] bought[2] the table[3] $\rightarrow$ Мен[1] үстелді[3] сатып алдым[2]), \todo{italics for examples please!}
+  placing of prepositional pharses before the verb  (They[1] played[2] on top of the tree [3] $\rightarrow$ Олар[1] ағаштың үстінде[3] ойнады [2]), etc.\ 
+\item Adding question word ма/ме/ба/бе in the end of question, if the question mark chunk ``?'' is detected (Did<VP\_ques> you<NP> watch<VP> last film<NP> ?<Q\_m> $\rightarrow$ Сіз<NP> соңғы фильмді<NP> көрдіңіз<VP> бе<ques> ?<Q\_m>).\todo{Notation: use brackets as above!}
+\item Place possessive for long noun + noun + noun structures(The university of city of Kazakhstan -- Қазақстан қаласының университеті).\todo{Explain better, as this is a rule to avoid repeating -NIng, -NIng, etc.
+All possessives are in place}
 \end{itemize}
@@
 
-\subsubsection{Postchunk}
-
-Additional cleaning stage was created for removing extra tags which did not determine through the chunker and 
-interchunk steps, for instance, noun was not acting as object and did not get accusative case, so not determined 
-tag was \texttt{<CD>} after ``cleaning'' it will be set on nominative case; at this step, also carried out 
-agreement between noun or adjective and copula ``е''(Мен дәрігер+е<p1 person singular>), vowel and consonants 
-agreement between last word of question and question particle (Сіз келдіңіз + ме? - Сіз келдіңіз + бе?).
+\paragraph{Postchunk:}\todo{Nothing is said about postchunk --- this talks about the additional cleaning stage}
+An additional cleaning stage was created for removing extra tags which were ot determined during the chunker and 
+interchunk steps, for instance, a noun was not determined to be acting as an object and therefore did not get accusative case; such a noun would have been assigned the  \texttt{<CD>} (``case to be determined'') tag, which after ``cleaning'' would be set to nominative.
+Other operations  carried out at this stage ensure
+the agreement between noun or adjective and the copula verb ``е''(\emph{Мен дәрігер+е}<p1 person singular>), \todo{add result for copula} or deciding the actual form of the question particle according to the preceding word (\emph{Сіз келдіңіз} + \emph{ме}?
+\(\to\) \emph{Сіз келдіңіз} + \emph{бе}?).
 
@@
 
-Current system can translate simple sentences and questions. The English-Kazakh bilingual dictionary 
-contains 13,135 entries, ``chunker'' rules are 168 and interchunk rules are 148. 
-Evaluating was tested against  revision 58998 in the Apertium SVN1. Lexical coverage is calculated 
+Current system can translate simple sentences and questions. The English--Kazakh bilingual dictionary 
+contains 13,135 entries, \todo{13130 before!} There are 168 ``chunker'' rules and 148 interchunk rules. 
+Evaluating was tested against revision 58998 in the Apertium Subversion repository. Lexical coverage is calculated 
 over EuroParl,\footnote{\url{http://www.statmt.org/europarl/v7/es-en.tgz}} SETimes,\footnote{\url{http://nlp.ffzg.hr/data/corpora/setimes/setimes.en-tr.txt.tgz}} NewsCommentary,\footnote{\url{http://www.statmt.org/wmt13/training-parallel-nc-v8.tgz}} Wikipedia.\footnote{\url{http://dumps.wikimedia.org/enwiki/20140402/enwiki-20140402-pages-articles.xml.bz2}; last checked 04/06/2014}
@@
 
-Evaluating the output of the system was measured with BLEU \citep{papineni02} and Word Error Rate 
+The output of the system was evaluated with BLEU \citep{papineni02} and the word error rate 
 metric \citep{levenshtein/1966}. We chose a short text in English, and by postediting output 
-of English-Kazakh system, was build parallel text. The output of each of the machine
-translation systems was posteditted independently to avoid biasing one particular system.
+of English--Kazakh system, a build parallel text was built. The output of each of the machine
+translation systems was postedited independently to avoid biasing one particular system.
 
-As can be seen,  coverage of dictionary is good, average score is 94\%. Our system outperforms the other 
-available RBMT system, but fall short of the state-of-the-art performance. However, unlike the state of 
+As can be seen,\todo{where!} the coverage of dictionary is good, average score is 94\%. Our system outperforms the other 
+available RBMT system, but fall short of the state-of-the-art performance.\todo{represented by...} However, unlike the state of 
 the art, which uses unavailable corpora, our system is free and open and can be improved.
 
-MT systems have some mistakes in translations, which was shown in <...> brackets. Google MT system 
+MT systems have some mistakes in translations, which was shown in <...> brackets.\todo{you used *!} Google MT system 
 has a higher BLEU score, but in translation separate phrases, it has common error: it does not assign 
@@
 
-We have presented the rule-based MT system from English to Kazakh. The current English-Kazakh machine translation 
+We have presented the design of a free/open-source rule-based MT system from English to Kazakh. The current English--Kazakh machine translation 
 system already successfully translates noun-phrases, verb-phrases, prepositional-phrases, and adjectival-phrases, and 
 contains a good vocabulary for testing purposes. 
-We plan to continue developing English-Kazakh pair, going to extend the coverage to 98\% on the corpora, 
+We plan to continue developing the English--Kazakh pair, going to extend the coverage to 98\% on the corpora, 
 which we used before. Additionally, we will improve the quality of translation by adding more rules, such like 
@@
 open-source MT system with Turkic languages or system with Kazakh language on right side to make transfer systems 
-between the Turkic or other language pairs. Related work is currently ongoing with Russian--Kazakh. 
+between the Turkic or other language pairs. Related work is currently ongoing with Russian--Kazakh.\todo{Say something about Kazakh--English.}
 Our system is available as free/open-source software and the whole system may be downloaded 
@@
 We also thank Prof.
-Tukeyev for his valuable input.
-
+Tukeyev for his valuable input.\todo{Some references didn't come out right.}
 \bibliographystyle{apalike}

### 2015-03-05 engkaz-eamt-2015/engkaz-eamt-2015.tex
added kazakh-english paper, answered to some comments
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59215 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 % Aida, Mikel, Fran, Inari, ... ?
+%Aida Sundetova
+%Al-Farabi Kazakh National University, Al-Farabi av., 71, 050040 Almaty, Kazakhstan
 
@@
 write up dictionaries and grammatical rules that are applied by an engine, in some 
-unusual case will produce better translation.
+case, where statistics not work well, will produce better translation.
 \todo{What does the last sentence mean? Clarify!}
@@
 For Kazakh, a finite-state transducer based on two-level rules \citep{washingtonsalimzyantyers14} is used for generation. 
-\item[Post-generator:] Performs some minor orthographical operations in the target language. This module is not used in the English--Kazakh system.\todo{Or is it?}
+\item[Post-generator:] Performs some minor orthographical operations in the target language. 
 \item[ Reformatter:] It places format tags back into the text so that its format is preserved.
@@
 preceding vowels and consonants, the plural suffix \emph{-LAr}\footnote{Uppercase Latin letters are used for archiphonemes (actually archigraphemes) that are realised as phonemes (actually graphemes) after morphophonological rules have been applied} could
-become \emph{-тар}, \emph{-тер}, \emph{-lar}, \emph{-ler}, \emph{-dar}, \emph{-ler} depending on vowel harmony: kitap+tar,
-mektep+ter, etc. \todo{Aida, please typeset this example in Cyrillic; braces are not used here, but are used later; homogeneize notation in paper.}
+become \emph{-тар}, \emph{-тер}, \emph{-лар}, \emph{-лер}, \emph{-дар}, \emph{-лер} depending on vowel harmony: кітап+тар,
+мектеп+тер, etc. \todo{Aida, please typeset this example in Cyrillic; braces are not used here, but are used later; DONE  homogeneize notation in paper.}
 
@@
 The English--Kazakh bilingual dictionary provides correspondences between English LFs and Kazakh LFs. Ambiguity is allowed: one of the LFs will be chosen by 
-lexical-selection rules depending on context. This dictionary currently contains 13,130 stems \citep{sundetova13a}.\todo{later you say 13,135}
+lexical-selection rules depending on context. This dictionary currently contains 13,135 stems \citep{sundetova13a}.\todo{later you say 13,135}
 
@@
 the following noun is a person: \emph{beautiful girl} \(\to\) \emph{сұлу қыз}; or as \emph{әдемі} or  \emph{көркем}, if the following noun means 
-place or \todo{or what?} \emph{beautiful mountains} \(\to\) \emph{әдемі таулар}.  Table~\ref{table:lexsel} shows some 
-lexical-selection rules:\todo{but this rule is not in the table!!!}
+place or \todo{or what? - mountains} \emph{beautiful mountains} \(\to\) \emph{әдемі таулар}.  Table~\ref{table:lexsel} shows some 
+lexical-selection rules:\todo{but this rule is not in the table!!! - should I add it or just remove upper example??}
 \begin{table*}
@@
 \end{table*}
-\todo{I do not understand the \emph{I see anything} example. Does it mean \emph{anything} as \emph{whatever}? I do not know how to put the \(w\) in this example}@
+\todo{Yes, it is mean I see whatever - I do not understand the \emph{I see anything} example. Does it mean \emph{anything} as \emph{whatever}? I do not know how to put the \(w\) in this example}@
 
@@
 
-This section describes the current structural transfer in \texttt{apertium-eng-kaz} (revision ..., ).\todo{Complete! 58998? Move to results?} English--Kazakh 
+This section describes the current structural transfer in \texttt{apertium-eng-kaz} (revision 58998, ).\todo{Complete! 58998? Move to results?} English--Kazakh 
 chunker rules, interchunk rules and an additional clean-up stage will be described in detail in the following 3 sections.
@@
   placing of prepositional pharses before the verb  (They[1] played[2] on top of the tree [3] $\rightarrow$ Олар[1] ағаштың үстінде[3] ойнады [2]), etc.\ 
-\item Adding question word ма/ме/ба/бе in the end of question, if the question mark chunk ``?'' is detected (Did<VP\_ques> you<NP> watch<VP> last film<NP> ?<Q\_m> $\rightarrow$ Сіз<NP> соңғы фильмді<NP> көрдіңіз<VP> бе<ques> ?<Q\_m>).\todo{Notation: use brackets as above!}
+\item Adding question word ма/ме/ба/бе in the end of question, if the question mark chunk ``?'' is detected (Did[VP\_ques] you[NP] watch[VP] last film[NP] ?[Q\_m] $\rightarrow$ Сіз[NP] соңғы фильмді[NP] көрдіңіз[VP] бе<ques> ?[Q\_m]).\todo{Notation: use brackets as above!}
 \item Place possessive for long noun + noun + noun structures(The university of city of Kazakhstan -- Қазақстан қаласының университеті).\todo{Explain better, as this is a rule to avoid repeating -NIng, -NIng, etc.
@@
 
-\paragraph{Postchunk:}\todo{Nothing is said about postchunk --- this talks about the additional cleaning stage}
+\paragraph{Postchunk and cleanup:}\todo{Can I do that?  --Nothing is said about postchunk --- this talks about the additional cleaning stage}
 An additional cleaning stage was created for removing extra tags which were ot determined during the chunker and 
@@
 Current system can translate simple sentences and questions. The English--Kazakh bilingual dictionary 
-contains 13,135 entries, \todo{13130 before!} There are 168 ``chunker'' rules and 148 interchunk rules. 
+contains 13,135 entries, \todo{13130 before!-changed} There are 168 ``chunker'' rules and 148 interchunk rules. 
 Evaluating was tested against revision 58998 in the Apertium Subversion repository. Lexical coverage is calculated 
@@
 
-As can be seen,\todo{where!} the coverage of dictionary is good, average score is 94\%. Our system outperforms the other 
+As can be seen from table of coverage,\todo{where!-table of coverage} the coverage of dictionary is good, average score is 94\%. Our system outperforms the other 
 available RBMT system, but fall short of the state-of-the-art performance.\todo{represented by...} However, unlike the state of 
@@
 open-source MT system with Turkic languages or system with Kazakh language on right side to make transfer systems 
-between the Turkic or other language pairs. Related work is currently ongoing with Russian--Kazakh.\todo{Say something about Kazakh--English.}
+between the Turkic or other language pairs. Related work is currently ongoing with Russian--Kazakh and Kazakh--English\citep{sundetova14}.\todo{Say something about Kazakh--English.}
 Our system is available as free/open-source software and the whole system may be downloaded 

### 2015-03-05 engkaz-eamt-2015/engkaz-eamt-2015.tex
acknowl.
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59219 72bbbca6-d526-0410-a7d9-f06f51895060
some fixes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59223 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-% Authors:
-% Aida, Mikel, Fran, Inari, ... ?
-%Aida Sundetova
-%Al-Farabi Kazakh National University, Al-Farabi av., 71, 050040 Almaty, Kazakhstan
+\author{Aida Sundetova\\
+  Kazakh National University,  \\
+  050040 Almaty, \\
+  {\tt xxx@xxx}\\
+  \And
+  \textbf{Francis M. Tyers}\\
+  HSL-fakultetet\\ 
+  UiT Norgga \'{a}rktala\v{s} universitehta \\
+  9017 Romsa, \\
+  {\tt francis.tyers@uit.no}
+  \And
+  Mikel L. Forcada\\
+  Dept. Lleng. i Sist. Inform. \\
+  Universitat d'Alacant, \\ 
+  Alacant \\
+  {\tt mlf@dlsi.ua.es}  
+}
 
@@
 
-\begin{abstract}\todo{Decide authors, and put them in phantom boxes}
-This paper presents the current state of development a shallow-transfer rule-based machine translation (MT) system from English to Kazakh. The main syntactic and morphological differences 
+\begin{abstract}
+This paper presents the current state of development a shallow-transfer rule-based machine 
+translation (MT) system from English to Kazakh. The main syntactic and morphological differences 
 between the two languages are presented, and it is showed how the MT 
@@
 for the English--Kazakh pair. An English--Kazakh statistical MT system has however recently been built by Google,\footnote{\url{http://translate.google.com}}; 
-other work in this language pair was performed by building a small parallel corpus and using Moses \citep{assylbekov14}.
-Existing commercial systems for English to Kazakh (Sanasoft\footnote{\url{http://www.sanasoft.kz/c/ru/node/47}} and Trident\footnote{\url{http://www.translate.ua/us/on-line}}) both appear to be rule-based.
-Rule-based machine translation, in which developers 
-write up dictionaries and grammatical rules that are applied by an engine, in some 
-case, where statistics not work well, will produce better translation.
-\todo{What does the last sentence mean? Clarify!}
-This paper describes work in progress of development of machine translation system 
-for English--Kazakh using the Apertium free/open-source machine 
+other work in this language pair was performed by building a small parallel corpus and using Moses \citep{assylbekov14}. Existing 
+commercial systems for English to Kazakh (Sanasoft\footnote{\url{http://www.sanasoft.kz/c/ru/node/47}} 
+and Trident\footnote{\url{http://www.translate.ua/us/on-line}}) both appear to be rule-based.
+
+This paper describes the development of a machine translation system 
+for English to Kazakh using the Apertium free/open-source machine 
 translation platform \citep{forcada11}. %(http://www.apertium.org).
@@
 %  choosing wrong one could produce errors in translation. 
-\todo{I have removed some text and  the statistical PoS tagging stuff, as it is not used}
 %POS tagger is based on a statistical model, which 
@@
 become \emph{-тар}, \emph{-тер}, \emph{-лар}, \emph{-лер}, \emph{-дар}, \emph{-лер} depending on vowel harmony: кітап+тар,
-мектеп+тер, etc. \todo{Aida, please typeset this example in Cyrillic; braces are not used here, but are used later; DONE  homogeneize notation in paper.}
+мектеп+тер, etc. 
+%\todo{Aida, please typeset this example in Cyrillic; braces are not used here, but are used later; DONE  homogeneize notation in paper.}
 
@@
 The English--Kazakh bilingual dictionary provides correspondences between English LFs and Kazakh LFs. Ambiguity is allowed: one of the LFs will be chosen by 
-lexical-selection rules depending on context. This dictionary currently contains 13,135 stems \citep{sundetova13a}.\todo{later you say 13,135}
+lexical-selection rules depending on context. This dictionary currently contains 13,135 stems \citep{sundetova13a}.
 
@@
 morphological case of noun phrases).
-Table~\ref{table:nps} describes what kind of noun phrases could appear in sentences and how it could be detected.\todo{This table does not add much.
+
+%Table~\ref{table:nps} describes what kind of noun phrases could appear in sentences and how it could be detected.\todo{This table does not add much.
 the first and second column of the table are essentially the same: it would be more interesting to give the Kazakh sequence of PoS tags, perhaps with an example}
 
-\begin{table*}
-  \centering
-  \begin{tabular}{|l|l|l|}
-  \hline
-  \textbf{Structure} & \textbf{Description} & \textbf{Examples} \\
-  \hline
-  \texttt{n} & Noun & house \\
-  \texttt{adj + n} & Adjective + noun & long road \\
-  \texttt{det + n} & Determiner + noun & an apple, the chance \\
-  \texttt{num + n} & Numeral + noun & five pieces \\
-  \texttt{num + adj + n} & Numeral + adjective + noun & five big pieces \\
-  \texttt{det + num + adj + n} & Determiner + numeral + adjective + noun & the five big pieces \\
-  \hline
-  \end{tabular}
-  \caption{Example of the most common structures of noun phrases}
-  \label{table:nps}
-\end{table*}
+% \begin{table*}
+%   \centering
+%   \begin{tabular}{|l|l|l|}
+%   \hline
+%   \textbf{Structure} & \textbf{Description} & \textbf{Examples} \\
+%   \hline
+%   \texttt{n} & Noun & house \\
+%   \texttt{adj + n} & Adjective + noun & long road \\
+%   \texttt{det + n} & Determiner + noun & an apple, the chance \\
+%   \texttt{num + n} & Numeral + noun & five pieces \\
+%   \texttt{num + adj + n} & Numeral + adjective + noun & five big pieces \\
+%   \texttt{det + num + adj + n} & Determiner + numeral + adjective + noun & the five big pieces \\
+%   \hline
+%   \end{tabular}
+%   \caption{Example of the most common structures of noun phrases}
+%   \label{table:nps}
+% \end{table*}
 
@@
   % which will choose right tense by watching context; 
-\todo{I removed text that pertained to kaz--eng}
 tenses expressing continued activity, such as the 
-  English present continuous or past continuous (I am playing, I was playing), have to be detected and 
-  mapped onto sets of two lexical units (Мен ойнап жатырмын, Мен ойнап отырдым) where the main verb is 
-  found in the -п participle form (ойнап), and a suitable finite form  (жатырмын, отырдым) of an auxiliary 
-  verb (жатыр, отыр) is used to express number and person agreement,\footnote{Actually, Kazakh language uses four auxiliary verbs: жатыр ('lie', 
-      used when the activity takes a long time), отыр ('sit', used when the activity appears to be done in a sitting position), тұр ('stand', 
-      when the takes a short time), and жүр (when the activity repeats regularly). Choosing the most adequate auxiliary verb is 
+  English present continuous or past continuous (\emph{I am playing}, \emph{I was playing}), have to be detected and 
+  mapped onto sets of two lexical units (\emph{Мен ойнап жатырмын}, \emph{Мен ойнап отырдым}) where the main verb is 
+  found in the -п participle form (\emph{ойнап}), and a suitable finite form  (\emph{жатырмын}, \emph{отырдым}) of an auxiliary 
+  verb (\emph{жатыр}, \emph{отыр}) is used to express number and person agreement,\footnote{Actually, Kazakh language uses four auxiliary verbs: \emph{жатыр} `lie', 
+      used when the activity takes a long time), отыр (`sit', used when the activity appears to be done in a sitting position), \emph{тұр} `stand', 
+      when the takes a short time), and \emph{жүр} (when the activity repeats regularly). Choosing the most adequate auxiliary verb is 
       hard without a semantic analysis, which is not easily available in Apertium. Our current choice (an approximation) 
-      is жатыр ('lie') for the present continuous and отыр ('sit') for the past continuous.} 
-  modal verbs are translated by adding adjectives, which means ``necessary'' (керек) or ``proper'' (жөн), \todo{All examples in italics.
-This part is hard to read.
-The choice of auxilliary is not justified.}
+      is \emph{жатыр} `lie' for the present continuous and отыр (`sit') for the past continuous.} 
+  modal verbs are translated by adding adjectives, which means \emph{керек} `necessary' or \emph{жөн} `proper', \todo{This part is hard to read. The choice of auxilliary is not justified.}
   and a form of the copula (absent in present tense); the subject receives the genitive or dative case; negative 
@@
 We also thank Prof.
-Tukeyev for his valuable input.\todo{Some references didn't come out right.}
+Tukeyev for his valuable input.
+Special thanks to Inari Listenmaa for her assistance in mentoring the Google Summer of Code project.\todo{Some references didn't come out right.}
 \bibliographystyle{apalike}

### 2015-03-05 engkaz-eamt-2015/engkaz-eamt-2015.tex
add
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59224 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-As can be seen from table of coverage,\todo{where!-table of coverage} the coverage of dictionary is good, average score is 94\%. Our system outperforms the other 
+As can be seen from Table~\ref{table:coverage}, the coverage of dictionaries is good, average score is 94\%. Our system outperforms the other 
 available RBMT system, but fall short of the state-of-the-art performance.\todo{represented by...} However, unlike the state of 
@@
 open-source MT system with Turkic languages or system with Kazakh language on right side to make transfer systems 
-between the Turkic or other language pairs. Related work is currently ongoing with Russian--Kazakh and Kazakh--English\citep{sundetova14}.\todo{Say something about Kazakh--English.}
+between the Turkic or other language pairs. Related work is currently ongoing with Russian--Kazakh and Kazakh--English \citep{sundetova14}.
 Our system is available as free/open-source software and the whole system may be downloaded 

### 2015-03-05 engkaz-eamt-2015/engkaz-eamt-2015.tex
changed <..>
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59226 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-MT systems have some mistakes in translations, which was shown in <...> brackets.\todo{you used *!} Google MT system 
+MT systems have some mistakes in translations, which was shown with *.\todo{DONE you used *!} Google MT system 
 has a higher BLEU score, but in translation separate phrases, it has common error: it does not assign 

### 2015-03-06 2015-eamt-assim/2015-eamt-assim.tex
more stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59236 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{graphicx}
-\usepackage{microtype} % This saves some space and makes nice typesetting choices
+%\usepackage{microtype} % This saves some space and makes nice typesetting choices
 
@@
 \newcommand{\comment}[1]{\marginpar{\scriptsize\sf \textcolor{blue}{#1}}}
+%\newcommand{\comment}[1]{}
 
@@
 \maketitle
-\renewcommand{\baselinestretch}{0.97} % the undetectable 3% squeeze
-\comment{EA: has not it first been used in trosterud12?}
+\renewcommand{\baselinestretch}{0.96} % the undetectable 3% squeeze
 \begin{abstract}
@@
   strategy independently proposed in 2012 and 2013 for evaluation of
-  machine translation for assimilation purposes. The strategy involves
-  human evaluators, who are asked to restore key words (to fill gaps) in
-  reference translations. The evaluation method is applied to two translation
-  directions.\comment{MLF: directions \(\to\) language pairs?} To reduce the amount of time required to prepare tasks
-  and analyze results, an open-source task management system is
-  introduced. The evaluation results suggest that the gap-filling task
-  reflects users' understanding of text,\comment{MLF: ``suggests''?
-I think this is what we assume to interpret our results: how about ``shows''?
-We don't show correlation with other evaluation.} and may be used to measure MT
+ MT for assimilation purposes. The strategy involves
+  human evaluators, who are asked to restore keywords (to fill gaps) in
+  reference translations. The evaluation method is applied to two language pairs, Basque--Spanish and Tatar--Russian. To reduce the amount of time required to prepare tasks
+  and analyse results, an open-source task management system is
+  introduced. The evaluation results show that the gap-filling task
+  reflects users' understanding of text, and may be used to measure MT
   quality for assimilation purposes.
@@
 
-\comment{EA: I'd like a citation for the ``most massive application'', could you suggest something? It is from Mikel's paper, but no reference there -- should I cite Mikel?}
-\comment{MLF: I added a footnote, please reduce and move citation to the reference}
-Despite the fact that, as a result of widespread usage of online MT, assimilation (or gisting) is currently the most frequent application of MT,\footnote{“In a given day we [Google] translate roughly as much text as you’d find in 1 million books.
-To put it another way: what all the professional human translators in the world produce in a year, our system translates in roughly a single day.”, Franz J.
-Och, \url{http://tinyurl.com/och-gisting}} few methodologies are established for assimilation evaluation of MT.
+Despite the fact that, as a result of widespread usage of online MT, assimilation (or gisting) is currently the most frequent application of MT (in 2012, daily output of Google Translate matched the yearly output of human translations \citep{google}), few methodologies are established for assimilation evaluation of MT.
 The methods include post-editing and comparison by bilingual experts \citep{ginesti09},\comment{MLF: should we also cite the original WMT reference cited by Ginestí et al.?} and multiple choice tests \citep{jones07,trosterud12}.
 These approaches are often costly and prone to subjectivity: see the discussion in \cite{oregan13}.
-As an alternative, the modification of \emph{cloze}\comment{\emph{cloze} stands for \emph{closure}.} testing \citep{taylor53} was introduced for assimilation evaluation, first by \citep{trosterud12} as a supplementary, and then by \citep{oregan13} as a stand-alone method.
+As an alternative, the modification of \emph{cloze}\comment{\emph{cloze} stands for \emph{closure}.} testing \citep{taylor53} was introduced for assimilation evaluation, first by \citet{trosterud12} as a supplementary, and then by \citet{oregan13} as a stand-alone method.
 Prior to this, cloze tests have been used to evaluate raw MT quality \citep{vanslype79,somers00}.
-While these authors ask informant to fill gaps in MT output, \citet{trosterud12} and \citet{oregan13} propose to fill gaps in the reference (human) translation.
-The gap-filling task models how well users comprehend the key points of the text, as it is roughly equivalent with answering questions.\comment{MLF: check if you agree with this} Thus, the method does not directly evaluate the quality of machine-produced text, but rather its usefulness in understanding the meaning of the original text.
+While these authors ask informants to fill gaps in MT output, \citet{trosterud12} and \citet{oregan13} propose to fill gaps in the reference (human) translation.
+The gap-filling task models how well users comprehend the key points of the text, as it is roughly equivalent with answering questions.
+Thus, the method does not directly evaluate the quality of machine-produced text, but rather its usefulness in understanding the meaning of the original text.
 
-\comment{EA: I understand we selected tat-rus and eng-kaz because they were in development. why did we select eu-es?}
 The gap-filling method has been successfully used to evaluate the Basque--English Apertium language pair.
 In this work we extend the evaluation to two more language pairs: Basque--Spanish and Tatar-Russian.
+The former pair, while not producing the output suitable for post-editing, is a good example of an assimilation MT system.
+In addition, Basque and Spanish are not mutually understandable, and therefore constitute a good pair for evaluation.
 For the latter pair, the evaluation served as a quality check in the period of active development during the Google Summer of Code 2014 programme.
 In addition to evaluating, we explore new aspects of the experiment: the correlation between evaluators' scores, and the influence of the texts' linguistic domain.
-To facilitate the evaluation, we introduce an automated system which creates task sets from parallel corpora given a range of parameters (number of gaps in a sentence, hint type, gap filler,\comment{MLF: gap filler or gap marker?} etc.), checks evaluators' answers, and calculates and reports generalized results.
-This system is integrated into the Appraise MT evaluation platform \citep{federmann12}; the code is open source and is available on GitHub.\footnote{\phantom{\url{https://github.com/Sereni/Appraise}}}\comment{MLF: check if phantom URL is ok}
+To facilitate the evaluation, we introduce an automated system which creates task sets from parallel corpora given a range of parameters (number of gaps in a sentence, hint type, gap filler,\comment{MLF: gap filler or gap marker?} \comment{EA: well, initially we had questions with lemmas in place of gaps, and also multiple choice questions where answer options got generated automatically.
+This is not used in this evaluation, but it's still in the toolkit.
+Should I remove?} etc.), checks evaluators' answers, and calculates and reports generalized results.
+This system is integrated into the Appraise MT evaluation platform \citep{federmann12}; the code is open source and is available on GitHub.\footnote{\phantom{\url{https://github.com/Sereni/Appraise}}}
 
-We anticipate that the assessed MT systems will contribute to the users' understanding of text, that is, the evaluation will show better results in MT-assisted task type.\comment{MLF: this last sentence is very hard to follow} We also expect to see different results depending on such parameters as text domain and the number of gaps in a sentence in relation to sentence length.
+We anticipate that the assessed MT systems will contribute to the users' understanding of text, that is, the users will show better results in tasks when assisted with MT.
+We also expect to see different results depending on text domain and the relative number of gaps in a sentence.
 
@@
 \begin{table*}
-  \begin{tabular}{|l|p{15cm}|}
-%@@@  \textbf{Src (Tatar)} & \rus{???? ?????? ???????; ????? 7 ??????????? ??? ??????? ?????, ????? ??? ????? 100 ???? ??????? ???????.}
-%  \hline
-%  \textbf{MT (Russian)} & \rus{Половина сказать часы; вам 7 тукталыштан поздний опуститься нужный , после его снова 100 метр сказать идти.} \\
+  \centering
+  \begin{tabular}{|l|l|}
      \hline
-     \textbf{Ref.} & \rus{Примерно полчаса; вам нужно выйти через 7 остановок, потом пройти ещё около 100 метров.} \\\hline
-     \textbf{10\%}   & \rus{Примерно полчаса; вам нужно выйти через 7 \{ \}, потом пройти ещё около 100 метров.} \\
-     \textbf{20\%}   & \rus{\{ \} полчаса; вам нужно \{ \} через 7 остановок, \{ \} пройти ещё около 100 метров.} \\
-     \textbf{30\%} & \rus{Примерно полчаса; вам нужно \{ \} через 7 \{ \}, потом пройти \{ \} около 100 \{ \}.} \\
+     \textbf{Ref} & \rus{{\small Примерно полчаса; вам нужно выйти через 7 остановок, потом пройти ещё около 100 метров.}} \\
+     \hline
+     \textbf{10\%} & \rus{{\small Примерно полчаса; вам нужно выйти через 7 \{ \}, потом пройти ещё около 100 метров.}} \\
+     \textbf{20\%} & \rus{{\small \{ \} полчаса; вам нужно \{ \} через 7 остановок, \{ \} пройти ещё около 100 метров.}} \\
+     \textbf{30\%} & \rus{{\small Примерно полчаса; вам нужно \{ \} через 7 \{ \}, потом пройти \{ \} около 100 \{ \}.}} \\
      \hline
@@
 of a given text correlates with the number of words they are able to correctly restore in the text. Therefore, the 
-base of an assimilation task is a (reference) sentence, where some of the words are blacked out, or removed. The sentence \comment{EA: I'm sorry, I'm not sure what else I could explain here...}
+base of an assimilation task is a (reference) sentence, where some of the words are blacked out, or removed. The sentence 
 is produced by a human (as opposed to machine-translated), and it is in the language known to evaluators, which is also 
@@
 also human-produced, but in the source language of the pair. The second type is the \emph{machine-translated} hint, which 
-comes from the machine translation of the \emph{source} sentence.
-Table~\ref{table:example} shows a sample task, and Figure~\ref{figure:screenshot} shows the task in the online evaluation environment.\comment{Use the highest possible resolution for the figure so that aliasing is imperceptible}
+comes from the machine translation of the \emph{source} sentence. Table~\ref{table:example} shows a sample task, and Figure~\ref{figure:screenshot} shows the task in the online evaluation environment.
+%\comment{Use the highest possible resolution for the figure so that aliasing is imperceptible}
 
@@
   \centering
-\includegraphics[width=0.9\textwidth,resolution=144]{appraise-scr}
- \caption{An example set of sentences in an online environment. The task is Russian legal text with 30\% gaps.}
+\includegraphics[width=1.0\textwidth,resolution=144]{appraise-scr}
+ \caption{An example set of sentences in the online environment. The task is Russian legal text with 30\% gaps.}
 \label{figure:screenshot}
@@
 
-\comment{EA: I don't really have a good explanation why we used source+mt tasks, except for being exhaustive. could you suggest anything?}
-In the course of the experiment, following \cite{oregan13}, the following hint combinations are offered:
+In the course of the experiment, following \cite{oregan13}, we offer these hint combinations:
 \begin{description}
@@
 in proper nouns or loan words.
-\item[Reference sentence and MT hint:] In addition to the task, the
+\item[Reference sentence and MT hint:] In addition to the reference sentence, the
 participants see the source sentence translated via the MT system Apertium \citep{forcada11}. This type of task is
@@
 \item[Reference sentence and both hints:] This task is
-added for completing the possible combination of task types, and to check
-for unexpected insights, should they arise.
+added to check whether MT and source provide complementary hints.
 \end{description}
 
-In order to prepare the evaluation questions, we determine and remove the keywords\comment{MLF: decide, \emph{key words} or \emph{keywords} everywhere} in the 
+In order to prepare the evaluation questions, we determine and remove the keywords in the 
 reference sentences. We consider two parameters: the list of allowed parts of speech (PoS), and the 
 number of gaps relative to sentence length (``gap density''). For the evaluations described in 
-this paper we use gap densities of 10, 20 and 30 percent, and the following parts of speech: noun, 
+this paper we use gap densities of 10, 20 and 30 percent (Table~\ref{table:percentage}), and the following parts of speech: noun, 
 proper noun, adjective, adverb and lexical verb (as opposed to auxiliary verb).
@@
 from the candidate list in such a manner that the gaps are distributed evenly throughout the sentence.
-We start at a random word in a sentence and check whether it is a keyword.
+We start at a random word in a sentence and check whether it is a keyword candidate.
 If yes, we remove it, and move \emph{n} words forward, going back to the beginning of sentence if necessary.
@@
 If the word is not a keyword, or has already been removed, we look at the next word instead.
-The process is repeated until the designated number of words has been removed, or until there are no more words in the keyword list.\comment{MLF: O'Regan and Forcada did something lots simpler: why not here?
-It is probably a bit late to change this}
+The process is repeated until the designated number of words has been removed, or until there are no more words in the keyword list.
+%\comment{MLF: O'Regan and Forcada did something lots simpler: why not here? It is probably a bit late to change this}
 
-Keyword removal is one of the most time-consuming steps in task preparation.
+%\comment{MLF: the procedure described so far seems automatic. What's manual and time consuming?} 
+Keyword removal could be one of the most time-consuming steps in task preparation.
 It normally requires human effort, because we would like to determine the words that contribute the most to understanding the text as opposed to removing random words.
-\comment{MLF: the procedure described so far seems automatic.
-What's manual and time consuming?} In our setup, the above procedure is performed by a script integrated into the task generation pipeline.
+In our automatic setup, the above procedure is performed by a script integrated into the task generation pipeline.
 Parts of speech are determined with Apertium morphological analyser.
@@
 Spanish and Russian) who had no command of source language of the pair (Basque
-and Tatar, respectively). Eleven evaluators participated in the in Basque--Spanish experiment, and 28 in Tatar-Russian (although not everyone completed the task
+and Tatar, respectively). Eleven evaluators participated in the Basque--Spanish experiment, and 28 in Tatar-Russian (although not everyone completed the task
 in full, see discussion).
 
-\comment{EA: maybe I should put this into the previous section?}
 By design, our gap-filling tasks require a human translation (reference) of source sentences.
@@
 \item  For Basque--Spanish, from the corpus of legal texts ``Memorias de traducci\'on del
-Servicio Oficial de Traductores del IVAP'';\footnote{\url{http://tinyurl.com/tm-ivap}}
+Servicio Oficial de Traductores del IVAP'';\footnote{\url{http://tinyurl.com/ivaptm2}} 
 \item  For Tatar-Russian, from the following sources on three different topics:
@@
 — 320 с.
-ISBN 5—298—00463—6 (стр.
+ISBN 5--298--00463--6 (стр.
 219, 220, 232, 233, 234)}} of spoken Tatar;
@@
 language pair in Apertium.
+The Basque--Spanish experiment featured 94, 181 and 272 gaps in the 10, 20 and 30 \% tasks, respectively.
+For Tatar--Russian these numbers are 272, 396 and 724, due to longer sentences used in task creation.
+\comment{EA to JA: Ilnar wanted to test the system on his own sentences, should I put it in the paper?}
 
@@
 The participants are given the instructions in their native language; these instructions are repeated above each task in the evaluation system.
-For the participants' convenience, the body of questions is split into smaller groups which allow indefinite time breaks\comment{MLF: time break OK?
-Possibility: \emph{multiple sessions}.} between them.
+For the participants' convenience, the body of questions is split into smaller groups which allow multiple evaluation sessions.
 The instructions are the
@@
 the answer word. We record each candidate synonym along with the answer key and the
-context sentence (see Table~\ref{table:syn} for an example). Based on this data, a native speaker of the target language decides
+context sentence.
+For example, the word \emph{asumir} is the original answer in the Spanish sentence \emph{Aprender a jugar y divertirse en el agua sin asumir riesgos}.
+However, two or more evaluators gave a different answer, \emph{corder}.
+Based on this data, a native speaker of the target language decides
 whether the candidate synonym is an acceptable replacement to the answer key in the given
@@
 
-\begin{table*}
-  \centering
-  \begin{tabular}{|l|l|}
-     \hline
-     \textbf{Sentence:}   & Aprender a jugar y divertirse en el agua sin asumir riesgos. \\
-     \textbf{Key:}   & asumir \\
-     \textbf{Synonym:} & correr \\
-     \hline
-  \end{tabular}
-  \caption{An example candidate synonym. The original word \emph{asumir} may be replaced with \emph{correr} in this context.} 
-  \label{table:syn}
-\end{table*}
-\comment{EA: added}
 The synonym lists for Basque--Spanish and Tatar--Russian contain 52 and 25 words, respectively.
 Time taken to compile each list depends on the number of candidate synonyms, and in our case was approximately 30 minutes.
-\comment{FMT: How long did making the synonym list take, how many synonyms were added? More details here. Also perhaps an example?}
 
@@
 
+%\comment{FMT: We should probably integrate the other tables, and refer to them using labels.} 
 The results are presented in this section. Tables \ref{table:res-eus-spa} and \ref{table:res-tat-rus} show
-the proportion and standard deviation of correct answers depending on evaluation mode and gap density.\comment{FMT: We should probably integrate the other tables, and refer to them using labels.} The evaluators' correct answer percentage is averaged over the number of evaluators.
+the proportion and standard deviation of correct answers depending on evaluation mode and gap density.
+The evaluators' correct answer percentage is averaged over the number of evaluators.
 In addition to the percentage of correct answers we kept a record of time taken to fill the gaps in one sentence.
@@
 We found that, as opposed to trying to devise the best word for no-hint gaps, the participants often resorted to filing these gaps with random words, which takes little time.
-\comment{EA: anything else I could write about times?}
 
@@
 Such an increase suggests that, in this case, the content words were restored by
-\comment{EA: to JA: I am not sure if this qualifies as common sense.
-What I meant is that people correctly guessed the meaning of the word in place, and expressed it in their own manner.
-Changed it to semantic context, is this a better rephrasing?}
 semantic context rather than through strong collocation.
@@
 \cite{oregan13} did random!
-Why was this changed?} The participants have not received any training tasks before the main evaluations.
+Why was this changed?}The participants have not received any training tasks before the main evaluations.
 Therefore, it is possible that the accommodation period is responsible for lower-than-expected scores in this mode of evaluation.
@@
 
-\subsection{Annotator agreement}\comment{MLF cannot get these tables to appear numbered later than the other ones!}
+\subsection{Annotator agreement}
 After obtaining the results we calculated Krippendorff's alpha~\citep{krippendorff70} measure to represent annotator agreement, shown in Tables~\ref{table:alpha-eus-spa} and~\ref{table:alpha-tat-rus}.
@@
 We selected this measure because of its compatibility with more than two annotators per task and the missing data (not all the gaps were evaluated).
-To calculate Krippendorff's alpha we used an algorithm implementation by Thomas Grill,\footnote{http://grrrr.org}\comment{MLF: Give a direct reference: I could not find the algorithm in such a wide website.
-Use tinyurl if it is too long.} dividing the answers in each gap into two categories: correct and incorrect.
+To calculate Krippendorff's alpha we used an algorithm implementation by Thomas Grill,\footnote{\url{http://grrrr.org/data/dev/krippendorff_alpha/}} dividing the answers in each gap into two categories: correct and incorrect.
 The previously obtained synonym lists were taken into account, i.e.
@@
 
-As the MT and MT+source hints are introduced, the agreement between annotators increases (measures closer to 1): the annotators are more consistently correct or incorrect in each given sentence.\comment{EA: I cite Juan Antonio here, because I am not completely sure what alpha shows about our experiment (lack of knowledge on my part)}  The agreement measure for the same sentences without hints is closer to zero, which contributes to the reliability of our methodology.
+As the MT and MT+source hints are introduced, the agreement between annotators increases (measures closer to 1): the annotators are more consistently correct or incorrect in each given sentence.
+The agreement measure for the same sentences without hints is closer to zero, which contributes to the reliability of our methodology.
 We note the outlier score in Tatar--Russian 10\% source tasks, which has the most contribution from the news texts (agreement of 0.917).
@@
 \subsection{Results for different domains}
-\comment{EA: and what would we like to say about those?
-The results depend heavily on the Tatar--Russian system.
-Do we have to provide info on what vocabulary it handles better, what it has and has not been trained on?
-Or do we just say that casual texts are easier to fill than legal texts?}
 For the Tatar--Russian language pair the participants were offered texts from three different domains (in equal proportions): casual conversations, legal texts and news.
@@
 \hline
-10\%&62.19 \(\pm\) 32.07&58.04 \(\pm\) 27.96&39.90 \(\pm\) 38.96&48.79 \(\pm\) 39.84\\
-20\%&64.81 \(\pm\) 30.03&70.18 \(\pm\) 26.98&31.37 \(\pm\) 28.42&30.91 \(\pm\) 30.22\\
-30\%&48.13 \(\pm\) 26.08&39.55 \(\pm\) 23.90&25.94 \(\pm\) 19.67&18.07 \(\pm\) 18.15\\
+10\%&62.19 \(\pm\) 32&58.04 \(\pm\) 28&39.90 \(\pm\) 39&48.79 \(\pm\) 40\\
+20\%&64.81 \(\pm\) 30&70.18 \(\pm\) 27&31.37 \(\pm\) 28&30.91 \(\pm\) 30\\
+30\%&48.13 \(\pm\) 26&39.55 \(\pm\) 24&25.94 \(\pm\) 20&18.07 \(\pm\) 18\\
 \hline
 \end{tabular}
-\caption {Basque--Spanish: Average number of gaps successfully filled (\%), using a synonym list} \label{table:res-eus-spa}
+\caption {Basque--Spanish: Average number of gaps successfully filled (\%), using a synonym list} 
+\label{table:res-eus-spa}
 \end{table*}
@@
 \hline
-10\%&63.76 \(\pm\) 41.15&57.02 \(\pm\) 41.57&53.52 \(\pm\) 43.57&45.96 \(\pm\) 40.86\\
-20\%&59.77 \(\pm\) 33.10&64.93 \(\pm\) 30.95&45.60 \(\pm\) 31.23&38.56 \(\pm\) 32.11\\
-30\%&55.83 \(\pm\) 26.13&58.77 \(\pm\) 28.07&40.00 \(\pm\) 28.16&35.45 \(\pm\) 29.76\\
+10\%&63.76 \(\pm\) 41&57.02 \(\pm\) 42&53.52 \(\pm\) 43&45.96 \(\pm\) 41\\
+20\%&59.77 \(\pm\) 33&64.93 \(\pm\) 31&45.60 \(\pm\) 31&38.56 \(\pm\) 32\\
+30\%&55.83 \(\pm\) 26&58.77 \(\pm\) 28&40.00 \(\pm\) 28&35.45 \(\pm\) 30\\
     \hline
@@
 \textbf{Gap percentage} & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
-10\%&63.64 \(\pm\) 44.84&63.64 \(\pm\) 43.12&62.12 \(\pm\) 46.11&53.03 \(\pm\) 44.28\\
-20\%&73.33 \(\pm\) 31.51&63.33 \(\pm\) 35.85&41.11 \(\pm\) 27.88&37.78 \(\pm\) 30.94\\
-30\%&70.28 \(\pm\) 31.12&59.58 \(\pm\) 23.89&39.44 \(\pm\) 27.10&37.71 \(\pm\) 33.32\\
+\hline
+10\%&63.64 \(\pm\) 45&63.64 \(\pm\) 43&62.12 \(\pm\) 46&53.03 \(\pm\) 44\\
+20\%&73.33 \(\pm\) 32&63.33 \(\pm\) 36&41.11 \(\pm\) 28&37.78 \(\pm\) 31\\
+30\%&70.28 \(\pm\) 31&59.58 \(\pm\) 24&39.44 \(\pm\) 27&37.71 \(\pm\) 33\\
 \hline
 \end{tabular}
-\caption {Tatar--Russian, casual conversations: Average number of gaps successfully filled (\%), using a synonym list} \label{table:rus-casual} 
+\caption {Tatar--Russian, casual conversations: Average number of gaps successfully filled (\%), using a synonym list} 
+\label{table:rus-casual} 
 \end{table*}
@@
 \textbf{Gap percentage} & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
-10\%&53.02 \(\pm\) 40.31&68.21 \(\pm\) 34.99&38.77 \(\pm\) 37.82&32.95 \(\pm\) 34.88\\
-20\%&60.57 \(\pm\) 24.98&65.76 \(\pm\) 24.16&50.28 \(\pm\) 33.55&48.31 \(\pm\) 33.52\\
-30\%&50.06 \(\pm\) 26.25&48.30 \(\pm\) 28.87&40.05 \(\pm\) 27.37&34.15 \(\pm\) 28.73\\
+\hline
+10\%&53.02 \(\pm\) 40&68.21 \(\pm\) 35&38.77 \(\pm\) 38&32.95 \(\pm\) 35\\
+20\%&60.57 \(\pm\) 25&65.76 \(\pm\) 24&50.28 \(\pm\) 34&48.31 \(\pm\) 34\\
+30\%&50.06 \(\pm\) 26&48.30 \(\pm\) 29&40.05 \(\pm\) 27&34.15 \(\pm\) 29\\
 \hline
@@
 \textbf{Gap percentage} & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
-10\%&53.33 \(\pm\) 37.86&59.77 \(\pm\) 43.66&57.47 \(\pm\) 42.12&49.44 \(\pm\) 39.08\\
-20\%&59.33 \(\pm\) 33.89&48.56 \(\pm\) 35.40&46.53 \(\pm\) 31.99&28.97 \(\pm\) 28.79\\
-30\%&58.14 \(\pm\) 22.42&61.49 \(\pm\) 22.11&40.51 \(\pm\) 30.10&34.81 \(\pm\) 27.01\\
+\hline
+10\%&53.33 \(\pm\) 38&59.77 \(\pm\) 44&57.47 \(\pm\) 42&49.44 \(\pm\) 39\\
+20\%&59.33 \(\pm\) 34&48.56 \(\pm\) 35&46.53 \(\pm\) 32&28.97 \(\pm\) 29\\
+30\%&58.14 \(\pm\) 22&61.49 \(\pm\) 22&40.51 \(\pm\) 30&34.81 \(\pm\) 27\\
 \hline
 \end{tabular}
-\caption {Tatar--Russian, news: Average number of gaps successfully filled (\%), using a synonym list} \label{table:rus-news} 
+\caption {Tatar--Russian, news: Average number of gaps successfully filled (\%), using a synonym list} 
+\label{table:rus-news} 
 \end{table*}
@@
 The version of the toolkit used in this
-\comment{EA to JA: by our repository I mean my github, is that ok?}
-experiment may be downloaded from our repository.\footnote{}\comment{MLF: Do the same as in the other GitHUb comment above}
+experiment may be downloaded from our repository.\footnote{\phantom{\url{https://github.com/Sereni/Appraise/tree/1e9d735faee64d1b97fb343ab111ace6a64509d7}}}
 
@@
 \begin{enumerate}
-\item If the experiment is conducted with different gap densities, each consecutive set of sentences should be derived from the set with lower gap density.
-To put another way, the sentence with e.g.\ 20\% gaps should contain all the gaps of the same sentence in 10\% mode, plus any additional gaps;\comment{MLF: no reason is given for this, I am not ready to agree.
-Consider removing this}
-\item Unless the evaluation is targeted at a specific text domain, it may be beneficial to include a stylistic variety of texts in the initial corpus.
-Neighboring sentences on the same topic may assist the users in gap-filling tasks;\comment{wasn't this already reported by \cite{oregan13}?}
+\item As reported in \cite{oregan13}, unless the evaluation is targeted at a specific text domain, it may be beneficial to include a stylistic variety of texts in the initial corpus.
+Neighboring sentences on the same topic may assist the users in gap-filling tasks;
 \item If possible, increase the number of evaluators, or reduce the number of questions per participant.
 In the above experiments each participant filled from 110 to 187 gaps, divided into small groups.
-Reducing the amount of work may increase task completion rate;\comment{MLF's crazy idea: recaptchas?}
+Reducing the amount of work may increase task completion rate;\comment{MLF's crazy idea: recaptchas?}\comment{EA: the participants did not leave the gaps empty, they just did not attempt the next portions of evaluation after finishing some and getting bored.
+or did I misunderstand the use of captchas?
+}
 \item To account for the adaptation period, provide training tasks before the main evaluations take place.
@@
 
+\comment{EA: urls do not show up in web citations, how can I fix that?}
 \bibliography{2015-eamt-assim}

### 2015-03-06 2015-eamt-assim/2015-eamt-assim.tex
fixes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59237 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-\newcommand{\comment}[1]{\marginpar{\scriptsize\sf \textcolor{blue}{#1}}}
-%\newcommand{\comment}[1]{}
+%\newcommand{\comment}[1]{\marginpar{\scriptsize\sf \textcolor{blue}{#1}}}
+\newcommand{\comment}[1]{}
 
@@
 
-Despite the fact that, as a result of widespread usage of online MT, assimilation (or gisting) is currently the most frequent application of MT (in 2012, daily output of Google Translate matched the yearly output of human translations \citep{google}), few methodologies are established for assimilation evaluation of MT.
+Despite the fact that, as a result of widespread usage of online MT, assimilation (or gisting) is currently the most frequent application of MT (in 2012, daily output of Google Translate matched the yearly output of human translations,\footnote{\url{http://googleblog.blogspot.co.uk/2012/04/breaking-down-language-barriersix-years.html}} few methodologies are established for assimilation evaluation of MT.
 The methods include post-editing and comparison by bilingual experts \citep{ginesti09},\comment{MLF: should we also cite the original WMT reference cited by Ginestí et al.?} and multiple choice tests \citep{jones07,trosterud12}.
@@
 
-\section{Acknowledgments}
+\section*{Acknowledgments}
 
-This work has been partly funded by the Spanish Ministerio de Econom{\'i}a y Competitividad through project TIN2012-32615 and by the Google Summer of Code 2014 programme.
+%This work has been partly funded by the Spanish Ministerio de Econom{\'i}a y Competitividad through project TIN2012-32615 and by the Google Summer of Code 2014 programme.
 We would like to thank the volunteers who participated in the evaluations.

### 2015-03-06 2015-eamt-assim/2015-eamt-assim.tex
I broke the paper; can't fit it in 8 pages, damn
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59239 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{eamt15}
+%\usepackage{float}
  \setlength\titlebox{6.5cm}    % Expanding the titlebox
@@
 \usepackage{graphicx}
-%\usepackage{microtype} % This saves some space and makes nice typesetting choices
+\usepackage{microtype} % This saves some space and makes nice typesetting choices
 
@@
 \newcommand{\comment}[1]{}
+\newcommand{\blankout}[1]{\phantom{#1}} % anonymous
+%\newcommand{\blankout}[1]{#1} % non-anonymous
 
@@
 
-\author{Ekaterina Ageeva\\
-  School of Linguistics\\
-  Higher School of Economics\\
-  Moscow, Russia\\
-  {\tt evageeva\_2@edu.hse.ru}\\[2ex]
-  \textbf{Francis M. Tyers}\\
-  HSL-fakultetet\\ 
-  UiT Norgga \'{a}rktala\v{s} universitehta \\
-  9017 Romsa, Norway \\
-  {\tt francis.tyers@uit.no}
+\author{\blankout{Ekaterina Ageeva}\\
+  \blankout{School of Linguistics}\\
+  \blankout{Higher School of Economics}\\
+  \blankout{Moscow, Russia}\\
+  \blankout{{\tt evageeva\_2@edu.hse.ru}}\\[2ex]
+  \blankout{\textbf{Francis M. Tyers}}\\
+  \blankout{HSL-fakultetet}\\ 
+  \blankout{UiT Norgga \'{a}rktala\v{s} universitehta} \\
+  \blankout{9017 Romsa, Norway} \\
+  \blankout{{\tt francis.tyers@uit.no}}
   \And
-  Mikel L. Forcada\\
-  Dept. Llenguatges i Sistemes Inform\`{a}tics\\
-  Universitat d'Alacant, Spain \\
-  {\tt mlf@dlsi.ua.es}  \\[2ex]
-  \textbf{Juan Antonio P\'{e}rez-Ortiz} \\
-  Dept. Llenguatges i Sistemes Inform\`{a}tics\\
-  Universitat d'Alacant, Spain \\
-  {\tt japerez@dlsi.ua.es}
+  \blankout{Mikel L. Forcada}\\
+  \blankout{Dept. Llenguatges i Sistemes Inform\`{a}tics}\\
+  \blankout{Universitat d'Alacant, Spain} \\
+  \blankout{{\tt mlf@dlsi.ua.es}}  \\[2ex]
+  \blankout{\textbf{Juan Antonio P\'{e}rez-Ortiz}} \\
+  \blankout{Dept. Llenguatges i Sistemes Inform\`{a}tics}\\
+  \blankout{Universitat d'Alacant, Spain} \\
+  \blankout{{\tt japerez@dlsi.ua.es}}
 }
@@
 \maketitle
-\renewcommand{\baselinestretch}{0.96} % the undetectable 3% squeeze
+\renewcommand{\baselinestretch}{0.97} % the undetectable 3% squeeze
 \begin{abstract}
@@
 
-Despite the fact that, as a result of widespread usage of online MT, assimilation (or gisting) is currently the most frequent application of MT (in 2012, daily output of Google Translate matched the yearly output of human translations,\footnote{\url{http://googleblog.blogspot.co.uk/2012/04/breaking-down-language-barriersix-years.html}} few methodologies are established for assimilation evaluation of MT.
+Despite the fact that, as a result of widespread usage of online MT, assimilation (or gisting) is currently the most frequent application of MT (in 2012, daily output of Google Translate matched the yearly output of human translations\footnote{\url{http://googleblog.blogspot.co.uk/2012/04/breaking-down-language-barriersix-years.html}}), few methodologies are established for assimilation evaluation of MT.
 The methods include post-editing and comparison by bilingual experts \citep{ginesti09},\comment{MLF: should we also cite the original WMT reference cited by Ginestí et al.?} and multiple choice tests \citep{jones07,trosterud12}.
@@
 In this work we extend the evaluation to two more language pairs: Basque--Spanish and Tatar-Russian.
-The former pair, while not producing the output suitable for post-editing, is a good example of an assimilation MT system.
+The former pair, while not producing output suitable for post-editing, is a good example of an assimilation MT system.
 In addition, Basque and Spanish are not mutually understandable, and therefore constitute a good pair for evaluation.
@@
 Should I remove?} etc.), checks evaluators' answers, and calculates and reports generalized results.
-This system is integrated into the Appraise MT evaluation platform \citep{federmann12}; the code is open source and is available on GitHub.\footnote{\phantom{\url{https://github.com/Sereni/Appraise}}}
+This system is integrated into the Appraise MT evaluation platform \citep{federmann12}; the code is open source and is available on GitHub.\footnote{\blankout{\url{https://github.com/Sereni/Appraise}}}
 
@@
 \label{sec:methodology}
+
+\comment{MLF the gap percentage table 2 is not referenced in the text; also, i have changed it to have the \textbf{Ref} row in the table?}
+
+This section discusses the reasoning behind the gap-filling method and task structure. The gap-filling method 
+of evaluating machine translation for assimilation purposes is based on the following hypothesis: a reader's understanding 
+of a given text correlates with the number of words they are able to correctly restore in the text. Therefore, the 
+base of an assimilation task is a (reference) sentence, where some of the words are blacked out, or removed. The sentence 
+is produced by a human (as opposed to machine-translated), and it is in the language known to evaluators, which is also 
+the \emph{target language} of the machine translation system.
+The additional elements of the task are what we call hints, or extra sentences that help the participant to understand 
+the main sentence. There are two types of hints: first, the \emph{source}, which is semantically equivalent to the \emph{reference}, 
+also human-produced, but in the source language of the pair. The second type is the \emph{machine-translated} hint, which 
+comes from the machine translation of the \emph{source} sentence. Table~\ref{table:example} shows a sample task, and Figure~\ref{figure:screenshot} shows the task in the online evaluation environment.
+%\comment{Use the highest possible resolution for the figure so that aliasing is imperceptible}
+\begin{figure*}
+  \centering
+\includegraphics[width=1.0\textwidth,resolution=144]{appraise-scr}
+ \caption{An example set of sentences in the online environment. The task is Russian legal text with 30\% gaps.}
+\label{figure:screenshot}
+\end{figure*}
 
@@
 
+In the course of the experiment, following \cite{oregan13}, we offer these hint combinations:
+\begin{description}\itemsep 0ex
+\item[Reference sentence only:] The participants are asked to fill the gaps without being given
+any context. This task serves as a baseline score and as an indicator of gaps
+that can be completed using common knowledge or language intuition (e.g.
+idioms and strong collocations).
+\item[Reference sentence and source sentence:] By setup, the
+participants have no command of the source language, however, it may help them to fill
+in proper nouns or loan words.
+\item[Reference sentence and MT hint:] In addition to the reference sentence, the
+participants see the source sentence translated via the MT system, in this case Apertium \citep{forcada11}. This type of task is
+used for measuring the contribution of machine translation to understanding the
+gist of the text.
+\item[Reference sentence and both hints:] This task is
+added to check whether MT and source provide complementary hints.
+\end{description}
+
+In order to prepare the evaluation questions, we determine and remove keywords from the 
+reference sentences. We consider two parameters: the list of allowed parts of speech (PoS), and the 
+number of gaps relative to sentence length (``gap density''). For the evaluations described in 
+this paper we use gap densities of 10, 20 and 30 percent (Table~\ref{table:percentage}), and the following parts of speech: noun, 
+proper noun, adjective, adverb and lexical verb (as opposed to auxiliary verb).
+
 \begin{table*}
@@
 \end{table*}
-\comment{MLF the gap percentage table 2 is not referenced in the text; also, i have changed it to have the \textbf{Ref} row in the table?}
-
-This section discusses the reasoning behind the gap-filling method and task structure. The gap-filling method 
-of evaluating machine translation for assimilation purposes is based on the following hypothesis: a reader's understanding 
-of a given text correlates with the number of words they are able to correctly restore in the text. Therefore, the 
-base of an assimilation task is a (reference) sentence, where some of the words are blacked out, or removed. The sentence 
-is produced by a human (as opposed to machine-translated), and it is in the language known to evaluators, which is also 
-the \emph{target language} of the machine translation system.
-The additional elements of the task are what we call hints, or extra sentences that help the participant to understand 
-the main sentence. There are two types of hints: first, the \emph{source}, which is semantically equivalent to the \emph{reference}, 
-also human-produced, but in the source language of the pair. The second type is the \emph{machine-translated} hint, which 
-comes from the machine translation of the \emph{source} sentence. Table~\ref{table:example} shows a sample task, and Figure~\ref{figure:screenshot} shows the task in the online evaluation environment.
-%\comment{Use the highest possible resolution for the figure so that aliasing is imperceptible}
-
-\begin{figure*}
-  \centering
-\includegraphics[width=1.0\textwidth,resolution=144]{appraise-scr}
- \caption{An example set of sentences in the online environment. The task is Russian legal text with 30\% gaps.}
-\label{figure:screenshot}
-\end{figure*}
-
-In the course of the experiment, following \cite{oregan13}, we offer these hint combinations:
-\begin{description}
-\item[Reference sentence only:] The participants are asked to fill the gaps without being given
-any context. This task serves as a baseline score and as an indicator of gaps
-that can be completed using common knowledge or language intuition (e.g.
-idioms and strong collocations).
-\item[Reference sentence and source sentence:] By setup, the
-participants have no command of the source language, however, it may help them to fill
-in proper nouns or loan words.
-\item[Reference sentence and MT hint:] In addition to the reference sentence, the
-participants see the source sentence translated via the MT system Apertium \citep{forcada11}. This type of task is
-used for measuring the contribution of machine translation to understanding the
-gist of text.
-\item[Reference sentence and both hints:] This task is
-added to check whether MT and source provide complementary hints.
-\end{description}
-
-In order to prepare the evaluation questions, we determine and remove the keywords in the 
-reference sentences. We consider two parameters: the list of allowed parts of speech (PoS), and the 
-number of gaps relative to sentence length (``gap density''). For the evaluations described in 
-this paper we use gap densities of 10, 20 and 30 percent (Table~\ref{table:percentage}), and the following parts of speech: noun, 
-proper noun, adjective, adverb and lexical verb (as opposed to auxiliary verb).
 
@@
 We therefore use parallel text sources, which provide the same sentence in two languages simultaneously:
-\begin{enumerate}
+\begin{enumerate}\itemsep 0ex
 \item  For Basque--Spanish, from the corpus of legal texts ``Memorias de traducci\'on del
-Servicio Oficial de Traductores del IVAP'';\footnote{\url{http://tinyurl.com/ivaptm2}} 
+Servicio Oficial de Traductores del IVAP''.\footnote{\url{http://tinyurl.com/ivaptm2}} 
 \item  For Tatar-Russian, from the following sources on three different topics:
-  \begin{enumerate}
+  \begin{enumerate}\itemsep 0ex
     \item  Casual conversations, from a textbook\footnote{\rus{Литвинов И.Л.
@@
 manually distributed between the participants by the following rules:
-\begin{enumerate}
+\begin{enumerate}\itemsep 0ex
 \item  Each participant evaluates every sentence (understood as a succession of words),
@@
 context sentence.
-For example, the word \emph{asumir} is the original answer in the Spanish sentence \emph{Aprender a jugar y divertirse en el agua sin asumir riesgos}.
-However, two or more evaluators gave a different answer, \emph{corder}.
+For example, the word \emph{asumir} is the original answer in the Spanish sentence \emph{Aprender a jugar y divertirse en el agua sin asumir riesgos} ('Learning to play and have fun in the water without taking risks').
+However, two or more evaluators gave a different answer, \emph{correr} (\emph{correr riesgos}, 'running risks').
 Based on this data, a native speaker of the target language decides
@@
 The synonym lists for Basque--Spanish and Tatar--Russian contain 52 and 25 words, respectively.
-Time taken to compile each list depends on the number of candidate synonyms, and in our case was approximately 30 minutes.
+The time taken to compile each list depends on the number of candidate synonyms, and in our case was approximately 30 minutes.
 
@@
 %\comment{FMT: We should probably integrate the other tables, and refer to them using labels.} 
-The results are presented in this section. Tables \ref{table:res-eus-spa} and \ref{table:res-tat-rus} show
+The results are presented in this section. Tables \ref{table:res-eus-spa} and \ref{table:res-tat-rus} 
+show
 the proportion and standard deviation of correct answers depending on evaluation mode and gap density.
-The evaluators' correct answer percentage is averaged over the number of evaluators.
-In addition to the percentage of correct answers we kept a record of time taken to fill the gaps in one sentence.
-To reduce the noise from participants who were distracted during evaluation, when
-calculating times we remove all the results over 6 minutes (the statistical mode is approximately
-two minutes). Typical time taken to complete one question varies from under one minute for tasks without hints and few gaps, to approximately two minutes for tasks with more hints and gaps.
-
-% expected results
-We expect to see the following trend: scores received in different task modes inside one gap density should be observed, from highest to lowest, in tasks with MT and source hint, then MT hint only, then source hint only, and finally, no hint.
-We also expect that with the increase in gap density, time taken to fill the gaps should also increase, and percentage of correct answers should decrease.
-
-% a few notes on time
-The latter trend holds: the average time taken to fill the gaps increases and the average percentage of correct answers decreases as the relative number of gaps goes up.
-The larger number of gaps in the sentence makes it more
-difficult to predict the answer based on the context, and also leaves more room for
-mistakes.
-Exploring different percentage-mode combinations, we may note that the 10\% no-hint tasks take the least time to complete.
-We would have expected longer completion time, since the participant must come up with their own answer unassisted.
-However, in the no-hint task the participant is required to read only one (reference) sentence, as opposed to two or three (reference and hints) in other tasks.
-Also, the number of gaps in 10\%-gap tasks is low, as it never exceeds three.
-We found that, as opposed to trying to devise the best word for no-hint gaps, the participants often resorted to filing these gaps with random words, which takes little time.
-
-% this section talks about percentages in eu-es and tat-rus, and how everything is nice.
-We will now discuss the percentage of correct answers based on task type.
-In general, tasks with MT hints score higher than tasks without MT hints.
-This aligns well with our expectations and suggests than machine translation helps to understand the provided text.
-In addition, tasks with source hints are completed better than tasks without hints, and the same relation holds between MT+source and MT-only types of tasks.
-
-Two records break this trend: the no-hint 10\% sentences in Basque--Spanish, which scored significantly higher than the source-hint in the same category, and MT+source 10\% sentences in Tatar--Russian, which we would have expected to score higher than the corresponding MT task.
-In the first case, this is largely due to the use of synonyms list.
-Before taking synonyms into account, the scores were 32 and 35 percent for source and no-hint tasks, respectively.
-This still shows a small difference in favor of no-hint tasks.
-However, the latter percentage increases significantly after we extend the answer list with synonyms.
-Such an increase suggests that, in this case, the content words were restored by
-semantic context rather than through strong collocation.
-The second pattern, low scores in Tatar--Russian 10\% MT+source, does not stem from the task content.
-Instead, it is the result of the fixed order of tasks: the participants have always been given MT+source 10\% sentences first, followed by other task types.\comment{MLF: Why?
-Not random?
-\cite{oregan13} did random!
-Why was this changed?}The participants have not received any training tasks before the main evaluations.
-Therefore, it is possible that the accommodation period is responsible for lower-than-expected scores in this mode of evaluation.
-
-It remains questionable whether we can compare results for different gap densities. The 10, 20 and 30\% sets were comprised
-of the same sentences. However, in each case different words were removed. It appears
-that some content words are easier to fill than the others. This may explain why in
-Basque--Spanish the 20\% MT tasks are completed with better accuracy than 10\% tasks.\comment{MLF: I think this shows that our statistics are weak, should we say something to excuse ourselves?}
-
-% participants and their suffering
-It is worth noting that many participants reported feeling frustrated in
-the course of evaluations, especially while working on the no-hint tasks. 6 out of 49
-participants quit the experiment before completing it.
-Considering the importance of receiving the full set of evaluations, we must address the issue of participant motivation in the upcoming experiments.
-
-\subsection{Annotator agreement}
-After obtaining the results we calculated Krippendorff's alpha~\citep{krippendorff70} measure to represent annotator agreement, shown in Tables~\ref{table:alpha-eus-spa} and~\ref{table:alpha-tat-rus}.
-\begin{table*}
-\centering
-  \begin{tabular}{|l|r|r|r|r|}
-  \hline
- \multirow{2}{*}{\textbf{Density}} & \multicolumn{4}{|c|}{Evaluation mode} \\\cline{2-5}
-                                            & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
-\hline
-10\%&0.496&0.517&0.400&0.124\\
-20\%&0.714&0.700&0.358&0.275\\
-30\%&0.559&0.430&0.406&0.300\\
-
-\hline
-\end{tabular}
-\caption {Basque--Spanish: Krippendorff Alpha measure} 
-\label{table:alpha-eus-spa}
-\end{table*}
-
-\begin{table*}
-\centering
-  \begin{tabular}{|l|r|r|r|r|}
-  \hline
- \multirow{2}{*}{\textbf{Density}} & \multicolumn{4}{|c|}{Evaluation mode} \\\cline{2-5}
-                                            & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
-\hline
-10\%&0.598&0.459 &0.711 &0.517 \\
-20\%&0.740&0.667 &0.473 &0.261 \\
-30\%&0.534&0.581 &0.411 &0.412 \\
-\hline
-\end{tabular}
-\caption {Tatar--Russian: Krippendorff Alpha measure} 
-\label{table:alpha-tat-rus}
-\end{table*}
-We selected this measure because of its compatibility with more than two annotators per task and the missing data (not all the gaps were evaluated).
-To calculate Krippendorff's alpha we used an algorithm implementation by Thomas Grill,\footnote{\url{http://grrrr.org/data/dev/krippendorff_alpha/}} dividing the answers in each gap into two categories: correct and incorrect.
-The previously obtained synonym lists were taken into account, i.e.
-if the two answers are different but both correct, they fall into one category.
-The measure was calculated separately for each hint and percentage combination.
-
-As the MT and MT+source hints are introduced, the agreement between annotators increases (measures closer to 1): the annotators are more consistently correct or incorrect in each given sentence.
-The agreement measure for the same sentences without hints is closer to zero, which contributes to the reliability of our methodology.
-We note the outlier score in Tatar--Russian 10\% source tasks, which has the most contribution from the news texts (agreement of 0.917).
-This set of sentences contains many loan words, which have similar form in Tatar and Russian (e.g.
-president, minister, championship), and are understood by Russian speakers.
-The gaps with loan words have mostly been filled correctly, while there was some disagreement in other gaps.
-
-% a little on categories
-\subsection{Results for different domains}
-For the Tatar--Russian language pair the participants were offered texts from three different domains (in equal proportions): casual conversations, legal texts and news.
-The results by domains are displayed in tables \ref{table:rus-casual}, \ref{table:rus-legal}, and \ref{table:rus-news}.
-The MT system used in the evaluation has been targeted to translate texts from all three of the domains.
-Taking into consideration the above discussion of 10\% MT+Source tasks, we observe similar results across the three categories.
-Note that, in casual and news texts, the source sentences paired with MT significantly improve participants' performance compared to MT-only task mode.
-This may be due to the fact that many words are borrowed from Russian into Tatar, and are in fact understood by Russian speakers.
-
-% results with synonyms
 
@@
 \end{table*}
+
+The evaluators' correct answer percentage is averaged over the number of evaluators.
+In addition to the percentage of correct answers we kept a record of the time taken to fill the gaps in one sentence.
+To reduce the noise from participants who were distracted during evaluation, when
+calculating times we remove all the results over 6 minutes (the statistical mode is approximately
+two minutes). The typical time taken to complete one question varies from under one minute for tasks without hints and few gaps, to approximately two minutes for tasks with more hints and gaps.
+
+% expected results
+We expect to see the following trend: scores received in different task modes inside one gap density should be observed, from highest to lowest, in tasks with MT and source hint, then MT hint only, then source hint only, and finally, no hint.
+We also expect that with the increase in gap density, the time taken to fill the gaps should also increase, and the percentage of correct answers should decrease.
+
+% a few notes on time
+The latter trend holds: the average time taken to fill the gaps increases and the average percentage of correct answers decreases as the relative number of gaps goes up.
+The larger number of gaps in the sentence makes it more
+difficult to predict the answer based on the context, and also leaves more room for
+mistakes.
+Exploring different percentage-mode combinations, we may note that the 10\% no-hint tasks take the least time to complete.
+We would have expected longer completion time, since the participant must come up with their own answer unassisted.
+However, in the no-hint task the participant is required to read only one (reference) sentence, as opposed to two or three (reference and hints) in other tasks.
+Also, the number of gaps in 10\%-gap tasks is low, as it never exceeds three.
+We found that, as opposed to trying to devise the best word for no-hint gaps, the participants often resorted to filing these gaps with random words, which takes little time.
+
+% this section talks about percentages in eu-es and tat-rus, and how everything is nice.
+We will now discuss the percentage of correct answers based on task type.
+In general, tasks with MT hints score higher than tasks without MT hints.
+This aligns well with our expectations and suggests than machine translation helps to understand the provided text.
+In addition, tasks with source hints are completed better than tasks without hints, and the same relation holds between MT+source and MT-only types of tasks.
+
+Two records break this trend: the no-hint 10\% sentences in Basque--Spanish, which scored significantly higher than the source-hint in the same category, and MT+source 10\% sentences in Tatar--Russian, which we would have expected to score higher than the corresponding MT task.
+In the first case, this is largely due to the use of synonyms list.
+Before taking synonyms into account, the scores were 32 and 35 percent for source and no-hint tasks, respectively.
+This still shows a small difference in favor of no-hint tasks.
+However, the latter percentage increases significantly after we extend the answer list with synonyms.
+Such an increase suggests that, in this case, the content words were restored by
+semantic context rather than through strong collocation.
+The second pattern, low scores in Tatar--Russian 10\% MT+source, does not stem from the task content.
+Instead, it is the result of the fixed order of tasks: the participants have always been given MT+source 10\% sentences first, followed by other task types.
+\comment{MLF: Why?
+Not random?
+\cite{oregan13} did random!
+Why was this changed?}The participants have not received any training tasks before the main evaluations.
+Therefore, it is possible that the accommodation period is responsible for lower-than-expected scores in this mode of evaluation.
+
+It remains questionable whether we can compare results for different gap densities. The 10, 20 and 30\% sets were comprised
+of the same sentences. However, in each case different words were removed. It appears
+that some content words are easier to fill than the others. This may explain why in
+Basque--Spanish the 20\% MT tasks are completed with better accuracy than 10\% tasks.\comment{MLF: I think this shows that our statistics are weak, should we say something to excuse ourselves?}
+
+% participants and their suffering
+It is worth noting that many participants reported feeling frustrated in
+the course of evaluations, especially while working on the no-hint tasks. 6 out of 49
+participants quit the experiment before completing it.
+Considering the importance of receiving the full set of evaluations, we must address the issue of participant motivation in the upcoming experiments.
+
+\subsection{Annotator agreement}
+After obtaining the results we calculated Krippendorff's alpha~\citep{krippendorff70} measure to represent annotator agreement, shown in Tables~\ref{table:alpha-eus-spa} and~\ref{table:alpha-tat-rus}.
+\begin{table}
+\centering
+  \begin{tabular}{|l|r|r|r|r|}
+  \hline
+ \multirow{2}{*}{\textbf{Density}} & \multicolumn{4}{|c|}{Evaluation mode} \\\cline{2-5}
+                                            & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
+\hline
+10\%&0.496&0.517&0.400&0.124\\
+20\%&0.714&0.700&0.358&0.275\\
+30\%&0.559&0.430&0.406&0.300\\
+
+\hline
+\end{tabular}
+\caption {Basque--Spanish: Krippendorff Alpha measure} 
+\label{table:alpha-eus-spa}
+\end{table}
+
+\begin{table}
+\centering
+  \begin{tabular}{|l|r|r|r|r|}
+  \hline
+ \multirow{2}{*}{\textbf{Density}} & \multicolumn{4}{|c|}{Evaluation mode} \\\cline{2-5}
+                                            & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
+\hline
+10\%&0.598&0.459 &0.711 &0.517 \\
+20\%&0.740&0.667 &0.473 &0.261 \\
+30\%&0.534&0.581 &0.411 &0.412 \\
+\hline
+\end{tabular}
+\caption {Tatar--Russian: Krippendorff Alpha measure} 
+\label{table:alpha-tat-rus}
+\end{table}
+We selected this measure because of its compatibility with more than two annotators per task and the missing data (not all the gaps were evaluated).
+To calculate Krippendorff's alpha we used an algorithm implementation by Thomas Grill,\footnote{\url{http://grrrr.org/data/dev/krippendorff_alpha/}} dividing the answers in each gap into two categories: correct and incorrect.
+The previously obtained synonym lists were taken into account, i.e.
+if the two answers are different but both correct, they fall into one category.
+The measure was calculated separately for each hint and percentage combination.
+
+As the MT and MT+source hints are introduced, the agreement between annotators increases (measures closer to 1): the annotators are more consistently correct or incorrect in each given sentence.
+The agreement measure for the same sentences without hints is closer to zero, which contributes to the reliability of our methodology.
+We note the outlier score in Tatar--Russian 10\% source tasks, which has the most contribution from the news texts (agreement of 0.917).
+This set of sentences contains many loan words, which have similar form in Tatar and Russian (e.g.
+president, minister, championship), and are understood by Russian speakers.
+The gaps with loan words have mostly been filled correctly, while there was some disagreement in other gaps.
+
+% a little on categories
+\subsection{Results for different domains}
+For the Tatar--Russian language pair the participants were offered texts from three different domains (in equal proportions): casual conversations, legal texts and news.
+The results by domains are displayed in tables \ref{table:rus-casual}, \ref{table:rus-legal}, and \ref{table:rus-news}.
+The MT system used in the evaluation has been targeted to translate texts from all three of the domains.
+Taking into consideration the above discussion of 10\% MT+Source tasks, we observe similar results across the three categories.
+Note that, in casual and news texts, the source sentences paired with MT significantly improve participants' performance compared to MT-only task mode.
+This may be due to the fact that many words are borrowed from Russian into Tatar, and are in fact understood by Russian speakers.
+
+% results with synonyms
 
@@
 The version of the toolkit used in this
-experiment may be downloaded from our repository.\footnote{\phantom{\url{https://github.com/Sereni/Appraise/tree/1e9d735faee64d1b97fb343ab111ace6a64509d7}}}
+experiment may be downloaded from our repository.\footnote{\blankout{\url{https://github.com/Sereni/Appraise/tree/1e9d735faee64d1b97fb343ab111ace6a64509d7}}}
 
@@
 any machine translation system. Based on our experience, we would like to suggest the following amendments to the procedure:
-\begin{enumerate}
+\begin{enumerate}\itemsep 0ex
 \item As reported in \cite{oregan13}, unless the evaluation is targeted at a specific text domain, it may be beneficial to include a stylistic variety of texts in the initial corpus.

### 2015-03-06 2015-eamt-assim/2015-eamt-assim.tex
back to 8 pages
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59240 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{eamt15}
-%\usepackage{float}
 \setlength\titlebox{6.5cm}    % Expanding the titlebox
@@
 
+\begin{table*}
+  \centering
+  \begin{tabular}{|l|l|}
+     \hline
+     \textbf{Ref}  & Ayudas econ\'{o}micas para el tratamiento de toxicoman\'{i}as en comunidades terap\'{e}uticas no concertadas. \\
+     \textbf{Task} & Ayudas econ\'{o}micas para el \{ \} de toxicoman\'{i}as en comunidades terap\'{e}uticas no concertadas. \\
+     \textbf{Src}  & Komunitate terapeutiko itundu gabeetan toxikomaniak tratatzeko diru-laguntzak ematea. \\
+     \textbf{MT}   & Comunidad terap\'{e}utico pactar gabeetan toxikomaniak las-ayudas de dinero para tratar dar. \\
+     \hline
+  \end{tabular}
+  \caption{An example group of sentences showing the gapped sentence and hint types. Reference, MT and task sentences are in Spanish, the source sentence is in Basque.} 
+  \label{table:example}
+\end{table*}
+
+\begin{table*}
+  \centering
+  \begin{tabular}{|l|l|}
+     \hline
+     \textbf{Ref} & \rus{{\small Примерно полчаса; вам нужно выйти через 7 остановок, потом пройти ещё около 100 метров.}} \\
+     \hline
+     \textbf{10\%} & \rus{{\small Примерно полчаса; вам нужно выйти через 7 \{ \}, потом пройти ещё около 100 метров.}} \\
+     \textbf{20\%} & \rus{{\small \{ \} полчаса; вам нужно \{ \} через 7 остановок, \{ \} пройти ещё около 100 метров.}} \\
+     \textbf{30\%} & \rus{{\small Примерно полчаса; вам нужно \{ \} через 7 \{ \}, потом пройти \{ \} около 100 \{ \}.}} \\
+     \hline
+  \end{tabular}
+  \caption{Example of different gap percentage settings for a Russian reference sentence.} 
+  \label{table:percentage}
+\end{table*}
 \comment{MLF the gap percentage table 2 is not referenced in the text; also, i have changed it to have the \textbf{Ref} row in the table?}
@@
 %\comment{Use the highest possible resolution for the figure so that aliasing is imperceptible}
+
 \begin{figure*}
@@
 \end{figure*}
-
-\begin{table*}
-  \centering
-  \begin{tabular}{|l|l|}
-     \hline
-     \textbf{Ref}  & Ayudas econ\'{o}micas para el tratamiento de toxicoman\'{i}as en comunidades terap\'{e}uticas no concertadas. \\
-     \textbf{Task} & Ayudas econ\'{o}micas para el \{ \} de toxicoman\'{i}as en comunidades terap\'{e}uticas no concertadas. \\
-     \textbf{Src}  & Komunitate terapeutiko itundu gabeetan toxikomaniak tratatzeko diru-laguntzak ematea. \\
-     \textbf{MT}   & Comunidad terap\'{e}utico pactar gabeetan toxikomaniak las-ayudas de dinero para tratar dar. \\
-     \hline
-  \end{tabular}
-  \caption{An example group of sentences showing the gapped sentence and hint types. Reference, MT and task sentences are in Spanish, the source sentence is in Basque.} 
-  \label{table:example}
-\end{table*}
 
@@
 proper noun, adjective, adverb and lexical verb (as opposed to auxiliary verb).
-
-\begin{table*}
-  \centering
-  \begin{tabular}{|l|l|}
-     \hline
-     \textbf{Ref} & \rus{{\small Примерно полчаса; вам нужно выйти через 7 остановок, потом пройти ещё около 100 метров.}} \\
-     \hline
-     \textbf{10\%} & \rus{{\small Примерно полчаса; вам нужно выйти через 7 \{ \}, потом пройти ещё около 100 метров.}} \\
-     \textbf{20\%} & \rus{{\small \{ \} полчаса; вам нужно \{ \} через 7 остановок, \{ \} пройти ещё около 100 метров.}} \\
-     \textbf{30\%} & \rus{{\small Примерно полчаса; вам нужно \{ \} через 7 \{ \}, потом пройти \{ \} около 100 \{ \}.}} \\
-     \hline
-  \end{tabular}
-  \caption{Example of different gap percentage settings for a Russian reference sentence.} 
-  \label{table:percentage}
-\end{table*}
 
@@
 We therefore use parallel text sources, which provide the same sentence in two languages simultaneously:
-\begin{enumerate}\itemsep 0ex
+\begin{enumerate}
 \item  For Basque--Spanish, from the corpus of legal texts ``Memorias de traducci\'on del
-Servicio Oficial de Traductores del IVAP''.\footnote{\url{http://tinyurl.com/ivaptm2}} 
+Servicio Oficial de Traductores del IVAP'';\footnote{\url{http://tinyurl.com/ivaptm2}} 
 \item  For Tatar-Russian, from the following sources on three different topics:
-  \begin{enumerate}\itemsep 0ex
+  \begin{enumerate}
     \item  Casual conversations, from a textbook\footnote{\rus{Литвинов И.Л.
@@
 manually distributed between the participants by the following rules:
-\begin{enumerate}\itemsep 0ex
+\begin{enumerate}
 \item  Each participant evaluates every sentence (understood as a succession of words),
@@
 the proportion and standard deviation of correct answers depending on evaluation mode and gap density.
-
-\begin{table*}
-\centering
-  \begin{tabular}{|l|r|r|r|r|}
-  \hline
- \multirow{2}{*}{\textbf{Density}} & \multicolumn{4}{|c|}{Evaluation mode} \\\cline{2-5}
-                                            & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
-\hline
-10\%&62.19 \(\pm\) 32&58.04 \(\pm\) 28&39.90 \(\pm\) 39&48.79 \(\pm\) 40\\
-20\%&64.81 \(\pm\) 30&70.18 \(\pm\) 27&31.37 \(\pm\) 28&30.91 \(\pm\) 30\\
-30\%&48.13 \(\pm\) 26&39.55 \(\pm\) 24&25.94 \(\pm\) 20&18.07 \(\pm\) 18\\
-\hline
-\end{tabular}
-\caption {Basque--Spanish: Average number of gaps successfully filled (\%), using a synonym list} 
-\label{table:res-eus-spa}
-\end{table*}
-
-\begin{table*}
-\centering
-  \begin{tabular}{|l|r|r|r|r|}
-    \hline
-    \multirow{2}{*}{\textbf{Density}} & \multicolumn{4}{|c|}{Evaluation mode} \\\cline{2-5}
-                                            & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
-\hline
-10\%&63.76 \(\pm\) 41&57.02 \(\pm\) 42&53.52 \(\pm\) 43&45.96 \(\pm\) 41\\
-20\%&59.77 \(\pm\) 33&64.93 \(\pm\) 31&45.60 \(\pm\) 31&38.56 \(\pm\) 32\\
-30\%&55.83 \(\pm\) 26&58.77 \(\pm\) 28&40.00 \(\pm\) 28&35.45 \(\pm\) 30\\
-    \hline
-  \end{tabular}
-  \caption{Tatar--Russian: Average number of gaps successfully filled (\%), using a synonym list}
-  \label{table:res-tat-rus}
-\end{table*}
-
 The evaluators' correct answer percentage is averaged over the number of evaluators.
@@
 The second pattern, low scores in Tatar--Russian 10\% MT+source, does not stem from the task content.
-Instead, it is the result of the fixed order of tasks: the participants have always been given MT+source 10\% sentences first, followed by other task types.
-\comment{MLF: Why?
+Instead, it is the result of the fixed order of tasks: the participants have always been given MT+source 10\% sentences first, followed by other task types.\comment{MLF: Why?
 Not random?
@@
 After obtaining the results we calculated Krippendorff's alpha~\citep{krippendorff70} measure to represent annotator agreement, shown in Tables~\ref{table:alpha-eus-spa} and~\ref{table:alpha-tat-rus}.
-\begin{table}
+\begin{table*}
 \centering
@@
 \label{table:alpha-eus-spa}
-\end{table}
+\end{table*}
 
-\begin{table}
+\begin{table*}
 \centering
@@
 \label{table:alpha-tat-rus}
-\end{table}
+\end{table*}
 We selected this measure because of its compatibility with more than two annotators per task and the missing data (not all the gaps were evaluated).
@@
 % results with synonyms
+
+\begin{table*}
+\centering
+  \begin{tabular}{|l|r|r|r|r|}
+  \hline
+ \multirow{2}{*}{\textbf{Density}} & \multicolumn{4}{|c|}{Evaluation mode} \\\cline{2-5}
+                                            & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
+\hline
+10\%&62.19 \(\pm\) 32&58.04 \(\pm\) 28&39.90 \(\pm\) 39&48.79 \(\pm\) 40\\
+20\%&64.81 \(\pm\) 30&70.18 \(\pm\) 27&31.37 \(\pm\) 28&30.91 \(\pm\) 30\\
+30\%&48.13 \(\pm\) 26&39.55 \(\pm\) 24&25.94 \(\pm\) 20&18.07 \(\pm\) 18\\
+\hline
+\end{tabular}
+\caption {Basque--Spanish: Average number of gaps successfully filled (\%), using a synonym list} 
+\label{table:res-eus-spa}
+\end{table*}
+
+\begin{table*}
+\centering
+  \begin{tabular}{|l|r|r|r|r|}
+    \hline
+    \multirow{2}{*}{\textbf{Density}} & \multicolumn{4}{|c|}{Evaluation mode} \\\cline{2-5}
+                                            & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
+\hline
+10\%&63.76 \(\pm\) 41&57.02 \(\pm\) 42&53.52 \(\pm\) 43&45.96 \(\pm\) 41\\
+20\%&59.77 \(\pm\) 33&64.93 \(\pm\) 31&45.60 \(\pm\) 31&38.56 \(\pm\) 32\\
+30\%&55.83 \(\pm\) 26&58.77 \(\pm\) 28&40.00 \(\pm\) 28&35.45 \(\pm\) 30\\
+    \hline
+  \end{tabular}
+  \caption{Tatar--Russian: Average number of gaps successfully filled (\%), using a synonym list}
+  \label{table:res-tat-rus}
+\end{table*}
 
@@
 The version of the toolkit used in this
-experiment may be downloaded from our repository.\footnote{\blankout{\url{https://github.com/Sereni/Appraise/tree/1e9d735faee64d1b97fb343ab111ace6a64509d7}}}
+experiment may be downloaded from our repository.\footnote{\phantom{\url{https://github.com/Sereni/Appraise/tree/1e9d735faee64d1b97fb343ab111ace6a64509d7}}}
 
@@
 any machine translation system. Based on our experience, we would like to suggest the following amendments to the procedure:
-\begin{enumerate}\itemsep 0ex
+\begin{enumerate}
 \item As reported in \cite{oregan13}, unless the evaluation is targeted at a specific text domain, it may be beneficial to include a stylistic variety of texts in the initial corpus.

### 2015-03-06 engkaz-eamt-2015/engkaz-eamt-2015.tex
foo
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59241 72bbbca6-d526-0410-a7d9-f06f51895060

@@
   050040 Almaty, \\
-  {\tt xxx@xxx}\\
+  {\tt sun27aida@gmail.com}\\
   \And
@@
 Current system can translate simple sentences and questions. The English--Kazakh bilingual dictionary 
-contains 13,135 entries, \todo{13130 before!-changed} There are 168 ``chunker'' rules and 148 interchunk rules. 
+contains 13,135 entries. There are 168 ``chunker'' rules and 148 interchunk rules. 
 Evaluating was tested against revision 58998 in the Apertium Subversion repository. Lexical coverage is calculated 
@@
 As can be seen from Table~\ref{table:coverage}, the coverage of dictionaries is good, average score is 94\%. Our system outperforms the other 
-available RBMT system, but fall short of the state-of-the-art performance.\todo{represented by...} However, unlike the state of 
+available RBMT system, but fall short of the state-of-the-art performance represented by Google Translate. However, unlike the state of 
 the art, which uses unavailable corpora, our system is free and open and can be improved.

### 2015-03-07 2015-eamt-assim/2015-eamt-assim.tex
A new version
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59245 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \end{table*}
+\subsection{Experiments with other language pairs}
+In addition to the mentioned language pairs, some preliminary experiments were carried out for the English--Kazakh pair.
+It remains however questionable whether we may draw reliable conclusions from this part of the experiment.
+By design, the evaluation tasks are distributed evenly among the participants, in such a manner that each sentence is evaluated in all modes or gap densities by several people.
+The English--Kazakh evaluations have been set up for 11 participants, however, only 8 participants attempted their tasks, and 6 of them completed it in full.
+As a result, the evaluation became unbalanced: out of 4 sets of sentences, three have been evaluated by three, two and one participant, respectively, and the fourth set has been partially completed by only one participant.
+A total of 40 (about 10\%) gaps have not been filled by any of the participants.
+Therefore, we do not report the English--Kazakh evaluation results here.
 
@@
 The version of the toolkit used in this
-experiment may be downloaded from our repository.\footnote{\phantom{\url{https://github.com/Sereni/Appraise/tree/1e9d735faee64d1b97fb343ab111ace6a64509d7}}}
+experiment may be downloaded from our repository.\footnote{\blankout{\url{https://github.com/Sereni/Appraise/tree/1e9d735faee64d1b97fb343ab111ace6a64509d7}}}
 
-In addition to the mentioned language pairs, some preliminary experiments were carried out for the English--Kazakh pair.
-It remains questionable whether we may draw reliable conclusions from this part of the experiment.
-By design, the evaluation tasks are distributed evenly among the participants, in such a manner that each sentence is evaluated in all modes or gap densities by several people.
-The English--Kazakh evaluations have been set up for 11 participants, however, only 8 participants attempted their tasks, and 6 of them completed it in full.
-As a result, the evaluation became unbalanced: out of 4 sets of sentences, three have been evaluated by three, two and one participant, respectively, and the fourth set has been partially completed by only one participant.
-A total of 40 (about 10\%) gaps have not been filled by any of the participants.
-Therefore, we do not report the English--Kazakh evaluation results here.
-
-The experiment may easily be repeated for any language pair (provided a parallel corpus) and
+The experiments may easily be repeated for any language pair (provided a parallel corpus) and
 any machine translation system. Based on our experience, we would like to suggest the following amendments to the procedure:
@@
 
-%This work has been partly funded by the Spanish Ministerio de Econom{\'i}a y Competitividad through project TIN2012-32615 and by the Google Summer of Code 2014 programme.
-We would like to thank the volunteers who participated in the evaluations.
+\phantom{This work has been partly funded by the Spanish Ministerio de Econom{\'i}a y Competitividad through project TIN2012-32615 and by the Google Summer of Code 2014 programme.
+We would like to thank the volunteers who participated in the evaluations.}
 

### 2015-03-07 engkaz-eamt-2015/engkaz-eamt-2015.tex
corrected number of rules
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59247 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 The second round of structural transfer is currently performed by a %proof-of-concept 
-set of 90 rules, representative of following operations:\todo{I have removed \emph{proof-of-concept} here}
+set of 140 rules, representative of following operations:\todo{I have removed \emph{proof-of-concept} here}
 \begin{itemize}
@@
 Current system can translate simple sentences and questions. The English--Kazakh bilingual dictionary 
-contains 13,135 entries. There are 168 ``chunker'' rules and 148 interchunk rules. 
+contains 13,135 entries. There are 168 ``chunker'' rules and 140 interchunk rules. 
 Evaluating was tested against revision 58998 in the Apertium Subversion repository. Lexical coverage is calculated 

### 2015-03-07 engkaz-eamt-2015/engkaz-eamt-2015.tex
delete revision at 3.2
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59248 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-This section describes the current structural transfer in \texttt{apertium-eng-kaz} (revision 58998, ).\todo{Complete! 58998? Move to results?} English--Kazakh 
+This section describes the current structural transfer in \texttt{apertium-eng-kaz}.\todo{Complete! 58998? Move to results? - moved} English--Kazakh 
 chunker rules, interchunk rules and an additional clean-up stage will be described in detail in the following 3 sections.

### 2015-03-07 engkaz-eamt-2015/engkaz-eamt-2015.tex
Mikel's ironing is done
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59254 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{microtype}
+\newcommand{\anonymize}[1]{\phantom{#1}}
+%\newcommand{\anonymize}[1]{#1}
 
@@
 
-\author{Aida Sundetova\\
-  Kazakh National University,  \\
-  050040 Almaty, \\
-  {\tt sun27aida@gmail.com}\\
+\author{\anonymize{Aida Sundetova}\\
+  \anonymize{Kazakh National University, } \\
+  \anonymize{050040 Almaty, Kazakhstan} \\
+  \anonymize{{\tt sun27aida@gmail.com}}\\
   \And
-  \textbf{Francis M. Tyers}\\
-  HSL-fakultetet\\ 
-  UiT Norgga \'{a}rktala\v{s} universitehta \\
-  9017 Romsa, \\
-  {\tt francis.tyers@uit.no}
+  \anonymize{\textbf{Francis M. Tyers}}\\
+  \anonymize{HSL-fakultehta}\\ 
+  \anonymize{UiT Norgga \'{a}rktala\v{s} universitehta} \\
+  \anonymize{N-9017 Romsa, Norway} \\
+  \anonymize{{\tt francis.tyers@uit.no}}
   \And
-  Mikel L. Forcada\\
-  Dept. Lleng. i Sist. Inform. \\
-  Universitat d'Alacant, \\ 
-  Alacant \\
-  {\tt mlf@dlsi.ua.es}  
+  \anonymize{Mikel L.\ Forcada}\\
+  \anonymize{Dept.\ Lleng.\ i Sist.\ Informàt.} \\
+  \anonymize{Universitat d'Alacant,} \\ 
+  \anonymize{E-03071 Alacant, Spain} \\
+  \anonymize{{\tt mlf@dlsi.ua.es}}
 }
@@
 This paper presents a shallow-transfer rule-based machine translation (MT) system from English 
-(a West Germanic language) to Kazakh (a Turkic language)
+(a West Germanic language) to Kazakh (a Turkic language). This is a very challenging language pair for MT.
 The complex agglutinative morphology of Kazakh 
@@
 system we need large sentence-aligned parallel corpora, which are not freely available
-for the English--Kazakh pair. An English--Kazakh statistical MT system has however recently been built by Google,\footnote{\url{http://translate.google.com}}; 
+for the English--Kazakh pair. An English--Kazakh statistical MT system has however recently been built by Google;\footnote{\url{http://translate.google.com}}
 other work in this language pair was performed by building a small parallel corpus and using Moses \citep{assylbekov14}. Existing 
@@
 \caption{The pipeline architecture of the Apertium system.}
-\label{fig:modules}
+\label{figure:pipeline}
 \vspace{-1em}
@@
 Apertium \citep{forcada11} is a free/open-source rule-based machine translation (MT) platform that 
-has been developed since 2005, started at the Universitat d'Alacant. At first, it was initially aimed at 
+has been developed since 2005, started at the Universitat d'Alacant. Apertium Was initially aimed at 
 translating texts between closely related languages, then it was extended to deal with less-related 
@@
 data for an increasing number of language pairs and they are licensed under the 
-free/open-source GNU General Public License (GPL, versions 2 and 3).
-\todo{If space allowed, it would be nice to add a running example as in the Apertium paper}
+free/open-source GNU General Public License (GPL, versions 2 and 3). The Apertium pipeline (see figure~\ref{figure:pipeline}) has the following components:
+%\todo{If space allowed, it would be nice to add a running example as in the Apertium paper}
 \begin{description}\itemsep 0ex
@@
 %  forms are delimited by ``/'', and tags(<...>) show grammatical attributes of lexical form. 
-\item[Part-of-speech (POS) tagger:] This module chooses one of the LF of an ambiguous SF. 
+\item[Part-of-speech (POS) tagger:] This module chooses one of the LFs of an ambiguous SF. 
 %As can be 
@@
 %\end{verbatim}
- Multiword units can be translated as a single word.
+ Multiword units can be translated as a single LF.
 \item[ Lexical selection:] It uses rules to select one of the target-language LFs, as described in \cite{tyers12a}.
-\item [Structural transfer:] This module uses pattern matching to identify sequences of LF (phrases or 
-  segments), which need syntactical processing 
+\item [Structural transfer:] This module uses pattern matching to identify sequences of LFs (phrases or 
+  segments), which need syntactical processing.
 %to deal 
@@
 %  number, gender, etc.). 
-It uses files with rules, which specify  the syntactic transformation such as word reorderings, 
-  lexical changes such as changes in prepositions and agreement between target language lexical forms. %Transfer rules can 
+It uses files with rules, which specify syntactic transformations such as word reorderings, 
+  lexical changes such as changes in prepositions, and agreement between target language lexical forms. %Transfer rules can 
 %  produce new sequences for the  target language, for instance, preposition-noun rule is used to build a sequence, where for 
@@
 % \end{verbatim} 
-\item[Morphological generator:] Transforms the sequence of target-language LFs forms, produced by the structural transfer, 
+\item[Morphological generator:] It transforms the sequence of target-language LFs, produced by the structural transfer, 
   to a corresponding sequence of target-language SFs. 
 % The morphological generator executes a finite-state transducer
-For Kazakh, a finite-state transducer based on two-level rules \citep{washingtonsalimzyantyers14} is used for generation. 
-\item[Post-generator:] Performs some minor orthographical operations in the target language. 
+
+\item[Post-generator:] Performs some minor orthographical operations in the target language.\footnote{This module is not used when the target language is Kazakh.} 
 \item[ Reformatter:] It places format tags back into the text so that its format is preserved.
@@
 
-Apertium mainly uses XML-based formats for linguistic data, which include bilingual and monolingual dictionaries, structural transfer and
-lexical selection rules.
+Apertium uses text-based (mainly XML-based) formats for linguistic
+data, which include bilingual and monolingual dictionaries, structural
+transfer and lexical selection rules.
 
@@
 
-The English dictionary is used to determine lexical forms of each surface
-forms, it contains of list of all lexical units, definition of
-source-language alphabet, grammatical categories, which is used to
-specify lexical forms, for example, like noun, verb, gender, etc.;
-each category has paradigms describing groups of correspondences
+Morphological dictionaries are used to define the correspondences between LFs and SFs; they contain a definition of the source-language alphabet and grammatical categories and attributes (such as noun, verb, plural, locative, etc.), a list of all lexical units, and inflection paradigms for all lexical categories; paradigms describe regularities in the correspondences
 between parts of SFs and LFs.
 
-The Kazakh dictionary is turned into a
-morphological transducer, using the Helsinki Finite
-State Toolkit \citep{hfst/2011}. It uses the \texttt{lexc} formalism
-for defining lexicons through word classes and subclasses, and the
-\texttt{twol} formalism for morphophonological rules such as vowel
-harmony, desonorization, nasalization, etc.  Depending on the
-preceding vowels and consonants, the plural suffix \emph{-LAr}\footnote{Uppercase Latin letters are used for archiphonemes (actually archigraphemes) that are realised as phonemes (actually graphemes) after morphophonological rules have been applied} could
-become \emph{-тар}, \emph{-тер}, \emph{-лар}, \emph{-лер}, \emph{-дар}, \emph{-лер} depending on vowel harmony: кітап+тар,
-мектеп+тер, etc. 
-%\todo{Aida, please typeset this example in Cyrillic; braces are not used here, but are used later; DONE  homogeneize notation in paper.}
+The English dictionary comes from existing data in Apertium, such as the English---Spanish language pair.
+
+The Kazakh dictionary \citep{washingtonsalimzyantyers14} is turned
+into a finite-state morphological transducer, using the Helsinki
+Finite State Toolkit \citep{hfst/2011}. The dictionary is based on
+two-level rules, and uses the \texttt{lexc} formalism for defining
+lexicons through word classes and subclasses, and the \texttt{twol}
+formalism for morphophonological rules such as vowel harmony,
+desonorization, nasalization, etc. An example of morphophonological
+rule: depending on the preceding vowels and consonants, the plural
+suffix \emph{-LAr}\footnote{Uppercase Latin letters are used for
+  archiphonemes (actually archigraphemes) that are realised as
+  phonemes (actually graphemes) after morphophonological rules have
+  been applied.} for noun could become \emph{-тар}, \emph{-тер}, \emph{-лар},
+\emph{-лер}, \emph{-дар}, \emph{-лер}:
+кітап+тар, мектеп+тер, etc.
 
@@
 The English--Kazakh bilingual dictionary provides correspondences between English LFs and Kazakh LFs. Ambiguity is allowed: one of the LFs will be chosen by 
-lexical-selection rules depending on context. This dictionary currently contains 13,135 stems \citep{sundetova13a}.
+lexical-selection rules depending on context (see below). This dictionary currently contains 13,135 stems \citep{sundetova13a}.
 
@@
 
-This kind of rules solves the problem of part-of-speech (PoS) ambiguity using the  Constraint Grammar (CG, \cite{karlsson95}) formalism.
-CG rules choose one of the LFs obtained for each SF.
+The problem of part-of-speech (PoS) ambiguity is solved using the  Constraint Grammar (CG, \cite{karlsson95}) formalism: CG rules choose one of the LFs obtained for each SF.
 %From English to Kazakh POS ambiguity could be only in English side, but also some choice also
@@
 
-Lexical-selection rules choose one of the alternative target-language LFs corresponding to one source-language LF, as described in \cite{tyers12a}.
+Lexical-selection rules choose one of the alternative target-language LFs corresponding to one source-language LF, as described by \cite{tyers12a}.
 Alternative translations are defined in bilingual dictionary by
@@
 the following noun is a person: \emph{beautiful girl} \(\to\) \emph{сұлу қыз}; or as \emph{әдемі} or  \emph{көркем}, if the following noun means 
-place or \todo{or what? - mountains} \emph{beautiful mountains} \(\to\) \emph{әдемі таулар}.  Table~\ref{table:lexsel} shows some 
-lexical-selection rules:\todo{but this rule is not in the table!!! - should I add it or just remove upper example??}
+place: \emph{beautiful mountains} \(\to\) \emph{әдемі таулар}.  Table~\ref{table:lexsel} shows examples of 
+lexical-selection rules:
+
 \begin{table*}
@@
 \end{table*}
-\todo{Yes, it is mean I see whatever - I do not understand the \emph{I see anything} example. Does it mean \emph{anything} as \emph{whatever}? I do not know how to put the \(w\) in this example}@
 
@@
 \item A first round of transformations (``chunker'') detects source language (SL) LF patterns and generates the 
-  corresponding sequences of TL LFs grouped in chunks representing simple constituents such as noun phrases, prepositional phrases, etc. 
+  corresponding sequences of target language (TL) LFs grouped in chunks representing simple constituents such as noun phrases, prepositional phrases, etc. 
 \item The second round (``interchunk'') reads patterns of chunks and produces a new sequence of chunks. This is the 
@@
 The structural transfer module in Apertium processes the stream of source-language lexical form -- target-language lexical 
-form pairs (SL LF–TL LF pairs) and transforms it into a new sequence of TL LFs after a series of structural transfer 
+form pairs (SL LF–TL LF pairs) and transforms it into a  sequence of TL LFs after a series of structural transfer 
 operations specified in a set of rules: reordering, elimination or insertion of TL LFs, agreement, etc.
 
-This section describes the current structural transfer in \texttt{apertium-eng-kaz}.\todo{Complete! 58998? Move to results? - moved} English--Kazakh 
+This section describes the current structural transfer in \texttt{apertium-eng-kaz}.  English--Kazakh 
 chunker rules, interchunk rules and an additional clean-up stage will be described in detail in the following 3 sections.
 
-\paragraph{Chunker:}\todo{This should have 4 numerals, but I don't know how to do this}
+\paragraph{Chunker:}
 In the first round of structural transfer, rules segment sentences into chunks, like short noun 
@@
     \hline
-    \textbf{Phrase type} & \textbf{Description} & \textbf{Examples} & \textbf{Chunk tags} \\
+    \textbf{Phrase type} & \textbf{Description} & \textbf{Examples} & \textbf{Kazakh chunk tags} \\
     \hline
-    \multirow{5}{*}{\texttt{NP}} & \multirow{5}{*}{Noun phrase}    & noun & number*, person, \\
-                                 &                                 & determiner-noun & possessor*, case\\
+    \multirow{5}{*}{\texttt{NP}} & \multirow{5}{*}{Noun phrase}    & noun & number, person, \\
+                                 &                                 & determiner-noun & possessor, case\\
                                  &                                 & numeral-noun & \\
@@
     \multirow{5}{*}{\texttt{VP}} & \multirow{5}{*}{Verb phrase}    & finite-verb & number, person, \\
-                                 &                                 & have-participle & tense/mood, possessive*, \\
-                                 &                                 & be-gerund & negation* \\
+                                 &                                 & have-participle & tense/mood, possessive, \\
+                                 &                                 & be-gerund & negation \\
                                  &                                 & must-infinitive & \\ 
@@
     \multirow{5}{*}{\texttt{PP}} & \multirow{5}{*}{Postpositional phrase} & preposition-noun & number, person, \\
-                                 &                                        & preposition-determiner-noun & possessor*, case \\
+                                 &                                        & preposition-determiner-noun & possessor, case \\
                                  &                                        & preposition-numeral-noun & \\
@@
     \multirow{2}{*}{\texttt{GenP}} & \multirow{2}{*}{Genitive phrase}  & As with \texttt{PP} except & number, person, \\
-                                   &                                   & with the preposition `of'. & possessor*, case \\
+                                   &                                   & with the preposition `of'. & possessor, case \\
     \hline
-    \multirow{2}{*}{\texttt{AdjP}} & \multirow{2}{*}{Adjective phrase}  & adjective & n/a \\
+    \multirow{2}{*}{\texttt{AdjP}} & \multirow{2}{*}{Adjective phrase}  & adjective & (no tags) \\
                                    &                                    & ``more''--adjective & \\
     \hline
-    \multirow{2}{*}{\texttt{SupP}} & \multirow{2}{*}{Superlative phrase} & adjective--``est'' & possessor*, case \\
+    \multirow{2}{*}{\texttt{SupP}} & \multirow{2}{*}{Superlative phrase} & adjective--``est'' & possessor, case \\
                                    &                                     & the--``most''--adjective & \\ 
@@
   \end{tabular}
-  \caption{Phrase types in the English-Kazakh chunker}
+  \caption{Phrase types in the English--Kazakh chunker. }
   \label{table:phrases} 
@@
 \begin{description}
-\item[Noun phrases:] Consider the following example: the chunker identifies the English sequence 
-  the large book  (determiner–adjective–noun) as a noun-phrase chunk. It translates into 
-  Kazakh, and assigns it four chunk-level tags: number (set to singular), person (set to 3rd), 
+\item[Noun phrases:] Consider the following example: the chunker identifies the English LF sequence for \emph{the large book}  (determiner--adjective--noun) as a noun-phrase chunk.
+It translates it into the Kazakh adjective--noun sequence \emph{үлкен кітап}, and assigns it four chunk-level tags: number (set to singular), person (set to 3rd),
   possessor (to be determined, as the noun \emph{кітап} ('book') could receive a 3rd-person possesive 
-  ending (\emph{кітабы}) later if the context were, for instance, \emph{the large book of animals}, \emph{аңдардың үлкен кітабы}), 
-  and case (to be determined as it could be, for instance, accusative in \emph{I saw the large book}, \emph{Мен үлкен кітапты көрдім} ). Also, the
-  noun form of the verb, for instance,  gerund form is determined as NP phrase (\emph{Her playing is good} – \emph{Оның ойыны жақсы}).\todo{Make sure all English and Kazakh examples are in italics}
+  ending (\emph{кітабы}) later if the context were, for instance, \emph{the large book of animals}, \emph{аңдардың үлкен кіта\textbf{бы}}), 
+  and case, to be determined later as it could be, for instance, accusative in \emph{I saw the large book}, \emph{Мен үлкен кітап\textbf{ты} көрдім}).
+Similar processing occurs for nominal forms of the verb: for instance, as in the NP \emph{her paying} in \emph{Her playing is good}, which is  \emph{Оның ойыны жақсы} (\emph{She}+\textsc{genitive} \emph{playing}+\textsc{poss-3rd} \emph{good}).
 \item[Prepositional phrases:] English prepositional phrases are translated into Kazakh as postpositional phrases, 
@@
    \begin{itemize}
-    \item Genitive -{N}{I}ң,\footnote{Upper case letters in braces (such as {N}) represent hypothetical archiphonemes (actually archigraphemes) that are realised as phonemes (actually graphemes) after morphophonological rules have been applied.
-For instance, in the genitive ending “-{N}{I}ң”, the archiphoneme {N} may be realised as т, д, or н and the archiphoneme {I} may be realised as  і  or ы depending on the previous phonological context.
-} which will be marked GenP:
-        [PP [P of ] [NP the beautiful garden] ] $\rightarrow$ [GenP [NP әдемі бақша] [P -ның] ] 
-    \item Translating into any of the morphological cases: locative ``-{D}{A}'',\footnote{{D} can be д or т, and {A} can be е or а, depending on the 
-        phonological context.} ablative ``-{D}{A}н'', etc., except the genitive ``-{N}{I}ң'' case: 
+    \item Genitive \emph{-NIң},\footnote{In the genitive ending \emph{-{N}{I}ң}, the archiphoneme {N} may be realised as т, д, or н and the archiphoneme {I} may be realised as  і  or ы depending on the previous phonological context.
+} in which will case the phrase will be marked GenP:
+        [PP [P \emph{of} ] [NP \emph{the beautiful garden}] ] $\rightarrow$ [GenP [NP \emph{әдемі бақша}] [P \emph{-ның}] ] 
+    \item Translating into any of the morphological cases: locative \emph{-{D}{A}}'',\footnote{\emph{D} can be \emph{д} or \emph{т}, and \emph{{A}} can be \emph{е} or \emph{а}, depending on the 
+        phonological context.} ablative \emph{-{D}{A}н}, etc. --- all except the genitive \emph{-{N}{I}ң} case: 
         [PP [P \emph{from} ] [NP \emph{the five cars}] ] $\rightarrow$ [PP [NP \emph{бес көлік}] [P \emph{-тен} ] ]
@@
     \end{itemize}
-\item[Verb phrases:] Translation of English verb phrases into Kazakh  are not always be straightforward, for instance: 
-  % present simple and future are rendered using the same tense in Kazakh (I play $\rightarrow$ Мен ойнаймын;  I 
-  % will play $\rightarrow$ Мен ойнаймын), this ambiguity problem may be solved by lexical selection, 
-  % which will choose right tense by watching context; 
+\item[Verb phrases:] Translation of English verb phrases into Kazakh is not always be straightforward. For instance,
 tenses expressing continued activity, such as the 
@@
   mapped onto sets of two lexical units (\emph{Мен ойнап жатырмын}, \emph{Мен ойнап отырдым}) where the main verb is 
-  found in the -п participle form (\emph{ойнап}), and a suitable finite form  (\emph{жатырмын}, \emph{отырдым}) of an auxiliary 
-  verb (\emph{жатыр}, \emph{отыр}) is used to express number and person agreement,\footnote{Actually, Kazakh language uses four auxiliary verbs: \emph{жатыр} `lie', 
+  found in the \emph{-п} participle form (\emph{ойна\textbf{п}}), and a suitable finite form  (\emph{жатырмын}, \emph{отырдым}) of an auxiliary 
+  verb (\emph{жатыр}, \emph{отыр}) is used to express number and person agreement\footnote{Actually, Kazakh uses four auxiliary verbs: \emph{жатыр} `lie', 
       used when the activity takes a long time), отыр (`sit', used when the activity appears to be done in a sitting position), \emph{тұр} `stand', 
@@
       hard without a semantic analysis, which is not easily available in Apertium. Our current choice (an approximation) 
-      is \emph{жатыр} `lie' for the present continuous and отыр (`sit') for the past continuous.} 
-  modal verbs are translated by adding adjectives, which means \emph{керек} `necessary' or \emph{жөн} `proper', \todo{This part is hard to read. The choice of auxilliary is not justified.}
-  and a form of the copula (absent in present tense); the subject receives the genitive or dative case; negative 
-  verbs for present continuous and past tense are translated by adding negative word ``емес'' for present 
-  perfect continuous (He has not been playing -- Ол ойнап отырған емес) and  ``жоқ'' for past tense and 
-  present continuous (see examples above).
+      is \emph{жатыр} `lie' for the present continuous and отыр (`sit') for the past continuous.}.
+Modal verbs are translated by adding adjectives, which means \emph{керек} `necessary' or \emph{жөн} `proper',
+  and a form of the copula (absent in present tense), with the subject receiving the genitive or dative case. Negative 
+  verbs for present continuous and past tense are translated by adding the negative word \emph{емес} for present 
+  perfect continuous (\emph{He has not been playing} \(\to\) \emph{Ол ойнап отырған емес}) and  \emph{жоқ} for past tense and 
+  present continuous.
 \item[Adjectival phrases:] In Kazakh noun phrases, adjectives come before nouns and do not show any agreement 
@@
 \item The adjective alone, marked AdjP: [AdjP \emph{big} ] $\rightarrow$ [AdjP \emph{үлкен}] 
-\item Comparative adjective phrases  (English more + adjective, or adjective-[e]r); the Kazakh translation 
-   chooses the comparative suffix ``-{I}р{A}{K}'': [AdjP  \emph{more  interesting} ]  $\rightarrow$ [AdjP \emph{қызығ\emph{ырақ}}]
-\item For superlative adjective phrases  ``the most + adjective''  or ``adjective-[e]st'', translation 
+\item Comparative adjective phrases  (English \emph{more} + adjective, or adjective-[\emph{e}]\emph{r}); the Kazakh translation 
+   chooses the comparative suffix \emph{{I}р{A}{K}}: [AdjP  \emph{more  interesting} ]  $\rightarrow$ [AdjP \emph{қызығ\textbf{ырақ}}]
+\item For superlative adjective phrases  (\emph{the most} + adjective  or adjective-[\emph{e}]\emph{st}, translation 
    is built using \emph{ең} + adjective:  [SupP  the \emph{biggest} ]  $\rightarrow$ [SupP \emph{ең үлкен}] 
-   [SupP  \emph{the most important}]   $\rightarrow$ [SupP \emph{ең маңызды}]
+   [SupP  \emph{the most important}]   $\rightarrow$ [SupP \emph{ең маңызды}].
 \end{itemize}
 Superlative adjective phrases have some properties of noun phrases (such as receiving possessive morphemes when 
-modified by a genitive phrase: \emph{the most beautiful of people} $\rightarrow$ \emph{адамдардың ең әдемі\textbf{сі}}); they are treated as a noun phrase with an implied noun (``the largest [book]'').
-\todo{The table of phrase types is not referred to in the text; also, the meaning of ``*'' is not explained}
+modified by a genitive phrase: \emph{the most beautiful of people} $\rightarrow$ \emph{адамдардың ең әдемі\textbf{сі}}); they are treated as a noun phrase with an implied noun (\emph{the largest} [\emph{book}]).
 
@@
   \centering
+  \begin{small}
   \begin{tabular}{|l|l|l|l|}
@@
     \hline
-    \multirow{2}{*}{Present simple} & \multirow{2}{*}{I play} & \multirow{2}{*}{ойна + <aorist> + <pers.>} & Мен ойнаймын \\
-                                    &                         &                                            & I play \\
+    \multirow{2}{*}{Present simple} & \multirow{2}{*}{\emph{I play}} & \multirow{2}{*}{\emph{ойна} + <aorist> + <pers.>} & \emph{Мен ойнаймын} \\
+                                    &                         &                                            & 'I play' \\
     \hline
-    \multirow{2}{*}{Present continuous} & \multirow{2}{*}{I am playing} & ойна + <perf. part.>  & Мен ойнап жатырмын \\
-                                    &                         &   жатыр + <pres.> + <pers.>     & I playing lie \\
+    \multirow{2}{*}{Present continuous} & \multirow{2}{*}{\emph{I am playing}} & \emph{ойна} + <perf. part.>  & \emph{Мен ойнап жатырмын} \\
+                                    &                         &   \emph{жатыр} + <pres.> + <pers.>     & 'I playing lie' \\
     \hline
-    \multirow{2}{*}{Past continuous} & \multirow{2}{*}{I was playing} & ойна + <perf. part.>  & Мен ойнап отырдым \\
-                                    &                         & отыр + д{I} <past>  + <pers.>     & I playing sat \\
+    \multirow{2}{*}{Past continuous} & \multirow{2}{*}{\emph{I was playing}} & \emph{ойна} + <perf. part.>  & \emph{Мен ойнап отырдым} \\
+                                    &                         & \emph{отыр} +  <past>  + <pers.>     & 'I playing sat' \\
     \hline
-    \multirow{2}{*}{Obligation} & \multirow{2}{*}{I must go} & Мен  + <gen.>  & Менің баруым керек \\
-                                    &                         & бар + <ger.> +  <px1sg> + керек    & My going necessary [is] \\
+    \multirow{2}{*}{Obligation} & \multirow{2}{*}{I must go} & \emph{Мен}  + <gen.>  & \emph{Менің баруым керек} \\
+                                    &                         & \emph{бар} + <ger.> +  <poss.\ p1 sg.> + \emph{керек}    & 'My going necessary [is]' \\
     \hline
-    \multirow{2}{*}{Should} & \multirow{2}{*}{I should go} & Мен  + <gen.>  & Менің Менің барғаным жөн \\
-                                    &                         & бар + <past.ger.> +  <px1sg> + жөн & My was.going proper [is] \\
+    \multirow{2}{*}{\emph{Should}} & \multirow{2}{*}{\emph{I should go}} & Мен  + <gen.>  & \emph{Менің барғаным жөн} \\
+                                    &                         & \emph{бар} + <past.ger.> +  <poss.\ p1 sg.> + \emph{жөн} & 'My having gone proper is' \\
     \hline
@@
     \hline
-    \multirow{2}{*}{Present cont.} & \multirow{2}{*}{I am not playing} & ойна + <perf. part.>  & Мен ойнап жатқан жоқпын \\
-                                    &                         &   жатыр + <pres.> + <pers.>     & I playing lying not \\
+    \multirow{2}{*}{Present cont.} & \multirow{2}{*}{\emph{I am not playing}} & \emph{ойна} + <perf. part.>  & \emph{Мен ойнап жатқан жоқпын} \\
+                                    &                         &   \emph{жатыр} + <pres.> + <pers.>     & 'I playing lying not' \\
     \hline
-    \multirow{2}{*}{Ability} & \multirow{2}{*}{I can not play} & ойна + <impf. part.>  & Мен ойнай алмаймын \\
-                                    &                         &   ал + <neg.> + <aor.> + <p1>     & I playing can.not \\
+    \multirow{2}{*}{Inability} & \multirow{2}{*}{\emph{I can not play}} & \emph{ойна} + <impf. part.>  & \emph{Мен ойнай алмаймын} \\
+                                    &                         &   \emph{ал} + <neg.> + <aor.> + <p1>     & 'I playing cannot' \\
     \hline
-    \multirow{3}{*}{Present perf. cont. } & \multirow{3}{*}{You have not been playing} & ойна + <perf. part.>  & Сіз ойнап отырған емессі \\
-                                    &                         & +  отыр + <past>  & You playing sat not  \\
-                                    &                         & + eмес <cop.neg> + <p2> & \\
+    \multirow{3}{*}{Present perf. cont. } & \multirow{3}{*}{\emph{You have not been playing}} & \emph{ойна} + <perf. part.>  & \emph{Сіз ойнап отырған емессі} \\
+                                    &                         & +  \emph{отыр} + <past>  & 'You playing sat not'  \\
+                                    &                         & + \emph{eмес} <cop.neg> + <p2> & \\
     \hline
@@
   \end{tabular}
-  \caption{Examples of Verb phrase translations}
+  \end{small}
+  \caption{Examples of verb-phrase translations}
   \label{table:vps}
@@
 \paragraph{Interchunk processing:}
-\todo{Table~\ref{table:vps} should be typeset with examples in italics and making sure that the margins are respected}
 
 The second round of structural transfer is currently performed by a %proof-of-concept 
-set of 140 rules, representative of following operations:\todo{I have removed \emph{proof-of-concept} here}
+set of 140 rules, representative of following operations:
 \begin{itemize}
-\item Inter-chunk agreement between subject noun phrase and verb phrase in their person and number.
+\item Persona and number agreement between the subject noun phrase and the verb phrase.
 \item Assigning case to noun phrases (which are generated without case by the chunker): for instance, 
-  accusative case for objects (\emph{I bought the table} $\rightarrow$ \emph{Мен үстелді  сатып алдым}), genitive case for 
-  obligatory constructs (\emph{I must see} $\rightarrow$ \emph{менің көруім керек}),  dative case for the verb \emph{to 
-  need} (\emph{I need a doctor} $\rightarrow$ \emph{Маған дәрігер керек}), locative case for 
-  possession (\emph{I have a book} $\rightarrow$ \emph{Менде кітап бар}), etc.\ 
-\item Reordering: placing of object before verb ( I[1] bought[2] the table[3] $\rightarrow$ Мен[1] үстелді[3] сатып алдым[2]), \todo{italics for examples please!}
-  placing of prepositional pharses before the verb  (They[1] played[2] on top of the tree [3] $\rightarrow$ Олар[1] ағаштың үстінде[3] ойнады [2]), etc.\ 
-\item Adding question word ма/ме/ба/бе in the end of question, if the question mark chunk ``?'' is detected (Did[VP\_ques] you[NP] watch[VP] last film[NP] ?[Q\_m] $\rightarrow$ Сіз[NP] соңғы фильмді[NP] көрдіңіз[VP] бе<ques> ?[Q\_m]).\todo{Notation: use brackets as above!}
-\item Place possessive for long noun + noun + noun structures(The university of city of Kazakhstan -- Қазақстан қаласының университеті).\todo{Explain better, as this is a rule to avoid repeating -NIng, -NIng, etc.
-All possessives are in place}
+  accusative case for objects (\emph{I bought the table} $\rightarrow$ \emph{Мен үстел\textbf{ді}  сатып алдым}), genitive case for 
+  obligatory constructs (\emph{I must see} $\rightarrow$ \emph{ме\textbf{нің} көруім керек}),  dative case for the verb \emph{to 
+  need} (\emph{I need a doctor} $\rightarrow$ \emph{Ма\textbf{ған} дәрігер керек}), locative case for 
+  possession (\emph{I have a book} $\rightarrow$ \emph{Мен\textbf{де} кітап бар}), etc.\ 
+ \item Reordering: placing of object before verb ( \emph{I}$_1$ \emph{bought}$_2$ \emph{the table}$_3$ $\rightarrow$ \emph{Мен}$_1$ \emph{үстелді}$_2$ \emph{сатып алдым}$_2$), 
+  placing of prepositional phrases before the verb  (\emph{They}$_1$ \emph{played}$_2$ \emph{on top of the tree}$_3$ $\rightarrow$ \emph{Олар}$_1$ \emph{ағаштың үстінде}$_3$ \emph{ойнады}$_2$), etc.\ 
+\item Adding the question word \emph{ма}/\emph{ме}/\emph{ба}/\emph{бе} at the end of interrogative sentences: if the question mark chunk ``?'' is detected (\emph{Did} \emph{you} \emph{watch} \emph{the latest film?} $\rightarrow$ \emph{Сіз соңғы фильмді көрдіңіз бе?}).
+\item Placing possessive marks but not genitive endings in long noun + noun + noun structures(\emph{The university of city of Kazakhstan} \(\to\) \emph{Қазақстан}-\(\emptyset\) \emph{қаласының университеті} instead of \emph{Қазақстан\textbf{ның} қаласының университеті}).
 \end{itemize}
@@
 
-\paragraph{Postchunk and cleanup:}\todo{Can I do that?  --Nothing is said about postchunk --- this talks about the additional cleaning stage}
-An additional cleaning stage was created for removing extra tags which were ot determined during the chunker and 
-interchunk steps, for instance, a noun was not determined to be acting as an object and therefore did not get accusative case; such a noun would have been assigned the  \texttt{<CD>} (``case to be determined'') tag, which after ``cleaning'' would be set to nominative.
+\paragraph{Postchunk:} English--Kazakh postchunk rules straightforwardly transfer chunk-level tags to the head word (for instance, if the noun phrase is in locative, the locative case is transferred to the head noun).
+
+\paragraph{Cleanup:} An additional phase was added to  English--Kazakh transfer to be able to remove or give default values to tags which were ot determined during the chunker and 
+interchunk steps.
+For instance, if a noun phrase was not determined to be acting as an object and therefore the head noun did not get accusative case, such a noun would be received with a  \texttt{<CD>} (``case to be determined'') tag, which after the cleanup would be set to nominative.
 Other operations carried out at this stage ensure
-the agreement between noun or adjective and the copula verb ``е''(\emph{Мен дәрігер+е}<p1 person singular>), \todo{add result for copula} or deciding the actual form of the question particle according to the preceding word (\emph{Сіз келдіңіз} + \emph{ме}?
+the agreement between noun or adjective and the copula verb \emph{е} (\emph{Мен дәрігер+е}<p1 person singular> \(\to\) \emph{Мен дәрігермен}), or deciding the actual form of the question particle according to the preceding word (\emph{Сіз келдіңіз} + \emph{ме}?
 \(\to\) \emph{Сіз келдіңіз} + \emph{бе}?).
@@
 
-Current system can translate simple sentences and questions. The English--Kazakh bilingual dictionary 
+The current system can translate simple sentences and questions. The English--Kazakh bilingual dictionary 
 contains 13,135 entries. There are 168 ``chunker'' rules and 140 interchunk rules. 
-Evaluating was tested against revision 58998 in the Apertium Subversion repository. Lexical coverage is calculated 
+Evaluation was performed on revision 58998 in the Apertium Subversion repository. Lexical coverage is calculated 
 over EuroParl,\footnote{\url{http://www.statmt.org/europarl/v7/es-en.tgz}} SETimes,\footnote{\url{http://nlp.ffzg.hr/data/corpora/setimes/setimes.en-tr.txt.tgz}} NewsCommentary,\footnote{\url{http://www.statmt.org/wmt13/training-parallel-nc-v8.tgz}} Wikipedia.\footnote{\url{http://dumps.wikimedia.org/enwiki/20140402/enwiki-20140402-pages-articles.xml.bz2}; last checked 04/06/2014}
@@
 As can be seen from Table~\ref{table:coverage}, the coverage of dictionaries is good, average score is 94\%. Our system outperforms the other 
-available RBMT system, but fall short of the state-of-the-art performance represented by Google Translate. However, unlike the state of 
-the art, which uses unavailable corpora, our system is free and open and can be improved.
+available RBMT system, but falls short of the state-of-the-art performance represented by Google Translate. However, unlike the state of 
+the art, which uses corpora which are unavailable to the general public, our system and the linguistic resources used in it are free and open/source and are open to improvement.
 
-MT systems have some mistakes in translations, which was shown with *.\todo{DONE you used *!} Google MT system 
-has a higher BLEU score, but in translation separate phrases, it has common error: it does not assign 
-the right possessive and case. Sanasoft system has more errors with choosing not right translations 
-and out of vocabulary words.
+MT systems have some mistakes in translations, which was shown with
+*. The Google MT system gets a higher BLEU score, but in translation
+of selected phrases, it makes common mistakes such as not assigning
+the right possessive and case. The Sanasoft system has more errors
+as regards the translation of words, and many out-of-vocabulary words.
 
@@
     \hline 
-    \multirow{6}{*}{Noun phrases} & \multirow{3}{*}{My difficult exercises} & Apertium & Менің қиын жаттығуларым \\
-                                  &                                         & Sanasoft & Менің қиын жаттықтырып *жатыр \\ % <not right POS>
-                                  &                                         & Google   & Менің қиын *жаттығулар \\ % <no 1 person possessive>
-                                  & \multirow{3}{*}{Conan Doyle}            & Apertium & *Дойлдың Конан \\
-                                  &                                         & Sanasoft & *Conan *Doyle \\ % <not right POS>
-                                  &                                         & Google   & Конан Дойл \\ % <no 1 person possessive>
+    \multirow{6}{*}{Noun phrases} & \multirow{3}{*}{\emph{My difficult exercises}} & Apertium & \emph{Менің қиын жаттығуларым} \\
+                                  &                                         & Sanasoft & \emph{Менің қиын жаттықтырып *жатыр} \\ % <not right POS>
+                                  &                                         & Google   & \emph{Менің қиын *жаттығулар} \\ % <no 1 person possessive>
+                                  & \multirow{3}{*}{\emph{Conan Doyle}}            & Apertium & \emph{*Дойлдың Конан} \\
+                                  &                                         & Sanasoft & \emph{*Conan *Doyle} \\ % <not right POS>
+                                  &                                         & Google   & \emph{Конан Дойл} \\ % <no 1 person possessive>
     \hline 
-    \multirow{3}{*}{Adpositional phrases}  & \multirow{3}{*}{Under three big trees}  & Apertium & үш үлкен ағаштың астында \\
-                                           &                                         & Sanasoft & *Three үлкен *ағаштар астында \\ % <unk> <no genitive>
-                                           &                                         & Google   & үш үлкен *ағаштарды астында \\ % <no genitive>
+    \multirow{3}{*}{Adpositional phrases}  & \multirow{3}{*}{\emph{Under three big trees}}  & Apertium & \emph{үш үлкен ағаштың астында} \\
+                                           &                                         & Sanasoft & \emph{*Three үлкен *ағаштар астында} \\ % <unk> <no genitive>
+                                           &                                         & Google   & \emph{үш үлкен *ағаштарды астында} \\ % <no genitive>
     \hline 
-    \multirow{3}{*}{Adjective phrases}    & \multirow{3}{*}{The most beautiful}      & Apertium & ең әдемі \\ 
-                                          &                                          & Sanasoft & *Көпшілік әдемі \\ % <not right translation> 
-                                          &                                          & Google   & ең әдемі \\
+    \multirow{3}{*}{Adjective phrases}    & \multirow{3}{*}{\emph{The most beautiful}}      & Apertium & \emph{ең әдемі} \\ 
+                                          &                                          & Sanasoft & *\emph{Көпшілік әдемі} \\ % <not right translation> 
+                                          &                                          & Google   & \emph{ең әдемі} \\
     \hline
-    \multirow{6}{*}{Modal verbs}  & \multirow{3}{*}{I must pay}             & Apertium & Мен жүргізе аламын \\
-                                  &                                         & Sanasoft & Мен must pay \\ % <no translation>
-                                  &                                         & Google   & *Мен *төлеуі тиіс \\ % <no genitive> <no possessive>
-                                  & \multirow{3}{*}{It must be James}       & Apertium & Бұл Джеймс *болады \\ % <should be "болуы керек"> 
-                                  &                                         & Sanasoft & *Оған Джеймс *must *болып *жатыр  \\ % <no translation>
-                                  &                                         & Google   & Джеймс болуы тиіс \\ % <no genitive> <no possessive>
+    \multirow{6}{*}{Modal verbs}  & \multirow{3}{*}{\emph{I must pay}}             & Apertium & \emph{Мен жүргізе аламын} \\
+                                  &                                         & Sanasoft & \emph{Мен *must *pay} \\ % <no translation>
+                                  &                                         & Google   & \emph{*Мен *төлеуі тиіс} \\ % <no genitive> <no possessive>
+                                  & \multirow{3}{*}{\emph{It must be James}}       & Apertium & \emph{Бұл Джеймс *болады} \\ % <should be "болуы керек"> 
+                                  &                                         & Sanasoft & \emph{*Оған Джеймс *must *болып *жатыр}  \\ % <no translation>
+                                  &                                         & Google   & \emph{Джеймс болуы тиіс} \\ % <no genitive> <no possessive>
     \hline 
-    \multirow{3}{*}{Questions}    & \multirow{3}{*}{Is it right?}           & Apertium & *Екен *дұрыстың ол? \\ % extra word + extra case 
-                                  &                                         & Sanasoft & Бұл *түзуi? \\ % <not right translation>
-                                  &                                         & Google   & Бұл дұрыс па? \\
+    \multirow{3}{*}{Questions}    & \multirow{3}{*}{\emph{Is it right?}}           & Apertium & \emph{*Екен *дұрыстың ол?} \\ % extra word + extra case 
+                                  &                                         & Sanasoft & \emph{Бұл *түзуi?} \\ % <not right translation>
+                                  &                                         & Google   & \emph{Бұл дұрыс па?} \\
     \hline 
@@
 constraint grammar rules, structural transfer and lexical rules. The future plan is to use the created data with other 
-open-source MT system with Turkic languages or system with Kazakh language on right side to make transfer systems 
+free/open-source MT systems involving Turkic languages or in systems having Kazakh as target language to make transfer systems 
 between the Turkic or other language pairs. Related work is currently ongoing with Russian--Kazakh and Kazakh--English \citep{sundetova14}.
@@
 Tukeyev for his valuable input.
-Special thanks to Inari Listenmaa for her assistance in mentoring the Google Summer of Code project.\todo{Some references didn't come out right.}
+Special thanks go to Inari Listenmaa for her assistance in mentoring the Google Summer of Code project and to Aidana Karibaeva for help in writing transfer rules.
+
+\todo{Some references didn't come out right.} 
 \bibliographystyle{apalike}

### 2015-03-07 engkaz-eamt-2015/engkaz-eamt-2015.tex
remove comments
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59256 72bbbca6-d526-0410-a7d9-f06f51895060
caught a minor mistake
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59257 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 translation (MT) system from English to Kazakh. The main syntactic and morphological differences 
-between the two languages are presented, and it is showed how the MT 
+between the two languages are presented, and it is shown how the MT 
 system was designed to tackle these challenges. We show an evaluation of system coverage and 
@@
 \item[Prepositional phrases:] English prepositional phrases are translated into Kazakh as postpositional phrases, 
-  there are three possible outcomes with different cases:\todo{remove braces for consistency with the previous notation for archiphonemes: avoid repeating the explanation in the footnote.}
+  there are three possible outcomes with different cases:
    \begin{itemize}
@@
 
-\todo{Some references didn't come out right.} 
+%\todo{Some references didn't come out right.} 
 \bibliographystyle{apalike}

### 2015-03-19 rusaccent-nodalida-2015/rusaccent.tex
Adding lots of comments for revisions from reviewers and CLEAR. Corrected some typos in bibtex.
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59426 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \newcommand{\rr}[1]{\marginpar{\scriptsize R: #1}} % Rob's comments
-\newcommand{\cw}[1]{\marginpar{\scriptsize W: #1}} % Cute wife's comments
 \newcommand{\dm}[1]{\marginpar{\scriptsize D: #1}} % Detmar's comments
@@
 
-\author{Author1 \\
-  Affiliation / Address line 1 \\
-  Affiliation / Address line 2 \\
-  Affiliation / Address line 3 \\
-  {\tt email@domain.com} \\\And
-  Author2 \\
-  Affiliation / Address line 1 \\
-  Affiliation / Address line 2 \\
-  Affiliation / Address line 3 \\
-  {\tt email@domain.com} \\}
+\author{Robert Reynolds\\
+  HSL Faculty\\
+  UiT The Arctic University of Norway\\
+  N-9018 Troms\o\\
+  {\tt robert.reynolds@uit.no} \\\And
+  Francis Tyers \\
+  HSL Faculty\\
+  UiT The Arctic University of Norway\\
+  N-9018 Troms\o\\
+  {\tt francis.tyers@uit.no} \\}
 
@@
   We evaluate the effectiveness of finite-state tools we developed for 
-  automatically annotating word stress in Russian unrestricted text. To our 
+  automatically annotating word stress in Russian unrestricted text. This task is
+  relevant for computer-assisted language learning and text-to-speech. To our 
   knowledge, this is the first study to empirically evaluate the results of this 
   task. Given an adequate lexicon with specified stress, the primary obstacle for 
-  correct stress placement is disambiguating homographic wordforms. Using a 
-  Constraint Grammar, we achieve 93.21\% accuracy with minimal errors, a 3.14\% 
-  improvement over the baseline undisambiguated performance with minimal errors. 
-  For applications with a higher threshold for errors, we combined our Constraint 
-  Grammar with frequency-based guessing and a simple
-  algorithm for unknown word stress position guessing, thereby achieving 96.15\%
-  accuracy. These results highlight the need for morphosyntactic disambiguation
-  in the wordstress placement task for Russian, and set a standard for future
-  research on this task.
+  correct stress placement is disambiguating homographic wordforms. The baseline
+  performance of this task is 90.07\%, (known words only, no morphosyntactic 
+  disambiguation). Using a Constraint Grammar to disambiguate homographs, we 
+  achieve 93.21\% accuracy with minimal errors. For applications with a higher 
+  threshold for errors, we achieved 96.15\% accuracy by incorporating frequency-
+  based guessing and a simple algorithm for guessing the stress position on 
+  unknown words. These results highlight the need for morphosyntactic 
+  disambiguation in the wordstress placement task for Russian, and set a standard 
+  for future research on this task.
 \end{abstract}
+\rr{language learning?}
 
-%\section{Credits}
-%
-%The template had a credits section for the history of the template, but we don't need
-%a Credits section, right?
-
+\rr{REV3: You should use "gold-standard" consistently throughout (sometimes the authors
+write "gold" to mean "gold-standard", which sounds strange.)}
+\rr{Anya: wordstress or word stress?}
 \section{Introduction}
-
-Russian word stress is variable and phonemic. The inflecting word classes include complex 
-patterns of shifting stress, and a lexeme's stress pattern cannot be predicted 
-from surface forms. Lexical stress and its attendant vowel reduction are 
-a prominent feature of spoken Russian; the incorrect placement of stress
-can render speech almost incomprehensible. However, standard written Russian does not 
+Lexical stress and its attendant vowel reduction are a prominent feature of 
+spoken Russian; the incorrect placement of stress can render speech almost 
+incomprehensible. This is because Russian word stress is phonemic, i.e. many 
+wordforms are distinguished from one another only by stress position. 
+This is the cause of considerable difficulty for learners, since 
+the inflecting word classes include 
+complex patterns of shifting stress, and a lexeme's stress pattern cannot be 
+predicted from surface forms. Furthermore, standard written Russian does not 
 typically mark word stress.\footnote{Texts intended for native speakers sometimes 
@@
 correctly converting written Russian text to speech is impossible. This is true 
-both for humans (e.g. foreign language students) and computers (e.g. text-to-
-speech).
+both for humans (e.g. foreign language students) and computers (e.g. 
+text-to-speech).\rr{add something about how TTS systems handle this right now. Just no one evaluates how well they are doing.}
 
@@
 \ex. Morphosyntactically incongruent homographs \label{ex:MSincongruent}
-\a. \rus{попад\'{а}ть} \emph{popadát'} `fall.\textsubscript{INF-IPFV}' \\
-    \rus{поп\'{а}дать} \emph{popádat'} `fall.\textsubscript{INF-PFV}'
-\b. \rus{узнаёт} \emph{uznaët} `finds out.\textsubscript{PRS-IPFV-3SG}' \\
-    \rus{узн\'{а}ет} \emph{uznáet} `finds out.\textsubscript{FUT-PFV-3SG}'
-\c. \rus{н\'{а}шей} \emph{nášej} `our.\textsubscript{F-SG-GEN/DAT/LOC...}'\\
+\a. \rus{н\'{а}шей} \emph{nášej} `our.\textsubscript{F-SG-GEN/DAT/LOC...}'\\
     \rus{наш\'{е}й} \emph{našéj} `sew on.\textsubscript{IMP-2SG}'
-\d. \rus{дор\'{о}га} \emph{doróga} `road.\textsubscript{N-F-SG-NOM}'\\
+\b. \rus{дор\'{о}га} \emph{doróga} `road.\textsubscript{N-F-SG-NOM}'\\
     \rus{дорог\'{а}} \emph{dorogá} `dear.\textsubscript{ADJ-F-SG-PRED}'
@@
 and whose morphosyntactic values are identical, as shown in
-Table~\ref{table:MScongruent}. This kind of stress ambiguity is relatively rare, and resolving this ambiguity would require the use of technologies such as Word Sense Disambiguation.
+\ref{ex:MScongruent}. This kind of stress ambiguity is relatively rare, and resolving this ambiguity would require the use of technologies such as Word Sense Disambiguation.
+%\rr{cite Tore's book on stress?}
 
-\begin{table}
-\begin{center}
-\begin{tabular}{c|ll}
-\bf Number-Case & \bf `castle' & \bf `lock' \\
-\hline
-\texttt{SG-NOM} & \rus{з\'{a}мок} & \rus{зам\'{о}к} \\
-\texttt{SG-GEN} & \rus{з\'{a}мка} & \rus{замк\'{а}} \\
-\texttt{PL-NOM} & \rus{з\'{a}мки} & \rus{замк\'{и}} \\
-\texttt{PL-GEN} & \rus{з\'{a}мков} & \rus{замк\'{о}в} \\
-... & \hskip 1em ... & \hskip 1em ... \\
-\end{tabular}
-\end{center}
-\caption{ \label{table:MScongruent} Morphosyntactically congruent homographs. }
-\end{table}
+\ex. Morphosyntactically congruent homographs \label{ex:MScongruent}
+\a. \rus{з\'{a}мок} \emph{z\'{a}mok} `castle.\textsubscript{SG-NOM}'\\
+	\rus{зам\'{о}к} \emph{zam\'{o}k} `lock.\textsubscript{SG-NOM}'
+\b. \rus{з\'{a}мков} \emph{z\'{a}mkov} `castle.\textsubscript{PL-GEN}'\\
+	\rus{замк\'{о}в} \emph{zamk\'{o}v} `lock.\textsubscript{PL-GEN}'
+\c. ...\\
+	...
+%\begin{table}
+%\begin{center}
+%\begin{tabular}{c|ll}
+%\bf Number-Case & \bf `castle' & \bf `lock' \\
+%\hline
+%\texttt{SG-NOM} & \rus{з\'{a}мок} & \rus{зам\'{о}к} \\
+%\texttt{SG-GEN} & \rus{з\'{a}мка} & \rus{замк\'{а}} \\
+%\texttt{PL-NOM} & \rus{з\'{a}мки} & \rus{замк\'{и}} \\
+%\texttt{PL-GEN} & \rus{з\'{a}мков} & \rus{замк\'{о}в} \\
+%... & \hskip 1em ... & \hskip 1em ... \\
+%\end{tabular}
+%\end{center}
+%\caption{ \label{table:MScongruent} Morphosyntactically congruent homographs. }
+%\end{table}
+
+%\rr{Yurij Zelenkov automatic stress Yandex, published?}
+% http://www.dialog-21.ru/digests/dialog2014/materials/pdf/ZelenkovYuG.pdf
+% Written by Zelenkov et al. about Segalovich's works (incl. stress placement in poetry)
 
 It should be noted that not all morphosyntactic ambiguity leads to stress 
-ambiguity. For example, all inflecting modifiers have the identical form for 
-F-SG-GEN, F-SG-LOC, F-SG-DAT and F-SG-INS. We refer to this as 
-\emph{stress-irrelevant} ambiguity, since all readings have
-the same stress placement. Stress-irrelevant ambiguity can be intraparadigmatic 
-and/or interlexical.
+ambiguity. For example, \emph{novyj} `new' (and every other adjective) has 
+identical forms for F-SG-GEN, F-SG-LOC, F-SG-DAT and F-SG-INS: \emph{n\'{o}voj}. 
+Likewise, the form \emph{vypej} has multiple possible readings 
+(including `drink.IMP', `bittern.PL-GEN'), but they all have the same stress 
+position: \emph{v\'{y}pej}. We refer to this as \emph{stress-irrelevant} 
+morphosyntactic ambiguity, since all readings have the same stress placement.
 
@@
 ambiguity to future work because it is by far the least common type of stress
-%\rr{Exactly how common is congruent stress ambiguity?}
+\rr{Exactly how common is congruent stress ambiguity?
+REVIEW 2: Since this type of ambiguity cannot be resolved with your approach, it imposes an upper bound.
+So, it would be very interesting to see how many examples of morphosyntactically congruent stress ambiguity you found in the annotated corpus.}
 ambiguity, and disambiguating morphosyntactically congruent stress requires 
-a fundamentally different technology from the other approaches of this study
-(i.e. Word Sense Disambiguation).
+fundamentally different technology from the other approaches of this study.
 
 \subsection{Background and task definition} \label{background}
-
-Automatic stress placement in Russian is similar to Diacritic Restoration, a task
-which has received increasing interest over the last 20 years. Generally 
+\rr{cite diacritic restoration}
+Automatic stress placement in Russian is similar to diacritic restoration, a task
+which has received increasing interest over the last 20 years. \rr{cite?}Generally 
 speaking, diacritics disambiguate otherwise homographic wordforms, so missing 
@@
 must first be normalized by restoring diacritics.
-
+\rr{Sasha has citations from a lab in Belorus. Lobanov.}
 A slightly different situation arises with languages whose standard orthography 
@@
 Russian orthography is similar to Arabic and Hebrew, since its vowel qualities 
-are underspecified in standard orthography.
+are underspecified in standard orthography. \rr{spell out why stress placement is important for pronunciation.}
 
-Many studies of Russian text-to-speech and automatic speech recognition make note of the 
-difficulties caused by the shortcomings of their stress-marking resources (e.g. 
-\newcite{krivnova_1998}), but very few studies have targeted stress-marking 
-itself, and they focus on placing stress on unknown words, with almost no 
-attention to the problems that arise with known words. For example, 
+Many studies of Russian text-to-speech and automatic speech recognition make note 
+of the difficulties caused by the shortcomings of their stress-marking resources 
+(e.g. \newcite{krivnova_1998}), but very few studies have targeted stress-marking 
+itself, and those that do focus on placing stress on unknown words, with almost 
+no attention to the problems that arise with running text. For example, 
 \newcite{Xomicevic_2008} developed a set of heuristics for guessing stress 
@@
 
+\rr{somehow make this point clearer. Machine learning is not possible for stress in running text because large corpora do not exist.}
 One major shortcoming of evaluating on wordlists is that they do not reflect 
@@
 possibility of disambiguating homographs based on syntactic context.
+\rr{REVIEWER 3: Why are results "inflated"?
+if the system does better on more rare forms, then results are "deflated".  if
+it does better on more frequent forms, they are "inflated".   This seems to be
+just a different kind of evaluation.
+It seems that the point of evaluating on running text ought to be precisely
+that the same text could have an ambiguous word with more than one stress
+pattern. So this is really evaluating a different task.}
 
@@
 \end{table*}
-
 \subsection{Stress corpus}
+\rr{imperative kosti is weird}
+\rr{Table 2 needs more explanation: what exactly these two columns mean and how you ended up with these two sets of readings}
 
@@
 free texts on Russian language-learning websites.\footnote{URLs can be found in
-the X data repository (URL will be added in the final version).} This small corpus (N=7689) is 
-representative of texts 
+the X data repository (URL will be added in the final version).} This small
+corpus (7689 tokens) is representative of texts 
 that learners of Russian are likely to encounter in their studies. These texts 
 include excerpts from well-known literary works, as well as dialogs and 
-prose and individual sentences that were written for learners.
+prose and individual sentences that were written for learners. \rr{1st and 2nd language learners? Sometimes children's textbooks have them, especially when syllables are separated.}
+\rr{Why not RNC corpus for gold corpus?}
 
@@
 \emph{not} mark stress on monosyllabic tokens, effectively assuming that
-all monosyllabics are stressed. However, this approach 
+all monosyllabics are stressed.\rr{citation?} However, this approach 
 is not well-motivated. Many words -- both monosyllabic and multisyllabic -- are 
 unstressed, especially among prepositions, conjunctions, and particles.
-% multisyllabics that can be unstressed: \rus{через, между} 
+%multisyllabics that can be unstressed: \rus{через, между} 
 Furthermore, there are many
-%\rr{examples}
+\rr{examples}
 high-frequency monosyllabic homographs that can be either stressed or unstressed, 
-depending on their part of speech, or particular collocations. Clearly, for
+depending on their part of speech, or particular collocations.\rr{examples} Clearly, for
 such words, one cannot assume that they are stressed on the basis of their 
 syllable count.
+\rr{monos are determining WHETHER stress is present. slightly different.}
 
@@
 In order to test our wordstress placement system, we removed all stress marks
-from the gold corpus, then marked stress on the plain text version, and then
+from the gold corpus, then marked stress on the unstressed version using our tools, and then
 compared with the original.
@@
 
+\rr{rename guess > guessRand}
+\rr{more about our corpus}
+\rr{explain guess better}
 The {\small {\tt bare}} approach is to not mark stress on words with more than 
@@
 stress position in Russian is the last syllable of the stem.} Based on this 
+\rr{what is the difference between last syllable and last vowel followed by consonant?}
 conclusion, our syllable-guessing method places the stress on the last vowel that 
-is followed by a consonant.\footnote{More sophisticated heuristics and machine-
-learning methods have been suggested, but unknown words are not the central focus 
-of this study, and so a simple method was judged to be sufficient.} This method 
+is followed by a consonant.\footnote{Although this approach is simplistic, 
+unknown words are not the central focus 
+of this study. More sophisticated heuristics and machine-
+learning approaches to unknown words are discussed in Section~\ref{discussion}.} This method 
 is applied in two approaches, {\small {\tt guessSyll}} 
@@
 {\small {\tt guess}} and {\small {\tt guessFreq}}, respectively.
+\rr{REVIEWER 2: However, one can imagine a token that gets some morphological readings, none of them being correct (for example, when the correct lexeme is not covered by the dictionary -- but an homonymous incorrect lexeme is).
+Do you have any estimations of how often this happens in the running text?}
 
@@
 
+\rr{REV3: The presentation of results is unclear: what exactly is  being measured?
+It would be best if the measure was presented on words that have a potential
+stress ambiguity.  (If this includes all words, it should be made clear, or
+more clear.)}
 Single-syllable words and words that are unstressed in the gold standard were
@@
 This number represents the ceiling for methods relying on the FST.
-Higher scores are only achievable by expanding the FST or by using 
+Higher scores are only achievable by expanding the FST \rr{not in dictionary}or by using 
 syllable-guessing algorithms. After running the Constraint Grammar, recall was 
@@
 removes some correct readings, or else there are errors in the gold corpus.
+\rr{sveta wanted more}
 
@@
 
+\rr{flow for Syll methods}
 The two right-most columns in 
@@
 (e.g. \rus{вер\'{и}м} \emph{ver\'{i}m}).
+\rr{don't count variation as errors, if it is acceptable}
+\rr{why errors?}
 
@@
 are the result of the fact that some words in the FST, especially proper nouns, 
-had not been assigned stress. Because the FST outputs a form -- albeit unstressed 
+had not been assigned stress. \rr{REV3: how often does FST block guessFreq?}Because the FST outputs a form -- albeit unstressed 
 -- the {\small {\tt Syll}} algorithm is not called. This means that 
@@
 {\small {\tt Syll}} guess was accurate 21\% of the time.
-%\rr{Examine errors and categorize/refute them.}
+\rr{Examine errors and categorize/refute them.}
 
@@
 of the 2~804~492 wordforms contained in our FST are stress-ambiguous.
-%\rr{These stress-ambiguous wordforms have an average of 2.01 possible stress positions, so a perfect stress placement tool has a 50/50 chance of getting them right.
+\rr{These stress-ambiguous wordforms have an average of 2.01 possible stress positions, so a perfect stress placement tool has a 50/50 chance of getting them right.
 This means that a perfect stress placement tool could achieve 99.48\% on their task.}
@@
 The experiment was limited by properties of the gold standard corpus, including
-its size, genre distribution, and quality. Ideally, a gold standard corpus for 
-Russian wordstress should mark 
-whether monosyllabic words are stressed, and of course its annotations should
-be accurate. Our gold corpus represents a broad variety of text genres, which 
+its size, genre distribution, and quality.
+%Ideally, a gold standard corpus for Russian wordstress should mark whether monosyllabic words are stressed, and of course its annotations should be accurate.
+Our gold corpus represents a broad variety of text genres, which 
 makes our results more generalizable, but a larger corpus would allow for
@@
 technical writing to have a lower proportion of words with stress ambiguity, 
-since it uses a higher proportion of borrowed words, calques, and neologisms with 
-simple stress patterns.
+since it contains a higher proportion of borrowed words, calques, and neologisms 
+with simple stress patterns.\lj{What about NomPl/AccPl -a spreading in tech jargon?}
 
@@
 for unknown words, \newcite{hall_sproat_russian_2013} achieve an accuracy of 
-83.9\%. Their model was trained on a full list of 
+83.9\%. Their model was trained on a full list of \rr{COULD?}
 Russian words, which is not representative of the words that would be unknown to 
@@
 (\rus{д\'{о}ктор}$\sim$\rus{доктор\'{а}} \emph{d\'{o}ktor$\sim$doktor\'{a}} 
-`doctor'$\sim$`doctors') are one exception to this generalization.},
+`doctor'$\sim$`doctors') is one exception to this generalization.},
 so we could exclude closed classes of words from the training data, leaving only
@@
 %\rr{What about using mystem to evaluate, convert to our tags, then generate?}
+\rr{morphosyntactic disambiguation}
 

### 2015-03-24 2015-nodalida-ruscg/ruscg.tex
add f-score
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59022 72bbbca6-d526-0410-a7d9-f06f51895060
add two types of eval
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59023 72bbbca6-d526-0410-a7d9-f06f51895060
section assignments
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59024 72bbbca6-d526-0410-a7d9-f06f51895060
more comments
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59292 72bbbca6-d526-0410-a7d9-f06f51895060
table
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59293 72bbbca6

--- 2015-02-18 2015-nodalida-ruscg/ruscg.tex
@@
 \section{Review of literature}
-% ROB + FRAN
+% ROB Russian POS tagging
+
+% FRAN
 % different CGs 
@@
 % estonian
-%
+
+State-of-the-art morphological analysis in Russian is primarily based on finite-
+state technology \cite{Nozhov-03,Segalovich-03}.\footnote{Machine-learning approaches have also been successfully applied to Russian, most notably by \newcite{sharoff08lrec-mocky}.} Almost without exception, all large-scale morphological transducers of Russian are based on the
+forward-looking \emph{Grammatical Dictionary of Russian} \cite{Zaliznjak-77}.
+This dictionary gives fine-grained morphological specifications for more than
+100~000 words, including inflectional endings, morphophonemic alternations, 
+stress patterns, exceptions, and idiosyncratic collocations.
+We base our work on the recently developed [anonymous] morphological transducer, which is also based on Zaliznjak's dictionary (Author, 2015).\footnote{The [anonymous] transducer is implemented using a two-level
+morphology \cite{Koskenniemi-84}, and can be compiled using either {\tt xfst}
+\cite{Beesley.Karttunen-03} or {\tt hfst} \cite{hfst-11}} This finite-state transducer (FST) generates all possible 
+morphosyntactic readings of each wordform, regardless of its frequency or probability.
+Because Russian is a relatively highly inflected language, broad coverage is important, but widespread homonymy leads to the generation of many spurious readings, as discussed in Section~\ref{ambiguity} below.
+Because of this, one of the foundational steps in Russian natural language processing is morphosyntactic disambiguation, or \emph{snjatie omonimii}.
+
+ -- and our Constraint 
+Grammar\footnote{Implemented using vislcg3 constraint grammar parser
+(http://beta.visl.sdu.dk/cg3.html).}
+\cite{Karlsson-90,Karlsson.Voutilainen.ea-95} then removes
+some readings based on syntactic context.
 
@@
 
-\section{Ambiguity in Russian}
+\section{Ambiguity in Russian} \label{ambiguity}
+
+% ROB
+% homography
+
+% sketch out the different kinds of ambiguity and show what percentage they are 
+% give examples
+% describe stress related ambiguity, e.g. where orthographic form is the same but diff. stress
+
+\begin{table}
+  \centering
+  \begin{tabular}{|l|r|r|}
+    \hline
+    \textbf{Type}     & \textbf{\% all tokens} & \textbf{\% ambig. tokens} \\
+    \hline
+    Intraparadigmatic & 59.0                   & 90.9   \\
+    Incongruent       & 27.7                   & 42.7   \\ 
+    Congruent         & 1.20                   & 1.8    \\ 
+    \hline
+  \end{tabular}
+  \caption{Types of ambiguity}
+  \label{table:ambiguity}
+\end{table}
+
+% numbers not perfect, e.g. лицо should be 'cong.' but will turn up as 'intra.'
 
 \section{Pipeline}
+
+% FRAN
 
@@
 
+% det/prn (e.g. pos... их, я, мой)
+% 'over'generation (e.g. all dets get all gen/nbr/case combinations)
+
 \subsection{Disambiguation rules}
+%FRAN
 
@@
 \section{Development process}
+%FRAN
 
@@
 
+\subsection{Intrinsic}
+%FRAN
+
 \begin{table*}
   \centering
-  \begin{tabular}{|l|r|r|r|r|}
+  \begin{tabular}{|l|r|r|r|r|r|}
     \hline
-    \textbf{Domain} & \textbf{Tokens} & \textbf{Precision} & \textbf{Recall} & \textbf{Ambig. solved} \\
+    \textbf{Domain} & \textbf{Tokens} & \textbf{Precision} & \textbf{Recall} & \textbf{F-score} & \textbf{Ambig. solved} \\
     \hline
-    Wikipedia       & 7,857      & 0.506        & 0.996    & 44.92\%  \\
-    News            & 642        & 0.471        & 0.990    & 41.60\%  \\
-    Literature      & 1,652      & 0.473        & 0.984    & 42.95\%  \\
+    Wikipedia       & 7,857      & 0.506        & 0.996    & 0.671 & 44.92\%  \\
+    Literature      & 1,652      & 0.473        & 0.984    & 0.638 & 42.95\%  \\
+    News            & 642        & 0.471        & 0.990    & 0.638 & 41.60\%  \\
     \hline
-    Total           & 10,150     & 0.498        &  0.994   &  44.39\% \\
+    \textbf{Average}& 10,150     & 0.498        &  0.994   & 0.663 & 44.39\% \\
     \hline
@@
 
+\subsection{Extrinsic}
+% ROB accent placement task
+
 \section{Future work}
 
+% ROB
+% FRAN
 % func analysis

### 2015-03-24 2015-nodalida-ruscg/ruscg.tex
added ambiguity intro
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59500 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \section{Review of literature}
-% ROB Russian POS tagging
 
-% FRAN
-% different CGs 
-% faroese
-% estonian
-
-State-of-the-art morphological analysis in Russian is primarily based on finite-
-state technology \cite{Nozhov-03,Segalovich-03}.\footnote{Machine-learning approaches have also been successfully applied to Russian, most notably by \newcite{sharoff08lrec-mocky}.} Almost without exception, all large-scale morphological transducers of Russian are based on the
+State-of-the-art morphological analysis in Russian is primarily based on
+finite-state technology
+\cite{Nozhov-03,Segalovich-03}.\footnote{Machine-learning approaches have also been successfully applied to Russian, most notably by 
+\newcite{sharoff08lrec-mocky}.} Almost without exception, all large-scale 
+morphological transducers of Russian are based on the
 forward-looking \emph{Grammatical Dictionary of Russian} \cite{Zaliznjak-77}.
@@
 
+% FRAN
+% different CGs 
+% faroese
+% estonian
+
  -- and our Constraint 
@@
 
-% ROB
-% homography
+We identify three different types of morphosyntactic ambiguity: 
+intraparadigmatic, morphosyntactically incongruent, and morphosyntactically 
+congruent. A recent study by Author (2015) found that about 7.5\% of 
+morphosyntactic ambiguity in a corpus of Russian resulted in stress placement 
+ambiguity. The following examples make use of stress ambiguity to illustrate 
+each kind of ambiguity. \emph{Intraparadigmatic} ambiguity refers to homographic 
+wordforms belonging to the same lexeme, as shown in \ref{ex:intrahom}.
 
-% sketch out the different kinds of ambiguity and show what percentage they are 
-% give examples
-% describe stress related ambiguity, e.g. where orthographic form is the same but diff. stress
+\needspace{4\baselineskip} % keeps example all together (no orphan line) 
+\ex. Intraparadigmatic homographs \label{ex:intrahom}
+\a. \rus{т\'{е}ла} \emph{t\'{e}la} `body.\textsubscript{SG-GEN}' 
+    \label{ex:bodySGGEN}
+\b. \rus{тел\'{а}} \emph{tel\'{a}} `body.\textsubscript{PL-NOM}' 
+    \label{ex:bodyPLNOM}
+
+The remaining two types of ambiguity occur between lexemes. 
+\emph{Morphosyntactically incongruent} ambiguity occurs between homographs that 
+belong to separate lexemes, and whose morphosyntactic values are different, as 
+shown in \ref{ex:MSincongruent}.
+
+\needspace{4\baselineskip} % keeps example all together (no orphan line) 
+\ex. Morphosyntactically incongruent homographs \label{ex:MSincongruent}
+\a. \rus{н\'{а}шей} \emph{nášej} `our.\textsubscript{F-SG-GEN/DAT/LOC...}'\\
+    \rus{наш\'{е}й} \emph{našéj} `sew on.\textsubscript{IMP-2SG}'
+\b. \rus{дор\'{о}га} \emph{doróga} `road.\textsubscript{N-F-SG-NOM}'\\
+    \rus{дорог\'{а}} \emph{dorogá} `dear.\textsubscript{ADJ-F-SG-PRED}'
+
+\emph{Morphosyntactically congruent} ambiguity occurs between homographs 
+that belong to separate lexemes, and whose morphosyntactic values are identical, 
+as shown in \ref{ex:MScongruent}. As indicated in Table~\ref{table:ambiguity}, 
+this kind of ambiguity is relatively rare, and resolving this ambiguity requires 
+the use of technologies such as Word Sense Disambiguation.
+
+\needspace{6\baselineskip} % keeps example all together (no orphan line) 
+\ex. Morphosyntactically congruent homographs \label{ex:MScongruent}
+\a. \rus{з\'{a}мок} \emph{z\'{a}mok} `castle.\textsubscript{SG-NOM}'\\
+	\rus{зам\'{о}к} \emph{zam\'{o}k} `lock.\textsubscript{SG-NOM}'
+\b. \rus{з\'{a}мков} \emph{z\'{a}mkov} `castle.\textsubscript{PL-GEN}'\\
+	\rus{замк\'{о}в} \emph{zamk\'{o}v} `lock.\textsubscript{PL-GEN}'
+\c. ...\\
+	...
+
+Table \ref{table:ambiguity} shows the proportional prevalence of each kind of 
+ambiguity. Note that these proportions do not sum to 1, since a given token may 
+exhibit more than one kind of ambiguity. The first column shows the proportion 
+of all tokens in a corpus that have each kind of ambiguity. The second column 
+shows what proportion of ambiguous tokens exhibit each kind of ambiguity.
+
+Table~\ref{table:ambiguity} shows that most morphosyntactic ambiguity in 
+unrestricted Russian text is rooted in intraparadigmatic and morphosyntactically 
+incongruent ambiguity. Detailed part-of-speech tagging with morphosyntactic 
+analysis can help disambiguate these forms. Morphosyntactically congruent stress 
+ambiguity cannot be disambiguated by means of detailed part-of-speech tagging, 
+and since congruent stress ambiguity represents only a small percentage of 
+ambiguous wordforms, we leave it to future work.
 
@@
   \centering
-  \begin{tabular}{|l|r|r|}
+  \begin{tabular}{l|rr}
     \hline
-    \textbf{Type}     & \textbf{\% all tokens} & \textbf{\% ambig. tokens} \\
+    \textbf{Type}  & \textbf{all tokens} & \textbf{ambig. tokens} \\
     \hline
-    Intraparadigmatic & 59.0                   & 90.9   \\
-    Incongruent       & 27.7                   & 42.7   \\ 
-    Congruent         & 1.20                   & 1.8    \\ 
+    Intraparadigm. & 0.590                   & 0.909   \\
+    Incongruent    & 0.277                   & 0.427   \\ 
+    Congruent      & 0.120                   & 0.180    \\ 
     \hline
   \end{tabular}
-  \caption{Types of ambiguity}
+  \caption{Proportional frequency of different types of morphosyntactic ambiguity in unrestricted text}
   \label{table:ambiguity}

### 2015-03-25 2015-nodalida-ruscg/ruscg.tex
a small amount of text about the corpus
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59512 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 this kind of ambiguity is relatively rare, and resolving this ambiguity requires 
-the use of technologies such as Word Sense Disambiguation.
+the use of technologies such as word sense disambiguation.
 
@@
 
+\subsection{Corpus}
+
+In order to evaluate the grammar we hand-annotated 10,150 words of Russian text
+from Wikipedia articles, public domain literature and freely-available news sources. The
+annotated texts are available online under the {\sc cc-by-sa} licence.\footnote{\url{...}}
+
+Hand-annotation proceeded as follows: The text was first morphologically analysed, 
+and then an annotator read through the output of the morphological analyser, commenting
+out the readings which were not appropriate in context. This annotated text was then checked
+by a second annotator.
+
+We chose our own texts as opposed to using a well-known hand-annotated corpus
+such as the Russian National Corpus (RNC) for two main reasons: the first was that the 
+RNC is not freely available; the second was that the standards for tokenisation, part-of-speech
+and morphological description are different from our morphological analyser.
+
 \subsection{Intrinsic}

### 2015-03-25 2015-nodalida-ruscg/ruscg.tex
a load of notes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59516 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
+% CGs for Slavic langs
+
+% The preliminary tagging performance is P:,96.1%, R:,99.8% for POS tagging and P:,88.2%, R:,98.1% for complete morphosyntactic tagging.
+\cite{peradin12}
+
 \section{Ambiguity in Russian} \label{ambiguity}
@@
 
+% we should perhaps mention something about systematic ambiguities and tagging guidelines,
+% and how we deal with stuffs
+% e.g. adj/participle
+%      det/pron
+%      acc2/nom
+%      lexicalised passive / pass
+
+% perhaps mention where our annotation guidelines are (do we mention stuff like один)?
+
 \section{Pipeline}
@@
 
+The constraint grammar is composed of 299 rules which
+
 \begin{table}
@@
                      & {\sc select} & {\sc remove} & {\sc map} \\
-    Safe & & & \\
-    Safe heuristic  & & & \\
-    Heuristic &  & & \\
-    Syntax labelling &  &   & \\ 
+    Safe             &   16         &   34         &  -- \\
+    Safe heuristic   &   89         &   76         &  -- \\
+    Heuristic        &   26         &   52         &  -- \\
+    Syntax labelling & --           & --           & 6 \\ 
     \hline
   \end{tabular}
-  \caption{Rules in the grammar are separated into four sections. }
+  \caption{The 299 rules in the grammar are separated into four sections depending
+      on rule reliability. }
 \end{table}
+
+% The philosophy is that safe rules should represent real constraints in
+% the language. Examples might be that a preposition cannot directly preceed a finite verb
+% or that prepositional case requires a preceeding preposition.
+
+% Examples where this might go wrong.
+
+% Safe heuristic rules should represent highly frequent tendencies in the language,
+% for example remove a genitive at the beginning of a sentence if it is capitalised
+% and there is no verb governing the genitive found to the right.
+
+% Examples where this might go wrong.
+
+Heuristic rules are those which we do not consider linguistic constraints, but express
+preferences, often dealing with overgeneration or overspecification in the morphological
+analyser. For example, remove the verbal adverb reading of \rus{такая}, which could be 
+the feminine singular nominative of \rus{такой} `such' or the verbal adverb 
+of \rus{такать} `say \emph{well}\ldots'.
+
+% Examples where this might go wrong.
+
+Given a large hand-annotated corpus we believe that most of the heuristic rules would be 
+better replaced with information learnt from the corpus through stochastic methods.
 
@@
 %FRAN
+
+A common approach taken when writing constraint grammar rules is to 
+apply the existing rule set to a new text, write new rules to deal with the 
+ambiguities, then apply the rules to a hand-annotated corpus to see
+how often the rule disambiguated correctly \cite{voutilainen99}.
+
+Due to the lack of a hand-annotated corpus compatible with our morphological
+analyser, we adopted a slightly modified technique. We picked a random text
+from the Russian Wikipedia, ran it through the morphological analyser, wrote 
+rules, and then ran the rules on the whole Wikipedia corpus. For each rule,
+we collected around 100 example applications and checked them. If a rule
+selected the appropriate reading in all cases, we included it in the \emph{safe}
+rule set, if it removed an appropriate reading in less then three cases, 
+then we included it in the \emph{safe heuristic} rule set. Otherwise we either
+discarded the rule or included it in the heuristic rule set.
 
@@
 from Wikipedia articles, public domain literature and freely-available news sources. The
-annotated texts are available online under the {\sc cc-by-sa} licence.\footnote{\url{...}}
+annotated texts are available online under the {\sc cc-by-sa} licence.\footnote{\url{http://...}}
 
@@
 
+% safe:
+% <REMOVE:382>	0	29	0	1
+% REMOVE Det IF (1 EOS) ;
+% Но, разумеется, никогда нам не исчерпать всего явления, не добраться до конца и начала его.
+% postposed determiner (or pronoun?)
+
+% <REMOVE:492>	0	12	0	1
+% REMOVE Det IF (0 Det OR Pron) (1C Ne) ;
+% Это я знал еще с 46-го года, когда начал писать, а может быть и раньше, - и факт этот не раз поражал меня и ставил меня в  недоумение  о  полезности искусства при таком видимом его бессилии.
+% postposed determiner
+
 \subsection{Extrinsic}
@@
 
+% better development workflow
+
 \section{Conclusions}
 
-%\section*{Acknowledgments}
-%
+\section*{Acknowledgments}
+\textit{Removed for review}
+
 %Do not number the acknowledgment section. Do not include this section

### 2015-03-25 2015-nodalida-ruscg/ruscg.tex
stuff}
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59523 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-\cite{yablonsky1999russian}
+The morphological transducer used in this study is described in Author (2015). 
+It is primarily based on Zaliznjak's ``Grammatical dictionary of Russian'', 
+including the 2001 version's appendix of proper nouns. It also includes approximately CITE Grishina and Lyashevskaya 2008. http://dict.ruslang.ru/gram.php .
 
+Example~\ref{ex:FSToutput} gives some examples of the output.
+
+\ex. \label{ex:FSToutput} 
+	\a. \rus{новый}+A+Msc+Inan+Sg+Nom
+	\b. \rus{автомат}+N+Msc+Inan+Sg+Nom
+	\c. 
 % det/prn (e.g. pos... их, я, мой)

### 2015-03-27 2015-nodalida-ruscg/ruscg.tex
comments
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59524 72bbbca6-d526-0410-a7d9-f06f51895060
stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59534 72bbbca6-d526-0410-a7d9-f06f51895060
more comments
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59544 72bbbca6-d526-0410-a7d9-f06f51895060
more tuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59545 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-% The philosophy is that safe rules should represent real constraints in
-% the language. Examples might be that a preposition cannot directly preceed a finite verb
-% or that prepositional case requires a preceeding preposition.
+The philosophy is that safe rules should represent real constraints in
+the language. Examples might be that a preposition cannot directly precede a finite verb
+or that prepositional case requires a preceeding preposition.
 
@@
 
-% Safe heuristic rules should represent highly frequent tendencies in the language,
-% for example remove a genitive at the beginning of a sentence if it is capitalised
-% and there is no verb governing the genitive found to the right.
+Safe heuristic rules should deal with highly frequent tendencies in the language.
+For example remove a genitive at the beginning of a sentence if it is capitalised
+and there is no verb governing the genitive found to the right and there is also no
+negated verb to the right. This rule relies on the fact that if the genitive is in first 
+position in the sentence it cannot modify anything before, and no preposition 
+can be governing it.
+
+This kind of rule often relies on completeness of sets, in this case the set of verbs
+that can take a genitive complement.
+
+%It also relies on accurate delimitation of sentence
+%boundaries, including rules for dealing with \ldots ellipsis.
 
@@
 preferences, often dealing with overgeneration or overspecification in the morphological
-analyser. For example, remove the verbal adverb reading of \rus{такая}, which could be 
+transducer. For example, remove the verbal adverb reading of \rus{такая}, which could be 
 the feminine singular nominative of \rus{такой} `such' or the verbal adverb 
@@
 
-% safe:
+%% safe:
+
 % <REMOVE:382>  0       29      0       1
@@
 % Но, разумеется, никогда нам не исчерпать всего явления, не добраться до конца и начала его.
-% postposed determiner (or pronoun?)
+%+ postposed determiner (or pronoun?)
 
@@
 % Это я знал еще с 46-го года, когда начал писать, а может быть и раньше, - и факт этот не раз поражал меня и ставил меня в недоумение о полезности искусства при таком видимом его бессилии.
-% postposed determiner
+%+ postposed determiner (факт этот)
+
+% <REMOVE:414>	0	0	0	1
+% REMOVE A + Short IF (-1C Fin OR Adv OR A) (0C Short OR Adv) ;
+% А если удастся, то я прошу только, чтоб схоронили меня, вполне убедясь, что я мертвая, потому что совсем неприятно проснуться в гробу под землею.
+%+ bad rule. adverb (maybe adj too) does not make sense here
+
+% <REMOVE:433>	0	992	0	1
+% REMOVE NGDAIP - Acc - Prp - Loc IF (-1C* Pr/V OR Pr/Na BARRIER (*) - Adv - Comp - DetIndecl - ModAcc - ModPrp) ;
+% В 1960-х электрифицированные высокоскоростные железные дороги появились в Японии и некоторых других странах.
+%+ incomplete barrier (ordinal in another case)
+
+% <REMOVE:494>	0	60	0	1
+% REMOVE Det IF (0 Det OR Pron) (1 Cm LINK 1 CC OR CS) ;
+% Но какие, однако же, два разные создания, точно обе с двух разных планет!
+%+ parentheticals
+
+%% safe-heuristic:
+
+% <SELECT:718>  0       18      0       1
+% SELECT A + $$NGDAIP IF (1C* N + $$NGDAIP BARRIER Punct OR Pr OR Lparen OR NGDAIP - $$NGDAIP LINK -1C* A + $$NGDAIP BARRIER Punct OR Pr OR Lparen OR NGDAIP - $$NGDAIP);
+% Движущей силой в поездах являются локомотивы, использующие электричество или производящие собственную мощность, обычно дизельными двигателями.
+%+ missing unification on number, rule too loose.
+
+% <REMOVE:769>  0       57      0       1
+% REMOVE IV IF (0 TV OR IV) (1C Acc) (NOT 1 AccAdv) ;
+% Их мама внутри дома с кошкой, она смотрит в окно и видит играющих Ваню и Машу.
+%+ participle is intransitive, because it is modifying Vanju and Mašu, but they are object of видит so look like object of играющих.
+both syntactically valid, but semantics would tell you that you can't play people.
+
+% <SELECT:692>	0	10	0	2
+% SELECT Pron IF (0C Prp) (0 Pron OR Det) (NOT 0 Num) (NOT 1* Prp BARRIER (*) - Adv - DetIndecl) ;
+% Наземные цели (танки, бронетранспортёры), характерные для штурмовиков и ударных вертолётов, имеют, в большинстве своём, значительно более тяжёлое и прочное бронирование, чем авиационные, что определяет использование бронебойных снарядов.
+%+ postposed determiner своём
+
+% <SELECT:864>	0	13	0	1
+% SELECT Nom IF (0C Nom OR Acc) (-1C* Nom BARRIER NGDAIP - Nom - Acc LINK -1C* IV BARRIER NGDAIP - Nom OR TV OR CLB OR Pr) ;
+% Если самоубийство не удастся, то пусть соберутся все отпраздновать мое воскресение из мертвых бокалами Клика.
+% [ only IV ] (anything not nom) [ only nom ] (any case but nom/acc) [nom or acc]
+%+ bad rule
+
+% <REMOVE:784>	0	67	0	2
+% REMOVE Acc IF (-1C Fin + IV) (NOT 0 AccAdv) ;
+% Я именно заметил ему перед этим, что я, чуть не сорок лет знающий "Горе от ума", только в этом году понял как следует один из самых ярких типов этой комедии, Молчалина, и понял именно, когда он же, то есть этот самый писатель, с которым я говорил, разъяснил мне Молчалина, вдруг выведя его в одном из своих сатирических очерков.
+%+ parenthetical (как следует)
+%!
+Было установлено, что после того, как 6 апреля 1994 года в авиакатастрофе погиб президент Хабиаримана, Багосора полностью захватил контроль над политической и военной ситуацией в стране, следовательно несёт ответственность за произошедшие затем события .
+%! нести = TV here, but IV ?
+
+% <REMOVE:656>	0	16	0	3
+% REMOVE TV - Pass IF (NOT 1* Acc) (NOT -1* Acc) ;
+% Она ничего не видит, она считает.
+% Люди никогда не получают ответа, когда говорят с собаками!
+% Она смотрит везде, но не может его найти.
+%+ genitive object
+
+%      REMOVE Acc IF (-1C Fin + IV) (NOT 0 AccAdv) ;
+%      bad interaction with auxiliary verb rule, we remove the acc reading on его
+
+%% heuristic:
+
+% <REMOVE:917>	0	533	0	2
+% REMOVE Dat IF (NOT 0 Prn/Sebe) (NOT 0 Anim OR Cog OR Ant) (NOT 0 Pron) (NOT 1* V/Dat) (NOT -1* V/Dat) (NOT -1* Prep/Dat) (NOT -1C A + Dat) ;
+% В связи с этим ортодоксальности стали противопоставлять ересь 
+%+ incomplete set (e.g. стали should appear in V/Dat)
+% В противоположность «юбилейной концепции» часть историков и археологов считает, как и прежде, что образование Киева как города проходило в VIII—X веках.
+%+ противоположность takes dative (против-)
+
+% <REMOVE:954>	0	272	0	7
+% REMOVE Adv/I IF (NOT -1* Adv/Ješë OR Jest OR Rel OR No BARRIER (*) - Adv) (NOT -1* CC BARRIER (*) - Adv) ;
+% А знаете ли вы, - вдруг сказал мне мой собеседник, видимо давно уже и глубоко пораженный своей идеей, - знаете ли, что, что бы вы ни написали, что бы ни вывели, что бы ни отметили в художественном произведении, - никогда вы не сравняетесь с действительностью.
+% Действительность тотчас же представит вам в этом же роде такой фазис какой вы и еще и не предлагали и превышающий все, что могло создать ваше собственное наблюдение и воображение!
+% Действительность тотчас же представит вам в этом же роде такой фазис какой вы и еще и не предлагали и превышающий все, что могло создать ваше собственное наблюдение и воображение!
+% Это я знал еще с 46-го года, когда начал писать, а может быть и раньше, - и факт этот не раз поражал меня и ставил меня в недоумение о полезности искусства при таком видимом его бессилии.
+% Но ведь в том-то и весь вопрос: чей глаз и кто в силах? 
+% Вот эта-то смерть и напомнила мне о сообщенном мне еще летом самоубийстве дочери эмигранта.
+% Зачастую и после убийств над телами жертв совершали издевательства — разрубали их на куски и т. д. 
+%+ this is a complete mess, in certain genres, и is much more frequent as an adverb
+% probably needs collocation info
+
+% <SELECT:1065>	0	4	0	1
+% SELECT $$NGDAIP IF (0 N + $$NGDAIP) (1 @CNP LINK 1 N + $$NGDAIP) ;
+% Однако после войны в Афинах, как и в Греции в целом, начался период ускоренного развития, который длился до 1980-х, когда впервые дали о себе знать проблема перенаселения столицы и проблема транспорта.
+%+ doesn't take into account genitive modifiers
+
+% <SELECT:1071>	0	37	0	1
+% SELECT $$NGDAIP IF (0 N + $$NGDAIP) (-1 @CNP LINK -1 N + $$NGDAIP) ;
+% Образ триединого бога есть сферическая поверхность, а именно: бог-отец в центре, бог-сын — на поверхности и святой дух — в симметричном отношении между центром и описанной вокруг него сферической поверхностью.
+%+ @CNP should be @CVP because of the тире
+
+% <SELECT:1013>	0	4	0	1
+% SELECT Sg IF (0 Prop) (0 Gen) (-1 N) (1 EOS) (NOT -1* Fin + Pl) ;
+% Старое здание Афинского университета, расположенное на проспекте Панепистимиу, является одним из самых стильных в городе, наряду с Национальной библиотекой и Академией Афин.
+%+ massive hack, not linguistically motivated
+%+ could be improved by only applying to ant/pat
 

### 2015-03-28 2015-nodalida-finnormal/finnormal2015.tex
deanonymise
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59506 72bbbca6-d526-0410-a7d9-f06f51895060
changes suggested by reviewers
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59547 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-\author{}
-%
-%Francis M. Tyers\\
-%  UiT Norgga \'{a}rktala\v{s} universitehta, \\
-%  N-9015 Norway \\
-%  {\tt francis.tyers@uit.no} \\\And
-%  Inari Listenmaa\\
-%  Chalmers tekniska h\"{o}gskola, \\
-%  Sweden \\
-%  {\tt inari@chalmers.se} \\}
-%
+\author{ Inari Listenmaa\\
+  Chalmers tekniska h\"{o}gskola, \\
+  Sweden \\
+  {\tt inari@chalmers.se} \\\And
+ Francis M. Tyers\\
+  HSL-fakultehta, \\
+  UiT Norgga \'{a}rktala\v{s} universitehta, \\
+  N-9015 Norway \\
+  {\tt francis.tyers@uit.no} }
+
 \date{}
@@
 \begin{abstract}
-  This paper presents an unsupervised method for converting between colloquial Finnish
+  This paper presents a rule-based method for converting between colloquial Finnish
   and standard Finnish. The method relies upon a small number of orthographical rules
@@
   corpus consisting of aligned sentences in colloquial Finnish, orthographically-standardised
-  colloquial Finnish and standard Finnish. The methods we present outperforms the baseline
-  of simply treating colloquial Finnish as standard Finnish and offers promise for the adaptation
-  of language-technology tools created for standard Finnish to colloquial Finnish. To this end
-  the paper also presents preliminary results which show promise for using normalisation in 
+  colloquial Finnish and standard Finnish. The method we present outperforms the baseline
+  of simply treating colloquial Finnish as standard Finnish, but is outperformed by a phrase-based 
+ MT system trained by the evaluation corpus.
+% Nevertheless, it offers promise for the adaptation of language-technology tools created for standard Finnish to colloquial Finnish. To this end
+  The paper also presents preliminary results which show promise for using normalisation in 
   the machine translation task.
@@
 forms, where they exist. The application of these tools to non-standard
-language can cause a substantial decrease in quality. For example in machine translation,
+language can cause a substantial decrease in quality for example in machine translation,
 parsing and part-of-speech tagging \cite{eisenstein2013}. Non-standard language can 
@@
 automatic tools. Another related area is machine translation between different written norms of 
-the same language. For example between Norwegian Bokm\aa{}l and Norwegian Nynorsk \cite{unhammer2009}.
+the same language, for example between Norwegian Bokm\aa{}l and Norwegian Nynorsk \cite{unhammer2009}.
 
@@
 \subsection{Colloquial Finnish}
-
-\newcite{viinikka2013} describe the common meaning of the terms colloquial (\emph{puhekieli}) and standard (\emph{yleiskieli} or \emph{kirjakieli}) Finnish: standard language is unified in morphology and vocabulary, following the regulations of a language board; colloquial language shows local and idiolectal variation, and has structures that are characteristic to spoken variety,  such as discourse particles and incomplete clauses.
-
-% Colloquial Finnish differs from standard Finnish in some key features:
-
-% Morphology shows dialectal variation, shortening and merging of morphemes, e.g. \emph{en min\"{a}} `not I' $\rightarrow$ \emph{emm\"{a}} (`I don't').
-
-% Syntax: due to the spontaneous nature of colloquial language, syntactic structures are common for spoken language. 
-%Prototypically, colloquial language is spoken and spontaneous, and standard is written, planned and structured.
-
-We illustrate the differences with the following example from our data set.
-Sentence 1 is the original colloquial version.
-The gloss shows the actual word-by-word translation, and the translation shows similar style and register in English.
-
-\ex
-\begingl
-\gla seiskakin oli vaan silleen et fonotaksista p\"{a}\"{a}ttelin //
-\glb seven-{\sc also} was just like.that that phonotactics.{\sc ela} I.deduced //
-\glft `also the seventh, it was like, I just deduced it from phonotactics' //
-\endgl
-\xe
-
-For the normalised version,\footnote{The normalised version is converted orthographically and lexically, but not syntactically or stylistically.} we changed only morphology and vocabulary.
-On the lexical level, the word \emph{seiska} `number 7' is colloquial style, and in the standard translation it is replaced by the ordinal \emph{seitsem\"{a}s} `seventh'.
-Other changes in the normalised version are common morphological or phonological, phenomena, such as reducing the diphthong in \emph{vain} $\rightarrow$ \emph{vaan} or shortening in  \emph{ett\"{a}} $\rightarrow$ \emph{et}.
-The original sentence and the normalised translation are shown below, aligned word by word.
-
-\ex
-\begingl
-\gla seiskakin oli vaan silleen et fonotaksista p\"{a}\"{a}ttelin //
-\glb seitsem\"{a}skin oli vain sill\"{a}\#lailla ett\"{a} fonotaksista p\"{a}\"{a}ttelin //
-%\glb seventh-{\sc also} was just like.that that phonotactics.{\sc ela} I.deduced //
-% \glft `also the seventh, it was like, I just deduced it from phonotactics' //
-\endgl
-\xe
-
- The syntactic structure of the original sentence is markedly spoken; the word \emph{seiska} is topicalised, and the main information ``deduced from phonotactics'' is in a subordinate clause. 
-The translation into standard Finnish is shorter and more precise, leaving just the main information.
-
-\ex
-\begingl
-\gla p\"{a}\"{a}ttelin seitsem\"{a}nnenkin fonotaksista //
-\glb I.deduced seventh-{\sc also} phonotactics.{\sc ela} //
-\glft `I deduced also the seventh from phonotactics' //
-\endgl
-\xe
-
-% ``Yleens\"{a} kirjakieli ymm\"{a}rret\"{a}\"{a}n samaksi kuin yleiskieli: kielenhuollon suositusten mukaiseksi standardoiduksi kieleksi, jonka sanasto on yleistajuista.
-Puhekieli puolestaan miellet\"{a}\"{a}n kirjakielt\"{a} arkisemmaksi ja tyylilt\"{a}\"{a}n vapaamuotoisemmaksi kielimuodoksi, jonka ei tarvitse noudattaa kirjakielen s\"{a}\"{a}ntöj\"{a} (esim.
-mie, h\"{a}nen koira, me kirjotetaan vs.
-min\"{a}, h\"{a}nen koiransa, me kirjoitamme).''
-
-\section{Corpus}
-
-\begin{table}
-  \centering
- \begin{tabular}{|l|r|r|r|}
-    \hline
-    \multirow{2}{*}{\textbf{Section}} & \multicolumn{3}{|c|}{\textbf{Tokens}} \\\cline{2-4}
-                                      & \texttt{train} & \texttt{test} & \texttt{dev} \\
-    \hline
-    Colloquial                           & 5,103 & 1,012 & 1,003 \\ 
-    Normalised                           & 5,105 & 1,012 & 1,003 \\ 
-    Standardised                         & 4,982 & 991 & 1,000 \\
-    \hline
- \end{tabular} 
-  \caption{Statistics on sentences from the parallel corpus.}
-  \label{table:corpsize}
-\end{table}
 
@@
 
+\newcite{viinikka2013} describe the common meaning of the terms colloquial (\emph{puhekieli}) and standard (\emph{yleiskieli} or \emph{kirjakieli}) Finnish: standard language is unified in morphology and vocabulary, following the regulations of a language board; colloquial language shows local and idiolectal variation, and has structures that are characteristic to spoken variety,  such as discourse particles and incomplete clauses.
+
+We illustrate the differences with the following example from our data set.
+Sentence 1 is the original colloquial version.
+The gloss shows the actual word-by-word translation, and the translation shows similar style and register in English.
+
+\ex
+\begingl
+\gla seiskakin oli vaan silleen et fonotaksista p\"{a}\"{a}ttelin //
+\glb seven-{\sc also} was just like.that that phonotactics.{\sc ela} I.deduced //
+\glft `also the seventh, it was like, I just deduced it from phonotactics' //
+\endgl
+\xe
+
+For the normalised version,\footnote{The normalised version is converted orthographically and lexically, but not syntactically or stylistically.} we changed only morphology and vocabulary.
+On the lexical level, the word \emph{seiska} `number 7' is colloquial style, and in the standard translation it is replaced by the ordinal \emph{seitsem\"{a}s} `seventh'.
+Other changes in the normalised version target common morphological or phonological phenomena, such as restoring the reduced diphthong in \emph{vaan} $\rightarrow$ \emph{vain}.
+The original sentence and the normalised translation are shown below, aligned word by word.
+
+\ex
+\begingl
+\gla seiskakin oli vaan silleen et fonotaksista p\"{a}\"{a}ttelin //
+\glb seitsem\"{a}skin oli vain sill\"{a}\#lailla ett\"{a} fonotaksista p\"{a}\"{a}ttelin //
+%\glb seventh-{\sc also} was just like.that that phonotactics.{\sc ela} I.deduced //
+% \glft `also the seventh, it was like, I just deduced it from phonotactics' //
+\endgl
+\xe
+
+ The syntactic structure of the original sentence is markedly spoken; the word \emph{seiska} is topicalised, and the main information ``deduced from phonotactics'' is in a subordinate clause. 
+The translation into standard Finnish is shorter and more precise, leaving just the main information.
+
+\ex
+\begingl
+\gla p\"{a}\"{a}ttelin seitsem\"{a}nnenkin fonotaksista //
+\glb I.deduced seventh-{\sc also} phonotactics.{\sc ela} //
+\glft `I deduced the seventh also from phonotactics' //
+\endgl
+\xe
+
+\section{Corpus}
+
+\begin{table}
+  \centering
+ \begin{tabular}{|l|r|r|r|}
+    \hline
+    \multirow{2}{*}{\textbf{Section}} & \multicolumn{3}{|c|}{\textbf{Tokens}} \\\cline{2-4}
+                                      & \texttt{dev} & \texttt{test} & \texttt{train} \\
+    \hline
+    Colloquial                           & 1,003 & 1,012 & 5,103 \\ 
+    Normalised                           & 1,003 & 1,012 & 5,103 \\ 
+    Standardised                         & 1,000 & 991 & 4,982 \\
+    \hline
+ \end{tabular} 
+  \caption{Statistics on sentences from the parallel corpus.}
+  \label{table:corpsize}
+\end{table}
+
 Our evaluation corpus was created by manually translating texts in colloquial Finnish
@@
 
-The corpus was split into three parts, testing, development and training. The testing
-and development portions contain approximately 1,000 words each, with the remaining approximately 5,000 words for training
-supervised models.\footnote{The corpus is split into 14 files of 500 words each. Files 01--02 
+The corpus was split into three parts, development, testing and training.
+The development and testing portions contain approximately 1,000 words each, with the remaining approximately 5,000 words for training
+phrase-based and character-based models.\footnote{The corpus is split into 14 files of 500 words each. Files 01--02 
   were used for development; 03--04 for testing and 05--14 for training.} Table~\ref{table:corpsize} gives statistics on the number of words in each section.
-
-\section{Experiments}
-
-\subsection{Unsupervised normalisation}
-
-For the unsupervised normalisation we applied a set of regular-expression based 
-replace rules to the input text to produce all the possible candidate sentences 
-in standard Finnish and then used a target-language model to rank the possible 
-candidates. The candidate with the highest rank was selected as the normalised sentence.
-For the target-language model we used the Finnish side of the English--Finnish EuroParl
-parallel corpus \cite{koehn2005}.
-
-We developed two sets of rules for the unsupervised normalisation approach:
-
-\begin{itemize}
-  \item \texttt{rules-1}: 273 rules from \newcite{karlsson2008}'s grammar of Finnish (\S95--97). The 
-    rules took around one hour to implement.
-  \item \texttt{rules-2}: 98 rules written by examining the development corpus, these rules also
-    took approximately one hour to implement.
-\end{itemize}
-
-The rules included both simple one-to-one (`m\"{a}' $\rightarrow$ `min\"{a}') and one-to-many (`emm\"{a}' $\rightarrow$ `en min\"{a}') word correspondences,
-and also regular expression substitutions which could match a prefix or a suffix (`(?+)nkaa' $\rightarrow$ `\textbackslash1n kanssa').
-
-Table~\ref{table:unsup-trace} gives an example trace of the system on a simple sentence
-using three replace rules.
 
@@
   \end{tabular}
-  \caption{Example trace of the unsupervised normalisation method. Rules are applied in order to each of 
+  \caption{Example trace of the normalisation method. Rules are applied in order to each of 
      the possible candidate translations in turn. The candidates are then ranked using an $n$-gram language model 
@@
      or the best candidate is output.}
-  \label{table:unsup-trace}
+  \label{table:normal-trace}
 \end{table*}
+  \newpage
+
+\section{Experiments}
+
+\subsection{Rule-based normalisation}
+
+For the rule-based normalisation we applied a set of regular-expression based 
+replace rules to the input text to produce all the possible candidate sentences 
+in standard Finnish and then used a target-language model to rank the possible 
+candidates. The candidate with the highest rank was selected as the normalised sentence.
+For the target-language model we used the Finnish side of the English--Finnish EuroParl
+parallel corpus \cite{koehn2005}.
+
+We developed two sets of rules:
+
+\begin{itemize}
+  \item \texttt{rules-1}: 273 rules from \newcite{karlsson2008}'s grammar of Finnish (\S95--97). The 
+    rules took around one hour to implement.
+  \item \texttt{rules-2}: 98 rules written by examining the development corpus, these rules also
+    took approximately one hour to implement.
+\end{itemize}
+
+The rules included both simple one-to-one (`m\"{a}' $\rightarrow$ `min\"{a}') and one-to-many (`emm\"{a}' $\rightarrow$ `en min\"{a}') word correspondences,
+and also regular expression substitutions which could match a prefix or a suffix (`(?+)nkaa' $\rightarrow$ `\textbackslash1n kanssa').
+
+Table~\ref{table:normal-trace} gives an example trace of the system on a simple sentence
+using three replace rules.
 
@@
 
-The target-language model corpus, trained using KenLM \cite{heafield2011}, used was the same as in the unsupervised experiments.
+The target-language model corpus, trained using KenLM \cite{heafield2011}, used was the same as in the rule-based experiments.
 
@@
      of the rules in \texttt{rules-1} and \texttt{rules-2}. The figures in bold are the 
-     best results for unsupervised and supervised methods.}
+     best results for rule-based and SMT methods.}
   \label{table:results-norm}
@@
 
-Our unsupervised method performs similarly to the character-based machine translation systems. Both 
-the character-based systems and the unsupervised system achieve around half of the performance of the 
-supervised phrase-based system.
+Our rule-based method performs similarly to the character-based machine translation systems. Both 
+the character-based systems and the rule-based system achieve around half of the performance of the phrase-based system.
 
-Out of the unsupervised systems, the set of rules which was created by examining the development corpus
+Out of the rule-based systems, the set of rules which was created by examining the development corpus
 outperforms both the set of rules from the grammar and the combined rules.
@@
 As for the methods, we would like to follow \newcite{scannell2014} in incorporating
-`translation' probabilities into our unsupervised normalisation model. Our current model relies 
+`translation' probabilities into our rule-based normalisation model. Our current model relies 
 exclusively on the target-language model probability, however some rules may be more
@@
 
-\textit{Removed for review}
-%Joonas Kylm\"{a}l\"{a}
+Joonas Kylm\"{a}l\"{a}
 

### 2015-03-28 2015-nodalida-finnormal/finnormal2015.tex
add speculation about the two rule sets
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59550 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 Our evaluation corpus was created by manually translating texts in colloquial Finnish
-to standard Finnish.\footnote{The corpus is freely available and published under the 
-CreativeCommons {\sc cc-by-sa} 3.0 licence.} The texts were extracts from internet 
+to standard Finnish. The corpus is freely available and published under the 
+CreativeCommons {\sc cc-by-sa} 3.0 licence\footnote{https://svn.code.sf.net/p/apertium/svn/languages/apertium-fin/texts/normalisation/}.
+The texts were extracts from internet 
 relay chat (IRC) conversations. We performed the conversion process in two steps, the first
@@
 outperforms both the set of rules from the grammar and the combined rules.
+The rules from the grammar capture more general tendencies, whereas the rules from the development corpus are more lexicalised.
+Since the testing corpus is small and only contains text from a single author, the higher performance of the second rule set could also be an due to overfitting.
 
-It is interesting to note that supervised systems trained on the normalised section of the corpus outperform
+It is interesting to note that MT systems trained on the normalised section of the corpus outperform
 those trained on the standardised corpus. One explanation for this could be that the corpus size is small, so 
@@
 wide-coverage experiments. We intend to expand the corpus size to at least 10,000
-words. Another weakness of the corpus is that it contains text from a single author,
-we would ideally like to add texts from other authors and other colloquial genres. Although
-the challenge here is finding text that is both free of privacy issues and available
-to release under a free/open-source licence.
+words. Another weakness of the corpus is that it contains text from a single author.
+We would ideally like to add texts from other authors and other colloquial genres---the 
+challenge here is finding text that is both free of privacy issues and available
+to release under a free/open-source sence.
 
@@
 
-Joonas Kylm\"{a}l\"{a}
+We thank Joonas Kylm\"{a}l\"{a} for translating the colloquial sentences into normalised and standardised versions.
 

### 2015-03-28 2015-nodalida-ruscg/ruscg.tex
learning curve for voting tagger
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59548 72bbbca6-d526-0410-a7d9-f06f51895060
måns' paper
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59549 72bbbca6-d526-0410-a7d9-f06f51895060
some intrinsic eval
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59551 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{url}
+\usepackage{graphicx}
 \usepackage[small,bf]{caption}
@@
 \subsection{Morphological analyser}
-% ROB
+\label{sec:morph}
 
@@
 \ex. \label{ex:FSToutput} 
-	\a. \rus{новый}+A+Msc+Inan+Sg+Nom
-	\b. \rus{автомат}+N+Msc+Inan+Sg+Nom
+	\a. \rus{новый}\texttt{{\small <adj><m><nn><sg><nom>}}
+	\b. \rus{автомат}\texttt{{\small <n><m><nn><sg><nom>}}
 	\c. 
@@
 
-\subsection{Combining with a statistical tagger}
-
 \section{Evaluation}
+\label{sec:eval}
 
 \subsection{Corpus}
-
 In order to evaluate the grammar we hand-annotated 10,150 words of Russian text
@@
 
-We chose our own texts as opposed to using a well-known hand-annotated corpus
+We chose to annotate our own texts as opposed to using a well-known hand-annotated corpus
 such as the Russian National Corpus (RNC) for two main reasons: the first was that the 
@@
   \caption{Results for the test corpora.}
+  \label{table:results}
 \end{table*}
@@
 % rule error analysis
+
+\begin{description}
+  \item[ Bad linguistics:] In some cases a rule did not take into account grammatical possibilities 
+    in the language. e.g. Two simple rules such as 
+    \begin{itemize}
+      \item \texttt{REMOVE Det IF (0 Det OR Pron) (1C Ne) ;}
+      \item \texttt{REMOVE Det IF (0 Det OR Pron) (1 Cm LINK 1 CC OR CS) ;}
+    \end{itemize}
+    did not take into account the possibility of having a postposed determiner as in
+    \begin{itemize}
+      \item \ldots \rus{а может быть и раньше, и факт этот не раз поражал меня} \ldots
+      \item \ldots and maybe even earlier, and fact \textbf{this} not once surprised me \ldots
+    \end{itemize}
+    or a interposed parenthetical as in
+    \begin{itemize}
+      \item \rus{Но какие, однако же, два разные создания, точно обе с двух разных планет!}
+      \item But what, \textbf{exactly} , two different creatures, just both from two different planets!
+    \end{itemize}
+  \item[ Bad rule:] In some cases a rule was simply incorrectly specified. For example, the 
+    following rule was designed to solve the ambiguity between short-form neuter adjectives and 
+    adverbs
+   \begin{itemize}
+     \item \texttt{REMOVE A + Short IF (-1C Fin OR Adv OR A) (0C Short OR Adv) ;}
+   \end{itemize}
+   However there is no reason why we should prefer an adverb over an adjective after an adverb, 
+   \begin{itemize}
+     \item \ldots \rus{потому что совсем неприятно проснуться в гробу под землею.}
+     \item \ldots because [it is] really \textbf{unpleasant} to wake up in a coffin under the ground.
+   \end{itemize}
+  \item[ Incomplete barrier:] xxx 
+   \begin{itemize} 
+     \item \texttt{REMOVE NGDAIP - Acc - Prp - Loc IF (-1C* Pr/V OR Pr/Na BARRIER (*) - Adv - Comp - DetIndecl - ModAcc - ModPrp) ;}
+   \end{itemize}
+   here the rule removes the nominative reading of the adjective to leave the accusative reading because the preposition \rus{в} `in'
+   is found preceeding. 
+   \begin{itemize}
+     \item \rus{В 1960-х электрифицированные высокоскоростные железные дороги появились в Японии и некоторых других странах.}
+     \item In the 1960's \textbf{electrified} high-speed railways appeared in Japan and some other countries.
+   \end{itemize} 
+  \item[ Incomplete set:] In some cases the rule was a good generalisation, but made use of a set
+    which was incomplete. For example:
+   \begin{itemize}
+     \item \texttt{REMOVE Dat IF (NOT 0 Prn/Sebe) (NOT 0 Anim OR Cog OR Ant) (NOT 0 Pron) (NOT 1* V/Dat) (NOT -1* V/Dat) (NOT -1* Prep/Dat) (NOT -1C A + Dat) ;}
+   \end{itemize} 
+   the set \texttt{V/Dat} does not contain the verb \rus{противопоставлять} `opposed to' which takes 
+   a dative argument.
+   \begin{itemize}
+     \item \rus{В связи с этим ортодоксальности стали противопоставлять ересь.}
+     \item In connection with this \textbf{orthodoxy} was opposed to heresy. 
+   \end{itemize}
+  \item[ Rule interaction:] The strong accusative rule below causes incorrect behaviour in the rule to remove transitivity readings
+  \begin{itemize}
+     \item \texttt{REMOVE TV - Pass IF (NOT 1* Acc) (NOT -1* Acc) ;}
+     \item \texttt{REMOVE Acc IF (-1C Fin + IV) (NOT 0 AccAdv) ;}
+  \end{itemize}
+  Consider the following example where \rus{может} `can' is tagged as intransitive, the second rule fires removing 
+  the accusative reading of \rus{его} `him', and thus given the lack of accusative reading, \rus{найти} `find'
+  is disambiguated as intransitive instead of transitive.
+  \begin{itemize}
+    \item \rus{Она смотрит везде, но не может его найти.}
+    \item She looks around, but she cannot \textbf{find him}.
+  \end{itemize}
+  \item[ Difficult linguistics:] Dealing with participles with arguments is challenging in the case that the arguments 
+    of the participle share the same government as the main verb.
+  \begin{itemize}
+    \item \texttt{REMOVE IV IF (0 TV OR IV) (1C Acc) (NOT 1 AccAdv) ;}
+  \end{itemize}
+    Here \rus{Ваню и Машу} `Vanja and Maša' are the object of \rus{видит} `sees' and not \rus{играющих} `playing', although
+    both verbs can take accusative object.
+  \begin{itemize}
+    \item \rus{Их мама внутри дома с кошкой, она смотрит в окно и видит играющих Ваню и Машу.}
+    \item Their mother is inside the house with the cat, she looks through the window and sees Vanja and Maša \textbf{playing}.
+  \end{itemize}
+    This kind of error would ideally be resolved with semantic knowledge. 
+\end{description}
 
@@
 
+\subsection{Combining with a statistical tagger}
+
+Given that just over half of all ambiguity remains after running our preliminary constraint
+grammar and that for many applications unambiguous output is necessary, we decided to 
+experiment with combining the constraint grammar with a statistical tagger to resolve remaining
+ambiguity.
+
+We follow the voting method described by \newcite{hulden12}. We used the freely available
+\texttt{hunpos} part-of-speech tagger \cite{halacsy07}. We performed 10-fold
+cross validation using our evaluation corpus, taking 10\% for testing and 90\% for training, and
+experimented with three configurations:
+
+\begin{itemize}
+  \item \texttt{HMM}: the \texttt{hunpos} part-of-speech tagger with its default options
+  \item \texttt{HMM+Morph}: as with \texttt{HMM} but incorporating the output of our 
+    morphological analyser (see section~\ref{sec:morph}) as a full form lexicon.
+  \item \texttt{HMM+Morph+CG}: we submitted the output from \texttt{HMM+Morph} and the 
+    constraint grammar to a voting procedure, whereby if the constraint grammar left
+    one valid reading, we chose that, otherwise if the constraint grammar left a word
+    with more than one reading, we chose the result from the \texttt{HMM+Morph} tagger.
+\end{itemize}
+
+As can be seen from Figure~\ref{fig:curve}, incorporating the constraint grammar 
+improves the performance of the HMM tagger, an improvement of nearly 5\% in accuracy,
+similar to that reported by \newcite{hulden12} for the same amount of training data. Our 
+constraint grammar also has a much lower precision as a result of the ambiguity remaining
+in the output.
+
+Note that the final accuracy is below the state of the art for Russian, however the state of
+the art employs unavailable training corpora of over five million tokens.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.45\textwidth]{graphics/learning-curve.pdf}
+  \caption{Learning curve for three taggers, \texttt{hunpos} with no lexicon, 
+    \texttt{hunpos} with a lexicon, and \texttt{hunpos} with a lexicon and the 
+    Russian constraint grammar in a voting set up.}
+  \label{fig:curve}
+\end{figure}
+
 \section{Future work}
@@
 % better development workflow
+
+% proof-of-concept for russian, but russian has loads of tagged corpora, for other langs, e.g. sorbian or belarusian 
+% a CG+hmm could be much faster than tagging hundreds of thousands of words of text.
 

### 2015-03-29 2015-nodalida-ruscg/ruscg.tex
some intrinsic eval
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59552 72bbbca6-d526-0410-a7d9-f06f51895060
more
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59553 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 Because Russian is a relatively highly inflected language, broad coverage is important, but widespread homonymy leads to the generation of many spurious readings, as discussed in Section~\ref{ambiguity} below.
-Because of this, one of the foundational steps in Russian natural language processing is morphosyntactic disambiguation, or \emph{snjatie omonimii}.
+Because of this, one of the foundational steps in Russian natural language processing is \rus{снятие омонимии} `morphosyntactic disambiguation'.
 
@@
 Grammar\footnote{Implemented using vislcg3 constraint grammar parser
-(http://beta.visl.sdu.dk/cg3.html).}
+(\url{http://beta.visl.sdu.dk/cg3.html}).}
 \cite{Karlsson-90,Karlsson.Voutilainen.ea-95} then removes
@@
 It is primarily based on Zaliznjak's ``Grammatical dictionary of Russian'', 
-including the 2001 version's appendix of proper nouns. It also includes approximately CITE Grishina and Lyashevskaya 2008. http://dict.ruslang.ru/gram.php .
+including the 2001 version's appendix of proper nouns. It also includes approximately CITE Grishina and Lyashevskaya 2008. \url{http://dict.ruslang.ru/gram.php} .
 
@@
                      & {\sc select} & {\sc remove} & {\sc map} \\
+    \hline
     Safe             &   16         &   34         &  -- \\

### 2015-03-30 2015-nodalida-ruscg/ruscg.tex
add some example output
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59566 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{url}
+\usepackage{alltt}
 \usepackage{graphicx}
@@
 discarded the rule or included it in the heuristic rule set.
+
+\begin{figure*}
+\begin{alltt}
+{\small
+\rus{
+"<В>"                                                         "<работу>"
+        "в" pr                                                     "работа" n f nn sg acc
+
+"<ноябре>"                                                    "<Международный>"
+        "ноябрь" n m nn sg prp                                     "международный" adj m an sg nom
+                                                                   "международный" adj m nn sg acc
+"<1994>"                                        
+        "1994" num                                            "<трибунал>"
+                                                                         "трибунал" n m nn sg acc
+"<года>"                                                           "трибунал" n m nn sg nom
+        "год" n m nn sg gen SELECT:462                             
+;        "год" n m nn pl nom fac SELECT:462                   "<по>"
+                                                                   "по" pr 
+"<в>"                                                             
+        "в" pr                                                "<Руанде>" 
+                                                                   "Руанда" np al f nn sg prp 
+"<Танзании>"                                                       "Руанда" np al f nn sg dat
+        "Танзания" np al f nn pl acc                           
+        "Танзания" np al f nn sg prp                          "<.>" 
+;        "Танзания" np al f nn pl nom REMOVE:424                   "." sent 
+;        "Танзания" np al f nn sg dat REMOVE:433 
+;        "Танзания" np al f nn sg gen REMOVE:433
+
+"<начал>"
+        "начало" n nt nn pl gen
+        "начать" vblex perf tv past m sg
+;        "начать" vblex perf iv past m sg REMOVE:769
+}}
+\end{alltt}
+  \caption{Example output from the morphological analyser and constraint grammar for the sentence \rus{В ноябре 1994 
+           года в Танзании начал работу Международный трибунал по Руанде.} ``The work of the International Tribunal 
+           for Rwanda started in Tanzania in November 1994.'' The input ambiguity is 1.76 readings per word and 
+           the output ambiguity is 1.38 readings per word. Recall is 1.0 and precision is 0.72. 
+ } 
+\end{figure*}
 

### 2015-03-30 2015-nodalida-ruscg/ruscg.tex
example ruls
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59567 72bbbca6-d526-0410-a7d9-f06f51895060

--- 2015-03-29 2015-nodalida-ruscg/ruscg.tex
@@
 discarded the rule or included it in the heuristic rule set.
+
+\begin{figure*}
+\centering
+\begin{alltt}
+{\small
+\rus{
+"<В>"                                                         "<работу>"
+        "в" pr                                                     "работа" n f nn sg acc
+
+"<ноябре>"                                                    "<Международный>"
+        "ноябрь" n m nn sg prp                                     "международный" adj m an sg nom
+                                                                   "международный" adj m nn sg acc
+"<1994>"                                        
+        "1994" num                                            "<трибунал>"
+                                                                         "трибунал" n m nn sg acc
+"<года>"                                                           "трибунал" n m nn sg nom
+        "год" n m nn sg gen SELECT:r462                             
+;        "год" n m nn pl nom fac SELECT:r462                   "<по>"
+                                                                   "по" pr 
+"<в>"                                                             
+        "в" pr                                                "<Руанде>" 
+                                                                   "Руанда" np al f nn sg prp 
+"<Танзании>"                                                       "Руанда" np al f nn sg dat
+        "Танзания" np al f nn pl acc                           
+        "Танзания" np al f nn sg prp                          "<.>" 
+;        "Танзания" np al f nn pl nom REMOVE:r424                   "." sent 
+;        "Танзания" np al f nn sg dat REMOVE:r433 
+;        "Танзания" np al f nn sg gen REMOVE:r433
+
+"<начал>"
+        "начало" n nt nn pl gen
+        "начать" vblex perf tv past m sg
+;        "начать" vblex perf iv past m sg REMOVE:r769
+}}
+\end{alltt}
+  \caption{Example output from the morphological analyser and constraint grammar for the sentence \rus{В ноябре 1994 
+           года в Танзании начал работу Международный трибунал по Руанде.} ``The work of the International Tribunal 
+           for Rwanda started in Tanzania in November 1994.'' The input ambiguity is 1.76 readings per word and 
+           the output ambiguity is 1.38 readings per word. Recall is 1.0 and precision is 0.72. Figure~\ref{fig:rules} 
+           shows the rules that fired for this example sentence.} 
+  \label{fig:output}
+\end{figure*}
+
+\begin{figure*}
+
+\begin{alltt}
+{\small
+### Safe
+
+SELECT:r462 Gen IF (0 Year) (-1 Num LINK -1 Months LINK -1 Pr/V);
+  # Select genitive reading of `\rus{года}' if there is a numeral immediately
+  # to the left, before that there is a month and before that there is 
+  # the preposition `\rus{в}'.
+
+REMOVE:r424 Nom IF (-1C Pr) ;
+  # Remove nominative case if there is a word which can only be a 
+  # preposition immediately to the left.
+
+REMOVE:r433 NGDAIP - Acc - Prp - Loc IF 
+                (-1C* Pr/V OR Pr/Na
+                  BARRIER (*) - Adv - Comp - DetIndecl - ModAcc - ModPrp);
+  # Remove all cases apart from accusative, preposition and locative
+  # if `\rus{в}' or `\rus{на}' are found to the left and are unambiguous. The barrier
+  # is anything that cannot be found inside a noun phrase.
+
+### Safe heuristic
+
+REMOVE:r769 IV IF (0 TV OR IV) (1C Acc) (NOT 1 AccAdv);
+  # Remove an intransitive reading of a verb if the next word can only
+  # be accusative and is not in the set of nouns which can be used 
+  # adverbially in accusative.
+}
+\end{alltt}
+  \caption{Some example rules from the grammar.}
+  \label{fig:rules}
+\end{figure*}
 

### 2015-03-30 2015-nodalida-ruscg/ruscg.tex
fixed ambiguity frequency table
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59568 72bbbca6-d526-0410-a7d9-f06f51895060
замков etc.
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59571 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{fixltx2e} %allows subscripts
+\usepackage{tipa} %allows IPA symbols
+\usepackage{linguex}
 %\usepackage{mathptmx}
@@
 Because Russian is a relatively highly inflected language, broad coverage is important, but widespread homonymy leads to the generation of many spurious readings, as discussed in Section~\ref{ambiguity} below.
-Because of this, one of the foundational steps in Russian natural language processing is \rus{снятие омонимии} `morphosyntactic disambiguation'.
+Because of this, one of the foundational steps in Russian natural language processing is homograph disambiguation.
 
@@
     \hline
-    Intraparadigm. & 0.590                   & 0.909   \\
-    Incongruent    & 0.277                   & 0.427   \\ 
-    Congruent      & 0.120                   & 0.180    \\ 
+    Intraparadigm. & 59.0\%                   & 90.9\%   \\
+    Incongruent    & 27.7\%                   & 42.7\%   \\ 
+    Congruent      & 1.2\%                   & 1.8\%    \\ 
     \hline
   \end{tabular}
-  \caption{Proportional frequency of different types of morphosyntactic ambiguity in unrestricted text}
+  \caption{Frequency of different types of morphosyntactic ambiguity in unrestricted text}
   \label{table:ambiguity}
@@
 
-% numbers not perfect, e.g. лицо should be 'cong.' but will turn up as 'intra.'
+Note that the percentages in Table~\ref{table:ambiguity} overlap, i.e. the readings of a given wordform
+can have more than one kind of ambiguity. For example, the wordform \emph{zamkov} has the readings given in \ref{ex:zamkov}.
+
+\ex. \label{ex:zamkov}
+	 \a. \rus{замок}$^1$+N+Msc+Inan+Pl+Gen \label{ex:zamkovA}
+	 \b. \rus{замок}$^2$+N+Msc+Inan+Pl+Gen \label{ex:zamkovB}
+	 \c. \rus{замковый}+A+Msc+Sg+Pred      \label{ex:zamkovC}
+
+The ambiguity between \ref{ex:zamkovA} and \ref{ex:zamkovB} is morphosyntactically congruent, and
+the ambiguity between \ref{ex:zamkovA}/\ref{ex:zamkovB} and \ref{ex:zamkovC} is
+morphosyntactically incongruent, so this wordform would be counted for both categories in
+Table~\ref{table:ambiguity} above.
 
@@
 \label{sec:morph}
+%ROB
 
 The morphological transducer used in this study is described in Author (2015). 
-It is primarily based on Zaliznjak's ``Grammatical dictionary of Russian'', 
+It is primarily based on Zaliznjak's \emph{Grammatical dictionary of Russian}, 
 including the 2001 version's appendix of proper nouns. It also includes approximately CITE Grishina and Lyashevskaya 2008. \url{http://dict.ruslang.ru/gram.php} .
@@
 % 'over'generation (e.g. all dets get all gen/nbr/case combinations)
+%^^^
+%ROB
 
@@
 
+The constraint grammar described in this paper has been applied to the task of automatic
+word stress placement (Author, 2015). This task is especially relevant for 
+Russian language learners, because vowels are pronounced differently depending on their position 
+relative to stress position. For example, the word \emph{molokó} `milk' is pronounced 
+/\textipa{m@l2kO}/, where each instance of the letter \emph{o} corresponds to a different vowel sound. 
+Russian has complicated patterns of shifting stress, which are difficult for learners to master.
+Almost 99\% of wordforms with ambiguous stress position can be disambiguated morphosyntactically,
+so a constraint grammar can potentially resolve most stress ambiguity indirectly. The results of Author 2015 show that our constraint grammar overcomes about 42\% of the ambiguity relevant to stress
+ambiguity in unrestricted text.
+
 \subsection{Combining with a statistical tagger}

### 2015-03-30 2015-nodalida-ruscg/ruscg.tex
more stuffs
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59573 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-Our purposes: Processing stressed wordforms and open-source machine translation
+This paper presents a preliminary constraint grammar for Russian. The principle 
+objective of the constraint grammar is to produce a high recall grammar to serve
+as input into other natural language processing tasks.
 
-\cite{Karlsson-90}
+The paper is laid out as follows: section~\ref{sec:lit} presents a review of 
+the literature on Russian language processing; section~\ref{ambiguity} gives
+an overview of ambiguity in Russian; section~\ref{sec:pipeline} describes
+our analysis pipeline; section~\ref{sec:devel} gives an account of our 
+development process; section~\ref{sec:eval} presents an evaluation of the 
+system, and sections~\ref{sec:future} and ~\ref{sec:conclusions} present future
+work and conclusions.
+
+%Our purposes: Processing stressed wordforms and open-source machine translation
+
+%\cite{Karlsson-90}
 
@@
 \section{Review of literature}
-
+\label{sec:lit}
 State-of-the-art morphological analysis in Russian is primarily based on
@@
 
- -- and our Constraint 
-Grammar\footnote{Implemented using vislcg3 constraint grammar parser
-(\url{http://beta.visl.sdu.dk/cg3.html}).}
-\cite{Karlsson-90,Karlsson.Voutilainen.ea-95} then removes
-some readings based on syntactic context.
+% -- and our Constraint 
+%Grammar\footnote{Implemented using vislcg3 constraint grammar parser
+%(\url{http://beta.visl.sdu.dk/cg3.html}).}
+%\cite{Karlsson-90,Karlsson.Voutilainen.ea-95} then removes
+%some readings based on syntactic context.
 
-\cite{trosterud2009}
+%\cite{trosterud2009}
 
@@
 % The preliminary tagging performance is P:,96.1%, R:,99.8% for POS tagging and P:,88.2%, R:,98.1% for complete morphosyntactic tagging.
-\cite{peradin12}
+% \cite{peradin12}
 
@@
 \section{Pipeline}
-
+\label{sec:pipeline}
 % FRAN
@@
 	\b. \rus{автомат}\texttt{{\small <n><m><nn><sg><nom>}}
-	\c. 
+%	\c. 
 % det/prn (e.g. pos... их, я, мой)
@@
 \section{Development process}
+\label{sec:devel}
 %FRAN
@@
 \section{Future work}
-
+\label{sec:future}
 % ROB

### 2015-03-30 2015-nodalida-ruscg/ruscg.tex
concl.
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59574 72bbbca6-d526-0410-a7d9-f06f51895060
fill out sections
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59575 72bbbca6-d526-0410-a7d9-f06f51895060

@@
    \end{itemize}
-  \item[ Incomplete barrier:] xxx 
+  \item[ Incomplete barrier:]  Some rules suffered from incomplete barriers, which is something that would
+     benefit from a more systematic treatment.
    \begin{itemize} 
@@
 
+We have a number of plans for future work, the first is increasing the precision of the grammar without
+decreasing recall. Secondly we would like to add syntactic function labelling and dependency analysis. For 
+the dependency analysis we would like to reuse the Giellatekno dependency grammar as in \cite{antonsen10}.
+
+The development workflow could also be improved, for example during the testing of each rule we could
+save the correct decisions of the grammar and gradually build up a tagged development corpus for the language.
+
+Also it is worth noting that although Russian has a great deal of non-free resources, this paper also presents
+a method which is promising for smaller or lesser-resourced Slavic languages such as Sorbian, Rusyn or Belarusian. Instead
+of hand-annotating a large quantity of text, it may be more efficient to work on grammatical resources --- such
+as a morphological analyser and constraint grammar --- and use 
+them alongside a smaller quantity of high-quality annotated text.
+
 \section{Conclusions}
 \label{sec:conclusions}
+
+This paper has presented a preliminary constraint grammar for Russian. The constraint grammar 
+is high recall and improves the performance over a simple HMM-based tagger. It also shows 
+state-of-the-art performance for the stress-placement task.
 

### 2015-03-30 2015-nodalida-ruscg/ruscg.tex
stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59577 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-\author{Francis M. Tyers \\
-  HSL-fakultehta, \\
-  UiT Norgga árktalaš universitehta, \\
-  N-9018 Romsa \\
-  {\tt francis.tyers@uit.no} \\\And
-  Robert Reynolds \\
-  HSL-fakultehta, \\
-  UiT Norgga árktalaš universitehta, \\
-  N-9018 Romsa \\
-  {\tt robert.reynolds@uit.no} \\}
+%\author{Francis M. Tyers \\
+%  HSL-fakultehta, \\
+%  UiT Norgga árktalaš universitehta, \\
+%  N-9018 Romsa \\
+%  {\tt francis.tyers@uit.no} \\\And
+%  Robert Reynolds \\
+%  HSL-fakultehta, \\
+%  UiT Norgga árktalaš universitehta, \\
+%  N-9018 Romsa \\
+%  {\tt robert.reynolds@uit.no} \\}
 
-% \author{Author1 \\
-%   Affiliation / Address line 1 \\
-%   Affiliation / Address line 2 \\
-%   Affiliation / Address line 3 \\
-%   {\tt email@domain.com} \\\And
-%   Author2 \\
-%   Affiliation / Address line 1 \\
-%   Affiliation / Address line 2 \\
-%   Affiliation / Address line 3 \\
-%   {\tt email@domain.com} \\}
-% 
+ \author{Author1 \\
+   Affiliation / Address line 1 \\
+   Affiliation / Address line 2 \\
+   Affiliation / Address line 3 \\
+   {\tt email@domain.com} \\\And
+   Author2 \\
+   Affiliation / Address line 1 \\
+   Affiliation / Address line 2 \\
+   Affiliation / Address line 3 \\
+   {\tt email@domain.com} \\}
+
 \date{2015}
@@
 Table \ref{table:ambiguity} shows the proportional prevalence of each kind of 
-ambiguity. Note that these proportions do not sum to 1, since a given token may 
-exhibit more than one kind of ambiguity. The first column shows the proportion 
+ambiguity. The first column shows the proportion 
 of all tokens in a corpus that have each kind of ambiguity. The second column 
 shows what proportion of ambiguous tokens exhibit each kind of ambiguity.
+Note that these proportions do not sum to 100\%, since a given token may 
+exhibit more than one kind of ambiguity. For example, the wordform 
+\emph{zamkov} has the readings given in \ref{ex:zamkov}.
+
+\ex. \label{ex:zamkov}
+	 \a. \rus{замок}$^1$+N+Msc+Inan+Pl+Gen \label{ex:zamkovA}
+	 \b. \rus{замок}$^2$+N+Msc+Inan+Pl+Gen \label{ex:zamkovB}
+	 \c. \rus{замковый}+A+Msc+Sg+Pred      \label{ex:zamkovC}
+
+The ambiguity between \ref{ex:zamkovA} and \ref{ex:zamkovB} is morphosyntactically congruent, and
+the ambiguity between \ref{ex:zamkovA}/\ref{ex:zamkovB} and \ref{ex:zamkovC} is
+morphosyntactically incongruent, so this wordform would be counted for both categories in
+Table~\ref{table:ambiguity}.
 
@@
 
-Note that the percentages in Table~\ref{table:ambiguity} overlap, i.e. the readings of a given wordform
-can have more than one kind of ambiguity. For example, the wordform \emph{zamkov} has the readings given in \ref{ex:zamkov}.
-
-\ex. \label{ex:zamkov}
-	 \a. \rus{замок}$^1$+N+Msc+Inan+Pl+Gen \label{ex:zamkovA}
-	 \b. \rus{замок}$^2$+N+Msc+Inan+Pl+Gen \label{ex:zamkovB}
-	 \c. \rus{замковый}+A+Msc+Sg+Pred      \label{ex:zamkovC}
-
-The ambiguity between \ref{ex:zamkovA} and \ref{ex:zamkovB} is morphosyntactically congruent, and
-the ambiguity between \ref{ex:zamkovA}/\ref{ex:zamkovB} and \ref{ex:zamkovC} is
-morphosyntactically incongruent, so this wordform would be counted for both categories in
-Table~\ref{table:ambiguity} above.
-
 % we should perhaps mention something about systematic ambiguities and tagging guidelines,
@@
 It is primarily based on Zaliznjak's \emph{Grammatical dictionary of Russian}, 
-including the 2001 version's appendix of proper nouns. It also includes approximately CITE Grishina and Lyashevskaya 2008. \url{http://dict.ruslang.ru/gram.php} .
+including the 2001 version's appendix of proper nouns. It also includes neologisms
+from Grishina and Lyashevskaya's \emph{Grammatical dictionary of new Russian words},
+which is intended to be a supplement to Zaliznjak's dictionary.\footnote{\url{http://dict.ruslang.ru/gram.php}}
 
-Example~\ref{ex:FSToutput} gives some examples of the output.
+Example~\ref{ex:FSToutput} gives some examples of the FST's output.
 

### 2015-04-03 2015-eamt-assim/2015-eamt-assim.tex
Deanonimized
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59615 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-\author{\blankout{Ekaterina Ageeva}\\
-  \blankout{School of Linguistics}\\
-  \blankout{Higher School of Economics}\\
-  \blankout{Moscow, Russia}\\
-  \blankout{{\tt evageeva\_2@edu.hse.ru}}\\[2ex]
-  \blankout{\textbf{Francis M. Tyers}}\\
-  \blankout{HSL-fakultetet}\\ 
-  \blankout{UiT Norgga \'{a}rktala\v{s} universitehta} \\
-  \blankout{9017 Romsa, Norway} \\
-  \blankout{{\tt francis.tyers@uit.no}}
+\author{{Ekaterina Ageeva}\\
+  {School of Linguistics}\\
+  {Higher School of Economics}\\
+  {Moscow, Russia}\\
+  {{\tt evageeva\_2@edu.hse.ru}}\\[2ex]
+  {\textbf{Francis M. Tyers}}\\
+  {HSL-fakultetet}\\ 
+  {UiT Norgga \'{a}rktala\v{s} universitehta} \\
+  {9017 Romsa, Norway} \\
+  {{\tt francis.tyers@uit.no}}
   \And
-  \blankout{Mikel L. Forcada}\\
-  \blankout{Dept. Llenguatges i Sistemes Inform\`{a}tics}\\
+  {Mikel L. Forcada}\\
+  {Dept. Llenguatges i Sistemes Inform\`{a}tics}\\
   \blankout{Universitat d'Alacant, Spain} \\
-  \blankout{{\tt mlf@dlsi.ua.es}}  \\[2ex]
-  \blankout{\textbf{Juan Antonio P\'{e}rez-Ortiz}} \\
-  \blankout{Dept. Llenguatges i Sistemes Inform\`{a}tics}\\
-  \blankout{Universitat d'Alacant, Spain} \\
-  \blankout{{\tt japerez@dlsi.ua.es}}
+  {{\tt mlf@dlsi.ua.es}}  \\[2ex]
+  {\textbf{Juan Antonio P\'{e}rez-Ortiz}} \\
+  {Dept. Llenguatges i Sistemes Inform\`{a}tics}\\
+  {Universitat d'Alacant, Spain} \\
+  {{\tt japerez@dlsi.ua.es}}
 }
@@
 Should I remove?} etc.), checks evaluators' answers, and calculates and reports generalized results.
-This system is integrated into the Appraise MT evaluation platform \citep{federmann12}; the code is open source and is available on GitHub.\footnote{\blankout{\url{https://github.com/Sereni/Appraise}}}
+This system is integrated into the Appraise MT evaluation platform \citep{federmann12}; the code is open source and is available on GitHub.\footnote{{\url{https://github.com/Sereni/Appraise}}}
 
@@
 The version of the toolkit used in this
-experiment may be downloaded from our repository.\footnote{\blankout{\url{https://github.com/Sereni/Appraise/tree/1e9d735faee64d1b97fb343ab111ace6a64509d7}}}
+experiment may be downloaded from our repository.\footnote{{\url{https://github.com/Sereni/Appraise/tree/1e9d735faee64d1b97fb343ab111ace6a64509d7}}}
 

### 2015-04-03 2015-eamt-assim/2015-eamt-assim.tex
Added a list of issues, fixed some
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59616 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-%\newcommand{\comment}[1]{\marginpar{\scriptsize\sf \textcolor{blue}{#1}}}
-\newcommand{\comment}[1]{}
+\newcommand{\comment}[1]{\marginpar{\scriptsize\sf \textcolor{blue}{#1}}}
+%\newcommand{\comment}[1]{}
 \newcommand{\blankout}[1]{\phantom{#1}} % anonymous
@@
 idioms and strong collocations).
+For example, in an English phrase 'Jack ordered <...> and chips', one of the natural answers would be 'fish'.
+Such an answer, however, may be unrelated to the meaning of text, and may be given on the basis of collocation only.
 \item[Reference sentence and source sentence:] By setup, the
@@
 number of gaps relative to sentence length (``gap density''). For the evaluations described in 
-this paper we use gap densities of 10, 20 and 30 percent (Table~\ref{table:percentage}), and the following parts of speech: noun, 
-proper noun, adjective, adverb and lexical verb (as opposed to auxiliary verb).
+this paper we use gap densities of 10, 20 and 30 percent (Table~\ref{table:percentage}), and the following parts of speech: noun (including proper nouns), adjective, adverb and lexical verb (as opposed to auxiliary verb).
 
@@
 Spanish and Russian) who had no command of source language of the pair (Basque
-and Tatar, respectively). Eleven evaluators participated in the Basque--Spanish experiment, and 28 in Tatar-Russian (although not everyone completed the task
+and Tatar, respectively).
+The knowledge was self-reported, and the participants were not asked about any other languages they may know.
+\comment{EA 12: should I elaborate on this?
+Write that Basque is not mutually understandable with anything?
+I am pretty sure we had no one from Tatarstan, but there may have been someone who knows some Bashkir, which might have helped them.
+Also, with Tatar, I think, Russian loanwords help more than anything else.
+} Eleven evaluators participated in the Basque--Spanish experiment, and 28 in Tatar-Russian (although not everyone completed the task
 in full, see discussion).
+The majority of Russian participants were aged 20-25, with university degrees or in the process of obtaining them.
+Although we have not asked the participants about their knowledge of languages other than Tatar and Russian, it is reasonable to assume that most Russian participants knew English to some extent.
+\comment{EA 12: could you tell me more about the Spanish evaluators?
+I would write something similar}
 
@@
 following: read all the available hints and fill each gap with one suitable word, guessing if unsure.
-Participants' answers are recorded and marked correct or incorrect automatically. In
+Participants' answers are recorded and marked correct or incorrect automatically. \comment{EA 11,16: should I try to clarify this? one of the reviewers thought it was manual work}In
 addition, the time taken to fill the gaps in one sentence is recorded.
@@
 However, two or more evaluators gave a different answer, \emph{correr} (\emph{correr riesgos}, 'running risks').
-Based on this data, a native speaker of the target language decides
+Based on this data, an expert, who is native speaker of the target language and who has not participated in the evaluations, decides
 whether the candidate synonym is an acceptable replacement to the answer key in the given
@@
 increase scores where appropriate.
-On average, the scores improved by three percent in all evaluation modes.
+On average, the scores improved by three percentage points in all evaluation modes.
 Candidate synonyms are extracted automatically from the evaluators' responses,  and each individual score is automatically updated according to the synonym list.
@@
 
-%\comment{FMT: We should probably integrate the other tables, and refer to them using labels.} 
 The results are presented in this section. Tables \ref{table:res-eus-spa} and \ref{table:res-tat-rus} 
@@
 
-% participants and their suffering
 It is worth noting that many participants reported feeling frustrated in
-the course of evaluations, especially while working on the no-hint tasks. 6 out of 49
+the course of evaluations, especially while working on the no-hint tasks.
+The latter required suggesting the words with very little context, which led some of the participants to giving random words for answers, or leaving the space blank.
+6 out of 49
 participants quit the experiment before completing it.
 Considering the importance of receiving the full set of evaluations, we must address the issue of participant motivation in the upcoming experiments.
+It may be beneficial to offer monetary compensation for the evaluators' efforts (in our case, they were volunteers).
 
@@
 After obtaining the results we calculated Krippendorff's alpha~\citep{krippendorff70} measure to represent annotator agreement, shown in Tables~\ref{table:alpha-eus-spa} and~\ref{table:alpha-tat-rus}.
-\begin{table*}
-\centering
-  \begin{tabular}{|l|r|r|r|r|}
-  \hline
- \multirow{2}{*}{\textbf{Density}} & \multicolumn{4}{|c|}{Evaluation mode} \\\cline{2-5}
-                                            & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
-\hline
-10\%&0.496&0.517&0.400&0.124\\
-20\%&0.714&0.700&0.358&0.275\\
-30\%&0.559&0.430&0.406&0.300\\
 
-\hline
-\end{tabular}
-\caption {Basque--Spanish: Krippendorff Alpha measure} 
-\label{table:alpha-eus-spa}
-\end{table*}
-
-\begin{table*}
-\centering
-  \begin{tabular}{|l|r|r|r|r|}
-  \hline
- \multirow{2}{*}{\textbf{Density}} & \multicolumn{4}{|c|}{Evaluation mode} \\\cline{2-5}
-                                            & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
-\hline
-10\%&0.598&0.459 &0.711 &0.517 \\
-20\%&0.740&0.667 &0.473 &0.261 \\
-30\%&0.534&0.581 &0.411 &0.412 \\
-\hline
-\end{tabular}
-\caption {Tatar--Russian: Krippendorff Alpha measure} 
-\label{table:alpha-tat-rus}
-\end{table*}
 We selected this measure because of its compatibility with more than two annotators per task and the missing data (not all the gaps were evaluated).
@@
 
+% actually, I think the following paragraph may be about kappa, will check once I get access to landis77
+%The interpretation of Krippendorff's alpha varies depending on the application.
+One of the general guidelines suggested by \citet{landis77} is as follows (cited after \citet{carletta97}: "[values of alpha] < 0 ``poor'' agreement, 0 to .2 ``slight'', .21 to .40 ``fair'', .41 to .60 ``moderate'', .61 - .80 ``substantial'', and .81 to 1 ``near perfect''"
+
 As the MT and MT+source hints are introduced, the agreement between annotators increases (measures closer to 1): the annotators are more consistently correct or incorrect in each given sentence.
 The agreement measure for the same sentences without hints is closer to zero, which contributes to the reliability of our methodology.
-We note the outlier score in Tatar--Russian 10\% source tasks, which has the most contribution from the news texts (agreement of 0.917).
-This set of sentences contains many loan words, which have similar form in Tatar and Russian (e.g.
+We note the outlier score in Tatar--Russian 10\% source tasks, which has the most contribution from the news texts (the agreement for this particular category reached 0.917).
+\comment{EA: this Alpha is from a table we do not report here, I added a couple of words to clarify}This set of sentences contains many loan words, which have similar form in Tatar and Russian (e.g.
 president, minister, championship), and are understood by Russian speakers.
@@
 \hline
-10\%&63.76 \(\pm\) 41&57.02 \(\pm\) 42&53.52 \(\pm\) 43&45.96 \(\pm\) 41\\
-20\%&59.77 \(\pm\) 33&64.93 \(\pm\) 31&45.60 \(\pm\) 31&38.56 \(\pm\) 32\\
-30\%&55.83 \(\pm\) 26&58.77 \(\pm\) 28&40.00 \(\pm\) 28&35.45 \(\pm\) 30\\
+10\%&57.02 \(\pm\) 42&63.76 \(\pm\) 41&53.52 \(\pm\) 43&45.96 \(\pm\) 41\\
+20\%&64.93 \(\pm\) 31&59.77 \(\pm\) 33&45.60 \(\pm\) 31&38.56 \(\pm\) 32\\
+30\%&58.77 \(\pm\) 28&55.83 \(\pm\) 26&40.00 \(\pm\) 28&35.45 \(\pm\) 30\\
     \hline
@@
   \label{table:res-tat-rus}
+\end{table*}
+
+\begin{table*}
+\centering
+  \begin{tabular}{|l|r|r|r|r|}
+  \hline
+ \multirow{2}{*}{\textbf{Density}} & \multicolumn{4}{|c|}{Evaluation mode} \\\cline{2-5}
+                                            & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
+\hline
+10\%&0.496&0.517&0.400&0.124\\
+20\%&0.714&0.700&0.358&0.275\\
+30\%&0.559&0.430&0.406&0.300\\
+
+\hline
+\end{tabular}
+\caption {Basque--Spanish: Krippendorff Alpha measure} 
+\label{table:alpha-eus-spa}
+\end{table*}
+
+\begin{table*}
+\centering
+  \begin{tabular}{|l|r|r|r|r|}
+  \hline
+ \multirow{2}{*}{\textbf{Density}} & \multicolumn{4}{|c|}{Evaluation mode} \\\cline{2-5}
+                                            & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
+\hline
+10\%&0.598&0.459 &0.711 &0.517 \\
+20\%&0.740&0.667 &0.473 &0.261 \\
+30\%&0.534&0.581 &0.411 &0.412 \\
+\hline
+\end{tabular}
+\caption {Tatar--Russian: Krippendorff Alpha measure} 
+\label{table:alpha-tat-rus}
 \end{table*}

### 2015-04-04 2015-eamt-assim/2015-eamt-assim.tex
More corrections by reviewers' comments
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59620 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-\newcommand{\comment}[1]{\marginpar{\scriptsize\sf \textcolor{blue}{#1}}}
-%\newcommand{\comment}[1]{}
-\newcommand{\blankout}[1]{\phantom{#1}} % anonymous
-%\newcommand{\blankout}[1]{#1} % non-anonymous
+%\newcommand{\comment}[1]{\marginpar{\scriptsize\sf \textcolor{blue}{#1}}}
+\newcommand{\comment}[1]{}
+%\newcommand{\blankout}[1]{\phantom{#1}} % anonymous
+\newcommand{\blankout}[1]{#1} % non-anonymous
 
@@
 Taking into consideration the above discussion of 10\% MT+Source tasks, we observe similar results across the three categories.
-Note that, in casual and news texts, the source sentences paired with MT significantly improve participants' performance compared to MT-only task mode.
+Note that the source sentences paired with MT improve participants' performance in casual texts, compared to MT-only task mode.
 This may be due to the fact that many words are borrowed from Russian into Tatar, and are in fact understood by Russian speakers.
@@
 
-\phantom{This work has been partly funded by the Spanish Ministerio de Econom{\'i}a y Competitividad through project TIN2012-32615 and by the Google Summer of Code 2014 programme.
-We would like to thank the volunteers who participated in the evaluations.}
+This work has been partly funded by the Spanish Ministerio de Econom{\'i}a y Competitividad through project TIN2012-32615 and by the Google Summer of Code 2014 programme.
+We would like to thank the volunteers who participated in the evaluations.
 

### 2015-04-05 2015-eamt-assim/2015-eamt-assim.tex
Fixed some more points from the reviews
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59621 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 a total of 36;
-\item  each participant evaluates 9 sentences in each of the four modes (see section~\ref{sec:methodology});
-\item  all sentences of the set are evaluated with 10, 20 and 30\% of words removed;
+\item  these sentences are divided into 4 groups of 9, one for each evaluation mode (see section~\ref{sec:methodology});
+\item in total, all sentences of the set are evaluated with 10, 20 and 30\% of words removed;
 \item each participant may encounter a given sentence in only one of the percentage variations;

### 2015-04-11 2015-eamt-assim/2015-eamt-assim.tex
More fixes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59684 72bbbca6-d526-0410-a7d9-f06f51895060

@@
   and analyse results, an open-source task management system is
-  introduced. The evaluation results show that the gap-filling task
-  reflects users' understanding of text, and may be used to measure MT
+  introduced. The evaluation results show that the gap-filling task may be suitable for measuring MT
   quality for assimilation purposes.
@@
 While these authors ask informants to fill gaps in MT output, \citet{trosterud12} and \citet{oregan13} propose to fill gaps in the reference (human) translation.
+A designated number of keywords is removed from the human-translated sentences.
+The evaluators are then asked to fill the gaps with suitable words.
 The gap-filling task models how well users comprehend the key points of the text, as it is roughly equivalent with answering questions.
@@
 For the latter pair, the evaluation served as a quality check in the period of active development during the Google Summer of Code 2014 programme.
-In addition to evaluating, we explore new aspects of the experiment: the correlation between evaluators' scores, and the influence of the texts' linguistic domain.
+In addition to evaluating, we explore the previously unconsidered aspects of the experiment: the correlation between evaluators' scores, and the effects of the texts' linguistic domain and the number of gaps in a sentence.
 To facilitate the evaluation, we introduce an automated system which creates task sets from parallel corpora given a range of parameters (number of gaps in a sentence, hint type, gap filler,\comment{MLF: gap filler or gap marker?} \comment{EA: well, initially we had questions with lemmas in place of gaps, and also multiple choice questions where answer options got generated automatically.
@@
 Finally, section~\ref{sec:conclusion} draws some conclusions.
+This paper is concerned primarily with assimilation evaluation; for a deeper discussion on evaluation see e.g.
+\citep{koehn2009}.
 
@@
 
-As the MT and MT+source hints are introduced, the agreement between annotators increases (measures closer to 1): the annotators are more consistently correct or incorrect in each given sentence.
+In general, the level of annotator agreement is relatively high.
+As the MT and MT+source hints are introduced, the agreement increases (measures closer to 1): the annotators are more consistently correct or incorrect in each given sentence.
 The agreement measure for the same sentences without hints is closer to zero, which contributes to the reliability of our methodology.
-We note the outlier score in Tatar--Russian 10\% source tasks, which has the most contribution from the news texts (the agreement for this particular category reached 0.917).
+We note the outlier score in Tatar--Russian 10\% source tasks, which has the most contribution from the news texts.
 \comment{EA: this Alpha is from a table we do not report here, I added a couple of words to clarify}This set of sentences contains many loan words, which have similar form in Tatar and Russian (e.g.
@@
 \hline
-10\%&62.19 \(\pm\) 32&58.04 \(\pm\) 28&39.90 \(\pm\) 39&48.79 \(\pm\) 40\\
-20\%&64.81 \(\pm\) 30&70.18 \(\pm\) 27&31.37 \(\pm\) 28&30.91 \(\pm\) 30\\
-30\%&48.13 \(\pm\) 26&39.55 \(\pm\) 24&25.94 \(\pm\) 20&18.07 \(\pm\) 18\\
+10\%&62 \(\pm\) 32&58 \(\pm\) 28&40 \(\pm\) 39&49 \(\pm\) 40\\
+20\%&65 \(\pm\) 30&70 \(\pm\) 27&31 \(\pm\) 28&31 \(\pm\) 30\\
+30\%&48 \(\pm\) 26&40 \(\pm\) 24&26 \(\pm\) 20&18 \(\pm\) 18\\
 \hline
@@
 \hline
-10\%&57.02 \(\pm\) 42&63.76 \(\pm\) 41&53.52 \(\pm\) 43&45.96 \(\pm\) 41\\
-20\%&64.93 \(\pm\) 31&59.77 \(\pm\) 33&45.60 \(\pm\) 31&38.56 \(\pm\) 32\\
-30\%&58.77 \(\pm\) 28&55.83 \(\pm\) 26&40.00 \(\pm\) 28&35.45 \(\pm\) 30\\
+10\%&57 \(\pm\) 42&64 \(\pm\) 41&54 \(\pm\) 43&46 \(\pm\) 41\\
+20\%&65 \(\pm\) 31&60 \(\pm\) 33&46 \(\pm\) 31&39 \(\pm\) 32\\
+30\%&59 \(\pm\) 28&56 \(\pm\) 26&40 \(\pm\) 28&35 \(\pm\) 30\\
     \hline
@@
 \hline
-10\%&63.64 \(\pm\) 45&63.64 \(\pm\) 43&62.12 \(\pm\) 46&53.03 \(\pm\) 44\\
-20\%&73.33 \(\pm\) 32&63.33 \(\pm\) 36&41.11 \(\pm\) 28&37.78 \(\pm\) 31\\
-30\%&70.28 \(\pm\) 31&59.58 \(\pm\) 24&39.44 \(\pm\) 27&37.71 \(\pm\) 33\\
+10\%&64 \(\pm\) 45&64 \(\pm\) 43&62 \(\pm\) 46&53 \(\pm\) 44\\
+20\%&73 \(\pm\) 32&63 \(\pm\) 36&41 \(\pm\) 28&38 \(\pm\) 31\\
+30\%&70 \(\pm\) 31&60 \(\pm\) 24&39 \(\pm\) 27&38 \(\pm\) 33\\
 \hline
@@
 \hline
-10\%&53.02 \(\pm\) 40&68.21 \(\pm\) 35&38.77 \(\pm\) 38&32.95 \(\pm\) 35\\
-20\%&60.57 \(\pm\) 25&65.76 \(\pm\) 24&50.28 \(\pm\) 34&48.31 \(\pm\) 34\\
-30\%&50.06 \(\pm\) 26&48.30 \(\pm\) 29&40.05 \(\pm\) 27&34.15 \(\pm\) 29\\
+10\%&53 \(\pm\) 40&68 \(\pm\) 35&39 \(\pm\) 38&33 \(\pm\) 35\\
+20\%&61 \(\pm\) 25&66 \(\pm\) 24&50 \(\pm\) 34&48 \(\pm\) 34\\
+30\%&50 \(\pm\) 26&48 \(\pm\) 29&40 \(\pm\) 27&34 \(\pm\) 29\\
 \hline
@@
 \hline
-10\%&53.33 \(\pm\) 38&59.77 \(\pm\) 44&57.47 \(\pm\) 42&49.44 \(\pm\) 39\\
-20\%&59.33 \(\pm\) 34&48.56 \(\pm\) 35&46.53 \(\pm\) 32&28.97 \(\pm\) 29\\
-30\%&58.14 \(\pm\) 22&61.49 \(\pm\) 22&40.51 \(\pm\) 30&34.81 \(\pm\) 27\\
+10\%&53 \(\pm\) 38&60 \(\pm\) 44&57 \(\pm\) 42&49 \(\pm\) 39\\
+20\%&59 \(\pm\) 34&49 \(\pm\) 35&47 \(\pm\) 32&29 \(\pm\) 29\\
+30\%&58 \(\pm\) 22&61 \(\pm\) 22&41 \(\pm\) 30&35 \(\pm\) 27\\
 \hline
@@
 \end{table*}
-\subsection{Experiments with other language pairs}
-In addition to the mentioned language pairs, some preliminary experiments were carried out for the English--Kazakh pair.
-It remains however questionable whether we may draw reliable conclusions from this part of the experiment.
-By design, the evaluation tasks are distributed evenly among the participants, in such a manner that each sentence is evaluated in all modes or gap densities by several people.
-The English--Kazakh evaluations have been set up for 11 participants, however, only 8 participants attempted their tasks, and 6 of them completed it in full.
-As a result, the evaluation became unbalanced: out of 4 sets of sentences, three have been evaluated by three, two and one participant, respectively, and the fourth set has been partially completed by only one participant.
-A total of 40 (about 10\%) gaps have not been filled by any of the participants.
-Therefore, we do not report the English--Kazakh evaluation results here.
 
@@
 
+As a consideration for future work, it may be beneficial to compare the results of evaluation by gap-filling method with the traditional evaluation metrics, as well as with human evaluation.
+
 \section*{Acknowledgments}

### 2015-04-12 2015-eamt-melsa/2015-eamt-melsa.tex
Changes made or indicated
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59722 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage[small,bf]{caption}
+\usepackage[backgroundcolor=lightgray]{todonotes}
+\newcommand{\mytodo}[1]{\todo[inline]{#1}}
+\newcommand{\todomlf}[1]{\renewcommand{\baselinestretch}{0.55}\todo{\scriptsize\sf MLF: #1}\renewcommand{\baselinestretch}{1.0}}
+
 \usepackage{microtype} % a.k.a. conference paper magic
@@
 \author{
-\phantom{Francis M.\ Tyers} \\
-\phantom{HSL-fakultehta,} \\
-\phantom{UiT Norgga \'{a}rktala\v{s} universitehta,} \\
-\phantom{N-9018 Romsa} \\
+Francis M.\ Tyers \\
+HSL-fakultehta, \\
+UiT Norgga \'{a}rktala\v{s} universitehta, \\
+N-9018 Romsa \\
  \And
-\phantom{Felipe S\'anchez-Mart\'inez} \\
-\phantom{Dept.\ Lleng.\ i Sist.\ Inform.,} \\
-\phantom{Universitat d'Alacant,}\\
-\phantom{E-03071 Alacant}\\
+Felipe S\'anchez-Mart\'inez \\
+Dept.\ Lleng.\ i Sist.\ Inform., \\
+Universitat d'Alacant,\\
+E-03071 Alacant\\
  \And
-\phantom{Mikel L.\ Forcada} \\
-\phantom{Dept.\ Lleng.\ i Sist.\ Inform.,} \\
-\phantom{Universitat d'Alacant,}\\
-\phantom{E-03071 Alacant}\\
+Mikel L.\ Forcada \\
+Dept.\ Lleng.\ i Sist.\ Inform., \\
+Universitat d'Alacant,\\
+E-03071 Alacant\\
 }
@@
 \section{Introduction}
-
+\todomlf{R1 has a general comment: ``watch out for long sentences''}
 Corpus-based MT has been the primary research direction in the field
@@
 refer to what is actually target word selection in machine
-translation; they used a parser to identify some syntactic relations
+translation; they used a parser to identify syntactic relations
 such as subject--object or subject--verb.  After generating all the
@@
 \section{Method}
-
+\todomlf{R3 says: ``The technique clearly resembles the "Deciphering Foreign Language," (S. Ravi and K. Knight) or “EM Decipherment for Large Vocabularies “
+(M. Nuhn, and H. Ney. EM) approaches, although these techniques were not meant for learning the disambiguation but for learning the translation dictionaries.
+The proposed method only executes one EM iteration in contrast to these works.
+It would be interesting to highlight differences and resemblances with these deciphering approaches.''.
+Check and cite?
+Also, single iteration WTF?
+}
 %\todo[inline]{FSM: In the experiments a development corpus is
@@
 Let the probability of a word $t$ being the translation of a
-word $s$ in an SL context $c$ be $p_s(t|c)$. In principle, this value
+word $s$ in a source-language context $c$ be $p_s(t|c)$. In principle, this value
 could be estimated directly from the available corpora for every
@@
 form $h_k^{s}(t, c)$, where $t$
-is a translation, and $c$ is an SL context. One difference is that \cite{berger1996} take $s$, $t$
+is a translation, and $c$ is a source-language (SL) context.\todomlf{Defined SL which had not been defined} One difference is that \cite{berger1996} take $s$, $t$
 and $c$ to be based on word forms, whereas in our method they are based
 on lemma forms.
-A feature where the Spanish word \emph{pez} (`fish' as a living animal) is seen as the
+An example would be the following feature where the Spanish word \emph{pez} (`fish' as a living animal) is seen as the
 translation of \emph{arrain} (`fish') in the context \emph{arrain
@@
 \begin{equation}
-h_k^{\mathrm{arrain}}(t, c) = \left \{ \begin{matrix}
+h_\mathrm{+handi}^{\mathrm{arrain}}(t, c) = \left \{ \begin{matrix}
                  ~~1 & \mbox{if } 
@@
 This feature considers a context of zero words to the left of the
-problem word and one word to the right of it.
+problem word and one word (+ \emph{handi}) to the right of it.\todomlf{Removed \(k\) and used a more informative subscript}
 
@@
 
-As a result of training, each feature in the classifier is assigned a
-weight $\lambda_k^s$. Combining these weights of active features
+As a result of training, each of the \(n_F\) features \(h^s_k(t,c)\) in the classifier is assigned a
+weight $\lambda_k^s$. \todomlf{Reworded to define \(n_F\)} Combining these weights of active features
 as in equation~(\ref{eq:max-ent-sum}) yields the probability of a
@@
 \end{equation}
+\todomlf{R3 complains: `` Probability notation is confusing.
+I would suggest the use of \(p(t | c,s)\) instead of \(p_s(t|c)\), that will simplify notation.
+For instance instead of \(Z^s (c)\) we would have \(Z(c,s)\).''.
+Ignore or take onboard?}
 
@@
 $g = (t_1, t_2, \ldots, t_{|S|})$ is a sequence of lexical-selection
-choices of those SL words, where \(t_i \in
-T_s(s_i)\). % and \(T_s(\cdot)\)
+choices of those SL words, where \(t_i\) is an element of \(T_s(s_i)\), the set of possible translations of \(s_i\).\todomlf{R1 asked for this, even if redundant, and says: ``can one source word be translated as multiple words?
+What about multi-word expressions that cannot be translated word-by-word or source words that do not have a TL equivalent'', etc.} % and \(T_s(\cdot)\)
 %is the set of all
@@
 schematically. %%%\todo{Can we try to bring the figure closer?}
-
+\todomlf{R3 says: ``it would have been nicer to see an example of one sentence going through the pipeline''. Can we add this in the text?}
 %Taking our previous example for the parallel corpus, Table~\ref{table:ex-monocorp} 
@@
 the fractional count for disambiguation path \(g_i\), $p(g_i|S)$ after
-suitable normalisation:
+suitable normalisation:\todomlf{R2 says ``Nevertheless, I've got problems to understand (4), in particular the paths in post-lexsel step.
+Some better explanation could be offered maybe by saying that (4) evaluates all possible solutions in the language model.''.
+Don't know what to do here.
+}
 \begin{equation}
@@
 
+\todomlf{R1 complains: ``the actual maximum entropy method itself is described very briefly (3 sentences at the end of Section 2)''}
 \section{Experimental setting}
@@
 --- although for all pairs there is a parallel corpus available
-for evaluation (see Section \ref{ss:measures}). 
+for evaluation (see Section~\ref{ss:measures}). 
 % Table~\ref{table:lang-pair-stats} 
@@
 %  and French. It has been developed part-time over a number of years.
+\todomlf{R1 says ``whole section is rather length especially the revision numbers and URLs of the specific translations systems need not be mentioned in my opinion.
+The space could rather be used to elaborate in the description of the method in Sec 2.''}
 
@@
 \item[English--Spanish:] The English--Spanish pair was developed from
-  a combination of the English--Catalan and Spanish--Catalan pairs
-  over a period of around four months. The pair
+  a combination of the English--Catalan and Spanish--Catalan pairs, and
+%  over a period of around four months. The pair
 %has been used in other
@@
   version used in this paper is SVN revision
-  41387.\footnote{\url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-en-es}}
+  41387.\footnote{\url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-en-es}}\todomlf{4-month mention removed in response to R1}
 \end{description}
@@
    error rate. The source sentence $S = (s_1, s_2, \ldots, s_{|S|})$
-   has two ambiguous words, \emph{estaci\'{o}} and \emph{llarg}. There is one difference
+   has two ambiguous words, \emph{estaci\'{o}} and \emph{llarg} (\(\mathrm{amb}(s_i)=1\), eq.~(\protect\ref{eq:amb})).
+There is one difference (\(\mathrm{diff}(T_r(s_i),T_t(s_i))=1\), eq.~(\protect\ref{eq:diff}))
    between the reference set $T_r(s_i)$ and the test set $T_t(s_i)$ of
@@
 single sentence, we define the lexical selection error rate (LER) of
-that sentence as
+that sentence as\todomlf{R3 says: ``The authors use LER, but F-score measures seem more appropriate to estimate the performance of the disambiguation module.
+Moreover, the constraint of the disambiguation module to generate only one word when evaluating seems extreme.'' He don't get it.}
 \begin{equation}
@@
 of reference translations $T_r(s_i)$ and the set of translations from
-the lexical selection module $T_t(s_i)$ is empty.  Recall that,
+the lexical selection module $T_t(s_i)$ is empty.  Recall that,\todomlf{R1 says ``this has already been mentioned before''. Don't know what to do}
 although $T_t(s_i)$ returns a set, this set will be a singleton, as
@@
 the first one will be selected.\footnote{In practice this does not happen
-as each ambiguous word has a \emph{default} translation.}
+as each ambiguous word has a \emph{default} translation.}\todomlf{R1 complains: ``why mention if it doesn't happen in practice? Relevance?'' What should we do?}
 
@@
 The table in Figure~\ref{fig:ler-example} gives an
-overview of the inputs. In the description it is assumed that the
+overview of the inputs.\todomlf{Caption of figure extended in response to R3} In the description it is assumed that the
 reference translation has been annotated by hand. However, hand
@@
 description of how the reference was built is given in
-section~\ref{sec:eval-corpus}.
+Section~\ref{sec:eval-corpus}.
 
@@
 translation quality in a real system. For this task, we use the
-widely-used {\sc bleu} metric (see
-Section~\ref{sec:eval-metrics}). This is not ideal for evaluating the
+widely-used {\sc bleu} metric \citep{papineni02}.\todomlf{BLEU reference advanced, and wrong section link removed} This is not ideal for evaluating the
 task of a lexical selection module as the performance of the module
@@
 % --- may also lead to the same problem.  
-The {\sc bleu} metric \citep{papineni02} is
+The {\sc bleu} metric  is
 included only as it is commonly used to evaluate MT systems.
@@
   input corpora, and how many sentences were left after processing for
-  testing, training and development. \todo{Why do ambiguity numbers
-    change so much from the paper?} Table~\ref{table:test-corp} gives
+  testing, training and development. %\todo{Why do ambiguity numbers
+%    change so much from the paper?} 
+Table~\ref{table:test-corp} gives
   information about the test corpora. 
@@
   \url{http://downloads.sourceforge.net/project/apertium/apertium-lex-tools/apertium-lex-tools-0.1.0.tar.gz}.}
-to calculate the feature weights.  After learning the feature sets and
+to calculate the feature weights.  
+\todomlf{R3 says: ``The proposed maximum entropy or log-linear model is trained in a surprising way.
+In the early days of maximum entropy models, those were trained using the GIS algorithm.
+However, it has been proved in the literature that GIS has clear disadvantages with respect to other batch gradient descent algorithms.
+For instance, it does not allow for the use of regularization [...], and it is extremely slow to converge, which yields worse models than other methods.
+The smoothing/regularization of the model is also poor since it is simply feature pruning instead of regularization.''}
+After learning the feature sets and
 weights, we compute the evaluation measures described in
-section~\ref{ss:measures}. There is an option to remove events 
+Section~\ref{ss:measures}. There is an option to remove events 
 $(s, t, c)$ which occur
@@
  the new approach with respect to the \emph{default behaviour} (the
- linguist-chosen defaults,\footnote{The high error rate
+ linguist-chosen defaults),\todomlf{R1 says the footnote is important and should be incorporated to the text; what should we do?}\footnote{The high error rate
    for the Breton--French pair may be as a result of having the
@@
 expensively obtained by scoring an exponential number of lexical
-selections for each sentence using the target-language model online.
+selections for each sentence using the target-language model online.\todomlf{R2 says: ``An error analysis would also help to understand the proposal.'' We don't have time now I think.}
 
@@
 
-\section*{Acknowledgements}
-
-\emph{Removed for review}
+\paragraph{Acknowledgements:} We acknowledge support from the Spanish Ministry of Industry and Competitiveness through project Ayutra (TIC2012-32615) and from the European Commission through project Abu-Matran (FP7-PEOPLE-2012-IAPP, ref.\ 324414) and thank all three anonymous referees for very useful comments on the paper.\todomlf{Fran: add your acknowledgements}
 

### 2015-04-12 2015-eamt-melsa/2015-eamt-melsa.tex
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59727 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \newcommand{\todomlf}[1]{\renewcommand{\baselinestretch}{0.55}\todo{\scriptsize\sf MLF: #1}\renewcommand{\baselinestretch}{1.0}}
+\newcommand{\todofsm}[1]{\renewcommand{\baselinestretch}{0.55}\todo{\scriptsize\sf FSM: #1}\renewcommand{\baselinestretch}{1.0}}
 
@@
 \todomlf{R1 has a general comment: ``watch out for long sentences''}
-Corpus-based MT has been the primary research direction in the field
-of machine translation in recent years. However, rule-based MT (RBMT)
-systems are still being developed, and there are many successful
-commercial and non-commercial systems. One reason for the continued
-development of RBMT systems is that in order to be successful,
-corpus-based MT requires parallel corpora in the order of tens of
-millions of words. Although for some language pairs these exist, they
-only exist for a fraction of the world's languages.
+Corpus-based machine translation (MT) has been the primary research
+direction in the field of MT in recent years. However, rule-based MT
+(RBMT) systems are still being developed, and there are many
+successful commercial and non-commercial systems. One reason for the
+continued development of RBMT systems is that in order to be
+successful, corpus-based MT requires parallel corpora in the order of
+tens of millions of words. Although for some language pairs these
+exist, they only exist for a fraction of the world's languages.
 
@@
 component it is necessary to make choices regarding words in the
-source language which may have more than one translation in the target
-language.
-
-\emph{Lexical selection} is the task of choosing, for a given
-source-language word, the most adequate translation in the target
-language among a known set of alternatives.  The task is related to
-the task of word-sense disambiguation \citep{ide98a}. However, it is
-different to word-sense disambiguation in that lexical selection is a
-bilingual problem, not a monolingual problem: its aim is to find the
-most adequate translation, not the most adequate sense. Thus, it is
-not necessary to choose between a series of fine-grained senses if all
-these senses result in the same final translation; however, it may
-sometimes be necessary to choose a different translation for the same
-sense, for example in a collocation.
+source language (SL) which may have more than one translation in the
+target language (TL).
+
+\emph{Lexical selection} is the task of choosing, for a given SL word,
+the most adequate translation in the TL among a known set of
+alternatives.  The task is related to the task of word-sense
+disambiguation \citep{ide98a}. However, it is different to word-sense
+disambiguation in that lexical selection is a bilingual problem, not a
+monolingual problem: its aim is to find the most adequate translation,
+not the most adequate sense. Thus, it is not necessary to choose
+between a series of fine-grained senses if all these senses result in
+the same final translation; however, it may sometimes be necessary to
+choose a different translation for the same sense, for example in a
+collocation.
 % It could also be the case that a single sense could have more than
@@
 \cite{dagan1994word} used the term \emph{word sense disambiguation} to
-refer to what is actually target word selection in machine
-translation; they used a parser to identify syntactic relations
-such as subject--object or subject--verb.  After generating all the
-possible translations for a given input sentence using an ambiguous
-bilingual dictionary, they extract the syntactic tuples from the
-target language and count the frequency in a previously-trained
-target-language model of tuples.  They use maximum-likelihood
-estimation to calculate the probability that a given target language
-tuple is the translation of a given source language tuple, with an
-automatically determined confidence threshold.
+refer to what is actually target word selection\todofsm{Should we call
+  it lexical selection instead of target word selection?} in MT; they
+used a parser to identify syntactic relations such as subject--object
+or subject--verb.  After generating all the possible translations for
+a given input sentence using an ambiguous bilingual dictionary, they
+extract the syntactic tuples from the TL and count the frequency in a
+previously-trained TL model of tuples.  They use maximum-likelihood
+estimation to calculate the probability that a given TL tuple is the
+translation of a given SL tuple, with an automatically determined
+confidence threshold.
 
@@
 classifiers on the specific problem of lexical selection in IBM-style
-word-based statistical machine translation. Other authors
-\citep{melero07a} have used target-language models to
-rank %@@@ carbonell06
+word-based statistical MT. Other authors \citep{melero07a} have used
+TL models to rank %@@@ carbonell06
 the translations resulting from all possible combinations of lexical
@@
 statistical % \todo{METIS too?}
-machine translation \citep{koehn2009statistical}, lexical selection is
-taken care of by a combination of the translation model and the
-language model. The translation model provides probabilities of
-translation between words or word sequences (often referred to as
-\emph{phrases}) in the source and target language. The target-language
-model provides probabilities of word sequences in the target language.
+MT \citep{koehn2009statistical}, lexical selection is taken care of by
+a combination of the translation model and the language model. The
+translation model provides probabilities of translation between words
+or word sequences (often referred to as \emph{phrases}) in the source
+and target language. The TL model provides probabilities of word
+sequences in the TL.
 % As regards rule-based machine translation, \cite{sanchez07a}
@@
 \cite{marechek10} trained a maximum-entropy lexical selector for their
-dependency-grammar-based transfer system TectoMT using a bilingual corpus.
-More recently, \cite{tyers12a} presented a method of lexical selection for rule-based
-machine translation based on rules which select or remove translations
+dependency-grammar-based transfer system TectoMT using a bilingual
+corpus.  More recently, \cite{tyers12a} presented a method of lexical
+selection for RBMT based on rules which select or remove translations
 in fixed-length contexts, along with a training method for learning
@@
 \section{Method}
-\todomlf{R3 says: ``The technique clearly resembles the "Deciphering Foreign Language," (S. Ravi and K. Knight) or “EM Decipherment for Large Vocabularies “
-(M. Nuhn, and H. Ney. EM) approaches, although these techniques were not meant for learning the disambiguation but for learning the translation dictionaries.
-The proposed method only executes one EM iteration in contrast to these works.
-It would be interesting to highlight differences and resemblances with these deciphering approaches.''.
-Check and cite?
-Also, single iteration WTF?
-}
+\todomlf{R3 says: ``The technique clearly resembles the "Deciphering
+  Foreign Language," (S. Ravi and K. Knight) or “EM Decipherment for
+  Large Vocabularies “ (M. Nuhn, and H. Ney. EM) approaches, although
+  these techniques were not meant for learning the disambiguation but
+  for learning the translation dictionaries.  The proposed method only
+  executes one EM iteration in contrast to these works. It would be
+  interesting to highlight differences and resemblances with these
+  deciphering approaches.''. Check and cite? Also, single iteration
+  WTF?}
 %\todo[inline]{FSM: In the experiments a development corpus is
@@
 %\comment{Improve this} 
-Lexical selection in this paper considers for
-each word a simple source-language context made up of neighbouring
-lemma+part-of-speech combinations. Contexts considered include up to
-two words to the left and up to two words to the right of the word to
-be translated. 
+Lexical selection in this paper considers for each word a simple SL
+context made up of neighbouring lemma+part-of-speech
+combinations. Contexts considered include up to two words to the left
+and up to two words to the right of the word to be translated.
 %\comment{FSM: Later, a development corpus is
@@
 
-Let the probability of a word $t$ being the translation of a
-word $s$ in a source-language context $c$ be $p_s(t|c)$. In principle, this value
-could be estimated directly from the available corpora for every
-combination of $(s, t, c)$.  This would however present two questions:
-(1) how should the relevant contexts be chosen? and (2) what should be
-done when $(s, t, c)$ is not found in the corpus?
-A maximum-entropy model answers both of these questions. It allows the
-contexts that we consider to be linguistically interesting to be
-defined \emph{a priori} and then integrate these seamlessly into a
-probabilistic model \citep{Manning99b}.  In answer to the second
-question, a maximum-entropy model assumes nothing about what is not
-present in the training data. That is, if there is no information in
-the training data, then it assumes that all outcomes ---that is, all
-possible translations--- are equally likely.  As previously mentioned,
-the principle of maximum entropy has been applied to the problem of
-lexical selection before; in particular, \cite{berger1996} cast the
-problem of lexical selection in statistical MT as a classification
-problem. They learn a separate maximum-entropy classifier for each SL
-word form,
+Let the probability of a word $t$ being the translation of a word $s$
+in a SL context $c$ be $p_s(t|c)$. In principle, this value could be
+estimated directly from the available corpora for every combination of
+$(s, t, c)$.  This would however present two questions: (1) how should
+the relevant contexts be chosen? and (2) what should be done when $(s,
+t, c)$ is not found in the corpus?  A maximum-entropy model answers
+both of these questions. It allows the contexts that we consider to be
+linguistically interesting to be defined \emph{a priori} and then
+integrate these seamlessly into a probabilistic model
+\citep{Manning99b}.  In answer to the second question, a
+maximum-entropy model maximise the entropy subject to match the
+expected counts of the designed features with those found in the
+training data. \todofsm{Solved R3: ``In answer to the second question, a
+  maximum-entropy model assumes nothing about what is not present in
+  the training data. That is, if there is no information in the
+  training data, then it assumes that all outcomes —that is, all
+  possible translations— are equally likely'' can be applied to a
+  uniform distribution for a discrete variable. I think the authors
+  are trying to restate the main property of the maximum entropy
+  model, i.e.  maximum entropy model maximize the entropy (which is
+  the uniformness of the distribution) subject to match the expected
+  counts of the designed features with those found in the training
+  data.  }
+%assumes nothing about what is not present in the
+%training data. 
+That is, if there is no information in the training data, then it
+assumes that all outcomes ---that is, all possible translations--- are
+equally likely.  As previously mentioned, the principle of maximum
+entropy has been applied to the problem of lexical selection before;
+in particular, \cite{berger1996} cast the problem of lexical selection
+in statistical MT as a classification problem. They learn a separate
+maximum-entropy classifier for each SL word form,
 %\comment{[:we do too, integrate this later]} 
-using SL context to
-distinguish between possible translations. These classifiers are then
-incorporated into the translation model of their word-based SMT
-system.
+using SL context to distinguish between possible translations. These
+classifiers are then incorporated into the translation model of their
+word-based statistical MT system.
 %\comment{[how is our work different?]}  
-In their approach, a
-classifier consists of a set of binary feature functions and
-corresponding weights for each feature. 
-In both \cite{berger1996} and our method, features are defined in the
-form $h_k^{s}(t, c)$, where $t$
-is a translation, and $c$ is a source-language (SL) context.\todomlf{Defined SL which had not been defined} One difference is that \cite{berger1996} take $s$, $t$
-and $c$ to be based on word forms, whereas in our method they are based
-on lemma forms.
-An example would be the following feature where the Spanish word \emph{pez} (`fish' as a living animal) is seen as the
+In their approach, a classifier consists of a set of binary feature
+functions and corresponding weights for each feature.  In both
+\cite{berger1996} and our method, features are defined in the form
+$h_k^{s}(t, c)$, where $t$ is a translation, and $c$ is a SL
+context. One difference is that \cite{berger1996} take $s$, $t$ and
+$c$ to be based on word forms, whereas in our method they are based on
+lemma forms.  An example would be the following feature where the
+Spanish word \emph{pez} (`fish' as a living animal) is seen as the
 translation of \emph{arrain} (`fish') in the context \emph{arrain
-  handi} `big fish' would therefore be defined as:
+  handi} `big fish' and would therefore be defined as:
 %\comment{FSM: I have introduced the index $k$
@@
 \end{equation}
-\todomlf{R3 complains: `` Probability notation is confusing.
-I would suggest the use of \(p(t | c,s)\) instead of \(p_s(t|c)\), that will simplify notation.
-For instance instead of \(Z^s (c)\) we would have \(Z(c,s)\).''.
-Ignore or take onboard?}
+\todomlf{R3 complains: `` Probability notation is confusing. I would
+  suggest the use of \(p(t | c,s)\) instead of \(p_s(t|c)\), that will
+  simplify notation. For instance instead of \(Z^s (c)\) we would have
+  \(Z(c,s)\).''. Ignore or take onboard? FSM: Mention that we follow
+  the notation of Berger?}
 
@@
 schematically. %%%\todo{Can we try to bring the figure closer?}
-\todomlf{R3 says: ``it would have been nicer to see an example of one sentence going through the pipeline''. Can we add this in the text?}
+\todomlf{R3 says: ``it would have been nicer to see an example of one
+  sentence going through the pipeline''. Can we add this in the text?
+  FSM: We do not have much space}
 %Taking our previous example for the parallel corpus, Table~\ref{table:ex-monocorp} 
@@
 As our method is unsupervised, and therefore the occurrences of
-specific lexical selection events \((s,t,c)\) cannot be counted, a
-target language model \(P_\mathrm{TL}(\cdot)\) is used to compute a value for
-the fractional count for disambiguation path \(g_i\), $p(g_i|S)$ after
-suitable normalisation:\todomlf{R2 says ``Nevertheless, I've got problems to understand (4), in particular the paths in post-lexsel step.
-Some better explanation could be offered maybe by saying that (4) evaluates all possible solutions in the language model.''.
-Don't know what to do here.
-}
+specific lexical selection events \((s,t,c)\) cannot be counted, a TL
+model \(P_\mathrm{TL}(\cdot)\) is used to compute a value for the
+fractional count for disambiguation path \(g_i\), $p(g_i|S)$ after
+suitable normalisation:\todomlf{R2 says ``Nevertheless, I've got
+  problems to understand (4), in particular the paths in post-lexsel
+  step. Some better explanation could be offered maybe by saying that
+  (4) evaluates all possible solutions in the language model.''. Don't
+  know what to do here. }
 \begin{equation}
@@
 count $p(g_i|S)$ for the events $(s,t,c)$ found in \(g_i\), that is,
-when in \(g_i\) $t$ is the translation for \(s\) in context
-\(c\). That is, as if event $(s,t,c)$ had been seen a fractional
-number $p(g_i|S)$ of times. We prune $(s, t, c)$
-occurring less than a certain number of
-times in the corpus, using a development corpus to guide pruning (see
-section~\ref{sec:eval-ref-results}).
+when in \(g_i\) the translation for \(s\) in context \(c\) is
+$t$. That is, as if event $(s,t,c)$ had been seen a fractional number
+$p(g_i|S)$ of times. We prune $(s, t, c)$ occurring less than a
+certain number of times in the corpus, using a development corpus to
+guide pruning (see section~\ref{sec:eval-ref-results}).
 %@@@ \comment{Reword this to avoid repetition} 
@@
 used by \cite{sanchez08b} to train a hidden-Markov-model-based
-part-of-speech tagger in a rule-based machine translation system.
+part-of-speech tagger in a RBMT system.
 %Table~\ref{tb:workedexample} shows how a collection of samples
@@
 transfer output is the result of looking up the translations of the 
-source-language \emph{lexical forms} --- lemmas and  % combinations of lemma and tags
+SL \emph{lexical forms} --- lemmas and  % combinations of lemma and tags
 tags --- in the bilingual dictionary.
@@
   but some entries were added later to start work on lexical
-  selection. The version used in this paper is SVN revision
-  41375.\footnote{\url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-br-fr}}
+  selection.\footnote{The version used in this paper is SVN revision
+  41375; \url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-br-fr}}
 %  This is the only available machine translation system between Breton
 %  and French. It has been developed part-time over a number of years.
-\todomlf{R1 says ``whole section is rather length especially the revision numbers and URLs of the specific translations systems need not be mentioned in my opinion.
-The space could rather be used to elaborate in the description of the method in Sec 2.''}
+  \todomlf{R1 says ``whole section is rather length especially the
+    revision numbers and URLs of the specific translations systems
+    need not be mentioned in my opinion. The space could rather be
+    used to elaborate in the description of the method in Sec
+    2.''. FSM: I have moved them to the footnote}
 
@@
 %  pair \citep{rangelov11}, and the SETimes parallel corpus
-  The lexical resources for the pair were tuned to the SETimes parallel corpus
-  \citep{tyers10}.  
+  The lexical resources for the pair were tuned to the SETimes
+  parallel corpus \citep{tyers10}.
 % The work was carried out over a period of eight
@@
     bilingual dictionary in Apertium \citep{forcada2011apertium} may
-    contain several target-language translation for a given SL
-    word. Where there is more than one possible translation, the
-    dictionary writer chooses the most general or most frequent
-    translation among the set of possible translations and marks it as
-    the \emph{linguistic default} translation.} As a result of
-  attempting to include all possible translations, the average number
-  of translations per word is much
-  higher % (see table~\ref{table:lang-pair-stats})
-  than in other pairs.
-% The transfer rules were written
-%  by hand to treat translation problems in the corpus. 
-The version of
-  the software used in this paper is SVN revision
-  41476.\footnote{\url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-mk-en}}
+    contain several TL translation for a given SL word. Where there is
+    more than one possible translation, the dictionary writer chooses
+    the most general or most frequent translation among the set of
+    possible translations and marks it as the \emph{linguistic
+      default} translation.} As a result of attempting to include all
+  possible translations, the average number of translations per word
+  is much higher % (see table~\ref{table:lang-pair-stats})
+  than in other pairs.\footnote{The version of the software used in this paper is SVN revision
+  41476; \url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-mk-en}}
 \item[Basque--Spanish:] The development of the Basque to Spanish pair
@@
   Alternative translations were included in the bilingual
-  dictionary. The version used is SVN revision 44846.\footnote{\url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-eu-es}}
+  dictionary.\footnote{The version used is SVN revision 44846; \url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-eu-es}}
 % , but non-default translations were marked with a
@@
   contains a number of entries in the bilingual dictionary with more
-  than one translation. %The most-frequent translation, as judged by
+  than one translation.\footnote{The
+  version used in this paper is SVN revision
+  41387; \url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-en-es}}
+
+%The most-frequent translation, as judged by
 %  the language pair developer, is marked as the \emph{default}
@@
 %  frequency. \todo{Default is explained for some LPs only, why?}  
-  The
-  version used in this paper is SVN revision
-  41387.\footnote{\url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-en-es}}\todomlf{4-month mention removed in response to R1}
+\todomlf{4-month mention removed in response to R1}
 \end{description}
@@
 system chooses a translation for a word which is not  the one found in an
-annotated reference. The process uses a source-language sentence, $S =
+annotated reference. The process uses a SL sentence, $S =
 (s_1, s_2, \ldots, s_{|S|})$ and three functions. The first function,
@@
 of reference translations $T_r(s_i)$ and the set of translations from
-the lexical selection module $T_t(s_i)$ is empty.  Recall that,\todomlf{R1 says ``this has already been mentioned before''. Don't know what to do}
-although $T_t(s_i)$ returns a set, this set will be a singleton, as
-when the lexical-selection module returns more than one translation
-the first one will be selected.\footnote{In practice this does not happen
-as each ambiguous word has a \emph{default} translation.}\todomlf{R1 complains: ``why mention if it doesn't happen in practice? Relevance?'' What should we do?}
+the lexical selection module $T_t(s_i)$ is empty.  Recall
+that,\todomlf{R1 says ``this has already been mentioned
+  before''. Don't know what to do FSM: I would do nothing, this is why
+the sentence says ``Recall ...''} although $T_t(s_i)$ returns a set,
+this set will be a singleton, as when the lexical-selection module
+returns more than one translation the first one will be
+selected.\footnote{In practice this does not happen as each ambiguous
+  word has a \emph{default} translation.}\todomlf{R1 complains: ``why
+  mention if it doesn't happen in practice? Relevance?'' What should
+  we do?}
 
@@
 
-For creating the test corpora, providing a source language corpus 
-for training, and a target language corpus for scoring, we used four parallel corpora: 
+For creating the test corpora, providing a SL corpus for training, and
+a TL corpus for scoring, we used four parallel corpora:
 \begin{itemize}
@@
 lexical selection applied to a simplified translation problem in
-SMT. They use word alignments from {\sc giza++} \citep{och03a} to
-annotate source language words with their translations from the
+statistical MT. They use word alignments from {\sc giza++}
+\citep{och03a} to annotate SL words with their translations from the
 reference translation in the parallel corpus. One disadvantage of this
-method is that only one translation is annotated per source language
-word, meaning that accuracies may be lower because of missing
-translations --- this happens when the system chooses a translation
-which is adequate, but is not found in the reference translation. A
-second disadvantage is that the word alignments may not be 100\%
-reliable, which decreases the accuracy of the annotated corpus. An
-alternative method is described by \cite{zinovjeva2000}, who manually
-tags ambiguous words in English sentences with their % a selection of
+method is that only one translation is annotated per SL word, meaning
+that accuracies may be lower because of missing translations --- this
+happens when the system chooses a translation which is adequate, but
+is not found in the reference translation. A second disadvantage is
+that the word alignments may not be 100\% reliable, which decreases
+the accuracy of the annotated corpus. An alternative method is
+described by \cite{zinovjeva2000}, who manually tags ambiguous words
+in English sentences with their % a selection of
 translation in Swedish. %The size of the annotated corpus is not
@@
     Moses} toolkit \citep{koehn07} in default configuration up to step
-  three of training.}  We then use
-these alignments along with the bilingual dictionary of the MT system
-in question to extract only those sentences where: (a) there is at least
-one ambiguous word; (b) that ambiguous word is aligned to a single word in
-the target language; and (c) the word it is aligned to in the target
-language is found in the bilingual dictionary of the MT
-system. Sentences where there are no ambiguous words (approximately
-90\%, see Table~\ref{table:input-corp}) are discarded. The source side
-of the extracted sentence is then passed through the lexical transfer
-module, which returns all the possible translations, and for each
-ambiguous word, the translation is selected which is found aligned in
-the reference.
+  three of training.}  We then use these alignments along with the
+bilingual dictionary of the MT system in question to extract only
+those sentences where: (a) there is at least one ambiguous word; (b)
+that ambiguous word is aligned to a single word in the TL; and (c) the
+word it is aligned to in the TL is found in the bilingual dictionary
+of the MT system. Sentences where there are no ambiguous words
+(approximately 90\%, see Table~\ref{table:input-corp}) are
+discarded. The source side of the extracted sentence is then passed
+through the lexical transfer module, which returns all the possible
+translations, and for each ambiguous word, the translation is selected
+which is found aligned in the reference.
 
@@
 \item \textbf{Target language model} (TLM). One method of lexical
-  selection is to use the existing machine translation system to
-  generate all the possible translations for an input sentence, and
-  then score these translations \emph{on-line} on a model of the
-  target language. The highest scoring sentence is then output. This
-  is the method used by %@@@ \cite{carbonell06} and
+  selection is to use the existing MT system to generate all the
+  possible translations for an input sentence, and then score these
+  translations \emph{on-line} on a model of the TL. The highest
+  scoring sentence is then output. This is the method used
+  by %@@@ \cite{carbonell06} and
   \cite{melero07a}.% \todo{Also METIS?}
@@
   \url{http://downloads.sourceforge.net/project/apertium/apertium-lex-tools/apertium-lex-tools-0.1.0.tar.gz}.}
-to calculate the feature weights.  
-\todomlf{R3 says: ``The proposed maximum entropy or log-linear model is trained in a surprising way.
-In the early days of maximum entropy models, those were trained using the GIS algorithm.
-However, it has been proved in the literature that GIS has clear disadvantages with respect to other batch gradient descent algorithms.
-For instance, it does not allow for the use of regularization [...], and it is extremely slow to converge, which yields worse models than other methods.
-The smoothing/regularization of the model is also poor since it is simply feature pruning instead of regularization.''}
-After learning the feature sets and
-weights, we compute the evaluation measures described in
-Section~\ref{ss:measures}. There is an option to remove events 
-$(s, t, c)$ which occur
-less than a certain number of times in the training corpus. This is
-referred to as the feature pruning frequency threshold --- features
-occuring less than the threshold are discarded. The value was set
-experimentally. Values of between two and seven were tested, and the
-ones which provided the best improvement on the development corpus
-were selected; they happen to come close to the rule-of-thumb value of
-five that \citet[p.~596]{Manning99b} found to be effective.
+to calculate the feature weights.  \todomlf{R3 says: ``The proposed
+  maximum entropy or log-linear model is trained in a surprising
+  way. In the early days of maximum entropy models, those were trained
+  using the GIS algorithm. However, it has been proved in the
+  literature that GIS has clear disadvantages with respect to other
+  batch gradient descent algorithms. For instance, it does not allow
+  for the use of regularization [...], and it is extremely slow to
+  converge, which yields worse models than other methods. The
+  smoothing/regularization of the model is also poor since it is
+  simply feature pruning instead of regularization.'' FSM: I would say
+  something in the discussion (future work).}  After learning the
+feature sets and weights, we compute the evaluation measures described
+in Section~\ref{ss:measures}. There is an option to remove events $(s,
+t, c)$ which occur less than a certain number of times in the training
+corpus. This is referred to as the feature pruning frequency threshold
+--- features occurring less than the threshold are discarded. The value
+was set experimentally. Values of between two and seven were tested,
+and the ones which provided the best improvement on the development
+corpus were selected; they happen to come close to the rule-of-thumb
+value of five that \citet[p.~596]{Manning99b} found to be effective.
 Table~\ref{table:maxent-features} shows the number of features that
@@
  the new approach with respect to the \emph{default behaviour} (the
- linguist-chosen defaults),\todomlf{R1 says the footnote is important and should be incorporated to the text; what should we do?}\footnote{The high error rate
-   for the Breton--French pair may be as a result of having the
-   linguistic defaults tuned to a different domain than that of the
-   corpus.}  with respect to the \emph{oracle} (which represents the
- upper bound to performance), and with respect to the results obtained
- by using the target-language model online, for each of the language
- pairs in Apertium with respect to our two evaluation metrics.
+ linguist-chosen defaults), with respect to the \emph{oracle} (which
+ represents the upper bound to performance), and with respect to the
+ results obtained by using the TL model online, for each of the
+ language pairs in Apertium with respect to our two evaluation
+ metrics. Note that the high error rate for the Breton--French pair
+ may be as a result of having the linguistic defaults tuned to a
+ different domain than that of the corpus.\todomlf{R1 says the
+   footnote is important and should be incorporated to the text; what
+   should we do? FSM: I have moved it to the main text and split the
+   paragraph into two pieces.}
+
  Significant improvements with respect to the results obtained using
- the target-language model online are apparent with the Breton--French
- ---the pair with the least data--- and the English--Spanish language
- pairs. In the remaining cases, the maximum-entropy method comes close
- to the target-language model performance in terms of similar or better BLEU and LER scores, at a much smaller
- computational cost.
+ the TL model online are apparent with the Breton--French ---the pair
+ with the least data--- and the English--Spanish language pairs. In
+ the remaining cases, the maximum-entropy method comes close to the TL
+ model performance in terms of similar or better BLEU and LER scores,
+ at a much smaller computational cost.
 % \todo{I miss a brief separate discussion of BLEU and LER, which are reported separately!}
@@
 
- Improvements with respect to the target-language-model performance
- are likely due to the effective use that the maximum-entropy model
- makes of information about the relevant source-language contexts and
- their translations, through the weighting of features representing
- those source-language contexts across the whole corpus.
+ Improvements with respect to the TL model performance are likely due
+ to the effective use that the maximum-entropy model makes of
+ information about the relevant SL contexts and their translations,
+ through the weighting of features representing those SL contexts
+ across the whole corpus.
 
@@
    fractional counts. The results in bold face show statistically significant
-   improvements for the maximum-entropy model compared to the target language model 
+   improvements for the maximum-entropy model compared to the TL model 
    according to pair-bootstrap resampling.}
@@
 without the need for an annotated corpus, (in this case a word-aligned
-bilingual corpus): one just needs a source-language corpus, a
-statistical target-language model, and the RBMT system itself. The
-input to the method is simply the part-of-speech tagged source text in
-which each word is annotated with all the translations provided by the
-bilingual dictionary in the system: this makes it applicable to almost
-any rule-based machine translation system. The system uses a
-maximum-entropy formalism for lexical selection, as \cite{berger1996}
-and \cite{marechek10}, but instead of counting actual lexical
-selection events in an annotated corpus, it counts fractional
-occurrences of these events as estimated by a target-language
-model. The method is evaluated both intrinsically (just looking at the
-actual lexical selection events) and extrinsically (measuring the
-quality of machine translation). Results on four language pairs using
-the Apertium \citep{forcada2011apertium} machine translation system
-show that the method obtains similar or better results than those
-expensively obtained by scoring an exponential number of lexical
-selections for each sentence using the target-language model online.\todomlf{R2 says: ``An error analysis would also help to understand the proposal.'' We don't have time now I think.}
+bilingual corpus): one just needs a SL corpus, a statistical TL model,
+and the RBMT system itself. The input to the method is simply the
+part-of-speech tagged source text in which each word is annotated with
+all the translations provided by the bilingual dictionary in the
+system: this makes it applicable to almost any RBMT system. The system
+uses a maximum-entropy formalism for lexical selection, as
+\cite{berger1996} and \cite{marechek10}, but instead of counting
+actual lexical selection events in an annotated corpus, it counts
+fractional occurrences of these events as estimated by a TL model. The
+method is evaluated both intrinsically (just looking at the actual
+lexical selection events) and extrinsically (measuring the quality of
+MT). Results on four language pairs using the Apertium
+\citep{forcada2011apertium} MT system show that the method obtains
+similar or better results than those expensively obtained by scoring
+an exponential number of lexical selections for each sentence using
+the TL model online.\todomlf{R2 says: ``An error analysis would also
+  help to understand the proposal.'' We don't have time now I think.}
 

### 2015-04-13 2015-eamt-melsa/2015-eamt-melsa.tex
some fixes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59731 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \cite{dagan1994word} used the term \emph{word sense disambiguation} to
-refer to what is actually target word selection\todofsm{Should we call
-  it lexical selection instead of target word selection?} in MT; they
+refer to what is actually lexical selection in MT; they
 used a parser to identify syntactic relations such as subject--object
@@
 \cite{berger1996} and our method, features are defined in the form
-$h_k^{s}(t, c)$, where $t$ is a translation, and $c$ is a SL
+$h_k^{s}(t, c)$,\footnote{We follow the notation of \cite{berger1996}} where $t$ is a translation, and $c$ is a SL
 context. One difference is that \cite{berger1996} take $s$, $t$ and
@@
 This feature considers a context of zero words to the left of the
-problem word and one word (+ \emph{handi}) to the right of it.\todomlf{Removed \(k\) and used a more informative subscript}
+problem word and one word (+ \emph{handi}) to the right of it.
 
@@
 As a result of training, each of the \(n_F\) features \(h^s_k(t,c)\) in the classifier is assigned a
-weight $\lambda_k^s$. \todomlf{Reworded to define \(n_F\)} Combining these weights of active features
+weight $\lambda_k^s$. Combining these weights of active features
 as in equation~(\ref{eq:max-ent-sum}) yields the probability of a
@@
 \end{equation}
-\todomlf{R3 complains: `` Probability notation is confusing. I would
-  suggest the use of \(p(t | c,s)\) instead of \(p_s(t|c)\), that will
-  simplify notation. For instance instead of \(Z^s (c)\) we would have
-  \(Z(c,s)\).''. Ignore or take onboard? FSM: Mention that we follow
-  the notation of Berger?}
+%\todomlf{R3 complains: `` Probability notation is confusing. I would
+%  suggest the use of \(p(t | c,s)\) instead of \(p_s(t|c)\), that will
+%  simplify notation. For instance instead of \(Z^s (c)\) we would have
+%  \(Z(c,s)\).''. Ignore or take onboard? FSM: Mention that we follow
+%  the notation of Berger?}
 

### 2015-04-13 2015-eamt-assim/2015-eamt-assim.tex
Added interpretation of Alpha values and a file with individual scores by each task mode (to calculate significance)
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59756 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-% actually, I think the following paragraph may be about kappa, will check once I get access to landis77
-%The interpretation of Krippendorff's alpha varies depending on the application.
-One of the general guidelines suggested by \citet{landis77} is as follows (cited after \citet{carletta97}: "[values of alpha] < 0 ``poor'' agreement, 0 to .2 ``slight'', .21 to .40 ``fair'', .41 to .60 ``moderate'', .61 - .80 ``substantial'', and .81 to 1 ``near perfect''"
+The interpretation of Krippendorff's alpha varies depending on the application.
+One of the general guidelines suggested by \citet{landis77} for kappa-like measures (which includes Krippendorff's Alpha) is as follows: k < 0.00 indicates ``poor'' agreement, 0.00 to 0.20 ``slight'', 0.21 to 0.40 ``fair'', 0.41 to 0.60 ``moderate'', 0.61 to 0.80 ``substantial'', and 0.81 to 1.00 ``near perfect''.
 

### 2015-04-15 2015-eamt-assim/2015-eamt-assim.tex
back to 8 pages
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59824 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \label{sec:results}
+\begin{table*}
+\centering
+  \begin{tabular}{|l|r|r|r|r|r|r|r|r|}
+  \hline
+ \multirow{2}{*}{\textbf{Density}} & \multicolumn{4}{|c|}{Basque--Spanish} & \multicolumn{4}{|c|}{Tatar--Russian} 
+\\\cline{2-9}
+                                            & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint}& \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} 
+\\
+\hline
+10\%&62 \(\pm\) 32&58 \(\pm\) 28&40 \(\pm\) 39&49 \(\pm\) 40
+&57 \(\pm\) 42&64 \(\pm\) 41&54 \(\pm\) 43&46 \(\pm\) 41\\
+20\%&65 \(\pm\) 30&70 \(\pm\) 27&31 \(\pm\) 28&31 \(\pm\) 30
+&65 \(\pm\) 31&60 \(\pm\) 33&46 \(\pm\) 31&39 \(\pm\) 32\\
+30\%&48 \(\pm\) 26&40 \(\pm\) 24&26 \(\pm\) 20&18 \(\pm\) 18
+&59 \(\pm\) 28&56 \(\pm\) 26&40 \(\pm\) 28&35 \(\pm\) 30\\
+\hline
+\end{tabular}
+\caption{Average number of gaps successfully filled (\%), using a synonym list} 
+\label{table:res}
+\end{table*}
 
-The results are presented in this section. Tables \ref{table:res-eus-spa} and \ref{table:res-tat-rus} 
-show
+The results are presented in this section. Table \ref{table:res} 
+shows
 the proportion and standard deviation of correct answers depending on evaluation mode and gap density.
@@
 \subsection{Annotator agreement}
-After obtaining the results we calculated Krippendorff's alpha~\citep{krippendorff70} measure to represent annotator agreement, shown in Tables~\ref{table:alpha-eus-spa} and~\ref{table:alpha-tat-rus}.
+After obtaining the results we calculated Krippendorff's alpha~\citep{krippendorff70} measure to represent annotator agreement, shown in Table~\ref{table:alpha}. 
+\begin{table*}
+\centering
+  \begin{tabular}{|l|r|r|r|r|r|r|r|r|}
+  \hline
+ \multirow{2}{*}{\textbf{Density}} & \multicolumn{4}{|c|}{Basque--Spanish}&\multicolumn{4}{|c|}{Tatar--Russian} \\\cline{2-9}
+                                            & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint}& \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
+\hline
+10\%&0.496&0.517&0.400&0.124&0.598&0.459 &0.711 &0.517\\
+20\%&0.714&0.700&0.358&0.275&0.740&0.667 &0.473 &0.261\\
+30\%&0.559&0.430&0.406&0.300&0.534&0.581 &0.411 &0.412\\
+
+\hline
+\end{tabular}
+\caption{Krippendorff Alpha measure} 
+\label{table:alpha}
+\end{table*}
 
@@
 \subsection{Results for different domains}
+\begin{table*}
+\centering
+\begin{tabular}{|r|r |*{4}{c}|}
+\hline
+\multicolumn{2}{|c|}{} &\multicolumn{4}{c|}{\textbf{Evaluation mode}}\\
+\hline
+\textbf{Domain} & \textbf{Gap percentage} & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
+\hline
+ \multirow{3}{*}{\textbf{Casual}} &
+10\%&64 \(\pm\) 45&64 \(\pm\) 43&62 \(\pm\) 46&53 \(\pm\) 44\\
+& 20\%&73 \(\pm\) 32&63 \(\pm\) 36&41 \(\pm\) 28&38 \(\pm\) 31\\
+& 30\%&70 \(\pm\) 31&60 \(\pm\) 24&39 \(\pm\) 27&38 \(\pm\) 33\\
+\hline
+ \multirow{3}{*}{\textbf{Legal}} &
+10\%&53 \(\pm\) 40&68 \(\pm\) 35&39 \(\pm\) 38&33 \(\pm\) 35\\
+&20\%&61 \(\pm\) 25&66 \(\pm\) 24&50 \(\pm\) 34&48 \(\pm\) 34\\
+&30\%&50 \(\pm\) 26&48 \(\pm\) 29&40 \(\pm\) 27&34 \(\pm\) 29\\
+\hline
+ \multirow{3}{*}{\textbf{News}} &
+10\%&53 \(\pm\) 38&60 \(\pm\) 44&57 \(\pm\) 42&49 \(\pm\) 39\\
+&20\%&59 \(\pm\) 34&49 \(\pm\) 35&47 \(\pm\) 32&29 \(\pm\) 29\\
+&30\%&58 \(\pm\) 22&61 \(\pm\) 22&41 \(\pm\) 30&35 \(\pm\) 27\\
+\hline
+
+\end{tabular}
+\caption{Tatar--Russian Average number of gaps successfully filled (\%), using a synonym list, for three different domains.} 
+\label{table:rus} 
+\end{table*}
+
 For the Tatar--Russian language pair the participants were offered texts from three different domains (in equal proportions): casual conversations, legal texts and news.
-The results by domains are displayed in tables \ref{table:rus-casual}, \ref{table:rus-legal}, and \ref{table:rus-news}.
+The results by domains are displayed in table~\ref{table:rus}.
 The MT system used in the evaluation has been targeted to translate texts from all three of the domains.
@@
 
-\begin{table*}
-\centering
-  \begin{tabular}{|l|r|r|r|r|}
-  \hline
- \multirow{2}{*}{\textbf{Density}} & \multicolumn{4}{|c|}{Evaluation mode} \\\cline{2-5}
-                                            & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
-\hline
-10\%&62 \(\pm\) 32&58 \(\pm\) 28&40 \(\pm\) 39&49 \(\pm\) 40\\
-20\%&65 \(\pm\) 30&70 \(\pm\) 27&31 \(\pm\) 28&31 \(\pm\) 30\\
-30\%&48 \(\pm\) 26&40 \(\pm\) 24&26 \(\pm\) 20&18 \(\pm\) 18\\
-\hline
-\end{tabular}
-\caption {Basque--Spanish: Average number of gaps successfully filled (\%), using a synonym list} 
-\label{table:res-eus-spa}
-\end{table*}
+% \begin{table*}
+% \centering
+%   \begin{tabular}{|l|r|r|r|r|}
+%   \hline
+%  \multirow{2}{*}{\textbf{Density}} & \multicolumn{4}{|c|}{Evaluation mode} \\\cline{2-5}
+%                                             & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
+% \hline
+% 10\%&62 \(\pm\) 32&58 \(\pm\) 28&40 \(\pm\) 39&49 \(\pm\) 40\\
+% 20\%&65 \(\pm\) 30&70 \(\pm\) 27&31 \(\pm\) 28&31 \(\pm\) 30\\
+% 30\%&48 \(\pm\) 26&40 \(\pm\) 24&26 \(\pm\) 20&18 \(\pm\) 18\\
+% \hline
+% \end{tabular}
+% \caption{Basque--Spanish: Average number of gaps successfully filled (\%), using a synonym list} 
+% \label{table:res-eus-spa}
+% \end{table*}
 
-\begin{table*}
-\centering
-  \begin{tabular}{|l|r|r|r|r|}
-    \hline
-    \multirow{2}{*}{\textbf{Density}} & \multicolumn{4}{|c|}{Evaluation mode} \\\cline{2-5}
-                                            & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
-\hline
-10\%&57 \(\pm\) 42&64 \(\pm\) 41&54 \(\pm\) 43&46 \(\pm\) 41\\
-20\%&65 \(\pm\) 31&60 \(\pm\) 33&46 \(\pm\) 31&39 \(\pm\) 32\\
-30\%&59 \(\pm\) 28&56 \(\pm\) 26&40 \(\pm\) 28&35 \(\pm\) 30\\
-    \hline
-  \end{tabular}
-  \caption{Tatar--Russian: Average number of gaps successfully filled (\%), using a synonym list}
-  \label{table:res-tat-rus}
-\end{table*}
+% \begin{table*}
+% \centering
+%   \begin{tabular}{|l|r|r|r|r|}
+%     \hline
+%     \multirow{2}{*}{\textbf{Density}} & \multicolumn{4}{|c|}{Evaluation mode} \\\cline{2-5}
+%                                             & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
+% \hline
+% 10\%&57 \(\pm\) 42&64 \(\pm\) 41&54 \(\pm\) 43&46 \(\pm\) 41\\
+% 20\%&65 \(\pm\) 31&60 \(\pm\) 33&46 \(\pm\) 31&39 \(\pm\) 32\\
+% 30\%&59 \(\pm\) 28&56 \(\pm\) 26&40 \(\pm\) 28&35 \(\pm\) 30\\
+%     \hline
+%   \end{tabular}
+%   \caption{Tatar--Russian: Average number of gaps successfully filled (\%), using a synonym list}
+%   \label{table:res-tat-rus}
+% \end{table*}
 
-\begin{table*}
-\centering
-  \begin{tabular}{|l|r|r|r|r|}
-  \hline
- \multirow{2}{*}{\textbf{Density}} & \multicolumn{4}{|c|}{Evaluation mode} \\\cline{2-5}
-                                            & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
-\hline
-10\%&0.496&0.517&0.400&0.124\\
-20\%&0.714&0.700&0.358&0.275\\
-30\%&0.559&0.430&0.406&0.300\\
+% \begin{table*}
+% \centering
+%   \begin{tabular}{|l|r|r|r|r|}
+%   \hline
+%  \multirow{2}{*}{\textbf{Density}} & \multicolumn{4}{|c|}{Evaluation mode} \\\cline{2-5}
+%                                             & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
+% \hline
+% 10\%&0.496&0.517&0.400&0.124\\
+% 20\%&0.714&0.700&0.358&0.275\\
+% 30\%&0.559&0.430&0.406&0.300\\
 
-\hline
-\end{tabular}
-\caption {Basque--Spanish: Krippendorff Alpha measure} 
-\label{table:alpha-eus-spa}
-\end{table*}
+% \hline
+% \end{tabular}
+% \caption {Basque--Spanish: Krippendorff Alpha measure} 
+% \label{table:alpha-eus-spa}
+% \end{table*}
 
-\begin{table*}
-\centering
-  \begin{tabular}{|l|r|r|r|r|}
-  \hline
- \multirow{2}{*}{\textbf{Density}} & \multicolumn{4}{|c|}{Evaluation mode} \\\cline{2-5}
-                                            & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
-\hline
-10\%&0.598&0.459 &0.711 &0.517 \\
-20\%&0.740&0.667 &0.473 &0.261 \\
-30\%&0.534&0.581 &0.411 &0.412 \\
-\hline
-\end{tabular}
-\caption {Tatar--Russian: Krippendorff Alpha measure} 
-\label{table:alpha-tat-rus}
-\end{table*}
+% \begin{table*}
+% \centering
+%   \begin{tabular}{|l|r|r|r|r|}
+%   \hline
+%  \multirow{2}{*}{\textbf{Density}} & \multicolumn{4}{|c|}{Evaluation mode} \\\cline{2-5}
+%                                             & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
+% \hline
+% 10\%&0.598&0.459 &0.711 &0.517 \\
+% 20\%&0.740&0.667 &0.473 &0.261 \\
+% 30\%&0.534&0.581 &0.411 &0.412 \\
+% \hline
+% \end{tabular}
+% \caption {Tatar--Russian: Krippendorff Alpha measure} 
+% \label{table:alpha-tat-rus}
+% \end{table*}
 
 % results by text mode in tat-rus, with synonyms
+% \begin{table*}
+% \centering
+% \begin{tabular}{|r|r |*{4}{c}|}
+% \hline
+%   &&\multicolumn{4}{c|}{\textbf{Evaluation mode}}\\
+% \hline
+% & \textbf{Gap percentage} & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
+% \hline
+%  \multirow{3}{*}{\textbf{Casual}} &
+% 10\%&64 \(\pm\) 45&64 \(\pm\) 43&62 \(\pm\) 46&53 \(\pm\) 44\\
+% & 20\%&73 \(\pm\) 32&63 \(\pm\) 36&41 \(\pm\) 28&38 \(\pm\) 31\\
+% & 30\%&70 \(\pm\) 31&60 \(\pm\) 24&39 \(\pm\) 27&38 \(\pm\) 33\\
+% \hline
+%  \multirow{3}{*}{\textbf{Legal}} &
+% 10\%&53 \(\pm\) 40&68 \(\pm\) 35&39 \(\pm\) 38&33 \(\pm\) 35\\
+% &20\%&61 \(\pm\) 25&66 \(\pm\) 24&50 \(\pm\) 34&48 \(\pm\) 34\\
+% &30\%&50 \(\pm\) 26&48 \(\pm\) 29&40 \(\pm\) 27&34 \(\pm\) 29\\
+% \hline
+%  \multirow{3}{*}{\textbf{News}} &
+% 10\%&53 \(\pm\) 38&60 \(\pm\) 44&57 \(\pm\) 42&49 \(\pm\) 39\\
+% &20\%&59 \(\pm\) 34&49 \(\pm\) 35&47 \(\pm\) 32&29 \(\pm\) 29\\
+% &30\%&58 \(\pm\) 22&61 \(\pm\) 22&41 \(\pm\) 30&35 \(\pm\) 27\\
+% \hline
 
-\begin{table*}
+% \end{tabular}
+% \caption{Tatar--Russian Average number of gaps successfully filled (\%), using a synonym list, for three different domains.} 
+% \label{table:rus} 
+% \end{table*}
 
-\centering
-\begin{tabular}{|r |*{4}{c}|}
-\hline
-  &\multicolumn{4}{c|}{\textbf{Evaluation mode}}\\
-\hline
-\textbf{Gap percentage} & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
-\hline
-10\%&64 \(\pm\) 45&64 \(\pm\) 43&62 \(\pm\) 46&53 \(\pm\) 44\\
-20\%&73 \(\pm\) 32&63 \(\pm\) 36&41 \(\pm\) 28&38 \(\pm\) 31\\
-30\%&70 \(\pm\) 31&60 \(\pm\) 24&39 \(\pm\) 27&38 \(\pm\) 33\\
-\hline
-\end{tabular}
-\caption {Tatar--Russian, casual conversations: Average number of gaps successfully filled (\%), using a synonym list} 
-\label{table:rus-casual} 
-\end{table*}
+% \begin{table*}
 
-\begin{table*}
-\centering
-\begin{tabular}{|r |*{4}{c}|}
-\hline
-  &\multicolumn{4}{c|}{\textbf{Evaluation mode}}\\
-\hline
-\textbf{Gap percentage} & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
-\hline
-10\%&53 \(\pm\) 40&68 \(\pm\) 35&39 \(\pm\) 38&33 \(\pm\) 35\\
-20\%&61 \(\pm\) 25&66 \(\pm\) 24&50 \(\pm\) 34&48 \(\pm\) 34\\
-30\%&50 \(\pm\) 26&48 \(\pm\) 29&40 \(\pm\) 27&34 \(\pm\) 29\\
-\hline
-\end{tabular}
-\caption {Tatar--Russian, legal texts: Average number of gaps successfully filled (\%), using a synonym list} \label{table:rus-legal} 
-\end{table*}
+% \centering
+% \begin{tabular}{|r |*{4}{c}|}
+% \hline
+%   &\multicolumn{4}{c|}{\textbf{Evaluation mode}}\\
+% \hline
+% \textbf{Gap percentage} & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
+% \hline
+% 10\%&64 \(\pm\) 45&64 \(\pm\) 43&62 \(\pm\) 46&53 \(\pm\) 44\\
+% 20\%&73 \(\pm\) 32&63 \(\pm\) 36&41 \(\pm\) 28&38 \(\pm\) 31\\
+% 30\%&70 \(\pm\) 31&60 \(\pm\) 24&39 \(\pm\) 27&38 \(\pm\) 33\\
+% \hline
+% \end{tabular}
+% \caption {Tatar--Russian, casual conversations: Average number of gaps successfully filled (\%), using a synonym list} 
+% \label{table:rus-casual} 
+% \end{table*}
 
-\begin{table*}
-\centering
-\begin{tabular}{|r |*{4}{c}|}
-\hline
-  &\multicolumn{4}{c|}{\textbf{Evaluation mode}}\\
-\hline
-\textbf{Gap percentage} & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
-\hline
-10\%&53 \(\pm\) 38&60 \(\pm\) 44&57 \(\pm\) 42&49 \(\pm\) 39\\
-20\%&59 \(\pm\) 34&49 \(\pm\) 35&47 \(\pm\) 32&29 \(\pm\) 29\\
-30\%&58 \(\pm\) 22&61 \(\pm\) 22&41 \(\pm\) 30&35 \(\pm\) 27\\
-\hline
-\end{tabular}
-\caption {Tatar--Russian, news: Average number of gaps successfully filled (\%), using a synonym list} 
-\label{table:rus-news} 
-\end{table*}
+% \begin{table*}
+% \centering
+% \begin{tabular}{|r |*{4}{c}|}
+% \hline
+%   &\multicolumn{4}{c|}{\textbf{Evaluation mode}}\\
+% \hline
+% \textbf{Gap percentage} & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
+% \hline
+% 10\%&53 \(\pm\) 40&68 \(\pm\) 35&39 \(\pm\) 38&33 \(\pm\) 35\\
+% 20\%&61 \(\pm\) 25&66 \(\pm\) 24&50 \(\pm\) 34&48 \(\pm\) 34\\
+% 30\%&50 \(\pm\) 26&48 \(\pm\) 29&40 \(\pm\) 27&34 \(\pm\) 29\\
+% \hline
+% \end{tabular}
+% \caption {Tatar--Russian, legal texts: Average number of gaps successfully filled (\%), using a synonym list} \label{table:rus-legal} 
+% \end{table*}
+
+% \begin{table*}
+% \centering
+% \begin{tabular}{|r |*{4}{c}|}
+% \hline
+%   &\multicolumn{4}{c|}{\textbf{Evaluation mode}}\\
+% \hline
+% \textbf{Gap percentage} & \textbf{MT \& Src} & \textbf{MT} & \textbf{Src} & \textbf{No hint} \\
+% \hline
+% 10\%&53 \(\pm\) 38&60 \(\pm\) 44&57 \(\pm\) 42&49 \(\pm\) 39\\
+% 20\%&59 \(\pm\) 34&49 \(\pm\) 35&47 \(\pm\) 32&29 \(\pm\) 29\\
+% 30\%&58 \(\pm\) 22&61 \(\pm\) 22&41 \(\pm\) 30&35 \(\pm\) 27\\
+% \hline
+% \end{tabular}
+% \caption {Tatar--Russian, news: Average number of gaps successfully filled (\%), using a synonym list} 
+% \label{table:rus-news} 
+% \end{table*}
 

### 2015-04-15 2015-eamt-assim/2015-eamt-assim.tex
Added significance to article text
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59825 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 In addition, tasks with source hints are completed better than tasks without hints, and the same relation holds between MT+source and MT-only types of tasks.
+In view of the relatively large standard deviations, the significance of the hints' contribution was tested using a linear regression model.
+The data points were represented as an individual evaluator's average score (the number of correct answers divided by the total number of answers) in each of the percentage-hint combinations.
+Two separate models were created: one for no-hint vs MT-hint tasks, and another for no-hint vs source-hint tasks.
+Given the null hypothesis of the regression line slope equal to zero, the contribution of MT hint is significant on the p~<~0.001 level.
+The contribution of the source hint is insignificant with p~=~0.162.
 
-Two records break this trend: the no-hint 10\% sentences in Basque--Spanish, which scored significantly higher than the source-hint in the same category, and MT+source 10\% sentences in Tatar--Russian, which we would have expected to score higher than the corresponding MT task.
+Two records in the data do not align with our expectations: the no-hint 10\% sentences in Basque--Spanish, which scored significantly higher than the source-hint in the same category, and MT+source 10\% sentences in Tatar--Russian, which we would have expected to score higher than the corresponding MT task.
 In the first case, this is largely due to the use of synonyms list.
@@
 This work has been partly funded by the Spanish Ministerio de Econom{\'i}a y Competitividad through project TIN2012-32615 and by the Google Summer of Code 2014 programme.
+Ekaterina Ageeva's work has been supported by Google Summer of Code 2014 through the Apertium project.
 We would like to thank the volunteers who participated in the evaluations.

### 2015-04-15 2015-eamt-assim/2015-eamt-assim.tex
Minor edits
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59840 72bbbca6-d526-0410-a7d9-f06f51895060

@@
   strategy independently proposed in 2012 and 2013 for evaluation of
- MT for assimilation purposes. The strategy involves
+ machine translation (MT) for assimilation purposes. The strategy involves
   human evaluators, who are asked to restore keywords (to fill gaps) in
@@
 The methods include post-editing and comparison by bilingual experts \citep{ginesti09},\comment{MLF: should we also cite the original WMT reference cited by Ginestí et al.?} and multiple choice tests \citep{jones07,trosterud12}.
-These approaches are often costly and prone to subjectivity: see the discussion in \cite{oregan13}.
+These approaches are often costly and prone to subjectivity: see the discussion by \cite{oregan13}.
 As an alternative, the modification of \emph{cloze}\comment{\emph{cloze} stands for \emph{closure}.} testing \citep{taylor53} was introduced for assimilation evaluation, first by \citet{trosterud12} as a supplementary, and then by \citet{oregan13} as a stand-alone method.
@@
 This paper is concerned primarily with assimilation evaluation; for a deeper discussion on evaluation see e.g.
-\citep{koehn2009}.
+\citep{koehn2010}.
 
@@
 Although we have not asked the participants about their knowledge of languages other than Tatar and Russian, it is reasonable to assume that most Russian participants knew English to some extent.
-\comment{EA 12: could you tell me more about the Spanish evaluators?
-I would write something similar}
+The Spanish participants were university staff with background in Computer Science.
 
@@
 Two separate models were created: one for no-hint vs MT-hint tasks, and another for no-hint vs source-hint tasks.
-Given the null hypothesis of the regression line slope equal to zero, the contribution of MT hint is significant on the p~<~0.001 level.
-The contribution of the source hint is insignificant with p~=~0.162.
+Given the null hypothesis of the regression line slope equal to zero, the contribution of MT hint is significant on the p~<~0.001 level, while the contribution of the source hint is significant only with p~<~0.162
 
@@
 
-We selected this measure because of its compatibility with more than two annotators per task and the missing data (not all the gaps were evaluated).
+We selected this measure because of its compatibility with more than two annotators per task and missing data (not all the gaps were evaluated).
 To calculate Krippendorff's alpha we used an algorithm implementation by Thomas Grill,\footnote{\url{http://grrrr.org/data/dev/krippendorff_alpha/}} dividing the answers in each gap into two categories: correct and incorrect.
@@
 The interpretation of Krippendorff's alpha varies depending on the application.
-One of the general guidelines suggested by \citet{landis77} for kappa-like measures (which includes Krippendorff's Alpha) is as follows: k < 0.00 indicates ``poor'' agreement, 0.00 to 0.20 ``slight'', 0.21 to 0.40 ``fair'', 0.41 to 0.60 ``moderate'', 0.61 to 0.80 ``substantial'', and 0.81 to 1.00 ``near perfect''.
+One of the general guidelines suggested by \citet{landis77} for kappa-like measures (which includes Krippendorff's Alpha) is as follows: k < 0 indicates ``poor'' agreement, 0 to 0.2 ``slight'', 0.21 to 0.4 ``fair'', 0.41 to 0.6 ``moderate'', 0.61 to 0.8 ``substantial'', and 0.81 to 1 ``near perfect''.
 
@@
 \begin{enumerate}
-\item As reported in \cite{oregan13}, unless the evaluation is targeted at a specific text domain, it may be beneficial to include a stylistic variety of texts in the initial corpus.
+\item As reported by \cite{oregan13}, unless the evaluation is targeted at a specific text domain, it may be beneficial to include a stylistic variety of texts in the initial corpus.
 Neighboring sentences on the same topic may assist the users in gap-filling tasks;

### 2015-04-16 2015-eamt-assim/2015-eamt-assim.tex
Mikel's last changes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59856 72bbbca6-d526-0410-a7d9-f06f51895060

@@
   {Dept. Llenguatges i Sistemes Inform\`{a}tics}\\
-  \blankout{Universitat d'Alacant, Spain} \\
+  \blankout{Universitat d'Alacant} \\
+  \blankout{E-03071 Alacant, Spain}\\
   {{\tt mlf@dlsi.ua.es}}  \\[2ex]
@@
   {Dept. Llenguatges i Sistemes Inform\`{a}tics}\\
-  {Universitat d'Alacant, Spain} \\
+  \blankout{Universitat d'Alacant} \\
+  \blankout{E-03071 Alacant, Spain}\\
   {{\tt japerez@dlsi.ua.es}}
@@
 These approaches are often costly and prone to subjectivity: see the discussion by \cite{oregan13}.
-As an alternative, the modification of \emph{cloze}\comment{\emph{cloze} stands for \emph{closure}.} testing \citep{taylor53} was introduced for assimilation evaluation, first by \citet{trosterud12} as a supplementary, and then by \citet{oregan13} as a stand-alone method.
+As an alternative, the modification of \emph{cloze}\comment{\emph{cloze} stands for \emph{closure}.} testing \citep{taylor53} was introduced for assimilation evaluation, first by \citet{trosterud12} as a supplementary technique, and then by \citet{oregan13} as a stand-alone method.
 Prior to this, cloze tests have been used to evaluate raw MT quality \citep{vanslype79,somers00}.
-While these authors ask informants to fill gaps in MT output, \citet{trosterud12} and \citet{oregan13} propose to fill gaps in the reference (human) translation.
+While these authors ask informants to fill gaps in MT output, \citet{trosterud12} and \citet{oregan13} ask informants to fill gaps in the reference (human) translation.
 A designated number of keywords is removed from the human-translated sentences.
-The evaluators are then asked to fill the gaps with suitable words.
+The evaluators are then asked to fill the gaps with suitable words with and without the help of MT output.
 The gap-filling task models how well users comprehend the key points of the text, as it is roughly equivalent with answering questions.
@@
 The gap-filling method has been successfully used to evaluate the Basque--English Apertium language pair.
-In this work we extend the evaluation to two more language pairs: Basque--Spanish and Tatar-Russian.
+In this work we extend the evaluation to two more language pairs: Basque--Spanish and Tatar--Russian.
 The former pair, while not producing output suitable for post-editing, is a good example of an assimilation MT system.
@@
 For the latter pair, the evaluation served as a quality check in the period of active development during the Google Summer of Code 2014 programme.
-In addition to evaluating, we explore the previously unconsidered aspects of the experiment: the correlation between evaluators' scores, and the effects of the texts' linguistic domain and the number of gaps in a sentence.
+In addition to evaluating, we explore the previously unconsidered aspects of the experiment: the correlation between evaluators' scores, and the effects of the linguistic domain of texts and the percentage of gaps in a sentence.
 To facilitate the evaluation, we introduce an automated system which creates task sets from parallel corpora given a range of parameters (number of gaps in a sentence, hint type, gap filler,\comment{MLF: gap filler or gap marker?} \comment{EA: well, initially we had questions with lemmas in place of gaps, and also multiple choice questions where answer options got generated automatically.
@@
 Should I remove?} etc.), checks evaluators' answers, and calculates and reports generalized results.
-This system is integrated into the Appraise MT evaluation platform \citep{federmann12}; the code is open source and is available on GitHub.\footnote{{\url{https://github.com/Sereni/Appraise}}}
+This system is integrated into the Appraise MT evaluation platform \citep{federmann12}; the code is open-source and is available on GitHub.\footnote{{\url{https://github.com/Sereni/Appraise}}}
 
-We anticipate that the assessed MT systems will contribute to the users' understanding of text, that is, the users will show better results in tasks when assisted with MT.
+We anticipate that the assessed MT systems will contribute to the users' understanding of text, that is, the users will show better results in gap-filling tasks when assisted with MT.
 We also expect to see different results depending on text domain and the relative number of gaps in a sentence.
@@
 This paper is concerned primarily with assimilation evaluation; for a deeper discussion on evaluation see e.g.
-\citep{koehn2010}.
+\citep[ch.~8]{koehn2010}.
 
@@
 idioms and strong collocations).
-For example, in an English phrase 'Jack ordered <...> and chips', one of the natural answers would be 'fish'.
-Such an answer, however, may be unrelated to the meaning of text, and may be given on the basis of collocation only.
+For example, in an English phrase `Jack ordered <...> and chips', one of the natural answers would be `fish'.
+Such an answer, however, may be unrelated to the meaning of the source text, and may be given on the basis of collocation only.
 \item[Reference sentence and source sentence:] By setup, the
@@
 In our automatic setup, the above procedure is performed by a script integrated into the task generation pipeline.
-Parts of speech are determined with Apertium morphological analyser.
+Parts of speech are determined with Apertium's morphological analysers.
 To control for homonymy, we only allow the word into the candidate list if all of its possible part of speech attributions are on the PoS list.
@@
 in full, see discussion).
-The majority of Russian participants were aged 20-25, with university degrees or in the process of obtaining them.
+The majority of Russian participants were aged 20--25, with university degrees or in the process of obtaining them.
 Although we have not asked the participants about their knowledge of languages other than Tatar and Russian, it is reasonable to assume that most Russian participants knew English to some extent.
-The Spanish participants were university staff with background in Computer Science.
+The Spanish participants were university staff with background in computer science.
 
@@
 Each set features 36 pairs of sentences. For the Basque--Spanish experiment the pairs were drawn
-randomly from the corpora, for Tatar--Russian, compiled by hand by the developer of the
+randomly from the corpora; for Tatar--Russian, compiled by hand by the developer of the
 language pair in Apertium.
@@
 manually distributed between the participants by the following rules:
-\begin{enumerate}
+\begin{enumerate}\itemsep -0.3ex
 \item  Each participant evaluates every sentence (understood as a succession of words),
@@
 increase scores where appropriate.
-On average, the scores improved by three percentage points in all evaluation modes.
+On average, the scores improve by three percentage points in all evaluation modes.
 Candidate synonyms are extracted automatically from the evaluators' responses,  and each individual score is automatically updated according to the synonym list.
@@
 \end{tabular}
-\caption{Average number of gaps successfully filled (\%), using a synonym list} 
+\caption{Average number of gaps successfully filled (\%), using a synonym list, for each language pair in all four task modes.} 
 \label{table:res}
@@
 % expected results
-We expect to see the following trend: scores received in different task modes inside one gap density should be observed, from highest to lowest, in tasks with MT and source hint, then MT hint only, then source hint only, and finally, no hint.
+We expect the scores obtained in different task modes inside one gap density to decrease when going from tasks with MT and source hint to tasks with MT hint only, to tasks with  source hint only, and finally, to tasks with no hint.
 We also expect that with the increase in gap density, the time taken to fill the gaps should also increase, and the percentage of correct answers should decrease.
@@
 In view of the relatively large standard deviations, the significance of the hints' contribution was tested using a linear regression model.
-The data points were represented as an individual evaluator's average score (the number of correct answers divided by the total number of answers) in each of the percentage-hint combinations.
-Two separate models were created: one for no-hint vs MT-hint tasks, and another for no-hint vs source-hint tasks.
-Given the null hypothesis of the regression line slope equal to zero, the contribution of MT hint is significant on the p~<~0.001 level, while the contribution of the source hint is significant only with p~<~0.162
+The data points (\(y\)) were represented as an individual evaluator's average score (the number of correct answers divided by the total number of answers) in each of the percentage-hint combinations.
+Two separate models were created: one for no-hint (\(x=0\)) vs MT-hint (\(x=1\)) tasks, and another for no-hint (\(x=0\)) vs source-hint (\(x=1\)) tasks.
+Given the null hypothesis that the slope \(b\) of the regression line \(y=a+bx\) equals zero, the contribution of MT hint is found to be significant on the \(p<0.001\) level, while the contribution of the source hint is significant only with \(p<0.162\).
 
-Two records in the data do not align with our expectations: the no-hint 10\% sentences in Basque--Spanish, which scored significantly higher than the source-hint in the same category, and MT+source 10\% sentences in Tatar--Russian, which we would have expected to score higher than the corresponding MT task.
+Two records in the data do not align with our expectations: the no-hint 10\% sentences in Basque--Spanish, which scored significantly higher than the source-hint in the same category, and MT+source 10\% sentences in Tatar--Russian, which we would have expected to score higher than the corresponding MT-only task.
 In the first case, this is largely due to the use of synonyms list.
@@
 The second pattern, low scores in Tatar--Russian 10\% MT+source, does not stem from the task content.
-Instead, it is the result of the fixed order of tasks: the participants have always been given MT+source 10\% sentences first, followed by other task types.\comment{MLF: Why?
-Not random?
-\cite{oregan13} did random!
-Why was this changed?}The participants have not received any training tasks before the main evaluations.
+Instead, it is the result of the fixed order of tasks: the participants have always been given MT+source 10\% sentences first, followed by other task types.
+The participants have not received any training tasks before the main evaluations.
 Therefore, it is possible that the accommodation period is responsible for lower-than-expected scores in this mode of evaluation.
 
-It remains questionable whether we can compare results for different gap densities. The 10, 20 and 30\% sets were comprised
-of the same sentences. However, in each case different words were removed. It appears
+It remains questionable whether we can compare results for different gap densities.
+The 10\%, 20\%, and 30\% sets contained the same sentences.
+However, in each case different words were removed.
+It appears
 that some content words are easier to fill than the others. This may explain why in
@@
 \end{tabular}
-\caption{Krippendorff Alpha measure} 
+\caption{Krippendorff Alpha measure of annotator agreement, for each language pair in all four task modes.} 
 \label{table:alpha}
@@
 The interpretation of Krippendorff's alpha varies depending on the application.
-One of the general guidelines suggested by \citet{landis77} for kappa-like measures (which includes Krippendorff's Alpha) is as follows: k < 0 indicates ``poor'' agreement, 0 to 0.2 ``slight'', 0.21 to 0.4 ``fair'', 0.41 to 0.6 ``moderate'', 0.61 to 0.8 ``substantial'', and 0.81 to 1 ``near perfect''.
+One of the general guidelines suggested by \citet{landis77} for kappa-like measures (which includes Krippendorff's Alpha) is as follows: \(k < 0\) indicates ``poor'' agreement, 0 to 0.2 ``slight'', 0.21 to 0.4 ``fair'', 0.41 to 0.6 ``moderate'', 0.61 to 0.8 ``substantial'', and 0.81 to 1 ``near perfect''.
 
@@
 As the MT and MT+source hints are introduced, the agreement increases (measures closer to 1): the annotators are more consistently correct or incorrect in each given sentence.
-The agreement measure for the same sentences without hints is closer to zero, which contributes to the reliability of our methodology.
+The agreement measure for the same sentences without hints is closer to zero, which attests to the reliability of our methodology.
 We note the outlier score in Tatar--Russian 10\% source tasks, which has the most contribution from the news texts.
@@
 \end{tabular}
-\caption{Tatar--Russian Average number of gaps successfully filled (\%), using a synonym list, for three different domains.} 
+\caption{Tatar--Russian Average number of gaps successfully filled (\%), using a synonym list, for three different domains, in all four task modes.} 
 \label{table:rus} 
@@
 any machine translation system. Based on our experience, we would like to suggest the following amendments to the procedure:
-\begin{enumerate}
+\begin{enumerate}\itemsep -0.3ex
 \item As reported by \cite{oregan13}, unless the evaluation is targeted at a specific text domain, it may be beneficial to include a stylistic variety of texts in the initial corpus.
@@
 
-\section*{Acknowledgments}
-
-This work has been partly funded by the Spanish Ministerio de Econom{\'i}a y Competitividad through project TIN2012-32615 and by the Google Summer of Code 2014 programme.
+\paragraph{Acknowledgments:} This work has been partly funded by the Spanish Ministerio de Econom{\'i}a y Competitividad through project TIN2012-32615.
 Ekaterina Ageeva's work has been supported by Google Summer of Code 2014 through the Apertium project.
-We would like to thank the volunteers who participated in the evaluations.
+We  thank the volunteers who participated in the evaluations.
 

### 2015-04-16 2015-eamt-melsa/2015-eamt-melsa.tex
Some stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59737 72bbbca6-d526-0410-a7d9-f06f51895060
added the decipherment references, removed comments, and ironed out text a bit
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59858 72bbbca6-d526-0410-a7d9-f06f51895060

@@
   source-language monolingual corpora, the machine translation (MT)
-  system in which the models are integrated and a statistical
+  system in which the models are integrated, and a statistical
   target-language model. Using the MT system, the sentences in 
@@
 \section{Introduction}
-\todomlf{R1 has a general comment: ``watch out for long sentences''}
+%\todomlf{R1 has a general comment: ``watch out for long sentences''}
 Corpus-based machine translation (MT) has been the primary research
@@
 not the most adequate sense. Thus, it is not necessary to choose
-between a series of fine-grained senses if all these senses result in
+among a series of fine-grained senses if all these senses result in
 the same final translation; however, it may sometimes be necessary to
@@
 in fixed-length contexts, along with a training method for learning
-the rules from a word-aligned parallel corpus.
+the rules from a word-aligned parallel corpus.\footnote{The work by \cite{ravi2011deciphering} and \cite{nuhn2014decipherment}, who decipher word-ciphered text using monolingual corpora only may be seen as a generalised version of the problem of lexical selection without parallel corpora.}
 
@@
 \section{Method}
-\todomlf{R3 says: ``The technique clearly resembles the "Deciphering
-  Foreign Language," (S. Ravi and K. Knight) or “EM Decipherment for
-  Large Vocabularies “ (M. Nuhn, and H. Ney. EM) approaches, although
-  these techniques were not meant for learning the disambiguation but
-  for learning the translation dictionaries.  The proposed method only
-  executes one EM iteration in contrast to these works. It would be
-  interesting to highlight differences and resemblances with these
-  deciphering approaches.''. Check and cite? Also, single iteration
-  WTF?}
+% \todomlf{R3 says: ``The technique clearly resembles the "Deciphering
+%   Foreign Language," (S. Ravi and K. Knight) or “EM Decipherment for
+%   Large Vocabularies “ (M. Nuhn, and H. Ney. EM) approaches, although
+%   these techniques were not meant for learning the disambiguation but
+%   for learning the translation dictionaries.  The proposed method only
+%   executes one EM iteration in contrast to these works. It would be
+%   interesting to highlight differences and resemblances with these
+%   deciphering approaches.''. Check and cite? Also, single iteration
+%   WTF?}
 %\todo[inline]{FSM: In the experiments a development corpus is
@@
 \citep{Manning99b}.  In answer to the second question, a
-maximum-entropy model maximise the entropy subject to match the
+maximum-entropy model maximises the entropy subject to match the
 expected counts of the designed features with those found in the
-training data. \todofsm{Solved R3: ``In answer to the second question, a
-  maximum-entropy model assumes nothing about what is not present in
-  the training data. That is, if there is no information in the
-  training data, then it assumes that all outcomes —that is, all
-  possible translations— are equally likely'' can be applied to a
-  uniform distribution for a discrete variable. I think the authors
-  are trying to restate the main property of the maximum entropy
-  model, i.e.  maximum entropy model maximize the entropy (which is
-  the uniformness of the distribution) subject to match the expected
-  counts of the designed features with those found in the training
-  data.  }
+training data. % \todofsm{Solved R3: ``In answer to the second question, a
+  % maximum-entropy model assumes nothing about what is not present in
+  % the training data. That is, if there is no information in the
+  % training data, then it assumes that all outcomes —that is, all
+  % possible translations— are equally likely'' can be applied to a
+  % uniform distribution for a discrete variable. I think the authors
+  % are trying to restate the main property of the maximum entropy
+  % model, i.e.  maximum entropy model maximize the entropy (which is
+  % the uniformness of the distribution) subject to match the expected
+  % counts of the designed features with those found in the training
+  % data.  }
 %assumes nothing about what is not present in the
@@
   \label{eq:steps}
-  S \to \mbox{\framebox{\parbox{0.8cm}{\textit{pre- lexsel}}}} \to (\{g_i\}_{i=1}^{i=|G|},S) \to \mbox{\framebox{\parbox{1.0cm}{\textit{\textbf{lexsel}}}}} \to (g^\star,S) \to \mbox{\framebox{\parbox{0.8cm}{\textit{post- lexsel}}}} \to \tau(g^\star,S)
+  S \to \mbox{\framebox{\parbox{0.9cm}{\textit{pre- lexsel}}}} \to (\{g_i\}_{i=1}^{i=|G|},S) \to \mbox{\framebox{\parbox{0.9cm}{\textit{\textbf{lexsel}}}}} \to (g^\star,S) \to \mbox{\framebox{\parbox{0.9cm}{\textit{post- lexsel}}}} \to \tau(g^\star,S)
 \end{displaymath}
- \caption{A schema of the training process}
+ \caption{A schema of the lexical selection process: source sentence \(S\) has \(|G|\) lexical selection paths \(g_i\): \emph{lexsel} selects one of them \(g^\star\), which is used to generate translation \(\tau(g^\star,S)\).}
  \label{fig:lexselschema}
@@
 $g = (t_1, t_2, \ldots, t_{|S|})$ is a sequence of lexical-selection
-choices of those SL words, where \(t_i\) is an element of \(T_s(s_i)\), the set of possible translations of \(s_i\).\todomlf{R1 asked for this, even if redundant, and says: ``can one source word be translated as multiple words?
+choices of those SL words, where \(t_i\) is an element of \(T_s(s_i)\), the set of possible translations of \(s_i\).\footnote{We deal only with single-word translations in this paper.}
+% \todomlf{R1 asked for this, even if redundant, and says: ``can one source word be translated as multiple words?
 What about multi-word expressions that cannot be translated word-by-word or source words that do not have a TL equivalent'', etc.} % and \(T_s(\cdot)\)
@@
 schematically. %%%\todo{Can we try to bring the figure closer?}
-\todomlf{R3 says: ``it would have been nicer to see an example of one
-  sentence going through the pipeline''. Can we add this in the text?
-  FSM: We do not have much space}
+% \todomlf{R3 says: ``it would have been nicer to see an example of one
+%   sentence going through the pipeline''. Can we add this in the text?
+%   FSM: We do not have much space}
 %Taking our previous example for the parallel corpus, Table~\ref{table:ex-monocorp} 
@@
 fractional count for disambiguation path \(g_i\), $p(g_i|S)$ after
-suitable normalisation:\todomlf{R2 says ``Nevertheless, I've got
-  problems to understand (4), in particular the paths in post-lexsel
-  step. Some better explanation could be offered maybe by saying that
-  (4) evaluates all possible solutions in the language model.''. Don't
-  know what to do here. }
+suitable normalisation:%\todomlf{R2 says ``Nevertheless, I've got
+  % problems to understand (4), in particular the paths in post-lexsel
+  % step. Some better explanation could be offered maybe by saying that
+  % (4) evaluates all possible solutions in the language model.''. Don't
+  % know what to do here. }
 \begin{equation}
@@
 % \end{table*}
-
-\todomlf{R1 complains: ``the actual maximum entropy method itself is described very briefly (3 sentences at the end of Section 2)''}
+% \todomlf{R1 complains: ``the actual maximum entropy method itself is described very briefly (3 sentences at the end of Section 2)''}
 \section{Experimental setting}
@@
 
-Evaluation will be performed using four language pairs. These pairs
+Evaluation will be performed using four Apertium \citep{forcada2011apertium} language pairs. These pairs
 have been selected as they include languages with different
@@
 --- although for all pairs there is a parallel corpus available
-for evaluation (see Section~\ref{ss:measures}). 
+for evaluation (see Section~\ref{ss:measures}).\footnote{The Apertium revision (version) used is given in footnotes.}
 % Table~\ref{table:lang-pair-stats} 
@@
 \begin{description}
-\addtolength{\itemsep}{-0.10in}
-
-\item[Breton--French:] Development of the Breton--French pair has been
-  described in \citep{tyers10b}. The bilingual % tyers09b
+\addtolength{\itemsep}{-1ex}
+\item[Breton--French]  \citep{tyers10b}: Bilingual
   dictionaries were not built with polysemy in mind from the outset,
   but some entries were added later to start work on lexical
-  selection.\footnote{The version used in this paper is SVN revision
+  selection.\footnote{Revision
   41375; \url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-br-fr}}
@@
 %  and French. It has been developed part-time over a number of years.
-  \todomlf{R1 says ``whole section is rather length especially the
-    revision numbers and URLs of the specific translations systems
-    need not be mentioned in my opinion. The space could rather be
-    used to elaborate in the description of the method in Sec
-    2.''. FSM: I have moved them to the footnote}
-
 \item[Macedonian--English:] The Macedonian--English pair
@@
   using {\sc giza++} \citep{och03a} was checked to ensure that it was
-  an adequate translation, and if so marked as the default.\footnote{A
-    bilingual dictionary in Apertium \citep{forcada2011apertium} may
-    contain several TL translation for a given SL word. Where there is
-    more than one possible translation, the dictionary writer chooses
-    the most general or most frequent translation among the set of
-    possible translations and marks it as the \emph{linguistic
-      default} translation.} As a result of attempting to include all
-  possible translations, the average number of translations per word
-  is much higher % (see table~\ref{table:lang-pair-stats})
-  than in other pairs.\footnote{The version of the software used in this paper is SVN revision
-  41476; \url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-mk-en}}
-\item[Basque--Spanish:] The development of the Basque to Spanish pair
-  is described in \cite{ginesti09}. 
+  an adequate translation, and if so marked as the
+  default.\footnote{Bilingual dictionaries in Apertium
+    \citep{forcada2011apertium} may contain several translations for a
+    given word. Dictionary writers may mark as \emph{linguistic default} the most
+    general or most frequent translation among the set of possible
+    translations.} As a result of attempting to include all possible
+  translations, the average number of translations per word is much
+  higher % (see table~\ref{table:lang-pair-stats})
+  than in other pairs.\footnote{Revision 41476;
+    \url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-mk-en}}
+\item[Basque--Spanish]  \citep{ginesti09}:
 % The bilingual dictionary was taken
 %  from the free/open-source Matxin system \citep{mayor2011}. 
-  Alternative translations were included in the bilingual
-  dictionary.\footnote{The version used is SVN revision 44846; \url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-eu-es}}
+  alternative translations were included in the bilingual
+  dictionary.\footnote{Revision 44846; \url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-eu-es}}
 % , but non-default translations were marked with a
@@
   contains a number of entries in the bilingual dictionary with more
-  than one translation.\footnote{The
-  version used in this paper is SVN revision
+  than one translation.\footnote{Revision
   41387; \url{https://svn.code.sf.net/p/apertium/svn/trunk/apertium-en-es}}
-
 %The most-frequent translation, as judged by
@@
 %  frequency. \todo{Default is explained for some LPs only, why?}  
-\todomlf{4-month mention removed in response to R1}
 \end{description}
@@
 single sentence, we define the lexical selection error rate (LER) of
-that sentence as\todomlf{R3 says: ``The authors use LER, but F-score measures seem more appropriate to estimate the performance of the disambiguation module.
+that sentence as% \todomlf{R3 says: ``The authors use LER, but F-score measures seem more appropriate to estimate the performance of the disambiguation module.
 Moreover, the constraint of the disambiguation module to generate only one word when evaluating seems extreme.'' He don't get it.}
@@
 the lexical selection module $T_t(s_i)$ is empty.  Recall
-that,\todomlf{R1 says ``this has already been mentioned
-  before''. Don't know what to do FSM: I would do nothing, this is why
-the sentence says ``Recall ...''} although $T_t(s_i)$ returns a set,
+that, %\todomlf{R1 says ``this has already been mentioned
+%  before''. Don't know what to do FSM: I would do nothing, this is why
+%the sentence says ``Recall ...''} 
+although $T_t(s_i)$ returns a set,
 this set will be a singleton, as when the lexical-selection module
-returns more than one translation the first one will be
-selected.\footnote{In practice this does not happen as each ambiguous
-  word has a \emph{default} translation.}\todomlf{R1 complains: ``why
-  mention if it doesn't happen in practice? Relevance?'' What should
-  we do?}
+returns more than one translation, Apertium will select the default one if marked or the first one of not.\footnote{In practice this does not happen as each ambiguous
+  word has a \emph{default} translation.}%\todomlf{R1 complains: ``why
+  % mention if it doesn't happen in practice? Relevance?'' What should
+  % we do?}
 
@@
 The table in Figure~\ref{fig:ler-example} gives an
-overview of the inputs.\todomlf{Caption of figure extended in response to R3} In the description it is assumed that the
+overview of the inputs.%\todomlf{Caption of figure extended in response to R3} 
+In the description it is assumed that the
 reference translation has been annotated by hand. However, hand
@@
 \subsubsection{Machine translation performance}
-%! \comment{This part of the text needs compressing and merging} 
+
 This is an extrinsic evaluation, which ideally would test how much the
@@
 translation quality in a real system. For this task, we use the
-widely-used {\sc bleu} metric \citep{papineni02}.\todomlf{BLEU reference advanced, and wrong section link removed} This is not ideal for evaluating the
+widely-used {\sc bleu} metric \citep{papineni02}.
+%\todomlf{BLEU reference advanced, and wrong section link removed} 
+This is not ideal for evaluating the
 task of a lexical selection module as the performance of the module
@@
   \url{http://downloads.sourceforge.net/project/apertium/apertium-lex-tools/apertium-lex-tools-0.1.0.tar.gz}.}
-to calculate the feature weights.  \todomlf{R3 says: ``The proposed
-  maximum entropy or log-linear model is trained in a surprising
-  way. In the early days of maximum entropy models, those were trained
-  using the GIS algorithm. However, it has been proved in the
-  literature that GIS has clear disadvantages with respect to other
-  batch gradient descent algorithms. For instance, it does not allow
-  for the use of regularization [...], and it is extremely slow to
-  converge, which yields worse models than other methods. The
-  smoothing/regularization of the model is also poor since it is
-  simply feature pruning instead of regularization.'' FSM: I would say
-  something in the discussion (future work).}  After learning the
+to calculate the feature weights.  %\todomlf{R3 says: ``The proposed
+  % maximum entropy or log-linear model is trained in a surprising
+  % way. In the early days of maximum entropy models, those were trained
+  % using the GIS algorithm. However, it has been proved in the
+  % literature that GIS has clear disadvantages with respect to other
+  % batch gradient descent algorithms. For instance, it does not allow
+  % for the use of regularization [...], and it is extremely slow to
+  % converge, which yields worse models than other methods. The
+  % smoothing/regularization of the model is also poor since it is
+  % simply feature pruning instead of regularization.'' FSM: I would say
+  % something in the discussion (future work).}
+After learning the
 feature sets and weights, we compute the evaluation measures described
@@
  may be as a result of having the linguistic defaults tuned to a
- different domain than that of the corpus.\todomlf{R1 says the
-   footnote is important and should be incorporated to the text; what
-   should we do? FSM: I have moved it to the main text and split the
-   paragraph into two pieces.}
+ different domain than that of the corpus.
 
@@
 an exponential number of lexical selections for each sentence using
-the TL model online.\todomlf{R2 says: ``An error analysis would also
-  help to understand the proposal.'' We don't have time now I think.}
+the TL model online.%\todomlf{R2 says: ``An error analysis would also
+%  help to understand the proposal.'' We don't have time now I think.}
 
@@
 
-\paragraph{Acknowledgements:} We acknowledge support from the Spanish Ministry of Industry and Competitiveness through project Ayutra (TIC2012-32615) and from the European Commission through project Abu-Matran (FP7-PEOPLE-2012-IAPP, ref.\ 324414) and thank all three anonymous referees for very useful comments on the paper.\todomlf{Fran: add your acknowledgements}
+\paragraph{Acknowledgements:} We acknowledge support from the Spanish Ministry of Industry and Competitiveness through project Ayutra (TIC2012-32615) and from the European Commission through project Abu-Matran (FP7-PEOPLE-2012-IAPP, ref.\ 324414) and thank all three anonymous referees for useful comments on the paper.%\todomlf{Fran: add your acknowledgements}
 

### 2015-04-28 2015-eamt-melsa/poster/2015-eamt-melsa-poster.tex
\dend{columns}
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60163 72bbbca6-d526-0410-a7d9-f06f51895060

\documentclass[25pt, a0paper, portrait, margin=0mm, innermargin=15mm,blockverticalspace=15mm, colspace=15mm, subcolspace=8mm]{tikzposter}

\usepackage{fontspec}
\usepackage{multirow}

\defaultfontfeatures{PunctuationSpace=3,Scale=MatchLowercase,Mapping=tex-text}

\setromanfont{Liberation Serif}
\setmonofont{DejaVu Sans Mono}

\settitle{ \centering \vbox{
%\@titlegraphic \\[\TP@titlegraphictotitledistance] \centering
\color{titlefgcolor} {\bfseries \Huge \sc \@title \par}
\vspace*{1em}
{\huge \@author \par} \vspace*{1em} {\LARGE \@institute}
}}

\makeatletter
\def\title#1{\gdef\@title{\scalebox{\TP@titletextscale}{%
\begin{minipage}[t]{0.7\linewidth}
\centering
#1
\par
\vspace{0.5em}
\end{minipage}%
}}}
\makeatother

\title{Unsupervised training of maximum-entropy models 
 for lexical selection in rule-based machine translation}
 \institute{$^\star$UiT Norgga árktalaš universitehta, $^\dagger$Universitat d'Alacant} % See Section 4.1
\author{Francis M. Tyers,$^\star$ Felipe Sánchez-Martínez,$^\dagger$ Mikel L. Forcada$^\dagger$} 
%\titlegraphic{
%\includegraphics[width=0.04\textwidth]{LogoSamisk.png} \hspace{20pt}
%\includegraphics[width=0.15\textwidth]{logodlsioficialcolor.eps} 
%}
\usetheme{Basic} % See Section 5
\begin{document}

\maketitle % See Section 4.1
\begin{columns} % See Section 4.4
\column{0.45} % See Section 4.4
\block{BlocktitleB}{Blocktext}
\column{0.55}
\block{BlocktitleC}{
\begin{center}
\includegraphics[width=0.5\textwidth]{architecture-lexsel.pdf}
\end{center}
}
%\note{Notetext} % See Section 4.3
\end{columns}
\begin{columns} % See Section 4.4
\column{0.45} % See Section 4.4
\block{BlocktitleB}{\begin{displaymath}
  \label{eq:steps}
  S \to \mbox{\framebox{\parbox{4.0cm}{\textit{pre-lexsel}}}} \to (\{g_i\}_{i=1}^{i=|G|},S) \to \mbox{\framebox{\parbox{3.5cm}{\textit{\textbf{lexsel}}}}} \to (g^\star,S) \to \mbox{\framebox{\parbox{4.0cm}{\textit{post- lexsel}}}} \to \tau(g^\star,S)
\end{displaymath}
}

\column{0.55}
\block{BlocktitleA}{ % See Section 4.2
\useinnerblockstyle{Table}
\begin{center}
\begin{tabular}{|c|l|r|}
     \hline
        & \textbf{Sentence} & $p(g_i|S)$ \\
     \hline % Los pescados grandes peque\~{n}o come
      \(S_1\) & \emph{Arrain handiak txikia jaten du}  &  \\
                         $\tau(g_{11},S_1)$ & El pez grande se come el peque\~{n}o &  0.830 \\
                         $\tau(g_{12},S_1)$ & El pescado grande se come el peque\~{n}o & 0.134 \\
                         $\tau(g_{13},S_1)$ & El pescado capaz se come el peque\~{n}o & 0.034\\
                         $\tau(g_{14},S_1)$ & El pez capaz se come el peque\~{n}o  & 0.001\\
     \hline % El padre el pescado nos ha preparado para cenar
      \(S_2\) & \emph{Aitak arraina prestatu digu afaltzeko} &  \\
                         $\tau(g_{21},S_2)$ & Padre nos ha preparado pescado para cenar & 0.846 \\
                         $\tau(g_{22},S_2)$ & Pap\'{a} nos ha preparado pescado para cenar & 0.134 \\
                         $\tau(g_{23},S_2)$ & Padre nos ha preparado pez para cenar & 0.015 \\
                         $\tau(g_{24},S_2)$ & Pap\'{a} nos ha preparado pez para cenar & 0.005 \\
     \hline % Aqu\'{\i} el pescado muy es dulce
     $S_3$  & \emph{Hemen arraina oso goxoa da}  & \\
                         $\tau(g_{31},S_3)$ & Aqu\'{\i} el pescado es muy rico & 0.912 \\
                         $\tau(g_{32},S_3)$ & Aqu\'{\i} el pescado es muy suave & 0.067 \\
                         $\tau(g_{33},S_3)$ & Aqu\'{\i} el pez es muy rico & 0.011 \\
                         $\tau(g_{34},S_3)$ & Aqu\'{\i} el pez es muy suave & 0.008 \\
                         $\tau(g_{35},S_3)$ & Aqu\'{\i} el pez es muy dulce & 0.001 \\
                         $\tau(g_{36},S_3)$ & Aqu\'{\i} el pescado es muy dulce & 0.001 \\
     \hline % Un pescado grande del mar es
     \(S_4\) & \emph{Itsasoko arrain handi bat da} & \\
                         $\tau(g_{41},S_4)$ & Es un pez grande del mar & 0.595 \\
                         $\tau(g_{42},S_4)$ & Es un pescado grande del mar & 0.403 \\
                         $\tau(g_{43},S_4)$ & Es un pez capaz del mar & 0.001 \\
                         $\tau(g_{44},S_4)$ & Es un pescado capaz del mar & 0.001 \\
     \hline % El pescado grande por esos serv\'{\i}an para alimentar
     $S_5$ &  \emph{Arrain handiak horiez baliatzen ziren elikatzeko} &  \\
                         $\tau(g_{51},S_5)$ & Los peces grandes se nutr\'{\i}an de esos & 0.862 \\
                         $\tau(g_{52},S_5)$ & Los pescados grandes se nutr\'{\i}an de esos & 0.121 \\
                         $\tau(g_{53},S_5)$ & Los peces grandes se alimentaban de esos & 0.012 \\
                         $\tau(g_{54},S_5)$ & Los peces capaces se alimentaban de esos & 0.001 \\
                         $\tau(g_{55},S_5)$ & Los pescados grandes se alimentaban de esos & 0.001 \\
                         $\tau(g_{56},S_5)$ & Los pescados capaces se alimentaban de esos & 0.001 \\
                         $\tau(g_{57},S_6)$ & Los peces capaces se nutr\'{\i}an de esos & 0.001 \\
                         $\tau(g_{58},S_7)$ & Los pescados capaces se nutr\'{\i}an de esos & 0.001 \\
     \hline

   \end{tabular}
\end{center}
}
\end{columns}
\block{Results}{ % See Section 4.2
 \begin{center}

  \begin{tabular}{|l|l|r|r|r||r|}
    \hline
    \multirow{2}{*}{{\bf Pair}}  & \multirow{2}{*}{{\bf Metric}} & \multicolumn{4}{|c|}{{\bf System}} \\ \cline{3-6}
                                 &              & {\tt Ling} & {\tt TLM} & \texttt{MaxEnt} & \texttt{Oracle} \\
    \hline % tlm-alig: [44.5, 50.2]
    \multirow{2}{*}{{\bf br-fr}} & {\sc ler} (\%)     & [54.8, 60.7] & [44.2, 50.5]  & {\bf [40.8, 46.9]} & [0.0, 0.0]      \\ 
                                 & {\sc bleu} (\%)    & [14.5, 16.4] & [15.4, 17.3]  & [14.8, 16.6] & [16.7, 18.6]     \\ 
    \hline % tlm-alig: 
    \multirow{2}{*}{{\bf mk-en}} & {\sc ler} (\%)     & [28.8, 32.6] & [26.8, 30.5]  & [25.2, 28.8] & [0.0, 0.0]    \\ 
                                 & {\sc bleu} (\%)    & [28.6, 31.0] & [30.7, 32.3]  & [29.1, 31.5] & [30.9, 33.3]    \\ 
    \hline % tlm-alig: 
    \multirow{2}{*}{{\bf eu-es}} & {\sc ler} (\%)      & [43.6, 48.8] & [38.8, 44.2]  & [40.9, 46.2] & [0.0, 0.0]     \\ 
                                 & {\sc bleu} (\%)     & [10.1, 12.0] & [10.6, 12.6]  & [10.3, 12.2] & [11.5, 13.5]     \\ 
    \hline % tlm-alig: [10.3, 13.9]
    \multirow{2}{*}{{\bf en-es}} & {\sc ler} (\%)      & [20.5, 24.9] & [15.1, 18.9]  & {\bf [10.4, 13.8]} & [0.0, 0.0]     \\ 
                                 & {\sc bleu} (\%)     & [21.5, 23.4] & [21.9, 23.8]  & {\bf [22.2, 24.1]} & [22.8, 24.7]     \\ 
    \hline
  \end{tabular}

 \end{center}

}
\end{document}

### 2015-04-29 2015-nodalida-ruscg/ruscg.tex
stuff for final draft
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60168 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-%\author{Francis M. Tyers \\
-%  HSL-fakultehta, \\
-%  UiT Norgga árktalaš universitehta, \\
-%  N-9018 Romsa \\
-%  {\tt francis.tyers@uit.no} \\\And
-%  Robert Reynolds \\
-%  HSL-fakultehta, \\
-%  UiT Norgga árktalaš universitehta, \\
-%  N-9018 Romsa \\
-%  {\tt robert.reynolds@uit.no} \\}
-
- \author{Author1 \\
-   Affiliation / Address line 1 \\
-   Affiliation / Address line 2 \\
-   Affiliation / Address line 3 \\
-   {\tt email@domain.com} \\\And
-   Author2 \\
-   Affiliation / Address line 1 \\
-   Affiliation / Address line 2 \\
-   Affiliation / Address line 3 \\
-   {\tt email@domain.com} \\}
+\author{Francis M. Tyers \\
+  HSL-fakultehta, \\
+  UiT Norgga árktalaš universitehta, \\
+  N-9018 Romsa \\
+  {\tt francis.tyers@uit.no} \\\And
+  Robert Reynolds \\
+  HSL-fakultehta, \\
+  UiT Norgga árktalaš universitehta, \\
+  N-9018 Romsa \\
+  {\tt robert.reynolds@uit.no} \\}
 
@@
  low precision.
+ \rr{more precise results here}
 \end{abstract}
@@
 stress patterns, exceptions, and idiosyncratic collocations.
-We base our work on the recently developed [anonymous] morphological transducer, which is also based on Zaliznjak's dictionary (Author, 2015).\footnote{The [anonymous] transducer is implemented using a two-level
+We developed a morphological transducer based on Zaliznjak's dictionary.\footnote{Our transducer is implemented using a two-level 
 morphology \cite{Koskenniemi-84}, and can be compiled using either {\tt xfst}
@@
 intraparadigmatic, morphosyntactically incongruent, and morphosyntactically 
-congruent. A recent study by Author (2015) found that about 7.5\% of 
+congruent. The following examples make use of stress ambiguity to illustrate 
+each kind of ambiguity.\footnote{A recent study by \newcite{reynolds.tyers-15} 
+found that about 7.5\% of 
 morphosyntactic ambiguity in a corpus of Russian resulted in stress placement 
-ambiguity. The following examples make use of stress ambiguity to illustrate 
-each kind of ambiguity. \emph{Intraparadigmatic} ambiguity refers to homographic 
+ambiguity.} \emph{Intraparadigmatic} ambiguity refers to homographic 
 wordforms belonging to the same lexeme, as shown in \ref{ex:intrahom}.
@@
 
-Table \ref{table:ambiguity} shows the proportional prevalence of each kind of 
+Table \ref{table:ambiguity} shows the prevalence of each kind of 
 ambiguity. The first column shows the proportion 
@@
 incongruent ambiguity. Detailed part-of-speech tagging with morphosyntactic 
-analysis can help disambiguate these forms. Morphosyntactically congruent stress 
+analysis can help disambiguate these forms. However, morphosyntactically congruent stress 
 ambiguity cannot be disambiguated by means of detailed part-of-speech tagging, 
@@
 
-\begin{table}
+\begin{table}[ht]
   \centering
@@
 
-The morphological transducer used in this study is described in Author (2015). 
-It is primarily based on Zaliznjak's \emph{Grammatical dictionary of Russian}, 
+The morphological transducer used in this study is 
+primarily based on Zaliznjak's \emph{Grammatical dictionary of Russian}, 
 including the 2001 version's appendix of proper nouns. It also includes neologisms
 from Grishina and Lyashevskaya's \emph{Grammatical dictionary of new Russian words},
-which is intended to be a supplement to Zaliznjak's dictionary.\footnote{\url{http://dict.ruslang.ru/gram.php}}
-
+which is intended to be a supplement to Zaliznjak's dictionary with words found in
+the Russian National Corpus.\footnote{\url{http://dict.ruslang.ru/gram.php}} 
 Example~\ref{ex:FSToutput} gives some examples of the FST's output.
@@
 
-The constraint grammar is composed of 299 rules which
+The constraint grammar is composed of 299 rules which are divided into four
+categories: Safe, Safe heuristic, Heuristic, and Syntax labelling. The distribution
+of rules is shown in Table~\ref{tab:ruleDist}.
 
-\begin{table}
+\begin{table}[hb]
   \centering
@@
       on rule reliability. }
+  \label{tab:ruleDist}
 \end{table}
@@
 
-% Examples where this might go wrong.
+% Examples where this might go wrong:
+% This rule can fail in dialogs where the preposition is explicit in the previous turn.
 
@@
 position in the sentence it cannot modify anything before, and no preposition 
-can be governing it.
-
-This kind of rule often relies on completeness of sets, in this case the set of verbs
-that can take a genitive complement.
+can be governing it. This kind of rule often relies on completeness of sets, 
+in this case the set of verbs that can take a genitive complement.
 
@@
 
-% Examples where this might go wrong.
+% Examples where this might go wrong:
+% Russian has relatively free word order. Technically there is no grammatical limitation
+% from having a genitive noun modifying a noun elsewhere in the sentence.
 
@@
 
-% Examples where this might go wrong.
+% Examples where this might go wrong:
+% Anywhere that a verbal adverb is possible, especially if there are no feminine nominals
+% in the sentence.
 
@@
 from Wikipedia articles, public domain literature and freely-available news sources. The
-annotated texts are available online under the {\sc cc-by-sa} licence.\footnote{\url{http://...}}
+annotated texts are available online under the {\sc cc-by-sa} licence.\footnote{\url{https://svn.code.sf.net/p/apertium/svn/languages/apertium-rus/texts/}}
+\rr{is this a good url? Do we need to put a COPYING file in that folder?}
 
@@
 The constraint grammar described in this paper has been applied to the task of automatic
-word stress placement (Author, 2015). This task is especially relevant for 
+word stress placement \cite{reynolds.tyers-15}. This task is especially relevant for 
 Russian language learners, because vowels are pronounced differently depending on their position 
@@
 Almost 99\% of wordforms with ambiguous stress position can be disambiguated morphosyntactically,
-so a constraint grammar can potentially resolve most stress ambiguity indirectly. The results of Author 2015 show that our constraint grammar overcomes about 42\% of the ambiguity relevant to stress
+so a constraint grammar can potentially resolve most stress ambiguity indirectly.
+The results of \newcite{reynolds.tyers-15} show that our constraint grammar overcomes about 42\% of the ambiguity relevant to stress
 ambiguity in unrestricted text.
@@
 Note that the final accuracy is below the state of the art for Russian, however the state of
-the art employs unavailable training corpora of over five million tokens.
+the art employs unavailable training corpora of over five million tokens.\rr{citations? This comes out of the blue.}
 
@@
 
-% better development workflow
-
-% proof-of-concept for russian, but russian has loads of tagged corpora, for other langs, e.g. sorbian or belarusian 
 % a CG+hmm could be much faster than tagging hundreds of thousands of words of text.
@@
 We have a number of plans for future work, the first is increasing the precision of the grammar without
-decreasing recall. Secondly we would like to add syntactic function labelling and dependency analysis. For 
-the dependency analysis we would like to reuse the Giellatekno dependency grammar as in \cite{antonsen10}.
+decreasing recall. Secondly we would like to add syntactic function labelling and dependency parsing. For 
+the dependency parse we plan to reuse the Giellatekno dependency grammar as in \cite{antonsen10}.
 
@@
 \section*{Acknowledgments}
-\textit{Removed for review}
 
-%Do not number the acknowledgment section. Do not include this section
-%when submitting your paper for review.
+We are grateful to Alan Turing, Richard Stallman, Steve Jobs, and Laura Janda. Without these
+people, this paper would not have been completed the way that it was. \rr{Revise acknowledgments!}
 

### 2015-04-30 2015-nodalida-ruscg/ruscg.tex
adding reviewer comments as marginpars
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60197 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{linguex}
+\usepackage[multiple]{footmisc} % separates consecutive footnotes with comma
 %\usepackage{mathptmx}
@@
  low precision.
- \rr{more precise results here}
+ \rr{add results here}
 \end{abstract}
@@
 
-This paper presents a preliminary constraint grammar for Russian. The principle 
+This paper presents a preliminary constraint grammar for Russian. The main 
 objective of the constraint grammar is to produce a high recall grammar to serve
 as input into other natural language processing tasks.
+\rr{R3: What is the rationale for choosing high recall at the expense of low precision?
+Was it a deliberate choice or a consequence of limited development time? In my
+experience, this strategy is good for corpus queries in linguistic research
+(fewer relevant constructions are missed) but might be problematic for language
+technology applications, which normally require full disambiguation. Since it is
+stated in the introduction that the principal objective of the grammar is to
+produce input to other language processing tasks, this question is especially
+relevant and warrants some discussion.}
 
@@
 congruent. The following examples make use of stress ambiguity to illustrate 
-each kind of ambiguity.\footnote{A recent study by \newcite{reynolds.tyers-15} 
+each kind of ambiguity.\footnote{Written standard Russian does not typically
+indicate stress position.}\footnote{A recent study by \newcite{reynolds.tyers-15} 
 found that about 7.5\% of 
@@
 	\rus{замк\'{о}в} \emph{zamk\'{o}v} `lock.\textsubscript{PL-GEN}'
-\c. ...\\
-	...
+\c. etc. \ldots\\ \rr{is this clear that the paradigm continues, but we don't show it?}
 
@@
 incongruent ambiguity. Detailed part-of-speech tagging with morphosyntactic 
-analysis can help disambiguate these forms. However, morphosyntactically congruent stress 
+analysis can help disambiguate these forms. However, morphosyntactically congruent
 ambiguity cannot be disambiguated by means of detailed part-of-speech tagging, 
@@
 from the Russian Wikipedia, ran it through the morphological analyser, wrote
+\rr{R1: Is ru-wikipedia representative of Russian texts on the whole?}
 rules, and then ran the rules on the whole Wikipedia corpus. For each rule,
@@
 %FRAN
+\rr{R1 does not understand "intrinsic/extrinsic" (and neither do I) Should we rename them to be qualitative/quantitative?}
+\rr{R1: Add some text introducing intrinsic evaluation. What is this section doing?}
 
@@
 
+\rr{R1: Tables 3 and 4 are not discussed enough.}
 % unknown words from test corpus, categorisation
@@
     \end{itemize}
+    \rr{R2: The first two rules in 6.2 are a bit odd - the effect would be exactly the same without the first context bracket. Did you mean 0C instead of 0?}
     did not take into account the possibility of having a postposed determiner as in
@@
 
+\rr{R1: Note that Johannessen et al (2012) and Lynum et al (2011) have written two articles about building a statistical tagger on top of a cg grammar.
+
+Johannessen, Janne Bond; Hagen, Kristin; Lynum, André; Nøklestad, Anders.
+2012.
+OBT+stat: A combined rule-based and statistical tagger.
+In Andersen, Gisle (ed.).
+Exploring Newspaper Language: Using the web to create and investigate a large corpus of modern Norwegian, s.
+51–66.
+
+Lynum, Andre; Hagen, Kristin; Johannessen, Janne Bondi; Nøklestad, Anders. OBT+Stat: Evaluation of a combined CG and statistical tagger. In NEALT Proceedings Series 2011 Vol. 14.}
+
 \begin{itemize}
@@
 \end{figure}
-
+\rr{R1: Figure 3 is small and hard to read. Shapes indistinguishable.}
+\rr{R2: If I interpret Fig.
+3 correctly, the HMM system alone is much more dependent on training corpus size than the voting set-up, where accuracies actually appear to become asymptotic already with 8-10,000 words in the training corpus, indicating that further improvement depends more on the CG than the HMM-component.}
 \section{Future work}
@@
 state-of-the-art performance for the stress-placement task.
+\rr{R2: I particularly like the idea of systematically assigning rules to heuristicity sections based on performance thresholds derived from annotation-inspection of a non-gold corpus.
+Another novel idea (at least in terms of actual implementation) is to go for 100\% recall and add a hybridized HMM-tagger in order to improve precision.}
 

### 2015-04-30 2015-eamt-melsa/poster/2015-eamt-melsa-poster.tex
more stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60200 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{multirow}
+\usepackage{url}
+\usepackage{alltt}
 
@@
 \column{0.55}
-\block{BlocktitleC}{
+\block{Apertium (\url{www.apertium.org})}{
   \begin{center}
@@
   \end{center}
+  \begin{itemize}
+    \item Free/open-source RBMT platform
+    \item Around 40 released language pairs
+  \end{itemize}
 }
@@
 \end{displaymath}
+}
+\block{Result comparison}{
+
+% Pertsona Fisikoen Errentaren gaineko Zerga (PFEZ).
+% Impuesto sobre la Renta de las Personas Físicas (IRPF).
+El impuesto superior de la renta de [-los mortal físicos-] \{+las personas físicas+\} (PFEZ) .  % 835
+
+% Beste langile baten ordez aritzea, eta laneko sarrera- eta irteera-erregistroak eta kontrolak aldaraztea.
+% La suplantación de otro trabajador, alterando los registros y controles de entrada y salida al trabajo.
+De un otro [-obrero-] {+trabajador+} reemplazar estar , y la entrada del [-curro-] {+trabajo+} y el registro de salida y el control aldaraztea .  % 781
+
+% 34. artikulua.- Babes Publikoko Etxebizitzen motak.
+% Artículo 34. Tipos de Vivienda de Protección Pública.
+34. El artículo. El tipo de las viviendas [-del amparo público-] {+de la protección pública+} . % 767
+
+% Hezkuntza-administrazioak ikastetxe publikoetako irakasleen plantilla, lege honetan ezarritako helburuen arabera egokituko du.
+% La Administración educativa adecuará la plantilla docente de los centros públicos a los objetivos establecidos en la presente Ley.
+La administración de educación de los profesores de los colegios públicos plantilla , según [-las metas puestas-] {+los objetivos establecidos+} en esta ley adecuará .  % 746
+
+% Hirugarren egunetik Lehen Hezkuntzan ingelesari eta euskarazi dagozkion bajak.
+% Desde el tercer día las bajas correspondientes a Inglés y Euskera en Educación Primaria.
+Desde el tercer día en la [-primera-] educación {+primaria+} al inglés y euskarazi le la baja que están . % 427
+
+% 16/1985 Legea, Espainiako Ondare Historikoarena.
+% Ley 16/1985 de Patrimonio Histórico Español
+Las 16/1985 leyes , el del [-acervo-] {+patrimonio+} histórico de España . % 317
+
+% Lehen hezkuntzan matrikulatuta dauden haurrek bai, badute eskubide hori; derrigorrezko bigarren hezkuntzan daudenek, ordea, ez.
+% Sí lo tienen los niños matriculados en Educación Primaria, no los de Educación Secundaria Obligatoria.
+En la [-primera-] educación {+primaria+} matrikulatuta los niños que están sí , si tienen ese derecho ; obligado en la [-segunda-] educación {+secundaria+} los que están , sin embargo , no .  % 265
+
+% Bigarren Hezkuntzan erabiltzen den hizkuntza Lehen Hezkuntzakoa baina zailagoa da.
+% que sería que el lenguaje que se utiliza en Secundaria es más difícil que el de Primaria.
+En la [-segunda-] educación {+secundaria+} el idioma que utiliza el de la [-primera-] educación {+primaria+} pero es más [-difícil-] {+complicado+} .  % 240
+
+}
+
+% -- weights should go on operations
+%  <rule weight="0.906306">
+%    <match lemma="ondare" tags="n"><select lemma="acervo" tags="n.*"/></match>
+%    <match lemma="historiko" tags="adj.*"/>
+%  </rule>
+%  <rule weight="1.08987">
+%    <match lemma="ondare" tags="n"><select lemma="patrimonio" tags="n.*"/></match>
+%    <match lemma="historiko" tags="adj.*"/>
+%  </rule>
+%
+
+\block{Training}{
+
+\begin{alltt}
+0 \$ 0.830 \# arrain:0 arrain\_handi:0 \# arrain:1 arrain\_handi:1  \# \\
+1 \$ 0.134 \# arrain:0 arrain\_handi:0 \# arrain:1 arrain\_handi:1  \# \\
+1 \$ 0.034 \# arrain:0 arrain\_handi:0 \# arrain:1 arrain\_handi:1  \# \\
+0 \$ 0.001 \# arrain:0 arrain\_handi:0 \# arrain:1 arrain\_handi:1  \# \\
+0 \$ 0.015 \# arrain:0 aita\_arrain:0 arrain\_prestatu:0 \# arrain:1 aita\_arrain:1 arrain\_prestatu:1 \# \\
+0 \$ 0.005 \# arrain:0 aita\_arrain:0 arrain\_prestatu:0 \# arrain:1 aita\_arrain:1 arrain\_prestatu:1 \# \\
+1 \$ 0.846 \# arrain:0 aita\_arrain:0 arrain\_prestatu:0 \# arrain:1 aita\_arrain:1 arrain\_prestatu:1 \# \\
+1 \$ 0.134 \# arrain:0 aita\_arrain:0 arrain\_prestatu:0 \# arrain:1 aita\_arrain:1 arrain\_prestatu:1 \# \\
+\end{alltt}
+\ldots
+}
+
+\block{Features}{
+\begin{alltt}
+1.1356   arrain:pez \\
+0.90723  arrain:pescado \\
+1.28196  aita\_arrain:pescado \\
+0.167523 aita\_arrain:pez \\
+0.49622  arrain:pescado\_handi \\
+1.37525  arrain:pez\_handi\\
+1.28196  arrain:pescado\_oso \\
+0.167523 arrain:pez\_oso \\
+1.28196  arrain:pescado\_prestatu \\
+0.167523 arrain:pez\_prestatu \\
+1.28196  hemen\_arrain:pescado \\
+0.167523 hemen\_arrain:pez \\
+1.40647  itsaso\_arrain:pescado \\
+0.8466   itsaso\_arrain:pez \\
+\end{alltt}
 }

### 2015-05-03 2015-eamt-melsa/poster/2015-eamt-melsa-poster.tex
update poster
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60210 72bbbca6-d526-0410-a7d9-f06f51895060

--- 2015-04-28 2015-eamt-melsa/poster/2015-eamt-melsa-poster.tex
@@
 \usepackage{multirow}
+\usepackage{url}
+\usepackage{alltt}
+\usepackage{color}
+%\usepackage[table]{xcolor}
 
@@
 
-\setromanfont{Liberation Serif}
+\setromanfont[Scale=1.1]{Liberation Serif}
 \setmonofont{DejaVu Sans Mono}
+
+\definecolor{darkgreen}{RGB}{0,127,0}
 
@@
 \vspace*{1em}
-{\huge \@author \par} \vspace*{1em} {\LARGE \@institute}
+{\LARGE \@author \par} \vspace*{1em} {\Large \@institute}
 }}
@@
 \def\title#1{\gdef\@title{\scalebox{\TP@titletextscale}{%
-\begin{minipage}[t]{0.7\linewidth}
+\begin{minipage}[t]{1.0\linewidth}
 \centering
@@
 %}
-\usetheme{Basic} % See Section 5
+%\usetheme{Basic} % See Section 5
 \begin{document}
@@
 \begin{columns} % See Section 4.4
-\column{0.45} % See Section 4.4
-\block{BlocktitleB}{Blocktext}
-\column{0.55}
-\block{BlocktitleC}{
+
+\column{0.50} % See Section 4.4
+
+\block[titleinnersep=10em]{Lexical selection}{
+
+  \begin{itemize}
+    \item Choosing the most adequate TL 
+translation for an SL word from a 
+set of possible translations.
+    \item b
+  \end{itemize}
+
+}
+
+  \block[titleinnersep=10em]{Apertium (\url{www.apertium.org})}{
     \begin{center}
-\includegraphics[width=0.5\textwidth]{architecture-lexsel.pdf}
+      \includegraphics[width=0.45\textwidth]{architecture-lexsel.pdf}
+    \end{center}
+    \begin{itemize}
+      \item Free/open-source RBMT platform
+      \item Around 40 released language pairs
+    \end{itemize}
+  }
+% -- weights should go on operations
+%  <rule weight="0.906306">
+%    <match lemma="ondare" tags="n"><select lemma="acervo" tags="n.*"/></match>
+%    <match lemma="historiko" tags="adj.*"/>
+%  </rule>
+%  <rule weight="1.08987">
+%    <match lemma="ondare" tags="n"><select lemma="patrimonio" tags="n.*"/></match>
+%    <match lemma="historiko" tags="adj.*"/>
+%  </rule>
+%
+
+  \block[titleinnersep=10em]{Training schema}{
+
+%\begin{displaymath}
+%    \label{eq:steps}
+%    S \to \mbox{\framebox{\parbox{4.0cm}{\textit{pre-lexsel}}}} \to (\{g_i\}_{i=1}^{i=|G|},S) \to \mbox{\framebox{\parbox{3.5cm}{\textit{\textbf{lexsel}}}}} \to (g^\star,S) \to \mbox{\framebox{\parbox{4.0cm}{\textit{post- lexsel}}}} \to \tau(g^\star,S)
+%  \end{displaymath}
+
+  \begin{center}
+    \includegraphics[width=0.45\textwidth]{schema.eps}
   \end{center}
   }
-%\note{Notetext} % See Section 4.3
-\end{columns}
-\begin{columns} % See Section 4.4
-\column{0.45} % See Section 4.4
-\block{BlocktitleB}{\begin{displaymath}
-  \label{eq:steps}
-  S \to \mbox{\framebox{\parbox{4.0cm}{\textit{pre-lexsel}}}} \to (\{g_i\}_{i=1}^{i=|G|},S) \to \mbox{\framebox{\parbox{3.5cm}{\textit{\textbf{lexsel}}}}} \to (g^\star,S) \to \mbox{\framebox{\parbox{4.0cm}{\textit{post- lexsel}}}} \to \tau(g^\star,S)
-\end{displaymath}
-}
 
-\column{0.55}
-\block{BlocktitleA}{ % See Section 4.2
+\column{0.5}
+
+\block[titleinnersep=10em]{Training}{ % See Section 4.2
+
+Translate SL sentences to TL using existing MT system and score the translations.
+
 \useinnerblockstyle{Table}
@@
       \(S_1\) & \emph{Arrain handiak txikia jaten du}  &  \\
-                         $\tau(g_{11},S_1)$ & El pez grande se come el peque\~{n}o &  0.830 \\
-                         $\tau(g_{12},S_1)$ & El pescado grande se come el peque\~{n}o & 0.134 \\
-                         $\tau(g_{13},S_1)$ & El pescado capaz se come el peque\~{n}o & 0.034\\
-                         $\tau(g_{14},S_1)$ & El pez capaz se come el peque\~{n}o  & 0.001\\
+                         $\tau(g_{1},S_1)$ & El pez grande se come el peque\~{n}o &  0.831 \\
+                         $\tau(g_{2},S_1)$ & El pescado grande se come el peque\~{n}o & 0.169 \\
      \hline % El padre el pescado nos ha preparado para cenar
       \(S_2\) & \emph{Aitak arraina prestatu digu afaltzeko} &  \\
-                         $\tau(g_{21},S_2)$ & Padre nos ha preparado pescado para cenar & 0.846 \\
-                         $\tau(g_{22},S_2)$ & Pap\'{a} nos ha preparado pescado para cenar & 0.134 \\
-                         $\tau(g_{23},S_2)$ & Padre nos ha preparado pez para cenar & 0.015 \\
-                         $\tau(g_{24},S_2)$ & Pap\'{a} nos ha preparado pez para cenar & 0.005 \\
+                         $\tau(g_{1},S_2)$ & Padre nos ha preparado pescado para cenar & 0.980 \\
+                         $\tau(g_{2},S_2)$ & Padre nos ha preparado pez para cenar & 0.020 \\
      \hline % Aqu\'{\i} el pescado muy es dulce
      $S_3$  & \emph{Hemen arraina oso goxoa da}  & \\
-                         $\tau(g_{31},S_3)$ & Aqu\'{\i} el pescado es muy rico & 0.912 \\
-                         $\tau(g_{32},S_3)$ & Aqu\'{\i} el pescado es muy suave & 0.067 \\
-                         $\tau(g_{33},S_3)$ & Aqu\'{\i} el pez es muy rico & 0.011 \\
-                         $\tau(g_{34},S_3)$ & Aqu\'{\i} el pez es muy suave & 0.008 \\
-                         $\tau(g_{35},S_3)$ & Aqu\'{\i} el pez es muy dulce & 0.001 \\
-                         $\tau(g_{36},S_3)$ & Aqu\'{\i} el pescado es muy dulce & 0.001 \\
+                         $\tau(g_{1},S_3)$ & Aqu\'{\i} el pescado es muy rico & 0.980 \\
+                         $\tau(g_{2},S_3)$ & Aqu\'{\i} el pez es muy rico & 0.020 \\
      \hline % Un pescado grande del mar es
      \(S_4\) & \emph{Itsasoko arrain handi bat da} & \\
-                         $\tau(g_{41},S_4)$ & Es un pez grande del mar & 0.595 \\
-                         $\tau(g_{42},S_4)$ & Es un pescado grande del mar & 0.403 \\
-                         $\tau(g_{43},S_4)$ & Es un pez capaz del mar & 0.001 \\
-                         $\tau(g_{44},S_4)$ & Es un pescado capaz del mar & 0.001 \\
+                         $\tau(g_{1},S_4)$ & Es un pez grande del mar & 0.596 \\
+                         $\tau(g_{2},S_4)$ & Es un pescado grande del mar & 0.404 \\
      \hline % El pescado grande por esos serv\'{\i}an para alimentar
      $S_5$ &  \emph{Arrain handiak horiez baliatzen ziren elikatzeko} &  \\
-                         $\tau(g_{51},S_5)$ & Los peces grandes se nutr\'{\i}an de esos & 0.862 \\
-                         $\tau(g_{52},S_5)$ & Los pescados grandes se nutr\'{\i}an de esos & 0.121 \\
-                         $\tau(g_{53},S_5)$ & Los peces grandes se alimentaban de esos & 0.012 \\
-                         $\tau(g_{54},S_5)$ & Los peces capaces se alimentaban de esos & 0.001 \\
-                         $\tau(g_{55},S_5)$ & Los pescados grandes se alimentaban de esos & 0.001 \\
-                         $\tau(g_{56},S_5)$ & Los pescados capaces se alimentaban de esos & 0.001 \\
-                         $\tau(g_{57},S_6)$ & Los peces capaces se nutr\'{\i}an de esos & 0.001 \\
-                         $\tau(g_{58},S_7)$ & Los pescados capaces se nutr\'{\i}an de esos & 0.001 \\
+                         $\tau(g_{1},S_5)$ & Los peces grandes se nutr\'{\i}an de esos & 0.876 \\
+                         $\tau(g_{2},S_5)$ & Los pescados grandes se nutr\'{\i}an de esos & 0.124 \\
      \hline
@@
 }
+
+\block[titleinnersep=10em]{Model}{
+
+Extract $n$-gram contexts and train MaxEnt model.
+
+\begin{center}
+\begin{tabular}{|l|r|l|r|}
+\hline
+$h^s(t,c)$ & $\lambda$ & $h^s(t,c)$ & $\lambda$ \\
+\hline
+arrain:pez & 1.1356 & arrain:pescado & 0.90723 \\
+aita arrain:pez & 0.167523 & aita arrain:pescado & 1.28196 \\
+arrain:pez handi & 1.37525 & arrain:pescado handi & 0.49622 \\ 
+arrain:pez oso & 0.167523 & arrain:pescado oso & 1.28196  \\ 
+arrain:pez prestatu & 0.167523 & arrain:pescado prestatu & 1.28196  \\
+hemen arrain:pez & 0.167523 & hemen arrain:pescado & 1.28196 \\ 
+itsaso arrain:pez & 0.8466  & itsaso arrain:pescado & 1.40647  \\
+
+\hline
+\end{tabular}
+\end{center}
+%1.1356   arrain:pez \\
+%0.90723  arrain:pescado \\
+%1.28196  aita\_arrain:pescado \\
+%0.167523 aita\_arrain:pez \\
+%0.49622  arrain:pescado\_handi \\
+%1.37525  arrain:pez\_handi\\
+%1.28196  arrain:pescado\_oso \\
+%0.167523 arrain:pez\_oso \\
+
+%1.28196  arrain:pescado\_prestatu \\
+%0.167523 arrain:pez\_prestatu \\
+%1.28196  hemen\_arrain:pescado \\
+%0.167523 hemen\_arrain:pez \\
+%1.40647  itsaso\_arrain:pescado \\
+%0.8466   itsaso\_arrain:pez \\
+%\end{alltt}
+}
+
 \end{columns}
-\block{Results}{ % See Section 4.2
+
+\block[titleinnersep=10em]{« Gaur arraina prestatuko dut. »}{
+
+Compile features into finite-state transducer.
+
+    \begin{center}
+      \includegraphics[width=0.70\textwidth]{transducer2.eps}
+    \end{center}
+\vspace{-20pt}
+    arrain:pez = 1.303123 //
+    arrain:pescado = 2.18919
+
+The translation with the highest probability is one with the highest sum of the weights.
+
+\begin{equation}
+  \underset{t \in T_s(s)}{\arg\max} ~ p_s(t|c) = \underset{t \in T_s(s)}{\arg\max} \sum_{k=1}^{n_F} \lambda_k^s h_k^s(t, c)
+\label{eq:max-ent-max}
+\end{equation}
+
+  }
+
+\begin{columns}
+
+%\block[titleinnersep=10em]{Training}{
+%
+%\begin{alltt}
+%0 \$ 0.830 \# arrain:0 arrain\_handi:0 \# arrain:1 arrain\_handi:1  \# \\
+%1 \$ 0.134 \# arrain:0 arrain\_handi:0 \# arrain:1 arrain\_handi:1  \# \\
+%1 \$ 0.034 \# arrain:0 arrain\_handi:0 \# arrain:1 arrain\_handi:1  \# \\
+%0 \$ 0.001 \# arrain:0 arrain\_handi:0 \# arrain:1 arrain\_handi:1  \# \\
+%0 \$ 0.015 \# arrain:0 aita\_arrain:0 arrain\_prestatu:0 \# arrain:1 aita\_arrain:1 arrain\_prestatu:1 \# \\
+%0 \$ 0.005 \# arrain:0 aita\_arrain:0 arrain\_prestatu:0 \# arrain:1 aita\_arrain:1 arrain\_prestatu:1 \# \\
+%1 \$ 0.846 \# arrain:0 aita\_arrain:0 arrain\_prestatu:0 \# arrain:1 aita\_arrain:1 arrain\_prestatu:1 \# \\
+%1 \$ 0.134 \# arrain:0 aita\_arrain:0 arrain\_prestatu:0 \# arrain:1 aita\_arrain:1 arrain\_prestatu:1 \# \\
+%\end{alltt}
+%\ldots
+%}
+%
+\column{0.55}
+
+\block[titleinnersep=10em]{Result comparison}{
+
+% Pertsona Fisikoen Errentaren gaineko Zerga (PFEZ).
+% Impuesto sobre la Renta de las Personas Físicas (IRPF).
+% El impuesto superior de la renta de [-los mortal físicos-] \{+las personas físicas+\} (PFEZ) .  % 835
+
+% Beste langile baten ordez aritzea, eta laneko sarrera- eta irteera-erregistroak eta kontrolak aldaraztea.
+\begin{center}
+\begin{tabular}{|l|}
+%De un otro \textcolor{red}{-obrero} reemplazar estar, y la entrada del \textcolor{red}{-curro} y el registro de salida y el control aldaraztea. \\  % 781 
+%De un otro \textcolor{blue}{+trabajador} reemplazar estar, y la entrada del \textcolor{blue}{+trabajo} y el registro de salida y el control aldaraztea.\\  % 781 
+%La suplantación de otro \textcolor{darkgreen}{trabajador}, alterando los registros y controles de entrada y salida al \textcolor{darkgreen}{trabajo}. \\
+%
+% Hezkuntza-administrazioak ikastetxe publikoetako irakasleen plantilla, lege honetan ezarritako helburuen arabera egokituko du.
+\hline
+%La administración de educación de los profesores de los colegios públicos plantilla, según \textcolor{red}{-las metas puestas} en esta ley adecuará. \\ % 746
+%La administración de educación de los profesores de los colegios públicos plantilla, según \textcolor{blue}{+los objetivos establecidos} en esta ley adecuará. \\ % 746
+%La Administración educativa adecuará la plantilla docente de los centros públicos a \textcolor{darkgreen}{los objetivos establecidos} en la presente Ley.
+%\ldots de los colegios públicos plantilla, según \textcolor{red}{-las metas puestas} en esta ley adecuará. \\ % 746
+%\ldots de los colegios públicos plantilla, según \textcolor{blue}{+los objetivos establecidos} en esta ley adecuará. \\ % 746
+%\ldots de los centros públicos a \textcolor{darkgreen}{los objetivos establecidos} en la presente Ley. \\
+
+%\hline
+% Lehen hezkuntzan matrikulatuta dauden haurrek bai, badute eskubide hori; derrigorrezko bigarren hezkuntzan daudenek, ordea, ez.
+En la \textcolor{red}{-primera educación} matrikulatuta los niños que están sí, si tienen ese derecho; \\ 
+~~~~~ obligado en la \textcolor{red}{-segunda educación} los que están, sin embargo, no.  \\  % 265
+En la \textcolor{blue}{+educación primaria} matrikulatuta los niños que están sí, si tienen ese derecho; \\ 
+~~~~~ obligado en la \textcolor{blue}{+educación secundaria} los que están, sin embargo, no. \\  % 265
+Sí lo tienen los niños matriculados en \textcolor{darkgreen}{Educación Primaria}, \\
+~~~~ no los de \textcolor{darkgreen}{Educación Secundaria} Obligatoria.\\
+\hline
+% Bigarren Hezkuntzan erabiltzen den hizkuntza Lehen Hezkuntzakoa baina zailagoa da.
+% que sería que el lenguaje que se utiliza en Secundaria es más difícil que el de Primaria.
+%En la [-segunda-] educación {+secundaria+} el idioma que utiliza el de la [-primera-] educación {+primaria+} pero es más [-difícil-] {+complicado+} .  % 240
+
+% 34. artikulua.- Babes Publikoko Etxebizitzen motak.
+34. El artículo. El tipo de las viviendas \textcolor{red}{-del amparo público}. \\ % 767
+34. El artículo. El tipo de las viviendas \textcolor{blue}{+de} la \textcolor{blue}{+protección pública}. \\ % 767
+Artículo 34. Tipos de Vivienda \textcolor{darkgreen}{de Protección Pública}.\\
+\hline
+% 16/1985 Legea, Espainiako Ondare Historikoarena.
+Las 16/1985 leyes , el del \textcolor{red}{-acervo} histórico de España. \\ % 317 
+Las 16/1985 leyes , el del \textcolor{blue}{+patrimonio} histórico de España.  \\ % 317 
+Ley 16/1985 de \textcolor{darkgreen}{Patrimonio} Histórico Español \\
+\hline
+\end{tabular}
+\end{center}
+
+}
+
+\column{0.45}
+
+\block[titleinnersep=10em]{Results}{ % See Section 4.2
  \begin{center}
@@
 }
+
+\end{columns}
+
 \end{document}

### 2015-05-03 2015-eamt-melsa/poster/2015-eamt-melsa-poster.tex
more
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60211 72bbbca6-d526-0410-a7d9-f06f51895060

@@
   \begin{itemize}
-    \item Choosing the most adequate TL 
-translation for an SL word from a 
-set of possible translations.
-    \item b
+    \item Lexical selection is:
+    \begin{itemize}
+    \item Choosing the most adequate 
+translation for a word from a 
+set of possibilities.
+    \end{itemize}
+    \item Why unsupervised ? 
+    \begin{itemize}
+      \item For many language pairs, parallel corpora are not available
+    \end{itemize}
+    \item Why maximum entropy ? 
+    \begin{itemize}
+      \item Allows linguists to define features for which weights are learnt.
+    \end{itemize}
   \end{itemize}

### 2015-05-04 2015-nodalida-ruscg/ruscg.tex
adding references
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60216 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 ambiguity.
+Similar approaches have been taken by previous researchers with Basque \cite{ezeiza.ea-98}, Czech \cite{hajic.ea-01,hajic.ea-07}, Norwegian \cite{johannessen.ea-11,johannessen.ea-12}, Spanish \cite{hulden12}, and Turkish \cite{oflazer.tur-96}.
 
@@
 experimented with three configurations:
-
-\rr{R1: Note that Johannessen et al (2012) and Lynum et al (2011) have written two articles about building a statistical tagger on top of a cg grammar.
-
-Johannessen, Janne Bond; Hagen, Kristin; Lynum, André; Nøklestad, Anders.
-2012.
-OBT+stat: A combined rule-based and statistical tagger.
-In Andersen, Gisle (ed.).
-Exploring Newspaper Language: Using the web to create and investigate a large corpus of modern Norwegian, s.
-51–66.
-
-Lynum, Andre; Hagen, Kristin; Johannessen, Janne Bondi; Nøklestad, Anders. OBT+Stat: Evaluation of a combined CG and statistical tagger. In NEALT Proceedings Series 2011 Vol. 14.}
 

### 2015-05-04 2015-nodalida-ruscg/ruscg.tex
revising
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60217 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 objective of the constraint grammar is to produce a high recall grammar to serve
-as input into other natural language processing tasks.
-\rr{R3: What is the rationale for choosing high recall at the expense of low precision?
-Was it a deliberate choice or a consequence of limited development time? In my
-experience, this strategy is good for corpus queries in linguistic research
-(fewer relevant constructions are missed) but might be problematic for language
-technology applications, which normally require full disambiguation. Since it is
-stated in the introduction that the principal objective of the grammar is to
-produce input to other language processing tasks, this question is especially
-relevant and warrants some discussion.}
+as input into other natural language processing tasks. There are two reasons to
+maintain high recall. First, one of the primary applications for this constraint grammar
+is computer-assisted language learning. In the domain, erroneous analyses can lead to
+significant frustration for learners. Therefore, it is important to limit disambiguation
+to cases that can be resolved with high confidence.
+Second, it is frequently the case that competing readings can
+be distinguished only by considering idiosyncratic collocational information. For such cases, 
+we expect that probabilistic approaches are both more effective and simpler to implement.
 
@@
 \b. \rus{з\'{a}мков} \emph{z\'{a}mkov} `castle.\textsubscript{PL-GEN}'\\
-	\rus{замк\'{о}в} \emph{zamk\'{o}v} `lock.\textsubscript{PL-GEN}'
-\c. etc. \ldots\\ \rr{is this clear that the paradigm continues, but we don't show it?}
+	\rus{замк\'{о}в} \emph{zamk\'{o}v} `lock.\textsubscript{PL-GEN}'\\
+    etc. \ldots
+
+\rr{is this clear that the paradigm continues, but we don't show it? One of the reviewers thought the example was incomplete}
 
@@
 from Wikipedia articles, public domain literature and freely-available news sources. The
-annotated texts are available online under the {\sc cc-by-sa} licence.\footnote{\url{https://svn.code.sf.net/p/apertium/svn/languages/apertium-rus/texts/}}
-\rr{is this a good url? Do we need to put a COPYING file in that folder?}
+annotated texts are available online under the {\sc cc-by-sa} licence.\footnote{\url{https://svn.code.sf.net/p/apertium/svn/languages/apertium-rus/texts/}}\rr{is this a good url?
+Do we need to put a COPYING file in that folder?}
 
@@
 
-\subsection{Intrinsic}
+\subsection{Qualitative evaluation}
 %FRAN
-\rr{R1 does not understand "intrinsic/extrinsic" (and neither do I) Should we rename them to be qualitative/quantitative?}
-\rr{R1: Add some text introducing intrinsic evaluation. What is this section doing?}
+
+In this section, we give a qualitative evaluation of errors made by the CG.
 
@@
 
-\subsection{Extrinsic}
+\subsection{Quantitative evaluation}
 % ROB accent placement task
@@
 Almost 99\% of wordforms with ambiguous stress position can be disambiguated morphosyntactically,
-so a constraint grammar can potentially resolve most stress ambiguity indirectly.
-The results of \newcite{reynolds.tyers-15} show that our constraint grammar overcomes about 42\% of the ambiguity relevant to stress
-ambiguity in unrestricted text.
+so a constraint grammar can potentially resolve most stress ambiguity indirectly. The results of 
+\newcite{reynolds.tyers-15} show that our constraint grammar overcomes about 42\% of the ambiguity 
+relevant to stress ambiguity in unrestricted text.
 
@@
 experiment with combining the constraint grammar with a statistical tagger to resolve remaining
-ambiguity.
-Similar approaches have been taken by previous researchers with Basque \cite{ezeiza.ea-98}, Czech \cite{hajic.ea-01,hajic.ea-07}, Norwegian \cite{johannessen.ea-11,johannessen.ea-12}, Spanish \cite{hulden12}, and Turkish \cite{oflazer.tur-96}.
+ambiguity. Similar approaches have been taken by previous researchers with Basque \cite{ezeiza.ea-98}, 
+Czech \cite{hajic.ea-01,hajic.ea-07}, Norwegian \cite{johannessen.ea-11,johannessen.ea-12}, 
+Spanish \cite{hulden12}, and Turkish \cite{oflazer.tur-96}.
 

### 2015-05-04 2015-nodalida-ruscg/ruscg.tex
revising more
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60220 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{linguex}
-\usepackage[multiple]{footmisc} % separates consecutive footnotes with comma
+%\usepackage[multiple]{footmisc} % separates consecutive footnotes with comma
 %\usepackage{mathptmx}
@@
 intraparadigmatic, morphosyntactically incongruent, and morphosyntactically 
-congruent. The following examples make use of stress ambiguity to illustrate 
+congruent. The following examples make use of word stress ambiguity to illustrate 
 each kind of ambiguity.\footnote{Written standard Russian does not typically
-indicate stress position.}\footnote{A recent study by \newcite{reynolds.tyers-15} 
-found that about 7.5\% of 
-morphosyntactic ambiguity in a corpus of Russian resulted in stress placement 
+indicate stress position, but knowing stress position is essential for pronunciation. 
+A recent study by \newcite{reynolds.tyers-15} found that about 7.5\% of 
+morphosyntactic ambiguity in a corpus of Russian resulted in stress position 
 ambiguity.} \emph{Intraparadigmatic} ambiguity refers to homographic 
@@
 that belong to separate lexemes, and whose morphosyntactic values are identical, 
-as shown in \ref{ex:MScongruent}. As indicated in Table~\ref{table:ambiguity}, 
-this kind of ambiguity is relatively rare, and resolving this ambiguity requires 
-the use of technologies such as word sense disambiguation.
+as shown in \ref{ex:MScongruent}.
 
@@
 incongruent ambiguity. Detailed part-of-speech tagging with morphosyntactic 
-analysis can help disambiguate these forms. However, morphosyntactically congruent
-ambiguity cannot be disambiguated by means of detailed part-of-speech tagging, 
-and since congruent stress ambiguity represents only a small percentage of 
-ambiguous wordforms, we leave it to future work.
+analysis can help disambiguate these forms. On the other hand, morphosyntactically congruent
+ambiguity represents only a very small percentage of ambiguous wordforms, and instead 
+of detailed part-of-speech tagging, it can be resolved by means of word sense disambiguation.
+Because of this difference, we leave morphosyntactically congruent ambiguity to future work.
 
@@
 \ex. \label{ex:FSToutput} 
-	\a. \rus{новый}\texttt{{\small <adj><m><nn><sg><nom>}}
-	\b. \rus{автомат}\texttt{{\small <n><m><nn><sg><nom>}}
+	\a. \rus{новый}\texttt{{\small <adj><m><nn><sg><nom>}}\\
+	    `new'
+	\b. \rus{автомат}\texttt{{\small <n><m><nn><sg><nom>}}\\
+	    `automaton, sub-machine gun'
 %	\c. 
@@
 The constraint grammar is composed of 299 rules which are divided into four
-categories: Safe, Safe heuristic, Heuristic, and Syntax labelling. The distribution
+categories: Safe, Safe heuristic, Heuristic, and Syntax labeling. The distribution
 of rules is shown in Table~\ref{tab:ruleDist}.
@@
 
-The philosophy is that safe rules should represent real constraints in
+The philosophy is that Safe rules should represent real constraints in
 the language. Examples might be that a preposition cannot directly precede a finite verb
-or that prepositional case requires a preceeding preposition.
+or that prepositional case requires a preceding preposition.
 

### 2015-05-04 2015-nodalida-ruscg/ruscg.tex
more stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60223 72bbbca6-d526-0410-a7d9-f06f51895060

@@
  low precision.
- \rr{add results here}
+ \rr{add results to abstract}
 \end{abstract}
@@
 state-of-the-art performance for the stress-placement task.
-\rr{R2: I particularly like the idea of systematically assigning rules to heuristicity sections based on performance thresholds derived from annotation-inspection of a non-gold corpus.
+\rr{Emphasize R2's thoughts more in conclusion: I particularly like the idea of systematically assigning rules to heuristicity sections based on performance thresholds derived from annotation-inspection of a non-gold corpus.
 Another novel idea (at least in terms of actual implementation) is to go for 100\% recall and add a hybridized HMM-tagger in order to improve precision.}
@@
 
-We are grateful to Alan Turing, Richard Stallman, Steve Jobs, and Laura Janda. Without these
-people, this paper would not have been completed the way that it was. \rr{Revise acknowledgments!}
+We are grateful to Koen Claessen for insightful discussion, as well as three anonymous reviewers who 
+gave thoughtful feedback on an earlier version of this paper. All remaining errors are our own. 
+\rr{Revise acknowledgments!}
 

### 2015-05-04 2015-nodalida-ruscg/ruscg.tex
more fixes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60226 72bbbca6-d526-0410-a7d9-f06f51895060

@@
  analyser and constraint grammar. The constraint 
- grammar is tuned to be high recall at the expense of 
+ grammar is tuned to be high recall (over 0.99) at the expense of 
  low precision.
- \rr{add results to abstract}
+% \rr{add results to abstract}
 \end{abstract}
@@
 
-\begin{table}[hb]
+\begin{table}
   \centering
@@
 analyser, we adopted a slightly modified technique. We picked a random text
-from the Russian Wikipedia, ran it through the morphological analyser, wrote
-\rr{R1: Is ru-wikipedia representative of Russian texts on the whole?}
+from the Russian Wikipedia,\footnote{The Russian Wikipedia was chosen as a testing corpus
+  as it is the largest, freely licensed corpus of Russian available on the internet. It is 
+  not representative of Russian texts as a whole.}
+ran it through the morphological analyser, wrote
+%\rr{R1: Is ru-wikipedia representative of Russian texts on the whole?}
 rules, and then ran the rules on the whole Wikipedia corpus. For each rule,
@@
 from Wikipedia articles, public domain literature and freely-available news sources. The
-annotated texts are available online under the {\sc cc-by-sa} licence.\footnote{\url{https://svn.code.sf.net/p/apertium/svn/languages/apertium-rus/texts/}}\rr{is this a good url?
-Do we need to put a COPYING file in that folder?}
+annotated texts are available online under the {\sc cc-by-sa} licence.\footnote{\url{https://svn.code.sf.net/p/apertium/svn/languages/apertium-rus/texts/}}
+%\rr{is this a good url? Do we need to put a COPYING file in that folder?}
 
@@
 
-\begin{figure}
+\begin{figure*}
   \centering
-  \includegraphics[width=0.45\textwidth]{graphics/learning-curve.pdf}
+  \includegraphics[width=0.70\textwidth]{graphics/learning-curve.pdf}
   \caption{Learning curve for three taggers, \texttt{hunpos} with no lexicon, 
@@
   \label{fig:curve}
-\end{figure}
-\rr{R1: Figure 3 is small and hard to read. Shapes indistinguishable.}
+\end{figure*}
 \rr{R2: If I interpret Fig.

### 2015-05-04 2015-eamt-melsa/poster/2015-eamt-melsa-poster.tex
update
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60212 72bbbca6-d526-0410-a7d9-f06f51895060
poster fixes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60231 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-\setromanfont[Scale=1.1]{Liberation Serif}
+\setromanfont[Scale=1.1]{FreeSans}
+%\setromanfont[Scale=1.1]{Liberation Serif}
 \setmonofont{DejaVu Sans Mono}
@@
 
-\settitle{ \centering \vbox{
-%\@titlegraphic \\[\TP@titlegraphictotitledistance] \centering
-\color{titlefgcolor} {\bfseries \Huge \sc \@title \par}
+%\settitle{ \centering \vbox{
+%\color{titlefgcolor} {\bfseries \Huge \sc \@title \par}
+%\vspace*{1em}
+%{\Large \@author \par} \vspace*{1em} {\large \@institute}
+%}\@titlegraphic \\[\TP@titlegraphictotitledistance] \centering
+%}
+
+\settitle{ 
+  \vbox{
+    \color{titlefgcolor} {\bfseries \Huge  \@title \par}
+    \vspace*{-0.5em}
+    \begin{tabular}{llr}
+    {\Large \@author \par} & ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~  &  \multirow{3}{*}{\@titlegraphic } \\ % [\TP@titlegraphictotitledistance]}  \\
+    & ~ & ~   \\
+    {\large \@institute}   & ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ &  \\
+    \end{tabular}
     \vspace*{1em}
-{\LARGE \@author \par} \vspace*{1em} {\Large \@institute}
-}}
+  } 
+}
 
@@
 \author{Francis M. Tyers,$^\star$ Felipe Sánchez-Martínez,$^\dagger$ Mikel L. Forcada$^\dagger$} 
-%\titlegraphic{
-%\includegraphics[width=0.04\textwidth]{LogoSamisk.png} \hspace{20pt}
-%\includegraphics[width=0.15\textwidth]{logodlsioficialcolor.eps} 
-%}
+\titlegraphic{
+\includegraphics[width=0.04\textwidth]{LogoSamisk.png} \hspace{25pt} \includegraphics[width=0.15\textwidth]{logodlsioficialcolor.eps} \hspace{10pt} \includegraphics[width=0.05\textwidth,angle=90]{apertium5a.pdf}
+}
 %\usetheme{Basic} % See Section 5
+%\usebackgroundstyle{Rays}
 \begin{document}
 
-\maketitle % See Section 4.1
+\maketitle[titletotopverticalspace=25pt,titletoblockverticalspace=25pt,width=0.96\textwidth] % See Section 4.1
 \begin{columns} % See Section 4.4
@@
 translation for a word from a 
-set of possibilities.
+set of possible translations from a bilingual dictionary.
     \end{itemize}
@@
     \begin{itemize}
-      \item Allows linguists to define features for which weights are learnt.
+      \item Allows linguists to freely define relevant features for which weights are automatically learnt.
     \end{itemize}
@@
     \begin{itemize}
-      \item Free/open-source RBMT platform
+      \item Free/open-source rule-based MT platform
       \item Around 40 released language pairs
@@
   \begin{center}
-    \includegraphics[width=0.45\textwidth]{schema.eps}
+    \includegraphics[width=0.34\textwidth]{schema.eps}
   \end{center}
@@
 
-Translate SL sentences to TL using existing MT system and score the translations.
+Translate SL sentences to TL using existing MT system and score the translations using an $n$-gram TL model.
 
@@
 
-\block[titleinnersep=10em]{« Gaur arraina prestatuko dut. »}{
+\block[titleinnersep=10em]{\emph{Gaur arraina prestatuko dut.} `Today I'll cook fish.'}{
 
@@
 \vspace{-20pt}
-    arrain:pez = 1.303123 //
-    arrain:pescado = 2.18919
 
@@
 \begin{equation}
-  \underset{t \in T_s(s)}{\arg\max} ~ p_s(t|c) = \underset{t \in T_s(s)}{\arg\max} \sum_{k=1}^{n_F} \lambda_k^s h_k^s(t, c)
-\label{eq:max-ent-max}
+p_s(t|c) = \frac{1}{Z} \exp{ \sum_{k=1}^{n_F} \lambda_k^s h_k^s(t, c)}
+\label{eq:max-ent-sum}
 \end{equation}
+
+    arrain:pez $ = 1.1356 + 0.16752 = 1.303123$
+
+    arrain:pescado $ = 0.90723 + 1.28196 = 2.18919$
+%\begin{equation}
+%  \underset{t \in T_s(s)}{\arg\max} ~ p_s(t|c) = \underset{t \in T_s(s)}{\arg\max} \sum_{k=1}^{n_F} \lambda_k^s h_k^s(t, c)
+%\label{eq:max-ent-max}
+%\end{equation}
 
@@
 \begin{center}
-\begin{tabular}{|l|}
+\begin{tabular}{|l|l|}
 %De un otro \textcolor{red}{-obrero} reemplazar estar, y la entrada del \textcolor{red}{-curro} y el registro de salida y el control aldaraztea. \\  % 781 
@@
 % Lehen hezkuntzan matrikulatuta dauden haurrek bai, badute eskubide hori; derrigorrezko bigarren hezkuntzan daudenek, ordea, ez.
-En la \textcolor{red}{-primera educación} matrikulatuta los niños que están sí, si tienen ese derecho; \\ 
-~~~~~ obligado en la \textcolor{red}{-segunda educación} los que están, sin embargo, no.  \\  % 265
-En la \textcolor{blue}{+educación primaria} matrikulatuta los niños que están sí, si tienen ese derecho; \\ 
-~~~~~ obligado en la \textcolor{blue}{+educación secundaria} los que están, sin embargo, no. \\  % 265
-Sí lo tienen los niños matriculados en \textcolor{darkgreen}{Educación Primaria}, \\
-~~~~ no los de \textcolor{darkgreen}{Educación Secundaria} Obligatoria.\\
+& En la \textcolor{red}{primera educación} matrikulatuta los niños que están sí, si tienen ese  \\ 
+& ~~ derecho; obligado en la \textcolor{red}{segunda educación} los que están, sin embargo, no.  \\  % 265
+& En la \textcolor{blue}{educación primaria} matrikulatuta los niños que están sí, si tienen ese  \\ 
+& ~~ derecho; obligado en la \textcolor{blue}{educación secundaria} los que están, sin embargo, no. \\  % 265
+ref & Sí lo tienen los niños matriculados en \textcolor{darkgreen}{Educación Primaria}, \\
+    &~~ no los de \textcolor{darkgreen}{Educación Secundaria} Obligatoria.\\
 \hline
@@
 % 34. artikulua.- Babes Publikoko Etxebizitzen motak.
-34. El artículo. El tipo de las viviendas \textcolor{red}{-del amparo público}. \\ % 767
-34. El artículo. El tipo de las viviendas \textcolor{blue}{+de} la \textcolor{blue}{+protección pública}. \\ % 767
-Artículo 34. Tipos de Vivienda \textcolor{darkgreen}{de Protección Pública}.\\
+& 34. El artículo. El tipo de las viviendas \textcolor{red}{del amparo público}. \\ % 767
+& 34. El artículo. El tipo de las viviendas \textcolor{blue}{de} la \textcolor{blue}{protección pública}. \\ % 767
+ref & Artículo 34. Tipos de Vivienda \textcolor{darkgreen}{de Protección Pública}.\\
 \hline
 % 16/1985 Legea, Espainiako Ondare Historikoarena.
-Las 16/1985 leyes , el del \textcolor{red}{-acervo} histórico de España. \\ % 317 
-Las 16/1985 leyes , el del \textcolor{blue}{+patrimonio} histórico de España.  \\ % 317 
-Ley 16/1985 de \textcolor{darkgreen}{Patrimonio} Histórico Español \\
+& Las 16/1985 leyes , el del \textcolor{red}{acervo} histórico de España. \\ % 317 
+& Las 16/1985 leyes , el del \textcolor{blue}{patrimonio} histórico de España.  \\ % 317 
+ref & Ley 16/1985 de \textcolor{darkgreen}{Patrimonio} Histórico Español \\
 \hline

### 2015-05-04 2015-eamt-melsa/poster/2015-eamt-melsa-poster.tex
poster stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60232 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{multirow}
+\usepackage{multicol}
 \usepackage{url}
@@
   \begin{center}
-    \includegraphics[width=0.34\textwidth]{schema.eps}
+    \includegraphics[width=0.40\textwidth]{schema.eps}
   \end{center}
@@
      \hline
-
    \end{tabular}
 \end{center}
+The scores are normalised to produce fractional counts.
 }
@@
 
-Extract $n$-gram contexts and train MaxEnt model.
+Extract $n$-gram contexts and train MaxEnt model using fractional counts.
 
@@
 \block[titleinnersep=10em]{\emph{Gaur arraina prestatuko dut.} `Today I'll cook fish.'}{
+%Compile features into finite-state transducer. 
+  \begin{multicols*}{2}
+\textbf{Input:} (SL $\rightarrow$ TL)
+~\\
+~\\
+\begin{tabular}{ll}
+~~  &  \texttt{gaur<adv>} $\rightarrow$ \texttt{hoy<adv>}  \\ 
+~~  &  \texttt{arrain<n>} $\rightarrow$ \texttt{pescado<n><m><ND> | pez<n><m><ND>} \\ 
+~~  &  \texttt{prestatu<vblex><pfut>} $\rightarrow$ \texttt{preparar<vblex><pfut>} \\
+~~  &  \texttt{ukan<vbsint><pri><NR\_HU><NK\_NI>} $\rightarrow$ \texttt{tener<vblex><pri>}
+\end{tabular}
+~\\
+~\\
+~\\
+Paths in the rule transducer are traversed and the weights of \\
+matching features are summed:
+\begin{displaymath}
+p_s(t|c) = \frac{1}{Z} \exp{ \sum_{k=1}^{n_F} \lambda_k^s h_k^s(t, c)}
+\label{eq:max-ent-sum}
+\end{displaymath}
 
-Compile features into finite-state transducer.
+    arrain:pez  = 1.1356 + 0.16752 = \textbf{1.303123}
+
+    arrain:pescado  = 0.90723 + 1.28196 = \textbf{2.18919}
+
+~\\
+
+The translation with the highest sum of weights is the most probable.
+
+      \includegraphics[width=0.40\textwidth]{transducer3.eps}
 
 \begin{center}
-      \includegraphics[width=0.70\textwidth]{transducer2.eps}
+      \textbf{Rule transducer}
 \end{center}
-\vspace{-20pt}
 
-The translation with the highest probability is one with the highest sum of the weights.
-
-\begin{equation}
-p_s(t|c) = \frac{1}{Z} \exp{ \sum_{k=1}^{n_F} \lambda_k^s h_k^s(t, c)}
-\label{eq:max-ent-sum}
-\end{equation}
-
-    arrain:pez $ = 1.1356 + 0.16752 = 1.303123$
-
-    arrain:pescado $ = 0.90723 + 1.28196 = 2.18919$
+\end{multicols*}
 %\begin{equation}
@@
 & ~~ derecho; obligado en la \textcolor{blue}{educación secundaria} los que están, sin embargo, no. \\  % 265
-ref & Sí lo tienen los niños matriculados en \textcolor{darkgreen}{Educación Primaria}, \\
+{\sc ref} & Sí lo tienen los niños matriculados en \textcolor{darkgreen}{Educación Primaria}, \\
     &~~ no los de \textcolor{darkgreen}{Educación Secundaria} Obligatoria.\\
@@
 & 34. El artículo. El tipo de las viviendas \textcolor{blue}{de} la \textcolor{blue}{protección pública}. \\ % 767
-ref & Artículo 34. Tipos de Vivienda \textcolor{darkgreen}{de Protección Pública}.\\
+{\sc ref} & Artículo 34. Tipos de Vivienda \textcolor{darkgreen}{de Protección Pública}.\\
 \hline
@@
 & Las 16/1985 leyes , el del \textcolor{blue}{patrimonio} histórico de España.  \\ % 317 
-ref & Ley 16/1985 de \textcolor{darkgreen}{Patrimonio} Histórico Español \\
+{\sc ref} & Ley 16/1985 de \textcolor{darkgreen}{Patrimonio} Histórico Español \\
 \hline
@@
  \end{center}
+ \texttt{Ling}: Linguistic defaults; \texttt{LER} = Lexical-selection error rate.
 

### 2015-05-04 2015-nodalida-ruscg/ruscg.tex
added sharoff reference
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60234 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-Note that the final accuracy is below the state of the art for Russian, however the state of
-the art employs unavailable training corpora of over five million tokens.\rr{citations? This comes out of the blue.}
+The final accuracy is below the state of the art for Russian. For instance, 
+\newcite{sharoff08lrec-mocky} report a maximum accuracy of 95.28\% using the TnT tagger.
+Note, however, that this model was trained on a much larger corpus -- over five million
+tokens -- which is not freely available.
 
@@
 3 correctly, the HMM system alone is much more dependent on training corpus size than the voting set-up, where accuracies actually appear to become asymptotic already with 8-10,000 words in the training corpus, indicating that further improvement depends more on the CG than the HMM-component.}
+
 \section{Future work}
@@
 
-We have a number of plans for future work, the first is increasing the precision of the grammar without
+We have a number of plans for future work, the first of which is increasing the precision of the grammar without
 decreasing recall. Secondly we would like to add syntactic function labelling and dependency parsing. For 

### 2015-05-04 2015-nodalida-ruscg/ruscg.tex
figure 1 alignment and discussing asymptoteness of fig 3
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60236 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 {\small
-\rus{
-"<В>"                                                         "<работу>"
-        "в" pr                                                     "работа" n f nn sg acc
+"<\rus{В}>"                                         "<\rus{работу}>"
+    "\rus{в}" pr                                        "\rus{работа}" n f nn sg acc
 
-"<ноябре>"                                                    "<Международный>"
-        "ноябрь" n m nn sg prp                                 "международный" adj m an sg nom
-                                                                         "международный" adj m nn sg acc
+"<\rus{ноябре}>"                                    "<\rus{Международный}>"
+    "\rus{ноябрь}" n m nn sg prp                        "\rus{международный}" adj m an sg nom
+                                                 "\rus{международный}" adj m nn sg acc
 "<1994>"                                        
-        "1994" num                                            "<трибунал>"
-                                                                         "трибунал" n m nn sg acc
-"<года>"                                                           "трибунал" n m nn sg nom
-        "год" n m nn sg gen SELECT:r462                             
-;        "год" n m nn pl nom fac SELECT:r462       "<по>"
-                                                                          "по" pr 
-"<в>"                                                             
-        "в" pr                                                    "<Руанде>" 
-                                                                          "Руанда" np al f nn sg prp 
-"<Танзании>"                                                     "Руанда" np al f nn sg dat
-        "Танзания" np al f nn pl acc                           
-        "Танзания" np al f nn sg prp                          "<.>" 
-;        "Танзания" np al f nn pl nom REMOVE:r424        "." sent 
-;        "Танзания" np al f nn sg dat REMOVE:r433 
-;        "Танзания" np al f nn sg gen REMOVE:r433
+    "1994" num                                      "<\rus{трибунал}>"
+                                                        "\rus{трибунал}" n m nn sg acc
+"<\rus{года}>"                                          "\rus{трибунал}" n m nn sg nom
+    "\rus{год}" n m nn sg gen SELECT:r462                             
+;   "\rus{год}" n m nn pl nom fac SELECT:r462       "<\rus{по}>"
+                                                        "\rus{по}" pr 
+"<\rus{в}>"                                                             
+    "\rus{в}" pr                                    "<\rus{Руанде}>" 
+                                                        "\rus{Руанда}" np al f nn sg prp 
+"<\rus{Танзании}>"                                      "\rus{Руанда}" np al f nn sg dat
+    "\rus{Танзания}" np al f nn pl acc                           
+    "\rus{Танзания}" np al f nn sg prp              "<.>" 
+;   "\rus{Танзания}" np al f nn pl nom REMOVE:r424      "." sent 
+;   "\rus{Танзания}" np al f nn sg dat REMOVE:r433 
+;   "\rus{Танзания}" np al f nn sg gen REMOVE:r433
 
-"<начал>"
-        "начало" n nt nn pl gen
-        "начать" vblex perf tv past m sg
-;        "начать" vblex perf iv past m sg REMOVE:r769
-}}
+"<\rus{начал}>"
+    "\rus{начало}" n nt nn pl gen
+    "\rus{начать}" vblex perf tv past m sg
+;   "\rus{начать}" vblex perf iv past m sg REMOVE:r769
+}
 \end{alltt}
@@
 improves the performance of the HMM tagger, an improvement of nearly 5\% in accuracy,
-similar to that reported by \newcite{hulden12} for the same amount of training data. Our 
-constraint grammar also has a much lower precision as a result of the ambiguity remaining
-in the output.
+similar to that reported by \newcite{hulden12} for the same amount of training data. 
+In Figure~\ref{fig:curve}, it appears that the HMM alone is much more dependent on 
+training corpus size than the voting setup, which improves very little between a training
+corpus size of 5,000 and 9,000. This seems to indicate that further improvements depend
+more on the CG than on training data.
 
-The final accuracy is below the state of the art for Russian. For instance, 
+Our constraint grammar also has a much lower precision as a result of the ambiguity remaining
+in the output. Similarly, the final accuracy is below the state of the art for Russian. For instance, 
 \newcite{sharoff08lrec-mocky} report a maximum accuracy of 95.28\% using the TnT tagger.

### 2015-05-04 2015-nodalida-ruscg/ruscg.tex
stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60237 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 	\rus{замк\'{о}в} \emph{zamk\'{o}v} `lock.\textsubscript{PL-GEN}'\\
-    etc. \ldots
-
-\rr{is this clear that the paradigm continues, but we don't show it? One of the reviewers thought the example was incomplete}
+    etc.
 
@@
     \end{itemize}
-    \rr{R2: The first two rules in 6.2 are a bit odd - the effect would be exactly the same without the first context bracket. Did you mean 0C instead of 0?}
     did not take into account the possibility of having a postposed determiner as in
@@
 training corpus size than the voting setup, which improves very little between a training
-corpus size of 5,000 and 9,000. This seems to indicate that further improvements depend
-more on the CG than on training data.
+corpus size of 5,000 and 9,000.
 
@@
 \end{figure*}
-\rr{R2: If I interpret Fig.
-3 correctly, the HMM system alone is much more dependent on training corpus size than the voting set-up, where accuracies actually appear to become asymptotic already with 8-10,000 words in the training corpus, indicating that further improvement depends more on the CG than the HMM-component.}
 

### 2015-05-04 2015-nodalida-ruscg/ruscg.tex
more
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60238 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 decreasing recall. Secondly we would like to add syntactic function labelling and dependency parsing. For 
-the dependency parse we plan to reuse the Giellatekno dependency grammar as in \cite{antonsen10}.
+the dependency parser we plan to reuse the Giellatekno dependency grammar as in \cite{antonsen10}.
 
 The development workflow could also be improved, for example during the testing of each rule we could
-save the correct decisions of the grammar and gradually build up a tagged development corpus for the language.
+save the correct decisions of the grammar. This would give us a partially-disambiguated development corpus,
+which could be gradually used to build up a gold-standard corpus, and which could also be used for regression
+testing to ensure that new rules added do not invalidate the correct decisions of previously written rules.
 
@@
 
-This paper has presented a preliminary constraint grammar for Russian. The constraint grammar 
-is high recall and improves the performance over a simple HMM-based tagger. It also shows 
-state-of-the-art performance for the stress-placement task.
+This paper has presented a preliminary constraint grammar for Russian, where rules have
+been assigned to sections based on observations of performance on a non-gold corpus. The constraint grammar 
+is high recall (over 0.99) and improves the performance over a trigram HMM-based tagger. It 
+also shows state-of-the-art performance for the stress-placement task.
+
 \rr{Emphasize R2's thoughts more in conclusion: I particularly like the idea of systematically assigning rules to heuristicity sections based on performance thresholds derived from annotation-inspection of a non-gold corpus.

### 2015-05-04 2015-nodalida-ruscg/ruscg.tex
multicols for fig 1
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60240 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{linguex}
+\usepackage{multicol}
 %\usepackage[multiple]{footmisc} % separates consecutive footnotes with comma
@@
 \centering
+\begin{multicols}{2}
 \begin{alltt}
 {\small
-"<\rus{В}>"                                         "<\rus{работу}>"
-    "\rus{в}" pr                                        "\rus{работа}" n f nn sg acc
 
-"<\rus{ноябре}>"                                    "<\rus{Международный}>"
-    "\rus{ноябрь}" n m nn sg prp                        "\rus{международный}" adj m an sg nom
-                                                 "\rus{международный}" adj m nn sg acc
+"<\rus{В}>"
+    "\rus{в}" pr
+
+"<\rus{ноябре}>"
+    "\rus{ноябрь}" n m nn sg prp
+
 "<1994>"
-    "1994" num                                      "<\rus{трибунал}>"
-                                                        "\rus{трибунал}" n m nn sg acc
-"<\rus{года}>"                                          "\rus{трибунал}" n m nn sg nom
+    "1994" num
+
+"<\rus{года}>"
     "\rus{год}" n m nn sg gen SELECT:r462
-;   "\rus{год}" n m nn pl nom fac SELECT:r462       "<\rus{по}>"
-                                                        "\rus{по}" pr 
+;   "\rus{год}" n m nn pl nom fac SELECT:r462
+
 "<\rus{в}>"
-    "\rus{в}" pr                                    "<\rus{Руанде}>" 
-                                                        "\rus{Руанда}" np al f nn sg prp 
-"<\rus{Танзании}>"                                      "\rus{Руанда}" np al f nn sg dat
+    "\rus{в}" pr
+
+"<\rus{Танзании}>"
     "\rus{Танзания}" np al f nn pl acc
-    "\rus{Танзания}" np al f nn sg prp              "<.>" 
-;   "\rus{Танзания}" np al f nn pl nom REMOVE:r424      "." sent 
+    "\rus{Танзания}" np al f nn sg prp
+;   "\rus{Танзания}" np al f nn pl nom REMOVE:r424
 ;   "\rus{Танзания}" np al f nn sg dat REMOVE:r433
@@
 ;   "\rus{начать}" vblex perf iv past m sg REMOVE:r769
+
+"<\rus{работу}>"
+    "\rus{работа}" n f nn sg acc
+
+"<\rus{Международный}>"
+    "\rus{международный}" adj m an sg nom
+    "\rus{международный}" adj m nn sg acc
+
+"<\rus{трибунал}>"
+    "\rus{трибунал}" n m nn sg acc
+    "\rus{трибунал}" n m nn sg nom
+
+"<\rus{по}>"
+    "\rus{по}" pr
+
+"<\rus{Руанде}>"
+    "\rus{Руанда}" np al f nn sg prp
+    "\rus{Руанда}" np al f nn sg dat
+"<.>"
+    "." sent
 }
 \end{alltt}
+\end{multicols}
   \caption{Example output from the morphological analyser and constraint grammar for the sentence \rus{В ноябре 1994 

### 2015-05-04 2015-nodalida-ruscg/ruscg.tex
some finishing touches
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60242 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \begin{multicols}{2}
+\noindent
 \begin{alltt}
@@
 
+Table~\ref{table:results} gives a quantitative evaluation of the performance of our CG on the test corpus.
+
 \subsection{Qualitative evaluation}
@@
 
-\begin{table}
-  \centering
-  \begin{tabular}{|l|r|r|}
-
-    \hline
-    Error in original        &   & 2  \\
-    Missing lexeme           &   & 8  \\   % безбалластный, FALSENEG: Кий<np><ant><m><sg><ins> ['кий<n><m><nn><sg><ins>']
-    Partial lexeme           &   & 2  \\   % франко-норманнское, рельс
-    Missing analysis         &   &    \\   % 
-    Equally valid analysis   &   &    \\   % большой<adj><comp><pred> ['больше<adv>'] (?)
-    Correct reading removed  &   & 34 \\   % 
-    \hline
-                             &   & 46 \\
-    \hline
-  \end{tabular}
-  \caption{Error analysis of false negatives}
-\end{table}
+%\begin{table}
+%  \centering
+%  \begin{tabular}{|l|r|r|}
+%    
+%    \hline
+%    Error in original        &   & 2  \\
+%    Missing lexeme           &   & 8  \\   % безбалластный, FALSENEG: Кий<np><ant><m><sg><ins> ['кий<n><m><nn><sg><ins>']
+%    Partial lexeme           &   & 2  \\   % франко-норманнское, рельс
+%    Missing analysis         &   &    \\   % 
+%    Equally valid analysis   &   &    \\   % большой<adj><comp><pred> ['больше<adv>'] (?)
+%    Correct reading removed  &   & 34 \\   % 
+%    \hline
+%                             &   & 46 \\
+%    \hline
+%  \end{tabular}
+%  \caption{Error analysis of false negatives}
+%\end{table}
 
@@
 
-\subsection{Quantitative evaluation}
+\subsection{Task-based evaluation}
 % ROB accent placement task
@@
 
-\rr{Emphasize R2's thoughts more in conclusion: I particularly like the idea of systematically assigning rules to heuristicity sections based on performance thresholds derived from annotation-inspection of a non-gold corpus.
-Another novel idea (at least in terms of actual implementation) is to go for 100\% recall and add a hybridized HMM-tagger in order to improve precision.}
-
 \section*{Acknowledgments}
@@
 gave thoughtful feedback on an earlier version of this paper. All remaining errors are our own.
-\rr{Revise acknowledgments!}
 

### 2015-05-06 2015-nodalida-ruscg/presentation_ruscg.tex
adding presentation
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60274 72bbbca6-d526-0410-a7d9-f06f51895060

%% Hey, emacs(1), this is -*- mode: latex; coding: utf-8; fill-column: 79 -*-

\documentclass{beamer}
\usepackage[T2A,T1]{fontenc}
\usepackage[russian,english]{babel}
\usepackage{relsize}
\usepackage{linguex}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}

\newcommand{\rus}[1]{\foreignlanguage{russian}{#1}}

\AtBeginSection[]
{
  \begin{frame}
    \frametitle{Table of Contents}
    \tableofcontents[currentsection,currentsubsection]
  \end{frame}
}

\usetheme{Montpellier}
\usecolortheme{beaver}

\title[A preliminary constraint grammar for Russian] % (optional, only for long titles)
{A preliminary constraint grammar for Russian}
%\subtitle{}
\author[Tyers, Reynolds] % (optional, for multiple authors)
{F.~Tyers \and R.~Reynolds}
\institute[University of Troms\o] % (optional)
{
  Department of Languages and Linguistics\\
  University of Troms\o
}
\date[NODALIDA-CG 2015] % (optional)
{Constraint Grammar Workshop - Methods, Tools and Applications (NODALIDA 2015)}
\subject{Constraint Grammar, Russian, Morphosyntactic disambiguation}

\begin{document}
\frame{\titlepage}

\section{Outline}

\begin{frame}
\frametitle{Presentation outline}
%\framesubtitle{}
\begin{itemize}
	\item Ambiguity in Russian
	\pause
	\item Finite-state transducer for Russian part-of-speech tagging
	\pause
	\item Constraint Grammar disambiguation rules
	\pause
	\item Development process
	\pause
	\item Evaluation
	\pause
	\begin{itemize}
		\item Qualitative
		\pause
		\item Task-based (word stress placement)
		\pause
		\item Combining with HMM tagger
	\end{itemize}
	\pause
	\item Future work
\end{itemize}
\end{frame}

\section{Ambiguity in Russian} %ROB
\begin{frame}
\frametitle{}
\framesubtitle{}
\begin{itemize}
	\item 
	\pause
	\item 
	\begin{itemize}
		\item 
		\pause
	\end{itemize}
	\item 
\end{itemize}
\end{frame}

\section{Finite-state transducer for Russian part-of-speech tagging} %ROB
\begin{frame}
\frametitle{New finite-state tools for Russian morphological analysis}
\framesubtitle{}
\begin{itemize}
	\item We developed a finite-state transducer using the two-level formalism \cite{Koskenniemi-83}
	\pause
	\begin{itemize}
		\item can be used to analyze and generate Russian wordforms, with or without stress marks
		\pause
	\end{itemize}
	\item Since Russian has systematic syncretism and widespread homonymy, we are developing a constraint grammar \cite{Karlsson.Voutilainen.ea-95} to disambiguate such cases
	\pause
	\item implemented in the freely available CG3 system (http://beta.visl.sdu.dk/cg3.html)
\end{itemize}
\end{frame}

\section{Constraint Grammar disambiguation rules} %FRAN

\begin{frame}
\frametitle{}
\framesubtitle{}
\begin{itemize}
	\item 
	\pause
	\item 
	\begin{itemize}
		\item 
		\pause
	\end{itemize}
	\item 
\end{itemize}
\end{frame}

\section{Development process} %FRAN
\begin{frame}
\frametitle{}
\framesubtitle{}
\begin{itemize}
	\item 
	\pause
	\item 
	\begin{itemize}
		\item 
		\pause
	\end{itemize}
	\item 
\end{itemize}
\end{frame}

\section{Evaluation}
\subsection{Qualitative evaluation} %FRAN
\begin{frame}
\frametitle{}
\framesubtitle{}
\begin{itemize}
	\item 
	\pause
	\item 
	\begin{itemize}
		\item 
		\pause
	\end{itemize}
	\item 
\end{itemize}
\end{frame}

\subsection{Task-based evaluation} %ROB

\begin{frame}
\frametitle{}
\framesubtitle{}
\begin{itemize}
	\item 
	\pause
	\item 
	\begin{itemize}
		\item 
		\pause
	\end{itemize}
	\item 
\end{itemize}
\end{frame}

\subsection{Combining with HMM tagger} %FRAN

\begin{frame}
\frametitle{}
\framesubtitle{}
\begin{itemize}
	\item 
	\pause
	\item 
	\begin{itemize}
		\item 
		\pause
	\end{itemize}
	\item 
\end{itemize}
\end{frame}

\section{Conclusions and future work} %FRAN
\begin{frame}
\frametitle{}
\framesubtitle{}
\begin{itemize}
	\item 
	\pause
	\item 
	\begin{itemize}
		\item 
		\pause
	\end{itemize}
	\item 
\end{itemize}
\end{frame}

\bibliographystyle{apalike}
\begin{tiny}
\bibliography{ruscg}
\end{tiny}
\end{document}

### 2015-05-06 2015-eamt-melsa/poster/2015-eamt-melsa-poster.tex
stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60275 72bbbca6-d526-0410-a7d9-f06f51895060

@@
     \color{titlefgcolor} {\bfseries \Huge  \@title \par}
-    \vspace*{-0.5em}
+%    \vspace*{-0.2em}
     \begin{tabular}{llr}
-    {\Large \@author \par} & ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~  &  \multirow{3}{*}{\@titlegraphic } \\ % [\TP@titlegraphictotitledistance]}  \\
+    {\Large \@author \par} & ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~  &  \multirow{3}{*}{\@titlegraphic } \\ % [\TP@titlegraphictotitledistance]}  \\
     & ~ & ~   \\
-    {\large \@institute}   & ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ &  \\
+    {\large \@institute}   & ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ &  \\
     \end{tabular}
@@
 \titlegraphic{
-\includegraphics[width=0.04\textwidth]{LogoSamisk.png} \hspace{25pt} \includegraphics[width=0.15\textwidth]{logodlsioficialcolor.eps} \hspace{10pt} \includegraphics[width=0.05\textwidth,angle=90]{apertium5a.pdf}
+\includegraphics[width=0.04\textwidth]{LogoSamisk.png} \hspace{35pt} \includegraphics[width=0.15\textwidth]{logodlsioficialcolor.eps} \hspace{15pt} \includegraphics[width=0.05\textwidth,angle=90]{apertium5a.pdf}
 }
@@
 
-\block[titleinnersep=10em]{Lexical selection}{
+\block[titleinnersep=14em]{Lexical selection}{
 
@@
 
-  \block[titleinnersep=10em]{Apertium (\url{www.apertium.org})}{
+  \block[titleinnersep=14em]{Apertium (\url{www.apertium.org})}{
+\vspace{-1pt}
     \begin{center}
@@
 
-  \block[titleinnersep=10em]{Training schema}{
+  \block[titleinnersep=14em]{Schema}{
 
-%\begin{displaymath}
-%    \label{eq:steps}
-%    S \to \mbox{\framebox{\parbox{4.0cm}{\textit{pre-lexsel}}}} \to (\{g_i\}_{i=1}^{i=|G|},S) \to \mbox{\framebox{\parbox{3.5cm}{\textit{\textbf{lexsel}}}}} \to (g^\star,S) \to \mbox{\framebox{\parbox{4.0cm}{\textit{post- lexsel}}}} \to \tau(g^\star,S)
-%  \end{displaymath}
+%This is produced in the first stages of RBMT, just after morphological
+%analysis, part-of-speech tagging, and bilingual dictionary lookup, and
+%before any structural transfer takes place (we will call this
+%\emph{pre-lexsel}). In our model, it is after these first stages that
+%lexical selection (\emph{lexsel}) occurs. After lexical selection,
+%structural transfer and generation take place; a function
+%$\tau(g_i,S)$ represents the result of these last stages, which we
+%will call \emph{post-lexsel}, and returns a finished translation of a
+%specific lexical-selection path $g_i$ of sentence
+%
 
-  \begin{center}
-    \includegraphics[width=0.40\textwidth]{schema.eps}
-  \end{center}
+\begin{displaymath}
+    S \to \mbox{\framebox{\parbox{4.0cm}{\textit{pre-lexsel}}}} \to (\{g_i\}_{i=1}^{i=|G|},S) \to \mbox{\framebox{\parbox{3.5cm}{\textit{\textbf{lexsel}}}}} \to (g^\star,S) \to \mbox{\framebox{\parbox{4.0cm}{\textit{post- lexsel}}}} \to \tau(g^\star,S)
+  \end{displaymath}
+\vspace*{0.1pt}
+\begin{itemize}
+ \item \textbf{pre-lexsel:} pipeline up to lexical transfer
+ \item \textbf{lexsel:} selects one of them \(g^\star\)
+ \item \textbf{post-lexsel:} returns finished translation \(\tau(g^\star,S)\)
+\end{itemize}
+%Source sentence \(S\) has \(|G|\) lexical selection paths \(g_i\): \emph{lexsel} selects one of them \(g^\star\), which is used to generate translation \(\tau(g^\star,S)\).
+
+%  \begin{center}
+%    \includegraphics[width=0.40\textwidth]{schema.eps}
+%  \end{center}
   }
@@
 
-\block[titleinnersep=10em]{Training}{ % See Section 4.2
+\block[titleinnersep=14em]{Training}{ % See Section 4.2
 
 Translate SL sentences to TL using existing MT system and score the translations using an $n$-gram TL model.
-
+~\\
 \useinnerblockstyle{Table}
@@
 
-\block[titleinnersep=10em]{Model}{
+\block[titleinnersep=14em]{Model}{
 
 Extract $n$-gram contexts and train MaxEnt model using fractional counts.
-
+\vspace{0.2pt}
 \begin{center}
@@
 
-\block[titleinnersep=10em]{\emph{Gaur arraina prestatuko dut.} `Today I'll cook fish.'}{
+\block[titleinnersep=14em]{\emph{Gaur arraina prestatuko dut.} `Today I'll cook fish.'}{
 %Compile features into finite-state transducer. 
@@
 
-    arrain:pez  = 1.1356 + 0.16752 = \textbf{1.303123}
+    arrain:pez  = 1.1356 + 0.16752 = 1.303123
 
@@
 \begin{center}
-      \textbf{Rule transducer}
+      \textbf{Rule transducer} \\
+{\small {\tt <:$n$>} denotes rule number}
 \end{center}
@@
 
-%\block[titleinnersep=10em]{Training}{
+%\block[titleinnersep=14em]{Training}{
 %
@@
 
-\block[titleinnersep=10em]{Result comparison}{
+\block[titleinnersep=14em]{Result comparison}{
 
@@
 
-\block[titleinnersep=10em]{Results}{ % See Section 4.2
+\block[titleinnersep=14em]{Results}{ % See Section 4.2
  \begin{center}
@@
  \end{center}
+~\\
  \texttt{Ling}: Linguistic defaults; \texttt{LER} = Lexical-selection error rate.

### 2015-05-07 2015-nodalida-ruscg/presentation_ruscg.tex
more slides, also fixed tag formatting in paper
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60292 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{graphicx}
+\usepackage{tipa} %allows IPA symbols
 
@@
 {F.~Tyers \and R.~Reynolds}
-\institute[University of Troms\o] % (optional)
+\institute[UiT, The Arctic University of Norway] % (optional)
 {
   Department of Languages and Linguistics\\
-  University of Troms\o
+  UiT, The Arctic University of Norway
 }
 \date[NODALIDA-CG 2015] % (optional)
-{Constraint Grammar Workshop - Methods, Tools and Applications (NODALIDA 2015)}
+{Constraint Grammar - Methods, Tools and Applications (NODALIDA workshop, 2015)}
 \subject{Constraint Grammar, Russian, Morphosyntactic disambiguation}
@@
 
-\section{Outline}
+\section{Goals of the paper}
 
 \begin{frame}
-\frametitle{Presentation outline}
+\frametitle{Goals}
 %\framesubtitle{}
 \begin{itemize}
-	\item Ambiguity in Russian
+	\item Preliminary work on morphosyntactic disambiguator for Russian
 	\pause
-	\item Finite-state transducer for Russian part-of-speech tagging
-	\pause
-	\item Constraint Grammar disambiguation rules
-	\pause
-	\item Development process
+	\item Tuned to be high recall at the expense of precision
 	\pause
 	\item Evaluation
-	\pause
 	\begin{itemize}
 		\item Qualitative
-		\pause
 		\item Task-based (word stress placement)
-		\pause
-		\item Combining with HMM tagger
+		\item Combining CG with HMM tagger
 	\end{itemize}
-	\pause
-	\item Future work
 \end{itemize}
@@
 \section{Ambiguity in Russian} %ROB
+
 \begin{frame}
-\frametitle{}
+\frametitle{Ambiguity in Russian}
 \framesubtitle{}
 \begin{itemize}
-	\item 
+	\item Russian is characterized by fusional morphology
+	\begin{itemize}
+		\item Nouns: 12+ paradigm cells (6 main cases, sg and pl)
+		\item Modifiers: 34+ paradigm cells (6 cases, 3 genders, etc.)
+		\item Verbs: 13+ paradigm cells (4 past, 6 non-past, imperatives, etc.)
+		% 200+ if you count participles
+
+		\item \rus{выпей} \emph{vypej} `drink.IMP' or `bittern.Pl-GEN'
+	\end{itemize}
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Three kinds of ambiguity}
+\framesubtitle{}
+\begin{itemize}
+	\item We identify three kinds of ambiguity
+	\begin{itemize}
+		\item Intraparadigmatic (same lemma, different morphosyntax)
+		\item Morphosyntactically incongruent (different lemma, different morphosyntax)		
+		\item Morphosyntactically congruent (different lemma, same morphosyntax)
+	\end{itemize}
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Intraparadigmatic ambiguity}
+\framesubtitle{}
+\begin{itemize}
+	\item \rus{тела} \emph{tela} `body'
+	\begin{itemize}
+		\item \rus{т\'{е}ла} \emph{téla} SG-GEN
+		\item \rus{тел\'{а}} \emph{telá} PL-NOM or PL-ACC
+	\end{itemize}
 	\pause
-	\item 
+	\item \rus{новой} \emph{novoj} `new'
 	\begin{itemize}
-		\item 
+		\item ADJ-SG-FEM-GEN
+		\item ADJ-SG-FEM-LOC
+		\item ADJ-SG-FEM-DAT
+		\item ADJ-SG-FEM-INS
+	\end{itemize}
 	\pause
+	\item \rus{пойдём} \emph{pojdëm} `go'
+	\begin{itemize}
+		\item FUT-1PL `we will go'
+		\item IMP-1PL `let's go!'
 	\end{itemize}
-	\item 
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Morphysyntactically incongruent ambiguity}
+\framesubtitle{}
+\begin{itemize}
+	\item \rus{нашей} \emph{našej}
+	\begin{itemize}
+		\item \rus{н\'{а}шей} \emph{nášej} `our.F-SG-GEN/DAT/LOC/INS'
+		\item \rus{наш\'{е}й} \emph{našéj} `sew on.IMP-2SG'
+	\end{itemize}
+	\pause
+	\item \rus{дорога} \emph{doroga}
+	\begin{itemize}
+		\item \rus{дор\'{о}га} \emph{doróga} `road.N-F-SG-NOM'
+    	\item \rus{дорог\'{а}} \emph{dorogá} `dear.ADJ-F-SG-PRED'
+	\end{itemize}
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Morphysyntactically congruent ambiguity}
+\framesubtitle{}
+\begin{itemize}
+	\item \rus{замок} \emph{zamok}
+	\begin{itemize}
+		\item \rus{з\'{a}мок} \emph{z\'{a}mok} `castle.SG-NOM'\\
+			  \rus{зам\'{о}к} \emph{zam\'{o}k} `lock.SG-NOM'
+		\item \rus{з\'{a}мков} \emph{z\'{a}mkov} `castle.PL-GEN'\\
+			  \rus{замк\'{о}в} \emph{zamk\'{o}v} `lock.PL-GEN'
+		\item etc.
+	\end{itemize}
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Frequency of each kind of ambiguity}
+\framesubtitle{}
+\begin{table}
+  \centering
+  \begin{tabular}{l|rr}
+    \hline
+    \textbf{Type}  & \textbf{all tokens} & \textbf{ambig. tokens} \\
+    \hline
+    Intraparadigm. & 59.0\%                   & 90.9\%   \\
+    Incongruent    & 27.7\%                   & 42.7\%   \\ 
+    Congruent      & 1.2\%                   & 1.8\%    \\ 
+    \hline
+  \end{tabular}
+  \caption{Frequency of different types of morphosyntactic ambiguity in unrestricted text}
+  %TODO caption says 'Table :' without a number
+  \label{table:ambiguity}
+\end{table}
+\begin{itemize}
+	\item \rus{з\'{a}мков} \emph{zámkov} `castle.PL-GEN'
+	\item \rus{замк\'{о}в} \emph{zamkóv} `lock.PL-GEN'
+	\item \rus{з\'{а}мков} \emph{zámkov} `having to do with castles.A-MSC-SG-PRED'
 \end{itemize}
@@
 \section{Finite-state transducer for Russian part-of-speech tagging} %ROB
+
 \begin{frame}
@@
 \begin{itemize}
-	\item We developed a finite-state transducer using the two-level formalism \cite{Koskenniemi-83}
+	\item We developed a finite-state transducer using the two-level formalism \cite{Koskenniemi-84}
+	\begin{itemize}	
+		\item analyze and generate Russian wordforms, with or without stress marks
 		\pause
+	\end{itemize}
+	\item Based on Zaliznjak's \emph{Grammatical dictionary of Russian} \cite{Zaliznjak-77}
 	\begin{itemize}
-		\item can be used to analyze and generate Russian wordforms, with or without stress marks
+		\item includes the proper noun appendix from 2001 version
+		\item supplemented with \emph{Grammatical dictionary of new Russian words} [Grishina \& Lyashevskaya 2008]
 		\pause
@@
 \begin{frame}
-\frametitle{}
+\frametitle{Automatic word stress placement}
 \framesubtitle{}
 \begin{itemize}
-	\item 
+	\item Russian word stress placement...
+	\begin{itemize}
+		\item ...includes complex patterns of shifting stress
 		\pause
-	\item 
+		\item ...is critical for language learners and text-to-speech\\
 		\begin{itemize}
-		\item 
+			\item \rus{догов\'{о}ром} \emph{dogovórom} > /\textipa{d@g2vOr@m}/
+		\end{itemize}
+		\pause
+		\item ...is lexically specified: there are no universal rules of stress pattern assignment
 		\pause
 	\end{itemize}
-	\item 
+	\item >7.48\% of tokens in Russian running text are stress-ambiguous
+	\item Theoretically, almost 99\% of stress ambiguity can be resolved using a constraint grammar
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Automatic word stress placement - results}
+\framesubtitle{}
+\begin{table}
+  \centering
+  \begin{tabular}{r | r r r }
+     & accuracy\% & error\% & abstention\%  \\
+    \hline
+    \hline
+    {\small {\tt RussianGram}} & 90.09 & 0.79 & 9.12 \\
+    {\small {\tt no CG}} & 90.07 & 0.49 & 9.44  \\
+    {\small {\tt CG}} & 93.21 & 0.74 & 6.05
+  \end{tabular}
+  \caption{Results of stress placement task evaluation}
+  \label{tab:results}
+\end{table}
+\begin{itemize}
+	\item CG makes gains of 3.14\% (42\% of ambiguity from FST)
+	\pause
+	\item CG's high recall leads to a low error rate
+	\pause
+	\item To be continued.... (tomorrow)
 \end{itemize}

### 2015-05-07 2015-nodalida-ruscg/ruscg.tex
removed last marginpar
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60243 72bbbca6-d526-0410-a7d9-f06f51895060
more slides, also fixed tag formatting in paper
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60292 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \ex. \label{ex:zamkov}
-	 \a. \rus{замок}$^1$+N+Msc+Inan+Pl+Gen \label{ex:zamkovA}
-	 \b. \rus{замок}$^2$+N+Msc+Inan+Pl+Gen \label{ex:zamkovB}
-	 \c. \rus{замковый}+A+Msc+Sg+Pred      \label{ex:zamkovC}
+	 \a. \rus{замок}$^1$ \texttt{{\small <n><m><nn><pl><gen>}} \label{ex:zamkovA}
+	 \b. \rus{замок}$^2$ \texttt{{\small <n><m><nn><pl><gen>}} \label{ex:zamkovB}
+	 \c. \rus{замковый} \texttt{{\small <adj><m><sg><pred>}}   \label{ex:zamkovC}
 
@@
 
-\rr{R1: Tables 3 and 4 are not discussed enough.}
 % unknown words from test corpus, categorisation

### 2015-05-07 2015-nodalida-ruscg/presentation_ruscg.tex
more slides
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60293 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \begin{frame}
-\frametitle{New finite-state tools for Russian morphological analysis}
+\frametitle{New finite-state transducer for Russian morphological analysis}
 \framesubtitle{}
@@
 	\begin{itemize}
+		\item lexc/twolc, can be compiled in xfst or hfst
 		\item analyze and generate Russian wordforms, with or without stress marks
@@
 	\end{itemize}
-	\item Since Russian has systematic syncretism and widespread homonymy, we are developing a constraint grammar \cite{Karlsson.Voutilainen.ea-95} to disambiguate such cases
-	\pause
-	\item implemented in the freely available CG3 system (http://beta.visl.sdu.dk/cg3.html)
+	\item Example output:
+	\begin{itemize}
+		\item \rus{новый}\texttt{{\small <adj><m><nn><sg><nom>}}\\
+	    `new'
+		\item \rus{автомат}\texttt{{\small <n><m><nn><sg><nom>}}\\
+	    `automaton, sub-machine gun'
+	\end{itemize}
 \end{itemize}
@@
 \begin{frame}
-\frametitle{}
+\frametitle{CG rules: distribution of rule across levels}
 \framesubtitle{}
+\begin{table}
+  \centering
+  \begin{tabular}{lrrr}
+    \hline
+                     & {\sc select} & {\sc remove} & {\sc map} \\
+    \hline
+    Safe             &   16         &   34         &  -- \\
+    Safe heuristic   &   89         &   76         &  -- \\
+    Heuristic        &   26         &   52         &  -- \\
+    Syntax labelling & --           & --           & 6 \\ 
+    \hline
+  \end{tabular}
+  \caption{The 299 rules in the grammar are separated into four sections depending
+      on rule reliability. }
+  \label{tab:ruleDist}
+\end{table}
 \begin{itemize}
-	\item 
+	\item Safe: real constraints in the language\\
+		{\small Ex: preposition cannot directly precede a finite verb}
 		\pause
-	\item 
-	\begin{itemize}
-		\item 
+	\item Safe heuristic: highly reliable tendencies\\ % Russian word order is very free
+		{\small Ex: remove GEN from sentence-init. if no verb governing GEN.}
 		\pause
-	\end{itemize}
-	\item 
+	\item Heuristic: tendencies that are true most of the time\\ % better replaced with machine-learning
+		{\small Ex: \rus{такая} is `such.FEM-SG-NOM', not `say~\rus{так}.V-PRS-ACT-ADV'}
 \end{itemize}
@@
 \section{Development process} %FRAN
+
 \begin{frame}
-\frametitle{}
+\frametitle{Writing rules and level assignment}
 \framesubtitle{}
 \begin{itemize}
-	\item 
+	\item No hand-annotated corpus was compatible with our analyser
 	\pause
-	\item 
+	\item We sorted new rules using Russian Wikipedia by collecting 100 example applications of the rule
+	\pause
 	\begin{itemize}
-		\item 
-		\pause
+		\item Safe: 100\% correct
+		\item Safe heuristic: > 97\% correct
+		\item Heuristic (or discard): $\leq$ 97\% correct 
 	\end{itemize}
-	\item 
 \end{itemize}

### 2015-05-07 2015-nodalida-ruscg/presentation_ruscg.tex
mroe slides for HMM
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60295 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \begin{frame}
-\frametitle{}
+\frametitle{Combining CG with HMM tagger}
 \framesubtitle{}
 \begin{itemize}
-	\item 
+	\item CG removes less than half of ambiguity
 	\pause
-	\item 
+	\item To resolve remaining ambiguity, we backoff to HMM tagger
 	\begin{itemize}
-		\item 
+		\item This approach has been applied with...\\
+		Basque \cite{ezeiza.ea-98},\\
+		Czech \cite{hajic.ea-01,hajic.ea-07},\\
+		Norwegian \cite{johannessen.ea-11,johannessen.ea-12},\\
+		Spanish \cite{hulden12},\\
+		and Turkish \cite{oflazer.tur-96}
 		\pause
+	\end{itemize}
+	\item We follow the voting method used by \cite{hulden12}
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Voting setup}
+\framesubtitle{}
+\begin{itemize}
+	\item Freely available hunpos tagger
+	\pause
+	\item 10-fold cross-validation (90\% training, 10\% testing)
+	\pause
+	\item 3 configurations
+	\begin{itemize}
+		\item HMM - hunpos with default options
+		\pause
+		\item HMM+Morph - incorporating FST output as full-form lexicon
+		\pause
+		\item HMM+Morph vs CG - Voting setup: if constraint grammar left 1 valid reading,
+		take that. Otherwise, take output of HMM+Morph.
 	\end{itemize}
@@
 
-\section{Conclusions and future work} %FRAN
 \begin{frame}
-\frametitle{}
+\frametitle{Voting results}
 \framesubtitle{}
+\begin{figure}
+  \centering
+  \includegraphics[width=0.70\textwidth]{graphics/learning-curve.pdf}
+  %\caption{Learning curve for three taggers, \texttt{hunpos} with no lexicon, 
+  %  \texttt{hunpos} with a lexicon, and \texttt{hunpos} with a lexicon and the 
+  %  Russian constraint grammar in a voting set up.}
+  \label{fig:curve}
+\end{figure}
 \begin{itemize}
+	\item hunpos tagger alone is more dependent on training-corpus size
+	\pause
 	\item 
@@
 	\item 
+	\item 
+\end{itemize}
+\end{frame}
+
+\section{Future work and conclusions} %FRAN
+
+\begin{frame}
+\frametitle{Future work}
+\framesubtitle{}
 \begin{itemize}
-		\item 
+	\item Improve precision while maintaining high recall
+	\pause
+	\item Add syntactic functions and dependency parsing
+	\begin{itemize}
+		\item We plan to use Giellatekno's dependency parser \cite{antonsen10}
 		\pause
 	\end{itemize}
-	\item 
+	\item Improve workflow by extending our gold-standard corpus while checking new rule accuracy
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Conclusions}
+\framesubtitle{}
+\begin{itemize}
+	\item Preliminary constraint grammar for Russian
+	\pause
+	\begin{itemize}
+		\item Tuned for high recall at expense of precision
+		\pause
+		\item Rules are assigned to levels, based on tested accuracy
+		\pause
+		\item Improves performance of trigram HMM-based tagger
+		\pause
+	\end{itemize}
+	\item Improve workflow by extending our gold-standard corpus while checking new rule accuracy
 \end{itemize}

### 2015-05-10 2015-nodalida-ruscg/presentation_ruscg.tex
revisions
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60326 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 		% 200+ if you count participles
-
-		\item \rus{выпей} \emph{vypej} `drink.IMP' or `bittern.Pl-GEN'
+	\end{itemize}
+	\pause
+	\item Both in-category and out-of-category ambiguity
+	\begin{itemize}
+		\item \rus{выпей} \emph{vypej} `drink.V-IMPER' or `bittern.N-PL-GEN'
+		\includegraphics[width=0.40\textwidth]{graphics/vypej3103.jpeg} %TODO add image
 	\end{itemize}
@@
   %TODO caption says 'Table :' without a number
+  %TODO numbers different from ERC
   \label{table:ambiguity}
 \end{table}
+\pause
+\begin{itemize}
+	\item Categories are not mutually exclusive
 	\begin{itemize}
@@
 		\item \rus{з\'{а}мков} \emph{zámkov} `having to do with castles.A-MSC-SG-PRED'
+	\end{itemize}
 \end{itemize}
@@
 		{\small Ex: \rus{такая} is `such.FEM-SG-NOM', not `say~\rus{так}.V-PRS-ACT-ADV'}
+		%TODO better example
 \end{itemize}
@@
 
-\section{Development process} %FRAN
+\subsection{Development process} %FRAN
 
@@
 		\item Heuristic (or discard): $\leq$ 97\% correct
+		%TODO put before asdf
 	\end{itemize}
@@
 \section{Evaluation}
+
+\begin{frame}
+\frametitle{Results}
+\framesubtitle{}
+\begin{table}
+  \centering
+  \begin{tabular}{l|rrrrr}
+    %\hline
+    \textbf{Domain} & \textbf{Tokens} & \textbf{Precision} & \textbf{Recall} & \textbf{F-score} & {Ambig. solved} \\
+    \hline
+    Wikipedia       & 7,857      & 0.506        & 0.996    & 0.671 & 44.92\%  \\
+    Literature      & 1,652      & 0.473        & 0.984    & 0.638 & 42.95\%  \\
+    News            & 642        & 0.471        & 0.990    & 0.638 & 41.60\%  \\
+    \hline
+    \textbf{Total}  & 10,150     & 0.498        &  0.994   & 0.663 & 44.39\% \\
+    %\hline
+  \end{tabular}
+  \caption{Results on the test corpora.}
+  \label{table:results}
+\end{table}
+\end{frame}
+
 \subsection{Qualitative evaluation} %FRAN
+
 \begin{frame}
-\frametitle{}
+\frametitle{Errors from `bad linguistics' (too limiting)}
 \framesubtitle{}
 \begin{itemize}
-	\item 
-	\pause
-	\item 
+	  \item In some cases a rule did not take into account grammatical possibilities 
+    in the language. e.g. Two simple rules such as\ldots
     \begin{itemize}
-		\item 
-		\pause
+      \item \texttt{REMOVE Det IF (0 Det OR Pron) (1C Ne) ;}
+      \item \texttt{REMOVE Det IF (0 Det OR Pron) (1 Cm LINK 1 CC OR CS) ;}
     \end{itemize}
-	\item 
+    \ldots did not take into account the possibility of having a postposed determiner as in\ldots
+    \begin{itemize}
+      \item \ldots \rus{а может быть и раньше, и факт этот не раз поражал меня} \ldots
+      \item \ldots and maybe even earlier, and fact \textbf{this} not once surprised me \ldots
+    \end{itemize}
+    \ldots or a interposed parenthetical as in\ldots
+    \begin{itemize}
+      \item \rus{Но какие, однако же, два разные создания, точно обе с двух разных планет!}
+      \item But what, \textbf{exactly} , two different creatures, just both from two different planets!
+    \end{itemize}
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Errors from bad rules}
+\framesubtitle{}
+\begin{itemize}
+	\item In some cases a rule was simply incorrectly specified. For example, the 
+	following rule was designed to solve the ambiguity between short-form neuter adjectives and 
+	adverbs
+   \begin{itemize}
+		\item \texttt{REMOVE A + Short IF (-1C Fin OR Adv OR A) (0C Short OR Adv) ;}
+   \end{itemize}
+   \item However there is no reason why we should prefer an adverb over an adjective after an adverb, 
+   \begin{itemize}
+     \item \ldots \rus{потому что совсем неприятно проснуться в гробу под землею.}
+     \item \ldots because [it is] really \textbf{unpleasant} to wake up in a coffin under the ground.
+   \end{itemize}
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Errors from incomplete barrier}
+\framesubtitle{}
+\begin{itemize}
+	\item Some rules suffered from incomplete barriers, which is something that would
+	benefit from a more systematic treatment.
+	\begin{itemize} 
+		\item \texttt{REMOVE NGDAIP - Acc - Prp - Loc IF (-1C* Pr/V OR Pr/Na BARRIER (*) - Adv - Comp - DetIndecl - ModAcc - ModPrp) ;}
+	\end{itemize}
+	here the rule removes the nominative reading of the adjective to leave the accusative 
+	reading because the preposition \rus{в} `in' is found preceding. 
+	\begin{itemize}
+     \item \rus{В 1960-х электрифицированные высокоскоростные железные дороги появились в Японии и некоторых других странах.}
+     \item In the 1960's \textbf{electrified} high-speed railways appeared in Japan and some other countries.
+   \end{itemize}
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Errors from incomplete sets}
+\framesubtitle{}
+\begin{itemize}
+	\item In some cases the rule was a good generalisation, but made use of a set
+	which was incomplete. For example:
+	\begin{itemize}
+		\item \texttt{REMOVE Dat IF (NOT 0 Prn/Sebe) (NOT 0 Anim OR Cog OR Ant) (NOT 0 Pron) (NOT 1* V/Dat) (NOT -1* V/Dat) (NOT -1* Prep/Dat) (NOT -1C A + Dat) ;}
+	\end{itemize} 
+	the set \texttt{V/Dat} does not contain the verb \rus{противопоставлять} `opposed to' which takes 
+	a dative argument.
+	\begin{itemize}
+		\item \rus{В связи с этим ортодоксальности стали противопоставлять ересь.}
+		\item In connection with this \textbf{orthodoxy} was opposed to heresy. 
+	\end{itemize}
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Errors from rule interaction}
+\framesubtitle{}
+\begin{itemize}
+	\item The strong accusative rule below causes incorrect behaviour in the rule to remove transitivity readings
+	\begin{itemize}
+		\item \texttt{REMOVE TV - Pass IF (NOT 1* Acc) (NOT -1* Acc) ;}
+		\item \texttt{REMOVE Acc IF (-1C Fin + IV) (NOT 0 AccAdv) ;}
+	\end{itemize}
+	Consider the following example where \rus{может} `can' is tagged as intransitive, the second rule fires removing 
+	the accusative reading of \rus{его} `him', and thus given the lack of accusative reading, \rus{найти} `find'
+	is disambiguated as intransitive instead of transitive.
+ 	\begin{itemize}
+		\item \rus{Она смотрит везде, но не может его найти.}
+		\item She looks around, but she cannot \textbf{find him}.
+	\end{itemize}
+\end{itemize}
+\end{frame}
+
+\begin{frame}
+\frametitle{Errors from difficult linguistics}
+\framesubtitle{}
+\begin{itemize}
+	\item Dealing with participles with arguments is challenging in the case that the arguments 
+	of the participle share the same government as the main verb.
+	\begin{itemize}
+		\item \texttt{REMOVE IV IF (0 TV OR IV) (1C Acc) (NOT 1 AccAdv);}
+	\end{itemize}
+		Here \rus{Ваню и Машу} `Vanja and Maša' are the object of \rus{видит} `sees' and not \rus{играющих} `playing', although
+		both verbs can take accusative object.
+	\begin{itemize}
+		\item \rus{Их мама внутри дома с кошкой, она смотрит в окно и видит играющих Ваню и Машу.}
+		\item Their mother is inside the house with the cat, she looks through the window and sees Vanja and Maša \textbf{playing}.
+	\end{itemize}
+	This kind of error would ideally be resolved with semantic knowledge.
 \end{itemize}
@@
 	\item CG's high recall leads to a low error rate
+	\begin{itemize}
+		\item Higher accuracy was achieved (96.15\%), but with higher error rate (3.14\%)
+	\end{itemize}
 	\pause
@@
 	\pause
-	\item To resolve remaining ambiguity, we backoff to HMM tagger
+	\item To resolve remaining ambiguity, we turn to stochastic methods
 	\begin{itemize}
@@
 	\end{itemize}
-	\item We follow the voting method used by \cite{hulden12}
+	\item We follow the voting method used by \cite{hulden12}, using the freely available hunpos tagger
 \end{itemize}
@@
 \begin{itemize}
-	\item Freely available hunpos tagger
-	\pause
-	\item 10-fold cross-validation (90\% training, 10\% testing)
-	\pause
 	\item 3 configurations
@@
 	\end{itemize}
-	\item 
+	\pause
+	\item 10-fold cross-validation
 \end{itemize}
@@
 	\item hunpos tagger alone is more dependent on training-corpus size
-	\pause
-	\item 
-	\pause
-	\item 
-	\item 
 \end{itemize}
@@
 \begin{itemize}
-	\item Improve precision while maintaining high recall
+	\item Improve precision (while maintaining high recall)
 	\pause
@@
 	\item Improve workflow by extending our gold-standard corpus while checking new rule accuracy
+	\begin{itemize}
+		\item Can be used for regression testing
+	\end{itemize}
 \end{itemize}

### 2015-06-17 2015-turklang-kazdep/paper.tex
beginnings of paper
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60801 72bbbca6-d526-0410-a7d9-f06f51895060

--- 2014-03-23 kipchak-lrec-2013/kipchak-paper.tex
@@
 
-\usepackage{lrec2014}
+\usepackage{turklang}
 
@@
 \usepackage{xunicode}
-\usepackage{xltxtra}
-\usepackage{natbib}
-\usepackage{multirow}
-%\usepackage{fullpage}
 \usepackage{multirow}
 \usepackage[colorlinks=true,citecolor=black,linkcolor=black,urlcolor=blue]{hyperref}
-\usepackage[small,bf]{caption}
-\usepackage{booktabs}
-\usepackage{jnwmaps}
 
-\defaultfontfeatures{Mapping=tex-text}
-\setromanfont{Times New Roman}
-\newfontfamily\symbl[Scale=MatchLowercase]{FreeSerif}
-\newfontfeature{IPA}{+mgrk}
-\newfontfamily\qipa[IPA,Scale=MatchLowercase]{FreeSerif}
-\newfontfamily\qgmk[Scale=MatchLowercase,Letters=SmallCaps]{Linux Libertine O}
-%
-\newcommand{\gmk}[1]{{\qgmk #1}}
-\newcommand{\tags}[1]{\texttt{#1}}
+\usepackage{natbib}
 
-\bibdata{kipchak-paper}
+\usepackage{times}
 
-\title{Finite-state morphological transducers for three Kypchak languages}
+\title{Dependency Treebank for Kazakh based on Universal Dependencies}
+\name{Francis Tyers$^a$ and Jonathan Washington$^b$}
+\address{$^a$ hargle,\\
+	$^b$ Departments of Linguistics and Central Eurasian Studies, Indiana University, Bloomington, IN 47405, USA}
 
-%\author{Jonathan North Washington \\
-%Departments of Linguistics and Central Eurasian Studies\\
-%Indiana University\\
-%Bloomington, IN 47405 (USA)\\
-%\texttt{jonwashi@indiana.edu} \and
-%Ilnar Salimzyanov  \\
-%Kazan Federal University \\
-%Kazan, Republic of Tatarstan\\
-%(Russian Federation) \\
-%\texttt{ilnar.salimzyan@gmail.com} \and 
-%Francis M. Tyers\\
-%5HSL-fakultehta, \\
-%UiT Norgga árktalaš universitehta, \\
-%9019 Romsa (Norway)\\
-%\texttt{francis.tyers@uit.no} 
-%}
-
-%FIXME: usually star comes first, then daggers
-\name{Jonathan North Washington$^\star$, Ilnar Salimzyanov$^\dagger$, Francis M. Tyers$^\ddagger$}
-
-\address{$^\star$Departments of Linguistics and Central Eurasian Studies\\
-Indiana University\\
-Bloomington, IN 47405 (USA)\\
-\texttt{jonwashi@indiana.edu} \and
-$^\dagger$Kazan Federal University \\
-Kazan, Republic of Tatarstan\\
-(Russian Federation) \\
-\texttt{ilnar.salimzyan@gmail.com} \and 
-$^\ddagger$HSL-fakultehta \\  
-UiT Norgga árktalaš universitehta \\
-9019 Romsa (Norway)\\
-\texttt{francis.tyers@uit.no} 
-}
-
-\abstract{This paper describes the development of free/open-source finite-state morphological transducers for three Turkic languages---Kazakh, Tatar, and Kumyk---representing one language from each of the three commonly distinguished sub-branches of the Kypchak branch of Turkic.
-The finite-state toolkit used for the work is the Helsinki Finite-State Toolkit (HFST).
-This paper describes how the development of a transducer for each subsequent closely-related language took less development time.
-An evaluation is presented which shows that the transducers all have a reasonable coverage---around 90\%---on freely available corpora of the languages, and high precision over a manually verified test set.
-\\
-	\Keywords{Kazakh, Tatar, Kumyk, morphology, transducer}
-}
+\abstract{hargle treebank bargle dependency grammar}
 
@@
 
-This paper describes the development of free/open-source morphological transducers for three closly related languages: Kazakh, Tatar, and Kumyk.
-Morphological transducers are computational models of the languages' morphology, and are used to output morphological analyses from word forms and vice-versa.
+\section{Background}
 
-These transducers were all developed as part of the Apertium project, which is aimed at creating rule-based machine translation (RBMT) systems for lesser resourced languages.
-As such, the transducers were developed with the intent that they be used as morphological analysers and generators in RBMT systems.
-However, this sort of finite-state transducer has a wide variety of usages, including for creating proofing tools such as spellcheckers and grammarcheckers, morphological annotation for linguistic research, creating resources for learners and CALL, and lemmatised dictionary lookup.
+\subsection{The Kazakh Language}
 
-%the potential uses of morphological transducers is much broader; for example, we have found that they are easily converted to spell checkers.
+\subsection{Treebanks}
 
-All three transducers were designed as ``ports'' of Apertium's Kyrgyz morphological transducer \citep{washington2012}.
-The Tatar and Kazakh transducers are used in a Kazakh--Tatar MT system \citep{salimzyanov2013}.
+\section{Methodology}
 
-%FIXME: brief background on how, when, and with what support were developed?
-% In Apertium project, Turkic transducers were developed either as part of Turkic-to-Turkic machine translators, or they started as a "port" (in particular of morphotactics) of already existing Turkic transducers.
-
-This paper gives an overview of the languages (§\ref{sec:lgs}), provides background on these morphological transducers and related previous work (§\ref{sec:bg}), describes the development effort and contents of the transducers (§\ref{sec:methodology}), evaluates the effectiveness of these transducers (§\ref{sec:evaluation}), and summarises the findings (§\ref{sec:conclusions}).
-
-\section{Languages}\label{sec:lgs}
-
-The three languages for which transducers were developed belong to the Northwestern branch of the Turkic family, which is often referred to as the Kypchak branch.
-This branch can be divided into three sub-branches.
-Kumyk is a member of the Western Kypchak group, Tatar is a member of the Northern Kypchak group, and Kazakh is a member of the Southern Kypchak group \citep[82-83]{histofturkic}.\footnote{It is the professional opinion of the authors of this paper that Kyrgyz constitutes a fourth branch of Kypchak.}   The geographic distribution of the languages is shown in map~\ref{map1}.
-%As such, each of these three languages represents one of the three branches of Kypchak.
-
-\begin{map*}[htbp]
-	\includegraphics[width=\textwidth]{map/map}
-	\caption{The three sub-branches of Kypchak (North, South, West), roughly divided with black lines, showing the geographic distribution of the three languages for which transducers were developed.
-The Kypchak languages shown on the map are Tatar (\texttt{tat}), Kazakh (\texttt{kaz}), and Kumyk (\texttt{kum}).
-The other codes represent Bashkir (\texttt{bak}), Kyrgyz (\texttt{kir}), Karakalpak (\texttt{kaa}), Nogay (\texttt{nog}), Karachay-Balkar (\texttt{krc}), Urum (\texttt{uum}), Crimean Tatar (\texttt{crh}), and Karaim (\texttt{kdr}).}
-	\label{map1}
-\end{map*}
-
-These languages display different amounts of linguistic influence from other Turkic branches (e.g., moderate Oghuz (SE) influence in the Western group, slight Oghuz influence in the Northern group) and from Mongolic languages (moderate influence on the Southern group, lighter in the other groups), and all display heavy influence from Persian.
-
-\subsection{Kazakh}
-Kazakh /q{\symbl ɑ}z{\symbl ɑ}q/ is spoken primarily in Kazakhstan, where it is the national language and is co-official with Russian.
-Large communities of native speakers also exist in China, neighbouring Central-Eurasian republics, and Mongolia.
-Ethnologue estimates
-the total number of speakers to be around 8 million \citep{ethnologue}.
-
-%Estimates of the total number of speakers range from 8 million \citep{ethnologue} to 11 million \citep{nationalencyklopedin} people.
-
-\subsection{Tatar}
-Tatar /t{\symbl ɒ}t{\symbl ɑ}r/ is spoken in and around Tatarstan, --- a republic of the Russian Federation, where it is co-official with Russian.
-The majority of native speakers are bilingual in Russian.
-Tatar is spoken by approximately 5.4 million people \citep{ethnologue}.
-%, and the standard varieties our system FIXMEs are written in Cyrillic.
-
-\subsection{Kumyk}
-Kumyk /qumuq/ is spoken in Dagestan, a republic of the Russian Federation, where it is co-official with a number of other national languages \citep{ethnologue}.
-There are approximately 430 thousand speakers \citep{ethnologue}.
-
-\section{Background}\label{sec:bg}
-\subsection{Morphological transducers}
-%FIXME: need a source, maybe something by Tommi?
-The objective of a morphological transducer is twofold: firstly to take surface forms (e.g., алдым) and generate all possible lexical forms, and secondly to take lexical forms (e.g.,  ал{\tt {\small <v><tv><ifi><p1><sg>}}, алд{\tt {\small <n><px1sg><nom>}},\footnote{For a description of the tags used throughout this paper, please see Appendix~\ref{app:glossary}.} etc.) and generate one or more surface forms.
-As they are implemented as finite-state transducers, they are reversible by default.
-For more information on using finite-state transducers for morphological analysis and generation, the reader is referred to \cite{beesley2003}.
-
-The transducers were designed based on the Helsinki Finite State Toolkit \citep{hfst/2011} which is a free/open-source reimplementation of the Xerox finite-state toolchain, popular in the field of morphological analysis.
-It implements both the \texttt{lexc} formalism for defining lexicons, and the \texttt{twol} and \texttt{xfst} formalisms for modeling morphophonological rules.
-It also supports other finite state transducer formalisms such as \texttt{sfst}.
-This toolkit has been chosen as it -- or the equivalent XFST \citep{beesley2003} -- has been widely used for other Turkic languages, such as Turkish \citep{coltekin2010}, Crimean~Tatar~\citep{altintas2001}, Turkmen \citep{tantug2006}, and Kyrgyz \citep{washington2012}, and is available under a free/open-source licence.
-
-The authors learnt of another Kazakh morphological transducer in existence \citep{bekmanova2013} only after this paper was submitted and our transducer was released.
-The system is unfortunately not freely available so we have not been able to evaluate this system or compare it to ours.
-
-Creating morphological transducers in the above-mentioned formalisms involves encoding linguistic knowledge about the language in the formalisms.
-The \texttt{lexc} and \texttt{twol} formalisms resemble linguistic formalisms, allowing the coders to work with abstractions resembling linguistic categories such as lexemes, morphemes, phonemes, and even archiphonemes.
-
-%\subsection{Description}
-
-The transducers are available / under development in apertium's subversion repository,\footnote{{\scriptsize \url{https://svn.code.sf.net/p/apertium/svn/languages/}}} in the directories \texttt{apertium-kaz}, \texttt{apertium-tat}, and \texttt{apertium-kum}, and compiled versions may be run at \url{http://turkic.apertium.org/}.
-The revision of each module in the subversion repository that the numbers in this paper (stem counts, evaluation, etc.) represent is \texttt{r50547} for Kazakh, \texttt{r50260} for Tatar, and \texttt{r50300} for Kumyk.
-
-\section{Methodology}\label{sec:methodology}
-% cite Hengeveld (1992) ?
-
-% Noun morphotactics -- is basically equivalent in all three, except for archiphonemes
-% Adjective categorisation
-% Adverb categorisation
-% "Irregular" harmony, caused by orthography
-% Non-finite verb form categorisation "converbs"
-
-% Handling numerals and acronyms.
-
-% Russian loans not obeying phonology N5, e.g. ё vs . самолёт
-
-\subsection{Development effort}
-
-% FIXME: This section needs to be totally rewritten I think.
-
-The three transducers discussed in this paper are for Kazakh, Tatar, and Kumyk.
-The Kazakh and Tatar transducers were originally created as part of an experimental Kazakh--Tatar machine translation system in December of 2010.
-The Kazakh transducer was expanded during Google Code-In 2010 and 2011, and the Tatar transducer was expanded as part of a prototype Tatar and Bashkir machine translation system \citep{tyerswashingtonsalimzyanbattalov12}.
-The Kazakh--Tatar machine translation system, along with the two transducers, was expanded as part of a Google Summer of Code project in 2012 \citep{salimzyanov2013}.
-
-The Kumyk transducer was developed starting at the beginning of October, 2013 as an experiment to see how difficult it would be to extend lessons learned from the development of the Tatar and Kazakh transducers to a related language.
-While the Kazakh and Tatar transducers took around six months work to reach their current coverage level, the Kumyk transducer only took a couple of weeks to reach the same level of coverage (see Figure~\ref{fig:kumcov}).
-This article explores how the development of the Kumyk transducer benefitted from knowledge gained from the development of the Tatar and Kazakh transducers.
-
-The morphotactics\footnote{The morphotactics of a language is the way in which morphemes can be combined to create words.} of Turkic languages are complex enough that even a linguist who is fluent in the language and has a good linguistic understanding of it may not understand how exactly all morphemes combine.
-Native speakers educated about the morphology of their languages also do not have an explicit knowledge of the complete morphotactics.
-Hence it often becomes necessary to use fieldwork methodology to elicit the full extent of the morphotactics, be this a linguist with little to no knowledge of a Turkic language working with a native speaker, or a native speaker who understands the extent of what knowledge is necessary to encode in the transducer.
-When there is no native speaker of a particular language available, the authors have found that information previously encoded about a closely related language or the intuitions of a speaker of a closely related language may be combined with the use of textual corpora to ``elicit'' information about the morphotactics of a language.
-Depending on the contents of corpus and chance, this may not result in a completely accurate model, but it is possible to be thorough.
-
-The Kazakh morphotactics were originally developed based on the Kyrgyz transducer, which was co-authored by the first author---who is fluent in and has a good linguistic knowledge of both Kyrgyz and Kazakh---and two native speakers of Kyrgyz.
-The morphotactics of Tatar were developed for the most part by the second author, a native speaker of Tatar, who also worked to polish off the morphotactics of the Kazakh transducer
-%The Kazakh morphotactics were originally developed based on the Kyrgyz transducer, which was co-authored by the first author, who is fluent in and has a good linguistic knowledge of Kyrgyz, along with two native speakers of Kyrgyz.
-The first author also developed the Kazakh morphotactics who is fluent in and has a good linguistic understanding Kazakh.
-The morphotactics of Tatar were developed for the most part by the third author, a native speaker of Tatar, who also worked to polish off the morphotactics of the Kazakh transducer.
-
-%As the authors found it difficult to locate native speakers of Kumyk, the morphotactics of the Kumyk transducer were developed based on the existing encoded morphotactics of Kazakh and Tatar (and occasionally Kyrgyz), with consultation of corpora, as described above.
-A dictionary \citep{bammatov1960} and a grammar \citep{olmesov2000} of Kumyk were also consulted as needed.
-
-In order to create the Kumyk transducer, we approached one of the major parts of speech at a time. We started with the nouns, copying the 
-continuation lexica\footnote{A continuation lexicon is a set of morphemes, for example, in the Turkic languages there is a continuation lexicon for `case' which includes the possible case suffixes.} (nominal morphotactics) from the Kazakh transducer.
-The suffixes were then replaced with the Kumyk suffixes according to the grammar \citep{olmesov2000}.
-Where the grammar was not explicit regarding a suffix form, we looked the corpus for possible forms and at their contexts.
-The same process was undertaken for verbs.
-Phonological rules were developed as needed.
-
-% FIXME: how long it took
-
-\begin{figure}
-%\centering
-%\hspace{-35pt}
-%\hspace{-1em}
-%    \rotatebox{-90}{
-%\hspace{-15pt}
-\hspace{-10pt}
-\includegraphics[width=0.37\textwidth,bb=50 50 554 770,angle=270]{graphs/kumyk-yoldash-coverage}
-%}
-%~\\
-%~\\
-
-\caption{Number of stems and coverage of a corpus over time for the Kumyk transducer. Time is measured in days 
-starting from the 2nd October 2013. The graph shows that to reach 80\% coverage took
-around a week, and to reach 90\% took another week.}
-\label{fig:kumcov}
-\vspace{-0.5em}
-\end{figure}
-
-\subsection{Transducer contents}
-
-Each transducer's \texttt{lexc} source consists of lists of stems, with each stem pointing at a complex continuation lexicon containing the appropriate morphology for the type of stem.
-
-%FIXME: update for the three languages
-The tagset for each transducer is designed to be compatible with the others.
-Each transducer consists of about 120 separate tags, of which close to 20 cover the main parts of speech (noun, verb, adjective, adverb, postposition, interjection, etc.).
-The remaining tags cover morphological subcategorisation for e.g.\ case, number, person, possession, transitivity, tense-aspect-mood, etc.
-The tags are represented as multicharacter symbols, between less-than \texttt{<} and greater-than \texttt{>} symbols.
-The tagset is quite extensive and still not entirely stabilised, so a full listing is not included here.
-However, the tags are listed in the source code of the transducers, along with comments describing their usage.
-
-%\subsection{Statistics}
-
-Table \ref{table:stems} lists the number of stems of the primary categories in each transducer.
-
-\begin{table}
-\begin{center}
-\begin{tabular}{lrrr}
-		\toprule
-\multirow{2}{*}{\textbf{Part of speech}} & \multicolumn{3}{c}{\textbf{Number of stems}} \\ \cline{2-4}
-                        & Kazakh & Tatar & Kumyk \\
-		\midrule
-		Noun & 2640 & 2795 & 2568 \\
-		Verb & 1470 & 1143 & 386 \\
-		Adjective & 754 & 816 & 219 \\
-		Proper noun & 5701 & 5361 & 1443 \\
-		Adverb & 171 & 177 & 63 \\
-		Numeral & 63 & 63 & 44 \\
-		Conjunction & 46 & 45 & 13 \\
-		Postposition & 50 & 43 & 12 \\
-		Pronoun & 32 & 28 & 17 \\
-		Determiner & 39 & 34 & 9 \\
-		\midrule
-		Total: & 11224 & 10737 & 4845 \\
-		\bottomrule
-\end{tabular}
- \caption{Number of stems in each of the categories. The large number of noun stems as compared
-    to other parts of speech in the Kumyk transducer can be explained by the relative ease of 
-    categorising nouns as opposed to adjectives, adverbs and verbs. Adding a noun essentially involves
-    choosing between loan word and native word. Adding stems from the other main categories 
-    requires more in depth categorisation.}
- \label{table:stems}
-\end{center}
-\vspace{-2em}
-\end{table}
-
-% FIXME: example output
-
-\begin{table*}
-
-% kaz: ^Құдай/Құдай<n><nom>/Құдай<n><attr>/Құдай<n><nom>+е<cop><p3><pl>/Құдай<n><nom>+е<cop><p3><sg>$ ^Өзінің/Өз<prn><ref><p3><sg><gen>$ ^жаратқандарының/жарат<v><tv><ger_past><pl><px3sp><gen>/жарат<v><tv><gpr_past><subst><pl><px3sp><gen>$ ^бәріне/бәрі<prn><qnt><px3sp><dat>$ ^қарап/қара<v><tv><prc_perf>/қара<v><tv><gna_perf>/қара<v><tv><prc_perf>$^,/,<cm>$ ^өте/өт<v><tv><prc_impf>/өте<adv>/өте<v><tv><imp><p2><sg>$ ^жақсы/жақсы<adj>/жақсы<adv>/жақсы<adj><advl>/жақсы<adj><subst><nom>/жақсы<adj>+е<cop><p3><pl>/жақсы<adj>+е<cop><p3><sg>/жақсы<adj><subst><nom>+е<cop><p3><pl>/жақсы<adj><subst><nom>+е<cop><p3><sg>$ ^екенін/е<cop><ger_past><px3sp><acc>$ ^көрді/көр<vaux><ifi><p3><pl>/көр<vaux><ifi><p3><sg>/көр<v><tv><ifi><p3><pl>/көр<v><tv><ifi><p3><sg>$^./.<sent>$
-% tat: ^Аллаһ/Аллаһ<n><nom>/Аллаһ<n><attr>/Аллаһ<n><nom>+и<cop><p3><sg>$ ^Үзе/Үз<prn><ref><p3><sg><nom>/Үз<prn><ref><p3><sg><nom>+и<cop><p3><sg>$ ^яраткан/ярат<v><tv><gpr_past>/яраткан<adj>/ярат<v><tv><ger_past><nom>/яраткан<adj><advl>/ярат<v><tv><past><p3><sg>/ярат<v><tv><gpr_past><subst><nom>/яраткан<adj><subst><nom>/яраткан<adj>+и<cop><p3><sg>/яраткан<adj><subst><nom>+и<cop><p3><sg>$ ^нәрсәләргә/нәрсә<prn><itg><pl><dat>$ ^карап/кара<vaux><prc_perf>/кара<vaux><gna_perf>/кара<v><tv><prc_perf>/кара<v><tv><gna_perf>$^,/,<cm>$ ^аларның/ул<prn><dem><pl><gen>/алар<prn><pers><p3><pl><gen>$ ^бик/бик<adv>/бик<n><nom>/бик<n><attr>/бик<n><nom>+и<cop><p3><sg>$ ^яхшы/яхшы<ij>/яхшы<adj>/яхшы<adj><advl>/яхшы<adj><subst><nom>/яхшы<adj>+и<cop><p3><sg>/яхшы<adj><subst><nom>+и<cop><p3><sg>$ ^икәнен/и<cop><ger_past><px3sp><acc>$ ^күрде/күр<v><tv><ifi><p3><sg>$^./.<sent>$
-% kum: ^Аллагь/аллагь<n><nom>/аллагь<n><attr>$ ^Оьзю/оьз<prn><ref><px3sp><nom>$ ^яратгъан/ярат<v><tv><gpr_past>/ярат<v><tv><gpr_past><subst><nom>/ярат<v><tv><ger_past><nom>/ярат<v><tv><past><p3><sg>$ ^затлагъа/зат<n><pl><dat>$ ^къарап/къара<v><tv><prc_perf>/къара<v><tv><gna_perf>$^,/,<cm>$ ^олар/о<prn><dem><pl><nom>/олар<prn><pers><p3><pl><nom>$ ^бек/бек<adv>$ ^яхшы/яхшы<adj>/яхшы<adj><subst><nom>$ ^экенин/э<cop><ger_past><px3sp><acc>$ ^гёрген/гёр<v><tv><gpr_past>/гёр<v><tv><gpr_past><subst><nom>/гёр<v><tv><ger_past><nom>/гёр<v><tv><past><p3><sg>$^../..<sent>$
-
-%
-%\textbf{Kazakh:} Құдай Өзінің жаратқандарының бәріне қарап, өте жақсы екенін көрді. \\
-%% (close to kaz) Ходай Үзенең яратканнарының барысына карап, бик яхшы икәнен күрде. \\
-%% (from the actual translation of the Bible) Аллаһ үзе яраткан нәрсәләрне күздән кичерде, һәм, менә, һәр нәрсәнең гаять яхшы икәнлеген күрде. \\
-%\textbf{Tatar:} Аллаһ Үзе яраткан нәрсәләргә карап, аларның бик яхшы икәнен күрде. \\
-%\textbf{Kumyk:} Аллагь Оьзю яратгъан затлагъа къарап, олар бек яхшы экенин гёрген. \\
-%
-\centering
-
-\scalebox{0.81}{
-\begin{tabular}{lll}
-\toprule
- \textbf{Kazakh} & \textbf{Tatar} & \textbf{Kumyk} \\
-\midrule
- Құдай Өзінің жаратқандарының  &  Аллаһ Үзе яраткан нәрсәләргә карап, & Аллагь Оьзю яратгъан затлагъа\\
- бәріне қарап, өте жақсы екенін көрді. &  аларның бик яхшы икәнен күрде. & къарап, олар бек яхшы экенин гёрген. \\
-\midrule
- Құдай\tags{<n><nom>} & Аллаһ\tags{<n><nom>} & Аллагь\tags{<n><nom>}  \\
- Өз\tags{<prn><ref><px3sp><gen>} & Үз\tags{<prn><ref><px3sp><nom>} & Оьз\tags{<prn><ref><px3sp><nom>} \\
- жарат\tags{<v><tv><ger\_past><pl><px3sp><gen>} & ярат\tags{<v><tv><gpr\_past>} & ярат\tags{<v><tv><gpr\_past>} \\
- бәрі\tags{<prn><qnt><px3sp><dat>} & нәрсә\tags{<prn><itg><pl><dat>} & зат\tags{<n><pl><dat>} \\
- қара\tags{<v><tv><gna\_perf>} & кара\tags{<v><tv><gna\_perf>} & къара\tags{<v><tv><gna\_perf>} \\ 
-,\tags{<cm>} & ,\tags{<cm>} & ,\tags{<cm>} \\
- ---      & алар\tags{<prn><pers><p3><pl><gen>} & олар\tags{<prn><pers><p3><pl><nom>} \\
- өте\tags{<adv>} & бик\tags{<adv>} & бек\tags{<adv>} \\
- жақсы\tags{<adj>} & яхшы\tags{<adj>} & яхшы\tags{<adj>} \\
- е\tags{<cop><ger\_past><px3sp><acc>} & и\tags{<cop><ger\_past><px3sp><acc>} & э\tags{<cop><ger\_past><px3sp><acc>} \\
- көр\tags{<v><tv><ifi><p3><sg>} & күр\tags{<v><tv><past><p3><sg>} & гёр\tags{<v><tv><past><p3><sg>} \\  
- .\tags{<sent>} & .\tags{<sent>} & .\tags{<sent>} \\
-\bottomrule
-\end{tabular}
-}
-\label{table:exoutput}
-\caption{An example of the output of each of the morphological transducers for the same 
-  sentence (``And God saw every thing that he had made, and, behold, it was 
-  very good.'', Genesis 1:31). The sentence may be very roughly glossed as ``God own-his created things-to looking,
-  their very good being saw''. The output has been 
-  abbreviated to only show the appropriate tag sequence in context, the actual output would give multiple interpretations
-  that would require further disambiguation. Refer to Appendix~\ref{app:glossary} for tag descriptions.}
-\vspace{-1.5em}
-\end{table*}
-
-\subsection{Categorisation and tagset}
-
-The open categories in Turkic languages can be broadly split into two groups: nominals
-and verbals. The nominals group can be further split into nouns, adjectives and adverbs.
-We define nouns as stems that can be subjects of a finite verb, adjectives as stems that principally qualify 
-nouns, and adverbs as stems that principally modify verbs.
-
-Most of the stems in the lexicon can be used, with no extra suffixes in any of these functions, for 
-example the adjective ‹яхшы› in Tatar can be used as attributively (e.g., ‹яхшы аучы› `good hunter'), adverbially
-(e.g., ‹яхшы бөгелә› `bends well'), and substantively, i.e. as a noun (e.g., ‹яхшы белән› `with a good [person]’). This
-is a productive process and relevant to the morphotactics, as an adjective used substantively may take the 
-full range of case and possessive suffixes, and adjectives used adverbially will take a different set of 
-clitics than adjectives used substantively and attributively.
-
-However, this is not completely productive, as not all adjectives can be used substantively or adverbially. A grammar which allows all adjectives to function
-as nouns and adverbs will overgenerate (and overanalyse). However, most grammatical descriptions of these languages do not 
-mention that there are exceptions to this process.
-
-Other analysers for Turkic languages---such as \texttt{TRmorph} \citep{coltekin2010}---approach this 
-problem by allowing a zero-derivation by which any noun can be `derived' into an adjective or 
-an adverb, any adjective can be `derived' into a noun or adverb, etc.
-
-Our approach is to describe this rather in terms of function (not unlike \cite{hengeveld92}). We 
-posit that nominals can be used either substantively, attributively or adverbially. Each part of speech 
-has a `default' function: nouns are by default substantive, adjectives are by default attributive, 
-and adverbs adverbial. When they are used outside this default use they receive a tag to mark 
-their function. Where these functions are ambiguous, they may be disambiguated 
-in context.
-
-The advantage of this approach as opposed to simply allowing one part of speech to derive into
-another is that it allows us to be more principled, and decreases overgeneration;
-for example, nouns used attributively will not take the whole adjective continuation (e.g.
-they will not take comparison, and cannot later substantivise). At the same time, it prevents 
-having to have two lexemes, one for each usage.
-
-There are certain lexemes that require more explicit categorisation, for example, the word 
-for `today'\footnote{Tatar: бүген; Kazakh: бүгін; Kumyk: бугюн. The word is etymologically composed of the words for `this' and `day'.} 
-cannot be used attributively in Kypchak languages without the presence of an extra morpheme -KI. Compound nouns
-using possessive morphology, such as Kazakh ауа райы `weather' (lit.
-`air condition'), do not take additional morphology the same way other substantives in the language do, due to the presence of a final possessive morpheme (e.g., ‹ауа райынан› `weather-\gmk{abl}', not *‹ауа райыдан›).
-%FIXME: Explain the compound 'weather'.
-
-%\begin{tabular}{lccc}
-%          &  Substantive & Attributive & Adverbial \\ 
-%Noun      &     X        &    ?        &  ?\\ 
-%Adjective &     ?        &   X         &  ? \\ 
-%Adverb    &     ?        &   ?         &  X \\ 
-%\end{tabular}
-
-% FIXME: how's the title now?
-\subsection{Orthography-phonology mapping issues}
-
-There is a class of phenomena encountered during the development of these transducers, united by the fact that the morphophonology needs information about the stem's phonological form that is not provided by the orthographic representation.
-These phenomena were left by \citet{washington2012} for future work, but many of them been implemented in the transducers described in the current paper.
-
-Types of these phenomena include single characters used to represent multiple phonological forms, loanwords, and acronyms and numerals.
-A brief overview of each follows, along with examples and details of our solutions.
-
-\subsubsection{Ambiguous characters}
-In Tatar, the `yoticised' vowel letters ‹е›, ‹я›, and ‹ю› each ambiguously represent a set of sounds.
-For example, ‹е› is the non-initial orthographic variant of ‹э› (as in ‹дәресләр› `lessons'), but can also represent /j/ followed by the phoneme of ‹э› (as in ‹егетләр› `boys') or ‹ы› (as in ‹еллар› `years').
-
-In Kumyk, ‹ё› and ‹ю› are used between consonants to represent front rounded vowels (as in ‹гёзлер› `eyes' and ‹гюнлер› `days'), but word-initially can represent ‹й› followed by either a front vowel (as in ‹юреклер› `hearts', ‹ёнкюлер› `darlings') or a back vowel (as in ‹юлдузлар› `stars', ‹ёллар› `roads').
-
-In Kazakh, the vowels /{\qipa ɘ}/ ‹і› and /ə/ ‹ы› followed by a glide /w/ ‹у› or /j/ ‹й› are written ‹у› and ‹и›, depending on the glide; for example.
-The character ‹ю› is used in turn to represent ‹й› followed by either multi-phoneme value of ‹у›.
-For example, ‹қиюда› /qəjəwd{\qipa ɑ}/ `in the process of chopping down' and ‹киюде› /k{\qipa ɘ}j{\qipa ɘ}wd{\qipa i̯ɘ}/ `in the process of getting dressed' have back- and front-harmonising values, respectively, for ‹и› and ‹ю›.
-
-These problems can often be solved by building the context into the morphophonological rules, though the addition of new contexts and exceptions to existing contexts results in rather complex \texttt{twol} rules.
-
-\subsubsection{Loanwords}
-In Kazakh, Tatar, and Kumyk, most Russian loanwords are spelled using Russian orthography, despite incompatibilities between Russian letter-to-phoneme mapping and that of the Turkic languages.
-
-In Kumyk, the character ‹ё› is used between consonants to represent a mid front rounded vowel (e.g., ‹сёзлер› `words'); however, in Russian words, ‹ё› is used after a consonant to represent the sound of ‹о› while indicating that the previous consonant is palatalised (e.g., ‹самолётлар› `aeroplanes').
-
-In Tatar, the character ‹и› normally represents a high front vowel (e.g., ‹галимнәр› `scientists'), but in Russian words it harmonises as a back vowel (e.g., ‹артистлар› `artists').
-
-In Kazakh, ‹е› is a front diphthong (e.g., ‹елдің› `country-\gmk{gen}'), but in e.g., family names with Russian morphology, it harmonises as a back vowel (e.g., ‹Назарбаевтың› `Nazarbayev-\gmk{gen}').
-
-Our solution for these consistent exceptions is to make a separate continuation lexicon for e.g., Russian nouns, which adds a character which the phonology deletes, but which triggers various phonological rules.
-
-Since \texttt{twol} ``phonological'' rules can be understood to be applied all at one stage (as opposed to being ordered), rules triggered by these characters must make reference to abstract characters that exist on the input tape but are null on the output tape.
-This complicates existing rules that have been designed to ignore null characters on the output tape.
-Hence, the \texttt{twol} rules resulting from the combination of processes simultaneously triggered by and ignoring output-null characters can quickly become quite unwieldy.
-
-\subsubsection{Acronyms and numerals}
-Acronyms and numerals are challenging as they will often be pronounced out loud
-   in their non-abbreviated forms, for example in Kazakh, ‹30-дан› `from thirty' would be spelled out ‹отыздан›, and 5-тен `from 5' would be spelled out ‹бестен›. The ablative
-   suffix -DAn alters for the phonology, but in the numeral string there is no indication of how they should alter.
-
-Dealing with these phenomena would not be necessary if we were setting out to develop a simple computational
-model of the phonology of the language. However, a wide-coverage morphological analyser and generator needs to be 
-able to deal with all phenomena that are found in corpora.
-
-Regarding numerals, work has been done for Finnish in the finite-state framework by \citep{karttunen06}; however, this relies on converting all numerals to their fully spelt out form, which would involve complex operations
-on the transducer.
-
-% FIXME: cyrillic in \texttt
-Our solution is to add phonological information at the end of morphemes that need it in the form of special ``abstract letters'' that trigger phonological processes at the morphophonological stage and are deleted.
-For example, the string \texttt{5<num><subst><abl>}\footnote{The resulting forms, 5-тен\super{\texttt{kaz}}, 5-тән\super{\texttt{tat}}, and 5-ден\super{\texttt{kum}} would be pronounced бестен\super{\texttt{kaz}}, биштән\super{\texttt{tat}}, and бешден\super{\texttt{kum}}.} has
-the morphotactic representation 5\{э\}\{с\}>-\{D\}\{A\}н, where \{э\} and \{с\} stand for phonological triggers, > represents a morpheme boundary, and \{D\}\{A\}н is the representation of the ablative morpheme at the morphotactic level.
-The symbol \{э\} signals that the following vowel needs to harmonise to a front unrounded vowel, and \{с\} signals that there is a final voiceless consonant.
-So, with 5\{э\}\{с\}\footnote{Note that there is not one extra symbol for every sound of the language, but a much reduced set.
-E.g., \{с\} is used to represent all voiceless consonants.} in the lexicon, and a consistent continuation lexicon specifying the underlying form of case affixes, the rules that operate on \{D\} and \{A\} are able to do produce the correct output form in each language, avoiding the incorrect default *5-дан.
-
-%FIXME: example of implementation
-
-This solution has the same issue as loanwords above.
-
-\section{Evaluation}\label{sec:evaluation}
-
-We have evaluated the morphological analysers in two ways.
-The first was by calculating the naïve coverage and mean ambiguity on freely available corpora.
-Naïve coverage refers to the percentage of surface forms in a given corpora that receive at least one analysis.
-Forms counted by this measure may have other analyses which are not delivered by the transducer.
-The mean ambiguity measure was calculated as the average number of analyses returned per token in the corpus.
-
-%by performing an evaluation of precision and recall on some smaller, hand-validated test sets.
-
-\subsection{Corpora}
-
-We tested the coverage of the Kazakh and Tatar analysers over three separate domains: encyclopaedic text,\footnote{The following Wikipedia dumps were used: \texttt{kkwiki-20131006-pages-articles.xml.bz2} and \texttt{ttwiki-20130225-pages-articles.xml.bz2}.}
- news,\footnote{All content from \url{http://www.azattyq.org/} for 2010 was used for Kazakh, as well as all content from 2005 to 2011 on \url{http://tat.tatar-inform.ru/} for Tatar.} and religion.\footnote{We used a Kazakh bible translation available from \url{https://kkitap.net/} and Quran translation available from \url{http://kuran.kz/}, as well as a Tatar translation of the New Testament available from \url{http://ibt.org.ru/} and Quran translation available from \url{http://tanzil.net/}.}  As there is currently no Wikipedia in Kumyk, we tested only news and religion.\footnote{The bible corpus consists of Genesis and the New Testament, as available from \url{http://ibt.org.ru/}, and the news corpus consists of all Kumyk content from \url{http://sh-tavisi.etnosmi.ru/}.}
-
-% Kazakh, religion: kaz.bible.kkitap.txt.bz2, kaz.quran.altay.txt.bz2
-% Kazakh, wp: kaz.wikipedia2.2013.txt.bz2
-% Kazakh, news: kaz.rferl.2010.xml.bz2, kaz.rferl2.2012.txt
-
-% Tatar, wp: tat.wikipedia.2013-02-25.txt.bz2
-% Tatar, news: tat.news.2005-2011_300K-sentences.txt.bz2
-% Tatar, religion: tat.nt.ibt.txt, tat.quran.nughmani.txt.bz2
-
-% Kumyk, news: kum.yoldash.2013.txt, kum.articles.kumukia.txt.bz2
-% Kumyk, religion: kum.nt.ibt.txt, (kum.genesis.ibt.txt)
-
-The coverage of each transducer over the various corpora is shown in table \ref{table:corpora}.
-
-\begin{table}
-\begin{center}
-\begin{tabular}{llrr}
-\toprule
-\textbf{Language} & \textbf{Corpus} & \textbf{Tokens} & \textbf{Coverage} (\%) \\
-\midrule
-\multirow{4}{*}{Kazakh} & Wikipedia & 25.6M & 85.61 $\pm$ 1.37 \\
-	& News & 3.8M & 92.12 $\pm$ 2.72  \\
-	& Religion & 851K & 92.49 $\pm$ 1.66 \\\cline{2-4}
-	& Average & -- & 90.07 $\pm$ 1.91 \\
-\midrule
-\multirow{4}{*}{Tatar} & Wikipedia & 159K & 86.35 $\pm$ 2.17 \\
-	& News & 5.2M & 89.75 $\pm$ 0.07 \\
-	& Religion & 382K & 91.25 $\pm$ 2.55 \\\cline{2-4}
-	& Average & -- & 89.12 $\pm$ 1.60 \\
-\midrule
-% news = kum.yoldash.2013.txt
-% religion = kum.genesis.ibt.txt, kum.nt.ibt.txt
-\multirow{4}{*}{Kumyk} & Wikipedia & -- & -- \\
-        & News & 286K &  91.10 $\pm$ 0.86 \\
-	& Religion & 227K & 92.47 $\pm$ 1.03 \\\cline{2-4}
-	& Average & -- & 91.78 $\pm$ 0.94 \\
-\bottomrule
-\end{tabular}
- \caption{Corpora used for naïve coverage tests}
- \label{table:corpora}
-\end{center}
-\vspace{-2.5em}
-\end{table}
-
-\subsection{Precision and recall}
-
-Precision and recall are measures of the average accuracy of analyses provided by a morphological transducer.
-Precision represents the number of the analyses given for a form that are correct.
-Recall is the percentage of analyses that are deemed correct for a form (by comparing against a gold standard) that are provided by the transducer.
-
-To calculate precision and recall, it was necessary to create a hand-verified list of surface forms and their analyses.
-We extracted 1,000 unique surface forms at random from a news corpus for each language, and checked that they were valid words in the languages and correctly spelt.
-Where a word was incorrectly spelt or deemed not to be a form used in the language, it was discarded and a new random word selected.
-
-%FIXME: footnote with urls
-This list of surface forms was then analysed with the most recent version of the analyser, and each analysis was checked.
-Where an analysis was erroneous, it was removed; where an analysis was missing, it was added.
-This process gave us a `gold standard' morphologically analysed word list of 1,000 forms for Kazakh and Tatar and 500 forms for Kumyk, along with their analyses.
-The list is publically available for each language in Apertium's svn repository.%\footnote{\url{https://apertium.svn.sourceforge.net/svnroot/apertium/branches/apertium-kir/eval/ref.1000.txt}}
-
-The lower number of forms for Kumyk was due to the limited access to native speakers of Kumyk. 
-The process for the Kumyk evaluation was also slightly different.
-They were first checked by a non-expert native speaker of Kumyk and then the result was proofread by the authors with the help of a dictionary.
-
-We then took the same list of surface forms and ran them through the morphological analyser once more.
-Precision was calculated as the number of analyses which were found in both the output from the morphological analyser and the gold standard, divided by the total number of analyses output by the morphological analyser.
-
-Recall was calculated as the total number of analyses found in both the output from the morphological analyser and the gold standard, divided by the number of analyses found in the morphological analyser plus the number of analyses found in the gold standard but not in the morphological analyser.
-
-The results for precision and recall are presented in table~\ref{table:precrecall}.
-
-The low recall for Kazakh can be explained by the fact that the corpus is much bigger, giving 
-more hapax words and proper names. There were 403 out-of-vocabulary words out of 1,000 in the Kazakh list.
-Of these 403, 160 were proper nouns, and 148 were common nouns. The lower precision for the 
-Tatar transducer can be partly explained by the less transparent orthography of Tatar.
-
-%Due to limited access to native Kumyk speakers, the authors performed precision themselves based on their admittedly incomplete knowledge of the language.  
-Recall is low for Kumyk due mostly to some systematic gaps in the morphotactics, such as lack of attributive analyses in several continuation lexica.
-There are also a higher number of missing verb stems (over half of all missing analyses were verb forms) in the Kumyk results than in the Kazakh and Tatar, where missing verb stems accounted for 10\% and 8\% respectively.
-
-% tat:
-%Precision: 95.0339558573854
-%Recall: 85.65416985462892
-%F-score: 90.10060362173037
-
-% kaz:
-% Precision: 98.61202908129543
-% Recall: 57.986785853089785
-% F-score: 73.02985805188449
-
-\begin{table}
-\begin{center}
-	\begin{tabular}{lrr}
-	\toprule
-		\textbf{Language} & \textbf{Precision (\%)} & \textbf{Recall (\%)} \\
-	\midrule
-		Kazakh & 98.61 &  57.98 \\
-		Tatar & 95.03 & 85.65 \\
-		Kumyk & 96.57 & 69.11 \\
-	\bottomrule
-	\end{tabular}
-	\caption{Precision and recall presented as percentages}
-	\label{table:precrecall}
-\end{center}
-\vspace{-1.5em}
-\end{table}
-
-\section{Future work}
-
-%Code switching - +example
-%More languages: Nogai, Bashkir, Karakalpak, Karachay-Balkar
-
-One direction for future work is to develop transducers for more languages.
-We have already constructed usable prototype transducers for three other Kypchak languages: Bashkir (North), Nogay (South), and Karakalpak (South).
-Since our ability to develop transducers is limited by availability of resources, including corpora in the languages and native-speaker consultants, the Western Kypchak languages (aside from Kumyk) have been more neglected by our team.
-However, these language communities would benefit from computational tools (such as spellcheckers) for their languages, and work on them may be bootstrapped from the existing transducers, so working on morphological transducers for these languages is also a priority.
-
-The principle obstacle to increasing coverage of the lexicons is the categorisation of stems.
-Future work would be investigating ways of automatically categorising stems by subcategory.
-For example, verbs stems by transitivity; adjective stems by whether  they can be used adverbially or substantively; etc.
-
-\section{Conclusions}\label{sec:conclusions}
-
-We have described morphological transducers for three Kypchak languages---one from each branch of Kypchak---including the development process and performance of the analysers.
-The development of the third transducer (for a related language) was substantially quicker than the first two as a result of being able to reuse large portions of the morphotactic description from the first two transducers.
-
-\section*{Acknowledgements}
-
-We would like to thank the Google Code-in (2011) for supporting the original development 
-of the Kazakh transducer, and in particular the effort by Nathan Maxson. We 
-would also like to thank the Google Summer of Code (2012) for supporting the 
-development of both the Kazakh and the Tatar transducers.
-
-The authors would also like to express their gratitude to Aida Sundetova 
-and Ağarhim Sultanmuradov
-for assistance in evaluating precision and recall.
-
-\bibliographystyle{lrec2012}
-\bibliography{kipchak-paper}
-
-\appendix
-
-\section{Glossary of symbols}
-\label{app:glossary}
-\centering
-\begin{tabular}{cl}
-\toprule
- \textbf{Symbol} & \textbf{Description}\\
-\midrule
- \tags{n} & Noun  \\
- \tags{v} & Verb  \\
- \tags{adj} & Adjective  \\
- \tags{adv} & Adverb  \\
- \tags{prn} & Pronoun  \\
- \tags{cm} &  Comma \\
- \tags{cop} & Copula  \\
- \tags{ifi} & Past definite  \\
- \tags{past} & Past  \\
- \tags{ger\_past} & Past gerund \\
- \tags{gna\_past} & Past verbal adverb  \\
- \tags{gpr\_past} & Past verbal adjective  \\
- \tags{itg} & Interrogative  \\
- \tags{nom} & Nominative \\
- \tags{gen} & Genitive  \\
- \tags{dat} & Dative  \\
- \tags{abl} & Ablative \\
- \tags{p1} & First person  \\
- \tags{p3} & Third person  \\
- \tags{px1sg} & First person, singular possessive  \\
- \tags{px3sp} & Third person possessive  \\
- \tags{pers} & Personal  \\
- \tags{qnt} & Quantifier  \\
- \tags{ref} & Reflexive  \\
- \tags{sg} & Singular  \\
- \tags{pl} &  Plural \\
- \tags{tv} & Transitive  \\
- \tags{sent} & End-of-sentence marker  \\
-
-\bottomrule
-
-\end{tabular}
 \end{document}

### 2015-06-17 2015-turklang-kazdep/paper.tex
stuffs
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60802 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-\title{Dependency Treebank for Kazakh based on Universal Dependencies}
-\name{Francis Tyers$^a$ and Jonathan Washington$^b$}
-\address{$^a$ hargle,\\
-	$^b$ Departments of Linguistics and Central Eurasian Studies, Indiana University, Bloomington, IN 47405, USA}
+\title{Towards a universal-dependency treebank for Kazakh}
+\name{Francis M. Tyers$^\dagger$ and Jonathan Washington$^\ast$}
+\address{$^\dagger$ HSL-fakultehta, UiT Norgga árktalaš universitehta, N-9015 Tromsø, Norway\\
+	$^\ast$ Departments of Linguistics and Central Eurasian Studies, Indiana University, Bloomington, IN 47405, USA}
 
@@
 
-\subsection{The Kazakh Language}
+\subsection{Kazakh}
 
@@
 
+\section*{Acknowledgements}
+
 \end{document}

### 2015-06-18 2015-turklang-kazdep/paper.tex
stuffs
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60809 72bbbca6-d526-0410-a7d9-f06f51895060
some stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60810 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \title{Towards a universal-dependency treebank for Kazakh}
-\name{Francis M. Tyers$^\dagger$ and Jonathan Washington$^\ast$}
-\address{$^\dagger$ HSL-fakultehta, UiT Norgga árktalaš universitehta, N-9015 Tromsø, Norway\\
-	$^\ast$ Departments of Linguistics and Central Eurasian Studies, Indiana University, Bloomington, IN 47405, USA}
+\name{Francis M. Tyers$^a$ and Jonathan Washington$^b$}
+\address{$^a$ HSL-fakultehta, UiT Norgga árktalaš universitehta, N-9015 Tromsø, Norway\\
+	$^b$ Departments of Linguistics and Central Eurasian Studies, Indiana University, Bloomington, IN 47405, USA}
+
+\keywords{Kazakh, treebank, dependency grammar, universal dependency}
 
@@
 
-\subsection{Treebanks}
+\subsection{Treebanks} % Related work
 
 \section{Methodology}
+
+\section{Conclusions}
 

### 2015-06-18 2015-turklang-kazdep/paper.tex
more stuff and bib
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60811 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{xunicode}
+
 \usepackage{multirow}
@@
 \usepackage{natbib}
+\bibdata{paper}
 
-\usepackage{times}
+%\usepackage{times}
+\setmainfont{Times New Roman}
 
@@
 \address{$^a$ HSL-fakultehta, UiT Norgga árktalaš universitehta, N-9015 Tromsø, Norway\\
-	$^b$ Departments of Linguistics and Central Eurasian Studies, Indiana University, Bloomington, IN 47405, USA}
+	$^b$ Departments of Linguistics and Central Eurasian Studies, Indiana University, Bloomington, IN 47405, USA\footnote{Corresponding author.  E-mail address: jonwashi@indiana.edu}}
 
-\keywords{Kazakh, treebank, dependency grammar, universal dependency}
+\keywords{Kazakh; treebank; dependency grammar; universal dependency}
 
@@
 \section{Introduction}
+
+\citet{hargle}
 
@@
 
+\bibliographystyle{apa}
+\bibliography{paper}
+
 \end{document}

### 2015-06-18 2015-turklang-kazdep/paper.tex
spacing more like what they're asking for
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60812 72bbbca6-d526-0410-a7d9-f06f51895060
floss
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60815 72bbbca6-d526-0410-a7d9-f06f51895060
some sources
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60818 72bbbca6-d526-0410-a7d9-f06f51895060
updates
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60824 72bbbca6-d526-0410-a7d9-f06f51895060
desc. corp
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60825 72

@@
-\documentclass[a4paper,11pt,twocolumn]{article}
+\documentclass[a4paper,11pt, onecolumn]{article}
 
@@
 
-\title{Towards a universal-dependency treebank for Kazakh}
+\title{Towards a free/open-source universal-dependency treebank for Kazakh}
 \name{Francis M. Tyers$^a$ and Jonathan Washington$^b$}
@@
 
-\abstract{hargle treebank bargle dependency grammar}
+\abstract{250-300-word ``extended abstract in English''}
 
@@
 
-\citet{hargle}
+\citet{Lynn12}, \citet{Atalay03}, \citet{Oflazer03}, \citet{DeMarneffe14}
 
@@
 
+\subsection{Corpus}
+
+\begin{table}
+  \centering
+  \begin{tabular}{|l|l|r|r|r|}
+    \hline
+    \textbf{Document} & \textbf{Description} & \textbf{Sentences} & \textbf{Tokens} & \textbf{Avg. length}\\
+    \hline
+    UDHR & Legal text on human rights & - & - & - \\
+    Phrasebook & Phrases from Wikitravel & 37 & 205 & 5.5 \\
+    Жиырма Бесінші Сөз & Philosophical text & 34 & 525 & 15.4 \\ 
+    Өлген қазан & Folk tale & 8 & 140 & 17.5 \\
+    Азамат къайда? & Story for children and language learners & 48 & 434 & 9.0 \\
+    Футболдан әлем чемпионаты 2014  & Wikipedia article (2014 World Cup) & 14 & 246 & 17.5 \\
+    Иран & Wikipedia article (Iran) & - & - & - \\
+    Радиан & Wikipedia article (Radian) & - & - & - \\
+    \hline
+  \end{tabular}
+  \caption{Composition of the corpus. The corpus covers a range of genres and text types.}
+\end{table}
+
+\subsection{Preprocessing}
+
+% tokenisation
+\cite{Washington14}
+
+\subsection{Annotation guidelines}
+
+\begin{table}
+  \centering
+  \begin{tabular}{|l|l|}
+    \hline
+    \textbf{Label} & \textbf{Description} \\
+
+    \hline
+  \end{tabular}
+  \caption{Universal dependency label set}
+\end{table}
+
+\section{Evaluation}
+
+\section{Future work}
+
+%crosslingual parsing: Tatar, Kumyk, Kyrgyz, Tuvan
+
 \section{Conclusions}

### 2015-06-18 2015-turklang-kazdep/paper.tex
desc. corp
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60826 72bbbca6-d526-0410-a7d9-f06f51895060

@@
     \hline
-    UDHR & Legal text on human rights & - & - & - \\
+    Declaration on Human Rights & Legal text on human rights & - & - & - \\
     Phrasebook & Phrases from Wikitravel & 37 & 205 & 5.5 \\

### 2015-06-19 2015-turklang-kazdep/paper.tex
harg
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60827 72bbbca6-d526-0410-a7d9-f06f51895060
another change
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60832 72bbbca6-d526-0410-a7d9-f06f51895060

--- 2015-06-18 2015-turklang-kazdep/paper.tex
@@
     \hline
-    UDHR & Legal text on human rights & - & - & - \\
+    Declaration on Human Rights & Legal text on human rights & 25 & 409 & 16.3 \\
     Phrasebook & Phrases from Wikitravel & 37 & 205 & 5.5 \\
@@
     Өлген қазан & Folk tale & 8 & 140 & 17.5 \\
-    Азамат къайда? & Story for children and language learners & 48 & 434 & 9.0 \\
+    Азамат қайда? & Story for children and language learners & 48 & 434 & 9.0 \\
     Футболдан әлем чемпионаты 2014  & Wikipedia article (2014 World Cup) & 14 & 246 & 17.5 \\
-    Иран & Wikipedia article (Iran) & - & - & - \\
-    Радиан & Wikipedia article (Radian) & - & - & - \\
+    Иран & Wikipedia article (Iran) & 111 & 1562 & 14.0 \\
+    Радиан & Wikipedia article (Radian) & 2 & 17 & 8.5 \\
     \hline
@@
 
-\subsection{Annotation guidelines}
+% multiword units
+
+\section{Annotation guidelines}
 
@@
 
+\subsection{Copula}
+
+\subsection{Coordination}
+
+\subsection{Complex nominals}
+
+% compounds + izafet
+
+\subsection{Non-finite clauses}
+
 \section{Evaluation}
@@
 %crosslingual parsing: Tatar, Kumyk, Kyrgyz, Tuvan
+
+%language-specific dependency relations --- (acl:relcl, ...)
 

### 2015-06-19 2015-turklang-kazdep/paper.tex
werw
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60841 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{multirow}
+\usepackage[small,bf]{caption} % CHECK IF THIS IS OK
 \usepackage[colorlinks=true,citecolor=black,linkcolor=black,urlcolor=blue]{hyperref}
@@
 
-\abstract{250-300-word ``extended abstract in English''}
+%250-300-word ``extended abstract in English''
+\abstract{
+This article describes the first steps towards a free/open-source dependency treebank
+for Kazakh based on universal dependency (UD) annotation standards. The treebank is 
+based on texts from a range of open-source and public domain sources ensuring its 
+free availability and extensibility. 
+}
 
@@
 
+% case and function
+% "zero" derivation
+
 \subsection{Treebanks} % Related work
+
+% what are treebanks used for ?
 
@@
   \centering
+\begin{small}
   \begin{tabular}{|l|l|r|r|r|}
@@
     \hline
-    Declaration on Human Rights & Legal text on human rights & 25 & 409 & 16.3 \\
+    UN Declaration on Human Rights & Legal text on human rights & 25 & 409 & 16.3 \\
     Phrasebook & Phrases from Wikitravel & 37 & 205 & 5.5 \\
@@
     Өлген қазан & Folk tale & 8 & 140 & 17.5 \\
-    Азамат қайда? & Story for children and language learners & 48 & 434 & 9.0 \\
+    Азамат қайда? & Story for language learners & 48 & 434 & 9.0 \\
     Футболдан әлем чемпионаты 2014  & Wikipedia article (2014 World Cup) & 14 & 246 & 17.5 \\
@@
     \hline
+    \multicolumn{2}{c|}{~} & - & - & - \\\cline{3-5}
   \end{tabular}
-  \caption{Composition of the corpus. The corpus covers a range of genres and text types.}
+\end{small}
+  \caption{Composition of the corpus. The corpus covers a range of genres and text types from 
+    free and public-domain sources.}
 \end{table}
@@
 % tokenisation
+% http://universaldependencies.github.io/docs/u/overview/tokenization.html
 \cite{Washington14}
@@
 \section{Annotation guidelines}
+
+% general procedure
+
+% - disambiguate
+% - annotate
+% - check the tree structure
+% - fix errors
+% - check for label consistency
+% - fix errors
 

### 2015-06-19 2015-turklang-kazdep/paper.tex
ergerg
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60843 72bbbca6-d526-0410-a7d9-f06f51895060

@@
     Өлген қазан                    & Folk tale                  & 8 & 140 & 17.5 \\
+    Ер Төстік                      & Folk tale                  & 23 & 203 & 8.8 \\
     Азамат қайда?                  & Story for language learners & 48 & 434 & 9.0 \\
@@
     \hline
-    \multicolumn{2}{c|}{~} & - & - & - \\\cline{3-5}
+    \multicolumn{2}{c|}{~} & 302 & 3741 & 12.38 \\\cline{3-5}
   \end{tabular}

### 2015-06-19 2015-turklang-kazdep/paper.tex
reger
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60844 72bbbca6-d526-0410-a7d9-f06f51895060

@@
     \hline
-    \multicolumn{2}{c|}{~} & 302 & 3741 & 12.38 \\\cline{3-5}
+    \multicolumn{2}{c|}{~} & 302 & 3741 & 12.3 \\\cline{3-5}
   \end{tabular}
@@
 
+Preprocessing the corpus consists of running the text through the Kazakh morphological
+analyser \citep{Washington14}, which also performs tokenisation of multiword units based 
+the longest match left-to-right. Tokenisation for Kazakh is a non-trivial task, and so
+we do not simply take space as a delimiter. The morphological analyser returns all 
+the possible morphological analyses for each word based on a lexicon of around 20,000 lexemes.
+After tokenisation and morphological analysis, the text is processed with a constraint-grammar 
+based disambiguator for Kazakh consisting of 113 rules which remove inappropriate 
+analyses in context. This reduces the average number of analyses per word from around 3.4
+to around 1.7.
+
 % tokenisation
 % http://universaldependencies.github.io/docs/u/overview/tokenization.html
-\cite{Washington14}
 
+\subsubsection{Tokens and words}
+
+% simple tokens
 % multiword units
+% copula
+% clitics
+% productive suffixes -lı (cite wafflelazor)
 
@@
 
+%stabilise guidelines
+
 %crosslingual parsing: Tatar, Kumyk, Kyrgyz, Tuvan

### 2015-06-19 2015-turklang-kazdep/paper.tex
more abs
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60845 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 This article describes the first steps towards a free/open-source dependency treebank
-for Kazakh based on universal dependency (UD) annotation standards. The treebank is 
-based on texts from a range of open-source and public domain sources ensuring its 
-free availability and extensibility. 
+for Kazakh based on universal dependency (UD) annotation standards. The treebank contains
+302 sentences and is based on texts from a range of open-source and public domain sources.
+This ensures its free availability and extensibility. Texts in the treebank are first morphologically 
+analysed and disambiguated and then annotated manually for dependency structure. 
 }

### 2015-06-19 2015-turklang-kazdep/paper.tex
more
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60846 72bbbca6-d526-0410-a7d9-f06f51895060
header fixes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60847 72bbbca6-d526-0410-a7d9-f06f51895060
beginnings of background
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60849 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \name{Francis M. Tyers$^a$ and Jonathan Washington$^b$}
+\shortauthor{Tyers \& Washington}
 \address{$^a$ HSL-fakultehta, UiT Norgga árktalaš universitehta, N-9015 Tromsø, Norway\\
@@
 \abstract{
-This article describes the first steps towards a free/open-source dependency treebank
-for Kazakh based on universal dependency (UD) annotation standards. The treebank contains
-302 sentences and is based on texts from a range of open-source and public domain sources.
-This ensures its free availability and extensibility. Texts in the treebank are first morphologically 
-analysed and disambiguated and then annotated manually for dependency structure. 
+This article describes the first steps towards a free/open-source dependency treebank for Kazakh based on universal dependency (UD) annotation standards.
+The treebank contains 302 sentences and is based on texts from a range of open-source and public domain sources.
+This ensures its free availability and extensibility.
+Texts in the treebank are first morphologically analysed and disambiguated and then annotated manually for dependency structure.
 }
@@
 \maketitleabstract{}
+\thispagestyle{firststyle}
 
@@
 
+Kazakh (қазақ тілі), a Turkic language of Central Asia with around 12 million speakers, hargle bargle. %FIXME
+
+In the process of designing a Kazakh transducer \citep{Washington14}, several decisions were made about how various inflectional morphemes affect the part of speech and syntactic function of various word forms.
+These decisions include the relations to other words defined by ``case'' morphemes, and processes of ``zero derivation''.
+
 % case and function
+%% indefinite acc/gen = "nom"
+%% attr/subst/advl function of cases
 % "zero" derivation
+%% attr/subst/etc.
+%% copulas?
 
@@
 
-\bibliographystyle{apa}
+\bibliographystyle{apalike}
 \bibliography{paper}

### 2015-06-20 2015-turklang-kazdep/paper.tex
abstract spacing fixed, part of background first draft
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60854 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \setmainfont{Times New Roman}
+
+\newcommand{\tag}[1]{\texttt{‹#1›}}
 
@@
 
+All of the traditionally defined ``cases'' in Kazakh have a variety of uses.
+The morphemes that mark the accusative and genitive cases are only used when the noun phrase is semantically definite.
+This means that indefinite accusative and genitive as well as nominative noun phrases are all unmarked for case, and hence are all ambiguous.
+Our Kazakh transducer marks these all as \tag{nom}, but their use may be disambiguated.
+Besides identifying the semantic role of an unmarked noun phrase, there are various agreement properties that make it clear what ``case'' the noun phrase ``truly'' receives: genitive noun phrases must have a corresponding noun phrase with possessive morphology that agrees in person with the genitive noun phrase; and nominative noun phrases must have a corresponding predicate that agrees in person with the nominative noun phrase, e.g., a verb or a copula.
+
+Copulas may also be non-overt.  In the present tense, all but third person copulas have realised forms, but the third person copula does not surface.
+
+Adjectives in Kazakh are by default attributive, but many may also be used either nominally (i.e., as nouns) or adverbially (i.e., as adverbs), or both.
+Our morphological transducer internally provides a series of classes; depending on class membership, an adjective may receive readings \tag{adj}\tag{advl} or \tag{adj}\tag{subst}, besides the default attributive reading of \tag{adj}.
+If an adjective may be used substantivally, then it will receive a range of substantive morphology as well.
+
 % case and function
@@
     Жиырма Бесінші Сөз             & Philosophical text        & 34 & 525 & 15.4 \\ 
-    Өлген қазан                    & Folk tale                  & 8 & 140 & 17.5 \\
-    Ер Төстік                      & Folk tale                  & 23 & 203 & 8.8 \\
+    Өлген қазан                    & Folk tale from Wikisource  & 8 & 140 & 17.5 \\
+    Ер Төстік                      & Folk tale from Wikisource  & 23 & 203 & 8.8 \\
     Азамат қайда?                  & Story for language learners & 48 & 434 & 9.0 \\

### 2015-06-20 2015-turklang-kazdep/paper.tex
correct title with genitive subject
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60856 72bbbca6-d526-0410-a7d9-f06f51895060

@@
     Жиырма Бесінші Сөз             & Philosophical text        & 34 & 525 & 15.4 \\ 
-    Өлген қазан                    & Folk tale from Wikisource  & 8 & 140 & 17.5 \\
+    Қожанасырдың тойға баруы       & Folk tale from Wikisource  & 8 & 140 & 17.5 \\
     Ер Төстік                      & Folk tale from Wikisource  & 23 & 203 & 8.8 \\

### 2015-06-20 2015-turklang-kazdep/paper.tex
tables repositioned
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60857 72bbbca6-d526-0410-a7d9-f06f51895060
what I've done so far
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60860 72bbbca6-d526-0410-a7d9-f06f51895060
Wefwef
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60862 72bbbca6-d526-0410-a7d9-f06f51895060
fixed makefile
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60863 72bbbca6-d526-0410-a7d9-f06f51895060
some more stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60867 72

@@
-\documentclass[a4paper,11pt, onecolumn]{article}
+\documentclass[a4paper,11pt, onecolumn,twoside]{article}
 
 \usepackage{turklang}
+
+%%%%% Language and font stuff %%%%%
 
@@
 \usepackage{xunicode}
+
+%\usepackage{times}
+\setmainfont{Times New Roman}
+\newfontfamily\udfont[Scale=MatchLowercase,Letters=SmallCaps]{Linux Libertine O}
+
+%%%%% General formatting stuff %%%%%
 
@@
 
-%\usepackage{times}
-\setmainfont{Times New Roman}
+%%%%% Stuff for this paper %%%%%
+\usepackage{tikz-dependency}
+\newcommand{\gmk}[1]{\texttt{#1}}
+\newcommand{\tag}[1]{\texttt{‹#1›}}
+\newcommand{\udtag}[1]{\texttt{@#1}}
+\newcommand{\udlabel}[1]{{\udfont #1}}
+%\newcommand{\corrauthor}[1]{\footnotemark\footnotetext{hargle}}
 
-\newcommand{\tag}[1]{\texttt{‹#1›}}
+%%%%% Paper metadata %%%%%
 
@@
 
+%%%%% End preamble %%%%%
+
 \begin{document}
 
+%\thanks{hargle}
 \maketitleabstract{}
@@
 
-\subsection{Kazakh}
+\subsection{Kazakh}\label{sec:kazakh}
 
@@
 
-\begin{table}
+\begin{table}[htbp]
   \centering
@@
 
+Some nominals occur together to act as a single nominal.
+
+One example of this is indefinite genitives.
+In such constructions, a ``modifying'' nominal is morphologically not marked (which means it will be tagged as \tag{nom}, as discussed in §\ref{sec:kazakh}), though should be considered an indefinite genitive.
+The modified element is a nominal with third-person possessive morphology, an example of which is provide in figure~\ref{fig:indefgen}.
+
+\begin{figure}[htbp]
+	\centering
+	\begin{dependency}[theme = simple, font = \small]
+	   \begin{deptext}[column sep=0.3cm]
+			әлем \& чемпионаты \\
+			\gmk{n.nom} \& \gmk{n.px3sg.nom} \\
+		\end{deptext}
+		\depedge{1}{2}{\udlabel{nmod}}
+	\end{dependency}
+	\caption{A nominal depending on an indefinite genitive.}\label{fig:indefgen}
+\end{figure}
+
+In these examples, the indefinite genitive is considered \udtag{nmod}.
+
 \subsection{Non-finite clauses}
@@
 \section{Evaluation}
+
+\cite{ballesteros15}, \cite{nivre07}
 
@@
 
+%Tolgonay?, Aida?
+
 \bibliographystyle{apalike}

### 2015-06-20 2015-turklang-kazdep/paper.tex
font size change
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60868 72bbbca6-d526-0410-a7d9-f06f51895060
add some eval stuffs
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60869 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \setmainfont{Times New Roman}
-\newfontfamily\udfont[Scale=MatchLowercase,Letters=SmallCaps]{Linux Libertine O}
+\newfontfamily\udfont[Scale=1.3,Letters=SmallCaps]{Linux Libertine O}
 
@@
 
-\cite{ballesteros15}, \cite{nivre07}
+In order to test the utility of the treebank in a real-world setting, we trained
+and evaluated a number of models using the popular MaltParser tool \citep{nivre07}.
+MaltParser is a toolkit for data-driven dependency parsing, it can learn a parsing
+model from treebank data and apply this model to parse unseen sentences. The parser
+has a large number of options and parameters that need to be optimised.
+To select the best parser configuration we relied on
+MaltOptimiser \citep{ballesteros15}. The optimiser recommended the \emph{covnonproj}
+parsing algorithm.
+
+As the treebank takes advantage of the new tokenisation standards in the CoNLL-U format,
+and MaltParser only supports CoNLL-X, certain transformations were needed to perform
+the experiments. The corpus was flattened with conjoined tokens receiving a dummy
+surface form. The converted corpus is available alongside the original.\footnote{\url{removed for review}}
+
+% LAS: 0.5529 [0.434 0.446 0.501 0.511 0.515 0.542 0.558 0.572 0.7 0.75]
+% UAS: 0.7376 [0.658 0.669 0.675 0.684 0.694 0.723 0.751 0.776 0.825 0.921]
+
+\begin{table}
+  \centering
+  \begin{tabular}{|l|r|r|}
+    \hline
+    \textbf{Features} & \textbf{LAS} & \textbf{UAS} \\
+    \hline
+     - & - & - \\
+    \hline
+  \end{tabular}
+  \caption{Preliminary parsing results from MaltParser using different models}
+  \label{table:eval}
+\end{table}
+
+To perform 10-fold cross validation we randomised the order of sentences in the corpus
+and split it into 10 parts. In each iteration we held out one part for testing and used
+the rest for training. We calculated the labelled-attachment score (LAS) and
+unlabelled-attachment score (UAS) for each of the models.
+
+The results we obtain for unlabelled attachment are similar to those obtained with similar 
+sized treebanks, for example the Irish treebank of \cite{Lynn12}, who report an LAS of
+63.3 and a UAS of 73.3 with the best model.
 

### 2015-06-20 2015-turklang-kazdep/paper.tex
examples of complex nominals
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60873 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
+\usepackage{subcaption}
+
 \usepackage{natbib}
@@
 \newcommand{\gmk}[1]{\texttt{#1}}
+\newcommand{\gloss}[1]{#1}
 \newcommand{\tag}[1]{\texttt{‹#1›}}
@@
 	\centering
+
+	\begin{subfigure}[b]{0.3\textwidth}
+		\begin{dependency}[theme = simple, font = \small]
+		   \begin{deptext}[column sep=0.3cm]
+				көрші \& елдер \\
+				\gmk{n.attr} \& \gmk{n.pl.nom} \\
+				\gloss{neighbour} \& \gloss{countries} \\
+			\end{deptext}
+			\depedge{1}{2}{\udlabel{compound}}
+		\end{dependency}
+		\caption{Compounding: a nominal depending on an attributive nominal.}\label{fig:compound}
+	\end{subfigure}
+	~
+	\begin{subfigure}[b]{0.3\textwidth}
 		\begin{dependency}[theme = simple, font = \small]
@@
 				\gmk{n.nom} \& \gmk{n.px3sg.nom} \\
+				\gloss{world} \& \gloss{championship} \\
 			\end{deptext}
@@
 		\end{dependency}
-	\caption{A nominal depending on an indefinite genitive.}\label{fig:indefgen}
+		\caption{Indefinite genitive: A nominal depending on an indefinite genitive.}\label{fig:indefgen}
+	\end{subfigure}
+	~
+	\begin{subfigure}[b]{0.3\textwidth}
+		\begin{dependency}[theme = simple, font = \small]
+		   \begin{deptext}[column sep=0.3cm]
+				Иранның \& экономиясы \\
+				\gmk{n.gen} \& \gmk{n.px3sg.nom} \\
+				\gloss{Iran's} \& \gloss{economy} \\
+			\end{deptext}
+			\depedge{1}{2}{\udlabel{nmod}}
+		\end{dependency}
+		\caption{Definite genitive: A nominal depending on an definite genitive.}\label{fig:indefgen}
+	\end{subfigure}
+
+	\caption{Dependency trees of complex nominal relations.}
 \end{figure}

### 2015-06-20 2015-turklang-kazdep/paper.tex
centering, wording
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60874 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 	\begin{subfigure}[b]{0.3\textwidth}
+		\centering
 		\begin{dependency}[theme = simple, font = \small]
@@
 		\end{dependency}
-		\caption{Compounding: a nominal depending on an attributive nominal.}\label{fig:compound}
+		\caption{Compounding: an attributive nominal depending on a nominal.}\label{fig:compound}
 	\end{subfigure}
-	~
+	\quad
 	\begin{subfigure}[b]{0.3\textwidth}
+		\centering
 		\begin{dependency}[theme = simple, font = \small]
@@
 		\end{dependency}
-		\caption{Indefinite genitive: A nominal depending on an indefinite genitive.}\label{fig:indefgen}
+		\caption{Indefinite genitive: An indefinite genitive depending on a nominal.}\label{fig:indefgen}
 	\end{subfigure}
-	~
+	\quad
 	\begin{subfigure}[b]{0.3\textwidth}
+		\centering
 		\begin{dependency}[theme = simple, font = \small]
@@
 		\end{dependency}
-		\caption{Definite genitive: A nominal depending on an definite genitive.}\label{fig:indefgen}
+		\caption{Definite genitive: A definite genitive depending on a nominal.}\label{fig:defgen}
 	\end{subfigure}

### 2015-06-20 2015-turklang-kazdep/paper.tex
arrows fixed
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60875 72bbbca6-d526-0410-a7d9-f06f51895060
some prose fixing in complex nominals section
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60876 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-Some nominals occur together to act as a single nominal.
+There are different ways in which two nominals may occur together to act as a single nominal.
+Compounds are formed by an attributive nominal (indistinguishable from the bare / nominative form, but tagged with \tag{attr}) preceding another nominal, as shown in \ref{fig:compound}.
+An indefinite genitive construction is formed by an indefinite genitive nominal (indistinguishable from the bare / nominative form, and tagged with \tag{nom}) preceding a nominal that has third-person possessive morphology, as shown in \ref{fig:indefgen}.
+A definite genetive construction is formed by a genitive-marked nominal (tagged \tag{gen}) precding a nominal that has third-person possessive morphology, as shown in \ref{fig:defgen}.
 
-One example of this is indefinite genitives.
+%One example of this is indefinite genitives.
 In such constructions, a ``modifying'' nominal is morphologically not marked (which means it will be tagged as \tag{nom}, as discussed in §\ref{sec:kazakh}), though should be considered an indefinite genitive.
@@
 			\end{deptext}
-			\depedge{1}{2}{\udlabel{compound}}
+			\depedge{2}{1}{\udlabel{compound}}
 		\end{dependency}
@@
 			\end{deptext}
-			\depedge{1}{2}{\udlabel{nmod}}
+			\depedge{2}{1}{\udlabel{nmod}}
 		\end{dependency}
@@
 
-In these examples, the indefinite genitive is considered \udtag{nmod}.
+As seen in the graphs, the compound relationship of an attributive nominal depending on another nominal is labelled \udlabel{compound}, and genitive relations are considered \udtag{nmod}, regardless of whether there is a definite or indefinite genitive construction.
 

### 2015-06-20 2015-turklang-kazdep/paper.tex
how's this
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60877 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \newcommand{\tag}[1]{\texttt{‹#1›}}
-\newcommand{\udtag}[1]{\texttt{@#1}}
+%\newcommand{\udtag}[1]{\texttt{@#1}}
+\newcommand{\udtag}[1]{\textsc{\texttt{#1}}}
 \newcommand{\udlabel}[1]{{\udfont #1}}
@@
 
-As seen in the graphs, the compound relationship of an attributive nominal depending on another nominal is labelled \udlabel{compound}, and genitive relations are considered \udtag{nmod}, regardless of whether there is a definite or indefinite genitive construction.
+As seen in the graphs, the compound relationship of an attributive nominal depending on another nominal is labelled \udtag{compound}, and genitive relations are considered \udtag{nmod}, regardless of whether there is a definite or indefinite genitive construction.
 

### 2015-06-20 2015-turklang-kazdep/paper.tex
better
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60878 72bbbca6-d526-0410-a7d9-f06f51895060
citation hargling
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60879 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \setmainfont{Times New Roman}
+\newfontfamily\ll[]{Linux Libertine O}
 \newfontfamily\udfont[Scale=1.3,Letters=SmallCaps]{Linux Libertine O}
@@
 
-\usepackage{natbib}
-\bibdata{paper}
+%\usepackage{natbib}
+%\bibdata{paper}
+%\usepackage[backend=bibtex]{biblatex}
+\usepackage[backend=biber,citestyle=apa,style=apa]{biblatex}
+\bibliography{paper}
 
@@
 %\newcommand{\udtag}[1]{\texttt{@#1}}
-\newcommand{\udtag}[1]{\textsc{\texttt{#1}}}
+\newcommand{\udtag}[1]{{\ll \textsc{#1}}}
 \newcommand{\udlabel}[1]{{\udfont #1}}
@@
 
-\citet{Lynn12}, \citet{Atalay03}, \citet{Oflazer03}, \citet{DeMarneffe14}
+\textcite{Lynn12}, \textcite{Atalay03}, \textcite{Oflazer03}, \textcite{DeMarneffe14}
 
@@
 
-Kazakh (қазақ тілі), a Turkic language of Central Asia with around 12 million speakers, hargle bargle. %FIXME
+Kazakh (қазақ тілі), a Turkic language of Central Asia with around 13 million speakers \parencite{ethnologue}, hargle bargle. %FIXME
 
-In the process of designing a Kazakh transducer \citep{Washington14}, several decisions were made about how various inflectional morphemes affect the part of speech and syntactic function of various word forms.
+While works like \textcite{Balaqayev54} provide decent syntactic descriptions, there is little to no work on the syntax of Kazakh within modern theoretical syntactic frameworks (though work on related languages, such as Turkish, exists).
+
+In the process of designing a Kazakh transducer \parencite{Washington14}, several decisions were made about how various inflectional morphemes affect the part of speech and syntactic function of various word forms.
 These decisions include the relations to other words defined by ``case'' morphemes, and processes of ``zero derivation''.
@@
 Preprocessing the corpus consists of running the text through the Kazakh morphological
-analyser \citep{Washington14}, which also performs tokenisation of multiword units based 
+analyser \parencite{Washington14}, which also performs tokenisation of multiword units based 
 the longest match left-to-right. Tokenisation for Kazakh is a non-trivial task, and so
@@
 In order to test the utility of the treebank in a real-world setting, we trained
-and evaluated a number of models using the popular MaltParser tool \citep{nivre07}.
+and evaluated a number of models using the popular MaltParser tool \parencite{nivre07}.
 MaltParser is a toolkit for data-driven dependency parsing, it can learn a parsing
@@
 To select the best parser configuration we relied on
-MaltOptimiser \citep{ballesteros15}. The optimiser recommended the \emph{covnonproj}
+MaltOptimiser \parencite{ballesteros15}. The optimiser recommended the \emph{covnonproj}
 parsing algorithm.
@@
 
-\bibliographystyle{apalike}
-\bibliography{paper}
+%\bibliographystyle{apalike}
+%\bibliography{paper}
+\printbibliography
 

### 2015-06-20 2015-turklang-kazdep/paper.tex
bibfixes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60880 72bbbca6-d526-0410-a7d9-f06f51895060
more bibfixes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60881 72bbbca6-d526-0410-a7d9-f06f51895060
more bibfixes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60883 72bbbca6-d526-0410-a7d9-f06f51895060
bibfixes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60884 72bbbca6-d526-0410-a7d9-f06f51895060
more
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60885 72bbbca6-d526-0410-a7d9-f06f518

@@
 %\usepackage[backend=bibtex]{biblatex}
-\usepackage[backend=biber,citestyle=apa,style=apa]{biblatex}
+\usepackage[backend=biber,citestyle=apa,style=authoryear,maxbibnames=99,maxcitenames=1]{biblatex}
 \bibliography{paper}
+\renewcommand*{\nameyeardelim}{\addcomma\space}
 
@@
 
+\begin{figure}[htbp]
+	\centering
+
+	\begin{subfigure}[b]{0.3\textwidth}
+		\centering
+		\begin{dependency}[theme = simple, font = \small]
+		   \begin{deptext}[column sep=0.3cm]
+				Айгүл \& студент \& · \\
+				\gmk{np} \& \gmk{n} \& \gmk{cop} \\
+				\gloss{Ayğul} \& \gloss{student} \& \gloss{is} \\
+			\end{deptext}
+			\depedge{2}{1}{\udlabel{subj}}
+			\depedge{3}{2}{\udlabel{cop}}
+		\end{dependency}
+		\caption{xxx }\label{fig:compound}
+	\end{subfigure}
+	\quad
+	\begin{subfigure}[b]{0.3\textwidth}
+		\centering
+		\begin{dependency}[theme = simple, font = \small]
+		   \begin{deptext}[column sep=0.3cm]
+				Айгүл \& студент \& - \\
+				\gmk{np} \& \gmk{-} \& \gmk{-} \\
+				\gloss{Ayğul} \& \gloss{-} \& \gloss{-}\\
+			\end{deptext}
+			\depedge{3}{2}{\udlabel{cop}}
+		\end{dependency}
+		\caption{xxx.}\label{fig:cop2}
+	\end{subfigure}
+	\quad
+	\begin{subfigure}[b]{0.3\textwidth}
+		\centering
+		\begin{dependency}[theme = simple, font = \small]
+		   \begin{deptext}[column sep=0.3cm]
+				Айгүл   \& - \& - \\
+				\gmk{n} \& \gmk{-} \& \gmk{-} \& \\
+				\gloss{Ayğul} \& \gloss{-} \& \gloss{-} \\
+			\end{deptext}
+			\depedge{3}{2}{\udlabel{cop}}
+		\end{dependency}
+		\caption{xxx.}\label{fig:cop3}
+	\end{subfigure}
+
+	\caption{Dependency trees of copula constructions.}
+\end{figure}
+
 \subsection{Coordination}
+
+\begin{figure}[htbp]
+    \centering
+
+	\begin{dependency}[theme = simple, font = \small]
+	   \begin{deptext}[column sep=0.08cm]
+%               1         2         3          4              5                6              7                8           9             10
+             Барлық \& адамдар \& тумысынан \& азат      \& және         \& қадір-қасиеті \& мен          \& құқықтары  \& тең       \& болып     \& дүниеге \& келеді \\
+                    \&         \&           \& \gmk{adj} \& \gmk{cnjcoo} \& \gmk{n}       \& \gmk{cnjcoo} \& \gmk{n.pl} \& \gmk{adj} \& \gmk{cop} \& \& \\
+                    \&         \&           \& \gloss{free} \& \gloss{and}  \& \gloss{dignity}  \&  \gloss{and} \& \gloss{rights} \&  \gloss{equal} \& \gloss{being} \& \& \\
+		\end{deptext}
+		\depedge{9}{4}{\udlabel{conj}}
+		\depedge{9}{5}{\udlabel{cc}}
+		\depedge{8}{6}{\udlabel{conj}}
+		\depedge{8}{7}{\udlabel{cc}}
+		\depedge{9}{8}{\udlabel{subj}}
+		\depedge{10}{9}{\udlabel{cop}}
+		\depedge{12}{10}{\udlabel{advcl}}
+	\end{dependency}
+	\caption{Coordination}\label{fig:coord}
+
+\end{figure}
 
@@
 To select the best parser configuration we relied on 
-MaltOptimiser \parencite{ballesteros15}. The optimiser recommended the \emph{covnonproj}
-parsing algorithm.
+MaltOptimiser \parencite{ballesteros15}. The optimiser was run separately for each of 
+the model configurations.
 
@@
 
+% system1:
+% nivreeager
+% LAS: 0.3341 [0.234 0.288 0.289 0.292 0.333 0.335 0.359 0.378 0.408 0.425]
+% UAS: 0.5093 [0.413 0.464 0.465 0.466 0.474 0.499 0.531 0.545 0.604 0.632]
+
+% system2:
+% nivreeager
+% LAS: 0.234+0.288+0.289+0.292+0.333+0.335+0.359+0.378+0.408+0.425
+% UAS: 0.413+0.464+0.465+0.466+0.474+0.499+0.531+0.545+0.604+0.632
+
+% system3:
+% nivreeager
 % LAS: 0.5529 [0.434 0.446 0.501 0.511 0.515 0.542 0.558 0.572 0.7 0.75]
@@
 
+% system4:
+% covnonproj
+% LAS: 0.6092 [0.519 0.535 0.544 0.569 0.569 0.62 0.621 0.649 0.683 0.783]
+% UAS: 0.7575 [0.674 0.686 0.69 0.714 0.727 0.746 0.795 0.803 0.812 0.928]
+
+%             [0.434+0.446+0.501+0.511+0.515+0.542+0.558+0.572+0.7+0.75]
+%             [0.658+0.669+0.675+0.684+0.694+0.723+0.751+0.776+0.825+0.921]
+
 \begin{table}
 \centering
-  \begin{tabular}{|l|r|r|}
+  \begin{tabular}{|l|l|r|r|}
     \hline
-    \textbf{Features} & \textbf{LAS} & \textbf{UAS} \\
+    \textbf{Features}       & \textbf{Algorithm} &\textbf{LAS} & \textbf{UAS} \\
     \hline
-     - & - & - \\
+     surface                & nivreeager  & 33.4 [23.4, 42.5] & 50.9 [41.3, 63.2] \\
+     surface+lemma          & nivreeager  & 33.4 [23.4, 42.5] & 50.9 [41.3, 63.2] \\
+     surface+lemma+POS      & nivreeager  & 55.2 [43.4, 75.0] & 73.7 [65.8, 92.1] \\
+     surface+lemma+POS+MSD  & nivreeager  & 55.2 [43.4, 75.0] & 73.7 [65.8, 92.1] \\
     \hline
   \end{tabular}
-  \caption{Preliminary parsing results from MaltParser using different models}
+  \caption{Preliminary parsing results from MaltParser using different models. The numbers in 
+    brackets denote the upper and lower bounds found during cross validation. Adding 
+    structural features to the model substantially improves the performance of the parser, although
+    we find no improvement in using full morphosyntactic description (MSD) over using
+    simply the first part-of-speech (POS) tag.}
   \label{table:eval}
@@
 To perform 10-fold cross validation we randomised the order of sentences in the corpus
-and split it into 10 parts. In each iteration we held out one part for testing and used
+and split it into 10 equally-sized parts. In each iteration we held out one part for testing and used
 the rest for training. We calculated the labelled-attachment score (LAS) and 
@@
 
-The results we obtain for unlabelled attachment are similar to those obtained with similar 
-sized treebanks, for example the Irish treebank of \cite{Lynn12}, who report an LAS of
+The results we obtain are similar to those obtained with similar 
+sized treebanks, for example the Irish treebank of \textcite{Lynn12}, who report an LAS of
 63.3 and a UAS of 73.3 with the best model.

### 2015-06-20 2015-turklang-kazdep/paper.tex
Kazakh section
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60889 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \newcommand{\gmk}[1]{\texttt{#1}}
-\newcommand{\gloss}[1]{#1}
+\newcommand{\gloss}[1]{{\em #1}}
 \newcommand{\tag}[1]{\texttt{‹#1›}}
@@
 
-Kazakh (қазақ тілі), a Turkic language of Central Asia with around 13 million speakers \parencite{ethnologue}, hargle bargle. %FIXME
+Kazakh (қазақ тілі), a Turkic language of Central Asia, is spoken by around 13 million people in Kazakhstan, China, Mongolia, and adjacent areas \parencite{ethnologue}.
+While works like \textcite{Balaqayev54} provide decent syntactic descriptions of the language, there is little to no work on the syntax of Kazakh within modern theoretical syntactic frameworks (though work on related languages, such as Turkish, exists).
 
-While works like \textcite{Balaqayev54} provide decent syntactic descriptions, there is little to no work on the syntax of Kazakh within modern theoretical syntactic frameworks (though work on related languages, such as Turkish, exists).
-
-In the process of designing a Kazakh transducer \parencite{Washington14}, several decisions were made about how various inflectional morphemes affect the part of speech and syntactic function of various word forms.
-These decisions include the relations to other words defined by ``case'' morphemes, and processes of ``zero derivation''.
+As an agglutinative language with rich morphology and agreement phenomena, Kazakh presents some interesting challenges for computational syntax.
+A morphological transducer has already been designed \parencite{Washington14}, which implements a number of decisions about how various inflectional morphemes affect the part of speech and syntactic function of various word forms.
+These decisions include how the relations ``case'' morphemes create between syntactic elements work, as well as a number of processes of ``zero derivation''.
 
 All of the traditionally defined ``cases'' in Kazakh have a variety of uses.
-The morphemes that mark the accusative and genitive cases are only used when the noun phrase is semantically definite.
-This means that indefinite accusative and genitive as well as nominative noun phrases are all unmarked for case, and hence are all ambiguous.
-Our Kazakh transducer marks these all as \tag{nom}, but their use may be disambiguated.
-Besides identifying the semantic role of an unmarked noun phrase, there are various agreement properties that make it clear what ``case'' the noun phrase ``truly'' receives: genitive noun phrases must have a corresponding noun phrase with possessive morphology that agrees in person with the genitive noun phrase; and nominative noun phrases must have a corresponding predicate that agrees in person with the nominative noun phrase, e.g., a verb or a copula.
+The morphemes that mark the accusative and genitive cases are only used when the noun phrase is semantically definite (e.g., үйді \gloss{the house} \gmk{acc}, үйдің \gloss{of the house} \gmk{gen}).
+This means that indefinite accusative and genitive as well as nominative noun phrases are all unmarked for case, and hence are all ambiguous (e.g., үй \gloss{house}).
+The Kazakh transducer marks these all as \tag{nom}, but their use may be disambiguated.
+Besides identifying the semantic role of an unmarked noun phrase, there are various agreement properties that make it clear what ``case'' the noun phrase ``truly'' receives: genitive noun phrases must have a corresponding noun phrase with possessive morphology that agrees in person with the genitive noun phrase; and truly nominative noun phrases must have a corresponding predicate that agrees in person with the nominative noun phrase---e.g., a verb or a copula.
 
-Copulas may also be non-overt.  In the present tense, all but third person copulas have realised forms, but the third person copula does not surface.
+Another category that may also be non-overt is the copula.
+In the present tense, all but third person copulas have realised forms (e.g., Мен үйде\textbf{мін}.
+\gloss{I'm at home.}), but the third person copula does not surface (e.g., Ол үйде.
+\gloss{S/he's at home.}).
 
-Adjectives in Kazakh are by default attributive, but many may also be used either nominally (i.e., as nouns) or adverbially (i.e., as adverbs), or both.
-Our morphological transducer internally provides a series of classes; depending on class membership, an adjective may receive readings \tag{adj}\tag{advl} or \tag{adj}\tag{subst}, besides the default attributive reading of \tag{adj}.
+Another type of ``zero-derivation'' involves the use of adjectives and nouns.
+In Kazakh, adjectives are all attributive, but many may also be used either nominally (i.e., as nouns) or adverbially (i.e., as adverbs), or both.
+The morphological transducer internally provides a series of classes; depending on class membership, an adjective may receive the readings \tag{adj}\tag{advl} or \tag{adj}\tag{subst}, besides the default attributive reading of \tag{adj}.
 If an adjective may be used substantivally, then it will receive a range of substantive morphology as well.
+Similarly, many nouns may be used attributively, besides the default substantive reading.
+Besides the nominative (and indefinite accusative/genitive) reading of \tag{nom}, the transducer provides an attributive reading for most nominals of \tag{nom}\tag{attr}.
+
+How each of these issues bears on a dependency analysis will be discussed in section~\ref{sec:annotation}.
 
@@
 
-\section{Annotation guidelines}
+\section{Annotation guidelines}\label{sec:annotation}
 
@@
 
-\begin{table}
+\begin{table}[htbp]
 \centering

### 2015-06-20 2015-turklang-kazdep/paper.tex
some minor fixes and notes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60890 72bbbca6-d526-0410-a7d9-f06f51895060
some cleanup of tables
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60891 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{subcaption}
+\usepackage{booktabs}
+\setlength{\cmidrulewidth}{\heavyrulewidth}
 
@@
 \begin{small}
-  \begin{tabular}{|l|l|r|r|r|}
-    \hline
+  \begin{tabular}{llrrr}
+    \toprule
     \textbf{Document} & \textbf{Description} & \textbf{Sentences} & \textbf{Tokens} & \textbf{Avg. length}\\
-    \hline
+    \midrule
     UN Declaration on Human Rights & Legal text on human rights & 25 & 409 & 16.3 \\
@@
     Радиан & Wikipedia article (Radian)                            & 2 & 17 & 8.5 \\
-    \hline
-    \multicolumn{2}{c|}{~} & 302 & 3741 & 12.3 \\\cline{3-5}
+    \midrule
+    \multicolumn{2}{c}{~} & 302 & 3741 & 12.3 \\\cmidrule{3-5}
   \end{tabular}
@@
 	\centering
-  \begin{tabular}{|l|l|}
-    \hline
+	\begin{tabular}{ll}
+		\toprule
 		\textbf{Label} & \textbf{Description} \\
-
-    \hline
+		\midrule
+				\\
+ 		\bottomrule
 	\end{tabular}
@@
 				\gmk{np} \& \gmk{n} \& \gmk{cop} \\
-				\gloss{Ayğul} \& \gloss{student} \& \gloss{is} \\
+				\gloss{Aygül} \& \gloss{student} \& \gloss{is} \\
 			\end{deptext}
@@
 				\gmk{np} \& \gmk{-} \& \gmk{-} \\
-				\gloss{Ayğul} \& \gloss{-} \& \gloss{-}\\
+				\gloss{Aygül} \& \gloss{-} \& \gloss{-}\\
 			\end{deptext}
@@
 				\gmk{n} \& \gmk{-} \& \gmk{-} \& \\
-				\gloss{Ayğul} \& \gloss{-} \& \gloss{-} \\
+				\gloss{Aygül} \& \gloss{-} \& \gloss{-} \\
 			\end{deptext}
@@
 
+% verbal adjectives → acl
+% verbal nouns (gerunds) → ccomp, no xcomp
+% verbal adverbs → advlc
+% * gna_cnd? → advlc?
+
 \section{Evaluation}
@@
 	\centering
-  \begin{tabular}{|l|l|r|r|}
-    \hline
-    \textbf{Features}       & \textbf{Algorithm} &\textbf{LAS} & \textbf{UAS} \\
-    \hline
+	\begin{tabular}{llrr}
+		\toprule
+		\textbf{Features}       & \textbf{Algorithm} &\textbf{LAS} [range] & \textbf{UAS} [range] \\
+		\midrule
 		surface                & nivreeager  & 33.4 [23.4, 42.5] & 50.9 [41.3, 63.2] \\
@@
 		surface+lemma+POS+MSD  & nivreeager  & 55.2 [43.4, 75.0] & 73.7 [65.8, 92.1] \\
-    \hline
+		\bottomrule
 	\end{tabular}
-  \caption{Preliminary parsing results from MaltParser using different models. The numbers in 
-    brackets denote the upper and lower bounds found during cross validation. Adding 
-    structural features to the model substantially improves the performance of the parser, although
-    we find no improvement in using full morphosyntactic description (MSD) over using
-    simply the first part-of-speech (POS) tag.}
+	\caption{Preliminary parsing results from MaltParser using different models.
+The numbers in brackets denote the upper and lower bounds found during cross validation.
+Adding structural features to the model substantially improves the performance of the parser, although we find no improvement in using full morphosyntactic description (MSD) over using simply the first part-of-speech (POS) tag.}
 	\label{table:eval}

### 2015-06-20 2015-turklang-kazdep/paper.tex
more table cleaning
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60892 72bbbca6-d526-0410-a7d9-f06f51895060
words
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60893 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{booktabs}
-\setlength{\cmidrulewidth}{\heavyrulewidth}
 
@@
 
-\textcite{Lynn12}, \textcite{Atalay03}, \textcite{Oflazer03}, \textcite{DeMarneffe14}
+This article describes the work towards the development of a dependency treebank for Kazakh, a
+Turkic language spoken in Central Asia.
+Despite its status as a \emph{core} Turkic
+language, little computational-linguistic research has been published on syntactic
+parsing for Kazakh.
+A valuable resource in the study of syntactic parsing is
+a treebank---a corpus of parsed text containing gold-standard syntactic annotation.
+
+Freely available treebanks exist for many languages, such as large languages like Finnish (x,y) 
+and Polish (xxx) and smaller languages like Irish \parencite{Lynn12}. To our knowledge only one treebank
+exists for another language, Turkish \parencite{Oflazer03}, which is unfortunately not freely available.
+
+In building our treebank we take advantage of existing work done on tokenisation,
+morphological analysis and part-of-speech tagging for Kazakh. We also take a pragmatic
+and iterative view of development of the treebank, in line with recent work 
+on cross-linguistic parsing with universal dependencies \parencite{DeMarneffe14}.
+
+The remainder of the paper is organised as follows. Section~\ref{sec:back} gives some
+background linguistic information on Kazakh, and outlines some special challenges in
+parsing Kazakh. In Section~\ref{sec:method} we describe the corpus itself and methodology
+used in annotating it. Section~\ref{sec:annotation} gives a sketch of some decisions
+we have made with respect to annotation guidelines, referring back to the discussion in
+Section~\ref{sec:back}. For reasons of space, these guidelines are not complete, but
+present a subset of guidelines which are of particular interest. A small experiment
+in statistical dependency parsing using the corpus is presented in Section~\ref{sec:eval},
+and in Sections~\ref{sec:future} and \ref{sec:conclusions} we give perspectives
+for future work and some concluding remarks.
+
+%\textcite{Atalay03}
 
@@
     \midrule
-    \multicolumn{2}{c}{~} & 302 & 3741 & 12.3 \\\cmidrule{3-5}
+	    \multicolumn{2}{c}{~} & 302 & 3741 & 12.3 \\\cmidrule[\heavyrulewidth]{3-5}
   \end{tabular}

### 2015-06-20 2015-turklang-kazdep/paper.tex
hargle
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60895 72bbbca6-d526-0410-a7d9-f06f51895060
refs
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60896 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-This article describes the work towards the development of a dependency treebank for Kazakh, a
-Turkic language spoken in Central Asia.
+This article describes work towards the development of a dependency treebank for Kazakh, a Turkic language spoken in Central Asia and Europe.
 Despite its status as a \emph{core} Turkic
@@
 
-Freely available treebanks exist for many languages, such as large languages like Finnish (x,y) 
+Freely available treebanks exist for many languages, such as large languages like Finnish \parencite{haverinen2013tdt,voutilainen11} 
 and Polish (xxx) and smaller languages like Irish \parencite{Lynn12}. To our knowledge only one treebank

### 2015-06-20 2015-turklang-kazdep/paper.tex
some fixes to introduction
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60897 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 This article describes work towards the development of a dependency treebank for Kazakh, a Turkic language spoken in Central Asia and Europe.
-Despite its status as a \emph{core} Turkic
+Despite its status as a \emph{core}\footnote{FIXME} Turkic %FIXME
 language, little computational-linguistic research has been published on syntactic
@@
 Freely available treebanks exist for many languages, such as large languages like Finnish \parencite{haverinen2013tdt,voutilainen11} 
-and Polish (xxx) and smaller languages like Irish \parencite{Lynn12}. To our knowledge only one treebank
-exists for another language, Turkish \parencite{Oflazer03}, which is unfortunately not freely available.
+and Polish (xxx) and smaller languages like Irish \parencite{Lynn12}.
+To our knowledge, however, a treebank exists for only one other Turkic language, Turkish \parencite{Oflazer03}, which is unfortunately not freely available.
 
@@
 background linguistic information on Kazakh, and outlines some special challenges in
-parsing Kazakh. In Section~\ref{sec:method} we describe the corpus itself and methodology
+parsing Kazakh. In Section~\ref{sec:method} we describe the corpus that we annotated and the methodology
 used in annotating it. Section~\ref{sec:annotation} gives a sketch of some decisions
@@
 
-\section{Background}
+\section{Background}\label{sec:back}
 
@@
 
-\section{Methodology}
+\section{Methodology}\label{sec:method}
 
@@
 
-\section{Evaluation}
+\section{Evaluation}\label{sec:eval}
 
@@
 
-\section{Future work}
+\section{Future work}\label{sec:future}
 

### 2015-06-20 2015-turklang-kazdep/paper.tex
fixes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60900 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 Freely available treebanks exist for many languages, such as large languages like Finnish \parencite{haverinen2013tdt,voutilainen11} 
-and Polish (xxx) and smaller languages like Irish \parencite{Lynn12}.
+and Polish \parencite{wolinksi11} and smaller languages like Irish \parencite{Lynn12}.
 To our knowledge, however, a treebank exists for only one other Turkic language, Turkish \parencite{Oflazer03}, which is unfortunately not freely available.

### 2015-06-20 2015-turklang-kazdep/paper.tex
bibfixes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60901 72bbbca6-d526-0410-a7d9-f06f51895060

--- 2015-06-20 2015-turklang-kazdep/paper.tex
@@
 Freely available treebanks exist for many languages, such as large languages like Finnish \parencite{haverinen2013tdt,voutilainen11} 
-and Polish (xxx) and smaller languages like Irish \parencite{Lynn12}.
+and Polish \parencite{wolinski11} and smaller languages like Irish \parencite{Lynn12}.
 To our knowledge, however, a treebank exists for only one other Turkic language, Turkish \parencite{Oflazer03}, which is unfortunately not freely available.

### 2015-06-20 2015-turklang-kazdep/paper.tex
more stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60902 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-% what are treebanks used for ?
+A treebank is a parsed corpus of sentences annotated syntactically following a particular
+syntactic theory. Two broad groups can be distinguished, phrase-structure treebanks
+which annotate constituency strucure, and dependency treebanks which annotate dependency
+structure. Some treebanks combine both.
+
+Treebanks can be used directly for linguistic and computational linguistic research, by
+performing search queries. For example, to extract a valency lexicon for verbs, or to
+study the frequency of various syntactic phenomena such as word order or nominal case usage
+and syntactic function.
+
+They can also be used to train statistical parsers which can
+be used to annotate previously unseen texts. These parsers can be used in end-user applications
+such as machine translation and computer-aided language learning.
+
+According to \textcite{nivre08}, a parser trained on a treebank of only 1,500 sentences 
+can provide reasonable parsing accuracy.
 

### 2015-06-20 2015-turklang-kazdep/paper.tex
some stuff in intro
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60904 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 %\usepackage[backend=bibtex]{biblatex}
-\usepackage[backend=biber,citestyle=apa,style=authoryear,maxbibnames=99,maxcitenames=1]{biblatex}
+\usepackage[backend=biber,citestyle=apa,style=authoryear,maxbibnames=99,maxcitenames=2]{biblatex}
 \bibliography{paper}
@@
 Kazakh (қазақ тілі), a Turkic language of Central Asia, is spoken by around 13 million people in Kazakhstan, China, Mongolia, and adjacent areas \parencite{ethnologue}.
-While works like \textcite{Balaqayev54} provide decent syntactic descriptions of the language, there is little to no work on the syntax of Kazakh within modern theoretical syntactic frameworks (though work on related languages, such as Turkish, exists).
+While works like \textcite{Balaqayev54} provide decent syntactic overviews of the language, there is little to no work on the syntax of Kazakh within modern theoretical syntactic frameworks.
+The authors are familiar with work on related languages, such as Turkish (e.g., \cite{Kornfilt97} and \cite{GokselKerslake05}); while not directly consulted in this work, such work has contributed to our understanding of Kazakh syntax.
 

### 2015-06-20 2015-turklang-kazdep/paper.tex
output
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60905 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
+\usepackage{alltt}
 \usepackage{polyglossia}
@@
 \subsection{Preprocessing}
+
+\begin{figure}
+\begin{tiny}
+\begin{alltt}
+"<Елде>"
+        "ел" n loc
+"<ағылшын>"
+        "ағылшын" n nom
+"<үстемдігіне>"
+        "үстемдік" n px3sp dat
+"<қарсы>"
+        "қарсы" post
+"<жаңа>"
+        "жаңа" adj
+"<толқулар>"
+        "толқу" n pl nom
+"<басталды>"
+        "баста" v tv pass ifi p3 pl
+"<.>"
+        "." sent
+\end{alltt}
+\end{tiny}
+\caption{Example output after preprocessing and manual disambiguation. Each token is annotated for lemma,
+   part of speech and morphosyntactic information.}
+\end{figure}
 

### 2015-06-20 2015-turklang-kazdep/paper.tex
minor fixes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60906 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 While works like \textcite{Balaqayev54} provide decent syntactic overviews of the language, there is little to no work on the syntax of Kazakh within modern theoretical syntactic frameworks.
-The authors are familiar with work on related languages, such as Turkish (e.g., \cite{Kornfilt97} and \cite{GokselKerslake05}); while not directly consulted in this work, such work has contributed to our understanding of Kazakh syntax.
+The authors are familiar with such work on related languages, such as Turkish (e.g., \cite{Kornfilt97} and \cite{GokselKerslake05}); while not directly consulted for this work, these works have contributed to our understanding of Kazakh syntax.
 

### 2015-06-20 2015-turklang-kazdep/paper.tex
hargle
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60908 72bbbca6-d526-0410-a7d9-f06f51895060

--- 2015-06-20 2015-turklang-kazdep/paper.tex
@@
 
-Kazakh (қазақ тілі), a Turkic language of Central Asia, is spoken by around 13 million people in Kazakhstan, China, Mongolia, and adjacent areas \parencite{ethnologue}.
+Kazakh (қазақ тілі), a Turkic language of Central Asia and Europe, is spoken by around 13 million people in Kazakhstan, China, Mongolia, and adjacent areas \parencite{ethnologue}.
 While works like \textcite{Balaqayev54} provide decent syntactic overviews of the language, there is little to no work on the syntax of Kazakh within modern theoretical syntactic frameworks.
-The authors are familiar with work on related languages, such as Turkish (e.g., \cite{Kornfilt97} and \cite{GokselKerslake05}); while not directly consulted in this work, such work has contributed to our understanding of Kazakh syntax.
+The authors are familiar with such work on related languages, especially Turkish (e.g., \cite{Kornfilt97} and \cite{GokselKerslake05}); while not directly consulted for this work, these works have contributed to our understanding of Kazakh syntax.
 
 As an agglutinative language with rich morphology and agreement phenomena, Kazakh presents some interesting challenges for computational syntax.
-A morphological transducer has already been designed \parencite{Washington14}, which implements a number of decisions about how various inflectional morphemes affect the part of speech and syntactic function of various word forms.
-These decisions include how the relations ``case'' morphemes create between syntactic elements work, as well as a number of processes of ``zero derivation''.
+These challenges include the syntactic functions of the various ``case'' morphemes, problems of ``zero derivation'', non-finite clauses,       These decisions include how the relations ``case'' morphemes create between syntactic elements work, as well as a number of processes of ``zero derivation''.A morphological transducer has already been designed \parencite{Washington14}, which implements a number of decisions about how various inflectional morphemes affect the part of speech and syntactic function of various word forms.
 

### 2015-06-20 2015-turklang-kazdep/paper.tex
updates
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60909 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-% simple tokens
-% multiword units
-% copula
-% clitics
+Tokenisation of the corpus is performed by our morphological analyser. This analyser
+performs tokenisation on the basis of a left-to-right longest match algorithm described
+in \textcite{garrido02}. Simple tokens such as \emph{толқулар} `riots' are maintained
+as a single token, and their lemma and morphological analysis is returned. Multiword
+units such as \emph{ауа райы} `weather' and \emph{ата-ананың} `of parents' are combined
+into a single token. Abbreviations and numerals which bear case, such as АҚШ-пен `with the USA'
+and \emph{90\%-ына} `to the 90\%' are analysed as a single token, as are light-verbs such as
+\emph{пайда бол} `to appear' and tense forms written with space like \emph{оқыған жоқ},
+the third-person negative past of \emph{оқы} `to learn'.
+
+In some cases a single token is split into two tokens, as with the aorist copula suffixes,
+\emph{үйдемін} `I am in the house' is tokenised as үй.{\sc loc} + е.{\sc cop.aor.sg1}. Furthermore
+two input tokens may result in three output tokens, \emph{бар ма?} `is there?' would be
+tokenised as бар.{\sc adj} + е.{\sc cop.aor.sg1} + ма.{\sc qst}.
+
 % productive suffixes -lı (cite wafflelazor)

### 2015-06-20 2015-turklang-kazdep/paper.tex
hargle
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60910 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 As an agglutinative language with rich morphology and agreement phenomena, Kazakh presents some interesting challenges for computational syntax.
-These challenges include the syntactic functions of the various ``case'' morphemes, problems of ``zero derivation'', non-finite clauses,       These decisions include how the relations ``case'' morphemes create between syntactic elements work, as well as a number of processes of ``zero derivation''.A morphological transducer has already been designed \parencite{Washington14}, which implements a number of decisions about how various inflectional morphemes affect the part of speech and syntactic function of various word forms.
+These challenges include the syntactic functions of the various ``case'' morphemes, problems of ``zero derivation'', non-finite clauses, and copulas and copula constructions.
+An existing morphological transducer of Kazakh \parencite{Washington14} implements analyses of how these various phenomena occur on the morphological level.
+These phenomena will be described in this section, and how they were dealt with in the annotation of the treebank will be described in section~\ref{sec:annotation}.
 

### 2015-06-20 2015-turklang-kazdep/paper.tex
section
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60911 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-\section{Annotation guidelines}\label{sec:annotation}
+\section{Annotation}\label{sec:annotation}
 
-% general procedure
+Annotation of the treebank followed an iterative approach. The first stage was to
+select a text and preprocess it using the morphological analyser and constraint grammar.
+The output of this process was then manually disambiguated by choosing the most appropriate
+analysis in context, or adding a new analysis if one was not returned by the morphological
+analyser. This latter step was mostly in the case of unknown proper nouns.\footnote{New lexemes
+found during the development of the corpus will contribute to the development of the
+morphological analyser.}
 
-% - disambiguate
-% - annotate
-% - check the tree structure
-% - fix errors
-% - check for label consistency
-% - fix errors
+After the text was disambiguated, it was annotated by a single annotator for dependency
+structure relying on the annotation guidelines for universal dependencies.\footnote{\url{http://universaldependencies.github.io/docs/u/overview/syntax.html}}
+
+Where it was not clear how to annotate a particular structure, it was left pending, that
+is marked with the {\sc dep} relation. After a text was annotated, graphical representations
+of the trees were produced and proofread by a second annotator. The final analysis for a
+sentence was decided on during discussions between the two annotators and in some cases
+through the universal-dependency issue tracker.\footnote{\url{https://github.com/UniversalDependencies/docs/issues/}}
+
+During the course of annotation, language-specific annotation guidelines for Kazakh are
+being produced and published online.\footnote{\url{http://wiki.apertium.org/wiki/Dependency_parsing_for_Turkic}} In 
+the remainder of this section we describe a number of Kazakh-specific decisions.
 

### 2015-06-21 2015-turklang-kazdep/paper.tex
hargle
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60914 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \newfontfamily\udfont[Scale=1.3,Letters=SmallCaps]{Linux Libertine O}
+\newfontfamily\fsfont[Scale=MatchLowercase]{FreeSerif}
 
@@
 \newcommand{\udlabel}[1]{{\udfont #1}}
+\newcommand{\tilda}{{\fsfont ∼}}
 %\newcommand{\corrauthor}[1]{\footnotemark\footnotetext{hargle}}
@@
 
-All of the traditionally defined ``cases'' in Kazakh have a variety of uses.
+In Kazakh, as in most languages with case, there is not a one-to-one relation between ``case'' morphemes and syntactic function (not to mention a wide range of semantic functions).
+The main syntactic functions of the traditionally defined cases in Kazakh are summarised in table~\ref{tab:cases}
+
+\begin{table}[htbp]
+	\centering
+	\caption{Primary syntactic functions of traditionally defined cases in Kazakh.}\label{tab:cases}
+	\begin{tabular}{l p{5em} p{8em} p{22em}}
+		\toprule
+			\textbf{Case} & \textbf{Morpheme} & \textbf{Functions} & \textbf{Examples} \\
+		\midrule
+			nominative & - & subject,\newline
+									attributive,\newline
+									indefinite object,\newline
+									indefinite genitival 
+								& hargle bargle \newline
+									jargal barble \newline 
+									\\
+			accusative & -/NI/ & object,\newline
+										embedded subject,\newline
+										causative subject
+									& foo \newline
+										bar \\
+			genitive & -/NIң/ & genitival,\newline
+										embedded subject \\
+			locative & -/DA/ & adverbial
+									& \\
+			ablative & -/DAн/ & adverbial,\newline
+										comparator \\
+			dative & -/GA/ & indirect object,\newline
+									attributive
+								& \\
+		\bottomrule
+	\end{tabular}
+\end{table}
+
+All of the traditionally defined ``cases'' in Kazakh have a variety of uses, .
 The morphemes that mark the accusative and genitive cases are only used when the noun phrase is semantically definite (e.g., үйді \gloss{the house} \gmk{acc}, үйдің \gloss{of the house} \gmk{gen}).
@@
   \centering
+  \caption{Composition of the corpus. The corpus covers a range of genres and text types from free and public-domain sources.}
 	\begin{small}
@@
 			\midrule
-	    \multicolumn{2}{c}{~} & 302 & 3741 & 12.3 \\\cmidrule[\heavyrulewidth]{3-5}
+			\multicolumn{2}{c}{\tilda{}} & 302 & 3741 & 12.3 \\\cmidrule[\heavyrulewidth]{3-5}
 		\end{tabular}
 	\end{small}
-  \caption{Composition of the corpus. The corpus covers a range of genres and text types from 
-    free and public-domain sources.}
 \end{table}

### 2015-06-21 2015-turklang-kazdep/paper.tex
small table
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60915 72bbbca6-d526-0410-a7d9-f06f51895060

--- 2015-06-20 2015-turklang-kazdep/paper.tex
@@
 
-All of the traditionally defined ``cases'' in Kazakh have a variety of uses.
+In Kazakh, as in most languages with case, there is not a one-to-one relation between ``case'' morphemes and syntactic function (not to mention a wide range of semantic functions).
+The main syntactic functions of the traditionally defined cases in Kazakh are summarised in table~\ref{tab:cases}.
+
+\begin{table}[htbp]
+	\centering
+	\caption{Primary syntactic functions of traditionally defined cases in Kazakh.}\label{tab:cases}
+	\begin{small}
+		\begin{tabular}{l p{5em} p{8em} p{22em}}
+			\toprule
+				\textbf{Case} & \textbf{Morpheme} & \textbf{Functions} & \textbf{Examples} \\
+			\midrule
+				nominative & - & subject,\newline
+										attributive,\newline
+										indefinite object,\newline
+										indefinite genitival 
+									& hargle bargle \newline
+										jargal barble \newline 
+										\\
+				accusative & -/NI/ & object,\newline
+											embedded subject,\newline
+											causative subject
+										& foo \newline
+											bar \\
+				genitive & -/NIң/ & genitival,\newline
+											embedded subject \\
+				locative & -/DA/ & adverbial
+										& \\
+				ablative & -/DAн/ & adverbial,\newline
+											comparator \\
+				dative & -/GA/ & indirect object,\newline
+										attributive
+									& \\
+			\bottomrule
+		\end{tabular}
+	\end{small}
+\end{table}
+
+All of the traditionally defined ``cases'' in Kazakh have a variety of uses, .
 The morphemes that mark the accusative and genitive cases are only used when the noun phrase is semantically definite (e.g., үйді \gloss{the house} \gmk{acc}, үйдің \gloss{of the house} \gmk{gen}).

### 2015-06-21 2015-turklang-kazdep/paper.tex
small refs
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60916 72bbbca6-d526-0410-a7d9-f06f51895060
some work on "nominative"
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60920 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 			\midrule
-				nominative & - & subject,\newline
-										attributive,\newline
-										indefinite object,\newline
+				nominative & - & subject\newline
+										attributive\newline
+										indefinite object\newline
 										indefinite genitival 
@@
 										\\
-				accusative & -/NI/ & object,\newline
-											embedded subject,\newline
+				accusative & -/NI/ & definite object\newline
+											embedded subject\newline
 											causative subject
@@
 											bar \\
-				genitive & -/NIң/ & genitival,\newline
+				genitive & -/NIң/ & definite genitival\newline
 											embedded subject \\
@@
 
-All of the traditionally defined ``cases'' in Kazakh have a variety of uses, .
-The morphemes that mark the accusative and genitive cases are only used when the noun phrase is semantically definite (e.g., үйді \gloss{the house} \gmk{acc}, үйдің \gloss{of the house} \gmk{gen}).
-This means that indefinite accusative and genitive as well as nominative noun phrases are all unmarked for case, and hence are all ambiguous (e.g., үй \gloss{house}).
-The Kazakh transducer marks these all as \tag{nom}, but their use may be disambiguated.
-Besides identifying the semantic role of an unmarked noun phrase, there are various agreement properties that make it clear what ``case'' the noun phrase ``truly'' receives: genitive noun phrases must have a corresponding noun phrase with possessive morphology that agrees in person with the genitive noun phrase; and truly nominative noun phrases must have a corresponding predicate that agrees in person with the nominative noun phrase---e.g., a verb or a copula.
+As seen in the table, the morphologically unmarked nominative case (e.g., үй \gloss{the/a house} \gmk{nom}) has a wide variety of uses, including indefinite object and genitival.
+Definite objects are marked with the accusative case (e.g., үйді \gloss{the house} \gmk{acc}), and definite genitivals are marked with the genitive case (e.g., үйдің \gloss{of the house} \gmk{gen}).
+While the Kazakh transducer marks these all as \tag{nom}, their use may be disambiguated.
+Genitival nominals, whether definite or not, must (NO EXCEPTIONS) have a corresponding nominal with possessive morphology that agrees in person and number (e.g., үйдің есіг\textbf{і} \gloss{the/a door to the house}), and subject nominals must have a corresponding predicate containing e.g., a verb or a copula that agrees in person and number with the nominative nominal (e.g., үй көшіріле\textbf{ді} \gloss{the house gets moved}).
 
@@
 %\bibliography{paper}
+\begin{small}
 \printbibliography
-
+\end{small}
 \end{document}

### 2015-06-21 2015-turklang-kazdep/paper.tex
update
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60921 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
+One difference in our annotation scheme compared to the standard universal dependency
+analysis is with coordination. While the universal dependency scheme takes the first
+conjunct as the head, we take the last. This decision was made based on the fact that
+Kazakh is a head-final language and morphological marking is only obligatory on the
+last conjunct in a series. Furthermore, experiments in representing coordination
+in other head-final languages have found that the final-conjunct head analysis 
+results in better parser accuracy \parencite{bengoetxea09}.
+
 \begin{figure}[htbp]

### 2015-06-21 2015-turklang-kazdep/paper.tex
new table
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60922 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 Definite objects are marked with the accusative case (e.g., үйді \gloss{the house} \gmk{acc}), and definite genitivals are marked with the genitive case (e.g., үйдің \gloss{of the house} \gmk{gen}).
-While the Kazakh transducer marks these all as \tag{nom}, their use may be disambiguated.
-Genitival nominals, whether definite or not, must (NO EXCEPTIONS) have a corresponding nominal with possessive morphology that agrees in person and number (e.g., үйдің есіг\textbf{і} \gloss{the/a door to the house}), and subject nominals must have a corresponding predicate containing e.g., a verb or a copula that agrees in person and number with the nominative nominal (e.g., үй көшіріле\textbf{ді} \gloss{the house gets moved}).
+The Kazakh transducer marks all bare nominals both as \tag{nom} and \tag{attr}, but the various functions may be disambiguated syntactically.
+Genitival nominals, whether definite or not, must (NO EXCEPTIONS) have a corresponding nominal with possessive morphology that agrees in person and number (e.g., үйдің есіг\textbf{і} \gloss{the/a door to the house}; үй тапсырма\textbf{сы} \gloss{homework}), and subject nominals must have a corresponding predicate containing e.g., a verb or a copula that agrees in person and number with the nominative nominal (e.g., үй көшіріле\textbf{ді} \gloss{the house gets moved}).
+When a nominative nominal depends on another nominal that does not have possessive case, the first one must be considered attributive (e.g., үй киім \gloss{house clothes}).
 
-Another category that may also be non-overt is the copula.
-In the present tense, all but third person copulas have realised forms (e.g., Мен үйде\textbf{мін}.
-\gloss{I'm at home.}), but the third person copula does not surface (e.g., Ол үйде.
-\gloss{S/he's at home.}).
+The attributive use of nominals is similar to the use of adjectives, and could even be thought of as a ``zero-derivation'' of nouns into adjectives.
+Interestingly, many adjectives can also be used substantively.
+Table~\ref{tab:zeroderiv} shows the various functions of nominals, adjectives, and adverbs.
+
+\begin{table}[htbp]
+	\centering
+	\caption{The default (first line of each) and ``zero-derived'' uses of nominals, adjectives, and adverbs}\label{tab:zeroderiv}
+	\begin{tabular}{lp{8em}p{22em}}
+		\toprule
+			\textbf{Category} & \textbf{Function} & \textbf{Example} \\
+		\midrule
+			Nominals & Substantive\newline
+							Attributive\newline
+							Adverbial
+						& hargle\newline 
+							jargal\newline
+							\\
+			Adjectives & Attributive\newline
+							Substantive\newline
+							Adverbial
+						& foo \newline
+						\newline
+						\\
+			Adverbs & Adverbial
+						&
+						\\
+		\bottomrule
+	\end{tabular}
+
+\end{table}
 
@@
 Similarly, many nouns may be used attributively, besides the default substantive reading.
-Besides the nominative (and indefinite accusative/genitive) reading of \tag{nom}, the transducer provides an attributive reading for most nominals of \tag{nom}\tag{attr}.
+Besides the nominative (and indefinite accusative/genitive) reading of \tag{nom}, the transducer provides an attributive reading for most nominals of \tag{attr}.
+
+Another category that may also be non-overt is the copula.
+In the present tense, all but third person copulas have realised forms (e.g., Мен үйде\textbf{мін}.
+\gloss{I'm at home.}), but the third person copula does not surface (e.g., Ол үйде.
+\gloss{S/he's at home.}).
 

### 2015-06-21 2015-turklang-kazdep/paper.tex
table2
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60923 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
+\usepackage{ulem}
 \usepackage{multirow}
@@
         \centering
-	\begin{tabular}{ll}
+\begin{small}
+        \begin{tabular}{ll|ll|ll}
                 \toprule
-		\textbf{Label} & \textbf{Description} \\
+                \textbf{Label} & \textbf{Description} & \textbf{Label} & \textbf{Description} & \textbf{Label} & \textbf{Description} \\
                 \midrule
-				\\
+\texttt{acl} & clausal modifier of noun & \texttt{csubjpass} & clausal passive subject & texttt{neg} & negation modifier\\
+\texttt{advcl} & adverbial clause modifier & \texttt{dep} & unspecified dependency & \texttt{nmod} & nominal modifier\\
+\texttt{advmod} & adverbial modifier & \texttt{det} & determiner & \texttt{nsubj} & nominal subject\\
+\texttt{amod} & adjectival modifier & \texttt{discourse} & discourse element & \texttt{nsubjpass} & passive nominal subject\\
+\texttt{appos} & appositional modifier & \texttt{dislocated} & dislocated elements & \texttt{nummod} & numeric modifier\\
+\texttt{aux} & auxiliary & \texttt{dobj} & direct object & \texttt{parataxis} & parataxis\\
+\texttt{\sout{auxpass}} & passive auxiliary & \texttt{expl} & expletive & \texttt{punct} & punctuation\\
+\texttt{case} & case marking & \texttt{foreign} & foreign words & \texttt{remnant} & remnant in ellipsis\\
+\texttt{cc} & coordinating conjunction & \texttt{goeswith} & goes with & \texttt{reparandum} & overridden disfluency\\
+\texttt{ccomp} & clausal complement & \texttt{iobj} & indirect object & \texttt{root} & root\\
+\texttt{compound} & compound & \texttt{list} & list & \texttt{vocative} & vocative\\
+\texttt{conj} & conjunct & \texttt{mark} & marker & \texttt{xcomp} & open clausal complement\\
+\texttt{cop} & copula & \texttt{mwe} & multi-word expression & \\
+\texttt{csubj} & clausal subject & \texttt{name} & name & \\
+
                 \bottomrule
         \end{tabular}
+\end{small}
         \caption{Universal dependency label set}

### 2015-06-21 2015-turklang-kazdep/paper.tex
updas
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60924 72bbbca6-d526-0410-a7d9-f06f51895060
update
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60925 72bbbca6-d526-0410-a7d9-f06f51895060

--- 2015-06-21 2015-turklang-kazdep/paper.tex
@@
 
+\usepackage[normalem]{ulem}
 \usepackage{multirow}
@@
 as a single token, and their lemma and morphological analysis is returned. Multiword
-units such as \emph{ауа райы} `weather' and \emph{ата-ананың} `of parents' are combined
+units such as \emph{ауа райы} `weather' and \emph{ата-анасының} `of their parents' are combined
 into a single token. Abbreviations and numerals which bear case, such as АҚШ-пен `with the USA'
@@
         \centering
-	\begin{tabular}{ll}
+\begin{small}
+        \begin{tabular}{ll|ll|ll}
                 \toprule
-		\textbf{Label} & \textbf{Description} \\
+                \textbf{Label} & \textbf{Description} & \textbf{Label} & \textbf{Description} & \textbf{Label} & \textbf{Description} \\
                 \midrule
-				\\
+\texttt{acl} & clausal modifier of noun & \texttt{\sout{csubjpass}} & clausal passive subject & texttt{\sout{neg}} & negation modifier\\
+\texttt{advcl} & adverbial clause modifier & \texttt{dep} & unspecified dependency & \texttt{nmod} & nominal modifier\\
+\texttt{advmod} & adverbial modifier & \texttt{det} & determiner & \texttt{\sout{nsubj}} & nominal subject\\
+\texttt{amod} & adjectival modifier & \texttt{discourse} & discourse element & \texttt{\sout{nsubjpass}} & passive nominal subject\\
+\texttt{appos} & appositional modifier & \texttt{\sout{dislocated}} & dislocated elements & \texttt{nummod} & numeric modifier\\
+\texttt{aux} & auxiliary & \texttt{dobj} & direct object & \texttt{parataxis} & parataxis\\
+\texttt{\sout{auxpass}} & passive auxiliary & \texttt{\sout{expl}} & expletive & \texttt{punct} & punctuation\\
+\texttt{case} & case marking & \texttt{foreign} & foreign words & \texttt{remnant} & remnant in ellipsis\\
+\texttt{cc} & coordinating conjunction & \texttt{\sout{goeswith}} & goes with & \texttt{\sout{reparandum}} & overridden disfluency\\
+\texttt{ccomp} & clausal complement & \texttt{iobj} & indirect object & \texttt{root} & root\\
+\texttt{compound} & compound & \texttt{list} & list & \texttt{vocative} & vocative\\
+\texttt{conj} & conjunct & \texttt{mark} & marker & \texttt{xcomp} & open clausal complement\\
+\texttt{cop} & copula & \texttt{\sout{mwe}} & multi-word expression & \\
+\texttt{\sout{csubj}} & clausal subject & \texttt{name} & name & \\
+
                 \bottomrule
         \end{tabular}
+\end{small}
         \caption{Universal dependency label set}

### 2015-06-21 2015-turklang-kazdep/paper.tex
dixW
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60926 72bbbca6-d526-0410-a7d9-f06f51895060
reformatting and stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60927 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{tikz-dependency}
-\newcommand{\gmk}[1]{\texttt{#1}}
-\newcommand{\gloss}[1]{{\em #1}}
+\newcommand{\gmk}[1]{{\rm {\sc \texttt{#1}}}}
+\newcommand{\kazakh}[1]{{\em #1}}
+\newcommand{\gloss}[1]{`#1'}
 \newcommand{\tag}[1]{\texttt{‹#1›}}
@@
 
-As seen in the table, the morphologically unmarked nominative case (e.g., үй \gloss{the/a house} \gmk{nom}) has a wide variety of uses, including indefinite object and genitival.
-Definite objects are marked with the accusative case (e.g., үйді \gloss{the house} \gmk{acc}), and definite genitivals are marked with the genitive case (e.g., үйдің \gloss{of the house} \gmk{gen}).
+As seen in the table, the morphologically unmarked nominative case (e.g., \kazakh{үй} \gloss{the/a house} \gmk{nom}) has a wide variety of uses, including indefinite object and genitival.
+Definite objects are marked with the accusative case (e.g., \kazakh{үйді} \gloss{the house} \gmk{acc}), and definite genitivals are marked with the genitive case (e.g., \kazakh{үйдің} \gloss{of the house} \gmk{gen}).
 The Kazakh transducer marks all bare nominals both as \tag{nom} and \tag{attr}, but the various functions may be disambiguated syntactically.
-Genitival nominals, whether definite or not, must (NO EXCEPTIONS) have a corresponding nominal with possessive morphology that agrees in person and number (e.g., үйдің есіг\textbf{і} \gloss{the/a door to the house}; үй тапсырма\textbf{сы} \gloss{homework}), and subject nominals must have a corresponding predicate containing e.g., a verb or a copula that agrees in person and number with the nominative nominal (e.g., үй көшіріле\textbf{ді} \gloss{the house gets moved}).
-When a nominative nominal depends on another nominal that does not have possessive case, the first one must be considered attributive (e.g., үй киім \gloss{house clothes}).
+Genitival nominals, whether definite or not, must (NO EXCEPTIONS) have a corresponding nominal with possessive morphology that agrees in person and number (e.g., \kazakh{үйдің есіг\textbf{і}} \gloss{the/a door to the house}; \kazakh{үй тапсырма\textbf{сы}} \gloss{homework}), and subject nominals must have a corresponding predicate containing e.g., a verb or a copula that agrees in person and number with the nominative nominal (e.g., \kazakh{үй көшіріле\textbf{ді}} \gloss{the house gets moved}).
+When a nominative nominal depends on another nominal that does not have possessive case, the first one must be considered attributive (e.g., \kazakh{үй киім} \gloss{house clothes}).
 
-The attributive use of nominals is similar to the use of adjectives, and could even be thought of as a ``zero-derivation'' of nouns into adjectives.
-Interestingly, many adjectives can also be used substantively.
-Table~\ref{tab:zeroderiv} shows the various functions of nominals, adjectives, and adverbs.
+The attributive use of nominals is similar to the use of adjectives, and could even be thought of as a ``zero-derivation'' of nominals into adjectives.
+Interestingly, many adjectives can also be used substantively, i.e., as nominals.
+Table~\ref{tab:zeroderiv} shows the various functions that of nominals, adjectives, and adverbs may take.
 
@@
 							jargal\newline
-							\\
+							бір рет \gloss{once} \\
 			Adjectives & Attributive\newline
@@
 
-Another type of ``zero-derivation'' involves the use of adjectives and nouns.
+Nominals (\tag{n}) default to substantive and adjectives (\tag{adj}) default to attributive.
+For the other readings, the transducer provides readings such as \tag{adj}\tag{advl} and \tag{n}\tag{attr}.
+
+%Another type of ``zero-derivation'' involves the use of adjectives and nouns.
 In Kazakh, adjectives are all attributive, but many may also be used either nominally (i.e., as nouns) or adverbially (i.e., as adverbs), or both.
@@
 Another category that may also be non-overt is the copula.
-In the present tense, all but third person copulas have realised forms (e.g., Мен үйде\textbf{мін}.
-\gloss{I'm at home.}), but the third person copula does not surface (e.g., Ол үйде.
-\gloss{S/he's at home.}).
+The primary strategy for copula constructions in Kazakh is the use of a defective verb \kazakh{е-}.
+In the present tense, all but third person copulas have realised forms (e.g., \kazakh{Мен үйде\textbf{мін}.} \gloss{I'm at home.}), but the third person copula does not surface (e.g., \kazakh{Ол үйде.} \gloss{S/he's at home.}).
 
@@
 performs tokenisation on the basis of a left-to-right longest match algorithm described
-in \textcite{garrido02}. Simple tokens such as \emph{толқулар} `riots' are maintained
+in \textcite{garrido02}. Simple tokens such as \kazakh{толқулар} \gloss{riots} are maintained
 as a single token, and their lemma and morphological analysis is returned. Multiword
-units such as \emph{ауа райы} `weather' and \emph{ата-анасының} `of their parents' are combined
-into a single token. Abbreviations and numerals which bear case, such as АҚШ-пен `with the USA'
-and \emph{90\%-ына} `to the 90\%' are analysed as a single token, as are light-verbs such as
-\emph{пайда бол} `to appear' and tense forms written with space like \emph{оқыған жоқ},
-the third-person negative past of \emph{оқы} `to learn'.
+units such as \kazakh{ауа райы} \gloss{weather} and \kazakh{ата-анасының} \gloss{of their parents} are combined
+into a single token. Abbreviations and numerals which bear case, such as \kazakh{АҚШ-пен} \gloss{with the USA}
+and \kazakh{90\%-ына} \gloss{to 90\%} are analysed as a single token, as are light-verbs such as
+\kazakh{пайда бол} \gloss{to appear} and tense forms written with space like \kazakh{оқыған жоқ},
+the third-person negative past of \kazakh{оқы-} \gloss{to study/read}.
 
 In some cases a single token is split into two tokens, as with the aorist copula suffixes,
-\emph{үйдемін} `I am in the house' is tokenised as үй.{\sc loc} + е.{\sc cop.aor.sg1}. Furthermore
-two input tokens may result in three output tokens, \emph{бар ма?} `is there?' would be
-tokenised as бар.{\sc adj} + е.{\sc cop.aor.sg1} + ма.{\sc qst}.
+\kazakh{үйдемін} \gloss{I am in the house} is tokenised as \kazakh{үй.\gmk{loc} + е.\gmk{cop.aor.sg1}}. Furthermore
+two input tokens may result in three output tokens, \kazakh{бар ма?} \gloss{is there?} would be
+tokenised as \kazakh{бар.\gmk{adj} + е.\gmk{cop.aor.sg1} + ма.\gmk{qst}}.
 
@@
                 \midrule
-\texttt{acl} & clausal modifier of noun & \texttt{\sout{csubjpass}} & clausal passive subject & texttt{\sout{neg}} & negation modifier\\
+\texttt{acl} & clausal modifier of noun & \texttt{\sout{csubjpass}} & clausal passive subject & \texttt{\sout{neg}} & negation modifier\\
 \texttt{advcl} & adverbial clause modifier & \texttt{dep} & unspecified dependency & \texttt{nmod} & nominal modifier\\

### 2015-06-21 2015-turklang-kazdep/paper.tex
update
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60928 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
+The copula (both \emph{е} and \emph{бол}) is a challenging problem for dependency 
+analysis of Kazakh. The universal dependency guidelines state that the copula should
+be the dependent of the lexical predicate. However, in many cases the copula in Kazakh
+is found in embedded clauses ...
+
+We have uniformly annotated the copula as a leaf node with the predicate, or adverbial
+as the head of the structure. For certain structures this is convenient, such as the
+bare copula in phrases like \ref{ex:cop1} or \ref{ex:cop2}, but for phrases where
+the copula is part of an embedded clause this is not necessarily the most  
+effective choice. In \ref{ex:cop3}, the copula holds all the morphological information,
+agreement with the subject and accusative marking for the embedded clause.
+
 \begin{figure}[htbp]
@@
 
-	\begin{subfigure}[b]{0.3\textwidth}
+        \begin{subfigure}[b]{0.25\textwidth}
                 \centering
@@
                         \depedge{2}{1}{\udlabel{subj}}
-			\depedge{3}{2}{\udlabel{cop}}
+                        \depedge{2}{3}{\udlabel{cop}}
                 \end{dependency}
@@
         \quad
-	\begin{subfigure}[b]{0.3\textwidth}
+        \begin{subfigure}[b]{0.25\textwidth}
                 \centering
@@
                    \begin{deptext}[column sep=0.3cm]
-				Айгүл \& студент \& - \\
-				\gmk{np} \& \gmk{-} \& \gmk{-} \\
-				\gloss{Aygül} \& \gloss{-} \& \gloss{-}\\
+                                Айгүл \& студент \& болған \\
+                                \gmk{np} \& \gmk{n} \& \gmk{past} \\
+                                \gloss{Aygül} \& \gloss{student} \& \gloss{was}\\
                         \end{deptext}
-			\depedge{3}{2}{\udlabel{cop}}
+                        \depedge{2}{1}{\udlabel{subj}}
+                        \depedge{2}{3}{\udlabel{cop}}
                 \end{dependency}
@@
         \quad
-	\begin{subfigure}[b]{0.3\textwidth}
+        \begin{subfigure}[b]{0.40\textwidth}
                 \centering
                 \begin{dependency}[theme = simple, font = \small]
-		   \begin{deptext}[column sep=0.3cm]
-				Айгүл   \& - \& - \\
-				\gmk{n} \& \gmk{-} \& \gmk{-} \& \\
-				\gloss{Aygül} \& \gloss{-} \& \gloss{-} \\
+                   \begin{deptext}[column sep=0.2cm]
+                                Айгүл \& Азаматтың \& қайда \& екенін \& білмейді. \\
+                                \gmk{np} \& \gmk{np.gen} \& \gmk{adv} \& \gmk{cop} \& \gmk{neg.aor} \\
+                                \gloss{Aygül} \& \gloss{Azamat} \& \gloss{where} \& \gloss{being} \& \gloss{knows not} \\
                         \end{deptext}
-			\depedge{3}{2}{\udlabel{cop}}
+                        \depedge{5}{1}{\udlabel{subj}}
+                        \depedge{3}{2}{\udlabel{subj}}
+                        \depedge{5}{3}{\udlabel{ccomp}}
+                        \depedge{3}{4}{\udlabel{cop}}
                 \end{dependency}

### 2015-06-21 2015-turklang-kazdep/paper.tex
some bibfixes, reformatted tree stuff, etc. and more stuff since then argh
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60931 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \newcommand{\gloss}[1]{`#1'}
+\newcommand{\tgloss}[1]{{\em #1}}
 \newcommand{\tag}[1]{\texttt{‹#1›}}
@@
 							Adverbial
-						& foo \newline
-						\newline
+						& \kazakh{\textbf{жақсы} үй} \gloss{nice house} \newline
+						\kazakh{ең \textbf{жақсы}лары} \gloss{\textbf{the nice}st \textbf{ones}} \newline
+						\kazakh{оны \textbf{жақсы} танимын} \gloss{I know him/her \textbf{well}.}
 						\\
@@
 The primary strategy for copula constructions in Kazakh is the use of a defective verb \kazakh{е-}.
-In the present tense, all but third person copulas have realised forms (e.g., \kazakh{Мен үйде\textbf{мін}.} \gloss{I'm at home.}), but the third person copula does not surface (e.g., \kazakh{Ол үйде.} \gloss{S/he's at home.}).
+In the present tense, the verb itself does not surface, but agreement morphology surfaces (cliticised to the previous word) in all but the third person forms (e.g., \kazakh{Мен үйде\textbf{мін}.} \gloss{I'm at home.} versus \kazakh{Ол үйде.} \gloss{S/he's at home.}).
+The defective copula verb also surfaces in the recent past tense (e.g., \kazakh{Мен үйде еді\textbf{м}.} \gloss{I was at home.}, \kazakh{Інім үйде еді.} \gloss{My younger brother was at home.}).
+Of particular interest are non-finite forms of the copula.
+Copula clauses can be attributive, as in \kazakh{үйі үлкен жігіт}
 
@@
                                 \gmk{np} \& \gmk{n} \& \gmk{cop} \\
-                                \gloss{Aygül} \& \gloss{student} \& \gloss{is} \\
+                                \tgloss{Aygül} \& \tgloss{student} \& \tgloss{is} \\
                         \end{deptext}
@@
                                 \gmk{np} \& \gmk{n} \& \gmk{past} \\
-                                \gloss{Aygül} \& \gloss{student} \& \gloss{was}\\
+                                \tgloss{Aygül} \& \tgloss{student} \& \tgloss{was}\\
                         \end{deptext}
@@
                                 \gmk{np} \& \gmk{np.gen} \& \gmk{adv} \& \gmk{cop} \& \gmk{neg.aor} \\
-                                \gloss{Aygül} \& \gloss{Azamat} \& \gloss{where} \& \gloss{being} \& \gloss{knows not} \\
+                                \tgloss{Aygül} \& \tgloss{Azamat} \& \tgloss{where} \& \tgloss{being} \& \tgloss{knows not} \\
                         \end{deptext}
@@
                     \&         \&           \& \gmk{adj} \& \gmk{cnjcoo} \& \gmk{n}       \& \gmk{cnjcoo} \& \gmk{n.pl} \& \gmk{adj} \& \gmk{cop} \& \& \\
-                    \&         \&           \& \gloss{free} \& \gloss{and}  \& \gloss{dignity}  \&  \gloss{and} \& \gloss{rights} \&  \gloss{equal} \& \gloss{being} \& \& \\
+                    \&         \&           \& \tgloss{free} \& \tgloss{and}  \& \tgloss{dignity}  \&  \tgloss{and} \& \tgloss{rights} \&  \tgloss{equal} \& \tgloss{being} \& \& \\
 		\end{deptext}
@@
 				\gmk{n.attr} \& \gmk{n.pl.nom} \\
-				\gloss{neighbour} \& \gloss{countries} \\
+				\tgloss{neighbour} \& \tgloss{countries} \\
 			\end{deptext}
@@
 				\gmk{n.nom} \& \gmk{n.px3sg.nom} \\
-				\gloss{world} \& \gloss{championship} \\
+				\tgloss{world} \& \tgloss{championship} \\
 			\end{deptext}
@@
 				\gmk{n.gen} \& \gmk{n.px3sg.nom} \\
-				\gloss{Iran's} \& \gloss{economy} \\
+				\tgloss{Iran's} \& \tgloss{economy} \\
 			\end{deptext}
@@
 \begin{table}[htbp]
+	\caption{Preliminary parsing results from MaltParser using different models.
+The numbers in brackets denote the upper and lower bounds found during cross validation.
+Adding structural features to the model substantially improves the performance of the parser, although we find no improvement in using full morphosyntactic description (MSD) over using simply the first part-of-speech (POS) tag.}
 	\centering
@@
 	\end{tabular}
-	\caption{Preliminary parsing results from MaltParser using different models.
-The numbers in brackets denote the upper and lower bounds found during cross validation.
-Adding structural features to the model substantially improves the performance of the parser, although we find no improvement in using full morphosyntactic description (MSD) over using simply the first part-of-speech (POS) tag.}
 	\label{table:eval}

### 2015-06-21 2015-turklang-kazdep/paper.tex
some table examples
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60932 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 	\caption{The default (first line of each) and ``zero-derived'' uses of nominals, adjectives, and adverbs}\label{tab:zeroderiv}
-	\begin{tabular}{lp{8em}p{22em}}
+	\begin{tabular}{lp{5em}p{22em}}
 		\toprule
@@
 							Adverbial
-						& hargle\newline 
-							jargal\newline
-							бір рет \gloss{once} \\
+						& \kazakh{\textbf{Қонақ}тарымыз бай.} \gloss{Our \textbf{guest}s are rich.}\newline 
+							\kazakh{\textbf{қонақ} үй} \gloss{hotel (lit., \textbf{guest} house)}\newline
+							\kazakh{бір \textbf{рет}} \gloss{one \textbf{time}} \\
 			Adjectives & Attributive\newline
@@
 							Adverbial
-						& \kazakh{\textbf{жақсы} үй} \gloss{nice house} \newline
+						& \kazakh{\textbf{жақсы} үй} \gloss{\textbf{nice} house} \newline
 						\kazakh{ең \textbf{жақсы}лары} \gloss{\textbf{the nice}st \textbf{ones}} \newline
-						\kazakh{оны \textbf{жақсы} танимын} \gloss{I know him/her \textbf{well}.}
+						\kazakh{Оны \textbf{жақсы} танимын.} \gloss{I know him/her \textbf{well}.}
 						\\
 			Adverbs & Adverbial
-						& 
+						& \kazakh{Ол \textbf{кеше} кетті} \gloss{S/he left \textbf{yesterday}.}
 						\\

### 2015-06-21 2015-turklang-kazdep/paper.tex
figs
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60933 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 as the head of the structure. For certain structures this is convenient, such as the
-bare copula in phrases like \ref{ex:cop1} or \ref{ex:cop2}, but for phrases where
+bare copula in phrases like \ref{fig:cop1} or \ref{fig:cop2}, but for phrases where
 the copula is part of an embedded clause this is not necessarily the most  
-effective choice. In \ref{ex:cop3}, the copula holds all the morphological information,
+effective choice. In \ref{fig:cop3}, the copula holds all the morphological information,
 agreement with the subject and accusative marking for the embedded clause.
@@
                 \begin{dependency}[theme = simple, font = \small]
-                   \begin{deptext}[column sep=0.3cm]
+                   \begin{deptext}[column sep=0.2cm]
                                 Айгүл \& студент \& · \\
-                                \gmk{np} \& \gmk{n} \& \gmk{cop} \\
-                                \tgloss{Aygül} \& \tgloss{student} \& \tgloss{is} \\
+                                \gmk{np} \& \gmk{n} \& \gmk{aor} \\
+                                \gloss{Aygül} \& \gloss{student} \& \gloss{is} \\
                         \end{deptext}
+                        \deproot[edge unit distance=1.3ex]{2}{\udlabel{root}}
                         \depedge{2}{1}{\udlabel{subj}}
@@
                 \end{dependency}
-                \caption{xxx }\label{fig:compound}
+                \caption{Non-surfaced copula.}\label{fig:cop1}
         \end{subfigure}
@@
                 \begin{dependency}[theme = simple, font = \small]
-                   \begin{deptext}[column sep=0.3cm]
+                   \begin{deptext}[column sep=0.2cm]
                                 Айгүл \& студент \& болған \\
                                 \gmk{np} \& \gmk{n} \& \gmk{past} \\
-                                \tgloss{Aygül} \& \tgloss{student} \& \tgloss{was}\\
+                                \gloss{Aygül} \& \gloss{student} \& \gloss{was}\\
                         \end{deptext}
+                        \deproot[edge unit distance=1.3ex]{2}{\udlabel{root}}
                         \depedge{2}{1}{\udlabel{subj}}
@@
                 \end{dependency}
-                \caption{xxx.}\label{fig:cop2}
+                \caption{Copula in past tense.}\label{fig:cop2}
         \end{subfigure}
@@
                    \begin{deptext}[column sep=0.2cm]
-                                Айгүл \& Азаматтың \& қайда \& екенін \& білмейді. \\
-                                \gmk{np} \& \gmk{np.gen} \& \gmk{adv} \& \gmk{cop} \& \gmk{neg.aor} \\
-                                \tgloss{Aygül} \& \tgloss{Azamat} \& \tgloss{where} \& \tgloss{being} \& \tgloss{knows not} \\
+                                Айгүл \& оның \& қайда \& екенін \& білмейді. \\
+                                \gmk{np} \& \gmk{gen} \& \gmk{adv} \& \gmk{cop} \& \gmk{neg.aor} \\
+                                \gloss{Aygül} \& \gloss{he} \& \gloss{where} \& \gloss{being} \& \gloss{knows not} \\
                         \end{deptext}
-                        \depedge{5}{1}{\udlabel{subj}}
+                        \deproot[edge unit distance=1.3ex]{5}{\udlabel{root}}
                         \depedge{3}{2}{\udlabel{subj}}
@@
                 \end{dependency}
-                \caption{xxx.}\label{fig:cop3}
+                \caption{Copula in an embedded clause.}\label{fig:cop3}
         \end{subfigure}
-
         \caption{Dependency trees of copula constructions.}
@@
            \begin{deptext}[column sep=0.08cm]
-%               1         2         3          4              5                6              7                8           9             10
-             Барлық \& адамдар \& тумысынан \& азат      \& және         \& қадір-қасиеті \& мен          \& құқықтары  \& тең       \& болып     \& дүниеге \& келеді \\
-                    \&         \&           \& \gmk{adj} \& \gmk{cnjcoo} \& \gmk{n}       \& \gmk{cnjcoo} \& \gmk{n.pl} \& \gmk{adj} \& \gmk{cop} \& \& \\
-                    \&         \&           \& \tgloss{free} \& \tgloss{and}  \& \tgloss{dignity}  \&  \tgloss{and} \& \tgloss{rights} \&  \tgloss{equal} \& \tgloss{being} \& \& \\
+%               1         2         3          4              5          6          7        8
+                Олар \& Финляндия, \& Швеция \& және \& Эстонияидан \& кейін \& Ресейге \& барды. \\
+                \gmk{prn} \& \gmk{nom} \& \gmk{nom} \& \gmk{cc} \& \gmk{abl} \& \gmk{post} \& \gmk{dat} \& \gmk{past} \\
+                \gloss{They} \& \gloss{Finland}, \& \gloss{Sweden} \& \gloss{and} \& \gloss{Estonia} \& \gloss{after} \& \gloss{Russia} \& \gloss{visited.} \\
                 \end{deptext}
-		\depedge{9}{4}{\udlabel{conj}}
-		\depedge{9}{5}{\udlabel{cc}}
-		\depedge{8}{6}{\udlabel{conj}}
-		\depedge{8}{7}{\udlabel{cc}}
-		\depedge{9}{8}{\udlabel{subj}}
-		\depedge{10}{9}{\udlabel{cop}}
-		\depedge{12}{10}{\udlabel{advcl}}
+                \depedge{5}{3}{\udlabel{conj}}
+                \depedge{5}{4}{\udlabel{cc}}
+                \depedge{5}{2}{\udlabel{conj}}
+                \depedge{5}{6}{\udlabel{case}}
+                \depedge{8}{5}{\udlabel{nmod}}
+                \depedge{8}{7}{\udlabel{nmod}}
         \end{dependency}
-	\caption{Coordination}\label{fig:coord}
+        \caption{Coordination: All conjucts are attached to the final conjunct, which is the head of the coordinated phrase.}\label{fig:coord}
 

### 2015-06-21 2015-turklang-kazdep/paper.tex
table fixes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60934 72bbbca6-d526-0410-a7d9-f06f51895060
future work from Fran
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60935 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 The Kazakh transducer marks all bare nominals both as \tag{nom} and \tag{attr}, but the various functions may be disambiguated syntactically.
-Genitival nominals, whether definite or not, must (NO EXCEPTIONS) have a corresponding nominal with possessive morphology that agrees in person and number (e.g., \kazakh{үйдің есіг\textbf{і}} \gloss{the/a door to the house}; \kazakh{үй тапсырма\textbf{сы}} \gloss{homework}), and subject nominals must have a corresponding predicate containing e.g., a verb or a copula that agrees in person and number with the nominative nominal (e.g., \kazakh{үй көшіріле\textbf{ді}} \gloss{the house gets moved}).
+Genitival nominals, whether definite or not, must have a corresponding nominal with possessive morphology that agrees in person and number (e.g., \kazakh{үйдің есіг\textbf{і}} \gloss{the/a door to the house}; \kazakh{үй тапсырма\textbf{сы}} \gloss{homework}), and subject nominals must have a corresponding predicate containing e.g., a verb or a copula that agrees in person and number with the nominative nominal (e.g., \kazakh{үй көшіріле\textbf{ді}} \gloss{the house gets moved}).
 When a nominative nominal depends on another nominal that does not have possessive case, the first one must be considered attributive (e.g., \kazakh{үй киім} \gloss{house clothes}).
@@
 	\caption{The default (first line of each) and ``zero-derived'' uses of nominals, adjectives, and adverbs}\label{tab:zeroderiv}
-	\begin{tabular}{lp{5em}p{22em}}
+	\begin{tabular}{lp{5em}p{20em}}
 		\toprule
@@
 			Adverbs & Adverbial
-						& \kazakh{Ол \textbf{кеше} кетті} \gloss{S/he left \textbf{yesterday}.}
+						& \kazakh{Ол \textbf{кеше} кетті.} \gloss{S/he left \textbf{yesterday}.}
 						\\
@@
 
+Non-finite
+
 Another category that may also be non-overt is the copula.
@@
 Of particular interest are non-finite forms of the copula.
-Copula clauses can be attributive, as in \kazakh{үйі үлкен жігіт}
+Copula clauses can be attributive, as in \kazakh{үйі жақсы дәрігер} \gloss{the doctor who has a nice house}, or literally \gloss{[his/her house is nice] the doctor}.
+While this construction never has overt copula marking, copula clauses which are complements of verbs always have an overt copula form, e.g.
+\kazakh{Дәрігердің үйі жақсы \textbf{екенін} білген жоқпын.} \gloss{I didn't know that the doctor's house was nice.}  Here, \kazakh{екен} is a suppletive gerundive form of the copula and is marked as accusative.
 
@@
 
+Future work will focus on improving the annotation guidelines and the consistency of
+annotation in the corpus. We will also study the possibility of deepening the annotation
+with Turkic-specific relations. When we have stable annotation guidelines we intend to 
+extend the corpus with more texts.
+We would also like to work on cross-lingual dependency parsing,
+that is, applying a model learnt on the Kazakh treebank to other Turkic languages
+such as Tatar, Kumyk and Tuvan. We have morphological analysers for these languages which have
+compatible tagsets for morphological features and as such it should be possible to learn a
+delexicalised model based on these features. As Turkic syntax is broadly homogenous, this
+presents a promising avenue for future work.
+
 \section{Conclusions}\label{sec:conclusions}

### 2015-06-21 2015-turklang-kazdep/paper.tex
commit
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60936 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-Non-finite
+Kazakh, like most Turkic languages, makes frequent use of non-finite verb forms by deriving verbal adjectives, verbal nouns, and verbal adverbs from verbs.
+Verbal adjective phrases modify nouns, as in \kazakh{мені \textbf{көрген} дәрігер} \gloss{the doctor that saw me}.
+Verbal nouns can function as subjects or complements of verbs or copulas, as in \kazakh{Дәрігер сені \textbf{көргенін} білген жоқпын.} \gloss{I didn't know that the doctor had seen you.} and \kazakh{Дәрігер сені \textbf{көргені} жақсы болыпты.} \gloss{It's good that doctor saw you.}
 
@@
 			\midrule
-			\multicolumn{2}{c}{\tilda{}} & 302 & 3741 & 12.3 \\\cmidrule[\heavyrulewidth]{3-5}
+			\multicolumn{2}{c}{~} & 302 & 3741 & 12.3 \\\cmidrule[\heavyrulewidth]{3-5}
 		\end{tabular}

### 2015-06-21 2015-turklang-kazdep/paper.tex
concluding remarks from Fran
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60937 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-\section{Conclusions}\label{sec:conclusions}
+\section{Concluding remarks}\label{sec:conclusions}
+
+We have presented the first steps towards a free/open-source dependency treebank for
+Kazakh with annotation based on the universal dependencies. The treebank is small, but
+provides a base for bootstrapping further. Performance of a state-of-the-art statistical
+parser trained on the treebank is comparable to other treebanks of similar size.
 

### 2015-06-21 2015-turklang-kazdep/paper.tex
kazakh section done?
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60940 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 Kazakh, like most Turkic languages, makes frequent use of non-finite verb forms by deriving verbal adjectives, verbal nouns, and verbal adverbs from verbs.
-Verbal adjective phrases modify nouns, as in \kazakh{мені \textbf{көрген} дәрігер} \gloss{the doctor that saw me}.
-Verbal nouns can function as subjects or complements of verbs or copulas, as in \kazakh{Дәрігер сені \textbf{көргенін} білген жоқпын.} \gloss{I didn't know that the doctor had seen you.} and \kazakh{Дәрігер сені \textbf{көргені} жақсы болыпты.} \gloss{It's good that doctor saw you.}
+Verbal adjective phrases modify nouns, as in \kazakh{\textbf{мені көрген} дәрігер} \gloss{the doctor \textbf{that saw me}}.
+Verbal nouns can function as subjects or complements of verbs or copulas, as in \kazakh{Дәрігер сені \textbf{көргенін} білген жоқпын.} \gloss{I didn't know that the doctor had seen you.} and \kazakh{Дәрігер сені \textbf{көргені} жақсы болыпты.} \gloss{It's good that doctor saw you.}  Verbal adverbs allow a verb phrase to function as a clausal verbal adjunct, as in \kazakh{Мен \textbf{дәрігерді көріп} қуанып кеттім.} \gloss{\textbf{Seeing the doctor}, I got happy.}  Verbal nouns with certain case morphology and/or occurring with certain postpositions can also function as a verbal adjunct much in the same way that verbal adverb clauses do, as in \kazakh{Мен \textbf{дәрігерді көргенде} қуанып кеттім.} \gloss{I got happy \textbf{when I saw the doctor}.}, here with a verbal noun in the locative case.
 
@@
 Of particular interest are non-finite forms of the copula.
-Copula clauses can be attributive, as in \kazakh{үйі жақсы дәрігер} \gloss{the doctor who has a nice house}, or literally \gloss{[his/her house is nice] the doctor}.
+Copula clauses can be attributive, as in \kazakh{\textbf{үйі жақсы} дәрігер} \gloss{the doctor \textbf{who has a nice house}}, or literally \gloss{\textbf{his/her house is nice} the doctor}.
 While this construction never has overt copula marking, copula clauses which are complements of verbs always have an overt copula form, e.g.
-\kazakh{Дәрігердің үйі жақсы \textbf{екенін} білген жоқпын.} \gloss{I didn't know that the doctor's house was nice.}  Here, \kazakh{екен} is a suppletive gerundive form of the copula and is marked as accusative.
+\kazakh{Дәрігердің үйі жақсы \textbf{екенін} білген жоқпын.} \gloss{I didn't know \textbf{that} the doctor's house \textbf{was} nice.}  Here, \kazakh{екен} is a suppletive gerundive form of the copula and is marked as accusative.
 

### 2015-06-21 2015-turklang-kazdep/paper.tex
stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60941 72bbbca6-d526-0410-a7d9-f06f51895060
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60943 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{multirow}
+\usepackage{multicol}
 \usepackage[small,bf]{caption} % CHECK IF THIS IS OK
@@
 \usepackage{booktabs}
+
+%\usepackage{supertabular}
+%\newcount\n
+%\n=0
+%\def\tablebody{}
+%\makeatletter
+%\loop\ifnum\n<100
+        %\advance\n by1
+        %\protected@edef\tablebody{\tablebody
+                %\textbf{\number\n.}& shortText
+                %\tabularnewline
+        %}
+%\repeat
+
+%\makeatletter
+%\let\mcnewpage=\newpage
+%\newcommand{\TrickSupertabularIntoMulticols}{%
+  %\renewcommand\newpage{%
+    %\if@firstcolumn
+      %\hrule width\linewidth height0pt
+      %\columnbreak
+    %\else
+      %\mcnewpage
+    %\fi
+  %}%
+%}
+%\makeatother
 
@@
 \abstract{
-This article describes the first steps towards a free/open-source dependency treebank for Kazakh based on universal dependency (UD) annotation standards.
-The treebank contains 302 sentences and is based on texts from a range of open-source and public domain sources.
-This ensures its free availability and extensibility.
-Texts in the treebank are first morphologically analysed and disambiguated and then annotated manually for dependency structure.
+This article describes the first steps towards a free/open-source dependency treebank for Kazakh based on
+universal dependency (UD) annotation standards. The treebank contains 302 sentences and is based on
+texts from a range of open-source and public domain sources. This ensures
+its free availability and extensibility. Texts in the treebank are first morphologically analysed
+and disambiguated and then annotated manually for dependency structure. In the article we present
+some issues in dependency syntax for Kazakh and how these are analysed in the universal-dependency
+framework. Preliminary results for statistical dependency parsing of Kazakh are reported, along with
+some directions for future research.
 }
@@
 
+%FIXME:
 % general procedure
@@
 \end{table}
+
+%\begin{small}
+%\begin{multicols*}{3}
+	%\tablehead{\toprule \textbf{Label} & \textbf{Description}\\\midrule}
+	%\tabletail{\bottomrule}
+%\TrickSupertabularIntoMulticols
+
+	%\begin{supertabular}{ll}
+%\texttt{acl} & clausal modifier of noun\\
+%\texttt{advcl} & adverbial clause modifier\\
+%\texttt{advmod} & adverbial modifier\\
+%\texttt{amod} & adjectival modifier \\
+%\texttt{appos} & appositional modifier\\
+%\texttt{aux} & auxiliary  \\
+%\texttt{\sout{auxpass}} & passive auxiliary \\
+%\texttt{case} & case marking \\
+%\texttt{cc} & coordinating conjunction \\
+%\texttt{ccomp} & clausal complement \\
+%\texttt{compound} & compound \\
+%\texttt{conj} & conjunct\\ 
+%\texttt{cop} & copula \\
+%\texttt{\sout{csubj}} & clausal subject \\
+%\texttt{\sout{csubjpass}} & clausal passive subject\\
+%\texttt{dep} & unspecified dependency\\
+%\texttt{det} & determiner \\
+%\texttt{discourse} & discourse element \\
+%\texttt{\sout{dislocated}} & dislocated elements \\
+%\texttt{dobj} & direct object \\
+%\texttt{\sout{expl}} & expletive \\
+%\texttt{foreign} & foreign words \\
+%\texttt{\sout{goeswith}} & goes with \\
+%\texttt{iobj} & indirect object \\
+%\texttt{list} & list \\
+%\texttt{mark} & marker \\
+%\texttt{\sout{mwe}} & multi-word expression \\ 
+%\texttt{name} & name \\
+%\texttt{\sout{neg}} & negation modifier\\
+%\texttt{nmod} & nominal modifier\\
+%\texttt{\sout{nsubj}} & nominal subject\\
+%\texttt{\sout{nsubjpass}} & passive nominal subject\\
+%\texttt{nummod} & numeric modifier\\
+%\texttt{parataxis} & parataxis\\
+%\texttt{punct} & punctuation\\
+%\texttt{remnant} & remnant in ellipsis\\
+%\texttt{\sout{reparandum}} & overridden disfluency\\
+%\texttt{root} & root\\
+%\texttt{vocative} & vocative\\
+%\texttt{xcomp} & open clausal complement\\
+	%\end{supertabular}
+%\end{multicols*}
+%\end{small}
 
@@
                 \gmk{prn} \& \gmk{nom} \& \gmk{nom} \& \gmk{cc} \& \gmk{abl} \& \gmk{post} \& \gmk{dat} \& \gmk{past} \\
-                \gloss{They} \& \gloss{Finland}, \& \gloss{Sweden} \& \gloss{and} \& \gloss{Estonia} \& \gloss{after} \& \gloss{Russia} \& \gloss{visited.} \\
+                \tgloss{They} \& \tgloss{Finland}, \& \tgloss{Sweden} \& \tgloss{and} \& \tgloss{Estonia} \& \tgloss{after} \& \tgloss{Russia} \& \tgloss{visited.} \\
                 \end{deptext}
@@
                 \depedge{8}{7}{\udlabel{nmod}}
+                \deproot[edge unit distance=1.3ex]{8}{\udlabel{root}}
         \end{dependency}
@@
 \subsection{Non-finite clauses}
+
+\begin{figure}[htbp]
+	\centering
+	\begin{dependency}[theme = simple, font = \small]
+		\begin{deptext}[column sep=0.08cm]
+			Бразилия \& өз \& жерінде \& чемпионатты \& екі \& рет \& өткізген \& бесінші \& ел \& болды. \\
+			\gmk{np.nom} \& \gmk{det} \& \gmk{nom.px3sp.loc} \& \gmk{n.acc} \& \gmk{num} \& \gmk{n} \& \gmk{v.tv.gpr} \& \gmk{num.ord} \& \gmk{n.nom} \& \gmk{v.iv.ifi} \\
+			\tgloss{Brasil} \& \tgloss{own} \& \tgloss{in its land} \& \tgloss{championship} \& \tgloss{two} \& \tgloss{times} \& \tgloss{having held} \& \tgloss{fifth} \& \tgloss{country} \& \tgloss{became} \\
+		\end{deptext}
+		\depedge{3}{2}{\udlabel{det}}
+		\depedge{7}{3}{\udlabel{nmod}}
+		\depedge{7}{4}{\udlabel{obj}}
+		\depedge{6}{5}{\udlabel{nummod}}
+		\depedge{7}{6}{\udlabel{nmod}}
+		\depedge{9}{7}{\udlabel{acl}}
+		\deproot[edge unit distance=1.3ex]{9}{\udlabel{root}}
+	\end{dependency}
+	\caption{Verbal adjectives: verbal adjectives are the head of everything in their own clause and are an \udtag{acl} dependency of the noun they modify.}\label{fig:gpr}
+\end{figure}
+
+Бразилия өз жерінде чемпионатты екі рет өткізген бесінші ел болды
+
+1960 жылдан бастап Батыс үлгісінде дами бастады.
+
+Ол терезеден Азамат пен Айгүлдің ойнағанына қарап тұр.
+
+Айгүл санап біткеннен кейін айналасына қарады.
 

### 2015-06-21 2015-turklang-kazdep/paper.tex
verbal adverbs
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60944 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 		\depedge{9}{7}{\udlabel{acl}}
+		\depedge{9}{8}{\udlabel{amod}}
+		\depedge{9}{10}{\udlabel{cop}}
 		\deproot[edge unit distance=1.3ex]{9}{\udlabel{root}}
@@
 	\caption{Verbal adjectives: verbal adjectives are the head of everything in their own clause and are an \udtag{acl} dependency of the noun they modify.}\label{fig:gpr}
+\end{figure}
+
+\begin{figure}[htbp]
+	\centering
+	\begin{dependency}[theme = simple, font = \small]
+		\begin{deptext}[column sep=0.08cm]
+			1960 \& жылдан \& бастап \& Батыс \& үлгісінде \& дами \& бастады. \\
+			\gmk{num} \& \gmk{n.abl} \& \gmk{v.tv.gna} \& \gmk{n.nom} \& \gmk{n.px3sp.loc} \& \gmk{v.iv.prc} \& \gmk{vaux.ifi} \\
+			\tgloss{1960} \& \tgloss{from year} \& \tgloss{starting} \& \tgloss{the West} \& \tgloss{on its model} \& \tgloss{developing} \& \tgloss{started} \\
+		\end{deptext}
+		\depedge{2}{1}{\udlabel{amod}}
+		\depedge{3}{2}{\udlabel{nmod}}
+		\depedge{6}{3}{\udlabel{advcl}}
+		\deproot[edge unit distance=1.3ex]{6}{\udlabel{root}}
+	\end{dependency}
+	\caption{Verbal adverbs: verbal adverbs are the head of everything in their own clause and are an \udtag{advcl} dependency of the verb they are subordinate to.}\label{fig:gpr}
 \end{figure}

### 2015-06-21 2015-turklang-kazdep/paper.tex
and this
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60947 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
+\begin{figure}[htbp]
+	\centering
+	\begin{dependency}[theme = simple, font = \small]
+		\begin{deptext}[column sep=0.08cm]
+			Ол \& терезеден \& Азамат \& пен \& Айгүлдің \& ойнағанына \& қарап \& тұр. \\
+			\gmk{prn} \& \gmk{n.abl} \& \gmk{np.nom} \& \gmk{cc} \& \gmk{np.gen} \& \gmk{v.iv.ger.px3sp.dat} \& \gmk{v.tv.prc} \& \gmk{vaux.pres} \\
+			\tgloss{she} \& \tgloss{from the window} \& \tgloss{Azamat} \& \tgloss{and} \& \tgloss{Aygül's} \& \tgloss{playing} \& \tgloss{watching} \& \tgloss{is} \\
+		\end{deptext}
+		\depedge{5}{3}{\udlabel{conj}}
+		\depedge{5}{4}{\udlabel{cc}}
+		\depedge{6}{5}{\udlabel{subj}}
+		\depedge{7}{6}{\udlabel{ccomp}}
+		\deproot[edge unit distance=1.3ex]{7}{\udlabel{root}}
+	\end{dependency}
+	\caption{Verbal nouns: verbal nouns are the head of everything in their own clause and can be a \udtag{ccomp} dependency of the verb they are the complement of.
+Note here that the subject noun phrase is in genitive case.}\label{fig:gpr}
+\end{figure}
+
 Бразилия өз жерінде чемпионатты екі рет өткізген бесінші ел болды

### 2015-06-21 2015-turklang-kazdep/paper.tex
Verbal nouns with case/postpositions
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60948 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 	\caption{Verbal nouns: verbal nouns are the head of everything in their own clause and can be a \udtag{ccomp} dependency of the verb they are the complement of.
-Note here that the subject noun phrase is in genitive case.}\label{fig:gpr}
+Note here that the subject noun phrase of the subordinate clause is in genitive case.}\label{fig:gpr}
+\end{figure}
+
+\begin{figure}[htbp]
+	\centering
+	\begin{dependency}[theme = simple, font = \small]
+		\begin{deptext}[column sep=0.08cm]
+			Айгүл \& санап \& біткеннен \& кейін \& айналасына \& қарады. \\
+			\gmk{n.nom} \& \gmk{v.iv.prc} \& \gmk{v.ger.abl} \& \gmk{post} \& \gmk{n.px3sp.dat} \& \gmk{v.tv.ifi} \\
+			\tgloss{Aygül} \& \tgloss{counting} \& \tgloss{having finished} \& \tgloss{after} \& \tgloss{to her surroundings} \& \tgloss{looked} \\
+		\end{deptext}
+		\depedge{5}{2}{\udlabel{advcl}}
+		\depedge{2}{3}{\udlabel{aux}}
+		\depedge{3}{4}{\udlabel{post}}
+		\depedge{6}{5}{\udlabel{obj}}
+		\deproot[edge unit distance=1.3ex]{6}{\udlabel{root}}
+	\end{dependency}
+	\caption{Verbal nouns with case/postpositions: verbal nouns with certain case morphology and/or postpositions are the head of everything in their own clause and can be an \udtag{advcl} dependency of the verb they are the complement of.}\label{fig:gpr}
 \end{figure}

### 2015-06-21 2015-turklang-kazdep/paper.tex
updates
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60949 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-\begin{figure}
-\begin{tiny}
-\begin{alltt}
-"<Елде>"
-        "ел" n loc
-"<ағылшын>"
-        "ағылшын" n nom
-"<үстемдігіне>"
-        "үстемдік" n px3sp dat
-"<қарсы>"
-        "қарсы" post
-"<жаңа>"
-        "жаңа" adj
-"<толқулар>"
-        "толқу" n pl nom
-"<басталды>"
-        "баста" v tv pass ifi p3 pl
-"<.>"
-        "." sent
-\end{alltt}
-\end{tiny}
-\caption{Example output after preprocessing and manual disambiguation. Each token is annotated for lemma,
-   part of speech and morphosyntactic information.}
-\end{figure}
+%\begin{figure}
+%\begin{tiny}
+%\begin{alltt}
+%"<Елде>"
+%        "ел" n loc
+%"<ағылшын>"
+%        "ағылшын" n nom
+%"<үстемдігіне>"
+%        "үстемдік" n px3sp dat
+%"<қарсы>"
+%        "қарсы" post
+%"<жаңа>"
+%        "жаңа" adj
+%"<толқулар>"
+%        "толқу" n pl nom
+%"<басталды>"
+%        "баста" v tv pass ifi p3 pl
+%"<.>"
+%        "." sent
+%\end{alltt}
+%\end{tiny}
+%\caption{Example output after preprocessing and manual disambiguation. Each token is annotated for lemma,
+%   part of speech and morphosyntactic information.}
+%\end{figure}
 
@@
 
-The copula (both \emph{е} and \emph{бол}) is a challenging problem for dependency 
+The copula (both \emph{е-} and \emph{бол-}) is a challenging problem for dependency 
 analysis of Kazakh. The universal dependency guidelines state that the copula should
@@
 			\gmk{np.nom} \& \gmk{det} \& \gmk{nom.px3sp.loc} \& \gmk{n.acc} \& \gmk{num} \& \gmk{n} \& \gmk{v.tv.gpr} \& \gmk{num.ord} \& \gmk{n.nom} \& \gmk{v.iv.ifi} \\
-			\tgloss{Brasil} \& \tgloss{own} \& \tgloss{in its land} \& \tgloss{championship} \& \tgloss{two} \& \tgloss{times} \& \tgloss{having held} \& \tgloss{fifth} \& \tgloss{country} \& \tgloss{became} \\
+			\tgloss{Brazil} \& \tgloss{own} \& \tgloss{in its land} \& \tgloss{championship} \& \tgloss{two} \& \tgloss{times} \& \tgloss{having held} \& \tgloss{fifth} \& \tgloss{country} \& \tgloss{became} \\
 		\end{deptext}

### 2015-06-21 2015-turklang-kazdep/paper.tex
aux fix
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60950 72bbbca6-d526-0410-a7d9-f06f51895060
tree fix
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60951 72bbbca6-d526-0410-a7d9-f06f51895060
nmod instead of obj
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60952 72bbbca6-d526-0410-a7d9-f06f51895060
now
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60953 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 			Айгүл \& санап \& біткеннен \& кейін \& айналасына \& қарады. \\
-			\gmk{n.nom} \& \gmk{v.iv.prc} \& \gmk{v.ger.abl} \& \gmk{post} \& \gmk{n.px3sp.dat} \& \gmk{v.tv.ifi} \\
+			\gmk{n.nom} \& \gmk{v.iv.prc} \& \gmk{vaux.ger.abl} \& \gmk{post} \& \gmk{n.px3sp.dat} \& \gmk{v.tv.ifi} \\
 			\tgloss{Aygül} \& \tgloss{counting} \& \tgloss{having finished} \& \tgloss{after} \& \tgloss{to her surroundings} \& \tgloss{looked} \\
 		\end{deptext}
-		\depedge{5}{2}{\udlabel{advcl}}
+		\depedge{6}{2}{\udlabel{advcl}}
 		\depedge{2}{3}{\udlabel{aux}}
 		\depedge{3}{4}{\udlabel{post}}
-		\depedge{6}{5}{\udlabel{obj}}
+		\depedge{6}{5}{\udlabel{nmod}}
 		\deproot[edge unit distance=1.3ex]{6}{\udlabel{root}}
@@
 \end{figure}
-
-Бразилия өз жерінде чемпионатты екі рет өткізген бесінші ел болды
-
-1960 жылдан бастап Батыс үлгісінде дами бастады.
-
-Ол терезеден Азамат пен Айгүлдің ойнағанына қарап тұр.
-
-Айгүл санап біткеннен кейін айналасына қарады.
 

### 2015-06-21 2015-turklang-kazdep/paper.tex
a little prose
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60954 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
+As discussed in section~\ref{sec:kazakh}, Kazakh makes extensive use of non-finite clauses, including verbal adjective clauses, verbal noun clauses, and verbal adverb clauses.
+
+Verbal adjective clauses modify a head nominal, effectively allowing a whole verb phrase to act as an adjective.
+The dependency relation between them is \udtag{acl}, per UD documentation.
+An example is provided in figure~\ref{fig:gpr}
+
 \begin{figure}[htbp]
@@
 	\end{dependency}
-	\caption{Verbal adverbs: verbal adverbs are the head of everything in their own clause and are an \udtag{advcl} dependency of the verb they are subordinate to.}\label{fig:gpr}
+	\caption{Verbal adverbs: verbal adverbs are the head of everything in their own clause and are an \udtag{advcl} dependency of the verb they are subordinate to.}\label{fig:gna}
 \end{figure}
@@
 	\caption{Verbal nouns: verbal nouns are the head of everything in their own clause and can be a \udtag{ccomp} dependency of the verb they are the complement of.
-Note here that the subject noun phrase of the subordinate clause is in genitive case.}\label{fig:gpr}
+Note here that the subject noun phrase of the subordinate clause is in genitive case.}\label{fig:gerdat}
 \end{figure}
@@
 	\end{dependency}
-	\caption{Verbal nouns with case/postpositions: verbal nouns with certain case morphology and/or postpositions are the head of everything in their own clause and can be an \udtag{advcl} dependency of the verb they are the complement of.}\label{fig:gpr}
+	\caption{Verbal nouns with case/postpositions: verbal nouns with certain case morphology and/or postpositions are the head of everything in their own clause and can be an \udtag{advcl} dependency of the verb they are the complement of.}\label{fig:gerabl}
 \end{figure}

### 2015-06-21 2015-turklang-kazdep/paper.tex
verbal nouns
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60955 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
+Verbal adverb clauses in Kazakh act as a clausal verbal adjunct, essentially allowing a whole verb phrase to act as an adverb.
+The dependency relation between the verbal adverb and the head verb is \udtag{advcl}, per UD documentation.
+An example is provided in figure~\ref{fig:gna}.
+
 \begin{figure}[htbp]
@@
 \end{figure}
+
+Verbal nouns allow a whole verb phrase to be used as a nominal, and the resulting verbal noun phrase may be used as a subject or object of a verb.
+When used as the subject of a verb, verbal noun phrases receive the UD label \udtag{subj} and when the object of a verb \udtag{ccomp}.
+An example of a verbal noun phrase used as a \udtag{ccomp} dependency of the main verb is given in table~\ref{fig:gerdat}.
 

### 2015-06-21 2015-turklang-kazdep/paper.tex
verbal nouns ad advcl
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60956 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 Verbal nouns allow a whole verb phrase to be used as a nominal, and the resulting verbal noun phrase may be used as a subject or object of a verb.
-When used as the subject of a verb, verbal noun phrases receive the UD label \udtag{subj} and when the object of a verb \udtag{ccomp}.
+When used as the subject of a verb, verbal noun phrases receive the UD label \udtag{subj}.
+When used as the object of a verb, verbal noun phrases receive the UD label \udtag{ccomp}, unless the subject of the subordinate clause is obligatorily identical to the subject of the main clause, in which case it receives \udtag{xcomp}.
 An example of a verbal noun phrase used as a \udtag{ccomp} dependency of the main verb is given in table~\ref{fig:gerdat}.
@@
 \end{figure}
+
+As mentioned earlier, verbal nouns combined with certain case morphology and/or postpositions can function as an adverbial adjunct to a main clause, similar to verbal adverb clauses.
+In such instances, they receive the UD label \udtag{advcl}, and example of which is given in figure~\label{fig:gerabl}.
 

### 2015-06-21 2015-turklang-kazdep/paper.tex
a few mistakes fixed
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60957 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 When used as the object of a verb, verbal noun phrases receive the UD label \udtag{ccomp}, unless the subject of the subordinate clause is obligatorily identical to the subject of the main clause, in which case it receives \udtag{xcomp}.
-An example of a verbal noun phrase used as a \udtag{ccomp} dependency of the main verb is given in table~\ref{fig:gerdat}.
+An example of a verbal noun phrase used as a \udtag{ccomp} dependency of the main verb is given in figure~\ref{fig:gerdat}.
 
@@
 As mentioned earlier, verbal nouns combined with certain case morphology and/or postpositions can function as an adverbial adjunct to a main clause, similar to verbal adverb clauses.
-In such instances, they receive the UD label \udtag{advcl}, and example of which is given in figure~\label{fig:gerabl}.
+In such instances, they receive the UD label \udtag{advcl}, and example of which is given in figure~\ref{fig:gerabl}.
 

### 2015-06-21 2015-turklang-kazdep/paper.tex
some of table 1
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60958 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 	\begin{small}
-		\begin{tabular}{l p{5em} p{8em} p{22em}}
+		\begin{tabular}{l l p{8em} p{27.15em}}
 			\toprule
@@
 										indefinite genitival 
-									& hargle bargle \newline
-										jargal barble \newline 
-										\\
+									& \kazakh{\textbf{Дәрігер} үйді көреді.} \gloss{\textbf{The doctor} sees the house.} \newline
+										\kazakh{\textbf{қонақ} үй} \gloss{hotel (lit., \textbf{guest} house)} \newline 
+										\kazakh{Дәрігер \textbf{үй} көреді.} \gloss{The doctor sees \textbf{a house}.}\newline
+										\kazakh{\textbf{үй} жануарлары} \gloss{\textbf{house} animals} \\\midrule
 				accusative & -/NI/ & definite object\newline
 											embedded subject\newline
-											causative subject
-										& foo \newline
-											bar \\
+											intr.\ causative subj.
+										& \kazakh{Дәрігер \textbf{үйді} көреді.} \gloss{The doctor sees \textbf{the house}.} \newline
+											bar \\\midrule
 				genitive & -/NIң/ & definite genitival\newline
-											embedded subject \\
+											embedded subject
+										& \kazakh{\textbf{үйдің} жануарлары} \gloss{the animals of \textbf{the house}} \newline
+											\kazakh{Ол \textbf{Айгүлдің} ойнағанына қарап тұр.} \gloss{She's watching \textbf{Aygül} play.} \\\midrule
 				locative & -/DA/ & adverbial
-										& \\
+										& \\\midrule
 				ablative & -/DAн/ & adverbial,\newline
-											comparator \\
-				dative & -/GA/ & indirect object,\newline
-										attributive
-									& \\
+											comparator \\\midrule
+				dative & -/GA/ & indirect object\newline
+										attributive\newline
+										tr.\ causative subj.
+									& \\\midrule
+				instrumental & -/Mен/ &  \\
 			\bottomrule

### 2015-06-21 2015-turklang-kazdep/paper.tex
table 2 cleanup
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60959 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 							\kazakh{\textbf{қонақ} үй} \gloss{hotel (lit., \textbf{guest} house)}\newline
-							\kazakh{бір \textbf{рет}} \gloss{one \textbf{time}} \\
+							\kazakh{бір \textbf{рет}} \gloss{one \textbf{time}} \\\midrule
 			Adjectives & Attributive\newline
@@
 						\kazakh{Оны \textbf{жақсы} танимын.} \gloss{I know him/her \textbf{well}.}
-						\\
+						\\\midrule
 			Adverbs & Adverbial

### 2015-06-21 2015-turklang-kazdep/paper.tex
no more table 4
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60960 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-\begin{table}[htbp]
-        \centering
-\begin{small}
-        \begin{tabular}{ll|ll|ll}
-                \toprule
-                \textbf{Label} & \textbf{Description} & \textbf{Label} & \textbf{Description} & \textbf{Label} & \textbf{Description} \\
-                \midrule
-\texttt{acl} & clausal modifier of noun & \texttt{\sout{csubjpass}} & clausal passive subject & \texttt{\sout{neg}} & negation modifier\\
-\texttt{advcl} & adverbial clause modifier & \texttt{dep} & unspecified dependency & \texttt{nmod} & nominal modifier\\
-\texttt{advmod} & adverbial modifier & \texttt{det} & determiner & \texttt{\sout{nsubj}} & nominal subject\\
-\texttt{amod} & adjectival modifier & \texttt{discourse} & discourse element & \texttt{\sout{nsubjpass}} & passive nominal subject\\
-\texttt{appos} & appositional modifier & \texttt{\sout{dislocated}} & dislocated elements & \texttt{nummod} & numeric modifier\\
-\texttt{aux} & auxiliary & \texttt{dobj} & direct object & \texttt{parataxis} & parataxis\\
-\texttt{\sout{auxpass}} & passive auxiliary & \texttt{\sout{expl}} & expletive & \texttt{punct} & punctuation\\
-\texttt{case} & case marking & \texttt{foreign} & foreign words & \texttt{remnant} & remnant in ellipsis\\
-\texttt{cc} & coordinating conjunction & \texttt{\sout{goeswith}} & goes with & \texttt{\sout{reparandum}} & overridden disfluency\\
-\texttt{ccomp} & clausal complement & \texttt{iobj} & indirect object & \texttt{root} & root\\
-\texttt{compound} & compound & \texttt{list} & list & \texttt{vocative} & vocative\\
-\texttt{conj} & conjunct & \texttt{mark} & marker & \texttt{xcomp} & open clausal complement\\
-\texttt{cop} & copula & \texttt{\sout{mwe}} & multi-word expression & \\
-\texttt{\sout{csubj}} & clausal subject & \texttt{name} & name & \\
+%\begin{table}[htbp]
+        %\centering
+%\begin{small}
+        %\begin{tabular}{ll|ll|ll}
+                %\toprule
+                %\textbf{Label} & \textbf{Description} & \textbf{Label} & \textbf{Description} & \textbf{Label} & \textbf{Description} \\
+                %\midrule
+%\texttt{acl} & clausal modifier of noun & \texttt{\sout{csubjpass}} & clausal passive subject & \texttt{\sout{neg}} & negation modifier\\
+%\texttt{advcl} & adverbial clause modifier & \texttt{dep} & unspecified dependency & \texttt{nmod} & nominal modifier\\
+%\texttt{advmod} & adverbial modifier & \texttt{det} & determiner & \texttt{\sout{nsubj}} & nominal subject\\
+%\texttt{amod} & adjectival modifier & \texttt{discourse} & discourse element & \texttt{\sout{nsubjpass}} & passive nominal subject\\
+%\texttt{appos} & appositional modifier & \texttt{\sout{dislocated}} & dislocated elements & \texttt{nummod} & numeric modifier\\
+%\texttt{aux} & auxiliary & \texttt{dobj} & direct object & \texttt{parataxis} & parataxis\\
+%\texttt{\sout{auxpass}} & passive auxiliary & \texttt{\sout{expl}} & expletive & \texttt{punct} & punctuation\\
+%\texttt{case} & case marking & \texttt{foreign} & foreign words & \texttt{remnant} & remnant in ellipsis\\
+%\texttt{cc} & coordinating conjunction & \texttt{\sout{goeswith}} & goes with & \texttt{\sout{reparandum}} & overridden disfluency\\
+%\texttt{ccomp} & clausal complement & \texttt{iobj} & indirect object & \texttt{root} & root\\
+%\texttt{compound} & compound & \texttt{list} & list & \texttt{vocative} & vocative\\
+%\texttt{conj} & conjunct & \texttt{mark} & marker & \texttt{xcomp} & open clausal complement\\
+%\texttt{cop} & copula & \texttt{\sout{mwe}} & multi-word expression & \\
+%\texttt{\sout{csubj}} & clausal subject & \texttt{name} & name & \\
 
-                \bottomrule
-        \end{tabular}
-\end{small}
-        \caption{Universal dependency label set}
-\end{table}
+                %\bottomrule
+        %\end{tabular}
+%\end{small}
+        %\caption{Universal dependency label set}
+%\end{table}
 

### 2015-06-21 2015-turklang-kazdep/paper.tex
more table 1
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60962 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 				locative & -/DA/ & adverbial
-										& \\\midrule
-				ablative & -/DAн/ & adverbial,\newline
-											comparator \\\midrule
+										& \kazakh{\textbf{Үйде} ұйықтап жатыр.} \gloss{S/he's sleeping \textbf{in the house}.} \\\midrule
+				ablative & -/DAн/ & adverbial\newline
+											comparator 
+										& \kazakh{Дәрігер \textbf{үйден} шықты.} \gloss{The doctor came out \textbf{of the house}.}\newline
+											\kazakh{\textbf{түйеден} үлкен} \gloss{bigger \textbf{than a camel}}\\\midrule
 				dative & -/GA/ & indirect object\newline

### 2015-06-21 2015-turklang-kazdep/paper.tex
examples
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60963 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 										attributive\newline
-										tr.\ causative subj.
-									& \\\midrule
-				instrumental & -/Mен/ &  \\
+										tr.\ causative subj.\newline
+										nom.\ complement \newline
+									& \\
+				instrumental & -/Mен/ &  adverbial & \kazakh{Олар \textbf{күшікпен} ойнап тұр.} \gloss{They're playing \textbf{with the dog}.} \\
 			\bottomrule

### 2015-06-21 2015-turklang-kazdep/paper.tex
oop
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60964 72bbbca6-d526-0410-a7d9-f06f51895060

--- 2015-06-21 2015-turklang-kazdep/paper.tex
@@
 										attributive\newline
-										tr.\ causative subj.
-									& \\\midrule
-				instrumental & -/Mен/ &  \\
+										tr.\ causative subj.\newline
+										nom.\ complement
+									& \\
+				instrumental & -/Mен/ &  adverbial & \kazakh{Олар \textbf{күшікпен} ойнап тұр.} \gloss{They're playing \textbf{with the dog}.} \\
 			\bottomrule

### 2015-06-21 2015-turklang-kazdep/paper.tex
complete table?
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60965 72bbbca6-d526-0410-a7d9-f06f51895060

--- 2015-06-21 2015-turklang-kazdep/paper.tex
@@
 										\kazakh{\textbf{үй} жануарлары} \gloss{\textbf{house} animals} \\\midrule
-				accusative & -/NI/ & definite object\newline
-											embedded subject\newline
-											intr.\ causative subj.
-										& \kazakh{Дәрігер \textbf{үйді} көреді.} \gloss{The doctor sees \textbf{the house}.} \newline
-											bar \\\midrule
+				accusative & -/NI/ & definite object %\newline
+										%	embedded subject\newline
+										%	intr.\ causative subj.
+										& \kazakh{Дәрігер \textbf{үйді} көреді.} \gloss{The doctor sees \textbf{the house}.}
+										%	bar
+										\\\midrule
 				genitive & -/NIң/ & definite genitival\newline
@@
 				dative & -/GA/ & indirect object\newline
-										attributive\newline
-										tr.\ causative subj.
-									& \\\midrule
-				instrumental & -/Mен/ &  \\
+										adverbial\newline
+									%	attributive\newline
+										tr.\ causative subj.%\newline
+									%	nom.\ complement
+									& \kazakh{\textbf{Ініме} кітап бердім.} \gloss{I gave \textbf{my younger brother} a book.} \newline
+										\kazakh{\textbf{Тойға} көп кісі келіпті.} \gloss{A lot of people came \textbf{to the feast.}}\newline
+										\kazakh{\textbf{Балаға} әнді тыңдаттым.} \gloss{I made \textbf{the child} listen to the song.}
+										\\\midrule
+				instrumental & -/Mен/ &  adverbial & \kazakh{Олар \textbf{күшікпен} ойнап тұр.} \gloss{They're playing \textbf{with the dog}.} \\
 			\bottomrule

### 2015-06-21 2015-turklang-kazdep/paper.tex
morphs
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60966 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 			\toprule
-				\textbf{Case} & \textbf{Morpheme} & \textbf{Functions} & \textbf{Examples} \\
+				\textbf{Case} & \textbf{Morph} & \textbf{Functions} & \textbf{Examples} \\
 			\midrule
-				nominative & - & subject\newline
+				nom. & - & subject\newline
 										attributive\newline
@@
 										\kazakh{\textbf{үй} жануарлары} \gloss{\textbf{house} animals} \\\midrule
-				accusative & -/NI/ & definite object %\newline
+				acc. & -/NI/ & definite object %\newline
 										%	embedded subject\newline
@@
 										\\\midrule
-				genitive & -/NIң/ & definite genitival\newline
+				gen. & -/NIң/ & definite genitival\newline
 											embedded subject
@@
 											\kazakh{Ол \textbf{Айгүлдің} ойнағанына қарап тұр.} \gloss{She's watching \textbf{Aygül} play.} \\\midrule
-				locative & -/DA/ & adverbial
+				loc. & -/DA/ & adverbial
 										& \kazakh{\textbf{Үйде} ұйықтап жатыр.} \gloss{S/he's sleeping \textbf{in the house}.} \\\midrule
-				ablative & -/DAн/ & adverbial\newline
+				abl. & -/DAн/ & adverbial\newline
 											comparator 
@@
 											\kazakh{\textbf{түйеден} үлкен} \gloss{bigger \textbf{than a camel}}\\\midrule
-				dative & -/GA/ & indirect object\newline
+				dat. & -/GA/ & indirect object\newline
 										adverbial\newline
@@
 										\\\midrule
-				instrumental & -/Mен/ &  adverbial & \kazakh{Олар \textbf{күшікпен} ойнап тұр.} \gloss{They're playing \textbf{with the dog}.} \\
+				inst. & -/Mен/ &  adverbial & \kazakh{Олар \textbf{күшікпен} ойнап тұр.} \gloss{They're playing \textbf{with the dog}.} \\
 			\bottomrule

### 2015-06-21 2015-turklang-kazdep/paper.tex
ups
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60967 72bbbca6-d526-0410-a7d9-f06f51895060
corresponding author and misc
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60968 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \address{$^a$ HSL-fakultehta, UiT Norgga árktalaš universitehta, N-9015 Tromsø, Norway\\
-	$^b$ Departments of Linguistics and Central Eurasian Studies, Indiana University, Bloomington, IN 47405, USA\footnote{Corresponding author.  E-mail address: jonwashi@indiana.edu}}
+	$^b$ Departments of Linguistics and Central Eurasian Studies, Indiana University, Bloomington, IN 47405, USA¹\footnote{Corresponding author.  E-mail address: jonwashi@indiana.edu}}
 
@@
 
+\footnotetext[1]{Corresponding author.  E-mail address: jonwashi@indiana.edu}\addtocounter{footnote}{1}
 This article describes work towards the development of a dependency treebank for Kazakh, a Turkic language spoken in Central Asia and Europe.
-Despite its status as a \emph{core}\footnote{FIXME} Turkic %FIXME
+Despite its status as a core Turkic
 language, little computational-linguistic research has been published on syntactic
@@
 such as machine translation and computer-aided language learning.
-
 According to \textcite{nivre08}, a parser trained on a treebank of only 1,500 sentences 

### 2015-06-21 2015-turklang-kazdep/paper.tex
sources
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60970 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
+To create the corpus we selected a range of texts from free and public-domain sources. These texts encompassed 
+texts from a variety of genres, such as: encyclopaedic articles, folk tales, legal texts, and phrases from a phrasebook.
+The corpus is not entirely balanced and leans more towards encyclopaedic text. Table~\ref{table:corpcomp} provides
+a breakdown of the sources.
+
 \begin{table}[htbp]
@@
   \caption{Composition of the corpus. The corpus covers a range of genres and text types from free and public-domain sources.}
+  \label{table:corpcomp}
 	\begin{small}

### 2015-06-21 2015-turklang-kazdep/paper.tex
fix
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60971 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 two input tokens may result in three output tokens, \kazakh{бар ма?} \gloss{is there?} would be
-tokenised as \kazakh{бар.\gmk{adj} + е.\gmk{cop.aor.sg1} + ма.\gmk{qst}}.
+tokenised as \kazakh{бар.\gmk{adj} + е.\gmk{cop.aor.sg3} + ма.\gmk{qst}}.
 

### 2015-06-21 2015-turklang-kazdep/paper.tex
fonts
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60972 72bbbca6-d526-0410-a7d9-f06f51895060
font stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60973 72bbbca6-d526-0410-a7d9-f06f51895060
fix gloss
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60974 72bbbca6-d526-0410-a7d9-f06f51895060
minor mistake
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60975 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{tikz-dependency}
-\newcommand{\gmk}[1]{{\rm {\sc \texttt{#1}}}}
+%\newcommand{\gmk}[1]{{\rm {\sc \texttt{#1}}}}
+\newcommand{\gmk}[1]{{\rm {\ll \textsc{#1}}}}
 \newcommand{\kazakh}[1]{{\em #1}}
@@
 \newcommand{\tgloss}[1]{{\em #1}}
-\newcommand{\tag}[1]{\texttt{‹#1›}}
+%\newcommand{\tag}[1]{\texttt{‹#1›}}
+\newcommand{\tag}[1]{{\ll \textsc{‹#1›}}}
 %\newcommand{\udtag}[1]{\texttt{@#1}}
@@
 											embedded subject
-										& \kazakh{\textbf{үйдің} жануарлары} \gloss{the animals of \textbf{the house}} \newline
+										& \kazakh{\textbf{үйдің} жануарлары} \gloss{the animals \textbf{of the house}} \newline
 											\kazakh{Ол \textbf{Айгүлдің} ойнағанына қарап тұр.} \gloss{She's watching \textbf{Aygül} play.} \\\midrule
@@
                                 \gmk{np} \& \gmk{n} \& \gmk{aor} \\
-                                \gloss{Aygül} \& \gloss{student} \& \gloss{is} \\
+                                \tgloss{Aygül} \& \tgloss{student} \& \tgloss{is} \\
                         \end{deptext}
@@
                                 \gmk{np} \& \gmk{gen} \& \gmk{adv} \& \gmk{cop} \& \gmk{neg.aor} \\
-                                \gloss{Aygül} \& \gloss{he} \& \gloss{where} \& \gloss{being} \& \gloss{knows not} \\
+                                \tgloss{Aygül} \& \tgloss{he} \& \tgloss{where} \& \tgloss{being} \& \tgloss{knows not} \\
                         \end{deptext}

### 2015-06-21 2015-turklang-kazdep/paper.tex
typo fix
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60976 72bbbca6-d526-0410-a7d9-f06f51895060
minor fixes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60977 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 	\begin{small}
-		\begin{tabular}{l l p{8em} p{27.15em}}
+		\begin{tabular}{l l p{9em} p{27.15em}}
 			\toprule
@@
 									%	attributive\newline
-										tr.\ causative subj.%\newline
+										trans.\ causative subj.%\newline
 									%	nom.\ complement
@@
 Interestingly, many adjectives can also be used substantively, i.e., as nominals.
-Table~\ref{tab:zeroderiv} shows the various functions that of nominals, adjectives, and adverbs may take.
+Table~\ref{tab:zeroderiv} shows the various functions that nominals, adjectives, and adverbs may take.
 
@@
 							Adverbial
-						& \kazakh{\textbf{Қонақ}тарымыз бай.} \gloss{Our \textbf{guest}s are rich.}\newline 
+						& \kazakh{\textbf{Қонақтарымыз} бай.} \gloss{\textbf{Our guests} are rich.}\newline 
 							\kazakh{\textbf{қонақ} үй} \gloss{hotel (lit., \textbf{guest} house)}\newline
-							\kazakh{бір \textbf{рет}} \gloss{one \textbf{time}} \\\midrule
+							\kazakh{Оны бір \textbf{рет} көрдім.} \gloss{I saw it one \textbf{time}} \\\midrule
 			Adjectives & Attributive\newline

### 2015-06-21 2015-turklang-kazdep/paper.tex
gloss
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60978 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 							\kazakh{\textbf{қонақ} үй} \gloss{hotel (lit., \textbf{guest} house)}\newline
-							\kazakh{Оны бір \textbf{рет} көрдім.} \gloss{I saw it one \textbf{time}} \\\midrule
+							\kazakh{Оны бір \textbf{рет} көрдім.} \gloss{I saw him/her/it one \textbf{time}.} \\\midrule
 			Adjectives & Attributive\newline

### 2015-06-21 2015-turklang-kazdep/paper.tex
gloss
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60979 72bbbca6-d526-0410-a7d9-f06f51895060
fixing Fran's wording and stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60980 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 						& \kazakh{\textbf{жақсы} үй} \gloss{\textbf{nice} house} \newline
-						\kazakh{ең \textbf{жақсы}лары} \gloss{\textbf{the nice}st \textbf{ones}} \newline
+						\kazakh{ең \textbf{жақсы}лары} \gloss{\textbf{the nice}st \textbf{one}s} \newline
 						\kazakh{Оны \textbf{жақсы} танимын.} \gloss{I know him/her \textbf{well}.}
@@
 
-Another category that may also be non-overt is the copula.
+Another category that may be non-overt is the copula.
 The primary strategy for copula constructions in Kazakh is the use of a defective verb \kazakh{е-}.
@@
 While this construction never has overt copula marking, copula clauses which are complements of verbs always have an overt copula form, e.g.
-\kazakh{Дәрігердің үйі жақсы \textbf{екенін} білген жоқпын.} \gloss{I didn't know \textbf{that} the doctor's house \textbf{was} nice.}  Here, \kazakh{екен} is a suppletive gerundive form of the copula and is marked as accusative.
+\kazakh{Дәрігердің үйі жақсы \textbf{екенін} білген жоқпын.} \gloss{I didn't know \textbf{that} the doctor's house \textbf{was} nice.}  Here, \kazakh{екен} is a suppletive verbal noun form of the copula and is marked as accusative.
 
@@
 A treebank is a parsed corpus of sentences annotated syntactically following a particular
-syntactic theory. Two broad groups can be distinguished, phrase-structure treebanks
+syntactic theory. Two broad groups can be distinguished: phrase-structure treebanks
 which annotate constituency strucure, and dependency treebanks which annotate dependency
@@
 
-Treebanks can be used directly for linguistic and computational linguistic research, by
-performing search queries. For example, to extract a valency lexicon for verbs, or to
-study the frequency of various syntactic phenomena such as word order or nominal case usage
+Treebanks can be used directly for linguistic and computational linguistic research by
+performing search queries---for example, to extract a valency lexicon for verbs, or to study the frequency of various syntactic phenomena such as word order or nominal case usage
 and syntactic function.
@@
 
-To create the corpus we selected a range of texts from free and public-domain sources. These texts encompassed 
-texts from a variety of genres, such as: encyclopaedic articles, folk tales, legal texts, and phrases from a phrasebook.
+To create the corpus, we selected a range of texts from free and public-domain sources. These texts encompassed 
+texts from a variety of genres, such as encyclopaedic articles, folk tales, legal texts, and phrases from a phrasebook.
 The corpus is not entirely balanced and leans more towards encyclopaedic text. Table~\ref{table:corpcomp} provides
@@
 into a single token. Abbreviations and numerals which bear case, such as \kazakh{АҚШ-пен} \gloss{with the USA}
-and \kazakh{90\%-ына} \gloss{to 90\%} are analysed as a single token, as are light-verbs such as
+and \kazakh{90\%-ына} \gloss{to 90\%} are analysed as a single token, as are light verb constructions such as
 \kazakh{пайда бол} \gloss{to appear} and tense forms written with space like \kazakh{оқыған жоқ},
@@
 In some cases a single token is split into two tokens, as with the aorist copula suffixes,
-\kazakh{үйдемін} \gloss{I am in the house} is tokenised as \kazakh{үй.\gmk{loc} + е.\gmk{cop.aor.sg1}}. Furthermore
-two input tokens may result in three output tokens, \kazakh{бар ма?} \gloss{is there?} would be
+e.g., \kazakh{үйдемін} \gloss{I am in the house} is tokenised as \kazakh{үй.\gmk{loc} + е.\gmk{cop.aor.sg1}}. Furthermore,
+two input tokens may result in three output tokens, e.g., \kazakh{бар ма?} \gloss{is there?} is
 tokenised as \kazakh{бар.\gmk{adj} + е.\gmk{cop.aor.sg3} + ма.\gmk{qst}}.
@@
 the copula is part of an embedded clause this is not necessarily the most  
-effective choice. In \ref{fig:cop3}, the copula holds all the morphological information,
+effective choice. In \ref{fig:cop3}, the copula holds all the morphological information, including
 agreement with the subject and accusative marking for the embedded clause.

### 2015-06-21 2015-turklang-kazdep/paper.tex
how about that
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60981 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 be the dependent of the lexical predicate. However, in many cases the copula in Kazakh
-is found in embedded clauses ...
+is found in embedded clauses, which morphologically acts much more like the head of the embedded clause.
 
@@
                                 \gmk{np} \& \gmk{gen} \& \gmk{adv} \& \gmk{cop} \& \gmk{neg.aor} \\
-                                \tgloss{Aygül} \& \tgloss{he} \& \tgloss{where} \& \tgloss{being} \& \tgloss{knows not} \\
+                                \tgloss{Aygül} \& \tgloss{his} \& \tgloss{where} \& \tgloss{being} \& \tgloss{knows not} \\
                         \end{deptext}

### 2015-06-21 2015-turklang-kazdep/paper.tex
updas
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60982 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 last conjunct in a series. Furthermore, experiments in representing coordination
-in other head-final languages have found that the final-conjunct head analysis 
+in other predominantly head-final languages have found that the final-conjunct head analysis 
 results in better parser accuracy \parencite{bengoetxea09}.
@@
 An indefinite genitive construction is formed by an indefinite genitive nominal (indistinguishable from the bare / nominative form, and tagged with \tag{nom}) preceding a nominal that has third-person possessive morphology, as shown in \ref{fig:indefgen}.
-A definite genetive construction is formed by a genitive-marked nominal (tagged \tag{gen}) precding a nominal that has third-person possessive morphology, as shown in \ref{fig:defgen}.
+A definite genitive construction is formed by a genitive-marked nominal (tagged \tag{gen}) precding a nominal that has third-person possessive morphology, as shown in \ref{fig:defgen}.
 

### 2015-06-21 2015-turklang-kazdep/paper.tex
removed for review
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60983 72bbbca6-d526-0410-a7d9-f06f51895060
coordination prrose
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60984 72bbbca6-d526-0410-a7d9-f06f51895060
misc
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60986 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 in other predominantly head-final languages have found that the final-conjunct head analysis 
-results in better parser accuracy \parencite{bengoetxea09}.
+results in better parser accuracy \parencite{bengoetxea09}.  Figure~\ref{fig:coord} shows an example of our coordination strategy.
 
@@
 There are different ways in which two nominals may occur together to act as a single nominal.
-Compounds are formed by an attributive nominal (indistinguishable from the bare / nominative form, but tagged with \tag{attr}) preceding another nominal, as shown in \ref{fig:compound}.
-An indefinite genitive construction is formed by an indefinite genitive nominal (indistinguishable from the bare / nominative form, and tagged with \tag{nom}) preceding a nominal that has third-person possessive morphology, as shown in \ref{fig:indefgen}.
+Compounds are formed by an attributive nominal (morphologically indistinguishable from the bare / nominative form, but tagged with \tag{attr}) preceding another nominal, as shown in \ref{fig:compound}.
+An indefinite genitive construction is formed by an indefinite genitive nominal (morphologically indistinguishable from the bare / nominative form, and tagged with \tag{nom}) preceding a nominal that has third-person possessive morphology, as shown in \ref{fig:indefgen}.
 A definite genitive construction is formed by a genitive-marked nominal (tagged \tag{gen}) precding a nominal that has third-person possessive morphology, as shown in \ref{fig:defgen}.
@@
 
-\begin{figure}[htbp]
+\begin{figure}[!htbp]
 	\centering
@@
 
+\emph{removed for review}
+
 %Tolgonay?, Aida?

### 2015-06-21 2015-turklang-kazdep/paper.tex
typo
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60987 72bbbca6-d526-0410-a7d9-f06f51895060
eval section fixes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60989 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 The dependency relation between them is \udtag{acl}, per UD documentation.
-An example is provided in figure~\ref{fig:gpr}
+An example is provided in figure~\ref{fig:gpr}.
 
@@
 and evaluated a number of models using the popular MaltParser tool \parencite{nivre07}.
-MaltParser is a toolkit for data-driven dependency parsing, it can learn a parsing
+MaltParser is a toolkit for data-driven dependency parsing; it can learn a parsing
 model from treebank data and apply this model to parse unseen sentences. The parser
@@
 \begin{table}[htbp]
-	\caption{Preliminary parsing results from MaltParser using different models.
-The numbers in brackets denote the upper and lower bounds found during cross validation.
-Adding structural features to the model substantially improves the performance of the parser, although we find no improvement in using full morphosyntactic description (MSD) over using simply the first part-of-speech (POS) tag.}
+	\caption{Preliminary parsing results from MaltParser using different models. The numbers in brackets denote the upper and lower bounds found during cross-validation.}
 	\centering
@@
 
-To perform 10-fold cross validation we randomised the order of sentences in the corpus
+To perform 10-fold cross-validation we randomised the order of sentences in the corpus
 and split it into 10 equally-sized parts. In each iteration we held out one part for testing and used
@@
 
-The results we obtain are similar to those obtained with similar 
-sized treebanks, for example the Irish treebank of \textcite{Lynn12}, who report an LAS of
-63.3 and a UAS of 73.3 with the best model.
+The results we obtain, shown in table~\ref{table:eval}, are similar to those obtained with similar 
+sized treebanks, for example the Irish treebank of \textcite{Lynn12}, for which an LAS of
+63.3 and a UAS of 73.3 were reported with the best model.
+Adding structural features to the model substantially improves the performance of the parser, although we find no improvement in using full morphosyntactic description (MSD) over using simply the first part-of-speech (POS) tag.
 

### 2015-06-21 2015-turklang-kazdep/paper.tex
no crossdressing Kumyks
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60990 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 that is, applying a model learnt on the Kazakh treebank to other Turkic languages
-such as Tatar, Kumyk and Tuvan. We have morphological analysers for these languages which have
+such as Tatar, Kumyk, and Tuvan. We have morphological analysers for these languages which have
 compatible tagsets for morphological features and as such it should be possible to learn a

### 2015-06-25 2015-turklang-kazdep/paper.tex
fix typo
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61052 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 %\usepackage[backend=bibtex]{biblatex}
-\usepackage[backend=biber,citestyle=apa,style=authoryear,maxbibnames=99,maxcitenames=2]{biblatex}
+\usepackage[backend=biber,citestyle=apa,style=authoryear-comp,maxbibnames=99,maxcitenames=2]{biblatex}
 \bibliography{paper}
 \renewcommand*{\nameyeardelim}{\addcomma\space}
+
+\DeclareNameAlias{sortname}{last-first}
 
@@
 		   \begin{deptext}[column sep=0.3cm]
-				Иранның \& экономиясы \\
+				Иранның \& экономикасы \\
 				\gmk{n.gen} \& \gmk{n.px3sg.nom} \\

### 2015-06-26 2015-compmorphon-morphver/poster.tex
draft poster, including Makefile and logo graphics
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61089 72bbbca6-d526-0410-a7d9-f06f51895060

--- 2014-06-20 kypchak-morphfest-2014/poster.tex
@@
-\documentclass[fontscale=0.3,landscape,paperwidth=72in,paperheight=36in]{baposter}  % fontscale=0.285, dvipdfm
+\documentclass[fontscale=0.3,landscape,paperwidth=48in,paperheight=36in]{baposter}  % fontscale=0.285, dvipdfm
 
@@
 
-%\usepackage{tikz}
+\usepackage{tikz}
 %\usepackage{pgfbaselayers}
@@
 
-%\definecolor{bordercol}{RGB}{40,40,40}
-%\definecolor{bordercol}{RGB}{100,0,0}
+\selectcolormodel{cmyk}
+
 \definecolor{bordercol}{RGB}{0,0,0}
-%\definecolor{headercol1}{RGB}{186,215,230}
-%\definecolor{headercol2}{RGB}{80,80,80}
-%\definecolor{headercol1}{RGB}{153,0,0}
-%\definecolor{headercol2}{RGB}{153,0,0}
-%\definecolor{headercol}{RGB}{6,130,22}
 %\definecolor{headercolone}{RGB}{6,130,22}
-\definecolor{headercolone}{RGB}{6,130,22}
-\definecolor{headercoltwo}{RGB}{6,130,22}
-%\definecolor{boxcolor}{RGB}{186,215,230}
-%\definecolor{boxcolor}{RGB}{110,230,209}
-%\definecolor{boxcolor}{RGB}{143,255,236}
-%\definecolor{boxcolor}{RGB}{171,255,241}
-%definecolor{boxcolor}{RGB}{176,232,223}
-\definecolor{boxcolor}{RGB}{176,208,223}
+\definecolor{headercol}{cmyk}{1,0.64,0,0.6}
+\definecolor{headercolone}{cmyk}{1,0.64,0,0.6}
+\definecolor{headercoltwo}{cmyk}{0.44,0,0.15,0.17}
+%\definecolor{boxcolor}{RGB}{176,208,223}
+\definecolor{boxcolor}{cmyk}{0.3,0.08,0.08,0.1}
 \definecolor{headerfontcol}{RGB}{255,255,255}
-%\definecolor{boxcolor}{RGB}{186,215,230}
-%\definecolor{boxcolor}{RGB}{226,200,200}
-%\newcommand{\hilitetwo}[1]{{\addfontfeature{Color=FF000099}#1}}
 \newcommand{\hilitetwo}[1]{{\addfontfeature{Color=99333399}#1}}
-%\newcommand{\hiliteone}[1]{{\addfontfeature{Color=33993399}#1}}
 \newcommand{\hiliteone}[1]{{\addfontfeature{Color=06821699}#1}}
@@
 \setromanfont[IPA]{Liberation Serif}
-%\setromanfont[IPA]{Times New Roman}
-%\setromanfont[IPA]{NimbusRomNo9L-Regu}
-%\setromanfont{Times New Roman}
-%\setmonofont[IPA]{Liberation Mono}
 \setmonofont[IPA]{DejaVu Sans Mono}
-%\renewcommand{\sfdefault}{phv}
-%\renewcommand{\rmdefault}{ptm}
-%\renewcommand{\ttdefault}{pcr}
 \usepackage[small,bf]{caption}
-%\newfontfamily\qipa[IPA]{NimbusRomNo9L-Regu}
 \newfontfamily\qipa[IPA,Scale=MatchLowercase]{FreeSerif}
@@
 \newcommand{\blank}[1]{\underline{\hspace{#1}}}
-% FIXME: Breaks baposter
-%\usepackage[novoc,fdf2alif]{arabxetex}
-%\newfontfamily\uighurfont[Script=Arabic,Scale=1.5]{Lateef}
-%\newfontfamily\arabicfont[Script=Arabic,Scale=1.5]{Lateef}
 
@@
 
-%\makeatletter
-%\renewenvironment{thebibliography}[1][]{\htwo{\bibname}}
-%     {\section*{\bibname}% <-- this line was changed from \chapter* to \section*
-%	  \htwo{\bibname}
-%      \@mkboth{\MakeUppercase\bibname}{\MakeUppercase\bibname}%
-%      \list{\@biblabel{\@arabic\c@enumiv}}%
-%           {\settowidth\labelwidth{\@biblabel{#1}}%
-%            \leftmargin\labelwidth
-%            \advance\leftmargin\labelsep
-%            \@openbib@code
-%            \usecounter{enumiv}%
-%            \let\p@enumiv\@empty
-%            \renewcommand\theenumiv{\@arabic\c@enumiv}}%
-%      \sloppy
-%      \clubpenalty4000
-%      \@clubpenalty \clubpenalty
-%      \widowpenalty4000%
-%      \sfcode`\.\@m}
-%     {\def\@noitemerr
-%       {\@latex@warning{Empty `thebibliography' environment}}%
-%      \endlist}%}
-%}
-%\makeatother
-%\usepackage[in]{fullpage}
 \usepackage[colorlinks=true,citecolor=black,linkcolor=black,urlcolor=black]{hyperref}
@@
 \begin{document}
-	% To get it to be A0 consistently on all machines..
-	%\special{papersize=1189mm,841mm}
-	%\special{papersize=841mm,1189mm}
-	\special{papersize=72in,36in}
+	% To get it to be the same size consistently on all machines..
+	\special{papersize=48in,36in}
 	\setlength{\pdfpageheight}{\paperheight}
@@
 
-%%% Setting Background Image %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\background{
-%	\includegraphics[width=0.99\textwidth]{flagkg2}
-%}
-	\background{{
-%		\begin{tikzpicture}[remember picture,overlay]%
-%			\draw (current page.north west)+(-4em,2em) node[anchor=north west] {\includegraphics[width=1.1\textwidth,height=1.1\textheight]{flagkg2}};
-%		\end{tikzpicture}%
-%%		\includegraphics[width=0.99\textwidth]{flagkg2}
-%%			\draw (current page.north west)+(-2em,2em) node[anchor=north west] {\includegraphics[width=0.9\textwidth]{flagkg2}};
-	}}
+	%%% Setting Background Image %%%
+	%\background{}
 
@@
 		headerColorTwo=headercoltwo,
-			headerFontColor=white,
+		%headerColorOne=blue,
+		%headerFontColor=white,
 		% Only simple background color used, no shading, so boxColorTwo isn't necessary
@@
 		textborder=roundedleft,
-			columns=6,
-		}{
-			%\hspace{-2em}\includegraphics[height=6.5em,bb=0 0 242 203]{apertium}
-			%\hspace{-2em}\includegraphics[height=6.5em,bb=0 0 612 792]{apertium2}
-			%\hspace{-2em}\includegraphics[height=6.5em,bb=0 0 203 242]{apertium3}
-			%\hspace{-2em}\includegraphics[height=6.5em,bb=0 0 242 203]{apertium3a}
-			%\hspace{-2em}\includegraphics[height=6.5em,bb=0 0 242 203]{apertium4}
-			%\hspace{-2em}
-			\includegraphics[angle=90,height=6.5em]{apertium5b}%\hspace{-2em}
-			%Eye Catcher, empty if option eyecatcher=false - unused
-			%\begin{minipage}[t]{0.2\textwidth}
-				\hspace{1cm}\includegraphics[height=6.5em]{22888536}
-				\hspace{1cm}\includegraphics[height=6.5em]{iu_tab}
-				\hspace{1cm}\includegraphics[height=6.5em]{kfu}
-				\hspace{1cm}\includegraphics[height=6.5em]{uitlogo}
-			%\end{minipage}
-		}{
+		columns=4,
+	}{ %Eye Catcher, empty if option eyecatcher=false - unused
+		\includegraphics[height=6.5em]{uitlogo}
+		\hspace{2cm}\begin{minipage}[t]{9em}
+			\vspace{-2cm}
+			\noindent\includegraphics[height=4.2em]{DCU_logo_2col}\\
+			\vspace{-0.5ex}\hspace{0.105cm}\includegraphics[height=2.15em]{DCU_2010_Name_Mark_289}
+			\end{minipage}
+			\hspace{2cm}\includegraphics[height=6.5em]{iu_tab}
+	}{ %Title (centered top)
 		{\vspace{0pt}\hspace{1cm}
-			{\titlefont \textsc{Designing finite-state morphological transducers\\for Kypchak languages}}}
-		}{
-			%Jonathan North Washington {\small (Indiana University, \texttt{jonwashi@indiana.edu})}, Mirlan Ipasov {\small (International Ataturk Alatoo University, \texttt{mipasov@gmail.com})}, Francis M.\ Tyers {\small (Universitat d'Alacant, \texttt{ftyers@dlsi.ua.es})}
-
+			{\titlefont \textsc{Finite-state morphologies and text corpora\\as resources for improving morphological descriptions}}}
+	}{ %Authors (centered top, below title)
 		\vspace{-0.5em}
 		\begin{center}
-			{\begin{minipage}[t]{12em}
+			\begin{minipage}[t]{9em}
+				\begin{spacing}{0.4}
+				{Francis M.\ Tyers}\\
+				{\footnotesize UiT Norgga Árktalaš Universitehta \\\texttt{francis.tyers@uit.no}}
+				\end{spacing}
+			\end{minipage}
+			\begin{minipage}[t]{9.5em}
+				\begin{spacing}{0.4}
+					{Tommi Pirinen\vphantom{y}}\\
+					{\footnotesize Ollscoil Chathair Bhaile Átha Cliath\\\texttt{FIXME Flammie}}
+				\end{spacing}
+			\end{minipage}
+			\begin{minipage}[t]{12em}
 				\begin{spacing}{0.4}
@@
 			\end{minipage}
-			\begin{minipage}[t]{11em}
-				\begin{spacing}{0.4}
-					{Ilnar Salimzyanov}\\
-					{\footnotesize Казан (Идел буе) федераль университеты \\\texttt{ilnar.salimzyan@gmail.com}}
-				\end{spacing}
-			\end{minipage}
-			\begin{minipage}[t]{9em}
-				\begin{spacing}{0.4}
-					{Francis M.\ Tyers}\\
-					{\footnotesize UiT Norgga Árktalaš Universitehta \\\texttt{francis.tyers@uit.no}}
-				\end{spacing}
-			\end{minipage}}
-			\begin{minipage}[t]{7em}
-				\vspace{-0.7em}
-				\begin{spacing}{0.4}
-					{\footnotesize Special thanks to:}\\
-					{\small Tolgonay Kubatova}\\
-					{\small Aida Sundetova}\\
-					%{\footnotesize \texttt{sun27aida@gmail.com}}
-					%{\small Ağarahim Sultanmuradov}\\
-					{\small Ağarahim Sultanmuradov}\\
-					%{\small Google Summer of Code}\\
-					%{\small Google Code-In}
-					%{\small GSoC, GCI}
-				\end{spacing}
-			\end{minipage}
 		\end{center}
-
-			%\begin{minipage}[c]{12em}
-			%	\centering
-			%	\begin{spacing}{0.4}
-			%		Jonathan Washington\\
-			%		Indiana University\\
-			%		\texttt{jonwashi@indiana.edu}
-			%	\end{spacing}
-			%\end{minipage}
-			%\begin{minipage}[c]{18em}
-			%	\centering
-			%	\begin{spacing}{0.4}
-			%		Mirlan Ipasov\\
-			%		International Ataturk Alatoo University\\
-			%		\texttt{mipasov@gmail.com}
-			%	\end{spacing}
-			%\end{minipage}
-			%\begin{minipage}[c]{10em}
-			%	\centering
-			%	\begin{spacing}{0.4}
-			%		Francis M.\ Tyers\\
-			%		Universitat d'Alacant\\
-			%		\texttt{ftyers@dlsi.ua.es}
-			%	\end{spacing}
-			%\end{minipage}
-
-			%Jonathan Washington, Mirlan Ipasov, Francis M.\ Tyers\\
-			%Indiana University, Universitat d'Alacant, International Ataturk Alatoo University\\
-			%\texttt{jonwashi@indiana.edu}, \texttt{mipasov@gmail.com}, \texttt{ftyers@dlsi.ua.es}
-		}{
-			%University Logo(s)
-			%{\begin{minipage}{19em}
-				%\hfill
-				%\includegraphics[height=6.5em]{apertium}
-				%\includegraphics[height=6.5em]{flags/flags}
-				\includegraphics[height=6.5em]{Flag_of_Dagestan}
-				\hspace{1cm}\includegraphics[height=6.5em]{Flag_of_Tatarstan}
-				\hspace{1cm}\includegraphics[height=6.5em]{Flag_of_Kyrgyzstan}
-				\hspace{1cm}\includegraphics[height=6.5em]{Flag_of_Kazakhstan}
-			%\end{minipage}}
+	}{ %More eye catchers, on the right side
+		\includegraphics[angle=90,height=6.5em]{apertium5b}%\hspace{-2em}
 	}
 
-%		\headerbox{Overview}{name=overview,column=0,row=1}{
-%			\begin{spacing}{0.8}
-%This paper describes the development of a free/open-source finite-state morphological transducer for Kyrgyz. The transducer 
-%has been developed for morphological generation for use within a prototype Turkish$\rightarrow$Kyrgyz machine translation system, but has also been extensively tested for analysis.
-The finite-state toolkit used for the work was the Helsinki Finite-State
-%Toolkit (HFST). The paper describes some issues in Kyrgyz morphology, the development of the tool, some linguistic issues encountered and how they were dealt with, and 
-%which issues are left to resolve.
-An evaluation is presented which shows that the transducer has medium-level coverage, between 82\% and 87\% on two freely available corpora of Kyrgyz, and high precision and recall over a manually verified test set.
-%			\end{spacing}\vspace{-0.25ex}
-%		}
-
-	\headerbox{Kypchak languages}{name=kypchaklgs,column=0,row=0}{
-		\vspace{0.1ex}
-		\includegraphics[width=\textwidth]{map/map}
-		\vspace{-1.25em}  % -1.25em
-		\begin{itemize}
-			\item Turkic languages (SOV, agglutinative, vowel harmony)
-		\end{itemize}
-
+	\headerbox{Random box 1}{name=boxone,column=0,row=0}{
 		\scalebox{0.78}{
-		\begin{tabular}{lllll}
-			\toprule
-				 & \textbf{Kyrgyz} & \textbf{Kazakh} & \textbf{Tatar} & \textbf{Kumyk} \\
-				 & /{\qipa qɯɾʁɯz}/ & /{\qipa qɑzɑq}/ & /{\qipa tɒtɑɾ}/ & /{\qipa qumuq}/ \\
-				classification & Eastern & Southern & Northern & Western \\
-			\midrule
-				\multicolumn{4}{l}{population of speakers} \\
-			\midrule
-				number & 3M & 8M-12M & 5.4M & 430K \\
-				primary & Kyrgyzstan & Kazakhstan & Tatarstan & Dagestan \\
-				secondary & China, etc. & China, Mongolia & Bashqortostan & --- \\
-			\midrule
-				\multicolumn{4}{l}{external influences} \\
-			\midrule
-				Mongolic & moderate & moderate & light & light \\
-				Oghuz & — & — & light & moderate \\
-				Persian & heavy & heavy & heavy & heavy \\
-				Russian & heavy & heavy & heavy & heavy \\
-			\bottomrule
-		\end{tabular}
+			hargle
 		}
-%			\begin{itemize}
-%				\item Similar historically to Southern Altay
-%				\item Similar by convergence to Kazakh, Uzbek
-%			\end{itemize}
-%			\item Spoken in
-%			\begin{itemize}
-%				\item Kyrgyzstan, as co-official language\\
-%				\textit{high levels of bilingualism with Russian}
-%				\item China, Tajikistan, Uzbekistan
-%			\end{itemize}
-%			\item Over 3 million native speakers\\(estimate based on data from Ethnologue) %\citep{lewis2009}
-%			\item Our transducer based on written Kyrgyz of the former Soviet Union (literary and colloquial standards)
-%		\end{itemize}
 	}
 
-	\headerbox{Morphological transducers}{name=morphtrans,below=kypchaklgs}{
+	\headerbox{Morphological transducers}{name=morphtrans,below=boxone}{
 		\htwo{Morphological transducers}
@@
 
-	\headerbox{Development effort}{name=devel,column=4,row=0}{
-	\scalebox{0.92}{
-		\begin{tabular}{lllll}
-			\toprule
-			 & \textbf{Kyrgyz} & \textbf{Kazakh} & \textbf{Tatar} & \textbf{Kumyk} \\
-			\midrule
-			begun & Apr.\ 2011 & Dec.\ 2010 & Dec.\ 2011 & Oct.\ 2013 \\
-			80\% cov.\ & Aug.?\ 2011 & Aug.\ 2012 & Aug.\ 2012 & Oct.\ 2013 \\\midrule
-			time & 4 months & 19 months & 7 months & 1 week \\
-			\bottomrule
-		\end{tabular}
-	}
-		\begin{itemize}
-			%\item Kyrgyz and Kazakh transducers begun in 2011
-			%\item Kyrgyz transducer reached 80\% coverage summer of 2011
-			\item (various periods of intermission, various rewrites)
-			\item Kazakh transducer based on Kyrgyz transducer
-			\item Kyrgyz transducer currently being rewritten based on insights gained while writing other Turkic transducers
-			\item Kumyk transducer based on Kazakh, Tatar transducers:\\
-			\tilda{}1 week to reach 80\% coverage, +1 week to reach 90\%
-		\end{itemize}
-		\vspace*{-2.5em}
-		\hspace*{-2em}\includegraphics[width=0.9\textwidth,angle=270]{../kipchak-lrec-2013/graphs/kumyk-yoldash-cov}
-		\vspace*{-2em}
-	}
-
-%		\headerbox{Available Turkic Transducers}{name=turkictrans,below=morphtrans}{
-%* Other Turkic morphological transducers\\
-%		}
-%		\headerbox{Framework: HFST / Apertium}{name=framework,below=turkictrans}{
-%* HFST/Apertium
-%		}
-
-%			\htwo{Morphological \& orthographical words}
-%
-%			\begin{itemize}
-%				\item{} {\qipa өнүктүрөбүзбү ?} \eng{will we develop [it]?}\\
-%				\texttt{ өнүк{\small <v><tv><caus><aor><p1><pl>+}бы{\small <qst>}}
-%				\item{} {\qipa келатсаң} \eng{if you come}\\
-%				\texttt{кел{\tt {\small <v><iv><prt\_impf>}}+жат{\small <vaux><gna\_cnd><p2><sg>}}
-%			\end{itemize}
-%
-%
-%			\codeex{
-%			{\small
-%			\texttt{LEXICON N-INFL-3PX-COMPOUND\\
-%			\vspace{-0.5ex}\%<n\%>:\%>\%\{S\%\}\%\{I\%\}\%\{n\%\} GEN-POS ;}\\
-%			
-%			\texttt{LEXICON Nouns\\
-%			\vspace{-0.5ex}аба\%\ ырайы:аба\%\ ырай\ N-INFL-3PX-COMPOUND ;\\\vspace{-0.5ex}\hspace{0.1pt}\hfill{}!\ "weather"\\
-%			\vspace{-0.5ex}чакыруу\%\ кагазы:чакыруу\%\ кагаз\ N-INFL-3PX-\\\vspace{-0.5ex}\hspace{0.1pt}\hfill{}COMPOUND ;\ !\ "invitation"}}
-%			}
-%				
-%		}
-
-					\renewcommand{\arraystretch}{0.65}
-
-		\headerbox{Example output}{name=exout,span=2,column=1,row=0}{
-
-		\htwo{Gloss}
-
-		\pex[aboveexskip=0pt,belowexskip=3pt,aboveglftskip=0pt]\label{ex:sent}
-			\begingl
-				\gla[everygla={\qipa }] Кудай Өзү жаратканынын баарына карап, {} өтө жакшы экенин көрдү. //
-				\glb[everyglb={\qipa }] Құдай Өзінің жаратқандарының бәріне қарап, {} өте жақсы екенін көрді. //
-				\glb[everyglb={\qipa }] Аллаһ Үзе яраткан нәрсәләргә карап, аларның бик яхшы икәнен күрде. //
-				\glb[everyglb={\qipa }] Аллагь Оьзю яратгъан затлагъа къарап, олар бек яхшы экенин гёрген. //
-				\glb God own-his created [everything/thing-s]-to looked.at, they/their very good being saw. // 
-				\glft \eng{God looked at everything he had created and saw that it was very good.} (Bible, Genesis 1:31) //
-			\endgl
-		\xe
-
-		\htwo{Output}
-\scalebox{0.68}{
-\begin{tabular}{llll}
-\toprule
- \textbf{Kyrgyz} (\textbf{kir}) & \textbf{Kazakh} (\textbf{kaz}) & \textbf{Tatar} (\textbf{tat}) & \textbf{Kumyk} (\textbf{kum}) \\
-\midrule
-\texttt{Кудай Өзү жаратканынын баарына} & \texttt{Құдай Өзінің жаратқандарының бәріне} & \texttt{Аллаһ Үзе яраткан нәрсәләргә карап,} & \texttt{Аллагь Оьзю яратгъан затлагъа} \\
-\texttt{карап, өтө жакшы экенин көрдү.} & \texttt{қарап, өте жақсы екенін көрді.} & \texttt{аларның бик яхшы икәнен күрде.} & \texttt{къарап, олар бек яхшы экенин гёрген.} \\
-\midrule
-	\texttt{Кудай\tag{n}\tag{nom}} & \texttt{Құдай\tag{n}\tag{nom}} & \texttt{Аллаһ\tag{n}\tag{nom}} & \texttt{Аллагь\tag{n}\tag{nom}} \\
-	\texttt{Өз\tag{prn}\tag{ref}\tag{px3sp}\tag{nom}} & \texttt{Өз\tag{prn}\tag{ref}\tag{px3sp}\tag{gen}} & \texttt{Үз\tag{prn}\tag{ref}\tag{px3sp}\tag{nom}} & \texttt{Оьз\tag{prn}\tag{ref}\tag{px3sp}\tag{nom}} \\
-	\texttt{жарат\tag{v}\tag{tv}\tag{ger\_past}\tag{px3sp}\tag{gen}} & \texttt{жарат\tag{v}\tag{tv}\tag{ger\_past}\tag{pl}\tag{px3sp}\tag{gen}} & \texttt{ярат\tag{v}\tag{tv}\tag{gpr\_past}} & \texttt{ярат\tag{v}\tag{tv}\tag{gpr\_past}} \\
-	\texttt{баары\tag{prn}\tag{qnt}\tag{px3sp}\tag{dat}} & \texttt{бәрі\tag{prn}\tag{qnt}\tag{px3sp}\tag{dat}} & \texttt{нәрсә\tag{n}\tag{pl}\tag{dat}} & \texttt{зат\tag{n}\tag{pl}\tag{dat}} \\
-	\texttt{кара\tag{v}\tag{tv}\tag{gna\_perf}} & \texttt{қара\tag{v}\tag{tv}\tag{gna\_perf}} & \texttt{кара\tag{v}\tag{tv}\tag{gna\_perf}} & \texttt{къара\tag{v}\tag{tv}\tag{gna\_perf}} \\ 
-	\texttt{,\tag{cm}} & \texttt{,\tag{cm}} & \texttt{,\tag{cm}} & \texttt{,\tag{cm}} \\
-	 --- & ---      & \texttt{алар\tag{prn}\tag{pers}\tag{p3}\tag{pl}\tag{gen}} & \texttt{олар\tag{prn}\tag{pers}\tag{p3}\tag{pl}\tag{nom}} \\
-	\texttt{өтө\tag{adv}} & \texttt{өте\tag{adv}} & \texttt{бик\tag{adv}} & \texttt{бек\tag{adv}} \\
-	\texttt{жакшы\tag{adj}} & \texttt{жақсы\tag{adj}} & \texttt{яхшы\tag{adj}} & \texttt{яхшы\tag{adj}} \\
-	\texttt{э\tag{cop}\tag{ger\_past}\tag{px3sp}\tag{acc}} & \texttt{е\tag{cop}\tag{ger\_past}\tag{px3sp}\tag{acc}} & \texttt{и\tag{cop}\tag{ger\_past}\tag{px3sp}\tag{acc}} & \texttt{э\tag{cop}\tag{ger\_past}\tag{px3sp}\tag{acc}} \\
-	\texttt{көр\tag{v}\tag{tv}\tag{ifi}\tag{p3}\tag{sg}} & \texttt{көр\tag{v}\tag{tv}\tag{ifi}\tag{p3}\tag{sg}} & \texttt{күр\tag{v}\tag{tv}\tag{past}\tag{p3}\tag{sg}} & \texttt{гёр\tag{v}\tag{tv}\tag{past}\tag{p3}\tag{sg}} \\
-	\texttt{.\tag{sent}} & \texttt{.\tag{sent}} & \texttt{.\tag{sent}} & \texttt{.\tag{sent}} \\
-\bottomrule
-\end{tabular}
-}
-
-		\htwo{Tagset}
-
-\vspace{-1.5em}
-\setlength{\columnsep}{-1em}
-\setlength{\topsep}{-\parskip}
-%\spacing{0.8}
-\begin{multicols}{5}
-\begin{tabbing}
-  \texttt{{\small \tag{n}}}\hspace{1.5em} \=  Noun\\[-0.5ex]
-%  \texttt{{\small <np>}} \> Proper noun\\[-0.5ex]
-  \texttt{{\small \tag{v}}} \>  Verb\\[-0.5ex]
-  \texttt{{\small \tag{prn}}} \> Pronoun \\[-0.5ex]
-  \texttt{{\small \tag{det}}} \>  Determiner\\[-0.5ex]
-%  \texttt{{\small <cnjcoo>}} \> Coord. conjunct.\\[-0.5ex]
-%  \texttt{{\small <cnjadv>}} \> Adv. conjunct.\\[-0.5ex]
-  \texttt{{\small \tag{adj}}} \> Adjective\\[-0.5ex]
-  \texttt{{\small \tag{adv}}} \> Adverb\\[-0.5ex]
-%  \texttt{{\small <vaux>}} \> Auxiliary verb\\[-0.5ex]
-%  \texttt{{\small <cop>}} \> Copula\\[-0.5ex]
-  \texttt{{\small \tag{iv}}} \> Intransitive\\[-0.5ex]
-  \texttt{{\small \tag{tv}}} \> Transitive\\[-0.5ex]
-%  \texttt{{\small <p2>}} \> Second person\\[-0.5ex]
-  \texttt{{\small \tag{p3}}} \> Third person\\[-0.5ex]
-%  \texttt{{\small <ant>}} \> Anthroponym\\[-0.5ex]
-%  \texttt{{\small <dem>}} \> Demonstrative\\[-0.5ex]
-%  \texttt{{\small <m>}} \> Masculine\\[-0.5ex]
-%  \texttt{{\small <sg>}} \> Singular\\[-0.5ex]
-  \texttt{{\small \tag{pl}}} \> Plural\\[-0.5ex]
-  \texttt{{\small \tag{ref}}} \> Reflexive \\[-0.5ex]
-  \texttt{{\small \tag{pers}}} \> Personal \\[-0.5ex]
-  \texttt{{\small \tag{nom}}} \> `Nominative'\\[-0.5ex]
-  \texttt{{\small \tag{gen}}} \> Genitive\\[-0.5ex]
-  \texttt{{\small \tag{acc}}} \> Accusative\\[-0.5ex]
-  \texttt{{\small \tag{dat}}} \> Dative\\[-0.5ex]
-  \texttt{{\small \tag{qnt}}} \> Quantifier \\[-0.5ex]
-%  \texttt{{\small \tag{itg}}} \> Interrogative \\[-0.5ex]
-%  \texttt{{\small <loc>}} \> Locative\\[-0.5ex]
-%  \texttt{{\small <px3sg>}} \> 3rd person poss. (Singular)\\[-0.5ex]
-  \texttt{{\small \tag{cm}}} \> Comma\\[-0.5ex]
-  \texttt{{\small \tag{sent}}} \hspace{0.2em} \= Sentence\\[-0.5ex]
-%  \texttt{{\small <px3pl>}} \> 3rd person poss. (Plural)\\[-0.5ex]
-%  \texttt{{\small <neg>}} \> Negative\\[-0.5ex]
-%  \texttt{{\small <aor>}} \> Aorist\\[-0.5ex]
-  \texttt{{\small \tag{past}}} \> Past (General) \\[-0.5ex]
-  \texttt{{\small \tag{ifi}}} \> Past \\[-0.5ex]
-  	\> \hspace{0.5em}(Eyewitness/Recent) \\[-0.5ex]
-%  \\[-0.5ex]
-%  \\[-0.5ex]
-  \texttt{{\small \tag{px3sp}}} \> 3rd person poss.\\[-0.5ex]
-  	\> \hspace{0.5em}(Singular/Plural)\\[-0.5ex]
-%  \texttt{{\small <imp>}} \> Imperative\\[-0.5ex]
-  \texttt{{\small \tag{gna\_perf}}} \hspace{0.05em} \= Verbal adverb \\[-0.5ex]
-  	\> \hspace{0.5em}(Perfect)\\[-0.5ex]
-  \texttt{{\small \tag{gpr\_past}}} \> Verbal adjective \\[-0.5ex]
-  	\> \hspace{0.5em}(Past)\\[-0.5ex]
-  \texttt{{\small \tag{ger\_past}}} \> Verbal noun \\[-0.5ex]
-	\> \hspace{0.5em} (Past)
-%  \texttt{{\small <prc\_impf>}} \> Participle (Imperfect)\\[-0.5ex]
-%  \texttt{{\small <prc\_irre>}} \> Participle (Irrealis)\\[-0.5ex]
-%  \texttt{{\small <prc\_real>}} \> Participle (Realis)\\[-0.5ex]
-
-\end{tabbing}
-\end{multicols}
-
-		}
-
-		\headerbox{Categorisation}{name=adjectives,column=4,below=devel}{
-		%\headerbox{Morphophonology}{name=morphophon,below=morphtrans}{
+	\headerbox{Categorisation}{name=adjectives,column=2}{
 		\htwo{Adjectives}
@@
 
-		\headerbox{Evaluation}{name=coverage,span=1,column=5,row=0}{
-
-			\htwo{Number of stems}
-			%\vspace{-1.5em}
-			\vspace{-0.5em}
-			\begin{center}
-			\begin{tabular}{lrrrr}
-				\toprule
-				\multirow{2}{*}{\textbf{Part of speech}} & \multicolumn{4}{c}{\textbf{Number of stems}} \\ \cmidrule{2-5}
-				& \textbf{Kyrgyz} & \textbf{Kazakh} & \textbf{Tatar} & \textbf{Kumyk} \\
-				\midrule
-				Noun & 4582 & 2640 & 2795 & 2568 \\
-				Verb & 1193 & 1470 & 1143 & 386 \\
-				Adjective & 1211 & 754 & 816 & 219 \\
-				Proper noun & 5887 & 5701 & 5361 & 1443 \\
-				Adverb & 312 & 171 & 177 & 63 \\
-				Numeral & 66 & 63 & 63 & 44 \\
-				Conjunction & 77 & 46 & 45 & 13 \\
-				Postposition & 50 & 50 & 43 & 12 \\
-				Pronoun & 51 & 32 & 28 & 17 \\
-				Determiner & 64 & 39 & 34 & 9 \\
-				\midrule
-				Total: & \hiliteone{13749} & \hiliteone{11224} & \hiliteone{10737} & \hiliteone{4845} \\
-				\bottomrule
-			\end{tabular}
-			\end{center}
-			%\vspace{2em}
-			\vspace{-0.5em}
-
-			\htwo{Test corpora}
-
-%		\scalebox{0.8}{
-%			\begin{tabular}{lll}
-%				\toprule
-%				\textbf{type} & \textbf{lang} & \textbf{contents} \\ % & \textbf{origin} \\
-%				\midrule
-%				Encyclopædic & kaz & Wikipedia \\ % wpdump & 20131006 \\
-%					& tat & Wikipedia \\ % wpdump & 20130225 \\
-%					& kum & — \\ % — & — \\
-%				\midrule
-%				News & kaz & RFE/RL (azattyq.org) \\ % RFE/RL & azattyq.org 2010 \\
-%					& tat & tat.tatar-inform.ru \\ % Татар-информ & tat.tatar-inform.ru 2005-2011 \\
-%					& kum & Ёлдаш (yoldash.etnosmi.ru) \\ % Ёлдаш & yoldash.etnosmi.ru \\
-%				\midrule
-%				Religion & kaz & Quran + Bible \\ % quran + bible & kkitap.net, kuran.kz \\
-%					& tat & Quran + New Testament \\ % quran + nt & ibt.org.ru, tanzil.net \\
-%					& kum & Genesis + New Testament \\ % genesis + nt & ibt.org.ru \\
-%				\bottomrule
-%			\end{tabular}
-%		}
-		\scalebox{0.8}{
-			\begin{tabular}{llll}
-				\toprule
-					& \textbf{Wikipedia} & \textbf{News} & \textbf{Religion} \\
-				\midrule
-					\textbf{Kyrgyz} & Wikipedia & azattyk.org & Bible \\
-					% 2013-05-30, 2010, 
-					\textbf{Kazakh} & Wikipedia & azattyq.org & Quran + Bible \\
-					\textbf{Tatar} & Wikipedia & tat.tatar-inform.ru & Quran + New Testament \\
-					\textbf{Kumyk} & --- & yoldash.etnosmi.ru & Genesis + New Testament \\
-				\bottomrule
-			\end{tabular}
-
-%			\begin{tabular}{llll}
-%				\toprule
-%					& \textbf{Kazakh} & \textbf{Tatar} & \textbf{Kumyk} \\
-%				\midrule
-%					Wikipedia & Wikipedia & Wikipedia & --- \\
-%					News & azattyq.org & tat.tatar-inform.ru & yoldash.etnosmi.ru \\
-%					Religion & Quran + Bible & Quran + New Testament & Genesis + New Testament \\
-%				\bottomrule
-%			\end{tabular}
-		}\vspace{0.55em}
-
-%			\begin{itemize}
-%				\item split into 10 equal parts; coverage calculated over each separately; standard deviation of mean calculated
-%			\end{itemize}
-			%both split into 10 equal parts; coverage calculated over each separately; standard deviation of the mean was calculated\\
-			\htwo{Evaluation measures}
+	\headerbox{Ongoing and future work}{name=nedostatki,column=3}{
 		\begin{itemize}
-				\item \textbf{Naïve coverage} - percentage of surface forms in a given corpus receiving ≥ 1 analysis%\\(surface forms may have missing analyses)
-				\item \textbf{Mean ambiguity} - average number of analyses for each surface form found in analysed corpus
-%				Precision is the probability that an analysis provided by the transducer is accurate.
-%				Recall is the probability that some correct analysis is among those provided.
-				\item \textbf{Precision} - probability that a provided analysis is valid
-				\item \textbf{Recall} - probability that a certain valid analysis is among those provided by the transducer
-				%\item \textbf{Precision} - of a form's analyses, \% correct
-				%\item \textbf{Recall} - \% of analyses provided by transducer that are correct for a form, by comparing against a gold standard
-			\end{itemize}
-			%\htwo{Coverage results}\\
-			\htwo{Evaluation results}
-%\centering
-%\vspace{1pt}\hspace{2em}
-			%%\begin{centering}
-			%{\large
-			%\begin{itemize}
-			%	\item each corpus split into 10 equal parts; coverage calculated over each separately; st.\ dev.\ of mean calculated
-			%\end{itemize}
-			\vspace{-0.5em}
-\begin{center}
-		\scalebox{0.94}{
-			\begin{tabular}{llrrr}
-				\toprule
-				 & \textbf{Corpus} & \textbf{Tokens} & \textbf{Coverage} (\%) & \textbf{Amb.} \\
-				\midrule
-
-				\multirow{4}{*}{\textbf{Kyrgyz}} & Wikipedia & 5.3M & 84.51 $\pm$ 2.27 & 3.56 \\
-					& News & 4.1M & 91.43 $\pm$ 0.51 & 4.19 \\
-					& Religion & 215K & 91.66 $\pm$ 1.81 & 3.99 \\\cmidrule{2-5}
-				 (r54474) & Average &  & \hiliteone{89.20} $\pm$ 3.48 & 3.91 \\
-				\midrule
-				\multirow{4}{*}{\textbf{Kazakh}} & Wikipedia & 25.6M & 85.61 $\pm$ 1.37 & 2.43 \\
-					& News & 3.8M & 92.12 $\pm$ 2.72  & 2.88 \\
-					& Religion & 851K & 92.49 $\pm$ 1.66  & 2.63 \\\cmidrule{2-5}
-				 (r50547) & Average &  & \hiliteone{90.07} $\pm$ 1.91  & 2.64\\
-				\midrule
-				\multirow{4}{*}{\textbf{Tatar}} & Wikipedia & 159K & 86.35 $\pm$ 2.17  & 2.24\\
-					& News & 5.2M & 89.75 $\pm$ 0.07  & 2.30 \\
-					& Religion & 382K & 91.25 $\pm$ 2.55 & 2.24 \\\cmidrule{2-5}
-				 (r50260) & Average &  & \hiliteone{89.12} $\pm$ 1.60 & 2.26 \\
-				\midrule
-				\multirow{3}{*}{\textbf{Kumyk}} & News & 286K &  91.10 $\pm$ 0.86  & 1.53 \\
-					& Religion & 227K & 92.47 $\pm$ 1.03  & 1.53 \\\cmidrule{2-5}
-				 (r50300) & Average &  & \hiliteone{91.78} $\pm$ 0.94 & 1.53 \\
-				\bottomrule
-			\end{tabular}\\
-		}
-\end{center}
-			%}
-			%%\end{centering}
-%		}
-
-%		\headerbox{Precision \& recall}{name=precrecall,column=1,below=coverage}{
-
-%			\htwo{Two measures}
-%			\htwo{Precision \& recall}
-			%\vspace{0.425em}
-			\vspace{-0.4em}
-			\begin{itemize}
-				\item {}{\small selected \& proofed unique random surface forms from news corpora}
-			\end{itemize}
-
-			\vspace{-1em}
-			\begin{center}
-			\scalebox{1.05}{
-				\begin{tabular}{lrrr}
-				\toprule
-					\textbf{Language} & \textbf{Forms} & \textbf{Precision} (\%) & \textbf{Recall} (\%) \\
-				\midrule
-					Kyrgyz & 200 & 90.77 & 69.15 \\
-					Kazakh & 1000 & 98.61 &  57.98 \\
-					Tatar & 1000 & 95.03 & 85.65 \\
-					Kumyk & 500 & 96.57 & 69.11 \\
-				\bottomrule
-				\end{tabular}
-			}
-			\end{center}
-
-			%\vspace{0.49em}
-%			\begin{itemize}
-%				\item selected \& proofed unique random surface forms from news corpus
-%				\item \textbf{Precision} (of a form's analyses \% correct): \hfill \hiliteone{\textbf{97.32\%}}
-%				\item \textbf{Recall} (percentage of analyses provided by the transducer that are correct for a form, by comparing against a gold standard): \hfill \hiliteone{\textbf{94.56\%}}
-%			\end{itemize}
-
-%			\htwo{Results}
-%
-%			\vspace{-0.5em}
-%			\begin{center}
-%				\begin{tabular}{ll}
-%					\toprule
-%					Precision & Recall \\
-%					\midrule 
-%					97.32\% & 94.56\%  \\
-%					\bottomrule
-%				\end{tabular}
-%			\end{center}
-		}
-
-		\headerbox{Orthography-phonology mapping issues}{name=twol,column=3}{ %,below=exout}{
-			\htwo{Ambiguous characters}
-			\begin{itemize}
-				\item Have front- and back-vowel readings in native words
-			\end{itemize}
-
-		%\scalebox{0.92}{
-			\begin{tabular}{llll}
-				\toprule
-					& \textbf{letters} & \textbf{values} & \textbf{examples} \\
-				\midrule
-					\multirow{2}{*}{\texttt{\textbf{kaz}}} & \multirow{2}{*}{{\qipa и, у, ю}} & {\qipa /əj, əw, jəw/} & {\qipa қ\hiliteone{ию}д\hilitetwo{а}} \eng{chopping down} \\
-						&	& {\qipa /ɘj, ɘw, jɘw/} &  {\qipa к\hiliteone{ию}д\hilitetwo{е}} \eng{getting dressed} \\
-				\midrule
-
-					\multirow{3}{*}{\texttt{\textbf{tat}}} & \multirow{3}{*}{{\qipa е}} & {\qipa э / C \_} & {\qipa дәр\hiliteone{е}сл\hilitetwo{ә}р} \eng{lessons} \\
-						&	& {\qipa /j/+ы} & {\qipa \hiliteone{е}лл\hilitetwo{а}р} \eng{years} \\
-						&	& {\qipa /j/+э} & {\qipa \hiliteone{е}гетл\hilitetwo{ә}р} \eng{boys} \\
-				\midrule
-
-					\multirow{6}{*}{\texttt{\textbf{kum}}} & \multirow{6}{*}{{\qipa ё, ю}} & \multirow{2}{*}{{\qipa /ø, y/ / C \_}} & {\qipa г\hiliteone{ю}нл\hilitetwo{е}р} \eng{days} \\
-						&	&	& {\qipa г\hiliteone{ё}зл\hilitetwo{е}р} \eng{eyes} \\
-						&	& \multirow{2}{*}{{\qipa /jø, jy/}} & {\qipa \hiliteone{ю}рекл\hilitetwo{е}р} \eng{hearts}\\
-						&	&	& {\qipa \hiliteone{ё}нкюл\hilitetwo{е}р} \eng{darlings} \\
-						&	& \multirow{2}{*}{{\qipa /jo, ju/}} & {\qipa \hiliteone{ю}лдузл\hilitetwo{а}р} \eng{stars} \\
-						&	&	& {\qipa \hiliteone{ё}лл\hilitetwo{а}р} \eng{roads} \\
-				\bottomrule
-			\end{tabular}
-		%}
-		\vspace{0.5em}
-			\begin{itemize}
-				\item solution:\ hairy \texttt{twol} rules cover majority of examples
-				\item unaccounted-for words get a harmony-forcing character
-				\item adjust rules for harmony-forcing characters
-			\end{itemize}
-
-			\htwo{Loanwords}
-				\begin{itemize}
-					\item Letters that represent front vowels in native words may represent ``back'' vowels in Russian words
-				\end{itemize}
-			\scalebox{0.92}{
-				\begin{tabular}{lll}
-					\toprule
-						 & \textbf{native word example} & \textbf{Russian word example} \\
-					\midrule
-						\texttt{\textbf{kaz}} & {\qipa \hiliteone{е}лд\hilitetwo{і}ң} \eng{country's} & {\qipa Назарба\hiliteone{е}вт\hilitetwo{ы}ң} \eng{Nazarbayev's} \\
-						\texttt{\textbf{tat}} & {\qipa гал\hiliteone{и}мн\hilitetwo{ә}р} \eng{scientists} & {\qipa арт\hiliteone{и}стл\hilitetwo{а}р} \eng{artists} \\
-						\texttt{\textbf{kum}} & {\qipa с\hiliteone{ё}зл\hilitetwo{е}р} \eng{words} & {\qipa самол\hiliteone{ё}тл\hilitetwo{а}р} \eng{airplanes} \\
-					\bottomrule
-				\end{tabular}
-			}
-			\begin{itemize}
-				\item solution:\ separate continuation lexicon (messy rules)
-			\end{itemize}
-				\codeex{
-					\begin{spacing}{0.9}
-					\texttt{LEXICON N1-RUS\\
-					:\%\{☭\%\} N1 ;\\
-					\vspace*{-0.5em}\\
-					LEXICON Nouns\\
-					артист:артист N1-RUS ; ! "artist"\\
-					галим:галим N1 ; ! "scientist"}\end{spacing}
-					\vspace{0em}
-				}\\
-
-			\vspace{-0.75ex}
-			\htwo{Acronyms and numerals}
-				\begin{itemize}
-					\item \texttt{twol} rules handle phonology for spelt-out words\\
-					{\qipa от\hiliteone{ыз}\hilitetwo{да}н} \eng{from thirty}, {\qipa б\hiliteone{ес}\hilitetwo{те}н} \eng{from five}
-					\item no phonological triggers available in numerals\vspace{-0.2em}\\
-					 (incorrect phonological triggers in acronyms)\\
-					{\qipa 30-\hilitetwo{да}н} \eng{from 30}, {\qipa 5-\hilitetwo{те}н} \eng{from 5}
-					\item solution:\ phonology-triggering characters
-					\item simplified:\ e.g., \archiphon{с} for all voiceless ostruents
-				\end{itemize}
-				\codeex{
-					\begin{spacing}{0.9}
-					\texttt{4:4\%\{э\%\}\%\{с\%\}           NUM-DIGIT ; ! "төрт"\\
-					5:5\%\{э\%\}\%\{с\%\}           NUM-DIGIT ; ! "бес"\\
-					3\%0:3\%0\%\{а\%\}\%\{з\%\}       NUM-DIGIT ; ! "отыз"}\end{spacing}
-					\vspace{0em}
-				}\\
-
-			\vspace{-0.75ex}
-			\htwo{й + vowel letters}
-				\begin{itemize}
-					%\item Stem-final й combines with following а, о, у to form я, ё, ю
-					\item{} [ а о у ] become [ я ё ю ] after й and й deletes
-					\item й incorporated into the context of many rules
-					\item additional rules to change the characters and delete original й
-				\end{itemize}
-%	\vspace{-1ex}
-%
-				\codeex{
-					\begin{spacing}{0.9}
-					\texttt{"Deletion of й before yoticised vowels"\\
-					й:0 <=> \blank{1em} [ :YotVow ]/[ :0 | \%>:\ ]\ ;}
-					\end{spacing}
-					\vspace{0em}
-				}
-%					\item 
-%				\end{itemize}
-
-			\htwo{A resulting messy \texttt{twol} rule}
-				\codeex{
-					\begin{spacing}{0.4}
-					RdYotVow = ё ю Ё Ю ; \\
-					AbstractVow = \%\{а\%\} \%\{э\%\} \%\{ү\%\} \%\{о\%\} ; \\
-					\\
-					"{A} front unrounded harmony" \\
-\%\{A\%\}:е <=> \hfill{} [ [ :FrontVow | [ :Vow :ь ] ] :Cns :Cns* ]/:0 \_ ; \\
-\hspace*{0.1em}\hfill{}                        [ [ :RdVow :ь ] :Cns :Cns* ]/:0 \_ ; \\
-\hspace*{0.1em}\hfill{}     [ [ [ \textbackslash{}[ .\#.\ | :Vow ] ] :RdYotVow ] :Cns :Cns* ]/:0 \_ ; \\
-\hspace*{0.1em}\hfill{}      [ :RdYotVow й:0 :RdYotVow :Cns :Cns* ]/[ :0 - й:0 ] \_ ; \\ % ! кююню \\
-\hspace*{0.1em}\hfill{} [[\%\{э\%\}:0|\%\{ү\%\}:0] :Cns* ]/[[ :0 - AbstractVow: ] | \%-: ]* \_ ; \\
-\hspace*{4em}        except \\
-\hspace*{0.1em}\hfill{}       [ :RdYotVow :Cns* \%\{☭\%\}:0 :Cns* ]/[ :0 - \%\{☭\%\}:0 ] \_ ; \\ % ! -ёрлар in Russian words \\
-\hspace*{0.1em}\hfill{}                          [ :Cns :р \%\{☭\%\}: \%>: :Cns* ]/:0 \_ ; \\% ! центр and тигр take back vowels \\
-\hspace*{0.1em}\hfill{}         [ [ :Vow - :RdYotVow ] :RdYotVow :Cns :Cns* ]/:0 \_ ; \\% ! for some reason the rule above doesn't exclude this condition  e.g., боюнгъа \\
-\hspace*{0.1em}\hfill{}                          [ :Vow ]/[ [ :0 - й:0 ] | \%>: ] \_ ;\end{spacing}
-			\vspace{1em}
-
-				}
-		}
-
-		\headerbox{Ongoing and future work}{name=nedostatki,column=5,below=coverage}{
-			\begin{itemize}
-%				\item case changes for words with one root\\
-%					\hiliteone{Ф}инландия \eng{Finland}, \hiliteone{ф}инландиялык \eng{Finnish}
-%				\item phonol.\ (vowel harmony, desonorisation) with abbrevs.\\
-%					АКШ [акышы] \eng{USA} → АКШ\hiliteone{н}ын / *АКШ\hiliteone{т}ын
-%				\item vowel harmony with numbers\\
-%					{\qipa 100 [жүз] → 100дүн [жүздүн] / *100нын}
-%				\item compound verbs (esp.\ ones with changeable parts)
-%				\item gerunds with mono-syllabic V-final verbs\\
-%					{\qipa иште- \eng{work} → иштеш / иштөө \eng{working}}\\
-%					{\qipa же- \eng{eat} → жеш / *жөө \eng{eating}}
 			\item Disambiguation, more stems, clean up transducers
@@
 		\end{itemize}
-			%\vspace{0.98em}
 	}
 
-		\headerbox{Further information}{name=getting,column=0,below=morphtrans}{
+	\headerbox{Further information}{name=getting,column=3,below=nedostatki}{
 		\begin{itemize}
@@
 			%info at {\small \url{http://wiki.apertium.org/wiki/apertium-kir}}
-
 				\item Turkic RBMT \textbf{mailing list} (>25 subscribers):\\
@@
 		\end{itemize}
-			%\vspace{2.8ex}
 	}
@@
 
-		\headerbox{Morphophonology}{name=morphophon,column=1,below=exout}{
-%			\htwo{Vowel harmony}
-%			\begin{itemize}
-%				\item two basic vowel archiphonemes:
-%				\begin{itemize}
-%					\item high vowel, \texttt{\{I\}}; [±back] and [±round] from prev.\ V\\
-%		\begin{tabular}{ccccc}
-%			\toprule
-%			after & result & & after & result \\
-%			\midrule
-%			{\qipa и} & {\qipa и} & & {\qipa ы} & {\qipa ы} \\
-%			{\qipa ү} & {\qipa ү} & & {\qipa у} & {\qipa у} \\
-%			{\qipa е} & {\qipa и} & & {\qipa а} & {\qipa ы} \\
-%			{\qipa ө} & {\qipa ү} & & {\qipa о} & {\qipa у} \\
-%			\bottomrule
-%		\end{tabular}
-%					\item low vowel, \texttt{\{A\}}; [±back] and [±round] from prev.\ V\\
-%					with one important exception: after \hilitetwo{у}\\
-%		\begin{tabular}{ccccc}
-%			\toprule
-%			after & result & & after & result \\
-%			\midrule
-%			{\qipa и} & {\qipa е} & & {\qipa ы} & {\qipa а} \\
-%			{\qipa ү} & {\qipa ө} & & {\qipa \hilitetwo{у}} & {\qipa \hilitetwo{а}} \\
-%			{\qipa е} & {\qipa е} & & {\qipa а} & {\qipa а} \\
-%			{\qipa ө} & {\qipa ө} & & {\qipa о} & {\qipa о} \\
-%			\bottomrule
-%		\end{tabular}
-%
-%
-%				\end{itemize}
-%				\item take backness and roundness of previous vowel\\
-%				exception: 
-%
-%			\end{itemize}
+		\headerbox{Morphophonology}{name=morphophon,column=1}{
 
@@
 				}
+			\vspace{0.1em}
 
@@
 }
-
-%			%\htwo{Nouns ending in /рн/}
-%
-%			\htwo{й+vowel letters}
-%			\begin{itemize}
-%				\item{} [ а о у ] become [ я ё ю ] after й and й deletes
-%				\item й incorporated into the context of many rules
-%				\item + separate rules to change the characters
-%				\item + a rule to delete the original й\\
-%			\end{itemize}
-%	\vspace{-1ex}
-%
-%				\codeex{
-%					\texttt{"Deletion of й before yoticised vowels"\\
-%					й:0 <=> \blank{1em} [ :YotVow ]/[ :0 | \%>:\ ]\ ;}
-%				}
-		}
-
-		\headerbox{Other uses}{name=uses,column=1,below=morphophon}{
-			\begin{itemize}
-				\item HFST transducers are trivially converted to \textbf{spell checkers}
-				\item Segmenter, e.g.\ {\qipa көргөзгөндөрдөнсүңбү} :\\
-				 {\small \texttt{көр\hilitegrey{>}\{G\}\{A\}з\hilitegrey{>}\{G\}\{A\}н\hilitegrey{>}\{L\}\{A\}р\hilitegrey{>}\{D\}\{A\}н\hilitegrey{>}с\{I\}ң\hilitegrey{>}\{B\}\{I\}}}
-
-				 %көр\hilitegrey{>}\{I\}ш\hilitegrey{>}\{G\}\{A\}н\hilitegrey{>}\{L\}\{A\}р\hilitegrey{>}\{I\}м\hilitegrey{>}\{D\}\{A\}н\hilitegrey{>}с\{I\}ң\hilitegrey{>}\{B\}\{I\}}}
-			\end{itemize}
-		}
-
-		\headerbox{Morphotactics}{name=lexc,column=2,below=exout}{
-			\htwo{Morphological \& orthographical words}
-
-			\begin{itemize}
-				\item{} {\qipa өнүктүрөбүзбү ?} \eng{will we develop [it]?}\\
-				\texttt{өнүк\tag{v}\tag{tv}\tag{caus}\tag{aor}\tag{p1}\tag{pl}+бы\tag{qst}}
-				\item{} {\qipa келатсаң} \eng{if you come}\\
-				\texttt{кел\tag{v}\tag{iv}\tag{prc\_impf}+жат\tag{vaux}\tag{gna\_cnd}\tag{p2}\tag{sg}}
-			\end{itemize}
-
-			%\htwo{Irregular negatives of finite verb forms}
-
-			\htwo{Irregular [noun + possessive + case] forms}
-
-			\begin{itemize}
-				\item Some combinations of possessive + case morphemes are unpredicted (i.e., not formed simply by concatenation and application of phonology):
-			\end{itemize}
-
-			\begin{center}
-			\vspace{-0.5em}
-{\small
-{\qipa		
-	\begin{tabular}{llllll}
-		\toprule
-		case & form & 1SG & 2SG & 3SP  \\
-		\midrule
-		nominative & — & -(I)м & -(I)ң & -(с)I  \\
-		accusative & -NI & -(I)мдI & -(I)ңдI & -\textbf{(с)Iн}  \\
-		genitive & -NIн & -(I)мдIн & -(I)ңдIн & -(с)IнIн  \\
-		locative & -DA & -(I)мдA & -(I)ңдA & -\textbf{(с)IндA}  \\
-		ablative & -DAн & -(I)мдAн, & -(I)ндAн, & -\textbf{(с)IнAн}  \\
-		~   &     & -\textbf{(I)мAн} & -\textbf{(I)ңAн} \\
-		dative & -GA & -\textbf{(I)мA} & -\textbf{(I)ңA} & -\textbf{(с)IнA} \\
-		\bottomrule
-	\end{tabular}\\
-}}
-	{\footnotesize A,I,N,D,G have various allophones; (I) null after vowels; (с) null after cons.}\vspace{-1em}\end{center}
-
-%				\item Trade-off:
-				\begin{itemize}
-					\item underlying \tag{px3sp} form used:\ \texttt{\archiphon{s}\archiphon{I}\archiphon{n}}
-					\item \texttt{\archiphon{s}} and \texttt{\archiphon{n}} default to с and н; rules map to null by context
-					\item morphophonology more complicated, morphotactics simpler
-%				\end{itemize}
-			\end{itemize}
-
-			\htwo{Noun-noun compounds}
-
-			\begin{itemize}
-				\item a N-N compund type:\ N2 has \tag{px3} and related morphology\\
-				e.g., \texttt{аба ырайы\tag{n}\tag{loc}:\ \hiliteone{аба ырайында}, *\hilitetwo{аба ырайыда}}
-			\end{itemize}
-			\vspace{0.6ex}
-			\codeex{
-			\begin{spacing}{0.9}
-			{\small
-			\texttt{LEXICON N-INFL-3PX-COMPOUND\\
-			\vspace{-0.5ex}\%<n\%>:\%>\%\{S\%\}\%\{I\%\}\%\{n\%\} GEN-POS ;}\\
-
-			\texttt{LEXICON Nouns\\
-			\vspace{-0.5ex}аба\%\ ырайы:аба\%\ ырай\ N-INFL-3PX-COMPOUND ;\\\vspace{-0.5ex}\hspace{0.1pt}\hfill{}!\ "weather"\\
-			\vspace{-0.5ex}чакыруу\%\ кагазы:чакыруу\%\ кагаз\ N-INFL-3PX-COMPOUND\\\vspace{-0.5ex}\hspace{0.1pt}\hfill{};\ !\ "invitation"}}\end{spacing}
-			\vspace{0.5ex}}
-
-			%\vspace{0.5ex}
-			%чакыруу кагазы
+\vspace{0.1em}
 		}

### 2015-06-27 2015-compmorphon-morphver/poster.tex
switch logo placement?
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61096 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 	}{ %Eye Catcher, empty if option eyecatcher=false - unused
-		\includegraphics[height=6.5em]{uitlogo}
-		\hspace{2cm}\begin{minipage}[t]{9em}
-			\vspace{-2cm}
-			\noindent\includegraphics[height=4.2em]{DCU_logo_2col}\\
-			\vspace{-0.5ex}\hspace{0.105cm}\includegraphics[height=2.15em]{DCU_2010_Name_Mark_289}
-			\end{minipage}
-			\hspace{2cm}\includegraphics[height=6.5em]{iu_tab}
+		\includegraphics[angle=90,height=6.5em]{apertium5b}%\hspace{-2em}
 	}{ %Title (centered top)
@@
 	}{ %More eye catchers, on the right side
-		\includegraphics[angle=90,height=6.5em]{apertium5b}%\hspace{-2em}
+		\includegraphics[height=6.5em]{uitlogo}
+		\hspace{2cm}\begin{minipage}[t]{9em}
+			\vspace{-2cm}
+			\noindent\includegraphics[height=4.2em]{DCU_logo_2col}\\
+			\vspace{-0.5ex}\hspace{0.105cm}\includegraphics[height=2.15em]{DCU_2010_Name_Mark_289}
+			\end{minipage}
+			\hspace{2cm}\includegraphics[height=6.5em]{iu_tab}
 	}

### 2015-07-03 2015-compmorphon-morphver/poster.tex
some real edits
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61166 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 %\definecolor{boxcolor}{RGB}{176,208,223}
-\definecolor{boxcolor}{cmyk}{0.3,0.08,0.08,0.1}
+%\definecolor{boxcolor}{cmyk}{0.3,0.08,0.08,0.1}
+\definecolor{boxcolor}{cmyk}{0.2,0.03,0.03,0.1}
 \definecolor{headerfontcol}{RGB}{255,255,255}
@@
 \usepackage{fontspec}
-\defaultfontfeatures{PunctuationSpace=3,Scale=MatchLowercase,Mapping=tex-text}
+\defaultfontfeatures{PunctuationSpace=2,Scale=MatchLowercase,Mapping=tex-text}
 \newfontfeature{IPA}{+mgrk}
@@
 %\newfontfamily\titlefont[IPA,Scale=0.52]{CMU Serif}
-\newfontfamily\titlefont[IPA,Scale=0.7]{CMU Serif}
+%\newfontfamily\titlefont[IPA,Scale=0.7]{CMU Serif}
+%\newfontfamily\titlefont[IPA,Scale=0.7]{CMU Sans Serif Demi Condensed}
+\newfontfamily\titlefont[
+	%SmallCapsFont={Andika},
+	%SmallCapsFont={Alegreya Sans SC},
+	SmallCapsFont={Andada SC},
+	%SmallCapsFont={Carrois Gothic SC},
+	SmallCapsFeatures={Letters=SmallCaps},
+	Scale=0.8,
+]{Andada SC}
+%\newfontfamily\titlefont[IPA,Scale=0.7]{Nimbus Sans L}
 %\newfontfamily\titlefont[IPA,Scale=0.55]{DejaVu Serif}
@@
 	}{ %Title (centered top)
-		{\vspace{0pt}\hspace{1cm}
-			{\titlefont \textsc{Finite-state morphologies and text corpora\\as resources for improving morphological descriptions}}}
+		{%\vspace{0pt}\hspace{1cm} \
+			{\titlefont \scshape{Finite-State Morphologies \& Text Corpora\\\vspace{0.2em}\scalebox{0.8}{as Resources for Improving Morphological Descriptions}}}}
 	}{ %Authors (centered top, below title)
-		\vspace{-0.5em}
+		\vspace{-0.6em}
 		\begin{center}
@@
 					{Tommi Pirinen\vphantom{y}}\\
-					{\footnotesize Ollscoil Chathair Bhaile Átha Cliath\\\texttt{FIXME Flammie}}
+					{\footnotesize Ollscoil Chathair Bhaile Átha Cliath\vphantom{y}\\\texttt{FIXME Flammie}}
 				\end{spacing}
 			\end{minipage}
-			\begin{minipage}[t]{12em}
+			\begin{minipage}[t]{11.2em}
 				\begin{spacing}{0.4}
@@
 
-	\headerbox{Random box 1}{name=boxone,column=0,row=0}{
+	\headerbox{Morphological Descriptions}{name=morphdesc,column=0,row=0}{
+		Morphological descriptions describe the \_morphotactics\_¹ and \_morphophonology\_² of a language.
+
+		These are a core part of grammatical descriptions, such as reference grammars and textbooks.
+Due to restrictions of the medium (such as number of pages), the limitations of introspection and working with native speakers, and the lack of automatic testing, such works are often incomplete.
+
+		¹ Morphotactics refers to the morphemes of a language and how they can be combined.
+
+² Morphophonology refers to the alternation between the various phonological and orthographical forms of a morpheme in a language.
 		\scalebox{0.78}{
@@
 
-	\headerbox{Morphological transducers}{name=morphtrans,below=boxone}{
+	\headerbox{Morphological transducers}{name=morphtrans,below=morphdesc}{
 		\htwo{Morphological transducers}
@@
 
-		\htwo{Framework: HFST}
+		\htwo{Framework:\ HFST}
 		\begin{itemize}
@@
 
-	\headerbox{Categorisation}{name=adjectives,column=2}{
+	\headerbox{Categorisation}{name=adjectives,column=3}{
+	For example, many grammars of Turkic languages (e.g., Somfai Kara, 2002) overgeneralise adjective morphology, stating that all adjectives may act like nouns (and take nominal morphology), or that all adjectives may be used as adverbs with no morphological changes.
+We have found that there is a range of “adjective classes” in most Turkic languages, including some which may not be substantivised and some which may not be used adverbially.
+This phenomenon is of particular interest because native-speaker intuitions on the morphological limitations of a given adjective can often be more restrictive than the range of possible uses found in a large text corpus.
 		\htwo{Adjectives}
@@
 
-	\headerbox{Ongoing and future work}{name=nedostatki,column=3}{
+	\headerbox{Ongoing and future work}{name=nedostatki,below=adjectives,column=3}{
 		\begin{itemize}
@@
 
-		\headerbox{Morphophonology}{name=morphophon,column=1}{
+		\headerbox{Example:\ Tuvan velar elision}{name=tuvanvelars,column=1,span=2}{
+			\begin{multicols}{2}
+				\htwo{Previous descriptions}
+				- Anderson and Harrison, 1999
+
+				- Исхаков, Ф.Г.; Пальмбах, 1961, p. 117-118
+				\htwo{Initial implementation}
+
+				\htwo{Initial frequency list}
+
+			\end{multicols}
+
+			\begin{multicols}{2}
+				\htwo{Examples of correct forms}
+
+				\htwo{More complete characterisation}
+
+				\htwo{Final implementation}
+
+				\htwo{Final frequency list}
+			\end{multicols}
+
+		}
+
+		\headerbox{Examples:\ ideas for other examples}{name=otherexamples,column=1,span=2,below=tuvanvelars}{
+			\begin{itemize}
+				\item Kyrgyz: conditions where л desonorises to д
+				\item Tuvan: рубль - рубльдиң *рубьдуң
+				\item clitics can go everywhere/anywhere
+				\item underformalised - “particles☹” can go anywhere
+				\item all adjs act like nouns (Somfai-Kara, p. 28)
+				\item all adjectives can be used as adverbs without any morphological changes (Somfai-Kara, p. 29)
+				\item Stuff found doing treebank
+			\end{itemize}
+
+		}
+
+		\headerbox{Morphophonology}{name=morphophon,column=1,below=otherexamples}{
 

### 2015-07-05 2015-compmorphon-morphver/poster.tex
Flammie's email address
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61168 72bbbca6-d526-0410-a7d9-f06f51895060
better scaled heaader graphics and stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61173 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 	}{ %Eye Catcher, empty if option eyecatcher=false - unused
-		\includegraphics[angle=90,height=6.5em]{apertium5b}%\hspace{-2em}
+		\includegraphics[angle=90,height=6.8em]{apertium5b}%\hspace{-2em}
 	}{ %Title (centered top)
@@
 					{Tommi Pirinen\vphantom{y}}\\
-					{\footnotesize Ollscoil Chathair Bhaile Átha Cliath\vphantom{y}\\\texttt{FIXME Flammie}}
+					{\footnotesize Ollscoil Chathair Bhaile Átha Cliath\vphantom{y}\\\texttt{tommi.pirinen@computing.dcu.ie}}
 				\end{spacing}
@@
 		\includegraphics[height=6.5em]{uitlogo}
-		\hspace{2cm}\begin{minipage}[t]{9em}
+		\hspace{2.5cm}\begin{minipage}[t]{7em}
 			\vspace{-2cm}
@@
 		\end{minipage}
-			\hspace{2cm}\includegraphics[height=6.5em]{iu_tab}
+		\hspace{2.5cm}\includegraphics[height=6.25em]{iu_tab-crop}
 	}
@@
 				\htwo{Previous descriptions}
-				- Anderson and Harrison, 1999
 
-				- Исхаков, Ф.Г.; Пальмбах, 1961, p. 117-118
+				- Anderson and Harrison (1999, 22-23)\\
+				\includegraphics[width=0.49\textwidth]{img/andersonharrison22-23}\\
+				- Исхаков, Ф.Г.; Пальмбах (1961, 117-118)\\
+				\includegraphics[width=0.49\textwidth]{img/isxakovpalmbax117}\\
+
+				\columnbreak{}
 				\htwo{Initial implementation}
@@
 				\htwo{Final frequency list}
+				\includegraphics[width=0.4\textwidth]{img/andersonharrison35}\\
 			\end{multicols}

### 2015-07-06 2015-compmorphon-morphver/poster.tex
more work on the Tuvan velar elision, fixed colour issues, updated eye-catchers / centred title
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61174 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \definecolor{headercol}{cmyk}{1,0.64,0,0.6}
-\definecolor{headercolone}{cmyk}{1,0.64,0,0.6}
-\definecolor{headercoltwo}{cmyk}{0.44,0,0.15,0.17}
+%\definecolor{headercolone}{cmyk}{1,0.64,0,0.6}
+%\definecolor{headercoltwo}{cmyk}{0.44,0,0.15,0.17}
+%\definecolor{headercolone}{cmyk}{0.3,0,0.15,0.17}
+%\definecolor{headercoltwo}{cmyk}{0.3,0,0.15,0.17}
+\definecolor{iumint}{cmyk}{0.40,0,0.23,0}
+\definecolor{iumidnight}{cmyk}{0.71,0.30,0.13,0.41}
 %\definecolor{boxcolor}{RGB}{176,208,223}
@@
 %\setromanfont[IPA,Scale=0.8]{CMU Serif}
-\setromanfont[IPA]{Liberation Serif}
+%\setromanfont[IPA]{Liberation Serif}
+\setromanfont[IPA]{Linux Libertine O}
 \setmonofont[IPA]{DejaVu Sans Mono}
@@
          \begin{minipage}[t]{0.91\textwidth}{\litamono
+				\begin{spacing}{0.4}
 	            #1
+				\vspace{-1em}\end{spacing}
          }\end{minipage}
@@
 		borderColor=bordercol,
-		headerColorOne=headercolone,
-		headerColorTwo=headercoltwo,
+		headerColorOne=iumidnight,
+		headerColorTwo=iumidnight,
 		%headerColorOne=blue,
-		%headerFontColor=white,
+		headerFontColor=white,
 		% Only simple background color used, no shading, so boxColorTwo isn't necessary
@@
 	}{ %Eye Catcher, empty if option eyecatcher=false - unused
-		\includegraphics[angle=90,height=6.8em]{apertium5b}%\hspace{-2em}
+		\includegraphics[height=6.5em]{uitlogo}
+		\hspace{2.5cm}\begin{minipage}[t]{7em}
+			\vspace{-2cm}
+			\noindent\includegraphics[height=4.2em]{DCU_logo_2col}\\
+			\vspace{-0.5ex}\hspace{0.105cm}\includegraphics[height=2.15em]{DCU_2010_Name_Mark_289}
+		\end{minipage}
+
 	}{ %Title (centered top)
@@
 			\end{minipage}
-			\begin{minipage}[t]{11.2em}
+			\begin{minipage}[t]{12em}
 				\begin{spacing}{0.4}
@@
 	}{ %More eye catchers, on the right side
-		\includegraphics[height=6.5em]{uitlogo}
-		\hspace{2.5cm}\begin{minipage}[t]{7em}
-			\vspace{-2cm}
-			\noindent\includegraphics[height=4.2em]{DCU_logo_2col}\\
-			\vspace{-0.5ex}\hspace{0.105cm}\includegraphics[height=2.15em]{DCU_2010_Name_Mark_289}
-		\end{minipage}
-		\hspace{2.5cm}\includegraphics[height=6.25em]{iu_tab-crop}
+		\includegraphics[height=6.25em]{iu_tab-crop}
+		\hspace{2.5cm}\includegraphics[angle=90,height=6.8em]{apertium5b}%\hspace{-2em}
 	}
@@
 
-				- Anderson and Harrison (1999, 22-23)\\
+					- Anderson and Harrison (1999, pp.\ 9, 22-23)\\
+					\includegraphics[width=0.49\textwidth]{img/andersonharrison9}\\
 					\includegraphics[width=0.49\textwidth]{img/andersonharrison22-23}\\
-				- Исхаков, Ф.Г.; Пальмбах (1961, 117-118)\\
+					- Исхаков, Ф.Г.; Пальмбах (1961, pp.\ 117-118)\\
 					\includegraphics[width=0.49\textwidth]{img/isxakovpalmbax117}\\
@@
 				\htwo{Initial implementation}
+					\begin{minipage}[t]{0.5\textwidth}
+					\codeex{"Intervocalic voiced velar deletion"\\
+г:0 <=> :Vow/:0* \_ [ \%>: :Vow ]/:0* ;\\
+\-~\\
+"Intervocalic voiceless velar deletion"\\
+к:0 <=> :Vow/:0* \_ [ \%>: :Vow ]/:0* ;\\
+\-~~~~except\\
+\-~~~~~~~~[ .\#.\ | \%- ] [ ( :Cns* ) ( :Vow* ) :Vow ]/:0 \_ [ \%>:\ :Vow ]/:0* ; %[ :0 - \%>:\ ]* ;
+					}
+					\end{minipage}
 
@@
 
+					(FIXME: has forms of соң in it)
+
+				\htwo{Second attempt at implementation}
+
+					\begin{minipage}[t]{0.5\textwidth}
+					\codeex{\hilitegrey{"Intervocalic voiced velar deletion"}\\
+Cx\hilitegrey{:0 <=> :Vow/:0* \_ [ \%>: :Vow ]/:0* ;}\\
+\-~~~~~where Cx in ( г ң ) ;\\
+\-~\\
+\hilitegrey{"Intervocalic voiceless velar deletion"\\
+к:0 <=> :Vow/:0* \_ [ \%>:\ :Vow ]/:0* ;\\
+\-~~~~except\\
+\-~~~~~~~~[ .\#.\ | \%- ] [ ( :Cns* ) ( :Vow* ) :Vow ]/:0 \_ [ \%>:\ :Vow ]/:0* ;} %[ :0 - \%>:\ ]* ;
+					}
+					\end{minipage}
+
+				\htwo{Resulting frequency list}
+
+					(FIXME: has forms of бажың, түң, чаң in it)
+
+				\htwo{Final implementation}
+
+					\begin{minipage}[t]{0.5\textwidth}
+					\codeex{\hilitegrey{"Intervocalic voiced velar deletion"\\
+Cx:0 <=> :Vow/:0* \_ [ \%>:\ :Vow ]/:0* ;}\\
+\-~~~~~except\\
+\-~~~~~~~~~:Vow \_ [ \%\{☭\%\}: :Vow ]/:0* ; \\
+\-~~~~~\hilitegrey{where Cx in ( г ң ) ;\\
+\-~\\
+"Intervocalic voiceless velar deletion"\\
+к:0 <=> :Vow/:0* \_ [ \%>:\ :Vow ]/:0* ;\\
+\-~~~~except\\
+\-~~~~~~~~[ .\#.\ | \%- ] [ ( :Cns* ) ( :Vow* ) :Vow ]/:0 \_ [ \%>:\ :Vow ]/:0* ;%\\
+%\-~~~~~~~~:Vow \_ [ \%\{☭\%\}:\ :Vow ]/:0* ;
+					}}
+					\end{minipage}
+
+				\htwo{Final frequency list}
+
 			\end{multicols}
 
-			\begin{multicols}{2}
-				\htwo{Examples of correct forms}
+			%\htwo{Examples of correct forms}
 
@@
 
-				\htwo{Final implementation}
-
-				\htwo{Final frequency list}
 			\includegraphics[width=0.4\textwidth]{img/andersonharrison35}\\
-			\end{multicols}
 

### 2015-07-06 2015-compmorphon-morphver/poster.tex
more work on Tuvan velar elision
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61175 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-					- Anderson and Harrison (1999, pp.\ 9, 22-23)\\
+					\begin{itemize}
+						\item Anderson and Harrison (1999, pp.\ 9, 22-23)
+					\end{itemize}
 					\includegraphics[width=0.49\textwidth]{img/andersonharrison9}\\
-					\includegraphics[width=0.49\textwidth]{img/andersonharrison22-23}\\
-					- Исхаков, Ф.Г.; Пальмбах (1961, pp.\ 117-118)\\
-					\includegraphics[width=0.49\textwidth]{img/isxakovpalmbax117}\\
+					\includegraphics[width=0.49\textwidth]{img/andersonharrison22-23}
+					\begin{itemize}
+						\item Исхаков, Ф.Г.; Пальмбах (1961, pp.\ 117-118)
+					\end{itemize}
+					\includegraphics[width=0.49\textwidth]{img/isxakovpalmbax117}
+
+				\htwo{More complete characterisation}
+				\begin{itemize}
+					\item /k/ ‹к› elides at the end of ≥1-{\qipa 𝛔} stems intervocalically;
+					\item /g/ ‹г› elides at the end of any stem intervocalically;
+					\item /ŋ/ ‹ң› elides at the end of \textit{some} stems intervocalically.
+				\end{itemize}
+				examples of correct forms in grammars:\\
+				\includegraphics[width=0.2\textwidth]{img/andersonharrison35}
 
@@
 
-			\htwo{More complete characterisation}
-
-			\includegraphics[width=0.4\textwidth]{img/andersonharrison35}\\
-
 		}

### 2015-07-06 2015-compmorphon-morphver/poster.tex
completed drafts of Morphological Descriptions box and Tuvan velar deletion box
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61186 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \newcommand{\hilitegrey}[1]{{\addfontfeature{Color=77777799}#1}}
+\newcommand{\hilitelightgrey}[1]{{\addfontfeature{Color=CCCCCC99}#1}}
 
@@
 \newfontfamily\litamono[IPA,Scale=0.5]{DejaVu Sans Mono}
+\newfontfamily\ortamono[IPA,Scale=0.625]{DejaVu Sans Mono}
 
@@
 				\vspace{-1em}\end{spacing}
+         }\end{minipage}
+      }
+   }
+}
+\newcommand{\outputex}[1]{
+   \fbox{\colorbox{iumidnight}{
+         \begin{minipage}[t]{0.91\textwidth}{\ortamono
+				\begin{spacing}{0.2}\vspace{-0.8em}
+	            \hilitelightgrey{#1}
+				\vspace{-2.5em}\end{spacing}
          }\end{minipage}
@@
 	\headerbox{Morphological Descriptions}{name=morphdesc,column=0,row=0}{
-		Morphological descriptions describe the \_morphotactics\_¹ and \_morphophonology\_² of a language.
+		\htwo{What are they?}
+		\begin{itemize}
+			\item core part of grammatical descriptions
+			\item found in reference grammars and textbooks
+		\end{itemize}
 
-		These are a core part of grammatical descriptions, such as reference grammars and textbooks.
-Due to restrictions of the medium (such as number of pages), the limitations of introspection and working with native speakers, and the lack of automatic testing, such works are often incomplete.
+		Descriptions of a language's
+		\begin{itemize}
+			\item \textbf{morphotactics} — the morphemes of a language and how they can be combined
+			\item \textbf{morphophonology} — the alternations between the various phonological and orthographical forms of each morpheme in a language
+		\end{itemize}
 
-		¹ Morphotactics refers to the morphemes of a language and how they can be combined.
-
-² Morphophonology refers to the alternation between the various phonological and orthographical forms of a morpheme in a language.
-		\scalebox{0.78}{
-			hargle
-		}
+		\htwo{Reasons for being \hilitetwo{incomplete}}
+		\begin{itemize}
+			\item restrictions of the medium (e.g., number of pages)
+			\item limitations of introspection, working w̄ native speakers
+			\item \textbf{lack of automatic testing}
+		\end{itemize}
 	}
@@
 				\begin{itemize}
-					\item /k/ ‹к› elides at the end of ≥1-{\qipa 𝛔} stems intervocalically;
-					\item /g/ ‹г› elides at the end of any stem intervocalically;
-					\item /ŋ/ ‹ң› elides at the end of \textit{some} stems intervocalically.
+					\item /k/ ‹к› elides at the end of >1-σ stems intervocalically\\
+						инэк+I → инээ, өк+I → өгү
+					\item /g/ ‹г› elides at the end of any stem intervocalically\\
+						өг+I → өө 
+					\item /ŋ/ ‹ң› elides at the end of \textbf{some} stems intervocalically.\\
+						\textbf{соң+I → соо}, чаң+I → чаңы, түң+I → түңү
 				\end{itemize}
-				examples of correct forms in grammars:\\
-				\includegraphics[width=0.2\textwidth]{img/andersonharrison35}
+
+				\htwo{Correct forms in the grammars}
+
+				But not part of the morphophonological description.\\
+				(A. \& H., p.\ 35)\hfill{}\includegraphics[width=0.2\textwidth]{img/andersonharrison35}\\
+				(И.\ \& П., p.\ 447)\hfill{}\includegraphics[width=0.35\textwidth]{img/isxakovpalmbax447}
 
 			\columnbreak{}
+				\htwo{Uanalysed forms before implementation}
+					\begin{itemize}
+						\item \hilitetwo{инээ}, \hiliteone{өгү}, \hilitetwo{өө}, \hilitetwo{соо}, \hiliteone{чаңы}
+					\end{itemize}
+					\begin{minipage}[t]{0.5\textwidth}
+					\outputex{
+						\begin{multicols}{4}
+							1590 соонда\vphantom{ү}\\
+							610 те\vphantom{ү}\\
+							153 өөнүң\vphantom{ү}\\
+							48 соондан\vphantom{ү}\\
+							46 өөм\vphantom{ү}\\
+							20 соон\vphantom{ү}\\
+							18 өө\vphantom{ү}\\
+							17 өөнге\vphantom{ү}\\
+							9 соондагы\vphantom{ү}\\
+							8 өөн\vphantom{ү}\columnbreak\\
+							6 инээниң\vphantom{ү}\\
+							6 соонче\vphantom{ү}\\
+							6 соонга\vphantom{ү}\\
+							3 инээм\vphantom{ү}\\
+							3 өөнден\vphantom{ү}\\
+							3 өөнде\vphantom{ү}\\
+							3 өөмнүң\vphantom{ү}\\
+							2 инээн\vphantom{ү}\\
+							2 инээ\vphantom{ү}\\
+							2 өөнче\vphantom{ү}\columnbreak\\
+							2 өөңерге\vphantom{ү}\\
+							2 өөңер\vphantom{ү}\\
+							2 өөмнү\vphantom{ү}\\
+							2 өөвүстен\vphantom{ү}\\
+							2 өөвүске\vphantom{ү}\\
+							2 өөвүс\vphantom{ү}\\
+							1 инээңни\vphantom{ү}\\
+							1 инээмниң\vphantom{ү}\\
+							1 Инээм\vphantom{ү}\\
+							1 инээвистиң\vphantom{ү}\\
+							1 өөңнүң\vphantom{ү}\\
+							1 өөң\vphantom{ү}\\
+							1 өөвүсче\vphantom{ү}\\
+							1 өөвүстүң\vphantom{ү}\\
+							1 өөвүстү\vphantom{ү}\\
+							1 өөвүсте
+						\end{multicols}}
+					\end{minipage}
+
 				\htwo{Initial implementation}
@@
 
-				\htwo{Initial frequency list}
-
-					(FIXME: has forms of соң in it)
+				\htwo{Unanalysed forms after initial implementation}
+					%(has forms of соң in it)
+					\begin{itemize}
+						\item \hiliteone{инээ}, \hiliteone{өгү}, \hiliteone{өө}, \hilitetwo{соо}, \hiliteone{чаңы}
+					\end{itemize}
+					\begin{minipage}[t]{0.5\textwidth}
+					\outputex{
+						\begin{multicols}{3}
+							1590 соонда\vphantom{ң}\\
+							610 те\vphantom{ң}\\
+							48 соондан\vphantom{ң}\\
+							20 соон\vphantom{ң}\\
+							6 соонче\vphantom{ң}\\
+							6 соонга
+						\end{multicols}
+					}
+					\end{minipage}
 
@@
 
-				\htwo{Resulting frequency list}
+				\htwo{Unanalysed forms after second implementation}
+					\begin{itemize}
+						\item \hiliteone{инээ}, \hiliteone{өгү}, \hiliteone{өө}, \hiliteone{соо}, \hilitetwo{чаңы}
+					\end{itemize}
+					\begin{minipage}[t]{0.5\textwidth}
+					\outputex{
+						\begin{multicols}{3}
+							610 те\vphantom{ү}\\
+							203 бажыңынга\vphantom{ү}\\
+							164 бажыңының\vphantom{ү}\\
+							102 бажыңы\vphantom{ү}\\
+							86 бажыңын\vphantom{ү}\\
+							56 түңү\vphantom{ү}\\
+							47 чаңы\vphantom{ү}\\
+							44 чаңын\vphantom{ү}\\
+							13 түңүн\vphantom{ү}\\
+							12 чаңының\vphantom{ү}\\
+							5 түңүнден\vphantom{ү}\\
+							4 чаңындан\vphantom{ү}\\
+							4 чаңында\vphantom{ү}\\
+							3 түңүнүң\vphantom{ү}\\
+							2 түңүнге
+						\end{multicols}}
+					\end{minipage}
 
-					(FIXME: has forms of бажың, түң, чаң in it)
+					%(has forms of бажың, түң, чаң in it)
 
@@
 
-				\htwo{Final frequency list}
-
+				\htwo{Unanalysed forms after final implementation}
+					\begin{itemize}
+						\item \hiliteone{инээ}, \hiliteone{өгү}, \hiliteone{өө}, \hiliteone{соо}, \hiliteone{чаңы}
+					\end{itemize}
+					\begin{minipage}[t]{0.5\textwidth}
+					\outputex{
+						\begin{multicols}{3}
+							610 те\\
+						\end{multicols}}
+					\end{minipage}
 			\end{multicols}

### 2015-07-07 2015-compmorphon-morphver/poster.tex
shrink the fonts just slightly
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61187 72bbbca6-d526-0410-a7d9-f06f51895060

@@
-\documentclass[fontscale=0.3,landscape,paperwidth=48in,paperheight=36in]{baposter}  % fontscale=0.285, dvipdfm
+\documentclass[fontscale=0.33,landscape,paperwidth=48in,paperheight=36in]{baposter}  % fontscale=0.285, dvipdfm
 
@@
 \setromanfont[IPA]{Linux Libertine O}
+%\setromanfont[IPA]{Nimbus Roman No9 L}
 \setmonofont[IPA]{DejaVu Sans Mono}
@@
 			\item restrictions of the medium (e.g., number of pages)
-			\item limitations of introspection, working w̄ native speakers
+			\item limitations of introspection, working with native speakers
 			\item \textbf{lack of automatic testing}

### 2015-07-07 2015-compmorphon-morphver/poster.tex
moving some stuff around, beginning of next example box
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61188 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 						\item \hilitetwo{инээ}, \hiliteone{өгү}, \hilitetwo{өө}, \hilitetwo{соо}, \hiliteone{чаңы}
-					\end{itemize}
+					\end{itemize}\vspace{-0.3ex}
 					\begin{minipage}[t]{0.5\textwidth}
@@
 						\item \hiliteone{инээ}, \hiliteone{өгү}, \hiliteone{өө}, \hilitetwo{соо}, \hiliteone{чаңы}
-					\end{itemize}
+					\end{itemize}\vspace{-0.3ex}
 					\begin{minipage}[t]{0.5\textwidth}
@@
 						\item \hiliteone{инээ}, \hiliteone{өгү}, \hiliteone{өө}, \hiliteone{соо}, \hilitetwo{чаңы}
-					\end{itemize}
+					\end{itemize}\vspace{-0.3ex}
 					\begin{minipage}[t]{0.5\textwidth}
@@
 						\item \hiliteone{инээ}, \hiliteone{өгү}, \hiliteone{өө}, \hiliteone{соо}, \hiliteone{чаңы}
-					\end{itemize}
+					\end{itemize}\vspace{-0.3ex}
 					\begin{minipage}[t]{0.5\textwidth}
@@
 
-		\headerbox{Examples:\ ideas for other examples}{name=otherexamples,column=1,span=2,below=tuvanvelars}{
+		\headerbox{Example: Tuvan vowel harmony after consonant clusters in loanwords}{name=otherexamples,column=1,span=2,below=tuvanvelars}{
+			\begin{minipage}{0.5\textwidth}
+				\codeex{"\{I\} harmony"\\
+\%\{I\%\}:Vy <=> :Vx [ :Cns* :RealCns ]/[ :0 | \%- ]* \_ ;\\
+\-~~~~~~~~~where Vx in ( ү и е э ө а о ы у я ё ю )\\
+\-~~~~~~~~~~~~~~ Vy in ( ү и и и ү ы у ы у ы у у )\\
+\-~~~~~~~~~matched ;}
+			\end{minipage}
+
+			\begin{minipage}{0.5\textwidth}
+				\codeex{"\{I\} harmony"\\
+\%\{I\%\}:Vy <=> :Vx [ :Cns* :RealCns ]/[ :0 | \%- ]* \_ ;\\
+\-~~~~~~~~~except\\
+\-~~~~~~~~~~~~~[ :BackVow :Cns* :Cns л: ь: :Cns* :RealCns ]/:0* \_ ;\\
+\-~~~~~~~~~~~~~[ :BackVow :Cns* :Cns л: ь:0 ]/:[ :0 - ь: ]* \_ ;\\
+\-~~~~~~~~~where Vx in ( ү и е э ө а о ы у я ё ю )\\
+\-~~~~~~~~~~~~~~ Vy in ( ү и и и ү ы у ы у ы у у )\\
+\-~~~~~~~~~matched ;\\
+\\
+"\{I\} always front when intervening Cль"\\
+\%\{I\%\}:и <=> [ :BackVow :Cns* :Cns л: ь: :Cns* :RealCns ]/:0* \_ ;\\
+\-~~~~~~~~~~~~[ :BackVow :Cns* :Cns л: ь:0 ]/:[ :0 - ь: ]* \_ ;}
+			\end{minipage}
+
+		}
+
+		\headerbox{Examples:\ ideas for other examples}{name=otherexamples,column=0,span=1,below=tuvanvelars}{ %{name=otherexamples,column=1,span=2,below=tuvanvelars}{
 			\begin{itemize}
 				\item Kyrgyz: conditions where л desonorises to д
-				\item Tuvan: рубль - рубльдиң *рубьдуң
+				\item Tuvan: рубль - рубльдиң *рубльдуң
 				\item clitics can go everywhere/anywhere
@@
 
-		\headerbox{Morphophonology}{name=morphophon,column=1,below=otherexamples}{
+		\headerbox{Morphophonology}{name=morphophon,column=0,below=otherexamples}{
 

### 2015-07-07 2015-compmorphon-morphver/poster.tex
some work on the loanword clusters
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61189 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-		Descriptions of a language's
+		\htwo{What they describe} %Describe a language's
 		\begin{itemize}
 			\item \textbf{morphotactics} — the morphemes of a language and how they can be combined
-			\item \textbf{morphophonology} — the alternations between the various phonological and orthographical forms of each morpheme in a language
+			\item \textbf{morphophonology} — the alternations between the various phonological and orthographical forms of each morpheme% in a language
 		\end{itemize}
@@
 			\item morphotactics implemented in \texttt{lexc}
-			\item morphophonology implemented in \texttt{twol}
+			\item morphophonology implemented in \texttt{twol} (SPE-style rules)
 			\item compiled separately; compose-intersected to single transducer\vspace{-0.1ex}\\
@@
 		\headerbox{Example: Tuvan vowel harmony after consonant clusters in loanwords}{name=otherexamples,column=1,span=2,below=tuvanvelars}{
+			\begin{multicols}{2}
+				\htwo{Tuvan vowel harmony}
+					\begin{itemize}
+						\item High and low vowels harmonise in backness
+						\item High vowels also harmonise in roundness
+					\end{itemize}
+				\htwo{Initial implementation (high-vowels)}
+
 				\begin{minipage}{0.5\textwidth}
@@
 
+				\htwo{Unanalysed forms in corpus}
+					\begin{itemize}
+						\item forms of loanwords ансамбль `ensemble' and рубль `rouble':\\\
+						\-\hfill{}the following vowel is always front unround
+					\end{itemize}
+					\begin{minipage}[t]{0.5\textwidth}
+					\outputex{
+						\begin{multicols}{3}
+							%790 рубль\\
+							347 рубльди\\
+							99 рубльге\\
+							83 ансамблиниң\vphantom{у}\\
+							60 ансамбли\vphantom{у}\\
+							54 рубльдиң\\
+							28 рубльден\\
+							17 ансамбльдиң\vphantom{у}\\
+							15 ансамбльдер\vphantom{у}\\
+							14 ансамбль\vphantom{у}\\
+							11 ансамблин\vphantom{у}\\
+							10 ансамблинге\vphantom{у}\\
+							8 ансамбльди\vphantom{у}\\
+							%2 рубльга\\
+							1 рубльдерни\\
+							1 рублин\\
+							1 рубли\\
+						\end{multicols}
+					}
+					\end{minipage}
+
+					\columnbreak
+				\htwo{Revised generalisation}
+					\begin{itemize}
+						\item After cons.\ cluster ending in ль, vowel always front unround
+						\item New discovery?  Appears to be not previously documented
+					\end{itemize}
+
+				\htwo{Revised implementation}
 				\begin{minipage}{0.5\textwidth}
-				\codeex{"\{I\} harmony"\\
-\%\{I\%\}:Vy <=> :Vx [ :Cns* :RealCns ]/[ :0 | \%- ]* \_ ;\\
+					\codeex{\hilitegrey{"\{I\} harmony"\\
+\%\{I\%\}:Vy <=> :Vx [ :Cns* :RealCns ]/[ :0 | \%- ]* \_ ;}\\
 \-~~~~~~~~~except\\
@@
 \-~~~~~~~~~~~~~[ :BackVow :Cns* :Cns л: ь:0 ]/:[ :0 - ь: ]* \_ ;\\
-\-~~~~~~~~~where Vx in ( ү и е э ө а о ы у я ё ю )\\
+\-~~~~~~~~~\hilitegrey{where Vx in ( ү и е э ө а о ы у я ё ю )\\
 \-~~~~~~~~~~~~~~ Vy in ( ү и и и ү ы у ы у ы у у )\\
-\-~~~~~~~~~matched ;\\
+\-~~~~~~~~~matched ;}\\
 \\
@@
 				\end{minipage}
+			\end{multicols}
 

### 2015-07-07 2015-compmorphon-morphver/poster.tex
Makefile updates and list of Turkic transducers
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61190 72bbbca6-d526-0410-a7d9-f06f51895060
minor fix
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61191 72bbbca6-d526-0410-a7d9-f06f51895060
some rearranging and more fixes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61207 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \newcommand{\hilitelightgrey}[1]{{\addfontfeature{Color=CCCCCC99}#1}}
+\newcommand{\hiliteblue}[1]{{\addfontfeature{Color=44697DFF}#1}}
 
@@
 \newcommand{\tags}[1]{\hiliteone{#1}}
-\newcommand{\tag}[1]{\hilitegrey{<}\hiliteone{#1}\hilitegrey{>}}
+\newcommand{\tag}[1]{{\ortamono \hilitegrey{<}\hiliteblue{#1}\hilitegrey{>}}}
 \newcommand{\archiphon}[1]{\hilitegrey{\{}\hiliteone{#1}\hilitegrey{\}}}
@@
 
-		\htwo{Transducers for Turkic languages}
+		\htwo{Turkic-language transducers we've made}
 		\begin{itemize}
-			\item Turkish (\hiliteone{Çöltekin, 2010 \& 2014}; \hilitetwo{Oflazer, 1994})
-			\item Crimean Tatar (\hilitetwo{Altıntaş, 2001})
-			\item Turkmen (\hilitetwo{Tantuğ et al., 2006})
-			\item Kazakh (\hilitetwo{Бекманова \& Махимов, 2013})
+			%\item Turkish (\hiliteone{Çöltekin, 2010 \& 2014}; \hilitetwo{Oflazer, 1994})
+			%\item Crimean Tatar (\hilitetwo{Altıntaş, 2001})
+			%\item Turkmen (\hilitetwo{Tantuğ et al., 2006})
+			%\item Kazakh (\hilitetwo{Бекманова \& Махимов, 2013})
 			%\vspace{1pt}\hfill{} \hiliteone{GPL (=free and open)}!
-			\item our Kyrgyz, Kazakh, Tatar, Kumyk:\ all \hiliteone{GPL (=free and open)}!
+			%\item our Kyrgyz, Kazakh, Tatar, Kumyk:\ all \hiliteone{GPL (=free and open)}!
+			\item Kyrgyz (Washington et al., 2012)
+			\item Tatar \& Bashkort (Tyers et al., 2012)
+			\item Kazakh (Salimzyanov et al., 2013)
+			\item Kumyk (Washington et al., 2014)
+			\item ongoing work on Sakha, Tuvan, Khakas, and more!
 		\end{itemize}
@@
 
-	\headerbox{Categorisation}{name=adjectives,column=3}{
-	For example, many grammars of Turkic languages (e.g., Somfai Kara, 2002) overgeneralise adjective morphology, stating that all adjectives may act like nouns (and take nominal morphology), or that all adjectives may be used as adverbs with no morphological changes.
-We have found that there is a range of “adjective classes” in most Turkic languages, including some which may not be substantivised and some which may not be used adverbially.
-This phenomenon is of particular interest because native-speaker intuitions on the morphological limitations of a given adjective can often be more restrictive than the range of possible uses found in a large text corpus.
-		\htwo{Adjectives}
+	\headerbox{Example: Categorisation}{name=adjectives,column=3}{
+		\htwo{Unrestricted 0-derivation}
 		\begin{itemize}
-				\item morphologically distinct adjective classes
-				\item most sources claim:\ adjectives can be used substantively and adverbially
-				\item Other Turkic transducers:\ \texttt{0}-derivation (\hilitetwo{overgenerates})
-				\item but not all adjectives have all of the following:\\
-				comparative forms, substantive readings, adverbial readings
-				\item Our approach:\ categorisation 
+			\item Common overgeneralisation of adjective morphology:
+			\begin{itemize}
+				\item all adjectives may act like nouns (i.e.\ take noun morphology)
+				\item all adjectives may be used as is as adverbs
+			\end{itemize}
+			\item Commonly reported in Turkic language grammars
+		\end{itemize}
+		Somfai Kara (2002, pp. 28-29):\ \includegraphics[width=0.53\textwidth]{img/kara28}\\
+		\-\hfill{}\includegraphics[width=0.8\textwidth]{img/kara29}
+
+		\htwo{Attested system}
+			\begin{itemize}
+				\item Some adjectives may not be substantivised
+				\item Some adjectives may not be used adverbially
+				\item Some adjectives do not have comparative forms
+				\item …a range of “adjective classes” in most Turkic languages
+			\end{itemize}
+		\htwo{Implementation}
+			\begin{itemize}
 				\item if properly categorised, only correct forms are analysed and generated
@@
 
-			\scalebox{0.7}{
+			\scalebox{0.83}{
 				\begin{tabular}{lllll}
@@
 				\end{tabular}
+			}\\
+			This phenomenon is of particular interest because native-speaker intuitions on the morphological limitations of a given adjective can often be more restrictive than the range of possible uses found in a large text corpus.
 	}
-
-		\htwo{Adverbs}
-			\begin{itemize}
-				\item Certain adverbs have special attributive and ablative forms
-				\item Mostly time adverbs
-				%\item Mostly time adverbs, e.g.\\
-				%{\qipa бүгүн} \eng{today}, {\qipa быйыл} \eng{this year}, {\qipa кечээ} \eng{yesterday}, {\qipa жана} \eng{just now}
-				%\item Attributive forms:\\
-				%{\qipa бүгүнкү} \eng{today's}, {\qipa быйылкы} \eng{this year's}, {\qipa кечээги} \eng{yesterday's}, {\qipa жанагы} \eng{just now}
-				%\item Ablative forms:\\
-				%{\qipa бүгүнтөн} \eng{from today}, {\qipa быйылтан} \eng{from this year}, {\qipa кечээтен} \eng{from yesterday}, {\qipa жанатан} \eng{from just now}
-				\item Some also have noun readings:\ regular ablative, other cases:
-				%{\qipa бүгүндөн} \eng{from today}, {\qipa быйылдан} \eng{from this year}
-			\end{itemize}
-			\scalebox{0.83}{
-			\begin{tabular}{lllll}
-				\toprule
-				word & {\qipa бүгүн} & {\qipa быйыл} & {\qipa кечээ} &  {\qipa жана}\\
-				gloss & \eng{today} & \eng{this year} & \eng{yesterday} & \eng{just now} \\
-				\midrule
-				\tag{attr} form & {\qipa бүгүнкү} & {\qipa быйылкы} & {\qipa кечээги} &  {\qipa жанагы}\\
-				\tag{adv}\tag{abl} form & {\qipa бүгүнтөн} & {\qipa быйылтан} & {\qipa кечээтен} &  {\qipa жанатан}\\
-				\tag{n}\tag{abl} form & {\qipa бүгүндөн} & {\qipa быйылдан} & — & — \\
-				\bottomrule
-			\end{tabular}
-			}\vspace{0.2em}
-				\codeex{
-					\begin{spacing}{0.8}
-					\texttt{LEXICON ADV-WITH-KI-ABL\\
-\\
-ADV ;\\
-ADV-KI ;\\
-ADV-ABL ;
-					}
-					\end{spacing}
-					\vspace{0em}
-				}
-	}
-
-	\headerbox{Ongoing and future work}{name=nedostatki,below=adjectives,column=3}{
-		\begin{itemize}
-			\item Disambiguation, more stems, clean up transducers
-			\item Machine translation between these languages
-			\item Bring other Kypchak transducers to comparable performance:\\
-			Qaraqalpaq, Bashqort, Nogay, Crimean Tatar
-			\item Other Turkic lgs:\ Uzbek, Uyghur, Chuvash, Yakut, Tuvan, etc.
-		\end{itemize}
-	}
-
-	\headerbox{Further information}{name=getting,column=3,below=nedostatki}{
-		\begin{itemize}
-			\item Part of \textbf{Apertium Turkic} project:\\
-			\url{http://wiki.apertium.org/wiki/Apertium\_Turkic}
-			\item Transducers available \textbf{live} at \href{http://turkic.apertium.org/}{\texttt{turkic.apertium.org}}
-			\item \textbf{Source code} available from Apertium's svn repo
-			%info at {\small \url{http://wiki.apertium.org/wiki/apertium-kir}}
-				\item Turkic RBMT \textbf{mailing list} (>25 subscribers):\\
-			\texttt{apertium-turkic@lists.sourceforge.net}\\
-			\vspace{-0.5ex}Feel free to post in any language!\\\vspace{-2.5ex}
-			\item See our papers in LREC proceedings\\
-			(2012:\ Kyrgyz, 2014:\ Kazakh, Tatar, Kumyk)
-			\item And feel free to contact the authors any time!
-		\end{itemize}
-	}
-
-%	\headerbox{References}{name=references,column=2,below=nedostatki}{
-%		%\begin{thebibliography}{1}\itemsep=-0.01em
-%		\setlength{\baselineskip}{0.4em}
-%	}
 
@@
 
-		\headerbox{Example: Tuvan vowel harmony after consonant clusters in loanwords}{name=otherexamples,column=1,span=2,below=tuvanvelars}{
-			\begin{multicols}{2}
+	\headerbox{Ongoing and future work}{name=nedostatki,below=tuvanvelars,column=2}{
+		\begin{itemize}
+			\item Disambiguation, more stems, clean up transducers
+			\item Machine translation between these languages
+			\item Bring other Kypchak transducers to comparable performance:\\
+			Qaraqalpaq, Bashqort, Nogay, Crimean Tatar
+			\item Other Turkic lgs:\ Uzbek, Uyghur, Chuvash, Yakut, Tuvan, etc.
+		\end{itemize}
+	}
+
+	\headerbox{Further information}{name=getting,column=2,below=nedostatki}{
+		\begin{itemize}
+			\item Part of \textbf{Apertium Turkic} project:\\
+			\url{http://wiki.apertium.org/wiki/Apertium\_Turkic}
+			\item Transducers available \textbf{live} at \href{http://turkic.apertium.org/}{\texttt{turkic.apertium.org}}
+			\item \textbf{Source code} available from Apertium's svn repo
+			%info at {\small \url{http://wiki.apertium.org/wiki/apertium-kir}}
+				\item Turkic RBMT \textbf{mailing list} (>25 subscribers):\\
+			\texttt{apertium-turkic@lists.sourceforge.net}\\
+			\vspace{-0.5ex}Feel free to post in any language!\\\vspace{-2.5ex}
+			\item See our papers in LREC proceedings\\
+			(2012:\ Kyrgyz, 2014:\ Kazakh, Tatar, Kumyk)
+			\item And feel free to contact the authors any time!
+		\end{itemize}
+	}
+
+%	\headerbox{References}{name=references,column=2,below=nedostatki}{
+%		%\begin{thebibliography}{1}\itemsep=-0.01em
+%		\setlength{\baselineskip}{0.4em}
+%	}
+
+		\headerbox{Example: Tuvan vowel harmony}{name=otherexamples,column=3,below=adjectives}{ %{name=otherexamples,column=1,span=2,below=tuvanvelars}{ % after consonant clusters in loanwords
+			%\begin{multicols}{2}
 				\htwo{Tuvan vowel harmony}
 					\begin{itemize}
-						\item High and low vowels harmonise in backness
-						\item High vowels also harmonise in roundness
+						\item High and low vowels harmonise in backness\\
+						киж\hiliteone{и}\tag{n}\tag{pl}\tag{acc}: киж\hiliteone{и}л\hiliteone{е}рн\hiliteone{и}, н\hilitetwo{о}м\tag{n}\tag{pl}\tag{acc}: н\hilitetwo{о}мн\hilitetwo{а}рн\hilitetwo{ы}
+						\item High vowels also harmonise in roundness\\
+						round: \hiliteone{ө}г\tag{n}\tag{acc}:\ \hiliteone{ө}гн\hiliteone{ү}, \hilitetwo{о}й\tag{n}\tag{acc}:\ \hilitetwo{о}йн\hilitetwo{у}\\
+						unround: ү\hiliteone{е}\tag{n}\tag{acc}:\ ү\hiliteone{е}н\hiliteone{и}, \hilitetwo{а}й\tag{n}\tag{acc}:\ \hilitetwo{а}йн\hilitetwo{ы}
 					\end{itemize}
@@
 
-				\begin{minipage}{0.5\textwidth}
+				%\begin{minipage}{0.5\textwidth}
 					\codeex{"\{I\} harmony"\\
@@
 \-~~~~~~~~~matched ;}
-				\end{minipage}
+				%\end{minipage}
 
@@
 					\end{itemize}
-					\begin{minipage}[t]{0.5\textwidth}
+					%\begin{minipage}[t]{0.5\textwidth}
 					\outputex{
@@
 					}
-					\end{minipage}
+					%\end{minipage}
 
-					\columnbreak
+					%\columnbreak
 				\htwo{Revised generalisation}
@@
 						\item After cons.\ cluster ending in ль, vowel always front unround
-						\item New discovery?  Appears to be not previously documented
+						\item New discovery?  Appears not to be previously documented
 					\end{itemize}
@@
 				\htwo{Revised implementation}
-				\begin{minipage}{0.5\textwidth}
+				%\begin{minipage}{0.5\textwidth}
 					\codeex{\hilitegrey{"\{I\} harmony"\\
@@
 \-~~~~~~~~~~~~[ :BackVow :Cns* :Cns л: ь:0 ]/:[ :0 - ь: ]* \_ ;}
-				\end{minipage}
-			\end{multicols}
+				%\end{minipage}
+			%\end{multicols}
 
@@
 				\item Kyrgyz: conditions where л desonorises to д
-				\item Tuvan: рубль - рубльдиң *рубльдуң
+				%\item Tuvan: рубль - рубльдиң *рубльдуң
 				\item clitics can go everywhere/anywhere
 				\item underformalised - “particles☹” can go anywhere
-				\item all adjs act like nouns (Somfai-Kara, p. 28)
-				\item all adjectives can be used as adverbs without any morphological changes (Somfai-Kara, p. 29)
+				%\item all adjs act like nouns (Somfai-Kara, p. 28)
+				%\item all adjectives can be used as adverbs without any morphological changes (Somfai-Kara, p. 29)
 				\item Stuff found doing treebank

### 2015-07-08 2015-compmorphon-morphver/poster.tex
more work
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61208 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-\usepackage{natbib}
+%\usepackage{natbib}
+\usepackage[backend=biber,citestyle=apa,style=authoryear,maxbibnames=99,maxcitenames=2]{biblatex}
+\bibliography{poster}
+\renewcommand*{\nameyeardelim}{\addcomma\space}
 
@@
 
-\newcommand{\citemultileft}[1]{(\citeauthor{#1}, \citeyear{#1}}
-\newcommand{\citemultimid}[1]{\citeauthor{#1}, \citeyear{#1}}
-\newcommand{\citemultiright}[1]{\citeauthor{#1}, \citeyear{#1})}
-\newcommand{\citetwoyears}[2]{\citeauthor{#1} (\citeyear{#1} and \citeyear{#2})}
+%\newcommand{\citemultileft}[1]{(\citeauthor{#1}, \citeyear{#1}}
+%\newcommand{\citemultimid}[1]{\citeauthor{#1}, \citeyear{#1}}
+%\newcommand{\citemultiright}[1]{\citeauthor{#1}, \citeyear{#1})}
+%\newcommand{\citetwoyears}[2]{\citeauthor{#1} (\citeyear{#1} and \citeyear{#2})}
 
@@
 
-	\headerbox{Morphological transducers}{name=morphtrans,below=morphdesc}{
+	\headerbox{Morphological Transducers}{name=morphtrans,below=morphdesc}{
 		\htwo{Morphological transducers}
@@
 			%\item our Kyrgyz, Kazakh, Tatar, Kumyk:\ all \hiliteone{GPL (=free and open)}!
-			\item Kyrgyz (Washington et al., 2012)
-			\item Tatar \& Bashkort (Tyers et al., 2012)
-			\item Kazakh (Salimzyanov et al., 2013)
-			\item Kumyk (Washington et al., 2014)
+			\item Kyrgyz \parencite{jnwmifmt2012lrecpaper}%  (Washington et al., 2012)
+			\item Tatar \& Bashkort \parencite{fmtjnwisrb2012lrecpaper} %(Tyers et al., 2012)
+			\item Kazakh \parencite{isjnwfmt2013mtsummitpaper} %(Salimzyanov et al., 2013)
+			\item Kumyk \parencite{jnwisfmt2014lrecpaper} %(Washington et al., 2014)
 			\item ongoing work on Sakha, Tuvan, Khakas, and more!
@@
 		\begin{itemize}
+			\item two-level method \parencite{koskenniemi83}
 			\item morphotactics implemented in \texttt{lexc}
@@
 
-	\headerbox{Example: Categorisation}{name=adjectives,column=3}{
-		\htwo{Unrestricted 0-derivation}
-		\begin{itemize}
-			\item Common overgeneralisation of adjective morphology:
-			\begin{itemize}
-				\item all adjectives may act like nouns (i.e.\ take noun morphology)
-				\item all adjectives may be used as is as adverbs
-			\end{itemize}
-			\item Commonly reported in Turkic language grammars
-		\end{itemize}
-		Somfai Kara (2002, pp. 28-29):\ \includegraphics[width=0.53\textwidth]{img/kara28}\\
-		\-\hfill{}\includegraphics[width=0.8\textwidth]{img/kara29}
-
-		\htwo{Attested system}
-			\begin{itemize}
-				\item Some adjectives may not be substantivised
-				\item Some adjectives may not be used adverbially
-				\item Some adjectives do not have comparative forms
-				\item …a range of “adjective classes” in most Turkic languages
-			\end{itemize}
-		\htwo{Implementation}
-			\begin{itemize}
-				\item if properly categorised, only correct forms are analysed and generated
-			\end{itemize}
-
-			\scalebox{0.83}{
-				\begin{tabular}{lllll}
-					\toprule
-					\textbf{Type} & \textbf{Gloss} & {\small \texttt{\tag{adj}(\tag{comp})}} & {\small \texttt{\tag{adj}(\tag{comp})\tag{subst}}} & {\small \texttt{\tag{adj}(\tag{comp})\tag{advl}}} \\
-					\midrule
-					\textbf{A1} & \eng{good} & {\qipa яхшы (яхшырак)} & {\qipa яхшы (яхшырак)} & {\qipa яхшы (яхшырак)} \\
-					\textbf{A2} & \eng{old} & {\qipa иске (искерәк)} & {\qipa иске (искерәк)} & {\qipa — (—)} \\
-					\textbf{A3} & \eng{dead} & {\qipa үле (—)} & {\qipa үле (—)} & {\qipa — (—)} \\
-					\textbf{A4} & \eng{basic} & {\qipa төп (—)} & {\qipa — (—)} & {\qipa — (—)} \\
-					\bottomrule
-				\end{tabular}
-			}\\
-			This phenomenon is of particular interest because native-speaker intuitions on the morphological limitations of a given adjective can often be more restrictive than the range of possible uses found in a large text corpus.
-	}
-
 		\headerbox{Example:\ Tuvan velar elision}{name=tuvanvelars,column=1,span=2}{
@@
 					\begin{itemize}
-						\item Anderson and Harrison (1999, pp.\ 9, 22-23)
+						\item \textcite[22-23]{andersonharrison99} %Anderson and Harrison (1999, pp.\ 9, 22-23)
 					\end{itemize}
@@
 					\begin{itemize}
-						\item Исхаков, Ф.Г.; Пальмбах (1961, pp.\ 117-118)
+						\item \textcite[117-118]{isxakovpalmbax61} %Исхаков, Ф.Г.; Пальмбах (1961, pp.\ 117-118)
 					\end{itemize}
@@
 
-	\headerbox{Ongoing and future work}{name=nedostatki,below=tuvanvelars,column=2}{
+	\headerbox{Example: Categorisation}{name=adjectives,column=1,below=tuvanvelars}{
+		\htwo{Unrestricted 0-derivation}
+		\begin{itemize}
+			\item Common overgeneralisation of adjective morphology:
+			\begin{itemize}
+				\item all adjectives may act like nouns (i.e.\ take noun morphology)
+				\item all adjectives may be used as is as adverbs
+			\end{itemize}
+			\item Commonly reported in Turkic language grammars
+		\end{itemize}
+		%Somfai Kara (2002, pp. 28-29):\ \includegraphics[width=0.53\textwidth]{img/kara28}\\
+		\textcite[28-29]{somfaikara02}:\ \includegraphics[width=0.53\textwidth]{img/kara28}\\
+		\-\hfill{}\includegraphics[width=0.8\textwidth]{img/kara29}
+
+		\htwo{Attested system}
+			\begin{itemize}
+				\item Some adjectives may not be substantivised
+				\item Some adjectives may not be used adverbially
+				\item Some adjectives do not have comparative forms
+				\item …a range of “adjective classes” in most Turkic languages
+			\end{itemize}
+		\htwo{Implementation: proper categorisation}
+			\begin{itemize}
+				\item only correct forms are analysed and generated
+			\end{itemize}
+
+			\scalebox{0.83}{
+				\begin{tabular}{lllll}
+					\toprule
+					\textbf{Type} & \textbf{Gloss} & {\small \texttt{\tag{adj}(\tag{comp})}} & {\small \texttt{\tag{adj}(\tag{comp})\tag{subst}}} & {\small \texttt{\tag{adj}(\tag{comp})\tag{advl}}} \\
+					\midrule
+					\textbf{A1} & \eng{good} & {\qipa яхшы (яхшырак)} & {\qipa яхшы (яхшырак)} & {\qipa яхшы (яхшырак)} \\
+					\textbf{A2} & \eng{old} & {\qipa иске (искерәк)} & {\qipa иске (искерәк)} & {\qipa — (—)} \\
+					\textbf{A3} & \eng{dead} & {\qipa үле (—)} & {\qipa үле (—)} & {\qipa — (—)} \\
+					\textbf{A4} & \eng{basic} & {\qipa төп (—)} & {\qipa — (—)} & {\qipa — (—)} \\
+					\bottomrule
+				\end{tabular}
+			}\\
+			This phenomenon is of particular interest because native-speaker intuitions on the morphological limitations of a given adjective can often be more restrictive than the range of possible uses found in a large text corpus.
+	}
+
+	\headerbox{Ongoing and future work}{name=nedostatki,column=2,below=tuvanvelars}{
 		\begin{itemize}
@@
 
-		\headerbox{Example: Tuvan vowel harmony}{name=otherexamples,column=3,below=adjectives}{ %{name=otherexamples,column=1,span=2,below=tuvanvelars}{ % after consonant clusters in loanwords
+		\headerbox{Example: Tuvan vowel harmony}{name=tuvanvowelharmony,column=3}{ %{name=otherexamples,column=1,span=2,below=tuvanvelars}{ % after consonant clusters in loanwords
 			%\begin{multicols}{2}
@@
 
-		\headerbox{Morphophonology}{name=morphophon,column=0,below=otherexamples}{
+%		\headerbox{Morphophonology}{name=morphophon,column=0,below=otherexamples}{
+%
+%			\htwo{Desonorisation (kaz \& kir)}
+%				\begin{itemize}
+%					\item {\texttt \{N\}} desonorises to д after a consonant\\
+%						алма-\hiliteone{\{N\}}\{I\} → алма\hiliteone{н}ы \eng{apple--\gmk{ACC}} \\
+%						сыр-\hiliteone{\{N\}}\{I\} → сыр\hiliteone{д}ы \eng{secret--\gmk{ACC}}
+%					\item {\texttt \{L\}} desonorises to д after cons.\ of sonority ≤ /l/ \\
+%						сыр-\hiliteone{\{L\}}\{A\}р → сыр\hiliteone{л}ар \eng{secret--\gmk{PL}} \\
+%						кыз-\hiliteone{\{L\}}\{A\}р → кыз\hiliteone{д}ар \eng{girl--\gmk{PL}} \\
+%				\end{itemize}
+%
+%				\vspace{-0.5ex}
+%				\codeex{
+%					\texttt{"L Desonorisation"\\
+%						\%\{L\%\}:д <=> :VoicedLowSonCns \%>:\ \blank{1em} ;}\\
+%			
+%					\texttt{"N Desonorisation"\\
+%						\%\{N\%\}:д <=> :VoicedCns \%>:\ \blank{1em} ;}
+%				}
+%			\vspace{0.1em}
+%
+%			\htwo{Epenthesis}
+%				\begin{itemize}
+%					\item Turn {\texttt \{y\}} into a harmonised high vowel when a vowel doesn't follow the following consonant:\\
+%					мур\archiphon{у}н → мурун \eng{nose}\\
+%					мур\archiphon{у}н+\archiphon{I}м → мурдум \eng{my nose}\\
+%				\end{itemize}
+%					\vspace{-0.5ex}
+%\codeex{
+%			\begin{spacing}{0.9}
+%{\small \texttt{\%\{y\%\}:Vy <=> [ :LastVowel :Cns* :Cns ]/[:0] \blank{1em}\\
+%	\vspace{-0.5ex}\hspace{1pt}\hfill{} [ :Cns [ .\#.\ | :Cns ] ]/[ :0 | \%>:]\ ;\\
+%  \vspace{-0.5ex}\hspace{1ex}where  Vy  in  (  и  ү  и  и  ү  ы  ы  у  у  ы  у  у )\\
+%  \vspace{-0.5ex}LastVowel  in  (  и  ү  е  э  ө  я  а  ё  о  ы  ю  у )\\
+%  \vspace{-0.5ex}\hspace{12ex}         matched ;}}\vspace{-1.5em}
+%  			\end{spacing}
+%}
+%\vspace{0.1em}
+%		}
 
-			\htwo{Desonorisation (kaz \& kir)}
-				\begin{itemize}
-					\item {\texttt \{N\}} desonorises to д after a consonant\\
-						алма-\hiliteone{\{N\}}\{I\} → алма\hiliteone{н}ы \eng{apple--\gmk{ACC}} \\
-						сыр-\hiliteone{\{N\}}\{I\} → сыр\hiliteone{д}ы \eng{secret--\gmk{ACC}}
-					\item {\texttt \{L\}} desonorises to д after cons.\ of sonority ≤ /l/ \\
-						сыр-\hiliteone{\{L\}}\{A\}р → сыр\hiliteone{л}ар \eng{secret--\gmk{PL}} \\
-						кыз-\hiliteone{\{L\}}\{A\}р → кыз\hiliteone{д}ар \eng{girl--\gmk{PL}} \\
-				\end{itemize}
-
-				\vspace{-0.5ex}
-				\codeex{
-					\texttt{"L Desonorisation"\\
-						\%\{L\%\}:д <=> :VoicedLowSonCns \%>:\ \blank{1em} ;}\\
-
-					\texttt{"N Desonorisation"\\
-						\%\{N\%\}:д <=> :VoicedCns \%>:\ \blank{1em} ;}
-				}
-			\vspace{0.1em}
-
-			\htwo{Epenthesis}
-				\begin{itemize}
-					\item Turn {\texttt \{y\}} into a harmonised high vowel when a vowel doesn't follow the following consonant:\\
-					мур\archiphon{у}н → мурун \eng{nose}\\
-					мур\archiphon{у}н+\archiphon{I}м → мурдум \eng{my nose}\\
-				\end{itemize}
-					\vspace{-0.5ex}
-\codeex{
-			\begin{spacing}{0.9}
-{\small \texttt{\%\{y\%\}:Vy <=> [ :LastVowel :Cns* :Cns ]/[:0] \blank{1em}\\
-	\vspace{-0.5ex}\hspace{1pt}\hfill{} [ :Cns [ .\#.\ | :Cns ] ]/[ :0 | \%>:]\ ;\\
-  \vspace{-0.5ex}\hspace{1ex}where  Vy  in  (  и  ү  и  и  ү  ы  ы  у  у  ы  у  у )\\
-  \vspace{-0.5ex}LastVowel  in  (  и  ү  е  э  ө  я  а  ё  о  ы  ю  у )\\
-  \vspace{-0.5ex}\hspace{12ex}         matched ;}}\vspace{-1.5em}
+		\headerbox{References}{name=references,column=3,below=tuvanvowelharmony}{
+			\begin{spacing}{0.85}
+				\printbibliography[heading=none]
 			\end{spacing}
-}
-\vspace{0.1em}
+			\vspace{-1em}
 		}

### 2015-07-08 2015-compmorphon-morphver/poster.tex
development cycle
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61214 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \usepackage{tikz}
+\usetikzlibrary{shapes,arrows} %for flowchart
+% Define block styles
+\tikzstyle{decision} = [diamond, draw, fill=blue!20, 
+    text width=4.5em, text badly centered, node distance=3cm, inner sep=0pt]
+\tikzstyle{block} = [rectangle, draw, fill=blue!20, 
+    text width=5em, text centered, rounded corners, minimum height=4em]
+\tikzstyle{line} = [draw, -latex']
+\tikzstyle{cloud} = [draw, ellipse,fill=red!20, node distance=3cm,
+    minimum height=2em]
 %\usepackage{pgfbaselayers}
@@
 
+		\headerbox{Development cycle}{name=development,below=morphtrans}{
+			%(1) look at frequency list --> check word --> if it's missing stem, add --> (1) ; if it's bad phonology --> fix phonology --> (1)
+			\begin{tikzpicture}[node distance = 2cm, auto]
+				% Place nodes
+				\node [block] (init) {analyse corpus};
+				\node [cloud, left of=init] (addstem) {add stem};
+				\node [cloud, right of=init] (fixphon) {fix phonology};
+				\node [block, below of=init] (freq) {examine frequency list};
+				%\node [block, below of=identify] (evaluate) {evaluate candidate models};
+				%\node [block, left of=evaluate, node distance=3cm] (update) {update model};
+				\node [decision, below of=freq] (decide) {cause for unanalysed forms?};
+				%\node [block, below of=decide, node distance=3cm] (stop) {stop};
+				% Draw edges
+				\path [line] (init) -- (freq);
+				\path [line] (freq) -- (decide);
+				\path [line] (decide) -| node [near start] {missing stem} (addstem);
+				\path [line] (decide.east) -| node [near start] {bad phonology} (fixphon);
+				\path [line,dashed] (addstem) -- (init);
+				\path [line,dashed] (fixphon) -- (init);
+			\end{tikzpicture}
+		}
+
 	\end{poster}

### 2015-07-08 2015-compmorphon-morphver/poster.tex
fixes to flowchart
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61218 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 % Define block styles
-\tikzstyle{decision} = [diamond, draw, fill=blue!20, 
-    text width=4.5em, text badly centered, node distance=3cm, inner sep=0pt]
-\tikzstyle{block} = [rectangle, draw, fill=blue!20, 
+\tikzstyle{decision} = [diamond, draw, fill=iumint, 
+    text width=5em, text badly centered, node distance=3cm, inner sep=0pt]
+\tikzstyle{block} = [rectangle, draw, fill=iulimestone, 
     text width=5em, text centered, rounded corners, minimum height=4em]
 \tikzstyle{line} = [draw, -latex']
-\tikzstyle{cloud} = [draw, ellipse,fill=red!20, node distance=3cm,
+\tikzstyle{cloud} = [draw, ellipse,fill=iumidnight, text=grey, node distance=3cm,
     minimum height=2em]
@@
 \definecolor{iumidnight}{cmyk}{0.71,0.30,0.13,0.41}
+\definecolor{iulimestone}{cmyk}{0.11,0.13,0.14,0.26}
 %\definecolor{boxcolor}{RGB}{176,208,223}
@@
 \renewcommand*{\nameyeardelim}{\addcomma\space}
+\renewcommand*{\bibfont}{\footnotesize}
 
@@
 		\headerbox{References}{name=references,column=3,below=tuvanvowelharmony}{
-			\begin{spacing}{0.85}
+			\begin{spacing}{0.7}
 				\printbibliography[heading=none]
 			\end{spacing}
-			\vspace{-1em}
+			\vspace{-0.5em}
 		}
@@
 			%(1) look at frequency list --> check word --> if it's missing stem, add --> (1) ; if it's bad phonology --> fix phonology --> (1)
+			\scalebox{0.8}{
 			\begin{tikzpicture}[node distance = 2cm, auto]
 				% Place nodes
-				\node [block] (init) {analyse corpus};
-				\node [cloud, left of=init] (addstem) {add stem};
-				\node [cloud, right of=init] (fixphon) {fix phonology};
-				\node [block, below of=init] (freq) {examine frequency list};
+				\node [block] (init) at (1,3) {analyse corpus} ;
+				\node [cloud, left of=init] (addstem) at (0,3) {add stem};
+				\node [cloud, right of=init] (fixphon) at (2,3) {fix phonology};
+				\node [block, below of=init] (freq) at (1,2) {examine frequency list};
+				\node [decision, right of=freq] (break) at (1.5,0) {did something break?};
 				%\node [block, below of=identify] (evaluate) {evaluate candidate models};
 				%\node [block, left of=evaluate, node distance=3cm] (update) {update model};
-				\node [decision, below of=freq] (decide) {cause for unanalysed forms?};
+				\node [decision, below right of=freq] (decide) at (0,-1) {cause for unanalysed forms?};
 				%\node [block, below of=decide, node distance=3cm] (stop) {stop};
@@
 				\path [line] (init) -- (freq);
-				\path [line] (freq) -- (decide);
+				\path [line] (freq) -- (break);
+				\path [line] (break.south) -| node [near start] {no} (decide.north) ;
+				\path [line] (break.north) -| node [near start] {yes}(fixphon.south) ;
 				\path [line] (decide) -| node [near start] {missing stem} (addstem);
-				\path [line] (decide.east) -| node [near start] {bad phonology} (fixphon);
+				\path [line] (decide.east) --(decide-|fixphon.east) node [below, midway] {bad phonology} |-(break.north-|fixphon.south) -- (fixphon.south);
 				\path [line,dashed] (addstem) -- (init);
 				\path [line,dashed] (fixphon) -- (init);
-			\end{tikzpicture}
+			\end{tikzpicture}}
 		}

### 2015-07-08 2015-compmorphon-morphver/poster.tex
updated dev cycle
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61219 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 				% Place nodes
+				\node [cloud, above of=fixphon] (assumed) at (5.5,1.5) {assumed phonology} ;
 				\node [block] (init) at (1,3) {analyse corpus} ;
-				\node [cloud, left of=init] (addstem) at (0,3) {add stem};
-				\node [cloud, right of=init] (fixphon) at (2,3) {fix phonology};
+				\node [block, left of=init] (addstem) at (0,3) {add stem};
+				\node [block, right of=init] (fixphon) at (3.5,3) {fix phonology};
 				\node [block, below of=init] (freq) at (1,2) {examine frequency list};
@@
 				\node [decision, below right of=freq] (decide) at (0,-1) {cause for unanalysed forms?};
+				\node [cloud] (revised) at (6.75,-4.5) {revised phonology} ;
 				%\node [block, below of=decide, node distance=3cm] (stop) {stop};
@@
 				\path [line] (decide.east) --(decide-|fixphon.east) node [below, midway] {bad phonology} |-(break.north-|fixphon.south) -- (fixphon.south);
-				\path [line,dashed] (addstem) -- (init);
-				\path [line,dashed] (fixphon) -- (init);
+				\path [line] (addstem) -- (init);
+				\path [line] (fixphon) -- (init);
+				\path [line,dashed] (assumed) -- (fixphon);
+				\path [line,dashed] (fixphon.east) --(fixphon.east-|assumed.east) |-(decide.east-|revised.north) -- (revised.north);
 			\end{tikzpicture}}

### 2015-07-08 2015-compmorphon-morphver/poster.tex
better flow chart?
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61221 72bbbca6-d526-0410-a7d9-f06f51895060

--- 2015-07-08 2015-compmorphon-morphver/poster.tex
@@
 % Define block styles
-\tikzstyle{decision} = [diamond, draw, fill=blue!20, 
-    text width=4.5em, text badly centered, node distance=3cm, inner sep=0pt]
-\tikzstyle{block} = [rectangle, draw, fill=blue!20, 
-    text width=5em, text centered, rounded corners, minimum height=4em]
-\tikzstyle{line} = [draw, -latex']
-\tikzstyle{cloud} = [draw, ellipse,fill=red!20, node distance=3cm,
-    minimum height=2em]
+\newcommand{\flowchartscale}{0.65}
+\tikzstyle{decision} = [diamond, draw, fill=iumint, 
+    text width=5em, text badly centered, node distance=3cm, inner sep=0pt,
+	 font=\diagsans, scale=\flowchartscale]
+\tikzstyle{block} = [rectangle, draw, fill=iulimestone, 
+    text width=5em, text centered, rounded corners, minimum height=4em,
+	 font=\diagsans, scale=\flowchartscale]
+\tikzstyle{line} = [draw, -latex', thick, font=\diagsans]
+\tikzstyle{cloud} = [draw, ellipse,fill=iumidnight, text=grey, text width=8em,
+	text centered, %node distance=3cm,
+	minimum height=2em, font=\diagsans, scale=\flowchartscale]
 %\usepackage{pgfbaselayers}
@@
 \newfontfamily\ortamono[IPA,Scale=0.625]{DejaVu Sans Mono}
+\newfontfamily\diagsans[IPA,Scale=1]{DejaVu Sans}
 
@@
 			\item Transducers available \textbf{live} at \href{http://turkic.apertium.org/}{\texttt{turkic.apertium.org}}
-			\item \textbf{Source code} available from Apertium's svn repo
+			\item \textbf{Source code} available from Apertium's SVN repo
 			%info at {\small \url{http://wiki.apertium.org/wiki/apertium-kir}}
@@
 			\vspace{-0.5ex}Feel free to post in any language!\\\vspace{-2.5ex}
-			\item See our papers in LREC proceedings\\
-			(2012:\ Kyrgyz, 2014:\ Kazakh, Tatar, Kumyk)
 			\item And feel free to contact the authors any time!
@@
 			%(1) look at frequency list --> check word --> if it's missing stem, add --> (1) ; if it's bad phonology --> fix phonology --> (1)
-			\begin{tikzpicture}[node distance = 2cm, auto]
-				% Place nodes
-				\node [block] (init) {analyse corpus};
-				\node [cloud, left of=init] (addstem) {add stem};
-				\node [cloud, right of=init] (fixphon) {fix phonology};
-				\node [block, below of=init] (freq) {examine frequency list};
-				%\node [block, below of=identify] (evaluate) {evaluate candidate models};
-				%\node [block, left of=evaluate, node distance=3cm] (update) {update model};
-				\node [decision, below of=freq] (decide) {cause for unanalysed forms?};
-				%\node [block, below of=decide, node distance=3cm] (stop) {stop};
-				% Draw edges
-				\path [line] (init) -- (freq);
-				\path [line] (freq) -- (decide);
-				\path [line] (decide) -| node [near start] {missing stem} (addstem);
-				\path [line] (decide.east) -| node [near start] {bad phonology} (fixphon);
-				\path [line,dashed] (addstem) -- (init);
-				\path [line,dashed] (fixphon) -- (init);
-			\end{tikzpicture}
+			%\scalebox{0.8}{
+			%\begin{tikzpicture}[node distance = 2cm, auto]
+			%	% Place nodes
+			%	\node [cloud, above of=fixphon] (assumed) at (5.5,1.5) {assumed phonology} ;
+			%	\node [block] (init) at (1,3) {analyse corpus} ;
+			%	\node [block, left of=init] (addstem) at (0,3) {add stem};
+			%	\node [block, right of=init] (fixphon) at (3.5,3) {fix phonology};
+			%	\node [block, below of=init] (freq) at (1,2) {examine frequency list};
+			%	\node [decision, right of=freq] (break) at (1.5,0) {did something break?};
+			%	%\node [block, below of=identify] (evaluate) {evaluate candidate models};
+			%	%\node [block, left of=evaluate, node distance=3cm] (update) {update model};
+			%	\node [decision, below right of=freq] (decide) at (0,-1) {cause for unanalysed forms?};
+			%	\node [cloud] (revised) at (6.75,-4.5) {revised phonology} ;
+			%	%\node [block, below of=decide, node distance=3cm] (stop) {stop};
+			%	% Draw edges
+			%	\path [line] (init) -- (freq);
+			%	\path [line] (freq) -- (break);
+			%	\path [line] (break.south) -| node [near start] {no} (decide.north) ;
+			%	\path [line] (break.north) -| node [near start] {yes}(fixphon.south) ;
+			%	\path [line] (decide) -| node [near start] {missing stem} (addstem);
+			%	\path [line] (decide.east) --(decide-|fixphon.east) node [below, midway] {bad phonology} |-(break.north-|fixphon.south) -- (fixphon.south);
+			%	\path [line] (addstem) -- (init);
+			%	\path [line] (fixphon) -- (init);
+			%	\path [line,dashed] (assumed) -- (fixphon);
+			%	\path [line,dashed] (fixphon.east) --(fixphon.east-|assumed.east) |-(decide.east-|revised.north) -- (revised.north);
+			%\scalebox{0.65}{
+			\begin{tikzpicture}[xscale=1.92,yscale=2.25]%, auto]
+
+				\node[cloud] (assumptions) at (1,2.75) {assumed generalisations};
+				\node[block] (analyse) at (1,2) {analyse corpus};
+				\node[block] (check) at (1,1) {examine freq.\ list};
+				\node[decision] (missingstem) at (1,0) {missing stem?};
+				\node[block] (addstem) at (0,2) {add stem};
+				\node[decision] (phonormorph) at (2,0) {type of problem?};
+				\node[block] (updatephon) at (3,0) {update phonol.};
+				\node[block] (updatemorph) at (3,1) {update morphol.};
+				\node[decision] (phonbetter) at( 4,0) {better?};
+				\node[decision] (morphbetter) at (4,1) {better?};
+				\node[cloud] (newgeneralisation) at (2.5,2) {new\\grammatical\\generalisation};
+
+				\path[line] (assumptions) -- (analyse);
+				\path[line] (analyse) -- (check);
+				\path[line] (check) -- node [right=0.75ex, scale=\flowchartscale, align=left] {unanalysed\\forms} (missingstem);
+				\path[line] (missingstem) -| node [below=1ex, near start, scale=\flowchartscale] {yes} (addstem);
+				\path[line] (addstem) -- (analyse);
+				\path[line] (missingstem) -- node [below=1ex, midway, scale=\flowchartscale] {no} (phonormorph);
+				\path[line] (phonormorph) -- node [below=1ex, near start, scale=\flowchartscale] {phon} (updatephon);
+				\path[line] (phonormorph) |- node [below=1ex, near end, scale=\flowchartscale] {morph} (updatemorph);
+				\path[line] (updatephon) -- node [above=0.5ex, scale=\flowchartscale, align=center] {diff \\ freq \\ list} (phonbetter) ;
+				\path[line] (updatemorph) -- node [above=0.5ex, scale=\flowchartscale, align=center] {diff \\ freq \\ list} (morphbetter) ;
+				\path[line] (phonbetter.east) -- node [below=1ex, near start, scale=\flowchartscale] {yes} (4.5,0) |- (newgeneralisation.east) ;
+				\path[line] (morphbetter.east) -- node [below=1ex, near start, scale=\flowchartscale] {yes} (4.5,1) |- (newgeneralisation.east) ;
+				\path[line] (phonbetter.north) |- node [right=1ex, near start, scale=\flowchartscale] {no} (4,0.5) -| (updatephon.north) ;
+				\path[line] (morphbetter.north) |- node [right=1ex, near start, scale=\flowchartscale] {no} (4,1.5) -| (updatemorph.north) ;
+				\path[line] (newgeneralisation) -- (analyse);
+
+			\end{tikzpicture}%}
+
 		}

### 2015-07-08 2015-compmorphon-morphver/poster.tex
one more step(!): compile
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61222 72bbbca6-d526-0410-a7d9-f06f51895060
stretch flow chart vertically
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61223 72bbbca6-d526-0410-a7d9-f06f51895060
smoe more space adjustments
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61224 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 			%\scalebox{0.65}{
-			\begin{tikzpicture}[xscale=1.92,yscale=2.25]%, auto]
+			\begin{tikzpicture}[xscale=1.92,yscale=2.65]%, auto]
 
 				\node[cloud] (assumptions) at (1,2.75) {assumed generalisations};
-				\node[block] (analyse) at (1,2) {analyse corpus};
+				\node[block] (compile) at (1,2) {compile analyser};
+				\node[block] (analyse) at (1,1.5) {analyse corpus};
 				\node[block] (check) at (1,1) {examine freq.\ list};
 				\node[decision] (missingstem) at (1,0) {missing stem?};
-				\node[block] (addstem) at (0,2) {add stem};
+				\node[block] (addstem) at (0.22,2) {add stem};
 				\node[decision] (phonormorph) at (2,0) {type of problem?};
@@
 
-				\path[line] (assumptions) -- (analyse);
+				\path[line] (assumptions) -- (compile);
+				\path[line] (compile) -- (analyse);
 				\path[line] (analyse) -- (check);
@@
 				\path[line] (missingstem) -| node [below=1ex, near start, scale=\flowchartscale] {yes} (addstem);
-				\path[line] (addstem) -- (analyse);
+				\path[line] (addstem) -- (compile);
 				\path[line] (missingstem) -- node [below=1ex, midway, scale=\flowchartscale] {no} (phonormorph);
@@
 				\path[line] (updatemorph) -- node [above=0.5ex, scale=\flowchartscale, align=center] {diff \\ freq \\ list} (morphbetter) ;
-				\path[line] (phonbetter.east) -- node [below=1ex, near start, scale=\flowchartscale] {yes} (4.5,0) |- (newgeneralisation.east) ;
-				\path[line] (morphbetter.east) -- node [below=1ex, near start, scale=\flowchartscale] {yes} (4.5,1) |- (newgeneralisation.east) ;
+				\path[line] (phonbetter.east) -- node [below=1ex, near start, scale=\flowchartscale] {yes} (4.6,0) |- (newgeneralisation.east) ;
+				\path[line] (morphbetter.east) -- node [below=1ex, near start, scale=\flowchartscale] {yes} (4.6,1) |- (newgeneralisation.east) ;
 				\path[line] (phonbetter.north) |- node [right=1ex, near start, scale=\flowchartscale] {no} (4,0.5) -| (updatephon.north) ;
 				\path[line] (morphbetter.north) |- node [right=1ex, near start, scale=\flowchartscale] {no} (4,1.5) -| (updatemorph.north) ;
-				\path[line] (newgeneralisation) -- (analyse);
+				\path[line] (newgeneralisation) -- (compile);
 

### 2015-07-09 2015-compmorphon-morphver/poster.tex
"preliminary development"
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61225 72bbbca6-d526-0410-a7d9-f06f51895060
some positioning and stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61227 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 %\pgfsetlayers{background,main,foreground}
+
+\tikzset{
+  every overlay node/.style={
+    draw=none,fill=none,rounded corners,anchor=north west,
+  },
+}
+% Usage:
+% \tikzoverlay at (-1cm,-5cm) {content};
+% or
+% \tikzoverlay[text width=5cm] at (-1cm,-5cm) {content};
+\def\tikzoverlay{%
+   \tikz[baseline,overlay]\node[every overlay node]
+}%
 
@@
 %\usepackage{natbib}
-\usepackage[backend=biber,citestyle=apa,style=authoryear,maxbibnames=99,maxcitenames=2]{biblatex}
+\usepackage[backend=biber,citestyle=apa,maxcitenames=2,style=authoryear,maxbibnames=99,uniquelist=false,uniquename=false]{biblatex}
 \bibliography{poster}
@@
 \renewcommand*{\bibfont}{\footnotesize}
+\renewcommand*{\intitlepunct}{\space}
 
@@
 		%\end{itemize}
-		\scalebox{0.98}{
-			алдым \hilitetwo{↔} \texttt{ал\tag{v}\tag{tv}\tag{ifi}\tag{p1}\tag{sg}}, \texttt{алд\tag{n}\tag{px1sg}\tag{nom}}
-		}
+		%\scalebox{0.98}{
+			алдым~~~\hilitetwo{↔}~~~\texttt{ал\tag{v}\tag{tv}\tag{ifi}\tag{p1}\tag{sg}},~~\texttt{алд\tag{n}\tag{px1sg}\tag{nom}}
+		%}
 		\end{itemize}
@@
 
-		\headerbox{Example:\ Tuvan velar elision}{name=tuvanvelars,column=1,span=2}{
+		\headerbox{Example: Tuvan velar elision}{name=tuvanvelars,column=1,span=2}{
 			\begin{multicols}{2}
@@
 			\columnbreak{}
-				\htwo{Uanalysed forms before implementation}
+				\htwo{Unanalysed forms before implementation}
 					\begin{itemize}
@@
 
-	\headerbox{Example: Categorisation}{name=adjectives,column=1,below=tuvanvelars}{
-		\htwo{Unrestricted 0-derivation}
-		\begin{itemize}
-			\item Common overgeneralisation of adjective morphology:
-			\begin{itemize}
-				\item all adjectives may act like nouns (i.e.\ take noun morphology)
-				\item all adjectives may be used as is as adverbs
-			\end{itemize}
-			\item Commonly reported in Turkic language grammars
-		\end{itemize}
-		%Somfai Kara (2002, pp. 28-29):\ \includegraphics[width=0.53\textwidth]{img/kara28}\\
-		\textcite[28-29]{somfaikara02}:\ \includegraphics[width=0.53\textwidth]{img/kara28}\\
-		\-\hfill{}\includegraphics[width=0.8\textwidth]{img/kara29}
-
-		\htwo{Attested system}
-			\begin{itemize}
-				\item Some adjectives may not be substantivised
-				\item Some adjectives may not be used adverbially
-				\item Some adjectives do not have comparative forms
-				\item …a range of “adjective classes” in most Turkic languages
-			\end{itemize}
-		\htwo{Implementation: proper categorisation}
-			\begin{itemize}
-				\item only correct forms are analysed and generated
-			\end{itemize}
-
-			\scalebox{0.83}{
-				\begin{tabular}{lllll}
-					\toprule
-					\textbf{Type} & \textbf{Gloss} & {\small \texttt{\tag{adj}(\tag{comp})}} & {\small \texttt{\tag{adj}(\tag{comp})\tag{subst}}} & {\small \texttt{\tag{adj}(\tag{comp})\tag{advl}}} \\
-					\midrule
-					\textbf{A1} & \eng{good} & {\qipa яхшы (яхшырак)} & {\qipa яхшы (яхшырак)} & {\qipa яхшы (яхшырак)} \\
-					\textbf{A2} & \eng{old} & {\qipa иске (искерәк)} & {\qipa иске (искерәк)} & {\qipa — (—)} \\
-					\textbf{A3} & \eng{dead} & {\qipa үле (—)} & {\qipa үле (—)} & {\qipa — (—)} \\
-					\textbf{A4} & \eng{basic} & {\qipa төп (—)} & {\qipa — (—)} & {\qipa — (—)} \\
-					\bottomrule
-				\end{tabular}
-			}\\
-			This phenomenon is of particular interest because native-speaker intuitions on the morphological limitations of a given adjective can often be more restrictive than the range of possible uses found in a large text corpus.
-	}
-
 	\headerbox{Ongoing and future work}{name=nedostatki,column=2,below=tuvanvelars}{
-		\begin{itemize}
-			\item Disambiguation, more stems, clean up transducers
-			\item Machine translation between these languages
-			\item Bring other Kypchak transducers to comparable performance:\\
-			Qaraqalpaq, Bashqort, Nogay, Crimean Tatar
-			\item Other Turkic lgs:\ Uzbek, Uyghur, Chuvash, Yakut, Tuvan, etc.
-		\end{itemize}
+		%\begin{itemize}
+			%\item Disambiguation, more stems, clean up transducers
+			%\item Machine translation between these languages
+			%\item Bring other Kypchak transducers to comparable performance:\\
+			%Qaraqalpaq, Bashqort, Nogay, Crimean Tatar
+			%\item Other Turkic lgs:\ Uzbek, Uyghur, Chuvash, Yakut, Tuvan, etc.
+		%\end{itemize}
 	}
@@
 		\end{itemize}
+		\tikzoverlay[text width=0.8in] at (2.85in, 0.77in) {
+			\includegraphics[height=0.8in]{qrcode}
+		};
 	}
-
-%	\headerbox{References}{name=references,column=2,below=nedostatki}{
-%		%\begin{thebibliography}{1}\itemsep=-0.01em
-%		\setlength{\baselineskip}{0.4em}
-%	}
 
@@
 
+	\headerbox{Example: Categorisation}{name=adjectives,column=3,below=tuvanvowelharmony}{
+		\htwo{Unrestricted 0-derivation}
+		\begin{itemize}
+			\item Common overgeneralisation of adjective morphology:
+			\begin{itemize}
+				\item all adjectives may act like nouns (i.e.\ take noun morphology)
+				\item all adjectives may be used as is as adverbs
+			\end{itemize}
+			\item Commonly reported in Turkic language grammars
+		\end{itemize}
+		%Somfai Kara (2002, pp. 28-29):\ \includegraphics[width=0.53\textwidth]{img/kara28}\\
+		\textcite[28-29]{somfaikara02}:\ \includegraphics[width=0.53\textwidth]{img/kara28}\\
+		\-\hfill{}\includegraphics[width=0.8\textwidth]{img/kara29}
+
+		\htwo{Attested system}
+			\begin{itemize}
+				\item Some adjectives may not be substantivised
+				\item Some adjectives may not be used adverbially
+				\item Some adjectives do not have comparative forms
+				\item …a range of “adjective classes” in most Turkic languages
+			\end{itemize}
+		\htwo{Implementation:\ proper categorisation}
+			\begin{itemize}
+				\item only correct forms are analysed and generated
+			\end{itemize}
+
+			\scalebox{0.83}{
+				\begin{tabular}{lllll}
+					\toprule
+					\textbf{Type} & \textbf{Gloss} & {\small \texttt{\tag{adj}(\tag{comp})}} & {\small \texttt{\tag{adj}(\tag{comp})\tag{subst}}} & {\small \texttt{\tag{adj}(\tag{comp})\tag{advl}}} \\
+					\midrule
+					\textbf{A1} & \eng{good} & {\qipa яхшы (яхшырак)} & {\qipa яхшы (яхшырак)} & {\qipa яхшы (яхшырак)} \\
+					\textbf{A2} & \eng{old} & {\qipa иске (искерәк)} & {\qipa иске (искерәк)} & {\qipa — (—)} \\
+					\textbf{A3} & \eng{dead} & {\qipa үле (—)} & {\qipa үле (—)} & {\qipa — (—)} \\
+					\textbf{A4} & \eng{basic} & {\qipa төп (—)} & {\qipa — (—)} & {\qipa — (—)} \\
+					\bottomrule
+				\end{tabular}
+			}\\
+			\begin{itemize}
+				\item Interestingly, native-speaker intuitions on the morphological limitations of a given adjective can often be more restrictive than the range of possible uses found in a large text corpus.
+			\end{itemize}
+	}
+
 		\headerbox{Examples:\ ideas for other examples}{name=otherexamples,column=0,span=1,below=tuvanvelars}{ %{name=otherexamples,column=1,span=2,below=tuvanvelars}{
@@
 
-		\headerbox{References}{name=references,column=3,below=tuvanvowelharmony}{
-			\begin{spacing}{0.7}
+		\headerbox{References}{name=references,column=1,below=tuvanvelars}{
+			\begin{spacing}{0.75}
 				\printbibliography[heading=none]
 			\end{spacing}
-			\vspace{-0.5em}
+			\vspace{-0.3em}
 		}
@@
 
-				\path[line] (assumptions) -- (compile);
+				\path[line] (assumptions) -- node [right=0.75ex, align=left, scale=\flowchartscale] {preliminary\\development} (compile);
 				\path[line] (compile) -- (analyse);

### 2015-07-09 2015-compmorphon-morphver/poster.tex
more layout and stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61228 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-	\headerbox{Ongoing and future work}{name=nedostatki,column=2,below=tuvanvelars}{
-		%\begin{itemize}
-			%\item Disambiguation, more stems, clean up transducers
-			%\item Machine translation between these languages
-			%\item Bring other Kypchak transducers to comparable performance:\\
-			%Qaraqalpaq, Bashqort, Nogay, Crimean Tatar
-			%\item Other Turkic lgs:\ Uzbek, Uyghur, Chuvash, Yakut, Tuvan, etc.
-		%\end{itemize}
-	}
-
-	\headerbox{Further information}{name=getting,column=2,below=nedostatki}{
-		\begin{itemize}
-			\item Part of \textbf{Apertium Turkic} project:\\
-			\url{http://wiki.apertium.org/wiki/Apertium\_Turkic}
-			\item Transducers available \textbf{live} at \href{http://turkic.apertium.org/}{\texttt{turkic.apertium.org}}
-			\item \textbf{Source code} available from Apertium's SVN repo
-			%info at {\small \url{http://wiki.apertium.org/wiki/apertium-kir}}
-				\item Turkic RBMT \textbf{mailing list} (>25 subscribers):\\
-			\texttt{apertium-turkic@lists.sourceforge.net}\\
-			\vspace{-0.5ex}Feel free to post in any language!\\\vspace{-2.5ex}
-			\item And feel free to contact the authors any time!
-		\end{itemize}
-		\tikzoverlay[text width=0.8in] at (2.85in, 0.77in) {
-			\includegraphics[height=0.8in]{qrcode}
-		};
-	}
-
 		\headerbox{Example: Tuvan vowel harmony}{name=tuvanvowelharmony,column=3}{ %{name=otherexamples,column=1,span=2,below=tuvanvelars}{ % after consonant clusters in loanwords
@@
 			\begin{itemize}
-				\item Interestingly, native-speaker intuitions on the morphological limitations of a given adjective can often be more restrictive than the range of possible uses found in a large text corpus.
+				%\item Interestingly, native-speaker intuitions on the morphological limitations of a given adjective can often be more restrictive than the range of possible uses found in a large text corpus.
+				\item Native speaker intuitions on morph.\ limitations often more restrictive than range of possible uses found in large text copora.
 			\end{itemize}
@@
 			\end{spacing}
-			\vspace{-0.3em}
+			\vspace{-0.2em}
+		}
+
+		\headerbox{Ongoing and future work}{name=nedostatki,column=2,below=tuvanvelars}{
+		\begin{itemize}
+				\item Fixing morphology and disambiguation by developing a dependency treebank for Kazakh
+				%\item Disambiguation, more stems, clean up transducers
+				%\item Machine translation between these languages
+				%\item Bring other Kypchak transducers to comparable performance:\\
+				%Qaraqalpaq, Bashqort, Nogay, Crimean Tatar
+				%\item Other Turkic lgs:\ Uzbek, Uyghur, Chuvash, Yakut, Tuvan, etc.
+			\end{itemize}
+		}
+
+		\headerbox{Further information}{name=getting,column=2,below=adjectives,span=2}{
+			%\vspace{-0.2em}
+			\tikzoverlay[text width=0.8in] at (6.73in, 0.04in) {
+				\includegraphics[height=0.8in]{qrcode}
+			};
+			\begin{itemize}
+				\item Part of \textbf{Apertium Turkic} project:\ \url{http://wiki.apertium.org/wiki/Apertium\_Turkic}
+				\item Transducers available \textbf{live} (\href{http://turkic.apertium.org/}{\texttt{turkic.apertium.org}}), \textbf{source code} under \hiliteone{GPL} from Apertium's SVN repo
+				%\item Transducers available \textbf{live} at \href{http://turkic.apertium.org/}{\texttt{turkic.apertium.org}}
+				%\item \textbf{Source code} of transducers available under \hiliteone{GPL} (free!) from Apertium's SVN repo
+				%info at {\small \url{http://wiki.apertium.org/wiki/apertium-kir}}
+				\item Multilingual Turkic RBMT \textbf{mailing list} (>25 subscribers):\ \texttt{apertium-turkic@lists.sourceforge.net}%\\
+				%\vspace{-0.5ex}Feel free to post in any language!\\\vspace{-2.5ex}
+				\item And feel free to contact the authors any time!
+			\end{itemize}
+			\vspace{0.3em}
 		}

### 2015-07-09 2015-compmorphon-morphver/poster.tex
some minor changes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61229 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 			\item restrictions of the medium (e.g., number of pages)
-			\item limitations of introspection, working with native speakers
+			\item limitations of introspection and working with native speakers
 			\item \textbf{lack of automatic testing}
@@
 			\item two-level method \parencite{koskenniemi83}
-			\item morphotactics implemented in \texttt{lexc}
-			\item morphophonology implemented in \texttt{twol} (SPE-style rules)
+			\item \textbf{morphotactics} implemented in \texttt{lexc}
+			\item \textbf{morphophonology} implemented in \texttt{twol} (SPE-style rules)
 			\item compiled separately; compose-intersected to single transducer\vspace{-0.1ex}\\
@@
 					\begin{itemize}
-						\item \textcite[22-23]{andersonharrison99} %Anderson and Harrison (1999, pp.\ 9, 22-23)
+						\item \textcite[9, 22-23]{andersonharrison99} %Anderson and Harrison (1999, pp.\ 9, 22-23)
 					\end{itemize}
@@
 				\begin{itemize}
-					\item /k/ ‹к› elides at the end of >1-σ stems intervocalically\\
+					\item /k/ ‹к› elides at the end of >1-σ stems intervocalically..\\
 						инэк+I → инээ, өк+I → өгү
@@
 						өг+I → өө 
-					\item /ŋ/ ‹ң› elides at the end of \textbf{some} stems intervocalically.\\
+					\item /ŋ/ ‹ң› elides at the end of \textbf{some} stems intervocalically\\
 						\textbf{соң+I → соо}, чаң+I → чаңы, түң+I → түңү
@@
 
-				But not part of the morphophonological description.\\
+				But not part of the morphophonological descriptions.\\
 				(A. \& H., p.\ 35)\hfill{}\includegraphics[width=0.2\textwidth]{img/andersonharrison35}\\
@@
 					\begin{itemize}
-						\item High and low vowels harmonise in backness\\
+						\item High and low vowels harmonise in \hilitetwo{back}ness\\
 						киж\hiliteone{и}\tag{n}\tag{pl}\tag{acc}: киж\hiliteone{и}л\hiliteone{е}рн\hiliteone{и}, н\hilitetwo{о}м\tag{n}\tag{pl}\tag{acc}: н\hilitetwo{о}мн\hilitetwo{а}рн\hilitetwo{ы}
@@
 			\begin{itemize}
-				\item only correct forms are analysed and generated
+				\item only correct forms are analysed and generated (Tatar exs.)
 			\end{itemize}

### 2015-07-09 2015-compmorphon-morphver/poster.tex
a few minor fixes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61230 72bbbca6-d526-0410-a7d9-f06f51895060
ongoing and future work
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61234 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 					\end{minipage}
+					\begin{itemize}
+						\item All forms of velar deletion are now analysed correctly!
+						\item Can now move on to other unanalysed forms\ldots{}
+					\end{itemize}
 			\end{multicols}
@@
 		\headerbox{References}{name=references,column=1,below=tuvanvelars}{
-			\begin{spacing}{0.75}
+			\vspace{-0.42em}
+			\begin{spacing}{0.3}
 				\printbibliography[heading=none]
 			\end{spacing}
-			\vspace{-0.2em}
+			\vspace{-0.05em}
 		}
@@
 		\begin{itemize}
-				\item Fixing morphology and disambiguation by developing a dependency treebank for Kazakh
+				\item Aannotating a dependency treebank for Kazakh:\\
+				a similar process that helps identify errors in morphological analyses and automatic disambiguation (constraint grammar)
+				%\item Leading to
+				\begin{itemize}
+					\item Improved rule-based morphological disambiguation% (constraint grammar)
+					\item Improved morphological analyser for Kazakh
+					\item A syntactic depedency treebank for Kazakh
+				\end{itemize}
+				%\item Work on rule-based morphological disambiguation with constraint grammar
+				\item Future work
+				\begin{itemize}
+					\item Annotate more gold standard text in Kazakh
+					\item Extend to other Turkic languages
+				\end{itemize}
 				%\item Disambiguation, more stems, clean up transducers

### 2015-07-09 2015-compmorphon-morphver/poster.tex
layout switch
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61235 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-		\headerbox{Example: Tuvan velar elision}{name=tuvanvelars,column=1,span=2}{
+		\headerbox{Example: Tuvan velar elision}{name=tuvanvelars,column=2,span=2}{
 			\begin{multicols}{2}
@@
 
-		\headerbox{Example: Tuvan vowel harmony}{name=tuvanvowelharmony,column=3}{ %{name=otherexamples,column=1,span=2,below=tuvanvelars}{ % after consonant clusters in loanwords
+		\headerbox{Example: Tuvan vowel harmony}{name=tuvanvowelharmony,column=1}{ %{name=otherexamples,column=1,span=2,below=tuvanvelars}{ % after consonant clusters in loanwords
 			%\begin{multicols}{2}
@@
 
-	\headerbox{Example: Categorisation}{name=adjectives,column=3,below=tuvanvowelharmony}{
+	\headerbox{Example: Categorisation}{name=adjectives,column=1,below=tuvanvowelharmony}{
 		\htwo{Unrestricted 0-derivation}
@@
 
-		\headerbox{References}{name=references,column=1,below=tuvanvelars}{
+		\headerbox{References}{name=references,column=3,below=tuvanvelars}{
 			\vspace{-0.42em}
@@
 
-		\headerbox{Further information}{name=getting,column=2,below=adjectives,span=2}{
+		\headerbox{Further information}{name=getting,column=1,below=adjectives,span=2}{
 			%\vspace{-0.2em}

### 2015-07-09 2015-compmorphon-morphver/poster.tex
maybe this?
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61236 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 		\begin{itemize}
-			\item core part of grammatical descriptions
-			\item found in reference grammars and textbooks
+			\item Core part of grammatical descriptions
+			\item Found in reference grammars and textbooks
 		\end{itemize}
@@
 		\begin{itemize}
-			\item \textbf{morphotactics} — the morphemes of a language and how they can be combined
-			\item \textbf{morphophonology} — the alternations between the various phonological and orthographical forms of each morpheme% in a language
+			\item \textbf{Morphotactics} — the morphemes of a language and how they can be combined
+			\item \textbf{Morphophonology} — the alternations between the various phonological and orthographical forms of each morpheme% in a language
 		\end{itemize}
@@
 		\begin{itemize}
-			\item restrictions of the medium (e.g., number of pages)
-			\item limitations of introspection and working with native speakers
-			\item \textbf{lack of automatic testing}
+			\item Restrictions of the medium (e.g., number of pages)
+			\item Limitations of introspection and working with native speakers
+			\item \textbf{Lack of automatic testing}
 		\end{itemize}
@@
 						\item forms of loanwords ансамбль `ensemble' and рубль `rouble':\\\
-						\-\hfill{}the following vowel is always front unround
+						the following vowel is always front unround
 					\end{itemize}
@@
 		\begin{itemize}
-				\item Aannotating a dependency treebank for Kazakh:\\
+				\item Current annotation of a dependency treebank for Kazakh:\\
 				a similar process that helps identify errors in morphological analyses and automatic disambiguation (constraint grammar)
@@
 				\begin{itemize}
-					\item Annotate more gold standard text in Kazakh
-					\item Extend to other Turkic languages
+					\item Annotate more gold standard text in more Turkic languages
+					\item Resaerch cross-lingual techniques for analysing Turkic
 				\end{itemize}

### 2015-07-09 2015-compmorphon-morphver/poster.tex
Fran's typo
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61237 72bbbca6-d526-0410-a7d9-f06f51895060
more Fran-os
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61238 72bbbca6-d526-0410-a7d9-f06f51895060
more minor fixes to future work
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61239 72bbbca6-d526-0410-a7d9-f06f51895060
more typoes agh
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61240 72bbbca6-d526-0410-a7d9-f06f51895060
typoooooes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@6124

@@
 				%\item Interestingly, native-speaker intuitions on the morphological limitations of a given adjective can often be more restrictive than the range of possible uses found in a large text corpus.
-				\item Native speaker intuitions on morph.\ limitations often more restrictive than range of possible uses found in large text copora.
+				\item Native speaker intuitions on morph.\ limitations often more restrictive than range of possible uses found in large text corpora.
 			\end{itemize}
@@
 					\item Improved rule-based morphological disambiguation% (constraint grammar)
-					\item Improved morphological analyser for Kazakh
-					\item A syntactic depedency treebank for Kazakh
+					\item Improved morphological analyser %for Kazakh
+					\item A syntactic dependency treebank %for Kazakh
 				\end{itemize}
 				%\item Work on rule-based morphological disambiguation with constraint grammar
-				\item Future work
+				\item Future work:
 				\begin{itemize}
 					\item Annotate more gold standard text in more Turkic languages
-					\item Resaerch cross-lingual techniques for analysing Turkic
+					\item Research cross-linguistic techniques for analysing Turkic
 				\end{itemize}

### 2015-08-23 2015-turklang-kazdep/paper.tex
better(?) bib fix?  Just turned off name disambiguation
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61534 72bbbca6-d526-0410-a7d9-f06f51895060
Acknowledgements
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61549 72bbbca6-d526-0410-a7d9-f06f51895060

--- 2015-06-21 2015-turklang-kazdep/paper.tex
@@
 %\usepackage[backend=bibtex]{biblatex}
-\usepackage[backend=biber,citestyle=apa,style=authoryear,maxbibnames=99,maxcitenames=2]{biblatex}
+\usepackage[backend=biber,citestyle=apa,style=authoryear-comp,maxbibnames=99,maxcitenames=2,uniquename=false]{biblatex}
 \bibliography{paper}
@@
 
-\emph{removed for review}
+The authors would like to acknowledge Tolgonay Kubatova; Zhenisbek Assylbekov, Aida Sundetova, and colleagues; and Aibek Makazhanov for the various ways in which they each contributed to this research.
 

### 2015-08-26 2015-turklang-kazdep/paper.tex
url
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61550 72bbbca6-d526-0410-a7d9-f06f51895060
current size numbers
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61587 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 This article describes the first steps towards a free/open-source dependency treebank for Kazakh based on
-universal dependency (UD) annotation standards. The treebank contains 302 sentences and is based on
+universal dependency (UD) annotation standards. The treebank contains 402 sentences and is based on
 texts from a range of open-source and public domain sources. This ensures
@@
   \centering
-  \caption{Composition of the corpus. The corpus covers a range of genres and text types from free and public-domain sources.}
+  \caption{Composition of the corpus. The corpus covers a range of genres and text types from free \& public-domain sources.}
   \label{table:corpcomp}
@@
 			\midrule
-				UN Declaration on Human Rights & Legal text on human rights & 25 & 409 & 16.3 \\
-				Phrasebook                     & Phrases from Wikitravel   & 37 & 205 & 5.5 \\
-				Жиырма Бесінші Сөз             & Philosophical text        & 34 & 525 & 15.4 \\ 
-				Қожанасырдың тойға баруы       & Folk tale from Wikisource  & 8 & 140 & 17.5 \\
-				Ер Төстік                      & Folk tale from Wikisource  & 23 & 203 & 8.8 \\
-				Азамат қайда?                  & Story for language learners & 48 & 434 & 9.0 \\
-				Футболдан әлем чемпионаты 2014 & Wikipedia article (2014 World Cup) & 14 & 246 & 17.5 \\
-				Иран & Wikipedia article (Iran)                                & 111 & 1562 & 14.0 \\
-				Радиан & Wikipedia article (Radian)                            & 2 & 17 & 8.5 \\
+				UN Declaration on Human Rights & Legal text on human rights & 26 & 417 & 16.0 \\ %
+				Phrasebook                     & Phrases from Wikitravel    & 38 & 204 & 5.4 \\ %
+				Жиырма Бесінші Сөз             & Philosophical text         & 33 & 467 & 14.2 \\ %
+				Қожанасырдың тойға баруы       & Folk tale from Wikisource  & 9 & 139 & 15.4 \\ %
+				Ер Төстік                      & Folk tale from Wikisource  & 25 & 209 & 8.4 \\ %
+				Азамат қайда?                  & Story for language learners & 49 & 433 & 8.8 \\ %
+				Футболдан әлем чемпионаты 2014 & Wikipedia article (2014 World Cup) & 12 & 191 & 15.9 \\ %
+				Иран & Wikipedia article (Iran)                             & 121 & 1714 & 14.2 \\ %
+				Радиан & Wikipedia article (Radian)                         & 2 & 17 & 8.5 \\ %
+				Шымкент & Wikipedia article (Shymkent)                      & 13 & 160 & 12.3 \\ %
+				Wikipedia misc. & Random sentences from Wikipedia           & 74 & 565 & 7.6 \\ %
 			\midrule
-			\multicolumn{2}{c}{~} & 302 & 3741 & 12.3 \\\cmidrule[\heavyrulewidth]{3-5}
+			\multicolumn{2}{c}{~} & 402 & 4516 & 11.2 \\\cmidrule[\heavyrulewidth]{3-5}
 		\end{tabular}
@@
 \end{table}
+\vspace{-0.5em}
 
@@
 \end{figure}
+\vspace{-0.5em}
 
@@
 
+\vspace{-0.5em}
 \begin{figure}[!htbp]
@@
 the experiments. The corpus was flattened with conjoined tokens receiving a dummy 
-surface form. The converted corpus is available alongside the original.\footnote{\url{removed for review}}
+surface form. The converted corpus is available alongside the original.\footnote{\url{http://svn.code.sf.net/p/apertium/svn/languages/apertium-kaz/texts/puupankki/}}
 

### 2015-08-26 2015-turklang-kazdep/paper.tex
nmod:poss in §4.3
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61588 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 			\end{deptext}
-			\depedge{2}{1}{\udlabel{nmod}}
+			\depedge{2}{1}{\udlabel{nmod:poss}}
 		\end{dependency}
@@
 
-As seen in the graphs, the compound relationship of an attributive nominal depending on another nominal is labelled \udtag{compound}, and genitive relations are considered \udtag{nmod}, regardless of whether there is a definite or indefinite genitive construction.
+As seen in the graphs, the compound relationship of an attributive nominal depending on another nominal is labelled \udtag{compound}, and genitive relations are considered \udtag{nmod:poss}, regardless of whether there is a definite or indefinite genitive construction.
 

### 2015-08-26 2015-turklang-kazdep/paper.tex
table of functions associated with each case
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61589 72bbbca6-d526-0410-a7d9-f06f51895060

@@
     %\fi
-  %}%
+  %}%%
 %}
@@
 \newcommand{\gloss}[1]{`#1'}
+\newcommand{\sgloss}[1]{\hspace{0.5em}`#1'}
+%\newcommand{\sgloss}[1]{\hfill{}`#1'}
 \newcommand{\tgloss}[1]{{\em #1}}
@@
 	\begin{small}
-		\begin{tabular}{l l p{9em} p{27.15em}}
+		\begin{tabular}{l l p{9em} p{28em}}
 			\toprule
@@
 										indefinite genitival 
-									& \kazakh{\textbf{Дәрігер} үйді көреді.} \gloss{\textbf{The doctor} sees the house.} \newline
-										\kazakh{\textbf{қонақ} үй} \gloss{hotel (lit., \textbf{guest} house)} \newline 
-										\kazakh{Дәрігер \textbf{үй} көреді.} \gloss{The doctor sees \textbf{a house}.}\newline
-										\kazakh{\textbf{үй} жануарлары} \gloss{\textbf{house} animals} \\\midrule
+									& \kazakh{\textbf{Дәрігер} үйді көреді.} \sgloss{\textbf{The doctor} sees the house.} \newline
+										\kazakh{\textbf{қонақ} үй} \sgloss{hotel (lit., \textbf{guest} house)} \newline 
+										\kazakh{Дәрігер \textbf{үй} көреді.} \sgloss{The doctor sees \textbf{a house}.}\newline
+										\kazakh{\textbf{үй} жануарлары} \sgloss{\textbf{house} animals} \\\midrule
 				acc. & -/NI/ & definite object %\newline
@@
 										%	intr.\ causative subj.
-										& \kazakh{Дәрігер \textbf{үйді} көреді.} \gloss{The doctor sees \textbf{the house}.}
+										& \kazakh{Дәрігер \textbf{үйді} көреді.} \sgloss{The doctor sees \textbf{the house}.}
 										%	bar
@@
 											embedded subject
-										& \kazakh{\textbf{үйдің} жануарлары} \gloss{the animals \textbf{of the house}} \newline
-											\kazakh{Ол \textbf{Айгүлдің} ойнағанына қарап тұр.} \gloss{She's watching \textbf{Aygül} play.} \\\midrule
+										& \kazakh{\textbf{үйдің} жануарлары} \sgloss{the animals \textbf{of the house}} \newline
+											\kazakh{Ол \textbf{Айгүлдің} ойнағанына қарап тұр.} \sgloss{She's watching \textbf{Aygül} play.} \\\midrule
 				loc. & -/DA/ & adverbial
-										& \kazakh{\textbf{Үйде} ұйықтап жатыр.} \gloss{S/he's sleeping \textbf{in the house}.} \\\midrule
+										& \kazakh{\textbf{Үйде} ұйықтап жатыр.} \sgloss{S/he's sleeping \textbf{in the house}.} \\\midrule
 				abl. & -/DAн/ & adverbial\newline
 											comparator 
-										& \kazakh{Дәрігер \textbf{үйден} шықты.} \gloss{The doctor came out \textbf{of the house}.}\newline
-											\kazakh{\textbf{түйеден} үлкен} \gloss{bigger \textbf{than a camel}}\\\midrule
+										& \kazakh{Дәрігер \textbf{үйден} шықты.} \sgloss{The doctor came out \textbf{of the house}.}\newline
+											\kazakh{\textbf{түйеден} үлкен} \sgloss{bigger \textbf{than a camel}}\\\midrule
 				dat. & -/GA/ & indirect object\newline
@@
 									%	nom.\ complement
-									& \kazakh{\textbf{Ініме} кітап бердім.} \gloss{I gave \textbf{my younger brother} a book.} \newline
-										\kazakh{\textbf{Тойға} көп кісі келіпті.} \gloss{A lot of people came \textbf{to the feast.}}\newline
-										\kazakh{\textbf{Балаға} әнді тыңдаттым.} \gloss{I made \textbf{the child} listen to the song.}
+									& \kazakh{\textbf{Ініме} кітап бердім.} \sgloss{I gave \textbf{my younger brother} a book.} \newline
+										\kazakh{\textbf{Тойға} көп кісі келіпті.} \sgloss{A lot of people came \textbf{to the feast.}}\newline
+										\kazakh{\textbf{Балаға} әнді тыңдаттым.} \sgloss{I made \textbf{the child} listen to the song.}
 										\\\midrule
-				inst. & -/Mен/ &  adverbial & \kazakh{Олар \textbf{күшікпен} ойнап тұр.} \gloss{They're playing \textbf{with the dog}.} \\
+				inst. & -/Mен/ &  adverbial & \kazakh{Олар \textbf{күшікпен} ойнап тұр.} \sgloss{They're playing \textbf{with the dog}.} \\
 			\bottomrule
@@
 	\caption{The default (first line of each) and ``zero-derived'' uses of nominals, adjectives, and adverbs}\label{tab:zeroderiv}
-	\begin{tabular}{lp{5em}p{20em}}
+	{\small
+	\begin{tabular}{lp{5em}p{21em}}
 		\toprule
@@
 							Adverbial
-						& \kazakh{\textbf{Қонақтарымыз} бай.} \gloss{\textbf{Our guests} are rich.}\newline 
-							\kazakh{\textbf{қонақ} үй} \gloss{hotel (lit., \textbf{guest} house)}\newline
-							\kazakh{Оны бір \textbf{рет} көрдім.} \gloss{I saw him/her/it one \textbf{time}.} \\\midrule
+						& \kazakh{\textbf{Қонақтарымыз} бай.} \sgloss{\textbf{Our guests} are rich.}\newline 
+							\kazakh{\textbf{қонақ} үй} \sgloss{hotel (lit., \textbf{guest} house)}\newline
+							\kazakh{Оны бір \textbf{рет} көрдім.} \sgloss{I saw him/her/it one \textbf{time}.} \\\midrule
 			Adjectives & Attributive\newline
@@
 							Adverbial
-						& \kazakh{\textbf{жақсы} үй} \gloss{\textbf{nice} house} \newline
-						\kazakh{ең \textbf{жақсы}лары} \gloss{\textbf{the nice}st \textbf{one}s} \newline
-						\kazakh{Оны \textbf{жақсы} танимын.} \gloss{I know him/her \textbf{well}.}
+						& \kazakh{\textbf{жақсы} үй} \sgloss{\textbf{nice} house} \newline
+						\kazakh{ең \textbf{жақсы}лары} \sgloss{\textbf{the nice}st \textbf{one}s} \newline
+						\kazakh{Оны \textbf{жақсы} танимын.} \sgloss{I know him/her \textbf{well}.}
 						\\\midrule
 			Adverbs & Adverbial
-						& \kazakh{Ол \textbf{кеше} кетті.} \gloss{S/he left \textbf{yesterday}.}
+						& \kazakh{Ол \textbf{кеше} кетті.} \sgloss{S/he left \textbf{yesterday}.}
 						\\
@@
 	\end{tabular}
+	}
 
@@
 \end{table}
-\vspace{-0.5em}
+
+Table~\ref{tab:casefunc} details the various tags (i.e., syntactic functions) associated with the traditionally defined morphological cases of Kazakh.
+It should be noted that there are several known uses which are currently unattested (partitive use of \gmk{abl} as \udtag{dobj}, embedded \udtag{subj} marked as \gmk{acc}) or rare (\gmk{dat} as \udtag{iobj}) in the corpus.
+
+\begin{table}[htbp]
+	\centering
+	\caption{The functions associated with each case in Kazakh as currently found in the corpus.}
+	\label{tab:casefunc}
+	\begin{tabular}{lrrrrrrrr}
+		\toprule
+			 & \udtag{iobj} & \udtag{cmpnd} & \udtag{nmod} & \udtag{nmod:poss} & \udtag{dobj} & \udtag{root} & \udtag{subj} & \udtag{voc} \\
+		\midrule
+			\gmk{nom} & — & 1 & 199 & 158 & 72 & 26 & 279 & 1 \\
+			\gmk{acc} & — & — & 1 & — & 101 & — & — & — \\
+			\gmk{gen} & — & — & 120 & 120 & — & — & 12 & — \\
+			\gmk{dat} & 1 & — & 113 & — & 13 & — & — & — \\
+			\gmk{loc} & — & — & 112 & 1 & — & 6 & — & — \\
+			\gmk{abl} & — & — & 64 & — & — & — & — & — \\
+			\gmk{ins} & — & — & 33 & — & — & — & — & — \\
+		\bottomrule
+	\end{tabular}
+\end{table}
+\vspace{-0.75em}
 
@@
 \end{figure}
-\vspace{-0.5em}
+\vspace{-0.75em}
 
@@
 In such instances, they receive the UD label \udtag{advcl}, and example of which is given in figure~\ref{fig:gerabl}.
+\vspace{-0.5em}
 

### 2015-08-26 2015-turklang-kazdep/paper.tex
some table 4 updates because of recent change
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61591 72bbbca6-d526-0410-a7d9-f06f51895060
a little more detail
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61592 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 
-Table~\ref{tab:casefunc} details the various tags (i.e., syntactic functions) associated with the traditionally defined morphological cases of Kazakh.
-It should be noted that there are several known uses which are currently unattested (partitive use of \gmk{abl} as \udtag{dobj}, embedded \udtag{subj} marked as \gmk{acc}) or rare (\gmk{dat} as \udtag{iobj}) in the corpus.
+Table~\ref{tab:casefunc} briefly details the various tags (i.e., syntactic functions) associated with the traditionally defined morphological cases of Kazakh as found in the corpus.
+Some mappings may be surprising, but will not be discussed in detail in this paper.
+It should also be noted that there are several known uses which are currently unattested (partitive use of \gmk{abl} as \udtag{dobj}, embedded \udtag{subj} marked as \gmk{acc}) or rare (\gmk{dat} as \udtag{iobj}) in the corpus.
 
@@
 		\midrule
-			\gmk{nom} & — & 1 & 199 & 158 & 72 & 26 & 279 & 1 \\
+			\gmk{nom} & — & 1 & 199 & 157 & 72 & 26 & 279 & 1 \\
 			\gmk{acc} & — & — & 1 & — & 101 & — & — & — \\
@@
 			\gmk{dat} & 1 & — & 113 & — & 13 & — & — & — \\
-			\gmk{loc} & — & — & 112 & 1 & — & 6 & — & — \\
+			\gmk{loc} & — & — & 113 & — & — & 6 & — & — \\
 			\gmk{abl} & — & — & 64 & — & — & — & — & — \\

### 2015-08-27 2015-turklang-kazdep/paper.tex
another fix to table 4
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61594 72bbbca6-d526-0410-a7d9-f06f51895060
cop → _, not .
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61596 72bbbca6-d526-0410-a7d9-f06f51895060
update numbers!
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61609 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 			\gmk{nom} & — & 1 & 199 & 157 & 72 & 26 & 279 & 1 \\
-			\gmk{acc} & — & — & 1 & — & 101 & — & — & — \\
+			\gmk{acc} & — & — & — & — & 102 & — & — & — \\
 			\gmk{gen} & — & — & 120 & 120 & — & — & 12 & — \\
@@
                    \begin{deptext}[column sep=0.2cm]
-                                Айгүл \& студент \& · \\
+                                Айгүл \& студент \& \_ \\
                                 \gmk{np} \& \gmk{n} \& \gmk{aor} \\
@@
 MaltOptimiser \parencite{ballesteros15}. The optimiser was run separately for each of 
-the model configurations.
+the model configurations, and the best combination of parameters was selected.
 
@@
 
-% system1:
+% system 1: 
+% LAS: 0.4285 [0.389+0.406+0.414+0.42+0.427+0.437+0.438+0.445+0.45+0.459]
+% UAS: 0.5842 [0.524+0.555+0.572+0.575+0.584+0.584+0.585+0.606+0.628+0.629]
+
+% system 2: 
 % nivreeager
-% LAS: 0.3341 [0.234 0.288 0.289 0.292 0.333 0.335 0.359 0.378 0.408 0.425]
-% UAS: 0.5093 [0.413 0.464 0.465 0.466 0.474 0.499 0.531 0.545 0.604 0.632]
+% LAS: 0.4285 [0.389+0.406+0.414+0.42+0.427+0.437+0.438+0.445+0.45+0.459]
+% UAS: 0.5842 [0.524+0.555+0.572+0.575+0.584+0.584+0.585+0.606+0.628+0.629]
 
-% system2:
+% system 3: 
 % nivreeager
-% LAS: 0.234+0.288+0.289+0.292+0.333+0.335+0.359+0.378+0.408+0.425
-% UAS: 0.413+0.464+0.465+0.466+0.474+0.499+0.531+0.545+0.604+0.632
+% LAS: 0.6499 [0.664+0.649+0.653+0.682+0.617+0.671+0.664+0.625+0.603+0.671]
+% UAS: 0.7703 [0.768+0.767+0.763+0.789+0.725+0.778+0.804+0.747+0.762+0.8]
 
-% system3:
+% system 4: 
 % nivreeager
-% LAS: 0.5529 [0.434 0.446 0.501 0.511 0.515 0.542 0.558 0.572 0.7 0.75]
-% UAS: 0.7376 [0.658 0.669 0.675 0.684 0.694 0.723 0.751 0.776 0.825 0.921]
-
-% system4:
-% covnonproj
-% LAS: 0.6092 [0.519 0.535 0.544 0.569 0.569 0.62 0.621 0.649 0.683 0.783]
-% UAS: 0.7575 [0.674 0.686 0.69 0.714 0.727 0.746 0.795 0.803 0.812 0.928]
-
-%             [0.434+0.446+0.501+0.511+0.515+0.542+0.558+0.572+0.7+0.75]
-%             [0.658+0.669+0.675+0.684+0.694+0.723+0.751+0.776+0.825+0.921]
+% LAS: 0.7688 [0.709+0.725+0.764+0.775+0.776+0.778+0.785+0.786+0.788+0.802]
+% UAS: 0.8143 [0.757+0.784+0.809+0.81+0.82+0.823+0.825+0.831+0.832+0.852]
 
@@
 		\midrule
-			surface                & nivreeager  & 33.4 [23.4, 42.5] & 50.9 [41.3, 63.2] \\
-			surface+lemma          & nivreeager  & 33.4 [23.4, 42.5] & 50.9 [41.3, 63.2] \\
-			surface+lemma+POS      & nivreeager  & 55.2 [43.4, 75.0] & 73.7 [65.8, 92.1] \\
-			surface+lemma+POS+MSD  & nivreeager  & 55.2 [43.4, 75.0] & 73.7 [65.8, 92.1] \\
+			surface                & nivreeager  & 42.8 [38.9, 45.9]  & 58.4 [52.4, 62.9] \\
+			surface+lemma          & nivreeager  & 42.8 [38.9, 45.9]  & 58.4 [52.4, 62.9]   \\
+			surface+lemma+POS      & nivreeager  & 64.9 [66.4, 67.1]  & 77.0 [76.8, 80.0] \\
+			surface+lemma+POS+MSD  & nivreeager  & 76.8 [70.9, 80.0]  & 81.4 [75.7, 85.2] \\
 		\bottomrule
@@
 sized treebanks, for example the Irish treebank of \textcite{Lynn12}, for which an LAS of
-63.3 and a UAS of 73.3 were reported with the best model.
-Adding structural features to the model substantially improves the performance of the parser, although we find no improvement in using full morphosyntactic description (MSD) over using simply the first part-of-speech (POS) tag.
+63.3 and a UAS of 73.3 were reported with the best model.  Adding structural features to the model 
+substantially improves the performance of the parser.
 
@@
 \end{document}
+
+%%%%%%%%%%%%%%%%%%%
+%%
+%% OLD RESULTS:
+
+% system1:
+% nivreeager
+% LAS: 0.3341 [0.234 0.288 0.289 0.292 0.333 0.335 0.359 0.378 0.408 0.425]
+% UAS: 0.5093 [0.413 0.464 0.465 0.466 0.474 0.499 0.531 0.545 0.604 0.632]
+
+% system2:
+% nivreeager
+% LAS: 0.234+0.288+0.289+0.292+0.333+0.335+0.359+0.378+0.408+0.425
+% UAS: 0.413+0.464+0.465+0.466+0.474+0.499+0.531+0.545+0.604+0.632
+
+% system3:
+% nivreeager
+% LAS: 0.5529 [0.434 0.446 0.501 0.511 0.515 0.542 0.558 0.572 0.7 0.75]
+% UAS: 0.7376 [0.658 0.669 0.675 0.684 0.694 0.723 0.751 0.776 0.825 0.921]
+
+% system4:
+% covnonproj
+% LAS: 0.6092 [0.519 0.535 0.544 0.569 0.569 0.62 0.621 0.649 0.683 0.783]
+% UAS: 0.7575 [0.674 0.686 0.69 0.714 0.727 0.746 0.795 0.803 0.812 0.928]
+
+%             [0.434+0.446+0.501+0.511+0.515+0.542+0.558+0.572+0.7+0.75]
+%             [0.658+0.669+0.675+0.684+0.694+0.723+0.751+0.776+0.825+0.921]

### 2015-08-27 2015-turklang-kazdep/paper.tex
nsubj and table 1 labels
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61613 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 			\midrule
-				nom. & - & subject\newline
+				\gmk{nom} & — & subject\newline
 										attributive\newline
@@
 										\kazakh{\textbf{үй} жануарлары} \sgloss{\textbf{house} animals} \\\midrule
-				acc. & -/NI/ & definite object %\newline
+				\gmk{acc} & -/NI/ & definite object %\newline
 										%	embedded subject\newline
@@
 										\\\midrule
-				gen. & -/NIң/ & definite genitival\newline
+				\gmk{gen} & -/NIң/ & definite genitival\newline
 											embedded subject
@@
 											\kazakh{Ол \textbf{Айгүлдің} ойнағанына қарап тұр.} \sgloss{She's watching \textbf{Aygül} play.} \\\midrule
-				loc. & -/DA/ & adverbial
+				\gmk{loc} & -/DA/ & adverbial
 										& \kazakh{\textbf{Үйде} ұйықтап жатыр.} \sgloss{S/he's sleeping \textbf{in the house}.} \\\midrule
-				abl. & -/DAн/ & adverbial\newline
+				\gmk{abl} & -/DAн/ & adverbial\newline
 											comparator 
@@
 											\kazakh{\textbf{түйеден} үлкен} \sgloss{bigger \textbf{than a camel}}\\\midrule
-				dat. & -/GA/ & indirect object\newline
+				\gmk{dat} & -/GA/ & indirect object\newline
 										adverbial\newline
@@
 										\\\midrule
-				inst. & -/Mен/ &  adverbial & \kazakh{Олар \textbf{күшікпен} ойнап тұр.} \sgloss{They're playing \textbf{with the dog}.} \\
+				\gmk{ins} & -/Mен/ &  adverbial & \kazakh{Олар \textbf{күшікпен} ойнап тұр.} \sgloss{They're playing \textbf{with the dog}.} \\
 			\bottomrule
@@
 Some mappings may be surprising, but will not be discussed in detail in this paper.
-It should also be noted that there are several known uses which are currently unattested (partitive use of \gmk{abl} as \udtag{dobj}, embedded \udtag{subj} marked as \gmk{acc}) or rare (\gmk{dat} as \udtag{iobj}) in the corpus.
+It should also be noted that there are several known uses which are currently unattested (partitive use of \gmk{abl} as \udtag{dobj}, embedded \udtag{nsubj} marked as \gmk{acc}) or rare (\gmk{dat} as \udtag{iobj}) in the corpus.
 
@@
 		\toprule
-			 & \udtag{iobj} & \udtag{cmpnd} & \udtag{nmod} & \udtag{nmod:poss} & \udtag{dobj} & \udtag{root} & \udtag{subj} & \udtag{voc} \\
+			 & \udtag{iobj} & \udtag{cmpnd} & \udtag{nmod} & \udtag{nmod:poss} & \udtag{dobj} & \udtag{root} & \udtag{nsubj} & \udtag{voc} \\
 		\midrule
@@
                         \deproot[edge unit distance=1.3ex]{2}{\udlabel{root}}
-                        \depedge{2}{1}{\udlabel{subj}}
+                        \depedge{2}{1}{\udlabel{nsubj}}
                         \depedge{2}{3}{\udlabel{cop}}
@@
                         \deproot[edge unit distance=1.3ex]{5}{\udlabel{root}}
-                        \depedge{3}{2}{\udlabel{subj}}
+                        \depedge{3}{2}{\udlabel{nsubj}}
                         \depedge{5}{3}{\udlabel{ccomp}}
@@
 Verbal nouns allow a whole verb phrase to be used as a nominal, and the resulting verbal noun phrase may be used as a subject or object of a verb.
-When used as the subject of a verb, verbal noun phrases receive the UD label \udtag{subj}.
+When used as the subject of a verb, verbal noun phrases receive the UD label \udtag{nsubj}.
 When used as the object of a verb, verbal noun phrases receive the UD label \udtag{ccomp}, unless the subject of the subordinate clause is obligatorily identical to the subject of the main clause, in which case it receives \udtag{xcomp}.
@@
 		\depedge{5}{4}{\udlabel{cc}}
-		\depedge{6}{5}{\udlabel{subj}}
+		\depedge{6}{5}{\udlabel{nsubj}}
 		\depedge{7}{6}{\udlabel{ccomp}}

### 2015-09-14 2015-turklang-kazdep/poster/reveal.sh
get reveal.js
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61887 72bbbca6-d526-0410-a7d9-f06f51895060

@@ -1,0 +1,4 @@
+#!/usr/bin/env bash
+
+wget https://github.com/hakimel/reveal.js/archive/3.1.0.tar.gz
+tar xzf 3.1.0.tar.gz

### 2015-09-17 2015-turklang-kazdep/slides/trees/minimal.tex
today's work
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@61928 72bbbca6-d526-0410-a7d9-f06f51895060

\documentclass[landscape,a5paper,12pt]{minimal}
\usepackage{tikz-dependency}

\usepackage{fontspec}
\usepackage{xunicode}
\usepackage{xltxtra}
\usepackage[margin=0.1in,landscape]{geometry}
\geometry{bmargin=0in}

\setromanfont{Times New Roman}

\newfontfamily\ll[]{Linux Libertine O}
%\newfontfamily\udfont[Scale=1.3,Letters=SmallCaps]{Linux Libertine O}
\newfontfamily\udfont[Scale=1.3,Color=FFFFFF99]{Alegreya Sans SC}
\newfontfamily\gmkfont[Scale=1,Color=FFFFFF99]{Alegreya Sans SC}
\newfontfamily\qgmk[Scale=MatchLowercase,Letters=SmallCaps,Color=FFFFFF99]{Linux Libertine O}

\newcommand{\white}[1]{{\addfontfeature{Color=FFFFFF99}#1}}

\newcommand{\gmk}[1]{{\gmkfont #1}}
\newcommand{\tgloss}[1]{{\em #1}}
\newcommand{\tag}[1]{{\ll \textsc{‹#1›}}}
\newcommand{\udtag}[1]{{\ll \textsc{#1}}}
\newcommand{\udlabel}[1]{{\udfont #1}}

\newcommand{\scriptsize}[1]{#1}

\begin{document}

\resizebox{\textwidth}{!}{%
	\centering
		\white{
		\begin{dependency}[theme = simple, edge style={white!60!white,very thick}] %, font = \small]
			\begin{deptext}[column sep=0.2cm]
				¶
		\end{dependency}}
}
\end{document}

### 2015-09-17 2015-turklang-kazdep/slides/trees/maketrees.py

@@ -1,0 +1,23 @@
+#!/usr/bin/env python3
+
+import glob, os
+
+with open('minimal.tex', 'r') as templateFile:
+	template = templateFile.read()
+
+for fn in glob.glob("*.dat"):
+	graphname=fn.replace(".dat", "")
+	with open(fn, 'r') as graphFile:
+		contents = graphFile.read()
+	output = template.replace("¶", contents)
+	with open(graphname+".tex", 'w') as outFile:
+		outFile.write(output)
+
+#print(a)
+
+#replacementPlace = re.compile("¶")
+#x = replacementPlace.sub(a, template)
+
+#with open('
+#
+#print(x)

### 2015-09-17 2015-turklang-kazdep/slides/trees/maketrees.sh

@@ -1,0 +1,11 @@
+#!/usr/bin/env bash
+
+python3 maketrees.py
+#for i in *.tex; do
+#	xelatex $i;
+#done;
+
+find *.tex ! -name minimal.tex -exec xelatex {} \;
+
+find *.pdf -exec inkscape {} -d 200 -D -e {}.png \;
+find *.pdf -exec inkscape {} -D -l {}.svg \;

### 2016-05-13 2016-lrec-tyvmorph/poster.tex
add stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@68334 72bbbca6-d526-0410-a7d9-f06f51895060
stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@68337 72bbbca6-d526-0410-a7d9-f06f51895060

--- 2015-07-09 2015-compmorphon-morphver/poster.tex
@@
 			\vspace{-2cm}
-			\noindent\includegraphics[height=4.2em]{DCU_logo_2col}\\
-			\vspace{-0.5ex}\hspace{0.105cm}\includegraphics[height=2.15em]{DCU_2010_Name_Mark_289}
+			\noindent\\
+			\vspace{-0.5ex}\hspace{0.105cm}
 		\end{minipage}
@@
 				\begin{spacing}{0.4}
-					{Tommi Pirinen\vphantom{y}}\\
-					{\footnotesize Ollscoil Chathair Bhaile Átha Cliath\vphantom{y}\\\texttt{tommi.pirinen@computing.dcu.ie}}
+					{Jonathan North Washington}\\
+					{\footnotesize Indiana University\\\texttt{jonwashi@indiana.edu}}
 				\end{spacing}
@@
 				\begin{spacing}{0.4}
-					{Jonathan North Washington}\\
-					{\footnotesize Indiana University\\\texttt{jonwashi@indiana.edu}}
+					{Aziyana Bayyr-ool}\\
+					{\footnotesize XXX\\\texttt{yyy@xxx}}
 				\end{spacing}
@@
 				\begin{itemize}
-					\item /k/ ‹к› elides at the end of >1-σ stems intervocalically..\\
+					\item /k/ ‹к› elides at the end of >1-σ stems intervocalically\\
 						инэк+I → инээ, өк+I → өгү

### 2016-05-13 2016-lrec-tyvmorph/poster.tex
A0 paper
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@68338 72bbbca6-d526-0410-a7d9-f06f51895060
some title and author formatting
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@68341 72bbbca6-d526-0410-a7d9-f06f51895060

@@
-\documentclass[fontscale=0.33,landscape,paperwidth=48in,paperheight=36in]{baposter}  % fontscale=0.285, dvipdfm
+\documentclass[fontscale=0.33,landscape,paperwidth=841mm,paperheight=1189mm]{baposter}  % fontscale=0.285, dvipdfm
 
@@
 	% To get it to be the same size consistently on all machines..
-	\special{papersize=48in,36in}
+	\special{papersize=841mm,1189mm}
 	\setlength{\pdfpageheight}{\paperheight}
@@
 		\includegraphics[height=6.5em]{uitlogo}
-		\hspace{2.5cm}\begin{minipage}[t]{7em}
-			\vspace{-2cm}
-			\noindent\\
-			\vspace{-0.5ex}\hspace{0.105cm}
-		\end{minipage}
+		%\hspace{2.5cm}\begin{minipage}[t]{7em}
+		%	\vspace{-2cm}
+		%	\noindent\\
+		%	\vspace{-0.5ex}\hspace{0.105cm}
+		%\end{minipage}
 
@@
 		{%\vspace{0pt}\hspace{1cm} \
-			{\titlefont \scshape{Finite-State Morphologies \& Text Corpora\\\vspace{0.2em}\scalebox{0.8}{as Resources for Improving Morphological Descriptions}}}}
+			{\titlefont \scshape{A finite-state morphological analyser for Tuvan}
+			}
+		}
 	}{ %Authors (centered top, below title)
@@
 			\end{minipage}
-			\begin{minipage}[t]{9.5em}
+			\begin{minipage}[t]{12em}
 				\begin{spacing}{0.4}
@@
 			\end{minipage}
-			\begin{minipage}[t]{12em}
+			\begin{minipage}[t]{9em}
 				\begin{spacing}{0.4}

### 2016-05-13 2016-lrec-tyvmorph/poster.tex
authors updates
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@68353 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 	}{ %Eye Catcher, empty if option eyecatcher=false - unused
-		\includegraphics[height=6.5em]{uitlogo}
+		\includegraphics[height=6.4em]{uitlogo}
+		\includegraphics[height=6.2em]{iu_tab-crop}
 		%\hspace{2.5cm}\begin{minipage}[t]{7em}
@@
 		\begin{center}
-			\begin{minipage}[t]{9em}
+			\begin{minipage}[t]{5.75em}
 				\begin{spacing}{0.4}
-				{Francis M.\ Tyers}\\
+				{Francis Tyers}\\
 				{\footnotesize UiT Norgga Árktalaš Universitehta \\\texttt{francis.tyers@uit.no}}
@@
 			\end{minipage}
-			\begin{minipage}[t]{12em}
+			\begin{minipage}[t]{9.25em}
 				\begin{spacing}{0.4}
-					{Jonathan North Washington}\\
+					{Jonathan Washington}\\
 					{\footnotesize Indiana University\\\texttt{jonwashi@indiana.edu}}
@@
 			\end{minipage}
-			\begin{minipage}[t]{9em}
+			\begin{minipage}[t]{8em}
 				\begin{spacing}{0.4}
 					{Aziyana Bayyr-ool}\\
-					{\footnotesize XXX\\\texttt{yyy@xxx}}
+					{\footnotesize Сибирское отделение Российской академии наук\\\texttt{azikoa@mail.ru}}
+				\end{spacing}
+			\end{minipage}
+			\begin{minipage}[t]{6em}
+				\begin{spacing}{0.4}
+					{Aelita Salchak}\\
+					{\footnotesize Тываның Күрүне Университеди\\\texttt{aelita\_74@mail.ru}}
 				\end{spacing}
@@
 	}{ %More eye catchers, on the right side
-		\includegraphics[height=6.25em]{iu_tab-crop}
-		\hspace{2.5cm}\includegraphics[angle=90,height=6.8em]{apertium5b}%\hspace{-2em}
+		\includegraphics[height=6.2em]{img/Tuvsu}
+		\includegraphics[height=6.2em]{img/soranif}
+		%\hspace{2.5cm}
+		\includegraphics[angle=90,height=6.7em]{apertium5b}%\hspace{-2em}
 	}

### 2016-05-13 2016-lrec-tyvmorph/poster.tex
a font adjustment that helps things fit
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@68354 72bbbca6-d526-0410-a7d9-f06f51895060

@@
-\documentclass[fontscale=0.33,landscape,paperwidth=841mm,paperheight=1189mm]{baposter}  % fontscale=0.285, dvipdfm
+\documentclass[fontscale=0.35,landscape,paperwidth=841mm,paperheight=1189mm]{baposter}  % fontscale=0.285, dvipdfm
 
@@
 		\begin{center}
-			\begin{minipage}[t]{5.75em}
+			\begin{minipage}[t]{7em}
 				\begin{spacing}{0.4}
-				{Francis Tyers}\\
+				{Francis M.\ Tyers}\\
 				{\footnotesize UiT Norgga Árktalaš Universitehta \\\texttt{francis.tyers@uit.no}}
@@
 			\end{minipage}
-			\begin{minipage}[t]{9.25em}
+			\begin{minipage}[t]{12em}
 				\begin{spacing}{0.4}
-					{Jonathan Washington}\\
+					{Jonathan North Washington}\\
 					{\footnotesize Indiana University\\\texttt{jonwashi@indiana.edu}}

### 2016-05-13 2016-lrec-tyvmorph/poster.tex
some more title hargling
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@68356 72bbbca6-d526-0410-a7d9-f06f51895060
more title etc. hargling
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@68357 72bbbca6-d526-0410-a7d9-f06f51895060

--- 2016-05-13 2016-lrec-tyvmorph/poster.tex
@@
 		headerborder=open,
-		headerheight=0.1\textheight,
+		headerheight=0.07\textheight,
 		%headershape=roundedright,
@@
 	}{ %Eye Catcher, empty if option eyecatcher=false - unused
-		\includegraphics[height=6.4em]{uitlogo}
-		\includegraphics[height=6.2em]{iu_tab-crop}
+		\setlength\fboxsep{0.5em}
+		\setlength\fboxrule{0pt}
+		\includegraphics[height=5.5em]{qrcode}
+		\includegraphics[height=5.8em]{uitlogo}\hspace{1em}
+		\fbox{\includegraphics[height=5.5em]{iu_tab-crop}}
 		%\hspace{2.5cm}\begin{minipage}[t]{7em}
@@
 	}{ %Authors (centered top, below title)
-		\vspace{-0.6em}
+		%\vspace{-0.6em}
 		\begin{center}
-			\begin{minipage}[t]{5.75em}
+			\begin{minipage}[t]{6em}
 				\begin{spacing}{0.4}
@@
 			\end{minipage}
-			\begin{minipage}[t]{8em}
+			\begin{minipage}[t]{8em} %{7.7em}
 				\begin{spacing}{0.4}
@@
 	}{ %More eye catchers, on the right side
-		\includegraphics[height=6.2em]{img/Tuvsu}
-		\includegraphics[height=6.2em]{img/soranif}
+		\setlength\fboxsep{0.5em}
+		\setlength\fboxrule{0pt}
+		\fbox{\includegraphics[height=5.5em]{img/Tuvsu}}
+		\includegraphics[height=5.4em]{img/soranif}
+		\includegraphics[angle=90,height=6.2em]{apertium5b}%\hspace{-2em}
 		%\hspace{2.5cm}
-		\includegraphics[angle=90,height=6.7em]{apertium5b}%\hspace{-2em}
 	}
@@
 			%\vspace{-0.2em}
-			\tikzoverlay[text width=0.8in] at (6.73in, 0.04in) {
-				\includegraphics[height=0.8in]{qrcode}
-			};
+			%\tikzoverlay[text width=0.8in] at (6.73in, 0.04in) {
+			%	\includegraphics[height=0.8in]{qrcode}
+			%};
 			\begin{itemize}

### 2016-05-14 2016-lrec-tyvmorph/poster.tex
fresh [mostly] empty poster
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@68415 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 		%headerColorOne=blue,
-		headerFontColor=white,
+		headerFontColor=yellow,
 		% Only simple background color used, no shading, so boxColorTwo isn't necessary
@@
 		textborder=roundedleft,
-		columns=4,
+		columns=3,
 	}{ %Eye Catcher, empty if option eyecatcher=false - unused
@@
 		\includegraphics[height=5.8em]{uitlogo}\hspace{1em}
-		\fbox{\includegraphics[height=5.5em]{iu_tab-crop}}
+		\fbox{\includegraphics[height=5.4em]{iu_tab-crop}}
 		%\hspace{2.5cm}\begin{minipage}[t]{7em}
@@
 
-	\headerbox{Morphological Descriptions}{name=morphdesc,column=0,row=0}{
-		\htwo{What are they?}
-		\begin{itemize}
-			\item Core part of grammatical descriptions
-			\item Found in reference grammars and textbooks
-		\end{itemize}
-
-		\htwo{What they describe} %Describe a language's
-		\begin{itemize}
-			\item \textbf{Morphotactics} — the morphemes of a language and how they can be combined
-			\item \textbf{Morphophonology} — the alternations between the various phonological and orthographical forms of each morpheme% in a language
-		\end{itemize}
-
-		\htwo{Reasons for being \hilitetwo{incomplete}}
-		\begin{itemize}
-			\item Restrictions of the medium (e.g., number of pages)
-			\item Limitations of introspection and working with native speakers
-			\item \textbf{Lack of automatic testing}
-		\end{itemize}
+	\headerbox{Tuvan}{name=tyvan,column=0,row=0}{
 	}
 
-	\headerbox{Morphological Transducers}{name=morphtrans,below=morphdesc}{
+	\headerbox{Morphological Transducers}{name=morphtrans,below=tyvan}{
 		\htwo{Morphological transducers}
@@
 	}
-
-		\headerbox{Example: Tuvan velar elision}{name=tuvanvelars,column=2,span=2}{
-			\begin{multicols}{2}
-				\htwo{Previous descriptions}
-
-					\begin{itemize}
-						\item \textcite[9, 22-23]{andersonharrison99} %Anderson and Harrison (1999, pp.\ 9, 22-23)
-					\end{itemize}
-					\includegraphics[width=0.49\textwidth]{img/andersonharrison9}\\
-					\includegraphics[width=0.49\textwidth]{img/andersonharrison22-23}
-					\begin{itemize}
-						\item \textcite[117-118]{isxakovpalmbax61} %Исхаков, Ф.Г.; Пальмбах (1961, pp.\ 117-118)
-					\end{itemize}
-					\includegraphics[width=0.49\textwidth]{img/isxakovpalmbax117}
-
-				\htwo{More complete characterisation}
-				\begin{itemize}
-					\item /k/ ‹к› elides at the end of >1-σ stems intervocalically\\
-						инэк+I → инээ, өк+I → өгү
-					\item /g/ ‹г› elides at the end of any stem intervocalically\\
-						өг+I → өө 
-					\item /ŋ/ ‹ң› elides at the end of \textbf{some} stems intervocalically\\
-						\textbf{соң+I → соо}, чаң+I → чаңы, түң+I → түңү
-				\end{itemize}
-
-				\htwo{Correct forms in the grammars}
-
-				But not part of the morphophonological descriptions.\\
-				(A. \& H., p.\ 35)\hfill{}\includegraphics[width=0.2\textwidth]{img/andersonharrison35}\\
-				(И.\ \& П., p.\ 447)\hfill{}\includegraphics[width=0.35\textwidth]{img/isxakovpalmbax447}
-
-			\columnbreak{}
-				\htwo{Unanalysed forms before implementation}
-					\begin{itemize}
-						\item \hilitetwo{инээ}, \hiliteone{өгү}, \hilitetwo{өө}, \hilitetwo{соо}, \hiliteone{чаңы}
-					\end{itemize}\vspace{-0.3ex}
-					\begin{minipage}[t]{0.5\textwidth}
-					\outputex{
-						\begin{multicols}{4}
-							1590 соонда\vphantom{ү}\\
-							610 те\vphantom{ү}\\
-							153 өөнүң\vphantom{ү}\\
-							48 соондан\vphantom{ү}\\
-							46 өөм\vphantom{ү}\\
-							20 соон\vphantom{ү}\\
-							18 өө\vphantom{ү}\\
-							17 өөнге\vphantom{ү}\\
-							9 соондагы\vphantom{ү}\\
-							8 өөн\vphantom{ү}\columnbreak\\
-							6 инээниң\vphantom{ү}\\
-							6 соонче\vphantom{ү}\\
-							6 соонга\vphantom{ү}\\
-							3 инээм\vphantom{ү}\\
-							3 өөнден\vphantom{ү}\\
-							3 өөнде\vphantom{ү}\\
-							3 өөмнүң\vphantom{ү}\\
-							2 инээн\vphantom{ү}\\
-							2 инээ\vphantom{ү}\\
-							2 өөнче\vphantom{ү}\columnbreak\\
-							2 өөңерге\vphantom{ү}\\
-							2 өөңер\vphantom{ү}\\
-							2 өөмнү\vphantom{ү}\\
-							2 өөвүстен\vphantom{ү}\\
-							2 өөвүске\vphantom{ү}\\
-							2 өөвүс\vphantom{ү}\\
-							1 инээңни\vphantom{ү}\\
-							1 инээмниң\vphantom{ү}\\
-							1 Инээм\vphantom{ү}\\
-							1 инээвистиң\vphantom{ү}\\
-							1 өөңнүң\vphantom{ү}\\
-							1 өөң\vphantom{ү}\\
-							1 өөвүсче\vphantom{ү}\\
-							1 өөвүстүң\vphantom{ү}\\
-							1 өөвүстү\vphantom{ү}\\
-							1 өөвүсте
-						\end{multicols}}
-					\end{minipage}
-
-				\htwo{Initial implementation}
-					\begin{minipage}[t]{0.5\textwidth}
-					\codeex{"Intervocalic voiced velar deletion"\\
-г:0 <=> :Vow/:0* \_ [ \%>: :Vow ]/:0* ;\\
-\-~\\
-"Intervocalic voiceless velar deletion"\\
-к:0 <=> :Vow/:0* \_ [ \%>: :Vow ]/:0* ;\\
-\-~~~~except\\
-\-~~~~~~~~[ .\#.\ | \%- ] [ ( :Cns* ) ( :Vow* ) :Vow ]/:0 \_ [ \%>:\ :Vow ]/:0* ; %[ :0 - \%>:\ ]* ;
-					}
-					\end{minipage}
-
-				\htwo{Unanalysed forms after initial implementation}
-					%(has forms of соң in it)
-					\begin{itemize}
-						\item \hiliteone{инээ}, \hiliteone{өгү}, \hiliteone{өө}, \hilitetwo{соо}, \hiliteone{чаңы}
-					\end{itemize}\vspace{-0.3ex}
-					\begin{minipage}[t]{0.5\textwidth}
-					\outputex{
-						\begin{multicols}{3}
-							1590 соонда\vphantom{ң}\\
-							610 те\vphantom{ң}\\
-							48 соондан\vphantom{ң}\\
-							20 соон\vphantom{ң}\\
-							6 соонче\vphantom{ң}\\
-							6 соонга
-						\end{multicols}
-					}
-					\end{minipage}
-
-				\htwo{Second attempt at implementation}
-
-					\begin{minipage}[t]{0.5\textwidth}
-					\codeex{\hilitegrey{"Intervocalic voiced velar deletion"}\\
-Cx\hilitegrey{:0 <=> :Vow/:0* \_ [ \%>: :Vow ]/:0* ;}\\
-\-~~~~~where Cx in ( г ң ) ;\\
-\-~\\
-\hilitegrey{"Intervocalic voiceless velar deletion"\\
-к:0 <=> :Vow/:0* \_ [ \%>:\ :Vow ]/:0* ;\\
-\-~~~~except\\
-\-~~~~~~~~[ .\#.\ | \%- ] [ ( :Cns* ) ( :Vow* ) :Vow ]/:0 \_ [ \%>:\ :Vow ]/:0* ;} %[ :0 - \%>:\ ]* ;
-					}
-					\end{minipage}
-
-				\htwo{Unanalysed forms after second implementation}
-					\begin{itemize}
-						\item \hiliteone{инээ}, \hiliteone{өгү}, \hiliteone{өө}, \hiliteone{соо}, \hilitetwo{чаңы}
-					\end{itemize}\vspace{-0.3ex}
-					\begin{minipage}[t]{0.5\textwidth}
-					\outputex{
-						\begin{multicols}{3}
-							610 те\vphantom{ү}\\
-							203 бажыңынга\vphantom{ү}\\
-							164 бажыңының\vphantom{ү}\\
-							102 бажыңы\vphantom{ү}\\
-							86 бажыңын\vphantom{ү}\\
-							56 түңү\vphantom{ү}\\
-							47 чаңы\vphantom{ү}\\
-							44 чаңын\vphantom{ү}\\
-							13 түңүн\vphantom{ү}\\
-							12 чаңының\vphantom{ү}\\
-							5 түңүнден\vphantom{ү}\\
-							4 чаңындан\vphantom{ү}\\
-							4 чаңында\vphantom{ү}\\
-							3 түңүнүң\vphantom{ү}\\
-							2 түңүнге
-						\end{multicols}}
-					\end{minipage}
-
-					%(has forms of бажың, түң, чаң in it)
-
-				\htwo{Final implementation}
-
-					\begin{minipage}[t]{0.5\textwidth}
-					\codeex{\hilitegrey{"Intervocalic voiced velar deletion"\\
-Cx:0 <=> :Vow/:0* \_ [ \%>:\ :Vow ]/:0* ;}\\
-\-~~~~~except\\
-\-~~~~~~~~~:Vow \_ [ \%\{☭\%\}: :Vow ]/:0* ; \\
-\-~~~~~\hilitegrey{where Cx in ( г ң ) ;\\
-\-~\\
-"Intervocalic voiceless velar deletion"\\
-к:0 <=> :Vow/:0* \_ [ \%>:\ :Vow ]/:0* ;\\
-\-~~~~except\\
-\-~~~~~~~~[ .\#.\ | \%- ] [ ( :Cns* ) ( :Vow* ) :Vow ]/:0 \_ [ \%>:\ :Vow ]/:0* ;%\\
-%\-~~~~~~~~:Vow \_ [ \%\{☭\%\}:\ :Vow ]/:0* ;
-					}}
-					\end{minipage}
-
-				\htwo{Unanalysed forms after final implementation}
-					\begin{itemize}
-						\item \hiliteone{инээ}, \hiliteone{өгү}, \hiliteone{өө}, \hiliteone{соо}, \hiliteone{чаңы}
-					\end{itemize}\vspace{-0.3ex}
-					\begin{minipage}[t]{0.5\textwidth}
-					\outputex{
-						\begin{multicols}{3}
-							610 те\\
-						\end{multicols}}
-					\end{minipage}
-					\begin{itemize}
-						\item All forms of velar deletion are now analysed correctly!
-						\item Can now move on to other unanalysed forms\ldots{}
-					\end{itemize}
-			\end{multicols}
-
-			%\htwo{Examples of correct forms}
-
-		}
-
-		\headerbox{Example: Tuvan vowel harmony}{name=tuvanvowelharmony,column=1}{ %{name=otherexamples,column=1,span=2,below=tuvanvelars}{ % after consonant clusters in loanwords
-			%\begin{multicols}{2}
-				\htwo{Tuvan vowel harmony}
-					\begin{itemize}
-						\item High and low vowels harmonise in \hilitetwo{back}ness\\
-						киж\hiliteone{и}\tag{n}\tag{pl}\tag{acc}: киж\hiliteone{и}л\hiliteone{е}рн\hiliteone{и}, н\hilitetwo{о}м\tag{n}\tag{pl}\tag{acc}: н\hilitetwo{о}мн\hilitetwo{а}рн\hilitetwo{ы}
-						\item High vowels also harmonise in roundness\\
-						round: \hiliteone{ө}г\tag{n}\tag{acc}:\ \hiliteone{ө}гн\hiliteone{ү}, \hilitetwo{о}й\tag{n}\tag{acc}:\ \hilitetwo{о}йн\hilitetwo{у}\\
-						unround: ү\hiliteone{е}\tag{n}\tag{acc}:\ ү\hiliteone{е}н\hiliteone{и}, \hilitetwo{а}й\tag{n}\tag{acc}:\ \hilitetwo{а}йн\hilitetwo{ы}
-					\end{itemize}
-				\htwo{Initial implementation (high-vowels)}
-
-				%\begin{minipage}{0.5\textwidth}
-					\codeex{"\{I\} harmony"\\
-\%\{I\%\}:Vy <=> :Vx [ :Cns* :RealCns ]/[ :0 | \%- ]* \_ ;\\
-\-~~~~~~~~~where Vx in ( ү и е э ө а о ы у я ё ю )\\
-\-~~~~~~~~~~~~~~ Vy in ( ү и и и ү ы у ы у ы у у )\\
-\-~~~~~~~~~matched ;}
-				%\end{minipage}
-
-				\htwo{Unanalysed forms in corpus}
-					\begin{itemize}
-						\item forms of loanwords ансамбль `ensemble' and рубль `rouble':\\\
-						the following vowel is always front unround
-					\end{itemize}
-					%\begin{minipage}[t]{0.5\textwidth}
-					\outputex{
-						\begin{multicols}{3}
-							%790 рубль\\
-							347 рубльди\\
-							99 рубльге\\
-							83 ансамблиниң\vphantom{у}\\
-							60 ансамбли\vphantom{у}\\
-							54 рубльдиң\\
-							28 рубльден\\
-							17 ансамбльдиң\vphantom{у}\\
-							15 ансамбльдер\vphantom{у}\\
-							14 ансамбль\vphantom{у}\\
-							11 ансамблин\vphantom{у}\\
-							10 ансамблинге\vphantom{у}\\
-							8 ансамбльди\vphantom{у}\\
-							%2 рубльга\\
-							1 рубльдерни\\
-							1 рублин\\
-							1 рубли\\
-						\end{multicols}
-					}
-					%\end{minipage}
-
-					%\columnbreak
-				\htwo{Revised generalisation}
-					\begin{itemize}
-						\item After cons.\ cluster ending in ль, vowel always front unround
-						\item New discovery?  Appears not to be previously documented
-					\end{itemize}
-
-				\htwo{Revised implementation}
-				%\begin{minipage}{0.5\textwidth}
-					\codeex{\hilitegrey{"\{I\} harmony"\\
-\%\{I\%\}:Vy <=> :Vx [ :Cns* :RealCns ]/[ :0 | \%- ]* \_ ;}\\
-\-~~~~~~~~~except\\
-\-~~~~~~~~~~~~~[ :BackVow :Cns* :Cns л: ь: :Cns* :RealCns ]/:0* \_ ;\\
-\-~~~~~~~~~~~~~[ :BackVow :Cns* :Cns л: ь:0 ]/:[ :0 - ь: ]* \_ ;\\
-\-~~~~~~~~~\hilitegrey{where Vx in ( ү и е э ө а о ы у я ё ю )\\
-\-~~~~~~~~~~~~~~ Vy in ( ү и и и ү ы у ы у ы у у )\\
-\-~~~~~~~~~matched ;}\\
-\\
-"\{I\} always front when intervening Cль"\\
-\%\{I\%\}:и <=> [ :BackVow :Cns* :Cns л: ь: :Cns* :RealCns ]/:0* \_ ;\\
-\-~~~~~~~~~~~~[ :BackVow :Cns* :Cns л: ь:0 ]/:[ :0 - ь: ]* \_ ;}
-				%\end{minipage}
-			%\end{multicols}
-
-		}
-
-	\headerbox{Example: Categorisation}{name=adjectives,column=1,below=tuvanvowelharmony}{
-		\htwo{Unrestricted 0-derivation}
-		\begin{itemize}
-			\item Common overgeneralisation of adjective morphology:
-			\begin{itemize}
-				\item all adjectives may act like nouns (i.e.\ take noun morphology)
-				\item all adjectives may be used as is as adverbs
-			\end{itemize}
-			\item Commonly reported in Turkic language grammars
-		\end{itemize}
-		%Somfai Kara (2002, pp. 28-29):\ \includegraphics[width=0.53\textwidth]{img/kara28}\\
-		\textcite[28-29]{somfaikara02}:\ \includegraphics[width=0.53\textwidth]{img/kara28}\\
-		\-\hfill{}\includegraphics[width=0.8\textwidth]{img/kara29}
-
-		\htwo{Attested system}
-			\begin{itemize}
-				\item Some adjectives may not be substantivised
-				\item Some adjectives may not be used adverbially
-				\item Some adjectives do not have comparative forms
-				\item …a range of “adjective classes” in most Turkic languages
-			\end{itemize}
-		\htwo{Implementation:\ proper categorisation}
-			\begin{itemize}
-				\item only correct forms are analysed and generated (Tatar exs.)
-			\end{itemize}
-
-			\scalebox{0.83}{
-				\begin{tabular}{lllll}
-					\toprule
-					\textbf{Type} & \textbf{Gloss} & {\small \texttt{\tag{adj}(\tag{comp})}} & {\small \texttt{\tag{adj}(\tag{comp})\tag{subst}}} & {\small \texttt{\tag{adj}(\tag{comp})\tag{advl}}} \\
-					\midrule
-					\textbf{A1} & \eng{good} & {\qipa яхшы (яхшырак)} & {\qipa яхшы (яхшырак)} & {\qipa яхшы (яхшырак)} \\
-					\textbf{A2} & \eng{old} & {\qipa иске (искерәк)} & {\qipa иске (искерәк)} & {\qipa — (—)} \\
-					\textbf{A3} & \eng{dead} & {\qipa үле (—)} & {\qipa үле (—)} & {\qipa — (—)} \\
-					\textbf{A4} & \eng{basic} & {\qipa төп (—)} & {\qipa — (—)} & {\qipa — (—)} \\
-					\bottomrule
-				\end{tabular}
-			}\\
-			\begin{itemize}
-				%\item Interestingly, native-speaker intuitions on the morphological limitations of a given adjective can often be more restrictive than the range of possible uses found in a large text corpus.
-				\item Native speaker intuitions on morph.\ limitations often more restrictive than range of possible uses found in large text corpora.
-			\end{itemize}
-	}
-
-		\headerbox{Examples:\ ideas for other examples}{name=otherexamples,column=0,span=1,below=tuvanvelars}{ %{name=otherexamples,column=1,span=2,below=tuvanvelars}{
-			\begin{itemize}
-				\item Kyrgyz: conditions where л desonorises to д
-				%\item Tuvan: рубль - рубльдиң *рубльдуң
-				\item clitics can go everywhere/anywhere
-				\item underformalised - “particles☹” can go anywhere
-				%\item all adjs act like nouns (Somfai-Kara, p. 28)
-				%\item all adjectives can be used as adverbs without any morphological changes (Somfai-Kara, p. 29)
-				\item Stuff found doing treebank
-			\end{itemize}
-
-		}
-
-%		\headerbox{Morphophonology}{name=morphophon,column=0,below=otherexamples}{
-%
-%			\htwo{Desonorisation (kaz \& kir)}
-%				\begin{itemize}
-%					\item {\texttt \{N\}} desonorises to д after a consonant\\
-%						алма-\hiliteone{\{N\}}\{I\} → алма\hiliteone{н}ы \eng{apple--\gmk{ACC}} \\
-%						сыр-\hiliteone{\{N\}}\{I\} → сыр\hiliteone{д}ы \eng{secret--\gmk{ACC}}
-%					\item {\texttt \{L\}} desonorises to д after cons.\ of sonority ≤ /l/ \\
-%						сыр-\hiliteone{\{L\}}\{A\}р → сыр\hiliteone{л}ар \eng{secret--\gmk{PL}} \\
-%						кыз-\hiliteone{\{L\}}\{A\}р → кыз\hiliteone{д}ар \eng{girl--\gmk{PL}} \\
-%				\end{itemize}
-%
-%				\vspace{-0.5ex}
-%				\codeex{
-%					\texttt{"L Desonorisation"\\
-%						\%\{L\%\}:д <=> :VoicedLowSonCns \%>:\ \blank{1em} ;}\\
-%			
-%					\texttt{"N Desonorisation"\\
-%						\%\{N\%\}:д <=> :VoicedCns \%>:\ \blank{1em} ;}
-%				}
-%			\vspace{0.1em}
-%
-%			\htwo{Epenthesis}
-%				\begin{itemize}
-%					\item Turn {\texttt \{y\}} into a harmonised high vowel when a vowel doesn't follow the following consonant:\\
-%					мур\archiphon{у}н → мурун \eng{nose}\\
-%					мур\archiphon{у}н+\archiphon{I}м → мурдум \eng{my nose}\\
-%				\end{itemize}
-%					\vspace{-0.5ex}
-%\codeex{
-%			\begin{spacing}{0.9}
-%{\small \texttt{\%\{y\%\}:Vy <=> [ :LastVowel :Cns* :Cns ]/[:0] \blank{1em}\\
-%	\vspace{-0.5ex}\hspace{1pt}\hfill{} [ :Cns [ .\#.\ | :Cns ] ]/[ :0 | \%>:]\ ;\\
-%  \vspace{-0.5ex}\hspace{1ex}where  Vy  in  (  и  ү  и  и  ү  ы  ы  у  у  ы  у  у )\\
-%  \vspace{-0.5ex}LastVowel  in  (  и  ү  е  э  ө  я  а  ё  о  ы  ю  у )\\
-%  \vspace{-0.5ex}\hspace{12ex}         matched ;}}\vspace{-1.5em}
-%  			\end{spacing}
-%}
-%\vspace{0.1em}
-%		}
-
 		\headerbox{References}{name=references,column=3,below=tuvanvelars}{
@@
 
-		\headerbox{Development cycle}{name=development,below=morphtrans}{
-			%(1) look at frequency list --> check word --> if it's missing stem, add --> (1) ; if it's bad phonology --> fix phonology --> (1)
-			%\scalebox{0.8}{
-			%\begin{tikzpicture}[node distance = 2cm, auto]
-			%	% Place nodes
-			%	\node [cloud, above of=fixphon] (assumed) at (5.5,1.5) {assumed phonology} ;
-			%	\node [block] (init) at (1,3) {analyse corpus} ;
-			%	\node [block, left of=init] (addstem) at (0,3) {add stem};
-			%	\node [block, right of=init] (fixphon) at (3.5,3) {fix phonology};
-			%	\node [block, below of=init] (freq) at (1,2) {examine frequency list};
-			%	\node [decision, right of=freq] (break) at (1.5,0) {did something break?};
-			%	%\node [block, below of=identify] (evaluate) {evaluate candidate models};
-			%	%\node [block, left of=evaluate, node distance=3cm] (update) {update model};
-			%	\node [decision, below right of=freq] (decide) at (0,-1) {cause for unanalysed forms?};
-			%	\node [cloud] (revised) at (6.75,-4.5) {revised phonology} ;
-			%	%\node [block, below of=decide, node distance=3cm] (stop) {stop};
-			%	% Draw edges
-			%	\path [line] (init) -- (freq);
-			%	\path [line] (freq) -- (break);
-			%	\path [line] (break.south) -| node [near start] {no} (decide.north) ;
-			%	\path [line] (break.north) -| node [near start] {yes}(fixphon.south) ;
-			%	\path [line] (decide) -| node [near start] {missing stem} (addstem);
-			%	\path [line] (decide.east) --(decide-|fixphon.east) node [below, midway] {bad phonology} |-(break.north-|fixphon.south) -- (fixphon.south);
-			%	\path [line] (addstem) -- (init);
-			%	\path [line] (fixphon) -- (init);
-			%	\path [line,dashed] (assumed) -- (fixphon);
-			%	\path [line,dashed] (fixphon.east) --(fixphon.east-|assumed.east) |-(decide.east-|revised.north) -- (revised.north);
-			%\scalebox{0.65}{
-			\begin{tikzpicture}[xscale=1.92,yscale=2.65]%, auto]
-
-				\node[cloud] (assumptions) at (1,2.75) {assumed generalisations};
-				\node[block] (compile) at (1,2) {compile analyser};
-				\node[block] (analyse) at (1,1.5) {analyse corpus};
-				\node[block] (check) at (1,1) {examine freq.\ list};
-				\node[decision] (missingstem) at (1,0) {missing stem?};
-				\node[block] (addstem) at (0.22,2) {add stem};
-				\node[decision] (phonormorph) at (2,0) {type of problem?};
-				\node[block] (updatephon) at (3,0) {update phonol.};
-				\node[block] (updatemorph) at (3,1) {update morphol.};
-				\node[decision] (phonbetter) at( 4,0) {better?};
-				\node[decision] (morphbetter) at (4,1) {better?};
-				\node[cloud] (newgeneralisation) at (2.5,2) {new\\grammatical\\generalisation};
-
-				\path[line] (assumptions) -- node [right=0.75ex, align=left, scale=\flowchartscale] {preliminary\\development} (compile);
-				\path[line] (compile) -- (analyse);
-				\path[line] (analyse) -- (check);
-				\path[line] (check) -- node [right=0.75ex, scale=\flowchartscale, align=left] {unanalysed\\forms} (missingstem);
-				\path[line] (missingstem) -| node [below=1ex, near start, scale=\flowchartscale] {yes} (addstem);
-				\path[line] (addstem) -- (compile);
-				\path[line] (missingstem) -- node [below=1ex, midway, scale=\flowchartscale] {no} (phonormorph);
-				\path[line] (phonormorph) -- node [below=1ex, near start, scale=\flowchartscale] {phon} (updatephon);
-				\path[line] (phonormorph) |- node [below=1ex, near end, scale=\flowchartscale] {morph} (updatemorph);
-				\path[line] (updatephon) -- node [above=0.5ex, scale=\flowchartscale, align=center] {diff \\ freq \\ list} (phonbetter) ;
-				\path[line] (updatemorph) -- node [above=0.5ex, scale=\flowchartscale, align=center] {diff \\ freq \\ list} (morphbetter) ;
-				\path[line] (phonbetter.east) -- node [below=1ex, near start, scale=\flowchartscale] {yes} (4.6,0) |- (newgeneralisation.east) ;
-				\path[line] (morphbetter.east) -- node [below=1ex, near start, scale=\flowchartscale] {yes} (4.6,1) |- (newgeneralisation.east) ;
-				\path[line] (phonbetter.north) |- node [right=1ex, near start, scale=\flowchartscale] {no} (4,0.5) -| (updatephon.north) ;
-				\path[line] (morphbetter.north) |- node [right=1ex, near start, scale=\flowchartscale] {no} (4,1.5) -| (updatemorph.north) ;
-				\path[line] (newgeneralisation) -- (compile);
-
-			\end{tikzpicture}%}
-
-		}
-
 	\end{poster}

### 2016-05-14 2016-lrec-tyvmorph/poster.tex
general structure
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@68416 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 		\setlength\fboxrule{0pt}
-		\includegraphics[height=5.5em]{qrcode}
+		\includegraphics[angle=90,height=6.2em]{apertium5b}%\hspace{-2em}
 		\includegraphics[height=5.8em]{uitlogo}\hspace{1em}
@@
 		\begin{center}
-			\begin{minipage}[t]{6em}
+			\begin{minipage}[t]{7em}
 				\begin{spacing}{0.4}
-				{Francis Tyers}\\
+				{Francis M.\ Tyers}\\
 				{\footnotesize UiT Norgga Árktalaš Universitehta \\\texttt{francis.tyers@uit.no}}
@@
 		\includegraphics[height=5.4em]{img/soranif}
-		\includegraphics[angle=90,height=6.2em]{apertium5b}%\hspace{-2em}
 		%\hspace{2.5cm}
@@
 	\headerbox{Tuvan}{name=tyvan,column=0,row=0}{
+		\includegraphics[width=\textwidth]{img/tuvanmap}
 	}
@@
 	}
-		\headerbox{References}{name=references,column=3,below=tuvanvelars}{
-			\vspace{-0.42em}
-			\begin{spacing}{0.3}
-				\printbibliography[heading=none]
-			\end{spacing}
-			\vspace{-0.05em}
+
+	\headerbox{Example}{name=example,span=2,column=1,row=0}{
+
 	}
 
-		\headerbox{Ongoing and future work}{name=nedostatki,column=2,below=tuvanvelars}{
+	\headerbox{Evaluation}{name=results,column=1,below=example}{
+
+	}
+
+		\headerbox{Ongoing and future work}{name=nedostatki,column=2,below=example}{
 		\begin{itemize}
@@
 
-		\headerbox{Further information}{name=getting,column=1,below=adjectives,span=2}{
-			%\vspace{-0.2em}
-			%\tikzoverlay[text width=0.8in] at (6.73in, 0.04in) {
-			%	\includegraphics[height=0.8in]{qrcode}
-			%};
+		\headerbox{Further information}{name=getting,column=1,below=results,span=2}{
+			\vspace{-0.2em}
+			\tikzoverlay[text width=0.8in] at (6.2in, 0.04in) {
+				\includegraphics[height=0.8in]{qrcode}
+			};
 			\begin{itemize}
@@
 
+	\headerbox{References}{name=references,below=nedostatki,column=2}{
+		\vspace{-0.42em}
+		\begin{spacing}{0.3}
+			\printbibliography[heading=none]
+		\end{spacing}
+		\vspace{-0.05em}
+	}
+
 	\end{poster}

### 2016-05-19 2016-lrec-tyvmorph/poster.tex
a questionable colour choice
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@68421 72bbbca6-d526-0410-a7d9-f06f51895060
some work on latex version of poster
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@68623 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \definecolor{iulimestone}{cmyk}{0.11,0.13,0.14,0.26}
-%\definecolor{boxcolor}{RGB}{176,208,223}
+\definecolor{tuvagold}{cmyk}{0.0,0.2,1.0,0.0}
+\definecolor{tuvablue}{cmyk}{0.57,0.23,0.0,0.1}
+\definecolor{boxcolor}{RGB}{176,208,223}
 %\definecolor{boxcolor}{cmyk}{0.3,0.08,0.08,0.1}
-\definecolor{boxcolor}{cmyk}{0.2,0.03,0.03,0.1}
+%\definecolor{boxcolor}{cmyk}{0.2,0.03,0.03,0.1}
+%\definecolor{boxcolor}{cmyk}{0.57,0.23,0.0,0.1}
 \definecolor{headerfontcol}{RGB}{255,255,255}
@@
 \newcommand{\hiliteblue}[1]{{\addfontfeature{Color=44697DFF}#1}}
+\newcommand{\hilitegold}[1]{{\addfontfeature{Color=tuvagold}#1}}
 
@@
 %\setromanfont[IPA]{Liberation Serif}
-\setromanfont[IPA]{Linux Libertine O}
+%\setromanfont[IPA]{Linux Libertine O}
+%\setromanfont[IPA]{Linux Biolinum O}
+\setromanfont[IPA,Scale=1.01]{PT Sans}
 %\setromanfont[IPA]{Nimbus Roman No9 L}
@@
 \usepackage[small,bf]{caption}
-\newfontfamily\qipa[IPA,Scale=MatchLowercase]{FreeSerif}
+%\newfontfamily\qipa[IPA,Scale=MatchLowercase]{FreeSerif}
+\newfontfamily\qipa[IPA,Scale=MatchLowercase]{FreeSans}
 \newfontfamily\qgmk[IPA,Scale=0.8]{CMU Serif}
@@
 \newfontfamily\ortamono[IPA,Scale=0.625]{DejaVu Sans Mono}
+\newfontfamily\ortaqmono[IPA,Scale=0.75]{DejaVu Sans Mono}
 \newfontfamily\diagsans[IPA,Scale=1]{DejaVu Sans}
@@
 \newcommand{\tags}[1]{\hiliteone{#1}}
-\newcommand{\tag}[1]{{\ortamono \hilitegrey{<}\hiliteblue{#1}\hilitegrey{>}}}
+\newcommand{\tag}[1]{{\ortaqmono \hilitegrey{<}\hiliteblue{#1}\hilitegrey{>}}}
 \newcommand{\archiphon}[1]{\hilitegrey{\{}\hiliteone{#1}\hilitegrey{\}}}
@@
 %\newcommand{\htwo}[1]{{\htwofont \textbf{:::#1:::}}}%\hrule} %\hline\\}
+%\newcommand{\htwo}[1]{{\htwofont \textbf{\hilitegold{\dotfill{}}#1\hilitegold{\dotfill{}}}}}
 \newcommand{\htwo}[1]{{\htwofont \textbf{\dotfill{}#1\dotfill{}}}}
+\newcommand{\dividerspace}{\vspace{0.5em}}
 
@@
 		%headerColorOne=blue,
-		headerFontColor=yellow,
+		headerFontColor=tuvagold,
 		% Only simple background color used, no shading, so boxColorTwo isn't necessary
@@
 		\setlength\fboxrule{0pt}
-		\includegraphics[angle=90,height=6.2em]{apertium5b}%\hspace{-2em}
+		%\includegraphics[angle=90,height=6.2em]{apertium5b}%\hspace{-2em}
 		\includegraphics[height=5.8em]{uitlogo}\hspace{1em}
@@
 		\includegraphics[width=\textwidth]{img/tuvanmap}
+		\tikzoverlay[text width=1in] at (2.5in, -0in) {
+			\includegraphics[width=1in]{img/Flag_of_Tuva}
+		};
+		\vspace{-1.2em}
+		\begin{itemize}
+			\item Sayan Turkic language
+			\item Spoken in the Tuva Republic
+			\item Approximately 300,000 speakers
+			\item Agglutinative morphology
+			\item Very little work on computational tools
+		\end{itemize}
 	}
@@
 		%\scalebox{0.98}{
-			алдым~~~\hilitetwo{↔}~~~\texttt{ал\tag{v}\tag{tv}\tag{ifi}\tag{p1}\tag{sg}},~~\texttt{алд\tag{n}\tag{px1sg}\tag{nom}}
+			алдым\hfill{}{\qipa \hilitetwo{↔}}\hfill{}\texttt{ал\tag{v}\tag{tv}\tag{ifi}\tag{p1}\tag{sg}},~~\texttt{алд\tag{n}\tag{px1sg}\tag{nom}}\\
+			өөнде\hfill{}{\qipa \hilitetwo{↔}}\hfill{}\texttt{өг\tag{n}\tag{px3sp}\tag{loc}}
 		%}
 		\end{itemize}
+		\dividerspace{}
 
-		\htwo{Turkic-language transducers we've made}
+		\htwo{Framework:\ HFST}
+		\begin{itemize}
+			\item Reimplements Xerox FST formalisms ({\tt lexc} \& {\tt twol})
+			\item Also provides a wrapper around popular free/open-source FST toolkits: SFST, OpenFST, and Foma
+		\end{itemize}
+		\dividerspace{}
+
+		\htwo{Approach}
+		\begin{itemize}
+			\item two-level method \parencite{koskenniemi83}
+			\item \textbf{morphotactics} implemented in \texttt{lexc}
+			\item \textbf{morphophonology} implemented in \texttt{twol} (SPE-style rules)
+			\item compiled separately; compose-intersected to single transducer\vspace{-0.1ex}\\
+			алдым\hfill{}{\qipa \hilitetwo{↔}}\hfill{}\texttt{ал\hilitegrey{>}\archiphon{D}\archiphon{I}\hilitegrey{>}м}\hfill{}{\qipa \hilitetwo{↔}}\hfill{}\texttt{ал\tag{v}\tag{tv}\tag{ifi}\tag{p1}\tag{sg}}\vspace{-0.1ex}\\
+			алдым\hfill{}{\qipa \hilitetwo{↔}}\hfill{}\texttt{алд\hilitegrey{>}\archiphon{I}м}\hfill{}{\qipa \hilitetwo{↔}}\hfill{}\texttt{алд\tag{n}\tag{px1sg}\tag{nom}}\\
+			өөнде\hfill{}{\qipa \hilitetwo{↔}}\hfill{}\texttt{өг\hilitegrey{>}\archiphon{z}\archiphon{I}\archiphon{n}\hilitegrey{>}\archiphon{D}\archiphon{A}}\hfill{}{\qipa \hilitetwo{↔}}\hfill{}\texttt{өг\tag{n}\tag{px3sp}\tag{loc}}
+		\end{itemize}
+
+	}
+
+		\headerbox{Development cycle}{name=development,below=morphtrans}{
+			\begin{tikzpicture}[xscale=1.92,yscale=2.65]%, auto]
+
+				\node[cloud] (assumptions) at (1,2.75) {assumed generalisations};
+				\node[block] (compile) at (1,2) {compile analyser};
+				\node[block] (analyse) at (1,1.5) {analyse corpus};
+				\node[block] (check) at (1,1) {examine freq.\ list};
+				\node[decision] (missingstem) at (1,0) {missing stem?};
+				\node[block] (addstem) at (0.22,2) {add stem};
+				\node[decision] (phonormorph) at (2,0) {type of problem?};
+				\node[block] (updatephon) at (3,0) {update phonol.};
+				\node[block] (updatemorph) at (3,1) {update morphol.};
+				\node[decision] (phonbetter) at( 4,0) {better?};
+				\node[decision] (morphbetter) at (4,1) {better?};
+				\node[cloud] (newgeneralisation) at (2.5,2) {new\\grammatical\\generalisation};
+
+				\path[line] (assumptions) -- node [right=0.75ex, align=left, scale=\flowchartscale] {preliminary\\development} (compile);
+				\path[line] (compile) -- (analyse);
+				\path[line] (analyse) -- (check);
+				\path[line] (check) -- node [right=0.75ex, scale=\flowchartscale, align=left] {unanalysed\\forms} (missingstem);
+				\path[line] (missingstem) -| node [below=1ex, near start, scale=\flowchartscale] {yes} (addstem);
+				\path[line] (addstem) -- (compile);
+				\path[line] (missingstem) -- node [below=1ex, midway, scale=\flowchartscale] {no} (phonormorph);
+				\path[line] (phonormorph) -- node [below=1ex, near start, scale=\flowchartscale] {phon} (updatephon);
+				\path[line] (phonormorph) |- node [below=1ex, near end, scale=\flowchartscale] {morph} (updatemorph);
+				\path[line] (updatephon) -- node [above=0.5ex, scale=\flowchartscale, align=center] {diff \\ freq \\ list} (phonbetter) ;
+				\path[line] (updatemorph) -- node [above=0.5ex, scale=\flowchartscale, align=center] {diff \\ freq \\ list} (morphbetter) ;
+				\path[line] (phonbetter.east) -- node [below=1ex, near start, scale=\flowchartscale] {yes} (4.6,0) |- (newgeneralisation.east) ;
+				\path[line] (morphbetter.east) -- node [below=1ex, near start, scale=\flowchartscale] {yes} (4.6,1) |- (newgeneralisation.east) ;
+				\path[line] (phonbetter.north) |- node [right=1ex, near start, scale=\flowchartscale] {no} (4,0.5) -| (updatephon.north) ;
+				\path[line] (morphbetter.north) |- node [right=1ex, near start, scale=\flowchartscale] {no} (4,1.5) -| (updatemorph.north) ;
+				\path[line] (newgeneralisation) -- (compile);
+
+			\end{tikzpicture}%}
+
+		}
+
+	\headerbox{Example}{name=example,span=2,column=1,row=0}{
+	« Бис кожууннуң соңгукчулары-биле ажылды чорутпастай бээривиске, көргүзүглер баксыраан. »\\
+
+		%\texttt{алд\hilitegrey{>}\archiphon{I}м}\hfill{}\hilitetwo{↔}\hfill{}\texttt{алд\tag{n}\tag{px1sg}\tag{nom}}
+
+		\texttt{\textasciicircum Бис/бис\tag{prn}\tag{pers}\tag{p1}\tag{pl}\tag{nom}\$}\\
+		\texttt{\textasciicircum кожууннуң/кожуун\tag{n}\tag{gen}\$}\\
+		\texttt{\textasciicircum соңгукчулары/соңгукчу\tag{n}\tag{pl}\tag{px3sp}\tag{nom}\$}\\
+		\texttt{\textasciicircum -биле/биле\tag{post}\$}\\
+		\texttt{\textasciicircum ажылды/ажыл\tag{n}\tag{acc}\$}\\
+		\texttt{\textasciicircum чорутпастай/чорут\tag{v}\tag{tv}\tag{cess}\tag{prc\_impf}\$}\\
+		\texttt{\textasciicircum бээривиске/бер\tag{vaux}\tag{ger\_aor}\tag{px1pl}\tag{dat}\$}\\
+		\texttt{\textasciicircum ,/,\tag{cm}\$}\\
+		\texttt{\textasciicircum көргүзүглер/көргүзүг\tag{n}\tag{pl}\tag{nom}\$}\\
+		\texttt{\textasciicircum баксыраан/баксыра\tag{v}\tag{iv}\tag{past}\tag{p3}\tag{pl}\$}\\
+		\texttt{\textasciicircum ./.\tag{sent}\$}
+
+	}
+
+	\headerbox{Evaluation}{name=results,column=1,below=example}{
+
+	}
+
+		\headerbox{Ongoing and future work}{name=nedostatki,column=2,below=example}{
+		\htwo{our other Turkic-language transducers}
 		\begin{itemize}
@@
 			\item Kumyk \parencite{jnwisfmt2014lrecpaper} %(Washington et al., 2014)
-			\item ongoing work on Sakha, Tuvan, Khakas, and more!
+			\item ongoing work on Sakha, Khakas, and more!
 		\end{itemize}
+		\dividerspace{}
 
-		\htwo{Framework:\ HFST}
+		%\begin{itemize}
+
+				%\item Current annotation of a dependency treebank for Kazakh:\\
+				%a similar process that helps identify errors in morphological analyses and automatic disambiguation (constraint grammar)
+				%%\item Leading to
+				%\begin{itemize}
+				%	\item Improved rule-based morphological disambiguation% (constraint grammar)
+				%	\item Improved morphological analyser %for Kazakh
+				%	\item A syntactic dependency treebank %for Kazakh
+				%\end{itemize}
+				%%\item Work on rule-based morphological disambiguation with constraint grammar
+			%\end{itemize}
+			\htwo{Future work}
 				\begin{itemize}
-			\item Reimplements Xerox FST formalisms ({\tt lexc} \& {\tt twol})
-			\item Also provides a wrapper around popular free/open-source FST toolkits: SFST, OpenFST, and Foma
-		\end{itemize}
-		\htwo{Approach}
-		\begin{itemize}
-			\item two-level method \parencite{koskenniemi83}
-			\item \textbf{morphotactics} implemented in \texttt{lexc}
-			\item \textbf{morphophonology} implemented in \texttt{twol} (SPE-style rules)
-			\item compiled separately; compose-intersected to single transducer\vspace{-0.1ex}\\
-			алдым\hfill{}\hilitetwo{↔}\hfill{}\texttt{ал\hilitegrey{>}\archiphon{D}\archiphon{I}\hilitegrey{>}м}\hfill{}\hilitetwo{↔}\hfill{}\texttt{ал\tag{v}\tag{tv}\tag{ifi}\tag{p1}\tag{sg}}\vspace{-0.1ex}\\
-			алдым\hfill{}\hilitetwo{↔}\hfill{}\texttt{алд\hilitegrey{>}\archiphon{I}м}\hfill{}\hilitetwo{↔}\hfill{}\texttt{алд\tag{n}\tag{px1sg}\tag{nom}}
-		\end{itemize}
-
-	}
-
-	\headerbox{Example}{name=example,span=2,column=1,row=0}{
-
-	}
-
-	\headerbox{Evaluation}{name=results,column=1,below=example}{
-
-	}
-
-		\headerbox{Ongoing and future work}{name=nedostatki,column=2,below=example}{
-		\begin{itemize}
-				\item Current annotation of a dependency treebank for Kazakh:\\
-				a similar process that helps identify errors in morphological analyses and automatic disambiguation (constraint grammar)
-				%\item Leading to
-				\begin{itemize}
-					\item Improved rule-based morphological disambiguation% (constraint grammar)
-					\item Improved morphological analyser %for Kazakh
-					\item A syntactic dependency treebank %for Kazakh
-				\end{itemize}
-				%\item Work on rule-based morphological disambiguation with constraint grammar
-				\item Future work:
-				\begin{itemize}
-					\item Annotate more gold standard text in more Turkic languages
-					\item Research cross-linguistic techniques for analysing Turkic
+					%\item Annotate more gold standard text in more Turkic languages
+					%\item Research cross-linguistic techniques for analysing Turkic
+					\item Dependency parsing (native and crosslingual)
+					\item Increase number of stems
+					\item Make more available to linguists
+					\item Integrate into end-user applications (spellchecking, MT)
 				\end{itemize}
@@
 				%\item Other Turkic lgs:\ Uzbek, Uyghur, Chuvash, Yakut, Tuvan, etc.
-			\end{itemize}
 		}
 
-		\headerbox{Further information}{name=getting,column=1,below=results,span=2}{
-			\vspace{-0.2em}
-			\tikzoverlay[text width=0.8in] at (6.2in, 0.04in) {
-				\includegraphics[height=0.8in]{qrcode}
+		\headerbox{Further information}{name=getting,column=0,below=development,span=3}{
+			\tikzoverlay[text height=1in] at (0in, 0.04in) {
+				\includegraphics[angle=90,height=1in]{apertium5b}
 			};
+			%\tikzoverlay[text width=0.8in] at (6.2in, 0.04in) {
+			\tikzoverlay[text height=1in] at (10in, 0.04in) {
+				\includegraphics[height=1in]{qrcode}
+			};
+			\tikzoverlay[text height=0.5in] at (8in, 0.04in) {
+				{\huge Thanks!}
+			};
+			\tikzoverlay[text width=6in] at (1in, 0.04in) {
 			\begin{itemize}
 				\item Part of \textbf{Apertium Turkic} project:\ \url{http://wiki.apertium.org/wiki/Apertium\_Turkic}
-				\item Transducers available \textbf{live} (\href{http://turkic.apertium.org/}{\texttt{turkic.apertium.org}}), \textbf{source code} under \hiliteone{GPL} from Apertium's SVN repo
+				\item Transducers available \textbf{live} on our website: {\Large \url{http://turkic.apertium.org/}} %\href{http://turkic.apertium.org/}{\texttt{turkic.apertium.org}}
+				\item \textbf{Source code} under \hiliteone{GPL} from Apertium's SVN repo
 				%\item Transducers available \textbf{live} at \href{http://turkic.apertium.org/}{\texttt{turkic.apertium.org}}
@@
 			\vspace{0.3em}
+			};
+			\vspace{1in}
 		}

### 2016-05-19 2016-lrec-tyvmorph/poster.tex
some more work on the latex poster
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@68624 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 % Define block styles
-\newcommand{\flowchartscale}{0.65}
+\newcommand{\flowchartscale}{0.85}
 \tikzstyle{decision} = [diamond, draw, fill=iumint, 
@@
 
-		\headerbox{Development cycle}{name=development,below=morphtrans}{
-			\begin{tikzpicture}[xscale=1.92,yscale=2.65]%, auto]
+		\headerbox{Development cycle}{name=development,below=morphtrans,span=2}{
+		\centering
+		\begin{tikzpicture}[xscale=3,yscale=2.8]%, auto]
 
@@
 			\node[block] (compile) at (1,2) {compile analyser};
-				\node[block] (analyse) at (1,1.5) {analyse corpus};
-				\node[block] (check) at (1,1) {examine freq.\ list};
+			\node[block] (analyse) at (1,1.4) {analyse corpus};
+			\node[block] (check) at (1,0.7) {examine freq.\ list};
 			\node[decision] (missingstem) at (1,0) {missing stem?};
-				\node[block] (addstem) at (0.22,2) {add stem};
-				\node[decision] (phonormorph) at (2,0) {type of problem?};
-				\node[block] (updatephon) at (3,0) {update phonol.};
-				\node[block] (updatemorph) at (3,1) {update morphol.};
-				\node[decision] (phonbetter) at( 4,0) {better?};
-				\node[decision] (morphbetter) at (4,1) {better?};
-				\node[cloud] (newgeneralisation) at (2.5,2) {new\\grammatical\\generalisation};
+			\node[block] (addstem) at (0,2) {add stem};
+			\node[decision] (phonormorph) at (2.25,0) {type of problem?};
+			\node[block] (updatephon) at (3.5,0) {update phonol.};
+			\node[block] (updatemorph) at (3.5,1) {update morphol.};
+			\node[decision] (phonbetter) at(5,0) {better?};
+			\node[decision] (morphbetter) at (5,1) {better?};
+			\node[cloud] (newgeneralisation) at (3,2) {new\\grammatical\\generalisation};
 
-				\path[line] (assumptions) -- node [right=0.75ex, align=left, scale=\flowchartscale] {preliminary\\development} (compile);
+			\path[line] (assumptions) -- node [right=0.75ex, align=left, scale=\flowchartscale] {initial\\development} (compile);
 			\path[line] (compile) -- (analyse);
-				\path[line] (analyse) -- (check);
-				\path[line] (check) -- node [right=0.75ex, scale=\flowchartscale, align=left] {unanalysed\\forms} (missingstem);
+			\path[line] (analyse) -- node [right=0.75ex, scale=\flowchartscale, align=left] {unanalysed\\forms} (check);
+			\path[line] (check) -- (missingstem);
 			\path[line] (missingstem) -| node [below=1ex, near start, scale=\flowchartscale] {yes} (addstem);
@@
 			\path[line] (phonormorph) |- node [below=1ex, near end, scale=\flowchartscale] {morph} (updatemorph);
-				\path[line] (updatephon) -- node [above=0.5ex, scale=\flowchartscale, align=center] {diff \\ freq \\ list} (phonbetter) ;
-				\path[line] (updatemorph) -- node [above=0.5ex, scale=\flowchartscale, align=center] {diff \\ freq \\ list} (morphbetter) ;
-				\path[line] (phonbetter.east) -- node [below=1ex, near start, scale=\flowchartscale] {yes} (4.6,0) |- (newgeneralisation.east) ;
-				\path[line] (morphbetter.east) -- node [below=1ex, near start, scale=\flowchartscale] {yes} (4.6,1) |- (newgeneralisation.east) ;
-				\path[line] (phonbetter.north) |- node [right=1ex, near start, scale=\flowchartscale] {no} (4,0.5) -| (updatephon.north) ;
-				\path[line] (morphbetter.north) |- node [right=1ex, near start, scale=\flowchartscale] {no} (4,1.5) -| (updatemorph.north) ;
+			\path[line] (updatephon) -- node [above=0.5ex, scale=\flowchartscale, align=center] {diff \\ freq list} (phonbetter) ;
+			\path[line] (updatemorph) -- node [above=0.5ex, scale=\flowchartscale, align=center] {diff \\ freq list} (morphbetter) ;
+			\path[line] (phonbetter.east) -- node [below=1ex, near start, scale=\flowchartscale] {yes} (5.6,0) |- (newgeneralisation.east) ;
+			\path[line] (morphbetter.east) -- node [below=1ex, near start, scale=\flowchartscale] {yes} (5.6,1) |- (newgeneralisation.east) ;
+			\path[line] (phonbetter.north) |- node [right=1ex, near start, scale=\flowchartscale] {no} (5,0.5) -| (updatephon.north) ;
+			\path[line] (morphbetter.north) |- node [right=1ex, near start, scale=\flowchartscale] {no} (5,1.5) -| (updatemorph.north) ;
 			\path[line] (newgeneralisation) -- (compile);
 
-			\end{tikzpicture}%}
+		\end{tikzpicture}
+%			\begin{tikzpicture}[xscale=1.92,yscale=2.65]%, auto]	
+%
+%				\node[cloud] (assumptions) at (1,2.75) {assumed generalisations};
+%				\node[block] (compile) at (1,2) {compile analyser};
+%				\node[block] (analyse) at (1,1.5) {analyse corpus};
+%				\node[block] (check) at (1,1) {examine freq.\ list};
+%				\node[decision] (missingstem) at (1,0) {missing stem?};
+%				\node[block] (addstem) at (0.22,2) {add stem};
+%				\node[decision] (phonormorph) at (2,0) {type of problem?};
+%				\node[block] (updatephon) at (3,0) {update phonol.};
+%				\node[block] (updatemorph) at (3,1) {update morphol.};
+%				\node[decision] (phonbetter) at( 4,0) {better?};
+%				\node[decision] (morphbetter) at (4,1) {better?};
+%				\node[cloud] (newgeneralisation) at (2.5,2) {new\\grammatical\\generalisation};
+%
+%				\path[line] (assumptions) -- node [right=0.75ex, align=left, scale=\flowchartscale] {preliminary\\development} (compile);
+%				\path[line] (compile) -- (analyse);
+%				\path[line] (analyse) -- (check);
+%				\path[line] (check) -- node [right=0.75ex, scale=\flowchartscale, align=left] {unanalysed\\forms} (missingstem);
+%				\path[line] (missingstem) -| node [below=1ex, near start, scale=\flowchartscale] {yes} (addstem);
+%				\path[line] (addstem) -- (compile);
+%				\path[line] (missingstem) -- node [below=1ex, midway, scale=\flowchartscale] {no} (phonormorph);
+%				\path[line] (phonormorph) -- node [below=1ex, near start, scale=\flowchartscale] {phon} (updatephon);
+%				\path[line] (phonormorph) |- node [below=1ex, near end, scale=\flowchartscale] {morph} (updatemorph);
+%				\path[line] (updatephon) -- node [above=0.5ex, scale=\flowchartscale, align=center] {diff \\ freq \\ list} (phonbetter) ;
+%				\path[line] (updatemorph) -- node [above=0.5ex, scale=\flowchartscale, align=center] {diff \\ freq \\ list} (morphbetter) ;
+%				\path[line] (phonbetter.east) -- node [below=1ex, near start, scale=\flowchartscale] {yes} (4.6,0) |- (newgeneralisation.east) ;
+%				\path[line] (morphbetter.east) -- node [below=1ex, near start, scale=\flowchartscale] {yes} (4.6,1) |- (newgeneralisation.east) ;
+%				\path[line] (phonbetter.north) |- node [right=1ex, near start, scale=\flowchartscale] {no} (4,0.5) -| (updatephon.north) ;
+%				\path[line] (morphbetter.north) |- node [right=1ex, near start, scale=\flowchartscale] {no} (4,1.5) -| (updatemorph.north) ;
+%				\path[line] (newgeneralisation) -- (compile);
+%			\end{tikzpicture}%}
 
@@
 
-	\headerbox{Evaluation}{name=results,column=1,below=example}{
+	\headerbox{Lexicon}{name=lexicon,column=1,below=example}{
+		\begin{tabular}{llr}
+			\toprule
+			Part of speech	&	Tag	& Stems \\
+			\midrule
+			Noun & \tag{n} & 4,226 \\
+			Proper noun & \tag{np} & 4,217 \\
+			Adjective & \tag{adj} & 1,603 \\
+			Verb & \tag{v} & 1,064 \\
+			Adverb & \tag{adv} & 136 \\
+			Numeral & \tag{num} & 85 \\
+			Conjunction & \tag{cnj*} & 70 \\
+			Postposition & \tag{post} & 28 \\
+			Pronoun & \tag{prn} & 35 \\
+			Determiner & \tag{det} & 26 \\
+			\midrule
+			Total &  & 11,490 \\
+			\bottomrule
+		\end{tabular}
+	}
+
+	\headerbox{Evaluation}{name=results,column=1,below=lexicon}{
 
@@
 			%\item our Kyrgyz, Kazakh, Tatar, Kumyk:\ all \hiliteone{GPL (=free and open)}!
-			\item Kyrgyz \parencite{jnwmifmt2012lrecpaper}%  (Washington et al., 2012)
-			\item Tatar \& Bashkort \parencite{fmtjnwisrb2012lrecpaper} %(Tyers et al., 2012)
-			\item Kazakh \parencite{isjnwfmt2013mtsummitpaper} %(Salimzyanov et al., 2013)
-			\item Kumyk \parencite{jnwisfmt2014lrecpaper} %(Washington et al., 2014)
+			\item \textbf{Kyrgyz} \parencite{jnwmifmt2012lrecpaper}%  (Washington et al., 2012)
+			\item \textbf{Tatar} \& \textbf{Bashkort} \parencite{fmtjnwisrb2012lrecpaper} %(Tyers et al., 2012)
+			\item \textbf{Kazakh} \parencite{isjnwfmt2013mtsummitpaper} %(Salimzyanov et al., 2013)
+			\item \textbf{Kumyk} \parencite{jnwisfmt2014lrecpaper} %(Washington et al., 2014)
 			\item ongoing work on Sakha, Khakas, and more!
 		\end{itemize}
+		\dividerspace{}
+
+		\htwo{Morphological disambiguation}
+		\begin{itemize}
+			\item \textbf{Kazakh} \parencite{zajnwftanasakbada2016ciclingproceedings}
+		\end{itemize}
+		\dividerspace{}
+
+		\htwo{Treebanks}
+		\vspace{-1em}
+		\newline
+			\tikzoverlay[text width=0.9\textwidth] at (0.55in, -0.5in) {
+				\includegraphics[width=0.95\textwidth]{img/dep-graph}
+			};
+		\begin{itemize}
+			\item \textbf{Kazakh} \parencite{fmtjnw2015kazdep}
+		\end{itemize}
+		\includegraphics[width=0.63\textwidth]{img/korp}
+		\begin{itemize}
+			\item work on \textbf{Tuvan} ongoing
+		\end{itemize}
+		\vspace{1.45in}
 		\dividerspace{}

### 2016-05-19 2016-lrec-tyvmorph/poster.tex
moar poster dev
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@68625 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 % Define block styles
-\newcommand{\flowchartscale}{0.85}
+\newcommand{\flowchartscale}{0.6}
 \tikzstyle{decision} = [diamond, draw, fill=iumint, 
@@
 
-		\headerbox{Development cycle}{name=development,below=morphtrans,span=2}{
+		\headerbox{Development cycle}{name=development,below=morphtrans,span=1}{
 		\centering
-		\begin{tikzpicture}[xscale=3,yscale=2.8]%, auto]
+		\begin{tikzpicture}[xscale=1.45,yscale=2.2]%, auto]
 
@@
 	\headerbox{Example}{name=example,span=2,column=1,row=0}{
-	« Бис кожууннуң соңгукчулары-биле ажылды чорутпастай бээривиске, көргүзүглер баксыраан. »\\
+	{\centering
+	{\fontsize{12pt}{12pt}\selectfont \em{« Бис кожууннуң соңгукчулары-биле ажылды чорутпастай бээривиске, көргүзүглер баксыраан. »}}\vspace{0.5em}\\
+	{\fontsize{12pt}{12pt}\selectfont ``Figures worsened when we stopped conducting business with the district's constituents.''}\\
+	%Когда мы перестали проводить работу с избирателями кожууна, показатели ухудшились
+	}
+	%\vspace{-1em}
 
 		%\texttt{алд\hilitegrey{>}\archiphon{I}м}\hfill{}\hilitetwo{↔}\hfill{}\texttt{алд\tag{n}\tag{px1sg}\tag{nom}}
-
+	\begin{multicols}{2}
+	{\fontsize{14pt}{14pt}\selectfont
 		\texttt{\textasciicircum Бис/бис\tag{prn}\tag{pers}\tag{p1}\tag{pl}\tag{nom}\$}\\
@@
 		\texttt{\textasciicircum баксыраан/баксыра\tag{v}\tag{iv}\tag{past}\tag{p3}\tag{pl}\$}\\
-		\texttt{\textasciicircum ./.\tag{sent}\$}
+		\texttt{\textasciicircum ./.\tag{sent}\$}}\\
+		\tikzoverlay[text width=10em] at (3.3in, 0in) {
+			Lexicon contents {\qipa ↗}
+		};
+		\columnbreak
 
-	}
-
-	\headerbox{Lexicon}{name=lexicon,column=1,below=example}{
-		\begin{tabular}{llr}
+		\hspace*{1in}\begin{tabular}{llr}
 			\toprule
@@
 			\bottomrule
-		\end{tabular}
-	}
+		\end{tabular}\\
 
-	\headerbox{Evaluation}{name=results,column=1,below=lexicon}{
+	\end{multicols}
 
@@
 
-		\headerbox{Ongoing and future work}{name=nedostatki,column=2,below=example}{
+	%\headerbox{Lexicon}{name=lexicon,column=1,below=example}{
+	%	%\resizebox{0.9\textwidth}{!}{
+	%	%}
+	%}
+
+	\headerbox{Evaluation}{name=results,column=2,below=example}{ %lexicon}{
+		\begin{multicols}{2}	
+		\begin{tabular}{lrr}
+			\toprule
+				& precision & recall \\
+			\midrule
+			known & 0.99 & 0.97 \\
+			all & 0.99 & 0.69 \\
+			\bottomrule
+		\end{tabular}\\
+		{\qipa ↑} quantitative eval\\
+		\newline
+		\vspace{0.01pt}\hfill{}qualitative eval {\qipa ↗}\\
+		\columnbreak
+		\begin{tabular}{lr}
+			\toprule
+			error type & count \\
+			\midrule
+			Missing stem & 364 \\
+			Other & 65 \\
+			Morphotactics & 19 \\
+			Phonology & 8 \\
+			Categorisation & 6 \\
+			\bottomrule
+		\end{tabular}
+		\end{multicols}
+		%\vspace{0.05em}
+
+	}
+
+		\headerbox{Ongoing and future work}{name=nedostatki,column=1,below=example}{
 		\htwo{our other Turkic-language transducers}
@@
 
-	\headerbox{References}{name=references,below=nedostatki,column=2}{
+	\headerbox{References}{name=references,below=results,column=2}{
 		\vspace{-0.42em}

### 2016-05-19 2016-lrec-tyvmorph/poster.tex
more hargling
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@68626 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \renewcommand*{\nameyeardelim}{\addcomma\space}
-\renewcommand*{\bibfont}{\footnotesize}
+\renewcommand*{\bibfont}{\scriptsize}
 \renewcommand*{\intitlepunct}{\space}
@@
 
-		\headerbox{Development cycle}{name=development,below=morphtrans,span=1}{
+	\headerbox{What should we put here}{name=harglebargle,column=0,below=morphtrans}{
+		\vspace{0.5in}
+
+	}
+
+		\headerbox{Development cycle}{name=development,below=harglebargle,span=1}{
 		\centering
-		\begin{tikzpicture}[xscale=1.45,yscale=2.2]%, auto]
+		\begin{tikzpicture}[xscale=1.45,yscale=2.6]%, auto]
 
@@
 	\headerbox{Evaluation}{name=results,column=2,below=example}{ %lexicon}{
-		\begin{multicols}{2}	
+		%\begin{multicols}{2}	
+		\begin{itemize}
+			\item Naïve coverage calculated on 5-part corpus
+		\end{itemize}
+		\begin{center}
+		\resizebox{0.9\textwidth}{!}{
 		\begin{tabular}{lrr}
 			\toprule
-				& precision & recall \\
+			 \textbf{Domain} & \textbf{Tokens} & \textbf{Coverage} (\%) \\
 			\midrule
-			known & 0.99 & 0.97 \\
-			all & 0.99 & 0.69 \\
+			 News       & 1,539,459 &  95.73  \\
+			 Religion   & 746,124 &  93.84 \\
+			 Literature & 297,830 &  91.96  \\
+			 Encyclopaedic  & 276,547 & 90.86 \\
+			 Folklore & 27,902 & 91.57  \\\cline{2-3}
+			 Average    & -- & 92.79 \\
 			\bottomrule
-		\end{tabular}\\
-		{\qipa ↑} quantitative eval\\
-		\newline
-		\vspace{0.01pt}\hfill{}qualitative eval {\qipa ↗}\\
-		\columnbreak
-		\begin{tabular}{lr}
+		\end{tabular}}
+		\end{center}
+
+		\begin{itemize}
+			\item Good results reflect high accuracy
+		\end{itemize}
+		\begin{center}
+		\resizebox{0.9\textwidth}{!}{
+		\begin{tabular}{llrr}
 			\toprule
-			error type & count \\
+				& \textbf{count} & \textbf{precision} & \textbf{recall} \\
 			\midrule
-			Missing stem & 364 \\
-			Other & 65 \\
-			Morphotactics & 19 \\
-			Phonology & 8 \\
-			Categorisation & 6 \\
+			known tokens & 1024 & 0.99 & 0.97 \\
+			all tokens & 1425 & 0.99 & 0.69 \\
 			\bottomrule
-		\end{tabular}
-		\end{multicols}
+		\end{tabular}}
+		\end{center}
+		%{\qipa ↑} quantitative eval\\
+		%\newline
+		%\vspace{0.01pt}\hfill{}qualitative eval {\qipa ↗}\\
+		%\columnbreak
+		\begin{itemize}
+			\item Qualitative evaluation
+		\end{itemize}
+
+		\begin{center}
+		\resizebox{0.9\textwidth}{!}{
+		\begin{tabular}{lrr}
+			\toprule
+			\textbf{error type} & \textbf{count} & \textbf{percentage} \\
+			\midrule
+			Missing stem &  364 & 78.8 \\
+			Other & 65 & 14.1  \\
+			Bad morphotactics & 19 & 4.1  \\
+			Bad phonology & 8 & 1.7  \\
+			Incorrect categorisation & 6 & 1.3  \\
+			\midrule
+			Total: & 462 & 100 \\
+			\bottomrule
+		\end{tabular}}
+		\end{center}
+
+	%	\begin{tabular}{lr}
+	%		\toprule
+	%		error type & count \\
+	%		\midrule
+	%		Missing stem & 364 \\
+	%		Other & 65 \\
+	%		Bad morphotactics & 19 \\
+	%		Bad phonology & 8 \\
+	%		Incorrect categorisation & 6 \\
+	%		\bottomrule
+	%	\end{tabular}
+		%\end{multicols}
 		%\vspace{0.05em}
@@
 
-		\headerbox{Ongoing and future work}{name=nedostatki,column=1,below=example}{
+		\headerbox{Morphotactics}{name=morphotactics,column=1,below=example}{
+			\centering
+			\includegraphics[width=0.9\textwidth]{img/dibuix}
+
+		}
+
+		\headerbox{Ongoing and future work}{name=nedostatki,column=1,below=morphotactics}{
 		\htwo{our other Turkic-language transducers}

### 2016-05-19 2016-lrec-tyvmorph/poster.tex
more hargling
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@68627 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 		\begin{center}
-			\begin{minipage}[t]{7em}
+			\begin{minipage}[t]{8.5em} %{7em}
 				\begin{spacing}{0.4}
@@
 			\end{minipage}
-			\begin{minipage}[t]{8em} %{7.7em}
+			\begin{minipage}[t]{9em} %{7.7em}
 				\begin{spacing}{0.4}
 					{Aziyana Bayyr-ool}\\
-					{\footnotesize Сибирское отделение Российской академии наук\\\texttt{azikoa@mail.ru}}
+					{\footnotesize Сибирское отделение Российской академии наук,~~\texttt{azikoa@mail.ru}}
 				\end{spacing}
 			\end{minipage}
-			\begin{minipage}[t]{6em}
+			\begin{minipage}[t]{8em} %{6em}
 				\begin{spacing}{0.4}
-					{Aelita Salchak}\\
+					{Aelita Salchak\vphantom{y}}\\
 					{\footnotesize Тываның Күрүне Университеди\\\texttt{aelita\_74@mail.ru}}
@@
 	\headerbox{What should we put here}{name=harglebargle,column=0,below=morphtrans}{
-		\vspace{0.5in}
+		\vspace{0.35in}
 
@@
 		\centering
-		\begin{tikzpicture}[xscale=1.45,yscale=2.6]%, auto]
+		\begin{tikzpicture}[xscale=1.45,yscale=2.5]%, auto]
 
@@
 		%\begin{multicols}{2}	
-		\begin{itemize}
-			\item Naïve coverage calculated on 5-part corpus
-		\end{itemize}
+		\htwo{5-part corpus for naïve coverage}
+		%\begin{itemize}
+		%	\item Naïve coverage calculated on 5-part corpus
+		%\end{itemize}
 		\begin{center}
@@
 
-		\begin{itemize}
-			\item Good results reflect high accuracy
-		\end{itemize}
+		\htwo{Naïve coverage results}
+		%\begin{itemize}
+		%	\item Good results reflect high accuracy
+		%\end{itemize}
 		\begin{center}
@@
 		%\columnbreak
-		\begin{itemize}
-			\item Qualitative evaluation
-		\end{itemize}
+
+		\htwo{Qualitative evaluation}
+		%\begin{itemize}
+		%	\item Qualitative evaluation
+		%\end{itemize}
 
@@
 		\headerbox{Morphotactics}{name=morphotactics,column=1,below=example}{
+			\begin{itemize}
+				\item Introduction of quasi-derivational verbal morphotactics
+			\end{itemize}
 			\centering

### 2016-05-19 2016-lrec-tyvmorph/poster.tex
pdf of latex poster
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@68629 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 			};
-			\tikzoverlay[text height=0.5in] at (8in, 0.04in) {
-				{\huge Thanks!}
+			\tikzoverlay[text height=0.5in] at (6.75in, 0.04in) {
+				\parbox{3in}{
+				\centering
+				{\huge \textbf{Четтирдивис!}}\\
+				\texttt{четтир\tag{v}\tag{tv}\tag{ifi}\tag{p1}\tag{pl}}\vspace{1em}\\
+				{\Large \url{http://turkic.apertium.org/}}
+				}
 			};
@@
 				\item Part of \textbf{Apertium Turkic} project:\ \url{http://wiki.apertium.org/wiki/Apertium\_Turkic}
-				\item Transducers available \textbf{live} on our website: {\Large \url{http://turkic.apertium.org/}} %\href{http://turkic.apertium.org/}{\texttt{turkic.apertium.org}}
+				\item Transducers available \textbf{live} on our website: \url{http://turkic.apertium.org/} %\href{http://turkic.apertium.org/}{\texttt{turkic.apertium.org}}
 				\item \textbf{Source code} under \hiliteone{GPL} from Apertium's SVN repo

### 2016-05-19 2016-lrec-tyvmorph/poster.tex
some fixes and stuff
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@68630 72bbbca6-d526-0410-a7d9-f06f51895060
some more hargling, but need more
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@68634 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \renewcommand*{\nameyeardelim}{\addcomma\space}
-\renewcommand*{\bibfont}{\scriptsize}
+\renewcommand*{\bibfont}{\tiny}
 \renewcommand*{\intitlepunct}{\space}
@@
 	%}
+		\headerbox{Morphotactics}{name=morphotactics,column=2,below=example}{
+			\begin{itemize}
+				\item Introduction of quasi-derivational verbal morphotactics
+				\item Combinatorics not previously documented
+			\end{itemize}
+			\centering
+			\includegraphics[width=0.9\textwidth]{img/dibuix}
 
-	\headerbox{Evaluation}{name=results,column=2,below=example}{ %lexicon}{
+		}
+	\headerbox{Evaluation}{name=results,column=2,below=morphotactics}{ %lexicon}{
 		%\begin{multicols}{2}	
@@
 
-		\htwo{Naïve coverage results}
+		\htwo{Precision and recall}
 		%\begin{itemize}
@@
 
-		\headerbox{Morphotactics}{name=morphotactics,column=1,below=example}{
-			\begin{itemize}
-				\item Introduction of quasi-derivational verbal morphotactics
-			\end{itemize}
-			\centering
-			\includegraphics[width=0.9\textwidth]{img/dibuix}
-
-		}
-
-		\headerbox{Ongoing and future work}{name=nedostatki,column=1,below=morphotactics}{
+		\headerbox{Ongoing and future work}{name=nedostatki,column=1,below=example}{
 		\htwo{our other Turkic-language transducers}
@@
 		\end{itemize}
-		\includegraphics[width=0.63\textwidth]{img/korp}
+		\includegraphics[width=0.63\textwidth]{img/korp-2}
 		\begin{itemize}
@@
 				%\item Other Turkic lgs:\ Uzbek, Uyghur, Chuvash, Yakut, Tuvan, etc.
+				\includegraphics[width=\textwidth]{img/tyv-spelling}
 		}
 
-		\headerbox{Further information}{name=getting,column=0,below=development,span=3}{
+	\headerbox{References}{name=references,below=results,column=2}{
+		\vspace{-0.42em}
+		\begin{spacing}{0.3}
+			\printbibliography[heading=none]
+		\end{spacing}
+		\vspace{-0.05em}
+	}
+
+		\headerbox{Further information}{name=getting,column=0,below=references,span=3}{
 			\tikzoverlay[text height=1in] at (0in, 0.04in) {
@@
 			};
-			\tikzoverlay[text height=0.5in] at (6.75in, 0.04in) {
+			\tikzoverlay[text height=0.5in] at (6.75in, 0in) {
 				\parbox{3in}{
@@
 
-	\headerbox{References}{name=references,below=results,column=2}{
-		\vspace{-0.42em}
-		\begin{spacing}{0.3}
-			\printbibliography[heading=none]
-		\end{spacing}
-		\vspace{-0.05em}
-	}
-
 	\end{poster}

### 2016-05-19 2016-lrec-tyvmorph/poster.tex
moar hargling
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@68635 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 	\headerbox{What should we put here}{name=harglebargle,column=0,below=morphtrans}{
-		\vspace{0.35in}
-
+		\begin{tabular}{lllll}
+		\toprule
+			stem & V & C & dative & genitive \\
+		\midrule
+			медаль & а & ль & медальга & медальдың \\
+			ансамбль & а & бль & ансамбльге & ансамбльдиң \\
+			руль & у & ль & рульга & рульдуң \\
+			рубль & у & бль & рубльге & рубльдиң \\
+		\bottomrule
+		\end{tabular}
+\codeex{\hilitegrey{"\{I\} harmony"\\
+	\%\{I\%\}:Vy <=> :Vx [ :Cns* :RealCns ]/[ :0 | \%- ]* \_ ;}\\
+	\-~~~~~~~~~except\\
+	\-~~~~~~~~~~~~~[ :BackVow :Cns* :Cns :л ь: :Cns* :RealCns ]/:0* \_ ;\\
+	\-~~~~~~~~~~~~~[ :BackVow :Cns* :Cns :л ь:0 ]/:[ :0 - ь: ]* \_ ;\\
+	\-~~~~~~~~~\hilitegrey{where Vx in ( ү и е э ө а о ы у я ё ю )\\
+	\-~~~~~~~~~~~~~~ Vy in ( ү и и и ү ы у ы у ы у у )\\
+	\-~~~~~~~~~matched ;}\\
+	\\
+	"\{I\} always front when intervening Cль"\\
+	\%\{I\%\}:и <=> [ :BackVow :Cns* :Cns :л ь: :Cns* :RealCns ]/:0* \_ ;\\
+	\-~~~~~~~~~~~~[ :BackVow :Cns* :Cns :л ь:0 ]/:[ :0 - ь: ]* \_ ;}
 	}
@@
 		\centering
-		\begin{tikzpicture}[xscale=1.45,yscale=2.5]%, auto]
+		\begin{tikzpicture}[xscale=1.45,yscale=2.2]%, auto]
 
@@
 
-		\headerbox{Further information}{name=getting,column=0,below=references,span=3}{
+		\headerbox{Further information}{name=getting,column=0,below=development,span=3}{
 			\tikzoverlay[text height=1in] at (0in, 0.04in) {

### 2016-05-19 2016-lrec-tyvmorph/poster.tex
some more hargling, getting there....
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@68636 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 \renewcommand*{\nameyeardelim}{\addcomma\space}
-\renewcommand*{\bibfont}{\tiny}
+\renewcommand*{\bibfont}{\fontsize{5.75pt}{5.75pt}\selectfont}
 \renewcommand*{\intitlepunct}{\space}
@@
 
-	\headerbox{What should we put here}{name=harglebargle,column=0,below=morphtrans}{
+	\headerbox{Finding out new linguistics things}{name=harglebargle,column=0,below=morphtrans}{
+		\begin{center}
 		\begin{tabular}{lllll}
@@
 		\midrule
-			медаль & а & ль & медальга & медальдың \\
-			ансамбль & а & бль & ансамбльге & ансамбльдиң \\
-			руль & у & ль & рульга & рульдуң \\
-			рубль & у & бль & рубльге & рубльдиң \\
+			медаль & а & ль & медальг\hiliteone{а} & медальд\hiliteone{ы}ң \\
+			ансамбль & а & бль & ансамбльг\hilitetwo{е} & ансамбльд\hilitetwo{и}ң \\\midrule
+			руль & у & ль & рульг\hiliteone{а} & рульд\hiliteone{у}ң \\
+			рубль & у & бль & рубльг\hilitetwo{е} & рубльд\hilitetwo{и}ң \\
 		\bottomrule
 		\end{tabular}
+		\end{center}\vspace{-0.5em}
 \codeex{\hilitegrey{"\{I\} harmony"\\
@@
 		\htwo{5-part corpus for naïve coverage}
+		\vspace{-0.2em}
 		%\begin{itemize}
@@
 		\htwo{Precision and recall}
+		\vspace{-0.2em}
 		%\begin{itemize}
@@
 		\htwo{Qualitative evaluation}
+		\vspace{-0.2em}
 		%\begin{itemize}
@@
 		\newline
-			\tikzoverlay[text width=0.9\textwidth] at (0.55in, -0.5in) {
+			\tikzoverlay[text width=0.9\textwidth] at (0.55in, -1in) {
 				\includegraphics[width=0.95\textwidth]{img/dep-graph}
@@
 			\item \textbf{Kazakh} \parencite{fmtjnw2015kazdep}
+			\item work on \textbf{Tuvan} ongoing
 		\end{itemize}
 		\includegraphics[width=0.63\textwidth]{img/korp-2}
-		\begin{itemize}
-			\item work on \textbf{Tuvan} ongoing
-		\end{itemize}
-		\vspace{1.45in}
+		%\begin{itemize}
+		%\end{itemize}
+		\vspace{1.55in}
 		\dividerspace{}

### 2016-05-19 2016-lrec-tyvmorph/poster.tex
and a good first draft of the final version?
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@68639 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 %\pgfsetlayers{background,main,foreground}
+\usepackage{tikz-dependency}
 
@@
 
+\newcommand{\tyv}[1]{\textit{#1}}
+\newcommand{\tgloss}[1]{\tyv{#1}}
+\newfontfamily\udfont[Scale=1.3,Letters=SmallCaps]{Linux Libertine O}
+\newcommand{\udtag}[1]{{\ll \textsc{#1}}}
+\newcommand{\udlabel}[1]{{\udfont #1}}
+
 %\newfontfamily\htwo[IPA,Scale=1.2]{FreeSans}}
@@
 \renewcommand*{\nameyeardelim}{\addcomma\space}
-\renewcommand*{\bibfont}{\fontsize{5.75pt}{5.75pt}\selectfont}
+\renewcommand*{\bibfont}{\fontsize{5.735pt}{5.735pt}\selectfont}
 \renewcommand*{\intitlepunct}{\space}
@@
 		\centering
-		\begin{tikzpicture}[xscale=1.45,yscale=2.2]%, auto]
+		\begin{tikzpicture}[xscale=1.45,yscale=2.212]%, auto]
 
@@
 		\newline
-			\tikzoverlay[text width=0.9\textwidth] at (0.55in, -1in) {
+			\tikzoverlay[text width=0.9\textwidth] at (0.55in, -2.25in) {
 				\includegraphics[width=0.95\textwidth]{img/dep-graph}
@@
 		\end{itemize}
-		\includegraphics[width=0.63\textwidth]{img/korp-2}
+		\resizebox{\textwidth}{!}{
+        \begin{dependency}[theme = simple, font = \small]
+           \begin{deptext}[column sep=0.08cm]
+%               1         2         3          4              5          6          7        8
+                %Мен \& тыва \& хүреш \& -биле \& сургуул \& кылбастай \& бээримге \& ачам \& хомудаан \& . \\
+                Бис \& кожууннуң \& соңгукчулары \& -биле \& ажылды \& чорутпастай \& бээривиске, \& көргүзүглер \& баксыраан \& . \\
+%                \gmk{prn} \& \gmk{n.attr} \& \gmk{n.nom} \& \gmk{post} \& \gmk{n} \& \gmk{v.tv.prc} \& \gmk{vaux} \& \gmk{n.px1sg.nom}  \& \gmk{v.iv.past} \&  \\
+                \gmk{prn} \& \gmk{n.gen} \& \gmk{n.pl.px3sp} \& \gmk{post} \& \gmk{n.acc} \& \gmk{v.tv.prc} \& \gmk{vaux} \& \gmk{n.pl.nom}  \& \gmk{v.iv.past} \&  \\
+                \tgloss{We} \& \tgloss{district} \& \tgloss{constituents} \& \tgloss{with} \& \tgloss{business} \& \tgloss{having stopped conducting} \& \tgloss{} \& \tgloss{figures} \& \tgloss{worsened} \& \\
+                \end{deptext}
+                \depedge{6}{1}{\udlabel{nsubj}}
+                \depedge{3}{2}{\udlabel{nmod:poss}}
+                \depedge{3}{4}{\udlabel{case}}
+                \depedge{6}{3}{\udlabel{nmod}}
+                \depedge{6}{5}{\udlabel{dobj}}
+                \depedge{6}{7}{\udlabel{aux}}
+                \depedge{9}{8}{\udlabel{nsubj}}
+                \depedge{9}{6}{\udlabel{advcl}}
+                \depedge{9}{10}{\udlabel{sent}}
+                \deproot[edge unit distance=3ex]{9}{\udlabel{root}}
+        \end{dependency}}\\
+		\setlength\fboxsep{0.5em}
+		\setlength\fboxrule{0pt}
+		\fbox{\includegraphics[width=0.63\textwidth]{img/korp-2}}
 		%\begin{itemize}
 		%\end{itemize}
-		\vspace{1.55in}
+		\vspace{1.8625in}
+
 		\dividerspace{}
@@
 				%\item Other Turkic lgs:\ Uzbek, Uyghur, Chuvash, Yakut, Tuvan, etc.
-				\includegraphics[width=\textwidth]{img/tyv-spelling}
+				\vspace{0.3em}
+				\includegraphics[width=\textwidth]{img/tyv-spelling-cropped}
+				\vspace{-0.1in}
 		}

### 2016-05-19 2016-lrec-tyvmorph/poster.tex
a few updates and changes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@68640 72bbbca6-d526-0410-a7d9-f06f51895060

@@
 			\item Sayan Turkic language
-			\item Spoken in the Tuva Republic
+			\item Spoken in Tuva Republic and nearby
 			\item Approximately 300,000 speakers
@@
 		\begin{itemize}
-			\item Efficient (in speed \& size) models of a language's morphology
+			\item Computational models of a language's morphology
+			\item Efficient in speed \& size
 			\item Take a surface form, and produce valid lexical form(s)
-			\item Take a lexical form, and produce valid surface form(s)\vspace{-0.1ex}\\
-		%\end{itemize}
+			\item Take a lexical form, and produce valid surface form(s)%\vspace{-0.1ex}\\
+		\end{itemize}
 		%\scalebox{0.98}{
@@
 		%}
-		\end{itemize}
+		%\end{itemize}
 		\dividerspace{}
@@
 			\item \textbf{morphotactics} implemented in \texttt{lexc}
-			\item \textbf{morphophonology} implemented in \texttt{twol} (SPE-style rules)
-			\item compiled separately; compose-intersected to single transducer\vspace{-0.1ex}\\
+			\item \textbf{morphophonology} implemented in \texttt{twol}
+			\item compiled separately; compose-intersected to single transducer%\vspace{-0.1ex}\\
+		\end{itemize}
 			алдым\hfill{}{\qipa \hilitetwo{↔}}\hfill{}\texttt{ал\hilitegrey{>}\archiphon{D}\archiphon{I}\hilitegrey{>}м}\hfill{}{\qipa \hilitetwo{↔}}\hfill{}\texttt{ал\tag{v}\tag{tv}\tag{ifi}\tag{p1}\tag{sg}}\vspace{-0.1ex}\\
@@
 			өөнде\hfill{}{\qipa \hilitetwo{↔}}\hfill{}\texttt{өг\hilitegrey{>}\archiphon{z}\archiphon{I}\archiphon{n}\hilitegrey{>}\archiphon{D}\archiphon{A}}\hfill{}{\qipa \hilitetwo{↔}}\hfill{}\texttt{өг\tag{n}\tag{px3sp}\tag{loc}}
+		%\end{itemize}
+		\begin{itemize}
+			\item mostly hand-written; some parts automatable
 		\end{itemize}
@@
 		\centering
-		\begin{tikzpicture}[xscale=1.45,yscale=2.212]%, auto]
+		\begin{tikzpicture}[xscale=1.45,yscale=2.19]%, auto]
 

### 2021-03-28 2015-rusaccent-nodalida/rusaccent.tex
Rename; Backup

--- 2015-03-19 rusaccent-nodalida-2015/rusaccent.tex
@@
 \usepackage{fixltx2e} %allows subscripts
+\usepackage{tipa} %allows IPA symbols
 %\usepackage{mathptmx}
@@
   unknown words. These results highlight the need for morphosyntactic 
-  disambiguation in the wordstress placement task for Russian, and set a standard 
+  disambiguation in the word stress placement task for Russian, and set a standard 
   for future research on this task.
@@
 
-\rr{REV3: You should use "gold-standard" consistently throughout (sometimes the authors
-write "gold" to mean "gold-standard", which sounds strange.)}
-\rr{Anya: wordstress or word stress?}
 \section{Introduction}
+
 Lexical stress and its attendant vowel reduction are a prominent feature of 
@@
 mark stress on words that cannot be disambiguated through context. Theoretically, 
-a perfect wordstress placement system could help an author identify
+a perfect word stress placement system could help an author identify
 tokens which should be stressed for natives: any token that cannot
@@
 Without information about lexical stress position, 
-correctly converting written Russian text to speech is impossible. This is true 
+correctly converting written Russian text to speech is impossible. Half of the
+vowel letters in Russian change their pronunciation significantly, depending on
+their position relative to the stress. For example the word \emph{dogovórom}
+`contract.SG-INS' is pronounced as /\textipa{d@g2vOr@m}/. Determining such vowel
+qualities is impossible without specifying the stress position. This is a problem 
 both for humans (e.g. foreign language students) and computers (e.g. 
-text-to-speech).\rr{add something about how TTS systems handle this right now. Just no one evaluates how well they are doing.}
+text-to-speech).
 
-We identify three different types of word stress ambiguity.
-\emph{Intraparadigmatic} stress ambiguity refers
-to homographic wordforms belonging to the same lexeme, as shown in 
-\ref{ex:intrahom}.
+We identify three different types relations between word stress ambiguity and
+morphosyntactic ambiguity. First, \emph{intraparadigmatic} stress ambiguity 
+refers to homographic wordforms belonging to the same lexeme, as shown in 
+\ref{ex:intrahom}.\footnote{Cyrillic is transliterated using the scientific transliteration
+scheme.}
 
@@
 \ex. Intraparadigmatic homographs \label{ex:intrahom}
-\a. \rus{т\'{е}ла} \emph{t\'{e}la} `body.\textsubscript{SG-GEN}' 
+\a. \emph{t\'{e}la} `body.\textsubscript{SG-GEN}' 
     \label{ex:bodySGGEN}
-\b. \rus{тел\'{а}} \emph{tel\'{a}} `body.\textsubscript{PL-NOM}' 
+\b. \emph{tel\'{a}} `body.\textsubscript{PL-NOM}' 
     \label{ex:bodyPLNOM}
@@
 \ex. Morphosyntactically incongruent homographs \label{ex:MSincongruent}
-\a. \rus{н\'{а}шей} \emph{nášej} `our.\textsubscript{F-SG-GEN/DAT/LOC...}'\\
-    \rus{наш\'{е}й} \emph{našéj} `sew on.\textsubscript{IMP-2SG}'
-\b. \rus{дор\'{о}га} \emph{doróga} `road.\textsubscript{N-F-SG-NOM}'\\
-    \rus{дорог\'{а}} \emph{dorogá} `dear.\textsubscript{ADJ-F-SG-PRED}'
+\a. \emph{nášej} `our.\textsubscript{F-SG-GEN/LOC/DAT/INS}'\\
+    \emph{našéj} `sew on.\textsubscript{IMP-2SG}'
+\b. \emph{doróga} `road.\textsubscript{N-F-SG-NOM}'\\
+    \emph{dorogá} `dear.\textsubscript{ADJ-F-SG-PRED}'
 
@@
 and whose morphosyntactic values are identical, as shown in
-\ref{ex:MScongruent}. This kind of stress ambiguity is relatively rare, and resolving this ambiguity would require the use of technologies such as Word Sense Disambiguation.
-%\rr{cite Tore's book on stress?}
+\ref{ex:MScongruent}. This kind of stress ambiguity is relatively rare, 
+and resolving this ambiguity would require the use of technologies such as Word Sense Disambiguation.
 
 \ex. Morphosyntactically congruent homographs \label{ex:MScongruent}
-\a. \rus{з\'{a}мок} \emph{z\'{a}mok} `castle.\textsubscript{SG-NOM}'\\
-	\rus{зам\'{о}к} \emph{zam\'{o}k} `lock.\textsubscript{SG-NOM}'
-\b. \rus{з\'{a}мков} \emph{z\'{a}mkov} `castle.\textsubscript{PL-GEN}'\\
-	\rus{замк\'{о}в} \emph{zamk\'{o}v} `lock.\textsubscript{PL-GEN}'
+\a. \emph{z\'{a}mok} `castle.\textsubscript{SG-NOM}'\\
+	\emph{zam\'{o}k} `lock.\textsubscript{SG-NOM}'
+\b. \emph{z\'{a}mkov} `castle.\textsubscript{PL-GEN}'\\
+	\emph{zamk\'{o}v} `lock.\textsubscript{PL-GEN}'
 \c. ...\\
@@
 
-It should be noted that not all morphosyntactic ambiguity leads to stress 
-ambiguity. For example, \emph{novyj} `new' (and every other adjective) has 
-identical forms for F-SG-GEN, F-SG-LOC, F-SG-DAT and F-SG-INS: \emph{n\'{o}voj}. 
-Likewise, the form \emph{vypej} has multiple possible readings 
-(including `drink.IMP', `bittern.PL-GEN'), but they all have the same stress 
-position: \emph{v\'{y}pej}. We refer to this as \emph{stress-irrelevant} 
+It should be noted that most morphosyntactic ambiguity in unrestricted text
+does not result in stress ambiguity. For example, \emph{novyj} `new' (and every 
+other adjective) has identical forms for F-SG-GEN, F-SG-LOC, F-SG-DAT and F-SG-
+INS: \emph{n\'{o}voj}. Likewise, the form \emph{vypej} has multiple possible 
+readings (including `drink.IMP', `bittern.PL-GEN'), but they all have the same 
+stress position: \emph{v\'{y}pej}. We refer to this as \emph{stress-irrelevant} 
 morphosyntactic ambiguity, since all readings have the same stress placement.
@@
 ambiguity to future work because it is by far the least common type of stress
-\rr{Exactly how common is congruent stress ambiguity?
+ambiguity ($\approx$~1.2\% of tokens in running text), and disambiguating morphosyntactically congruent stress requires 
+fundamentally different technology from the other approaches of this study.
+%\rr{Exactly how common is congruent stress ambiguity?
 REVIEW 2: Since this type of ambiguity cannot be resolved with your approach, it imposes an upper bound.
 So, it would be very interesting to see how many examples of morphosyntactically congruent stress ambiguity you found in the annotated corpus.}
-ambiguity, and disambiguating morphosyntactically congruent stress requires 
-fundamentally different technology from the other approaches of this study.
 
 \subsection{Background and task definition} \label{background}
-\rr{cite diacritic restoration}
+%\rr{cite diacritic restoration}
 Automatic stress placement in Russian is similar to diacritic restoration, a task
@@
 must first be normalized by restoring diacritics.
-\rr{Sasha has citations from a lab in Belorus. Lobanov.}
+%\rr{Sasha has citations from a lab in Belorus. Lobanov.}
 A slightly different situation arises with languages whose standard orthography 
@@
 Russian orthography is similar to Arabic and Hebrew, since its vowel qualities 
-are underspecified in standard orthography. \rr{spell out why stress placement is important for pronunciation.}
+are underspecified in standard orthography.
 
+\rr{add something about how TTS systems handle this right now. Just no one evaluates how well they are doing.}
 Many studies of Russian text-to-speech and automatic speech recognition make note 
 of the difficulties caused by the shortcomings of their stress-marking resources 
-(e.g. \newcite{krivnova_1998}), but very few studies have targeted stress-marking 
+(e.g. \newcite{krivnova-98}), but very few studies have targeted stress-marking 
 itself, and those that do focus on placing stress on unknown words, with almost 
-no attention to the problems that arise with running text. For example, 
-\newcite{Xomicevic_2008} developed a set of heuristics for guessing stress 
+no attention to the problems that arise with known words in running text. For example, 
+\newcite{Xomicevic.Rybin.ea-08} developed a set of heuristics for guessing stress 
 placement on unknown words in Russian. More recently, 
-\newcite{hall_sproat_russian_2013} trained a maximum entropy model on a 
+\newcite{Hall.Sproat-13} trained a maximum entropy model on a 
 dictionary of Russian words, and evaluated on wordlists containing `known' and 
-`unknown' wordforms.\footnote{\newcite{hall_sproat_russian_2013} randomly 
+`unknown' wordforms.\footnote{\newcite{Hall.Sproat-13} randomly 
 selected their training and test data from a list of wordforms, and so a number 
@@
 test data whose sibling wordforms from the same lexeme were in the training set 
-were categorised as `known' wordforms.} Their model achieved 98.7\% accuracy on 
-known words, and 83.9\% accuracy on unknown words.
-
-\rr{somehow make this point clearer. Machine learning is not possible for stress in running text because large corpora do not exist.}
-One major shortcoming of evaluating on wordlists is that they do not reflect 
-wordform frequency.
-Many of the most problematic stress ambiguities in Russian occur in 
-high-frequency wordforms, so evaluations based on wordlists - as opposed to 
-running text - have inflated results (see discussion in 
-Section~\ref{discussion}). Furthermore, working with running text allows for the
+were categorized as `known' wordforms.} Their model achieved 98.7\% accuracy on 
+known words, and 83.9\% accuracy on unknown words. The task of training and 
+evaluating on wordlists is different from that of placing
+stress in running text, so these results are not comparable to the results reported
+in this paper. Since many of the most problematic stress ambiguities in 
+Russian occur in high-frequency wordforms, evaluations of wordform lists encounter
+stress ambiguity seven times less frequently than in running text (see discussion in 
+Section~\ref{discussion}). Furthermore, working with running text includes the
 possibility of disambiguating homographs based on syntactic context.
-\rr{REVIEWER 3: Why are results "inflated"?
-if the system does better on more rare forms, then results are "deflated".  if
-it does better on more frequent forms, they are "inflated".   This seems to be
-just a different kind of evaluation.
-It seems that the point of evaluating on running text ought to be precisely
-that the same text could have an ambiguous word with more than one stress
-pattern. So this is really evaluating a different task.}
 
@@
   \begin{tabular}{r|c|c}
-    & \rus{кость}-N-F-SG-GEN\hskip 1em\rus{к\'{о}сти} & \rus{кость}-N-F-SG-GEN\hskip 1em\rus{к\'{о}сти} \\
-    & \rus{кость}-N-F-SG-DAT\hskip 1em\rus{к\'{о}сти} & \rus{кость}-N-F-SG-DAT\hskip 1em\rus{к\'{о}сти} \\
-    & \rus{костить}-V-IPFV-IMP\hskip 1em\rus{кост\'{и}} & \\
+%    & Reading set 1 & Reading set 2 \\
+%    \hline
+   Readings: & \rus{кость}-N-F-SG-GEN\hskip 1em\emph{kósti} & \rus{кость}-N-F-SG-GEN\hskip 1em\emph{kósti} \\
+    & \rus{кость}-N-F-SG-DAT\hskip 1em\emph{kósti} & \rus{кость}-N-F-SG-DAT\hskip 1em\emph{kósti} \\
+    & \rus{костить}-V-IPFV-IMP~~\emph{kostí} & \\
     \hline
-    {\small {\tt bare}} & \rus{кости} & \rus{кости} \\
-    {\small {\tt safe}} & \rus{кости} & \rus{к\'{о}сти} \\
-    {\small {\tt guess}} & \rus{к\'{о}сти} ($p$=0.67) or \rus{кост\'{и}} ($p$=0.33) & \rus{к\'{о}сти}\\
-    {\small {\tt guessFreq}} & \rus{к\'{о}сти} & \rus{к\'{о}сти}
+    {\small {\tt bare}} & \emph{kosti} & \emph{kosti} \\
+    {\small {\tt safe}} & \emph{kosti} & \emph{kósti} \\
+    {\small {\tt randReading}} & \emph{kósti} ($p$=0.67) or \emph{kostí} ($p$=0.33) & \emph{kósti}\\
+    {\small {\tt freqReading}} & \emph{kósti} & \emph{kósti}
   \end{tabular}
-  \caption{Example output of each stress placement approach, given a particular set of readings for the token \rus{кости} \emph{kosti}.}
+  \caption{Example output of each stress placement approach, given a particular set of readings for the token \emph{kosti}}
   \label{tab:conditions}
 \end{table*}
+
 \subsection{Stress corpus}
-\rr{imperative kosti is weird}
-\rr{Table 2 needs more explanation: what exactly these two columns mean and how you ended up with these two sets of readings}
 
 Russian texts with marked word stress are relatively rare, except in materials
-for language learners, which are predominantly proprietary. Our gold corpus was 
+for second language learners, which are predominantly proprietary. Our gold-standard corpus was 
 collected from
@@
 the X data repository (URL will be added in the final version).} This small
+\rr{ADD TO TROLLING!}
 corpus (7689 tokens) is representative of texts 
 that learners of Russian are likely to encounter in their studies. These texts 
-include excerpts from well-known literary works, as well as dialogs and 
-prose and individual sentences that were written for learners. \rr{1st and 2nd language learners? Sometimes children's textbooks have them, especially when syllables are separated.}
-\rr{Why not RNC corpus for gold corpus?}
+include excerpts from well-known literary works, as well as dialogs, 
+prose, and individual sentences that were written for learners. 
+%\rr{Why not RNC corpus for gold-standard corpus?}
 
@@
 \emph{not} mark stress on monosyllabic tokens, effectively assuming that
-all monosyllabics are stressed.\rr{citation?} However, this approach 
+all monosyllabics are stressed. However, this approach 
 is not well-motivated. Many words -- both monosyllabic and multisyllabic -- are 
@@
 Furthermore, there are many
-\rr{examples}
 high-frequency monosyllabic homographs that can be either stressed or unstressed, 
-depending on their part of speech, or particular collocations.\rr{examples} Clearly, for
-such words, one cannot assume that they are stressed on the basis of their 
-syllable count.
-\rr{monos are determining WHETHER stress is present. slightly different.}
+depending on their part of speech, or particular collocations. For example,
+the token \emph{čto} is stressed when it means `what' and unstressed in the
+conjunction \emph{potomu čto} `because'. For such words, one cannot simply
+assume that they are stressed on the basis of their syllable count.
+%\rr{monos are determining WHETHER stress is present. slightly different. This is very relevant for syllable-guessing algorithms.}
 
 Based on these considerations, we built our tools to mark stress on every
-word, both monosyllabic and multisyllabic. However, because our gold corpus texts 
+word, both monosyllabic and multisyllabic. However, because our gold-standard corpus texts 
 do not mark stress on monosyllabic words, we cannot evaluate our annotation of 
@@
 tools are built to mark secondary stress, we cannot evaluate secondary stress
-marks, since they are absent in our gold corpus.
+marks, since they are absent in our gold-standard corpus.
 
-In order to test our wordstress placement system, we removed all stress marks
-from the gold corpus, then marked stress on the unstressed version using our tools, and then
+In order to test our word stress placement system, we removed all stress marks
+from the gold-standard corpus, then marked stress on the unstressed version using our tools, and then
 compared with the original.
@@
 Our Finite-State Transducer\footnote{Using two-level 
-morphology \cite{Koskenniemi-84}, implemented in both xfst
-\cite{Beesley.Karttunen-03} and hfst \cite{hfst-11}} (FST) generates all possible 
+morphology \cite{Koskenniemi-83,Koskenniemi-84}, implemented in both xfst
+\cite{Beesley.Karttunen-03} and hfst \cite{Linden.Silfverberg.ea-11}} (FST) generates all possible 
 morphosyntactic readings of each wordform, and our Constraint 
@@
 we also evaluate various approaches to deal with the remaining ambiguity,
-as described below.
+as described below. Table~\ref{tab:conditions} shows two possible sets of readings
+for the token \emph{kosti}, as well as the output of each approach described below.
+The first column exhibits stress ambiguity between the noun readings and the imperative
+verb reading. The second column shows a similar set of readings, after the
+constraint grammar has removed the imperative verb reading. This results in only
+stress-irrelevant ambiguity.
 
-\rr{rename guess > guessRand}
 \rr{more about our corpus}
-\rr{explain guess better}
 The {\small {\tt bare}} approach is to not mark stress on words with more than 
@@
 
-The {\small {\tt guess}} approach is to randomly select one of the available 
+The {\small {\tt randReading}} approach is to randomly select one of the available 
 readings. In the first column of
-Table~\ref{tab:conditions}, a random selection means that \rus{к\'{о}сти} 
-\emph{k\'{o}sti} is twice as likely as \rus{кост\'{и}} \emph{kost\'{i}}. The
+Table~\ref{tab:conditions}, a random selection means that \emph{kósti} 
+is twice as likely as \emph{kostí}. The
 second column of Table~\ref{tab:conditions} contains stress-irrelevant
@@
 
-The {\small {\tt guessFreq}} 
+The {\small {\tt freqReading}} 
 approach is to select the reading that is most frequent, with frequency data
 taken from a separate hand-disambiguated corpus. If none of the readings are 
-found in the corpus, then {\small {\tt guessFreq}} selects the reading
+found in the corpus, then {\small {\tt freqReading}} selects the reading
 with the tag sequence (lemma removed) that is most frequent in our corpus. If the 
 tag sequence is not found
-in our frequency list, then {\small {\tt guessFreq}} backs off to the 
-{\small {\tt guess}} algorithm. In the first column of
-Table~\ref{tab:conditions}, {\small {\tt guessFreq}} selects \rus{к\'{о}сти} 
-\emph{k\'{o}sti} because the tag sequence N-F-SG-GEN is more frequent than the
+in our frequency list, then {\small {\tt freqReading}} backs off to the 
+{\small {\tt randReading}} algorithm. In the first column of
+Table~\ref{tab:conditions}, {\small {\tt freqReading}} selects \emph{kósti} 
+because the tag sequence N-F-SG-GEN is more frequent than the
 other alternatives. Note 
 that for tokens with stress-irrelevant ambiguity (e.g. the second column of
-Table~\ref{tab:conditions}), {\small {\tt guess}} and 
-{\small {\tt guessFreq}} produce the same result as the {\small {\tt safe}} 
+Table~\ref{tab:conditions}), {\small {\tt randReading}} and 
+{\small {\tt freqReading}} produce the same result as the {\small {\tt safe}} 
 method.
@@
 in vowel-final words.\footnote{There is some controversy over how to define 
-default stress in Russian. \newcite{Crosswhite.ea-03} conclude that the default 
-stress position in Russian is the last syllable of the stem.} Based on this 
-\rr{what is the difference between last syllable and last vowel followed by consonant?}
-conclusion, our syllable-guessing method places the stress on the last vowel that 
+default stress in Russian, cf. \newcite{Crosswhite.Alderete.ea-03}.} Based on this 
+%\rr{what is the difference between last syllable and last vowel followed by consonant?}
+conclusion, the {\tt guessSyll} method places the stress on the last vowel that 
 is followed by a consonant.\footnote{Although this approach is simplistic, 
@@
 learning approaches to unknown words are discussed in Section~\ref{discussion}.} This method 
-is applied in two approaches, {\small {\tt guessSyll}} 
-and {\small {\tt guessFreqSyll}}, which are otherwise identical to 
-{\small {\tt guess}} and {\small {\tt guessFreq}}, respectively.
-\rr{REVIEWER 2: However, one can imagine a token that gets some morphological readings, none of them being correct (for example, when the correct lexeme is not covered by the dictionary -- but an homonymous incorrect lexeme is).
-Do you have any estimations of how often this happens in the running text?}
+is applied in two approaches, {\small {\tt randReading+guessSyll}} 
+and {\small {\tt freqReading+guessSyll}}, which are otherwise identical to 
+{\small {\tt randReading}} and {\small {\tt freqReading}}, respectively.
 
@@
 Constraint Grammar) in combination with the {\small {\tt bare}}, 
-{\small {\tt safe}}, {\small {\tt guess}}, {\small {\tt guessFreq}}, 
-{\small {\tt guessSyll}}, and {\small {\tt guessFreqSyll}} approaches. We also 
+{\small {\tt safe}}, {\small {\tt randReading}}, {\small {\tt freqReading}}, 
+{\small {\tt randReading+guessSyll}}, and {\small {\tt freqReading+guessSyll}} approaches. We also 
 compare our outcomes with the RussianGram\footnote{http://russiangram.com/} 
@@
     \hline
-    {\small {\tt noCGbare}} & 30.43 & 0.17 & 69.39 & 30.61 & 69.57 \\
-    {\small {\tt noCGsafe}} & 90.07 & 0.49 & 9.44 & 90.56 & 9.93 \\
-    {\small {\tt noCGguess}} & 94.34 & 3.36 & 2.30 & 97.70 & 5.66 \\
-    {\small {\tt noCGguessFreq}} & 95.53 & 2.59 & 1.88 & 98.12 & 4.47 \\
-    {\small {\tt noCGguessSyll}} & 94.99 & 4.05 & 0.96 & 99.04 & 5.01 \\
-    {\small {\tt noCGguessFreqSyll}} & 95.83 & 3.46 & 0.72 & 99.28 & 4.17 \\
+    {\small {\tt noCG+bare}} & 30.43 & 0.17 & 69.39 & 30.61 & 69.57 \\
+    {\small {\tt noCG+safe}} & 90.07 & 0.49 & 9.44 & 90.56 & 9.93 \\
+    {\small {\tt noCG+randReading}} & 94.34 & 3.36 & 2.30 & 97.70 & 5.66 \\
+    {\small {\tt noCG+freqReading}} & 95.53 & 2.59 & 1.88 & 98.12 & 4.47 \\
+    {\small {\tt noCG+randReading+guessSyll}} & 94.99 & 4.05 & 0.96 & 99.04 & 5.01 \\
+    {\small {\tt noCG+freqReading+guessSyll}} & 95.83 & 3.46 & 0.72 & 99.28 & 4.17 \\
     \hline
-    {\small {\tt CGbare}} & 45.78 & 0.44 & 53.78 & 46.22 & 54.22 \\
-    {\small {\tt CGsafe}} & 93.21 & 0.74 & 6.05 & 93.95 & 6.79 \\
-    {\small {\tt CGguess}} & 95.50 & 2.59 & 1.90 & 98.10 & 4.50 \\
-    {\small {\tt CGguessFreq}} & 95.73 & 2.40 & 1.88 & 98.12 & 4.27 \\
-    {\small {\tt CGguessSyll}} & 95.92 & 3.33 & 0.74 & 99.26 & 4.08 \\
-    {\small {\tt CGguessFreqSyll}} & 96.15 & 3.14 & 0.72 & 99.28 & 3.85 \\
+    {\small {\tt CG+bare}} & 45.78 & 0.44 & 53.78 & 46.22 & 54.22 \\
+    {\small {\tt CG+safe}} & 93.21 & 0.74 & 6.05 & 93.95 & 6.79 \\
+    {\small {\tt CG+randReading}} & 95.50 & 2.59 & 1.90 & 98.10 & 4.50 \\
+    {\small {\tt CG+freqReading}} & 95.73 & 2.40 & 1.88 & 98.12 & 4.27 \\
+    {\small {\tt CG+randReading+guessSyll}} & 95.92 & 3.33 & 0.74 & 99.26 & 4.08 \\
+    {\small {\tt CG+freqReading+guessSyll}} & 96.15 & 3.14 & 0.72 & 99.28 & 3.85 \\
     \hline
@@
   \end{tabular}
-  \caption{Results of stress placement task evaluation. (N = 4048)}
+  \caption{Results of stress placement task evaluation}
   \label{tab:results}
@@
 
-\rr{REV3: The presentation of results is unclear: what exactly is  being measured?
-It would be best if the measure was presented on words that have a potential
-stress ambiguity.  (If this includes all words, it should be made clear, or
-more clear.)}
-Single-syllable words and words that are unstressed in the gold standard were
-not evaluated for reasons discussed above, leaving 4048 tokens for evaluation.
+%\rr{REV3: The presentation of results is unclear: what exactly is  being measured?
+It would be best if the measure was presented on words that have a potential stress ambiguity.
+(If this includes all words, it should be made clear, or more clear.)}
+We evaluated all multisyllabic words with marked stress in the gold-standard corpus (N = 4048).
 Since our approach is lexicon-based, some of our results should be interpreted
-with respect to how many of the stressed wordforms in the gold standard corpus
+with respect to how many of the stressed wordforms in the gold-standard corpus
 can be found in the output of the finite-state transducer. We refer to this 
@@
 This number represents the ceiling for methods relying on the FST.
-Higher scores are only achievable by expanding the FST \rr{not in dictionary}or by using 
+Higher scores are only achievable by expanding the FST's lexicon or by using 
 syllable-guessing algorithms. After running the Constraint Grammar, recall was 
-97.35\%, a reduction of 0.20\%. Apparently the Constraint Grammar erroneously 
-removes some correct readings, or else there are errors in the gold corpus.
-\rr{sveta wanted more}
+97.35\%, a reduction of 0.20\%. Our method of computing recall assumes that if 
+even one reading is output by the FST, then all possible readings are present.
+We have not attempted to formally estimate how frequently this assumption fails, 
+but we expect such cases to be quite rare.
+
+%\rr{REVIEWER 2: However, one can imagine a token that gets some morphological readings, none of them being correct (for example, when the correct lexeme is not covered by the dictionary -- but an homonymous incorrect lexeme is).
+Do you have any estimations of how often this happens in the running text?}
 
@@
 and RussianGram ({\small {\tt RussianGram}}). Results are given in 
-Table~\ref{tab:results}. Each token was categorised as either an accurate output, 
+Table~\ref{tab:results}. Each token was categorized as either an accurate output, 
 or one of two categories of failures: errors and abstentions. 
@@
 it is counted as an `abstention'. Abstentions can be result of either unknown
-wordforms, or known wordforms with unspecified stress.
+wordforms, or known wordforms with stress unspecified in our lexicon.
 
-\rr{flow for Syll methods}
 The two right-most columns in 
@@
 proportion of tokens on which our system output a stressed wordform. In the case 
-of {\small {\tt noCGbare}}, the accuracy\% (30.43) and error\% (0.17) sum to the 
+of {\small {\tt noCG+bare}}, the accuracy\% (30.43) and error\% (0.17) sum to the 
 totTry value of 30.61. The term `totFail' refers to the sum of error rate 
@@
 proportion of tokens for which the system failed to output the correct stressed 
-form. In the case of {\small {\tt noCGbare}}, the error\% (0.17) 
+form. In the case of {\small {\tt noCG+bare}}, the error\% (0.17) 
 and abstention\% (69.39) sum to the totFail value of of 69.57 (rounded). 
@@
 
-The {\small {\tt noCGbare}} approach achieves a baseline accuracy of
+The {\small {\tt noCG+bare}} approach achieves a baseline accuracy of
 30.43\%, so roughly two thirds of the tokens in our corpus are
@@
 forms whose stress position varies from speaker to speaker (e.g.
-\rus{зав\'{и}л\'{и}сь} \emph{zav\'{i}l\'{i}s'}), or errors in the gold corpus 
-(e.g. \rus{вер\'{и}м} \emph{ver\'{i}m}).
-\rr{don't count variation as errors, if it is acceptable}
-\rr{why errors?}
+\emph{zavílis'} vs. \emph{zavilís'} `they crinkled'), or errors in the gold-standard corpus 
+(e.g. \emph{verím} `we believe').
+%\rr{don't count variation as errors, if it is acceptable}
+%\rr{why errors?}
 
-The {\small {\tt noCGsafe}} approach
+The {\small {\tt noCG+safe}} approach
 achieves a 60\% improvement in accuracy (90.07\%), which means that 89.39\% of 
@@
 RussianGram web service achieves results that are very close 
-to the {\small {\tt noCGsafe}} approach.
+to the {\small {\tt noCG+safe}} approach.
 
 Since the ceiling recall for the FST is 97.55\%, and since the 
-{\small {\tt noCGsafe}} approach achieves 90.07\%, the maximum improvement 
+{\small {\tt noCG+safe}} approach achieves 90.07\%, the maximum improvement 
 that a Constraint Grammar could theoretically achieve is 7.48\%. A comparison
-of {\small {\tt noCGsafe}} and {\small {\tt CGsafe}} reveals an improvement
+of {\small {\tt noCG+safe}} and {\small {\tt CG+safe}} reveals an improvement
 of 3.14\%, which is about 42\% of the way to the ceiling recall.
@@
 
-The {\small {\tt CGguess}} and {\small {\tt CGguessFreq}} approaches are also 
+The {\small {\tt CG+randReading}} and {\small {\tt CG+freqReading}} approaches are also 
 limited by the 97.55\% ceiling from the FST, and their accuracies achieve
-improvements of 2.29\% and 2.52\%, respectively, over {\small {\tt CGsafe}}.
+improvements of 2.29\% and 2.52\%, respectively, over {\small {\tt CG+safe}}.
 However, these gains come at the cost of error rates as much as 3.5 times 
-higher than {\small {\tt CGsafe}}: +1.85\% and +1.66\%, respectively. It is
-not surprising that {\small {\tt CGguessFreq}} has higher accuracy and a lower
-error rate than {\small {\tt CGguess}}, since frequency-based guesses are by
+higher than {\small {\tt CG+safe}}: +1.85\% and +1.66\%, respectively. It is
+not surprising that {\small {\tt CG+freqReading}} has higher accuracy and a lower
+error rate than {\small {\tt CG+randReading}}, since frequency-based guesses are by
 definition more likely to occur. The frequency data were taken from a very 
@@
 
-The two {\small {\tt Syll}} approaches were designed to make a blind guess on 
+The {\small {\tt guessSyll}} approach was designed to make a blind guess on 
 every wordform that is not found in the FST, which would ideally result in an
 abstention rate of 0\%. However, the abstention rates of approximately 0.7\%
-are the result of the fact that some words in the FST, especially proper nouns, 
-had not been assigned stress. \rr{REV3: how often does FST block guessFreq?}Because the FST outputs a form -- albeit unstressed 
--- the {\small {\tt Syll}} algorithm is not called. This means that 
-{\small {\tt Syll}} is only guessing on about 2\% of the tokens. The improvement
-on overall accuracy from {\small {\tt CGguessFreq}} to 
-{\small {\tt CGguessFreqSyll}} is 0.42\%, which means that the 
-{\small {\tt Syll}} guess was accurate 21\% of the time.
-\rr{Examine errors and categorize/refute them.}
+are a manifestation of the fact that some words in the FST, especially proper nouns, 
+have not been assigned stress. Because the FST outputs a form -- albeit unstressed 
+-- the {\small {\tt guessSyll}} algorithm is not called. This means that 
+{\small {\tt guessSyll}} is only guessing on about 2\% of the tokens. The improvement
+on overall accuracy from {\small {\tt CG+freqReading}} to 
+{\small {\tt CG+freqReading+guessSyll}} is 0.42\%, which means that the 
+{\small {\tt guessSyll}} method guess was accurate 21\% of the time.
+%\rr{Examine errors and categorize/refute them.}
 
@@
 One of the main points of this paper is to highlight the importance of 
-syntactic context in the Russian wordstress placement task. If your intended 
+syntactic context in the Russian word stress placement task. If your intended 
 application has a low tolerance for error, the 
-{\small {\tt noCGsafe}} approach represents the highest accuracy that is possible 
+{\small {\tt noCG+safe}} approach represents the highest accuracy that is possible 
 without leveraging syntactic information for disambiguation (90.07\%). In other words, a 
 system that is blind to morphosyntax and contextual disambiguation cannot
-significantly outperform {\small {\tt noCGsafe}}. It would appear that this is
+significantly outperform {\small {\tt noCG+safe}}. It would appear that this is
 the method used by {\small {\tt RussianGram}}, since its results are so similar
-to {\small {\tt noCGsafe}}. Indeed, this result can be achieved most efficiently
+to {\small {\tt noCG+safe}}. Indeed, this result can be achieved most efficiently
 without any part-of-speech tagging, but through simple dictionary lookup.
 
-We noted in Section~\ref{background} that \newcite{hall_sproat_russian_2013}
+We noted in Section~\ref{background} that \newcite{Hall.Sproat-13}
 achieved 98.7\% accuracy on stress placement for individual wordforms 
 in a list (i.e. \emph{without} syntax). This result is 8.63\% higher than
-{\small {\tt noCGsafe}}, but it is also a fundamentally different task.
+{\small {\tt noCG+safe}}, but it is also a fundamentally different task.
 Based on the surface 
 forms in our FST -- which is based on the same dictionary used for 
-\newcite{hall_sproat_russian_2013} -- 
+\newcite{Hall.Sproat-13} -- 
 we calculate that only 29~518 (1.05\%) 
 of the 2~804~492 wordforms contained in our FST are stress-ambiguous.
-\rr{These stress-ambiguous wordforms have an average of 2.01 possible stress positions, so a perfect stress placement tool has a 50/50 chance of getting them right.
+%\rr{These stress-ambiguous wordforms have an average of 2.01 possible stress positions, so a perfect stress placement tool has a 50/50 chance of getting them right.
 This means that a perfect stress placement tool could achieve 99.48\% on their task.}
 In our corpus of unrestricted text, at least 7.5\% of the tokens are 
-stress-ambiguous. Stress ambiguity is therefore seven times more prevalent in our 
+stress-ambiguous. Therefore, stress ambiguity is more than seven times more prevalent in our 
 corpus of unrestricted text than it is in our wordform dictionary.
-Since the task of wordstress placement is virtually always performed on
+Since the task of word stress placement is virtually always performed on
 running text, it seems prudent to make use of surrounding contextual information.
@@
 
-We are unaware of any other empirical evaluations of Russian wordstress placement 
+We are unaware of any other empirical evaluations of Russian word stress placement 
 in unrestricted text. The results of our experiment are promising, but many 
 questions remain unanswered.
-The experiment was limited by properties of the gold standard corpus, including
+The experiment was limited by properties of the gold-standard corpus, including
 its size, genre distribution, and quality.
-%Ideally, a gold standard corpus for Russian wordstress should mark whether monosyllabic words are stressed, and of course its annotations should be accurate.
-Our gold corpus represents a broad variety of text genres, which 
+%Ideally, a gold-standard corpus for Russian word stress should mark whether monosyllabic words are stressed, and of course its annotations should be accurate.
+Our gold-standard corpus represents a broad variety of text genres, which 
 makes our results more generalizable, but a larger corpus would allow for
@@
 since it contains a higher proportion of borrowed words, calques, and neologisms 
-with simple stress patterns.\lj{What about NomPl/AccPl -a spreading in tech jargon?}
+with simple stress patterns.
+%\lj{What about NomPl/AccPl -a spreading in tech jargon?}
 
@@
 
-The syllable-guessing algorithm {\small {\tt Syll}} used in this 
+The syllable-guessing algorithm {\small {\tt guessSyll}} used in this 
 experiment was overly simplistic, and so it was not surprising that it was only 
 moderately successful. More rigorous rule-based approaches have been suggested in 
-other studies \cite{Church-85,Williams-87,Xomicevic_2008}. For example, 
-\newcite{Xomicevic_2008} attempt to parse the unknown token by matching known 
+other studies \cite{Church-85,Williams-87,Xomicevic.Rybin.ea-08}. For example, 
+\newcite{Xomicevic.Rybin.ea-08} attempt to parse the unknown token by matching known 
 prefixes and suffixes.
@@
 Other studies have applied machine-learning to guessing stress of unknown words
-\cite{Pearson.ea-00,Webster-04,Dou.ea-09,hall_sproat_russian_2013}. For example,
-for unknown words, \newcite{hall_sproat_russian_2013} achieve an accuracy of 
-83.9\%. Their model was trained on a full list of \rr{COULD?}
+\cite{Pearson.Kuhn.ea-00,Webster-04,Dou.Bergsma.ea-09,Hall.Sproat-13}. For example,
+for unknown words, \newcite{Hall.Sproat-13} achieve an accuracy of 
+83.9\%. Their model was trained on a full list of
 Russian words, which is not representative of the words that would be unknown to 
-a system like ours, so we could modify their approach to fit our application. 
-Most of the complicated wordstress patterns are closed 
+a system like ours, so it would be possible modify their approach to fit our application. 
+Most of the complicated word stress patterns are closed 
 classes\footnote{The growing number of masculine nouns with shifting stress 
-(\rus{д\'{о}ктор}$\sim$\rus{доктор\'{а}} \emph{d\'{o}ktor$\sim$doktor\'{a}} 
-`doctor'$\sim$`doctors') is one exception to this generalization.},
+(\emph{d\'{o}ktor$\sim$doktor\'{a}} `doctor'$\sim$`doctors') is one exception 
+to this generalization.},
 so we could exclude closed classes of words from the training data, leaving only
@@
 We have demonstrated the effectiveness of using a Constraint Grammar to improve
-the results of a Russian wordstress placement task in unrestricted text by 
-resolving 42\% of the stress ambiguity in our gold corpus. We showed
+the results of a Russian word stress placement task in unrestricted text by 
+resolving 42\% of the stress ambiguity in our gold-standard corpus. We showed
 that stress ambiguity is seven times more prevalent in our corpus of running text
@@
 collecting stressed corpora of varying text complexity and/or genre, as well as
-implementing and/or adapting established wordstress-guessing methods for unknown 
+implementing and/or adapting established word stress-guessing methods for unknown 
 words.
 
-%\section*{Acknowledgments}
-%
-%Do not number the acknowledgment section. Do not include this section
-%when submitting your paper for review.
+\section*{Acknowledgments}
+
+We are very grateful to Laura Janda, Tore Nesset and other members of the CLEAR
+research group at UiT The Arctic University of Norway for comments on an
+earlier draft of this paper. We are also grateful to three anonymous reviewers
+for thoughtful criticism. All remaining errors are our own.
 
 \bibliographystyle{acl}
-\bibliography{rusaccent}
+\bibliography{icall}
 

### 2021-05-27 2015-compmorphon-morphver/poster.tex
OFTC, http -> https, url fixups

@@
 			\begin{itemize}
-				\item Part of \textbf{Apertium Turkic} project:\ \url{http://wiki.apertium.org/wiki/Apertium\_Turkic}
+				\item Part of \textbf{Apertium Turkic} project:\ \url{https://wiki.apertium.org/wiki/Apertium\_Turkic}
 				\item Transducers available \textbf{live} (\href{http://turkic.apertium.org/}{\texttt{turkic.apertium.org}}), \textbf{source code} under \hiliteone{GPL} from Apertium's SVN repo
@@
 				%\item \textbf{Source code} of transducers available under \hiliteone{GPL} (free!) from Apertium's SVN repo
-				%info at {\small \url{http://wiki.apertium.org/wiki/apertium-kir}}
+				%info at {\small \url{https://wiki.apertium.org/wiki/apertium-kir}}
 				\item Multilingual Turkic RBMT \textbf{mailing list} (>25 subscribers):\ \texttt{apertium-turkic@lists.sourceforge.net}%\\

### 2022-06-09 2022-lrec-sahmorph/img/graphs.r
2022-lrec-sahmorph

@@ -1,0 +1,65 @@
+require(data.table)
+library(ggplot2)
+#library(dplyr)
+
+make_graph <- function(name, xax=T, yax=T, leg=T, asp_ratio=0.8) {
+	lgs1<-as.data.frame(fread(paste(name, ".tsv", sep="")))
+	print(lgs1)
+
+	types = c("Nouns", "Proper Nouns", "Verbs", "Adjectives", "Other")
+	allLgs = c("tat", "sah", "kaz", "tur", "kir", "crh", "tyv",
+		"chv","uzb","bak","kaa","uig","krc","gag","kum",
+		"aze","tki","tuk","nog","kjh","alt","ota")
+
+	lgs1_new <- data.frame(language=factor(levels=allLgs), type=factor(levels=rev(types)), value=integer(),stringsAsFactors=FALSE)
+	#print(lgs1_new)
+
+	for (lg in lgs1$language) {
+		for (type in types) {
+			#print(lgs1[[type]])
+			value = lgs1[[type]][lgs1$language==lg]
+			#new_row = c(lg,type,value)
+			new_row = list(language=c(lg),type=type,value=value)
+			lgs1_new[nrow(lgs1_new) + 1,] <- new_row
+		}
+	}
+	warnings()
+	print(lgs1_new)
+
+	#lgs1_new <- lgs1_new %>% arrange(n)
+	
+	#lgs1_new$type <- as.character(lgs1_new$type)
+	#lgs1_new$type <- factor(lgs1_new$type, levels=types)
+	
+	aspect_ratio <- asp_ratio
+	height <- 4
+
+	if (yax) {
+		yaxTh <- theme(axis.title.y=element_text(size=rel(1.4)))
+	} else {
+		yaxTh <- theme(axis.title.y=element_blank())
+	}
+	if (xax) {
+		xaxTh <- theme(axis.title.x=element_text(size=rel(1.4)))
+	} else {
+		xaxTh <- theme(axis.title.x=element_blank())
+	}
+	if (leg) {
+		legTh <- theme(legend.position = c(0.6, 0.75))
+	} else {
+		legTh <- theme(legend.position = "none")
+	}
+
+	p<-ggplot(data=lgs1_new)+ aes(fill=type,x=language,y=value/1000) +#, aes(fill=types))+
+		geom_bar(position="stack", stat="identity") +
+		labs(y="stems (×1000)",fill="category") +
+		legTh +
[…]

### 2022-06-09 2022-lrec-sahmorph/img/qrcode.sh

@@ -1,0 +1,6 @@
+#qrencode -o qrcode.png -l M "https://github.com/jonorthwash/ultrasound-processing"
+
+qrencode -o qrcode.svg -t SVG -l M "https://github.com/apertium/apertium-sah"
+inkscape qrcode.svg --export-filename=qrcode.pdf
+
+
